---
title: "Inflate Match Parameter Analysis"
author: "test_susine"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.show = "hold")
library(ggplot2)
library(dplyr)
library(tidyr)
```

# User Options

```{r user-options}
# ==== USER OPTIONS ====

# Save plots to files instead of displaying in workbook?
# Set to TRUE when running on HPC or when you want PNG outputs
save_plots_to_file <- FALSE

# Output directory for saved plots (relative to workbook location)
plot_output_dir <- here::here("output", "inflate_match_analysis_plots")

# Create output directory if saving plots
if (save_plots_to_file) {

  dir.create(plot_output_dir, showWarnings = FALSE, recursive = TRUE)
  cat("Plots will be saved to:", plot_output_dir, "\n")
}

# Helper function to save or display a plot
save_or_show <- function(p, filename, width = 10, height = 7, dpi = 300) {
  if (save_plots_to_file) {
    filepath <- file.path(plot_output_dir, filename)
    ggsave(filepath, p, width = width, height = height, dpi = dpi, bg = "white")
    cat("Saved:", filename, "\n")
  } else {
    print(p)
  }
}
```

---

# Overview

This workbook analyzes the `inflate_match` parameter in `simulate_data.R` to verify it works as intended.

**Goal of `inflate_match`**: Scale up the variance of non-causal SNP annotations (mu_0) to approach the same scale as causal SNP annotations. This prevents the model from trivially distinguishing causal from non-causal SNPs based on annotation magnitude alone.

---

# Section 1: Code Review

## 1.1 The `simulate_priors()` Function

Let's examine the key logic:
```r
# For causal SNPs:
mu_0[causal_idx] <- beta[causal_idx] + rnorm(length(causal_idx), sd = sqrt(noise_var))

# For non-causal SNPs:
target_var <- noise_var + inflate_match * causal_var
mu_0[noncausal_idx] <- rnorm(length(noncausal_idx), sd = sqrt(target_var))
```

**Key variables:**
- `causal_var` = Var(beta[causal]) - variance of the true causal effects
- `noise_var` = causal_var * (1 - annotation_r2) / annotation_r2 - noise added to annotations
- `target_var` = noise_var + inflate_match * causal_var - variance for non-causal mu_0

**Expected behavior:**

For **causal SNPs**, the variance of mu_0 is:
$$\text{Var}(\mu_0^{causal}) = \text{Var}(\beta_{causal}) + \text{noise\_var} = \text{causal\_var} + \text{noise\_var}$$

For **non-causal SNPs**, the variance of mu_0 is:
$$\text{Var}(\mu_0^{noncausal}) = \text{target\_var} = \text{noise\_var} + \text{inflate\_match} \times \text{causal\_var}$$

**When `inflate_match = 1`:**
$$\text{Var}(\mu_0^{noncausal}) = \text{noise\_var} + \text{causal\_var} = \text{Var}(\mu_0^{causal})$$

So at `inflate_match = 1`, both causal and non-causal SNPs should have the same mu_0 variance!

---

# Section 2: Simulation Experiment

## 2.1 Setup

```{r simulation-setup}
# Load the package functions
devtools::load_all(here::here())

# Simulation parameters
set.seed(42)
n_snps <- 1000      # Number of SNPs
n_causal <- 50      # Number of causal SNPs
effect_sd <- 1      # SD of causal effects
annotation_r2_values <- c(0.2, 0.4, 0.6, 0.8)  # Annotation quality levels
inflate_match_values <- seq(0, 1.5, by = 0.25)  # Range of inflate_match to test
n_reps <- 100       # Number of replications per setting
```

## 2.2 Run Simulation

```{r run-simulation}
# Storage for results
results <- list()

for (annot_r2 in annotation_r2_values) {
  for (inflate in inflate_match_values) {
    for (rep in seq_len(n_reps)) {
      # Simulate effect sizes
      effects <- simulate_effect_sizes(
        p = n_snps,
        p_star = n_causal,
        effect_sd = effect_sd,
        seed = rep * 1000 + round(annot_r2 * 100) + round(inflate * 100)
      )
      
      # Simulate priors (this is where inflate_match is used)
      priors <- simulate_priors(
        beta = effects$beta,
        annotation_r2 = annot_r2,
        inflate_match = inflate,
        gamma_shrink = NA,  # No shrinkage for this test
        base_sigma2 = NULL,
        effect_sd = effect_sd
      )
      
      # Extract mu_0 for causal and non-causal
      mu_0 <- priors$mu_0
      causal_idx <- effects$causal_idx
      noncausal_idx <- setdiff(seq_len(n_snps), causal_idx)
      
      # Compute statistics
      results[[length(results) + 1]] <- data.frame(
        annotation_r2 = annot_r2,
        inflate_match = inflate,
        rep = rep,
        # Causal SNP statistics
        causal_mu0_mean = mean(mu_0[causal_idx]),
        causal_mu0_var = var(mu_0[causal_idx]),
        causal_mu0_abs_mean = mean(abs(mu_0[causal_idx])),
        causal_beta_var = var(effects$beta[causal_idx]),
        # Non-causal SNP statistics
        noncausal_mu0_mean = mean(mu_0[noncausal_idx]),
        noncausal_mu0_var = var(mu_0[noncausal_idx]),
        noncausal_mu0_abs_mean = mean(abs(mu_0[noncausal_idx])),
        # Ratio
        var_ratio = var(mu_0[noncausal_idx]) / var(mu_0[causal_idx])
      )
    }
  }
}

# Combine results
sim_results <- dplyr::bind_rows(results)

cat("Simulation complete:", nrow(sim_results), "rows\n")
cat("Settings tested:\n")
cat("  annotation_r2:", paste(annotation_r2_values, collapse = ", "), "\n")
cat("  inflate_match:", paste(inflate_match_values, collapse = ", "), "\n")
```

## 2.3 Summarize Results

```{r summarize-results}
# Aggregate across replications
summary_stats <- sim_results %>%
  group_by(annotation_r2, inflate_match) %>%
  summarise(
    # Means across reps
    causal_var_mean = mean(causal_mu0_var),
    causal_var_se = sd(causal_mu0_var) / sqrt(n()),
    noncausal_var_mean = mean(noncausal_mu0_var),
    noncausal_var_se = sd(noncausal_mu0_var) / sqrt(n()),
    var_ratio_mean = mean(var_ratio),
    var_ratio_se = sd(var_ratio) / sqrt(n()),
    causal_abs_mean = mean(causal_mu0_abs_mean),
    noncausal_abs_mean = mean(noncausal_mu0_abs_mean),
    .groups = "drop"
  )

print(summary_stats, n = 50)
```

---

# Section 3: Visualizations

## 3.1 Variance Comparison: Causal vs Non-Causal

```{r plot-variance-comparison, fig.width=12, fig.height=8}
# Reshape for plotting
var_long <- summary_stats %>%
  select(annotation_r2, inflate_match, causal_var_mean, noncausal_var_mean) %>%
  pivot_longer(
    cols = c(causal_var_mean, noncausal_var_mean),
    names_to = "snp_type",
    values_to = "variance"
  ) %>%
  mutate(
    snp_type = ifelse(snp_type == "causal_var_mean", "Causal", "Non-causal")
  )

plt <- ggplot(var_long, aes(x = inflate_match, y = variance, color = snp_type)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ annotation_r2, labeller = label_both, scales = "free_y") +
  labs(
    title = "Variance of mu_0 by SNP Type and inflate_match",
    subtitle = "At inflate_match=1, variances should be equal",
    x = "inflate_match",
    y = "Var(mu_0)",
    color = "SNP Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  geom_vline(xintercept = 1, linetype = "dashed", alpha = 0.5)

save_or_show(plt, "01_variance_comparison.png", width = 12, height = 8)
```

## 3.2 Variance Ratio (Non-causal / Causal)

```{r plot-variance-ratio, fig.width=10, fig.height=6}
plt <- ggplot(summary_stats, aes(x = inflate_match, y = var_ratio_mean, color = factor(annotation_r2))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_ribbon(
    aes(ymin = var_ratio_mean - 2*var_ratio_se, ymax = var_ratio_mean + 2*var_ratio_se, fill = factor(annotation_r2)),
    alpha = 0.2, color = NA
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  labs(
    title = "Variance Ratio: Var(mu_0 non-causal) / Var(mu_0 causal)",
    subtitle = "Red dashed line = 1.0 (equal variance). Should cross ~1 at inflate_match=1",
    x = "inflate_match",
    y = "Variance Ratio",
    color = "annotation_r2",
    fill = "annotation_r2"
  ) +
  theme_minimal() +
  theme(legend.position = "right") +
  scale_y_continuous(breaks = seq(0, 2, by = 0.25))

save_or_show(plt, "02_variance_ratio.png", width = 10, height = 6)
```

## 3.3 Mean Absolute mu_0 Comparison

```{r plot-abs-mean, fig.width=12, fig.height=8}
abs_long <- summary_stats %>%
  select(annotation_r2, inflate_match, causal_abs_mean, noncausal_abs_mean) %>%
  pivot_longer(
    cols = c(causal_abs_mean, noncausal_abs_mean),
    names_to = "snp_type",
    values_to = "abs_mean"
  ) %>%
  mutate(
    snp_type = ifelse(snp_type == "causal_abs_mean", "Causal", "Non-causal")
  )

plt <- ggplot(abs_long, aes(x = inflate_match, y = abs_mean, color = snp_type)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ annotation_r2, labeller = label_both) +
  labs(
    title = "Mean |mu_0| by SNP Type and inflate_match",
    subtitle = "Average magnitude of annotations",
    x = "inflate_match",
    y = "Mean |mu_0|",
    color = "SNP Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  geom_vline(xintercept = 1, linetype = "dashed", alpha = 0.5)

save_or_show(plt, "03_abs_mean_comparison.png", width = 12, height = 8)
```

## 3.4 Distribution Overlap at Different inflate_match Levels

```{r plot-distributions, fig.width=14, fig.height=10}
# Pick a single annotation_r2 and show distributions at key inflate_match values
annot_r2_focus <- 0.4
inflate_focus <- c(0, 0.5, 1.0, 1.5)

# Run one rep for each setting to get raw data
dist_data <- list()
for (inflate in inflate_focus) {
  effects <- simulate_effect_sizes(p = n_snps, p_star = n_causal, effect_sd = effect_sd, seed = 999)
  priors <- simulate_priors(
    beta = effects$beta,
    annotation_r2 = annot_r2_focus,
    inflate_match = inflate,
    gamma_shrink = NA,
    effect_sd = effect_sd
  )
  
  causal_idx <- effects$causal_idx
  noncausal_idx <- setdiff(seq_len(n_snps), causal_idx)
  
  dist_data[[length(dist_data) + 1]] <- data.frame(
    inflate_match = inflate,
    snp_type = "Causal",
    mu_0 = priors$mu_0[causal_idx]
  )
  dist_data[[length(dist_data) + 1]] <- data.frame(
    inflate_match = inflate,
    snp_type = "Non-causal",
    mu_0 = priors$mu_0[noncausal_idx]
  )
}

dist_df <- dplyr::bind_rows(dist_data)

plt <- ggplot(dist_df, aes(x = mu_0, fill = snp_type)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ inflate_match, labeller = label_both, scales = "free_y") +
  labs(
    title = sprintf("Distribution of mu_0 at annotation_r2 = %.1f", annot_r2_focus),
    subtitle = "As inflate_match increases, non-causal distribution widens to match causal",
    x = "mu_0 (annotation value)",
    y = "Density",
    fill = "SNP Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

save_or_show(plt, "04_distribution_overlap.png", width = 14, height = 10)
```

---

# Section 4: Theoretical Derivation Check

## 4.1 Expected vs Observed Variance

Let's verify the formulas match:

```{r theoretical-check}
# For a specific setting, compute theoretical vs observed variance
test_annot_r2 <- 0.4
test_inflate <- 1.0

# Theoretical values
causal_var_true <- effect_sd^2  # Var(beta) for causal
noise_var_theory <- causal_var_true * (1 - test_annot_r2) / test_annot_r2

# Causal mu_0 variance = Var(beta) + noise_var
causal_mu0_var_theory <- causal_var_true + noise_var_theory

# Non-causal mu_0 variance = noise_var + inflate_match * causal_var
noncausal_mu0_var_theory <- noise_var_theory + test_inflate * causal_var_true

cat("=== Theoretical Calculation ===\n")
cat(sprintf("annotation_r2 = %.2f, inflate_match = %.2f\n\n", test_annot_r2, test_inflate))
cat(sprintf("causal_var (Var of beta[causal]) = %.4f\n", causal_var_true))
cat(sprintf("noise_var = causal_var * (1 - r2) / r2 = %.4f * %.2f / %.2f = %.4f\n", 
            causal_var_true, 1 - test_annot_r2, test_annot_r2, noise_var_theory))
cat(sprintf("\nCausal mu_0 variance (theory):     %.4f + %.4f = %.4f\n", 
            causal_var_true, noise_var_theory, causal_mu0_var_theory))
cat(sprintf("Non-causal mu_0 variance (theory): %.4f + %.2f * %.4f = %.4f\n", 
            noise_var_theory, test_inflate, causal_var_true, noncausal_mu0_var_theory))
cat(sprintf("\nVariance ratio (theory): %.4f / %.4f = %.4f\n",
            noncausal_mu0_var_theory, causal_mu0_var_theory, 
            noncausal_mu0_var_theory / causal_mu0_var_theory))

# Compare to observed
observed <- summary_stats %>%
  filter(annotation_r2 == test_annot_r2, inflate_match == test_inflate)

cat("\n=== Observed (from simulation) ===\n")
cat(sprintf("Causal mu_0 variance (observed):     %.4f\n", observed$causal_var_mean))
cat(sprintf("Non-causal mu_0 variance (observed): %.4f\n", observed$noncausal_var_mean))
cat(sprintf("Variance ratio (observed):           %.4f\n", observed$var_ratio_mean))
```

---

# Section 5: What to Check in Model Outputs

## 5.1 SNPs Parquet Files

---

# Section 5: Validate Against Historical Run Data

This section loads actual simulation outputs from a completed job and verifies
that `inflate_match` behaves as expected in practice.

## 5.1 Configure Job to Analyze

```{r historical-job-config}
# ==== USER INPUT: Specify the job to analyze ====
historical_job_name <- "prior_default_testing"
historical_job_id <- "46997729"
output_root <- here::here("output")

# Set to TRUE to run this analysis (requires the job outputs to exist)
run_historical_analysis <- TRUE
```

## 5.2 Load Run Table and SNPs Dataset

```{r historical-load-data, eval=run_historical_analysis}
# Resolve paths
paths <- test_susine::resolve_job_paths(

  job_name = historical_job_name,
  parent_job_id = historical_job_id,
  output_root = output_root
)

cat("Analyzing job:", historical_job_name, "/", historical_job_id, "\n")
cat("Run history dir:", paths$run_history_dir, "\n")
cat("SNPs dataset dir:", paths$snps_dataset_dir, "\n")

# Check paths exist
if (!dir.exists(paths$run_history_dir)) {
  stop("Run history directory not found: ", paths$run_history_dir)
}
if (!dir.exists(paths$snps_dataset_dir)) {
  stop("SNPs dataset not found: ", paths$snps_dataset_dir)
}

# Load run table
run_table_hist <- readr::read_csv(
  file.path(paths$run_history_dir, "run_table.csv"),
  show_col_types = FALSE
)

cat("\nRun table loaded:", nrow(run_table_hist), "runs\n")
cat("Columns:", paste(names(run_table_hist), collapse = ", "), "\n")

# Check for required columns
required_cols <- c("run_id", "inflate_match", "annotation_r2", "use_case_id")
missing_cols <- setdiff(required_cols, names(run_table_hist))
if (length(missing_cols) > 0) {
  warning("Missing columns in run_table: ", paste(missing_cols, collapse = ", "))
}

# Show inflate_match and annotation_r2 values
cat("\ninflate_match values:", paste(sort(unique(run_table_hist$inflate_match)), collapse = ", "), "\n")
cat("annotation_r2 values:", paste(sort(unique(run_table_hist$annotation_r2)), collapse = ", "), "\n")
cat("use_case_id values:", paste(unique(run_table_hist$use_case_id), collapse = ", "), "\n")
```

## 5.3 Aggregate SNPs Across Runs (Memory-Efficient)

The SNPs dataset is partitioned by use_case_id. We aggregate **all SNPs** across runs
by (causal, inflate_match, annotation_r2, p_star) to get proper variance estimates.

**Memory-efficient approach**: Compute sufficient statistics at the Arrow level,
then derive variance and R² from those statistics. This avoids loading the full
dataset into memory.

```{r historical-compute-stats, eval=run_historical_analysis}
# Open the SNPs dataset with Hive partitioning
snps_ds <- arrow::open_dataset(paths$snps_dataset_dir, partitioning = "use_case_id")

cat("SNPs dataset schema:\n")
print(snps_ds$schema)

# Prepare run metadata for Arrow join
# Ensure run_id is integer to match parquet schema
run_metadata <- run_table_hist %>%
  dplyr::select(run_id, use_case_id, inflate_match, annotation_r2, p_star) %>%
  dplyr::mutate(run_id = as.integer(run_id))

cat("\nRun metadata:", nrow(run_metadata), "runs\n")

# Convert to Arrow table for efficient join
run_metadata_arrow <- arrow::as_arrow_table(run_metadata)

# Compute sufficient statistics at Arrow level:
# - For variance: need n, sum(x), sum(x^2)
# - For correlation/R²: need n, sum(x), sum(y), sum(x^2), sum(y^2), sum(xy)
# 
# Var(X) = E[X²] - E[X]² = (sum_x2/n) - (sum_x/n)²
# Cor(X,Y) = [E[XY] - E[X]E[Y]] / [SD(X) * SD(Y)]

cat("\nComputing sufficient statistics at Arrow level...\n")
cat("(This aggregates before collecting - memory efficient)\n")

# Join SNPs with run metadata at Arrow level, then aggregate
# Group by (causal, inflate_match, annotation_r2, p_star)
snps_agg_raw <- snps_ds %>%
  dplyr::select(run_id, mu_0, beta, causal) %>%
  # Join with run metadata
  dplyr::left_join(run_metadata_arrow, by = "run_id") %>%
  # Group and compute sufficient statistics
  dplyr::group_by(causal, inflate_match, annotation_r2, p_star) %>%
  dplyr::summarise(
    n = n(),
    n_runs = dplyr::n_distinct(run_id),
    # For mu_0
    sum_mu0 = sum(mu_0, na.rm = TRUE),
    sum_mu0_sq = sum(mu_0^2, na.rm = TRUE),
    # For beta
    sum_beta = sum(beta, na.rm = TRUE),
    sum_beta_sq = sum(beta^2, na.rm = TRUE),
    # For correlation (mu_0, beta)
    sum_mu0_beta = sum(mu_0 * beta, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::collect()

cat("Aggregated data collected:", nrow(snps_agg_raw), "rows\n")
cat("(Much smaller than raw SNPs!)\n")

# Also compute overall stats (all SNPs together, ignoring causal status)
# for actual_overall_r2
snps_overall_raw <- snps_ds %>%
  dplyr::select(run_id, mu_0, beta, causal) %>%
  dplyr::left_join(run_metadata_arrow, by = "run_id") %>%
  dplyr::group_by(inflate_match, annotation_r2, p_star) %>%
  dplyr::summarise(
    n_overall = n(),
    sum_mu0_overall = sum(mu_0, na.rm = TRUE),
    sum_mu0_sq_overall = sum(mu_0^2, na.rm = TRUE),
    sum_beta_overall = sum(beta, na.rm = TRUE),
    sum_beta_sq_overall = sum(beta^2, na.rm = TRUE),
    sum_mu0_beta_overall = sum(mu_0 * beta, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::collect()

cat("Overall aggregated data:", nrow(snps_overall_raw), "rows\n")
```

## 5.4 Create Aggregated Summary Table

Derive variance, R², and other statistics from sufficient statistics.

```{r historical-aggregate-table, eval=run_historical_analysis}
# Helper function to compute variance from sufficient statistics
# Var(X) = E[X²] - E[X]² = (sum_x2/n) - (sum_x/n)²
compute_var <- function(n, sum_x, sum_x2) {
  mean_x <- sum_x / n
  (sum_x2 / n) - mean_x^2
}

# Helper function to compute correlation from sufficient statistics
# Cor(X,Y) = Cov(X,Y) / (SD(X) * SD(Y))
# Cov(X,Y) = E[XY] - E[X]E[Y]
compute_cor <- function(n, sum_x, sum_y, sum_x2, sum_y2, sum_xy) {
  mean_x <- sum_x / n
  mean_y <- sum_y / n
  var_x <- (sum_x2 / n) - mean_x^2
  var_y <- (sum_y2 / n) - mean_y^2
  cov_xy <- (sum_xy / n) - mean_x * mean_y
  
  # Handle edge cases
  if (var_x <= 0 || var_y <= 0) return(NA_real_)
  cov_xy / sqrt(var_x * var_y)
}

# Compute derived statistics for the main table
hist_agg_table <- snps_agg_raw %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    n_snps = n,
    mu_0_mean = sum_mu0 / n,
    mu_0_var = compute_var(n, sum_mu0, sum_mu0_sq),
    mu_0_sd = sqrt(pmax(0, mu_0_var)),
    beta_mean = sum_beta / n,
    beta_var = compute_var(n, sum_beta, sum_beta_sq),
    beta_sd = sqrt(pmax(0, beta_var)),
    # Correlation for this group
    cor_mu0_beta = compute_cor(n, sum_mu0, sum_beta, sum_mu0_sq, sum_beta_sq, sum_mu0_beta),
    actual_causal_r2 = cor_mu0_beta^2
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(-n, -sum_mu0, -sum_mu0_sq, -sum_beta, -sum_beta_sq, 
                -sum_mu0_beta, -cor_mu0_beta)

# Compute overall R² (all SNPs, regardless of causal status)
overall_r2 <- snps_overall_raw %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    cor_overall = compute_cor(n_overall, sum_mu0_overall, sum_beta_overall, 
                               sum_mu0_sq_overall, sum_beta_sq_overall, sum_mu0_beta_overall),
    actual_overall_r2 = cor_overall^2
  ) %>%
  dplyr::ungroup() %>%
  dplyr::select(inflate_match, annotation_r2, p_star, actual_overall_r2)

# Join overall R² and clean up
hist_agg_table <- hist_agg_table %>%
  dplyr::left_join(overall_r2, by = c("inflate_match", "annotation_r2", "p_star")) %>%
  # Set actual_causal_r2 to NA for non-causal rows
  dplyr::mutate(
    actual_causal_r2 = ifelse(causal, actual_causal_r2, NA_real_)
  ) %>%
  dplyr::arrange(inflate_match, annotation_r2, p_star, desc(causal))

cat("=== Aggregated SNP Statistics Table ===\n")
cat("(Computed from sufficient statistics - memory efficient)\n\n")
print(hist_agg_table, n = 50)

# Save table for reference
if (save_plots_to_file) {
  table_path <- file.path(plot_output_dir, "historical_aggregated_stats.csv")
  readr::write_csv(hist_agg_table, table_path)
  cat("\nSaved table to:", table_path, "\n")
}
```

## 5.5 Compute Variance Ratios from Aggregated Data

```{r historical-variance-ratios, eval=run_historical_analysis}
# Pivot to get causal/noncausal side by side for each setting
hist_variance_ratios <- hist_agg_table %>%
  dplyr::select(causal, inflate_match, annotation_r2, p_star, n_snps, n_runs, mu_0_var) %>%
  tidyr::pivot_wider(
    id_cols = c(inflate_match, annotation_r2, p_star),
    names_from = causal,
    values_from = c(n_snps, n_runs, mu_0_var),
    names_glue = "{.value}_{ifelse(causal, 'causal', 'noncausal')}"
  )

# Fix column names
names(hist_variance_ratios) <- gsub("_TRUE", "_causal", names(hist_variance_ratios))
names(hist_variance_ratios) <- gsub("_FALSE", "_noncausal", names(hist_variance_ratios))

# Compute variance ratio
hist_variance_ratios <- hist_variance_ratios %>%
  dplyr::mutate(
    var_ratio = mu_0_var_noncausal / mu_0_var_causal
  ) %>%
  dplyr::arrange(inflate_match, annotation_r2, p_star)

cat("=== Variance Ratios by Setting ===\n")
print(hist_variance_ratios, n = 50)

# Summary: average variance ratio at each inflate_match level
hist_summary <- hist_variance_ratios %>%
  dplyr::group_by(inflate_match, annotation_r2) %>%
  dplyr::summarise(
    n_settings = n(),
    var_ratio_mean = mean(var_ratio, na.rm = TRUE),
    var_ratio_sd = sd(var_ratio, na.rm = TRUE),
    .groups = "drop"
  )

cat("\n=== Summary: Mean Variance Ratio by (inflate_match, annotation_r2) ===\n")
print(hist_summary, n = 30)
```

## 5.6 Visualize Historical Results

```{r historical-plot-variance-ratio, fig.width=12, fig.height=8, eval=run_historical_analysis}
# Plot variance ratio vs inflate_match, colored by annotation_r2, faceted by p_star
plt <- ggplot(hist_variance_ratios, 
              aes(x = inflate_match, y = var_ratio, color = factor(annotation_r2))) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 1) +
  facet_wrap(~ p_star, labeller = label_both) +
  labs(
    title = "Historical Run Data: Var(mu_0 noncausal) / Var(mu_0 causal)",
    subtitle = sprintf("Job: %s / %s | Red line = 1.0 (equal variance)", 
                       historical_job_name, historical_job_id),
    x = "inflate_match",
    y = "Variance Ratio",
    color = "annotation_r2"
  ) +
  theme_minimal() +
  theme(legend.position = "right") +
  scale_y_continuous(breaks = seq(0, 2, by = 0.25))

save_or_show(plt, "05_historical_variance_ratio.png", width = 12, height = 8)
```

```{r historical-plot-r2-causal, fig.width=12, fig.height=8, eval=run_historical_analysis}
# Plot actual_causal_r2 (causal SNPs only)
# Reshape to get causal-only rows
r2_causal_data <- hist_agg_table %>%
  dplyr::filter(causal == TRUE) %>%
  dplyr::select(inflate_match, annotation_r2, p_star, actual_causal_r2)

plt <- ggplot(r2_causal_data, 
              aes(x = inflate_match, y = actual_causal_r2, color = factor(annotation_r2))) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  facet_wrap(~ p_star, labeller = label_both) +
  labs(
    title = "Historical Run Data: R² of mu_0 Predicting Beta (Causal SNPs Only)",
    subtitle = sprintf("Job: %s / %s", historical_job_name, historical_job_id),
    x = "inflate_match",
    y = "R² (causal SNPs)",
    color = "annotation_r2"
  ) +
  theme_minimal() +
  theme(legend.position = "right") +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.2))

save_or_show(plt, "06_historical_r2_causal.png", width = 12, height = 8)
```

```{r historical-plot-variance-comparison, fig.width=14, fig.height=10, eval=run_historical_analysis}
# Plot Var(mu_0) by SNP type (causal vs non-causal) across all settings
# This shows how causal and non-causal variance change with inflate_match
var_comparison <- hist_agg_table %>%
  dplyr::mutate(
    snp_type = ifelse(causal, "Causal", "Non-causal")
  )

plt <- ggplot(var_comparison, 
              aes(x = inflate_match, y = mu_0_var, color = snp_type, linetype = snp_type)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_grid(p_star ~ annotation_r2, scales = "free_y", 
             labeller = labeller(annotation_r2 = label_both, p_star = label_both)) +
  labs(
    title = "Historical Run Data: Var(mu_0) by SNP Type (Aggregated Across Runs)",
    subtitle = sprintf("Job: %s / %s", historical_job_name, historical_job_id),
    x = "inflate_match",
    y = "Var(mu_0)",
    color = "SNP Type",
    linetype = "SNP Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

save_or_show(plt, "06b_historical_variance_comparison.png", width = 14, height = 10)
```

```{r historical-plot-r2-overall, fig.width=12, fig.height=8, eval=run_historical_analysis}
# Plot actual_overall_r2 (all SNPs: causal + non-causal)
# Just need one row per (inflate_match, annotation_r2, p_star) for overall R²
r2_overall_data <- hist_agg_table %>%
  dplyr::filter(causal == TRUE) %>%  # Get one row per setting (causal rows have the overall R²)
  dplyr::select(inflate_match, annotation_r2, p_star, actual_overall_r2)

plt <- ggplot(r2_overall_data, 
              aes(x = inflate_match, y = actual_overall_r2, color = factor(annotation_r2))) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  facet_wrap(~ p_star, labeller = label_both, scales = "fixed") +
  labs(
    title = "Historical Run Data: R² of mu_0 Predicting Beta (All SNPs)",
    subtitle = sprintf("Job: %s / %s | Includes both causal and non-causal SNPs", 
                       historical_job_name, historical_job_id),
    x = "inflate_match",
    y = "R² (all SNPs)",
    color = "annotation_r2"
  ) +
  theme_minimal() +
  theme(legend.position = "right")

save_or_show(plt, "07_historical_r2_overall.png", width = 12, height = 8)
```

## 5.7 Compare Theoretical vs Historical

```{r historical-vs-theoretical, eval=run_historical_analysis}
# Compare historical variance ratios to our simulation predictions
cat("=== Comparison: Simulation vs Historical ===\n\n")

# Get overlapping settings (hist_summary is now aggregated across runs)
common_settings <- hist_summary %>%
  dplyr::inner_join(
    summary_stats %>% dplyr::select(annotation_r2, inflate_match, var_ratio_mean),
    by = c("annotation_r2", "inflate_match"),
    suffix = c("_historical", "_simulation")
  )

if (nrow(common_settings) > 0) {
  cat("Settings with both simulation and historical data:\n")
  print(common_settings %>% 
          dplyr::select(annotation_r2, inflate_match, 
                        var_ratio_mean_historical, var_ratio_mean_simulation) %>%
          dplyr::mutate(
            diff = var_ratio_mean_historical - var_ratio_mean_simulation,
            pct_diff = 100 * diff / var_ratio_mean_simulation
          ))
  
  # Check correlation
  cor_val <- cor(common_settings$var_ratio_mean_historical, 
                 common_settings$var_ratio_mean_simulation, 
                 use = "complete.obs")
  cat(sprintf("\nCorrelation between historical and simulation variance ratios: %.3f\n", cor_val))
  
  # Mean absolute difference
  mad <- mean(abs(common_settings$var_ratio_mean_historical - 
                   common_settings$var_ratio_mean_simulation), na.rm = TRUE)
  cat(sprintf("Mean absolute difference: %.4f\n", mad))
} else {
  cat("No overlapping settings found between simulation and historical data.\n")
  cat("This may be because the historical job used different inflate_match values.\n")
}
```

---

# Section 6: Summary & Conclusions

```{r summary}
cat("=== Key Findings ===\n\n")

# Check if variance ratio crosses 1 at inflate_match = 1
ratio_at_1 <- summary_stats %>%
  filter(inflate_match == 1) %>%
  summarise(mean_ratio = mean(var_ratio_mean), sd_ratio = sd(var_ratio_mean))

cat(sprintf("At inflate_match = 1.0:\n"))
cat(sprintf("  Average variance ratio across all annotation_r2: %.4f (SD: %.4f)\n", 
            ratio_at_1$mean_ratio, ratio_at_1$sd_ratio))

if (abs(ratio_at_1$mean_ratio - 1) < 0.05) {
  cat("  ✓ Variance ratio is close to 1.0 as expected!\n\n")
} else {
  cat("  ✗ Variance ratio deviates from 1.0 - investigate further\n\n")
}

cat("Interpretation:\n")
cat("- inflate_match = 0: Non-causal SNPs have smaller mu_0 variance than causal\n")
cat("- inflate_match = 1: Non-causal and causal SNPs have ~equal mu_0 variance\n")
cat("- inflate_match > 1: Non-causal SNPs have LARGER mu_0 variance than causal\n")
```

