---
title: "SuSiNE SNPs Dataset Preview"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: false
---

# Overview

Quick peek at the Parquet-based `snps_dataset` without loading the full table.

## Setup

```{r setup, message=FALSE}
library(here)
library(ggplot2)
here::i_am("vignettes/snps_dataset_preview.Rmd")
here()

knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 6,
  dpi = 300,
  fig.retina = 2,
  out.width = "100%"
)
```

## Load a small sample

```{r load-snps-sample, message=FALSE}
# Adjust these to point at the desired run
job_name <- "susie_scen_1_recreation"
job_id <- "46880273"

snps_path <- file.path(
  here(),
  "output", "slurm_output",
  job_name, job_id,
  "combined", "aggregated",
  "snps_dataset"
)

stopifnot(dir.exists(snps_path))

# Create output directory for plots
plot_dir <- file.path(dirname(snps_path), "plots")
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
message("Plots will be saved to: ", plot_dir)

library(dplyr)
library(arrow)

# Read only the first few rows; arrow will push the slice down so it does not scan the whole dataset
snps_preview <- arrow::open_dataset(snps_path) %>%
  slice_head(n = 20) %>%
  collect()

snps_preview
```

# Temporarily repartition existing dataset by use_case_id

```{r repartition-snps-dataset, message=FALSE}
# # This writes a partitioned copy alongside the original without loading everything into memory.
# # Remove or disable after use; future runs should emit partitioned data natively.
# partitioned_path <- file.path(dirname(snps_path), "snps_dataset_by_use_case")
# 
# if (!dir.exists(partitioned_path)) {
#   # Map run_id -> use_case_id (small table) then join in-arrow and write partitioned
#   run_table_path <- file.path(here(), "output", "run_history", job_name, job_id, "run_table.csv")
#   run_map <- readr::read_csv(
#     run_table_path,
#     show_col_types = FALSE,
#     col_select = c(run_id, use_case_id),
#     col_types = readr::cols(
#       run_id = readr::col_integer(),
#       use_case_id = readr::col_character()
#     )
#   ) %>%
#     dplyr::mutate(run_id = as.integer(run_id))
# 
#   snps_ds <- arrow::open_dataset(snps_path)
#   run_map_ds <- arrow::InMemoryDataset$create(run_map)
# 
#   snps_with_uc <- snps_ds %>%
#     dplyr::left_join(run_map_ds, by = "run_id") %>%
#     dplyr::select(use_case_id, dplyr::everything())
# 
#   arrow::write_dataset(
#     snps_with_uc,
#     path = partitioned_path,
#     format = "parquet",
#     partitioning = "use_case_id"
#   )
# 
#   rm(run_map, run_map_ds, snps_ds, snps_with_uc)
# }
# 
# partitioned_path
```

# Power vs. FDR (Fig. 2D/E style) from SNP PIPs

```{r power-fdr-setup}
# User inputs
split_by_var_list <- c("use_case_id") # options: use_case_id, p_star, L, y_noise, annotation_r2, inflate_match, gamma_shrink
filter_p_star <- 5                   # set to 5 to focus on p_star = 5, or NA to include all
filter_annotation_r2 <- NA            # set to 0.8 to focus on high-quality annotations (includes NA), or NA to include all
pip_bucket_width <- 0.01               # bucket width for PIP sweeping; smaller = finer curves
```

```{r power-fdr-curves, message=FALSE}
split_cols <- c("run_id", "use_case_id", "p_star", "L", "y_noise", "annotation_r2", "inflate_match", "gamma_shrink")

run_table_path <- file.path(here(), "output", "run_history", job_name, job_id, "run_table.csv")
use_cases_path <- file.path(here(), "output", "run_history", job_name, job_id, "use_cases.csv")

run_map <- readr::read_csv(
  run_table_path,
  show_col_types = FALSE,
  col_select = split_cols,
  col_types = readr::cols(
    run_id = readr::col_integer(),
    use_case_id = readr::col_character(),
    p_star = readr::col_integer(),
    L = readr::col_integer(),
    y_noise = readr::col_double(),
    annotation_r2 = readr::col_double(),
    inflate_match = readr::col_double(),
    gamma_shrink = readr::col_double()
  )
) %>%
  dplyr::mutate(run_id = as.integer(run_id))

use_case_labels <- readr::read_csv(
  use_cases_path,
  show_col_types = FALSE,
  col_select = c(use_case_id, label),
  col_types = readr::cols(
    use_case_id = readr::col_character(),
    label = readr::col_character()
  )
)

run_map <- dplyr::left_join(run_map, use_case_labels, by = "use_case_id")

# Guard split vars
group_vars <- split_by_var_list
stopifnot(length(group_vars) >= 1)
if (!all(group_vars %in% colnames(run_map))) {
  stop("split_by_var_list contains unsupported columns.")
}

# Open SNPs dataset and check what columns are already present
snps_ds <- arrow::open_dataset(snps_path)
snps_cols <- names(snps_ds)

# If use_case_id is already in the SNPs dataset (from partitioning or embedded),
# exclude it from run_map to avoid column collision during join.
# But keep 'label' since it's not in snps and we need it for plotting.
join_cols_to_exclude <- intersect(setdiff(names(run_map), c("run_id", "label")), snps_cols)
if (length(join_cols_to_exclude)) {
  message("SNPs dataset already has columns: ", paste(join_cols_to_exclude, collapse = ", "),
          " — excluding from join to avoid collision.")
  run_map_for_join <- dplyr::select(run_map, -all_of(join_cols_to_exclude))
} else {
  run_map_for_join <- run_map
}

run_map_ds <- arrow::InMemoryDataset$create(run_map_for_join)

snps_joined <- snps_ds %>%
  dplyr::left_join(run_map_ds, by = "run_id")

if (!is.na(filter_p_star)) {
  snps_joined <- snps_joined %>% dplyr::filter(p_star == filter_p_star)
}
if (!is.na(filter_annotation_r2)) {
  snps_joined <- snps_joined %>% dplyr::filter(annotation_r2 == filter_annotation_r2 | is.na(annotation_r2))
}

# Include label in grouping when we split by use_case_id so legends show names
label_group <- if ("use_case_id" %in% group_vars) "label" else character(0)

# Bucket PIPs once, then accumulate to build Power vs. FDR curves
pip_counts <- snps_joined %>%
  dplyr::mutate(
    pip_bucket = pmax(0, pmin(1, floor(pip / pip_bucket_width) * pip_bucket_width))
  ) %>%
  dplyr::group_by(dplyr::across(all_of(c(group_vars, label_group))), pip_bucket) %>%
  dplyr::summarise(
    n = dplyr::n(),
    pos = sum(causal == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::collect()

pip_curve_data <- pip_counts %>%
  dplyr::group_by(dplyr::across(all_of(c(group_vars, label_group)))) %>%
  dplyr::arrange(dplyr::desc(pip_bucket), .by_group = TRUE) %>%
  dplyr::mutate(
    total_pos = sum(pos, na.rm = TRUE),
    cum_tp = cumsum(pos),
    cum_pred = cumsum(n),
    precision = if_else(cum_pred > 0, cum_tp / cum_pred, NA_real_),
    recall = if_else(total_pos > 0, cum_tp / total_pos, NA_real_),
    fdr = 1 - precision
  ) %>%
  dplyr::filter(total_pos > 0, !is.na(precision), !is.na(recall)) %>%
  dplyr::ungroup()

# Build human-friendly labels: prefer use_case label when available; append other group fields
extra_vars <- setdiff(group_vars, "use_case_id")
pip_curve_data <- pip_curve_data %>%
  dplyr::mutate(
    base_label = dplyr::coalesce(label, use_case_id),
    extra_label = if (length(extra_vars)) {
      interaction(dplyr::across(all_of(extra_vars)), drop = TRUE, lex.order = TRUE)
    } else {
      NA_character_
    },
    curve_label = ifelse(is.na(extra_label), base_label, paste0(base_label, " [", extra_label, "]"))
  )

# Persist the curve data for reuse
pip_curve_data_path <- file.path(dirname(snps_path), "pip_power_fdr.csv")
readr::write_csv(pip_curve_data, pip_curve_data_path)

pip_curve_data_path
```

```{r power-fdr-plot, fig.width=15, fig.height=7, message=FALSE}
p_power_fdr <- ggplot(pip_curve_data, aes(x = fdr, y = recall, color = curve_label)) +
  geom_line() +
  geom_point(size = 0.8) +
  labs(
    title = "Power vs. FDR (PIP sweep)",
    subtitle = paste0(
      "Grouped by: ", paste(split_by_var_list, collapse = ", "),
      if (!is.na(filter_p_star)) paste0("; p_star = ", filter_p_star) else "",
      if (!is.na(filter_annotation_r2)) paste0("; annotation_r2 = ", filter_annotation_r2, " (including NA)") else ""
    ),
    x = "False Discovery Rate (1 - Precision)",
    y = "Power (Recall)",
    color = "Group"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave(file.path(plot_dir, "power_vs_fdr.png"), p_power_fdr, width = 15, height = 7, dpi = 300)
message("Saved: power_vs_fdr.png")
# p_power_fdr  # Uncomment to display in viewer
```

# Notes

- The preview uses `slice_head()` on the Arrow dataset, so it avoids loading the full Parquet collection.
- Bump `n` in `slice_head()` if you want to see more rows; keep it small to stay memory-light.
