---
title: "Real-Data Annotation Low-PIP Diagnostic"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    dev: ragg_png
editor_options:
  markdown:
    wrap: 72
---

This one-off workbook computes the reviewer-requested real-data diagnostic:
the correlation between marginal z-scores and annotations among variants with
low SuSiE-anchor posterior inclusion probability.

The workbook is intended to run after the real-data ensemble collect step, using
the existing aggregated outputs only. It does not refit any model.

```{r setup, message=FALSE, fig.keep='none'}
here::i_am("vignettes/real data pipeline/real_data_annotation_low_pip_diagnostic.Rmd")
library(here)
library(dplyr)
library(readr)
library(purrr)
library(ggplot2)
library(arrow)
library(knitr)

theme_set(theme_bw(base_size = 12))

# Headless HPC has no X11/Cairo, so base png()/svg() fail. During preview/paper
# renders (env var set), route knitted figures and every .png save through ragg.
.headless_preview <- nzchar(Sys.getenv("RD_PREVIEW_FIG_DIR")) ||
  nzchar(Sys.getenv("RD_PREVIEW_PAPER_DIR"))
if (.headless_preview && !requireNamespace("ragg", quietly = TRUE)) {
  stop("Headless preview render requires the ragg package for PNG output.")
}
if (.headless_preview) {
  knitr::opts_chunk$set(dev = "ragg_png")
  ggsave <- function(filename, plot = ggplot2::last_plot(), ..., device = NULL) {
    if (is.null(device) && grepl("\\.png$", filename, ignore.case = TRUE)) {
      device <- ragg::agg_png
    }
    ggplot2::ggsave(filename, plot = plot, ..., device = device)
  }
}
```

# 1. User Configuration

```{r config}
job_name <- "real_data_ensemble_geometric_n20"
parent_job_id <- "52906940"
low_pip_threshold <- 0.01

if (identical(parent_job_id, "REPLACE_WITH_SLURM_JOB_ID")) {
  stop("Set parent_job_id to the SLURM job id returned by sbatch before running.")
}

aggregated_dir <- {
  ov <- Sys.getenv("RD_PREVIEW_AGG_DIR", "")
  if (nzchar(ov)) ov else here("output", "slurm_output", job_name, parent_job_id, "aggregated")
}
figure_dir <- {
  ov <- Sys.getenv("RD_PREVIEW_FIG_DIR", "")
  if (nzchar(ov)) ov else here("output", "slurm_output", job_name, parent_job_id, "figures", "real_data_ensemble")
}
overall_figure_dir <- file.path(figure_dir, "overall")
paper_real_data_case_study_dir <- {
  ov <- Sys.getenv("RD_PREVIEW_PAPER_DIR", "")
  if (nzchar(ov)) ov else here("..", "..", "Writings", "plots", "real_data_case_study")
}

dir.create(overall_figure_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Load Aggregated Outputs

```{r load}
paper_summary_path <- file.path(aggregated_dir, "paper_real_data_ensemble_summary.csv")
variant_posteriors_path <- file.path(aggregated_dir, "variant_posteriors_dataset")

if (!file.exists(paper_summary_path)) {
  stop("Missing paper summary: ", paper_summary_path)
}
if (!dir.exists(variant_posteriors_path)) {
  stop("Missing variant posterior dataset: ", variant_posteriors_path)
}

paper_summary <- readr::read_csv(paper_summary_path, show_col_types = FALSE)
required_summary_cols <- c("locus_id", "susie_anchor_run_id")
missing_summary_cols <- setdiff(required_summary_cols, names(paper_summary))
if (length(missing_summary_cols) > 0) {
  stop("paper_real_data_ensemble_summary.csv is missing columns: ",
       paste(missing_summary_cols, collapse = ", "))
}
if (!"gene_name" %in% names(paper_summary)) {
  paper_summary$gene_name <- NA_character_
}
if (nrow(paper_summary) == 0) {
  stop("paper_real_data_ensemble_summary.csv has zero rows.")
}

variant_posteriors <- arrow::open_dataset(variant_posteriors_path)
required_variant_cols <- c(
  "locus_id", "run_id", "variant_id", "pip", "z_score", "annotation_a"
)
missing_variant_cols <- setdiff(required_variant_cols, names(variant_posteriors))
if (length(missing_variant_cols) > 0) {
  stop("variant_posteriors_dataset is missing columns: ",
       paste(missing_variant_cols, collapse = ", "))
}

tibble::tibble(
  aggregated_dir = aggregated_dir,
  n_loci = nrow(paper_summary),
  low_pip_threshold = low_pip_threshold
)
```

# 3. Compute Diagnostic

```{r helpers}
safe_cor <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  if (sum(ok) < 2L) return(NA_real_)
  suppressWarnings(stats::cor(x[ok], y[ok]))
}

safe_cor_p_value <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  if (sum(ok) < 3L) return(NA_real_)
  suppressWarnings(
    tryCatch(
      stats::cor.test(x[ok], y[ok])$p.value,
      error = function(e) NA_real_
    )
  )
}

pull_anchor_variants <- function(locus_id, anchor_run_id) {
  variant_posteriors %>%
    dplyr::filter(
      .data$locus_id == !!locus_id,
      .data$run_id == !!anchor_run_id
    ) %>%
    dplyr::select(
      .data$locus_id,
      .data$run_id,
      .data$variant_id,
      .data$pip,
      .data$z_score,
      .data$annotation_a
    ) %>%
    dplyr::collect()
}
```

```{r compute}
annotation_low_pip_diagnostic <- purrr::map_dfr(seq_len(nrow(paper_summary)), function(i) {
  loc <- paper_summary$locus_id[[i]]
  gene <- paper_summary$gene_name[[i]]
  anchor_run_id <- paper_summary$susie_anchor_run_id[[i]]

  if (is.na(anchor_run_id)) {
    return(tibble::tibble(
      locus_id = loc,
      gene_name = gene,
      susie_anchor_run_id = NA_integer_,
      n_variants = 0L,
      n_anchor_low_pip = 0L,
      n_finite_low_pip = 0L,
      low_pip_fraction = NA_real_,
      cor_a_z_anchor_low_pip = NA_real_,
      cor_a_z_anchor_low_pip_p_value = NA_real_,
      abs_cor_a_z_anchor_low_pip = NA_real_,
      mean_abs_z_anchor_low_pip = NA_real_,
      mean_abs_a_anchor_low_pip = NA_real_
    ))
  }

  anchor_variants <- pull_anchor_variants(loc, anchor_run_id)
  low_pip <- anchor_variants %>%
    dplyr::filter(.data$pip < low_pip_threshold)
  finite_low_pip <- is.finite(low_pip$z_score) & is.finite(low_pip$annotation_a)

  cor_val <- safe_cor(low_pip$annotation_a, low_pip$z_score)
  tibble::tibble(
    locus_id = loc,
    gene_name = gene,
    susie_anchor_run_id = anchor_run_id,
    n_variants = nrow(anchor_variants),
    n_anchor_low_pip = nrow(low_pip),
    n_finite_low_pip = sum(finite_low_pip),
    low_pip_fraction = nrow(low_pip) / nrow(anchor_variants),
    cor_a_z_anchor_low_pip = cor_val,
    cor_a_z_anchor_low_pip_p_value = safe_cor_p_value(low_pip$annotation_a, low_pip$z_score),
    abs_cor_a_z_anchor_low_pip = abs(cor_val),
    mean_abs_z_anchor_low_pip = mean(abs(low_pip$z_score[finite_low_pip]), na.rm = TRUE),
    mean_abs_a_anchor_low_pip = mean(abs(low_pip$annotation_a[finite_low_pip]), na.rm = TRUE)
  )
})

if (nrow(annotation_low_pip_diagnostic) == 0) {
  stop("Annotation low-PIP diagnostic produced zero rows.")
}
if (any(annotation_low_pip_diagnostic$n_finite_low_pip < 2L)) {
  warning("At least one locus has fewer than two finite low-PIP variants; correlation is NA.")
}

annotation_low_pip_summary <- annotation_low_pip_diagnostic %>%
  dplyr::summarise(
    n_loci = dplyr::n(),
    n_loci_with_finite_cor = sum(is.finite(.data$cor_a_z_anchor_low_pip)),
    median_cor = median(.data$cor_a_z_anchor_low_pip, na.rm = TRUE),
    max_abs_cor = max(.data$abs_cor_a_z_anchor_low_pip, na.rm = TRUE),
    median_abs_cor = median(.data$abs_cor_a_z_anchor_low_pip, na.rm = TRUE),
    median_n_anchor_low_pip = median(.data$n_anchor_low_pip, na.rm = TRUE),
    min_n_anchor_low_pip = min(.data$n_anchor_low_pip, na.rm = TRUE)
  )

annotation_low_pip_summary
```

# 4. Save Outputs

```{r save}
diagnostic_csv <- file.path(
  overall_figure_dir,
  "paper_real_data_annotation_low_pip_alignment.csv"
)
summary_csv <- file.path(
  overall_figure_dir,
  "paper_real_data_annotation_low_pip_alignment_summary.csv"
)

readr::write_csv(annotation_low_pip_diagnostic, diagnostic_csv)
readr::write_csv(annotation_low_pip_summary, summary_csv)

if (dir.exists(paper_real_data_case_study_dir)) {
  file.copy(
    diagnostic_csv,
    file.path(paper_real_data_case_study_dir, basename(diagnostic_csv)),
    overwrite = TRUE
  )
  file.copy(
    summary_csv,
    file.path(paper_real_data_case_study_dir, basename(summary_csv)),
    overwrite = TRUE
  )
}

tibble::tibble(
  diagnostic_csv = diagnostic_csv,
  summary_csv = summary_csv,
  copied_to_paper_dir = dir.exists(paper_real_data_case_study_dir)
)
```

# 5. Visualize

```{r plot, fig.width=8, fig.height=5}
plot_tbl <- annotation_low_pip_diagnostic %>%
  dplyr::mutate(
    locus_label = dplyr::if_else(
      is.na(.data$gene_name) | .data$gene_name == "",
      .data$locus_id,
      paste0(.data$gene_name, " (", .data$locus_id, ")")
    ),
    locus_label = stats::reorder(.data$locus_label, .data$cor_a_z_anchor_low_pip)
  )

p_alignment <- ggplot(
  plot_tbl,
  aes(x = .data$locus_label, y = .data$cor_a_z_anchor_low_pip)
) +
  geom_hline(yintercept = 0, linewidth = 0.3, color = "grey55") +
  geom_point(aes(size = .data$n_anchor_low_pip), alpha = 0.85) +
  coord_flip() +
  scale_size_continuous(name = "Variants with\nanchor PIP < 0.01", range = c(1.8, 5.5)) +
  labs(
    x = NULL,
    y = "cor(annotation, marginal z) among low-anchor-PIP variants",
    title = "Real-data annotation alignment on SuSiE-anchor low-PIP variants",
    subtitle = paste0("Low-PIP threshold: PIP < ", low_pip_threshold)
  ) +
  theme(
    legend.position = "right",
    plot.title.position = "plot"
  )

p_alignment

alignment_png <- file.path(
  overall_figure_dir,
  "paper_real_data_annotation_low_pip_alignment.png"
)
ggsave(alignment_png, p_alignment, width = 8, height = 5, dpi = 300)

if (dir.exists(paper_real_data_case_study_dir)) {
  ggsave(
    file.path(paper_real_data_case_study_dir, basename(alignment_png)),
    p_alignment,
    width = 8,
    height = 5,
    dpi = 300
  )
}
```

# 6. Table

```{r table}
annotation_low_pip_diagnostic %>%
  dplyr::arrange(dplyr::desc(.data$abs_cor_a_z_anchor_low_pip)) %>%
  knitr::kable(digits = 4)
```
