---
title: "Real-Data Ensemble — Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

This workbook reads consolidated real-data ensemble outputs and renders the
main locus-level visual diagnostics.

```{r setup, message=FALSE}
here::i_am("vignettes/real data pipeline/visualize_results_workbook_real_data_ensemble.Rmd")
library(here)
library(devtools)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(arrow)
library(knitr)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")
theme_set(theme_bw(base_size = 12))
```

# 1. User Configuration

```{r config}
job_name <- "real_data_ensemble_geometric_n20"
parent_job_id <- "52906940"
force_recompute_basin_r2_drift <- FALSE

if (identical(parent_job_id, "REPLACE_WITH_SLURM_JOB_ID")) {
  stop("Set parent_job_id to the SLURM job id returned by sbatch before visualizing.")
}

aggregated_dir <- here("output", "slurm_output", job_name, parent_job_id, "aggregated")
figure_dir <- here("output", "slurm_output", job_name, parent_job_id, "figures", "real_data_ensemble")
overall_figure_dir <- file.path(figure_dir, "overall")
locus_figure_dir <- file.path(figure_dir, "per_locus")
dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(overall_figure_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(locus_figure_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Load Outputs

```{r load}
functional_grid_summary <- readr::read_csv(
  file.path(aggregated_dir, "functional_grid_summary.csv"),
  show_col_types = FALSE
)
multimodal_metrics <- readr::read_csv(
  file.path(aggregated_dir, "multimodal_metrics.csv"),
  show_col_types = FALSE
)
run_comparisons <- readr::read_csv(
  file.path(aggregated_dir, "run_comparisons.csv"),
  show_col_types = FALSE
)
anchor_summary <- readr::read_csv(
  file.path(aggregated_dir, "susie_anchor_summary.csv"),
  show_col_types = FALSE
)
highest_weight_refit_summary <- if (file.exists(file.path(aggregated_dir, "highest_weight_refit_summary.csv"))) {
  readr::read_csv(file.path(aggregated_dir, "highest_weight_refit_summary.csv"), show_col_types = FALSE)
} else {
  tibble::tibble(locus_id = character(), refit_pve_postmean_std = numeric())
}
highest_weight_refit_basin <- if (file.exists(file.path(aggregated_dir, "highest_weight_refit_basin_r2_drift.csv"))) {
  readr::read_csv(file.path(aggregated_dir, "highest_weight_refit_basin_r2_drift.csv"), show_col_types = FALSE)
} else {
  tibble::tibble(
    locus_id = character(),
    basin_r2_drift_highest_weight_vs_susie_anchor = numeric(),
    basin_r2_drift_refit_vs_susie_anchor = numeric()
  )
}
paper_summary <- if (file.exists(file.path(aggregated_dir, "paper_real_data_ensemble_summary.csv"))) {
  readr::read_csv(file.path(aggregated_dir, "paper_real_data_ensemble_summary.csv"), show_col_types = FALSE)
} else {
  tibble::tibble(
    locus_id = character(),
    susie_anchor_run_id = integer(),
    highest_weight_source_run_id = integer(),
    highest_weight_source_agg_weight_run = numeric(),
    jsd_susie_anchor_vs_susine_ensemble = numeric(),
    jsd_susie_anchor_vs_refit = numeric(),
    jsd_susine_ensemble_vs_refit = numeric()
  )
}
for (nm in c("susie_anchor_run_id", "highest_weight_source_run_id")) {
  if (!nm %in% names(paper_summary)) paper_summary[[nm]] <- NA_integer_
}
if (!"highest_weight_source_agg_weight_run" %in% names(paper_summary)) {
  paper_summary$highest_weight_source_agg_weight_run <- NA_real_
}
cluster_summary <- readr::read_csv(
  file.path(aggregated_dir, "cluster_summary.csv"),
  show_col_types = FALSE
)
top_variants <- readr::read_csv(
  file.path(aggregated_dir, "top_variants_cluster_weight.csv"),
  show_col_types = FALSE
)
aggregated_variants <- arrow::read_parquet(
  file.path(aggregated_dir, "aggregated_variant_pips_cluster_weight.parquet")
) %>%
  tibble::as_tibble()
highest_weight_refit_variants <- if (file.exists(file.path(aggregated_dir, "highest_weight_refit_variant_posteriors.parquet"))) {
  arrow::read_parquet(file.path(aggregated_dir, "highest_weight_refit_variant_posteriors.parquet")) %>%
    tibble::as_tibble()
} else {
  tibble::tibble(locus_id = character(), variant_id = character(), pip = numeric())
}
variant_posteriors <- arrow::open_dataset(
  file.path(aggregated_dir, "variant_posteriors_dataset")
) %>%
  dplyr::collect()

loci <- sort(unique(functional_grid_summary$locus_id))
loci
```

# 3. Plot Helpers

```{r helpers}
format_grid_value <- function(x) {
  out <- sprintf("%.2f", as.numeric(x))
  out <- sub("0+$", "", out)
  sub("\\.$", "", out)
}

add_grid_labels <- function(df) {
  c_vals <- sort(unique(df$c_value))
  sigma_vals <- sort(unique(df$sigma_0_2_scalar))
  df %>%
    dplyr::mutate(
      c_label = factor(format_grid_value(.data$c_value),
                       levels = format_grid_value(c_vals)),
      sigma_label = factor(format_grid_value(.data$sigma_0_2_scalar),
                           levels = format_grid_value(sigma_vals))
    )
}

nearest_reference_run <- function(locus_id, target_c = 0, target_sigma = 0.2) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  is.finite(.data$c_value),
                  is.finite(.data$sigma_0_2_scalar)) %>%
    dplyr::mutate(
      c_distance = abs(.data$c_value - target_c),
      sigma_distance = abs(log(.data$sigma_0_2_scalar / target_sigma))
    ) %>%
    dplyr::arrange(.data$c_distance, .data$sigma_distance, .data$run_id)
  if (!nrow(df)) return(NA_integer_)
  as.integer(df$run_id[[1]])
}

reference_pip_tbl <- function(locus_id) {
  ref_run_id <- nearest_reference_run(locus_id)
  if (is.na(ref_run_id)) {
    return(tibble::tibble(variant_id = character(), baseline_pip = numeric()))
  }
  variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  .data$run_id == !!ref_run_id) %>%
    dplyr::select(.data$variant_id, baseline_pip = .data$pip)
}

refit_pip_tbl <- function(locus_id) {
  if (!nrow(highest_weight_refit_variants)) {
    return(tibble::tibble(variant_id = character(), refit_pip = numeric()))
  }
  highest_weight_refit_variants %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    dplyr::select(.data$variant_id, refit_pip = .data$pip)
}

run_pip_vec <- function(locus_id, run_id) {
  if (length(run_id) != 1L || is.na(run_id)) return(NULL)
  x <- variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  .data$run_id == !!as.integer(run_id)) %>%
    dplyr::arrange(.data$ld_matrix_index) %>%
    dplyr::pull(.data$pip)
  if (!length(x)) return(NULL)
  x
}

safe_pip_jsd <- function(x, y) {
  if (is.null(x) || is.null(y) || length(x) != length(y) || !length(x)) {
    return(NA_real_)
  }
  as.numeric(js_distance(x, y))
}

run_pip_jsd <- function(locus_id, run_id_a, run_id_b) {
  safe_pip_jsd(run_pip_vec(locus_id, run_id_a), run_pip_vec(locus_id, run_id_b))
}

core95_entropy <- function(prob, rho = 0.95) {
  p <- as.numeric(prob)
  p[!is.finite(p) | p < 0] <- 0
  s <- sum(p)
  if (!is.finite(s) || s <= 0) return(NA_real_)
  p <- p / s
  ord <- order(p, decreasing = TRUE)
  cum_p <- cumsum(p[ord])
  keep_n <- which(cum_p >= rho)[[1]]
  if (!length(keep_n) || is.na(keep_n)) keep_n <- length(ord)
  core <- p[ord[seq_len(keep_n)]]
  core <- core / sum(core)
  -sum(core[core > 0] * log(core[core > 0]))
}

geom_locus_text <- function(mapping = NULL, data = NULL, ..., size = 2.4,
                            seed = NA, box.padding = 0.35,
                            point.padding = 0.35, force = 1.4,
                            force_pull = 1, max.iter = 10000,
                            max.time = 1, segment.size = 0.3,
                            show.legend = FALSE) {
  if (requireNamespace("ggrepel", quietly = TRUE)) {
    return(ggrepel::geom_text_repel(
      mapping = mapping,
      data = data,
      size = size,
      min.segment.length = 0,
      box.padding = box.padding,
      point.padding = point.padding,
      force = force,
      force_pull = force_pull,
      max.iter = max.iter,
      max.time = max.time,
      segment.size = segment.size,
      max.overlaps = Inf,
      seed = seed,
      show.legend = show.legend,
      ...
    ))
  }
  ggplot2::geom_text(
    mapping = mapping,
    data = data,
    size = size,
    check_overlap = TRUE,
    show.legend = show.legend,
    ...
  )
}

plot_cluster_grid <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)
  sigma_vals <- sort(unique(df$sigma_0_2_scalar))

  ggplot(df, aes(x = c_value, y = sigma_0_2_scalar, color = factor(cluster_id))) +
    geom_point(size = 3) +
    geom_vline(aes(xintercept = baseline_c_l), linetype = "dashed", color = "grey40") +
    scale_y_log10(breaks = sigma_vals, labels = format_grid_value(sigma_vals)) +
    labs(
      title = paste("Clustered Functional Grid:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      color = "Cluster"
    )
}

plot_weight_heatmap <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    add_grid_labels()

  ggplot(df, aes(x = c_label, y = sigma_label, fill = agg_weight_run)) +
    geom_tile(color = "white") +
    scale_fill_gradient(low = "white", high = "#0B6E4F") +
    labs(
      title = paste("Cluster-Weight Aggregation Weights:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      fill = "Weight"
    )
}

plot_drift <- function(locus_id) {
  ref_run_id <- nearest_reference_run(locus_id)
  ref_pip <- reference_pip_tbl(locus_id)
  run_tbl <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    dplyr::arrange(.data$sigma_0_2_scalar, .data$c_value)
  pip_tbl <- variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  .data$run_id %in% run_tbl$run_id) %>%
    dplyr::select(.data$run_id, .data$variant_id, .data$pip) %>%
    dplyr::inner_join(ref_pip, by = "variant_id")
  df <- run_tbl %>%
    dplyr::mutate(
      jsd_from_reference_c0_sigma02 = purrr::map_dbl(.data$run_id, function(rid) {
        x <- pip_tbl %>%
          dplyr::filter(.data$run_id == .env$rid) %>%
          dplyr::arrange(.data$variant_id)
        if (!nrow(x)) return(NA_real_)
        js_distance(x$pip, x$baseline_pip)
      })
    )

  ggplot(df, aes(x = c_value, y = jsd_from_reference_c0_sigma02, color = factor(sigma_0_2_scalar))) +
    geom_line(linewidth = 0.7) +
    geom_point(size = 2) +
    labs(
      title = paste("PIP Drift From c = 0, sigma0^2 = 0.2 Reference:", locus_id),
      subtitle = paste("Reference run_id:", ref_run_id),
      x = "Annotation scale c",
      y = "JSD from nearest c = 0, sigma0^2 = 0.2 baseline",
      color = expression(sigma[0]^2)
    )
}

plot_elbo_heatmap <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    add_grid_labels()

  ggplot(df, aes(x = c_label, y = sigma_label, fill = elbo_final)) +
    geom_tile(color = "white") +
    scale_fill_viridis_c() +
    labs(
      title = paste("ELBO Heatmap:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      fill = "ELBO"
    )
}

plot_anchor_comparison <- function(locus_id) {
  row <- anchor_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)

  status <- if ("susie_anchor_status" %in% names(row) && nrow(row))
    row$susie_anchor_status[[1]] else "ok"
  err_msg <- if ("susie_anchor_error" %in% names(row) && nrow(row))
    row$susie_anchor_error[[1]] else NA_character_

  df <- row %>%
    dplyr::transmute(
      locus_id,
      `susieR anchor` = susie_anchor_pve_postmean_std,
      `susine c=0` = baseline_pve_postmean_std
    ) %>%
    tidyr::pivot_longer(-.data$locus_id, names_to = "fit_type", values_to = "pve_postmean_std")

  p <- ggplot(df, aes(x = fit_type, y = pve_postmean_std, fill = fit_type)) +
    geom_col(width = 0.6, na.rm = TRUE) +
    labs(
      title = paste("susine vs susieR posterior-mean PVE:", locus_id),
      x = NULL,
      y = "Posterior-mean PVE proxy"
    ) +
    theme(legend.position = "none")

  if (identical(status, "failed")) {
    p <- p + labs(
      subtitle = paste0("susieR anchor fit FAILED",
                        if (!is.na(err_msg)) paste0(": ", err_msg) else "")
    )
  }
  p
}

plot_aggregated_lollipop <- function(locus_id) {
  agg_df <- aggregated_variants %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    dplyr::arrange(dplyr::desc(.data$aggregated_pip), .data$ld_matrix_index) %>%
    dplyr::slice_head(n = 20)

  anchor_row <- anchor_summary %>% dplyr::filter(.data$locus_id == !!locus_id)
  anchor_id <- anchor_row %>% dplyr::pull(.data$susie_anchor_run_id)
  baseline_id <- anchor_row %>% dplyr::pull(.data$baseline_run_id)
  status <- if ("susie_anchor_status" %in% names(anchor_row) && nrow(anchor_row))
    anchor_row$susie_anchor_status[[1]] else "ok"
  err_msg <- if ("susie_anchor_error" %in% names(anchor_row) && nrow(anchor_row))
    anchor_row$susie_anchor_error[[1]] else NA_character_

  overlay <- variant_posteriors %>%
    dplyr::filter(.data$run_id %in% c(anchor_id, baseline_id), .data$variant_id %in% agg_df$variant_id) %>%
    dplyr::mutate(fit_type = dplyr::case_when(
      .data$run_id %in% anchor_id ~ "susieR anchor",
      .data$run_id %in% baseline_id ~ "susine c=0",
      TRUE ~ "other"
    ))

  base_plot <- ggplot(agg_df, aes(x = reorder(variant_id, aggregated_pip), y = aggregated_pip)) +
    geom_segment(aes(xend = variant_id, y = 0, yend = aggregated_pip), color = "grey70") +
    geom_point(size = 2.5, color = "#0B6E4F") +
    coord_flip() +
    labs(
      title = paste("Aggregated Variant PIPs:", locus_id),
      x = "Variant",
      y = "Aggregated PIP"
    )

  if (identical(status, "failed")) {
    base_plot <- base_plot + labs(
      subtitle = paste0("susieR anchor fit FAILED — no susieR overlay",
                        if (!is.na(err_msg)) paste0(" (", err_msg, ")") else "")
    )
  }

  if (!nrow(overlay)) {
    return(base_plot)
  }

  base_plot +
    geom_point(
      data = overlay,
      aes(x = variant_id, y = pip, color = fit_type),
      size = 2,
      inherit.aes = FALSE
    )
}
```

```{r lollipop-helper-override}
plot_aggregated_lollipop <- function(locus_id) {
  plot_df <- aggregated_variants %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    dplyr::left_join(reference_pip_tbl(locus_id), by = "variant_id") %>%
    dplyr::left_join(refit_pip_tbl(locus_id), by = "variant_id") %>%
    dplyr::mutate(
      max_display_pip = pmax(
        .data$baseline_pip,
        .data$aggregated_pip,
        .data$refit_pip,
        na.rm = TRUE
      )
    )

  display_threshold <- 0.1
  threshold_candidates <- plot_df %>%
    dplyr::filter(is.finite(.data$max_display_pip), .data$max_display_pip >= display_threshold)
  if (nrow(threshold_candidates) > 15L) {
    display_threshold <- sort(threshold_candidates$max_display_pip, decreasing = TRUE)[[15L]]
  }

  plot_df <- plot_df %>%
    dplyr::filter(is.finite(.data$max_display_pip), .data$max_display_pip >= display_threshold) %>%
    dplyr::arrange(
      dplyr::desc(.data$max_display_pip),
      dplyr::desc(.data$aggregated_pip),
      .data$ld_matrix_index
    ) %>%
    dplyr::slice_head(n = 15)

  if (!nrow(plot_df)) {
    plot_df <- aggregated_variants %>%
      dplyr::filter(.data$locus_id == !!locus_id) %>%
      dplyr::arrange(dplyr::desc(.data$aggregated_pip), .data$ld_matrix_index) %>%
      dplyr::slice_head(n = 5) %>%
      dplyr::left_join(reference_pip_tbl(locus_id), by = "variant_id") %>%
      dplyr::left_join(refit_pip_tbl(locus_id), by = "variant_id")
    display_threshold <- NA_real_
  }

  plot_df <- plot_df %>%
    dplyr::arrange(dplyr::desc(.data$aggregated_pip), .data$ld_matrix_index) %>%
    dplyr::mutate(variant_f = factor(.data$variant_id, levels = rev(.data$variant_id)))
  points_df <- dplyr::bind_rows(
    plot_df %>%
      dplyr::transmute(variant_f, pip = baseline_pip, fit_type = "Baseline"),
    plot_df %>%
      dplyr::transmute(variant_f, pip = aggregated_pip, fit_type = "SuSiNE ensemble"),
    plot_df %>%
      dplyr::transmute(variant_f, pip = refit_pip, fit_type = "Warm refit")
  ) %>%
    dplyr::filter(is.finite(.data$pip)) %>%
    dplyr::mutate(fit_type = factor(.data$fit_type, levels = c("Baseline", "SuSiNE ensemble", "Warm refit")))

  ggplot(plot_df, aes(y = variant_f)) +
    geom_segment(
      aes(x = baseline_pip, xend = aggregated_pip, yend = variant_f),
      color = "grey35",
      linewidth = 0.8,
      alpha = 0.75,
      na.rm = TRUE
    ) +
    geom_point(
      data = points_df,
      aes(x = pip, y = variant_f, color = fit_type),
      size = 2.8,
      inherit.aes = FALSE
    ) +
    scale_color_manual(
      values = c("Baseline" = "#666666", "SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"),
      breaks = c("Baseline", "SuSiNE ensemble", "Warm refit"),
      name = NULL
    ) +
    labs(
      title = paste("Baseline vs Ensemble Variant PIPs:", locus_id),
      subtitle = if (is.finite(display_threshold)) {
        sprintf("Variants shown when baseline, ensemble, or warm-refit PIP >= %.3g", display_threshold)
      } else {
        "No variants reached PIP >= 0.1; showing top 5 ensemble PIPs"
      },
      x = "PIP",
      y = "Variant"
    ) +
    theme(legend.position = "bottom")
}
```

# 4. Save Locus Figures

```{r save-figures, fig.width=10, fig.height=6}
unlink(file.path(figure_dir, "*_anchor_comparison.png"))
unlink(file.path(locus_figure_dir, "*.png"))

purrr::walk(loci, function(locus_id) {
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_cluster_grid.png")),
    plot_cluster_grid(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_weight_heatmap.png")),
    plot_weight_heatmap(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_drift.png")),
    plot_drift(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_elbo_heatmap.png")),
    plot_elbo_heatmap(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  if (locus_id %in% anchor_summary$locus_id) {
    ggsave(
      file.path(locus_figure_dir, paste0(locus_id, "_aggregated_lollipop.png")),
      plot_aggregated_lollipop(locus_id),
      width = 9,
      height = 6,
      dpi = 150
    )
  }
})
```

# 5. Study-Level Multimodality Summary

```{r multimodality}
multimodal_plot <- ggplot(
  multimodal_metrics,
  aes(x = reorder(locus_id, mean_jsd), y = mean_jsd, fill = n_clusters)
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Mean PIP JSD by Locus",
    x = "Locus",
    y = "Mean JSD",
    fill = "Clusters"
  )

ggsave(
  file.path(overall_figure_dir, "multimodality_summary.png"),
  multimodal_plot,
  width = 9,
  height = 6,
  dpi = 150
)

multimodal_plot
```

# 6. Study-Level Ensemble Effects

```{r overall-ensemble-effects, fig.width=8, fig.height=6}
overall_weight_tbl <- functional_grid_summary %>%
  add_grid_labels() %>%
  dplyr::group_by(.data$c_label, .data$sigma_label) %>%
  dplyr::summarise(
    mean_weight = mean(.data$agg_weight_run, na.rm = TRUE),
    n_loci = dplyr::n_distinct(.data$locus_id),
    .groups = "drop"
  )

overall_weight_plot <- ggplot(overall_weight_tbl, aes(x = c_label, y = sigma_label, fill = mean_weight)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "#0B6E4F") +
  labs(
    title = "Mean Ensemble Weight Across Loci",
    x = "Annotation scale c",
    y = expression(sigma[0]^2),
    fill = "Mean weight"
  )

ggsave(
  file.path(overall_figure_dir, "overall_mean_aggregation_weight_heatmap.png"),
  overall_weight_plot,
  width = 8,
  height = 5,
  dpi = 150
)

overall_pip_change_tbl <- purrr::map_dfr(loci, function(loc) {
  reference_pip_tbl(loc) %>%
    dplyr::inner_join(
      aggregated_variants %>%
        dplyr::filter(.data$locus_id == !!loc) %>%
        dplyr::select(.data$variant_id, .data$aggregated_pip),
      by = "variant_id"
    ) %>%
    dplyr::left_join(
      refit_pip_tbl(loc),
      by = "variant_id"
    ) %>%
    dplyr::mutate(locus_id = loc, .before = 1L)
})

overall_pip_long_tbl <- overall_pip_change_tbl %>%
  tidyr::pivot_longer(
    cols = dplyr::any_of(c("aggregated_pip", "refit_pip")),
    names_to = "comparison",
    values_to = "comparison_pip"
  ) %>%
  dplyr::mutate(
    comparison = dplyr::recode(
      .data$comparison,
      aggregated_pip = "SuSiNE ensemble",
      refit_pip = "Warm refit"
    ),
    comparison = factor(.data$comparison, levels = c("SuSiNE ensemble", "Warm refit"))
  ) %>%
  dplyr::filter(is.finite(.data$comparison_pip))

overall_pip_change_plot <- ggplot(
  overall_pip_long_tbl,
  aes(x = baseline_pip, y = comparison_pip, color = comparison)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(alpha = 0.18, size = 0.8) +
  scale_color_manual(values = c("SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"), name = NULL) +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    title = "Variant PIP Changes Across Loci",
    x = "Baseline PIP",
    y = "Comparison PIP"
  )

ggsave(
  file.path(overall_figure_dir, "overall_variant_pip_baseline_vs_ensemble_and_refit.png"),
  overall_pip_change_plot,
  width = 6,
  height = 6,
  dpi = 150
)

overall_pve_tbl <- anchor_summary %>%
  dplyr::left_join(
    highest_weight_refit_summary %>%
      dplyr::select(.data$locus_id, refit_pve_postmean_std),
    by = "locus_id"
  ) %>%
  dplyr::transmute(
    locus_id,
    susie_anchor_run_id,
    susie_anchor_pve_postmean_std,
    ensemble_pve_postmean_std,
    refit_pve_postmean_std
  )

pve_lim <- range(c(overall_pve_tbl$susie_anchor_pve_postmean_std,
                   overall_pve_tbl$ensemble_pve_postmean_std,
                   overall_pve_tbl$refit_pve_postmean_std), na.rm = TRUE)
if (!all(is.finite(pve_lim))) pve_lim <- c(0, 1)
pve_pad <- max(0.02, diff(pve_lim) * 0.05)
pve_lim <- c(max(0, pve_lim[[1]] - pve_pad), min(1, pve_lim[[2]] + pve_pad))

overall_pve_long_tbl <- overall_pve_tbl %>%
  tidyr::pivot_longer(
    cols = dplyr::any_of(c("ensemble_pve_postmean_std", "refit_pve_postmean_std")),
    names_to = "comparison",
    values_to = "comparison_pve_postmean_std"
  ) %>%
  dplyr::mutate(
    comparison = dplyr::recode(
      .data$comparison,
      ensemble_pve_postmean_std = "SuSiNE ensemble",
      refit_pve_postmean_std = "Warm refit"
    ),
    comparison = factor(.data$comparison, levels = c("SuSiNE ensemble", "Warm refit"))
  ) %>%
  dplyr::filter(
    is.finite(.data$susie_anchor_pve_postmean_std),
    is.finite(.data$comparison_pve_postmean_std)
  )

overall_pve_change_plot <- ggplot(
  overall_pve_long_tbl,
  aes(x = susie_anchor_pve_postmean_std, y = comparison_pve_postmean_std, color = comparison)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(size = 2.4) +
  geom_text(aes(label = locus_id), size = 2.4, vjust = -0.55, check_overlap = TRUE, show.legend = FALSE) +
  scale_color_manual(values = c("SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"), name = NULL) +
  coord_equal(xlim = pve_lim, ylim = pve_lim) +
  labs(
    title = "Posterior-Mean PVE Proxy: SuSiE vs Ensemble and Refit",
    x = "SuSiE anchor posterior-mean PVE",
    y = "Comparison posterior-mean PVE"
  )

ggsave(
  file.path(overall_figure_dir, "overall_pve_postmean_susie_vs_ensemble_and_refit.png"),
  overall_pve_change_plot,
  width = 6,
  height = 6,
  dpi = 150
)

overall_jsd_triangle_tbl <- paper_summary %>%
  dplyr::transmute(
    locus_id,
    jsd_susie_to_susine = .data$jsd_susie_anchor_vs_susine_ensemble,
    jsd_susie_to_refit = .data$jsd_susie_anchor_vs_refit,
    jsd_susine_to_refit = .data$jsd_susine_ensemble_vs_refit
  ) %>%
  dplyr::filter(
    is.finite(.data$jsd_susie_to_susine),
    is.finite(.data$jsd_susie_to_refit),
    is.finite(.data$jsd_susine_to_refit)
  )

jsd_lim <- range(c(overall_jsd_triangle_tbl$jsd_susie_to_susine,
                   overall_jsd_triangle_tbl$jsd_susie_to_refit), na.rm = TRUE)
if (!all(is.finite(jsd_lim))) jsd_lim <- c(0, 1)
jsd_pad <- max(0.01, diff(jsd_lim) * 0.05)
jsd_lim <- c(max(0, jsd_lim[[1]] - jsd_pad), jsd_lim[[2]] + jsd_pad)

overall_jsd_triangle_plot <- ggplot(
  overall_jsd_triangle_tbl,
  aes(
    x = jsd_susie_to_susine,
    y = jsd_susie_to_refit,
    color = jsd_susine_to_refit,
    size = jsd_susine_to_refit
  )
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(alpha = 0.85, na.rm = TRUE) +
  geom_text(aes(label = locus_id), size = 2.4, vjust = -0.55, check_overlap = TRUE, show.legend = FALSE, na.rm = TRUE) +
  scale_color_viridis_c(option = "C", name = "SuSiNE vs refit JSD") +
  scale_size_continuous(range = c(1.8, 5), name = "SuSiNE vs refit JSD") +
  coord_equal(xlim = jsd_lim, ylim = jsd_lim) +
  labs(
    title = "PIP JSD Triangle: SuSiE, SuSiNE, and Warm Refit",
    x = "SuSiE baseline to SuSiNE ensemble JSD",
    y = "SuSiE baseline to warm refit JSD"
  )

ggsave(
  file.path(overall_figure_dir, "overall_pip_jsd_triangle_susie_susine_refit.png"),
  overall_jsd_triangle_plot,
  width = 6.8,
  height = 6,
  dpi = 150
)

refit_basin_r2_drift_plot <- ggplot(
  highest_weight_refit_basin,
  aes(
    x = basin_r2_drift_highest_weight_vs_susie_anchor,
    y = basin_r2_drift_refit_vs_susie_anchor
  )
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(size = 2.4, color = "#C65D00", na.rm = TRUE) +
  geom_text(aes(label = locus_id), size = 2.4, vjust = -0.55, check_overlap = TRUE, na.rm = TRUE) +
  labs(
    title = "Basin r2 Drift From SuSiE Baseline",
    x = "Highest-weight SuSiNE fit vs SuSiE baseline",
    y = "Warm refit vs SuSiE baseline"
  )

ggsave(
  file.path(overall_figure_dir, "overall_basin_r2_drift_highest_weight_vs_refit.png"),
  refit_basin_r2_drift_plot,
  width = 6,
  height = 6,
  dpi = 150
)

overall_weight_plot
overall_pip_change_plot
overall_pve_change_plot
overall_jsd_triangle_plot
refit_basin_r2_drift_plot
```

# 7. Paper Overall Figure

```{r paper-overall-figure, fig.width=7.52, fig.height=7.8}
theme_real_data_paper <- function(base_size = 8) {
  theme_bw(base_size = base_size) +
    theme(
      panel.grid.major = element_line(color = "grey91", linewidth = 0.22),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "grey35", fill = NA, linewidth = 0.35),
      axis.title = element_text(size = base_size + 0.5),
      axis.text = element_text(size = base_size - 1, color = "grey25"),
      plot.title = element_text(face = "bold", size = base_size + 1.2,
                                margin = margin(b = 3)),
      strip.background = element_rect(fill = "grey88", color = "grey35",
                                      linewidth = 0.35),
      strip.text = element_text(size = base_size - 0.8, color = "grey20"),
      legend.title = element_text(size = base_size - 0.4),
      legend.text = element_text(size = base_size - 1.1),
      legend.key.height = grid::unit(3.2, "mm"),
      legend.key.width = grid::unit(6.0, "mm"),
      plot.margin = margin(2.5, 2.5, 2.5, 2.5)
    )
}

paper_panel_theme <- theme_real_data_paper(base_size = 8) +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.margin = margin(t = -2, b = -4),
    legend.box.margin = margin(t = -2, b = -4),
    axis.title = element_text(size = 8.5, face = "plain"),
    axis.text = element_text(size = 7, face = "plain", color = "grey25"),
    plot.title = element_text(face = "bold", size = 9.2, margin = margin(b = 3)),
    legend.title = element_text(size = 7.6, face = "plain"),
    legend.text = element_text(size = 6.9, face = "plain")
  )

weighted_annotation_tbl <- functional_grid_summary %>%
  dplyr::filter(
    is.finite(.data$c_value),
    is.finite(.data$sigma_0_2_scalar),
    is.finite(.data$agg_weight_run)
  ) %>%
  dplyr::group_by(.data$locus_id, .data$gene_name) %>%
  dplyr::summarise(
    weight_total = sum(.data$agg_weight_run, na.rm = TRUE),
    weighted_c_value = sum(.data$agg_weight_run * .data$c_value, na.rm = TRUE) /
      sum(.data$agg_weight_run, na.rm = TRUE),
    weighted_sigma_0_2_scalar = sum(.data$agg_weight_run * .data$sigma_0_2_scalar, na.rm = TRUE) /
      sum(.data$agg_weight_run, na.rm = TRUE),
    off_c0_weight = sum(.data$agg_weight_run[abs(.data$c_value) > 1e-12], na.rm = TRUE) /
      sum(.data$agg_weight_run, na.rm = TRUE),
    max_model_weight = max(.data$agg_weight_run, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(
    is.finite(.data$weight_total),
    .data$weight_total > 0,
    is.finite(.data$weighted_c_value),
    is.finite(.data$weighted_sigma_0_2_scalar)
  ) %>%
  dplyr::mutate(
    locus_label = dplyr::if_else(
      .data$weighted_c_value > 0.1 | .data$weighted_sigma_0_2_scalar > 0.1,
      sub("_lung$", "", .data$locus_id),
      ""
    )
  )

paper_annotation_weight_plot <- ggplot(
  weighted_annotation_tbl,
  aes(x = weighted_c_value, y = weighted_sigma_0_2_scalar)
) +
  geom_hline(yintercept = 0, color = "grey75", linewidth = 0.4) +
  geom_vline(xintercept = 0, color = "grey75", linewidth = 0.4) +
  geom_point(
    aes(fill = off_c0_weight),
    shape = 21,
    color = "grey20",
    size = 2.6,
    alpha = 0.9
  ) +
  geom_locus_text(
    data = weighted_annotation_tbl,
    aes(label = locus_label),
    size = 2.2,
    seed = 11,
    box.padding = 0.65,
    point.padding = 0.85,
    force = 4,
    force_pull = 0.03,
    max.iter = 20000,
    max.time = 2,
    segment.size = 0.25
  ) +
  scale_fill_viridis_c(option = "D", limits = c(0, 1), name = "Total c>0 weight") +
  guides(
    fill = guide_colorbar(
      title.position = "left",
      title.hjust = 0.5,
      barwidth = grid::unit(27, "mm"),
      barheight = grid::unit(2.6, "mm"),
      label.theme = element_text(size = 6.9),
      title.theme = element_text(size = 7.6, margin = margin(r = 8))
    )
  ) +
  coord_cartesian(ylim = c(-0.08, 1.05), clip = "off") +
  labs(
    title = "Annotation influence varies across loci",
    x = "Weighted annotation scale c",
    y = expression("Weighted " * sigma[0]^2)
  ) +
  paper_panel_theme

source_jsd_tbl <- paper_summary %>%
  dplyr::transmute(
    locus_id,
    highest_weight_source_run_id,
    susie_anchor_run_id,
    # Match the fitted-model basin drift x-axis: this is highest-weight model
    # PIP JSD, not cluster-weight aggregated ensemble PIP JSD.
    jsd_susie_anchor_vs_highest_weight_source = purrr::pmap_dbl(
      list(.data$locus_id, .data$susie_anchor_run_id, .data$highest_weight_source_run_id),
      run_pip_jsd
    )
  )

drift_jsd_tbl <- highest_weight_refit_basin %>%
  dplyr::left_join(
    paper_summary %>%
      dplyr::select(
        .data$locus_id,
        .data$highest_weight_source_agg_weight_run,
        .data$jsd_susie_anchor_vs_refit
      ),
    by = "locus_id"
  ) %>%
  dplyr::left_join(source_jsd_tbl, by = "locus_id")

drift_jsd_long_tbl <- dplyr::bind_rows(
  drift_jsd_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "Highest-weight model",
      basin_r2_drift = .data$basin_r2_drift_highest_weight_vs_susie_anchor,
      pip_jsd = .data$jsd_susie_anchor_vs_highest_weight_source,
      source_weight = .data$highest_weight_source_agg_weight_run
    ),
  drift_jsd_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "Warm refit",
      basin_r2_drift = .data$basin_r2_drift_refit_vs_susie_anchor,
      pip_jsd = .data$jsd_susie_anchor_vs_refit,
      source_weight = .data$highest_weight_source_agg_weight_run
    )
  ) %>%
  dplyr::mutate(
    fit_type = factor(.data$fit_type, levels = c("Highest-weight model", "Warm refit")),
    locus_label = dplyr::if_else(
      .data$fit_type == "Highest-weight model" &
        (.data$basin_r2_drift > 0.1 | .data$pip_jsd > 0.3),
      sub("_lung$", "", .data$locus_id),
      ""
    )
  ) %>%
  dplyr::filter(is.finite(.data$basin_r2_drift), is.finite(.data$pip_jsd))

paper_drift_jsd_plot <- ggplot(
  drift_jsd_long_tbl,
  aes(x = basin_r2_drift, y = pip_jsd)
) +
  geom_path(
    aes(group = locus_id),
    color = "grey70",
    linewidth = 0.5,
    na.rm = TRUE
  ) +
  geom_point(
    aes(color = fit_type),
    size = 2.4,
    alpha = 0.9,
    na.rm = TRUE
  ) +
  geom_locus_text(
    data = drift_jsd_long_tbl,
    aes(label = locus_label),
    seed = 2
  ) +
  scale_color_manual(
    values = c("Highest-weight model" = "#0B6E4F", "Warm refit" = "#C65D00"),
    name = NULL
  ) +
  labs(
    title = "Shifts to PIPs and effect bases can decouple",
    x = expression("Basis " * r^2 * " drift from SuSiE baseline"),
    y = "PIP JSD from SuSiE baseline"
  ) +
  paper_panel_theme

paper_pve_tbl <- anchor_summary %>%
  dplyr::left_join(
    highest_weight_refit_summary %>%
      dplyr::select(.data$locus_id, .data$refit_pve_postmean_std),
    by = "locus_id"
  ) %>%
  dplyr::transmute(
    locus_id,
    susie_anchor_pve = .data$susie_anchor_pve_postmean_std,
    susine_ensemble_pve = .data$ensemble_pve_postmean_std,
    warm_refit_pve = .data$refit_pve_postmean_std
  )

paper_pve_long_tbl <- dplyr::bind_rows(
  paper_pve_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "SuSiNE ensemble",
      susie_anchor_pve,
      comparison_pve = .data$susine_ensemble_pve
    ),
  paper_pve_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "Warm refit",
      susie_anchor_pve,
      comparison_pve = .data$warm_refit_pve
    )
) %>%
  dplyr::mutate(
    fit_type = factor(.data$fit_type, levels = c("SuSiNE ensemble", "Warm refit")),
    locus_label = dplyr::if_else(
      .data$fit_type == "SuSiNE ensemble" &
        (.data$susie_anchor_pve > 0.15 | .data$comparison_pve > 0.15),
      sub("_lung$", "", .data$locus_id),
      ""
    )
  ) %>%
  dplyr::filter(is.finite(.data$susie_anchor_pve), is.finite(.data$comparison_pve))

paper_pve_segment_tbl <- paper_pve_tbl %>%
  dplyr::filter(
    is.finite(.data$susie_anchor_pve),
    is.finite(.data$susine_ensemble_pve),
    is.finite(.data$warm_refit_pve)
  )

paper_pve_lim <- range(
  c(paper_pve_long_tbl$susie_anchor_pve, paper_pve_long_tbl$comparison_pve),
  na.rm = TRUE
)
if (!all(is.finite(paper_pve_lim))) paper_pve_lim <- c(0, 1)
paper_pve_pad <- max(0.02, diff(paper_pve_lim) * 0.05)
paper_pve_lim <- c(
  max(0, paper_pve_lim[[1]] - paper_pve_pad),
  min(1, paper_pve_lim[[2]] + paper_pve_pad)
)

paper_pve_plot <- ggplot(
  paper_pve_long_tbl,
  aes(x = susie_anchor_pve, y = comparison_pve)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey55", linewidth = 0.5) +
  geom_segment(
    data = paper_pve_segment_tbl,
    aes(
      x = susie_anchor_pve,
      xend = susie_anchor_pve,
      y = susine_ensemble_pve,
      yend = warm_refit_pve
    ),
    color = "grey70",
    linewidth = 0.5,
    inherit.aes = FALSE
  ) +
  geom_point(
    data = dplyr::filter(paper_pve_long_tbl, .data$fit_type == "Warm refit"),
    aes(color = fit_type),
    size = 2.4,
    alpha = 0.9
  ) +
  geom_point(
    data = dplyr::filter(paper_pve_long_tbl, .data$fit_type == "SuSiNE ensemble"),
    aes(color = fit_type),
    size = 2.4,
    alpha = 0.9
  ) +
  geom_locus_text(
    data = paper_pve_long_tbl,
    aes(label = locus_label),
    seed = 3
  ) +
  scale_color_manual(
    values = c("SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"),
    name = NULL
  ) +
  coord_equal(xlim = paper_pve_lim, ylim = paper_pve_lim) +
  labs(
    title = "Annotations can boost explanatory signal",
    x = expression("SuSiE baseline PVE (" * h[g]^2 * " proxy)"),
    y = expression("Comparison PVE (" * h[g]^2 * " proxy)")
  ) +
  paper_panel_theme

effect_diffuseness_run_tbl <- paper_summary %>%
  dplyr::select(
    .data$locus_id,
    .data$susie_anchor_run_id,
    .data$highest_weight_source_run_id
  ) %>%
  tidyr::pivot_longer(
    cols = c("susie_anchor_run_id", "highest_weight_source_run_id"),
    names_to = "fit_type",
    values_to = "run_id"
  ) %>%
  dplyr::mutate(
    fit_type = dplyr::recode(
      .data$fit_type,
      susie_anchor_run_id = "SuSiE baseline",
      highest_weight_source_run_id = "Highest-weight model"
    ),
    fit_type = factor(.data$fit_type, levels = c("SuSiE baseline", "Highest-weight model"))
  ) %>%
  dplyr::filter(!is.na(.data$run_id)) %>%
  dplyr::distinct(.data$locus_id, .data$fit_type, .data$run_id)

effect_diffuseness_tbl <- if (nrow(effect_diffuseness_run_tbl)) {
  effect_posteriors_path <- file.path(aggregated_dir, "effect_posteriors_dataset")
  if (dir.exists(effect_posteriors_path)) {
    arrow::open_dataset(effect_posteriors_path) %>%
      dplyr::filter(.data$run_id %in% !!effect_diffuseness_run_tbl$run_id) %>%
      dplyr::select(.data$run_id, .data$locus_id, .data$effect_l, .data$alpha) %>%
      dplyr::collect() %>%
      dplyr::inner_join(effect_diffuseness_run_tbl, by = c("run_id", "locus_id")) %>%
      dplyr::group_by(.data$locus_id, .data$run_id, .data$fit_type, .data$effect_l) %>%
      dplyr::summarise(
        effect_pip_entropy_core95 = core95_entropy(.data$alpha),
        effect_k_eff_core95 = exp(.data$effect_pip_entropy_core95),
        alpha_mass = sum(.data$alpha, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::filter(
        is.finite(.data$effect_k_eff_core95),
        .data$alpha_mass > 1e-8
      )
  } else {
    warning("effect_posteriors_dataset is missing; skipping effect diffuseness panel.")
    tibble::tibble(
      locus_id = character(),
      run_id = integer(),
      fit_type = factor(levels = c("SuSiE baseline", "Highest-weight model")),
      effect_l = integer(),
      effect_pip_entropy_core95 = numeric(),
      effect_k_eff_core95 = numeric(),
      alpha_mass = numeric()
    )
  }
} else {
  tibble::tibble(
    locus_id = character(),
    run_id = integer(),
    fit_type = factor(levels = c("SuSiE baseline", "Highest-weight model")),
    effect_l = integer(),
    effect_pip_entropy_core95 = numeric(),
    effect_k_eff_core95 = numeric(),
    alpha_mass = numeric()
  )
}

diffuseness_breaks <- c(0, 1.5, 5, 20, 100, 2500, 4000, Inf)
diffuseness_labels <- c(
  "0-1.5", "1.5-5", "5-20", "20-100",
  "100-2.5k", "2.5k-4k", "4k+"
)
effect_diffuseness_count_tbl <- effect_diffuseness_tbl %>%
  dplyr::mutate(
    diffuseness_bin = cut(
      .data$effect_k_eff_core95,
      breaks = diffuseness_breaks,
      labels = diffuseness_labels,
      include.lowest = TRUE,
      right = FALSE
    )
  ) %>%
  dplyr::filter(!is.na(.data$diffuseness_bin)) %>%
  dplyr::count(.data$fit_type, .data$diffuseness_bin, name = "n") %>%
  tidyr::complete(
    fit_type = factor(c("SuSiE baseline", "Highest-weight model"),
                      levels = c("SuSiE baseline", "Highest-weight model")),
    diffuseness_bin = factor(diffuseness_labels, levels = diffuseness_labels),
    fill = list(n = 0L)
  ) %>%
  dplyr::mutate(
    diffuseness_bin = factor(.data$diffuseness_bin, levels = diffuseness_labels)
  )

paper_effect_diffuseness_plot <- ggplot(
  effect_diffuseness_count_tbl,
  aes(x = diffuseness_bin, y = n, color = fit_type, fill = fit_type)
) +
  geom_col(
    position = "identity",
    width = 0.82,
    alpha = 0.38,
    linewidth = 0.25,
    na.rm = TRUE
  ) +
  scale_color_manual(
    values = c("SuSiE baseline" = "grey20", "Highest-weight model" = "#0B6E4F"),
    name = NULL
  ) +
  scale_fill_manual(
    values = c("SuSiE baseline" = "grey20", "Highest-weight model" = "#0B6E4F"),
    name = NULL
  ) +
  labs(
    title = "Annotations sharpen effects",
    x = expression("Effective top-95% support size  " * K[l]),
    y = "Effect count"
  ) +
  paper_panel_theme +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_a.png"),
  paper_annotation_weight_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)
ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_b.png"),
  paper_drift_jsd_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)
ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_c.png"),
  paper_pve_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)
ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_d.png"),
  paper_effect_diffuseness_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)

if (requireNamespace("patchwork", quietly = TRUE)) {
  paper_overall_figure <- (
    (paper_annotation_weight_plot | paper_drift_jsd_plot) /
      patchwork::plot_spacer() /
      (paper_pve_plot | paper_effect_diffuseness_plot)
  ) +
    patchwork::plot_layout(heights = c(1, 0.05, 1)) +
    patchwork::plot_annotation(tag_levels = "A") &
    theme(
      plot.tag = element_text(face = "bold", size = 12,
                              margin = margin(0, 1, 0, 0)),
      plot.tag.position = c(0.002, 0.995),
      plot.margin = margin(4.5, 4.5, 4.5, 4.5),
      panel.spacing.x = grid::unit(3.2, "mm"),
      panel.spacing.y = grid::unit(16.0, "mm")
    )

  ggsave(
    file.path(overall_figure_dir, "paper_overall_annotation_drift_pve.png"),
    paper_overall_figure,
    width = 7.52,
    height = 7.8,
    dpi = 300,
    bg = "white",
    limitsize = FALSE
  )
  ggsave(
    file.path(overall_figure_dir, "paper_overall_annotation_drift_pve.pdf"),
    paper_overall_figure,
    width = 7.52,
    height = 7.8,
    bg = "white",
    limitsize = FALSE
  )

  paper_overall_figure
} else {
  warning("patchwork is not installed; saved individual paper panels only.")
  list(
    annotation = paper_annotation_weight_plot,
    drift_jsd = paper_drift_jsd_plot,
    pve = paper_pve_plot,
    effect_diffuseness = paper_effect_diffuseness_plot
  )
}
```

# 8. Selected Locus Zoom Figure

```{r paper-selected-locus-zoom, fig.width=7.52, fig.height=9.2}
resolve_locus_token <- function(token, locus_ids) {
  hit <- locus_ids[grepl(paste0("^", token), locus_ids, ignore.case = TRUE)]
  if (!length(hit)) {
    hit <- locus_ids[grepl(token, locus_ids, ignore.case = TRUE)]
  }
  if (!length(hit)) {
    stop("Could not resolve selected locus token: ", token)
  }
  hit[[1]]
}

selected_locus_tokens <- c("rrp7a", "arsa", "prpf38b")
selected_zoom_loci <- vapply(
  selected_locus_tokens,
  resolve_locus_token,
  character(1),
  locus_ids = loci
)

paper_locus_label <- function(locus_id) sub("_lung$", "", locus_id)

selected_lollipop_xlim <- c(0, 1)
selected_drift_tbl <- purrr::map_dfr(selected_zoom_loci, function(loc) {
  ref_pip <- reference_pip_tbl(loc)
  run_tbl <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!loc) %>%
    dplyr::arrange(.data$sigma_0_2_scalar, .data$c_value)
  pip_tbl <- variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!loc,
                  .data$run_id %in% run_tbl$run_id) %>%
    dplyr::select(.data$run_id, .data$variant_id, .data$pip) %>%
    dplyr::inner_join(ref_pip, by = "variant_id")
  run_tbl %>%
    dplyr::mutate(
      jsd_from_reference_c0_sigma02 = purrr::map_dbl(.data$run_id, function(rid) {
        x <- pip_tbl %>%
          dplyr::filter(.data$run_id == .env$rid) %>%
          dplyr::arrange(.data$variant_id)
        if (!nrow(x)) return(NA_real_)
        js_distance(x$pip, x$baseline_pip)
      }),
      locus_id = loc,
      .before = 1L
    )
})
selected_drift_xlim <- range(selected_drift_tbl$c_value, na.rm = TRUE)
if (!all(is.finite(selected_drift_xlim))) selected_drift_xlim <- c(0, 1)
selected_drift_ylim <- range(selected_drift_tbl$jsd_from_reference_c0_sigma02, na.rm = TRUE)
if (!all(is.finite(selected_drift_ylim))) selected_drift_ylim <- c(0, 1)
selected_drift_ylim <- c(0, selected_drift_ylim[[2]] * 1.05)

paper_lollipop_plot <- function(locus_id, panel_letter, show_legend = FALSE) {
  plot_aggregated_lollipop(locus_id) +
    scale_x_continuous(limits = selected_lollipop_xlim) +
    labs(
      title = paste0(panel_letter, "   ", paper_locus_label(locus_id), ": variant PIPs"),
      subtitle = NULL
    ) +
    paper_panel_theme +
    theme(
      legend.position = if (show_legend) "bottom" else "none",
      axis.text.y = element_text(size = 5.8)
    )
}

paper_pip_drift_plot <- function(locus_id, panel_letter, show_legend = FALSE) {
  plot_drift(locus_id) +
    coord_cartesian(xlim = selected_drift_xlim, ylim = selected_drift_ylim) +
    labs(
      title = paste0(panel_letter, "   ", paper_locus_label(locus_id), ": PIP drift"),
      subtitle = NULL,
      x = "Annotation scale c",
      y = "PIP JSD from baseline",
      color = expression("Prior variance " * sigma[0]^2)
    ) +
    paper_panel_theme +
    theme(legend.position = if (show_legend) "bottom" else "none")
}

if (requireNamespace("patchwork", quietly = TRUE)) {
  zoom_panel_letters <- LETTERS[seq_len(2L * length(selected_zoom_loci))]
  selected_locus_zoom_rows <- purrr::map2(selected_zoom_loci, seq_along(selected_zoom_loci), function(loc, i) {
    show_legend <- i == length(selected_zoom_loci)
    left_letter <- zoom_panel_letters[[2L * i - 1L]]
    right_letter <- zoom_panel_letters[[2L * i]]
    paper_lollipop_plot(loc, left_letter, show_legend = show_legend) |
      paper_pip_drift_plot(loc, right_letter, show_legend = show_legend)
  })

  selected_locus_zoom_figure <- patchwork::wrap_plots(
    selected_locus_zoom_rows,
    ncol = 1
  ) +
    patchwork::plot_layout(heights = rep(1, length(selected_locus_zoom_rows))) &
    theme(
      plot.margin = margin(4.5, 4.5, 4.5, 4.5),
      panel.spacing.x = grid::unit(3.2, "mm"),
      panel.spacing.y = grid::unit(5.5, "mm")
    )

  ggsave(
    file.path(overall_figure_dir, "paper_selected_locus_zoom_lollipop_drift.png"),
    selected_locus_zoom_figure,
    width = 7.52,
    height = 9.2,
    dpi = 300,
    bg = "white",
    limitsize = FALSE
  )
  ggsave(
    file.path(overall_figure_dir, "paper_selected_locus_zoom_lollipop_drift.pdf"),
    selected_locus_zoom_figure,
    width = 7.52,
    height = 9.2,
    bg = "white",
    limitsize = FALSE
  )

  selected_locus_zoom_figure
} else {
  warning("patchwork is not installed; selected-locus zoom figure skipped.")
  invisible(NULL)
}
```

# 9. Basin r² Drift From c = 0 (one panel per locus)

Drift metric: weighted (1 - r²) on max-weight 1-1 matched per-effect coefficient
vectors, weighted by `min(var_y_explained_a, var_y_explained_b)`. RSS analog of
`weighted_r2_drift` from `effect_matching_drift_analysis_5arm.Rmd`. Uses the
R-weighted cosine `(b_i' R b_j) / sqrt((b_i' R b_i)(b_j' R b_j))` since X is
not available in the SS regime.

```{r basin-r2-drift-compute, message=TRUE}
# Need run_metrics to find fit_rds_paths and locus / sigma / c keys.
run_metrics <- readr::read_csv(
  file.path(aggregated_dir, "run_metrics_full.csv"),
  show_col_types = FALSE
)

# Manifest path for load_real_data_locus_bundle. Read from the job_config that
# real_case_slurm copies into run_history alongside the slurm output.
job_config_path <- file.path(
  here("output", "run_history", job_name, parent_job_id, "job_config.json")
)
job_cfg <- jsonlite::fromJSON(job_config_path, simplifyVector = FALSE)
manifest_path <- job_cfg$job$manifest_path
repo_root <- job_cfg$paths$repo_root %||% here()

functional_runs_for_drift <- run_metrics %>%
  dplyr::filter(.data$backend == "susine_rss",
                .data$run_family == "functional_grid",
                !is.na(.data$fit_rds_path)) %>%
  dplyr::select(locus_id, run_id, c_value, sigma_0_2_scalar, fit_rds_path)

format_duration <- function(seconds) {
  if (!is.finite(seconds)) return("unknown")
  seconds <- max(0, as.integer(round(seconds)))
  hh <- seconds %/% 3600L
  mm <- (seconds %% 3600L) %/% 60L
  ss <- seconds %% 60L
  if (hh > 0L) sprintf("%dh %02dm %02ds", hh, mm, ss) else sprintf("%dm %02ds", mm, ss)
}

emit_progress_line <- function(x) {
  cat(x, "\n")
  try(flush(stdout()), silent = TRUE)
  if (interactive()) flush.console()
  invisible(x)
}

make_basin_r2_progress <- function(total_units, report_every_pct = 1) {
  total_units <- max(1L, as.integer(total_units))
  done_units <- 0L
  next_pct <- 0
  start_time <- Sys.time()
  function(label = NULL, units = 1L, force = FALSE) {
    done_units <<- min(total_units, done_units + as.integer(units))
    pct <- 100 * done_units / total_units
    if (isTRUE(force) || done_units == total_units || pct >= next_pct) {
      elapsed <- as.numeric(difftime(Sys.time(), start_time, units = "secs"))
      eta <- if (done_units > 0L) elapsed * (total_units - done_units) / done_units else NA_real_
      label_text <- if (!is.null(label) && nzchar(label)) paste0(" | ", label) else ""
      emit_progress_line(sprintf(
        "[%s] basin r2 drift %.1f%% (%d/%d units), elapsed %s, ETA %s%s",
        format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
        pct, done_units, total_units,
        format_duration(elapsed), format_duration(eta),
        label_text
      ))
      while (next_pct <= pct) {
        next_pct <<- next_pct + report_every_pct
      }
    }
    invisible(done_units)
  }
}

basin_r2_drift_path <- file.path(overall_figure_dir, "basin_r2_drift.csv")
if (file.exists(basin_r2_drift_path) && !isTRUE(force_recompute_basin_r2_drift)) {
  basin_r2_drift_tbl <- readr::read_csv(basin_r2_drift_path, show_col_types = FALSE)
} else {
  drift_loci <- sort(unique(functional_runs_for_drift$locus_id))
  basin_r2_progress <- make_basin_r2_progress(
    total_units = length(drift_loci) + 2L * nrow(functional_runs_for_drift),
    report_every_pct = 1
  )
  basin_r2_progress("starting", units = 0L, force = TRUE)
  basin_r2_drift_tbl <- purrr::map_dfr(
    drift_loci,
    function(loc) {
      runs_tbl <- dplyr::filter(functional_runs_for_drift, .data$locus_id == loc)
      basin_r2_progress(
        sprintf("%s: loading bundle (%d runs)", loc, nrow(runs_tbl)),
        units = 0L,
        force = TRUE
      )
      # apply_sample_size_filter = FALSE so R matches the dimensions of fits
      # produced before the filter was introduced. Set TRUE if you have re-run
      # the SLURM job under the filter and want filter-consistent diagnostics.
      bundle <- tryCatch(
        load_real_data_locus_bundle(loc, manifest_path = manifest_path,
                                    repo_root = repo_root,
                                    apply_sample_size_filter = FALSE),
        error = function(e) NULL
      )
      basin_r2_progress(paste0(loc, ": loaded bundle"))
      if (is.null(bundle)) {
        emit_progress_line(paste0("Skipping ", loc, ": bundle load failed."))
        basin_r2_progress(
          paste0(loc, ": skipped remaining work after bundle load failure"),
          units = 2L * nrow(runs_tbl)
        )
        return(tibble::tibble())
      }
      locus_drift_units_done <- 0L
      drift <- real_data_basin_r2_drift_locus(
        bundle$R,
        runs_tbl,
        progress = function(stage, index, total) {
          locus_drift_units_done <<- locus_drift_units_done + 1L
          basin_r2_progress(sprintf("%s: %s %d/%d", loc, stage, index, total))
        }
      )
      expected_locus_drift_units <- 2L * nrow(runs_tbl)
      if (locus_drift_units_done < expected_locus_drift_units) {
        basin_r2_progress(
          paste0(loc, ": accounted for skipped drift work"),
          units = expected_locus_drift_units - locus_drift_units_done
        )
      }
      if (!nrow(drift)) return(tibble::tibble())
      drift %>% dplyr::mutate(locus_id = loc, .before = 1L)
    }
  )
  basin_r2_progress("finished", units = 0L, force = TRUE)
  readr::write_csv(basin_r2_drift_tbl, basin_r2_drift_path)
}
basin_r2_drift_tbl
```

```{r basin-r2-drift-plot, fig.width=10, fig.height=6}
plot_basin_r2_drift <- function(loc) {
  df <- dplyr::filter(basin_r2_drift_tbl, .data$locus_id == loc)
  flag_row <- dataset_metrics_for_flag %>% dplyr::filter(.data$locus_id == loc)
  flagged <- isTRUE(nrow(flag_row) == 1L &&
                    isTRUE(flag_row$ld_z_inconsistency_flag[[1]]))
  ggplot(df, aes(x = c_value, y = basin_r2_drift,
                 color = factor(sigma_0_2_scalar))) +
    geom_line() +
    geom_point(size = 2) +
    labs(
      title = paste0("Basin r² Drift From c = 0: ", loc,
                     if (flagged) "  [LD/z inconsistent]" else ""),
      subtitle = "Weighted (1 - r²) on matched per-effect b_l, R-weighted cosine",
      x = "Annotation scale c",
      y = "Weighted (1 - r²) drift",
      color = expression(sigma[0]^2)
    )
}

# Optional: pull the inconsistency flag from dataset_metrics if present so we
# can annotate flagged loci in the title.
dataset_metrics_path <- file.path(aggregated_dir, "dataset_metrics_full.csv")
dataset_metrics_for_flag <- if (file.exists(dataset_metrics_path)) {
  readr::read_csv(dataset_metrics_path, show_col_types = FALSE)
} else {
  tibble::tibble(locus_id = character(), ld_z_inconsistency_flag = logical())
}

purrr::walk(sort(unique(basin_r2_drift_tbl$locus_id)), function(loc) {
  p <- plot_basin_r2_drift(loc)
  ggsave(
    file.path(locus_figure_dir, paste0(loc, "_basin_r2_drift.png")),
    p, width = 9, height = 5, dpi = 150
  )
  print(p)
})
```
