---
title: "Real-Data Ensemble — Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

This workbook reads consolidated real-data ensemble outputs and renders the
main locus-level visual diagnostics.

```{r setup, message=FALSE}
here::i_am("vignettes/real data pipeline/visualize_results_workbook_real_data_ensemble.Rmd")
library(here)
library(devtools)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(arrow)
library(knitr)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")
theme_set(theme_bw(base_size = 12))
```

# 1. User Configuration

```{r config}
job_name <- "real_data_ensemble_selected10"
parent_job_id <- "REPLACE_WITH_SLURM_PARENT_ID"

aggregated_dir <- here("output", "slurm_output", job_name, parent_job_id, "aggregated")
figure_dir <- here("output", "slurm_output", job_name, parent_job_id, "figures", "real_data_ensemble")
dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Load Outputs

```{r load}
functional_grid_summary <- readr::read_csv(
  file.path(aggregated_dir, "functional_grid_summary.csv"),
  show_col_types = FALSE
)
multimodal_metrics <- readr::read_csv(
  file.path(aggregated_dir, "multimodal_metrics.csv"),
  show_col_types = FALSE
)
run_comparisons <- readr::read_csv(
  file.path(aggregated_dir, "run_comparisons.csv"),
  show_col_types = FALSE
)
anchor_summary <- readr::read_csv(
  file.path(aggregated_dir, "susie_anchor_summary.csv"),
  show_col_types = FALSE
)
cluster_summary <- readr::read_csv(
  file.path(aggregated_dir, "cluster_summary.csv"),
  show_col_types = FALSE
)
top_variants <- readr::read_csv(
  file.path(aggregated_dir, "top_variants_cluster_weight.csv"),
  show_col_types = FALSE
)
aggregated_variants <- arrow::read_parquet(
  file.path(aggregated_dir, "aggregated_variant_pips_cluster_weight.parquet")
) %>%
  tibble::as_tibble()
variant_posteriors <- arrow::open_dataset(
  file.path(aggregated_dir, "variant_posteriors_dataset")
) %>%
  dplyr::collect()

loci <- sort(unique(functional_grid_summary$locus_id))
loci
```

# 3. Plot Helpers

```{r helpers}
plot_cluster_grid <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)

  ggplot(df, aes(x = c_value, y = sigma_0_2_scalar, color = factor(cluster_id))) +
    geom_point(size = 3) +
    geom_vline(aes(xintercept = baseline_c_l), linetype = "dashed", color = "grey40") +
    labs(
      title = paste("Clustered Functional Grid:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      color = "Cluster"
    )
}

plot_weight_heatmap <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)

  ggplot(df, aes(x = c_value, y = sigma_0_2_scalar, fill = agg_weight_run)) +
    geom_tile(color = "white") +
    scale_fill_gradient(low = "white", high = "#0B6E4F") +
    labs(
      title = paste("Cluster-Weight Aggregation Weights:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      fill = "Weight"
    )
}

plot_drift <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)

  ggplot(df, aes(x = c_value, y = jsd_from_same_sigma_c0, color = factor(sigma_0_2_scalar))) +
    geom_line() +
    geom_point(size = 2) +
    labs(
      title = paste("PIP Drift From c = 0:", locus_id),
      x = "Annotation scale c",
      y = "JSD from same-sigma c = 0 baseline",
      color = expression(sigma[0]^2)
    )
}

plot_elbo_heatmap <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)

  ggplot(df, aes(x = c_value, y = sigma_0_2_scalar, fill = elbo_final)) +
    geom_tile(color = "white") +
    scale_fill_viridis_c() +
    labs(
      title = paste("ELBO Heatmap:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      fill = "ELBO"
    )
}

plot_anchor_comparison <- function(locus_id) {
  row <- anchor_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)

  status <- if ("susie_anchor_status" %in% names(row) && nrow(row))
    row$susie_anchor_status[[1]] else "ok"
  err_msg <- if ("susie_anchor_error" %in% names(row) && nrow(row))
    row$susie_anchor_error[[1]] else NA_character_

  df <- row %>%
    dplyr::transmute(
      locus_id,
      `susieR anchor` = susie_anchor_h2_proxy_std,
      `susine c=0` = baseline_h2_proxy_std
    ) %>%
    tidyr::pivot_longer(-.data$locus_id, names_to = "fit_type", values_to = "h2_proxy_std")

  p <- ggplot(df, aes(x = fit_type, y = h2_proxy_std, fill = fit_type)) +
    geom_col(width = 0.6, na.rm = TRUE) +
    labs(
      title = paste("susine vs susieR h2 Proxy:", locus_id),
      x = NULL,
      y = "Standardized h2 proxy"
    ) +
    theme(legend.position = "none")

  if (identical(status, "failed")) {
    p <- p + labs(
      subtitle = paste0("susieR anchor fit FAILED",
                        if (!is.na(err_msg)) paste0(": ", err_msg) else "")
    )
  }
  p
}

plot_aggregated_lollipop <- function(locus_id) {
  agg_df <- aggregated_variants %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    dplyr::arrange(dplyr::desc(.data$aggregated_pip), .data$ld_matrix_index) %>%
    dplyr::slice_head(n = 20)

  anchor_row <- anchor_summary %>% dplyr::filter(.data$locus_id == !!locus_id)
  anchor_id <- anchor_row %>% dplyr::pull(.data$susie_anchor_run_id)
  baseline_id <- anchor_row %>% dplyr::pull(.data$baseline_run_id)
  status <- if ("susie_anchor_status" %in% names(anchor_row) && nrow(anchor_row))
    anchor_row$susie_anchor_status[[1]] else "ok"
  err_msg <- if ("susie_anchor_error" %in% names(anchor_row) && nrow(anchor_row))
    anchor_row$susie_anchor_error[[1]] else NA_character_

  overlay <- variant_posteriors %>%
    dplyr::filter(.data$run_id %in% c(anchor_id, baseline_id), .data$variant_id %in% agg_df$variant_id) %>%
    dplyr::mutate(fit_type = dplyr::case_when(
      .data$run_id %in% anchor_id ~ "susieR anchor",
      .data$run_id %in% baseline_id ~ "susine c=0",
      TRUE ~ "other"
    ))

  base_plot <- ggplot(agg_df, aes(x = reorder(variant_id, aggregated_pip), y = aggregated_pip)) +
    geom_segment(aes(xend = variant_id, y = 0, yend = aggregated_pip), color = "grey70") +
    geom_point(size = 2.5, color = "#0B6E4F") +
    coord_flip() +
    labs(
      title = paste("Aggregated Variant PIPs:", locus_id),
      x = "Variant",
      y = "Aggregated PIP"
    )

  if (identical(status, "failed")) {
    base_plot <- base_plot + labs(
      subtitle = paste0("susieR anchor fit FAILED — no susieR overlay",
                        if (!is.na(err_msg)) paste0(" (", err_msg, ")") else "")
    )
  }

  if (!nrow(overlay)) {
    return(base_plot)
  }

  base_plot +
    geom_point(
      data = overlay,
      aes(x = variant_id, y = pip, color = fit_type),
      size = 2,
      inherit.aes = FALSE
    )
}
```

# 4. Save Locus Figures

```{r save-figures, fig.width=10, fig.height=6}
purrr::walk(loci, function(locus_id) {
  ggsave(
    file.path(figure_dir, paste0(locus_id, "_cluster_grid.png")),
    plot_cluster_grid(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(figure_dir, paste0(locus_id, "_weight_heatmap.png")),
    plot_weight_heatmap(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(figure_dir, paste0(locus_id, "_drift.png")),
    plot_drift(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(figure_dir, paste0(locus_id, "_elbo_heatmap.png")),
    plot_elbo_heatmap(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  if (locus_id %in% anchor_summary$locus_id) {
    ggsave(
      file.path(figure_dir, paste0(locus_id, "_anchor_comparison.png")),
      plot_anchor_comparison(locus_id),
      width = 7,
      height = 4,
      dpi = 150
    )
    ggsave(
      file.path(figure_dir, paste0(locus_id, "_aggregated_lollipop.png")),
      plot_aggregated_lollipop(locus_id),
      width = 9,
      height = 6,
      dpi = 150
    )
  }
})
```

# 5. Study-Level Multimodality Summary

```{r multimodality}
multimodal_plot <- ggplot(
  multimodal_metrics,
  aes(x = reorder(locus_id, mean_jsd), y = mean_jsd, fill = n_clusters)
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Mean PIP JSD by Locus",
    x = "Locus",
    y = "Mean JSD",
    fill = "Clusters"
  )

ggsave(
  file.path(figure_dir, "multimodality_summary.png"),
  multimodal_plot,
  width = 9,
  height = 6,
  dpi = 150
)

multimodal_plot
```

# 6. Basin r² Drift From c = 0 (one panel per locus)

Drift metric: weighted (1 - r²) on max-weight 1-1 matched per-effect coefficient
vectors, weighted by `min(var_y_explained_a, var_y_explained_b)`. RSS analog of
`weighted_r2_drift` from `effect_matching_drift_analysis_5arm.Rmd`. Uses the
R-weighted cosine `(b_i' R b_j) / sqrt((b_i' R b_i)(b_j' R b_j))` since X is
not available in the SS regime.

```{r basin-r2-drift-compute, message=FALSE}
# Need run_metrics to find fit_rds_paths and locus / sigma / c keys.
run_metrics <- readr::read_csv(
  file.path(aggregated_dir, "run_metrics_full.csv"),
  show_col_types = FALSE
)

# Manifest path for load_real_data_locus_bundle. Read from the job_config that
# real_case_slurm copies into run_history alongside the slurm output.
job_config_path <- file.path(
  here("output", "run_history", job_name, parent_job_id, "job_config.json")
)
job_cfg <- jsonlite::fromJSON(job_config_path, simplifyVector = FALSE)
manifest_path <- job_cfg$job$manifest_path
repo_root <- job_cfg$paths$repo_root %||% here()

functional_runs_for_drift <- run_metrics %>%
  dplyr::filter(.data$backend == "susine_rss",
                .data$run_family == "functional_grid",
                !is.na(.data$fit_rds_path)) %>%
  dplyr::select(locus_id, run_id, c_value, sigma_0_2_scalar, fit_rds_path)

basin_r2_drift_tbl <- purrr::map_dfr(
  sort(unique(functional_runs_for_drift$locus_id)),
  function(loc) {
    runs_tbl <- dplyr::filter(functional_runs_for_drift, .data$locus_id == loc)
    bundle <- tryCatch(
      load_real_data_locus_bundle(loc, manifest_path = manifest_path, repo_root = repo_root),
      error = function(e) NULL
    )
    if (is.null(bundle)) {
      message("Skipping ", loc, ": bundle load failed.")
      return(tibble::tibble())
    }
    drift <- real_data_basin_r2_drift_locus(bundle$R, runs_tbl)
    if (!nrow(drift)) return(tibble::tibble())
    drift %>% dplyr::mutate(locus_id = loc, .before = 1L)
  }
)

readr::write_csv(basin_r2_drift_tbl, file.path(figure_dir, "basin_r2_drift.csv"))
basin_r2_drift_tbl
```

```{r basin-r2-drift-plot, fig.width=10, fig.height=6}
plot_basin_r2_drift <- function(loc) {
  df <- dplyr::filter(basin_r2_drift_tbl, .data$locus_id == loc)
  flag_row <- dataset_metrics_for_flag %>% dplyr::filter(.data$locus_id == loc)
  flagged <- isTRUE(nrow(flag_row) == 1L &&
                    isTRUE(flag_row$ld_z_inconsistency_flag[[1]]))
  ggplot(df, aes(x = c_value, y = basin_r2_drift,
                 color = factor(sigma_0_2_scalar))) +
    geom_line() +
    geom_point(size = 2) +
    labs(
      title = paste0("Basin r² Drift From c = 0: ", loc,
                     if (flagged) "  [LD/z inconsistent]" else ""),
      subtitle = "Weighted (1 - r²) on matched per-effect b_l, R-weighted cosine",
      x = "Annotation scale c",
      y = "Weighted (1 - r²) drift",
      color = expression(sigma[0]^2)
    )
}

# Optional: pull the inconsistency flag from dataset_metrics if present so we
# can annotate flagged loci in the title.
dataset_metrics_path <- file.path(aggregated_dir, "dataset_metrics_full.csv")
dataset_metrics_for_flag <- if (file.exists(dataset_metrics_path)) {
  readr::read_csv(dataset_metrics_path, show_col_types = FALSE)
} else {
  tibble::tibble(locus_id = character(), ld_z_inconsistency_flag = logical())
}

purrr::walk(sort(unique(basin_r2_drift_tbl$locus_id)), function(loc) {
  p <- plot_basin_r2_drift(loc)
  ggsave(
    file.path(figure_dir, paste0(loc, "_basin_r2_drift.png")),
    p, width = 9, height = 5, dpi = 150
  )
  print(p)
})
```
