---
title: "Real-Data Ensemble — Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

This workbook reads consolidated real-data ensemble outputs and renders the
main locus-level visual diagnostics.

```{r setup, message=FALSE}
here::i_am("vignettes/real data pipeline/visualize_results_workbook_real_data_ensemble.Rmd")
library(here)
library(devtools)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(arrow)
library(knitr)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")
theme_set(theme_bw(base_size = 12))
```

# 1. User Configuration

```{r config}
job_name <- "real_data_ensemble_geometric_n20"
parent_job_id <- "52906940"
force_recompute_basin_r2_drift <- FALSE

if (identical(parent_job_id, "REPLACE_WITH_SLURM_JOB_ID")) {
  stop("Set parent_job_id to the SLURM job id returned by sbatch before visualizing.")
}

aggregated_dir <- here("output", "slurm_output", job_name, parent_job_id, "aggregated")
figure_dir <- here("output", "slurm_output", job_name, parent_job_id, "figures", "real_data_ensemble")
overall_figure_dir <- file.path(figure_dir, "overall")
locus_figure_dir <- file.path(figure_dir, "per_locus")
dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(overall_figure_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(locus_figure_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Load Outputs

```{r load}
functional_grid_summary <- readr::read_csv(
  file.path(aggregated_dir, "functional_grid_summary.csv"),
  show_col_types = FALSE
)
multimodal_metrics <- readr::read_csv(
  file.path(aggregated_dir, "multimodal_metrics.csv"),
  show_col_types = FALSE
)
run_comparisons <- readr::read_csv(
  file.path(aggregated_dir, "run_comparisons.csv"),
  show_col_types = FALSE
)
anchor_summary <- readr::read_csv(
  file.path(aggregated_dir, "susie_anchor_summary.csv"),
  show_col_types = FALSE
)
highest_weight_refit_summary <- if (file.exists(file.path(aggregated_dir, "highest_weight_refit_summary.csv"))) {
  readr::read_csv(file.path(aggregated_dir, "highest_weight_refit_summary.csv"), show_col_types = FALSE)
} else {
  tibble::tibble(locus_id = character(), refit_pve_postmean_std = numeric())
}
highest_weight_refit_basin <- if (file.exists(file.path(aggregated_dir, "highest_weight_refit_basin_r2_drift.csv"))) {
  readr::read_csv(file.path(aggregated_dir, "highest_weight_refit_basin_r2_drift.csv"), show_col_types = FALSE)
} else {
  tibble::tibble(
    locus_id = character(),
    basin_r2_drift_highest_weight_vs_susie_anchor = numeric(),
    basin_r2_drift_refit_vs_susie_anchor = numeric(),
    basis_drift_var_y_highest_weight_vs_susie_anchor = numeric(),
    basis_drift_var_y_refit_vs_susie_anchor = numeric()
  )
}
for (nm in c(
  "basis_drift_var_y_highest_weight_vs_susie_anchor",
  "basis_drift_var_y_refit_vs_susie_anchor",
  "basis_drift_var_y_refit_vs_highest_weight"
)) {
  if (!nm %in% names(highest_weight_refit_basin)) highest_weight_refit_basin[[nm]] <- NA_real_
}
paper_summary <- if (file.exists(file.path(aggregated_dir, "paper_real_data_ensemble_summary.csv"))) {
  readr::read_csv(file.path(aggregated_dir, "paper_real_data_ensemble_summary.csv"), show_col_types = FALSE)
} else {
  tibble::tibble(
    locus_id = character(),
    susie_anchor_run_id = integer(),
    highest_weight_source_run_id = integer(),
    highest_weight_source_agg_weight_run = numeric(),
    jsd_susie_anchor_vs_susine_ensemble = numeric(),
    jsd_susie_anchor_vs_refit = numeric(),
    jsd_susine_ensemble_vs_refit = numeric()
  )
}
for (nm in c("susie_anchor_run_id", "highest_weight_source_run_id")) {
  if (!nm %in% names(paper_summary)) paper_summary[[nm]] <- NA_integer_
}
if (!"highest_weight_source_agg_weight_run" %in% names(paper_summary)) {
  paper_summary$highest_weight_source_agg_weight_run <- NA_real_
}
cluster_summary <- readr::read_csv(
  file.path(aggregated_dir, "cluster_summary.csv"),
  show_col_types = FALSE
)
top_variants <- readr::read_csv(
  file.path(aggregated_dir, "top_variants_cluster_weight.csv"),
  show_col_types = FALSE
)
aggregated_variants <- arrow::read_parquet(
  file.path(aggregated_dir, "aggregated_variant_pips_cluster_weight.parquet")
) %>%
  tibble::as_tibble()
highest_weight_refit_variants <- if (file.exists(file.path(aggregated_dir, "highest_weight_refit_variant_posteriors.parquet"))) {
  arrow::read_parquet(file.path(aggregated_dir, "highest_weight_refit_variant_posteriors.parquet")) %>%
    tibble::as_tibble()
} else {
  tibble::tibble(locus_id = character(), variant_id = character(), pip = numeric())
}
# Kept lazy on purpose: the variant_posteriors_dataset is ~ (loci x runs x p)
# rows wide and collecting it up front pushed the workbook over RAM limits on
# HPC. All downstream consumers below filter on locus_id (and usually run_id)
# first, so Arrow can push the predicate into the parquet scan. Per-consumer
# `dplyr::collect()` materializes only the slice that consumer needs.
variant_posteriors <- arrow::open_dataset(
  file.path(aggregated_dir, "variant_posteriors_dataset")
)

loci <- sort(unique(functional_grid_summary$locus_id))
loci
```

# 3. Plot Helpers

```{r helpers}
format_grid_value <- function(x) {
  out <- sprintf("%.2f", as.numeric(x))
  out <- sub("0+$", "", out)
  sub("\\.$", "", out)
}

add_grid_labels <- function(df) {
  c_vals <- sort(unique(df$c_value))
  sigma_vals <- sort(unique(df$sigma_0_2_scalar))
  df %>%
    dplyr::mutate(
      c_label = factor(format_grid_value(.data$c_value),
                       levels = format_grid_value(c_vals)),
      sigma_label = factor(format_grid_value(.data$sigma_0_2_scalar),
                           levels = format_grid_value(sigma_vals))
    )
}

nearest_reference_run <- function(locus_id, target_c = 0, target_sigma = 0.2) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  is.finite(.data$c_value),
                  is.finite(.data$sigma_0_2_scalar)) %>%
    dplyr::mutate(
      c_distance = abs(.data$c_value - target_c),
      sigma_distance = abs(log(.data$sigma_0_2_scalar / target_sigma))
    ) %>%
    dplyr::arrange(.data$c_distance, .data$sigma_distance, .data$run_id)
  if (!nrow(df)) return(NA_integer_)
  as.integer(df$run_id[[1]])
}

reference_pip_tbl <- function(locus_id) {
  ref_run_id <- nearest_reference_run(locus_id)
  if (is.na(ref_run_id)) {
    return(tibble::tibble(
      variant_id = character(),
      baseline_pip = numeric(),
      baseline_posterior_mean = numeric(),
      baseline_conditional_effect = numeric()
    ))
  }
  df <- variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  .data$run_id == !!ref_run_id) %>%
    dplyr::collect()
  if (!"posterior_mean" %in% names(df)) df$posterior_mean <- NA_real_
  # Variant-level conditional effect mean: E[beta_j | variant j is included].
  df %>%
    dplyr::transmute(
      .data$variant_id,
      baseline_pip = .data$pip,
      baseline_posterior_mean = .data$posterior_mean,
      # Hard-zero (not NA) below PIP < 0.001: variants the model gives
      # negligible inclusion probability to do not carry a meaningful
      # "effect given inclusion" estimate, and a numeric 0 lets them
      # render at the natural reference point in the lollipop instead of
      # leaving orphan ensemble markers with no baseline counterpart.
      baseline_conditional_effect = dplyr::if_else(
        is.finite(.data$pip) & .data$pip >= 0.001,
        .data$posterior_mean / .data$pip,
        0
      )
    )
}

refit_pip_tbl <- function(locus_id) {
  if (!nrow(highest_weight_refit_variants)) {
    return(tibble::tibble(
      variant_id = character(),
      refit_pip = numeric(),
      refit_posterior_mean = numeric(),
      refit_conditional_effect = numeric()
    ))
  }
  df <- highest_weight_refit_variants %>%
    dplyr::filter(.data$locus_id == !!locus_id)
  if (!"posterior_mean" %in% names(df)) df$posterior_mean <- NA_real_
  df %>%
    dplyr::transmute(
      .data$variant_id,
      refit_pip = .data$pip,
      refit_posterior_mean = .data$posterior_mean,
      refit_conditional_effect = dplyr::if_else(
        is.finite(.data$pip) & .data$pip >= 0.001,
        .data$posterior_mean / .data$pip,
        0
      )
    )
}

real_data_job_context_cache <- NULL
real_data_job_context <- function() {
  if (!is.null(real_data_job_context_cache)) return(real_data_job_context_cache)
  job_config_path <- file.path(
    here("output", "run_history", job_name, parent_job_id, "job_config.json")
  )
  job_cfg <- jsonlite::fromJSON(job_config_path, simplifyVector = FALSE)
  real_data_job_context_cache <<- list(
    manifest_path = job_cfg$job$manifest_path,
    repo_root = job_cfg$paths$repo_root %||% here()
  )
  real_data_job_context_cache
}

selected_locus_R_cache <- new.env(parent = emptyenv())
selected_locus_R_matrix <- function(locus_id) {
  key <- as.character(locus_id)
  if (exists(key, envir = selected_locus_R_cache, inherits = FALSE)) {
    return(get(key, envir = selected_locus_R_cache, inherits = FALSE))
  }
  ctx <- real_data_job_context()
  bundle <- load_real_data_locus_bundle(
    locus_id,
    manifest_path = ctx$manifest_path,
    repo_root = ctx$repo_root,
    apply_sample_size_filter = FALSE
  )
  assign(key, bundle$R, envir = selected_locus_R_cache)
  bundle$R
}

# Floored-alpha-weighted per-variant conditional effect E[b_j | gamma_j = 1].
# For a single fit:
#   mu_lj         = b_hat[l, j] / alpha[l, j]                  (per-component conditional)
#   alpha_lj_clip = max(alpha[l, j], alpha_floor)              (floor at 1e-4 by default)
#   conditional_j = sum_l(alpha_lj_clip * mu_lj) / sum_l(alpha_lj_clip)
#
# For the SuSiNE cluster-weight ensemble (Option B aggregation): aggregate the
# (numerator, denominator) vectors across nominee runs r weighted by w_r:
#   agg_conditional_j = sum_r w_r * num_r(j) / sum_r w_r * denom_r(j)
#
# Cached per locus. Loads the baseline susine c=0 fit, every cluster-weight
# nominee in the ensemble, and the warm refit from their .rds files.
floored_conditional_alpha_floor <- 0.0001

floored_conditional_cache <- new.env(parent = emptyenv())

# Edge cases encountered while computing floored conditionals are appended
# here as one row each and reported by report_floored_conditional_diagnostics()
# at the bottom of the workbook. kind values:
#   "baseline_run_missing"       : nearest_reference_run returned NA
#   "baseline_fit_unreadable"    : fit_rds_path missing on disk or readRDS failed
#   "baseline_fit_invalid"       : fit lacks alpha / b_hat slots or wrong dims
#   "refit_file_missing"         : no .rds under highest_weight_refits/<locus>/
#   "refit_fit_unreadable"       : refit .rds present but readRDS failed
#   "refit_fit_invalid"          : refit fit lacks alpha / b_hat slots or wrong dims
#   "ensemble_no_nominees"       : no rows with agg_weight_run > 0 for this locus
#   "ensemble_nominee_skipped"   : a single nominee fit was unreadable or invalid
#   "ensemble_all_nominees_failed" : every nominee was skipped
#   "denom_zero_variants"        : per-locus count of variants whose ensemble
#                                  denom was 0 (so conditional is NA)
floored_conditional_diagnostics <- new.env(parent = emptyenv())
floored_conditional_diagnostics$rows <- list()

floored_conditional_log <- function(locus_id, kind, detail = NA_character_) {
  floored_conditional_diagnostics$rows[[
    length(floored_conditional_diagnostics$rows) + 1L
  ]] <- tibble::tibble(
    locus_id = as.character(locus_id),
    kind = as.character(kind),
    detail = as.character(detail)
  )
  invisible(NULL)
}

floored_conditional_diagnostics_tbl <- function() {
  if (!length(floored_conditional_diagnostics$rows)) {
    return(tibble::tibble(locus_id = character(), kind = character(), detail = character()))
  }
  dplyr::bind_rows(floored_conditional_diagnostics$rows)
}

report_floored_conditional_diagnostics <- function() {
  diag_tbl <- floored_conditional_diagnostics_tbl()
  cat("\n", strrep("=", 72), "\n", sep = "")
  cat("Floored-conditional edge-case report\n")
  cat(strrep("=", 72), "\n", sep = "")
  if (!nrow(diag_tbl)) {
    cat("No edge cases encountered. All loci had baseline, ensemble, and refit",
        "\nfloored conditionals computed cleanly.\n", sep = "")
    return(invisible(diag_tbl))
  }
  cat(sprintf("%d edge-case row(s) across %d locus/loci.\n\n",
              nrow(diag_tbl), dplyr::n_distinct(diag_tbl$locus_id)))
  summary_tbl <- diag_tbl %>%
    dplyr::count(.data$kind, name = "n") %>%
    dplyr::arrange(dplyr::desc(.data$n))
  print(summary_tbl, n = Inf)
  cat("\nDetail (one row per event):\n")
  print(diag_tbl, n = Inf)
  invisible(diag_tbl)
}

floored_conditional_run_metrics_cache <- NULL
floored_conditional_agg_weights_cache <- NULL

floored_conditional_load_run_metrics <- function() {
  if (!is.null(floored_conditional_run_metrics_cache)) {
    return(floored_conditional_run_metrics_cache)
  }
  tbl <- readr::read_csv(
    file.path(aggregated_dir, "run_metrics_full.csv"),
    show_col_types = FALSE
  ) %>%
    dplyr::select(.data$locus_id, .data$run_id, .data$fit_rds_path)
  floored_conditional_run_metrics_cache <<- tbl
  tbl
}

floored_conditional_load_agg_weights <- function() {
  if (!is.null(floored_conditional_agg_weights_cache)) {
    return(floored_conditional_agg_weights_cache)
  }
  tbl <- readr::read_csv(
    file.path(aggregated_dir, "aggregation_weights_cluster_weight.csv"),
    show_col_types = FALSE
  ) %>%
    dplyr::select(.data$locus_id, .data$run_id, .data$agg_weight_run)
  floored_conditional_agg_weights_cache <<- tbl
  tbl
}

# Returns list(num = colSums(alpha_clip * mu), denom = colSums(alpha_clip)) or
# NULL if the fit lacks the required slots. Handles both susine fits
# (alpha and alpha-weighted b_hat live in fit$effect_fits$) and susieR fits
# (alpha at fit$alpha, raw mu at fit$mu); see real_data_backend_matrices().
floored_conditional_for_fit_object <- function(fit, alpha_floor = floored_conditional_alpha_floor) {
  if (is.null(fit)) return(NULL)
  alpha <- NULL
  mu <- NULL
  if (!is.null(fit$effect_fits$alpha) && !is.null(fit$effect_fits$b_hat)) {
    # susine fit: b_hat is already alpha-weighted (alpha * mu), so recover mu.
    alpha <- as.matrix(fit$effect_fits$alpha)
    alpha_b_hat <- as.matrix(fit$effect_fits$b_hat)
    if (!all(dim(alpha) == dim(alpha_b_hat))) return(NULL)
    mu <- ifelse(alpha > 0, alpha_b_hat / alpha, 0)
  } else if (!is.null(fit$alpha) && !is.null(fit$mu)) {
    # susieR fit: mu is the raw per-component conditional mean.
    alpha <- as.matrix(fit$alpha)
    mu <- as.matrix(fit$mu)
    if (!all(dim(alpha) == dim(mu))) return(NULL)
  } else {
    return(NULL)
  }
  alpha_clip <- pmax(alpha, alpha_floor)
  list(
    num   = colSums(alpha_clip * mu, na.rm = TRUE),
    denom = colSums(alpha_clip,      na.rm = TRUE),
    p     = ncol(alpha)
  )
}

floored_conditional_load_fit <- function(path) {
  if (length(path) != 1L || is.na(path) || !nzchar(path) || !file.exists(path)) return(NULL)
  tryCatch(readRDS(path), error = function(e) NULL)
}

floored_conditional_for_locus <- function(locus_id) {
  key <- as.character(locus_id)
  if (exists(key, envir = floored_conditional_cache, inherits = FALSE)) {
    return(get(key, envir = floored_conditional_cache, inherits = FALSE))
  }

  run_metrics <- floored_conditional_load_run_metrics()
  agg_weights <- floored_conditional_load_agg_weights()

  # Variant ordering: pull from variant_posteriors so the output aligns to
  # variant_id by ld_matrix_index (fit columns are indexed by ld_matrix_index).
  baseline_run_id <- nearest_reference_run(locus_id)
  if (is.na(baseline_run_id)) {
    floored_conditional_log(locus_id, "baseline_run_missing",
                            "nearest_reference_run returned NA (no c=0 functional grid run)")
  }
  variant_order <- if (!is.na(baseline_run_id)) {
    variant_posteriors %>%
      dplyr::filter(.data$locus_id == !!locus_id,
                    .data$run_id == !!baseline_run_id) %>%
      dplyr::select(.data$variant_id, .data$ld_matrix_index) %>%
      dplyr::collect() %>%
      dplyr::arrange(.data$ld_matrix_index)
  } else {
    tibble::tibble(variant_id = character(), ld_matrix_index = integer())
  }
  p <- nrow(variant_order)
  na_vec <- rep(NA_real_, p)

  empty_result <- tibble::tibble(
    variant_id = variant_order$variant_id,
    baseline_conditional_effect_floored = na_vec,
    aggregated_conditional_effect_floored = na_vec,
    refit_conditional_effect_floored = na_vec
  )

  if (!p) {
    assign(key, empty_result, envir = floored_conditional_cache)
    return(empty_result)
  }

  path_for <- function(rid) {
    if (is.na(rid)) return(NA_character_)
    out <- run_metrics %>%
      dplyr::filter(.data$locus_id == !!locus_id, .data$run_id == !!rid) %>%
      dplyr::pull(.data$fit_rds_path)
    if (!length(out)) NA_character_ else out[[1]]
  }

  # Returns either a $num/$denom list or NULL, and reports the specific reason
  # for failure via the floored_conditional_log so the end-of-workbook report
  # can attribute every NA to a concrete cause.
  compute_for_path_with_diag <- function(path, role, run_id = NA) {
    detail_run <- if (!is.na(run_id)) paste0("run_id=", run_id, ", path=", path) else paste0("path=", path)
    if (length(path) != 1L || is.na(path) || !nzchar(path)) {
      floored_conditional_log(locus_id, paste0(role, "_fit_unreadable"),
                              paste0("no fit_rds_path. ", detail_run))
      return(NULL)
    }
    if (!file.exists(path)) {
      floored_conditional_log(locus_id, paste0(role, "_fit_unreadable"),
                              paste0("file does not exist. ", detail_run))
      return(NULL)
    }
    fit <- tryCatch(readRDS(path), error = function(e) {
      floored_conditional_log(locus_id, paste0(role, "_fit_unreadable"),
                              paste0("readRDS failed: ", conditionMessage(e), ". ", detail_run))
      NULL
    })
    if (is.null(fit)) return(NULL)
    fc <- floored_conditional_for_fit_object(fit)
    if (is.null(fc)) {
      floored_conditional_log(locus_id, paste0(role, "_fit_invalid"),
                              paste0("fit lacks priors$alpha or effect_fits$b_hat. ", detail_run))
      return(NULL)
    }
    if (fc$p != p) {
      floored_conditional_log(locus_id, paste0(role, "_fit_invalid"),
                              sprintf("variant-count mismatch: fit has p=%d, locus expects p=%d. %s",
                                      fc$p, p, detail_run))
      return(NULL)
    }
    fc
  }

  # Baseline (susine c = 0)
  baseline_fc <- if (!is.na(baseline_run_id)) {
    compute_for_path_with_diag(path_for(baseline_run_id),
                               role = "baseline", run_id = baseline_run_id)
  } else NULL

  # Warm refit (separate file naming)
  refit_dir <- file.path(aggregated_dir, "highest_weight_refits", locus_id)
  refit_files <- if (dir.exists(refit_dir)) {
    list.files(refit_dir, pattern = "\\.rds$", full.names = TRUE)
  } else {
    character(0)
  }
  refit_fc <- if (length(refit_files)) {
    compute_for_path_with_diag(refit_files[[1]], role = "refit")
  } else {
    floored_conditional_log(locus_id, "refit_file_missing",
                            paste0("no .rds files under ", refit_dir))
    NULL
  }

  # Ensemble: weighted aggregate across cluster-weight nominees
  nominees <- agg_weights %>%
    dplyr::filter(.data$locus_id == !!locus_id, .data$agg_weight_run > 0)
  ens_num <- rep(0, p)
  ens_denom <- rep(0, p)
  ens_valid <- FALSE
  ensemble_nominees_skipped <- 0L
  if (!nrow(nominees)) {
    floored_conditional_log(locus_id, "ensemble_no_nominees",
                            "agg_weight_run > 0 returned 0 rows")
  } else {
    for (i in seq_len(nrow(nominees))) {
      rid <- nominees$run_id[[i]]
      w <- nominees$agg_weight_run[[i]]
      fc <- compute_for_path_with_diag(path_for(rid),
                                       role = "ensemble_nominee", run_id = rid)
      if (is.null(fc)) {
        ensemble_nominees_skipped <- ensemble_nominees_skipped + 1L
        next
      }
      ens_num <- ens_num + w * fc$num
      ens_denom <- ens_denom + w * fc$denom
      ens_valid <- TRUE
    }
    if (!ens_valid) {
      floored_conditional_log(locus_id, "ensemble_all_nominees_failed",
                              sprintf("%d nominee(s) attempted, all skipped",
                                      nrow(nominees)))
    }
  }

  safe_div <- function(n, d) {
    out <- ifelse(d > 0, n / d, NA_real_)
    as.numeric(out)
  }

  if (ens_valid) {
    n_zero_denom <- sum(ens_denom <= 0)
    if (n_zero_denom > 0L) {
      floored_conditional_log(locus_id, "denom_zero_variants",
                              sprintf("%d variant(s) had ensemble denom <= 0 -> conditional NA",
                                      n_zero_denom))
    }
  }

  result <- tibble::tibble(
    variant_id = variant_order$variant_id,
    baseline_conditional_effect_floored = if (!is.null(baseline_fc)) safe_div(baseline_fc$num, baseline_fc$denom) else na_vec,
    aggregated_conditional_effect_floored = if (ens_valid) safe_div(ens_num, ens_denom) else na_vec,
    refit_conditional_effect_floored = if (!is.null(refit_fc)) safe_div(refit_fc$num, refit_fc$denom) else na_vec
  )

  assign(key, result, envir = floored_conditional_cache)
  result
}

signal_contribution_vec <- function(b, R) {
  b <- as.numeric(b)
  if (!length(b) || is.null(R) || nrow(R) != length(b) || ncol(R) != length(b)) {
    return(rep(NA_real_, length(b)))
  }
  as.numeric(b * as.vector(R %*% b))
}

add_signal_contribution_cols <- function(df, locus_id) {
  if (!nrow(df)) return(df)
  R <- selected_locus_R_matrix(locus_id)
  if (is.null(R) || nrow(R) != nrow(df) || ncol(R) != nrow(df)) {
    df$baseline_signal_contribution <- NA_real_
    df$aggregated_signal_contribution <- NA_real_
    df$refit_signal_contribution <- NA_real_
    return(df)
  }
  ord <- order(df$ld_matrix_index)
  inv_ord <- order(ord)
  df$baseline_signal_contribution <- signal_contribution_vec(df$baseline_posterior_mean[ord], R)[inv_ord]
  df$aggregated_signal_contribution <- signal_contribution_vec(df$aggregated_posterior_mean[ord], R)[inv_ord]
  df$refit_signal_contribution <- signal_contribution_vec(df$refit_posterior_mean[ord], R)[inv_ord]
  df
}

run_pip_vec <- function(locus_id, run_id) {
  if (length(run_id) != 1L || is.na(run_id)) return(NULL)
  # Arrow does not implement arrange()+pull() server-side; collect first.
  x <- variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  .data$run_id == !!as.integer(run_id)) %>%
    dplyr::select(.data$ld_matrix_index, .data$pip) %>%
    dplyr::collect() %>%
    dplyr::arrange(.data$ld_matrix_index) %>%
    dplyr::pull(.data$pip)
  if (!length(x)) return(NULL)
  x
}

safe_pip_jsd <- function(x, y) {
  if (is.null(x) || is.null(y) || length(x) != length(y) || !length(x)) {
    return(NA_real_)
  }
  as.numeric(js_distance(x, y))
}

run_pip_jsd <- function(locus_id, run_id_a, run_id_b) {
  safe_pip_jsd(run_pip_vec(locus_id, run_id_a), run_pip_vec(locus_id, run_id_b))
}

core95_entropy <- function(prob, rho = 0.95) {
  p <- as.numeric(prob)
  p[!is.finite(p) | p < 0] <- 0
  s <- sum(p)
  if (!is.finite(s) || s <= 0) return(NA_real_)
  p <- p / s
  ord <- order(p, decreasing = TRUE)
  cum_p <- cumsum(p[ord])
  keep_n <- which(cum_p >= rho)[[1]]
  if (!length(keep_n) || is.na(keep_n)) keep_n <- length(ord)
  core <- p[ord[seq_len(keep_n)]]
  core <- core / sum(core)
  -sum(core[core > 0] * log(core[core > 0]))
}

geom_locus_text <- function(mapping = NULL, data = NULL, ..., size = 2.4,
                            seed = NA, box.padding = 0.35,
                            point.padding = 0.35, force = 1.4,
                            force_pull = 1, max.iter = 10000,
                            max.time = 1, segment.size = 0.3,
                            show.legend = FALSE) {
  if (requireNamespace("ggrepel", quietly = TRUE)) {
    return(ggrepel::geom_text_repel(
      mapping = mapping,
      data = data,
      size = size,
      min.segment.length = 0,
      box.padding = box.padding,
      point.padding = point.padding,
      force = force,
      force_pull = force_pull,
      max.iter = max.iter,
      max.time = max.time,
      segment.size = segment.size,
      max.overlaps = Inf,
      seed = seed,
      show.legend = show.legend,
      ...
    ))
  }
  ggplot2::geom_text(
    mapping = mapping,
    data = data,
    size = size,
    check_overlap = TRUE,
    show.legend = show.legend,
    ...
  )
}

plot_cluster_grid <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)
  sigma_vals <- sort(unique(df$sigma_0_2_scalar))

  ggplot(df, aes(x = c_value, y = sigma_0_2_scalar, color = factor(cluster_id))) +
    geom_point(size = 3) +
    geom_vline(aes(xintercept = baseline_c_l), linetype = "dashed", color = "grey40") +
    scale_y_log10(breaks = sigma_vals, labels = format_grid_value(sigma_vals)) +
    labs(
      title = paste("Clustered Functional Grid:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      color = "Cluster"
    )
}

plot_weight_heatmap <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    add_grid_labels()

  ggplot(df, aes(x = c_label, y = sigma_label, fill = agg_weight_run)) +
    geom_tile(color = "white") +
    scale_fill_gradient(low = "white", high = "#0B6E4F") +
    labs(
      title = paste("Cluster-Weight Aggregation Weights:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      fill = "Weight"
    )
}

plot_drift <- function(locus_id) {
  ref_run_id <- nearest_reference_run(locus_id)
  ref_pip <- reference_pip_tbl(locus_id)
  run_tbl <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    dplyr::arrange(.data$sigma_0_2_scalar, .data$c_value)
  pip_tbl <- variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  .data$run_id %in% run_tbl$run_id) %>%
    dplyr::select(.data$run_id, .data$variant_id, .data$pip) %>%
    dplyr::collect() %>%
    dplyr::inner_join(ref_pip, by = "variant_id")
  df <- run_tbl %>%
    dplyr::mutate(
      jsd_from_reference_c0_sigma02 = purrr::map_dbl(.data$run_id, function(rid) {
        x <- pip_tbl %>%
          dplyr::filter(.data$run_id == .env$rid) %>%
          dplyr::arrange(.data$variant_id)
        if (!nrow(x)) return(NA_real_)
        js_distance(x$pip, x$baseline_pip)
      })
    )

  ggplot(df, aes(x = c_value, y = jsd_from_reference_c0_sigma02, color = factor(sigma_0_2_scalar))) +
    geom_line(linewidth = 0.7) +
    geom_point(size = 2) +
    labs(
      title = paste("PIP Drift From c = 0, sigma0^2 = 0.2 Reference:", locus_id),
      subtitle = paste("Reference run_id:", ref_run_id),
      x = "Annotation scale c",
      y = "JSD from nearest c = 0, sigma0^2 = 0.2 baseline",
      color = expression(sigma[0]^2)
    )
}

plot_elbo_heatmap <- function(locus_id) {
  df <- functional_grid_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    add_grid_labels()

  ggplot(df, aes(x = c_label, y = sigma_label, fill = elbo_final)) +
    geom_tile(color = "white") +
    scale_fill_viridis_c() +
    labs(
      title = paste("ELBO Heatmap:", locus_id),
      x = "Annotation scale c",
      y = expression(sigma[0]^2),
      fill = "ELBO"
    )
}

plot_anchor_comparison <- function(locus_id) {
  row <- anchor_summary %>%
    dplyr::filter(.data$locus_id == !!locus_id)

  status <- if ("susie_anchor_status" %in% names(row) && nrow(row))
    row$susie_anchor_status[[1]] else "ok"
  err_msg <- if ("susie_anchor_error" %in% names(row) && nrow(row))
    row$susie_anchor_error[[1]] else NA_character_

  df <- row %>%
    dplyr::transmute(
      locus_id,
      `susieR anchor` = susie_anchor_pve_postmean_std,
      `susine c=0` = baseline_pve_postmean_std
    ) %>%
    tidyr::pivot_longer(-.data$locus_id, names_to = "fit_type", values_to = "pve_postmean_std")

  p <- ggplot(df, aes(x = fit_type, y = pve_postmean_std, fill = fit_type)) +
    geom_col(width = 0.6, na.rm = TRUE) +
    labs(
      title = paste("susine vs susieR posterior-mean PVE:", locus_id),
      x = NULL,
      y = "Posterior-mean PVE proxy"
    ) +
    theme(legend.position = "none")

  if (identical(status, "failed")) {
    p <- p + labs(
      subtitle = paste0("susieR anchor fit FAILED",
                        if (!is.na(err_msg)) paste0(": ", err_msg) else "")
    )
  }
  p
}

plot_aggregated_lollipop <- function(locus_id) {
  agg_df <- aggregated_variants %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    dplyr::arrange(dplyr::desc(.data$aggregated_pip), .data$ld_matrix_index) %>%
    dplyr::slice_head(n = 20)

  anchor_row <- anchor_summary %>% dplyr::filter(.data$locus_id == !!locus_id)
  anchor_id <- anchor_row %>% dplyr::pull(.data$susie_anchor_run_id)
  baseline_id <- anchor_row %>% dplyr::pull(.data$baseline_run_id)
  status <- if ("susie_anchor_status" %in% names(anchor_row) && nrow(anchor_row))
    anchor_row$susie_anchor_status[[1]] else "ok"
  err_msg <- if ("susie_anchor_error" %in% names(anchor_row) && nrow(anchor_row))
    anchor_row$susie_anchor_error[[1]] else NA_character_

  overlay <- variant_posteriors %>%
    dplyr::filter(.data$locus_id == !!locus_id,
                  .data$run_id %in% c(anchor_id, baseline_id),
                  .data$variant_id %in% agg_df$variant_id) %>%
    dplyr::collect() %>%
    dplyr::mutate(fit_type = dplyr::case_when(
      .data$run_id %in% anchor_id ~ "susieR anchor",
      .data$run_id %in% baseline_id ~ "susine c=0",
      TRUE ~ "other"
    ))

  base_plot <- ggplot(agg_df, aes(x = reorder(variant_id, aggregated_pip), y = aggregated_pip)) +
    geom_segment(aes(xend = variant_id, y = 0, yend = aggregated_pip), color = "grey70") +
    geom_point(size = 2.5, color = "#0B6E4F") +
    coord_flip() +
    labs(
      title = paste("Aggregated Variant PIPs:", locus_id),
      x = "Variant",
      y = "Aggregated PIP"
    )

  if (identical(status, "failed")) {
    base_plot <- base_plot + labs(
      subtitle = paste0("susieR anchor fit FAILED — no susieR overlay",
                        if (!is.na(err_msg)) paste0(" (", err_msg, ")") else "")
    )
  }

  if (!nrow(overlay)) {
    return(base_plot)
  }

  base_plot +
    geom_point(
      data = overlay,
      aes(x = variant_id, y = pip, color = fit_type),
      size = 2,
      inherit.aes = FALSE
    )
}
```

```{r lollipop-helper-override}
selected_lollipop_variant_tbl <- function(
    locus_id,
    n_variants = 15L,
    selection_metric = c("pip_delta", "posterior_mean_delta", "signal_contribution_delta")
) {
  n_variants <- as.integer(n_variants)
  selection_metric <- match.arg(selection_metric)
  locus_aggregated_variants <- aggregated_variants %>%
    dplyr::filter(.data$locus_id == !!locus_id) %>%
    {
      if (!"aggregated_posterior_mean" %in% names(.)) {
        dplyr::mutate(., aggregated_posterior_mean = NA_real_)
      } else {
        .
      }
    }

  # Floored-alpha-weighted conditional effect (Option B aggregation across the
  # SuSiNE ensemble). Overrides the legacy posterior_mean / pip values in
  # reference_pip_tbl / refit_pip_tbl, and replaces the pip<0.001 hardcode
  # for aggregated_conditional_effect entirely. See
  # floored_conditional_for_locus() for the recipe.
  floored_cond_tbl <- floored_conditional_for_locus(locus_id)

  plot_df <- locus_aggregated_variants %>%
    dplyr::left_join(reference_pip_tbl(locus_id), by = "variant_id") %>%
    dplyr::left_join(refit_pip_tbl(locus_id), by = "variant_id") %>%
    dplyr::left_join(floored_cond_tbl, by = "variant_id") %>%
    add_signal_contribution_cols(locus_id) %>%
    dplyr::mutate(
      baseline_conditional_effect = .data$baseline_conditional_effect_floored,
      refit_conditional_effect = .data$refit_conditional_effect_floored,
      aggregated_conditional_effect = .data$aggregated_conditional_effect_floored,
      max_display_pip = pmax(
        .data$baseline_pip,
        .data$aggregated_pip,
        .data$refit_pip,
        na.rm = TRUE
      ),
      abs_delta_baseline_ensemble = abs(.data$aggregated_pip - .data$baseline_pip),
      abs_delta_posterior_mean = abs(.data$aggregated_posterior_mean - .data$baseline_posterior_mean),
      abs_delta_signal_contribution = abs(.data$aggregated_signal_contribution - .data$baseline_signal_contribution),
      selection_score = dplyr::case_when(
        .env$selection_metric == "posterior_mean_delta" ~ .data$abs_delta_posterior_mean,
        .env$selection_metric == "signal_contribution_delta" ~ .data$abs_delta_signal_contribution,
        TRUE ~ .data$abs_delta_baseline_ensemble
      )
    ) %>%
    dplyr::filter(is.finite(.data$selection_score)) %>%
    dplyr::arrange(
      dplyr::desc(.data$selection_score),
      .data$ld_matrix_index
    ) %>%
    dplyr::slice_head(n = n_variants)

  if (!nrow(plot_df)) {
    plot_df <- locus_aggregated_variants %>%
      dplyr::arrange(dplyr::desc(.data$aggregated_pip), .data$ld_matrix_index) %>%
      dplyr::slice_head(n = n_variants) %>%
      dplyr::left_join(reference_pip_tbl(locus_id), by = "variant_id") %>%
      dplyr::left_join(refit_pip_tbl(locus_id), by = "variant_id") %>%
      dplyr::left_join(floored_cond_tbl, by = "variant_id") %>%
      add_signal_contribution_cols(locus_id) %>%
      dplyr::mutate(
        baseline_conditional_effect = .data$baseline_conditional_effect_floored,
        refit_conditional_effect = .data$refit_conditional_effect_floored,
        aggregated_conditional_effect = .data$aggregated_conditional_effect_floored,
        abs_delta_posterior_mean = abs(.data$aggregated_posterior_mean - .data$baseline_posterior_mean),
        abs_delta_signal_contribution = abs(.data$aggregated_signal_contribution - .data$baseline_signal_contribution)
      )
  }

  plot_df %>%
    dplyr::arrange(dplyr::desc(.data$aggregated_pip), .data$ld_matrix_index) %>%
    dplyr::mutate(variant_f = factor(.data$variant_id, levels = rev(.data$variant_id)))
}

plot_aggregated_lollipop <- function(locus_id, n_variants = 15L, selection_metric = "pip_delta") {
  plot_df <- selected_lollipop_variant_tbl(
    locus_id,
    n_variants = n_variants,
    selection_metric = selection_metric
  )
  points_df <- dplyr::bind_rows(
    plot_df %>%
      dplyr::transmute(variant_f, pip = baseline_pip, fit_type = "Baseline"),
    plot_df %>%
      dplyr::transmute(variant_f, pip = aggregated_pip, fit_type = "SuSiNE ensemble"),
    plot_df %>%
      dplyr::transmute(variant_f, pip = refit_pip, fit_type = "Warm refit")
  ) %>%
    dplyr::filter(is.finite(.data$pip)) %>%
    dplyr::mutate(fit_type = factor(.data$fit_type, levels = c("Baseline", "SuSiNE ensemble", "Warm refit")))

  ggplot(plot_df, aes(y = variant_f)) +
    geom_segment(
      aes(x = baseline_pip, xend = aggregated_pip, yend = variant_f),
      color = "grey35",
      linewidth = 0.8,
      alpha = 0.75,
      na.rm = TRUE
    ) +
    geom_point(
      data = points_df,
      aes(x = pip, y = variant_f, color = fit_type),
      size = 2.8,
      inherit.aes = FALSE
    ) +
    scale_color_manual(
      values = c("Baseline" = "#666666", "SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"),
      breaks = c("Baseline", "SuSiNE ensemble", "Warm refit"),
      name = NULL
    ) +
    labs(
      title = paste("Baseline vs Ensemble Variant PIPs:", locus_id),
      subtitle = paste0("Top ", n_variants, " variants by absolute baseline-to-ensemble PIP difference"),
      x = "PIP",
      y = "Variant"
    ) +
    theme(legend.position = "bottom")
}
```

# 4. Save Locus Figures

```{r save-figures, fig.width=10, fig.height=6}
unlink(file.path(figure_dir, "*_anchor_comparison.png"))
unlink(file.path(locus_figure_dir, "*.png"))

purrr::walk(loci, function(locus_id) {
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_cluster_grid.png")),
    plot_cluster_grid(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_weight_heatmap.png")),
    plot_weight_heatmap(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_drift.png")),
    plot_drift(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  ggsave(
    file.path(locus_figure_dir, paste0(locus_id, "_elbo_heatmap.png")),
    plot_elbo_heatmap(locus_id),
    width = 8,
    height = 5,
    dpi = 150
  )
  if (locus_id %in% anchor_summary$locus_id) {
    ggsave(
      file.path(locus_figure_dir, paste0(locus_id, "_aggregated_lollipop.png")),
      plot_aggregated_lollipop(locus_id),
      width = 9,
      height = 6,
      dpi = 150
    )
  }
})
```

# 5. Study-Level Multimodality Summary

```{r multimodality}
multimodal_plot <- ggplot(
  multimodal_metrics,
  aes(x = reorder(locus_id, mean_jsd), y = mean_jsd, fill = n_clusters)
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Mean PIP JSD by Locus",
    x = "Locus",
    y = "Mean JSD",
    fill = "Clusters"
  )

ggsave(
  file.path(overall_figure_dir, "multimodality_summary.png"),
  multimodal_plot,
  width = 9,
  height = 6,
  dpi = 150
)

multimodal_plot
```

# 6. Study-Level Ensemble Effects

```{r overall-ensemble-effects, fig.width=8, fig.height=6}
overall_weight_tbl <- functional_grid_summary %>%
  add_grid_labels() %>%
  dplyr::group_by(.data$c_label, .data$sigma_label) %>%
  dplyr::summarise(
    mean_weight = mean(.data$agg_weight_run, na.rm = TRUE),
    n_loci = dplyr::n_distinct(.data$locus_id),
    .groups = "drop"
  )

overall_weight_plot <- ggplot(overall_weight_tbl, aes(x = c_label, y = sigma_label, fill = mean_weight)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "#0B6E4F") +
  labs(
    title = "Mean Ensemble Weight Across Loci",
    x = "Annotation scale c",
    y = expression(sigma[0]^2),
    fill = "Mean weight"
  )

ggsave(
  file.path(overall_figure_dir, "overall_mean_aggregation_weight_heatmap.png"),
  overall_weight_plot,
  width = 8,
  height = 5,
  dpi = 150
)

overall_pip_change_tbl <- purrr::map_dfr(loci, function(loc) {
  reference_pip_tbl(loc) %>%
    dplyr::inner_join(
      aggregated_variants %>%
        dplyr::filter(.data$locus_id == !!loc) %>%
        dplyr::select(.data$variant_id, .data$aggregated_pip),
      by = "variant_id"
    ) %>%
    dplyr::left_join(
      refit_pip_tbl(loc),
      by = "variant_id"
    ) %>%
    dplyr::mutate(locus_id = loc, .before = 1L)
})

overall_pip_long_tbl <- overall_pip_change_tbl %>%
  tidyr::pivot_longer(
    cols = dplyr::any_of(c("aggregated_pip", "refit_pip")),
    names_to = "comparison",
    values_to = "comparison_pip"
  ) %>%
  dplyr::mutate(
    comparison = dplyr::recode(
      .data$comparison,
      aggregated_pip = "SuSiNE ensemble",
      refit_pip = "Warm refit"
    ),
    comparison = factor(.data$comparison, levels = c("SuSiNE ensemble", "Warm refit"))
  ) %>%
  dplyr::filter(is.finite(.data$comparison_pip))

overall_pip_change_plot <- ggplot(
  overall_pip_long_tbl,
  aes(x = baseline_pip, y = comparison_pip, color = comparison)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(alpha = 0.18, size = 0.8) +
  scale_color_manual(values = c("SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"), name = NULL) +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    title = "Variant PIP Changes Across Loci",
    x = "Baseline PIP",
    y = "Comparison PIP"
  )

ggsave(
  file.path(overall_figure_dir, "overall_variant_pip_baseline_vs_ensemble_and_refit.png"),
  overall_pip_change_plot,
  width = 6,
  height = 6,
  dpi = 150
)

overall_pve_tbl <- anchor_summary %>%
  dplyr::left_join(
    highest_weight_refit_summary %>%
      dplyr::select(.data$locus_id, refit_pve_postmean_std),
    by = "locus_id"
  ) %>%
  dplyr::transmute(
    locus_id,
    susie_anchor_run_id,
    susie_anchor_pve_postmean_std,
    ensemble_pve_postmean_std,
    refit_pve_postmean_std
  )

pve_lim <- range(c(overall_pve_tbl$susie_anchor_pve_postmean_std,
                   overall_pve_tbl$ensemble_pve_postmean_std,
                   overall_pve_tbl$refit_pve_postmean_std), na.rm = TRUE)
if (!all(is.finite(pve_lim))) pve_lim <- c(0, 1)
pve_pad <- max(0.02, diff(pve_lim) * 0.05)
pve_lim <- c(max(0, pve_lim[[1]] - pve_pad), min(1, pve_lim[[2]] + pve_pad))

overall_pve_long_tbl <- overall_pve_tbl %>%
  tidyr::pivot_longer(
    cols = dplyr::any_of(c("ensemble_pve_postmean_std", "refit_pve_postmean_std")),
    names_to = "comparison",
    values_to = "comparison_pve_postmean_std"
  ) %>%
  dplyr::mutate(
    comparison = dplyr::recode(
      .data$comparison,
      ensemble_pve_postmean_std = "SuSiNE ensemble",
      refit_pve_postmean_std = "Warm refit"
    ),
    comparison = factor(.data$comparison, levels = c("SuSiNE ensemble", "Warm refit"))
  ) %>%
  dplyr::filter(
    is.finite(.data$susie_anchor_pve_postmean_std),
    is.finite(.data$comparison_pve_postmean_std)
  )

overall_pve_change_plot <- ggplot(
  overall_pve_long_tbl,
  aes(x = susie_anchor_pve_postmean_std, y = comparison_pve_postmean_std, color = comparison)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(size = 2.4) +
  geom_text(aes(label = locus_id), size = 2.4, vjust = -0.55, check_overlap = TRUE, show.legend = FALSE) +
  scale_color_manual(values = c("SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"), name = NULL) +
  coord_equal(xlim = pve_lim, ylim = pve_lim) +
  labs(
    title = "Posterior-Mean PVE Proxy: SuSiE vs Ensemble and Refit",
    x = "SuSiE anchor posterior-mean PVE",
    y = "Comparison posterior-mean PVE"
  )

ggsave(
  file.path(overall_figure_dir, "overall_pve_postmean_susie_vs_ensemble_and_refit.png"),
  overall_pve_change_plot,
  width = 6,
  height = 6,
  dpi = 150
)

overall_jsd_triangle_tbl <- paper_summary %>%
  dplyr::transmute(
    locus_id,
    jsd_susie_to_susine = .data$jsd_susie_anchor_vs_susine_ensemble,
    jsd_susie_to_refit = .data$jsd_susie_anchor_vs_refit,
    jsd_susine_to_refit = .data$jsd_susine_ensemble_vs_refit
  ) %>%
  dplyr::filter(
    is.finite(.data$jsd_susie_to_susine),
    is.finite(.data$jsd_susie_to_refit),
    is.finite(.data$jsd_susine_to_refit)
  )

jsd_lim <- range(c(overall_jsd_triangle_tbl$jsd_susie_to_susine,
                   overall_jsd_triangle_tbl$jsd_susie_to_refit), na.rm = TRUE)
if (!all(is.finite(jsd_lim))) jsd_lim <- c(0, 1)
jsd_pad <- max(0.01, diff(jsd_lim) * 0.05)
jsd_lim <- c(max(0, jsd_lim[[1]] - jsd_pad), jsd_lim[[2]] + jsd_pad)

overall_jsd_triangle_plot <- ggplot(
  overall_jsd_triangle_tbl,
  aes(
    x = jsd_susie_to_susine,
    y = jsd_susie_to_refit,
    color = jsd_susine_to_refit,
    size = jsd_susine_to_refit
  )
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(alpha = 0.85, na.rm = TRUE) +
  geom_text(aes(label = locus_id), size = 2.4, vjust = -0.55, check_overlap = TRUE, show.legend = FALSE, na.rm = TRUE) +
  scale_color_viridis_c(option = "C", name = "SuSiNE vs refit JSD") +
  scale_size_continuous(range = c(1.8, 5), name = "SuSiNE vs refit JSD") +
  coord_equal(xlim = jsd_lim, ylim = jsd_lim) +
  labs(
    title = "PIP JSD Triangle: SuSiE, SuSiNE, and Warm Refit",
    x = "SuSiE baseline to SuSiNE ensemble JSD",
    y = "SuSiE baseline to warm refit JSD"
  )

ggsave(
  file.path(overall_figure_dir, "overall_pip_jsd_triangle_susie_susine_refit.png"),
  overall_jsd_triangle_plot,
  width = 6.8,
  height = 6,
  dpi = 150
)

refit_basin_r2_drift_plot <- ggplot(
  highest_weight_refit_basin,
  aes(
    x = basin_r2_drift_highest_weight_vs_susie_anchor,
    y = basin_r2_drift_refit_vs_susie_anchor
  )
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50", linewidth = 0.7) +
  geom_point(size = 2.4, color = "#C65D00", na.rm = TRUE) +
  geom_text(aes(label = locus_id), size = 2.4, vjust = -0.55, check_overlap = TRUE, na.rm = TRUE) +
  labs(
    title = "Basin r2 Drift From SuSiE Baseline",
    x = "Highest-weight SuSiNE fit vs SuSiE baseline",
    y = "Warm refit vs SuSiE baseline"
  )

ggsave(
  file.path(overall_figure_dir, "overall_basin_r2_drift_highest_weight_vs_refit.png"),
  refit_basin_r2_drift_plot,
  width = 6,
  height = 6,
  dpi = 150
)

overall_weight_plot
overall_pip_change_plot
overall_pve_change_plot
overall_jsd_triangle_plot
refit_basin_r2_drift_plot
```

# 7. Paper Overall Figure

```{r paper-overall-figure, fig.width=7.52, fig.height=7.8}
theme_real_data_paper <- function(base_size = 8) {
  theme_bw(base_size = base_size) +
    theme(
      panel.grid.major = element_line(color = "grey91", linewidth = 0.22),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "grey35", fill = NA, linewidth = 0.35),
      axis.title = element_text(size = base_size + 0.5),
      axis.text = element_text(size = base_size - 1, color = "grey25"),
      plot.title = element_text(face = "bold", size = base_size + 1.2,
                                margin = margin(b = 3)),
      strip.background = element_rect(fill = "grey88", color = "grey35",
                                      linewidth = 0.35),
      strip.text = element_text(size = base_size - 0.8, color = "grey20"),
      legend.title = element_text(size = base_size - 0.4),
      legend.text = element_text(size = base_size - 1.1),
      legend.key.height = grid::unit(3.2, "mm"),
      legend.key.width = grid::unit(6.0, "mm"),
      plot.margin = margin(2.5, 2.5, 2.5, 2.5)
    )
}

paper_panel_theme <- theme_real_data_paper(base_size = 8) +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.margin = margin(t = -2, b = -4),
    legend.box.margin = margin(t = -2, b = -4),
    axis.title = element_text(size = 8.5, face = "plain"),
    axis.text = element_text(size = 7, face = "plain", color = "grey25"),
    plot.title = element_text(face = "bold", size = 9.2, margin = margin(b = 3)),
    legend.title = element_text(size = 7.6, face = "plain"),
    legend.text = element_text(size = 6.9, face = "plain")
  )

weighted_annotation_tbl <- functional_grid_summary %>%
  dplyr::filter(
    is.finite(.data$c_value),
    is.finite(.data$sigma_0_2_scalar),
    is.finite(.data$agg_weight_run)
  ) %>%
  dplyr::group_by(.data$locus_id, .data$gene_name) %>%
  dplyr::summarise(
    weight_total = sum(.data$agg_weight_run, na.rm = TRUE),
    weighted_c_value = sum(.data$agg_weight_run * .data$c_value, na.rm = TRUE) /
      sum(.data$agg_weight_run, na.rm = TRUE),
    weighted_sigma_0_2_scalar = sum(.data$agg_weight_run * .data$sigma_0_2_scalar, na.rm = TRUE) /
      sum(.data$agg_weight_run, na.rm = TRUE),
    off_c0_weight = sum(.data$agg_weight_run[abs(.data$c_value) > 1e-12], na.rm = TRUE) /
      sum(.data$agg_weight_run, na.rm = TRUE),
    max_model_weight = max(.data$agg_weight_run, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(
    is.finite(.data$weight_total),
    .data$weight_total > 0,
    is.finite(.data$weighted_c_value),
    is.finite(.data$weighted_sigma_0_2_scalar)
  ) %>%
  dplyr::mutate(
    locus_label = dplyr::if_else(
      .data$weighted_c_value > 0.1 | .data$weighted_sigma_0_2_scalar > 0.1,
      sub("_lung$", "", .data$locus_id),
      ""
    )
  )

paper_annotation_weight_plot <- ggplot(
  weighted_annotation_tbl,
  aes(x = weighted_c_value, y = weighted_sigma_0_2_scalar)
) +
  geom_hline(yintercept = 0, color = "grey75", linewidth = 0.4) +
  geom_vline(xintercept = 0, color = "grey75", linewidth = 0.4) +
  geom_point(
    aes(fill = off_c0_weight),
    shape = 21,
    color = "grey20",
    size = 2.6,
    alpha = 0.9
  ) +
  geom_locus_text(
    data = weighted_annotation_tbl,
    aes(label = locus_label),
    size = 2.2,
    seed = 11,
    box.padding = 0.65,
    point.padding = 0.85,
    force = 4,
    force_pull = 0.03,
    max.iter = 20000,
    max.time = 2,
    segment.size = 0.25
  ) +
  scale_fill_viridis_c(option = "D", limits = c(0, 1), name = "Total c>0 weight") +
  guides(
    fill = guide_colorbar(
      title.position = "left",
      title.hjust = 0.5,
      barwidth = grid::unit(27, "mm"),
      barheight = grid::unit(2.6, "mm"),
      label.theme = element_text(size = 6.9),
      title.theme = element_text(size = 7.6, margin = margin(r = 8))
    )
  ) +
  coord_cartesian(ylim = c(-0.08, 1.05), clip = "off") +
  labs(
    title = "Annotation influence varies across loci",
    x = "Weighted annotation scale c",
    y = expression("Weighted " * sigma[0]^2)
  ) +
  paper_panel_theme

source_jsd_tbl <- paper_summary %>%
  dplyr::transmute(
    locus_id,
    highest_weight_source_run_id,
    susie_anchor_run_id,
    # Match the fitted-model basin drift x-axis: this is highest-weight model
    # PIP JSD, not cluster-weight aggregated ensemble PIP JSD.
    jsd_susie_anchor_vs_highest_weight_source = purrr::pmap_dbl(
      list(.data$locus_id, .data$susie_anchor_run_id, .data$highest_weight_source_run_id),
      run_pip_jsd
    )
  )

drift_jsd_tbl <- highest_weight_refit_basin %>%
  dplyr::left_join(
    paper_summary %>%
      dplyr::select(
        .data$locus_id,
        .data$highest_weight_source_agg_weight_run,
        .data$jsd_susie_anchor_vs_refit
      ),
    by = "locus_id"
  ) %>%
  dplyr::left_join(source_jsd_tbl, by = "locus_id")

drift_jsd_long_tbl <- dplyr::bind_rows(
  drift_jsd_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "Highest-weight model",
      basis_drift_var_y = .data$basis_drift_var_y_highest_weight_vs_susie_anchor,
      pip_jsd = .data$jsd_susie_anchor_vs_highest_weight_source,
      source_weight = .data$highest_weight_source_agg_weight_run
    ),
  drift_jsd_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "Warm refit",
      basis_drift_var_y = .data$basis_drift_var_y_refit_vs_susie_anchor,
      pip_jsd = .data$jsd_susie_anchor_vs_refit,
      source_weight = .data$highest_weight_source_agg_weight_run
    )
  ) %>%
  dplyr::mutate(
    fit_type = factor(.data$fit_type, levels = c("Highest-weight model", "Warm refit")),
    locus_label = dplyr::if_else(
      .data$fit_type == "Highest-weight model" &
        (.data$basis_drift_var_y > 0.05 | .data$pip_jsd > 0.3),
      sub("_lung$", "", .data$locus_id),
      ""
    )
  ) %>%
  dplyr::filter(is.finite(.data$basis_drift_var_y), is.finite(.data$pip_jsd))

paper_drift_jsd_plot <- ggplot(
  drift_jsd_long_tbl,
  aes(x = basis_drift_var_y, y = pip_jsd)
) +
  geom_path(
    aes(group = locus_id),
    color = "grey70",
    linewidth = 0.5,
    na.rm = TRUE
  ) +
  geom_point(
    aes(color = fit_type),
    size = 2.4,
    alpha = 0.9,
    na.rm = TRUE
  ) +
  geom_locus_text(
    data = drift_jsd_long_tbl,
    aes(label = locus_label),
    seed = 2
  ) +
  scale_color_manual(
    values = c("Highest-weight model" = "#0B6E4F", "Warm refit" = "#C65D00"),
    name = NULL
  ) +
  labs(
    title = "Annotations refine signal within and across LD groups",
    x = expression("Effect-basis drift from SuSiE baseline (fraction of var(" * y * "))"),
    y = "PIP JSD from SuSiE baseline"
  ) +
  paper_panel_theme

paper_pve_tbl <- anchor_summary %>%
  dplyr::left_join(
    highest_weight_refit_summary %>%
      dplyr::select(.data$locus_id, .data$refit_pve_postmean_std),
    by = "locus_id"
  ) %>%
  dplyr::transmute(
    locus_id,
    susie_anchor_pve = .data$susie_anchor_pve_postmean_std,
    susine_ensemble_pve = .data$ensemble_pve_postmean_std,
    warm_refit_pve = .data$refit_pve_postmean_std
  )

paper_pve_long_tbl <- dplyr::bind_rows(
  paper_pve_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "SuSiNE ensemble",
      susie_anchor_pve,
      comparison_pve = .data$susine_ensemble_pve
    ),
  paper_pve_tbl %>%
    dplyr::transmute(
      locus_id,
      fit_type = "Warm refit",
      susie_anchor_pve,
      comparison_pve = .data$warm_refit_pve
    )
) %>%
  dplyr::mutate(
    fit_type = factor(.data$fit_type, levels = c("SuSiNE ensemble", "Warm refit")),
    locus_label = dplyr::if_else(
      .data$fit_type == "SuSiNE ensemble" &
        (.data$susie_anchor_pve > 0.15 | .data$comparison_pve > 0.15),
      sub("_lung$", "", .data$locus_id),
      ""
    )
  ) %>%
  dplyr::filter(is.finite(.data$susie_anchor_pve), is.finite(.data$comparison_pve))

paper_pve_segment_tbl <- paper_pve_tbl %>%
  dplyr::filter(
    is.finite(.data$susie_anchor_pve),
    is.finite(.data$susine_ensemble_pve),
    is.finite(.data$warm_refit_pve)
  )

paper_pve_lim <- range(
  c(paper_pve_long_tbl$susie_anchor_pve, paper_pve_long_tbl$comparison_pve),
  na.rm = TRUE
)
if (!all(is.finite(paper_pve_lim))) paper_pve_lim <- c(0, 1)
paper_pve_pad <- max(0.02, diff(paper_pve_lim) * 0.05)
paper_pve_lim <- c(
  max(0, paper_pve_lim[[1]] - paper_pve_pad),
  min(1, paper_pve_lim[[2]] + paper_pve_pad)
)

paper_pve_plot <- ggplot(
  paper_pve_long_tbl,
  aes(x = susie_anchor_pve, y = comparison_pve)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey55", linewidth = 0.5) +
  geom_segment(
    data = paper_pve_segment_tbl,
    aes(
      x = susie_anchor_pve,
      xend = susie_anchor_pve,
      y = susine_ensemble_pve,
      yend = warm_refit_pve
    ),
    color = "grey70",
    linewidth = 0.5,
    inherit.aes = FALSE
  ) +
  geom_point(
    data = dplyr::filter(paper_pve_long_tbl, .data$fit_type == "Warm refit"),
    aes(color = fit_type),
    size = 2.4,
    alpha = 0.9
  ) +
  geom_point(
    data = dplyr::filter(paper_pve_long_tbl, .data$fit_type == "SuSiNE ensemble"),
    aes(color = fit_type),
    size = 2.4,
    alpha = 0.9
  ) +
  geom_locus_text(
    data = paper_pve_long_tbl,
    aes(label = locus_label),
    seed = 3
  ) +
  scale_color_manual(
    values = c("SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"),
    name = NULL
  ) +
  coord_equal(xlim = paper_pve_lim, ylim = paper_pve_lim) +
  labs(
    title = "Annotations can boost explanatory signal",
    x = expression("SuSiE baseline PVE (" * h[g]^2 * " proxy)"),
    y = expression("Comparison PVE (" * h[g]^2 * " proxy)")
  ) +
  paper_panel_theme

effect_diffuseness_run_tbl <- paper_summary %>%
  dplyr::select(
    .data$locus_id,
    .data$susie_anchor_run_id,
    .data$highest_weight_source_run_id
  ) %>%
  tidyr::pivot_longer(
    cols = c("susie_anchor_run_id", "highest_weight_source_run_id"),
    names_to = "fit_type",
    values_to = "run_id"
  ) %>%
  dplyr::mutate(
    fit_type = dplyr::recode(
      .data$fit_type,
      susie_anchor_run_id = "SuSiE baseline",
      highest_weight_source_run_id = "Highest-weight model"
    ),
    fit_type = factor(.data$fit_type, levels = c("SuSiE baseline", "Highest-weight model"))
  ) %>%
  dplyr::filter(!is.na(.data$run_id)) %>%
  dplyr::distinct(.data$locus_id, .data$fit_type, .data$run_id)

effect_diffuseness_tbl <- if (nrow(effect_diffuseness_run_tbl)) {
  effect_posteriors_path <- file.path(aggregated_dir, "effect_posteriors_dataset")
  if (dir.exists(effect_posteriors_path)) {
    arrow::open_dataset(effect_posteriors_path) %>%
      dplyr::filter(.data$run_id %in% !!effect_diffuseness_run_tbl$run_id) %>%
      dplyr::select(.data$run_id, .data$locus_id, .data$effect_l, .data$alpha) %>%
      dplyr::collect() %>%
      dplyr::inner_join(effect_diffuseness_run_tbl, by = c("run_id", "locus_id")) %>%
      dplyr::group_by(.data$locus_id, .data$run_id, .data$fit_type, .data$effect_l) %>%
      dplyr::summarise(
        effect_pip_entropy_core95 = core95_entropy(.data$alpha),
        effect_k_eff_core95 = exp(.data$effect_pip_entropy_core95),
        alpha_mass = sum(.data$alpha, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::filter(
        is.finite(.data$effect_k_eff_core95),
        .data$alpha_mass > 1e-8
      )
  } else {
    warning("effect_posteriors_dataset is missing; skipping effect diffuseness panel.")
    tibble::tibble(
      locus_id = character(),
      run_id = integer(),
      fit_type = factor(levels = c("SuSiE baseline", "Highest-weight model")),
      effect_l = integer(),
      effect_pip_entropy_core95 = numeric(),
      effect_k_eff_core95 = numeric(),
      alpha_mass = numeric()
    )
  }
} else {
  tibble::tibble(
    locus_id = character(),
    run_id = integer(),
    fit_type = factor(levels = c("SuSiE baseline", "Highest-weight model")),
    effect_l = integer(),
    effect_pip_entropy_core95 = numeric(),
    effect_k_eff_core95 = numeric(),
    alpha_mass = numeric()
  )
}

diffuseness_breaks <- c(0, 1.5, 5, 20, 100, 2500, 4000, Inf)
diffuseness_labels <- c(
  "0-1.5", "1.5-5", "5-20", "20-100",
  "100-2.5k", "2.5k-4k", "4k+"
)
effect_diffuseness_count_tbl <- effect_diffuseness_tbl %>%
  dplyr::mutate(
    diffuseness_bin = cut(
      .data$effect_k_eff_core95,
      breaks = diffuseness_breaks,
      labels = diffuseness_labels,
      include.lowest = TRUE,
      right = FALSE
    )
  ) %>%
  dplyr::filter(!is.na(.data$diffuseness_bin)) %>%
  dplyr::count(.data$fit_type, .data$diffuseness_bin, name = "n") %>%
  tidyr::complete(
    fit_type = factor(c("SuSiE baseline", "Highest-weight model"),
                      levels = c("SuSiE baseline", "Highest-weight model")),
    diffuseness_bin = factor(diffuseness_labels, levels = diffuseness_labels),
    fill = list(n = 0L)
  ) %>%
  dplyr::mutate(
    diffuseness_bin = factor(.data$diffuseness_bin, levels = diffuseness_labels)
  )

paper_effect_diffuseness_plot <- ggplot(
  effect_diffuseness_count_tbl,
  aes(x = diffuseness_bin, y = n, color = fit_type, fill = fit_type)
) +
  geom_col(
    position = "identity",
    width = 0.82,
    alpha = 0.38,
    linewidth = 0.25,
    na.rm = TRUE
  ) +
  scale_color_manual(
    values = c("SuSiE baseline" = "grey20", "Highest-weight model" = "#0B6E4F"),
    name = NULL
  ) +
  scale_fill_manual(
    values = c("SuSiE baseline" = "grey20", "Highest-weight model" = "#0B6E4F"),
    name = NULL
  ) +
  labs(
    title = "Annotations sharpen effects",
    x = expression("Effective top-95% support size  " * K[l]),
    y = "Effect count"
  ) +
  paper_panel_theme +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_a.png"),
  paper_annotation_weight_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)
ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_b.png"),
  paper_drift_jsd_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)
ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_c.png"),
  paper_pve_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)
ggsave(
  file.path(overall_figure_dir, "paper_overall_annotation_drift_pve_panel_d.png"),
  paper_effect_diffuseness_plot,
  width = 3.76,
  height = 3.9,
  dpi = 300,
  bg = "white"
)

if (requireNamespace("patchwork", quietly = TRUE)) {
  paper_overall_figure <- (
    (paper_annotation_weight_plot | paper_drift_jsd_plot) /
      patchwork::plot_spacer() /
      (paper_pve_plot | paper_effect_diffuseness_plot)
  ) +
    patchwork::plot_layout(heights = c(1, 0.05, 1)) +
    patchwork::plot_annotation(tag_levels = "A") &
    theme(
      plot.tag = element_text(face = "bold", size = 12,
                              margin = margin(0, 1, 0, 0)),
      plot.tag.position = c(0.002, 0.995),
      plot.margin = margin(4.5, 4.5, 4.5, 4.5),
      panel.spacing.x = grid::unit(3.2, "mm"),
      panel.spacing.y = grid::unit(16.0, "mm")
    )

  ggsave(
    file.path(overall_figure_dir, "paper_overall_annotation_drift_pve.png"),
    paper_overall_figure,
    width = 7.52,
    height = 7.8,
    dpi = 300,
    bg = "white",
    limitsize = FALSE
  )
  ggsave(
    file.path(overall_figure_dir, "paper_overall_annotation_drift_pve.pdf"),
    paper_overall_figure,
    width = 7.52,
    height = 7.8,
    bg = "white",
    limitsize = FALSE
  )

  paper_overall_figure
} else {
  warning("patchwork is not installed; saved individual paper panels only.")
  list(
    annotation = paper_annotation_weight_plot,
    drift_jsd = paper_drift_jsd_plot,
    pve = paper_pve_plot,
    effect_diffuseness = paper_effect_diffuseness_plot
  )
}
```

# 8. Selected Locus Zoom Figure

```{r paper-selected-locus-zoom, fig.width=7.52, fig.height=9.2}
resolve_locus_token <- function(token, locus_ids) {
  hit <- locus_ids[grepl(paste0("^", token), locus_ids, ignore.case = TRUE)]
  if (!length(hit)) {
    hit <- locus_ids[grepl(token, locus_ids, ignore.case = TRUE)]
  }
  if (!length(hit)) {
    stop("Could not resolve selected locus token: ", token)
  }
  hit[[1]]
}

selected_locus_tokens <- c("rrp7a", "arsa", "prpf38b")
selected_zoom_loci <- vapply(
  selected_locus_tokens,
  resolve_locus_token,
  character(1),
  locus_ids = loci
)

paper_locus_label <- function(locus_id) sub("_lung$", "", locus_id)

selected_lollipop_xlim <- c(0, 1)
selected_l2_locus_tokens <- c("ydjc", "arsa", "rrp7a", "tmtc1", "lgals9", "znf280b")
selected_l2_n_variants <- 6L
selected_zoom_l2_loci <- vapply(
  selected_l2_locus_tokens,
  resolve_locus_token,
  character(1),
  locus_ids = loci
)

ensure_basin_r2_drift_tbl <- function() {
  basin_r2_drift_path <- file.path(overall_figure_dir, "basin_r2_drift.csv")
  if (file.exists(basin_r2_drift_path) && !isTRUE(force_recompute_basin_r2_drift)) {
    return(readr::read_csv(basin_r2_drift_path, show_col_types = FALSE))
  }

  run_metrics <- readr::read_csv(
    file.path(aggregated_dir, "run_metrics_full.csv"),
    show_col_types = FALSE
  )
  job_config_path <- file.path(
    here("output", "run_history", job_name, parent_job_id, "job_config.json")
  )
  job_cfg <- jsonlite::fromJSON(job_config_path, simplifyVector = FALSE)
  manifest_path <- job_cfg$job$manifest_path
  repo_root <- job_cfg$paths$repo_root %||% here()

  functional_runs_for_drift <- run_metrics %>%
    dplyr::filter(.data$backend == "susine_rss",
                  .data$run_family == "functional_grid",
                  !is.na(.data$fit_rds_path)) %>%
    dplyr::select(locus_id, run_id, c_value, sigma_0_2_scalar, fit_rds_path)

  format_duration <- function(seconds) {
    if (!is.finite(seconds)) return("unknown")
    seconds <- max(0, as.integer(round(seconds)))
    hh <- seconds %/% 3600L
    mm <- (seconds %% 3600L) %/% 60L
    ss <- seconds %% 60L
    if (hh > 0L) sprintf("%dh %02dm %02ds", hh, mm, ss) else sprintf("%dm %02ds", mm, ss)
  }
  emit_progress_line <- function(x) {
    cat(x, "\n")
    try(flush(stdout()), silent = TRUE)
    if (interactive()) flush.console()
    invisible(x)
  }
  make_basin_r2_progress <- function(total_units, report_every_pct = 1) {
    total_units <- max(1L, as.integer(total_units))
    done_units <- 0L
    next_pct <- 0
    start_time <- Sys.time()
    function(label = NULL, units = 1L, force = FALSE) {
      done_units <<- min(total_units, done_units + as.integer(units))
      pct <- 100 * done_units / total_units
      if (isTRUE(force) || done_units == total_units || pct >= next_pct) {
        elapsed <- as.numeric(difftime(Sys.time(), start_time, units = "secs"))
        eta <- if (done_units > 0L) elapsed * (total_units - done_units) / done_units else NA_real_
        label_text <- if (!is.null(label) && nzchar(label)) paste0(" | ", label) else ""
        emit_progress_line(sprintf(
          "[%s] basin r2 drift %.1f%% (%d/%d units), elapsed %s, ETA %s%s",
          format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
          pct, done_units, total_units,
          format_duration(elapsed), format_duration(eta),
          label_text
        ))
        while (next_pct <= pct) {
          next_pct <<- next_pct + report_every_pct
        }
      }
      invisible(done_units)
    }
  }

  drift_loci <- sort(unique(functional_runs_for_drift$locus_id))
  basin_r2_progress <- make_basin_r2_progress(
    total_units = length(drift_loci) + 2L * nrow(functional_runs_for_drift),
    report_every_pct = 1
  )
  basin_r2_progress("starting", units = 0L, force = TRUE)
  basin_r2_drift_tbl <- purrr::map_dfr(
    drift_loci,
    function(loc) {
      runs_tbl <- dplyr::filter(functional_runs_for_drift, .data$locus_id == loc)
      basin_r2_progress(
        sprintf("%s: loading bundle (%d runs)", loc, nrow(runs_tbl)),
        units = 0L,
        force = TRUE
      )
      bundle <- tryCatch(
        load_real_data_locus_bundle(loc, manifest_path = manifest_path,
                                    repo_root = repo_root,
                                    apply_sample_size_filter = FALSE),
        error = function(e) NULL
      )
      basin_r2_progress(paste0(loc, ": loaded bundle"))
      if (is.null(bundle)) {
        emit_progress_line(paste0("Skipping ", loc, ": bundle load failed."))
        basin_r2_progress(
          paste0(loc, ": skipped remaining work after bundle load failure"),
          units = 2L * nrow(runs_tbl)
        )
        return(tibble::tibble())
      }
      locus_drift_units_done <- 0L
      drift <- real_data_basin_r2_drift_locus(
        bundle$R,
        runs_tbl,
        progress = function(stage, index, total) {
          locus_drift_units_done <<- locus_drift_units_done + 1L
          basin_r2_progress(sprintf("%s: %s %d/%d", loc, stage, index, total))
        }
      )
      expected_locus_drift_units <- 2L * nrow(runs_tbl)
      if (locus_drift_units_done < expected_locus_drift_units) {
        basin_r2_progress(
          paste0(loc, ": accounted for skipped drift work"),
          units = expected_locus_drift_units - locus_drift_units_done
        )
      }
      if (!nrow(drift)) return(tibble::tibble())
      drift %>% dplyr::mutate(locus_id = loc, .before = 1L)
    }
  )
  basin_r2_progress("finished", units = 0L, force = TRUE)
  if (!nrow(basin_r2_drift_tbl) && !all(c("locus_id", "run_id", "c_value", "sigma_0_2_scalar", "basin_r2_drift") %in% names(basin_r2_drift_tbl))) {
    basin_r2_drift_tbl <- tibble::tibble(
      locus_id = character(),
      run_id = integer(),
      c_value = numeric(),
      sigma_0_2_scalar = numeric(),
      basin_r2_drift = numeric()
    )
  }
  readr::write_csv(basin_r2_drift_tbl, basin_r2_drift_path)
  basin_r2_drift_tbl
}

build_selected_drift_tbl <- function(selected_loci) {
  purrr::map_dfr(selected_loci, function(loc) {
    ref_pip <- reference_pip_tbl(loc)
    run_tbl <- functional_grid_summary %>%
      dplyr::filter(.data$locus_id == !!loc) %>%
      dplyr::arrange(.data$sigma_0_2_scalar, .data$c_value)
    # Pull only the needed rows from the parquet, join refs once, then split
    # by run_id so the per-cell drift computations are O(p) hash lookups
    # instead of O(runs * p) repeated filters.
    pip_tbl <- variant_posteriors %>%
      dplyr::filter(.data$locus_id == !!loc,
                    .data$run_id %in% run_tbl$run_id) %>%
      dplyr::select(.data$run_id, .data$variant_id, .data$pip) %>%
      dplyr::collect() %>%
      dplyr::inner_join(ref_pip, by = "variant_id") %>%
      dplyr::arrange(.data$variant_id)
    pip_tbl_by_run <- split(pip_tbl, pip_tbl$run_id)
    drift_for_run <- function(rid, kind) {
      x <- pip_tbl_by_run[[as.character(rid)]]
      if (is.null(x) || !nrow(x)) return(NA_real_)
      if (identical(kind, "jsd")) {
        js_distance(x$pip, x$baseline_pip)
      } else {
        sqrt(sum((x$pip - x$baseline_pip)^2, na.rm = TRUE))
      }
    }
    run_tbl %>%
      dplyr::mutate(
        jsd_from_reference_c0_sigma02 = purrr::map_dbl(.data$run_id, drift_for_run, kind = "jsd"),
        l2_from_reference_c0_sigma02 = purrr::map_dbl(.data$run_id, drift_for_run, kind = "l2"),
        locus_id = loc,
        .before = 1L
      )
  })
}

selected_drift_tbl <- build_selected_drift_tbl(selected_zoom_loci)
selected_l2_drift_tbl <- build_selected_drift_tbl(selected_zoom_l2_loci)
selected_l2_effect_tbl <- purrr::map_dfr(
  selected_zoom_l2_loci,
  ~ selected_lollipop_variant_tbl(
    .x,
    n_variants = selected_l2_n_variants,
    selection_metric = "posterior_mean_delta"
  )
)
# Shared signed-conditional-effect xlim across the selected loci. Conditional
# effects E[b_j | gamma_j = 1] live on a roughly common scale across loci
# (bounded by ~ sqrt(var(y) / var(X_j)) ~ 1 on standardized data), so a common
# axis is honest and lets the reader compare effect-size magnitudes across
# rows. Symmetric around zero so positive and negative directions read evenly.
selected_l2_effect_xlim_raw <- range(c(
  selected_l2_effect_tbl$baseline_conditional_effect,
  selected_l2_effect_tbl$aggregated_conditional_effect,
  selected_l2_effect_tbl$refit_conditional_effect
), na.rm = TRUE)
if (!all(is.finite(selected_l2_effect_xlim_raw))) selected_l2_effect_xlim_raw <- c(-1, 1)
selected_l2_effect_xlim_abs <- max(abs(selected_l2_effect_xlim_raw))
if (!is.finite(selected_l2_effect_xlim_abs) || selected_l2_effect_xlim_abs <= 0) {
  selected_l2_effect_xlim_abs <- 1
}
selected_l2_effect_xlim <- c(-1, 1) * selected_l2_effect_xlim_abs * 1.08
# Make sure the axis range covers the hardcoded breaks below; otherwise large
# breaks (e.g. +/-50) would silently fall outside the panel and not render.
selected_l2_effect_xlim <- range(c(
  selected_l2_effect_xlim,
  c(-50, -5, -1, 0, 1, 5, 50)
))

# Signed-log axis for the conditional-effect column. With sigma = 1 and
# base = 2 below, the transform is sign(x) * log2(|x| + 1): asinh-like,
# smoothly passes through zero, linear near the origin (slope 1 / ln 2),
# and log-compressed at large magnitudes. Equal axis spacing now reflects
# equal multiplicative factors (every doubling of |x| moves the same
# distance). Labels stay on the raw conditional-effect scale; only tick
# spacing is warped. Breaks are hardcoded at {0, 1, 5, 50} and mirrored to
# the negative side since conditional effects are signed.
selected_l2_effect_signed_log_sigma <- 1
selected_l2_effect_breaks <- c(-50, -5, -1, 0, 1, 5, 50)
selected_l2_effect_break_labels <- as.character(selected_l2_effect_breaks)
selected_drift_xlim <- range(selected_drift_tbl$c_value, na.rm = TRUE)
if (!all(is.finite(selected_drift_xlim))) selected_drift_xlim <- c(0, 1)
selected_drift_ylim <- range(selected_drift_tbl$jsd_from_reference_c0_sigma02, na.rm = TRUE)
if (!all(is.finite(selected_drift_ylim))) selected_drift_ylim <- c(0, 1)
selected_drift_ylim <- c(0, selected_drift_ylim[[2]] * 1.05)
selected_highest_weight_drift_tbl <- selected_drift_tbl %>%
  dplyr::group_by(.data$locus_id) %>%
  dplyr::arrange(
    dplyr::desc(.data$agg_weight_run),
    dplyr::desc(.data$elbo_final),
    .data$run_id,
    .by_group = TRUE
  ) %>%
  dplyr::slice_head(n = 1) %>%
  dplyr::ungroup()
selected_l2_highest_weight_drift_tbl <- selected_l2_drift_tbl %>%
  dplyr::group_by(.data$locus_id) %>%
  dplyr::arrange(
    dplyr::desc(.data$agg_weight_run),
    dplyr::desc(.data$elbo_final),
    .data$run_id,
    .by_group = TRUE
  ) %>%
  dplyr::slice_head(n = 1) %>%
  dplyr::ungroup()

paper_lollipop_plot <- function(locus_id, panel_letter, show_legend = FALSE) {
  plot_aggregated_lollipop(locus_id) +
    scale_x_continuous(limits = selected_lollipop_xlim) +
    labs(
      title = paste0(panel_letter, "   ", paper_locus_label(locus_id), ": most shifted PIPs"),
      subtitle = NULL
    ) +
    paper_panel_theme +
    theme(
      legend.position = if (show_legend) "bottom" else "none",
      axis.text.y = element_text(size = 5.8)
    )
}

paper_pip_drift_plot <- function(locus_id, panel_letter, show_legend = FALSE) {
  plot_drift(locus_id) +
    geom_point(
      data = dplyr::filter(selected_highest_weight_drift_tbl, .data$locus_id == !!locus_id),
      aes(x = c_value, y = jsd_from_reference_c0_sigma02),
      inherit.aes = FALSE,
      shape = 21,
      size = 4.0,
      stroke = 1.1,
      fill = NA,
      color = "black",
      na.rm = TRUE
    ) +
    coord_cartesian(xlim = selected_drift_xlim, ylim = selected_drift_ylim) +
    labs(
      title = paste0(panel_letter, "   ", paper_locus_label(locus_id), ": PIP drift"),
      subtitle = NULL,
      x = "Annotation scale c",
      y = "PIP JSD from baseline",
      color = expression("Prior variance " * sigma[0]^2)
    ) +
    paper_panel_theme +
    theme(legend.position = if (show_legend) "bottom" else "none")
}

paper_column_header_plot <- function(label, x = 0.5) {
  ggplot() +
    annotate("text", x = x, y = 0.46, label = label,
             fontface = "bold", size = 3.2) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void() +
    theme(plot.margin = margin(2, 0, -1, 0))
}

paper_locus_row_label_plot <- function(locus_id) {
  ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = paper_locus_label(locus_id),
             angle = 90, fontface = "bold", size = 3.0) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE) +
    theme_void()
}

paper_lollipop_l2_plot <- function(locus_id, show_x_axis = FALSE, show_legend = FALSE,
                                   column_title = NULL) {
  plot_aggregated_lollipop(
    locus_id,
    n_variants = selected_l2_n_variants,
    selection_metric = "posterior_mean_delta"
  ) +
    scale_x_continuous(limits = selected_lollipop_xlim) +
    labs(
      title = column_title,
      subtitle = NULL,
      x = if (show_x_axis) "PIP" else NULL,
      y = NULL
    ) +
    paper_panel_theme +
    theme(
      legend.position = if (show_legend) "bottom" else "none",
      axis.text.y = element_text(size = 5.8),
      plot.title = if (is.null(column_title)) element_blank() else element_text(
        face = "bold", size = 9, hjust = 0.5, margin = margin(b = 4)
      ),
      plot.subtitle = element_blank(),
      axis.title.x = if (show_x_axis) element_text(size = 8.5) else element_blank(),
      axis.text.x = if (show_x_axis) element_text(size = 7, color = "grey25") else element_blank(),
      axis.ticks.x = if (show_x_axis) element_line(color = "grey35") else element_blank()
    )
}

paper_pip_drift_l2_plot <- function(locus_id, show_x_axis = FALSE, show_legend = FALSE,
                                    column_title = NULL) {
  df <- dplyr::filter(selected_l2_drift_tbl, .data$locus_id == !!locus_id)
  highlight_df <- dplyr::filter(selected_l2_highest_weight_drift_tbl, .data$locus_id == !!locus_id) %>%
    dplyr::mutate(weight_label = sprintf("%.2f", .data$agg_weight_run))
  y_span <- diff(range(df$l2_from_reference_c0_sigma02, na.rm = TRUE))
  if (!is.finite(y_span) || y_span <= 0) y_span <- 1
  label_nudge_x <- if (nrow(highlight_df) && highlight_df$c_value[[1]] > mean(selected_drift_xlim)) -0.12 else 0.12
  label_nudge_y <- 0.08 * y_span

  p <- ggplot(df, aes(x = c_value, y = l2_from_reference_c0_sigma02,
                      color = factor(sigma_0_2_scalar))) +
    geom_line(linewidth = 0.7, na.rm = TRUE) +
    geom_point(size = 2, na.rm = TRUE) +
    geom_point(
      data = highlight_df,
      aes(x = c_value, y = l2_from_reference_c0_sigma02),
      inherit.aes = FALSE,
      shape = 21,
      size = 4.0,
      stroke = 1.1,
      fill = NA,
      color = "black",
      na.rm = TRUE
    )

  if (nrow(highlight_df)) {
    if (requireNamespace("ggrepel", quietly = TRUE)) {
      p <- p +
        ggrepel::geom_label_repel(
          data = highlight_df,
          aes(x = c_value, y = l2_from_reference_c0_sigma02, label = weight_label),
          inherit.aes = FALSE,
          color = "black",
          fill = "white",
          label.size = 0.18,
          size = 2.5,
          box.padding = 0.35,
          point.padding = 0.55,
          min.segment.length = 0,
          segment.color = "black",
          segment.size = 0.25,
          seed = 17,
          nudge_x = label_nudge_x,
          nudge_y = label_nudge_y,
          max.overlaps = Inf,
          show.legend = FALSE
        )
    } else {
      p <- p +
        geom_label(
          data = highlight_df,
          aes(x = c_value, y = l2_from_reference_c0_sigma02, label = weight_label),
          inherit.aes = FALSE,
          color = "black",
          fill = "white",
          label.size = 0.18,
          size = 2.5,
          nudge_x = label_nudge_x,
          nudge_y = label_nudge_y,
          show.legend = FALSE
        )
    }
  }

  p +
    scale_y_continuous(
      labels = scales::label_number(accuracy = 0.1),
      breaks = scales::breaks_extended(n = 4)
    ) +
    coord_cartesian(xlim = selected_drift_xlim) +
    labs(
      title = column_title,
      x = if (show_x_axis) "Annotation scale c" else NULL,
      y = NULL,
      color = expression("Prior variance " * sigma[0]^2)
    ) +
    paper_panel_theme +
    theme(
      legend.position = if (show_legend) "bottom" else "none",
      plot.title = if (is.null(column_title)) element_blank() else element_text(
        face = "bold", size = 9, hjust = 0.5, margin = margin(b = 4)
      ),
      axis.title.x = if (show_x_axis) element_text(size = 8.5) else element_blank(),
      axis.text.x = if (show_x_axis) element_text(size = 7, color = "grey25") else element_blank(),
      axis.ticks.x = if (show_x_axis) element_line(color = "grey35") else element_blank()
    )
}

paper_signal_contribution_l2_plot <- function(locus_id, show_x_axis = FALSE, show_legend = FALSE,
                                              column_title = NULL) {
  # Despite the legacy name, this panel plots the signed standardized
  # conditional posterior-mean effect E[b_j | gamma_j = 1] for the same
  # variants as the left lollipop column. Using the *conditional* effect
  # (rather than the unconditional E[b_j] = pip_j * E[b_j | gamma_j = 1])
  # decouples the middle column from the left: left = inclusion probability,
  # middle = direction/magnitude given inclusion. The shared x-axis across
  # rows is intentional and meaningful because conditional effects live on a
  # roughly common standardized scale across loci. Variants with pip < 0.001
  # have their conditional effect hard-coded to 0 upstream (in
  # reference_pip_tbl / refit_pip_tbl / selected_lollipop_variant_tbl).
  plot_df <- selected_lollipop_variant_tbl(
    locus_id,
    n_variants = selected_l2_n_variants,
    selection_metric = "posterior_mean_delta"
  )
  points_df <- dplyr::bind_rows(
    plot_df %>%
      dplyr::transmute(variant_f, conditional_effect = baseline_conditional_effect, fit_type = "Baseline"),
    plot_df %>%
      dplyr::transmute(variant_f, conditional_effect = aggregated_conditional_effect, fit_type = "SuSiNE ensemble"),
    plot_df %>%
      dplyr::transmute(variant_f, conditional_effect = refit_conditional_effect, fit_type = "Warm refit")
  ) %>%
    dplyr::filter(is.finite(.data$conditional_effect)) %>%
    dplyr::mutate(fit_type = factor(.data$fit_type, levels = c("Baseline", "SuSiNE ensemble", "Warm refit")))

  ggplot(plot_df, aes(y = variant_f)) +
    geom_vline(xintercept = 0, color = "grey78", linewidth = 0.35) +
    geom_segment(
      aes(x = baseline_conditional_effect, xend = aggregated_conditional_effect, yend = variant_f),
      color = "grey35",
      linewidth = 0.8,
      alpha = 0.75,
      na.rm = TRUE
    ) +
    geom_point(
      data = points_df,
      aes(x = conditional_effect, y = variant_f, color = fit_type),
      size = 2.8,
      inherit.aes = FALSE
    ) +
    scale_color_manual(
      values = c("Baseline" = "#666666", "SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"),
      breaks = c("Baseline", "SuSiNE ensemble", "Warm refit"),
      name = NULL
    ) +
    scale_x_continuous(
      trans = scales::pseudo_log_trans(sigma = selected_l2_effect_signed_log_sigma, base = 2),
      breaks = selected_l2_effect_breaks,
      labels = selected_l2_effect_break_labels,
      limits = selected_l2_effect_xlim,
      oob = scales::oob_squish
    ) +
    labs(
      title = column_title,
      subtitle = NULL,
      x = if (show_x_axis) expression("Conditional effect  " * E * "[" * b[j] * " | " * gamma[j] * " = 1]") else NULL,
      y = NULL
    ) +
    paper_panel_theme +
    theme(
      legend.position = if (show_legend) "bottom" else "none",
      plot.title = if (is.null(column_title)) element_blank() else element_text(
        face = "bold", size = 9, hjust = 0.5, margin = margin(b = 4)
      ),
      plot.subtitle = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.title.x = if (show_x_axis) element_text(size = 8.5) else element_blank(),
      axis.text.x = if (show_x_axis) element_text(size = 7, color = "grey25") else element_blank(),
      axis.ticks.x = if (show_x_axis) element_line(color = "grey35") else element_blank()
    )
}

paper_lollipop_l2_legend_plot <- function() {
  df <- tibble::tibble(
    fit_type = factor(
      c("Baseline", "SuSiNE ensemble", "Warm refit"),
      levels = c("Baseline", "SuSiNE ensemble", "Warm refit")
    ),
    x = c(0.07, 0.43, 0.07),
    y = c(0.68, 0.68, 0.30)
  )
  ggplot(df, aes(x = x, y = y, color = fit_type)) +
    geom_point(size = 2.8) +
    geom_text(aes(label = fit_type), x = df$x + 0.055, hjust = 0,
              size = 2.55, color = "grey15") +
    scale_color_manual(
      values = c("Baseline" = "#666666", "SuSiNE ensemble" = "#0B6E4F", "Warm refit" = "#C65D00"),
      guide = "none"
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void() +
    theme(plot.margin = margin(0, 4, 0, 4))
}

paper_highlight_l2_legend_plot <- function() {
  ggplot() +
    geom_point(
      aes(x = 0.10, y = 0.64),
      shape = 21,
      size = 4.0,
      stroke = 1.1,
      fill = NA,
      color = "black"
    ) +
    geom_label(
      aes(x = 0.10, y = 0.28, label = "0.42"),
      size = 2.5,
      label.size = 0.18,
      fill = "white",
      color = "black"
    ) +
    geom_segment(
      aes(x = 0.10, xend = 0.10, y = 0.39, yend = 0.53),
      color = "black",
      linewidth = 0.25
    ) +
    annotate(
      "text",
      x = 0.25,
      y = 0.48,
      label = "circle: highest-weight setting\nnumber: aggregation weight",
      hjust = 0,
      size = 2.55,
      color = "grey15"
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void() +
    theme(plot.margin = margin(0, 4, 0, 4))
}

paper_sigma_l2_legend_plot <- function() {
  sigma_levels <- sort(unique(selected_l2_drift_tbl$sigma_0_2_scalar))
  sigma_labels <- format_grid_value(sigma_levels)
  sigma_cols <- scales::hue_pal()(length(sigma_levels))
  df <- tibble::tibble(
    sigma_label = sigma_labels,
    color = sigma_cols,
    col = rep(seq_len(4), length.out = length(sigma_levels)),
    row = rep(c(2, 1), each = 4, length.out = length(sigma_levels))
  ) %>%
    dplyr::mutate(
      x = 0.21 + (.data$col - 1) * 0.18,
      xend = .data$x + 0.045,
      text_x = .data$xend + 0.018,
      y = 0.28 + (.data$row - 1) * 0.35
    )
  ggplot(df) +
    geom_segment(aes(x = x, xend = xend, y = y, yend = y, color = sigma_label),
                 linewidth = 0.7) +
    geom_point(aes(x = xend, y = y, color = sigma_label), size = 2.1) +
    geom_text(aes(x = text_x, y = y, label = sigma_label),
              hjust = 0, size = 2.45, color = "grey15") +
    scale_color_manual(values = stats::setNames(df$color, df$sigma_label), guide = "none") +
    annotate(
      "text",
      x = 0.10,
      y = 0.50,
      label = expression(sigma[0]^2),
      hjust = 0.5,
      size = 3.35,
      color = "grey15"
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE, clip = "off") +
    theme_void() +
    theme(plot.margin = margin(0, 4, 0, 4))
}

if (requireNamespace("patchwork", quietly = TRUE)) {
  zoom_panel_letters <- LETTERS[seq_len(2L * length(selected_zoom_loci))]
  selected_locus_zoom_rows <- purrr::map2(selected_zoom_loci, seq_along(selected_zoom_loci), function(loc, i) {
    show_legend <- i == length(selected_zoom_loci)
    left_letter <- zoom_panel_letters[[2L * i - 1L]]
    right_letter <- zoom_panel_letters[[2L * i]]
    paper_lollipop_plot(loc, left_letter, show_legend = show_legend) |
      paper_pip_drift_plot(loc, right_letter, show_legend = show_legend)
  })

  selected_locus_zoom_figure <- patchwork::wrap_plots(
    selected_locus_zoom_rows,
    ncol = 1
  ) +
    patchwork::plot_layout(heights = rep(1, length(selected_locus_zoom_rows))) &
    theme(
      plot.margin = margin(4.5, 4.5, 4.5, 4.5),
      panel.spacing.x = grid::unit(3.2, "mm"),
      panel.spacing.y = grid::unit(5.5, "mm")
    )

  # Column titles are now set as plot.title on the row-1 plots themselves
  # rather than as a separate header strip above. plot.title anchors to the
  # plot panel (data region), so titles auto-center over their column
  # regardless of how much width each plot's y-axis labels consume. This
  # replaces the previous per-header `x =` nudges, which were fragile and
  # had to be re-tuned whenever axis labels or panel widths changed.
  selected_locus_zoom_l2_rows <- purrr::map2(selected_zoom_l2_loci, seq_along(selected_zoom_l2_loci), function(loc, i) {
    show_x_axis <- i == length(selected_zoom_l2_loci)
    is_first_row <- i == 1L
    title_lollipop <- if (is_first_row) "PIPs for shifted effects" else NULL
    title_effect   <- if (is_first_row) "Fitted signal contribution" else NULL
    title_drift    <- if (is_first_row) "PIP L2 drift from baseline" else NULL
    (
        paper_locus_row_label_plot(loc) |
        paper_lollipop_l2_plot(loc, show_x_axis = show_x_axis, show_legend = FALSE,
                               column_title = title_lollipop) |
        paper_signal_contribution_l2_plot(loc, show_x_axis = show_x_axis, show_legend = FALSE,
                                          column_title = title_effect) |
        paper_pip_drift_l2_plot(loc, show_x_axis = show_x_axis, show_legend = FALSE,
                                column_title = title_drift)
    ) +
      patchwork::plot_layout(widths = c(0.07, 1, 1, 1))
  })

  selected_locus_zoom_l2_legend <- (
    paper_lollipop_l2_legend_plot() |
      paper_highlight_l2_legend_plot() |
      paper_sigma_l2_legend_plot()
  ) +
    patchwork::plot_layout(widths = c(1, 0.95, 1.35))

  # First row is a touch taller than the others to absorb the plot.title
  # vertical space without compressing its panel area below the other rows.
  selected_locus_zoom_l2_heights <- c(
    0.98,
    rep(0.90, length(selected_locus_zoom_l2_rows) - 1L),
    0.42
  )
  selected_locus_zoom_l2_figure <- patchwork::wrap_plots(
    c(selected_locus_zoom_l2_rows, list(selected_locus_zoom_l2_legend)),
    ncol = 1
  ) +
    patchwork::plot_layout(heights = selected_locus_zoom_l2_heights) &
    theme(
      plot.margin = margin(4.5, 4.5, 2.5, 4.5),
      panel.spacing.x = grid::unit(3.2, "mm"),
      panel.spacing.y = grid::unit(2.0, "mm")
    )

  ggsave(
    file.path(overall_figure_dir, "paper_selected_locus_zoom_lollipop_drift.png"),
    selected_locus_zoom_figure,
    width = 7.52,
    height = 9.2,
    dpi = 300,
    bg = "white",
    limitsize = FALSE
  )
  ggsave(
    file.path(overall_figure_dir, "paper_selected_locus_zoom_lollipop_drift.pdf"),
    selected_locus_zoom_figure,
    width = 7.52,
    height = 9.2,
    bg = "white",
    limitsize = FALSE
  )
  ggsave(
    file.path(overall_figure_dir, "paper_selected_locus_zoom_lollipop_l2_drift.png"),
    selected_locus_zoom_l2_figure,
    width = 7.52,
    height = 8.4,
    dpi = 300,
    bg = "white",
    limitsize = FALSE
  )
  ggsave(
    file.path(overall_figure_dir, "paper_selected_locus_zoom_lollipop_l2_drift.pdf"),
    selected_locus_zoom_l2_figure,
    width = 7.52,
    height = 8.4,
    bg = "white",
    limitsize = FALSE
  )

  list(
    jsd = selected_locus_zoom_figure,
    l2 = selected_locus_zoom_l2_figure
  )
} else {
  warning("patchwork is not installed; selected-locus zoom figure skipped.")
  invisible(NULL)
}
```

```{r free-locus-R-cache}
# The lollipop helpers cache per-locus R matrices (each ~ p^2 doubles, around
# 0.3-0.5 GB at p = 6-7k) in selected_locus_R_cache. Nothing downstream needs
# them, so evict before section 9 reloads its own bundles per locus.
rm(list = ls(envir = selected_locus_R_cache), envir = selected_locus_R_cache)
invisible(gc(full = TRUE))
```

# 9. Basin r² Drift From c = 0 (one panel per locus)

Drift metric: weighted (1 - r²) on max-weight 1-1 matched per-effect coefficient
vectors, weighted by `min(var_y_explained_a, var_y_explained_b)`. RSS analog of
`weighted_r2_drift` from `effect_matching_drift_analysis_5arm.Rmd`. Uses the
R-weighted cosine `(b_i' R b_j) / sqrt((b_i' R b_i)(b_j' R b_j))` since X is
not available in the SS regime.

```{r basin-r2-drift-compute, message=TRUE}
basin_r2_drift_tbl <- ensure_basin_r2_drift_tbl()
basin_r2_drift_tbl
```

```{r basin-r2-drift-plot, fig.width=10, fig.height=6}
plot_basin_r2_drift <- function(loc) {
  df <- dplyr::filter(basin_r2_drift_tbl, .data$locus_id == loc)
  flag_row <- dataset_metrics_for_flag %>% dplyr::filter(.data$locus_id == loc)
  flagged <- isTRUE(nrow(flag_row) == 1L &&
                    isTRUE(flag_row$ld_z_inconsistency_flag[[1]]))
  ggplot(df, aes(x = c_value, y = basin_r2_drift,
                 color = factor(sigma_0_2_scalar))) +
    geom_line() +
    geom_point(size = 2) +
    labs(
      title = paste0("Basin r² Drift From c = 0: ", loc,
                     if (flagged) "  [LD/z inconsistent]" else ""),
      subtitle = "Weighted (1 - r²) on matched per-effect b_l, R-weighted cosine",
      x = "Annotation scale c",
      y = "Weighted (1 - r²) drift",
      color = expression(sigma[0]^2)
    )
}

# Optional: pull the inconsistency flag from dataset_metrics if present so we
# can annotate flagged loci in the title.
dataset_metrics_path <- file.path(aggregated_dir, "dataset_metrics_full.csv")
dataset_metrics_for_flag <- if (file.exists(dataset_metrics_path)) {
  readr::read_csv(dataset_metrics_path, show_col_types = FALSE)
} else {
  tibble::tibble(locus_id = character(), ld_z_inconsistency_flag = logical())
}

purrr::walk(sort(unique(basin_r2_drift_tbl$locus_id)), function(loc) {
  p <- plot_basin_r2_drift(loc)
  ggsave(
    file.path(locus_figure_dir, paste0(loc, "_basin_r2_drift.png")),
    p, width = 9, height = 5, dpi = 150
  )
  print(p)
})
```

# 10. Floored-Conditional Edge-Case Report

```{r floored-conditional-edge-case-report, message=TRUE}
# Reports any locus/fit/variant edge case that left a floored-conditional
# entry as NA. Kinds are documented above the floored_conditional_log helper.
# Prints "No edge cases encountered." when everything went cleanly.
report_floored_conditional_diagnostics()
```
