---
title: "Analyze Local Nonlinear Difficulty Models"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

One-off workbook to search baseline-only nonlinear difficulty models
using the local `ensemble_full_prelim` files. Each additive candidate
model includes:

1. `M1`
2. one thresholded LD alpha metric
3. any two z metrics

After selecting the best additive four-variable model, the workbook runs
an AIC-based search over all subsets of two-way interactions among those
four variables.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/analyze_local_nonlinear_difficulty_models.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(knitr)
library(purrr)
library(mgcv)

theme_set(theme_bw(base_size = 12))
```

## 1. Inputs

```{r inputs}
input_dir <- here("output", "slurm_output", "ensemble_full_prelim")

difficulty_data_path <- file.path(input_dir, "difficulty_model_data.csv")
dataset_bundles_path <- file.path(input_dir, "dataset_bundles.csv")
ld_metrics_path <- here("data", "sampled_simulated_genotypes", "scenario_1_ld_alpha_metrics.csv")

fig_dir <- file.path(input_dir, "figures", "dataset_difficulty_local")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

cat("Input dir:", input_dir, "\n")
cat("Difficulty data:", difficulty_data_path, "\n")
cat("Dataset bundles:", dataset_bundles_path, "\n")
cat("LD metrics:", ld_metrics_path, "\n")
cat("Figure dir:", fig_dir, "\n")
```

## 2. Load and Join Data

```{r load}
stopifnot(dir.exists(input_dir))
stopifnot(file.exists(difficulty_data_path))
stopifnot(file.exists(dataset_bundles_path))
stopifnot(file.exists(ld_metrics_path))

difficulty_df <- readr::read_csv(difficulty_data_path, show_col_types = FALSE) %>%
  dplyr::mutate(dataset_bundle_id = as.integer(.data$dataset_bundle_id))

dataset_bundles <- readr::read_csv(dataset_bundles_path, show_col_types = FALSE) %>%
  dplyr::mutate(
    dataset_bundle_id = as.integer(.data$dataset_bundle_id),
    matrix_id = as.integer(.data$matrix_id)
  ) %>%
  dplyr::distinct(.data$dataset_bundle_id, .keep_all = TRUE)

ld_metrics_df <- readr::read_csv(ld_metrics_path, show_col_types = FALSE) %>%
  dplyr::mutate(matrix_id = as.integer(.data$matrix_id))

difficulty_model_data <- difficulty_df %>%
  dplyr::left_join(
    dataset_bundles %>% dplyr::select("dataset_bundle_id", "matrix_id"),
    by = "dataset_bundle_id"
  ) %>%
  dplyr::left_join(
    ld_metrics_df %>%
      dplyr::select(
        "matrix_id",
        "ld_abs_alpha_2_thr_0p5",
        "ld_abs_alpha_4_thr_0p5"
      ),
    by = "matrix_id"
  ) %>%
  dplyr::arrange(.data$dataset_bundle_id)

required_cols <- c(
  "dataset_bundle_id",
  "matrix_id",
  "M1",
  "ld_abs_alpha_2_thr_0p5",
  "ld_abs_alpha_4_thr_0p5",
  "z_topk_ratio",
  "z_max_abs",
  "z_count_abs_gt_3",
  "z_eff_signals",
  "auprc_baseline"
)

missing_required <- setdiff(required_cols, names(difficulty_model_data))
if (length(missing_required)) {
  stop(
    "Input data are missing required columns: ",
    paste(missing_required, collapse = ", ")
  )
}

cat("Rows:", nrow(difficulty_model_data), "\n")
cat("Columns:", ncol(difficulty_model_data), "\n")
```

## 3. Helpers

```{r helpers}
html_escape <- function(x) {
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x
}

format_html_table <- function(df, digits = 4L) {
  if (is.null(df) || !nrow(df)) {
    return("<p><em>No rows.</em></p>")
  }
  df_fmt <- df
  is_num <- vapply(df_fmt, is.numeric, logical(1))
  df_fmt[is_num] <- lapply(df_fmt[is_num], function(col) round(col, digits))
  header <- paste0(
    "<tr>",
    paste(sprintf("<th>%s</th>", html_escape(names(df_fmt))), collapse = ""),
    "</tr>"
  )
  rows <- apply(as.data.frame(df_fmt, stringsAsFactors = FALSE), 1, function(row) {
    paste0(
      "<tr>",
      paste(sprintf("<td>%s</td>", html_escape(as.character(row))), collapse = ""),
      "</tr>"
    )
  })
  paste0(
    "<table class='report-table'>",
    "<thead>", header, "</thead>",
    "<tbody>", paste(rows, collapse = ""), "</tbody>",
    "</table>"
  )
}

write_html_report <- function(path, title, sections) {
  html_sections <- vapply(sections, function(section) {
    header_html <- if (!is.null(section$header) && nzchar(section$header)) {
      sprintf("<h2>%s</h2>", html_escape(section$header))
    } else {
      ""
    }
    text_html <- if (!is.null(section$text) && length(section$text)) {
      paste(sprintf("<p>%s</p>", html_escape(section$text)), collapse = "")
    } else {
      ""
    }
    digits <- if (!is.null(section$digits)) section$digits else 4L
    table_html <- if (!is.null(section$table)) {
      format_html_table(section$table, digits = digits)
    } else {
      ""
    }
    paste0(header_html, text_html, table_html)
  }, character(1))

  html <- paste0(
    "<!DOCTYPE html><html><head><meta charset='utf-8'>",
    "<title>", html_escape(title), "</title>",
    "<style>",
    "body{font-family:Arial,sans-serif;margin:24px;line-height:1.4;}",
    "h1,h2{margin-bottom:0.4em;}",
    ".report-table{border-collapse:collapse;width:100%;margin:12px 0 24px 0;}",
    ".report-table th,.report-table td{border:1px solid #cccccc;padding:6px 8px;text-align:left;}",
    ".report-table th{background:#f3f3f3;}",
    "code{background:#f5f5f5;padding:2px 4px;}",
    "</style></head><body>",
    "<h1>", html_escape(title), "</h1>",
    paste(html_sections, collapse = ""),
    "</body></html>"
  )

  writeLines(html, con = path, useBytes = TRUE)
}

enumerate_metric_pairs <- function(metrics) {
  pairs <- utils::combn(metrics, 2, simplify = FALSE)
  tibble::tibble(
    metric1 = vapply(pairs, `[`, character(1), 1),
    metric2 = vapply(pairs, `[`, character(1), 2)
  )
}

choose_k <- function(x, max_k = 5L) {
  n_unique <- length(unique(x[is.finite(x)]))
  max(3L, min(max_k, n_unique - 1L))
}

as_term_table <- function(mat, first_col_name) {
  if (is.null(mat) || !length(mat)) {
    return(tibble::tibble())
  }
  out <- tibble::as_tibble(mat, rownames = first_col_name)
  numeric_cols <- vapply(out, is.numeric, logical(1))
  out[numeric_cols] <- lapply(out[numeric_cols], as.numeric)
  out
}

fit_gam_formula <- function(data, outcome_col, predictors, formula_txt, meta = list()) {
  vars <- c(outcome_col, predictors)
  model_df <- data %>%
    dplyr::select(dplyr::all_of(c("dataset_bundle_id", "matrix_id", vars))) %>%
    dplyr::filter(dplyr::if_all(dplyr::all_of(vars), ~ is.finite(.x)))

  n_used <- nrow(model_df)
  base_summary <- tibble::tibble(
    outcome = outcome_col,
    formula = formula_txt,
    n_used = n_used,
    r_squared = NA_real_,
    approx_adj_r_squared = NA_real_,
    deviance_explained = NA_real_,
    rmse = NA_real_,
    aic = NA_real_,
    status = "insufficient_rows"
  )
  for (nm in names(meta)) {
    base_summary[[nm]] <- meta[[nm]]
  }

  if (n_used < 20L) {
    return(list(
      summary = base_summary,
      parametric = tibble::tibble(),
      smooth = tibble::tibble(),
      predictions = model_df
    ))
  }

  fit <- tryCatch(
    mgcv::gam(stats::as.formula(formula_txt), data = model_df, method = "REML"),
    error = function(e) e
  )

  if (inherits(fit, "error")) {
    base_summary$status <- paste0("fit_error: ", fit$message)
    return(list(
      summary = base_summary,
      parametric = tibble::tibble(),
      smooth = tibble::tibble(),
      predictions = model_df
    ))
  }

  sm <- summary(fit)
  pred <- stats::predict(fit, newdata = model_df)
  resid <- model_df[[outcome_col]] - pred
  edf_total <- sum(sm$edf, na.rm = TRUE)
  approx_adj_r2 <- if (is.finite(sm$r.sq) && n_used > (edf_total + 1)) {
    1 - (1 - sm$r.sq) * ((n_used - 1) / (n_used - edf_total - 1))
  } else {
    NA_real_
  }

  summary_tbl <- tibble::tibble(
    outcome = outcome_col,
    formula = formula_txt,
    n_used = n_used,
    r_squared = unname(sm$r.sq),
    approx_adj_r_squared = approx_adj_r2,
    deviance_explained = unname(sm$dev.expl),
    rmse = sqrt(mean(resid^2)),
    aic = stats::AIC(fit),
    status = "ok"
  )
  for (nm in names(meta)) {
    summary_tbl[[nm]] <- meta[[nm]]
  }

  list(
    summary = summary_tbl,
    parametric = as_term_table(sm$p.table, "term"),
    smooth = as_term_table(sm$s.table, "smooth_term"),
    predictions = model_df %>%
      dplyr::mutate(
        predicted = as.numeric(pred),
        residual = as.numeric(resid)
      )
  )
}

fit_additive_gam_model <- function(data, outcome_col, ld_metric, z_metric1, z_metric2) {
  predictors <- c("M1", ld_metric, z_metric1, z_metric2)
  k_vals <- vapply(predictors, function(var) choose_k(data[[var]]), integer(1))
  formula_txt <- paste0(
    outcome_col, " ~ ",
    paste(
      sprintf("s(%s, k = %d)", predictors, k_vals),
      collapse = " + "
    )
  )

  fit_gam_formula(
    data = data,
    outcome_col = outcome_col,
    predictors = predictors,
    formula_txt = formula_txt,
    meta = list(
      ld_metric = ld_metric,
      z_metric1 = z_metric1,
      z_metric2 = z_metric2
    )
  )
}

enumerate_refinement_models <- function(fixed_var, optional_vars) {
  optional_subsets <- list(character(0))
  for (m in seq_along(optional_vars)) {
    optional_subsets <- c(optional_subsets, utils::combn(optional_vars, m, simplify = FALSE))
  }

  rows <- vector("list", length = 0L)
  row_id <- 0L

  for (kept_optional in optional_subsets) {
    included_vars <- c(fixed_var, kept_optional)
    dropped_vars <- setdiff(optional_vars, kept_optional)
    pair_list <- if (length(included_vars) >= 2L) utils::combn(included_vars, 2, simplify = FALSE) else list()
    pair_labels <- vapply(pair_list, function(x) paste(x, collapse = ":"), character(1))
    pair_idx <- seq_along(pair_list)
    interaction_subsets <- list(integer(0))
    for (m in seq_along(pair_idx)) {
      interaction_subsets <- c(interaction_subsets, utils::combn(pair_idx, m, simplify = FALSE))
    }

    for (idx in interaction_subsets) {
      row_id <- row_id + 1L
      rows[[row_id]] <- tibble::tibble(
        refinement_id = row_id,
        included_vars = list(included_vars),
        dropped_vars = list(dropped_vars),
        kept_optional_label = if (length(kept_optional)) paste(kept_optional, collapse = ", ") else "none",
        dropped_label = if (length(dropped_vars)) paste(dropped_vars, collapse = ", ") else "none",
        interaction_terms = list(pair_list[idx]),
        interaction_label = if (length(idx)) paste(pair_labels[idx], collapse = " + ") else "additive_only",
        n_optional_kept = length(kept_optional),
        n_interactions = length(idx)
      )
    }
  }

  dplyr::bind_rows(rows)
}

fit_refined_gam_model <- function(data, outcome_col, fixed_var, optional_vars, interaction_terms, interaction_label) {
  vars <- c(fixed_var, optional_vars)
  k_vals <- stats::setNames(vapply(vars, function(var) choose_k(data[[var]]), integer(1)), vars)
  main_terms <- sprintf("s(%s, k = %d)", vars, unname(k_vals))
  interaction_txt <- if (!length(interaction_terms)) {
    character(0)
  } else {
    vapply(interaction_terms, function(pair) {
      sprintf(
        "ti(%s, %s, k = c(%d, %d))",
        pair[[1]], pair[[2]], k_vals[[pair[[1]]]], k_vals[[pair[[2]]]]
      )
    }, character(1))
  }
  formula_txt <- paste(outcome_col, "~", paste(c(main_terms, interaction_txt), collapse = " + "))

  fit_gam_formula(
    data = data,
    outcome_col = outcome_col,
    predictors = vars,
    formula_txt = formula_txt,
    meta = list(
      ld_metric = dplyr::first(optional_vars[grepl("^ld_", optional_vars)], default = NA_character_),
      z_metric1 = dplyr::nth(optional_vars[grepl("^z_", optional_vars)], 1L, default = NA_character_),
      z_metric2 = dplyr::nth(optional_vars[grepl("^z_", optional_vars)], 2L, default = NA_character_),
      kept_optional_label = if (length(optional_vars)) paste(optional_vars, collapse = ", ") else "none",
      dropped_label = NA_character_,
      interaction_label = interaction_label,
      n_optional_kept = length(optional_vars),
      n_interactions = length(interaction_terms)
    )
  )
}
```

## 4. Additive Search

```{r model-search}
candidate_ld_metrics <- c("ld_abs_alpha_2_thr_0p5", "ld_abs_alpha_4_thr_0p5")
candidate_ld_metrics <- candidate_ld_metrics[candidate_ld_metrics %in% names(difficulty_model_data)]

candidate_z_metrics <- c(
  "z_topk_ratio",
  "z_max_abs",
  "z_count_abs_gt_3",
  "z_eff_signals"
)
candidate_z_metrics <- candidate_z_metrics[candidate_z_metrics %in% names(difficulty_model_data)]
z_pairs <- enumerate_metric_pairs(candidate_z_metrics)

if (!length(candidate_ld_metrics) || nrow(z_pairs) == 0L) {
  stop("Need at least one candidate LD metric and one z-metric pair.")
}

search_grid <- tidyr::expand_grid(
  ld_metric = candidate_ld_metrics,
  z_metric1 = z_pairs$metric1,
  z_metric2 = z_pairs$metric2
)

additive_fits <- purrr::pmap(
  search_grid,
  function(ld_metric, z_metric1, z_metric2) {
    fit_additive_gam_model(
      difficulty_model_data,
      outcome_col = "auprc_baseline",
      ld_metric = ld_metric,
      z_metric1 = z_metric1,
      z_metric2 = z_metric2
    )
  }
)

additive_search <- dplyr::bind_rows(lapply(additive_fits, `[[`, "summary")) %>%
  dplyr::arrange(
    dplyr::desc(.data$approx_adj_r_squared),
    dplyr::desc(.data$r_squared),
    dplyr::desc(.data$deviance_explained),
    .data$aic,
    .data$ld_metric,
    .data$z_metric1,
    .data$z_metric2
  )

best_additive_row <- additive_search %>%
  dplyr::filter(.data$status == "ok") %>%
  dplyr::slice_head(n = 1)

if (!nrow(best_additive_row)) {
  stop("No successful additive baseline model fits.")
}

selected_vars <- c(
  "M1",
  best_additive_row$ld_metric[[1]],
  best_additive_row$z_metric1[[1]],
  best_additive_row$z_metric2[[1]]
)

fixed_var <- "M1"
optional_vars <- setdiff(selected_vars, fixed_var)

refinement_grid <- enumerate_refinement_models(
  fixed_var = fixed_var,
  optional_vars = optional_vars
)

refinement_fits <- lapply(seq_len(nrow(refinement_grid)), function(i) {
  fit <- fit_refined_gam_model(
    difficulty_model_data,
    outcome_col = "auprc_baseline",
    fixed_var = fixed_var,
    optional_vars = refinement_grid$included_vars[[i]][refinement_grid$included_vars[[i]] != fixed_var],
    interaction_terms = refinement_grid$interaction_terms[[i]],
    interaction_label = refinement_grid$interaction_label[[i]]
  )
  fit$summary$dropped_label <- refinement_grid$dropped_label[[i]]
  fit
})

refinement_search <- dplyr::bind_rows(lapply(refinement_fits, `[[`, "summary")) %>%
  dplyr::arrange(
    .data$aic,
    dplyr::desc(.data$approx_adj_r_squared),
    dplyr::desc(.data$r_squared),
    dplyr::desc(.data$deviance_explained),
    .data$n_optional_kept,
    .data$interaction_label
  )

best_refinement_row <- refinement_search %>%
  dplyr::filter(.data$status == "ok") %>%
  dplyr::slice_head(n = 1)

if (!nrow(best_refinement_row)) {
  stop("No successful refinement baseline model fits.")
}

kable(best_additive_row)
kable(head(refinement_search, 10))
```

## 5. Save Reports and Prediction Tables

```{r reports}
additive_fit_idx <- which(
  search_grid$ld_metric == best_additive_row$ld_metric[[1]] &
    search_grid$z_metric1 == best_additive_row$z_metric1[[1]] &
    search_grid$z_metric2 == best_additive_row$z_metric2[[1]]
)[[1]]
best_additive_fit <- additive_fits[[additive_fit_idx]]

interaction_fit_idx <- which(
  refinement_search$interaction_label == best_refinement_row$interaction_label[[1]] &
    refinement_search$formula == best_refinement_row$formula[[1]] &
    refinement_search$kept_optional_label == best_refinement_row$kept_optional_label[[1]]
)[[1]]
best_refinement_fit <- refinement_fits[[interaction_fit_idx]]

search_html <- file.path(fig_dir, "difficulty_model_search_baseline_local.html")
best_fit_html <- file.path(fig_dir, "difficulty_model_refit_baseline_local.html")
additive_search_csv <- file.path(fig_dir, "difficulty_model_search_baseline_additive.csv")
refinement_search_csv <- file.path(fig_dir, "difficulty_model_search_baseline_refinement.csv")
summary_csv <- file.path(fig_dir, "difficulty_model_best_fit_baseline_summary.csv")
param_csv <- file.path(fig_dir, "difficulty_model_best_fit_baseline_parametric.csv")
smooth_csv <- file.path(fig_dir, "difficulty_model_best_fit_baseline_smooth.csv")
data_csv <- file.path(fig_dir, "difficulty_model_data_baseline_local.csv")

prediction_tbl <- best_refinement_fit$predictions %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    predicted_baseline_local = as.numeric(.data$predicted),
    residual_baseline_local = as.numeric(.data$residual)
  )

baseline_model_data <- difficulty_model_data %>%
  dplyr::left_join(prediction_tbl, by = "dataset_bundle_id")

write_html_report(
  search_html,
  title = "Baseline AUPRC difficulty model search",
  sections = list(
    list(
      header = "Additive Search Summary",
      text = c(
        "Outcome: auprc_baseline",
        paste0("Candidate LD metrics: ", paste(candidate_ld_metrics, collapse = ", ")),
        paste0("Candidate z metrics: ", paste(candidate_z_metrics, collapse = ", ")),
        paste0("Z-metric pairs searched: ", nrow(z_pairs)),
        paste0("Selected additive LD metric: ", best_additive_row$ld_metric[[1]]),
        paste0("Selected additive z metrics: ", best_additive_row$z_metric1[[1]], " and ", best_additive_row$z_metric2[[1]]),
        paste0("Selected additive formula: ", best_additive_row$formula[[1]])
      ),
      table = best_additive_row
    ),
    list(
      header = "Refinement Search Summary",
      text = c(
        paste0("Starting variables after additive search: ", paste(selected_vars, collapse = ", ")),
        paste0("Refinement models searched with M1 fixed: ", nrow(refinement_grid)),
        paste0("Retained optional variables: ", best_refinement_row$kept_optional_label[[1]]),
        paste0("Dropped optional variables: ", best_refinement_row$dropped_label[[1]]),
        paste0("Selected interaction set: ", best_refinement_row$interaction_label[[1]]),
        paste0("Selected final formula: ", best_refinement_row$formula[[1]])
      ),
      table = best_refinement_row
    ),
    list(
      header = "All Additive Candidate Models",
      table = additive_search
    ),
    list(
      header = "All Refinement Candidate Models",
      table = refinement_search
    )
  )
)

write_html_report(
  best_fit_html,
  title = "Baseline AUPRC difficulty model refit",
  sections = list(
    list(
      header = "Final Model Summary",
      text = c(
        paste0("Selected LD metric: ", best_additive_row$ld_metric[[1]]),
        paste0("Selected z metrics: ", best_additive_row$z_metric1[[1]], " and ", best_additive_row$z_metric2[[1]]),
        paste0("Retained optional variables: ", best_refinement_row$kept_optional_label[[1]]),
        paste0("Dropped optional variables: ", best_refinement_row$dropped_label[[1]]),
        paste0("Selected interaction set: ", best_refinement_row$interaction_label[[1]]),
        paste0("Formula: ", best_refinement_row$formula[[1]])
      ),
      table = best_refinement_fit$summary
    ),
    list(
      header = "Parametric Terms",
      table = best_refinement_fit$parametric
    ),
    list(
      header = "Smooth Terms",
      table = best_refinement_fit$smooth
    )
  )
)

readr::write_csv(additive_search, additive_search_csv)
readr::write_csv(refinement_search, refinement_search_csv)
readr::write_csv(best_refinement_fit$summary, summary_csv)
readr::write_csv(best_refinement_fit$parametric, param_csv)
readr::write_csv(best_refinement_fit$smooth, smooth_csv)
readr::write_csv(baseline_model_data, data_csv)

cat("Saved search HTML:", search_html, "\n")
cat("Saved refit HTML:", best_fit_html, "\n")
cat("Saved additive search CSV:", additive_search_csv, "\n")
cat("Saved refinement search CSV:", refinement_search_csv, "\n")
cat("Saved summary CSV:", summary_csv, "\n")
cat("Saved parametric CSV:", param_csv, "\n")
cat("Saved smooth CSV:", smooth_csv, "\n")
cat("Saved data CSV:", data_csv, "\n")
```

## 6. Save Prediction Plots

```{r prediction-plots}
plot_df <- best_refinement_fit$predictions %>%
  dplyr::filter(is.finite(.data$predicted), is.finite(.data$auprc_baseline))

p <- ggplot(
  plot_df,
  aes(x = .data$predicted, y = .data$auprc_baseline)
) +
  geom_point(alpha = 0.65, size = 2.0, color = "steelblue4") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray40") +
  labs(
    x = "Predicted baseline AUPRC",
    y = "Observed baseline AUPRC",
    title = "Prediction plot: Baseline AUPRC",
    subtitle = paste0(
      "kept: ", best_refinement_row$kept_optional_label[[1]],
      " | dropped: ", best_refinement_row$dropped_label[[1]],
      " | ", best_refinement_row$interaction_label[[1]]
    )
  )

plot_path <- file.path(fig_dir, "difficulty_model_predictions_baseline_local.png")
ggsave(plot_path, p, width = 8, height = 6)
cat("Saved plot:", plot_path, "\n")

invisible(NULL)
```
