---
title: "Baseline Vanilla Effect Purity Diagnostics"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

One-off validation workbook for baseline `susine_vanilla` effect rows.

This workbook:

1. Loads the raw `run_table.csv` from run history.
2. Identifies baseline vanilla runs.
3. Reads raw `_effect_metrics.csv` and `_model_metrics.csv` flush files from
   the task output tree.
4. Keeps `explore_method == "base"` and the selected baseline run IDs.
5. Joins run-level power from `model_metrics`.
6. Saves diagnostic plots and the filtered analysis table.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/baseline_vanilla_effect_purity_diagnostics.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
library(dplyr)
library(readr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

## 1. Inputs

```{r inputs}
job_name <- "ensemble_scaling_full"
job_id <- "50339226"

output_root <- here("output")
run_history_dir <- file.path(output_root, "run_history", job_name, job_id)
job_dir <- file.path(output_root, "slurm_output", job_name, job_id)
fig_dir <- file.path(job_dir, "figures", "one_off_validations")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

run_table_path <- file.path(run_history_dir, "run_table.csv")

effect_out_csv <- file.path(fig_dir, "baseline_vanilla_effect_metrics_joined.csv")
plot_power_purity <- file.path(fig_dir, "baseline_vanilla_power_vs_purity.png")
plot_size_purity <- file.path(fig_dir, "baseline_vanilla_size_vs_purity.png")
plot_size_power <- file.path(fig_dir, "baseline_vanilla_size_vs_power_colored_by_purity.png")

cat("Run history dir:", run_history_dir, "\n")
cat("Job dir:", job_dir, "\n")
cat("Figure dir:", fig_dir, "\n")
cat("Job name:", job_name, "\n")
cat("Job id:", job_id, "\n")
cat("Run table:", run_table_path, "\n")
```

## 2. Run Selection

```{r run-selection}
stopifnot(dir.exists(job_dir))
stopifnot(file.exists(run_table_path))

run_table <- readr::read_csv(run_table_path, show_col_types = FALSE)

baseline_run_ids <- run_table %>%
  dplyr::filter(
    .data$spec_name == "baseline-single",
    .data$use_case_id == "susine_vanilla"
  ) %>%
  dplyr::distinct(.data$run_id) %>%
  dplyr::pull(.data$run_id)

cat("Selected baseline vanilla runs:", length(baseline_run_ids), "\n")
```

## 3. Load Flush Files

```{r load-flush}
effect_files <- list.files(
  job_dir,
  pattern = "_effect_metrics\\.csv$",
  recursive = TRUE,
  full.names = TRUE
)

model_files <- list.files(
  job_dir,
  pattern = "_model_metrics\\.csv$",
  recursive = TRUE,
  full.names = TRUE
)

cat("Effect flush files found:", length(effect_files), "\n")
cat("Model flush files found:", length(model_files), "\n")

if (!length(effect_files)) {
  stop("No effect_metrics flush files found under: ", job_dir)
}
if (!length(model_files)) {
  stop("No model_metrics flush files found under: ", job_dir)
}

read_flush_csvs <- function(paths, col_select = NULL) {
  tbls <- lapply(paths, function(path) {
    readr::read_csv(path, show_col_types = FALSE, col_select = dplyr::any_of(col_select))
  })
  dplyr::bind_rows(tbls)
}

effect_metrics <- read_flush_csvs(
  effect_files,
  col_select = c(
    "effect", "size", "purity", "coverage", "run_id",
    "explore_method", "variant_id", "agg_method"
  )
)

model_metrics <- read_flush_csvs(
  model_files,
  col_select = c(
    "run_id", "power", "filtering", "explore_method",
    "variant_id", "agg_method"
  )
)

cat("Effect rows loaded:", nrow(effect_metrics), "\n")
cat("Model rows loaded:", nrow(model_metrics), "\n")
```

## 4. Build Analysis Table

```{r analysis-table}
baseline_effects <- effect_metrics %>%
  dplyr::filter(
    .data$explore_method == "base",
    .data$run_id %in% baseline_run_ids
  ) %>%
  dplyr::mutate(
    effect = as.integer(.data$effect),
    size = as.numeric(.data$size),
    purity = as.numeric(.data$purity),
    coverage = as.numeric(.data$coverage)
  )

baseline_power <- model_metrics %>%
  dplyr::filter(
    .data$explore_method == "base",
    .data$filtering == "purity_filtered",
    .data$run_id %in% baseline_run_ids
  ) %>%
  dplyr::mutate(power = as.numeric(.data$power)) %>%
  dplyr::distinct(.data$run_id, .keep_all = TRUE) %>%
  dplyr::select(.data$run_id, .data$power)

analysis_df <- baseline_effects %>%
  dplyr::left_join(baseline_power, by = "run_id") %>%
  dplyr::arrange(.data$run_id, .data$effect)

readr::write_csv(analysis_df, effect_out_csv)

cat("Analysis rows:", nrow(analysis_df), "\n")
cat("Unique runs represented:", dplyr::n_distinct(analysis_df$run_id), "\n")
cat("Saved analysis CSV:", effect_out_csv, "\n")

kable(head(analysis_df, 20))
```

## 5. Save Diagnostic Plots

```{r plots}
plot_df <- analysis_df %>%
  dplyr::filter(
    is.finite(.data$size),
    is.finite(.data$purity),
    is.finite(.data$power)
  )

if (!nrow(plot_df)) {
  stop("No complete rows available for plotting.")
}

p_power_purity <- ggplot(
  plot_df,
  aes(x = .data$purity, y = .data$power)
) +
  geom_point(alpha = 0.6, size = 1.8, color = "steelblue4",
             position = position_jitter(width = 0.005, height = 0.01)) +
  geom_smooth(method = "lm", se = FALSE, color = "firebrick", linewidth = 0.8) +
  labs(
    x = "Effect purity",
    y = "Run-level power",
    title = "Baseline vanilla: power vs purity"
  )

p_size_purity <- ggplot(
  plot_df,
  aes(x = .data$purity, y = .data$size)
) +
  geom_point(alpha = 0.65, size = 1.8, color = "darkgreen") +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.8) +
  labs(
    x = "Effect purity",
    y = "Effect size",
    title = "Baseline vanilla: size vs purity"
  )

p_size_power <- ggplot(
  plot_df,
  aes(x = .data$power, y = .data$size, color = .data$purity)
) +
  geom_point(alpha = 0.7, size = 2,
             position = position_jitter(width = 0.01, height = 0)) +
  scale_color_viridis_c(option = "C") +
  labs(
    x = "Run-level power",
    y = "Effect size",
    color = "Purity",
    title = "Baseline vanilla: size vs power, colored by purity"
  )

ggsave(plot_power_purity, p_power_purity, width = 8, height = 6)
ggsave(plot_size_purity, p_size_purity, width = 8, height = 6)
ggsave(plot_size_power, p_size_power, width = 8, height = 6)

cat("Saved plot:", plot_power_purity, "\n")
cat("Saved plot:", plot_size_purity, "\n")
cat("Saved plot:", plot_size_power, "\n")
invisible(NULL)
```
