---
title: "Causal Basis Alignment vs AUPRC Decomposition (5-arm)"
output:
  html_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: 72
---

This workbook is a sibling to
`effect_matching_drift_paper_figure_5arm.Rmd`. It replaces one-to-one
Hungarian truth-basis drift with a truth-audited causal basis alignment score:
for each top-K causal explanatory signal, find the fitted model effect that
best covers it, then average the coverage over causal signals.

The goal is to separate two mean-level channels:

1. explanatory basis alignment: whether the model contains fitted effects that
   cover each top-K causal explanatory signal;
2. residual AUPRC movement: the AUPRC delta not predicted by the baseline
   alignment curve. This is a proxy for within-basis or SNP-ranking confusion,
   not a causal variance decomposition.

The workbook still plots explanatory basis size as context, but the main
decomposition uses causal basis alignment only because cross-validation showed
that size adds little reliable baseline AUPRC signal.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/causal_basis_alignment_auprc_decomposition_5arm.Rmd")

library(here)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(scales)
library(tibble)
```

# Inputs

```{r paths}
output_root <- here("output")
study_name <- "baseline_sims_5arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
plot_dir <- file.path(study_dir, "drift_analysis_plots")
alignment_plot_dir <- file.path(plot_dir, "causal_basis_alignment")
dir.create(alignment_plot_dir, recursive = TRUE, showWarnings = FALSE)

records_rds <- file.path(study_dir, "effect_refit_records.rds")
auprc_per_dataset_csv <- file.path(study_dir, "auprc_per_dataset_5arm.csv")
dataset_bundles_csv <- file.path(study_dir, "dataset_bundles.csv")

truth_top_k <- 23L
truth_fitted_y_rds <- file.path(
  study_dir, sprintf("causal_basis_truth_fitted_y_5arm_top%d.rds", truth_top_k)
)
alignment_scores_csv <- file.path(
  study_dir, sprintf("causal_basis_alignment_scores_5arm_top%d.csv", truth_top_k)
)
decomp_rows_csv <- file.path(
  study_dir, sprintf("causal_basis_alignment_auprc_decomposition_rows_top%d.csv",
                     truth_top_k)
)
decomp_summary_csv <- file.path(
  study_dir,
  sprintf("causal_basis_alignment_auprc_decomposition_summary_top%d.csv",
          truth_top_k)
)

baseline_alignment_png <- file.path(
  alignment_plot_dir, "causal_basis_alignment_baseline_only.png"
)
decomposition_png <- file.path(
  alignment_plot_dir, "causal_basis_alignment_auprc_basis_decomposition.png"
)
delta_alignment_mupi_png <- file.path(
  alignment_plot_dir, "causal_basis_alignment_delta_mu_vs_pi.png"
)

required_files <- c(records_rds, auprc_per_dataset_csv, dataset_bundles_csv)
missing_files <- required_files[!file.exists(required_files)]
if (length(missing_files)) {
  stop("Missing required inputs:\n", paste(missing_files, collapse = "\n"))
}

cat("Study dir:", study_dir, "\n")
cat("Alignment plot dir:", alignment_plot_dir, "\n")
```

# Options

```{r options}
baseline_method <- "susie_vanilla"
mu_method <- "susine_functional_mu"
pi_method <- "susie_functional_pi"
warm_mu_method <- "susie_vanilla_warm_from_mu"
warm_pi_method <- "susie_vanilla_warm_from_pi"

methods_in_order <- c(
  baseline_method, mu_method, warm_mu_method, pi_method, warm_pi_method
)
method_labels <- c(
  susie_vanilla = "Cold vanilla",
  susine_functional_mu = "Annotated mu",
  susie_vanilla_warm_from_mu = "Warm vanilla from mu",
  susie_functional_pi = "Annotated pi",
  susie_vanilla_warm_from_pi = "Warm vanilla from pi"
)
chain_colors <- c("mu chain" = "#2b8cbe", "pi chain" = "#74c476")

# This is intentionally smooth. We want a baseline mean curve, not an
# individual-dataset predictor.
baseline_curve_spline_df <- 5
```

# Helpers

```{r helpers}
theme_alignment <- function(base_size = 9) {
  theme_bw(base_size = base_size) +
    theme(
      panel.grid.major = element_line(color = "grey91", linewidth = 0.22),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "grey35", fill = NA, linewidth = 0.35),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size - 1, color = "grey25"),
      plot.title = element_text(face = "bold", size = base_size + 1.1,
                                margin = margin(b = 3)),
      strip.background = element_rect(fill = "grey88", color = "grey35",
                                      linewidth = 0.35),
      strip.text = element_text(size = base_size - 0.8, color = "grey20"),
      legend.title = element_text(size = base_size - 0.6),
      legend.text = element_text(size = base_size - 1.2),
      plot.margin = margin(3, 3, 3, 3)
    )
}

safe_corr <- function(x) {
  out <- pmax(-1, pmin(1, x))
  if (!is.null(dim(x))) {
    dim(out) <- dim(x)
    dimnames(out) <- dimnames(x)
  }
  out
}

records_to_fitted_y_matrix <- function(rows) {
  rows <- rows %>% dplyr::arrange(.data$effect)
  mat <- do.call(cbind, rows$fitted_y_effect)
  storage.mode(mat) <- "double"
  colnames(mat) <- as.character(rows$effect)
  mat
}

safe_cor_matrix <- function(mat_a, mat_b, sd_eps = 1e-12) {
  if (nrow(mat_a) != nrow(mat_b)) {
    stop("Mismatched n (", nrow(mat_a), " vs ", nrow(mat_b), ").")
  }
  sd_a <- apply(mat_a, 2L, stats::sd)
  sd_b <- apply(mat_b, 2L, stats::sd)
  active_a <- is.finite(sd_a) & sd_a > sd_eps
  active_b <- is.finite(sd_b) & sd_b > sd_eps

  cor_mat <- matrix(0, nrow = ncol(mat_a), ncol = ncol(mat_b),
                    dimnames = list(colnames(mat_a), colnames(mat_b)))
  if (any(active_a) && any(active_b)) {
    sub <- suppressWarnings(stats::cor(
      mat_a[, active_a, drop = FALSE],
      mat_b[, active_b, drop = FALSE],
      use = "pairwise.complete.obs"
    ))
    sub[!is.finite(sub)] <- 0
    cor_mat[active_a, active_b] <- sub
  }
  cor_mat
}

clip01 <- function(x) {
  pmin(pmax(x, 0), 1)
}

clamp <- function(x, lim) {
  pmin(pmax(x, lim[[1]]), lim[[2]])
}

smooth_line_df <- function(df, x_col, y_col, out_x = NULL,
                           spline_df = baseline_curve_spline_df) {
  d <- df %>%
    dplyr::filter(is.finite(.data[[x_col]]), is.finite(.data[[y_col]])) %>%
    dplyr::arrange(.data[[x_col]])
  if (nrow(d) < spline_df + 2L) {
    return(tibble::tibble())
  }
  fit <- stats::smooth.spline(
    x = d[[x_col]],
    y = d[[y_col]],
    df = spline_df
  )
  if (is.null(out_x)) {
    out_x <- seq(min(d[[x_col]], na.rm = TRUE),
                 max(d[[x_col]], na.rm = TRUE),
                 length.out = 300L)
  }
  tibble::tibble(
    x = out_x,
    y = clip01(stats::predict(fit, x = out_x)$y)
  )
}
```

# Read Cached Fits

```{r read-cache}
records <- readRDS(records_rds) %>%
  dplyr::mutate(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    matrix_id = as.character(.data$matrix_id),
    effect = as.integer(.data$effect),
    model_method = as.character(.data$model_method)
  ) %>%
  dplyr::filter(.data$model_method %in% methods_in_order)

auprc_per_dataset <- readr::read_csv(
  auprc_per_dataset_csv, show_col_types = FALSE
) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

auprc_long <- auprc_per_dataset %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    !!baseline_method := .data$auprc_susie_vanilla,
    !!mu_method := .data$auprc_susine_functional_mu,
    !!warm_mu_method := .data$auprc_susie_vanilla_warm_from_mu,
    !!pi_method := .data$auprc_susie_functional_pi,
    !!warm_pi_method := .data$auprc_susie_vanilla_warm_from_pi
  ) %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(methods_in_order),
    names_to = "model_method",
    values_to = "auprc"
  )

dataset_bundles <- readr::read_csv(
  dataset_bundles_csv, show_col_types = FALSE
) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

records_by_dataset_method <- records %>%
  split(list(.$dataset_bundle_id, .$model_method))

cat("Effect records:", nrow(records), "\n")
cat("Datasets:", dplyr::n_distinct(records$dataset_bundle_id), "\n")
cat("Methods:", paste(sort(unique(records$model_method)), collapse = ", "), "\n")
```

# Build Top-K Causal Fitted-y Signals

```{r build-truth-fitted-y}
if (file.exists(truth_fitted_y_rds)) {
  truth_fitted_y_lookup <- readRDS(truth_fitted_y_rds)
  cat("Loaded cached top-K causal fitted-y lookup from ",
      truth_fitted_y_rds, "\n", sep = "")
} else {
  if (!requireNamespace("devtools", quietly = TRUE)) {
    stop("devtools is needed to regenerate top-K causal fitted-y signals.")
  }
  suppressMessages(devtools::load_all(here()))

  susine_sister_path <- here("..", "susine")
  if (file.exists(file.path(susine_sister_path, "DESCRIPTION"))) {
    options(test_susine.local_susine_path = normalizePath(
      susine_sister_path, winslash = "/", mustWork = TRUE
    ))
    suppressMessages(devtools::load_all(susine_sister_path))
  }

  source(here("vignettes", "one_off_validations", "_5arm_cfg.R"))
  cfg_bundle <- build_5arm_cfg(
    here_root = here(),
    output_root = output_root,
    baseline_screening_csv = file.path(
      output_root, "slurm_output", baseline_job_name_5arm,
      baseline_parent_job_id_5arm, "consolidated", "screening_summary.csv"
    ),
    baseline_cfg_path = file.path(
      output_root, "run_history", baseline_job_name_5arm,
      baseline_parent_job_id_5arm, "job_config.json"
    ),
    settings_cache_csv = file.path(study_dir, "selected_5arm_settings.csv"),
    prefer_settings_cache = TRUE,
    allow_manual_fallback = TRUE
  )
  cfg_for_truth <- cfg_bundle$cfg

  ds_ids <- sort(unique(records$dataset_bundle_id))
  truth_fitted_y_lookup <- vector("list", length(ds_ids))
  names(truth_fitted_y_lookup) <- ds_ids
  started <- proc.time()[["elapsed"]]

  for (i in seq_along(ds_ids)) {
    ds_id <- ds_ids[[i]]
    bundle_row <- dataset_bundles %>%
      dplyr::filter(.data$dataset_bundle_id == !!ds_id)
    if (nrow(bundle_row) != 1L) {
      stop("Expected one dataset bundle row for ", ds_id,
           "; found ", nrow(bundle_row), ".")
    }

    data_bundle <- test_susine:::generate_data_for_bundle(
      bundle_row, cfg_for_truth
    )
    causal <- as.integer(data_bundle$causal_idx)
    beta <- as.numeric(data_bundle$beta)
    if (!length(causal)) {
      truth_fitted_y_lookup[[ds_id]] <- NULL
      next
    }

    abs_beta_at_causals <- abs(beta[causal])
    ord <- order(abs_beta_at_causals, decreasing = TRUE)
    top_k_i <- min(truth_top_k, length(causal))
    top_idx <- causal[ord[seq_len(top_k_i)]]

    var_y <- stats::var(data_bundle$y)
    fitted_y_effect <- lapply(top_idx, function(j) {
      as.numeric(data_bundle$X[, j] * beta[j])
    })
    var_y_explained_frac <- vapply(
      fitted_y_effect,
      function(v) stats::var(v) / var_y,
      numeric(1)
    )

    truth_fitted_y_lookup[[ds_id]] <- list(
      top_idx = as.integer(top_idx),
      beta_top = as.numeric(beta[top_idx]),
      fitted_y_effect = fitted_y_effect,
      var_y_explained_frac = as.numeric(var_y_explained_frac),
      n_obs = length(data_bundle$y)
    )

    if (i %% 50L == 0L || i == length(ds_ids)) {
      elapsed <- proc.time()[["elapsed"]] - started
      cat(sprintf("Built truth fitted-y %d/%d | elapsed %.1fs\n",
                  i, length(ds_ids), elapsed))
    }
  }

  truth_fitted_y_lookup <- truth_fitted_y_lookup[
    !vapply(truth_fitted_y_lookup, is.null, logical(1))
  ]
  saveRDS(truth_fitted_y_lookup, truth_fitted_y_rds)
  cat("Saved top-K causal fitted-y lookup to ",
      truth_fitted_y_rds, "\n", sep = "")
}
```

# Causal Basis Alignment Score

```{r compute-alignment}
# For each top-K causal signal t_j, score coverage_j = max_l corr(t_j, f_l)^2,
# where f_l ranges over fitted model effects. The weighted score audits the
# truth side directly and does not require a one-to-one effect assignment.
compute_alignment_one_fit <- function(rows, truth_obj) {
  if (is.null(rows) || !nrow(rows) || is.null(truth_obj)) {
    return(NULL)
  }

  fit_mat <- records_to_fitted_y_matrix(rows)
  truth_mat <- do.call(cbind, truth_obj$fitted_y_effect)
  storage.mode(truth_mat) <- "double"
  colnames(truth_mat) <- paste0("truth_", seq_len(ncol(truth_mat)))

  cor_mat <- safe_cor_matrix(fit_mat, truth_mat)
  coverage_by_truth <- apply(safe_corr(cor_mat)^2, 2L, max, na.rm = TRUE)
  coverage_by_truth[!is.finite(coverage_by_truth)] <- 0

  truth_w <- pmax(tidyr::replace_na(
    as.numeric(truth_obj$var_y_explained_frac), 0
  ), 0)
  model_w <- pmax(tidyr::replace_na(
    as.numeric(rows$var_y_explained_frac), 0
  ), 0)

  truth_pve <- sum(truth_w, na.rm = TRUE)
  model_pve <- sum(model_w, na.rm = TRUE)
  alignment_weighted <- if (truth_pve > 0) {
    sum(truth_w * coverage_by_truth, na.rm = TRUE) / truth_pve
  } else {
    NA_real_
  }

  tibble::tibble(
    dataset_bundle_id = as.character(rows$dataset_bundle_id[[1]]),
    matrix_id = as.character(rows$matrix_id[[1]]),
    model_method = as.character(rows$model_method[[1]]),
    n_model_effects = nrow(rows),
    n_truth_signals = length(truth_w),
    causal_basis_alignment_score = alignment_weighted,
    causal_basis_misalignment = 1 - alignment_weighted,
    causal_basis_alignment_unweighted = mean(coverage_by_truth, na.rm = TRUE),
    min_causal_signal_coverage = min(coverage_by_truth, na.rm = TRUE),
    model_pve_frac = model_pve,
    truth_topk_pve_frac = truth_pve,
    pve_ratio_model_to_truth = model_pve / truth_pve,
    pve_log2_ratio_model_to_truth = log2(model_pve / truth_pve),
    pve_abs_log2_ratio_model_to_truth = abs(log2(model_pve / truth_pve))
  )
}

if (file.exists(alignment_scores_csv)) {
  alignment_scores <- readr::read_csv(
    alignment_scores_csv, show_col_types = FALSE
  ) %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))
  cat("Loaded cached causal basis alignment scores from ",
      alignment_scores_csv, "\n", sep = "")
} else {
  rows_out <- list()
  ds_ids <- intersect(
    sort(unique(records$dataset_bundle_id)),
    names(truth_fitted_y_lookup)
  )
  started <- proc.time()[["elapsed"]]
  counter <- 0L
  for (ds_id in ds_ids) {
    truth_obj <- truth_fitted_y_lookup[[ds_id]]
    for (m in methods_in_order) {
      rows <- records_by_dataset_method[[paste(ds_id, m, sep = ".")]]
      res <- compute_alignment_one_fit(rows, truth_obj)
      if (!is.null(res)) {
        rows_out[[length(rows_out) + 1L]] <- res
      }
      counter <- counter + 1L
    }
    if (length(rows_out) %% 250L == 0L) {
      elapsed <- proc.time()[["elapsed"]] - started
      cat(sprintf("Computed alignment for %d fits | elapsed %.1fs\n",
                  length(rows_out), elapsed))
    }
  }

  alignment_scores <- dplyr::bind_rows(rows_out) %>%
    dplyr::left_join(auprc_long, by = c("dataset_bundle_id", "model_method"))
  readr::write_csv(alignment_scores, alignment_scores_csv)
  cat("Wrote causal basis alignment scores to ",
      alignment_scores_csv, "\n", sep = "")
}

print(
  alignment_scores %>%
    dplyr::group_by(.data$model_method) %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean_alignment = mean(.data$causal_basis_alignment_score, na.rm = TRUE),
      mean_misalignment = mean(.data$causal_basis_misalignment, na.rm = TRUE),
      mean_pve_ratio = mean(.data$pve_ratio_model_to_truth, na.rm = TRUE),
      mean_auprc = mean(.data$auprc, na.rm = TRUE),
      .groups = "drop"
    )
)
```

# Baseline Relationships

```{r baseline-relationships, fig.width=10.2, fig.height=4.8}
baseline_alignment <- alignment_scores %>%
  dplyr::filter(.data$model_method == baseline_method) %>%
  dplyr::filter(
    is.finite(.data$auprc),
    is.finite(.data$causal_basis_alignment_score),
    is.finite(.data$pve_log2_ratio_model_to_truth)
  )

alignment_range <- range(
  baseline_alignment$causal_basis_alignment_score, na.rm = TRUE
)
baseline_alignment_curve_fit <- stats::smooth.spline(
  x = baseline_alignment$causal_basis_alignment_score,
  y = baseline_alignment$auprc,
  df = baseline_curve_spline_df
)
predict_baseline_auprc_from_alignment <- function(alignment_score) {
  alignment_score <- clamp(alignment_score, alignment_range)
  clip01(stats::predict(
    baseline_alignment_curve_fit,
    x = alignment_score
  )$y)
}

misalignment_curve <- smooth_line_df(
  baseline_alignment,
  x_col = "causal_basis_misalignment",
  y_col = "auprc"
)
size_curve <- smooth_line_df(
  baseline_alignment,
  x_col = "pve_log2_ratio_model_to_truth",
  y_col = "auprc"
)
alignment_curve <- tibble::tibble(
  causal_basis_alignment_score = seq(
    alignment_range[[1]], alignment_range[[2]], length.out = 300L
  )
) %>%
  dplyr::mutate(
    causal_basis_misalignment = 1 - .data$causal_basis_alignment_score,
    auprc_hat = predict_baseline_auprc_from_alignment(
      .data$causal_basis_alignment_score
    )
  ) %>%
  dplyr::arrange(.data$causal_basis_misalignment)

p_baseline_misalignment <- ggplot(
  baseline_alignment,
  aes(x = .data$causal_basis_misalignment, y = .data$auprc,
      color = .data$pve_log2_ratio_model_to_truth)
) +
  geom_point(alpha = 0.48, size = 1.45, stroke = 0) +
  geom_line(
    data = alignment_curve,
    aes(x = .data$causal_basis_misalignment, y = .data$auprc_hat),
    inherit.aes = FALSE,
    color = "#c44e52", linewidth = 0.75
  ) +
  scale_color_viridis_c(option = "C", name = "log2(model PVE /\ntruth PVE)") +
  labs(
    x = "Causal basis misalignment (1 - weighted coverage)",
    y = "AUPRC (cold vanilla)",
    title = "Baseline AUPRC vs causal basis misalignment"
  ) +
  theme_alignment(base_size = 9)

p_baseline_size <- ggplot(
  baseline_alignment,
  aes(x = .data$pve_log2_ratio_model_to_truth, y = .data$auprc,
      color = .data$causal_basis_alignment_score)
) +
  geom_vline(xintercept = 0, color = "grey60", linewidth = 0.25) +
  geom_point(alpha = 0.48, size = 1.45, stroke = 0) +
  geom_line(
    data = size_curve,
    aes(x = .data$x, y = .data$y),
    inherit.aes = FALSE,
    color = "#c44e52", linewidth = 0.75
  ) +
  scale_color_viridis_c(option = "D", name = "Causal basis\nalignment") +
  labs(
    x = "log2(model PVE / top-K causal PVE)",
    y = "AUPRC (cold vanilla)",
    title = "Baseline AUPRC vs explanatory basis size"
  ) +
  theme_alignment(base_size = 9)

p_baseline_combined <- p_baseline_misalignment | p_baseline_size
ggsave(
  baseline_alignment_png,
  p_baseline_combined, width = 10.2, height = 4.8, dpi = 220, bg = "white"
)
print(p_baseline_combined)
```

# Mean Delta Attribution

```{r transition-decomposition, fig.width=11.5, fig.height=7.2}
state_method_map <- tibble::tribble(
  ~chain, ~state,       ~model_method,
  "mu",   "cold",       baseline_method,
  "mu",   "annotated",  mu_method,
  "mu",   "warm",       warm_mu_method,
  "pi",   "cold",       baseline_method,
  "pi",   "annotated",  pi_method,
  "pi",   "warm",       warm_pi_method
)

state_metrics <- alignment_scores %>%
  dplyr::inner_join(state_method_map, by = "model_method",
                    relationship = "many-to-many") %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    chain = .data$chain,
    state = .data$state,
    alignment = .data$causal_basis_alignment_score,
    misalignment = .data$causal_basis_misalignment,
    size_log2_ratio = .data$pve_log2_ratio_model_to_truth,
    auprc = .data$auprc
  )

state_wide <- state_metrics %>%
  tidyr::pivot_wider(
    names_from = .data$state,
    values_from = c(.data$alignment, .data$misalignment,
                    .data$size_log2_ratio, .data$auprc)
  )

transition_rows <- dplyr::bind_rows(
  state_wide %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = .data$chain,
      transition = "cold_ann",
      start_alignment = .data$alignment_cold,
      end_alignment = .data$alignment_annotated,
      start_misalignment = .data$misalignment_cold,
      end_misalignment = .data$misalignment_annotated,
      start_auprc = .data$auprc_cold,
      end_auprc = .data$auprc_annotated
    ),
  state_wide %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = .data$chain,
      transition = "cold_warm",
      start_alignment = .data$alignment_cold,
      end_alignment = .data$alignment_warm,
      start_misalignment = .data$misalignment_cold,
      end_misalignment = .data$misalignment_warm,
      start_auprc = .data$auprc_cold,
      end_auprc = .data$auprc_warm
    )
) %>%
  dplyr::mutate(
    f_start = predict_baseline_auprc_from_alignment(.data$start_alignment),
    f_end = predict_baseline_auprc_from_alignment(.data$end_alignment),
    observed_delta = .data$end_auprc - .data$start_auprc,
    alignment_pred_delta = .data$f_end - .data$f_start,
    residual_component = .data$observed_delta - .data$alignment_pred_delta,
    delta_alignment = .data$end_alignment - .data$start_alignment,
    delta_misalignment = .data$end_misalignment - .data$start_misalignment,
    chain = factor(.data$chain, levels = c("mu", "pi"),
                   labels = c("mu chain", "pi chain")),
    transition = factor(
      .data$transition,
      levels = c("cold_ann", "cold_warm"),
      labels = c("Cold -> annotated", "Cold -> warm vanilla")
    )
  ) %>%
  dplyr::filter(
    is.finite(.data$observed_delta),
    is.finite(.data$alignment_pred_delta),
    is.finite(.data$residual_component)
  )

transition_summary <- transition_rows %>%
  dplyr::group_by(.data$chain, .data$transition) %>%
  dplyr::summarise(
    n = dplyr::n(),
    observed_mean = mean(.data$observed_delta, na.rm = TRUE),
    alignment_pred_mean = mean(.data$alignment_pred_delta, na.rm = TRUE),
    residual_mean = mean(.data$residual_component, na.rm = TRUE),
    alignment_share_of_mean = .data$alignment_pred_mean /
      .data$observed_mean,
    mean_delta_alignment = mean(.data$delta_alignment, na.rm = TRUE),
    .groups = "drop"
  )

readr::write_csv(transition_rows, decomp_rows_csv)
readr::write_csv(transition_summary, decomp_summary_csv)
cat("Causal basis alignment decomposition summary:\n")
print(transition_summary)

decomp_lim <- max(abs(c(
  transition_rows$observed_delta,
  transition_rows$alignment_pred_delta
)), na.rm = TRUE)
decomp_lim <- max(0.05, decomp_lim)

p_alignment_scatter <- ggplot(
  transition_rows,
  aes(x = .data$alignment_pred_delta, y = .data$observed_delta,
      color = .data$start_misalignment)
) +
  geom_hline(yintercept = 0, color = "grey65", linewidth = 0.25) +
  geom_vline(xintercept = 0, color = "grey65", linewidth = 0.25) +
  geom_abline(slope = 1, intercept = 0, color = "grey35",
              linetype = "dashed", linewidth = 0.35) +
  geom_point(alpha = 0.58, size = 1.45, stroke = 0) +
  facet_grid(chain ~ transition) +
  scale_color_viridis_c(option = "C", name = "Baseline causal\nbasis misalignment") +
  coord_cartesian(xlim = c(-decomp_lim, decomp_lim),
                  ylim = c(-decomp_lim, decomp_lim)) +
  labs(
    x = "AUPRC delta predicted from causal basis alignment",
    y = "Observed AUPRC delta",
    title = "How much observed AUPRC gain is predicted by causal basis alignment?"
  ) +
  theme_alignment(base_size = 9) +
  theme(legend.position = "right")

decomp_components <- transition_summary %>%
  dplyr::select(.data$chain, .data$transition,
                .data$alignment_pred_mean,
                .data$residual_mean) %>%
  tidyr::pivot_longer(
    cols = c("alignment_pred_mean", "residual_mean"),
    names_to = "component",
    values_to = "mean_delta"
  ) %>%
  dplyr::mutate(
    component = factor(
      .data$component,
      levels = c("alignment_pred_mean", "residual_mean"),
      labels = c("Causal basis alignment", "Residual / within-basis")
    )
  )

p_decomp_bar <- ggplot(
  decomp_components,
  aes(x = .data$chain, y = .data$mean_delta, fill = .data$component)
) +
  geom_hline(yintercept = 0, color = "grey45", linewidth = 0.25) +
  geom_col(width = 0.62) +
  geom_point(
    data = transition_summary,
    aes(x = .data$chain, y = .data$observed_mean),
    inherit.aes = FALSE,
    shape = 21, fill = "white", color = "black", size = 2.3,
    stroke = 0.35
  ) +
  facet_wrap(~ transition, nrow = 1) +
  scale_fill_manual(
    values = c(
      "Causal basis alignment" = "#4c72b0",
      "Residual / within-basis" = "#dd8452"
    )
  ) +
  labs(
    x = NULL,
    y = "Mean AUPRC delta",
    fill = NULL,
    title = "Mean observed gain split by the baseline alignment curve",
    subtitle = "White dot marks observed mean delta. Residual is not predicted by causal basis alignment."
  ) +
  theme_alignment(base_size = 9) +
  theme(legend.position = "bottom")

p_metric_shift <- ggplot(
  transition_rows,
  aes(x = .data$start_misalignment, y = -.data$delta_misalignment,
      color = .data$observed_delta)
) +
  geom_hline(yintercept = 0, color = "grey65", linewidth = 0.25) +
  geom_point(alpha = 0.58, size = 1.45, stroke = 0) +
  facet_grid(chain ~ transition) +
  scale_color_gradient2(
    low = "#2b6cb0", mid = "white", high = "#c53030",
    midpoint = 0, name = "Observed\nAUPRC delta"
  ) +
  labs(
    x = "Baseline causal basis misalignment",
    y = "Improvement in causal basis misalignment",
    title = "Where alignment improves across each transition"
  ) +
  theme_alignment(base_size = 9) +
  theme(legend.position = "right")

p_decomp_combined <- p_alignment_scatter /
  (p_decomp_bar | p_metric_shift) +
  plot_layout(heights = c(1.18, 1))

ggsave(
  decomposition_png,
  p_decomp_combined, width = 11.5, height = 7.2, dpi = 220, bg = "white"
)
print(p_decomp_combined)
```

```{r delta-alignment-mu-vs-pi, fig.width=8.8, fig.height=4.2}
if (exists("transition_rows")) {
  delta_alignment_mupi <- transition_rows %>%
    dplyr::mutate(
      chain = as.character(.data$chain),
      transition = as.character(.data$transition)
    ) %>%
    dplyr::select(.data$dataset_bundle_id, .data$chain, .data$transition,
                  .data$delta_alignment, .data$observed_delta,
                  .data$start_misalignment) %>%
    tidyr::pivot_wider(
      names_from = .data$chain,
      values_from = c(.data$delta_alignment, .data$observed_delta,
                      .data$start_misalignment)
    ) %>%
    dplyr::mutate(
      delta_alignment_mu_minus_pi =
        .data$`delta_alignment_mu chain` -
        .data$`delta_alignment_pi chain`,
      observed_delta_mu_minus_pi =
        .data$`observed_delta_mu chain` -
        .data$`observed_delta_pi chain`,
      baseline_misalignment = rowMeans(
        cbind(.data$`start_misalignment_mu chain`,
              .data$`start_misalignment_pi chain`),
        na.rm = TRUE
      ),
      transition = factor(
        .data$transition,
        levels = c("Cold -> annotated", "Cold -> warm vanilla")
      )
    )

  alignment_lim <- max(abs(c(
    delta_alignment_mupi$`delta_alignment_mu chain`,
    delta_alignment_mupi$`delta_alignment_pi chain`
  )), na.rm = TRUE)
  alignment_lim <- max(0.01, alignment_lim)

  p_delta_alignment_mupi <- ggplot(
    delta_alignment_mupi,
    aes(x = .data$`delta_alignment_pi chain`,
        y = .data$`delta_alignment_mu chain`,
        color = .data$observed_delta_mu_minus_pi)
  ) +
    geom_hline(yintercept = 0, color = "grey65", linewidth = 0.25) +
    geom_vline(xintercept = 0, color = "grey65", linewidth = 0.25) +
    geom_abline(slope = 1, intercept = 0, color = "grey35",
                linetype = "dashed", linewidth = 0.35) +
    geom_point(alpha = 0.62, size = 1.55, stroke = 0) +
    facet_wrap(~ transition, nrow = 1) +
    scale_color_gradient2(
      low = "#2b6cb0", mid = "white", high = "#c53030",
      midpoint = 0,
      name = expression(Delta * "AUPRC"[mu] - Delta * "AUPRC"[pi])
    ) +
    coord_fixed(xlim = c(-alignment_lim, alignment_lim),
                ylim = c(-alignment_lim, alignment_lim)) +
    labs(
      x = "Delta causal basis alignment (pi chain)",
      y = "Delta causal basis alignment (mu chain)",
      title = "Mu vs pi movement in causal basis alignment",
      subtitle = "Dashed line: equal alignment movement. Positive delta means better causal signal coverage."
    ) +
    theme_alignment(base_size = 9) +
    theme(legend.position = "right")

  ggsave(
    delta_alignment_mupi_png,
    p_delta_alignment_mupi, width = 8.8, height = 4.2, dpi = 220,
    bg = "white"
  )
  print(p_delta_alignment_mupi)

  cat("Mu vs pi delta-alignment summary:\n")
  print(
    delta_alignment_mupi %>%
      dplyr::group_by(.data$transition) %>%
      dplyr::summarise(
        n = dplyr::n(),
        mean_delta_alignment_mu =
          mean(.data$`delta_alignment_mu chain`, na.rm = TRUE),
        mean_delta_alignment_pi =
          mean(.data$`delta_alignment_pi chain`, na.rm = TRUE),
        mean_mu_minus_pi =
          mean(.data$delta_alignment_mu_minus_pi, na.rm = TRUE),
        frac_mu_gt_pi =
          mean(.data$delta_alignment_mu_minus_pi > 0, na.rm = TRUE),
        .groups = "drop"
      )
  )
} else {
  message("Skipping mu-vs-pi delta alignment plot; transition_rows missing.")
}
```

# Interpretation Notes

The alignment score is deliberately truth-side:

```text
coverage_j = max_l cor(top-K causal signal j, fitted effect l)^2
alignment = weighted mean_j coverage_j
```

This differs from one-to-one effect matching because every top-K causal signal
is audited directly, and several causal signals may select the same fitted
effect if the model has merged or rotated components. The score is still
correlation-based, so correlated causal signals can inflate apparent coverage;
that is why the decomposition should be read as a diagnostic for explanatory
basis alignment, not as an oracle causal mechanism.

The attribution is a mean-shift calculation. It fits the cold-vanilla baseline
mean curve

```text
E[AUPRC | causal basis alignment]
```

and asks how much each transition moves along that curve. The residual term
contains within-basis SNP-ranking effects, model-specific effects not present
in the cold baseline curve, and noise.

The PVE totals are component-wise sums of `var(fitted_y_component) / var(y)`.
They remain in the baseline context plot as size diagnostics, not joint
variance decompositions. When fitted or causal components are non-orthogonal,
the summed component PVE can differ from the variance explained by the span of
the components.
