---
title: "Compute Scenario 1 LD Alpha Metrics"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

One-off workbook to compute continuous LD-tail metrics for the 150
locally saved scenario-1 genotype matrices.

For each matrix, this workbook computes:

- `mean(|r|^alpha)` over the strict upper triangle for `alpha = 1, ..., 8`
- the same quantity after zeroing out all pairs with `|r| < 0.5`

It writes one row per `matrix_id`, using the same `matrix_id` assignment
logic as the run-control code for `scenario_1`.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/compute_scenario1_ld_alpha_metrics.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
library(dplyr)
library(readr)
library(knitr)
```

## 1. Inputs

```{r inputs}
repo_root <- here()
summary_path <- here("data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv")
output_csv <- here("data", "sampled_simulated_genotypes", "scenario_1_ld_alpha_metrics.csv")

alphas <- 1:8
ld_threshold <- 0.5

cat("Repo root:", repo_root, "\n")
cat("Summary path:", summary_path, "\n")
cat("Output CSV:", output_csv, "\n")
```

## 2. Matrix Catalog

```{r catalog}
stopifnot(file.exists(summary_path))

matrix_catalog <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = repo_root,
  summary_path = summary_path
) %>%
  dplyr::filter(.data$data_scenario == "scenario_1") %>%
  dplyr::arrange(.data$matrix_id)

cat("Scenario-1 matrices:", nrow(matrix_catalog), "\n")
kable(head(matrix_catalog %>% dplyr::select(.data$matrix_id, .data$dataset_label, .data$matrix_path), 10))
```

## 3. Helpers

```{r helpers}
compute_ld_alpha_metrics_one <- function(matrix_path,
                                         repo_root,
                                         alphas = 1:8,
                                         ld_threshold = 0.5) {
  matrix_path_abs <- if (grepl("^[A-Za-z]:/|^/", matrix_path)) {
    matrix_path
  } else {
    file.path(repo_root, matrix_path)
  }

  x_tbl <- readr::read_tsv(matrix_path_abs, show_col_types = FALSE)
  if (ncol(x_tbl) < 2L) {
    stop("Expected at least one ID column plus genotype columns in: ", matrix_path_abs)
  }

  x_mat <- as.matrix(x_tbl[, -1, drop = FALSE])
  storage.mode(x_mat) <- "numeric"

  r_mat <- suppressWarnings(stats::cor(x_mat, use = "pairwise.complete.obs"))
  abs_r <- abs(r_mat[upper.tri(r_mat)])
  abs_r_thr <- ifelse(abs_r >= ld_threshold, abs_r, 0)

  raw_vals <- stats::setNames(
    vapply(alphas, function(alpha) mean(abs_r^alpha, na.rm = TRUE), numeric(1)),
    paste0("ld_abs_alpha_", alphas)
  )

  thr_vals <- stats::setNames(
    vapply(alphas, function(alpha) mean(abs_r_thr^alpha, na.rm = TRUE), numeric(1)),
    paste0("ld_abs_alpha_", alphas, "_thr_", gsub("\\.", "p", as.character(ld_threshold)))
  )

  out <- tibble::as_tibble(as.list(c(raw_vals, thr_vals)))
  rm(x_tbl, x_mat, r_mat, abs_r, abs_r_thr)
  gc(verbose = FALSE)
  out
}
```

## 4. Compute Metrics

```{r compute}
metric_rows <- vector("list", nrow(matrix_catalog))

for (i in seq_len(nrow(matrix_catalog))) {
  row <- matrix_catalog[i, , drop = FALSE]
  cat(sprintf("[%d/%d] matrix_id=%s %s\n",
              i, nrow(matrix_catalog), row$matrix_id[[1]], row$dataset_label[[1]]))

  metric_rows[[i]] <- row %>%
    dplyr::select(
      "matrix_id", "data_scenario", "dataset_label",
      "matrix_index", "matrix_path", "manifest_path",
      "participant_count", "snps_post", "snp_set"
    ) %>%
    dplyr::bind_cols(
      compute_ld_alpha_metrics_one(
        matrix_path = row$matrix_path[[1]],
        repo_root = repo_root,
        alphas = alphas,
        ld_threshold = ld_threshold
      )
    )

  gc(verbose = FALSE)
}

ld_alpha_metrics <- dplyr::bind_rows(metric_rows) %>%
  dplyr::arrange(.data$matrix_id)

readr::write_csv(ld_alpha_metrics, output_csv)

cat("Wrote:", output_csv, "\n")
cat("Rows:", nrow(ld_alpha_metrics), "\n")
cat("Columns:", ncol(ld_alpha_metrics), "\n")

kable(head(ld_alpha_metrics, 10))
```
