---
title: "C-CSR Ensemble Smoke Run Control"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

One-off local smoke test for the `C-CSR` ensemble only. This builds a
single-dataset-bundle config and includes a bottom cell to run task `1`
locally.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/csr_ensemble_smoke_run_control.Rmd")
library(here)
here()

library(devtools)

susine_path <- here("..", "susine")
if (!file.exists(file.path(susine_path, "DESCRIPTION"))) {
  stop("susine package not found at: ", susine_path)
}

options(
  test_susine.local_susine_path =
    normalizePath(susine_path, winslash = "/", mustWork = TRUE)
)

devtools::load_all(susine_path, reset = TRUE)
devtools::load_all(".", reset = TRUE)

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 1. Minimal C-CSR Grid

Use a tiny `c x sigma_0_2` grid so the refit path is exercised without
turning this into a real compute run.

```{r grids}
c_grid_smoke <- c(0.0, 0.5)
sigma_grid_smoke <- c(0.1, 0.2)

stopifnot(length(c_grid_smoke) * length(sigma_grid_smoke) == 4L)
stopifnot(0.2 %in% sigma_grid_smoke)

cat("c grid:", c_grid_smoke, "\n")
cat("sigma grid:", sigma_grid_smoke, "\n")
```

# 2. C-CSR Spec Only

```{r spec}
spec_C_CSR <- list(
  name = "C-CSR",
  use_case_ids = "susine_functional_mu",
  exploration_methods = "cs_grid_refit",
  exploration_mode = "separate",
  K = 4L,
  c_grid_values = c_grid_smoke,
  sigma_0_2_grid_values = sigma_grid_smoke
)

specs <- list(spec_C_CSR)

str(specs)
```

# 3. Build One-Bundle Config

```{r build-config}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)

data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::slice_head(n = 1)

cat("Selected matrix rows:", nrow(data_matrix_catalog), "\n")
knitr::kable(data_matrix_catalog)

cfg <- make_job_config(
  job_name = "csr_ensemble_smoke",
  HPC = FALSE,
  time = "00:20:00",
  mem = "4G",
  email = "mgc5166@psu.edu",
  output_root = here("output"),
  grid_mode = "full",

  use_case_ids = "susine_functional_mu",
  exploration_specs = specs,

  L_grid = 10L,
  y_noise_grid = NA_real_,
  p_star_grid = NA_integer_,
  architecture_grid = "susie2_oligogenic",
  seeds = 1L,

  prior_quality = tibble::tibble(
    annotation_r2 = 0.2,
    inflate_match = 1
  ),

  sigma_0_2_scalars = 0.2,

  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),

  task_unit = "dataset",
  bundles_per_task = 1L,
  task_assignment_seed = 1L,

  max_iter = 50L,

  aggregation_methods = c(
    "cluster_weight",
    "cluster_weight_jsd_050",
    "elbo_softmax",
    "max_elbo",
    "uniform"
  ),
  include_overall_pool = FALSE,
  softmax_temperature = 1,

  credible_set_rho = 0.95,
  purity_threshold = 0.50,
  verbose_file_output = FALSE,
  write_snps_parquet = TRUE,
  write_confusion_bins = TRUE,
  write_tier_cs_metrics = TRUE,

  metrics_settings = list(
    pip_breaks = sort(unique(c(
      seq(0.00, 0.05, by = 0.005),
      seq(0.05, 0.10, by = 0.01),
      seq(0.10, 1.00, by = 0.02)
    ))),
    z_top_k = 10,
    jsd_threshold = 0.15
  )
)

stopifnot(setequal(unique(cfg$tables$runs$spec_name), "C-CSR"))
stopifnot(cfg$job$n_tasks == 1L)
```

# 4. Smoke Budget

```{r budget}
runs <- cfg$tables$runs

cat("Total runs:", nrow(runs), "\n")
cat("Total dataset bundles:", n_distinct(runs$dataset_bundle_id), "\n")
cat("Total tasks:", cfg$job$n_tasks, "\n")

runs %>%
  dplyr::count(
    spec_name,
    use_case_id,
    annotation_r2,
    exploration_methods,
    name = "n_runs"
  ) %>%
  knitr::kable(caption = "Expected C-CSR smoke runs")
```

# 5. Write Artifacts

```{r write-artifacts}
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)

arts
```

# 6. Local Smoke Run

This is the one-cell local test run. It executes the single dataset task
for `C-CSR`.

```{r run-local-smoke, eval=FALSE}
stopifnot(cfg$job$n_tasks == 1L)

Sys.setenv(
  SUSINE_JOB_NAME = cfg$job$name,
  SUSINE_PARENT_ID = "local_csr_smoke"
)

run_task(
  job_name = cfg$job$name,
  task_id = 1L,
  job_root = here("output"),
  config_path = arts$job_config,
  quiet = FALSE
)
```

# 7. Quick Output Check

```{r validate-local-smoke, eval=FALSE}
staging_dir <- file.path(
  cfg$paths$slurm_output_dir,
  cfg$job$name,
  "local_csr_smoke",
  "task-001"
)
stopifnot(dir.exists(staging_dir))

staging_files <- list.files(
  staging_dir,
  recursive = TRUE,
  full.names = TRUE
)

cat("Staging files:", length(staging_files), "\n")
print(table(tools::file_ext(staging_files)))

model_metrics_files <- list.files(
  staging_dir,
  pattern = "model_metrics",
  recursive = TRUE,
  full.names = TRUE
)

cat("Model metrics files:", length(model_metrics_files), "\n")

if (length(model_metrics_files) > 0) {
  mm <- readr::read_csv(model_metrics_files[1], show_col_types = FALSE)
  if ("spec_name" %in% names(mm)) {
    print(table(mm$spec_name))
  }
  if ("agg_method" %in% names(mm)) {
    print(table(mm$agg_method))
  }
}
```
