---
title: "EB vs Vanilla (mu=0) Diagnostic Workbook"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(forcats)
```

## Goal

This workbook isolates the `mu=0` comparison:

- EB prior variance (`susie_eb`, `susine_eb`, `susie_ash_eb`)
- fixed prior variance counterparts (`susie_vanilla`, `susine_vanilla`, `susie_ash_fixed`)

It provides both:

1. cell-level summaries (`p_star x y_noise`),
2. dataset-level head-to-head comparisons (EB vs fixed 0.2 and EB vs best fixed in grid).

## Inputs

```{r paths}
job_id <- "48988192"
run_name <- "pilot_phase_a"

agg_dir <- file.path("test_susine", "output", "slurm_output", run_name, job_id, "aggregated")
run_hist_dir <- file.path("test_susine", "output", "run_history", run_name, job_id)

conf_path <- file.path(agg_dir, "confusion_matrix_detailed.csv")
model_path <- file.path(agg_dir, "model_metrics.csv")
run_table_path <- file.path(run_hist_dir, "run_table.csv")
use_cases_path <- file.path(run_hist_dir, "use_cases.csv")

stopifnot(file.exists(conf_path), file.exists(model_path), file.exists(run_table_path), file.exists(use_cases_path))

confusion <- read_csv(conf_path, show_col_types = FALSE)
model_metrics <- read_csv(model_path, show_col_types = FALSE)
run_table <- read_csv(run_table_path, show_col_types = FALSE)
use_cases <- read_csv(use_cases_path, show_col_types = FALSE)

cat("confusion rows:", nrow(confusion), "\n")
cat("model_metrics rows:", nrow(model_metrics), "\n")
cat("run_table rows:", nrow(run_table), "\n")
```

## Helpers

```{r helpers}
compute_power_fdr_curves <- function(confusion_df, group_vars) {
  confusion_df %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars)), pip_threshold) %>%
    dplyr::summarise(
      n_causal_at_bucket = sum(n_causal_at_bucket, na.rm = TRUE),
      n_noncausal_at_bucket = sum(n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::arrange(dplyr::desc(pip_threshold), .by_group = TRUE) %>%
    dplyr::mutate(
      TP = cumsum(n_causal_at_bucket),
      FP = cumsum(n_noncausal_at_bucket),
      total_causal = sum(n_causal_at_bucket),
      precision = dplyr::if_else(TP + FP > 0, TP / (TP + FP), NA_real_),
      recall = dplyr::if_else(total_causal > 0, TP / total_causal, NA_real_)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::filter(!is.na(precision), !is.na(recall))
}

compute_auprc_single <- function(precision, recall) {
  df <- data.frame(precision = precision, recall = recall) %>%
    dplyr::filter(!is.na(precision), !is.na(recall)) %>%
    dplyr::arrange(recall)
  if (nrow(df) < 2) return(NA_real_)
  if (min(df$recall) > 0) {
    df <- dplyr::bind_rows(data.frame(recall = 0, precision = 1), df)
  }
  sum(diff(df$recall) * (head(df$precision, -1) + tail(df$precision, -1)) / 2)
}

compute_auprc_from_confusion <- function(confusion_df, group_vars) {
  curves <- compute_power_fdr_curves(confusion_df, group_vars)
  curves %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::summarise(AUPRC = compute_auprc_single(precision, recall), .groups = "drop")
}
```

## Use Case Definitions

```{r use-cases}
pairs <- tibble::tribble(
  ~family,  ~eb_id,          ~fixed_id,
  "susie",  "susie_eb",      "susie_vanilla",
  "susine", "susine_eb",     "susine_vanilla",
  "ash",    "susie_ash_eb",  "susie_ash_fixed"
)

ids <- unique(c(pairs$eb_id, pairs$fixed_id))

labels <- use_cases %>%
  dplyr::filter(use_case_id %in% ids) %>%
  dplyr::select(use_case_id, label)

labels
```

## 1) Cell-Level AUPRC (p_star x y_noise)

```{r cell-level-auprc}
confusion_base <- confusion %>%
  dplyr::filter(is.na(agg_method) | agg_method == "none") %>%
  dplyr::filter(use_case_id %in% ids)

auprc_cell <- compute_auprc_from_confusion(
  confusion_base,
  group_vars = c("use_case_id", "p_star", "y_noise")
) %>%
  dplyr::left_join(labels, by = "use_case_id")

auprc_cell %>% dplyr::arrange(y_noise, use_case_id, p_star) %>% head(20)
```

```{r heatmaps, fig.width=12, fig.height=7}
ggplot(
  auprc_cell %>% dplyr::mutate(use_case_id = forcats::fct_reorder(use_case_id, AUPRC, .fun = mean)),
  aes(x = factor(p_star), y = use_case_id, fill = AUPRC)
) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.3f", AUPRC)), size = 2.8, color = "white") +
  facet_wrap(~paste0("y_noise=", y_noise), ncol = 3) +
  scale_fill_viridis_c(option = "magma", direction = -1) +
  labs(title = "Cell-Level AUPRC", x = "p_star", y = NULL, fill = "AUPRC") +
  theme_minimal(base_size = 11)
```

## 2) Cell-Level EB Deltas

```{r cell-deltas}
cell_delta <- pairs %>%
  dplyr::left_join(
    auprc_cell %>% dplyr::select(use_case_id, p_star, y_noise, AUPRC) %>%
      dplyr::rename(eb_id = use_case_id, auprc_eb = AUPRC),
    by = "eb_id"
  ) %>%
  dplyr::left_join(
    auprc_cell %>% dplyr::select(use_case_id, p_star, y_noise, AUPRC) %>%
      dplyr::rename(fixed_id = use_case_id, auprc_fixed = AUPRC),
    by = c("fixed_id", "p_star", "y_noise")
  ) %>%
  dplyr::mutate(delta = auprc_eb - auprc_fixed)

cell_delta %>%
  dplyr::group_by(family) %>%
  dplyr::summarise(
    mean_delta = mean(delta, na.rm = TRUE),
    median_delta = median(delta, na.rm = TRUE),
    pct_positive = mean(delta > 0, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r cell-delta-plot, fig.width=10, fig.height=4}
ggplot(cell_delta, aes(x = factor(p_star), y = delta, fill = family)) +
  geom_hline(yintercept = 0, linetype = 2, color = "gray40") +
  geom_col(position = "dodge") +
  facet_wrap(~y_noise, ncol = 3) +
  labs(title = "EB - Fixed AUPRC (cell-level)", x = "p_star", y = "Delta AUPRC") +
  theme_minimal(base_size = 11)
```

## 3) Dataset-Level Head-to-Head

```{r dataset-level}
# one model-level row per run_id (base, unfiltered)
run_auprc <- model_metrics %>%
  dplyr::filter(filtering == "unfiltered", is.na(agg_method) | agg_method == "none") %>%
  dplyr::select(run_id, dataset_bundle_id, use_case_id, AUPRC) %>%
  dplyr::left_join(
    run_table %>% dplyr::select(run_id, dataset_bundle_id, use_case_id, p_star, y_noise, sigma_0_2_scalar),
    by = c("run_id", "dataset_bundle_id", "use_case_id")
  )

eb_runs <- run_auprc %>%
  dplyr::filter(use_case_id %in% pairs$eb_id) %>%
  dplyr::rename(eb_id = use_case_id, auprc_eb = AUPRC)

fixed_runs <- run_auprc %>%
  dplyr::filter(use_case_id %in% pairs$fixed_id) %>%
  dplyr::rename(fixed_id = use_case_id)

dataset_cmp_02 <- pairs %>%
  dplyr::left_join(eb_runs, by = "eb_id") %>%
  dplyr::left_join(
    fixed_runs %>%
      dplyr::filter(sigma_0_2_scalar == 0.2) %>%
      dplyr::rename(auprc_fixed_02 = AUPRC),
    by = c("fixed_id", "dataset_bundle_id", "p_star", "y_noise")
  ) %>%
  dplyr::mutate(delta_eb_vs_fixed02 = auprc_eb - auprc_fixed_02)

fixed_best <- fixed_runs %>%
  dplyr::group_by(fixed_id, dataset_bundle_id, p_star, y_noise) %>%
  dplyr::summarise(
    best_fixed = max(AUPRC, na.rm = TRUE),
    best_sigma = sigma_0_2_scalar[which.max(AUPRC)[1]],
    .groups = "drop"
  )

dataset_cmp_best <- pairs %>%
  dplyr::left_join(eb_runs, by = "eb_id") %>%
  dplyr::left_join(
    fixed_best,
    by = c("fixed_id", "dataset_bundle_id", "p_star", "y_noise")
  ) %>%
  dplyr::mutate(delta_eb_vs_best = auprc_eb - best_fixed)

dataset_cmp_02 %>%
  dplyr::group_by(family) %>%
  dplyr::summarise(
    mean_delta = mean(delta_eb_vs_fixed02, na.rm = TRUE),
    median_delta = median(delta_eb_vs_fixed02, na.rm = TRUE),
    pct_positive = mean(delta_eb_vs_fixed02 > 0, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r dataset-best-summary}
dataset_cmp_best %>%
  dplyr::group_by(family) %>%
  dplyr::summarise(
    mean_delta = mean(delta_eb_vs_best, na.rm = TRUE),
    median_delta = median(delta_eb_vs_best, na.rm = TRUE),
    pct_positive = mean(delta_eb_vs_best > 0, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r sigma-frequency}
dataset_cmp_best %>%
  dplyr::count(family, best_sigma) %>%
  dplyr::group_by(family) %>%
  dplyr::mutate(freq = n / sum(n))
```

## 4) Optional Deeper Probe (Single Matrix CSV)

If this file exists, load the custom probe:

- `test_susine/output/eb_probe_mu0_singlematrix_compare.csv`

```{r optional-probe}
probe_path <- file.path("test_susine", "output", "eb_probe_mu0_singlematrix_compare.csv")
if (file.exists(probe_path)) {
  probe <- read_csv(probe_path, show_col_types = FALSE)
  print(
    probe %>%
      dplyr::group_by(p_star, y_noise) %>%
      dplyr::summarise(
        mean_delta_02 = mean(delta_vs_02, na.rm = TRUE),
        pct_pos_02 = mean(delta_vs_02 > 0, na.rm = TRUE),
        mean_delta_best = mean(delta_vs_best, na.rm = TRUE),
        .groups = "drop"
      )
  )
} else {
  cat("Probe CSV not found:", probe_path, "\n")
}
```

## 5) What To Do Next

1. Add convergence and prior-variance traces to model-level exports for EB runs.
2. In next large simulation, stratify reporting by:
   `family x p_star x y_noise x architecture`.
3. Include both comparisons in the paper:
   - EB vs fixed default (`sigma_0_2 = 0.2`)
   - EB vs best fixed in grid (oracle upper bound).
