---
title: "Investigation: EB vs. Vanilla Prior Variance in SuSiE"
date: "2026-03-05"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: show
editor_options:
  markdown:
    wrap: 72
---

# Background and Motivation

## The Prior Variance as Effective Temperature

In the SER, the log Bayes factor for SNP $j$ is:

$$
\log \text{BF}_j = \log \mathcal{N}(\hat\beta_j;\, \mu_{0j},\, \sigma_0^2 + \hat{v}_j) - \log \mathcal{N}(\hat\beta_j;\, 0,\, \hat{v}_j)
$$

where $\hat\beta_j = X_j^T y / d_j$ and $\hat{v}_j = \sigma^2/d_j$.
The posterior inclusion probability is $\alpha_j \propto \pi_j \exp(\log
\text{BF}_j)$.

$\sigma_0^2$ controls the **effective temperature** of this softmax:

-   **Large $\sigma_0^2$**: the prior accommodates extreme effects;
    high-signal SNPs get sharper alpha concentration
    (winner-take-all).
-   **Small $\sigma_0^2$**: the prior penalizes large effects; alpha
    spreads more evenly across LD-correlated SNPs.
-   **$\sigma_0^2 = 0$**: log BF = 0 for all SNPs → alpha reverts to
    prior $\pi$ (automatic sparsity / null effect).

Crucially, $\sigma_0^2$ changes the *contrast* of the BFs without
changing their *ranking* (approximately, when $\hat{v}_j$ is
constant). This is why EB for $\sigma_0^2$ is relatively stable.

## What EB Does

The EB update (`estimate_prior_variance = TRUE` in susieR) finds
$\sigma_0^2$ per effect $\ell$ that maximizes the SER marginal
log-likelihood. Three regimes:

1.  **Typical EB**: converges to a $\sigma_0^2$ that reflects the
    signal strength in the partial residuals. Reasonable, data-adaptive.
2.  **Great EB**: collapses $\sigma_0^2 \to 0$ for null effects
    (automatic sparsity), inflates for real signals. This is the
    per-effect shrinkage advantage.
3.  **Bad EB**: $\sigma_0^2$ estimate is distorted by LD contamination
    in partial residuals; oscillates between iterations; or collapses
    a real signal to zero due to numerical instability.

## Prior Belief vs. Observation

**Prior belief**: EB is unstable because it introduces circularity.
Estimating $\sigma_0^2$ from the data changes alpha, which changes
residuals, which changes the next $\sigma_0^2$ estimate. Under
multimodal posteriors (multiple basins), EB can distort ELBO values
asymmetrically, breaking ensemble aggregation (see `susie_pathology`
notebook, Scenario 3).

**Observation from pilot_phase_a**: EB (`susie_eb`) outperforms vanilla
(`susie_vanilla`, fixed $\sigma_0^2 = 0.2 \times \text{var}(y)$).
This workbook investigates **where** and **why**.

## When is Fixed $\sigma_0^2 = 0.2$ "Right"?

The default scaled prior variance 0.2 assumes each causal SNP explains
about 20% of the trait variance *per effect*. This is:

-   Too large when p\* is high (many weak effects, each explaining
    $\ll 20\%$)
-   Too small when p\* is low and signal is strong (one SNP explaining
    $> 20\%$)
-   About right when L=10, p\*=2, h\^2=0.5 → each effect ~5-25% of
    variance

EB adapts $\sigma_0^2$ per-effect, potentially finding the right
temperature for each component. This is the fundamental advantage.

```{r setup, message=FALSE, warning=FALSE}
here::i_am(
  "vignettes/one_off_validations/eb_vs_vanilla_investigation.Rmd"
)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(readr)
library(knitr)
library(stringr)

theme_set(theme_bw(base_size = 12))
```

# Data Loading

```{r load-data}
# -------------------------------------------------------------------
# Point this to your pilot_phase_a aggregated results directory.
# Expect: model_metrics.csv (or .parquet), dataset_bundles.csv
# -------------------------------------------------------------------
results_path <- here(
  "output", "slurm_output", "pilot_phase_a", "48997267"
)
stopifnot(
  dir.exists(results_path) ||
    stop("Results directory not found. Update results_path above.")
)
```

```{r read-metrics}
# Try parquet first, fall back to CSV
read_safe <- function(dir, stem) {
  pq <- file.path(dir, paste0(stem, ".parquet"))
  csv <- file.path(dir, paste0(stem, ".csv"))
  if (file.exists(pq)) {
    arrow::read_parquet(pq)
  } else if (file.exists(csv)) {
    readr::read_csv(csv, show_col_types = FALSE)
  } else {
    stop("Cannot find ", stem, " in ", dir)
  }
}

model_metrics <- read_safe(results_path, "model_metrics")
```

```{r peek}
glimpse(model_metrics)
```

Depending on the column names in your output, adjust the filtering
below. The key columns are:

-   `prior_spec_id` or `use_case` (identifies susie_vanilla vs
    susie_eb)
-   `p_star`, `y_noise` (simulation regime)
-   `AUPRC` or `auprc` (primary metric)
-   Power, CS metrics as available

```{r identify-columns}
# Auto-detect column naming conventions
col_names <- names(model_metrics)
cat("Available columns:\n")
cat(str_wrap(paste(col_names, collapse = ", "), width = 80), "\n")
```

# Filter to EB vs. Vanilla

```{r filter}
# Adjust these filters based on actual column names in your output.
# This assumes a 'prior_spec_id' or 'use_case_id' column.
use_case_col <- intersect(
  c("prior_spec_id", "use_case_id", "use_case", "use_case_label"),
  col_names
)[1]

if (is.na(use_case_col)) {
  stop("Cannot find use case column. Available: ", paste(col_names, collapse = ", "))
}

cat("Using use case column:", use_case_col, "\n")
cat("Unique values:\n")
print(sort(unique(model_metrics[[use_case_col]])))
```

```{r filter-eb-vanilla}
# Filter to the two use cases of interest
# Adjust the string matching as needed based on your catalog
eb_label <- grep("eb", unique(model_metrics[[use_case_col]]),
                 value = TRUE, ignore.case = TRUE)
vanilla_label <- grep("vanilla", unique(model_metrics[[use_case_col]]),
                      value = TRUE, ignore.case = TRUE)

# Keep only susie-backend (not susine) variants
eb_label <- grep("susie", eb_label, value = TRUE, ignore.case = TRUE)
vanilla_label <- grep("susie", vanilla_label, value = TRUE, ignore.case = TRUE)

# Exclude susine variants (had a bug)
eb_label <- setdiff(eb_label, grep("susine", eb_label, value = TRUE,
                                    ignore.case = TRUE))
vanilla_label <- setdiff(vanilla_label, grep("susine", vanilla_label,
                                              value = TRUE,
                                              ignore.case = TRUE))

cat("EB labels:", paste(eb_label, collapse = ", "), "\n")
cat("Vanilla labels:", paste(vanilla_label, collapse = ", "), "\n")

df <- model_metrics %>%
  filter(.data[[use_case_col]] %in% c(eb_label, vanilla_label)) %>%
  mutate(
    method = case_when(
      .data[[use_case_col]] %in% eb_label ~ "EB",
      .data[[use_case_col]] %in% vanilla_label ~ "Vanilla",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(method))

cat("Rows after filtering:", nrow(df), "\n")
cat("By method:\n")
print(count(df, method))
```

# Paired Comparison Construction

We need to pair each EB run with its vanilla counterpart on the same
dataset. The pairing key should include everything that defines a
dataset + exploration setting *except* the use case.

```{r pair-key}
# Identify the pairing columns -- adjust as needed
# Typical: dataset_bundle_id (or dataset_id) + exploration columns
pair_candidates <- intersect(
  c("dataset_bundle_id", "dataset_id", "matrix_id", "seed",
    "p_star", "y_noise", "exploration_method", "exploration_id",
    "annotation_r2", "inflate_match"),
  col_names
)
cat("Pairing on:", paste(pair_candidates, collapse = ", "), "\n")
```

```{r build-paired}
# Identify the metric columns
metric_cols <- intersect(
  c("AUPRC", "auprc", "power", "fdr", "FDR",
    "mean_purity", "mean_size", "mean_coverage", "cs_power",
    "n_cs", "cross_entropy"),
  col_names
)
cat("Metric columns:", paste(metric_cols, collapse = ", "), "\n")

paired <- df %>%
  select(all_of(c(pair_candidates, "method", metric_cols))) %>%
  pivot_wider(
    id_cols = all_of(pair_candidates),
    names_from = method,
    values_from = all_of(metric_cols),
    names_glue = "{metric_cols}_{method}"
  )

# Compute deltas (EB - Vanilla) for each metric
for (m in metric_cols) {
  eb_col <- paste0(m, "_EB")
  van_col <- paste0(m, "_Vanilla")
  delta_col <- paste0("delta_", m)
  if (eb_col %in% names(paired) && van_col %in% names(paired)) {
    paired[[delta_col]] <- paired[[eb_col]] - paired[[van_col]]
  }
}

cat("Paired dataset rows:", nrow(paired), "\n")
```

# Where Does EB Win?

## Overall Summary

```{r overall-summary}
# Identify delta columns
delta_cols <- grep("^delta_", names(paired), value = TRUE)

summary_tbl <- paired %>%
  summarize(
    across(
      all_of(delta_cols),
      list(
        mean = ~mean(.x, na.rm = TRUE),
        median = ~median(.x, na.rm = TRUE),
        pct_positive = ~mean(.x > 0, na.rm = TRUE)
      ),
      .names = "{.col}__{.fn}"
    )
  ) %>%
  pivot_longer(everything(), names_to = "stat", values_to = "value") %>%
  separate(stat, into = c("metric", "fn"), sep = "__") %>%
  pivot_wider(names_from = fn, values_from = value)

kable(summary_tbl, digits = 4, caption = "EB - Vanilla: overall summary")
```

## AUPRC by Regime

```{r auprc-by-regime, fig.width=10, fig.height=6}
# Find the AUPRC column name
auprc_col <- intersect(c("AUPRC", "auprc"), col_names)[1]
delta_auprc <- paste0("delta_", auprc_col)

if (!is.null(auprc_col) && delta_auprc %in% names(paired)) {

  if ("p_star" %in% names(paired) && "y_noise" %in% names(paired)) {
    p1 <- paired %>%
      ggplot(aes(x = factor(p_star), y = .data[[delta_auprc]])) +
      geom_boxplot(outlier.alpha = 0.3) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
      facet_wrap(~y_noise, labeller = label_both) +
      labs(
        title = "EB advantage (delta AUPRC) by p* and noise level",
        x = "p* (causal count)",
        y = "AUPRC(EB) - AUPRC(Vanilla)"
      )
    print(p1)
  }
}
```

## Heatmap: Mean Delta AUPRC

```{r heatmap-delta, fig.width=8, fig.height=5}
if (!is.null(auprc_col) && delta_auprc %in% names(paired) &&
    "p_star" %in% names(paired) && "y_noise" %in% names(paired)) {

  heat_df <- paired %>%
    group_by(p_star, y_noise) %>%
    summarize(
      mean_delta = mean(.data[[delta_auprc]], na.rm = TRUE),
      n = n(),
      .groups = "drop"
    )

  p2 <- heat_df %>%
    ggplot(aes(
      x = factor(p_star), y = factor(y_noise), fill = mean_delta
    )) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(aes(label = sprintf("%.3f\n(n=%d)", mean_delta, n)),
              size = 3) +
    scale_fill_gradient2(
      midpoint = 0, low = "steelblue", high = "firebrick",
      name = "Mean delta\nAUPRC"
    ) +
    labs(
      title = "Mean EB advantage over Vanilla",
      x = "p* (causal count)", y = "y_noise"
    )
  print(p2)
}
```

## CS Metrics Comparison

```{r cs-metrics, fig.width=10, fig.height=8}
cs_deltas <- intersect(
  c("delta_mean_purity", "delta_mean_size",
    "delta_mean_coverage", "delta_cs_power"),
  names(paired)
)

if (length(cs_deltas) > 0 && "p_star" %in% names(paired)) {
  paired %>%
    select(all_of(c("p_star", "y_noise", cs_deltas))) %>%
    pivot_longer(
      cols = all_of(cs_deltas),
      names_to = "metric", values_to = "delta"
    ) %>%
    mutate(metric = str_remove(metric, "^delta_")) %>%
    ggplot(aes(x = factor(p_star), y = delta)) +
    geom_boxplot(outlier.alpha = 0.3) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    facet_grid(metric ~ y_noise, scales = "free_y",
               labeller = label_both) +
    labs(
      title = "CS metric deltas (EB - Vanilla)",
      x = "p*", y = "Delta"
    ) +
    theme(strip.text.y = element_text(angle = 0))
}
```

# Where Does EB Lose?

```{r eb-loses}
if (!is.null(auprc_col) && delta_auprc %in% names(paired)) {

  losers <- paired %>%
    filter(.data[[delta_auprc]] < -0.01)

  cat("Datasets where EB loses by > 0.01 AUPRC:", nrow(losers),
      "out of", nrow(paired), "\n")

  if (nrow(losers) > 0 && "p_star" %in% names(losers)) {
    cat("\nLosing regime breakdown:\n")
    losers %>%
      count(p_star, y_noise) %>%
      arrange(desc(n)) %>%
      kable(caption = "Regimes where EB underperforms Vanilla")
  }
}
```

# Scatter Diagnostics

## AUPRC: EB vs. Vanilla

```{r scatter-auprc, fig.width=7, fig.height=7}
eb_auprc <- paste0(auprc_col, "_EB")
van_auprc <- paste0(auprc_col, "_Vanilla")

if (eb_auprc %in% names(paired) && van_auprc %in% names(paired)) {

  p3 <- paired %>%
    ggplot(aes(
      x = .data[[van_auprc]],
      y = .data[[eb_auprc]],
      color = factor(p_star)
    )) +
    geom_point(alpha = 0.5, size = 1.5) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    coord_equal() +
    labs(
      title = "AUPRC: EB vs. Vanilla",
      x = "Vanilla AUPRC", y = "EB AUPRC",
      color = "p*"
    )
  print(p3)
}
```

## Delta AUPRC vs. Dataset Difficulty (M1)

If dataset-level M1 metrics are available, join and plot.

```{r delta-vs-m1, fig.width=8, fig.height=5}
# Try to load dataset metrics for M1
dataset_metrics_file <- file.path(results_path, "dataset_metrics.csv")
x_metrics_file <- file.path(results_path, "x_metrics.csv")

m1_available <- FALSE

if (file.exists(dataset_metrics_file)) {
  dataset_metrics <- read_csv(dataset_metrics_file, show_col_types = FALSE)
  m1_col <- intersect(c("M1", "m1", "ld_complexity"), names(dataset_metrics))
  if (length(m1_col) > 0) m1_available <- TRUE
} else if (file.exists(x_metrics_file)) {
  dataset_metrics <- read_csv(x_metrics_file, show_col_types = FALSE)
  m1_col <- intersect(c("M1", "m1", "ld_complexity"), names(dataset_metrics))
  if (length(m1_col) > 0) m1_available <- TRUE
}

if (m1_available) {
  # Join on matrix_id or dataset_bundle_id
  join_col <- intersect(
    c("matrix_id", "dataset_bundle_id"),
    intersect(names(paired), names(dataset_metrics))
  )[1]

  if (!is.na(join_col)) {
    paired_m1 <- paired %>%
      left_join(
        dataset_metrics %>% select(all_of(c(join_col, m1_col[1]))),
        by = join_col
      )

    p4 <- paired_m1 %>%
      ggplot(aes(
        x = .data[[m1_col[1]]],
        y = .data[[delta_auprc]],
        color = factor(p_star)
      )) +
      geom_point(alpha = 0.5) +
      geom_smooth(method = "loess", se = TRUE, color = "black",
                  linewidth = 0.8) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
      labs(
        title = "EB advantage vs. LD complexity (M1)",
        x = "M1 (LD complexity)", y = "delta AUPRC (EB - Vanilla)",
        color = "p*"
      )
    print(p4)
  }
} else {
  cat("M1 metrics not found. Skipping M1 analysis.\n")
}
```

# Theoretical Interpretation

## When Fixed 0.2 is Misspecified

The default `scaled_prior_variance = 0.2` means each causal SNP's
prior effect size has variance $0.2 \times \text{var}(y)$. This is
the "right" value when each causal SNP explains about 20% of
variance.

With L=10 effects and varying p\*:

-   **p\*=1, y_noise=0.5 (h2=0.5)**: 1 causal SNP explaining 50% →
    true per-SNP variance $\approx 0.5 \times \text{var}(y)$.
    Fixed 0.2 is **too small** → alpha too diffuse → power loss.
-   **p\*=20, y_noise=0.95 (h2=0.05)**: 20 causal SNPs sharing 5% →
    per-SNP variance $\approx 0.0025 \times \text{var}(y)$.
    Fixed 0.2 is **way too large** → model over-fits noise.

EB adapts: sets large V for strong effects, small V for weak ones.
The advantage should be largest when the fixed value is most wrong.

## Prediction: EB Advantage Increases with p\*

At high p\*, each effect is weak. Fixed 0.2 over-estimates the per-SNP
variance, making alpha overly concentrated on noise. EB shrinks V
toward the true (small) value, providing better calibration.

At low p\*, each effect is strong. Fixed 0.2 under-estimates the
per-SNP variance, making alpha insufficiently concentrated. EB inflates
V to match the strong signal.

**Net prediction:** EB advantage should be positive across the board,
but largest at the extremes of p\* (where misspecification is worst)
and smallest near p\*=2-3 (where 0.2 happens to be approximately
correct for moderate h2).

## When EB Should Fail

EB should struggle when:

1.  **High M1 (strong LD)**: partial residuals for effect $\ell$ are
    contaminated by LD with effect $\ell'$, inflating V estimates
2.  **Competing effects**: multiple effects fight for the same
    signal, causing V to oscillate
3.  **Multimodal posteriors**: different basins prefer different V
    values; EB locks into one basin's V

**Check against data:** Does delta_AUPRC decrease or reverse at
high M1?

```{r interpretation-check}
if (m1_available && exists("paired_m1")) {
  # Tercile analysis
  paired_m1 <- paired_m1 %>%
    mutate(
      m1_tercile = ntile(.data[[m1_col[1]]], 3) %>%
        factor(labels = c("Low M1", "Med M1", "High M1"))
    )

  paired_m1 %>%
    group_by(m1_tercile) %>%
    summarize(
      mean_delta = mean(.data[[delta_auprc]], na.rm = TRUE),
      median_delta = median(.data[[delta_auprc]], na.rm = TRUE),
      pct_eb_wins = mean(.data[[delta_auprc]] > 0, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    ) %>%
    kable(digits = 4,
          caption = "EB advantage by M1 tercile")
}
```

# Conclusions

**Fill in after running the workbook on actual data.**

Key questions to answer:

1.  Does EB win across the board, or only in specific regimes?
2.  Is the advantage largest at extreme p\* (where fixed 0.2 is most
    misspecified)?
3.  Does EB fail at high M1 (LD contamination)?
4.  What are the implications for Phase B ensemble design?
    -   Should sigma_0^2-grid ensembles use EB as baseline, or fixed?
    -   If EB is reliable, the ensemble's job is only multimodality
        resolution (random restarts), not hyperparameter search.

```{r session-info}
sessionInfo()
```
