---
title: "Effect Matching Refit Export"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Standalone one-off workbook for the fitted-y effect-matching study. This
does not read previous baseline simulation outputs. It recreates the
same 150 matrix x 4 seed baseline-sims dataset grid, then fits exactly:

- `susie_vanilla`
- `susine_functional_mu` with `annotation_r2 = 0.3`,
  `inflate_match = 0.9`, `c = 0.6`, `sigma_0_2 = 0.01`

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_refit_export.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(knitr)
```

# 1. User Parameters

```{r user-parameters}
study_name <- "baseline_sims_standalone_effect_matching"
output_root <- here("output")
study_dir <- file.path(output_root, "effect_matching_study", study_name)
dir.create(study_dir, recursive = TRUE, showWarnings = FALSE)

# Set to a small integer, e.g. 5L, for a local smoke run.
max_datasets <- Inf
force_refit <- FALSE
checkpoint_every <- 25L

# Baseline-sims dataset recipe.
n_matrices_target <- 150L
seeds <- 1:4
L_value <- 10L
y_noise_value <- NA_real_
p_star_value <- NA_integer_
architecture_value <- "susie2_oligogenic"
data_scenario <- "scenario_1"

# Matching-study model settings.
annotation_r2_value <- 0.3
inflate_match_value <- 0.9
c_value_mu <- 0.6
sigma_0_2_value <- 0.01
max_iter <- 100L
credible_set_rho <- 0.95
purity_threshold <- 0.50

records_rds <- file.path(study_dir, "effect_refit_records.rds")
metadata_csv <- file.path(study_dir, "effect_refit_metadata.csv")
selected_runs_csv <- file.path(study_dir, "selected_run_rows.csv")
dataset_bundles_csv <- file.path(study_dir, "dataset_bundles.csv")
progress_csv <- file.path(study_dir, "refit_progress_log.csv")

cat("Study dir:", study_dir, "\n")
```

# 2. Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

format_duration <- function(seconds) {
  seconds <- as.numeric(seconds)
  if (!is.finite(seconds)) {
    return("unknown")
  }
  h <- floor(seconds / 3600)
  m <- floor((seconds %% 3600) / 60)
  s <- round(seconds %% 60)
  sprintf("%02d:%02d:%02d", h, m, s)
}

save_checkpoint <- function(records, progress_rows) {
  saveRDS(records, records_rds)

  metadata <- records
  if ("indices" %in% names(metadata)) {
    metadata <- metadata %>%
      dplyr::mutate(
        cs_indices = vapply(
          .data$indices,
          function(idx) paste(as.integer(idx), collapse = ";"),
          character(1)
        )
      )
  }
  metadata <- metadata %>%
    dplyr::select(-dplyr::where(is.list))
  readr::write_csv(metadata, metadata_csv)

  if (length(progress_rows)) {
    readr::write_csv(dplyr::bind_rows(progress_rows), progress_csv)
  }
}

first_existing_numeric <- function(df, cols, default = NA_real_) {
  for (nm in cols) {
    if (nm %in% names(df)) {
      val <- suppressWarnings(as.numeric(df[[nm]][[1]]))
      if (length(val)) {
        return(val)
      }
    }
  }
  default
}

effect_coef_matrix <- function(fit, X) {
  alpha <- fit$effect_fits$alpha
  if (is.list(alpha)) {
    alpha <- do.call(rbind, alpha)
  }
  alpha <- as.matrix(alpha)
  L <- nrow(alpha)
  p <- ncol(X)

  b_hat <- fit$effect_fits$b_hat %||% NULL
  if (!is.null(b_hat)) {
    b_hat <- as.matrix(b_hat)
    if (nrow(b_hat) == p && ncol(b_hat) == L) {
      b_hat <- t(b_hat)
    }
    if (nrow(b_hat) == L && ncol(b_hat) == p) {
      return(b_hat)
    }
  }

  mu <- fit$effect_fits$mu %||% fit$raw$mu %||% NULL
  if (!is.null(mu)) {
    mu <- as.matrix(mu)
    if (nrow(mu) == p && ncol(mu) == L) {
      mu <- t(mu)
    }
    if (nrow(mu) == L && ncol(mu) == p) {
      return(alpha * mu)
    }
  }

  stop("Could not recover an L x p per-effect coefficient matrix from the fit.")
}

extract_effect_records <- function(fit, eval_res, data_bundle, run_row, model_method, fit_meta) {
  X <- data_bundle$X
  y <- as.numeric(data_bundle$y)
  b_hat <- effect_coef_matrix(fit, X)
  fitted_y_mat <- X %*% t(b_hat)
  L <- ncol(fitted_y_mat)
  var_y <- stats::var(y)

  effect_tbl <- eval_res$effects_unfiltered %>%
    tibble::as_tibble() %>%
    dplyr::mutate(effect = as.integer(.data$effect))

  tibble::tibble(effect = seq_len(L)) %>%
    dplyr::left_join(effect_tbl, by = "effect") %>%
    dplyr::mutate(
      fitted_y_effect = lapply(seq_len(L), function(l) as.numeric(fitted_y_mat[, l])),
      var_y_explained_frac = vapply(
        .data$fitted_y_effect,
        function(fy) {
          val <- stats::var(fy) / var_y
          ifelse(is.finite(val), val, NA_real_)
        },
        numeric(1)
      ),
      var_y_explained_pct = 100 * .data$var_y_explained_frac,
      dataset_bundle_id = as.character(run_row$dataset_bundle_id[[1]]),
      run_id = as.integer(run_row$run_id[[1]]),
      model_method = model_method,
      use_case_id = as.character(run_row$use_case_id[[1]]),
      spec_name = as.character(run_row$spec_name[[1]]),
      setting_label = as.character(run_row$setting_label[[1]]),
      annotation_label = as.character(run_row$annotation_label[[1]]),
      annotation_r2 = as.numeric(run_row$annotation_r2[[1]]),
      inflate_match = as.numeric(run_row$inflate_match[[1]]),
      c_value = as.numeric(run_row$c_value[[1]]),
      tau_value = as.numeric(run_row$tau_value[[1]]),
      sigma_0_2_scalar = as.character(run_row$sigma_0_2_scalar[[1]]),
      phenotype_seed = as.integer(run_row$phenotype_seed[[1]]),
      matrix_id = as.character(run_row$matrix_id[[1]]),
      architecture = as.character(run_row$architecture[[1]]),
      p_star = as.integer(run_row$p_star[[1]]),
      y_noise = as.numeric(run_row$y_noise[[1]]),
      fit_wall_time_sec = as.numeric(fit_meta$wall_time_sec %||% NA_real_),
      model_AUPRC = as.numeric(eval_res$model_unfiltered$AUPRC[[1]]),
      model_tpr_fpr05 = first_existing_numeric(
        eval_res$model_unfiltered,
        c("tpr_fpr05", "tpr_at_fpr_05")
      )
    ) %>%
    dplyr::relocate(
      dplyr::all_of(c(
        "dataset_bundle_id", "model_method", "run_id",
        "effect", "setting_label", "annotation_label"
      ))
    )
}

fit_one_method <- function(run_row, data_bundle, job_config) {
  run_row <- tibble::as_tibble(run_row)
  use_case <- test_susine:::lookup_use_case(job_config, run_row$use_case_id[[1]])
  if (!nrow(use_case)) {
    stop("Unknown use case: ", run_row$use_case_id[[1]])
  }

  model_result <- test_susine:::run_use_case(
    use_case = use_case,
    run_row = run_row,
    data_bundle = data_bundle,
    job_config = job_config
  )
  fit <- model_result$fits[[1]]
  fit_meta <- model_result$fit_meta[[1]] %||% list()

  eval_res <- test_susine:::evaluate_model(
    fit = fit,
    X = data_bundle$X,
    y = data_bundle$y,
    causal_idx = data_bundle$causal_idx,
    rho = job_config$job$credible_set_rho %||% credible_set_rho,
    purity_threshold = job_config$job$purity_threshold %||% purity_threshold,
    compute_curves = FALSE
  )

  extract_effect_records(
    fit = fit,
    eval_res = eval_res,
    data_bundle = data_bundle,
    run_row = run_row,
    model_method = run_row$method_family[[1]],
    fit_meta = fit_meta
  )
}
```

# 3. Build Standalone Baseline Dataset Grid

```{r build-standalone-config}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = data_scenario,
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)

m1_summary_path <- here(
  "data", "sampled_simulated_genotypes", "scenario_sampling_summary_with_m1.csv"
)
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>%
      dplyr::select(.data$matrix_id, .data$matrix_path),
    base_dir = here()
  ) %>%
    dplyr::select(.data$matrix_id, .data$M1)
  catalog_with_m1 <- data_matrix_catalog_full %>%
    dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}

if (!"matrix_id" %in% names(catalog_with_m1)) {
  if ("gene" %in% names(catalog_with_m1)) {
    catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = .data$gene)
  }
  join_col <- intersect(c("dataset_label", "matrix_path"), names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(
      data_matrix_catalog_full %>%
        dplyr::select(.data$matrix_id, dplyr::all_of(join_col)),
      by = join_col
    )
}

catalog_with_m1 <- catalog_with_m1 %>%
  dplyr::filter(!is.na(.data$M1)) %>%
  dplyr::arrange(.data$M1)

n_target <- min(n_matrices_target, nrow(catalog_with_m1))
idx <- unique(round(seq(1, nrow(catalog_with_m1), length.out = n_target)))
selected_matrices <- catalog_with_m1[idx, , drop = FALSE]

data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(
    selected_matrices %>% dplyr::select(.data$matrix_id),
    by = "matrix_id"
  )

spec_susie_vanilla <- list(
  name = "effect_match_susie_vanilla",
  use_case_ids = "susie_vanilla",
  exploration_methods = "sigma_0_2_grid",
  exploration_mode = "separate",
  K = 1L,
  sigma_0_2_grid_values = sigma_0_2_value
)

spec_susine_functional_mu <- list(
  name = "effect_match_susine_functional_mu",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 1L,
  c_grid_values = c_value_mu,
  sigma_0_2_grid_values = sigma_0_2_value
)

cfg <- test_susine::make_job_config(
  job_name = study_name,
  HPC = FALSE,
  time = "02:00:00",
  mem = "8G",
  output_root = output_root,
  grid_mode = "full",
  use_case_ids = c("susie_vanilla", "susine_functional_mu"),
  exploration_specs = list(spec_susie_vanilla, spec_susine_functional_mu),
  L_grid = L_value,
  y_noise_grid = y_noise_value,
  p_star_grid = p_star_value,
  architecture_grid = architecture_value,
  seeds = seeds,
  prior_quality = tibble::tibble(
    annotation_r2 = annotation_r2_value,
    inflate_match = inflate_match_value
  ),
  sigma_0_2_scalars = sigma_0_2_value,
  data_scenarios = data_scenario,
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),
  task_unit = "dataset",
  bundles_per_task = 1L,
  task_assignment_seed = 42L,
  max_iter = max_iter,
  aggregation_methods = character(0),
  include_overall_pool = FALSE,
  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = FALSE,
  write_snps_parquet = FALSE,
  write_confusion_bins = FALSE,
  write_tier_cs_metrics = TRUE,
  write_scaling_confusion_bins = FALSE,
  metrics_settings = list(
    z_top_k = 10L,
    jsd_threshold = 0.15
  )
)

run_table <- cfg$tables$runs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
    method_family = .data$use_case_id,
    annotation_label = dplyr::if_else(
      is.na(.data$annotation_r2) | is.na(.data$inflate_match),
      "no_annotation",
      paste0("r2=", .data$annotation_r2, " | inflate=", .data$inflate_match)
    ),
    setting_label = dplyr::case_when(
      .data$use_case_id == "susie_vanilla" ~
        paste0("susie_vanilla | sigma=", .data$sigma_0_2_scalar),
      .data$use_case_id == "susine_functional_mu" ~
        paste0(
          "susine_functional_mu | c=", .data$c_value,
          " | sigma=", .data$sigma_0_2_scalar
        ),
      TRUE ~ .data$use_case_id
    )
  )
cfg$tables$runs <- run_table

selected_runs <- run_table %>%
  dplyr::filter(.data$use_case_id %in% c("susie_vanilla", "susine_functional_mu")) %>%
  dplyr::arrange(.data$dataset_bundle_id, .data$use_case_id)

readr::write_csv(selected_runs, selected_runs_csv)
readr::write_csv(cfg$tables$dataset_bundles, dataset_bundles_csv)

cat("Matrices selected:", nrow(data_matrix_catalog), "\n")
cat("Dataset bundles:", nrow(cfg$tables$dataset_bundles), "\n")
cat("Selected run rows:", nrow(selected_runs), "\n")
cat("Methods:", paste(sort(unique(selected_runs$use_case_id)), collapse = ", "), "\n")
cat("Functional-mu annotation_r2:", annotation_r2_value, "\n")
cat("Functional-mu inflate_match:", inflate_match_value, "\n")
cat("Functional-mu c:", c_value_mu, "\n")
cat("Sigma_0_2 scalar:", sigma_0_2_value, "\n")

stopifnot(nrow(cfg$tables$dataset_bundles) == n_matrices_target * length(seeds))
stopifnot(nrow(selected_runs) == 2L * nrow(cfg$tables$dataset_bundles))
```

# 4. Refit And Export

```{r refit-export}
dataset_ids <- sort(as.character(unique(selected_runs$dataset_bundle_id)))
if (is.finite(max_datasets)) {
  dataset_ids <- head(dataset_ids, as.integer(max_datasets))
}

existing_records <- tibble::tibble()
if (!force_refit && file.exists(records_rds)) {
  existing_records <- readRDS(records_rds)
  cat("Loaded existing records:", nrow(existing_records), "\n")
}

completed_ids <- character(0)
if (nrow(existing_records)) {
  completed_ids <- existing_records %>%
    dplyr::count(.data$dataset_bundle_id, .data$model_method, name = "n_effects") %>%
    dplyr::filter(.data$n_effects > 0L) %>%
    dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
    dplyr::filter(.data$n_methods == 2L) %>%
    dplyr::pull(.data$dataset_bundle_id)
}

todo_ids <- setdiff(dataset_ids, completed_ids)
cat("Datasets queued:", length(dataset_ids), "\n")
cat("Already completed datasets:", length(intersect(dataset_ids, completed_ids)), "\n")
cat("Datasets remaining:", length(todo_ids), "\n")

records <- existing_records
progress_rows <- list()
if (file.exists(progress_csv)) {
  old_progress <- readr::read_csv(progress_csv, show_col_types = FALSE)
  progress_rows <- split(old_progress, seq_len(nrow(old_progress)))
}

started <- proc.time()[["elapsed"]]
total_to_run <- length(todo_ids)

for (i in seq_along(todo_ids)) {
  dataset_id <- todo_ids[[i]]
  dataset_started <- proc.time()[["elapsed"]]

  bundle_row <- cfg$tables$dataset_bundles %>%
    tibble::as_tibble() %>%
    dplyr::filter(as.character(.data$dataset_bundle_id) == !!dataset_id)
  if (nrow(bundle_row) != 1L) {
    stop("Expected one dataset bundle row for dataset_bundle_id=", dataset_id, "; found ", nrow(bundle_row), ".")
  }

  data_bundle <- test_susine:::generate_data_for_bundle(bundle_row, cfg)

  dataset_run_rows <- selected_runs %>%
    dplyr::filter(as.character(.data$dataset_bundle_id) == !!dataset_id) %>%
    dplyr::arrange(factor(.data$method_family, levels = c("susie_vanilla", "susine_functional_mu")))

  dataset_records <- purrr::map_dfr(
    seq_len(nrow(dataset_run_rows)),
    function(j) {
      fit_one_method(dataset_run_rows[j, , drop = FALSE], data_bundle, cfg)
    }
  )

  records <- dplyr::bind_rows(records, dataset_records)

  elapsed <- proc.time()[["elapsed"]] - started
  avg_sec <- elapsed / i
  eta <- avg_sec * (total_to_run - i)
  dataset_elapsed <- proc.time()[["elapsed"]] - dataset_started

  progress_row <- tibble::tibble(
    completed_this_session = i,
    total_this_session = total_to_run,
    completed_overall = length(intersect(dataset_ids, completed_ids)) + i,
    total_overall = length(dataset_ids),
    dataset_bundle_id = dataset_id,
    dataset_elapsed_sec = dataset_elapsed,
    elapsed_sec = elapsed,
    eta_sec = eta,
    timestamp = as.character(Sys.time())
  )
  progress_rows[[length(progress_rows) + 1L]] <- progress_row

  cat(sprintf(
    "%d/%d datasets fitted | elapsed: %s | avg: %.1f sec/dataset | ETA: %s | dataset_id: %s\n",
    length(intersect(dataset_ids, completed_ids)) + i,
    length(dataset_ids),
    format_duration(elapsed),
    avg_sec,
    format_duration(eta),
    dataset_id
  ))

  if (i %% checkpoint_every == 0L || i == total_to_run) {
    save_checkpoint(records, progress_rows)
    cat("Checkpoint saved:", records_rds, "\n")
  }
}

if (!length(todo_ids)) {
  save_checkpoint(records, progress_rows)
}

cat("Final record rows:", nrow(records), "\n")
cat("Final datasets:", dplyr::n_distinct(records$dataset_bundle_id), "\n")
cat("Records RDS:", records_rds, "\n")
cat("Metadata CSV:", metadata_csv, "\n")
```

# 5. Sanity Checks

```{r sanity}
records_check <- readRDS(records_rds)

summary_counts <- records_check %>%
  dplyr::count(.data$dataset_bundle_id, .data$model_method, name = "n_effects") %>%
  dplyr::count(.data$model_method, .data$n_effects, name = "n_datasets")

kable(summary_counts)

vector_lengths <- records_check %>%
  dplyr::mutate(fitted_y_length = vapply(.data$fitted_y_effect, length, integer(1))) %>%
  dplyr::count(.data$model_method, .data$fitted_y_length, name = "n_effect_rows")

kable(vector_lengths)

if (any(vector_lengths$fitted_y_length <= 0L)) {
  stop("At least one fitted_y_effect vector is empty.")
}
```
