---
title: "Effect Matching Threshold Calibration"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

One-off workbook for calibrating fitted-y effect matching rules. It reads
the refit export, compares true same-dataset `susie_vanilla` versus
`susine_functional_mu` effect pairs, builds a deranged null distribution,
and finds threshold tuples that keep null matches at or below 1%.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_threshold_calibration.Rmd")
library(here)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

# 1. Inputs

```{r inputs}
output_root <- here("output")

study_name <- "baseline_sims_standalone_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
records_rds <- file.path(study_dir, "effect_refit_records.rds")

n_null_derangements <- 50L
null_seed <- 20260419L

corr_min_grid <- seq(0.50, 0.95, by = 0.05)
margin_min_grid <- c(0, 0.025, 0.05, 0.075, 0.10, 0.125, 0.15, 0.20, 0.25, 0.30)
var_min_grid <- c(0, 0.001, 0.0025, 0.005, 0.0075, 0.01, 0.015, 0.02, 0.03, 0.05)

near_zero_sd <- 1e-12
null_rate_cutoff <- 0.01

real_candidates_csv <- file.path(study_dir, "real_match_candidates.csv")
null_candidates_csv <- file.path(study_dir, "null_match_candidates.csv")
threshold_grid_csv <- file.path(study_dir, "threshold_grid_results.csv")
frontier_csv <- file.path(study_dir, "frontier_1pct_null.csv")
winner_csv <- file.path(study_dir, "winner_threshold.csv")
accepted_real_csv <- file.path(study_dir, "accepted_real_matches_at_winner.csv")
accepted_delta_csv <- file.path(study_dir, "accepted_real_match_delta_summary.csv")
frontier_plot_png <- file.path(study_dir, "frontier_real_vs_null.png")

cat("Study dir:", study_dir, "\n")
cat("Records RDS:", records_rds, "\n")
cat("Null derangements:", n_null_derangements, "\n")
```

# 2. Helpers

```{r helpers}
format_duration <- function(seconds) {
  seconds <- as.numeric(seconds)
  if (!is.finite(seconds)) {
    return("unknown")
  }
  h <- floor(seconds / 3600)
  m <- floor((seconds %% 3600) / 60)
  s <- round(seconds %% 60)
  sprintf("%02d:%02d:%02d", h, m, s)
}

make_derangement <- function(ids) {
  if (length(ids) < 2L) {
    stop("Need at least two datasets to create a derangement.")
  }
  repeat {
    perm <- sample(ids, length(ids), replace = FALSE)
    if (!any(perm == ids)) {
      return(perm)
    }
  }
}

records_to_matrix <- function(rows) {
  rows <- rows %>% dplyr::arrange(.data$effect)
  mat <- do.call(cbind, rows$fitted_y_effect)
  storage.mode(mat) <- "double"
  colnames(mat) <- as.character(rows$effect)
  mat
}

safe_cor_matrix <- function(vanilla_mat, mu_mat, sd_eps = near_zero_sd) {
  v_sd <- apply(vanilla_mat, 2L, stats::sd)
  m_sd <- apply(mu_mat, 2L, stats::sd)
  active_v <- is.finite(v_sd) & v_sd > sd_eps
  active_m <- is.finite(m_sd) & m_sd > sd_eps

  cor_mat <- matrix(
    0,
    nrow = ncol(vanilla_mat),
    ncol = ncol(mu_mat),
    dimnames = list(colnames(vanilla_mat), colnames(mu_mat))
  )

  if (any(active_v) && any(active_m)) {
    sub_cor <- suppressWarnings(stats::cor(
      vanilla_mat[, active_v, drop = FALSE],
      mu_mat[, active_m, drop = FALSE],
      use = "pairwise.complete.obs"
    ))
    sub_cor[!is.finite(sub_cor)] <- 0
    cor_mat[active_v, active_m] <- sub_cor
  }

  list(cor = cor_mat, active_vanilla = active_v, active_mu = active_m)
}

second_best_margin <- function(values, chosen_idx) {
  if (length(values) <= 1L) {
    return(Inf)
  }
  alternatives <- values[-chosen_idx]
  best_alt <- max(alternatives, na.rm = TRUE)
  values[[chosen_idx]] - best_alt
}

safe_min_numeric <- function(...) {
  x <- c(...)
  x <- x[is.finite(x)]
  if (!length(x)) {
    return(NA_real_)
  }
  min(x)
}

effect_metric_cols <- c(
  "size", "purity", "coverage",
  "effect_pip_entropy", "effect_pip_entropy_core95",
  "effect_k_eff_signal", "effect_k_eff_signal_core95",
  "tail_inflation_ratio", "tail_inflation_log",
  "accuracy_ratio", "var_y_explained_frac", "var_y_explained_pct"
)

add_effect_metric_columns <- function(out, vanilla_rows, mu_rows, vanilla_idx, mu_idx) {
  available <- intersect(effect_metric_cols, intersect(names(vanilla_rows), names(mu_rows)))
  for (nm in available) {
    out[[paste0(nm, "_vanilla")]] <- vanilla_rows[[nm]][vanilla_idx]
    out[[paste0(nm, "_mu")]] <- mu_rows[[nm]][mu_idx]
    out[[paste0(nm, "_delta_mu_minus_vanilla")]] <-
      mu_rows[[nm]][mu_idx] - vanilla_rows[[nm]][vanilla_idx]
  }
  out
}

compare_dataset_pair <- function(vanilla_rows,
                                 mu_rows,
                                 vanilla_dataset_id,
                                 mu_dataset_id,
                                 comparison_type,
                                 comparison_id,
                                 null_replicate = NA_integer_,
                                 include_effect_metrics = FALSE) {
  vanilla_rows <- vanilla_rows %>% dplyr::arrange(.data$effect)
  mu_rows <- mu_rows %>% dplyr::arrange(.data$effect)

  vanilla_mat <- records_to_matrix(vanilla_rows)
  mu_mat <- records_to_matrix(mu_rows)
  cor_info <- safe_cor_matrix(vanilla_mat, mu_mat)
  cor_mat <- cor_info$cor

  row_best <- max.col(cor_mat, ties.method = "first")
  col_best <- max.col(t(cor_mat), ties.method = "first")

  candidate_rows <- vector("list", nrow(cor_mat))
  n_candidates <- 0L

  for (i in seq_len(nrow(cor_mat))) {
    j <- row_best[[i]]
    if (col_best[[j]] != i) {
      next
    }

    n_candidates <- n_candidates + 1L
    row_margin <- second_best_margin(cor_mat[i, ], j)
    col_margin <- second_best_margin(cor_mat[, j], i)

    out <- tibble::tibble(
      comparison_type = comparison_type,
      comparison_id = comparison_id,
      null_replicate = as.integer(null_replicate),
      vanilla_dataset_bundle_id = vanilla_dataset_id,
      mu_dataset_bundle_id = mu_dataset_id,
      vanilla_effect = as.integer(rownames(cor_mat)[[i]]),
      mu_effect = as.integer(colnames(cor_mat)[[j]]),
      corr = as.numeric(cor_mat[i, j]),
      row_margin = as.numeric(row_margin),
      col_margin = as.numeric(col_margin),
      margin = min(row_margin, col_margin),
      var_y_explained_frac_vanilla = as.numeric(vanilla_rows$var_y_explained_frac[[i]]),
      var_y_explained_frac_mu = as.numeric(mu_rows$var_y_explained_frac[[j]]),
      pair_var_explained_frac = safe_min_numeric(
        as.numeric(vanilla_rows$var_y_explained_frac[[i]]),
        as.numeric(mu_rows$var_y_explained_frac[[j]])
      ),
      active_vanilla = as.logical(cor_info$active_vanilla[[i]]),
      active_mu = as.logical(cor_info$active_mu[[j]]),
      mutual_best = TRUE
    )

    if (include_effect_metrics) {
      out <- add_effect_metric_columns(out, vanilla_rows, mu_rows, i, j)
    }

    candidate_rows[[n_candidates]] <- out
  }

  if (!n_candidates) {
    return(tibble::tibble())
  }

  dplyr::bind_rows(candidate_rows[seq_len(n_candidates)])
}

score_threshold_grid <- function(candidates, total_comparisons, grid, label) {
  if (!nrow(candidates)) {
    return(grid %>%
      dplyr::mutate(
        "{label}_matches_total" := 0L,
        "{label}_matches_per_dataset_pair" := 0,
        "{label}_any_match_rate" := 0
      ))
  }

  corr <- candidates$corr
  margin <- candidates$margin
  var_frac <- candidates$pair_var_explained_frac
  comparison_id <- candidates$comparison_id

  started <- proc.time()[["elapsed"]]
  rows <- vector("list", nrow(grid))

  for (i in seq_len(nrow(grid))) {
    keep <- corr >= grid$corr_min[[i]] &
      margin >= grid$margin_min[[i]] &
      var_frac >= grid$var_min[[i]]
    keep[is.na(keep)] <- FALSE

    n_matches <- sum(keep, na.rm = TRUE)
    any_rate <- if (n_matches > 0L) {
      length(unique(comparison_id[keep])) / total_comparisons
    } else {
      0
    }

    rows[[i]] <- tibble::tibble(
      corr_min = grid$corr_min[[i]],
      margin_min = grid$margin_min[[i]],
      var_min = grid$var_min[[i]],
      matches_total = n_matches,
      matches_per_dataset_pair = n_matches / total_comparisons,
      any_match_rate = any_rate
    )

    if (i %% 100L == 0L || i == nrow(grid)) {
      elapsed <- proc.time()[["elapsed"]] - started
      eta <- elapsed / i * (nrow(grid) - i)
      cat(sprintf(
        "%s threshold scoring %d/%d | elapsed: %s | ETA: %s\n",
        label,
        i,
        nrow(grid),
        format_duration(elapsed),
        format_duration(eta)
      ))
    }
  }

  dplyr::bind_rows(rows) %>%
    dplyr::rename_with(
      .fn = ~ paste0(label, "_", .x),
      .cols = c("matches_total", "matches_per_dataset_pair", "any_match_rate")
    )
}
```

# 3. Load Records

```{r load-records}
stopifnot(file.exists(records_rds))

records <- readRDS(records_rds) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

needed_methods <- c("susie_vanilla", "susine_functional_mu")
method_counts <- records %>%
  dplyr::filter(.data$model_method %in% needed_methods) %>%
  dplyr::count(.data$dataset_bundle_id, .data$model_method, name = "n_effects")

complete_ids <- method_counts %>%
  dplyr::filter(.data$n_effects > 0L) %>%
  dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
  dplyr::filter(.data$n_methods == length(needed_methods)) %>%
  dplyr::pull(.data$dataset_bundle_id) %>%
  sort()

if (length(complete_ids) < 2L) {
  stop("Need at least two complete datasets for real/null matching; found ", length(complete_ids), ".")
}

records <- records %>%
  dplyr::filter(.data$dataset_bundle_id %in% complete_ids, .data$model_method %in% needed_methods)

vanilla_by_dataset <- records %>%
  dplyr::filter(.data$model_method == "susie_vanilla") %>%
  split(.$dataset_bundle_id)

mu_by_dataset <- records %>%
  dplyr::filter(.data$model_method == "susine_functional_mu") %>%
  split(.$dataset_bundle_id)

cat("Complete datasets:", length(complete_ids), "\n")
kable(method_counts %>% dplyr::count(.data$model_method, .data$n_effects, name = "n_datasets"))
```

# 4. Real Match Candidates

```{r real-candidates}
real_started <- proc.time()[["elapsed"]]

real_candidates <- purrr::map_dfr(
  seq_along(complete_ids),
  function(i) {
    dataset_id <- complete_ids[[i]]
    out <- compare_dataset_pair(
      vanilla_rows = vanilla_by_dataset[[dataset_id]],
      mu_rows = mu_by_dataset[[dataset_id]],
      vanilla_dataset_id = dataset_id,
      mu_dataset_id = dataset_id,
      comparison_type = "real",
      comparison_id = paste0("real_", dataset_id),
      null_replicate = NA_integer_,
      include_effect_metrics = TRUE
    )

    if (i %% 50L == 0L || i == length(complete_ids)) {
      elapsed <- proc.time()[["elapsed"]] - real_started
      eta <- elapsed / i * (length(complete_ids) - i)
      cat(sprintf(
        "Real candidates %d/%d | elapsed: %s | ETA: %s\n",
        i,
        length(complete_ids),
        format_duration(elapsed),
        format_duration(eta)
      ))
    }

    out
  }
)

readr::write_csv(real_candidates, real_candidates_csv)

cat("Real candidate rows:", nrow(real_candidates), "\n")
cat("Real comparisons:", length(complete_ids), "\n")
cat("Saved:", real_candidates_csv, "\n")
```

# 5. Null Match Candidates

```{r null-candidates}
set.seed(null_seed)
null_started <- proc.time()[["elapsed"]]
null_candidate_chunks <- vector("list", n_null_derangements)
derangement_checks <- vector("list", n_null_derangements)

for (rep_i in seq_len(n_null_derangements)) {
  mu_ids <- make_derangement(complete_ids)

  derangement_checks[[rep_i]] <- tibble::tibble(
    null_replicate = rep_i,
    n_pairs = length(complete_ids),
    n_fixed_points = sum(mu_ids == complete_ids)
  )

  null_candidate_chunks[[rep_i]] <- purrr::map2_dfr(
    complete_ids,
    mu_ids,
    function(vanilla_id, mu_id) {
      compare_dataset_pair(
        vanilla_rows = vanilla_by_dataset[[vanilla_id]],
        mu_rows = mu_by_dataset[[mu_id]],
        vanilla_dataset_id = vanilla_id,
        mu_dataset_id = mu_id,
        comparison_type = "null",
        comparison_id = paste0("null_", rep_i, "_", vanilla_id, "_", mu_id),
        null_replicate = rep_i,
        include_effect_metrics = FALSE
      )
    }
  )

  elapsed <- proc.time()[["elapsed"]] - null_started
  eta <- elapsed / rep_i * (n_null_derangements - rep_i)
  cat(sprintf(
    "Null derangement %d/%d | elapsed: %s | ETA: %s | candidate rows so far: %d\n",
    rep_i,
    n_null_derangements,
    format_duration(elapsed),
    format_duration(eta),
    sum(vapply(null_candidate_chunks[seq_len(rep_i)], nrow, integer(1)))
  ))
}

derangement_check <- dplyr::bind_rows(derangement_checks)
if (any(derangement_check$n_fixed_points != 0L)) {
  print(derangement_check)
  stop("At least one null derangement contained a true dataset pair.")
}

null_candidates <- dplyr::bind_rows(null_candidate_chunks)
readr::write_csv(null_candidates, null_candidates_csv)

cat("Null candidate rows:", nrow(null_candidates), "\n")
cat("Null comparisons:", n_null_derangements * length(complete_ids), "\n")
cat("Saved:", null_candidates_csv, "\n")
```

# 6. Threshold Grid

```{r threshold-grid}
threshold_grid <- tidyr::expand_grid(
  corr_min = corr_min_grid,
  margin_min = margin_min_grid,
  var_min = var_min_grid
)

real_scores <- score_threshold_grid(
  candidates = real_candidates,
  total_comparisons = length(complete_ids),
  grid = threshold_grid,
  label = "real"
)

null_scores <- score_threshold_grid(
  candidates = null_candidates,
  total_comparisons = n_null_derangements * length(complete_ids),
  grid = threshold_grid,
  label = "null"
)

threshold_results <- threshold_grid %>%
  dplyr::left_join(real_scores, by = c("corr_min", "margin_min", "var_min")) %>%
  dplyr::left_join(null_scores, by = c("corr_min", "margin_min", "var_min")) %>%
  dplyr::mutate(
    passes_1pct_null = .data$null_matches_per_dataset_pair <= null_rate_cutoff
  ) %>%
  dplyr::arrange(
    .data$passes_1pct_null,
    dplyr::desc(.data$real_matches_total),
    .data$null_matches_per_dataset_pair,
    dplyr::desc(.data$corr_min),
    dplyr::desc(.data$margin_min),
    .by_group = FALSE
  )

frontier <- threshold_results %>%
  dplyr::filter(.data$passes_1pct_null) %>%
  dplyr::arrange(
    dplyr::desc(.data$real_matches_total),
    .data$null_matches_per_dataset_pair,
    dplyr::desc(.data$corr_min),
    dplyr::desc(.data$margin_min)
  )

if (!nrow(frontier)) {
  stop("No threshold tuple satisfied the <=1% null match criterion.")
}

winner <- frontier %>%
  dplyr::slice(1L)

readr::write_csv(threshold_results, threshold_grid_csv)
readr::write_csv(frontier, frontier_csv)
readr::write_csv(winner, winner_csv)

cat("Threshold grid rows:", nrow(threshold_results), "\n")
cat("1% null frontier rows:", nrow(frontier), "\n")
cat("Winner:\n")
print(winner)
```

# 7. Accepted Real Matches At Winner

```{r winner-matches}
accepted_real <- real_candidates %>%
  dplyr::filter(
    .data$corr >= winner$corr_min[[1]],
    .data$margin >= winner$margin_min[[1]],
    .data$pair_var_explained_frac >= winner$var_min[[1]]
  ) %>%
  dplyr::arrange(.data$vanilla_dataset_bundle_id, dplyr::desc(.data$corr))

readr::write_csv(accepted_real, accepted_real_csv)

cat("Selected threshold tuple:\n")
cat("  corr_min:", winner$corr_min[[1]], "\n")
cat("  margin_min:", winner$margin_min[[1]], "\n")
cat("  var_min:", winner$var_min[[1]], "\n")
cat("  null matches per dataset pair:", winner$null_matches_per_dataset_pair[[1]], "\n")
cat("  null any-match rate:", winner$null_any_match_rate[[1]], "\n")
cat("  real matches:", winner$real_matches_total[[1]], "\n")
cat("  real matches per dataset pair:", winner$real_matches_per_dataset_pair[[1]], "\n")

matches_per_dataset <- accepted_real %>%
  dplyr::count(.data$vanilla_dataset_bundle_id, name = "n_matches")

cat("\nReal matches per dataset summary:\n")
print(summary(matches_per_dataset$n_matches))

diagnostic_delta_cols <- grep("_delta_mu_minus_vanilla$", names(accepted_real), value = TRUE)
delta_summary <- purrr::map_dfr(
  diagnostic_delta_cols,
  function(nm) {
    x <- accepted_real[[nm]]
    tibble::tibble(
      metric = sub("_delta_mu_minus_vanilla$", "", nm),
      n = sum(is.finite(x)),
      median_delta = stats::median(x, na.rm = TRUE),
      q05_delta = as.numeric(stats::quantile(x, 0.05, na.rm = TRUE)),
      q25_delta = as.numeric(stats::quantile(x, 0.25, na.rm = TRUE)),
      q75_delta = as.numeric(stats::quantile(x, 0.75, na.rm = TRUE)),
      q95_delta = as.numeric(stats::quantile(x, 0.95, na.rm = TRUE))
    )
  }
)

corr_summary <- accepted_real %>%
  dplyr::summarise(
    n = dplyr::n(),
    median_corr = stats::median(.data$corr, na.rm = TRUE),
    q05_corr = as.numeric(stats::quantile(.data$corr, 0.05, na.rm = TRUE)),
    q25_corr = as.numeric(stats::quantile(.data$corr, 0.25, na.rm = TRUE)),
    q75_corr = as.numeric(stats::quantile(.data$corr, 0.75, na.rm = TRUE)),
    q95_corr = as.numeric(stats::quantile(.data$corr, 0.95, na.rm = TRUE)),
    median_margin = stats::median(.data$margin, na.rm = TRUE),
    median_pair_var_pct = 100 * stats::median(.data$pair_var_explained_frac, na.rm = TRUE)
  )

readr::write_csv(delta_summary, accepted_delta_csv)

cat("\nAccepted match correlation summary:\n")
print(corr_summary)
cat("\nAccepted paired diagnostic deltas:\n")
print(delta_summary, n = Inf)

kable(head(accepted_real, 20), digits = 4)
```

# 8. Frontier Plot

```{r frontier-plot, fig.width=9, fig.height=6}
frontier_plot <- threshold_results %>%
  dplyr::filter(.data$null_matches_per_dataset_pair <= 0.05)

p <- ggplot(
  frontier_plot,
  aes(
    x = .data$null_matches_per_dataset_pair,
    y = .data$real_matches_total,
    color = .data$var_min,
    shape = .data$passes_1pct_null
  )
) +
  geom_point(alpha = 0.75, size = 2) +
  geom_vline(xintercept = null_rate_cutoff, linetype = "dashed", color = "black") +
  geom_point(
    data = winner,
    aes(
      x = .data$null_matches_per_dataset_pair,
      y = .data$real_matches_total
    ),
    inherit.aes = FALSE,
    color = "red",
    size = 4
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    x = "Null matches per dataset pair",
    y = "Real accepted matches",
    color = "Minimum var(y) explained",
    shape = "<=1% null",
    title = "Effect matching threshold frontier"
  ) +
  theme(legend.position = "bottom")

ggsave(frontier_plot_png, p, width = 9, height = 6, dpi = 150)
print(p)

cat("Frontier plot:", frontier_plot_png, "\n")
```
