---
title: "Effect Drift Paper Figure (5-arm, cached plotting only)"
output:
  html_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: 72
---

This notebook is the lightweight plotting companion to
`effect_matching_drift_analysis_5arm.Rmd`. It assumes the heavy notebook has
already generated the cached CSV outputs in `study_dir`, then rebuilds only the
combined paper figure. Use this notebook for figure layout, titles, labels,
legends, and styling tweaks.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_drift_paper_figure_5arm.Rmd")

library(here)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(scales)
```

# Inputs

```{r paths}
output_root <- here("output")
study_name <- "baseline_sims_5arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
plot_dir <- file.path(study_dir, "drift_analysis_plots")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

drift_summary_csv <- file.path(study_dir, "drift_summary_5arm.csv")
auprc_per_dataset_csv <- file.path(study_dir, "auprc_per_dataset_5arm.csv")
pip_drift_summary_csv <- file.path(study_dir, "pip_drift_summary_5arm.csv")
pr_curves_pooled_csv <- file.path(study_dir, "pr_curves_pooled_top8_5arm.csv")
pr_auc_pooled_csv <- file.path(study_dir, "pr_auc_pooled_top8_5arm.csv")
effect_quality_pair_rows_csv <- file.path(study_dir, "effect_quality_pair_rows_5arm.csv")
effect_quality_density_shift_csv <- file.path(study_dir, "effect_quality_density_shift_5arm.csv")
pip_records_rds <- file.path(study_dir, "pip_records_5arm.rds")
causal_truth_rds <- file.path(study_dir, "causal_truth_5arm.rds")

paper_figure_png <- file.path(plot_dir, "paper_figure_effect_drift.png")
paper_basis_divergence_png <- file.path(
  plot_dir, "paper_figure_truth_drift_basis_divergence.png"
)

required_cache_files <- c(
  drift_summary_csv,
  auprc_per_dataset_csv,
  pip_drift_summary_csv,
  pr_curves_pooled_csv
)
missing_cache_files <- required_cache_files[!file.exists(required_cache_files)]
if (length(missing_cache_files)) {
  stop(
    "Missing required cached plotting inputs. Run ",
    "effect_matching_drift_analysis_5arm.Rmd once to generate them:\n",
    paste(missing_cache_files, collapse = "\n")
  )
}

cat("Study dir:", study_dir, "\n")
cat("Plot dir:", plot_dir, "\n")
```

# Plot Options

```{r options}
baseline_method <- "susie_vanilla"
mu_method <- "susine_functional_mu"
pi_method <- "susie_functional_pi"
warm_mu_method <- "susie_vanilla_warm_from_mu"
warm_pi_method <- "susie_vanilla_warm_from_pi"

method_labels <- c(
  susie_vanilla = "SuSiE-vanilla (cold)",
  susine_functional_mu = "SuSiNE-functional-mu (cold)",
  susie_vanilla_warm_from_mu = "SuSiE-vanilla (warm from mu)",
  susie_functional_pi = "SuSiE-functional-pi (cold)",
  susie_vanilla_warm_from_pi = "SuSiE-vanilla (warm from pi)"
)
method_legend_short <- c(
  susie_vanilla = "Cold SuSiE",
  susine_functional_mu = "Annotated mu0",
  susie_vanilla_warm_from_mu = "Warm refit (mu0)",
  susie_functional_pi = "Annotated pi",
  susie_vanilla_warm_from_pi = "Warm refit (pi)"
)
method_color <- c(
  susie_vanilla = "black",
  susine_functional_mu = "#2b8cbe",
  susie_vanilla_warm_from_mu = "#9ecae1",
  susie_functional_pi = "#74c476",
  susie_vanilla_warm_from_pi = "#bae4b3"
)
methods_in_order <- c(
  baseline_method, mu_method, warm_mu_method, pi_method, warm_pi_method
)

pr_top_k <- 8L

# Toggle: how many top-|beta| causals define the truth basin used by the
# basin-drift metric. Decoupled from pr_top_k (the AUPRC scoring K, locked
# at 8). Must match the truth_drift_top_k used to build the cache in
# effect_matching_drift_analysis_5arm.Rmd.
truth_drift_top_k <- 23L
truth_drift_spline_df <- 5L
divergence_breaks <- c(-Inf, 0.05, 0.10, Inf)
divergence_labels <- c("[0, 0.05]", "(0.05, 0.10]", "(0.10, inf)")
baseline_bin_palette <- c(
  "[0, 0.05]"    = "#4e79a7",
  "(0.05, 0.10]" = "#f28e2b",
  "(0.10, inf)"  = "#e15759"
)

x_keff_breaks <- c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000)
x_limits <- NULL
y_limits <- c(0, 1)
density_grid_n <- 140L
density_bandwidth <- c(0.45, 0.10)
density_trim_quantile <- c(0.005, 0.995)
density_fill_quantile <- 0.95
```

# Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

resolve_limits <- function(values, user_limits, trim = density_trim_quantile) {
  if (!is.null(user_limits)) {
    return(user_limits)
  }
  values <- values[is.finite(values)]
  if (!length(values)) {
    stop("Cannot infer limits from an empty/non-finite vector.")
  }
  stats::quantile(values, probs = trim, na.rm = TRUE, names = FALSE)
}

keff_x_scale <- function(breaks = x_keff_breaks, xlim = NULL) {
  finite_breaks <- breaks[breaks > 0]
  scale_x_continuous(
    breaks = log(finite_breaks),
    labels = formatC(finite_breaks, format = "fg", big.mark = ","),
    limits = xlim
  )
}

basis_divergence_bin <- function(x) {
  cut(x, breaks = divergence_breaks, labels = divergence_labels, right = TRUE)
}

basis_divergence_axis <- "Explanatory basis divergence"
basis_reduction_axis <- "Reduction in basis divergence"
transition_label_map <- c(
  cold_ann = "Cold -> annotated",
  cold_warm = "Cold -> warm refit"
)
chain_label_map <- c(
  mu = "mu0 channel",
  pi = "pi channel"
)

theme_effect_paper <- function(base_size = 8) {
  theme_bw(base_size = base_size) +
    theme(
      panel.grid.major = element_line(color = "grey91", linewidth = 0.22),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "grey35", fill = NA, linewidth = 0.35),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size - 1, color = "grey25"),
      plot.title = element_text(face = "bold", size = base_size + 1.2,
                                margin = margin(b = 3)),
      strip.background = element_rect(fill = "grey88", color = "grey35",
                                      linewidth = 0.35),
      strip.text = element_text(size = base_size - 0.8, color = "grey20"),
      legend.title = element_text(size = base_size - 0.4),
      legend.text = element_text(size = base_size - 1.1),
      legend.key.height = grid::unit(3.2, "mm"),
      legend.key.width = grid::unit(6.0, "mm"),
      plot.margin = margin(2.5, 2.5, 2.5, 2.5)
    )
}

weighted_kde2d_df <- function(df, xlim, ylim, n, h) {
  group_values <- c("source", "destination")
  dens <- lapply(group_values, function(g) {
    sub <- df %>%
      dplyr::filter(.data$arm_role == g) %>%
      dplyr::filter(
        is.finite(.data$x_value),
        is.finite(.data$y_value),
        .data$x_value >= xlim[[1]],
        .data$x_value <= xlim[[2]],
        .data$y_value >= ylim[[1]],
        .data$y_value <= ylim[[2]]
      )
    if (nrow(sub) < 3L) return(NULL)
    w <- pmax(tidyr::replace_na(sub$pve_weight, 0), 0)
    if (sum(w) <= 0) {
      w <- rep(1 / nrow(sub), nrow(sub))
    } else {
      w <- w / sum(w)
    }
    x_grid <- seq(xlim[[1]], xlim[[2]], length.out = n)
    y_grid <- seq(ylim[[1]], ylim[[2]], length.out = n)
    kx <- outer(x_grid, sub$x_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[1]]) / h[[1]]
    })
    ky <- outer(y_grid, sub$y_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[2]]) / h[[2]]
    })
    ky <- sweep(ky, 2L, w, `*`)
    z <- kx %*% t(ky)
    list(group = g, x = x_grid, y = y_grid, z = z,
         n_rows = nrow(sub), pve_mass = sum(sub$pve_weight, na.rm = TRUE))
  })
  names(dens) <- group_values
  if (is.null(dens$source) || is.null(dens$destination)) {
    return(tibble::tibble())
  }
  tidyr::expand_grid(x = dens$source$x, y = dens$source$y) %>%
    dplyr::mutate(
      density_source = as.vector(t(dens$source$z)),
      density_destination = as.vector(t(dens$destination$z)),
      density_diff = .data$density_destination - .data$density_source,
      n_source = dens$source$n_rows,
      n_destination = dens$destination$n_rows,
      pve_source = dens$source$pve_mass,
      pve_destination = dens$destination$pve_mass
    )
}

build_density_shift_from_pair_rows <- function(pair_rows) {
  x_limits_resolved <- resolve_limits(pair_rows$x_value, x_limits)
  y_limits_resolved <- resolve_limits(pair_rows$y_value, y_limits)
  density_shift_rows <- list()
  for (ch in levels(pair_rows$chain)) {
    for (cmp in levels(pair_rows$comparison_kind)) {
      sub <- pair_rows %>%
        dplyr::filter(.data$chain == ch, .data$comparison_kind == cmp)
      dens <- weighted_kde2d_df(
        sub,
        xlim = x_limits_resolved,
        ylim = y_limits_resolved,
        n = density_grid_n,
        h = density_bandwidth
      )
      if (!nrow(dens)) next
      density_shift_rows[[length(density_shift_rows) + 1L]] <- dens %>%
        dplyr::mutate(chain = ch, comparison_kind = cmp)
    }
  }
  dplyr::bind_rows(density_shift_rows)
}
```

# Read Cached Data

```{r read-cache}
drift_summary <- readr::read_csv(drift_summary_csv, show_col_types = FALSE)
auprc_per_dataset <- readr::read_csv(auprc_per_dataset_csv, show_col_types = FALSE) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))
pip_drift_summary <- readr::read_csv(pip_drift_summary_csv, show_col_types = FALSE) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

pr_curve_df <- readr::read_csv(pr_curves_pooled_csv, show_col_types = FALSE) %>%
  dplyr::mutate(
    model_method = factor(.data$model_method, levels = methods_in_order)
  )
if (file.exists(pr_auc_pooled_csv)) {
  pr_auc_df <- readr::read_csv(pr_auc_pooled_csv, show_col_types = FALSE) %>%
    dplyr::mutate(
      model_method = factor(.data$model_method, levels = methods_in_order)
    )
  prevalence <- unique(pr_auc_df$prevalence)
  if (!length(prevalence) || !is.finite(prevalence[[1]])) {
    prevalence <- 0.008
  } else {
    prevalence <- prevalence[[1]]
  }
} else {
  message("PR-AUC cache missing; estimating AUPRC labels from cached PR curves.")
  pr_auc_df <- pr_curve_df %>%
    dplyr::arrange(.data$model_method, .data$recall) %>%
    dplyr::group_by(.data$model_method) %>%
    dplyr::summarise(
      auprc_integral = sum(
        diff(.data$recall) *
          0.5 * (dplyr::lag(.data$precision)[-1] + .data$precision[-1]),
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      model_method = factor(.data$model_method, levels = methods_in_order)
    )
  prevalence <- 0.008
}

if (file.exists(effect_quality_density_shift_csv)) {
  density_shift_df <- readr::read_csv(
    effect_quality_density_shift_csv, show_col_types = FALSE
  )
} else {
  if (!file.exists(effect_quality_pair_rows_csv)) {
    stop(
      "Missing both density-shift cache and pair-row cache:\n",
      effect_quality_density_shift_csv, "\n", effect_quality_pair_rows_csv
    )
  }
  message("Density-shift cache missing; rebuilding once from effect-quality pair rows.")
  effect_quality_pair_rows <- readr::read_csv(
    effect_quality_pair_rows_csv, show_col_types = FALSE
  ) %>%
    dplyr::mutate(
      chain = factor(.data$chain, levels = c("mu", "pi")),
      arm_role = factor(.data$arm_role, levels = c("source", "destination")),
      comparison_kind = factor(
        .data$comparison_kind,
        levels = c("vanilla_to_treatment",
                   "treatment_to_warm",
                   "vanilla_to_warm")
      )
    )
  density_shift_df <- build_density_shift_from_pair_rows(effect_quality_pair_rows)
  readr::write_csv(density_shift_df, effect_quality_density_shift_csv)
}

density_shift_df <- density_shift_df %>%
  dplyr::mutate(
    chain = factor(.data$chain, levels = c("mu", "pi")),
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    )
  )

x_limits_resolved <- range(density_shift_df$x, na.rm = TRUE)
y_limits_resolved <- range(density_shift_df$y, na.rm = TRUE)
shared_abs_diff <- abs(density_shift_df$density_diff)
shared_abs_diff <- shared_abs_diff[is.finite(shared_abs_diff)]
shared_fill_lim <- stats::quantile(
  shared_abs_diff, probs = density_fill_quantile, na.rm = TRUE, names = FALSE
)
if (!is.finite(shared_fill_lim) || shared_fill_lim <= 0) {
  shared_fill_lim <- max(shared_abs_diff, na.rm = TRUE)
}
if (!is.finite(shared_fill_lim) || shared_fill_lim <= 0) {
  shared_fill_lim <- 1
}
```

# Combined Paper Figure

```{r paper-figure, fig.width=8.8, fig.height=9.9}
expr_basin_x <- expression(paste("Basin drift (", mu, " chain)"))
expr_basin_y <- expression(paste("Basin drift (", pi, " chain)"))
expr_jsd_x   <- expression(paste("PIP JSD (", mu, " chain)"))
expr_jsd_y   <- expression(paste("PIP JSD (", pi, " chain)"))
expr_dauprc  <- expression(paste(Delta, "AUPRC vs cold SuSiE"))

chain_labeller <- c(mu = "mu0 channel", pi = "pi channel")
pair_kind_labeller <- c(
  "cold -> annotated" = "Cold -> annotated",
  "cold -> vanilla_warm" = "Cold -> warm refit"
)

method_legend_labels <- vapply(
  levels(pr_curve_df$model_method),
  function(m) {
    auprc_val <- pr_auc_df$auprc_integral[
      match(m, as.character(pr_auc_df$model_method))
    ]
    sprintf("%s (%.3f)",
            method_legend_short[[m]] %||% method_labels[[m]] %||% m,
            auprc_val)
  },
  character(1)
)

p_pr <- ggplot(pr_curve_df,
               aes(x = .data$recall, y = .data$precision,
                   color = .data$model_method)) +
  geom_hline(yintercept = prevalence, color = "grey60",
             linetype = "dashed", linewidth = 0.4) +
  geom_step(direction = "vh", linewidth = 0.7) +
  scale_color_manual(
    values = method_color[levels(pr_curve_df$model_method)],
    labels = method_legend_labels
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(x = "Recall", y = "Precision", color = "Method (pooled AUPRC)") +
  theme(legend.position = "right")

strip_chrome <- function(p) {
  p + labs(title = NULL, subtitle = NULL) +
    theme(plot.title = element_blank(), plot.subtitle = element_blank())
}

panel_A <- strip_chrome(p_pr) +
  labs(title = paste0(
    "A. Precision-recall on top-", pr_top_k,
    " causal SNPs (600 datasets)"
  )) +
  guides(color = guide_legend(
    nrow = 2, byrow = TRUE, title.position = "top",
    override.aes = list(linewidth = 0.7)
  )) +
  theme_effect_paper(base_size = 8) +
  theme(
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.margin = margin(t = -4),
    legend.key.height = grid::unit(3.0, "mm"),
    legend.key.width = grid::unit(6, "mm"),
    legend.title = element_text(size = 6.8),
    legend.text = element_text(size = 6.2),
    panel.grid.minor = element_line(color = "grey95", linewidth = 0.18)
  )

basin_long_pf <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "vanilla__pi",
    "vanilla__vanilla_warm_from_mu", "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::if_else(
      .data$pair_id %in% c("vanilla__mu", "vanilla__pi"),
      "cold_ann", "cold_warm"
    ),
    basin_drift = .data$weighted_r2_drift
  )

jsd_long_pf <- pip_drift_summary %>%
  dplyr::filter(.data$pair_kind %in% c("vanilla_to_treatment",
                                       "vanilla_to_warm")) %>%
  dplyr::transmute(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    chain = .data$chain,
    pair_kind = dplyr::if_else(
      .data$pair_kind == "vanilla_to_treatment", "cold_ann", "cold_warm"
    ),
    pip_jsd = .data$pip_jsd
  )

basin_wide_pf <- basin_long_pf %>%
  tidyr::pivot_wider(names_from = chain, values_from = basin_drift,
                     names_prefix = "basin_")
jsd_wide_pf <- jsd_long_pf %>%
  tidyr::pivot_wider(names_from = chain, values_from = pip_jsd,
                     names_prefix = "jsd_")
joined_pf <- basin_wide_pf %>%
  dplyr::inner_join(jsd_wide_pf, by = c("dataset_bundle_id", "pair_kind"))

basin_max_pf <- max(0.05,
  max(joined_pf$basin_mu, joined_pf$basin_pi, na.rm = TRUE)) + 0.01
jsd_max_pf <- max(0.05,
  max(joined_pf$jsd_mu, joined_pf$jsd_pi, na.rm = TRUE)) + 0.02
basin_color_max <- max(joined_pf$basin_mu, na.rm = TRUE)
jsd_color_max <- max(joined_pf$jsd_mu, na.rm = TRUE)

make_b_panel <- function(df, x_col, y_col, color_col,
                         x_label, y_label, color_title, title_lab,
                         color_lim_max, max_val,
                         show_y = TRUE,
                         show_legend = TRUE) {
  ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]],
                 color = .data[[color_col]])) +
    geom_abline(slope = 1, intercept = 0,
                color = "grey50", linetype = "dashed", linewidth = 0.35) +
    geom_point(alpha = 0.62, size = 2.25, stroke = 0) +
    scale_color_viridis_c(
      option = "C",
      limits = c(0, color_lim_max),
      breaks = c(0, color_lim_max),
      labels = c("low", "high"),
      guide = if (show_legend) {
        guide_colorbar(
          title.position = "top",
          title.hjust = 0.5,
          label.position = "bottom",
          barwidth  = grid::unit(2.3, "cm"),
          barheight = grid::unit(0.24, "cm"),
          direction = "horizontal"
        )
      } else {
        "none"
      }
    ) +
    scale_x_continuous(expand = expansion(mult = c(0.015, 0.035))) +
    scale_y_continuous(expand = expansion(mult = c(0.015, 0.035))) +
    coord_cartesian(xlim = c(-0.02, max_val), ylim = c(-0.01, max_val),
                    expand = FALSE) +
    labs(x = x_label, y = y_label, color = color_title,
         title = title_lab) +
    theme_effect_paper(base_size = 7.2) +
    theme(
      plot.title = element_text(size = 7.6, face = "bold",
                                margin = margin(b = 0.5)),
      axis.title.x = element_text(size = 7.0, margin = margin(t = 3)),
      axis.title.y = element_text(size = 7.0, margin = margin(r = 3)),
      axis.text = element_text(size = 6.2, color = "grey25"),
      legend.margin = margin(t = -3, b = -4),
      legend.box.margin = margin(t = -4),
      plot.margin = margin(0.5, 2.5, 1.0, 2.5),
      aspect.ratio = 1
    ) +
    {
      if (show_y) {
        theme()
      } else {
        theme(axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank())
      }
    }
}

joined_pf_cold_ann  <- joined_pf %>% dplyr::filter(pair_kind == "cold_ann")
joined_pf_cold_warm <- joined_pf %>% dplyr::filter(pair_kind == "cold_warm")

basin_color_title <- expression(paste("PIP JSD (", mu, ")"))
jsd_color_title   <- expression(paste("Basin drift (", mu, ")"))

pB1 <- make_b_panel(
  joined_pf_cold_ann, "basin_mu", "basin_pi", "jsd_mu",
  expr_basin_x, expr_basin_y, basin_color_title,
  "Cold -> annotated",
  color_lim_max = jsd_color_max, max_val = basin_max_pf,
  show_y = TRUE,
  show_legend = TRUE
)
pB2 <- make_b_panel(
  joined_pf_cold_warm, "basin_mu", "basin_pi", "jsd_mu",
  expr_basin_x, expr_basin_y, basin_color_title,
  "Cold -> warm refit",
  color_lim_max = jsd_color_max, max_val = basin_max_pf,
  show_y = FALSE,
  show_legend = FALSE
)
pB3 <- make_b_panel(
  joined_pf_cold_ann, "jsd_mu", "jsd_pi", "basin_mu",
  expr_jsd_x, expr_jsd_y, jsd_color_title,
  "Cold -> annotated",
  color_lim_max = basin_color_max, max_val = jsd_max_pf,
  show_y = TRUE,
  show_legend = TRUE
)
pB4 <- make_b_panel(
  joined_pf_cold_warm, "jsd_mu", "jsd_pi", "basin_mu",
  expr_jsd_x, expr_jsd_y, jsd_color_title,
  "Cold -> warm refit",
  color_lim_max = basin_color_max, max_val = jsd_max_pf,
  show_y = FALSE,
  show_legend = FALSE
)

b_pair_header <- function(lbl) {
  ggplot() +
    annotate("text", x = 0.5, y = 0.38, label = lbl,
             size = 2.5, fontface = "bold") +
    xlim(0, 1) + ylim(0, 1) +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0))
}

b_panel_title <- function(lbl) {
  ggplot() +
    annotate("text", x = -0.055, y = 0.70, label = lbl,
             hjust = 0, size = 3.05, fontface = "bold") +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0))
}

basin_pair_B <- (
  b_pair_header("Basin drift") /
    (pB1 | pB2) /
    guide_area()
) +
  plot_layout(
    heights = c(0.055, 1, 0.070),
    guides = "collect"
  ) &
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.title = element_text(size = 6.8),
    legend.text = element_text(size = 6.0),
    legend.margin = margin(t = -9, r = 0, b = -7, l = 0),
    legend.box.margin = margin(t = -12, r = 0, b = -6, l = 0)
  )

jsd_pair_B <- (
  b_pair_header("PIP JSD") /
    (pB3 | pB4) /
    guide_area()
) +
  plot_layout(
    heights = c(0.055, 1, 0.070),
    guides = "collect"
  ) &
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.title = element_text(size = 6.8),
    legend.text = element_text(size = 6.0),
    legend.margin = margin(t = -9, r = 0, b = -7, l = 0),
    legend.box.margin = margin(t = -12, r = 0, b = -6, l = 0)
  )

panel_C <- (
  b_panel_title("C. Effect-level drift diagnostics (mu vs pi)") /
    (basin_pair_B | jsd_pair_B)
) +
  plot_layout(heights = c(0.060, 1))

auprc_lookup_pf <- auprc_per_dataset %>%
  dplyr::transmute(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    delta_mu_treat = .data$auprc_delta_mu,
    delta_mu_warm  = .data$auprc_delta_warm_mu_vs_vanilla,
    delta_pi_treat = .data$auprc_delta_pi,
    delta_pi_warm  = .data$auprc_delta_warm_pi_vs_vanilla
  )

panel_B_long <- basin_long_pf %>%
  dplyr::left_join(auprc_lookup_pf, by = "dataset_bundle_id") %>%
  dplyr::mutate(
    auprc_delta = dplyr::case_when(
      chain == "mu" & pair_kind == "cold_ann"  ~ delta_mu_treat,
      chain == "mu" & pair_kind == "cold_warm" ~ delta_mu_warm,
      chain == "pi" & pair_kind == "cold_ann"  ~ delta_pi_treat,
      chain == "pi" & pair_kind == "cold_warm" ~ delta_pi_warm
    ),
    drift_bucket = cut(
      basin_drift,
      breaks = c(-Inf, 0.05, Inf),
      labels = c("<0.05", ">=0.05"),
      right = FALSE
    ),
    chain = factor(chain, levels = c("mu", "pi")),
    pair_kind = factor(
      pair_kind,
      levels = c("cold_ann", "cold_warm"),
      labels = c("cold -> annotated", "cold -> vanilla_warm")
    )
  ) %>%
  dplyr::filter(is.finite(.data$auprc_delta), !is.na(.data$drift_bucket))

bucket_counts <- panel_B_long %>%
  dplyr::count(.data$chain, .data$pair_kind, .data$drift_bucket)
y_top <- max(panel_B_long$auprc_delta, na.rm = TRUE) * 1.08

panel_B <- ggplot(
  panel_B_long,
  aes(x = .data$drift_bucket,
      y = .data$auprc_delta,
      fill = .data$drift_bucket)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_boxplot(alpha = 0.65, outlier.size = 0.6, width = 0.65) +
  geom_text(
    data = bucket_counts,
    aes(x = .data$drift_bucket, y = y_top,
        label = sprintf("n=%d", .data$n)),
    inherit.aes = FALSE, size = 2.1, color = "grey25"
  ) +
  facet_grid(
    chain ~ pair_kind,
    labeller = labeller(chain = chain_labeller,
                        pair_kind = pair_kind_labeller)
  ) +
  scale_fill_manual(values = c("<0.05" = "#878787", ">=0.05" = "#c44e52")) +
  labs(
    x = "Basin drift bucket",
    y = expr_dauprc,
    fill = NULL,
    title = "B. AUPRC change by basin-drift bucket"
  ) +
  guides(fill = "none") +
  theme_effect_paper(base_size = 8) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 6.8),
    plot.title = element_text(face = "bold", size = 9.2,
                              margin = margin(b = 3))
  )

panel_D_df <- density_shift_df %>%
  dplyr::filter(.data$comparison_kind %in% c(
    "vanilla_to_treatment", "vanilla_to_warm"
  )) %>%
  dplyr::mutate(
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment", "vanilla_to_warm"),
      labels = c("cold -> annotated", "cold -> vanilla_warm")
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

x_label_paper <- expression(paste(
  "Effect diffuseness: effective top-95% support size  ", K[l]
))
y_label_paper <- expression(paste(
  "Effect accuracy: ", A[l], " = max causal PIP / max SNP PIP"
))

panel_D_core <- ggplot(panel_D_df,
                  aes(x = .data$x, y = .data$y,
                      fill = .data$density_diff)) +
  geom_raster(interpolate = TRUE) +
  geom_contour(aes(z = .data$density_diff),
               color = "grey20", alpha = 0.45, linewidth = 0.18) +
  facet_grid(
    chain ~ comparison_kind,
    labeller = labeller(chain = chain_labeller,
                        comparison_kind = pair_kind_labeller)
  ) +
  scale_fill_gradient2(
    low = "#2b6cb0", mid = "white", high = "#c53030",
    midpoint = 0, limits = c(-shared_fill_lim, shared_fill_lim),
    oob = scales::squish
  ) +
  keff_x_scale(xlim = x_limits_resolved) +
  coord_cartesian(xlim = x_limits_resolved,
                  ylim = y_limits_resolved,
                  expand = FALSE) +
  labs(
    x = x_label_paper, y = y_label_paper,
    fill = "destination - source\n(PVE-weighted density)",
    title = "D. Per-effect quality shifts"
  ) +
  guides(fill = guide_colorbar(
    title.position = "top",
    title.hjust = 0.5,
    barwidth = grid::unit(0.34, "cm"),
    barheight = grid::unit(2.4, "cm"),
    direction = "vertical"
  )) +
  theme_effect_paper(base_size = 8) +
  theme(
    legend.position = "right",
    legend.justification = "center",
    legend.margin = margin(0, 0, 0, 2),
    legend.box.margin = margin(0, 0, 0, 2),
    panel.spacing.y = grid::unit(1.25, "mm"),
    plot.title = element_text(face = "bold", size = 9.2,
                              margin = margin(b = 3))
  )

panel_D <- (panel_D_core + guide_area()) +
  plot_layout(widths = c(1, 0.13), guides = "collect") &
  theme(legend.position = "right")

top_summary_row <- (wrap_elements(full = panel_A) | wrap_elements(full = panel_B)) +
  plot_layout(widths = c(0.52, 0.48))

combined_fig <- (top_summary_row / panel_C / panel_D) +
  plot_layout(heights = c(1, 1, 1))

ggsave(
  paper_figure_png, combined_fig,
  width = 8.8, height = 9.9, dpi = 300, bg = "white",
  limitsize = FALSE
)
cat("Wrote combined paper figure to ", paper_figure_png, "\n", sep = "")
print(combined_fig)
```

# Ad Hoc Diagnostics For Panel C

```{r panel-c-ad-hoc-diagnostics, fig.width=10, fig.height=4.8}
diagnostic_dir <- file.path(plot_dir, "panel_c_ad_hoc_diagnostics")
dir.create(diagnostic_dir, recursive = TRUE, showWarnings = FALSE)

basin_quadrant_threshold <- 0.10

classify_basin_quadrant <- function(basin_mu, basin_pi,
                                    threshold = basin_quadrant_threshold) {
  dplyr::case_when(
    basin_mu >= threshold & basin_pi >= threshold ~ "both",
    basin_mu >= threshold & basin_pi <  threshold ~ "mu_only",
    basin_mu <  threshold & basin_pi >= threshold ~ "pi_only",
    TRUE ~ "neither"
  )
}

panel_c_diag <- joined_pf %>%
  dplyr::left_join(auprc_lookup_pf, by = "dataset_bundle_id") %>%
  dplyr::mutate(
    comparison = dplyr::recode(
      .data$pair_kind,
      cold_ann = "cold -> annotated",
      cold_warm = "cold -> warm refit"
    ),
    delta_mu = dplyr::if_else(.data$pair_kind == "cold_ann",
                             .data$delta_mu_treat, .data$delta_mu_warm),
    delta_pi = dplyr::if_else(.data$pair_kind == "cold_ann",
                             .data$delta_pi_treat, .data$delta_pi_warm),
    delta_mu_minus_pi = .data$delta_mu - .data$delta_pi,
    basin_quadrant = classify_basin_quadrant(.data$basin_mu, .data$basin_pi),
    basin_quadrant = factor(
      .data$basin_quadrant,
      levels = c("neither", "mu_only", "pi_only", "both")
    ),
    comparison = factor(
      .data$comparison,
      levels = c("cold -> annotated", "cold -> warm refit")
    )
  )

quadrant_summary <- panel_c_diag %>%
  dplyr::group_by(.data$comparison, .data$basin_quadrant) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean_basin_mu = mean(.data$basin_mu, na.rm = TRUE),
    mean_basin_pi = mean(.data$basin_pi, na.rm = TRUE),
    mean_jsd_mu = mean(.data$jsd_mu, na.rm = TRUE),
    mean_jsd_pi = mean(.data$jsd_pi, na.rm = TRUE),
    mean_delta_mu = mean(.data$delta_mu, na.rm = TRUE),
    mean_delta_pi = mean(.data$delta_pi, na.rm = TRUE),
    mean_delta_mu_minus_pi = mean(.data$delta_mu_minus_pi, na.rm = TRUE),
    median_delta_mu_minus_pi = median(.data$delta_mu_minus_pi, na.rm = TRUE),
    .groups = "drop"
  )
readr::write_csv(
  quadrant_summary,
  file.path(diagnostic_dir, "panel_c_basin_quadrant_summary.csv")
)
print(quadrant_summary)

quadrant_membership <- panel_c_diag %>%
  dplyr::select(
    .data$dataset_bundle_id, .data$pair_kind, .data$basin_quadrant
  ) %>%
  tidyr::pivot_wider(names_from = .data$pair_kind,
                     values_from = .data$basin_quadrant) %>%
  dplyr::rename(
    cold_annotated_quadrant = .data$cold_ann,
    cold_warm_quadrant = .data$cold_warm
  )
readr::write_csv(
  quadrant_membership,
  file.path(diagnostic_dir, "panel_c_basin_quadrant_membership.csv")
)

quadrant_transition <- quadrant_membership %>%
  dplyr::count(.data$cold_annotated_quadrant, .data$cold_warm_quadrant,
               name = "n") %>%
  tidyr::complete(
    cold_annotated_quadrant = factor(
      c("neither", "mu_only", "pi_only", "both"),
      levels = c("neither", "mu_only", "pi_only", "both")
    ),
    cold_warm_quadrant = factor(
      c("neither", "mu_only", "pi_only", "both"),
      levels = c("neither", "mu_only", "pi_only", "both")
    ),
    fill = list(n = 0L)
  )
readr::write_csv(
  quadrant_transition,
  file.path(diagnostic_dir, "panel_c_basin_quadrant_transition.csv")
)

p_quadrant_transition <- ggplot(
  quadrant_transition,
  aes(x = .data$cold_annotated_quadrant,
      y = .data$cold_warm_quadrant,
      fill = .data$n)
) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = .data$n), size = 3.2, color = "grey15") +
  scale_fill_gradient(low = "grey96", high = "#2b6cb0") +
  labs(
    x = "Basin quadrant: cold -> annotated",
    y = "Basin quadrant: cold -> warm refit",
    fill = "datasets",
    title = "Panel C diagnostic: basin-drift quadrant membership"
  ) +
  coord_equal() +
  theme_effect_paper(base_size = 10) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
ggsave(
  file.path(diagnostic_dir, "panel_c_basin_quadrant_transition.png"),
  p_quadrant_transition,
  width = 6.2, height = 5.0, dpi = 220, bg = "white"
)
print(p_quadrant_transition)

delta_diff_lim <- max(abs(panel_c_diag$delta_mu_minus_pi), na.rm = TRUE)
p_delta_diff_basin <- ggplot(
  panel_c_diag,
  aes(x = .data$basin_mu, y = .data$basin_pi,
      color = .data$delta_mu_minus_pi)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50",
              linetype = "dashed", linewidth = 0.35) +
  geom_hline(yintercept = basin_quadrant_threshold, color = "grey75",
             linewidth = 0.25) +
  geom_vline(xintercept = basin_quadrant_threshold, color = "grey75",
             linewidth = 0.25) +
  geom_point(alpha = 0.72, size = 1.9) +
  facet_wrap(~comparison, nrow = 1) +
  scale_color_gradient2(
    low = "#2166ac", mid = "grey96", high = "#b2182b",
    midpoint = 0, limits = c(-delta_diff_lim, delta_diff_lim),
    oob = scales::squish
  ) +
  coord_cartesian(xlim = c(-0.02, basin_max_pf),
                  ylim = c(-0.01, basin_max_pf),
                  expand = FALSE) +
  labs(
    x = expr_basin_x,
    y = expr_basin_y,
    color = expression(Delta * "AUPRC"[mu] - Delta * "AUPRC"[pi]),
    title = "Panel C diagnostic: relative AUPRC gain over basin-drift geometry"
  ) +
  theme_effect_paper(base_size = 10)
ggsave(
  file.path(diagnostic_dir, "panel_c_basin_scatter_delta_mu_minus_pi.png"),
  p_delta_diff_basin,
  width = 9.5, height = 4.3, dpi = 220, bg = "white"
)
print(p_delta_diff_basin)

metric_triangle_long <- basin_long_pf %>%
  dplyr::inner_join(
    jsd_long_pf,
    by = c("dataset_bundle_id", "chain", "pair_kind")
  ) %>%
  dplyr::left_join(auprc_lookup_pf, by = "dataset_bundle_id") %>%
  dplyr::mutate(
    auprc_delta = dplyr::case_when(
      .data$chain == "mu" & .data$pair_kind == "cold_ann" ~
        .data$delta_mu_treat,
      .data$chain == "mu" & .data$pair_kind == "cold_warm" ~
        .data$delta_mu_warm,
      .data$chain == "pi" & .data$pair_kind == "cold_ann" ~
        .data$delta_pi_treat,
      .data$chain == "pi" & .data$pair_kind == "cold_warm" ~
        .data$delta_pi_warm,
      TRUE ~ NA_real_
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"),
                   labels = c("mu0 channel", "pi channel")),
    comparison = factor(
      dplyr::recode(
        .data$pair_kind,
        cold_ann = "cold -> annotated",
        cold_warm = "cold -> warm refit"
      ),
      levels = c("cold -> annotated", "cold -> warm refit")
    )
  ) %>%
  dplyr::select(
    .data$dataset_bundle_id, .data$chain, .data$comparison,
    .data$basin_drift, .data$pip_jsd, .data$auprc_delta
  )

readr::write_csv(
  metric_triangle_long,
  file.path(diagnostic_dir, "panel_c_metric_triangle_long.csv")
)

metric_axis_labels <- c(
  basin_drift = "Basin drift",
  pip_jsd = "PIP JSD",
  auprc_delta = "AUPRC delta vs cold SuSiE"
)

make_metric_triangle_plot <- function(df, x_metric, y_metric, color_metric,
                                      filename) {
  p <- ggplot(
    df,
    aes(x = .data[[x_metric]], y = .data[[y_metric]],
        color = .data[[color_metric]])
  ) +
    geom_hline(
      yintercept = 0,
      color = if (y_metric == "auprc_delta") "grey70" else NA,
      linewidth = 0.25
    ) +
    geom_vline(
      xintercept = 0,
      color = if (x_metric == "auprc_delta") "grey70" else NA,
      linewidth = 0.25
    ) +
    geom_point(alpha = 0.70, size = 1.55, stroke = 0) +
    facet_grid(chain ~ comparison) +
    scale_color_viridis_c(option = "C", oob = scales::squish) +
    labs(
      x = metric_axis_labels[[x_metric]],
      y = metric_axis_labels[[y_metric]],
      color = metric_axis_labels[[color_metric]],
      title = paste(
        "Panel C diagnostic:",
        metric_axis_labels[[x_metric]], "vs",
        metric_axis_labels[[y_metric]],
        "colored by", metric_axis_labels[[color_metric]]
      )
    ) +
    theme_effect_paper(base_size = 9) +
    theme(
      legend.position = "right",
      panel.grid.minor = element_line(color = "grey95", linewidth = 0.18)
    )

  ggsave(
    file.path(diagnostic_dir, filename),
    p,
    width = 8.8, height = 5.4, dpi = 220, bg = "white"
  )
  print(p)
  invisible(p)
}

make_metric_triangle_plot(
  metric_triangle_long,
  x_metric = "basin_drift",
  y_metric = "pip_jsd",
  color_metric = "auprc_delta",
  filename = "panel_c_metric_triangle_color_auprc_delta.png"
)

make_metric_triangle_plot(
  metric_triangle_long,
  x_metric = "pip_jsd",
  y_metric = "auprc_delta",
  color_metric = "basin_drift",
  filename = "panel_c_metric_triangle_color_basin_drift.png"
)

make_metric_triangle_plot(
  metric_triangle_long,
  x_metric = "auprc_delta",
  y_metric = "basin_drift",
  color_metric = "pip_jsd",
  filename = "panel_c_metric_triangle_color_pip_jsd.png"
)

representative_datasets <- panel_c_diag %>%
  dplyr::mutate(
    representative_score = abs(.data$delta_mu_minus_pi) +
      0.05 * (.data$basin_mu + .data$basin_pi)
  ) %>%
  dplyr::group_by(.data$comparison, .data$basin_quadrant) %>%
  dplyr::slice_max(.data$representative_score, n = 2L, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(.data$comparison, .data$basin_quadrant,
                 dplyr::desc(.data$representative_score)) %>%
  dplyr::select(
    .data$dataset_bundle_id, .data$comparison, .data$basin_quadrant,
    .data$basin_mu, .data$basin_pi, .data$jsd_mu, .data$jsd_pi,
    .data$delta_mu, .data$delta_pi, .data$delta_mu_minus_pi
  )
readr::write_csv(
  representative_datasets,
  file.path(diagnostic_dir, "panel_c_representative_datasets.csv")
)
print(representative_datasets)

if (file.exists(pip_records_rds) && file.exists(causal_truth_rds)) {
  pip_records_diag <- readRDS(pip_records_rds) %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))
  causal_truth_diag <- readRDS(causal_truth_rds)
  rep_ids <- unique(representative_datasets$dataset_bundle_id)
  top_pip_rows <- lapply(rep_ids, function(ds_id) {
    truth_i <- causal_truth_diag[[as.character(ds_id)]]
    if (is.null(truth_i)) return(NULL)
    pip_records_diag %>%
      dplyr::filter(.data$dataset_bundle_id == !!ds_id) %>%
      dplyr::filter(.data$model_method %in% methods_in_order) %>%
      dplyr::group_by(.data$model_method) %>%
      dplyr::group_modify(function(.x, .y) {
        pip_vec <- as.numeric(.x$pip[[1]])
        tibble::tibble(
          snp_index = seq_along(pip_vec),
          pip = pip_vec,
          is_top8_causal = seq_along(pip_vec) %in% truth_i$top_idx,
          is_any_causal = seq_along(pip_vec) %in% truth_i$causal_idx
        ) %>%
          dplyr::arrange(dplyr::desc(.data$pip)) %>%
          dplyr::slice_head(n = 12L)
      }) %>%
      dplyr::ungroup() %>%
      dplyr::mutate(dataset_bundle_id = ds_id, .before = 1L)
  }) %>%
    dplyr::bind_rows() %>%
    dplyr::left_join(representative_datasets, by = "dataset_bundle_id")
  if (nrow(top_pip_rows) > 0L) {
    top_pip_rows <- top_pip_rows %>%
      dplyr::relocate(
        .data$dataset_bundle_id, .data$comparison, .data$basin_quadrant,
        .data$model_method, .data$snp_index, .data$pip,
        .data$is_top8_causal, .data$is_any_causal
      )
  }
  readr::write_csv(
    top_pip_rows,
    file.path(diagnostic_dir, "panel_c_representative_top_pips.csv")
  )
  cat("Wrote representative top-PIP table to ",
      file.path(diagnostic_dir, "panel_c_representative_top_pips.csv"),
      "\n", sep = "")
} else {
  message("Skipping representative top-PIP table; missing PIP/truth RDS cache.")
}

# ---- Swap-pair plots: basin chord drift vs PIP JSD ----
# Two complementary views of the same (basin, JSD) plane.
#   facet = pair_kind, color = chain   -> chain difference within transition
#   facet = chain,     color = pair_kind -> transition difference within chain
# Together these visualize the four basin/JSD coupling cells from the
# pairwise correlation heatmap (mu/pi x annotated/warm).
swap_pair_long <- joined_pf %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    pair_kind = .data$pair_kind,
    chain_mu_basin = .data$basin_mu, chain_mu_jsd = .data$jsd_mu,
    chain_pi_basin = .data$basin_pi, chain_pi_jsd = .data$jsd_pi
  ) %>%
  tidyr::pivot_longer(
    cols = dplyr::starts_with("chain_"),
    names_to = c("chain", ".value"),
    names_pattern = "chain_(mu|pi)_(basin|jsd)"
  ) %>%
  dplyr::mutate(
    chain = factor(.data$chain, levels = c("mu", "pi")),
    pair_kind = factor(
      .data$pair_kind, levels = c("cold_ann", "cold_warm"),
      labels = c("cold -> annotated", "cold -> warm refit")
    )
  )

chain_color_diag <- c(mu = "#2b8cbe", pi = "#74c476")
pair_color_diag  <- c(`cold -> annotated`     = "#c44e52",
                      `cold -> warm refit`  = "#4c72b0")
axis_lim_swap <- c(0, max(swap_pair_long$basin, swap_pair_long$jsd,
                          na.rm = TRUE) + 0.01)

p_swap_facet_pair <- ggplot(
  swap_pair_long,
  aes(x = .data$basin, y = .data$jsd, color = .data$chain)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50",
              linetype = "dashed", linewidth = 0.3) +
  geom_point(alpha = 0.55, size = 1.6, stroke = 0) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
  facet_wrap(~ pair_kind, nrow = 1) +
  scale_color_manual(values = chain_color_diag,
                     labels = chain_label_map) +
  coord_cartesian(xlim = axis_lim_swap, ylim = axis_lim_swap) +
  labs(
    x = "Basin drift", y = "PIP JSD", color = NULL,
    title = "Panel C diagnostic: basin/JSD coupling by chain within transition"
  ) +
  theme_effect_paper(base_size = 10) +
  theme(legend.position = "bottom")
ggsave(
  file.path(diagnostic_dir, "panel_c_swap_facet_pair_color_chain.png"),
  p_swap_facet_pair, width = 8.4, height = 4.2, dpi = 220, bg = "white"
)
print(p_swap_facet_pair)

p_swap_facet_chain <- ggplot(
  swap_pair_long,
  aes(x = .data$basin, y = .data$jsd, color = .data$pair_kind)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey50",
              linetype = "dashed", linewidth = 0.3) +
  geom_point(alpha = 0.55, size = 1.6, stroke = 0) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
  facet_wrap(~ chain, nrow = 1,
             labeller = labeller(chain = chain_label_map)) +
  scale_color_manual(values = pair_color_diag) +
  coord_cartesian(xlim = axis_lim_swap, ylim = axis_lim_swap) +
  labs(
    x = "Basin drift", y = "PIP JSD", color = NULL,
    title = "Panel C diagnostic: basin/JSD coupling by transition within chain"
  ) +
  theme_effect_paper(base_size = 10) +
  theme(legend.position = "bottom")
ggsave(
  file.path(diagnostic_dir, "panel_c_swap_facet_chain_color_pair.png"),
  p_swap_facet_chain, width = 8.4, height = 4.2, dpi = 220, bg = "white"
)
print(p_swap_facet_chain)

cat("Wrote Panel C ad hoc diagnostics to ", diagnostic_dir, "\n", sep = "")
```

```{r panel-c-correlation-diagnostics, fig.width=9, fig.height=4.6}
correlation_dir <- file.path(diagnostic_dir, "correlation_summaries")
dir.create(correlation_dir, recursive = TRUE, showWarnings = FALSE)

metric_pairs <- tibble::tribble(
  ~metric_x,       ~metric_y,      ~pair_label,
  "basin_drift",  "pip_jsd",      "basin drift vs PIP JSD",
  "basin_drift",  "auprc_delta",  "basin drift vs AUPRC delta",
  "pip_jsd",      "auprc_delta",  "PIP JSD vs AUPRC delta"
)

pairwise_cor_rows <- metric_triangle_long %>%
  tidyr::nest(data = -c(chain, comparison)) %>%
  tidyr::expand_grid(metric_pairs) %>%
  dplyr::mutate(
    stats = purrr::pmap(
      list(data, metric_x, metric_y),
      function(d, metric_x, metric_y) {
        complete <- is.finite(d[[metric_x]]) & is.finite(d[[metric_y]])
        n_complete <- sum(complete)
        if (n_complete < 3L) {
          return(tibble::tibble(n = n_complete, pearson_r = NA_real_,
                                p_value = NA_real_))
        }
        test <- stats::cor.test(
          d[[metric_x]][complete],
          d[[metric_y]][complete],
          method = "pearson"
        )
        tibble::tibble(
          n = n_complete,
          pearson_r = unname(test$estimate),
          p_value = test$p.value
        )
      }
    )
  ) %>%
  dplyr::select(-data) %>%
  tidyr::unnest(cols = stats) %>%
  dplyr::mutate(
    pair_label = factor(
      .data$pair_label,
      levels = metric_pairs$pair_label
    )
  )

readr::write_csv(
  pairwise_cor_rows,
  file.path(correlation_dir, "panel_c_pairwise_correlations.csv")
)
print(pairwise_cor_rows)

cor_lim <- max(abs(pairwise_cor_rows$pearson_r), na.rm = TRUE)
if (!is.finite(cor_lim) || cor_lim <= 0) cor_lim <- 1
p_pairwise_cor <- ggplot(
  pairwise_cor_rows,
  aes(x = .data$pair_label, y = .data$chain, fill = .data$pearson_r)
) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(
    aes(label = sprintf("r=%.2f\nn=%d", .data$pearson_r, .data$n)),
    size = 3.0,
    lineheight = 0.9,
    color = "grey12"
  ) +
  facet_wrap(~comparison, nrow = 1) +
  scale_fill_gradient2(
    low = "#2166ac", mid = "grey96", high = "#b2182b",
    midpoint = 0, limits = c(-cor_lim, cor_lim), oob = scales::squish
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Pearson r",
    title = "Panel C diagnostic: pairwise metric correlations"
  ) +
  theme_effect_paper(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    panel.grid = element_blank()
  )
ggsave(
  file.path(correlation_dir, "panel_c_pairwise_correlations.png"),
  p_pairwise_cor,
  width = 8.8, height = 4.2, dpi = 220, bg = "white"
)
print(p_pairwise_cor)

partial_cor_one <- function(d, x, y, z) {
  d <- d %>%
    dplyr::filter(
      is.finite(.data[[x]]),
      is.finite(.data[[y]]),
      is.finite(.data[[z]])
    )
  if (nrow(d) < 4L) return(tibble::tibble(n = nrow(d), partial_r = NA_real_))
  rx <- stats::resid(stats::lm(d[[x]] ~ d[[z]]))
  ry <- stats::resid(stats::lm(d[[y]] ~ d[[z]]))
  tibble::tibble(n = nrow(d), partial_r = stats::cor(rx, ry))
}

partial_cor_specs <- tibble::tribble(
  ~metric_x,       ~metric_y,      ~control_metric, ~partial_label,
  "auprc_delta",  "pip_jsd",      "basin_drift",  "AUPRC delta vs PIP JSD | basin drift",
  "auprc_delta",  "basin_drift",  "pip_jsd",      "AUPRC delta vs basin drift | PIP JSD",
  "basin_drift",  "pip_jsd",      "auprc_delta",  "basin drift vs PIP JSD | AUPRC delta"
)

partial_cor_rows <- metric_triangle_long %>%
  tidyr::nest(data = -c(chain, comparison)) %>%
  tidyr::expand_grid(partial_cor_specs) %>%
  dplyr::mutate(
    stats = purrr::pmap(
      list(data, metric_x, metric_y, control_metric),
      partial_cor_one
    )
  ) %>%
  dplyr::select(-data) %>%
  tidyr::unnest(cols = stats) %>%
  dplyr::mutate(
    partial_label = factor(
      .data$partial_label,
      levels = partial_cor_specs$partial_label
    )
  )

readr::write_csv(
  partial_cor_rows,
  file.path(correlation_dir, "panel_c_partial_correlations.csv")
)
print(partial_cor_rows)

partial_lim <- max(abs(partial_cor_rows$partial_r), na.rm = TRUE)
if (!is.finite(partial_lim) || partial_lim <= 0) partial_lim <- 1
p_partial_cor <- ggplot(
  partial_cor_rows,
  aes(x = .data$partial_label, y = .data$chain, fill = .data$partial_r)
) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(
    aes(label = sprintf("r=%.2f\nn=%d", .data$partial_r, .data$n)),
    size = 2.8,
    lineheight = 0.9,
    color = "grey12"
  ) +
  facet_wrap(~comparison, nrow = 1) +
  scale_fill_gradient2(
    low = "#2166ac", mid = "grey96", high = "#b2182b",
    midpoint = 0, limits = c(-partial_lim, partial_lim), oob = scales::squish
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Partial r",
    title = "Panel C diagnostic: partial metric correlations"
  ) +
  theme_effect_paper(base_size = 9) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    panel.grid = element_blank()
  )
ggsave(
  file.path(correlation_dir, "panel_c_partial_correlations.png"),
  p_partial_cor,
  width = 9.6, height = 4.2, dpi = 220, bg = "white"
)
print(p_partial_cor)
```

# Explanatory Basis Divergence Sanity Check

```{r explanatory-basis-divergence-sanity, fig.width=9, fig.height=4.4}
truth_drift_csv <- file.path(
  study_dir,
  sprintf("truth_drift_summary_5arm_top%d.csv", truth_drift_top_k)
)

# Backward-compat: if only the legacy un-suffixed cache exists, fall back to
# it when truth_drift_top_k == 8 (matches the legacy build's K).
legacy_truth_drift_csv <- file.path(study_dir, "truth_drift_summary_5arm.csv")
if (!file.exists(truth_drift_csv) &&
    truth_drift_top_k == 8L &&
    file.exists(legacy_truth_drift_csv)) {
  truth_drift_csv <- legacy_truth_drift_csv
  cat("Using legacy un-suffixed truth_drift_summary cache: ",
      truth_drift_csv, "\n", sep = "")
}

if (!file.exists(truth_drift_csv)) {
  message(
    "truth_drift_summary_5arm.csv not yet cached. Run the explanatory-basis ",
    "section of effect_matching_drift_analysis_5arm.Rmd to generate it."
  )
} else {
  truth_drift_summary <- readr::read_csv(truth_drift_csv,
                                         show_col_types = FALSE) %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

  # The main paper figure (joined_pf / panel_c_diag) uses weighted_r2_drift,
  # not weighted_chord_drift, so the apples-to-apples sanity comparison
  # against auprc_delta is on the r2 axis. The CSV exposes both columns; swap
  # to truth_chord_drift below if you want to compare on chord instead.
  truth_drift_column <- "truth_r2_drift"

  # delta_truth per (dataset, chain, pair_kind): baseline explanatory-basis
  # divergence minus treatment divergence. Positive = treatment moved the fit
  # closer to the top-k truth basis.
  delta_truth_for_pair <- function(treat_method, chain_label, pair_kind_label) {
    truth_drift_summary %>%
      dplyr::filter(.data$model_method %in% c(baseline_method, treat_method)) %>%
      tidyr::pivot_wider(
        id_cols = dataset_bundle_id,
        names_from = model_method,
        values_from = dplyr::all_of(truth_drift_column)
      ) %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        chain = chain_label,
        pair_kind = pair_kind_label,
        delta_truth = .data[[baseline_method]] - .data[[treat_method]]
      )
  }

  delta_truth_long <- dplyr::bind_rows(
    delta_truth_for_pair(mu_method,      "mu", "cold_ann"),
    delta_truth_for_pair(warm_mu_method, "mu", "cold_warm"),
    delta_truth_for_pair(pi_method,      "pi", "cold_ann"),
    delta_truth_for_pair(warm_pi_method, "pi", "cold_warm")
  )

  # Pair up against per-dataset AUPRC delta from panel_c_diag.
  panel_c_diag_sanity <- panel_c_diag %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      pair_kind = dplyr::recode(
        as.character(.data$comparison),
        `cold -> annotated` = "cold_ann",
        `cold -> warm refit` = "cold_warm"
      ),
      delta_mu = .data$delta_mu,
      delta_pi = .data$delta_pi
    ) %>%
    tidyr::pivot_longer(
      cols = c(delta_mu, delta_pi),
      names_to = "chain",
      values_to = "auprc_delta",
      names_prefix = "delta_"
    )

  sanity_join <- delta_truth_long %>%
    dplyr::inner_join(panel_c_diag_sanity,
                      by = c("dataset_bundle_id", "pair_kind", "chain"))

  sanity_corr <- sanity_join %>%
    dplyr::group_by(.data$chain, .data$pair_kind) %>%
    dplyr::summarise(
      r = stats::cor(.data$delta_truth, .data$auprc_delta,
                     use = "complete.obs"),
      n = sum(is.finite(.data$delta_truth) & is.finite(.data$auprc_delta)),
      .groups = "drop"
    )
  readr::write_csv(
    sanity_corr,
    file.path(diagnostic_dir, "truth_drift_sanity_correlations.csv")
  )
  cat("Explanatory-basis-divergence vs AUPRC-delta sanity correlations:\n")
  print(sanity_corr)
  cat(
    "\nDecision rule:\n",
    "  mu cold_ann r > 0.95 -> divergence is a reparameterization of AUPRC; abandon\n",
    "  mu cold_ann r in [0.6, 0.85] -> proceed to directionality scatter\n",
    "  mu cold_ann r < 0.5 -> carries independent information; verify per-dataset\n",
    "                          divergence distribution for pathological values\n",
    sep = ""
  )

  p_sanity <- ggplot(
    sanity_join,
    aes(x = .data$delta_truth, y = .data$auprc_delta,
        color = .data$chain)
  ) +
    geom_hline(yintercept = 0, color = "grey60", linewidth = 0.25) +
    geom_vline(xintercept = 0, color = "grey60", linewidth = 0.25) +
    geom_point(alpha = 0.55, size = 1.5, stroke = 0) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
    facet_wrap(~ pair_kind, nrow = 1,
               labeller = labeller(pair_kind = c(
                 cold_ann  = "Cold -> annotated",
                 cold_warm = "Cold -> warm refit"
               ))) +
    scale_color_manual(values = c(mu = "#2b8cbe", pi = "#74c476"),
                       labels = chain_label_map) +
    labs(
      x = "Reduction in basis divergence",
      y = "Delta AUPRC vs cold SuSiE",
      color = NULL,
      title = "Basis movement and AUPRC gain"
    ) +
    theme_effect_paper(base_size = 10) +
    theme(legend.position = "bottom")
  ggsave(
    file.path(diagnostic_dir, "truth_drift_sanity_scatter_simple.png"),
    p_sanity, width = 8.4, height = 4.2, dpi = 220, bg = "white"
  )
  print(p_sanity)
}
```

```{r explanatory-basis-divergence-sanity-3col-scatter, fig.width=12, fig.height=6.8}
# 2-row x 3-col diagnostic. Refinements:
#   - Robust regression (MASS::rlm, Huber M-estimator) instead of OLS so the
#     col-1 slope isn't dominated by 1-2 extreme-leverage outliers.
#   - Axes trimmed to 1%/99% quantiles per column, then expanded with an
#     annotation gutter. Dots are winsorized to the inner trim limits; count
#     labels sit just outside those inner limits, so the labels do not cover
#     dots.
#   - Cell-count labels positioned at the outer edge of each zone. The center
#     cell is moved into the top annotation gutter because its natural position
#     at the origin would sit on top of the densest points.
#   - Bigger dots (size 1.0 -> 1.7) for visibility through the overplotted
#     center.
#   - Col 1 keeps chain color; col 2 colors by baseline AUPRC; col 3 colors
#     by baseline explanatory-basis divergence bin.
if (file.exists(truth_drift_csv) && exists("sanity_join")) {
  delta_truth_thresh <- 0.05
  delta_auprc_thresh <- 0.02
  display_trim_q <- c(0.01, 0.99)
  display_axis_pad <- 0.03

  baseline_context_lookup <- truth_drift_summary %>%
    dplyr::filter(.data$model_method == baseline_method) %>%
    dplyr::transmute(
      dataset_bundle_id = as.character(.data$dataset_bundle_id),
      baseline_basis_divergence = .data[[truth_drift_column]],
      baseline_bin = basis_divergence_bin(.data[[truth_drift_column]])
    ) %>%
    dplyr::inner_join(
      auprc_per_dataset %>%
        dplyr::transmute(
          dataset_bundle_id = as.character(.data$dataset_bundle_id),
          baseline_auprc = .data$auprc_susie_vanilla
        ),
      by = "dataset_bundle_id"
    )

  sanity_join_lab <- sanity_join %>%
    dplyr::mutate(
      pair_kind = factor(.data$pair_kind,
                         levels = c("cold_ann", "cold_warm"),
                         labels = transition_label_map)
    )

  pair_within_dataset <- sanity_join %>%
    dplyr::select(.data$dataset_bundle_id, .data$pair_kind, .data$chain,
                  .data$delta_truth, .data$auprc_delta) %>%
    tidyr::pivot_wider(
      names_from = .data$chain,
      values_from = c(.data$delta_truth, .data$auprc_delta)
    ) %>%
    dplyr::mutate(
      pair_kind = factor(.data$pair_kind,
                         levels = c("cold_ann", "cold_warm"),
                         labels = transition_label_map)
    ) %>%
    dplyr::inner_join(baseline_context_lookup, by = "dataset_bundle_id")

  quantile_lim <- function(x, probs = display_trim_q) {
    x <- x[is.finite(x)]
    if (!length(x)) {
      return(c(-1, 1))
    }
    lim <- stats::quantile(x, probs = probs, na.rm = TRUE, names = FALSE)
    if (!all(is.finite(lim)) || lim[[1]] == lim[[2]]) {
      center <- mean(x, na.rm = TRUE)
      spread <- max(abs(x - center), na.rm = TRUE)
      spread <- if (is.finite(spread) && spread > 0) spread else 1
      lim <- c(center - spread, center + spread)
    }
    lim
  }
  expand_axis_lim <- function(lim, pad = display_axis_pad) {
    width <- diff(lim)
    if (!is.finite(width) || width <= 0) {
      width <- max(abs(lim), 1, na.rm = TRUE)
    }
    lim + c(-pad, pad) * width
  }
  col1_xlim <- quantile_lim(sanity_join_lab$delta_truth)
  col1_ylim <- quantile_lim(sanity_join_lab$auprc_delta)
  col2_axislim <- quantile_lim(c(pair_within_dataset$auprc_delta_pi,
                                 pair_within_dataset$auprc_delta_mu))
  col3_axislim <- quantile_lim(c(pair_within_dataset$delta_truth_pi,
                                 pair_within_dataset$delta_truth_mu))
  col1_plot_xlim <- expand_axis_lim(col1_xlim)
  col1_plot_ylim <- expand_axis_lim(col1_ylim)
  col2_plot_axislim <- expand_axis_lim(col2_axislim)
  col3_plot_axislim <- expand_axis_lim(col3_axislim)

  # Winsorize col-1 data to the trim limits before fitting OLS. The smoother
  # is then fit on the same values that are visually shown at the panel
  # edges; isolated high-leverage outliers (sparse density) get pulled to
  # the trim boundary and lose their disproportionate pull on the slope,
  # while the dense upper-tail points (which already sit inside the limits)
  # retain full influence.
  winsorize <- function(x, lim) {
    pmin(pmax(x, lim[[1]]), lim[[2]])
  }
  sanity_join_lab_wins <- sanity_join_lab %>%
    dplyr::mutate(
      delta_truth = winsorize(.data$delta_truth, col1_xlim),
      auprc_delta = winsorize(.data$auprc_delta, col1_ylim)
    )
  pair_within_dataset_wins <- pair_within_dataset %>%
    dplyr::mutate(
      auprc_delta_pi_plot = winsorize(.data$auprc_delta_pi, col2_axislim),
      auprc_delta_mu_plot = winsorize(.data$auprc_delta_mu, col2_axislim),
      delta_truth_pi_plot = winsorize(.data$delta_truth_pi, col3_axislim),
      delta_truth_mu_plot = winsorize(.data$delta_truth_mu, col3_axislim)
    )

  chain_color_3col <- c(mu = "#2b8cbe", pi = "#74c476")
  zero_lines <- function() {
    list(
      geom_vline(xintercept = 0, color = "grey55", linewidth = 0.25),
      geom_hline(yintercept = 0, color = "grey55", linewidth = 0.25)
    )
  }

  diagonal_count_labels <- function(df, x_col, y_col, lim) {
    df %>%
      dplyr::group_by(.data$pair_kind) %>%
      dplyr::summarise(
        mu_gt_pi = sum(.data[[y_col]] > .data[[x_col]], na.rm = TRUE),
        pi_gt_mu = sum(.data[[y_col]] < .data[[x_col]], na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::mutate(
        x = lim[[1]] * 0.94,
        y = lim[[2]] * 0.92,
        label = sprintf("mu0 > pi: %d\npi > mu0: %d",
                        .data$mu_gt_pi, .data$pi_gt_mu)
      )
  }
  col2_diag_labels <- diagonal_count_labels(
    pair_within_dataset,
    "auprc_delta_pi",
    "auprc_delta_mu",
    col2_axislim
  )
  col3_diag_labels <- diagonal_count_labels(
    pair_within_dataset,
    "delta_truth_pi",
    "delta_truth_mu",
    col3_axislim
  )

  diagonal_count_layer <- function(label_df) {
    geom_label(
      data = label_df,
      aes(x = .data$x, y = .data$y, label = .data$label),
      inherit.aes = FALSE, hjust = 0, vjust = 1,
      size = 2.55, color = "grey20", fill = "white",
      alpha = 0.88, label.size = 0.15,
      label.padding = grid::unit(1.2, "pt")
    )
  }

  p_col1 <- ggplot(
    sanity_join_lab_wins,
    aes(x = .data$delta_truth, y = .data$auprc_delta,
        color = .data$chain)
  ) +
    zero_lines() +
    geom_point(alpha = 0.45, size = 1.7, stroke = 0) +
    geom_smooth(method = "lm", se = TRUE, linewidth = 0.6,
                fill = "grey75", alpha = 0.18) +
    facet_wrap(~ pair_kind, ncol = 1) +
    scale_x_continuous(limits = col1_plot_xlim, oob = scales::squish,
                       expand = expansion(mult = 0)) +
    scale_y_continuous(limits = col1_plot_ylim, oob = scales::squish,
                       expand = expansion(mult = 0)) +
    scale_color_manual(values = chain_color_3col,
                       labels = chain_label_map) +
    labs(
      x = basis_reduction_axis,
      y = "Delta AUPRC vs cold SuSiE",
      color = NULL,
      title = "Basis movement predicts AUPRC gain"
    ) +
    theme_effect_paper(base_size = 9)

  p_col2 <- ggplot(
    pair_within_dataset_wins,
    aes(x = .data$auprc_delta_pi_plot, y = .data$auprc_delta_mu_plot,
        color = .data$baseline_bin)
  ) +
    zero_lines() +
    geom_abline(slope = 1, intercept = 0, color = "grey45",
                linetype = "dotted", linewidth = 0.3) +
    geom_point(alpha = 0.55, size = 1.7, stroke = 0) +
    diagonal_count_layer(col2_diag_labels) +
    facet_wrap(~ pair_kind, ncol = 1) +
    scale_x_continuous(limits = col2_plot_axislim, oob = scales::squish,
                       expand = expansion(mult = 0)) +
    scale_y_continuous(limits = col2_plot_axislim, oob = scales::squish,
                       expand = expansion(mult = 0)) +
    scale_color_manual(
      values = baseline_bin_palette,
      name = "Baseline basis divergence",
      guide = guide_legend(title.position = "top", nrow = 1)
    ) +
    labs(
      x = "Delta AUPRC (pi channel)",
      y = "Delta AUPRC (mu0 channel)",
      title = "AUPRC gain: mu0 vs pi"
    ) +
    theme_effect_paper(base_size = 9)

  p_col3 <- ggplot(
    pair_within_dataset_wins,
    aes(x = .data$delta_truth_pi_plot, y = .data$delta_truth_mu_plot,
        color = .data$baseline_bin)
  ) +
    zero_lines() +
    geom_abline(slope = 1, intercept = 0, color = "grey45",
                linetype = "dotted", linewidth = 0.3) +
    geom_point(alpha = 0.55, size = 1.7, stroke = 0) +
    diagonal_count_layer(col3_diag_labels) +
    facet_wrap(~ pair_kind, ncol = 1) +
    scale_x_continuous(limits = col3_plot_axislim, oob = scales::squish,
                       expand = expansion(mult = 0)) +
    scale_y_continuous(limits = col3_plot_axislim, oob = scales::squish,
                       expand = expansion(mult = 0)) +
    scale_color_manual(
      values = baseline_bin_palette,
      name = "Baseline basis divergence",
      guide = guide_legend(title.position = "top", nrow = 1)
    ) +
    labs(
      x = "Basis-divergence reduction (pi channel)",
      y = "Basis-divergence reduction (mu0 channel)",
      title = "Basin movement: mu0 vs pi"
    ) +
    theme_effect_paper(base_size = 9)

  combined_3col <- patchwork::wrap_plots(
    p_col2,
    p_col3,
    p_col1,
    ncol = 3,
    widths = c(1, 1, 1),
    guides = "collect"
  ) &
    theme(legend.position = "bottom",
          legend.justification = "center",
          legend.box = "horizontal")

  ggsave(
    file.path(diagnostic_dir, "truth_drift_sanity_3col_scatter.png"),
    combined_3col, width = 12, height = 6.8, dpi = 220, bg = "white"
  )
  ggsave(
    file.path(diagnostic_dir, "truth_drift_sanity_scatter.png"),
    combined_3col, width = 12, height = 6.8, dpi = 220, bg = "white"
  )
  print(combined_3col)

  cat("Display winsorization limits (1st/99th percentiles):\n")
  cat(sprintf("  Panel E x (basis-divergence reduction): [%.3f, %.3f]\n",
              col1_xlim[1], col1_xlim[2]))
  cat(sprintf("  Panel E y (delta AUPRC): [%.3f, %.3f]\n",
              col1_ylim[1], col1_ylim[2]))
  cat(sprintf("  Panel C axes (delta AUPRC): [%.3f, %.3f]\n",
              col2_axislim[1], col2_axislim[2]))
  cat(sprintf("  Panel D axes (basis-divergence reduction): [%.3f, %.3f]\n",
              col3_axislim[1], col3_axislim[2]))
} else {
  if (exists("combined_3col")) {
    rm(combined_3col)
  }
  message("Skipping 3-col scatter; sanity_join not yet built.")
}
```

```{r explanatory-basis-divergence-sanity-3col-violin, fig.width=12, fig.height=6.8}
# Variant of the 3-col layout above. Col 1 unchanged; cols 2 and 3 swap the
# per-dataset paired scatters for violin + boxplot summaries by chain. Loses
# pairing info, gains a cleaner read on per-chain marginal distributions.
if (file.exists(truth_drift_csv) && exists("sanity_join")) {
  sanity_join_lab2 <- sanity_join %>%
    dplyr::mutate(
      pair_kind = factor(.data$pair_kind,
                         levels = c("cold_ann", "cold_warm"),
                         labels = transition_label_map),
      chain = factor(.data$chain, levels = c("mu", "pi"))
    )

  chain_color_3col <- c(mu = "#2b8cbe", pi = "#74c476")

  p_col1_v <- ggplot(
    sanity_join_lab2,
    aes(x = .data$delta_truth, y = .data$auprc_delta,
        color = .data$chain)
  ) +
    geom_hline(yintercept = 0, color = "grey60", linewidth = 0.25) +
    geom_vline(xintercept = 0, color = "grey60", linewidth = 0.25) +
    geom_point(alpha = 0.5, size = 1.2, stroke = 0) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.5) +
    facet_wrap(~ pair_kind, ncol = 1) +
    scale_color_manual(values = chain_color_3col,
                       labels = chain_label_map) +
    labs(
      x = basis_reduction_axis,
      y = "Delta AUPRC vs cold SuSiE",
      color = NULL,
      title = "Basis movement and AUPRC gain"
    ) +
    theme_effect_paper(base_size = 9)

  p_col2_v <- ggplot(
    sanity_join_lab2,
    aes(x = .data$chain, y = .data$auprc_delta, fill = .data$chain)
  ) +
    geom_hline(yintercept = 0, color = "grey60", linewidth = 0.25) +
    geom_violin(alpha = 0.55, color = NA, scale = "width") +
    geom_boxplot(width = 0.18, alpha = 0.4, fill = "white",
                 color = "grey20", outlier.size = 0.4) +
    facet_wrap(~ pair_kind, ncol = 1) +
    scale_fill_manual(values = chain_color_3col, guide = "none") +
    scale_x_discrete(labels = c(mu = "mu0", pi = "pi")) +
    labs(x = NULL, y = "Delta AUPRC vs cold SuSiE",
         title = "AUPRC gain by channel") +
    theme_effect_paper(base_size = 9)

  p_col3_v <- ggplot(
    sanity_join_lab2,
    aes(x = .data$chain, y = .data$delta_truth, fill = .data$chain)
  ) +
    geom_hline(yintercept = 0, color = "grey60", linewidth = 0.25) +
    geom_violin(alpha = 0.55, color = NA, scale = "width") +
    geom_boxplot(width = 0.18, alpha = 0.4, fill = "white",
                 color = "grey20", outlier.size = 0.4) +
    facet_wrap(~ pair_kind, ncol = 1) +
    scale_fill_manual(values = chain_color_3col, guide = "none") +
    scale_x_discrete(labels = c(mu = "mu0", pi = "pi")) +
    labs(x = NULL,
         y = basis_reduction_axis,
         title = "Basis movement by channel") +
    theme_effect_paper(base_size = 9)

  combined_3col_v <- (p_col1_v | p_col2_v | p_col3_v) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")

  ggsave(
    file.path(diagnostic_dir, "truth_drift_sanity_3col_violin.png"),
    combined_3col_v, width = 12, height = 6.8, dpi = 220, bg = "white"
  )
  print(combined_3col_v)
} else {
  message("Skipping 3-col violin; sanity_join not yet built.")
}
```

# Baseline Cold Vanilla on (explanatory basis divergence, AUPRC)

```{r explanatory-basis-divergence-baseline-only, fig.width=6.4, fig.height=4.4}
# Foundation plot for the trajectory and curve-residual analyses: cold
# SuSiE-vanilla AUPRC vs explanatory basis divergence, one point per dataset.
# The low-df smoothing spline is reused below to ask how much AUPRC movement
# is predicted by movement along the baseline divergence -> AUPRC curve.
if (file.exists(truth_drift_csv)) {
  baseline_points <- truth_drift_summary %>%
    dplyr::filter(.data$model_method == baseline_method) %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      truth_drift = .data[[truth_drift_column]]
    ) %>%
    dplyr::inner_join(
      auprc_per_dataset %>%
        dplyr::transmute(
          dataset_bundle_id = .data$dataset_bundle_id,
          auprc = .data$auprc_susie_vanilla
        ),
      by = "dataset_bundle_id"
    )

  cat("Baseline (cold vanilla) points: ", nrow(baseline_points), "\n", sep = "")
  cat(sprintf(
    "  explanatory_basis_divergence: median %.3f, mean %.3f, range [%.3f, %.3f]\n",
    stats::median(baseline_points$truth_drift, na.rm = TRUE),
    mean(baseline_points$truth_drift, na.rm = TRUE),
    min(baseline_points$truth_drift, na.rm = TRUE),
    max(baseline_points$truth_drift, na.rm = TRUE)
  ))
  cat(sprintf(
    "  auprc:       median %.3f, mean %.3f, range [%.3f, %.3f]\n",
    stats::median(baseline_points$auprc, na.rm = TRUE),
    mean(baseline_points$auprc, na.rm = TRUE),
    min(baseline_points$auprc, na.rm = TRUE),
    max(baseline_points$auprc, na.rm = TRUE)
  ))

  baseline_curve_df <- baseline_points %>%
    dplyr::filter(is.finite(.data$truth_drift), is.finite(.data$auprc)) %>%
    dplyr::arrange(.data$truth_drift) %>%
    dplyr::mutate(divergence_bin = basis_divergence_bin(.data$truth_drift))
  baseline_curve_fit <- stats::smooth.spline(
    x = baseline_curve_df$truth_drift,
    y = baseline_curve_df$auprc,
    df = truth_drift_spline_df
  )
  baseline_curve_fitted <- stats::predict(
    baseline_curve_fit,
    x = baseline_curve_df$truth_drift
  )$y
  baseline_curve_r2 <- 1 - sum(
    (baseline_curve_df$auprc - baseline_curve_fitted)^2,
    na.rm = TRUE
  ) / sum(
    (baseline_curve_df$auprc - mean(baseline_curve_df$auprc, na.rm = TRUE))^2,
    na.rm = TRUE
  )
  predict_baseline_auprc_from_divergence <- function(x) {
    pred <- stats::predict(baseline_curve_fit, x = x)$y
    pmin(pmax(pred, 0), 1)
  }
  baseline_curve_grid <- tibble::tibble(
    truth_drift = seq(
      min(baseline_curve_df$truth_drift, na.rm = TRUE),
      max(baseline_curve_df$truth_drift, na.rm = TRUE),
      length.out = 400L
    )
  ) %>%
    dplyr::mutate(
      auprc_hat = predict_baseline_auprc_from_divergence(.data$truth_drift)
    )

  p_baseline_truth <- ggplot(
    baseline_curve_df,
    aes(x = .data$truth_drift, y = .data$auprc,
        color = .data$divergence_bin)
  ) +
    geom_point(alpha = 0.52, size = 1.5, stroke = 0) +
    geom_line(
      data = baseline_curve_grid,
      aes(x = .data$truth_drift, y = .data$auprc_hat),
      inherit.aes = FALSE,
      color = "#c44e52", linewidth = 0.7
    ) +
    annotate(
      "label",
      x = stats::quantile(baseline_curve_df$truth_drift, 0.72,
                          na.rm = TRUE, names = FALSE),
      y = stats::quantile(baseline_curve_df$auprc, 0.88,
                          na.rm = TRUE, names = FALSE),
      label = sprintf("Smoothing spline df=%d\nR^2 = %.3f",
                      truth_drift_spline_df, baseline_curve_r2),
      hjust = 0, vjust = 1, size = 3.0, label.size = 0.2,
      fill = "white", color = "grey20", alpha = 0.9
    ) +
    scale_color_manual(
      values = baseline_bin_palette,
      name = "Baseline basis divergence"
    ) +
    labs(
      x = basis_divergence_axis,
      y = "AUPRC (cold SuSiE)",
      title = "Baseline AUPRC vs basis divergence",
      subtitle = sprintf("n = %d datasets; spline reused for decomposition.",
                         nrow(baseline_points)),
      color = NULL
    ) +
    theme_effect_paper(base_size = 10) +
    theme(legend.position = "bottom")

  ggsave(
    file.path(diagnostic_dir, "truth_drift_baseline_only.png"),
    p_baseline_truth, width = 6.4, height = 4.4, dpi = 220, bg = "white"
  )
  print(p_baseline_truth)
} else {
  message("Skipping baseline-only plot; truth_drift_summary not cached yet.")
}
```

```{r explanatory-basis-divergence-baseline-boxplot, fig.width=6.8, fig.height=4.4}
# Cold vanilla AUPRC distribution stratified into three basis-divergence bins.
# Bin edges match the bins used for centroid stratification in the quadrant
# trajectory plot below.
if (exists("baseline_points")) {
  baseline_binned <- baseline_points %>%
    dplyr::mutate(
      divergence_bin = basis_divergence_bin(.data$truth_drift)
    )

  bin_counts <- baseline_binned %>%
    dplyr::count(.data$divergence_bin, name = "n")
  bin_stats <- baseline_binned %>%
    dplyr::group_by(.data$divergence_bin) %>%
    dplyr::summarise(
      n = dplyr::n(),
      median = stats::median(.data$auprc, na.rm = TRUE),
      mean = mean(.data$auprc, na.rm = TRUE),
      q25 = stats::quantile(.data$auprc, 0.25, na.rm = TRUE,
                            names = FALSE),
      q75 = stats::quantile(.data$auprc, 0.75, na.rm = TRUE,
                            names = FALSE),
      .groups = "drop"
    )
  cat("Cold baseline AUPRC by basis divergence bin:\n")
  print(bin_stats)

  y_label_top <- max(baseline_binned$auprc, na.rm = TRUE) * 1.06
  p_baseline_boxplot <- ggplot(
    baseline_binned,
    aes(x = .data$divergence_bin, y = .data$auprc,
        fill = .data$divergence_bin)
  ) +
    geom_boxplot(alpha = 0.7, outlier.size = 0.7, width = 0.6) +
    geom_text(
      data = bin_counts,
      aes(x = .data$divergence_bin, y = y_label_top,
          label = sprintf("n=%d", .data$n)),
      inherit.aes = FALSE, size = 3.2, color = "grey25"
    ) +
    scale_fill_manual(values = baseline_bin_palette, guide = "none") +
    scale_color_manual(values = baseline_bin_palette, guide = "none") +
    labs(
      x = "Baseline basis-divergence bin",
      y = "AUPRC (cold SuSiE)",
      title = "Baseline AUPRC by divergence bin",
      subtitle = "Same bins used throughout this figure"
    ) +
    theme_effect_paper(base_size = 10)

  ggsave(
    file.path(diagnostic_dir, "truth_drift_baseline_boxplot.png"),
    p_baseline_boxplot, width = 6.8, height = 4.4, dpi = 220,
    bg = "white"
  )
  print(p_baseline_boxplot)
} else {
  message("Skipping baseline boxplot; baseline_points not built yet.")
}
```

# Explanatory Basis Divergence Trajectory (cold -> annotated -> warm)

```{r explanatory-basis-divergence-trajectory, fig.width=10, fig.height=4.8}
# Per-dataset trajectories on the (explanatory basis divergence, AUPRC) plane. Each dataset
# contributes three points per chain (cold, annotated, warm) connected by
# a path. Faded individual paths show the cloud; bold centroid arrows show
# the mean trajectory per chain.
if (file.exists(truth_drift_csv)) {
  # AUPRC at each state per chain. auprc_per_dataset already has raw
  # per-method AUPRC columns from the analysis workbook.
  trajectory_auprc <- dplyr::bind_rows(
    auprc_per_dataset %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        chain = "mu",
        cold = .data$auprc_susie_vanilla,
        annotated = .data$auprc_susine_functional_mu,
        warm = .data$auprc_susie_vanilla_warm_from_mu
      ),
    auprc_per_dataset %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        chain = "pi",
        cold = .data$auprc_susie_vanilla,
        annotated = .data$auprc_susie_functional_pi,
        warm = .data$auprc_susie_vanilla_warm_from_pi
      )
  ) %>%
    tidyr::pivot_longer(
      cols = c(cold, annotated, warm),
      names_to = "state",
      values_to = "auprc"
    )

  # Explanatory basis divergence at each state per chain, using the same metric as the
  # sanity scatter (controlled by truth_drift_column above).
  state_method_map <- tibble::tribble(
    ~chain, ~state,       ~model_method,
    "mu",   "cold",       baseline_method,
    "mu",   "annotated",  mu_method,
    "mu",   "warm",       warm_mu_method,
    "pi",   "cold",       baseline_method,
    "pi",   "annotated",  pi_method,
    "pi",   "warm",       warm_pi_method
  )

  trajectory_drift <- truth_drift_summary %>%
    dplyr::select(.data$dataset_bundle_id, .data$model_method,
                  dplyr::all_of(truth_drift_column)) %>%
    dplyr::rename(truth_drift = dplyr::all_of(truth_drift_column)) %>%
    dplyr::inner_join(state_method_map, by = "model_method",
                      relationship = "many-to-many")

  trajectory_long <- trajectory_auprc %>%
    dplyr::inner_join(
      trajectory_drift %>%
        dplyr::select(.data$dataset_bundle_id, .data$chain, .data$state,
                      .data$truth_drift),
      by = c("dataset_bundle_id", "chain", "state")
    ) %>%
    dplyr::mutate(
      state = factor(.data$state, levels = c("cold", "annotated", "warm")),
      chain = factor(.data$chain, levels = c("mu", "pi"))
    )

  # Centroid trajectory per chain (mean across datasets at each state).
  centroids <- trajectory_long %>%
    dplyr::group_by(.data$chain, .data$state) %>%
    dplyr::summarise(
      truth_drift = mean(.data$truth_drift, na.rm = TRUE),
      auprc = mean(.data$auprc, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::arrange(.data$chain, .data$state)

  chain_color_diag2 <- c(mu = "#2b8cbe", pi = "#74c476")
  state_fill <- c(cold = "grey20", annotated = "#c44e52", warm = "#4c72b0")
  state_shape <- c(cold = 21, annotated = 22, warm = 23)

  p_trajectory <- ggplot() +
    geom_path(
      data = trajectory_long,
      aes(x = .data$truth_drift, y = .data$auprc,
          group = interaction(.data$dataset_bundle_id, .data$chain),
          color = .data$chain),
      alpha = 0.05, linewidth = 0.2
    ) +
    geom_point(
      data = trajectory_long,
      aes(x = .data$truth_drift, y = .data$auprc,
          color = .data$chain, shape = .data$state),
      alpha = 0.10, size = 0.7, stroke = 0
    ) +
    geom_path(
      data = centroids,
      aes(x = .data$truth_drift, y = .data$auprc, group = .data$chain),
      arrow = grid::arrow(length = grid::unit(2.5, "mm"), type = "closed"),
      linewidth = 0.9, color = "black"
    ) +
    geom_point(
      data = centroids,
      aes(x = .data$truth_drift, y = .data$auprc,
          fill = .data$state, shape = .data$state),
      size = 3.2, color = "black", stroke = 0.4
    ) +
    facet_wrap(~ chain, nrow = 1,
               labeller = labeller(chain = chain_label_map)) +
    scale_color_manual(values = chain_color_diag2, guide = "none") +
    scale_fill_manual(values = state_fill, name = "State (centroid)") +
    scale_shape_manual(values = state_shape, name = "State") +
    labs(
      x = basis_divergence_axis,
      y = "AUPRC",
      title = "Per-dataset trajectories on the basis-divergence axis",
      subtitle = paste(
        "Faded paths = individual datasets (cold -> annotated -> warm refit);",
        "black arrow = chain centroid trajectory"
      )
    ) +
    theme_effect_paper(base_size = 10) +
    theme(legend.position = "bottom")

  ggsave(
    file.path(diagnostic_dir, "truth_drift_trajectory_static.png"),
    p_trajectory, width = 10, height = 4.8, dpi = 220, bg = "white"
  )
  print(p_trajectory)

  # Centroid coordinates printed so the directional reading is unambiguous
  # without zooming on the plot.
  cat("Centroid (explanatory_basis_divergence, AUPRC) per (chain, state):\n")
  print(centroids %>%
          dplyr::mutate(
            truth_drift = round(.data$truth_drift, 4),
            auprc = round(.data$auprc, 4)
          ))
} else {
  message("Skipping trajectory plot; truth_drift_summary not cached yet.")
}
```

```{r explanatory-basis-divergence-auprc-decomposition, fig.width=11.5, fig.height=7.2}
# Nonlinear decomposition of AUPRC changes. The basis-driven component is the
# AUPRC change predicted by moving along the cold-vanilla baseline curve:
#   f(end divergence) - f(cold divergence).
# The residual is the remaining AUPRC delta after that nonlinear baseline
# relationship is accounted for.
if (exists("trajectory_long") &&
    exists("predict_baseline_auprc_from_divergence")) {
  trajectory_wide <- trajectory_long %>%
    dplyr::select(.data$dataset_bundle_id, .data$chain, .data$state,
                  .data$truth_drift, .data$auprc) %>%
    tidyr::pivot_wider(
      names_from = .data$state,
      values_from = c(.data$truth_drift, .data$auprc)
    )

  auprc_basis_decomp <- dplyr::bind_rows(
    trajectory_wide %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        chain = .data$chain,
        transition = "cold_ann",
        start_divergence = .data$truth_drift_cold,
        end_divergence = .data$truth_drift_annotated,
        start_auprc = .data$auprc_cold,
        end_auprc = .data$auprc_annotated
      ),
    trajectory_wide %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        chain = .data$chain,
        transition = "cold_warm",
        start_divergence = .data$truth_drift_cold,
        end_divergence = .data$truth_drift_warm,
        start_auprc = .data$auprc_cold,
        end_auprc = .data$auprc_warm
      )
  ) %>%
    dplyr::mutate(
      observed_delta = .data$end_auprc - .data$start_auprc,
      divergence_improvement = .data$start_divergence - .data$end_divergence,
      curve_start = predict_baseline_auprc_from_divergence(
        .data$start_divergence
      ),
      curve_end = predict_baseline_auprc_from_divergence(
        .data$end_divergence
      ),
      basis_pred_delta = .data$curve_end - .data$curve_start,
      residual_delta = .data$observed_delta - .data$basis_pred_delta,
      start_divergence_bin = basis_divergence_bin(.data$start_divergence),
      basis_share = dplyr::if_else(
        abs(.data$observed_delta) >= 0.01,
        .data$basis_pred_delta / .data$observed_delta,
        NA_real_
      ),
      chain = factor(.data$chain, levels = c("mu", "pi"),
                     labels = c("mu0 channel", "pi channel")),
      transition = factor(
        .data$transition,
        levels = c("cold_ann", "cold_warm"),
        labels = transition_label_map
      )
    ) %>%
    dplyr::filter(
      is.finite(.data$observed_delta),
      is.finite(.data$basis_pred_delta),
      is.finite(.data$residual_delta)
    ) %>%
    dplyr::group_by(.data$chain, .data$transition) %>%
    dplyr::mutate(
      transition_intercept = mean(.data$residual_delta, na.rm = TRUE),
      basis_plus_intercept_delta = .data$basis_pred_delta +
        .data$transition_intercept,
      residual_after_intercept = .data$observed_delta -
        .data$basis_plus_intercept_delta
    ) %>%
    dplyr::ungroup()

  auprc_basis_decomp_summary <- auprc_basis_decomp %>%
    dplyr::group_by(.data$chain, .data$transition) %>%
    dplyr::summarise(
      n = dplyr::n(),
      observed_mean = mean(.data$observed_delta, na.rm = TRUE),
      basis_mean = mean(.data$basis_pred_delta, na.rm = TRUE),
      transition_intercept = dplyr::first(.data$transition_intercept),
      fitted_mean = mean(.data$basis_plus_intercept_delta, na.rm = TRUE),
      residual_after_intercept_mean = mean(
        .data$residual_after_intercept,
        na.rm = TRUE
      ),
      basis_share_of_mean = .data$basis_mean / .data$observed_mean,
      median_basis_share_when_gain = stats::median(
        .data$basis_share[.data$observed_delta >= 0.01],
        na.rm = TRUE
      ),
      mean_divergence_improvement = mean(
        .data$divergence_improvement,
        na.rm = TRUE
      ),
      basis_r2_with_transition_intercept = {
        sst <- sum(
          (.data$observed_delta - mean(.data$observed_delta, na.rm = TRUE))^2,
          na.rm = TRUE
        )
        sse <- sum(
          (.data$observed_delta - .data$basis_plus_intercept_delta)^2,
          na.rm = TRUE
        )
        if (is.finite(sst) && sst > 0) 1 - sse / sst else NA_real_
      },
      basis_observed_cor = stats::cor(
        .data$basis_pred_delta,
        .data$observed_delta,
        use = "complete.obs"
      ),
      .groups = "drop"
    )
  readr::write_csv(
    auprc_basis_decomp,
    file.path(diagnostic_dir,
              "truth_drift_auprc_basis_decomposition_rows.csv")
  )
  readr::write_csv(
    auprc_basis_decomp_summary,
    file.path(diagnostic_dir,
              "truth_drift_auprc_basis_decomposition_summary.csv")
  )
  cat("AUPRC decomposition using baseline nonlinear divergence curve:\n")
  print(auprc_basis_decomp_summary)

  decomp_lim <- max(abs(c(
    auprc_basis_decomp$observed_delta,
    auprc_basis_decomp$basis_pred_delta
  )), na.rm = TRUE)
  decomp_lim <- max(0.05, decomp_lim)
  decomp_plot_lim <- 0.4

  auprc_basis_decomp_plot <- auprc_basis_decomp %>%
    dplyr::mutate(
      basis_pred_delta_plot = pmin(
        pmax(.data$basis_pred_delta, -decomp_plot_lim),
        decomp_plot_lim
      ),
      observed_delta_plot = pmin(
        pmax(.data$observed_delta, -decomp_plot_lim),
        decomp_plot_lim
      )
    )

  decomp_annotation <- auprc_basis_decomp_summary %>%
    dplyr::mutate(
      x = -0.96 * decomp_plot_lim,
      y = 0.94 * decomp_plot_lim,
      label = sprintf(
        paste0(
          "R^2 = %.2f\n",
          "observed mean = %.3f\n",
          "basis shift = %.3f\n",
          "non-basis shift = %.3f"
        ),
        .data$basis_r2_with_transition_intercept,
        .data$observed_mean,
        .data$basis_mean,
        .data$transition_intercept
      )
    )

  p_basis_explain_scatter <- ggplot(
    auprc_basis_decomp_plot,
    aes(x = .data$basis_pred_delta_plot, y = .data$observed_delta_plot,
        color = .data$start_divergence_bin)
  ) +
    geom_hline(yintercept = 0, color = "grey65", linewidth = 0.25) +
    geom_vline(xintercept = 0, color = "grey65", linewidth = 0.25) +
    geom_abline(slope = 1, intercept = 0, color = "grey35",
                linetype = "dashed", linewidth = 0.35) +
    geom_abline(
      data = auprc_basis_decomp_summary,
      aes(intercept = .data$transition_intercept, slope = 1),
      inherit.aes = FALSE, color = "#c44e52", linewidth = 0.45
    ) +
    geom_point(alpha = 0.58, size = 1.45, stroke = 0) +
    geom_label(
      data = decomp_annotation,
      aes(x = .data$x, y = .data$y, label = .data$label),
      inherit.aes = FALSE, hjust = 0, vjust = 1,
      size = 2.55, label.size = 0.16, fill = "white",
      color = "grey18", alpha = 0.9
    ) +
    facet_grid(chain ~ transition) +
    scale_color_manual(
      values = baseline_bin_palette,
      name = "Baseline basis\ndivergence"
    ) +
    coord_cartesian(xlim = c(-decomp_plot_lim, decomp_plot_lim),
                    ylim = c(-decomp_plot_lim, decomp_plot_lim)) +
    labs(
      x = "Predicted delta AUPRC from basis movement",
      y = "Observed delta AUPRC",
      title = "Decomposing AUPRC gains"
    ) +
    theme_effect_paper(base_size = 9) +
    theme(legend.position = "bottom")

  ggsave(
    file.path(diagnostic_dir,
              "truth_drift_auprc_basis_decomposition.png"),
    p_basis_explain_scatter, width = 11.5, height = 6.2, dpi = 220,
    bg = "white"
  )
  print(p_basis_explain_scatter)
} else {
  message(
    "Skipping AUPRC basis decomposition; need trajectory_long and ",
    "baseline divergence smoother."
  )
}
```

```{r explanatory-basis-divergence-paper-figure, fig.width=12, fig.height=16.5}
if (exists("p_baseline_truth") &&
    exists("p_baseline_boxplot") &&
    exists("combined_3col") &&
    exists("p_basis_explain_scatter")) {
  paper_basis_divergence_fig <- patchwork::wrap_plots(
    (p_baseline_truth | p_baseline_boxplot),
    combined_3col,
    p_basis_explain_scatter,
    ncol = 1,
    heights = c(0.78, 1.18, 1.08)
  ) +
    plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(face = "bold", size = 11))

  ggsave(
    paper_basis_divergence_png,
    paper_basis_divergence_fig,
    width = 12, height = 16.5, dpi = 300, bg = "white",
    limitsize = FALSE
  )
  cat("Wrote basis-divergence paper figure to ",
      paper_basis_divergence_png, "\n", sep = "")
  print(paper_basis_divergence_fig)
} else {
  message(
    "Skipping basis-divergence paper figure; need baseline, 3-column, ",
    "and decomposition panels."
  )
}
```

```{r explanatory-basis-divergence-trajectory-quadrant, fig.width=11.5, fig.height=5.2}
# Revised trajectory plot:
#   - thicker, less faded per-dataset paths
#   - paths colored by the cold -> warm net (AUPRC, divergence) quadrant
#   - centroid arrows stratified by the dataset's cold (baseline) basis
#     divergence bin (3 centroid trajectories per chain facet)
if (exists("trajectory_long")) {
  # Per (dataset, chain): the cold -> warm net (AUPRC, divergence) quadrant
  # and the dataset's baseline (cold) divergence bin.
  net_movement <- trajectory_long %>%
    dplyr::filter(.data$state %in% c("cold", "warm")) %>%
    dplyr::select(.data$dataset_bundle_id, .data$chain, .data$state,
                  .data$truth_drift, .data$auprc) %>%
    tidyr::pivot_wider(
      names_from = .data$state,
      values_from = c(.data$truth_drift, .data$auprc)
    ) %>%
    dplyr::mutate(
      delta_auprc = .data$auprc_warm - .data$auprc_cold,
      delta_divergence = .data$truth_drift_warm - .data$truth_drift_cold,
      quadrant = dplyr::case_when(
        .data$delta_auprc >= 0 & .data$delta_divergence <= 0 ~
          "+AUPRC, -divergence",
        .data$delta_auprc >= 0 & .data$delta_divergence > 0  ~
          "+AUPRC, +divergence",
        .data$delta_auprc < 0  & .data$delta_divergence <= 0 ~
          "-AUPRC, -divergence",
        .data$delta_auprc < 0  & .data$delta_divergence > 0  ~
          "-AUPRC, +divergence"
      ),
      baseline_bin = basis_divergence_bin(.data$truth_drift_cold)
    )

  trajectory_long_q <- trajectory_long %>%
    dplyr::inner_join(
      net_movement %>%
        dplyr::select(.data$dataset_bundle_id, .data$chain,
                      .data$quadrant, .data$baseline_bin),
      by = c("dataset_bundle_id", "chain")
    ) %>%
    dplyr::mutate(
      quadrant = factor(.data$quadrant, levels = c(
        "+AUPRC, -divergence", "+AUPRC, +divergence",
        "-AUPRC, -divergence", "-AUPRC, +divergence"
      ))
    )

  # Quadrant counts per (chain, baseline_bin).
  quad_bin_counts <- trajectory_long_q %>%
    dplyr::filter(.data$state == "cold") %>%
    dplyr::count(.data$chain, .data$baseline_bin, .data$quadrant)
  cat("Quadrant counts per (chain, baseline_bin) [cold -> warm net]:\n")
  print(quad_bin_counts)

  # Centroid trajectory per (chain, baseline_bin).
  centroids_binned <- trajectory_long_q %>%
    dplyr::group_by(.data$chain, .data$baseline_bin, .data$state) %>%
    dplyr::summarise(
      truth_drift = mean(.data$truth_drift, na.rm = TRUE),
      auprc = mean(.data$auprc, na.rm = TRUE),
      n = dplyr::n(),
      .groups = "drop"
    ) %>%
    dplyr::arrange(.data$chain, .data$baseline_bin, .data$state)

  quadrant_palette <- c(
    "+AUPRC, -divergence" = "#1b9e77",
    "+AUPRC, +divergence" = "#7570b3",
    "-AUPRC, -divergence" = "#e7298a",
    "-AUPRC, +divergence" = "#d95f02"
  )
  bin_linetype <- c(
    "[0, 0.05]"    = "solid",
    "(0.05, 0.10]" = "longdash",
    "(0.10, inf)"  = "dotted"
  )
  state_fill <- c(cold = "grey20", annotated = "#c44e52",
                  warm = "#4c72b0")

  p_trajectory_quadrant <- ggplot() +
    # Per-dataset paths, colored by cold->warm quadrant
    geom_path(
      data = trajectory_long_q,
      aes(x = .data$truth_drift, y = .data$auprc,
          group = interaction(.data$dataset_bundle_id, .data$chain),
          color = .data$quadrant),
      alpha = 0.18, linewidth = 0.5
    ) +
    scale_color_manual(values = quadrant_palette,
                       name = "Cold -> warm quadrant") +
    # Centroid arrows per baseline_bin (linetype distinguishes bin so the
    # color channel stays reserved for the quadrant scale above).
    geom_path(
      data = centroids_binned,
      aes(x = .data$truth_drift, y = .data$auprc,
          group = .data$baseline_bin,
          linetype = .data$baseline_bin),
      arrow = grid::arrow(length = grid::unit(2.5, "mm"),
                          type = "closed"),
      linewidth = 0.95,
      color = "black"
    ) +
    scale_linetype_manual(values = bin_linetype,
                          name = "Baseline bin (centroid)") +
    # Centroid points: shape = state, fill = state
    geom_point(
      data = centroids_binned,
      aes(x = .data$truth_drift, y = .data$auprc,
          fill = .data$state),
      size = 3.0, shape = 21, color = "black", stroke = 0.4
    ) +
    scale_fill_manual(values = state_fill, name = "State (centroid)") +
    facet_wrap(~ chain, nrow = 1,
               labeller = labeller(chain = chain_label_map)) +
    labs(
      x = basis_divergence_axis,
      y = "AUPRC",
      title = paste(
        "Trajectories colored by cold-to-warm outcome;",
        "centroids stratified by baseline divergence bin"
      ),
      subtitle = paste(
        "Faded paths: per-dataset cold -> annotated -> warm refit, colored by",
        "net (AUPRC, divergence) sign.",
        "Black arrows: chain centroid trajectory per baseline bin (linetype)."
      )
    ) +
    theme_effect_paper(base_size = 10) +
    theme(legend.position = "right",
          legend.box = "vertical")

  ggsave(
    file.path(diagnostic_dir, "truth_drift_trajectory_quadrant.png"),
    p_trajectory_quadrant, width = 11.5, height = 5.2, dpi = 220,
    bg = "white"
  )
  print(p_trajectory_quadrant)

  cat("Centroid coords per (chain, baseline_bin, state):\n")
  print(centroids_binned %>%
          dplyr::mutate(
            truth_drift = round(.data$truth_drift, 4),
            auprc = round(.data$auprc, 4)
          ))
} else {
  message(
    "Skipping quadrant trajectory; trajectory_long not in scope. ",
    "Run the explanatory-basis-divergence trajectory chunk first."
  )
}
```
