---
title: "Effect Drift Analysis (5-arm: cold + warm refits)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Reads the 5-arm refit records and computes per-dataset **effect drift**
between every interesting arm pair, plus an AUPRC trajectory across the
two refit chains:

```
mu chain:                              pi chain:
  vanilla -> mu                          vanilla -> pi
  mu -> vanilla_warm_from_mu             pi -> vanilla_warm_from_pi
  vanilla -> vanilla_warm_from_mu        vanilla -> vanilla_warm_from_pi
```

**Effect drift** between two fits A and B is computed by:

1. Build the `L x L` correlation matrix of `fitted_y_effect` vectors
   (inactive effects with sd=0 contribute correlation 0).
2. Solve the max-weight 1-1 assignment via `clue::solve_LSAP(cor_mat,
   maximum = TRUE)`.
3. Record all `L` assignment correlations.
4. Derive per-pair summaries: `mean_corr`, `min_corr`, `n_below_0p9`
   (count of assignment edges with correlation < 0.9), and
   `weighted_drift` (1 - assignment correlations weighted by
   `min(var_y_explained_A_i, var_y_explained_B_j)` so inactive pairs
   contribute little).

This drift framework + AUPRC trajectory operationalizes the
**refinement-vs-exploration test**:

| D(vanilla, treatment) | D(vanilla, vanilla_warm) | dAUPRC(treatment) | dAUPRC(vanilla_warm) | Mechanism |
|---|---|---|---|---|
| Low | Low | + | + | Within-basin PIP weighting; basin unchanged |
| Low | Low | + | ~0 | PIP-weighting only; reverts when prior removed |
| High | High | + | + | **Exploration win**: new basin, vanilla holds it |
| High | Low | + | ~0 | New basin but unstable; vanilla snaps back |

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_drift_analysis_5arm.Rmd")
library(here)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)
library(clue)         # solve_LSAP for max-weight 1-1 assignment

theme_set(theme_bw(base_size = 12))
```

# 1. User Parameters

```{r user-parameters}
output_root <- here("output")
study_name <- "baseline_sims_5arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
records_rds <- file.path(study_dir, "effect_refit_records.rds")

baseline_method <- "susie_vanilla"
mu_method <- "susine_functional_mu"
pi_method <- "susie_functional_pi"
warm_mu_method <- "susie_vanilla_warm_from_mu"
warm_pi_method <- "susie_vanilla_warm_from_pi"

# Pretty labels for plotting.
method_labels <- c(
  susie_vanilla = "vanilla (cold)",
  susine_functional_mu = "mu (cold)",
  susie_vanilla_warm_from_mu = "vanilla (warm-from-mu)",
  susie_functional_pi = "pi (cold)",
  susie_vanilla_warm_from_pi = "vanilla (warm-from-pi)"
)
method_color <- c(
  susie_vanilla = "black",
  susine_functional_mu = "#2b8cbe",
  susie_vanilla_warm_from_mu = "#9ecae1",
  susie_functional_pi = "#74c476",
  susie_vanilla_warm_from_pi = "#bae4b3"
)

# Drift threshold (the user's "unpaired" cutoff).
drift_corr_cutoff <- 0.9

# Active-effect threshold for the weighted/filtered drift summaries.
active_var_threshold <- 0.005

# Plot output directory.
plot_dir <- file.path(study_dir, "drift_analysis_plots")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Output filenames.
drift_assignments_csv <- file.path(study_dir, "drift_assignments_5arm.csv")
drift_summary_csv <- file.path(study_dir, "drift_summary_5arm.csv")
auprc_per_dataset_csv <- file.path(study_dir, "auprc_per_dataset_5arm.csv")

drift_profile_hist_png <- file.path(plot_dir, "drift_profile_histograms.png")
chain_trajectory_png <- file.path(plot_dir, "chain_auprc_trajectory.png")
headline_scatter_png <- file.path(plot_dir, "headline_drift_vs_warm_auprc.png")
unpaired_vs_delta_png <- file.path(plot_dir, "unpaired_count_vs_auprc_delta.png")
mu_vs_pi_drift_png <- file.path(plot_dir, "mu_vs_pi_drift_comparison.png")
basin_pip_drift_png <- file.path(plot_dir, "basin_and_pip_drift_summary.png")
pip_records_rds <- file.path(study_dir, "pip_records_5arm.rds")
pip_drift_summary_csv <- file.path(study_dir, "pip_drift_summary_5arm.csv")

cat("Study dir:", study_dir, "\n")
cat("Plot dir:", plot_dir, "\n")
cat("Drift threshold for `n_below_<cut>`:", drift_corr_cutoff, "\n")
cat("Active-effect var-explained threshold:", active_var_threshold, "\n")
```

# 2. Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

format_duration <- function(seconds) {
  seconds <- as.numeric(seconds)
  if (!is.finite(seconds)) return("unknown")
  h <- floor(seconds / 3600)
  m <- floor((seconds %% 3600) / 60)
  s <- round(seconds %% 60)
  sprintf("%02d:%02d:%02d", h, m, s)
}

# Stack effect rows for one (dataset, model) into an n x L matrix of
# fitted_y_effect vectors, in effect order.
records_to_fitted_y_matrix <- function(rows) {
  rows <- rows %>% dplyr::arrange(.data$effect)
  mat <- do.call(cbind, rows$fitted_y_effect)
  storage.mode(mat) <- "double"
  colnames(mat) <- as.character(rows$effect)
  mat
}

# Correlation matrix robust to inactive (sd=0) effects.
safe_cor_matrix <- function(mat_a, mat_b, sd_eps = 1e-12) {
  if (nrow(mat_a) != nrow(mat_b)) {
    stop("Mismatched n (", nrow(mat_a), " vs ", nrow(mat_b), ").")
  }
  sd_a <- apply(mat_a, 2L, stats::sd)
  sd_b <- apply(mat_b, 2L, stats::sd)
  active_a <- is.finite(sd_a) & sd_a > sd_eps
  active_b <- is.finite(sd_b) & sd_b > sd_eps

  cor_mat <- matrix(0, nrow = ncol(mat_a), ncol = ncol(mat_b),
                    dimnames = list(colnames(mat_a), colnames(mat_b)))
  if (any(active_a) && any(active_b)) {
    sub <- suppressWarnings(stats::cor(
      mat_a[, active_a, drop = FALSE],
      mat_b[, active_b, drop = FALSE],
      use = "pairwise.complete.obs"
    ))
    sub[!is.finite(sub)] <- 0
    cor_mat[active_a, active_b] <- sub
  }
  list(cor = cor_mat, active_a = active_a, active_b = active_b)
}

# For one (dataset, A, B) pair, run Hungarian on the L x L correlation
# matrix and return per-edge data.
compute_drift_pair <- function(rows_a, rows_b,
                               method_a, method_b,
                               dataset_bundle_id, matrix_id) {
  rows_a <- rows_a %>% dplyr::arrange(.data$effect)
  rows_b <- rows_b %>% dplyr::arrange(.data$effect)

  mat_a <- records_to_fitted_y_matrix(rows_a)
  mat_b <- records_to_fitted_y_matrix(rows_b)
  cor_info <- safe_cor_matrix(mat_a, mat_b)
  cor_mat <- cor_info$cor

  # clue::solve_LSAP() requires nonnegative entries even with
  # maximum = TRUE. Shift the full matrix if needed; adding a constant
  # to every edge preserves the optimal 1-1 assignment.
  assignment_weights <- cor_mat
  min_weight <- suppressWarnings(min(assignment_weights, na.rm = TRUE))
  if (!is.finite(min_weight)) {
    min_weight <- 0
  }
  if (min_weight < 0) {
    assignment_weights <- assignment_weights - min_weight
  }
  assignment <- as.integer(clue::solve_LSAP(
    assignment_weights,
    maximum = TRUE
  ))
  L <- ncol(mat_a)

  effect_a_ids <- as.integer(rows_a$effect)
  effect_b_ids <- as.integer(rows_b$effect)
  var_y_a <- as.numeric(rows_a$var_y_explained_frac)
  var_y_b <- as.numeric(rows_b$var_y_explained_frac)

  tibble::tibble(
    dataset_bundle_id = dataset_bundle_id,
    matrix_id = matrix_id,
    method_a = method_a,
    method_b = method_b,
    edge = seq_len(L),
    effect_a = effect_a_ids,
    effect_b = effect_b_ids[assignment],
    corr = cor_mat[cbind(seq_len(L), assignment)],
    var_y_explained_a = var_y_a,
    var_y_explained_b = var_y_b[assignment],
    active_a = cor_info$active_a,
    active_b = cor_info$active_b[assignment]
  )
}

# Summarize per (dataset, A, B) from the per-edge assignment table.
summarize_drift <- function(edges_df, drift_corr_cutoff, active_var_threshold) {
  edges_df %>%
    dplyr::group_by(
      .data$dataset_bundle_id, .data$matrix_id,
      .data$method_a, .data$method_b
    ) %>%
    dplyr::summarise(
      n_edges = dplyr::n(),
      mean_corr = mean(.data$corr, na.rm = TRUE),
      median_corr = stats::median(.data$corr, na.rm = TRUE),
      min_corr = suppressWarnings(min(.data$corr, na.rm = TRUE)),
      n_below_cutoff = sum(.data$corr < drift_corr_cutoff, na.rm = TRUE),
      n_active_pairs = sum(
        .data$var_y_explained_a > active_var_threshold &
          .data$var_y_explained_b > active_var_threshold,
        na.rm = TRUE
      ),
      n_active_below_cutoff = sum(
        .data$corr < drift_corr_cutoff &
          .data$var_y_explained_a > active_var_threshold &
          .data$var_y_explained_b > active_var_threshold,
        na.rm = TRUE
      ),
      mean_corr_active = {
        keep <- .data$var_y_explained_a > active_var_threshold &
                  .data$var_y_explained_b > active_var_threshold
        if (any(keep)) mean(.data$corr[keep], na.rm = TRUE) else NA_real_
      },
      weighted_drift = {
        w <- pmin(
          tidyr::replace_na(.data$var_y_explained_a, 0),
          tidyr::replace_na(.data$var_y_explained_b, 0)
        )
        if (sum(w) > 0) {
          # 1 - weighted-mean correlation, so larger = more drift
          1 - sum(w * .data$corr, na.rm = TRUE) / sum(w)
        } else {
          NA_real_
        }
      },
      .groups = "drop"
    )
}
```

# 3. Load Records

```{r load-records}
stopifnot(file.exists(records_rds))
records <- readRDS(records_rds) %>%
  dplyr::mutate(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    matrix_id = as.character(.data$matrix_id),
    effect = as.integer(.data$effect)
  )

methods_present <- sort(unique(records$model_method))
cat("Methods in records:", paste(methods_present, collapse = ", "), "\n")

required_methods <- c(baseline_method, mu_method, warm_mu_method,
                      pi_method, warm_pi_method)
missing_methods <- setdiff(required_methods, methods_present)
if (length(missing_methods)) {
  stop("Missing methods in records: ", paste(missing_methods, collapse = ", "),
       ". Re-run effect_matching_refit_export_5arm.Rmd first.")
}

records <- records %>% dplyr::filter(.data$model_method %in% required_methods)

# Index records by (dataset, model_method) for fast pair access.
records_by_dataset_method <- records %>% split(list(.$dataset_bundle_id, .$model_method))
```

# 4. Pairwise Drift

```{r compute-drift}
# Pairs we want (label, method_a, method_b).
pair_specs <- tibble::tribble(
  ~pair_id,                        ~chain, ~method_a,         ~method_b,
  "vanilla__mu",                   "mu",   baseline_method,   mu_method,
  "mu__vanilla_warm_from_mu",      "mu",   mu_method,         warm_mu_method,
  "vanilla__vanilla_warm_from_mu", "mu",   baseline_method,   warm_mu_method,
  "vanilla__pi",                   "pi",   baseline_method,   pi_method,
  "pi__vanilla_warm_from_pi",      "pi",   pi_method,         warm_pi_method,
  "vanilla__vanilla_warm_from_pi", "pi",   baseline_method,   warm_pi_method
)
print(pair_specs)

# Datasets with all 5 arms present.
datasets_complete <- records %>%
  dplyr::count(.data$dataset_bundle_id, .data$model_method) %>%
  dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
  dplyr::filter(.data$n_methods == length(required_methods)) %>%
  dplyr::pull(.data$dataset_bundle_id)
cat("Datasets with all 5 arms present:", length(datasets_complete), "\n")

dataset_meta <- records %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$matrix_id) %>%
  dplyr::filter(.data$dataset_bundle_id %in% datasets_complete)

started <- proc.time()[["elapsed"]]
edge_chunks <- list()
n_total <- length(datasets_complete) * nrow(pair_specs)
counter <- 0L

for (ds_idx in seq_along(datasets_complete)) {
  ds_id <- datasets_complete[[ds_idx]]
  matrix_id_i <- dataset_meta$matrix_id[
    match(ds_id, dataset_meta$dataset_bundle_id)
  ]
  for (pi_idx in seq_len(nrow(pair_specs))) {
    counter <- counter + 1L
    spec <- pair_specs[pi_idx, ]
    rows_a <- records_by_dataset_method[[paste(ds_id, spec$method_a, sep = ".")]]
    rows_b <- records_by_dataset_method[[paste(ds_id, spec$method_b, sep = ".")]]
    if (is.null(rows_a) || is.null(rows_b)) {
      next
    }
    edge_chunks[[length(edge_chunks) + 1L]] <- compute_drift_pair(
      rows_a = rows_a, rows_b = rows_b,
      method_a = spec$method_a, method_b = spec$method_b,
      dataset_bundle_id = ds_id,
      matrix_id = matrix_id_i
    ) %>% dplyr::mutate(pair_id = spec$pair_id, chain = spec$chain)
  }
  if (ds_idx %% 100L == 0L || ds_idx == length(datasets_complete)) {
    elapsed <- proc.time()[["elapsed"]] - started
    eta <- elapsed / counter * (n_total - counter)
    cat(sprintf(
      "Drift %d/%d datasets (%d/%d pairs) | elapsed %s | ETA %s\n",
      ds_idx, length(datasets_complete), counter, n_total,
      format_duration(elapsed), format_duration(eta)
    ))
  }
}

drift_assignments <- dplyr::bind_rows(edge_chunks)
readr::write_csv(drift_assignments, drift_assignments_csv)
cat("Drift assignment rows:", nrow(drift_assignments), "\n")

drift_summary <- summarize_drift(drift_assignments, drift_corr_cutoff, active_var_threshold) %>%
  dplyr::left_join(pair_specs, by = c("method_a", "method_b"))
readr::write_csv(drift_summary, drift_summary_csv)
cat("Drift summary rows:", nrow(drift_summary), "\n")
```

# 5. Drift Sanity (D(arm, arm) == 0)

```{r drift-sanity, fig.width=8, fig.height=4}
# As a self-test, compute drift of vanilla against itself for a small subset.
# All correlations should be 1, mean_corr ~ 1, n_below_0p9 == 0.
sanity_ids <- head(datasets_complete, 10L)
sanity_rows <- list()
for (ds_id in sanity_ids) {
  rows_v <- records_by_dataset_method[[paste(ds_id, baseline_method, sep = ".")]]
  if (is.null(rows_v)) next
  matrix_id_i <- dataset_meta$matrix_id[match(ds_id, dataset_meta$dataset_bundle_id)]
  sanity_rows[[length(sanity_rows) + 1L]] <- compute_drift_pair(
    rows_a = rows_v, rows_b = rows_v,
    method_a = baseline_method, method_b = baseline_method,
    dataset_bundle_id = ds_id, matrix_id = matrix_id_i
  )
}
sanity_edges <- dplyr::bind_rows(sanity_rows)
sanity_summary <- summarize_drift(sanity_edges, drift_corr_cutoff, active_var_threshold)
cat("Self-drift sanity (vanilla vs vanilla):\n")
print(sanity_summary %>%
        dplyr::summarise(
          mean_of_mean_corr = mean(.data$mean_corr, na.rm = TRUE),
          mean_n_below_cutoff = mean(.data$n_below_cutoff),
          max_n_below_cutoff = max(.data$n_below_cutoff)
        ))
# Note: inactive (sd=0) effects against themselves give correlation 0 by
# construction, so n_below_cutoff > 0 is expected. mean_corr_active should
# be close to 1.
cat("Self-drift mean corr (active-only):", mean(sanity_summary$mean_corr_active, na.rm = TRUE), "\n")
```

# 6. Per-Dataset AUPRC Trajectory

```{r per-dataset-auprc}
auprc_per_dataset <- records %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$model_method, .data$model_AUPRC) %>%
  tidyr::pivot_wider(
    names_from = .data$model_method,
    values_from = .data$model_AUPRC,
    names_prefix = "auprc_"
  ) %>%
  dplyr::mutate(
    auprc_delta_mu = .data$auprc_susine_functional_mu - .data$auprc_susie_vanilla,
    auprc_delta_pi = .data$auprc_susie_functional_pi - .data$auprc_susie_vanilla,
    auprc_delta_warm_mu_vs_vanilla =
      .data$auprc_susie_vanilla_warm_from_mu - .data$auprc_susie_vanilla,
    auprc_delta_warm_pi_vs_vanilla =
      .data$auprc_susie_vanilla_warm_from_pi - .data$auprc_susie_vanilla,
    auprc_round_trip_mu =
      .data$auprc_susie_vanilla_warm_from_mu - .data$auprc_susine_functional_mu,
    auprc_round_trip_pi =
      .data$auprc_susie_vanilla_warm_from_pi - .data$auprc_susie_functional_pi
  )
readr::write_csv(auprc_per_dataset, auprc_per_dataset_csv)
cat("Per-dataset AUPRC rows:", nrow(auprc_per_dataset), "\n")

# Long format for spaghetti / facet plots.
auprc_long <- auprc_per_dataset %>%
  tidyr::pivot_longer(
    cols = dplyr::starts_with("auprc_"),
    names_to = "metric",
    values_to = "value"
  )
```

# 7. Drift Profile Histograms

```{r drift-profile-hists, fig.width=12, fig.height=7}
drift_assignments_for_plot <- drift_assignments %>%
  dplyr::mutate(
    pair_id = factor(.data$pair_id, levels = pair_specs$pair_id)
  )

p_drift_hists <- ggplot(drift_assignments_for_plot,
                        aes(x = .data$corr, fill = .data$chain)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 60, position = "identity", alpha = 0.55, color = NA
  ) +
  geom_vline(xintercept = drift_corr_cutoff, linetype = "dashed", color = "grey25") +
  facet_wrap(~ pair_id, ncol = 3) +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  coord_cartesian(xlim = c(-0.05, 1.0)) +
  labs(
    x = "Hungarian-assigned fitted_y correlation per effect",
    y = "Density",
    fill = "Chain",
    title = "Distribution of per-effect assignment correlations across all pairs",
    subtitle = sprintf(
      "Dashed line = drift_corr_cutoff (%.2f). All 10 effects per (dataset, pair) included.",
      drift_corr_cutoff
    )
  ) +
  theme(legend.position = "bottom")

ggsave(drift_profile_hist_png, p_drift_hists, width = 12, height = 7, dpi = 150)
print(p_drift_hists)
```

# 8. Chain AUPRC Trajectory

```{r chain-trajectory, fig.width=11, fig.height=5}
chain_long <- dplyr::bind_rows(
  auprc_per_dataset %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step1 = .data$auprc_susie_vanilla,
      step2 = .data$auprc_susine_functional_mu,
      step3 = .data$auprc_susie_vanilla_warm_from_mu
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step1 = .data$auprc_susie_vanilla,
      step2 = .data$auprc_susie_functional_pi,
      step3 = .data$auprc_susie_vanilla_warm_from_pi
    )
) %>%
  tidyr::pivot_longer(
    cols = c("step1", "step2", "step3"),
    names_to = "step",
    values_to = "auprc"
  ) %>%
  dplyr::mutate(
    step_label = dplyr::case_when(
      step == "step1" ~ "vanilla (cold)",
      step == "step2" & chain == "mu" ~ "mu",
      step == "step2" & chain == "pi" ~ "pi",
      step == "step3" ~ "vanilla (warm)",
      TRUE ~ step
    ),
    step_label = factor(step_label,
                        levels = c("vanilla (cold)", "mu", "pi", "vanilla (warm)"))
  )

p_chain <- ggplot(chain_long, aes(x = .data$step_label, y = .data$auprc)) +
  geom_line(aes(group = .data$dataset_bundle_id), alpha = 0.06) +
  geom_violin(aes(fill = .data$chain), alpha = 0.55, color = NA, scale = "width") +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.7) +
  facet_wrap(~ chain, scales = "free_x") +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  labs(
    x = NULL, y = "AUPRC",
    fill = "Chain",
    title = "Per-dataset AUPRC trajectory along each refit chain",
    subtitle = "Each thin line = one dataset; violin / box = marginal distribution per step."
  ) +
  theme(legend.position = "bottom")

ggsave(chain_trajectory_png, p_chain, width = 11, height = 5, dpi = 150)
print(p_chain)
```

# 9. Headline 4-Quadrant Test

```{r headline-scatter, fig.width=12, fig.height=5.5}
# Join drift(vanilla, vanilla_warm) with AUPRC delta(vanilla_warm - vanilla)
headline_df_mu <- drift_summary %>%
  dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_mu") %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    chain = "mu",
    drift_mean_corr = .data$mean_corr,
    drift_mean_corr_active = .data$mean_corr_active,
    n_below_cutoff = .data$n_below_cutoff
  ) %>%
  dplyr::left_join(
    auprc_per_dataset %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        auprc_delta_warm_vs_vanilla = .data$auprc_delta_warm_mu_vs_vanilla,
        auprc_delta_treatment = .data$auprc_delta_mu
      ),
    by = "dataset_bundle_id"
  )

headline_df_pi <- drift_summary %>%
  dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_pi") %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    chain = "pi",
    drift_mean_corr = .data$mean_corr,
    drift_mean_corr_active = .data$mean_corr_active,
    n_below_cutoff = .data$n_below_cutoff
  ) %>%
  dplyr::left_join(
    auprc_per_dataset %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        auprc_delta_warm_vs_vanilla = .data$auprc_delta_warm_pi_vs_vanilla,
        auprc_delta_treatment = .data$auprc_delta_pi
      ),
    by = "dataset_bundle_id"
  )

headline_df <- dplyr::bind_rows(headline_df_mu, headline_df_pi)

p_headline <- ggplot(
  headline_df,
  aes(x = .data$drift_mean_corr_active,
      y = .data$auprc_delta_warm_vs_vanilla,
      color = .data$auprc_delta_treatment)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_vline(xintercept = drift_corr_cutoff, color = "grey45", linetype = "dotted") +
  geom_point(alpha = 0.55, size = 1.4) +
  scale_color_gradient2(
    low = "#2b6cb0", mid = "white", high = "#c53030", midpoint = 0
  ) +
  facet_wrap(~ chain, labeller = labeller(chain = c(mu = "mu chain", pi = "pi chain"))) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(
    x = "Mean active-effect drift D(vanilla, vanilla_warm) (1 = identical)",
    y = "AUPRC(vanilla_warm) - AUPRC(vanilla_cold)",
    color = "AUPRC(treatment)\n- AUPRC(vanilla)",
    title = "Headline: refinement vs exploration — does the warm vanilla refit retain treatment basin?",
    subtitle = "Top-right quadrant (high drift + high warm-vanilla ΔAUPRC) = exploration win. Top-left = pure within-basin PIP weighting."
  )

ggsave(headline_scatter_png, p_headline, width = 12, height = 5.5, dpi = 150)
print(p_headline)
```

# 10. Unpaired Count vs AUPRC Delta (User's Plot)

```{r unpaired-vs-delta, fig.width=12, fig.height=8}
# Build a long table: per (dataset, chain, step_label), report n_below_cutoff
# (drift) and the relevant AUPRC delta for that step.
step_drift_delta <- dplyr::bind_rows(
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step_label = "step 1: vanilla -> mu",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_delta_mu),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "mu__vanilla_warm_from_mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step_label = "step 2: mu -> vanilla_warm",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_round_trip_mu),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step_label = "step 1: vanilla -> pi",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_delta_pi),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "pi__vanilla_warm_from_pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step_label = "step 2: pi -> vanilla_warm",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_round_trip_pi),
      by = "dataset_bundle_id"
    )
)

step_rho <- step_drift_delta %>%
  dplyr::group_by(.data$chain, .data$step_label) %>%
  dplyr::summarise(
    n = dplyr::n(),
    rho = suppressWarnings(stats::cor(
      .data$n_below_cutoff, .data$auprc_delta,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    label = sprintf("rho = %.3f (n=%d)", .data$rho, .data$n),
    .groups = "drop"
  )

p_unpaired <- ggplot(
  step_drift_delta,
  aes(x = .data$n_below_cutoff, y = .data$auprc_delta)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_jitter(width = 0.25, height = 0, alpha = 0.45, size = 1.1,
              aes(color = .data$mean_corr)) +
  geom_smooth(method = "loess", formula = y ~ x, se = FALSE,
              color = "black", linetype = "dashed", linewidth = 0.5) +
  geom_text(
    data = step_rho,
    aes(x = -Inf, y = Inf, label = .data$label),
    hjust = -0.05, vjust = 1.4, size = 3.3, inherit.aes = FALSE
  ) +
  facet_grid(chain ~ step_label) +
  scale_x_continuous(breaks = 0:10) +
  scale_color_viridis_c(option = "C") +
  labs(
    x = sprintf("# unpaired effects (assignment corr < %.2f)", drift_corr_cutoff),
    y = "AUPRC delta for this step",
    color = "Mean active\nassignment corr",
    title = sprintf(
      "User's hypothesis test: does # of unpaired effects (cutoff %.2f) predict AUPRC change at each step?",
      drift_corr_cutoff
    ),
    subtitle = "Step 1: cold->treatment (positive y => treatment wins). Step 2: treatment->vanilla_warm (positive y => warm refit IMPROVES on treatment, rare; negative => snaps back to vanilla)."
  )

ggsave(unpaired_vs_delta_png, p_unpaired, width = 12, height = 8, dpi = 150)
print(p_unpaired)
```

# 11. Mu vs Pi Drift Comparison

```{r mu-vs-pi-drift, fig.width=11, fig.height=5}
# Side-by-side: drift(vanilla, treatment) for mu vs pi, per dataset.
mu_vs_pi_join <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c("vanilla__mu", "vanilla__pi")) %>%
  dplyr::select(.data$dataset_bundle_id, .data$pair_id,
                .data$mean_corr_active, .data$n_below_cutoff,
                .data$weighted_drift) %>%
  tidyr::pivot_wider(
    id_cols = .data$dataset_bundle_id,
    names_from = .data$pair_id,
    values_from = c("mean_corr_active", "n_below_cutoff", "weighted_drift")
  )

p_mupi <- ggplot(
  mu_vs_pi_join,
  aes(x = .data$mean_corr_active_vanilla__mu,
      y = .data$mean_corr_active_vanilla__pi)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey45", linetype = "dashed") +
  geom_point(alpha = 0.5, size = 1.2) +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    x = "Mean active-effect drift: vanilla -> mu",
    y = "Mean active-effect drift: vanilla -> pi",
    title = "How much does each annotation mechanism move the model?",
    subtitle = "y=x line: equal drift. Below diagonal: pi explores less than mu. Above: pi explores more."
  )

ggsave(mu_vs_pi_drift_png, p_mupi, width = 11, height = 5, dpi = 150)
print(p_mupi)
```

# 12. Per-Dataset PIP JSD

Loads the per-fit model-wide PIP sidecar (`pip_records_5arm.rds`, written by
`effect_matching_refit_export_5arm.Rmd`) and computes Jensen-Shannon
divergence between matched (vanilla, treatment) and (vanilla, vanilla_warm)
PIP vectors for each dataset. JSD definition mirrors
`R/run_model.R::js_distance`: a categorical JSD on raw model-wide PIP
vectors with an additive epsilon, summed (not averaged) across SNPs. See
the study plan note "JSD ... currently computed on raw model-wide PIPs".

```{r pip-jsd}
# Inline copy of test_susine:::js_distance (R/run_model.R:2021) so the Rmd
# does not depend on devtools::load_all() for the package.
js_distance <- function(p, q, eps = 1e-12) {
  p <- as.numeric(p); q <- as.numeric(q)
  p <- p + eps; q <- q + eps
  m <- 0.5 * (p + q)
  kl <- function(a, b) sum(a * log(a / b))
  0.5 * kl(p, m) + 0.5 * kl(q, m)
}

stopifnot(file.exists(pip_records_rds))
pip_records <- readRDS(pip_records_rds) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

cat("PIP records rows:", nrow(pip_records), "\n")
cat("Datasets in PIP records:", dplyr::n_distinct(pip_records$dataset_bundle_id), "\n")

pip_by_dataset_method <- pip_records %>%
  dplyr::select(.data$dataset_bundle_id, .data$model_method, .data$pip)

# Sanity: self-JSD (vanilla vs vanilla) should be exactly 0.
self_check_ids <- head(unique(pip_records$dataset_bundle_id), 5L)
self_jsd <- vapply(self_check_ids, function(ds_id) {
  v <- pip_by_dataset_method %>%
    dplyr::filter(.data$dataset_bundle_id == ds_id,
                  .data$model_method == baseline_method) %>%
    dplyr::pull(.data$pip) %>% .[[1]]
  js_distance(v, v)
}, numeric(1))
cat("Self-JSD (vanilla vs vanilla) on first 5 datasets:",
    paste(sprintf("%.3e", self_jsd), collapse = ", "), "\n")

# Pairs we want for the master figure: (chain, pair_kind) -> (method_a, method_b)
pip_pair_specs <- tibble::tribble(
  ~chain, ~pair_kind,             ~method_a,         ~method_b,
  "mu",   "vanilla_to_treatment", baseline_method,   mu_method,
  "mu",   "vanilla_to_warm",      baseline_method,   warm_mu_method,
  "pi",   "vanilla_to_treatment", baseline_method,   pi_method,
  "pi",   "vanilla_to_warm",      baseline_method,   warm_pi_method
)

# Index PIPs by (dataset, method) for fast lookup.
pip_lookup <- pip_records %>%
  dplyr::group_by(.data$dataset_bundle_id, .data$model_method) %>%
  dplyr::summarise(pip = list(.data$pip[[1L]]), .groups = "drop")
pip_index <- setNames(
  pip_lookup$pip,
  paste(pip_lookup$dataset_bundle_id, pip_lookup$model_method, sep = "||")
)

datasets_with_all_pips <- pip_records %>%
  dplyr::count(.data$dataset_bundle_id, .data$model_method) %>%
  dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
  dplyr::filter(.data$n_methods == length(required_methods)) %>%
  dplyr::pull(.data$dataset_bundle_id)

pip_jsd_rows <- list()
for (ds_id in datasets_with_all_pips) {
  for (k in seq_len(nrow(pip_pair_specs))) {
    spec <- pip_pair_specs[k, ]
    pa <- pip_index[[paste(ds_id, spec$method_a, sep = "||")]]
    pb <- pip_index[[paste(ds_id, spec$method_b, sep = "||")]]
    if (is.null(pa) || is.null(pb) || length(pa) != length(pb)) next
    pip_jsd_rows[[length(pip_jsd_rows) + 1L]] <- tibble::tibble(
      dataset_bundle_id = ds_id,
      chain = spec$chain,
      pair_kind = spec$pair_kind,
      pip_jsd = js_distance(pa, pb)
    )
  }
}
pip_drift_summary <- dplyr::bind_rows(pip_jsd_rows)
readr::write_csv(pip_drift_summary, pip_drift_summary_csv)
cat("PIP drift summary rows:", nrow(pip_drift_summary), "\n")

# Sanity: rho between basin weighted_drift and pip_jsd on matched (dataset,
# pair). Both metrics measure "how much did the model move", so a strongly
# positive rank correlation is expected.
basin_for_join <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "vanilla__vanilla_warm_from_mu",
    "vanilla__pi", "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    .data$dataset_bundle_id,
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::if_else(grepl("warm", .data$pair_id),
                               "vanilla_to_warm", "vanilla_to_treatment"),
    weighted_drift = .data$weighted_drift
  )

basin_pip_join <- basin_for_join %>%
  dplyr::inner_join(
    pip_drift_summary,
    by = c("dataset_bundle_id", "chain", "pair_kind")
  )

rho_check <- basin_pip_join %>%
  dplyr::group_by(.data$chain, .data$pair_kind) %>%
  dplyr::summarise(
    n = dplyr::n(),
    rho = suppressWarnings(stats::cor(
      .data$weighted_drift, .data$pip_jsd,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    .groups = "drop"
  )
print(rho_check)
```

# 13. Master 4-Panel Drift Summary Figure

Two metrics x two pair kinds. mu and pi overlaid in each subplot.

```{r master-drift-figure, fig.width=11, fig.height=7}
basin_long <- basin_for_join %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$chain, .data$pair_kind,
    metric = "basin_drift",
    value = .data$weighted_drift
  )

pip_long <- pip_drift_summary %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$chain, .data$pair_kind,
    metric = "pip_drift",
    value = .data$pip_jsd
  )

combined <- dplyr::bind_rows(basin_long, pip_long) %>%
  dplyr::mutate(
    metric = factor(.data$metric, levels = c("basin_drift", "pip_drift")),
    pair_kind = factor(
      .data$pair_kind,
      levels = c("vanilla_to_treatment", "vanilla_to_warm")
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

p_master <- ggplot(combined, aes(x = .data$value, fill = .data$chain)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 50, position = "identity", alpha = 0.55, color = NA
  ) +
  facet_grid(
    metric ~ pair_kind,
    scales = "free",
    labeller = labeller(
      metric = c(
        basin_drift = "Basin drift (1 - weighted corr)",
        pip_drift   = "PIP drift (JSD)"
      ),
      pair_kind = c(
        vanilla_to_treatment = "vanilla -> annotated",
        vanilla_to_warm      = "vanilla -> vanilla_warm"
      )
    )
  ) +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  labs(
    x = "Drift metric value", y = "Density", fill = "Chain",
    title = "Basin and PIP drift along the 5-arm refit chain",
    subtitle = "One observation per dataset per chain. Top row: effect-level basin drift (weighted_drift). Bottom row: model-wide PIP-vector JSD."
  ) +
  theme(legend.position = "bottom")

ggsave(basin_pip_drift_png, p_master, width = 11, height = 7, dpi = 150)
print(p_master)
```

# 14. Outputs

```{r output-index}
outputs <- tibble::tibble(
  artifact = c(
    "per-effect drift assignments",
    "per-(dataset, pair) drift summary",
    "per-dataset AUPRC trajectory",
    "drift profile histograms",
    "chain AUPRC trajectory",
    "headline 4-quadrant scatter",
    "unpaired-count vs AUPRC delta scatter",
    "mu vs pi drift comparison",
    "per-dataset PIP JSD summary",
    "master basin + PIP drift figure"
  ),
  path = c(
    drift_assignments_csv,
    drift_summary_csv,
    auprc_per_dataset_csv,
    drift_profile_hist_png,
    chain_trajectory_png,
    headline_scatter_png,
    unpaired_vs_delta_png,
    mu_vs_pi_drift_png,
    pip_drift_summary_csv,
    basin_pip_drift_png
  )
)
kable(outputs)
```
