---
title: "Effect Matching Refit Export (5-arm: cold + warm refit chains)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Extends the 3-arm refit study with two warm-started vanilla refits — one
chained off each treatment fit. Per dataset, fits these arms in this order
(so each warm refit has its source fit live in memory):

1. `susie_vanilla` (cold)
2. `susine_functional_mu` (cold, c = 0.6, annotation_r2 = 0.5)
3. `susie_vanilla_warm_from_mu` (vanilla prior, warm-started from arm 2's fit)
4. `susie_functional_pi` (cold, tau = 0.464, annotation_r2 = 0.5)
5. `susie_vanilla_warm_from_pi` (vanilla prior, warm-started from arm 4's fit)

Warm refits use `mu_0 = 0`, `sigma_0_2 = 0.01` (matched to cold vanilla, NOT
the harness default 0.2), uniform inclusion weights, and `init_effect_fits`
synthesized from the source fit's `(alpha, b_hat, b_2_hat)` slots — the
same mechanism `run_default_prior_refit()` uses in the production harness
([R/run_model.R:1587-1619](R/run_model.R#L1587)).

Output goes to a fresh study directory; the 3-arm outputs are preserved.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_refit_export_5arm.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(knitr)
```

# 1. User Parameters

```{r user-parameters}
study_name <- "baseline_sims_5arm_effect_matching"
output_root <- here("output")
study_dir <- file.path(output_root, "effect_matching_study", study_name)
dir.create(study_dir, recursive = TRUE, showWarnings = FALSE)

# Set to a small integer, e.g. 5L, for a local smoke run.
max_datasets <- Inf
force_refit <- FALSE
checkpoint_every <- 25L

# Baseline-sims dataset recipe (must match the 3-arm refit so dataset bundle
# ids align across studies for cross-study comparison).
n_matrices_target <- 150L
seeds <- 1:4
L_value <- 10L
y_noise_value <- NA_real_
p_star_value <- NA_integer_
architecture_value <- "susie2_oligogenic"
data_scenario <- "scenario_1"

# Treatment-arm settings (harmonized annotations; only prior mechanism differs).
annotation_r2_value <- 0.5
inflate_match_value <- 0.9
c_value_mu <- 0.6
tau_value_pi <- 0.464
sigma_0_2_value <- 0.01
max_iter <- 100L
credible_set_rho <- 0.95
purity_threshold <- 0.50

# Method ordering. The two warm-refit method labels are NOT in the use_case
# catalog; they're handled by the bespoke fit_warm_refit() helper below.
cold_method_levels <- c("susie_vanilla", "susine_functional_mu", "susie_functional_pi")
warm_method_levels <- c("susie_vanilla_warm_from_mu", "susie_vanilla_warm_from_pi")
all_method_levels  <- c(cold_method_levels, warm_method_levels)
n_methods_expected <- length(all_method_levels)

# Offset added to source run_id to make warm-refit run_ids globally unique.
warm_run_id_offset <- 10000000L

records_rds <- file.path(study_dir, "effect_refit_records.rds")
metadata_csv <- file.path(study_dir, "effect_refit_metadata.csv")
selected_runs_csv <- file.path(study_dir, "selected_run_rows.csv")
dataset_bundles_csv <- file.path(study_dir, "dataset_bundles.csv")
progress_csv <- file.path(study_dir, "refit_progress_log.csv")

cat("Study dir:", study_dir, "\n")
cat("Methods (in fit order):", paste(all_method_levels, collapse = ", "), "\n")
```

# 2. Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

format_duration <- function(seconds) {
  seconds <- as.numeric(seconds)
  if (!is.finite(seconds)) {
    return("unknown")
  }
  h <- floor(seconds / 3600)
  m <- floor((seconds %% 3600) / 60)
  s <- round(seconds %% 60)
  sprintf("%02d:%02d:%02d", h, m, s)
}

save_checkpoint <- function(records, progress_rows) {
  saveRDS(records, records_rds)

  metadata <- records
  if ("indices" %in% names(metadata)) {
    metadata <- metadata %>%
      dplyr::mutate(
        cs_indices = vapply(
          .data$indices,
          function(idx) paste(as.integer(idx), collapse = ";"),
          character(1)
        )
      )
  }
  metadata <- metadata %>%
    dplyr::select(-dplyr::where(is.list))
  readr::write_csv(metadata, metadata_csv)

  if (length(progress_rows)) {
    readr::write_csv(dplyr::bind_rows(progress_rows), progress_csv)
  }
}

first_existing_numeric <- function(df, cols, default = NA_real_) {
  for (nm in cols) {
    if (nm %in% names(df)) {
      val <- suppressWarnings(as.numeric(df[[nm]][[1]]))
      if (length(val)) {
        return(val)
      }
    }
  }
  default
}

effect_coef_matrix <- function(fit, X) {
  alpha <- fit$effect_fits$alpha
  if (is.list(alpha)) {
    alpha <- do.call(rbind, alpha)
  }
  alpha <- as.matrix(alpha)
  L <- nrow(alpha)
  p <- ncol(X)

  b_hat <- fit$effect_fits$b_hat %||% NULL
  if (!is.null(b_hat)) {
    b_hat <- as.matrix(b_hat)
    if (nrow(b_hat) == p && ncol(b_hat) == L) {
      b_hat <- t(b_hat)
    }
    if (nrow(b_hat) == L && ncol(b_hat) == p) {
      return(b_hat)
    }
  }

  mu <- fit$effect_fits$mu %||% fit$raw$mu %||% NULL
  if (!is.null(mu)) {
    mu <- as.matrix(mu)
    if (nrow(mu) == p && ncol(mu) == L) {
      mu <- t(mu)
    }
    if (nrow(mu) == L && ncol(mu) == p) {
      return(alpha * mu)
    }
  }

  stop("Could not recover an L x p per-effect coefficient matrix from the fit.")
}

extract_effect_records <- function(fit, eval_res, data_bundle, run_row, model_method, fit_meta) {
  X <- data_bundle$X
  y <- as.numeric(data_bundle$y)
  b_hat <- effect_coef_matrix(fit, X)
  fitted_y_mat <- X %*% t(b_hat)
  L <- ncol(fitted_y_mat)
  var_y <- stats::var(y)

  effect_tbl <- eval_res$effects_unfiltered %>%
    tibble::as_tibble() %>%
    dplyr::mutate(effect = as.integer(.data$effect))

  tibble::tibble(effect = seq_len(L)) %>%
    dplyr::left_join(effect_tbl, by = "effect") %>%
    dplyr::mutate(
      fitted_y_effect = lapply(seq_len(L), function(l) as.numeric(fitted_y_mat[, l])),
      var_y_explained_frac = vapply(
        .data$fitted_y_effect,
        function(fy) {
          val <- stats::var(fy) / var_y
          ifelse(is.finite(val), val, NA_real_)
        },
        numeric(1)
      ),
      var_y_explained_pct = 100 * .data$var_y_explained_frac,
      dataset_bundle_id = as.character(run_row$dataset_bundle_id[[1]]),
      run_id = as.integer(run_row$run_id[[1]]),
      model_method = model_method,
      use_case_id = as.character(run_row$use_case_id[[1]]),
      spec_name = as.character(run_row$spec_name[[1]]),
      setting_label = as.character(run_row$setting_label[[1]]),
      annotation_label = as.character(run_row$annotation_label[[1]]),
      annotation_r2 = as.numeric(run_row$annotation_r2[[1]]),
      inflate_match = as.numeric(run_row$inflate_match[[1]]),
      c_value = as.numeric(run_row$c_value[[1]]),
      tau_value = as.numeric(run_row$tau_value[[1]]),
      sigma_0_2_scalar = as.character(run_row$sigma_0_2_scalar[[1]]),
      phenotype_seed = as.integer(run_row$phenotype_seed[[1]]),
      matrix_id = as.character(run_row$matrix_id[[1]]),
      architecture = as.character(run_row$architecture[[1]]),
      p_star = as.integer(run_row$p_star[[1]]),
      y_noise = as.numeric(run_row$y_noise[[1]]),
      fit_wall_time_sec = as.numeric(fit_meta$wall_time_sec %||% NA_real_),
      model_AUPRC = as.numeric(eval_res$model_unfiltered$AUPRC[[1]]),
      model_tpr_fpr05 = first_existing_numeric(
        eval_res$model_unfiltered,
        c("tpr_fpr05", "tpr_at_fpr_05")
      )
    ) %>%
    dplyr::relocate(
      dplyr::all_of(c(
        "dataset_bundle_id", "model_method", "run_id",
        "effect", "setting_label", "annotation_label"
      ))
    )
}

# Cold fit: dispatch through the use-case catalog like the 3-arm notebook,
# but ALSO return the fit object so it can seed a warm refit downstream.
fit_one_method_with_fit <- function(run_row, data_bundle, job_config) {
  run_row <- tibble::as_tibble(run_row)
  use_case <- test_susine:::lookup_use_case(job_config, run_row$use_case_id[[1]])
  if (!nrow(use_case)) {
    stop("Unknown use case: ", run_row$use_case_id[[1]])
  }

  model_result <- test_susine:::run_use_case(
    use_case = use_case,
    run_row = run_row,
    data_bundle = data_bundle,
    job_config = job_config
  )
  fit <- model_result$fits[[1]]
  fit_meta <- model_result$fit_meta[[1]] %||% list()

  eval_res <- test_susine:::evaluate_model(
    fit = fit,
    X = data_bundle$X,
    y = data_bundle$y,
    causal_idx = data_bundle$causal_idx,
    rho = job_config$job$credible_set_rho %||% credible_set_rho,
    purity_threshold = job_config$job$purity_threshold %||% purity_threshold,
    compute_curves = FALSE
  )

  records <- extract_effect_records(
    fit = fit,
    eval_res = eval_res,
    data_bundle = data_bundle,
    run_row = run_row,
    model_method = run_row$method_family[[1]],
    fit_meta = fit_meta
  )

  list(records = records, fit = fit, eval_res = eval_res)
}

# Schema-detecting init builder. Three layouts encountered:
#   (1) susine fit   : fit$effect_fits has alpha + b_hat + b_2_hat (matrices).
#   (2) susieR fit, harness-wrapped: fit$effect_fits has only alpha; the
#       original susieR fit (with alpha, mu, mu2) is stashed under fit$raw.
#       Convert via b_hat = alpha * mu and b_2_hat = alpha * mu2 (the SuSiE
#       variational identity: mu/mu2 are conditional-on-selection moments).
#   (3) raw susieR fit (no wrap): alpha / mu / mu2 sit at the top level.
build_init_effect_fits_robust <- function(source_fit) {
  ef <- source_fit$effect_fits
  raw <- source_fit$raw

  pull <- function(field) {
    if (!is.null(ef) && !is.null(ef[[field]]))   return(ef[[field]])
    if (!is.null(raw) && !is.null(raw[[field]])) return(raw[[field]])
    source_fit[[field]]
  }

  alpha   <- pull("alpha")
  b_hat   <- pull("b_hat")
  b_2_hat <- pull("b_2_hat")

  if (is.null(alpha)) {
    stop("Source fit has no alpha matrix (checked effect_fits, raw, and top-level).")
  }
  if (is.list(alpha) && !is.data.frame(alpha)) alpha <- do.call(rbind, alpha)
  alpha <- as.matrix(alpha)

  if (is.null(b_hat) || is.null(b_2_hat)) {
    mu  <- pull("mu")
    mu2 <- pull("mu2")
    if (is.null(mu) || is.null(mu2)) {
      stop(
        "Source fit lacks (b_hat, b_2_hat) AND (mu, mu2); cannot construct init_effect_fits.\n",
        "  effect_fits fields: ", paste(names(ef %||% list()), collapse = ", "), "\n",
        "  raw fields:        ", paste(names(raw %||% list()), collapse = ", "), "\n",
        "  top-level fields:  ", paste(names(source_fit), collapse = ", ")
      )
    }
    if (is.list(mu) && !is.data.frame(mu))   mu  <- do.call(rbind, mu)
    if (is.list(mu2) && !is.data.frame(mu2)) mu2 <- do.call(rbind, mu2)
    mu  <- as.matrix(mu)
    mu2 <- as.matrix(mu2)
    b_hat   <- alpha * mu
    b_2_hat <- alpha * mu2
  } else {
    if (is.list(b_hat) && !is.data.frame(b_hat))     b_hat   <- do.call(rbind, b_hat)
    if (is.list(b_2_hat) && !is.data.frame(b_2_hat)) b_2_hat <- do.call(rbind, b_2_hat)
    b_hat   <- as.matrix(b_hat)
    b_2_hat <- as.matrix(b_2_hat)
  }

  list(alpha = alpha, b_hat = b_hat, b_2_hat = b_2_hat)
}

# Warm refit: vanilla prior, init_effect_fits derived from source via the
# schema-detecting builder above. Synthesizes a run_row that distinguishes
# the warm fit's records from the source's.
fit_warm_refit <- function(source_fit, source_run_row, data_bundle, job_config,
                           model_method_label,
                           run_id_offset = warm_run_id_offset) {
  source_run_row <- tibble::as_tibble(source_run_row)
  L <- as.integer(source_run_row$L[[1]] %||% L_value)
  p <- ncol(data_bundle$X)
  iter_max <- as.integer(job_config$job$max_iter %||% max_iter)
  tol <- as.numeric(job_config$job$tol %||% 1e-3)

  susine_fn <- test_susine:::resolve_susine_fn(require_init_effect_fits = TRUE)
  init_eff <- build_init_effect_fits_robust(source_fit)

  started <- proc.time()[["elapsed"]]
  fit <- susine_fn(
    L = L,
    X = data_bundle$X,
    y = data_bundle$y,
    mu_0 = 0,
    sigma_0_2 = sigma_0_2_value,             # match cold vanilla, NOT 0.2
    prior_inclusion_weights = rep(1 / p, p),
    prior_update_method = "none",
    max_iter = iter_max,
    tol = tol,
    verbose = FALSE,
    init_effect_fits = init_eff
  )
  wall_time_sec <- proc.time()[["elapsed"]] - started

  eval_res <- test_susine:::evaluate_model(
    fit = fit,
    X = data_bundle$X,
    y = data_bundle$y,
    causal_idx = data_bundle$causal_idx,
    rho = job_config$job$credible_set_rho %||% credible_set_rho,
    purity_threshold = job_config$job$purity_threshold %||% purity_threshold,
    compute_curves = FALSE
  )

  # Build a synthetic run_row for record extraction: keep dataset / phenotype
  # metadata from the source row, override identity / annotation fields.
  warm_run_row <- source_run_row
  warm_run_row$run_id <- as.integer(source_run_row$run_id[[1]]) + run_id_offset
  warm_run_row$method_family <- model_method_label
  warm_run_row$use_case_id <- "susie_vanilla"
  warm_run_row$spec_name <- paste0("warm_refit_from_", source_run_row$method_family[[1]])
  warm_run_row$setting_label <- paste0(
    model_method_label, " | warm-from=", source_run_row$method_family[[1]],
    " | sigma=", sigma_0_2_value
  )
  warm_run_row$annotation_label <- "no_annotation"
  warm_run_row$annotation_r2 <- NA_real_
  warm_run_row$inflate_match <- NA_real_
  warm_run_row$c_value <- NA_real_
  warm_run_row$tau_value <- NA_real_
  warm_run_row$sigma_0_2_scalar <- as.character(sigma_0_2_value)

  records <- extract_effect_records(
    fit = fit,
    eval_res = eval_res,
    data_bundle = data_bundle,
    run_row = warm_run_row,
    model_method = model_method_label,
    fit_meta = list(wall_time_sec = wall_time_sec)
  )

  list(records = records, fit = fit, eval_res = eval_res)
}
```

# 3. Build Standalone Baseline Dataset Grid

```{r build-standalone-config}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = data_scenario,
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)

m1_summary_path <- here(
  "data", "sampled_simulated_genotypes", "scenario_sampling_summary_with_m1.csv"
)
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>%
      dplyr::select(.data$matrix_id, .data$matrix_path),
    base_dir = here()
  ) %>%
    dplyr::select(.data$matrix_id, .data$M1)
  catalog_with_m1 <- data_matrix_catalog_full %>%
    dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}

if (!"matrix_id" %in% names(catalog_with_m1)) {
  if ("gene" %in% names(catalog_with_m1)) {
    catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = .data$gene)
  }
  join_col <- intersect(c("dataset_label", "matrix_path"), names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(
      data_matrix_catalog_full %>%
        dplyr::select(.data$matrix_id, dplyr::all_of(join_col)),
      by = join_col
    )
}

catalog_with_m1 <- catalog_with_m1 %>%
  dplyr::filter(!is.na(.data$M1)) %>%
  dplyr::arrange(.data$M1)

n_target <- min(n_matrices_target, nrow(catalog_with_m1))
idx <- unique(round(seq(1, nrow(catalog_with_m1), length.out = n_target)))
selected_matrices <- catalog_with_m1[idx, , drop = FALSE]

data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(
    selected_matrices %>% dplyr::select(.data$matrix_id),
    by = "matrix_id"
  )

spec_susie_vanilla <- list(
  name = "effect_match_susie_vanilla",
  use_case_ids = "susie_vanilla",
  exploration_methods = "sigma_0_2_grid",
  exploration_mode = "separate",
  K = 1L,
  sigma_0_2_grid_values = sigma_0_2_value
)

spec_susine_functional_mu <- list(
  name = "effect_match_susine_functional_mu",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 1L,
  c_grid_values = c_value_mu,
  sigma_0_2_grid_values = sigma_0_2_value
)

spec_susie_functional_pi <- list(
  name = "effect_match_susie_functional_pi",
  use_case_ids = "susie_functional_pi",
  exploration_methods = c("tau_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 1L,
  tau_grid_values = tau_value_pi,
  sigma_0_2_grid_values = sigma_0_2_value
)

cfg <- test_susine::make_job_config(
  job_name = study_name,
  HPC = FALSE,
  time = "02:00:00",
  mem = "8G",
  output_root = output_root,
  grid_mode = "full",
  use_case_ids = cold_method_levels,
  exploration_specs = list(
    spec_susie_vanilla,
    spec_susine_functional_mu,
    spec_susie_functional_pi
  ),
  L_grid = L_value,
  y_noise_grid = y_noise_value,
  p_star_grid = p_star_value,
  architecture_grid = architecture_value,
  seeds = seeds,
  prior_quality = tibble::tibble(
    annotation_r2 = annotation_r2_value,
    inflate_match = inflate_match_value
  ),
  sigma_0_2_scalars = sigma_0_2_value,
  data_scenarios = data_scenario,
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),
  task_unit = "dataset",
  bundles_per_task = 1L,
  task_assignment_seed = 42L,
  max_iter = max_iter,
  aggregation_methods = character(0),
  include_overall_pool = FALSE,
  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = FALSE,
  write_snps_parquet = FALSE,
  write_confusion_bins = FALSE,
  write_tier_cs_metrics = TRUE,
  write_scaling_confusion_bins = FALSE,
  metrics_settings = list(
    z_top_k = 10L,
    jsd_threshold = 0.15
  )
)

run_table <- cfg$tables$runs %>%
  tibble::as_tibble() %>%
  dplyr::mutate(
    method_family = .data$use_case_id,
    annotation_label = dplyr::if_else(
      is.na(.data$annotation_r2) | is.na(.data$inflate_match),
      "no_annotation",
      paste0("r2=", .data$annotation_r2, " | inflate=", .data$inflate_match)
    ),
    setting_label = dplyr::case_when(
      .data$use_case_id == "susie_vanilla" ~
        paste0("susie_vanilla | sigma=", .data$sigma_0_2_scalar),
      .data$use_case_id == "susine_functional_mu" ~
        paste0(
          "susine_functional_mu | c=", .data$c_value,
          " | sigma=", .data$sigma_0_2_scalar
        ),
      .data$use_case_id == "susie_functional_pi" ~
        paste0(
          "susie_functional_pi | tau=", .data$tau_value,
          " | sigma=", .data$sigma_0_2_scalar
        ),
      TRUE ~ .data$use_case_id
    )
  )
cfg$tables$runs <- run_table

selected_runs <- run_table %>%
  dplyr::filter(.data$use_case_id %in% cold_method_levels) %>%
  dplyr::arrange(.data$dataset_bundle_id, .data$use_case_id)

readr::write_csv(selected_runs, selected_runs_csv)
readr::write_csv(cfg$tables$dataset_bundles, dataset_bundles_csv)

cat("Matrices selected:", nrow(data_matrix_catalog), "\n")
cat("Dataset bundles:", nrow(cfg$tables$dataset_bundles), "\n")
cat("Selected (cold) run rows:", nrow(selected_runs), "\n")
cat("Cold methods:", paste(sort(unique(selected_runs$use_case_id)), collapse = ", "), "\n")
cat("Annotation r2:", annotation_r2_value, " | inflate_match:", inflate_match_value, "\n")
cat("Mu c:", c_value_mu, " | Pi tau:", tau_value_pi, " | sigma_0_2:", sigma_0_2_value, "\n")

stopifnot(nrow(cfg$tables$dataset_bundles) == n_matrices_target * length(seeds))
stopifnot(nrow(selected_runs) == length(cold_method_levels) * nrow(cfg$tables$dataset_bundles))
```

# 4. Refit And Export

```{r refit-export}
dataset_ids <- sort(as.character(unique(selected_runs$dataset_bundle_id)))
if (is.finite(max_datasets)) {
  dataset_ids <- head(dataset_ids, as.integer(max_datasets))
}

existing_records <- tibble::tibble()
if (!force_refit && file.exists(records_rds)) {
  existing_records <- readRDS(records_rds)
  cat("Loaded existing records:", nrow(existing_records), "\n")
}

completed_ids <- character(0)
if (nrow(existing_records)) {
  completed_ids <- existing_records %>%
    dplyr::count(.data$dataset_bundle_id, .data$model_method, name = "n_effects") %>%
    dplyr::filter(.data$n_effects > 0L) %>%
    dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
    dplyr::filter(.data$n_methods == n_methods_expected) %>%
    dplyr::pull(.data$dataset_bundle_id)
}

todo_ids <- setdiff(dataset_ids, completed_ids)
cat("Datasets queued:", length(dataset_ids), "\n")
cat("Already completed datasets:", length(intersect(dataset_ids, completed_ids)), "\n")
cat("Datasets remaining:", length(todo_ids), "\n")

records <- existing_records
progress_rows <- list()
if (file.exists(progress_csv)) {
  old_progress <- readr::read_csv(progress_csv, show_col_types = FALSE) %>%
    dplyr::mutate(
      dataset_bundle_id = as.character(.data$dataset_bundle_id),
      timestamp = as.character(.data$timestamp)
    )
  progress_rows <- split(old_progress, seq_len(nrow(old_progress)))
}

started <- proc.time()[["elapsed"]]
total_to_run <- length(todo_ids)

for (i in seq_along(todo_ids)) {
  dataset_id <- todo_ids[[i]]
  dataset_started <- proc.time()[["elapsed"]]

  bundle_row <- cfg$tables$dataset_bundles %>%
    tibble::as_tibble() %>%
    dplyr::filter(as.character(.data$dataset_bundle_id) == !!dataset_id)
  if (nrow(bundle_row) != 1L) {
    stop("Expected one dataset bundle row for dataset_bundle_id=", dataset_id, "; found ", nrow(bundle_row), ".")
  }

  data_bundle <- test_susine:::generate_data_for_bundle(bundle_row, cfg)

  dataset_run_rows <- selected_runs %>%
    dplyr::filter(as.character(.data$dataset_bundle_id) == !!dataset_id)

  pick_run_row <- function(method) {
    rr <- dataset_run_rows %>%
      dplyr::filter(.data$method_family == !!method)
    if (nrow(rr) != 1L) {
      stop(
        "Expected exactly one run row for dataset ", dataset_id,
        " and method ", method, "; got ", nrow(rr), "."
      )
    }
    rr
  }

  vanilla_row <- pick_run_row("susie_vanilla")
  mu_row      <- pick_run_row("susine_functional_mu")
  pi_row      <- pick_run_row("susie_functional_pi")

  # Cold fits.
  vanilla_res <- fit_one_method_with_fit(vanilla_row, data_bundle, cfg)
  mu_res      <- fit_one_method_with_fit(mu_row,      data_bundle, cfg)
  warm_mu_res <- fit_warm_refit(
    source_fit = mu_res$fit,
    source_run_row = mu_row,
    data_bundle = data_bundle,
    job_config = cfg,
    model_method_label = "susie_vanilla_warm_from_mu"
  )
  pi_res      <- fit_one_method_with_fit(pi_row,      data_bundle, cfg)
  warm_pi_res <- fit_warm_refit(
    source_fit = pi_res$fit,
    source_run_row = pi_row,
    data_bundle = data_bundle,
    job_config = cfg,
    model_method_label = "susie_vanilla_warm_from_pi"
  )

  dataset_records <- dplyr::bind_rows(
    vanilla_res$records,
    mu_res$records,
    warm_mu_res$records,
    pi_res$records,
    warm_pi_res$records
  )

  records <- dplyr::bind_rows(records, dataset_records)

  elapsed <- proc.time()[["elapsed"]] - started
  avg_sec <- elapsed / i
  eta <- avg_sec * (total_to_run - i)
  dataset_elapsed <- proc.time()[["elapsed"]] - dataset_started

  progress_row <- tibble::tibble(
    completed_this_session = i,
    total_this_session = total_to_run,
    completed_overall = length(intersect(dataset_ids, completed_ids)) + i,
    total_overall = length(dataset_ids),
    dataset_bundle_id = dataset_id,
    dataset_elapsed_sec = dataset_elapsed,
    elapsed_sec = elapsed,
    eta_sec = eta,
    timestamp = as.character(Sys.time())
  )
  progress_rows[[length(progress_rows) + 1L]] <- progress_row

  cat(sprintf(
    "%d/%d datasets fitted | elapsed: %s | avg: %.1f sec/dataset | ETA: %s | dataset_id: %s\n",
    length(intersect(dataset_ids, completed_ids)) + i,
    length(dataset_ids),
    format_duration(elapsed),
    avg_sec,
    format_duration(eta),
    dataset_id
  ))

  if (i %% checkpoint_every == 0L || i == total_to_run) {
    save_checkpoint(records, progress_rows)
    cat("Checkpoint saved:", records_rds, "\n")
  }
}

if (!length(todo_ids)) {
  save_checkpoint(records, progress_rows)
}

cat("Final record rows:", nrow(records), "\n")
cat("Final datasets:", dplyr::n_distinct(records$dataset_bundle_id), "\n")
cat("Records RDS:", records_rds, "\n")
cat("Metadata CSV:", metadata_csv, "\n")
```

# 5. Sanity Checks

```{r sanity}
records_check <- readRDS(records_rds)

summary_counts <- records_check %>%
  dplyr::count(.data$dataset_bundle_id, .data$model_method, name = "n_effects") %>%
  dplyr::count(.data$model_method, .data$n_effects, name = "n_datasets")

kable(summary_counts)

vector_lengths <- records_check %>%
  dplyr::mutate(fitted_y_length = vapply(.data$fitted_y_effect, length, integer(1))) %>%
  dplyr::count(.data$model_method, .data$fitted_y_length, name = "n_effect_rows")

kable(vector_lengths)

if (any(vector_lengths$fitted_y_length <= 0L)) {
  stop("At least one fitted_y_effect vector is empty.")
}

methods_present <- sort(unique(records_check$model_method))
cat("Distinct model_method values in records:",
    paste(methods_present, collapse = ", "), "\n")
if (!setequal(methods_present, all_method_levels)) {
  warning(
    "Records do not contain exactly the expected five methods. Expected: ",
    paste(all_method_levels, collapse = ", "),
    ". Got: ",
    paste(methods_present, collapse = ", ")
  )
}

# AUPRC sanity: per-dataset, susine_vanilla_warm_from_mu should NOT always
# match the cold susie_vanilla AUPRC exactly (otherwise the warm refit isn't
# actually re-running IBSS). And it should NOT always match susine_functional_mu
# either (otherwise vanilla isn't being applied as the prior).
auprc_wide <- records_check %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$model_method, .data$model_AUPRC) %>%
  tidyr::pivot_wider(names_from = .data$model_method, values_from = .data$model_AUPRC)

if (all(c("susie_vanilla", "susie_vanilla_warm_from_mu") %in% names(auprc_wide))) {
  identical_to_cold <- mean(
    abs(auprc_wide$susie_vanilla - auprc_wide$susie_vanilla_warm_from_mu) < 1e-9,
    na.rm = TRUE
  )
  cat(sprintf(
    "Fraction of datasets where vanilla_warm_from_mu AUPRC == cold vanilla AUPRC exactly: %.3f\n",
    identical_to_cold
  ))
  if (identical_to_cold > 0.95) {
    warning("Almost all warm-from-mu refits produced AUPRC identical to cold vanilla. ",
            "Check that init_effect_fits is taking effect (warm refit may not be re-running IBSS).")
  }
}
```
