---
title: "Effect Drift Analysis (5-arm: cold + warm refits)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Reads the 5-arm refit records and computes per-dataset **effect drift**
between every interesting arm pair, plus an AUPRC trajectory across the
two refit chains:

```
mu chain:                              pi chain:
  vanilla -> mu                          vanilla -> pi
  mu -> vanilla_warm_from_mu             pi -> vanilla_warm_from_pi
  vanilla -> vanilla_warm_from_mu        vanilla -> vanilla_warm_from_pi
```

**Effect drift** between two fits A and B is computed by:

1. Build the `L x L` correlation matrix of `fitted_y_effect` vectors
   (inactive effects with sd=0 contribute correlation 0).
2. Solve the max-weight 1-1 assignment via `clue::solve_LSAP(cor_mat,
   maximum = TRUE)`.
3. Record all `L` assignment correlations.
4. Derive per-pair summaries: `mean_corr`, `min_corr`, `n_below_0p9`
   (count of assignment edges with correlation < 0.9), legacy
   `weighted_drift` (`1 - corr`), and the main basin-drift metric
   `weighted_chord_drift` (`sqrt(2 * (1 - corr))`), with both weighted
   by `min(var_y_explained_A_i, var_y_explained_B_j)` so inactive pairs
   contribute little.

This drift framework + AUPRC trajectory operationalizes the
**refinement-vs-exploration test**:

| D(vanilla, treatment) | D(vanilla, vanilla_warm) | dAUPRC(treatment) | dAUPRC(vanilla_warm) | Mechanism |
|---|---|---|---|---|
| Low | Low | + | + | Within-basin PIP weighting; basin unchanged |
| Low | Low | + | ~0 | PIP-weighting only; reverts when prior removed |
| High | High | + | + | **Exploration win**: new basin, vanilla holds it |
| High | Low | + | ~0 | New basin but unstable; vanilla snaps back |

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_drift_analysis_5arm.Rmd")
library(here)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)
library(clue)         # solve_LSAP for max-weight 1-1 assignment

theme_set(theme_bw(base_size = 12))
```

# 1. User Parameters

```{r user-parameters}
output_root <- here("output")
study_name <- "baseline_sims_5arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
records_rds <- file.path(study_dir, "effect_refit_records.rds")

baseline_method <- "susie_vanilla"
mu_method <- "susine_functional_mu"
pi_method <- "susie_functional_pi"
warm_mu_method <- "susie_vanilla_warm_from_mu"
warm_pi_method <- "susie_vanilla_warm_from_pi"

# Pretty labels for plotting.
method_labels <- c(
  susie_vanilla = "vanilla (cold)",
  susine_functional_mu = "mu (cold)",
  susie_vanilla_warm_from_mu = "vanilla (warm-from-mu)",
  susie_functional_pi = "pi (cold)",
  susie_vanilla_warm_from_pi = "vanilla (warm-from-pi)"
)
method_color <- c(
  susie_vanilla = "black",
  susine_functional_mu = "#2b8cbe",
  susie_vanilla_warm_from_mu = "#9ecae1",
  susie_functional_pi = "#74c476",
  susie_vanilla_warm_from_pi = "#bae4b3"
)

# Drift threshold (the user's "unpaired" cutoff).
drift_corr_cutoff <- 0.9

# Active-effect threshold for the weighted/filtered drift summaries.
active_var_threshold <- 0.005

# Effect-pair threshold used for matched/unmatched diagnostic density plots.
effect_match_corr_min <- 0.99

# Joint effect-quality axes. `effect_pip_entropy_core95` is log effective k;
# axis labels are shown on the effective-k scale.
x_metric <- "effect_pip_entropy_core95"
y_metric <- "accuracy_ratio"
x_label <- "Core-95 effective k (= exp(entropy))"
y_label <- "Accuracy ratio"
x_keff_breaks <- c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000)
x_limits <- NULL
y_limits <- c(0, 1)

density_grid_n <- 140L
density_bandwidth <- c(0.45, 0.10)
density_trim_quantile <- c(0.005, 0.995)
density_fill_quantile <- 0.95

# Plot output directory.
plot_dir <- file.path(study_dir, "drift_analysis_plots")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Output filenames.
drift_assignments_csv <- file.path(study_dir, "drift_assignments_5arm.csv")
drift_summary_csv <- file.path(study_dir, "drift_summary_5arm.csv")
auprc_per_dataset_csv <- file.path(study_dir, "auprc_per_dataset_5arm.csv")

drift_profile_hist_png <- file.path(plot_dir, "drift_profile_histograms.png")
chain_trajectory_png <- file.path(plot_dir, "chain_auprc_trajectory.png")
headline_scatter_png <- file.path(plot_dir, "headline_drift_vs_warm_auprc.png")
unpaired_vs_delta_png <- file.path(plot_dir, "unpaired_count_vs_auprc_delta.png")
mu_vs_pi_drift_png <- file.path(plot_dir, "mu_vs_pi_drift_comparison.png")
basin_pip_drift_png <- file.path(plot_dir, "basin_and_pip_drift_summary.png")
effect_quality_density_mu_png <- file.path(
  plot_dir, "effect_quality_density_matched_unmatched_mu_5arm.png"
)
effect_quality_density_pi_png <- file.path(
  plot_dir, "effect_quality_density_matched_unmatched_pi_5arm.png"
)
effect_quality_pair_rows_csv <- file.path(
  study_dir, "effect_quality_pair_rows_5arm.csv"
)
pip_records_rds <- file.path(study_dir, "pip_records_5arm.rds")
pip_drift_summary_csv <- file.path(study_dir, "pip_drift_summary_5arm.csv")
causal_idx_rds <- file.path(study_dir, "causal_idx_5arm.rds")
pr_curves_pooled_png <- file.path(plot_dir, "pr_curves_pooled.png")
pr_curves_pooled_csv <- file.path(study_dir, "pr_curves_pooled_5arm.csv")

cat("Study dir:", study_dir, "\n")
cat("Plot dir:", plot_dir, "\n")
cat("Drift threshold for `n_below_<cut>`:", drift_corr_cutoff, "\n")
cat("Active-effect var-explained threshold:", active_var_threshold, "\n")
cat("Effect-quality matched threshold:", effect_match_corr_min, "\n")
```

# 2. Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

format_duration <- function(seconds) {
  seconds <- as.numeric(seconds)
  if (!is.finite(seconds)) return("unknown")
  h <- floor(seconds / 3600)
  m <- floor((seconds %% 3600) / 60)
  s <- round(seconds %% 60)
  sprintf("%02d:%02d:%02d", h, m, s)
}

require_metric <- function(df, metric) {
  if (!metric %in% names(df)) {
    stop("Metric column not found: ", metric)
  }
}

resolve_limits <- function(values, user_limits, trim = density_trim_quantile) {
  if (!is.null(user_limits)) {
    return(user_limits)
  }
  values <- values[is.finite(values)]
  if (!length(values)) {
    stop("Cannot infer limits from an empty/non-finite vector.")
  }
  stats::quantile(values, probs = trim, na.rm = TRUE, names = FALSE)
}

keff_x_scale <- function(breaks = x_keff_breaks, xlim = NULL) {
  finite_breaks <- breaks[breaks > 0]
  scale_x_continuous(
    breaks = log(finite_breaks),
    labels = formatC(finite_breaks, format = "fg", big.mark = ","),
    limits = xlim
  )
}

safe_corr <- function(x) {
  pmax(-1, pmin(1, x))
}

chord_distance <- function(corr) {
  sqrt(pmax(0, 2 * (1 - safe_corr(corr))))
}

# Stack effect rows for one (dataset, model) into an n x L matrix of
# fitted_y_effect vectors, in effect order.
records_to_fitted_y_matrix <- function(rows) {
  rows <- rows %>% dplyr::arrange(.data$effect)
  mat <- do.call(cbind, rows$fitted_y_effect)
  storage.mode(mat) <- "double"
  colnames(mat) <- as.character(rows$effect)
  mat
}

# Correlation matrix robust to inactive (sd=0) effects.
safe_cor_matrix <- function(mat_a, mat_b, sd_eps = 1e-12) {
  if (nrow(mat_a) != nrow(mat_b)) {
    stop("Mismatched n (", nrow(mat_a), " vs ", nrow(mat_b), ").")
  }
  sd_a <- apply(mat_a, 2L, stats::sd)
  sd_b <- apply(mat_b, 2L, stats::sd)
  active_a <- is.finite(sd_a) & sd_a > sd_eps
  active_b <- is.finite(sd_b) & sd_b > sd_eps

  cor_mat <- matrix(0, nrow = ncol(mat_a), ncol = ncol(mat_b),
                    dimnames = list(colnames(mat_a), colnames(mat_b)))
  if (any(active_a) && any(active_b)) {
    sub <- suppressWarnings(stats::cor(
      mat_a[, active_a, drop = FALSE],
      mat_b[, active_b, drop = FALSE],
      use = "pairwise.complete.obs"
    ))
    sub[!is.finite(sub)] <- 0
    cor_mat[active_a, active_b] <- sub
  }
  list(cor = cor_mat, active_a = active_a, active_b = active_b)
}

# For one (dataset, A, B) pair, run Hungarian on the L x L correlation
# matrix and return per-edge data.
compute_drift_pair <- function(rows_a, rows_b,
                               method_a, method_b,
                               dataset_bundle_id, matrix_id) {
  rows_a <- rows_a %>% dplyr::arrange(.data$effect)
  rows_b <- rows_b %>% dplyr::arrange(.data$effect)

  mat_a <- records_to_fitted_y_matrix(rows_a)
  mat_b <- records_to_fitted_y_matrix(rows_b)
  cor_info <- safe_cor_matrix(mat_a, mat_b)
  cor_mat <- cor_info$cor

  # clue::solve_LSAP() requires nonnegative entries even with
  # maximum = TRUE. Shift the full matrix if needed; adding a constant
  # to every edge preserves the optimal 1-1 assignment.
  assignment_weights <- cor_mat
  min_weight <- suppressWarnings(min(assignment_weights, na.rm = TRUE))
  if (!is.finite(min_weight)) {
    min_weight <- 0
  }
  if (min_weight < 0) {
    assignment_weights <- assignment_weights - min_weight
  }
  assignment <- as.integer(clue::solve_LSAP(
    assignment_weights,
    maximum = TRUE
  ))
  L <- ncol(mat_a)

  effect_a_ids <- as.integer(rows_a$effect)
  effect_b_ids <- as.integer(rows_b$effect)
  var_y_a <- as.numeric(rows_a$var_y_explained_frac)
  var_y_b <- as.numeric(rows_b$var_y_explained_frac)

  tibble::tibble(
    dataset_bundle_id = dataset_bundle_id,
    matrix_id = matrix_id,
    method_a = method_a,
    method_b = method_b,
    edge = seq_len(L),
    effect_a = effect_a_ids,
    effect_b = effect_b_ids[assignment],
    corr = cor_mat[cbind(seq_len(L), assignment)],
    var_y_explained_a = var_y_a,
    var_y_explained_b = var_y_b[assignment],
    active_a = cor_info$active_a,
    active_b = cor_info$active_b[assignment]
  )
}

# Summarize per (dataset, A, B) from the per-edge assignment table.
summarize_drift <- function(edges_df, drift_corr_cutoff, active_var_threshold) {
  edges_df %>%
    dplyr::group_by(
      .data$dataset_bundle_id, .data$matrix_id,
      .data$method_a, .data$method_b
    ) %>%
    dplyr::summarise(
      n_edges = dplyr::n(),
      mean_corr = mean(.data$corr, na.rm = TRUE),
      median_corr = stats::median(.data$corr, na.rm = TRUE),
      min_corr = suppressWarnings(min(.data$corr, na.rm = TRUE)),
      n_below_cutoff = sum(.data$corr < drift_corr_cutoff, na.rm = TRUE),
      n_active_pairs = sum(
        .data$var_y_explained_a > active_var_threshold &
          .data$var_y_explained_b > active_var_threshold,
        na.rm = TRUE
      ),
      n_active_below_cutoff = sum(
        .data$corr < drift_corr_cutoff &
          .data$var_y_explained_a > active_var_threshold &
          .data$var_y_explained_b > active_var_threshold,
        na.rm = TRUE
      ),
      mean_corr_active = {
        keep <- .data$var_y_explained_a > active_var_threshold &
                  .data$var_y_explained_b > active_var_threshold
        if (any(keep)) mean(.data$corr[keep], na.rm = TRUE) else NA_real_
      },
      weighted_drift = {
        w <- pmin(
          tidyr::replace_na(.data$var_y_explained_a, 0),
          tidyr::replace_na(.data$var_y_explained_b, 0)
        )
        if (sum(w) > 0) {
          # 1 - weighted-mean correlation, so larger = more drift
          1 - sum(w * .data$corr, na.rm = TRUE) / sum(w)
        } else {
          NA_real_
        }
      },
      weighted_chord_drift = {
        w <- pmin(
          tidyr::replace_na(.data$var_y_explained_a, 0),
          tidyr::replace_na(.data$var_y_explained_b, 0)
        )
        if (sum(w) > 0) {
          sum(w * chord_distance(.data$corr), na.rm = TRUE) / sum(w)
        } else {
          NA_real_
        }
      },
      .groups = "drop"
    )
}
```

# 3. Load Records

```{r load-records}
stopifnot(file.exists(records_rds))
records <- readRDS(records_rds) %>%
  dplyr::mutate(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    matrix_id = as.character(.data$matrix_id),
    effect = as.integer(.data$effect)
  )

methods_present <- sort(unique(records$model_method))
cat("Methods in records:", paste(methods_present, collapse = ", "), "\n")

required_methods <- c(baseline_method, mu_method, warm_mu_method,
                      pi_method, warm_pi_method)
missing_methods <- setdiff(required_methods, methods_present)
if (length(missing_methods)) {
  stop("Missing methods in records: ", paste(missing_methods, collapse = ", "),
       ". Re-run effect_matching_refit_export_5arm.Rmd first.")
}

records <- records %>% dplyr::filter(.data$model_method %in% required_methods)
require_metric(records, x_metric)
require_metric(records, y_metric)

# Index records by (dataset, model_method) for fast pair access.
records_by_dataset_method <- records %>% split(list(.$dataset_bundle_id, .$model_method))
```

# 4. Pairwise Drift

```{r compute-drift}
# Pairs we want (label, method_a, method_b).
pair_specs <- tibble::tribble(
  ~pair_id,                        ~chain, ~method_a,         ~method_b,
  "vanilla__mu",                   "mu",   baseline_method,   mu_method,
  "mu__vanilla_warm_from_mu",      "mu",   mu_method,         warm_mu_method,
  "vanilla__vanilla_warm_from_mu", "mu",   baseline_method,   warm_mu_method,
  "vanilla__pi",                   "pi",   baseline_method,   pi_method,
  "pi__vanilla_warm_from_pi",      "pi",   pi_method,         warm_pi_method,
  "vanilla__vanilla_warm_from_pi", "pi",   baseline_method,   warm_pi_method
)
print(pair_specs)

# Datasets with all 5 arms present.
datasets_complete <- records %>%
  dplyr::count(.data$dataset_bundle_id, .data$model_method) %>%
  dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
  dplyr::filter(.data$n_methods == length(required_methods)) %>%
  dplyr::pull(.data$dataset_bundle_id)
cat("Datasets with all 5 arms present:", length(datasets_complete), "\n")

dataset_meta <- records %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$matrix_id) %>%
  dplyr::filter(.data$dataset_bundle_id %in% datasets_complete)

started <- proc.time()[["elapsed"]]
edge_chunks <- list()
n_total <- length(datasets_complete) * nrow(pair_specs)
counter <- 0L

for (ds_idx in seq_along(datasets_complete)) {
  ds_id <- datasets_complete[[ds_idx]]
  matrix_id_i <- dataset_meta$matrix_id[
    match(ds_id, dataset_meta$dataset_bundle_id)
  ]
  for (pi_idx in seq_len(nrow(pair_specs))) {
    counter <- counter + 1L
    spec <- pair_specs[pi_idx, ]
    rows_a <- records_by_dataset_method[[paste(ds_id, spec$method_a, sep = ".")]]
    rows_b <- records_by_dataset_method[[paste(ds_id, spec$method_b, sep = ".")]]
    if (is.null(rows_a) || is.null(rows_b)) {
      next
    }
    edge_chunks[[length(edge_chunks) + 1L]] <- compute_drift_pair(
      rows_a = rows_a, rows_b = rows_b,
      method_a = spec$method_a, method_b = spec$method_b,
      dataset_bundle_id = ds_id,
      matrix_id = matrix_id_i
    ) %>% dplyr::mutate(pair_id = spec$pair_id, chain = spec$chain)
  }
  if (ds_idx %% 100L == 0L || ds_idx == length(datasets_complete)) {
    elapsed <- proc.time()[["elapsed"]] - started
    eta <- elapsed / counter * (n_total - counter)
    cat(sprintf(
      "Drift %d/%d datasets (%d/%d pairs) | elapsed %s | ETA %s\n",
      ds_idx, length(datasets_complete), counter, n_total,
      format_duration(elapsed), format_duration(eta)
    ))
  }
}

drift_assignments <- dplyr::bind_rows(edge_chunks)
readr::write_csv(drift_assignments, drift_assignments_csv)
cat("Drift assignment rows:", nrow(drift_assignments), "\n")

drift_summary <- summarize_drift(drift_assignments, drift_corr_cutoff, active_var_threshold) %>%
  dplyr::left_join(pair_specs, by = c("method_a", "method_b"))
readr::write_csv(drift_summary, drift_summary_csv)
cat("Drift summary rows:", nrow(drift_summary), "\n")
```

# 5. Drift Sanity (D(arm, arm) == 0)

```{r drift-sanity, fig.width=8, fig.height=4}
# As a self-test, compute drift of vanilla against itself for a small subset.
# All correlations should be 1, mean_corr ~ 1, n_below_0p9 == 0.
sanity_ids <- head(datasets_complete, 10L)
sanity_rows <- list()
for (ds_id in sanity_ids) {
  rows_v <- records_by_dataset_method[[paste(ds_id, baseline_method, sep = ".")]]
  if (is.null(rows_v)) next
  matrix_id_i <- dataset_meta$matrix_id[match(ds_id, dataset_meta$dataset_bundle_id)]
  sanity_rows[[length(sanity_rows) + 1L]] <- compute_drift_pair(
    rows_a = rows_v, rows_b = rows_v,
    method_a = baseline_method, method_b = baseline_method,
    dataset_bundle_id = ds_id, matrix_id = matrix_id_i
  )
}
sanity_edges <- dplyr::bind_rows(sanity_rows)
sanity_summary <- summarize_drift(sanity_edges, drift_corr_cutoff, active_var_threshold)
cat("Self-drift sanity (vanilla vs vanilla):\n")
print(sanity_summary %>%
        dplyr::summarise(
          mean_of_mean_corr = mean(.data$mean_corr, na.rm = TRUE),
          mean_n_below_cutoff = mean(.data$n_below_cutoff),
          max_n_below_cutoff = max(.data$n_below_cutoff)
        ))
# Note: inactive (sd=0) effects against themselves give correlation 0 by
# construction, so n_below_cutoff > 0 is expected. mean_corr_active should
# be close to 1.
cat("Self-drift mean corr (active-only):", mean(sanity_summary$mean_corr_active, na.rm = TRUE), "\n")
```

# 6. Per-Dataset AUPRC Trajectory

```{r per-dataset-auprc}
auprc_per_dataset <- records %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$model_method, .data$model_AUPRC) %>%
  tidyr::pivot_wider(
    names_from = .data$model_method,
    values_from = .data$model_AUPRC,
    names_prefix = "auprc_"
  ) %>%
  dplyr::mutate(
    auprc_delta_mu = .data$auprc_susine_functional_mu - .data$auprc_susie_vanilla,
    auprc_delta_pi = .data$auprc_susie_functional_pi - .data$auprc_susie_vanilla,
    auprc_delta_warm_mu_vs_vanilla =
      .data$auprc_susie_vanilla_warm_from_mu - .data$auprc_susie_vanilla,
    auprc_delta_warm_pi_vs_vanilla =
      .data$auprc_susie_vanilla_warm_from_pi - .data$auprc_susie_vanilla,
    auprc_round_trip_mu =
      .data$auprc_susie_vanilla_warm_from_mu - .data$auprc_susine_functional_mu,
    auprc_round_trip_pi =
      .data$auprc_susie_vanilla_warm_from_pi - .data$auprc_susie_functional_pi
  )
readr::write_csv(auprc_per_dataset, auprc_per_dataset_csv)
cat("Per-dataset AUPRC rows:", nrow(auprc_per_dataset), "\n")

# Long format for spaghetti / facet plots.
auprc_long <- auprc_per_dataset %>%
  tidyr::pivot_longer(
    cols = dplyr::starts_with("auprc_"),
    names_to = "metric",
    values_to = "value"
  )
```

# 7. Drift Profile Histograms

```{r drift-profile-hists, fig.width=12, fig.height=7}
drift_assignments_for_plot <- drift_assignments %>%
  dplyr::mutate(
    pair_id = factor(.data$pair_id, levels = pair_specs$pair_id)
  )

p_drift_hists <- ggplot(drift_assignments_for_plot,
                        aes(x = .data$corr, fill = .data$chain)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 60, position = "identity", alpha = 0.55, color = NA
  ) +
  geom_vline(xintercept = drift_corr_cutoff, linetype = "dashed", color = "grey25") +
  facet_wrap(~ pair_id, ncol = 3) +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  coord_cartesian(xlim = c(-0.05, 1.0)) +
  labs(
    x = "Hungarian-assigned fitted_y correlation per effect",
    y = "Density",
    fill = "Chain",
    title = "Distribution of per-effect assignment correlations across all pairs",
    subtitle = sprintf(
      "Dashed line = drift_corr_cutoff (%.2f). All 10 effects per (dataset, pair) included.",
      drift_corr_cutoff
    )
  ) +
  theme(legend.position = "bottom")

ggsave(drift_profile_hist_png, p_drift_hists, width = 12, height = 7, dpi = 150)
print(p_drift_hists)
```

# 8. Chain AUPRC Trajectory

```{r chain-trajectory, fig.width=11, fig.height=5}
chain_long <- dplyr::bind_rows(
  auprc_per_dataset %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step1 = .data$auprc_susie_vanilla,
      step2 = .data$auprc_susine_functional_mu,
      step3 = .data$auprc_susie_vanilla_warm_from_mu
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step1 = .data$auprc_susie_vanilla,
      step2 = .data$auprc_susie_functional_pi,
      step3 = .data$auprc_susie_vanilla_warm_from_pi
    )
) %>%
  tidyr::pivot_longer(
    cols = c("step1", "step2", "step3"),
    names_to = "step",
    values_to = "auprc"
  ) %>%
  dplyr::mutate(
    step_label = dplyr::case_when(
      step == "step1" ~ "vanilla (cold)",
      step == "step2" & chain == "mu" ~ "mu",
      step == "step2" & chain == "pi" ~ "pi",
      step == "step3" ~ "vanilla (warm)",
      TRUE ~ step
    ),
    step_label = factor(step_label,
                        levels = c("vanilla (cold)", "mu", "pi", "vanilla (warm)"))
  )

p_chain <- ggplot(chain_long, aes(x = .data$step_label, y = .data$auprc)) +
  geom_line(aes(group = .data$dataset_bundle_id), alpha = 0.06) +
  geom_violin(aes(fill = .data$chain), alpha = 0.55, color = NA, scale = "width") +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.7) +
  facet_wrap(~ chain, scales = "free_x") +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  labs(
    x = NULL, y = "AUPRC",
    fill = "Chain",
    title = "Per-dataset AUPRC trajectory along each refit chain",
    subtitle = "Each thin line = one dataset; violin / box = marginal distribution per step."
  ) +
  theme(legend.position = "bottom")

ggsave(chain_trajectory_png, p_chain, width = 11, height = 5, dpi = 150)
print(p_chain)
```

# 9. Headline 4-Quadrant Test

```{r headline-scatter, fig.width=12, fig.height=5.5}
# Join drift(vanilla, vanilla_warm) with AUPRC delta(vanilla_warm - vanilla)
headline_df_mu <- drift_summary %>%
  dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_mu") %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    chain = "mu",
    drift_mean_corr = .data$mean_corr,
    drift_mean_corr_active = .data$mean_corr_active,
    n_below_cutoff = .data$n_below_cutoff
  ) %>%
  dplyr::left_join(
    auprc_per_dataset %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        auprc_delta_warm_vs_vanilla = .data$auprc_delta_warm_mu_vs_vanilla,
        auprc_delta_treatment = .data$auprc_delta_mu
      ),
    by = "dataset_bundle_id"
  )

headline_df_pi <- drift_summary %>%
  dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_pi") %>%
  dplyr::transmute(
    dataset_bundle_id = .data$dataset_bundle_id,
    chain = "pi",
    drift_mean_corr = .data$mean_corr,
    drift_mean_corr_active = .data$mean_corr_active,
    n_below_cutoff = .data$n_below_cutoff
  ) %>%
  dplyr::left_join(
    auprc_per_dataset %>%
      dplyr::transmute(
        dataset_bundle_id = .data$dataset_bundle_id,
        auprc_delta_warm_vs_vanilla = .data$auprc_delta_warm_pi_vs_vanilla,
        auprc_delta_treatment = .data$auprc_delta_pi
      ),
    by = "dataset_bundle_id"
  )

headline_df <- dplyr::bind_rows(headline_df_mu, headline_df_pi)

p_headline <- ggplot(
  headline_df,
  aes(x = .data$drift_mean_corr_active,
      y = .data$auprc_delta_warm_vs_vanilla,
      color = .data$auprc_delta_treatment)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_vline(xintercept = drift_corr_cutoff, color = "grey45", linetype = "dotted") +
  geom_point(alpha = 0.55, size = 1.4) +
  scale_color_gradient2(
    low = "#2b6cb0", mid = "white", high = "#c53030", midpoint = 0
  ) +
  facet_wrap(~ chain, labeller = labeller(chain = c(mu = "mu chain", pi = "pi chain"))) +
  scale_x_continuous(limits = c(0, 1)) +
  labs(
    x = "Mean active-effect drift D(vanilla, vanilla_warm) (1 = identical)",
    y = "AUPRC(vanilla_warm) - AUPRC(vanilla_cold)",
    color = "AUPRC(treatment)\n- AUPRC(vanilla)",
    title = "Headline: refinement vs exploration — does the warm vanilla refit retain treatment basin?",
    subtitle = "Top-right quadrant (high drift + high warm-vanilla ΔAUPRC) = exploration win. Top-left = pure within-basin PIP weighting."
  )

ggsave(headline_scatter_png, p_headline, width = 12, height = 5.5, dpi = 150)
print(p_headline)
```

# 10. Unpaired Count vs AUPRC Delta (User's Plot)

```{r unpaired-vs-delta, fig.width=12, fig.height=8}
# Build a long table: per (dataset, chain, step_label), report n_below_cutoff
# (drift) and the relevant AUPRC delta for that step.
step_drift_delta <- dplyr::bind_rows(
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step_label = "step 1: vanilla -> mu",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_delta_mu),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "mu__vanilla_warm_from_mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step_label = "step 2: mu -> vanilla_warm",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_round_trip_mu),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step_label = "step 1: vanilla -> pi",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_delta_pi),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "pi__vanilla_warm_from_pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step_label = "step 2: pi -> vanilla_warm",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_round_trip_pi),
      by = "dataset_bundle_id"
    )
)

step_rho <- step_drift_delta %>%
  dplyr::group_by(.data$chain, .data$step_label) %>%
  dplyr::summarise(
    n = dplyr::n(),
    rho = suppressWarnings(stats::cor(
      .data$n_below_cutoff, .data$auprc_delta,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    label = sprintf("rho = %.3f (n=%d)", .data$rho, .data$n),
    .groups = "drop"
  )

p_unpaired <- ggplot(
  step_drift_delta,
  aes(x = .data$n_below_cutoff, y = .data$auprc_delta)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_jitter(width = 0.25, height = 0, alpha = 0.45, size = 1.1,
              aes(color = .data$mean_corr)) +
  geom_smooth(method = "loess", formula = y ~ x, se = FALSE,
              color = "black", linetype = "dashed", linewidth = 0.5) +
  geom_text(
    data = step_rho,
    aes(x = -Inf, y = Inf, label = .data$label),
    hjust = -0.05, vjust = 1.4, size = 3.3, inherit.aes = FALSE
  ) +
  facet_grid(chain ~ step_label) +
  scale_x_continuous(breaks = 0:10) +
  scale_color_viridis_c(option = "C") +
  labs(
    x = sprintf("# unpaired effects (assignment corr < %.2f)", drift_corr_cutoff),
    y = "AUPRC delta for this step",
    color = "Mean active\nassignment corr",
    title = sprintf(
      "User's hypothesis test: does # of unpaired effects (cutoff %.2f) predict AUPRC change at each step?",
      drift_corr_cutoff
    ),
    subtitle = "Step 1: cold->treatment (positive y => treatment wins). Step 2: treatment->vanilla_warm (positive y => warm refit IMPROVES on treatment, rare; negative => snaps back to vanilla)."
  )

ggsave(unpaired_vs_delta_png, p_unpaired, width = 12, height = 8, dpi = 150)
print(p_unpaired)
```

# 11. Mu vs Pi Drift Comparison

Two panels per dataset: the cold treatment fit (vanilla -> annotated)
answers "did the prior move the model at all?", and the warm-vanilla refit
(vanilla -> vanilla_warm) answers "does the new basin survive when the
prior is removed?". Stacking them side-by-side lets the reader read
exploration vs basin-retention off a single figure.

```{r mu-vs-pi-drift, fig.width=11, fig.height=5.5}
mu_vs_pi_long <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "vanilla__pi",
    "vanilla__vanilla_warm_from_mu", "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    .data$dataset_bundle_id,
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::if_else(
      .data$pair_id %in% c("vanilla__mu", "vanilla__pi"),
      "vanilla -> annotated",
      "vanilla -> vanilla_warm"
    ),
    weighted_chord_drift = .data$weighted_chord_drift
  )

mu_vs_pi_wide <- mu_vs_pi_long %>%
  tidyr::pivot_wider(
    id_cols = c("dataset_bundle_id", "pair_kind"),
    names_from = "chain",
    values_from = "weighted_chord_drift"
  ) %>%
  dplyr::mutate(
    pair_kind = factor(
      .data$pair_kind,
      levels = c("vanilla -> annotated", "vanilla -> vanilla_warm")
    )
  )

max_mupi_chord <- max(
  mu_vs_pi_wide$mu, mu_vs_pi_wide$pi, na.rm = TRUE
)
max_mupi_chord <- max(0.05, max_mupi_chord)

p_mupi <- ggplot(
  mu_vs_pi_wide,
  aes(x = .data$mu, y = .data$pi)
) +
  geom_abline(slope = 1, intercept = 0,
              color = "grey45", linetype = "dashed") +
  geom_point(alpha = 0.5, size = 1.2) +
  facet_wrap(~ pair_kind, ncol = 2) +
  coord_fixed(xlim = c(0, max_mupi_chord),
              ylim = c(0, max_mupi_chord)) +
  labs(
    x = "Weighted chord basin drift: mu chain",
    y = "Weighted chord basin drift: pi chain",
    title = "How much does each annotation mechanism move the model? (cold treatment vs warm-vanilla refit)",
    subtitle = "Left: cold treatment fit (exploration). Right: warm-vanilla refit (basin retention). y=x line marks equal mu/pi drift."
  )

ggsave(mu_vs_pi_drift_png, p_mupi, width = 11, height = 5.5, dpi = 150)
print(p_mupi)
```

# 12. Per-Dataset PIP JSD

Loads the per-fit model-wide PIP sidecar (`pip_records_5arm.rds`, written by
`effect_matching_refit_export_5arm.Rmd`) and computes Jensen-Shannon
divergence between matched (vanilla, treatment) and (vanilla, vanilla_warm)
PIP vectors for each dataset. JSD definition mirrors
`R/run_model.R::js_distance`: a categorical JSD on raw model-wide PIP
vectors with an additive epsilon, summed (not averaged) across SNPs. See
the study plan note "JSD ... currently computed on raw model-wide PIPs".

```{r pip-jsd}
# Inline copy of test_susine:::js_distance (R/run_model.R:2021) so the Rmd
# does not depend on devtools::load_all() for the package.
js_distance <- function(p, q, eps = 1e-12) {
  p <- as.numeric(p); q <- as.numeric(q)
  p <- p + eps; q <- q + eps
  m <- 0.5 * (p + q)
  kl <- function(a, b) sum(a * log(a / b))
  0.5 * kl(p, m) + 0.5 * kl(q, m)
}

# Pairs we want for the master figure: (chain, pair_kind) -> (method_a, method_b)
pip_pair_specs <- tibble::tribble(
  ~chain, ~pair_kind,             ~method_a,         ~method_b,
  "mu",   "vanilla_to_treatment", baseline_method,   mu_method,
  "mu",   "treatment_to_warm",    mu_method,         warm_mu_method,
  "mu",   "vanilla_to_warm",      baseline_method,   warm_mu_method,
  "pi",   "vanilla_to_treatment", baseline_method,   pi_method,
  "pi",   "treatment_to_warm",    pi_method,         warm_pi_method,
  "pi",   "vanilla_to_warm",      baseline_method,   warm_pi_method
)

if (file.exists(pip_records_rds)) {
  pip_records <- readRDS(pip_records_rds) %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

  cat("PIP records rows:", nrow(pip_records), "\n")
  cat("Datasets in PIP records:",
      dplyr::n_distinct(pip_records$dataset_bundle_id), "\n")

  pip_by_dataset_method <- pip_records %>%
    dplyr::select(.data$dataset_bundle_id, .data$model_method, .data$pip)

  # Sanity: self-JSD (vanilla vs vanilla) should be exactly 0.
  self_check_ids <- head(unique(pip_records$dataset_bundle_id), 5L)
  self_jsd <- vapply(self_check_ids, function(ds_id) {
    v <- pip_by_dataset_method %>%
      dplyr::filter(.data$dataset_bundle_id == ds_id,
                    .data$model_method == baseline_method) %>%
      dplyr::pull(.data$pip) %>% .[[1]]
    js_distance(v, v)
  }, numeric(1))
  cat("Self-JSD (vanilla vs vanilla) on first 5 datasets:",
      paste(sprintf("%.3e", self_jsd), collapse = ", "), "\n")

  # Index PIPs by (dataset, method) for fast lookup.
  pip_lookup <- pip_records %>%
    dplyr::group_by(.data$dataset_bundle_id, .data$model_method) %>%
    dplyr::summarise(pip = list(.data$pip[[1L]]), .groups = "drop")
  pip_index <- setNames(
    pip_lookup$pip,
    paste(pip_lookup$dataset_bundle_id, pip_lookup$model_method, sep = "||")
  )

  datasets_with_all_pips <- pip_records %>%
    dplyr::count(.data$dataset_bundle_id, .data$model_method) %>%
    dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
    dplyr::filter(.data$n_methods == length(required_methods)) %>%
    dplyr::pull(.data$dataset_bundle_id)

  pip_jsd_rows <- list()
  for (ds_id in datasets_with_all_pips) {
    for (k in seq_len(nrow(pip_pair_specs))) {
      spec <- pip_pair_specs[k, ]
      pa <- pip_index[[paste(ds_id, spec$method_a, sep = "||")]]
      pb <- pip_index[[paste(ds_id, spec$method_b, sep = "||")]]
      if (is.null(pa) || is.null(pb) || length(pa) != length(pb)) next
      pip_jsd_rows[[length(pip_jsd_rows) + 1L]] <- tibble::tibble(
        dataset_bundle_id = ds_id,
        chain = spec$chain,
        pair_kind = spec$pair_kind,
        pip_jsd = js_distance(pa, pb)
      )
    }
  }
  pip_drift_summary <- dplyr::bind_rows(pip_jsd_rows)
} else {
  warning(
    "Missing ", pip_records_rds,
    "; PIP JSD panels will be empty. Re-run ",
    "effect_matching_refit_export_5arm.Rmd to generate the PIP sidecar."
  )
  pip_drift_summary <- tidyr::expand_grid(
    dataset_bundle_id = datasets_complete,
    pip_pair_specs %>% dplyr::select(.data$chain, .data$pair_kind)
  ) %>%
    dplyr::mutate(pip_jsd = NA_real_)
}
readr::write_csv(pip_drift_summary, pip_drift_summary_csv)
cat("PIP drift summary rows:", nrow(pip_drift_summary), "\n")

# Sanity: rho between basin weighted_chord_drift and pip_jsd on matched (dataset,
# pair). Both metrics measure "how much did the model move", so a strongly
# positive rank correlation is expected.
basin_for_join <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "mu__vanilla_warm_from_mu",
    "vanilla__vanilla_warm_from_mu",
    "vanilla__pi", "pi__vanilla_warm_from_pi",
    "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    .data$dataset_bundle_id,
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::case_when(
      .data$pair_id %in% c("vanilla__mu", "vanilla__pi") ~
        "vanilla_to_treatment",
      .data$pair_id %in% c("mu__vanilla_warm_from_mu",
                           "pi__vanilla_warm_from_pi") ~
        "treatment_to_warm",
      TRUE ~ "vanilla_to_warm"
    ),
    weighted_chord_drift = .data$weighted_chord_drift
  )

basin_pip_join <- basin_for_join %>%
  dplyr::inner_join(
    pip_drift_summary,
    by = c("dataset_bundle_id", "chain", "pair_kind")
  )

rho_check <- basin_pip_join %>%
  dplyr::group_by(.data$chain, .data$pair_kind) %>%
  dplyr::summarise(
    n = dplyr::n(),
    rho = suppressWarnings(stats::cor(
      .data$weighted_chord_drift, .data$pip_jsd,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    .groups = "drop"
  )
print(rho_check)
```

# 13. Master Drift, PIP, and AUPRC Summary Figure

Three metrics x three comparisons. mu and pi are overlaid in each subplot.

```{r master-drift-figure, fig.width=13, fig.height=8.5}
basin_long <- basin_for_join %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$chain, .data$pair_kind,
    metric = "basin_drift",
    value = .data$weighted_chord_drift
  )

pip_long <- pip_drift_summary %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$chain, .data$pair_kind,
    metric = "pip_drift",
    value = .data$pip_jsd
  )

auprc_delta_long <- dplyr::bind_rows(
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "mu",
      pair_kind = "vanilla_to_treatment",
      metric = "auprc_delta",
      value = .data$auprc_delta_mu
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "mu",
      pair_kind = "treatment_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_round_trip_mu
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "mu",
      pair_kind = "vanilla_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_delta_warm_mu_vs_vanilla
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "pi",
      pair_kind = "vanilla_to_treatment",
      metric = "auprc_delta",
      value = .data$auprc_delta_pi
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "pi",
      pair_kind = "treatment_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_round_trip_pi
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "pi",
      pair_kind = "vanilla_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_delta_warm_pi_vs_vanilla
    )
)

combined <- dplyr::bind_rows(basin_long, pip_long, auprc_delta_long) %>%
  dplyr::mutate(
    metric = factor(
      .data$metric,
      levels = c("basin_drift", "pip_drift", "auprc_delta")
    ),
    pair_kind = factor(
      .data$pair_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

p_master <- ggplot(combined, aes(x = .data$value, fill = .data$chain)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 50, position = "identity", alpha = 0.55, color = NA
  ) +
  facet_grid(
    pair_kind ~ metric,
    scales = "free_x",
    labeller = labeller(
      metric = c(
        basin_drift = "Basin drift\n(weighted chord)",
        pip_drift   = "PIP drift\n(JSD)",
        auprc_delta = "AUPRC delta"
      ),
      pair_kind = c(
        vanilla_to_treatment = "vanilla -> annotated",
        treatment_to_warm    = "annotated -> vanilla_warm",
        vanilla_to_warm      = "vanilla -> vanilla_warm"
      )
    )
  ) +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  labs(
    x = NULL, y = "Density", fill = "Chain",
    title = "Basin, PIP, and AUPRC movement along the 5-arm refit chain",
    subtitle = "Basin drift uses Hungarian-matched fitted-y weighted chord distance; PIP drift is model-wide PIP-vector JSD."
  ) +
  theme(legend.position = "bottom")

ggsave(basin_pip_drift_png, p_master, width = 13, height = 8.5, dpi = 150)
print(p_master)
```

# 14. Matched and Unmatched Effect-Quality Density Shifts

For each chain and comparison, effects are classified as matched if their
Hungarian-assigned fitted-y correlation is at least `effect_match_corr_min`.
The KDEs below are weighted by per-effect explained-variance fraction.

```{r effect-quality-density-5arm, fig.width=13, fig.height=7.5}
effect_metric_rows <- records %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$model_method, .data$effect,
    x_value = .data[[x_metric]],
    y_value = .data[[y_metric]],
    .data$var_y_explained_frac
  )

effect_quality_pair_specs <- tibble::tribble(
  ~pair_id,                        ~chain, ~comparison_kind,        ~source_method,   ~destination_method,
  "vanilla__mu",                   "mu",   "vanilla_to_treatment",  baseline_method,  mu_method,
  "mu__vanilla_warm_from_mu",      "mu",   "treatment_to_warm",     mu_method,        warm_mu_method,
  "vanilla__vanilla_warm_from_mu", "mu",   "vanilla_to_warm",       baseline_method,  warm_mu_method,
  "vanilla__pi",                   "pi",   "vanilla_to_treatment",  baseline_method,  pi_method,
  "pi__vanilla_warm_from_pi",      "pi",   "treatment_to_warm",     pi_method,        warm_pi_method,
  "vanilla__vanilla_warm_from_pi", "pi",   "vanilla_to_warm",       baseline_method,  warm_pi_method
)

build_effect_quality_rows <- function(spec) {
  edges <- drift_assignments %>%
    dplyr::filter(.data$pair_id == spec$pair_id) %>%
    dplyr::mutate(
      match_status = dplyr::if_else(
        .data$corr >= effect_match_corr_min, "matched", "unmatched"
      )
    )

  source_rows <- edges %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = spec$chain,
      comparison_kind = spec$comparison_kind,
      match_status = .data$match_status,
      arm_role = "source",
      model_method = spec$source_method,
      effect = .data$effect_a,
      assigned_corr = .data$corr
    )

  destination_rows <- edges %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = spec$chain,
      comparison_kind = spec$comparison_kind,
      match_status = .data$match_status,
      arm_role = "destination",
      model_method = spec$destination_method,
      effect = .data$effect_b,
      assigned_corr = .data$corr
    )

  dplyr::bind_rows(source_rows, destination_rows) %>%
    dplyr::left_join(
      effect_metric_rows,
      by = c("dataset_bundle_id", "model_method", "effect")
    ) %>%
    dplyr::mutate(
      pve_weight = pmax(tidyr::replace_na(.data$var_y_explained_frac, 0), 0)
    )
}

effect_quality_pair_rows <- purrr::map_dfr(
  seq_len(nrow(effect_quality_pair_specs)),
  ~ build_effect_quality_rows(effect_quality_pair_specs[.x, ])
) %>%
  dplyr::mutate(
    chain = factor(.data$chain, levels = c("mu", "pi")),
    match_status = factor(.data$match_status,
                          levels = c("matched", "unmatched")),
    arm_role = factor(.data$arm_role,
                      levels = c("source", "destination")),
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    )
  )
readr::write_csv(effect_quality_pair_rows, effect_quality_pair_rows_csv)

x_limits_resolved <- resolve_limits(effect_quality_pair_rows$x_value, x_limits)
y_limits_resolved <- resolve_limits(effect_quality_pair_rows$y_value, y_limits)

weighted_kde2d_df <- function(df, xlim, ylim, n, h) {
  if (!requireNamespace("MASS", quietly = TRUE)) {
    stop("Package MASS is required for weighted KDE plots.")
  }
  group_values <- c("source", "destination")
  dens <- lapply(group_values, function(g) {
    sub <- df %>%
      dplyr::filter(.data$arm_role == g) %>%
      dplyr::filter(
        is.finite(.data$x_value),
        is.finite(.data$y_value),
        .data$x_value >= xlim[[1]],
        .data$x_value <= xlim[[2]],
        .data$y_value >= ylim[[1]],
        .data$y_value <= ylim[[2]]
      )

    if (nrow(sub) < 3L) {
      return(NULL)
    }

    w <- pmax(tidyr::replace_na(sub$pve_weight, 0), 0)
    if (sum(w) <= 0) {
      w <- rep(1 / nrow(sub), nrow(sub))
    } else {
      w <- w / sum(w)
    }

    x_grid <- seq(xlim[[1]], xlim[[2]], length.out = n)
    y_grid <- seq(ylim[[1]], ylim[[2]], length.out = n)
    kx <- outer(x_grid, sub$x_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[1]]) / h[[1]]
    })
    ky <- outer(y_grid, sub$y_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[2]]) / h[[2]]
    })
    ky <- sweep(ky, 2L, w, `*`)
    z <- kx %*% t(ky)

    list(group = g, x = x_grid, y = y_grid, z = z, n_rows = nrow(sub),
         pve_mass = sum(sub$pve_weight, na.rm = TRUE))
  })
  names(dens) <- group_values

  if (is.null(dens$source) || is.null(dens$destination)) {
    return(tibble::tibble())
  }

  tidyr::expand_grid(x = dens$source$x, y = dens$source$y) %>%
    dplyr::mutate(
      density_source = as.vector(t(dens$source$z)),
      density_destination = as.vector(t(dens$destination$z)),
      density_diff = .data$density_destination - .data$density_source,
      n_source = dens$source$n_rows,
      n_destination = dens$destination$n_rows,
      pve_source = dens$source$pve_mass,
      pve_destination = dens$destination$pve_mass
    )
}

density_shift_rows <- list()
for (ch in levels(effect_quality_pair_rows$chain)) {
  for (cmp in levels(effect_quality_pair_rows$comparison_kind)) {
    for (status in levels(effect_quality_pair_rows$match_status)) {
      sub <- effect_quality_pair_rows %>%
        dplyr::filter(
          .data$chain == ch,
          .data$comparison_kind == cmp,
          .data$match_status == status
        )
      dens <- weighted_kde2d_df(
        sub,
        xlim = x_limits_resolved,
        ylim = y_limits_resolved,
        n = density_grid_n,
        h = density_bandwidth
      )
      if (!nrow(dens)) {
        next
      }
      density_shift_rows[[length(density_shift_rows) + 1L]] <- dens %>%
        dplyr::mutate(
          chain = ch,
          comparison_kind = cmp,
          match_status = status
        )
    }
  }
}

density_shift_df <- dplyr::bind_rows(density_shift_rows) %>%
  dplyr::mutate(
    chain = factor(.data$chain, levels = c("mu", "pi")),
    match_status = factor(.data$match_status,
                          levels = c("matched", "unmatched")),
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    )
  )

plot_density_shift_chain <- function(density_df, chain_value, out_path) {
  df <- density_df %>% dplyr::filter(.data$chain == chain_value)
  abs_diff <- abs(df$density_diff)
  abs_diff <- abs_diff[is.finite(abs_diff)]
  fill_lim <- stats::quantile(
    abs_diff, probs = density_fill_quantile, na.rm = TRUE, names = FALSE
  )
  if (!is.finite(fill_lim) || fill_lim <= 0) {
    fill_lim <- max(abs_diff, na.rm = TRUE)
  }
  if (!is.finite(fill_lim) || fill_lim <= 0) {
    fill_lim <- 1
  }

  p <- ggplot(df, aes(x = .data$x, y = .data$y,
                      fill = .data$density_diff)) +
    geom_raster(interpolate = TRUE) +
    geom_contour(aes(z = .data$density_diff),
                 color = "grey20", alpha = 0.45, linewidth = 0.18) +
    facet_grid(
      match_status ~ comparison_kind,
      labeller = labeller(
        match_status = c(
          matched = "matched effects",
          unmatched = "unmatched effects"
        ),
        comparison_kind = c(
          vanilla_to_treatment = "vanilla -> annotated",
          treatment_to_warm = "annotated -> vanilla_warm",
          vanilla_to_warm = "vanilla -> vanilla_warm"
        )
      )
    ) +
    scale_fill_gradient2(
      low = "#2b6cb0", mid = "white", high = "#c53030",
      midpoint = 0, limits = c(-fill_lim, fill_lim),
      oob = scales::squish
    ) +
    keff_x_scale(xlim = x_limits_resolved) +
    coord_cartesian(xlim = x_limits_resolved,
                    ylim = y_limits_resolved,
                    expand = FALSE) +
    labs(
      x = x_label, y = y_label,
      fill = "destination - source\nPVE-weighted density",
      title = sprintf(
        "Effect-quality density shifts along the %s refit chain",
        chain_value
      ),
      subtitle = sprintf(
        "Red: higher destination density; blue: higher source density. Matched if fitted-y corr >= %.2f.",
        effect_match_corr_min
      )
    ) +
    theme(legend.position = "right")

  ggsave(out_path, p, width = 13, height = 7.5, dpi = 150)
  p
}

p_effect_density_mu <- plot_density_shift_chain(
  density_shift_df, "mu", effect_quality_density_mu_png
)
p_effect_density_pi <- plot_density_shift_chain(
  density_shift_df, "pi", effect_quality_density_pi_png
)
print(p_effect_density_mu)
print(p_effect_density_pi)
```

# 15. Pooled Precision-Recall Curves

Pooled precision-recall curves across all 600 datasets for the 5 arms
(vanilla, mu, pi, vanilla_warm_from_mu, vanilla_warm_from_pi). Inputs are
the model-wide PIP vectors in `pip_records_5arm.rds` and the per-dataset
causal indices, which are regenerated from `dataset_bundles.csv` via
`test_susine:::generate_data_for_bundle()` (no fits are re-run; the
hot loop is the deterministic phenotype simulation).

The shared 5-arm config used here is identical to the one in
`effect_matching_refit_export_5arm.Rmd` and lives in `_5arm_cfg.R`.

```{r pr-curves-setup, message=FALSE}
have_pip_records <- file.exists(pip_records_rds)
if (!have_pip_records) {
  warning(
    "Missing ", pip_records_rds,
    "; PR curves will be skipped. Re-run effect_matching_refit_export_5arm.Rmd ",
    "(or its PIP-export tail) to regenerate the sidecar."
  )
}

if (have_pip_records && !requireNamespace("PRROC", quietly = TRUE)) {
  warning("PRROC not installed; PR curves chunk skipped. install.packages('PRROC').")
  have_pip_records <- FALSE
}
```

```{r pr-curves-causal-idx, eval=have_pip_records}
# `here()` resolves to the test_susine package root (set by here::i_am()
# in section 1). Load the harness via devtools so we can call
# test_susine:::generate_data_for_bundle().
suppressMessages(devtools::load_all(here()))

# Optionally load the sister susine package the way refit_export_5arm.Rmd
# does, so the cfg builder's downstream call sites stay parallel.
susine_sister_path <- here("..", "susine")
if (file.exists(file.path(susine_sister_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(
    susine_sister_path, winslash = "/", mustWork = TRUE
  ))
  suppressMessages(devtools::load_all(susine_sister_path))
}

source(here("vignettes", "one_off_validations", "_5arm_cfg.R"))

if (file.exists(causal_idx_rds)) {
  causal_idx_lookup <- readRDS(causal_idx_rds)
  cat("Loaded cached causal_idx for ", length(causal_idx_lookup),
      " datasets from ", causal_idx_rds, "\n", sep = "")
} else {
  cfg_bundle <- build_5arm_cfg(
    here_root = here(),
    output_root = output_root
  )
  cfg_for_pr <- cfg_bundle$cfg

  pip_records_for_idx <- readRDS(pip_records_rds) %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

  datasets_with_pips <- sort(unique(pip_records_for_idx$dataset_bundle_id))
  cat("Regenerating causal_idx for ", length(datasets_with_pips),
      " datasets (no fits, just data simulation).\n", sep = "")

  bundle_rows_all <- cfg_for_pr$tables$dataset_bundles %>%
    tibble::as_tibble() %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

  causal_idx_lookup <- list()
  started <- proc.time()[["elapsed"]]
  for (i in seq_along(datasets_with_pips)) {
    ds_id <- datasets_with_pips[[i]]
    bundle_row <- bundle_rows_all %>%
      dplyr::filter(.data$dataset_bundle_id == !!ds_id)
    if (nrow(bundle_row) != 1L) {
      stop("Expected one bundle row for ", ds_id, "; got ", nrow(bundle_row))
    }
    data_bundle <- test_susine:::generate_data_for_bundle(
      bundle_row, cfg_for_pr
    )
    causal_idx_lookup[[ds_id]] <- as.integer(data_bundle$causal_idx)
    if (i %% 50L == 0L || i == length(datasets_with_pips)) {
      elapsed <- proc.time()[["elapsed"]] - started
      eta <- elapsed / i * (length(datasets_with_pips) - i)
      cat(sprintf(
        "Regenerated %d/%d datasets | elapsed %s | ETA %s\n",
        i, length(datasets_with_pips),
        format_duration(elapsed), format_duration(eta)
      ))
    }
  }
  saveRDS(causal_idx_lookup, causal_idx_rds)
  cat("Saved causal_idx lookup to ", causal_idx_rds, "\n", sep = "")
}
```

```{r pr-curves-build, eval=have_pip_records, fig.width=9, fig.height=6}
pip_records_for_pr <- readRDS(pip_records_rds) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

# Pool PIPs and binary causal labels across all (dataset, variant)
# observations, separately for each of the 5 methods.
pooled_rows <- list()
for (i in seq_len(nrow(pip_records_for_pr))) {
  ds_id <- pip_records_for_pr$dataset_bundle_id[[i]]
  method <- pip_records_for_pr$model_method[[i]]
  pip_vec <- pip_records_for_pr$pip[[i]]
  causal <- causal_idx_lookup[[ds_id]]
  if (is.null(causal)) {
    next
  }
  is_causal <- integer(length(pip_vec))
  causal_in_range <- causal[causal >= 1L & causal <= length(pip_vec)]
  is_causal[causal_in_range] <- 1L
  pooled_rows[[length(pooled_rows) + 1L]] <- tibble::tibble(
    dataset_bundle_id = ds_id,
    model_method = method,
    pip = as.numeric(pip_vec),
    is_causal = is_causal
  )
}
pooled_df <- dplyr::bind_rows(pooled_rows)

cat("Pooled rows:", nrow(pooled_df), "\n")
cat("Per-method counts:\n")
print(pooled_df %>% dplyr::count(.data$model_method, .data$is_causal))

prevalence <- mean(pooled_df$is_causal)
cat(sprintf("Overall causal prevalence: %.4f\n", prevalence))

# One PR curve per method.
methods_in_order <- c(
  baseline_method, mu_method, warm_mu_method, pi_method, warm_pi_method
)
methods_in_order <- intersect(methods_in_order,
                              unique(pooled_df$model_method))

pr_curve_rows <- list()
pr_auc_rows <- list()
for (m in methods_in_order) {
  sub <- pooled_df %>% dplyr::filter(.data$model_method == m)
  if (!nrow(sub) || sum(sub$is_causal) == 0L) next
  pr <- PRROC::pr.curve(
    scores.class0 = sub$pip,
    weights.class0 = sub$is_causal,
    curve = TRUE,
    sorted = FALSE,
    rand.compute = FALSE
  )
  curve_mat <- pr$curve  # columns: recall, precision, threshold
  pr_curve_rows[[length(pr_curve_rows) + 1L]] <- tibble::tibble(
    model_method = m,
    recall = curve_mat[, 1L],
    precision = curve_mat[, 2L],
    threshold = curve_mat[, 3L]
  )
  pr_auc_rows[[length(pr_auc_rows) + 1L]] <- tibble::tibble(
    model_method = m,
    auprc_integral = as.numeric(pr$auc.integral),
    auprc_dg = as.numeric(pr$auc.davis.goadrich)
  )
}
pr_curve_df <- dplyr::bind_rows(pr_curve_rows) %>%
  dplyr::mutate(model_method = factor(.data$model_method,
                                      levels = methods_in_order))
pr_auc_df <- dplyr::bind_rows(pr_auc_rows) %>%
  dplyr::mutate(model_method = factor(.data$model_method,
                                      levels = methods_in_order))

readr::write_csv(pr_curve_df, pr_curves_pooled_csv)
cat("Wrote PR curve rows:", nrow(pr_curve_df), "->", pr_curves_pooled_csv, "\n")
print(pr_auc_df)

method_legend_labels <- vapply(
  levels(pr_curve_df$model_method),
  function(m) {
    auprc_val <- pr_auc_df$auprc_integral[
      match(m, as.character(pr_auc_df$model_method))
    ]
    sprintf("%s (AUPRC = %.3f)",
            method_labels[[m]] %||% m,
            auprc_val)
  },
  character(1)
)

p_pr <- ggplot(pr_curve_df,
               aes(x = .data$recall, y = .data$precision,
                   color = .data$model_method)) +
  geom_hline(yintercept = prevalence, color = "grey60",
             linetype = "dashed", linewidth = 0.4) +
  geom_step(direction = "vh", linewidth = 0.7) +
  scale_color_manual(
    values = method_color[levels(pr_curve_df$model_method)],
    labels = method_legend_labels
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    x = "Recall", y = "Precision", color = "Method (pooled AUPRC)",
    title = "Pooled precision-recall across all 600 datasets, by 5-arm method",
    subtitle = sprintf(
      "Pooled across %d datasets x %d variants. Dashed line = causal prevalence (%.3f).",
      length(unique(pooled_df$dataset_bundle_id)),
      max(table(pooled_df$dataset_bundle_id, pooled_df$model_method)),
      prevalence
    )
  ) +
  theme(legend.position = "right")

ggsave(pr_curves_pooled_png, p_pr, width = 9, height = 6, dpi = 150)
print(p_pr)
```

# 16. Outputs

```{r output-index}
outputs <- tibble::tibble(
  artifact = c(
    "per-effect drift assignments",
    "per-(dataset, pair) drift summary",
    "per-dataset AUPRC trajectory",
    "drift profile histograms",
    "chain AUPRC trajectory",
    "headline 4-quadrant scatter",
    "unpaired-count vs AUPRC delta scatter",
    "mu vs pi drift comparison (cold + warm panels)",
    "per-dataset PIP JSD summary",
    "master basin + PIP + AUPRC drift figure",
    "effect-quality matched/unmatched density (mu)",
    "effect-quality matched/unmatched density (pi)",
    "effect-quality pair rows",
    "pooled PR curves (5 arms)",
    "pooled PR curve rows (csv)",
    "regenerated causal_idx cache"
  ),
  path = c(
    drift_assignments_csv,
    drift_summary_csv,
    auprc_per_dataset_csv,
    drift_profile_hist_png,
    chain_trajectory_png,
    headline_scatter_png,
    unpaired_vs_delta_png,
    mu_vs_pi_drift_png,
    pip_drift_summary_csv,
    basin_pip_drift_png,
    effect_quality_density_mu_png,
    effect_quality_density_pi_png,
    effect_quality_pair_rows_csv,
    pr_curves_pooled_png,
    pr_curves_pooled_csv,
    causal_idx_rds
  )
)
kable(outputs)
```
