---
title: "Effect Matching Diagnostic Plots (3-arm)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

3-arm extension of `effect_matching_diagnostic_plots.Rmd`. Reads:

- `effect_refit_records.rds` from the 3-arm refit study
- `real_match_candidates_vanilla_mu_same_X.csv` (vanilla<->mu pairing)
- `real_match_candidates_vanilla_pi_same_X.csv` (vanilla<->pi pairing)

and produces three categories of plots:

- **Global 3-arm overlays** (scatter, univariate histograms): all three
  arms shown together to characterize the marginal effect distributions.
- **Per-pairing plots** (differential density, flow, delta density): the
  same vanilla<->treatment plots as the 2-arm notebook, generated twice
  (once per treatment arm).
- **Cross-arm AUPRC attribution and confound diagnostics**: each plot
  faceted by treatment arm so mu and pi can be compared side-by-side.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_diagnostic_plots_3arm.Rmd")
library(here)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

# 1. User Parameters

```{r user-parameters}
output_root <- here("output")

study_name <- "baseline_sims_3arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
records_rds <- file.path(study_dir, "effect_refit_records.rds")

baseline_method <- "susie_vanilla"
treatment_methods <- c("susine_functional_mu", "susie_functional_pi")
treatment_short <- c(
  susine_functional_mu = "mu",
  susie_functional_pi = "pi"
)
treatment_color <- c(
  susie_vanilla = "black",
  susine_functional_mu = "#2b8cbe",
  susie_functional_pi = "#74c476"
)

real_candidates_csv <- c(
  mu = file.path(study_dir, "real_match_candidates_vanilla_mu_same_X.csv"),
  pi = file.path(study_dir, "real_match_candidates_vanilla_pi_same_X.csv")
)

# Operating threshold (locked from the 2-arm same-X calibration; reused here).
manual_corr_min <- 0.99
manual_margin_min <- 0
manual_var_min <- 0

plot_dir <- file.path(study_dir, "diagnostic_plots")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Joint diagnostic axes.
x_metric <- "effect_pip_entropy_core95"
y_metric <- "accuracy_ratio"
x_label <- "Core-95 effective k (= exp(entropy))"
y_label <- "Accuracy ratio"
x_keff_breaks <- c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000)
delta_keff_ratio_breaks <- c(1/16, 1/8, 1/4, 1/2, 1, 2, 4, 8, 16)
delta_keff_ratio_labels <- c("1/16", "1/8", "1/4", "1/2", "1", "2", "4", "8", "16")

x_limits <- NULL
y_limits <- c(0, 1)

density_grid_n <- 160L
density_bandwidth <- c(0.45, 0.10)
density_trim_quantile <- c(0.005, 0.995)
density_fill_limit <- NULL
density_fill_quantile <- 0.95

hist_bins <- 60L

delta_density_contour_breaks <- c(
  0, 0.005, 0.015, 0.04, 0.08, 0.15, 0.25, 0.4, 0.6, 0.8, 1.0
)

max_arrows_to_plot <- Inf
arrow_alpha <- 0.40
arrow_linewidth <- 0.25
arrow_head_length_inches <- 0.055
label_top_discrepancies <- 25L

# Output filenames. Per-pairing plots get a `_vs_<short>` suffix.
scatter_all_png <- file.path(plot_dir, "effect_diagnostic_scatter_all_effects_3arm.png")
hist_diffuseness_png <- file.path(plot_dir, "effect_diagnostic_univariate_hist_diffuseness_3arm.png")
hist_accuracy_png <- file.path(plot_dir, "effect_diagnostic_univariate_hist_accuracy_3arm.png")
auprc_delta_hist_png <- file.path(plot_dir, "effect_diagnostic_auprc_delta_hist_3arm.png")
match_count_hist_png <- file.path(plot_dir, "effect_diagnostic_match_count_hist_3arm.png")
auprc_match_scatter_png <- file.path(plot_dir, "effect_diagnostic_auprc_vs_match_scatter_3arm.png")
baseline_vs_treatment_png <- file.path(plot_dir, "effect_diagnostic_baseline_vs_treatment_scatter_3arm.png")
scatter_stratified_png <- file.path(plot_dir, "effect_diagnostic_auprc_vs_match_stratified_3arm.png")
confound_check_png <- file.path(plot_dir, "effect_diagnostic_vanilla_auprc_vs_match_3arm.png")
per_dataset_summary_csv <- file.path(plot_dir, "per_dataset_auprc_match_summary_3arm.csv")

per_pair_filename <- function(stem, short_label, ext = "png") {
  file.path(plot_dir, sprintf("%s_vs_%s.%s", stem, short_label, ext))
}

cat("Study dir:", study_dir, "\n")
cat("Plot dir:", plot_dir, "\n")
cat("Treatment arms:", paste(treatment_methods, collapse = ", "), "\n")
```

# 2. Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

resolve_limits <- function(values, user_limits, trim = density_trim_quantile) {
  if (!is.null(user_limits)) {
    return(user_limits)
  }
  values <- values[is.finite(values)]
  if (!length(values)) {
    stop("Cannot infer limits from an empty/non-finite vector.")
  }
  stats::quantile(values, probs = trim, na.rm = TRUE, names = FALSE)
}

require_metric <- function(df, metric) {
  if (!metric %in% names(df)) {
    stop("Metric column not found: ", metric)
  }
}

keff_x_scale <- function(breaks = x_keff_breaks, xlim = NULL) {
  finite_breaks <- breaks[breaks > 0]
  scale_x_continuous(
    breaks = log(finite_breaks),
    labels = formatC(finite_breaks, format = "fg", big.mark = ","),
    limits = xlim
  )
}

keff_ratio_x_scale <- function(breaks = delta_keff_ratio_breaks,
                               labels = delta_keff_ratio_labels,
                               xlim = NULL) {
  scale_x_continuous(
    breaks = log(breaks),
    labels = labels,
    limits = xlim
  )
}

kde2d_df <- function(df, x_col, y_col, group_col, group_values, xlim, ylim, n, h) {
  if (!requireNamespace("MASS", quietly = TRUE)) {
    stop("Package MASS is required for kde2d density plots.")
  }
  if (length(group_values) != 2L) {
    stop("kde2d_df expects exactly two group values; got ", length(group_values))
  }

  dens <- lapply(group_values, function(g) {
    sub <- df %>%
      dplyr::filter(.data[[group_col]] == g) %>%
      dplyr::filter(
        is.finite(.data[[x_col]]),
        is.finite(.data[[y_col]]),
        .data[[x_col]] >= xlim[[1]],
        .data[[x_col]] <= xlim[[2]],
        .data[[y_col]] >= ylim[[1]],
        .data[[y_col]] <= ylim[[2]]
      )

    if (nrow(sub) < 3L) {
      stop("Not enough rows for density group ", g, ": ", nrow(sub))
    }

    kde <- MASS::kde2d(
      x = sub[[x_col]],
      y = sub[[y_col]],
      n = n,
      lims = c(xlim, ylim),
      h = h
    )
    list(group = g, kde = kde, n_rows = nrow(sub))
  })

  base <- dens[[1]]$kde
  z1 <- dens[[1]]$kde$z
  z2 <- dens[[2]]$kde$z

  tidyr::expand_grid(x = base$x, y = base$y) %>%
    dplyr::mutate(
      density_baseline = as.vector(t(z1)),
      density_treatment = as.vector(t(z2)),
      density_diff = .data$density_treatment - .data$density_baseline,
      group_baseline = dens[[1]]$group,
      group_treatment = dens[[2]]$group,
      n_baseline = dens[[1]]$n_rows,
      n_treatment = dens[[2]]$n_rows
    )
}

plot_density_diff <- function(density_df, title, subtitle, out_path,
                              xlim_resolved, ylim_resolved) {
  abs_diff <- abs(density_df$density_diff)
  abs_diff <- abs_diff[is.finite(abs_diff)]

  fill_lim <- density_fill_limit
  if (is.null(fill_lim)) {
    if (length(abs_diff)) {
      fill_lim <- as.numeric(stats::quantile(
        abs_diff, probs = density_fill_quantile, na.rm = TRUE
      ))
      max_abs <- max(abs_diff, na.rm = TRUE)
    } else {
      fill_lim <- 1
      max_abs <- 1
    }
    if (!is.finite(fill_lim) || fill_lim <= 0) {
      fill_lim <- max(max_abs, .Machine$double.eps)
    }
  }

  fill_legend_title <- sprintf(
    "treatment - baseline\ndensity\n(clipped at\n+/-%s of max)",
    if (is.null(density_fill_limit)) {
      paste0(round(100 * density_fill_quantile), "%-tile")
    } else {
      "manual"
    }
  )

  p <- ggplot(density_df, aes(x = .data$x, y = .data$y, fill = .data$density_diff)) +
    geom_raster(interpolate = TRUE) +
    geom_contour(aes(z = .data$density_diff), color = "grey20", alpha = 0.45, linewidth = 0.2) +
    scale_fill_gradient2(
      low = "#2b6cb0", mid = "white", high = "#c53030",
      midpoint = 0, limits = c(-fill_lim, fill_lim),
      oob = scales::squish
    ) +
    keff_x_scale(xlim = xlim_resolved) +
    coord_cartesian(xlim = xlim_resolved, ylim = ylim_resolved, expand = FALSE) +
    labs(
      x = x_label, y = y_label,
      fill = fill_legend_title,
      title = title, subtitle = subtitle
    ) +
    theme(legend.position = "right")

  ggsave(out_path, p, width = 9, height = 6, dpi = 150)
  p
}

plot_scatter_3arm <- function(df, title, out_path,
                              xlim_resolved, ylim_resolved,
                              palette = treatment_color) {
  p <- ggplot(
    df,
    aes(x = .data[[x_metric]], y = .data[[y_metric]], color = .data$model_method)
  ) +
    geom_point(alpha = 0.32, size = 0.85) +
    scale_color_manual(
      values = palette,
      breaks = names(palette)
    ) +
    keff_x_scale(xlim = xlim_resolved) +
    coord_cartesian(xlim = xlim_resolved, ylim = ylim_resolved) +
    labs(
      x = x_label, y = y_label,
      color = "Model",
      title = title
    ) +
    theme(legend.position = "bottom")

  ggsave(out_path, p, width = 10, height = 7, dpi = 150)
  p
}

plot_univariate_hist_3arm <- function(df, metric, xlab_text, out_path, title,
                                      bins = hist_bins, use_keff_x = FALSE,
                                      xlim = NULL, palette = treatment_color) {
  df_clean <- df %>%
    dplyr::filter(is.finite(.data[[metric]]))
  if (!is.null(xlim)) {
    df_clean <- df_clean %>%
      dplyr::filter(.data[[metric]] >= xlim[[1]], .data[[metric]] <= xlim[[2]])
  }

  panel_counts <- df_clean %>%
    dplyr::count(.data$model_method, name = "n_effects")
  cat(title, "panel counts:\n")
  print(panel_counts)

  p <- ggplot(df_clean, aes(x = .data[[metric]], fill = .data$model_method)) +
    geom_histogram(
      aes(y = after_stat(density)),
      position = "identity", bins = bins, alpha = 0.45, color = NA
    ) +
    scale_fill_manual(
      values = palette,
      breaks = names(palette)
    ) +
    labs(
      x = xlab_text, y = "Density",
      fill = "Model", title = title
    ) +
    theme(legend.position = "bottom")

  if (use_keff_x) {
    p <- p + keff_x_scale(xlim = xlim)
  } else if (!is.null(xlim)) {
    p <- p + coord_cartesian(xlim = xlim)
  }

  ggsave(out_path, p, width = 10, height = 5, dpi = 150)
  p
}

# Per-treatment match label helpers.
treatment_label <- function(treatment) {
  switch(
    treatment,
    susine_functional_mu = "susine functional-mu",
    susie_functional_pi = "susie functional-pi",
    treatment
  )
}
```

# 3. Load And Classify Effects

```{r load-classify}
stopifnot(file.exists(records_rds))
for (treatment in treatment_methods) {
  short <- treatment_short[[treatment]]
  csv_path <- real_candidates_csv[[short]]
  if (!file.exists(csv_path)) {
    stop(
      "Missing real candidates CSV for ", treatment, " at ", csv_path,
      ". Run effect_matching_threshold_calibration_3arm.Rmd first."
    )
  }
}

records <- readRDS(records_rds) %>%
  dplyr::mutate(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    matrix_id = as.character(.data$matrix_id),
    effect = as.integer(.data$effect)
  )

require_metric(records, x_metric)
require_metric(records, y_metric)

methods_present <- sort(unique(records$model_method))
cat("Methods in records:", paste(methods_present, collapse = ", "), "\n")
required_methods <- c(baseline_method, treatment_methods)
missing_methods <- setdiff(required_methods, methods_present)
if (length(missing_methods)) {
  stop("Missing methods in records: ", paste(missing_methods, collapse = ", "))
}

# Load both candidate CSVs and apply the manual threshold to each.
load_accepted <- function(short) {
  readr::read_csv(real_candidates_csv[[short]], show_col_types = FALSE) %>%
    dplyr::mutate(
      vanilla_dataset_bundle_id = as.character(.data$vanilla_dataset_bundle_id),
      mu_dataset_bundle_id = as.character(.data$mu_dataset_bundle_id),
      vanilla_effect = as.integer(.data$vanilla_effect),
      mu_effect = as.integer(.data$mu_effect)
    ) %>%
    dplyr::filter(
      .data$corr >= manual_corr_min,
      .data$margin >= manual_margin_min,
      .data$pair_var_explained_frac >= manual_var_min
    ) %>%
    dplyr::arrange(.data$vanilla_dataset_bundle_id, dplyr::desc(.data$corr))
}

accepted_by_arm <- list(
  mu = load_accepted("mu"),
  pi = load_accepted("pi")
)

cat("Accepted matched pairs per arm:\n")
for (short in names(accepted_by_arm)) {
  cat(sprintf("  %s: %d\n", short, nrow(accepted_by_arm[[short]])))
}

# Build per-arm match-status keys for the records table. For each row in
# records, we tag whether it participates in an accepted match for its
# corresponding pairing. Vanilla rows can be matched in mu and/or pi pairings;
# treatment rows only in their own pairing.
build_match_keys <- function(accepted_pair, treatment_method, treatment_short) {
  vanilla_keys <- accepted_pair %>%
    dplyr::transmute(
      dataset_bundle_id = .data$vanilla_dataset_bundle_id,
      effect = .data$vanilla_effect,
      model_method = baseline_method,
      match_id = paste0(
        treatment_short, "_",
        .data$vanilla_dataset_bundle_id, "_",
        .data$vanilla_effect, "_", .data$mu_effect
      )
    )
  treatment_keys <- accepted_pair %>%
    dplyr::transmute(
      dataset_bundle_id = .data$mu_dataset_bundle_id,
      effect = .data$mu_effect,
      model_method = treatment_method,
      match_id = paste0(
        treatment_short, "_",
        .data$vanilla_dataset_bundle_id, "_",
        .data$vanilla_effect, "_", .data$mu_effect
      )
    )
  dplyr::bind_rows(vanilla_keys, treatment_keys)
}

match_keys_by_arm <- lapply(treatment_methods, function(treatment) {
  short <- treatment_short[[treatment]]
  build_match_keys(accepted_by_arm[[short]], treatment, short)
})
names(match_keys_by_arm) <- treatment_short[treatment_methods]

# For each treatment, attach a `matched_<short>` boolean to records: TRUE if
# this (dataset, model_method, effect) participates in any accepted pair
# for that pairing.
records_labeled <- records
for (short in names(match_keys_by_arm)) {
  keys <- match_keys_by_arm[[short]] %>%
    dplyr::distinct(.data$dataset_bundle_id, .data$model_method, .data$effect) %>%
    dplyr::mutate("matched_{short}" := TRUE)
  records_labeled <- records_labeled %>%
    dplyr::left_join(
      keys,
      by = c("dataset_bundle_id", "model_method", "effect")
    ) %>%
    dplyr::mutate(
      "matched_{short}" := tidyr::replace_na(.data[[paste0("matched_", short)]], FALSE)
    )
}

x_limits_resolved <- resolve_limits(records_labeled[[x_metric]], x_limits)
y_limits_resolved <- resolve_limits(records_labeled[[y_metric]], y_limits)

cat("Records labeled rows:", nrow(records_labeled), "\n")
cat("Per-arm matched effect counts (records_labeled):\n")
for (short in names(match_keys_by_arm)) {
  col <- paste0("matched_", short)
  print(
    records_labeled %>%
      dplyr::count(.data$model_method, .data[[col]], name = "n_effects")
  )
}
cat("Resolved x limits:", paste(x_limits_resolved, collapse = ", "), "\n")
cat("Resolved y limits:", paste(y_limits_resolved, collapse = ", "), "\n")
```

# 4. Global 3-Arm Scatter

```{r global-scatter, fig.width=10, fig.height=7}
p_scatter_all <- plot_scatter_3arm(
  records_labeled,
  title = "Effect diagnostic scatter: all 3 arms (all effects)",
  out_path = scatter_all_png,
  xlim_resolved = x_limits_resolved,
  ylim_resolved = y_limits_resolved
)
print(p_scatter_all)
cat("Saved:", scatter_all_png, "\n")
```

# 5. Univariate Histograms (3-arm overlay)

```{r univariate-histograms, fig.width=10, fig.height=5}
p_hist_diff <- plot_univariate_hist_3arm(
  records_labeled,
  metric = x_metric,
  xlab_text = x_label,
  out_path = hist_diffuseness_png,
  title = "Effect diffuseness: Core-95 effective k by model arm",
  use_keff_x = TRUE,
  xlim = x_limits_resolved
)
print(p_hist_diff)

p_hist_acc <- plot_univariate_hist_3arm(
  records_labeled,
  metric = y_metric,
  xlab_text = y_label,
  out_path = hist_accuracy_png,
  title = "Effect accuracy ratio by model arm",
  use_keff_x = FALSE,
  xlim = y_limits_resolved
)
print(p_hist_acc)

cat("Saved:", hist_diffuseness_png, "\n")
cat("Saved:", hist_accuracy_png, "\n")
```

# 6. Per-Pairing Plots (Differential Density, Flow, Delta Density)

```{r per-pair-helper}
make_pair_plots <- function(treatment_method, accepted_pair) {
  short <- treatment_short[[treatment_method]]
  treatment_pretty <- treatment_label(treatment_method)
  pair_methods <- c(baseline_method, treatment_method)
  matched_col <- paste0("matched_", short)

  pair_records <- records_labeled %>%
    dplyr::filter(.data$model_method %in% pair_methods) %>%
    dplyr::mutate(
      match_status = dplyr::if_else(.data[[matched_col]], "matched", "unmatched")
    )

  # File paths
  density_all_png <- per_pair_filename("effect_diagnostic_density_diff_all_effects", short)
  density_matched_png <- per_pair_filename("effect_diagnostic_density_diff_matched_effects", short)
  density_unmatched_png <- per_pair_filename("effect_diagnostic_density_diff_unmatched_effects", short)
  flow_png <- per_pair_filename("effect_diagnostic_matched_flow", short)
  delta_density_png <- per_pair_filename("effect_diagnostic_matched_delta_density", short)
  matched_pairs_csv <- per_pair_filename("accepted_matched_pairs_for_plots", short, "csv")
  matched_delta_csv <- per_pair_filename("accepted_matched_pair_deltas_for_plots", short, "csv")

  # Differential 2D densities
  density_all_df <- kde2d_df(
    pair_records, x_col = x_metric, y_col = y_metric,
    group_col = "model_method", group_values = pair_methods,
    xlim = x_limits_resolved, ylim = y_limits_resolved,
    n = density_grid_n, h = density_bandwidth
  )
  p_density_all <- plot_density_diff(
    density_all_df,
    title = sprintf("Differential joint density: %s vs vanilla, all effects", treatment_pretty),
    subtitle = sprintf("Red: higher %s density; blue: higher vanilla density", treatment_pretty),
    out_path = density_all_png,
    xlim_resolved = x_limits_resolved,
    ylim_resolved = y_limits_resolved
  )
  print(p_density_all)

  for (status in c("matched", "unmatched")) {
    sub <- pair_records %>% dplyr::filter(.data$match_status == status)
    if (dplyr::n_distinct(sub$model_method) < 2L || nrow(sub) < 6L) {
      cat("Skipping density for status", status, "(", short, ") - not enough rows.\n")
      next
    }
    dens <- kde2d_df(
      sub, x_col = x_metric, y_col = y_metric,
      group_col = "model_method", group_values = pair_methods,
      xlim = x_limits_resolved, ylim = y_limits_resolved,
      n = density_grid_n, h = density_bandwidth
    )
    out_path <- if (status == "matched") density_matched_png else density_unmatched_png
    p <- plot_density_diff(
      dens,
      title = sprintf("Differential joint density: %s vs vanilla, %s effects",
                      treatment_pretty, status),
      subtitle = sprintf("Red: higher %s density; blue: higher vanilla density",
                         treatment_pretty),
      out_path = out_path,
      xlim_resolved = x_limits_resolved,
      ylim_resolved = y_limits_resolved
    )
    print(p)
  }

  # Flow plot --- requires matched pairs
  matched_vanilla <- pair_records %>%
    dplyr::filter(.data$model_method == baseline_method, .data$match_status == "matched") %>%
    dplyr::transmute(
      dataset_bundle_id_left = .data$dataset_bundle_id,
      effect_left = .data$effect,
      vanilla_x = .data[[x_metric]],
      vanilla_y = .data[[y_metric]]
    )
  matched_treatment <- pair_records %>%
    dplyr::filter(.data$model_method == treatment_method, .data$match_status == "matched") %>%
    dplyr::transmute(
      dataset_bundle_id_right = .data$dataset_bundle_id,
      effect_right = .data$effect,
      mu_x = .data[[x_metric]],
      mu_y = .data[[y_metric]]
    )

  flow_df <- accepted_pair %>%
    dplyr::transmute(
      vanilla_dataset_bundle_id = .data$vanilla_dataset_bundle_id,
      mu_dataset_bundle_id = .data$mu_dataset_bundle_id,
      vanilla_effect = .data$vanilla_effect,
      mu_effect = .data$mu_effect,
      corr = .data$corr,
      margin = .data$margin,
      pair_var_explained_frac = .data$pair_var_explained_frac
    ) %>%
    dplyr::left_join(
      matched_vanilla,
      by = c("vanilla_dataset_bundle_id" = "dataset_bundle_id_left",
             "vanilla_effect" = "effect_left")
    ) %>%
    dplyr::left_join(
      matched_treatment,
      by = c("mu_dataset_bundle_id" = "dataset_bundle_id_right",
             "mu_effect" = "effect_right")
    ) %>%
    dplyr::mutate(
      delta_x = .data$mu_x - .data$vanilla_x,
      delta_y = .data$mu_y - .data$vanilla_y,
      movement_distance = sqrt(.data$delta_x^2 + .data$delta_y^2),
      delta_direction = dplyr::case_when(
        .data$delta_y > 0 & .data$delta_x > 0 ~ "more accurate + more diffuse",
        .data$delta_y > 0 & .data$delta_x <= 0 ~ "more accurate + less diffuse",
        .data$delta_y <= 0 & .data$delta_x > 0 ~ "less accurate + more diffuse",
        TRUE ~ "less accurate + less diffuse"
      )
    ) %>%
    dplyr::arrange(dplyr::desc(.data$movement_distance))

  if (is.finite(max_arrows_to_plot)) {
    flow_plot_df <- flow_df %>% dplyr::slice_head(n = as.integer(max_arrows_to_plot))
  } else {
    flow_plot_df <- flow_df
  }

  readr::write_csv(accepted_pair, matched_pairs_csv)
  readr::write_csv(flow_df, matched_delta_csv)

  top_labels <- flow_df %>% dplyr::slice_head(n = label_top_discrepancies)

  p_flow <- ggplot() +
    geom_point(
      data = flow_plot_df,
      aes(x = .data$vanilla_x, y = .data$vanilla_y),
      color = "black", alpha = 0.35, size = 1.0
    ) +
    geom_segment(
      data = flow_plot_df,
      aes(x = .data$vanilla_x, y = .data$vanilla_y,
          xend = .data$mu_x, yend = .data$mu_y,
          color = .data$delta_direction),
      alpha = arrow_alpha,
      linewidth = arrow_linewidth,
      arrow = grid::arrow(length = grid::unit(arrow_head_length_inches, "inches"), type = "closed")
    ) +
    geom_point(
      data = flow_plot_df,
      aes(x = .data$mu_x, y = .data$mu_y, color = .data$delta_direction),
      alpha = 0.55, size = 1.1
    ) +
    geom_point(
      data = top_labels,
      aes(x = .data$mu_x, y = .data$mu_y),
      color = "red", size = 1.8, alpha = 0.75
    ) +
    keff_x_scale(xlim = x_limits_resolved) +
    coord_cartesian(xlim = x_limits_resolved, ylim = y_limits_resolved) +
    labs(
      x = x_label, y = y_label,
      color = "Movement",
      title = sprintf("Matched effect movement: vanilla -> %s", treatment_pretty),
      subtitle = sprintf("Arrows start at vanilla, end at %s", treatment_pretty)
    ) +
    theme(legend.position = "bottom")

  ggsave(flow_png, p_flow, width = 10, height = 8, dpi = 150)
  print(p_flow)

  # Matched delta density
  if (nrow(flow_df)) {
    p_delta <- ggplot(flow_df, aes(x = .data$delta_x, y = .data$delta_y)) +
      geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
      geom_vline(xintercept = 0, color = "grey45", linetype = "dashed") +
      geom_density_2d_filled(
        alpha = 0.75, contour_var = "ndensity",
        breaks = delta_density_contour_breaks
      ) +
      geom_point(alpha = 0.25, size = 0.8, color = "black") +
      keff_ratio_x_scale() +
      labs(
        x = "Core-95 effective k ratio (treatment / vanilla)",
        y = paste0("Delta ", y_label, " (treatment - vanilla)"),
        fill = "Relative\ndensity",
        title = sprintf("Matched displacement density: vanilla -> %s", treatment_pretty)
      ) +
      theme(legend.position = "right")

    ggsave(delta_density_png, p_delta, width = 10, height = 6, dpi = 150)
    print(p_delta)
  }

  list(
    treatment = treatment_method,
    short = short,
    flow_df = flow_df,
    accepted_pair = accepted_pair,
    files = list(
      density_all = density_all_png,
      density_matched = density_matched_png,
      density_unmatched = density_unmatched_png,
      flow = flow_png,
      delta_density = delta_density_png,
      matched_pairs_csv = matched_pairs_csv,
      matched_delta_csv = matched_delta_csv
    )
  )
}
```

```{r per-pair-mu, fig.width=9, fig.height=6}
pair_results <- list()
pair_results$mu <- make_pair_plots(
  treatment_method = "susine_functional_mu",
  accepted_pair = accepted_by_arm$mu
)
```

```{r per-pair-pi, fig.width=9, fig.height=6}
pair_results$pi <- make_pair_plots(
  treatment_method = "susie_functional_pi",
  accepted_pair = accepted_by_arm$pi
)
```

# 7. Per-Dataset Summary

```{r per-dataset-summary}
per_dataset_auprc <- records_labeled %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$model_method, .data$model_AUPRC) %>%
  tidyr::pivot_wider(
    names_from = .data$model_method,
    values_from = .data$model_AUPRC,
    names_prefix = "auprc_"
  ) %>%
  dplyr::mutate(
    auprc_delta_mu = .data$auprc_susine_functional_mu - .data$auprc_susie_vanilla,
    auprc_delta_pi = .data$auprc_susie_functional_pi - .data$auprc_susie_vanilla
  )

build_per_arm_summary <- function(short, accepted_pair, flow_df) {
  matched_per_dataset <- accepted_pair %>%
    dplyr::count(.data$vanilla_dataset_bundle_id, name = paste0("n_matched_pairs_", short)) %>%
    dplyr::rename(dataset_bundle_id = .data$vanilla_dataset_bundle_id)

  movement_per_match <- flow_df %>%
    dplyr::transmute(
      dataset_bundle_id = .data$vanilla_dataset_bundle_id,
      vanilla_effect = .data$vanilla_effect,
      mu_effect = .data$mu_effect,
      movement_distance = .data$movement_distance
    )

  movement_per_dataset <- movement_per_match %>%
    dplyr::group_by(.data$dataset_bundle_id) %>%
    dplyr::summarise(
      "mean_movement_distance_{short}" := mean(.data$movement_distance, na.rm = TRUE),
      "median_movement_distance_{short}" := stats::median(.data$movement_distance, na.rm = TRUE),
      .groups = "drop"
    )

  matched_per_dataset %>%
    dplyr::full_join(movement_per_dataset, by = "dataset_bundle_id")
}

per_dataset_summary <- per_dataset_auprc
for (short in names(pair_results)) {
  arm_summary <- build_per_arm_summary(
    short = short,
    accepted_pair = pair_results[[short]]$accepted_pair,
    flow_df = pair_results[[short]]$flow_df
  )
  per_dataset_summary <- per_dataset_summary %>%
    dplyr::left_join(arm_summary, by = "dataset_bundle_id") %>%
    dplyr::mutate(
      "n_matched_pairs_{short}" := tidyr::replace_na(
        .data[[paste0("n_matched_pairs_", short)]], 0L
      )
    )
}

readr::write_csv(per_dataset_summary, per_dataset_summary_csv)
cat("Saved per-dataset summary:", per_dataset_summary_csv, "\n")

# Long-format for downstream faceting.
per_dataset_long <- per_dataset_summary %>%
  dplyr::select(
    .data$dataset_bundle_id,
    .data$auprc_susie_vanilla,
    .data$auprc_susine_functional_mu,
    .data$auprc_susie_functional_pi,
    .data$auprc_delta_mu, .data$auprc_delta_pi,
    .data$n_matched_pairs_mu, .data$n_matched_pairs_pi,
    .data$mean_movement_distance_mu, .data$mean_movement_distance_pi
  ) %>%
  tidyr::pivot_longer(
    cols = c(
      "auprc_delta_mu", "auprc_delta_pi",
      "n_matched_pairs_mu", "n_matched_pairs_pi",
      "mean_movement_distance_mu", "mean_movement_distance_pi"
    ),
    names_to = c(".value", "treatment_short"),
    names_pattern = "(.+)_(mu|pi)$"
  ) %>%
  dplyr::mutate(
    treatment_method = dplyr::recode(
      .data$treatment_short,
      mu = "susine_functional_mu",
      pi = "susie_functional_pi"
    ),
    auprc_treatment = dplyr::if_else(
      .data$treatment_short == "mu",
      .data$auprc_susine_functional_mu,
      .data$auprc_susie_functional_pi
    )
  )
```

# 8. AUPRC Attribution

```{r auprc-delta-hist, fig.width=10, fig.height=5}
delta_summary <- per_dataset_long %>%
  dplyr::group_by(.data$treatment_method) %>%
  dplyr::summarise(
    n = sum(is.finite(.data$auprc_delta)),
    n_pos = sum(.data$auprc_delta > 0, na.rm = TRUE),
    n_neg = sum(.data$auprc_delta < 0, na.rm = TRUE),
    median_delta = stats::median(.data$auprc_delta, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    label = sprintf(
      "%s: %d gains, %d losses, median = %.4f",
      .data$treatment_method, .data$n_pos, .data$n_neg, .data$median_delta
    )
  )
cat("AUPRC delta summary by treatment:\n")
print(delta_summary)

p_auprc_delta <- ggplot(per_dataset_long, aes(x = .data$auprc_delta,
                                              fill = .data$treatment_method)) +
  geom_vline(xintercept = 0, color = "grey45", linetype = "dashed") +
  geom_histogram(
    aes(y = after_stat(density)),
    position = "identity", bins = hist_bins, alpha = 0.45, color = NA
  ) +
  scale_fill_manual(
    values = treatment_color[treatment_methods],
    breaks = treatment_methods
  ) +
  labs(
    x = "AUPRC delta (treatment - vanilla)",
    y = "Density",
    fill = "Treatment",
    title = "Per-dataset AUPRC improvement by treatment arm",
    subtitle = paste(delta_summary$label, collapse = " | ")
  ) +
  theme(legend.position = "bottom")

ggsave(auprc_delta_hist_png, p_auprc_delta, width = 10, height = 5, dpi = 150)
print(p_auprc_delta)
```

```{r match-count-hist, fig.width=10, fig.height=5}
p_match_count <- ggplot(per_dataset_long, aes(x = .data$n_matched_pairs)) +
  geom_histogram(binwidth = 1, fill = "#2b8cbe", alpha = 0.7, color = NA) +
  scale_x_continuous(breaks = 0:10) +
  facet_wrap(~ treatment_method) +
  labs(
    x = "Number of accepted matched pairs per dataset (out of L = 10)",
    y = "Number of datasets",
    title = "Per-dataset matched-pair count by treatment arm"
  )

ggsave(match_count_hist_png, p_match_count, width = 10, height = 5, dpi = 150)
print(p_match_count)
```

```{r auprc-vs-match-scatter, fig.width=11, fig.height=5}
arm_rho <- per_dataset_long %>%
  dplyr::group_by(.data$treatment_method) %>%
  dplyr::summarise(
    rho = suppressWarnings(stats::cor(
      .data$n_matched_pairs, .data$auprc_delta,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    n = dplyr::n(),
    label = sprintf("rho = %.3f (n=%d)", .data$rho, .data$n),
    .groups = "drop"
  )
cat("Per-arm match-count vs AUPRC-delta Spearman rho:\n")
print(arm_rho)

p_auprc_match_scatter <- ggplot(
  per_dataset_long,
  aes(x = .data$n_matched_pairs, y = .data$auprc_delta,
      color = .data$mean_movement_distance)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_jitter(width = 0.25, height = 0, alpha = 0.55, size = 1.3) +
  geom_smooth(
    method = "loess", formula = y ~ x, se = FALSE,
    color = "black", linetype = "dashed", linewidth = 0.6
  ) +
  geom_text(
    data = arm_rho,
    aes(x = -Inf, y = Inf, label = .data$label),
    hjust = -0.05, vjust = 1.4, size = 3.5,
    inherit.aes = FALSE
  ) +
  facet_wrap(~ treatment_method, scales = "free_y") +
  scale_x_continuous(breaks = 0:10) +
  scale_color_viridis_c(option = "C", na.value = "grey60") +
  labs(
    x = "Number of accepted matched pairs per dataset",
    y = "AUPRC delta (treatment - vanilla)",
    color = "Mean matched\npair movement",
    title = "Per-dataset: matching density vs AUPRC improvement, by treatment"
  )

ggsave(auprc_match_scatter_png, p_auprc_match_scatter, width = 11, height = 5, dpi = 150)
print(p_auprc_match_scatter)
```

# 9. Confound Diagnostics

```{r baseline-vs-treatment, fig.width=11, fig.height=6}
n_above_diag <- per_dataset_long %>%
  dplyr::group_by(.data$treatment_method) %>%
  dplyr::summarise(
    n = sum(is.finite(.data$auprc_delta)),
    n_above = sum(.data$auprc_delta > 0, na.rm = TRUE),
    pearson_r = suppressWarnings(stats::cor(
      .data$auprc_susie_vanilla, .data$auprc_treatment,
      use = "pairwise.complete.obs"
    )),
    label = sprintf(
      "Pearson r = %.3f | %d/%d above y=x",
      .data$pearson_r, .data$n_above, .data$n
    ),
    .groups = "drop"
  )
cat("Baseline vs treatment AUPRC summary:\n")
print(n_above_diag)

p_baseline_vs_treatment <- ggplot(
  per_dataset_long,
  aes(x = .data$auprc_susie_vanilla, y = .data$auprc_treatment,
      color = .data$n_matched_pairs)
) +
  geom_abline(slope = 1, intercept = 0, color = "grey45", linetype = "dashed") +
  geom_point(alpha = 0.6, size = 1.3) +
  geom_smooth(
    method = "lm", formula = y ~ x, se = FALSE,
    color = "black", linetype = "dashed", linewidth = 0.5
  ) +
  geom_text(
    data = n_above_diag,
    aes(x = -Inf, y = Inf, label = .data$label),
    hjust = -0.05, vjust = 1.4, size = 3.5,
    inherit.aes = FALSE
  ) +
  facet_wrap(~ treatment_method) +
  scale_color_viridis_c(option = "C") +
  coord_fixed() +
  labs(
    x = "Vanilla SuSiE AUPRC",
    y = "Treatment AUPRC",
    color = "Matched pairs",
    title = "Per-dataset AUPRC: vanilla vs treatment (per arm)"
  )

ggsave(baseline_vs_treatment_png, p_baseline_vs_treatment, width = 11, height = 6, dpi = 150)
print(p_baseline_vs_treatment)
```

```{r stratified-confound, fig.width=12, fig.height=10}
quart_breaks <- stats::quantile(
  per_dataset_summary$auprc_susie_vanilla,
  probs = c(0, 0.25, 0.5, 0.75, 1.0),
  na.rm = TRUE, names = FALSE
)
quart_breaks <- unique(quart_breaks)

if (length(quart_breaks) >= 3L) {
  per_dataset_long_strat <- per_dataset_long %>%
    dplyr::mutate(
      vanilla_auprc_quartile = cut(
        .data$auprc_susie_vanilla,
        breaks = quart_breaks,
        include.lowest = TRUE
      )
    ) %>%
    dplyr::filter(!is.na(.data$vanilla_auprc_quartile))

  quartile_rho <- per_dataset_long_strat %>%
    dplyr::group_by(.data$treatment_method, .data$vanilla_auprc_quartile) %>%
    dplyr::summarise(
      n = dplyr::n(),
      rho = suppressWarnings(stats::cor(
        .data$n_matched_pairs, .data$auprc_delta,
        method = "spearman", use = "pairwise.complete.obs"
      )),
      label = sprintf("rho = %.3f (n=%d)", .data$rho, .data$n),
      .groups = "drop"
    )

  cat("Per-(treatment, quartile) Spearman rho:\n")
  print(quartile_rho)

  p_strat <- ggplot(
    per_dataset_long_strat,
    aes(x = .data$n_matched_pairs, y = .data$auprc_delta,
        color = .data$mean_movement_distance)
  ) +
    geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
    geom_jitter(width = 0.25, height = 0, alpha = 0.55, size = 1.1) +
    geom_smooth(
      method = "loess", formula = y ~ x, se = FALSE,
      color = "black", linetype = "dashed", linewidth = 0.5
    ) +
    geom_text(
      data = quartile_rho,
      aes(x = -Inf, y = Inf, label = .data$label),
      hjust = -0.05, vjust = 1.4, size = 3,
      inherit.aes = FALSE
    ) +
    facet_grid(treatment_method ~ vanilla_auprc_quartile, scales = "free_y") +
    scale_x_continuous(breaks = 0:10) +
    scale_color_viridis_c(option = "C", na.value = "grey60") +
    labs(
      x = "Number of accepted matched pairs per dataset",
      y = "AUPRC delta (treatment - vanilla)",
      color = "Mean matched\npair movement",
      title = "Match count vs AUPRC delta by treatment x vanilla AUPRC quartile",
      subtitle = "Lower quartiles = harder datasets; persistent negative rho within panels => trend is real, not confounded"
    )

  ggsave(scatter_stratified_png, p_strat, width = 12, height = 10, dpi = 150)
  print(p_strat)
} else {
  cat("Not enough distinct vanilla AUPRC values to form quartiles; skipping.\n")
}
```

```{r confound-check, fig.width=11, fig.height=5}
confound_rho <- per_dataset_long %>%
  dplyr::group_by(.data$treatment_method) %>%
  dplyr::summarise(
    rho = suppressWarnings(stats::cor(
      .data$auprc_susie_vanilla, .data$n_matched_pairs,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    label = sprintf("rho = %.3f", .data$rho),
    .groups = "drop"
  )
cat("Per-arm vanilla AUPRC vs match count Spearman rho:\n")
print(confound_rho)

p_confound <- ggplot(
  per_dataset_long,
  aes(x = .data$auprc_susie_vanilla, y = .data$n_matched_pairs,
      color = .data$auprc_delta)
) +
  geom_jitter(width = 0, height = 0.25, alpha = 0.6, size = 1.3) +
  geom_smooth(
    method = "loess", formula = y ~ x, se = FALSE,
    color = "black", linetype = "dashed", linewidth = 0.6
  ) +
  geom_text(
    data = confound_rho,
    aes(x = -Inf, y = Inf, label = .data$label),
    hjust = -0.05, vjust = 1.4, size = 3.5,
    inherit.aes = FALSE
  ) +
  scale_color_gradient2(
    low = "#2b6cb0", mid = "white", high = "#c53030", midpoint = 0
  ) +
  scale_y_continuous(breaks = 0:10) +
  facet_wrap(~ treatment_method) +
  labs(
    x = "Vanilla AUPRC (low = hard, high = easy)",
    y = "Number of accepted matched pairs",
    color = "AUPRC delta\n(treatment - vanilla)",
    title = "Confound check: vanilla AUPRC vs match count, by treatment"
  )

ggsave(confound_check_png, p_confound, width = 11, height = 5, dpi = 150)
print(p_confound)
```

# 10. Output Index

```{r output-index}
per_pair_files <- function(short) {
  paths <- pair_results[[short]]$files
  tibble::tibble(
    artifact = paste0(c(
      "all-effect differential density",
      "matched differential density",
      "unmatched differential density",
      "matched flow plot",
      "matched delta density",
      "accepted matched pairs",
      "matched pair deltas"
    ), " (vs ", short, ")"),
    path = c(
      paths$density_all,
      paths$density_matched,
      paths$density_unmatched,
      paths$flow,
      paths$delta_density,
      paths$matched_pairs_csv,
      paths$matched_delta_csv
    )
  )
}

global_outputs <- tibble::tibble(
  artifact = c(
    "3-arm scatter (all effects)",
    "3-arm univariate diffuseness histogram",
    "3-arm univariate accuracy histogram",
    "AUPRC delta histogram (overlay)",
    "match count histogram (faceted)",
    "AUPRC delta vs match count scatter (faceted)",
    "baseline vs treatment AUPRC scatter (faceted)",
    "AUPRC delta vs match count stratified by vanilla AUPRC quartile",
    "vanilla AUPRC vs match count (confound check, faceted)",
    "per-dataset AUPRC and match summary"
  ),
  path = c(
    scatter_all_png,
    hist_diffuseness_png,
    hist_accuracy_png,
    auprc_delta_hist_png,
    match_count_hist_png,
    auprc_match_scatter_png,
    baseline_vs_treatment_png,
    scatter_stratified_png,
    confound_check_png,
    per_dataset_summary_csv
  )
)

outputs <- dplyr::bind_rows(
  global_outputs,
  per_pair_files("mu"),
  per_pair_files("pi")
)

kable(outputs)
```
