---
title: "Effect Drift Analysis (5-arm: cold + warm refits)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Reads the 5-arm refit records and computes per-dataset **effect drift**
between every interesting arm pair, plus an AUPRC trajectory across the
two refit chains:

```
mu chain:                              pi chain:
  vanilla -> mu                          vanilla -> pi
  mu -> vanilla_warm_from_mu             pi -> vanilla_warm_from_pi
  vanilla -> vanilla_warm_from_mu        vanilla -> vanilla_warm_from_pi
```

**Effect drift** between two fits A and B is computed by:

1. Build the `L x L` correlation matrix of `fitted_y_effect` vectors
   (inactive effects with sd=0 contribute correlation 0).
2. Solve the max-weight 1-1 assignment via `clue::solve_LSAP(cor_mat,
   maximum = TRUE)`.
3. Record all `L` assignment correlations.
4. Derive per-pair summaries: `mean_corr`, `min_corr`, `n_below_0p9`
   (count of assignment edges with correlation < 0.9), legacy
   `weighted_drift` (`1 - corr`), `weighted_chord_drift`
   (`sqrt(2 * (1 - corr))`, geometric L2 distance between unit-normalized
   fitted-y vectors), and `weighted_r2_drift` (`1 - corr^2`,
   `sin^2(theta)` between matched 1D fitted-y subspaces, the explanatory-
   subspace residual variance). All weighted by
   `min(var_y_explained_A_i, var_y_explained_B_j)` so inactive pairs
   contribute little. Downstream plots are produced in parallel for the
   chord and 1 - r^2 versions; see the `_r2.png` siblings of the basin
   figures.

This drift framework + AUPRC trajectory operationalizes the
**refinement-vs-exploration test**:

| D(vanilla, treatment) | D(vanilla, vanilla_warm) | dAUPRC(treatment) | dAUPRC(vanilla_warm) | Mechanism |
|---|---|---|---|---|
| Low | Low | + | + | Within-basin PIP weighting; basin unchanged |
| Low | Low | + | ~0 | PIP-weighting only; reverts when prior removed |
| High | High | + | + | **Exploration win**: new basin, vanilla holds it |
| High | Low | + | ~0 | New basin but unstable; vanilla snaps back |

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_drift_analysis_5arm.Rmd")
library(here)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)
library(clue)         # solve_LSAP for max-weight 1-1 assignment

theme_set(theme_bw(base_size = 12))
```

# 1. User Parameters

```{r user-parameters}
output_root <- here("output")
study_name <- "baseline_sims_5arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
records_rds <- file.path(study_dir, "effect_refit_records.rds")

baseline_method <- "susie_vanilla"
mu_method <- "susine_functional_mu"
pi_method <- "susie_functional_pi"
warm_mu_method <- "susie_vanilla_warm_from_mu"
warm_pi_method <- "susie_vanilla_warm_from_pi"

# Pretty labels for plotting.
method_labels <- c(
  susie_vanilla = "SuSiE-vanilla (cold)",
  susine_functional_mu = "SuSiNE-functional-mu (cold)",
  susie_vanilla_warm_from_mu = "SuSiE-vanilla (warm from mu)",
  susie_functional_pi = "SuSiE-functional-pi (cold)",
  susie_vanilla_warm_from_pi = "SuSiE-vanilla (warm from pi)"
)
method_color <- c(
  susie_vanilla = "black",
  susine_functional_mu = "#2b8cbe",
  susie_vanilla_warm_from_mu = "#9ecae1",
  susie_functional_pi = "#74c476",
  susie_vanilla_warm_from_pi = "#bae4b3"
)

# Drift threshold (the user's "unpaired" cutoff).
drift_corr_cutoff <- 0.9

# Active-effect threshold for the weighted/filtered drift summaries.
active_var_threshold <- 0.005

# Effect-pair threshold used for matched/unmatched diagnostic density plots.
effect_match_corr_min <- 0.99

# Joint effect-quality axes. `effect_pip_entropy_core95` is log effective k;
# axis labels are shown on the effective-k scale.
x_metric <- "effect_pip_entropy_core95"
y_metric <- "accuracy_ratio"
x_label <- expression(
  paste("Effect diffuseness: effective top-95% support size  ",
        K[l], " = exp(", italic(H), "(", alpha[l]^{(95)}, "))")
)
y_label <- expression(
  paste("Effect accuracy: ", A[l],
        " = max[true causal] PIP / max[any] PIP")
)
x_keff_breaks <- c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000)
x_limits <- NULL
y_limits <- c(0, 1)

density_grid_n <- 140L
density_bandwidth <- c(0.45, 0.10)
density_trim_quantile <- c(0.005, 0.995)
density_fill_quantile <- 0.95

# Plot output directory.
plot_dir <- file.path(study_dir, "drift_analysis_plots")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Output filenames.
drift_assignments_csv <- file.path(study_dir, "drift_assignments_5arm.csv")
drift_summary_csv <- file.path(study_dir, "drift_summary_5arm.csv")
auprc_per_dataset_csv <- file.path(study_dir, "auprc_per_dataset_5arm.csv")

drift_profile_hist_png <- file.path(plot_dir, "drift_profile_histograms.png")
chain_trajectory_png <- file.path(plot_dir, "chain_auprc_trajectory.png")
headline_scatter_png <- file.path(plot_dir, "headline_drift_vs_warm_auprc.png")
unpaired_vs_delta_png <- file.path(plot_dir, "unpaired_count_vs_auprc_delta.png")
mu_vs_pi_drift_png <- file.path(plot_dir, "mu_vs_pi_drift_comparison.png")
basin_pip_drift_png <- file.path(plot_dir, "basin_and_pip_drift_summary.png")
effect_quality_density_mu_png <- file.path(
  plot_dir, "effect_quality_density_pooled_mu_5arm.png"
)
effect_quality_density_pi_png <- file.path(
  plot_dir, "effect_quality_density_pooled_pi_5arm.png"
)
effect_quality_pair_rows_csv <- file.path(
  study_dir, "effect_quality_pair_rows_5arm.csv"
)
pip_records_rds <- file.path(study_dir, "pip_records_5arm.rds")
pip_drift_summary_csv <- file.path(study_dir, "pip_drift_summary_5arm.csv")
causal_idx_rds <- file.path(study_dir, "causal_idx_5arm.rds")
causal_truth_rds <- file.path(study_dir, "causal_truth_5arm.rds")
pr_curves_pooled_png <- file.path(plot_dir, "pr_curves_pooled_top8.png")
pr_curves_pooled_csv <- file.path(study_dir, "pr_curves_pooled_top8_5arm.csv")
paper_figure_png <- file.path(plot_dir, "paper_figure_effect_drift.png")

# Parallel outputs for the 1 - r^2 ("explanatory subspace distance") version
# of the basin-drift metric. The chord and r2 versions consume the same
# drift_summary table; only the column read and the axis labels differ.
headline_scatter_r2_png <- file.path(
  plot_dir, "headline_drift_vs_warm_auprc_r2.png"
)
mu_vs_pi_drift_r2_png <- file.path(
  plot_dir, "mu_vs_pi_drift_comparison_r2.png"
)
basin_pip_drift_r2_png <- file.path(
  plot_dir, "basin_and_pip_drift_summary_r2.png"
)
paper_figure_r2_png <- file.path(
  plot_dir, "paper_figure_effect_drift_r2.png"
)

# Basin-drift-vs-PIP-JSD scatter, one figure per basin metric. Each figure
# is a 2 x 2 grid: rows = chain (mu/pi), columns = pair_kind (cold->ann /
# cold->warm). x = basin drift, y = PIP JSD.
basin_vs_pip_jsd_chord_png <- file.path(
  plot_dir, "basin_vs_pip_jsd_chord.png"
)
basin_vs_pip_jsd_r2_png <- file.path(
  plot_dir, "basin_vs_pip_jsd_r2.png"
)

# Number of causals counted as positives in pooled PR scoring (top-K by
# |beta| per dataset). This matches the baseline simulation study, which
# uses the top 8 = 3 sparse + 5 oligogenic tier of the SuSiE 2.0
# oligogenic architecture.
pr_top_k <- 8L

cat("Study dir:", study_dir, "\n")
cat("Plot dir:", plot_dir, "\n")
cat("Drift threshold for `n_below_<cut>`:", drift_corr_cutoff, "\n")
cat("Active-effect var-explained threshold:", active_var_threshold, "\n")
cat("Effect-quality matched threshold:", effect_match_corr_min, "\n")
```

# 2. Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

format_duration <- function(seconds) {
  seconds <- as.numeric(seconds)
  if (!is.finite(seconds)) return("unknown")
  h <- floor(seconds / 3600)
  m <- floor((seconds %% 3600) / 60)
  s <- round(seconds %% 60)
  sprintf("%02d:%02d:%02d", h, m, s)
}

require_metric <- function(df, metric) {
  if (!metric %in% names(df)) {
    stop("Metric column not found: ", metric)
  }
}

resolve_limits <- function(values, user_limits, trim = density_trim_quantile) {
  if (!is.null(user_limits)) {
    return(user_limits)
  }
  values <- values[is.finite(values)]
  if (!length(values)) {
    stop("Cannot infer limits from an empty/non-finite vector.")
  }
  stats::quantile(values, probs = trim, na.rm = TRUE, names = FALSE)
}

keff_x_scale <- function(breaks = x_keff_breaks, xlim = NULL) {
  finite_breaks <- breaks[breaks > 0]
  scale_x_continuous(
    breaks = log(finite_breaks),
    labels = formatC(finite_breaks, format = "fg", big.mark = ","),
    limits = xlim
  )
}

safe_corr <- function(x) {
  pmax(-1, pmin(1, x))
}

chord_distance <- function(corr) {
  sqrt(pmax(0, 2 * (1 - safe_corr(corr))))
}

# 1 - r^2 distance: residual variance after linear projection (sin^2 of the
# principal angle between the two 1D fitted_y subspaces). Bounded in [0, 1],
# so 0.95 -> 0.0975 ("near-identical basis") and 0.5 -> 0.75 ("most of the
# way to a fully different basis"). See methods write-up.
r2_distance <- function(corr) {
  1 - safe_corr(corr)^2
}

# Stack effect rows for one (dataset, model) into an n x L matrix of
# fitted_y_effect vectors, in effect order.
records_to_fitted_y_matrix <- function(rows) {
  rows <- rows %>% dplyr::arrange(.data$effect)
  mat <- do.call(cbind, rows$fitted_y_effect)
  storage.mode(mat) <- "double"
  colnames(mat) <- as.character(rows$effect)
  mat
}

# Correlation matrix robust to inactive (sd=0) effects.
safe_cor_matrix <- function(mat_a, mat_b, sd_eps = 1e-12) {
  if (nrow(mat_a) != nrow(mat_b)) {
    stop("Mismatched n (", nrow(mat_a), " vs ", nrow(mat_b), ").")
  }
  sd_a <- apply(mat_a, 2L, stats::sd)
  sd_b <- apply(mat_b, 2L, stats::sd)
  active_a <- is.finite(sd_a) & sd_a > sd_eps
  active_b <- is.finite(sd_b) & sd_b > sd_eps

  cor_mat <- matrix(0, nrow = ncol(mat_a), ncol = ncol(mat_b),
                    dimnames = list(colnames(mat_a), colnames(mat_b)))
  if (any(active_a) && any(active_b)) {
    sub <- suppressWarnings(stats::cor(
      mat_a[, active_a, drop = FALSE],
      mat_b[, active_b, drop = FALSE],
      use = "pairwise.complete.obs"
    ))
    sub[!is.finite(sub)] <- 0
    cor_mat[active_a, active_b] <- sub
  }
  list(cor = cor_mat, active_a = active_a, active_b = active_b)
}

# For one (dataset, A, B) pair, run Hungarian on the L x L correlation
# matrix and return per-edge data.
compute_drift_pair <- function(rows_a, rows_b,
                               method_a, method_b,
                               dataset_bundle_id, matrix_id) {
  rows_a <- rows_a %>% dplyr::arrange(.data$effect)
  rows_b <- rows_b %>% dplyr::arrange(.data$effect)

  mat_a <- records_to_fitted_y_matrix(rows_a)
  mat_b <- records_to_fitted_y_matrix(rows_b)
  cor_info <- safe_cor_matrix(mat_a, mat_b)
  cor_mat <- cor_info$cor

  # clue::solve_LSAP() requires nonnegative entries even with
  # maximum = TRUE. Shift the full matrix if needed; adding a constant
  # to every edge preserves the optimal 1-1 assignment.
  assignment_weights <- cor_mat
  min_weight <- suppressWarnings(min(assignment_weights, na.rm = TRUE))
  if (!is.finite(min_weight)) {
    min_weight <- 0
  }
  if (min_weight < 0) {
    assignment_weights <- assignment_weights - min_weight
  }
  assignment <- as.integer(clue::solve_LSAP(
    assignment_weights,
    maximum = TRUE
  ))
  L <- ncol(mat_a)

  effect_a_ids <- as.integer(rows_a$effect)
  effect_b_ids <- as.integer(rows_b$effect)
  var_y_a <- as.numeric(rows_a$var_y_explained_frac)
  var_y_b <- as.numeric(rows_b$var_y_explained_frac)

  tibble::tibble(
    dataset_bundle_id = dataset_bundle_id,
    matrix_id = matrix_id,
    method_a = method_a,
    method_b = method_b,
    edge = seq_len(L),
    effect_a = effect_a_ids,
    effect_b = effect_b_ids[assignment],
    corr = cor_mat[cbind(seq_len(L), assignment)],
    var_y_explained_a = var_y_a,
    var_y_explained_b = var_y_b[assignment],
    active_a = cor_info$active_a,
    active_b = cor_info$active_b[assignment]
  )
}

# Summarize per (dataset, A, B) from the per-edge assignment table.
summarize_drift <- function(edges_df, drift_corr_cutoff, active_var_threshold) {
  edges_df %>%
    dplyr::group_by(
      .data$dataset_bundle_id, .data$matrix_id,
      .data$method_a, .data$method_b
    ) %>%
    dplyr::summarise(
      n_edges = dplyr::n(),
      mean_corr = mean(.data$corr, na.rm = TRUE),
      median_corr = stats::median(.data$corr, na.rm = TRUE),
      min_corr = suppressWarnings(min(.data$corr, na.rm = TRUE)),
      n_below_cutoff = sum(.data$corr < drift_corr_cutoff, na.rm = TRUE),
      n_active_pairs = sum(
        .data$var_y_explained_a > active_var_threshold &
          .data$var_y_explained_b > active_var_threshold,
        na.rm = TRUE
      ),
      n_active_below_cutoff = sum(
        .data$corr < drift_corr_cutoff &
          .data$var_y_explained_a > active_var_threshold &
          .data$var_y_explained_b > active_var_threshold,
        na.rm = TRUE
      ),
      mean_corr_active = {
        keep <- .data$var_y_explained_a > active_var_threshold &
                  .data$var_y_explained_b > active_var_threshold
        if (any(keep)) mean(.data$corr[keep], na.rm = TRUE) else NA_real_
      },
      weighted_drift = {
        w <- pmin(
          tidyr::replace_na(.data$var_y_explained_a, 0),
          tidyr::replace_na(.data$var_y_explained_b, 0)
        )
        if (sum(w) > 0) {
          # 1 - weighted-mean correlation, so larger = more drift
          1 - sum(w * .data$corr, na.rm = TRUE) / sum(w)
        } else {
          NA_real_
        }
      },
      weighted_chord_drift = {
        w <- pmin(
          tidyr::replace_na(.data$var_y_explained_a, 0),
          tidyr::replace_na(.data$var_y_explained_b, 0)
        )
        if (sum(w) > 0) {
          sum(w * chord_distance(.data$corr), na.rm = TRUE) / sum(w)
        } else {
          NA_real_
        }
      },
      weighted_r2_drift = {
        w <- pmin(
          tidyr::replace_na(.data$var_y_explained_a, 0),
          tidyr::replace_na(.data$var_y_explained_b, 0)
        )
        if (sum(w) > 0) {
          sum(w * r2_distance(.data$corr), na.rm = TRUE) / sum(w)
        } else {
          NA_real_
        }
      },
      .groups = "drop"
    )
}
```

# 3. Load Records

```{r load-records}
stopifnot(file.exists(records_rds))
records <- readRDS(records_rds) %>%
  dplyr::mutate(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    matrix_id = as.character(.data$matrix_id),
    effect = as.integer(.data$effect)
  )

methods_present <- sort(unique(records$model_method))
cat("Methods in records:", paste(methods_present, collapse = ", "), "\n")

required_methods <- c(baseline_method, mu_method, warm_mu_method,
                      pi_method, warm_pi_method)
missing_methods <- setdiff(required_methods, methods_present)
if (length(missing_methods)) {
  stop("Missing methods in records: ", paste(missing_methods, collapse = ", "),
       ". Re-run effect_matching_refit_export_5arm.Rmd first.")
}

records <- records %>% dplyr::filter(.data$model_method %in% required_methods)
require_metric(records, x_metric)
require_metric(records, y_metric)

# Index records by (dataset, model_method) for fast pair access.
records_by_dataset_method <- records %>% split(list(.$dataset_bundle_id, .$model_method))
```

# 4. Pairwise Drift

```{r compute-drift}
# Pairs we want (label, method_a, method_b).
pair_specs <- tibble::tribble(
  ~pair_id,                        ~chain, ~method_a,         ~method_b,
  "vanilla__mu",                   "mu",   baseline_method,   mu_method,
  "mu__vanilla_warm_from_mu",      "mu",   mu_method,         warm_mu_method,
  "vanilla__vanilla_warm_from_mu", "mu",   baseline_method,   warm_mu_method,
  "vanilla__pi",                   "pi",   baseline_method,   pi_method,
  "pi__vanilla_warm_from_pi",      "pi",   pi_method,         warm_pi_method,
  "vanilla__vanilla_warm_from_pi", "pi",   baseline_method,   warm_pi_method
)
print(pair_specs)

# Datasets with all 5 arms present.
datasets_complete <- records %>%
  dplyr::count(.data$dataset_bundle_id, .data$model_method) %>%
  dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
  dplyr::filter(.data$n_methods == length(required_methods)) %>%
  dplyr::pull(.data$dataset_bundle_id)
cat("Datasets with all 5 arms present:", length(datasets_complete), "\n")

dataset_meta <- records %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$matrix_id) %>%
  dplyr::filter(.data$dataset_bundle_id %in% datasets_complete)

started <- proc.time()[["elapsed"]]
edge_chunks <- list()
n_total <- length(datasets_complete) * nrow(pair_specs)
counter <- 0L

for (ds_idx in seq_along(datasets_complete)) {
  ds_id <- datasets_complete[[ds_idx]]
  matrix_id_i <- dataset_meta$matrix_id[
    match(ds_id, dataset_meta$dataset_bundle_id)
  ]
  for (pi_idx in seq_len(nrow(pair_specs))) {
    counter <- counter + 1L
    spec <- pair_specs[pi_idx, ]
    rows_a <- records_by_dataset_method[[paste(ds_id, spec$method_a, sep = ".")]]
    rows_b <- records_by_dataset_method[[paste(ds_id, spec$method_b, sep = ".")]]
    if (is.null(rows_a) || is.null(rows_b)) {
      next
    }
    edge_chunks[[length(edge_chunks) + 1L]] <- compute_drift_pair(
      rows_a = rows_a, rows_b = rows_b,
      method_a = spec$method_a, method_b = spec$method_b,
      dataset_bundle_id = ds_id,
      matrix_id = matrix_id_i
    ) %>% dplyr::mutate(pair_id = spec$pair_id, chain = spec$chain)
  }
  if (ds_idx %% 100L == 0L || ds_idx == length(datasets_complete)) {
    elapsed <- proc.time()[["elapsed"]] - started
    eta <- elapsed / counter * (n_total - counter)
    cat(sprintf(
      "Drift %d/%d datasets (%d/%d pairs) | elapsed %s | ETA %s\n",
      ds_idx, length(datasets_complete), counter, n_total,
      format_duration(elapsed), format_duration(eta)
    ))
  }
}

drift_assignments <- dplyr::bind_rows(edge_chunks)
readr::write_csv(drift_assignments, drift_assignments_csv)
cat("Drift assignment rows:", nrow(drift_assignments), "\n")

drift_summary <- summarize_drift(drift_assignments, drift_corr_cutoff, active_var_threshold) %>%
  dplyr::left_join(pair_specs, by = c("method_a", "method_b"))
readr::write_csv(drift_summary, drift_summary_csv)
cat("Drift summary rows:", nrow(drift_summary), "\n")
```

# 5. Drift Sanity (D(arm, arm) == 0)

```{r drift-sanity, fig.width=8, fig.height=4}
# As a self-test, compute drift of vanilla against itself for a small subset.
# All correlations should be 1, mean_corr ~ 1, n_below_0p9 == 0.
sanity_ids <- head(datasets_complete, 10L)
sanity_rows <- list()
for (ds_id in sanity_ids) {
  rows_v <- records_by_dataset_method[[paste(ds_id, baseline_method, sep = ".")]]
  if (is.null(rows_v)) next
  matrix_id_i <- dataset_meta$matrix_id[match(ds_id, dataset_meta$dataset_bundle_id)]
  sanity_rows[[length(sanity_rows) + 1L]] <- compute_drift_pair(
    rows_a = rows_v, rows_b = rows_v,
    method_a = baseline_method, method_b = baseline_method,
    dataset_bundle_id = ds_id, matrix_id = matrix_id_i
  )
}
sanity_edges <- dplyr::bind_rows(sanity_rows)
sanity_summary <- summarize_drift(sanity_edges, drift_corr_cutoff, active_var_threshold)
cat("Self-drift sanity (vanilla vs vanilla):\n")
print(sanity_summary %>%
        dplyr::summarise(
          mean_of_mean_corr = mean(.data$mean_corr, na.rm = TRUE),
          mean_n_below_cutoff = mean(.data$n_below_cutoff),
          max_n_below_cutoff = max(.data$n_below_cutoff)
        ))
# Note: inactive (sd=0) effects against themselves give correlation 0 by
# construction, so n_below_cutoff > 0 is expected. mean_corr_active should
# be close to 1.
cat("Self-drift mean corr (active-only):", mean(sanity_summary$mean_corr_active, na.rm = TRUE), "\n")
```

# 6. Per-Dataset AUPRC Trajectory

```{r per-dataset-auprc}
auprc_per_dataset <- records %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$model_method, .data$model_AUPRC) %>%
  tidyr::pivot_wider(
    names_from = .data$model_method,
    values_from = .data$model_AUPRC,
    names_prefix = "auprc_"
  ) %>%
  dplyr::mutate(
    auprc_delta_mu = .data$auprc_susine_functional_mu - .data$auprc_susie_vanilla,
    auprc_delta_pi = .data$auprc_susie_functional_pi - .data$auprc_susie_vanilla,
    auprc_delta_warm_mu_vs_vanilla =
      .data$auprc_susie_vanilla_warm_from_mu - .data$auprc_susie_vanilla,
    auprc_delta_warm_pi_vs_vanilla =
      .data$auprc_susie_vanilla_warm_from_pi - .data$auprc_susie_vanilla,
    auprc_round_trip_mu =
      .data$auprc_susie_vanilla_warm_from_mu - .data$auprc_susine_functional_mu,
    auprc_round_trip_pi =
      .data$auprc_susie_vanilla_warm_from_pi - .data$auprc_susie_functional_pi
  )
readr::write_csv(auprc_per_dataset, auprc_per_dataset_csv)
cat("Per-dataset AUPRC rows:", nrow(auprc_per_dataset), "\n")

# Long format for spaghetti / facet plots.
auprc_long <- auprc_per_dataset %>%
  tidyr::pivot_longer(
    cols = dplyr::starts_with("auprc_"),
    names_to = "metric",
    values_to = "value"
  )
```

# 7. Drift Profile Histograms

```{r drift-profile-hists, fig.width=12, fig.height=7}
drift_assignments_for_plot <- drift_assignments %>%
  dplyr::mutate(
    pair_id = factor(.data$pair_id, levels = pair_specs$pair_id)
  )

p_drift_hists <- ggplot(drift_assignments_for_plot,
                        aes(x = .data$corr, fill = .data$chain)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 60, position = "identity", alpha = 0.55, color = NA
  ) +
  geom_vline(xintercept = drift_corr_cutoff, linetype = "dashed", color = "grey25") +
  facet_wrap(~ pair_id, ncol = 3) +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  coord_cartesian(xlim = c(-0.05, 1.0)) +
  labs(
    x = "Hungarian-assigned fitted_y correlation per effect",
    y = "Density",
    fill = "Chain",
    title = "Distribution of per-effect assignment correlations across all pairs",
    subtitle = sprintf(
      "Dashed line = drift_corr_cutoff (%.2f). All 10 effects per (dataset, pair) included.",
      drift_corr_cutoff
    )
  ) +
  theme(legend.position = "bottom")

ggsave(drift_profile_hist_png, p_drift_hists, width = 12, height = 7, dpi = 150)
print(p_drift_hists)
```

# 8. Chain AUPRC Trajectory

```{r chain-trajectory, fig.width=11, fig.height=5}
chain_long <- dplyr::bind_rows(
  auprc_per_dataset %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step1 = .data$auprc_susie_vanilla,
      step2 = .data$auprc_susine_functional_mu,
      step3 = .data$auprc_susie_vanilla_warm_from_mu
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step1 = .data$auprc_susie_vanilla,
      step2 = .data$auprc_susie_functional_pi,
      step3 = .data$auprc_susie_vanilla_warm_from_pi
    )
) %>%
  tidyr::pivot_longer(
    cols = c("step1", "step2", "step3"),
    names_to = "step",
    values_to = "auprc"
  ) %>%
  dplyr::mutate(
    step_label = dplyr::case_when(
      step == "step1" ~ "vanilla (cold)",
      step == "step2" & chain == "mu" ~ "mu",
      step == "step2" & chain == "pi" ~ "pi",
      step == "step3" ~ "vanilla (warm)",
      TRUE ~ step
    ),
    step_label = factor(step_label,
                        levels = c("vanilla (cold)", "mu", "pi", "vanilla (warm)"))
  )

p_chain <- ggplot(chain_long, aes(x = .data$step_label, y = .data$auprc)) +
  geom_line(aes(group = .data$dataset_bundle_id), alpha = 0.06) +
  geom_violin(aes(fill = .data$chain), alpha = 0.55, color = NA, scale = "width") +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.7) +
  facet_wrap(~ chain, scales = "free_x") +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  labs(
    x = NULL, y = "AUPRC",
    fill = "Chain",
    title = "Per-dataset AUPRC trajectory along each refit chain",
    subtitle = "Each thin line = one dataset; violin / box = marginal distribution per step."
  ) +
  theme(legend.position = "bottom")

ggsave(chain_trajectory_png, p_chain, width = 11, height = 5, dpi = 150)
print(p_chain)
```

# 9. Headline 4-Quadrant Test

```{r headline-scatter, fig.width=12, fig.height=5.5}
# For each chain, x is the basin chord drift between cold vanilla and the
# warm-vanilla refit (i.e. how much the basin moved AND survived prior
# removal). Two facets show two different AUPRC deltas with the same x:
#   - "vanilla -> annotated":  treatment AUPRC - cold vanilla AUPRC
#   - "vanilla -> vanilla_warm": warm-vanilla AUPRC - cold vanilla AUPRC
# Both should track x positively if a non-trivial basin movement is
# producing genuine fine-mapping improvement.
headline_long <- dplyr::bind_rows(
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      basin_drift = .data$weighted_chord_drift
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(
          dataset_bundle_id = .data$dataset_bundle_id,
          delta_auprc_treatment = .data$auprc_delta_mu,
          delta_auprc_warm = .data$auprc_delta_warm_mu_vs_vanilla
        ),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      basin_drift = .data$weighted_chord_drift
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(
          dataset_bundle_id = .data$dataset_bundle_id,
          delta_auprc_treatment = .data$auprc_delta_pi,
          delta_auprc_warm = .data$auprc_delta_warm_pi_vs_vanilla
        ),
      by = "dataset_bundle_id"
    )
) %>%
  tidyr::pivot_longer(
    cols = c("delta_auprc_treatment", "delta_auprc_warm"),
    names_to = "delta_kind",
    values_to = "delta_auprc"
  ) %>%
  dplyr::mutate(
    delta_kind = factor(
      .data$delta_kind,
      levels = c("delta_auprc_treatment", "delta_auprc_warm"),
      labels = c(
        "DAUPRC(annotated cold) - vanilla",
        "DAUPRC(vanilla, warm refit) - vanilla"
      )
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

headline_max_drift <- max(0.05,
                           max(headline_long$basin_drift, na.rm = TRUE))

p_headline <- ggplot(
  headline_long,
  aes(x = .data$basin_drift, y = .data$delta_auprc, color = .data$chain)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_point(alpha = 0.5, size = 1.2) +
  scale_color_manual(
    values = c(mu = "#2b8cbe", pi = "#74c476"),
    labels = c(mu = "mu chain (signed mu_0)",
               pi = "pi chain (unsigned pi)")
  ) +
  facet_wrap(~ delta_kind, ncol = 2) +
  scale_x_continuous(limits = c(0, headline_max_drift)) +
  labs(
    x = expression(
      paste("Basin chord drift  ",
            italic(D)["chord"],
            "(vanilla, vanilla-warm)")
    ),
    y = expression(paste(Delta, "AUPRC vs cold vanilla")),
    color = "Annotation channel",
    title = "AUPRC change versus basin chord drift, by chain and arm",
    subtitle = "x in both facets: basin chord drift between cold vanilla and the warm-vanilla refit. Left: AUPRC gain of the cold annotated fit over cold vanilla. Right: AUPRC gain of the warm-vanilla refit over cold vanilla. y=0 line: arm matches cold vanilla."
  ) +
  theme(legend.position = "bottom")

ggsave(headline_scatter_png, p_headline, width = 12, height = 5.5, dpi = 150)
print(p_headline)
```

```{r headline-scatter-r2, fig.width=12, fig.height=5.5}
# 1 - r^2 sibling of the headline scatter. Same shape, same data sources;
# only the basin metric column and the axis label change.
headline_long_r2 <- dplyr::bind_rows(
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      basin_drift = .data$weighted_r2_drift
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(
          dataset_bundle_id = .data$dataset_bundle_id,
          delta_auprc_treatment = .data$auprc_delta_mu,
          delta_auprc_warm = .data$auprc_delta_warm_mu_vs_vanilla
        ),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__vanilla_warm_from_pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      basin_drift = .data$weighted_r2_drift
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(
          dataset_bundle_id = .data$dataset_bundle_id,
          delta_auprc_treatment = .data$auprc_delta_pi,
          delta_auprc_warm = .data$auprc_delta_warm_pi_vs_vanilla
        ),
      by = "dataset_bundle_id"
    )
) %>%
  tidyr::pivot_longer(
    cols = c("delta_auprc_treatment", "delta_auprc_warm"),
    names_to = "delta_kind",
    values_to = "delta_auprc"
  ) %>%
  dplyr::mutate(
    delta_kind = factor(
      .data$delta_kind,
      levels = c("delta_auprc_treatment", "delta_auprc_warm"),
      labels = c(
        "DAUPRC(annotated cold) - vanilla",
        "DAUPRC(vanilla, warm refit) - vanilla"
      )
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

headline_max_drift_r2 <- max(0.05,
                              max(headline_long_r2$basin_drift, na.rm = TRUE))

p_headline_r2 <- ggplot(
  headline_long_r2,
  aes(x = .data$basin_drift, y = .data$delta_auprc, color = .data$chain)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_point(alpha = 0.5, size = 1.2) +
  scale_color_manual(
    values = c(mu = "#2b8cbe", pi = "#74c476"),
    labels = c(mu = "mu chain (signed mu_0)",
               pi = "pi chain (unsigned pi)")
  ) +
  facet_wrap(~ delta_kind, ncol = 2) +
  scale_x_continuous(limits = c(0, headline_max_drift_r2)) +
  labs(
    x = expression(
      paste("Basin drift  ",
            italic(D), "(vanilla, vanilla-warm)")
    ),
    y = expression(paste(Delta, "AUPRC vs cold vanilla")),
    color = "Annotation channel",
    title = "AUPRC change versus basin drift, by chain and arm",
    subtitle = "x in both facets: basin drift between cold vanilla and the warm-vanilla refit. Same data and axes as the chord-distance sibling; the metric is sin^2(theta) of the principal angle between matched fitted-y components."
  ) +
  theme(legend.position = "bottom")

ggsave(headline_scatter_r2_png, p_headline_r2,
       width = 12, height = 5.5, dpi = 150)
print(p_headline_r2)
```

# 10. Unpaired Count vs AUPRC Delta (User's Plot)

```{r unpaired-vs-delta, fig.width=12, fig.height=8}
# Build a long table: per (dataset, chain, step_label), report n_below_cutoff
# (drift) and the relevant AUPRC delta for that step.
step_drift_delta <- dplyr::bind_rows(
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step_label = "step 1: vanilla -> mu",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_delta_mu),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "mu__vanilla_warm_from_mu") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "mu",
      step_label = "step 2: mu -> vanilla_warm",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_round_trip_mu),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "vanilla__pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step_label = "step 1: vanilla -> pi",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_delta_pi),
      by = "dataset_bundle_id"
    ),
  drift_summary %>%
    dplyr::filter(.data$pair_id == "pi__vanilla_warm_from_pi") %>%
    dplyr::transmute(
      dataset_bundle_id = .data$dataset_bundle_id,
      chain = "pi",
      step_label = "step 2: pi -> vanilla_warm",
      n_below_cutoff = .data$n_below_cutoff,
      mean_corr = .data$mean_corr_active
    ) %>%
    dplyr::left_join(
      auprc_per_dataset %>%
        dplyr::transmute(.data$dataset_bundle_id,
                         auprc_delta = .data$auprc_round_trip_pi),
      by = "dataset_bundle_id"
    )
)

step_rho <- step_drift_delta %>%
  dplyr::group_by(.data$chain, .data$step_label) %>%
  dplyr::summarise(
    n = dplyr::n(),
    rho = suppressWarnings(stats::cor(
      .data$n_below_cutoff, .data$auprc_delta,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    label = sprintf("rho = %.3f (n=%d)", .data$rho, .data$n),
    .groups = "drop"
  )

p_unpaired <- ggplot(
  step_drift_delta,
  aes(x = .data$n_below_cutoff, y = .data$auprc_delta)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_jitter(width = 0.25, height = 0, alpha = 0.45, size = 1.1,
              aes(color = .data$mean_corr)) +
  geom_smooth(method = "loess", formula = y ~ x, se = FALSE,
              color = "black", linetype = "dashed", linewidth = 0.5) +
  geom_text(
    data = step_rho,
    aes(x = -Inf, y = Inf, label = .data$label),
    hjust = -0.05, vjust = 1.4, size = 3.3, inherit.aes = FALSE
  ) +
  facet_grid(chain ~ step_label) +
  scale_x_continuous(breaks = 0:10) +
  scale_color_viridis_c(option = "C") +
  labs(
    x = sprintf("# unpaired effects (assignment corr < %.2f)", drift_corr_cutoff),
    y = "AUPRC delta for this step",
    color = "Mean active\nassignment corr",
    title = sprintf(
      "User's hypothesis test: does # of unpaired effects (cutoff %.2f) predict AUPRC change at each step?",
      drift_corr_cutoff
    ),
    subtitle = "Step 1: cold->treatment (positive y => treatment wins). Step 2: treatment->vanilla_warm (positive y => warm refit IMPROVES on treatment, rare; negative => snaps back to vanilla)."
  )

ggsave(unpaired_vs_delta_png, p_unpaired, width = 12, height = 8, dpi = 150)
print(p_unpaired)
```

# 11. Per-Dataset PIP JSD

Loads the per-fit model-wide PIP sidecar (`pip_records_5arm.rds`, written by
`effect_matching_refit_export_5arm.Rmd`) and computes Jensen-Shannon
divergence between matched (vanilla, treatment) and (vanilla, vanilla_warm)
PIP vectors for each dataset. JSD definition mirrors
`R/run_model.R::js_distance`: a categorical JSD on raw model-wide PIP
vectors with an additive epsilon, summed (not averaged) across SNPs. See
the study plan note "JSD ... currently computed on raw model-wide PIPs".

This block runs before the mu-vs-pi figure so the figure can plot basin
drift and PIP JSD side-by-side.

```{r pip-jsd}
# Inline copy of test_susine:::js_distance (R/run_model.R:2021) so the Rmd
# does not depend on devtools::load_all() for the package.
js_distance <- function(p, q, eps = 1e-12) {
  p <- as.numeric(p); q <- as.numeric(q)
  p <- p + eps; q <- q + eps
  m <- 0.5 * (p + q)
  kl <- function(a, b) sum(a * log(a / b))
  0.5 * kl(p, m) + 0.5 * kl(q, m)
}

# Pairs we want for the master figure: (chain, pair_kind) -> (method_a, method_b)
pip_pair_specs <- tibble::tribble(
  ~chain, ~pair_kind,             ~method_a,         ~method_b,
  "mu",   "vanilla_to_treatment", baseline_method,   mu_method,
  "mu",   "treatment_to_warm",    mu_method,         warm_mu_method,
  "mu",   "vanilla_to_warm",      baseline_method,   warm_mu_method,
  "pi",   "vanilla_to_treatment", baseline_method,   pi_method,
  "pi",   "treatment_to_warm",    pi_method,         warm_pi_method,
  "pi",   "vanilla_to_warm",      baseline_method,   warm_pi_method
)

if (file.exists(pip_records_rds)) {
  pip_records <- readRDS(pip_records_rds) %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

  cat("PIP records rows:", nrow(pip_records), "\n")
  cat("Datasets in PIP records:",
      dplyr::n_distinct(pip_records$dataset_bundle_id), "\n")

  pip_by_dataset_method <- pip_records %>%
    dplyr::select(.data$dataset_bundle_id, .data$model_method, .data$pip)

  # Sanity: self-JSD (vanilla vs vanilla) should be exactly 0.
  self_check_ids <- head(unique(pip_records$dataset_bundle_id), 5L)
  self_jsd <- vapply(self_check_ids, function(ds_id) {
    v <- pip_by_dataset_method %>%
      dplyr::filter(.data$dataset_bundle_id == ds_id,
                    .data$model_method == baseline_method) %>%
      dplyr::pull(.data$pip) %>% .[[1]]
    js_distance(v, v)
  }, numeric(1))
  cat("Self-JSD (vanilla vs vanilla) on first 5 datasets:",
      paste(sprintf("%.3e", self_jsd), collapse = ", "), "\n")

  # Index PIPs by (dataset, method) for fast lookup.
  pip_lookup <- pip_records %>%
    dplyr::group_by(.data$dataset_bundle_id, .data$model_method) %>%
    dplyr::summarise(pip = list(.data$pip[[1L]]), .groups = "drop")
  pip_index <- setNames(
    pip_lookup$pip,
    paste(pip_lookup$dataset_bundle_id, pip_lookup$model_method, sep = "||")
  )

  datasets_with_all_pips <- pip_records %>%
    dplyr::count(.data$dataset_bundle_id, .data$model_method) %>%
    dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
    dplyr::filter(.data$n_methods == length(required_methods)) %>%
    dplyr::pull(.data$dataset_bundle_id)

  pip_jsd_rows <- list()
  for (ds_id in datasets_with_all_pips) {
    for (k in seq_len(nrow(pip_pair_specs))) {
      spec <- pip_pair_specs[k, ]
      pa <- pip_index[[paste(ds_id, spec$method_a, sep = "||")]]
      pb <- pip_index[[paste(ds_id, spec$method_b, sep = "||")]]
      if (is.null(pa) || is.null(pb) || length(pa) != length(pb)) next
      pip_jsd_rows[[length(pip_jsd_rows) + 1L]] <- tibble::tibble(
        dataset_bundle_id = ds_id,
        chain = spec$chain,
        pair_kind = spec$pair_kind,
        pip_jsd = js_distance(pa, pb)
      )
    }
  }
  pip_drift_summary <- dplyr::bind_rows(pip_jsd_rows)
} else {
  warning(
    "Missing ", pip_records_rds,
    "; PIP JSD panels will be empty. Re-run ",
    "effect_matching_refit_export_5arm.Rmd to generate the PIP sidecar."
  )
  pip_drift_summary <- tidyr::expand_grid(
    dataset_bundle_id = datasets_complete,
    pip_pair_specs %>% dplyr::select(.data$chain, .data$pair_kind)
  ) %>%
    dplyr::mutate(pip_jsd = NA_real_)
}
readr::write_csv(pip_drift_summary, pip_drift_summary_csv)
cat("PIP drift summary rows:", nrow(pip_drift_summary), "\n")

# Sanity: rho between basin drift and pip_jsd on matched (dataset, pair).
# Both metrics measure "how much did the model move", so a strongly
# positive rank correlation is expected. Carry both basin metrics
# (weighted_chord_drift and weighted_r2_drift) so downstream chunks can
# read either column.
basin_for_join <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "mu__vanilla_warm_from_mu",
    "vanilla__vanilla_warm_from_mu",
    "vanilla__pi", "pi__vanilla_warm_from_pi",
    "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    .data$dataset_bundle_id,
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::case_when(
      .data$pair_id %in% c("vanilla__mu", "vanilla__pi") ~
        "vanilla_to_treatment",
      .data$pair_id %in% c("mu__vanilla_warm_from_mu",
                           "pi__vanilla_warm_from_pi") ~
        "treatment_to_warm",
      TRUE ~ "vanilla_to_warm"
    ),
    weighted_chord_drift = .data$weighted_chord_drift,
    weighted_r2_drift = .data$weighted_r2_drift
  )

basin_pip_join <- basin_for_join %>%
  dplyr::inner_join(
    pip_drift_summary,
    by = c("dataset_bundle_id", "chain", "pair_kind")
  )

rho_check <- basin_pip_join %>%
  dplyr::group_by(.data$chain, .data$pair_kind) %>%
  dplyr::summarise(
    n = dplyr::n(),
    rho_chord = suppressWarnings(stats::cor(
      .data$weighted_chord_drift, .data$pip_jsd,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    rho_r2 = suppressWarnings(stats::cor(
      .data$weighted_r2_drift, .data$pip_jsd,
      method = "spearman", use = "pairwise.complete.obs"
    )),
    .groups = "drop"
  )
print(rho_check)
```

# 12. Mu vs Pi Drift Comparison (basin + PIP JSD, 4-panel)

Four panels per dataset, all with mu chain on x and pi chain on y. Left
two panels show the **basin drift** (fitted-y subspace displacement,
the *explanatory* axis): cold annotated fit and the warm-vanilla refit.
Right two panels show the **PIP JSD** between the same fits (the
*causal* axis: PIP-mass redistribution). Stacking basin and PIP
side-by-side lets the reader compare the two semantic axes for each
chain.

The chord-distance and 1 - r^2 versions of the basin metric are produced
in separate chunks below, so the reader can compare scales side by side.

```{r mu-vs-pi-drift, fig.width=18, fig.height=5}
mu_vs_pi_basin_chord <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "vanilla__pi",
    "vanilla__vanilla_warm_from_mu", "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    .data$dataset_bundle_id,
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::if_else(
      .data$pair_id %in% c("vanilla__mu", "vanilla__pi"),
      "vanilla -> annotated",
      "vanilla -> vanilla_warm"
    ),
    metric_value = .data$weighted_chord_drift
  ) %>%
  tidyr::pivot_wider(
    id_cols = c("dataset_bundle_id", "pair_kind"),
    names_from = "chain",
    values_from = "metric_value"
  )

mu_vs_pi_pip_jsd <- pip_drift_summary %>%
  dplyr::filter(.data$pair_kind %in% c("vanilla_to_treatment",
                                        "vanilla_to_warm")) %>%
  dplyr::transmute(
    .data$dataset_bundle_id,
    .data$chain,
    pair_kind = dplyr::if_else(
      .data$pair_kind == "vanilla_to_treatment",
      "vanilla -> annotated",
      "vanilla -> vanilla_warm"
    ),
    metric_value = .data$pip_jsd
  ) %>%
  tidyr::pivot_wider(
    id_cols = c("dataset_bundle_id", "pair_kind"),
    names_from = "chain",
    values_from = "metric_value"
  )

basin_chord_max <- max(0.05,
  max(mu_vs_pi_basin_chord$mu, mu_vs_pi_basin_chord$pi, na.rm = TRUE)
)
jsd_max <- max(0.05,
  max(mu_vs_pi_pip_jsd$mu, mu_vs_pi_pip_jsd$pi, na.rm = TRUE)
)

# Panel builder: scatter of mu (x) vs pi (y), with y=x dashed reference
# and a coord_fixed scale shared within each metric pair.
make_mupi_panel <- function(df, kind, max_val, axis_label, title_lab) {
  d <- df %>% dplyr::filter(.data$pair_kind == kind)
  ggplot(d, aes(x = .data$mu, y = .data$pi)) +
    geom_abline(slope = 1, intercept = 0,
                color = "grey45", linetype = "dashed") +
    geom_point(alpha = 0.5, size = 1.0) +
    coord_fixed(xlim = c(0, max_val), ylim = c(0, max_val)) +
    labs(
      x = paste0(axis_label, " (mu chain)"),
      y = paste0(axis_label, " (pi chain)"),
      title = title_lab
    ) +
    theme(plot.title = element_text(size = 11))
}

p_basin_chord_cold <- make_mupi_panel(
  mu_vs_pi_basin_chord, "vanilla -> annotated", basin_chord_max,
  "Basin chord drift",
  "Basin chord: vanilla -> annotated"
)
p_basin_chord_warm <- make_mupi_panel(
  mu_vs_pi_basin_chord, "vanilla -> vanilla_warm", basin_chord_max,
  "Basin chord drift",
  "Basin chord: vanilla -> vanilla_warm"
)
p_jsd_cold <- make_mupi_panel(
  mu_vs_pi_pip_jsd, "vanilla -> annotated", jsd_max,
  "PIP JSD",
  "PIP JSD: vanilla -> annotated"
)
p_jsd_warm <- make_mupi_panel(
  mu_vs_pi_pip_jsd, "vanilla -> vanilla_warm", jsd_max,
  "PIP JSD",
  "PIP JSD: vanilla -> vanilla_warm"
)

if (requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
  p_mupi <- p_basin_chord_cold + p_basin_chord_warm +
            p_jsd_cold + p_jsd_warm +
            plot_layout(nrow = 1) +
            plot_annotation(
              title = "Mu vs Pi: basin chord drift (explanatory) and PIP JSD (causal)",
              subtitle = "Each point is one of 600 datasets. Below y=x: pi moved less than mu on the metric in question. Left two: basin chord drift (fitted-y subspace displacement). Right two: PIP JSD drift (causal-PIP redistribution)."
            )
  ggsave(mu_vs_pi_drift_png, p_mupi, width = 18, height = 5, dpi = 150)
  print(p_mupi)
} else {
  warning("patchwork not installed; skipping 4-panel mu-vs-pi figure (chord). install.packages('patchwork').")
}
```

```{r mu-vs-pi-drift-r2, fig.width=18, fig.height=5}
# 4-panel mu-vs-pi using 1 - r^2 weighted basin distance instead of chord.
# The right two panels (PIP JSD) are identical to the chord version; only
# the left two basin panels change.
mu_vs_pi_basin_r2 <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "vanilla__pi",
    "vanilla__vanilla_warm_from_mu", "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    .data$dataset_bundle_id,
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::if_else(
      .data$pair_id %in% c("vanilla__mu", "vanilla__pi"),
      "vanilla -> annotated",
      "vanilla -> vanilla_warm"
    ),
    metric_value = .data$weighted_r2_drift
  ) %>%
  tidyr::pivot_wider(
    id_cols = c("dataset_bundle_id", "pair_kind"),
    names_from = "chain",
    values_from = "metric_value"
  )

basin_r2_max <- max(0.05,
  max(mu_vs_pi_basin_r2$mu, mu_vs_pi_basin_r2$pi, na.rm = TRUE)
)

p_basin_r2_cold <- make_mupi_panel(
  mu_vs_pi_basin_r2, "vanilla -> annotated", basin_r2_max,
  "Basin drift",
  "Basin drift: vanilla -> annotated"
)
p_basin_r2_warm <- make_mupi_panel(
  mu_vs_pi_basin_r2, "vanilla -> vanilla_warm", basin_r2_max,
  "Basin drift",
  "Basin drift: vanilla -> vanilla_warm"
)

if (requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
  p_mupi_r2 <- p_basin_r2_cold + p_basin_r2_warm +
               p_jsd_cold + p_jsd_warm +
               plot_layout(nrow = 1) +
               plot_annotation(
                 title = "Mu vs Pi: basin drift (explanatory) and PIP JSD (causal)",
                 subtitle = "Each point is one of 600 datasets. 1 - r^2 = sin^2(theta) = explanatory-subspace residual variance per fitted-y component, weighted by min var-explained."
               )
  ggsave(mu_vs_pi_drift_r2_png, p_mupi_r2, width = 18, height = 5, dpi = 150)
  print(p_mupi_r2)
} else {
  warning("patchwork not installed; skipping 4-panel mu-vs-pi figure (1 - r^2). install.packages('patchwork').")
}
```

# 13. Basin Drift vs PIP JSD Scatter

For each (chain, pair_kind) combination, plot per-dataset basin drift on
the x-axis against PIP JSD on the y-axis. Reveals the joint distribution
of the explanatory and causal axes: a tight diagonal => fitted-y movement
and PIP redistribution co-occur; off-diagonal points expose datasets
where one axis moves without the other. 2 x 2 grid: rows = chain
(mu top, pi bottom), columns = pair_kind (cold->annotated left,
cold->warm right). Produced separately for the chord and 1 - r^2
versions of the basin metric.

```{r basin-vs-pip-jsd-chord, fig.width=11, fig.height=10}
basin_jsd_data <- basin_pip_join %>%
  dplyr::filter(.data$pair_kind %in% c("vanilla_to_treatment",
                                        "vanilla_to_warm")) %>%
  dplyr::mutate(
    pair_kind = factor(
      .data$pair_kind,
      levels = c("vanilla_to_treatment", "vanilla_to_warm"),
      labels = c("vanilla -> annotated", "vanilla -> vanilla_warm")
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

# Per-axis caps shared across all 4 panels for fair comparison.
chord_drift_max <- max(0.05,
  max(basin_jsd_data$weighted_chord_drift, na.rm = TRUE)
)
r2_drift_max <- max(0.05,
  max(basin_jsd_data$weighted_r2_drift, na.rm = TRUE)
)
jsd_drift_max <- max(0.05,
  max(basin_jsd_data$pip_jsd, na.rm = TRUE)
)

chain_color <- c(mu = "#2b8cbe", pi = "#74c476")

make_basin_jsd_panel <- function(df, chain_val, pair_kind_val, drift_col,
                                  drift_label, drift_max, jsd_max) {
  d <- df %>% dplyr::filter(
    .data$chain == chain_val,
    .data$pair_kind == pair_kind_val
  )
  rho_val <- suppressWarnings(stats::cor(
    d[[drift_col]], d$pip_jsd,
    method = "spearman", use = "pairwise.complete.obs"
  ))
  rho_label <- sprintf("rho_s = %.3f (n=%d)",
                       rho_val, sum(is.finite(d[[drift_col]] * d$pip_jsd)))
  ggplot(d, aes(x = .data[[drift_col]], y = .data$pip_jsd)) +
    geom_point(alpha = 0.5, size = 1.0, color = chain_color[[chain_val]]) +
    geom_smooth(method = "loess", formula = y ~ x, se = FALSE,
                color = "grey25", linetype = "dashed", linewidth = 0.5) +
    annotate("text", x = -Inf, y = Inf, hjust = -0.05, vjust = 1.4,
             label = rho_label, size = 3.3) +
    scale_x_continuous(limits = c(0, drift_max)) +
    scale_y_continuous(limits = c(0, jsd_max)) +
    labs(
      x = drift_label,
      y = "PIP JSD",
      title = paste0(chain_val, " chain: ", pair_kind_val)
    ) +
    theme(plot.title = element_text(size = 11))
}

p_bj_mu_cold <- make_basin_jsd_panel(
  basin_jsd_data, "mu", "vanilla -> annotated",
  "weighted_chord_drift", "Basin chord drift",
  chord_drift_max, jsd_drift_max
)
p_bj_mu_warm <- make_basin_jsd_panel(
  basin_jsd_data, "mu", "vanilla -> vanilla_warm",
  "weighted_chord_drift", "Basin chord drift",
  chord_drift_max, jsd_drift_max
)
p_bj_pi_cold <- make_basin_jsd_panel(
  basin_jsd_data, "pi", "vanilla -> annotated",
  "weighted_chord_drift", "Basin chord drift",
  chord_drift_max, jsd_drift_max
)
p_bj_pi_warm <- make_basin_jsd_panel(
  basin_jsd_data, "pi", "vanilla -> vanilla_warm",
  "weighted_chord_drift", "Basin chord drift",
  chord_drift_max, jsd_drift_max
)

if (requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
  p_basin_jsd_chord <-
    (p_bj_mu_cold | p_bj_mu_warm) /
    (p_bj_pi_cold | p_bj_pi_warm) +
    plot_annotation(
      title = "Basin chord drift vs PIP JSD, per (chain, pair_kind)",
      subtitle = "Each point is one of 600 datasets. x: weighted Hungarian chord distance between matched fitted-y components. y: PIP JSD between model-wide PIP vectors. Dashed line: per-panel loess. Tight diagonal => explanatory and causal axes co-move."
    )
  ggsave(basin_vs_pip_jsd_chord_png, p_basin_jsd_chord,
         width = 11, height = 10, dpi = 150)
  print(p_basin_jsd_chord)
} else {
  warning("patchwork not installed; skipping basin-vs-jsd figure (chord). install.packages('patchwork').")
}
```

```{r basin-vs-pip-jsd-r2, fig.width=11, fig.height=10}
# 1 - r^2 sibling. Same data, x-axis is the principal-angle-based metric.
p_bj_mu_cold_r2 <- make_basin_jsd_panel(
  basin_jsd_data, "mu", "vanilla -> annotated",
  "weighted_r2_drift", "Basin drift",
  r2_drift_max, jsd_drift_max
)
p_bj_mu_warm_r2 <- make_basin_jsd_panel(
  basin_jsd_data, "mu", "vanilla -> vanilla_warm",
  "weighted_r2_drift", "Basin drift",
  r2_drift_max, jsd_drift_max
)
p_bj_pi_cold_r2 <- make_basin_jsd_panel(
  basin_jsd_data, "pi", "vanilla -> annotated",
  "weighted_r2_drift", "Basin drift",
  r2_drift_max, jsd_drift_max
)
p_bj_pi_warm_r2 <- make_basin_jsd_panel(
  basin_jsd_data, "pi", "vanilla -> vanilla_warm",
  "weighted_r2_drift", "Basin drift",
  r2_drift_max, jsd_drift_max
)

if (requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
  p_basin_jsd_r2 <-
    (p_bj_mu_cold_r2 | p_bj_mu_warm_r2) /
    (p_bj_pi_cold_r2 | p_bj_pi_warm_r2) +
    plot_annotation(
      title = "Basin drift vs PIP JSD, per (chain, pair_kind)",
      subtitle = "Same data as the chord version, x-axis swapped to 1 - r^2 = sin^2(theta) of the principal angle between matched 1D fitted-y subspaces."
    )
  ggsave(basin_vs_pip_jsd_r2_png, p_basin_jsd_r2,
         width = 11, height = 10, dpi = 150)
  print(p_basin_jsd_r2)
} else {
  warning("patchwork not installed; skipping basin-vs-jsd figure (1 - r^2). install.packages('patchwork').")
}
```

# 14. Master Drift, PIP, and AUPRC Summary Figure

Three metrics x three comparisons. mu and pi are overlaid in each subplot.

```{r master-drift-figure, fig.width=13, fig.height=8.5}
basin_long <- basin_for_join %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$chain, .data$pair_kind,
    metric = "basin_drift",
    value = .data$weighted_chord_drift
  )

pip_long <- pip_drift_summary %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$chain, .data$pair_kind,
    metric = "pip_drift",
    value = .data$pip_jsd
  )

auprc_delta_long <- dplyr::bind_rows(
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "mu",
      pair_kind = "vanilla_to_treatment",
      metric = "auprc_delta",
      value = .data$auprc_delta_mu
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "mu",
      pair_kind = "treatment_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_round_trip_mu
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "mu",
      pair_kind = "vanilla_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_delta_warm_mu_vs_vanilla
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "pi",
      pair_kind = "vanilla_to_treatment",
      metric = "auprc_delta",
      value = .data$auprc_delta_pi
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "pi",
      pair_kind = "treatment_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_round_trip_pi
    ),
  auprc_per_dataset %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = "pi",
      pair_kind = "vanilla_to_warm",
      metric = "auprc_delta",
      value = .data$auprc_delta_warm_pi_vs_vanilla
    )
)

combined <- dplyr::bind_rows(basin_long, pip_long, auprc_delta_long) %>%
  dplyr::mutate(
    metric = factor(
      .data$metric,
      levels = c("basin_drift", "pip_drift", "auprc_delta")
    ),
    pair_kind = factor(
      .data$pair_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

p_master <- ggplot(combined, aes(x = .data$value, fill = .data$chain)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 50, position = "identity", alpha = 0.55, color = NA
  ) +
  facet_grid(
    pair_kind ~ metric,
    scales = "free_x",
    labeller = labeller(
      metric = c(
        basin_drift = "Basin drift\n(weighted chord)",
        pip_drift   = "PIP drift\n(JSD)",
        auprc_delta = "AUPRC delta"
      ),
      pair_kind = c(
        vanilla_to_treatment = "vanilla -> annotated",
        treatment_to_warm    = "annotated -> vanilla_warm",
        vanilla_to_warm      = "vanilla -> vanilla_warm"
      )
    )
  ) +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  labs(
    x = NULL, y = "Density", fill = "Chain",
    title = "Basin, PIP, and AUPRC movement along the 5-arm refit chain",
    subtitle = "Basin drift uses Hungarian-matched fitted-y weighted chord distance; PIP drift is model-wide PIP-vector JSD."
  ) +
  theme(legend.position = "bottom")

ggsave(basin_pip_drift_png, p_master, width = 13, height = 8.5, dpi = 150)
print(p_master)
```

```{r master-drift-figure-r2, fig.width=13, fig.height=8.5}
# 1 - r^2 sibling of the master figure. Reuses pip_long and auprc_delta_long
# from the chord chunk; only the basin_long source column changes.
basin_long_r2 <- basin_for_join %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$chain, .data$pair_kind,
    metric = "basin_drift",
    value = .data$weighted_r2_drift
  )

combined_r2 <- dplyr::bind_rows(basin_long_r2, pip_long, auprc_delta_long) %>%
  dplyr::mutate(
    metric = factor(
      .data$metric,
      levels = c("basin_drift", "pip_drift", "auprc_delta")
    ),
    pair_kind = factor(
      .data$pair_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

p_master_r2 <- ggplot(combined_r2, aes(x = .data$value, fill = .data$chain)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 50, position = "identity", alpha = 0.55, color = NA
  ) +
  facet_grid(
    pair_kind ~ metric,
    scales = "free_x",
    labeller = labeller(
      metric = c(
        basin_drift = "Basin drift\n(weighted 1 - r^2)",
        pip_drift   = "PIP drift\n(JSD)",
        auprc_delta = "AUPRC delta"
      ),
      pair_kind = c(
        vanilla_to_treatment = "vanilla -> annotated",
        treatment_to_warm    = "annotated -> vanilla_warm",
        vanilla_to_warm      = "vanilla -> vanilla_warm"
      )
    )
  ) +
  scale_fill_manual(values = c(mu = "#2b8cbe", pi = "#74c476")) +
  labs(
    x = NULL, y = "Density", fill = "Chain",
    title = "Basin drift, PIP, and AUPRC movement along the 5-arm refit chain",
    subtitle = "Basin drift uses Hungarian-matched fitted-y weighted 1 - r^2 distance (sin^2 of the principal angle); PIP drift is model-wide PIP-vector JSD."
  ) +
  theme(legend.position = "bottom")

ggsave(basin_pip_drift_r2_png, p_master_r2,
       width = 13, height = 8.5, dpi = 150)
print(p_master_r2)
```

# 15. Pooled Effect-Quality Density Shifts

For each chain and comparison, the joint (effective k, accuracy ratio)
density of all effects is estimated by a 2D PVE-weighted KDE separately
for the source arm and the destination arm; the panel shade is the
destination density minus the source density (red = effects accumulating;
blue = effects vacating). Effects are pooled across all (dataset, fit)
pairs without splitting by Hungarian match correlation -- the per-effect
matched/unmatched split was previously available but turned out not to
add a useful axis to this plot.

The `match_status` column is still recorded in the CSV
(`effect_quality_pair_rows_5arm.csv`) for downstream slicing if needed;
the plotted PNGs no longer split on it.

```{r effect-quality-density-5arm, fig.width=13, fig.height=7.5}
effect_metric_rows <- records %>%
  dplyr::transmute(
    .data$dataset_bundle_id, .data$model_method, .data$effect,
    x_value = .data[[x_metric]],
    y_value = .data[[y_metric]],
    .data$var_y_explained_frac
  )

effect_quality_pair_specs <- tibble::tribble(
  ~pair_id,                        ~chain, ~comparison_kind,        ~source_method,   ~destination_method,
  "vanilla__mu",                   "mu",   "vanilla_to_treatment",  baseline_method,  mu_method,
  "mu__vanilla_warm_from_mu",      "mu",   "treatment_to_warm",     mu_method,        warm_mu_method,
  "vanilla__vanilla_warm_from_mu", "mu",   "vanilla_to_warm",       baseline_method,  warm_mu_method,
  "vanilla__pi",                   "pi",   "vanilla_to_treatment",  baseline_method,  pi_method,
  "pi__vanilla_warm_from_pi",      "pi",   "treatment_to_warm",     pi_method,        warm_pi_method,
  "vanilla__vanilla_warm_from_pi", "pi",   "vanilla_to_warm",       baseline_method,  warm_pi_method
)

build_effect_quality_rows <- function(spec) {
  edges <- drift_assignments %>%
    dplyr::filter(.data$pair_id == spec$pair_id) %>%
    dplyr::mutate(
      match_status = dplyr::if_else(
        .data$corr >= effect_match_corr_min, "matched", "unmatched"
      )
    )

  source_rows <- edges %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = spec$chain,
      comparison_kind = spec$comparison_kind,
      match_status = .data$match_status,
      arm_role = "source",
      model_method = spec$source_method,
      effect = .data$effect_a,
      assigned_corr = .data$corr
    )

  destination_rows <- edges %>%
    dplyr::transmute(
      .data$dataset_bundle_id,
      chain = spec$chain,
      comparison_kind = spec$comparison_kind,
      match_status = .data$match_status,
      arm_role = "destination",
      model_method = spec$destination_method,
      effect = .data$effect_b,
      assigned_corr = .data$corr
    )

  dplyr::bind_rows(source_rows, destination_rows) %>%
    dplyr::left_join(
      effect_metric_rows,
      by = c("dataset_bundle_id", "model_method", "effect")
    ) %>%
    dplyr::mutate(
      pve_weight = pmax(tidyr::replace_na(.data$var_y_explained_frac, 0), 0)
    )
}

effect_quality_pair_rows <- purrr::map_dfr(
  seq_len(nrow(effect_quality_pair_specs)),
  ~ build_effect_quality_rows(effect_quality_pair_specs[.x, ])
) %>%
  dplyr::mutate(
    chain = factor(.data$chain, levels = c("mu", "pi")),
    match_status = factor(.data$match_status,
                          levels = c("matched", "unmatched")),
    arm_role = factor(.data$arm_role,
                      levels = c("source", "destination")),
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    )
  )
readr::write_csv(effect_quality_pair_rows, effect_quality_pair_rows_csv)

x_limits_resolved <- resolve_limits(effect_quality_pair_rows$x_value, x_limits)
y_limits_resolved <- resolve_limits(effect_quality_pair_rows$y_value, y_limits)

weighted_kde2d_df <- function(df, xlim, ylim, n, h) {
  if (!requireNamespace("MASS", quietly = TRUE)) {
    stop("Package MASS is required for weighted KDE plots.")
  }
  group_values <- c("source", "destination")
  dens <- lapply(group_values, function(g) {
    sub <- df %>%
      dplyr::filter(.data$arm_role == g) %>%
      dplyr::filter(
        is.finite(.data$x_value),
        is.finite(.data$y_value),
        .data$x_value >= xlim[[1]],
        .data$x_value <= xlim[[2]],
        .data$y_value >= ylim[[1]],
        .data$y_value <= ylim[[2]]
      )

    if (nrow(sub) < 3L) {
      return(NULL)
    }

    w <- pmax(tidyr::replace_na(sub$pve_weight, 0), 0)
    if (sum(w) <= 0) {
      w <- rep(1 / nrow(sub), nrow(sub))
    } else {
      w <- w / sum(w)
    }

    x_grid <- seq(xlim[[1]], xlim[[2]], length.out = n)
    y_grid <- seq(ylim[[1]], ylim[[2]], length.out = n)
    kx <- outer(x_grid, sub$x_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[1]]) / h[[1]]
    })
    ky <- outer(y_grid, sub$y_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[2]]) / h[[2]]
    })
    ky <- sweep(ky, 2L, w, `*`)
    z <- kx %*% t(ky)

    list(group = g, x = x_grid, y = y_grid, z = z, n_rows = nrow(sub),
         pve_mass = sum(sub$pve_weight, na.rm = TRUE))
  })
  names(dens) <- group_values

  if (is.null(dens$source) || is.null(dens$destination)) {
    return(tibble::tibble())
  }

  tidyr::expand_grid(x = dens$source$x, y = dens$source$y) %>%
    dplyr::mutate(
      density_source = as.vector(t(dens$source$z)),
      density_destination = as.vector(t(dens$destination$z)),
      density_diff = .data$density_destination - .data$density_source,
      n_source = dens$source$n_rows,
      n_destination = dens$destination$n_rows,
      pve_source = dens$source$pve_mass,
      pve_destination = dens$destination$pve_mass
    )
}

density_shift_rows <- list()
for (ch in levels(effect_quality_pair_rows$chain)) {
  for (cmp in levels(effect_quality_pair_rows$comparison_kind)) {
    sub <- effect_quality_pair_rows %>%
      dplyr::filter(
        .data$chain == ch,
        .data$comparison_kind == cmp
      )
    dens <- weighted_kde2d_df(
      sub,
      xlim = x_limits_resolved,
      ylim = y_limits_resolved,
      n = density_grid_n,
      h = density_bandwidth
    )
    if (!nrow(dens)) {
      next
    }
    density_shift_rows[[length(density_shift_rows) + 1L]] <- dens %>%
      dplyr::mutate(
        chain = ch,
        comparison_kind = cmp
      )
  }
}

density_shift_df <- dplyr::bind_rows(density_shift_rows) %>%
  dplyr::mutate(
    chain = factor(.data$chain, levels = c("mu", "pi")),
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    )
  )

# Use the 95th-percentile |density_diff| from BOTH chains pooled to set a
# shared symmetric fill scale; that way the two chains' panels are
# directly comparable in magnitude, which the previous independent-scale
# version did not allow.
shared_abs_diff <- abs(density_shift_df$density_diff)
shared_abs_diff <- shared_abs_diff[is.finite(shared_abs_diff)]
shared_fill_lim <- stats::quantile(
  shared_abs_diff, probs = density_fill_quantile, na.rm = TRUE, names = FALSE
)
if (!is.finite(shared_fill_lim) || shared_fill_lim <= 0) {
  shared_fill_lim <- max(shared_abs_diff, na.rm = TRUE)
}
if (!is.finite(shared_fill_lim) || shared_fill_lim <= 0) {
  shared_fill_lim <- 1
}

plot_density_shift_chain <- function(density_df, chain_value, out_path,
                                     fill_lim = shared_fill_lim) {
  df <- density_df %>% dplyr::filter(.data$chain == chain_value)

  p <- ggplot(df, aes(x = .data$x, y = .data$y,
                      fill = .data$density_diff)) +
    geom_raster(interpolate = TRUE) +
    geom_contour(aes(z = .data$density_diff),
                 color = "grey20", alpha = 0.45, linewidth = 0.18) +
    facet_wrap(
      ~ comparison_kind,
      ncol = 3,
      labeller = labeller(
        comparison_kind = c(
          vanilla_to_treatment = "vanilla -> annotated",
          treatment_to_warm = "annotated -> vanilla_warm",
          vanilla_to_warm = "vanilla -> vanilla_warm"
        )
      )
    ) +
    scale_fill_gradient2(
      low = "#2b6cb0", mid = "white", high = "#c53030",
      midpoint = 0, limits = c(-fill_lim, fill_lim),
      oob = scales::squish
    ) +
    keff_x_scale(xlim = x_limits_resolved) +
    coord_cartesian(xlim = x_limits_resolved,
                    ylim = y_limits_resolved,
                    expand = FALSE) +
    labs(
      x = x_label, y = y_label,
      fill = "destination minus source\n(PVE-weighted density)",
      title = sprintf(
        "Per-effect quality shifts along the %s refit chain (effects pooled, weighted by PVE)",
        chain_value
      ),
      subtitle = "Red: effects accumulating at this (diffuseness, accuracy). Blue: effects vacating it. Top-left = peaked + accurate; bottom-right = diffuse + inaccurate. Shared fill scale across mu and pi panels."
    ) +
    theme(legend.position = "right")

  ggsave(out_path, p, width = 13, height = 4.4, dpi = 150)
  p
}

p_effect_density_mu <- plot_density_shift_chain(
  density_shift_df, "mu", effect_quality_density_mu_png
)
p_effect_density_pi <- plot_density_shift_chain(
  density_shift_df, "pi", effect_quality_density_pi_png
)
print(p_effect_density_mu)
print(p_effect_density_pi)
```

# 16. Pooled Precision-Recall Curves

Pooled precision-recall curves across all 600 datasets for the 5 arms
(vanilla, mu, pi, vanilla_warm_from_mu, vanilla_warm_from_pi). Inputs are
the model-wide PIP vectors in `pip_records_5arm.rds` and the per-dataset
causal indices, which are regenerated from `dataset_bundles.csv` via
`test_susine:::generate_data_for_bundle()` (no fits are re-run; the
hot loop is the deterministic phenotype simulation).

The shared 5-arm config used here is identical to the one in
`effect_matching_refit_export_5arm.Rmd` and lives in `_5arm_cfg.R`.

```{r pr-curves-setup, message=FALSE}
have_pip_records <- file.exists(pip_records_rds)
if (!have_pip_records) {
  warning(
    "Missing ", pip_records_rds,
    "; PR curves will be skipped. Re-run effect_matching_refit_export_5arm.Rmd ",
    "(or its PIP-export tail) to regenerate the sidecar."
  )
}

if (have_pip_records && !requireNamespace("PRROC", quietly = TRUE)) {
  warning("PRROC not installed; PR curves chunk skipped. install.packages('PRROC').")
  have_pip_records <- FALSE
}
```

```{r pr-curves-causal-idx, eval=have_pip_records}
# `here()` resolves to the test_susine package root (set by here::i_am()
# in section 1). Load the harness via devtools so we can call
# test_susine:::generate_data_for_bundle().
suppressMessages(devtools::load_all(here()))

# Optionally load the sister susine package the way refit_export_5arm.Rmd
# does, so the cfg builder's downstream call sites stay parallel.
susine_sister_path <- here("..", "susine")
if (file.exists(file.path(susine_sister_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(
    susine_sister_path, winslash = "/", mustWork = TRUE
  ))
  suppressMessages(devtools::load_all(susine_sister_path))
}

source(here("vignettes", "one_off_validations", "_5arm_cfg.R"))

if (file.exists(causal_truth_rds)) {
  causal_truth_lookup <- readRDS(causal_truth_rds)
  cat("Loaded cached causal_truth for ", length(causal_truth_lookup),
      " datasets from ", causal_truth_rds, "\n", sep = "")
} else {
  cfg_bundle <- build_5arm_cfg(
    here_root = here(),
    output_root = output_root
  )
  cfg_for_pr <- cfg_bundle$cfg

  pip_records_for_idx <- readRDS(pip_records_rds) %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

  datasets_with_pips <- sort(unique(pip_records_for_idx$dataset_bundle_id))
  cat("Regenerating causal truth (idx + top-", pr_top_k, " by |beta|) for ",
      length(datasets_with_pips),
      " datasets (no fits, just data simulation).\n", sep = "")

  bundle_rows_all <- cfg_for_pr$tables$dataset_bundles %>%
    tibble::as_tibble() %>%
    dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

  causal_truth_lookup <- list()
  started <- proc.time()[["elapsed"]]
  for (i in seq_along(datasets_with_pips)) {
    ds_id <- datasets_with_pips[[i]]
    bundle_row <- bundle_rows_all %>%
      dplyr::filter(.data$dataset_bundle_id == !!ds_id)
    if (nrow(bundle_row) != 1L) {
      stop("Expected one bundle row for ", ds_id, "; got ", nrow(bundle_row))
    }
    data_bundle <- test_susine:::generate_data_for_bundle(
      bundle_row, cfg_for_pr
    )
    causal <- as.integer(data_bundle$causal_idx)
    beta <- as.numeric(data_bundle$beta)

    # Rank causal positions by |beta| descending; take top pr_top_k.
    abs_beta_at_causals <- abs(beta[causal])
    ord <- order(abs_beta_at_causals, decreasing = TRUE)
    top_k <- min(pr_top_k, length(causal))
    top_idx <- causal[ord[seq_len(top_k)]]

    causal_truth_lookup[[ds_id]] <- list(
      causal_idx = causal,
      top_idx = as.integer(top_idx),
      beta_at_causal = beta[causal][ord]
    )

    if (i %% 50L == 0L || i == length(datasets_with_pips)) {
      elapsed <- proc.time()[["elapsed"]] - started
      eta <- elapsed / i * (length(datasets_with_pips) - i)
      cat(sprintf(
        "Regenerated %d/%d datasets | elapsed %s | ETA %s\n",
        i, length(datasets_with_pips),
        format_duration(elapsed), format_duration(eta)
      ))
    }
  }
  saveRDS(causal_truth_lookup, causal_truth_rds)
  cat("Saved causal truth lookup to ", causal_truth_rds, "\n", sep = "")
}

# Backward-compatible alias for any downstream code that expected the
# original causal_idx_lookup (full 23-tier truth).
causal_idx_lookup <- lapply(causal_truth_lookup, `[[`, "causal_idx")
```

```{r pr-curves-build, eval=have_pip_records, fig.width=9, fig.height=6}
pip_records_for_pr <- readRDS(pip_records_rds) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

# Pool PIPs and binary positive labels across all (dataset, variant)
# observations, separately for each of the 5 methods. Positives are the
# top pr_top_k causal positions per dataset, ranked by |beta|; the
# remaining causals are excluded from the negative set so that polygenic
# background variants are not counted as false positives. (This matches
# the baseline-sims AUPRC convention.)
pooled_rows <- list()
for (i in seq_len(nrow(pip_records_for_pr))) {
  ds_id <- pip_records_for_pr$dataset_bundle_id[[i]]
  method <- pip_records_for_pr$model_method[[i]]
  pip_vec <- pip_records_for_pr$pip[[i]]
  truth <- causal_truth_lookup[[ds_id]]
  if (is.null(truth)) {
    next
  }
  p_len <- length(pip_vec)
  top_idx <- truth$top_idx[truth$top_idx >= 1L & truth$top_idx <= p_len]
  causal <- truth$causal_idx[truth$causal_idx >= 1L &
                              truth$causal_idx <= p_len]
  is_top <- integer(p_len)
  is_top[top_idx] <- 1L
  is_other_causal <- integer(p_len)
  is_other_causal[setdiff(causal, top_idx)] <- 1L
  keep <- is_other_causal == 0L  # drop polygenic-background causals
  pooled_rows[[length(pooled_rows) + 1L]] <- tibble::tibble(
    dataset_bundle_id = ds_id,
    model_method = method,
    pip = as.numeric(pip_vec[keep]),
    is_causal = is_top[keep]
  )
}
pooled_df <- dplyr::bind_rows(pooled_rows)

cat("Pooled rows:", nrow(pooled_df), "\n")
cat("Per-method counts:\n")
print(pooled_df %>% dplyr::count(.data$model_method, .data$is_causal))

prevalence <- mean(pooled_df$is_causal)
cat(sprintf("Overall causal prevalence: %.4f\n", prevalence))

# One PR curve per method.
methods_in_order <- c(
  baseline_method, mu_method, warm_mu_method, pi_method, warm_pi_method
)
methods_in_order <- intersect(methods_in_order,
                              unique(pooled_df$model_method))

pr_curve_rows <- list()
pr_auc_rows <- list()
for (m in methods_in_order) {
  sub <- pooled_df %>% dplyr::filter(.data$model_method == m)
  if (!nrow(sub) || sum(sub$is_causal) == 0L) next
  pr <- PRROC::pr.curve(
    scores.class0 = sub$pip,
    weights.class0 = sub$is_causal,
    curve = TRUE,
    sorted = FALSE,
    rand.compute = FALSE
  )
  curve_mat <- pr$curve  # columns: recall, precision, threshold
  pr_curve_rows[[length(pr_curve_rows) + 1L]] <- tibble::tibble(
    model_method = m,
    recall = curve_mat[, 1L],
    precision = curve_mat[, 2L],
    threshold = curve_mat[, 3L]
  )
  pr_auc_rows[[length(pr_auc_rows) + 1L]] <- tibble::tibble(
    model_method = m,
    auprc_integral = as.numeric(pr$auc.integral),
    auprc_dg = as.numeric(pr$auc.davis.goadrich)
  )
}
pr_curve_df <- dplyr::bind_rows(pr_curve_rows) %>%
  dplyr::mutate(model_method = factor(.data$model_method,
                                      levels = methods_in_order))
pr_auc_df <- dplyr::bind_rows(pr_auc_rows) %>%
  dplyr::mutate(model_method = factor(.data$model_method,
                                      levels = methods_in_order))

readr::write_csv(pr_curve_df, pr_curves_pooled_csv)
cat("Wrote PR curve rows:", nrow(pr_curve_df), "->", pr_curves_pooled_csv, "\n")
print(pr_auc_df)

method_legend_labels <- vapply(
  levels(pr_curve_df$model_method),
  function(m) {
    auprc_val <- pr_auc_df$auprc_integral[
      match(m, as.character(pr_auc_df$model_method))
    ]
    sprintf("%s (AUPRC = %.3f)",
            method_labels[[m]] %||% m,
            auprc_val)
  },
  character(1)
)

p_pr <- ggplot(pr_curve_df,
               aes(x = .data$recall, y = .data$precision,
                   color = .data$model_method)) +
  geom_hline(yintercept = prevalence, color = "grey60",
             linetype = "dashed", linewidth = 0.4) +
  geom_step(direction = "vh", linewidth = 0.7) +
  scale_color_manual(
    values = method_color[levels(pr_curve_df$model_method)],
    labels = method_legend_labels
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    x = "Recall",
    y = "Precision",
    color = "Method (pooled AUPRC)",
    title = sprintf(
      "Pooled precision-recall across %d datasets at the AlphaGenome calibration regime",
      length(unique(pooled_df$dataset_bundle_id))
    ),
    subtitle = sprintf(
      "Positives = top %d causal SNPs by |beta| per dataset (sparse + oligogenic tiers); polygenic tier excluded from negatives. Prevalence = %.4f. Dashed line = random-classifier baseline.",
      pr_top_k, prevalence
    )
  ) +
  theme(legend.position = "right")

ggsave(pr_curves_pooled_png, p_pr, width = 9, height = 6, dpi = 150)
print(p_pr)
```

# 17. Combined Paper Figure

A single full-width figure that walks the reader from the headline
performance result (PR curves) down through the drift mechanics and the
per-effect mechanism. Component plots are stripped of their per-chunk
titles/subtitles so the paper caption can carry the framing.

```{r paper-figure, eval=have_pip_records, fig.width=16, fig.height=22}
# Headline 4-panel paper figure. Locked in to the 1 - r^2 basin metric
# (chord version retained as diagnostic in §12/§14/§17 sibling chunks).
#
#   A. Pooled PR curves (5 arms, top-K causals)
#   B. Mu vs pi scatters with cross-axis coloring (4 sub-panels):
#        B(i):  basin drift cold->ann,  colored by mu PIP JSD (cold->ann)
#        B(ii): basin drift cold->warm, colored by mu PIP JSD (cold->warm)
#        B(iii): PIP JSD cold->ann,     colored by mu basin drift (cold->ann)
#        B(iv): PIP JSD cold->warm,     colored by mu basin drift (cold->warm)
#      Coloring is on a linear viridis scale (no log). The four sub-panels
#      are grouped as two pairs (basin row | JSD row), each pair sharing
#      a single horizontal colorbar legend below the pair so the
#      individual scatters can use their full width.
#   C. Boxplots of dAUPRC by 3-bucket basin-drift cut (<0.05 / 0.05-0.10
#      / >=0.10), faceted chain x pair_kind. Per-cell n annotated above
#      each box. Boxplots (not densities) chosen because the high-drift
#      bucket has only ~10-25 datasets per cell.
#   D. Per-effect quality shifts on the (K_l, A_l) plane, 2 chains x 2
#      contrasts (cold->ann, cold->warm). Drops the annotated->warm
#      intermediate. Mu and pi shown side-by-side so the reader can
#      visually confirm both chains do the same per-effect work.

if (!requireNamespace("patchwork", quietly = TRUE)) {
  warning(
    "patchwork not installed; combined paper figure skipped. ",
    "install.packages('patchwork')."
  )
} else {
  library(patchwork)

  # Greek-symbol shortcuts to keep label code tight.
  expr_basin_x <- expression(paste("Basin drift (", mu, " chain)"))
  expr_basin_y <- expression(paste("Basin drift (", pi, " chain)"))
  expr_jsd_x   <- expression(paste("PIP JSD (", mu, " chain)"))
  expr_jsd_y   <- expression(paste("PIP JSD (", pi, " chain)"))
  expr_dauprc  <- expression(paste(Delta, "AUPRC vs cold vanilla"))

  # Common labellers (chain rows, pair_kind cols) used in C and D.
  chain_labeller <- c(
    mu = "μ chain (signed prior mean, μ₀)",
    pi = "π chain (unsigned inclusion log-odds)"
  )
  pair_kind_labeller <- c(
    "cold -> annotated"    = "Cold annotated fit (treatment)",
    "cold -> vanilla_warm" = "Warm-vanilla refit (durability check)"
  )

  # ============================================================
  # PANEL A. PR curves (existing p_pr)
  # ============================================================
  strip_chrome <- function(p) {
    p + labs(title = NULL, subtitle = NULL) +
      theme(plot.title = element_blank(),
            plot.subtitle = element_blank())
  }
  panel_A <- strip_chrome(p_pr) +
    labs(title = paste0(
      "A. Pooled precision-recall on top-", pr_top_k,
      " causal SNPs (5 arms, 600 datasets)"
    ))

  # ============================================================
  # PANEL B. Mu vs pi scatters with cross-axis coloring
  # ============================================================
  basin_long_pf <- drift_summary %>%
    dplyr::filter(.data$pair_id %in% c(
      "vanilla__mu", "vanilla__pi",
      "vanilla__vanilla_warm_from_mu", "vanilla__vanilla_warm_from_pi"
    )) %>%
    dplyr::transmute(
      dataset_bundle_id = as.character(.data$dataset_bundle_id),
      chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
      pair_kind = dplyr::if_else(
        .data$pair_id %in% c("vanilla__mu", "vanilla__pi"),
        "cold_ann", "cold_warm"
      ),
      basin_drift = .data$weighted_r2_drift
    )

  jsd_long_pf <- pip_drift_summary %>%
    dplyr::filter(.data$pair_kind %in% c("vanilla_to_treatment",
                                          "vanilla_to_warm")) %>%
    dplyr::transmute(
      dataset_bundle_id = as.character(.data$dataset_bundle_id),
      chain = .data$chain,
      pair_kind = dplyr::if_else(
        .data$pair_kind == "vanilla_to_treatment", "cold_ann", "cold_warm"
      ),
      pip_jsd = .data$pip_jsd
    )

  basin_wide_pf <- basin_long_pf %>%
    tidyr::pivot_wider(names_from = chain, values_from = basin_drift,
                       names_prefix = "basin_")

  jsd_wide_pf <- jsd_long_pf %>%
    tidyr::pivot_wider(names_from = chain, values_from = pip_jsd,
                       names_prefix = "jsd_")

  joined_pf <- basin_wide_pf %>%
    dplyr::inner_join(jsd_wide_pf, by = c("dataset_bundle_id", "pair_kind"))

  # Axis maxes (square plots, both chains share scale).
  basin_max_pf <- max(0.05,
    max(joined_pf$basin_mu, joined_pf$basin_pi, na.rm = TRUE))
  jsd_max_pf <- max(0.05,
    max(joined_pf$jsd_mu, joined_pf$jsd_pi, na.rm = TRUE))

  # Color limits (must match within each pair so the collected legend is
  # consistent). Color uses mu-side values across BOTH cold->ann and
  # cold->warm subsets, so we take the max across the full joined table.
  basin_color_max <- max(joined_pf$basin_mu, na.rm = TRUE)
  jsd_color_max   <- max(joined_pf$jsd_mu,   na.rm = TRUE)

  # Sub-panel builder. Color is linear viridis (no log). Endpoint labels
  # carry the intuition ("no change" -> "most change") so the reader
  # doesn't need to read raw values off the legend bar.
  #
  # coord_cartesian + theme(aspect.ratio = 1) decouples the panel's
  # on-screen square size from the data range, so basin (0-0.8) and
  # JSD (0-1.7) panels render at the same physical size.
  make_b_panel <- function(df, x_col, y_col, color_col,
                           x_label, y_label, color_title, title_lab,
                           color_lim_max, max_val,
                           show_y = TRUE) {
    ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]],
                   color = .data[[color_col]])) +
      geom_abline(slope = 1, intercept = 0,
                  color = "grey50", linetype = "dashed", linewidth = 0.35) +
      geom_point(alpha = 0.58, size = 1.35, stroke = 0) +
      scale_color_viridis_c(
        option = "C",
        limits = c(0, color_lim_max),
        breaks = c(0, color_lim_max),
        labels = c("no change", "most change"),
        guide = guide_colorbar(
          title.position = "top",
          title.hjust = 0.5,
          barwidth  = grid::unit(5.3, "cm"),
          barheight = grid::unit(0.35, "cm"),
          direction = "horizontal"
        )
      ) +
      scale_x_continuous(expand = expansion(mult = c(0.015, 0.035))) +
      scale_y_continuous(expand = expansion(mult = c(0.015, 0.035))) +
      coord_cartesian(xlim = c(0, max_val), ylim = c(0, max_val),
                      expand = FALSE) +
      labs(x = x_label, y = y_label, color = color_title,
           title = title_lab) +
      theme(
        plot.title = element_text(size = 10.5, face = "bold",
                                  margin = margin(b = 4)),
        axis.title.x = element_text(size = 10, margin = margin(t = 5)),
        axis.title.y = element_text(size = 10, margin = margin(r = 5)),
        axis.text = element_text(size = 8.5, color = "grey25"),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.3),
        panel.grid.minor = element_line(color = "grey95", linewidth = 0.2),
        legend.margin = margin(t = -2, b = -2),
        plot.margin = margin(3, 6, 3, 6),
        aspect.ratio = 1
      ) +
      {
        if (show_y) {
          theme()
        } else {
          theme(axis.title.y = element_blank(),
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank())
        }
      }
  }

  joined_pf_cold_ann  <- joined_pf %>% dplyr::filter(pair_kind == "cold_ann")
  joined_pf_cold_warm <- joined_pf %>% dplyr::filter(pair_kind == "cold_warm")

  # Color title is shared across each pair via plot_layout(guides="collect").
  basin_color_title <- expression(paste("PIP JSD (", mu, ")"))
  jsd_color_title   <- expression(paste("Basin drift (", mu, ")"))

  pB1 <- make_b_panel(
    joined_pf_cold_ann, "basin_mu", "basin_pi", "jsd_mu",
    expr_basin_x, expr_basin_y, basin_color_title,
    "B(i). Cold -> annotated",
    color_lim_max = jsd_color_max, max_val = basin_max_pf,
    show_y = TRUE
  )
  pB2 <- make_b_panel(
    joined_pf_cold_warm, "basin_mu", "basin_pi", "jsd_mu",
    expr_basin_x, expr_basin_y, basin_color_title,
    "B(ii). Cold -> warm vanilla",
    color_lim_max = jsd_color_max, max_val = basin_max_pf,
    show_y = FALSE
  )
  pB3 <- make_b_panel(
    joined_pf_cold_ann, "jsd_mu", "jsd_pi", "basin_mu",
    expr_jsd_x, expr_jsd_y, jsd_color_title,
    "B(iii). Cold -> annotated",
    color_lim_max = basin_color_max, max_val = jsd_max_pf,
    show_y = TRUE
  )
  pB4 <- make_b_panel(
    joined_pf_cold_warm, "jsd_mu", "jsd_pi", "basin_mu",
    expr_jsd_x, expr_jsd_y, jsd_color_title,
    "B(iv). Cold -> warm vanilla",
    color_lim_max = basin_color_max, max_val = jsd_max_pf,
    show_y = FALSE
  )

  # A 2 x 2 two-pair layout gives the four scatters much more physical area
  # than the old 1 x 4 strip. Within each row-pair, the left plot carries
  # the shared y-axis label and the right plot drops duplicate y text.
  basin_pair_B <- (pB1 | pB2) +
    plot_layout(guides = "collect", widths = c(1, 1)) +
    plot_annotation(title = "B. Basin drift: mu vs pi") &
    theme(
      legend.position  = "bottom",
      legend.direction = "horizontal",
      legend.box       = "horizontal",
      legend.title     = element_text(size = 9.5),
      legend.text      = element_text(size = 8.5),
      plot.title       = element_text(face = "bold", size = 11,
                                      margin = margin(b = 5))
    )

  jsd_pair_B <- (pB3 | pB4) +
    plot_layout(guides = "collect", widths = c(1, 1)) +
    plot_annotation(title = "PIP JSD: mu vs pi") &
    theme(
      legend.position  = "bottom",
      legend.direction = "horizontal",
      legend.box       = "horizontal",
      legend.title     = element_text(size = 9.5),
      legend.text      = element_text(size = 8.5),
      plot.title       = element_text(face = "bold", size = 11,
                                      margin = margin(b = 5))
    )

  panel_B <- (basin_pair_B / jsd_pair_B) +
    plot_layout(heights = c(1, 1))

  # ============================================================
  # PANEL C. dAUPRC boxplots by 3-bucket basin-drift cut
  # ============================================================
  auprc_lookup_pf <- auprc_per_dataset %>%
    dplyr::transmute(
      dataset_bundle_id = as.character(.data$dataset_bundle_id),
      delta_mu_treat = .data$auprc_delta_mu,
      delta_mu_warm  = .data$auprc_delta_warm_mu_vs_vanilla,
      delta_pi_treat = .data$auprc_delta_pi,
      delta_pi_warm  = .data$auprc_delta_warm_pi_vs_vanilla
    )

  panel_C_long <- basin_long_pf %>%
    dplyr::left_join(auprc_lookup_pf, by = "dataset_bundle_id") %>%
    dplyr::mutate(
      auprc_delta = dplyr::case_when(
        chain == "mu" & pair_kind == "cold_ann"  ~ delta_mu_treat,
        chain == "mu" & pair_kind == "cold_warm" ~ delta_mu_warm,
        chain == "pi" & pair_kind == "cold_ann"  ~ delta_pi_treat,
        chain == "pi" & pair_kind == "cold_warm" ~ delta_pi_warm
      ),
      drift_bucket = cut(
        basin_drift,
        breaks = c(-Inf, 0.05, 0.10, Inf),
        labels = c("low (<0.05)", "mid (0.05-0.10)", "high (>=0.10)"),
        right = FALSE
      ),
      chain = factor(chain, levels = c("mu", "pi")),
      pair_kind = factor(
        pair_kind,
        levels = c("cold_ann", "cold_warm"),
        labels = c("cold -> annotated", "cold -> vanilla_warm")
      )
    ) %>%
    dplyr::filter(is.finite(.data$auprc_delta), !is.na(.data$drift_bucket))

  # Per-cell counts for n=X annotations above each box.
  bucket_counts <- panel_C_long %>%
    dplyr::count(.data$chain, .data$pair_kind, .data$drift_bucket)
  y_top <- max(panel_C_long$auprc_delta, na.rm = TRUE) * 1.08

  panel_C <- ggplot(
    panel_C_long,
    aes(x = .data$drift_bucket,
        y = .data$auprc_delta,
        fill = .data$drift_bucket)
  ) +
    geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
    geom_boxplot(alpha = 0.65, outlier.size = 0.6, width = 0.65) +
    geom_text(
      data = bucket_counts,
      aes(x = .data$drift_bucket, y = y_top,
          label = sprintf("n=%d", .data$n)),
      inherit.aes = FALSE, size = 3.0, color = "grey25"
    ) +
    facet_grid(
      chain ~ pair_kind,
      labeller = labeller(chain = chain_labeller,
                          pair_kind = pair_kind_labeller)
    ) +
    scale_fill_manual(values = c(
      "low (<0.05)"     = "#878787",
      "mid (0.05-0.10)" = "#fd8d3c",
      "high (>=0.10)"   = "#a50f15"
    )) +
    labs(
      x = "Basin drift bucket",
      y = expr_dauprc,
      fill = "Basin drift",
      title = "C. ΔAUPRC by basin-drift bucket (3-bucket cut at 0.05 / 0.10)"
    ) +
    theme(plot.title = element_text(size = 11),
          legend.position = "right",
          axis.text.x = element_text(angle = 0, hjust = 0.5))

  # ============================================================
  # PANEL D. Per-effect (K_l, A_l) density shifts: 2 x 2
  #          rows = chain (mu, pi); cols = (cold->ann, cold->warm)
  # ============================================================
  panel_D_df <- density_shift_df %>%
    dplyr::filter(.data$comparison_kind %in% c(
      "vanilla_to_treatment", "vanilla_to_warm"
    )) %>%
    dplyr::mutate(
      comparison_kind = factor(
        .data$comparison_kind,
        levels = c("vanilla_to_treatment", "vanilla_to_warm"),
        labels = c("cold -> annotated", "cold -> vanilla_warm")
      ),
      chain = factor(.data$chain, levels = c("mu", "pi"))
    )

  panel_D <- ggplot(panel_D_df,
                    aes(x = .data$x, y = .data$y,
                        fill = .data$density_diff)) +
    geom_raster(interpolate = TRUE) +
    geom_contour(aes(z = .data$density_diff),
                 color = "grey20", alpha = 0.45, linewidth = 0.18) +
    facet_grid(
      chain ~ comparison_kind,
      labeller = labeller(chain = chain_labeller,
                          comparison_kind = pair_kind_labeller)
    ) +
    scale_fill_gradient2(
      low = "#2b6cb0", mid = "white", high = "#c53030",
      midpoint = 0, limits = c(-shared_fill_lim, shared_fill_lim),
      oob = scales::squish
    ) +
    keff_x_scale(xlim = x_limits_resolved) +
    coord_cartesian(xlim = x_limits_resolved,
                    ylim = y_limits_resolved,
                    expand = FALSE) +
    labs(
      x = x_label, y = y_label,
      fill = "destination - source\n(PVE-weighted density)",
      title = "D. Per-effect quality shifts on the (Kℓ, Aℓ) plane"
    ) +
    theme(plot.title = element_text(size = 11),
          legend.position = "right")

  # ============================================================
  # Assemble
  # ============================================================
  combined_fig <- panel_A / panel_B / panel_C / panel_D +
    plot_layout(heights = c(0.58, 2.15, 1.0, 1.1)) &
    theme(plot.title = element_text(face = "bold", size = 12))

  ggsave(
    paper_figure_png, combined_fig,
    width = 16.5, height = 22, dpi = 160, limitsize = FALSE
  )
  cat("Wrote combined paper figure to ", paper_figure_png, "\n", sep = "")
  print(combined_fig)
}
```

# 18. Outputs

```{r output-index}
outputs <- tibble::tibble(
  artifact = c(
    "per-effect drift assignments",
    "per-(dataset, pair) drift summary",
    "per-dataset AUPRC trajectory",
    "drift profile histograms",
    "chain AUPRC trajectory",
    "headline scatter (chord, diagnostic)",
    "headline scatter (basin drift)",
    "unpaired-count vs AUPRC delta scatter",
    "mu vs pi drift comparison: chord + JSD (diagnostic)",
    "mu vs pi drift comparison: basin drift + JSD (4-panel diagnostic)",
    "basin chord drift vs PIP JSD scatter (diagnostic)",
    "basin drift vs PIP JSD scatter (2x2 by chain x pair_kind)",
    "per-dataset PIP JSD summary",
    "master basin + PIP + AUPRC drift figure (chord, diagnostic)",
    "master basin + PIP + AUPRC drift figure (basin drift)",
    "effect-quality density pooled (mu)",
    "effect-quality density pooled (pi)",
    "effect-quality pair rows",
    "pooled PR curves top-K by |beta| (5 arms)",
    "pooled PR curve rows (csv)",
    "regenerated causal truth cache (idx + top-K)",
    "combined paper figure (4-panel headline, basin drift)"
  ),
  path = c(
    drift_assignments_csv,
    drift_summary_csv,
    auprc_per_dataset_csv,
    drift_profile_hist_png,
    chain_trajectory_png,
    headline_scatter_png,
    headline_scatter_r2_png,
    unpaired_vs_delta_png,
    mu_vs_pi_drift_png,
    mu_vs_pi_drift_r2_png,
    basin_vs_pip_jsd_chord_png,
    basin_vs_pip_jsd_r2_png,
    pip_drift_summary_csv,
    basin_pip_drift_png,
    basin_pip_drift_r2_png,
    effect_quality_density_mu_png,
    effect_quality_density_pi_png,
    effect_quality_pair_rows_csv,
    pr_curves_pooled_png,
    pr_curves_pooled_csv,
    causal_truth_rds,
    paper_figure_png
  )
)
kable(outputs)
```
