---
title: "Effect Matching Diagnostic Plots"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

One-off plotting workbook for the fitted-y effect matching study. It
uses the refit export plus the real match candidates from threshold
calibration to visualize joint effect diagnostics before and after
matching.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_diagnostic_plots.Rmd")
library(here)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

# 1. User Parameters

```{r user-parameters}
output_root <- here("output")

study_name <- "baseline_sims_standalone_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
records_rds <- file.path(study_dir, "effect_refit_records.rds")
real_candidates_csv <- file.path(study_dir, "real_match_candidates_same_X.csv")
winner_csv <- file.path(study_dir, "winner_threshold_same_X.csv")

plot_dir <- file.path(study_dir, "diagnostic_plots")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Threshold source.
#
# Manual thresholds are locked to a Pareto-frontier point from the same-X
# null calibration (see effect_matching_threshold_calibration_same_X.Rmd):
# corr_min = 0.99, margin_min = 0, var_min = 0. At this point the calibration
# returned 1934 real matches with null rate 2.22% per dataset pair, giving
# implied FDR ~0.7%. The auto-winner at the 5% null cutoff was looser
# (corr=0.95) with more matches but FDR ~1.4%; we trade ~10% of matches for
# ~2x cleaner pairs because false pairs would add isotropic noise to the
# joint-density plots and we already have plenty of real pairs.
use_winner_threshold <- FALSE
manual_corr_min <- 0.99
manual_margin_min <- 0
manual_var_min <- 0

# Joint diagnostic axes. The diffuseness axis stays in entropy units (cleanly
# behaved for KDE / arrows), but its tick labels are rendered as the
# corresponding effective k = exp(entropy), which is more interpretable.
x_metric <- "effect_pip_entropy_core95"
y_metric <- "accuracy_ratio"
x_label <- "Core-95 effective k (= exp(entropy))"
y_label <- "Accuracy ratio"
x_keff_breaks <- c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000)
delta_keff_ratio_breaks <- c(1/16, 1/8, 1/4, 1/2, 1, 2, 4, 8, 16)
delta_keff_ratio_labels <- c("1/16", "1/8", "1/4", "1/2", "1", "2", "4", "8", "16")

# Axis limits. Set either to NULL for automatic limits from the data.
x_limits <- NULL
y_limits <- c(0, 1)

# 2D kernel density controls.
#  - density_bandwidth: half-widths for kde2d. Bumped relative to the original
#    (0.30, 0.06) -> (0.45, 0.10) for less granular speckling on the dense
#    regions while keeping enough resolution to see local mu-vs-vanilla shifts.
#  - density_fill_quantile: instead of letting the color scale span the full
#    range of |density_diff| (which lets one extreme outlier dominate and washes
#    everything else to white), clip the scale at this quantile of |density_diff|.
#    Anything beyond is squished to the saturation extremes.
density_grid_n <- 160L
density_bandwidth <- c(0.45, 0.10)
density_trim_quantile <- c(0.005, 0.995)
density_fill_limit <- NULL
density_fill_quantile <- 0.95

# Univariate histogram controls.
hist_bins <- 60L

# Matched-delta 2D density contour breaks. Default geom_density_2d_filled uses
# 10 evenly-spaced bins on [0, 1], which makes the lowest band dominate by area.
# These uneven breaks give more resolution in the low-density tail so the
# off-peak structure becomes visible without changing the underlying KDE.
delta_density_contour_breaks <- c(
  0, 0.005, 0.015, 0.04, 0.08, 0.15, 0.25, 0.4, 0.6, 0.8, 1.0
)

# Arrow/flow plot controls for matched effects.
max_arrows_to_plot <- Inf
arrow_alpha <- 0.40
arrow_linewidth <- 0.25
arrow_head_length_inches <- 0.055
label_top_discrepancies <- 25L

# Output filenames.
scatter_all_png <- file.path(plot_dir, "effect_diagnostic_scatter_all_effects.png")
scatter_matched_unmatched_png <- file.path(plot_dir, "effect_diagnostic_scatter_matched_unmatched.png")
density_all_png <- file.path(plot_dir, "effect_diagnostic_density_diff_all_effects.png")
density_matched_png <- file.path(plot_dir, "effect_diagnostic_density_diff_matched_effects.png")
density_unmatched_png <- file.path(plot_dir, "effect_diagnostic_density_diff_unmatched_effects.png")
flow_png <- file.path(plot_dir, "effect_diagnostic_matched_flow.png")
delta_density_png <- file.path(plot_dir, "effect_diagnostic_matched_delta_density.png")
hist_diffuseness_png <- file.path(plot_dir, "effect_diagnostic_univariate_hist_diffuseness.png")
hist_accuracy_png <- file.path(plot_dir, "effect_diagnostic_univariate_hist_accuracy.png")
auprc_delta_hist_png <- file.path(plot_dir, "effect_diagnostic_auprc_delta_hist.png")
match_count_hist_png <- file.path(plot_dir, "effect_diagnostic_match_count_hist.png")
auprc_match_scatter_png <- file.path(plot_dir, "effect_diagnostic_auprc_vs_match_scatter.png")
scatter_stratified_png <- file.path(plot_dir, "effect_diagnostic_auprc_vs_match_stratified.png")
confound_check_png <- file.path(plot_dir, "effect_diagnostic_vanilla_auprc_vs_match.png")
baseline_vs_susine_png <- file.path(plot_dir, "effect_diagnostic_baseline_auprc_vs_susine_auprc.png")
matched_pairs_csv <- file.path(plot_dir, "accepted_matched_pairs_for_plots.csv")
matched_delta_csv <- file.path(plot_dir, "accepted_matched_pair_deltas_for_plots.csv")
per_dataset_summary_csv <- file.path(plot_dir, "per_dataset_auprc_match_summary.csv")

cat("Study dir:", study_dir, "\n")
cat("Plot dir:", plot_dir, "\n")
```

# 2. Helpers

```{r helpers}
metric_label <- function(metric) {
  labels <- c(
    effect_pip_entropy = "PIP entropy",
    effect_pip_entropy_core95 = "Core-95 PIP entropy",
    effect_k_eff_signal = "k-eff signal",
    effect_k_eff_signal_core95 = "Core-95 k-eff signal",
    accuracy_ratio = "Accuracy ratio",
    size = "CS size",
    purity = "CS purity",
    coverage = "CS coverage",
    var_y_explained_pct = "% var(y) explained",
    var_y_explained_frac = "Fraction var(y) explained"
  )
  labels[[metric]] %||% metric
}

`%||%` <- function(x, y) if (is.null(x)) y else x

read_thresholds <- function() {
  if (use_winner_threshold && file.exists(winner_csv)) {
    winner <- readr::read_csv(winner_csv, show_col_types = FALSE)
    if (nrow(winner) != 1L) {
      stop("Expected one row in winner_threshold.csv; found ", nrow(winner), ".")
    }
    return(tibble::tibble(
      corr_min = winner$corr_min[[1]],
      margin_min = winner$margin_min[[1]],
      var_min = winner$var_min[[1]],
      source = "winner_threshold.csv"
    ))
  }

  tibble::tibble(
    corr_min = manual_corr_min,
    margin_min = manual_margin_min,
    var_min = manual_var_min,
    source = "manual"
  )
}

resolve_limits <- function(values, user_limits, trim = density_trim_quantile) {
  if (!is.null(user_limits)) {
    return(user_limits)
  }
  values <- values[is.finite(values)]
  if (!length(values)) {
    stop("Cannot infer limits from an empty/non-finite vector.")
  }
  stats::quantile(values, probs = trim, na.rm = TRUE, names = FALSE)
}

require_metric <- function(df, metric) {
  if (!metric %in% names(df)) {
    stop("Metric column not found: ", metric)
  }
}

kde2d_df <- function(df, x_col, y_col, group_col, xlim, ylim, n, h) {
  if (!requireNamespace("MASS", quietly = TRUE)) {
    stop("Package MASS is required for kde2d density plots.")
  }

  preferred_groups <- c("susie_vanilla", "susine_functional_mu")
  groups <- intersect(preferred_groups, unique(df[[group_col]]))
  if (length(groups) < 2L) {
    groups <- sort(unique(df[[group_col]]))
  }
  if (length(groups) != 2L) {
    stop("Expected exactly two density groups; found: ", paste(groups, collapse = ", "))
  }

  dens <- lapply(groups, function(g) {
    sub <- df %>%
      dplyr::filter(.data[[group_col]] == g) %>%
      dplyr::filter(
        is.finite(.data[[x_col]]),
        is.finite(.data[[y_col]]),
        .data[[x_col]] >= xlim[[1]],
        .data[[x_col]] <= xlim[[2]],
        .data[[y_col]] >= ylim[[1]],
        .data[[y_col]] <= ylim[[2]]
      )

    if (nrow(sub) < 3L) {
      stop("Not enough rows for density group ", g, ": ", nrow(sub))
    }

    kde <- MASS::kde2d(
      x = sub[[x_col]],
      y = sub[[y_col]],
      n = n,
      lims = c(xlim, ylim),
      h = h
    )

    list(group = g, kde = kde, n_rows = nrow(sub))
  })

  base <- dens[[1]]$kde
  z1 <- dens[[1]]$kde$z
  z2 <- dens[[2]]$kde$z

  tidyr::expand_grid(
    x = base$x,
    y = base$y
  ) %>%
    dplyr::mutate(
      density_group_1 = as.vector(t(z1)),
      density_group_2 = as.vector(t(z2)),
      density_diff = .data$density_group_2 - .data$density_group_1,
      group_1 = dens[[1]]$group,
      group_2 = dens[[2]]$group,
      n_group_1 = dens[[1]]$n_rows,
      n_group_2 = dens[[2]]$n_rows
    )
}

keff_x_scale <- function(breaks = x_keff_breaks, xlim = NULL) {
  finite_breaks <- breaks[breaks > 0]
  scale_x_continuous(
    breaks = log(finite_breaks),
    labels = formatC(finite_breaks, format = "fg", big.mark = ","),
    limits = xlim
  )
}

keff_ratio_x_scale <- function(breaks = delta_keff_ratio_breaks,
                               labels = delta_keff_ratio_labels,
                               xlim = NULL) {
  scale_x_continuous(
    breaks = log(breaks),
    labels = labels,
    limits = xlim
  )
}

plot_density_diff <- function(density_df, title, subtitle, out_path) {
  abs_diff <- abs(density_df$density_diff)
  abs_diff <- abs_diff[is.finite(abs_diff)]

  fill_lim <- density_fill_limit
  if (is.null(fill_lim)) {
    if (length(abs_diff)) {
      fill_lim <- as.numeric(stats::quantile(
        abs_diff,
        probs = density_fill_quantile,
        na.rm = TRUE
      ))
      max_abs <- max(abs_diff, na.rm = TRUE)
    } else {
      fill_lim <- 1
      max_abs <- 1
    }
    if (!is.finite(fill_lim) || fill_lim <= 0) {
      fill_lim <- max(max_abs, .Machine$double.eps)
    }
  }

  fill_legend_title <- sprintf(
    "mu - vanilla\ndensity\n(clipped at\n+/-%s of max)",
    if (is.null(density_fill_limit)) {
      paste0(round(100 * density_fill_quantile), "%-tile")
    } else {
      "manual"
    }
  )

  p <- ggplot(density_df, aes(x = .data$x, y = .data$y, fill = .data$density_diff)) +
    geom_raster(interpolate = TRUE) +
    geom_contour(aes(z = .data$density_diff), color = "grey20", alpha = 0.45, linewidth = 0.2) +
    scale_fill_gradient2(
      low = "#2b6cb0",
      mid = "white",
      high = "#c53030",
      midpoint = 0,
      limits = c(-fill_lim, fill_lim),
      oob = scales::squish
    ) +
    keff_x_scale(xlim = x_limits_resolved) +
    coord_cartesian(xlim = x_limits_resolved, ylim = y_limits_resolved, expand = FALSE) +
    labs(
      x = x_label,
      y = y_label,
      fill = fill_legend_title,
      title = title,
      subtitle = subtitle
    ) +
    theme(legend.position = "right")

  ggsave(out_path, p, width = 9, height = 6, dpi = 150)
  p
}

plot_scatter <- function(df, title, out_path, facet_var = NULL) {
  p <- ggplot(
    df,
    aes(x = .data[[x_metric]], y = .data[[y_metric]], color = .data$model_method)
  ) +
    geom_point(alpha = 0.35, size = 0.9) +
    scale_color_manual(
      values = c(susie_vanilla = "black", susine_functional_mu = "#2b8cbe"),
      breaks = c("susie_vanilla", "susine_functional_mu")
    ) +
    keff_x_scale(xlim = x_limits_resolved) +
    coord_cartesian(xlim = x_limits_resolved, ylim = y_limits_resolved) +
    labs(
      x = x_label,
      y = y_label,
      color = "Model",
      title = title
    ) +
    theme(legend.position = "bottom")

  if (!is.null(facet_var)) {
    p <- p + facet_wrap(stats::as.formula(paste("~", facet_var)))
  }

  ggsave(out_path, p, width = 10, height = 7, dpi = 150)
  p
}

plot_univariate_hist <- function(df,
                                 metric,
                                 xlab_text,
                                 out_path,
                                 title,
                                 bins = hist_bins,
                                 use_keff_x = FALSE,
                                 xlim = NULL) {
  panel_levels <- c("All effects", "Matched", "Unmatched")
  df_panels <- dplyr::bind_rows(
    df %>% dplyr::mutate(panel = "All effects"),
    df %>% dplyr::filter(.data$match_status == "matched") %>%
      dplyr::mutate(panel = "Matched"),
    df %>% dplyr::filter(.data$match_status == "unmatched") %>%
      dplyr::mutate(panel = "Unmatched")
  ) %>%
    dplyr::filter(is.finite(.data[[metric]])) %>%
    dplyr::mutate(panel = factor(.data$panel, levels = panel_levels))

  if (!is.null(xlim)) {
    df_panels <- df_panels %>%
      dplyr::filter(
        .data[[metric]] >= xlim[[1]],
        .data[[metric]] <= xlim[[2]]
      )
  }

  panel_counts <- df_panels %>%
    dplyr::count(.data$panel, .data$model_method, name = "n_effects")
  cat(title, "panel counts:\n")
  print(panel_counts)

  p <- ggplot(df_panels, aes(x = .data[[metric]], fill = .data$model_method)) +
    geom_histogram(
      aes(y = after_stat(density)),
      position = "identity",
      bins = bins,
      alpha = 0.45,
      color = NA
    ) +
    facet_wrap(~ panel, ncol = 3, scales = "free_y") +
    scale_fill_manual(
      values = c(susie_vanilla = "black", susine_functional_mu = "#2b8cbe"),
      breaks = c("susie_vanilla", "susine_functional_mu")
    ) +
    labs(
      x = xlab_text,
      y = "Density",
      fill = "Model",
      title = title
    ) +
    theme(legend.position = "bottom")

  if (use_keff_x) {
    p <- p + keff_x_scale(xlim = xlim)
  } else if (!is.null(xlim)) {
    p <- p + coord_cartesian(xlim = xlim)
  }

  ggsave(out_path, p, width = 11, height = 5, dpi = 150)
  p
}
```

# 3. Load And Classify Effects

```{r load-classify}
stopifnot(file.exists(records_rds))
stopifnot(file.exists(real_candidates_csv))

records <- readRDS(records_rds) %>%
  dplyr::mutate(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    effect = as.integer(.data$effect)
  )

require_metric(records, x_metric)
require_metric(records, y_metric)

thresholds <- read_thresholds()
cat("Matching thresholds:\n")
print(thresholds)

real_candidates <- readr::read_csv(real_candidates_csv, show_col_types = FALSE) %>%
  dplyr::mutate(
    vanilla_dataset_bundle_id = as.character(.data$vanilla_dataset_bundle_id),
    mu_dataset_bundle_id = as.character(.data$mu_dataset_bundle_id),
    vanilla_effect = as.integer(.data$vanilla_effect),
    mu_effect = as.integer(.data$mu_effect)
  )

accepted_matches <- real_candidates %>%
  dplyr::filter(
    .data$corr >= thresholds$corr_min[[1]],
    .data$margin >= thresholds$margin_min[[1]],
    .data$pair_var_explained_frac >= thresholds$var_min[[1]]
  ) %>%
  dplyr::arrange(.data$vanilla_dataset_bundle_id, dplyr::desc(.data$corr))

vanilla_match_keys <- accepted_matches %>%
  dplyr::transmute(
    dataset_bundle_id = .data$vanilla_dataset_bundle_id,
    effect = .data$vanilla_effect,
    match_id = paste0(.data$vanilla_dataset_bundle_id, "_", .data$vanilla_effect, "_", .data$mu_effect)
  )

mu_match_keys <- accepted_matches %>%
  dplyr::transmute(
    dataset_bundle_id = .data$mu_dataset_bundle_id,
    effect = .data$mu_effect,
    match_id = paste0(.data$vanilla_dataset_bundle_id, "_", .data$vanilla_effect, "_", .data$mu_effect)
  )

records_labeled <- records %>%
  dplyr::left_join(
    vanilla_match_keys %>% dplyr::mutate(model_method = "susie_vanilla"),
    by = c("dataset_bundle_id", "effect", "model_method")
  ) %>%
  dplyr::left_join(
    mu_match_keys %>% dplyr::mutate(model_method = "susine_functional_mu"),
    by = c("dataset_bundle_id", "effect", "model_method"),
    suffix = c("_vanilla_key", "_mu_key")
  ) %>%
  dplyr::mutate(
    match_id = dplyr::coalesce(.data$match_id_vanilla_key, .data$match_id_mu_key),
    match_status = dplyr::if_else(is.na(.data$match_id), "unmatched", "matched")
  ) %>%
  dplyr::select(-dplyr::any_of(c("match_id_vanilla_key", "match_id_mu_key")))

x_limits_resolved <- resolve_limits(records_labeled[[x_metric]], x_limits)
y_limits_resolved <- resolve_limits(records_labeled[[y_metric]], y_limits)

readr::write_csv(accepted_matches, matched_pairs_csv)

cat("Accepted matches:", nrow(accepted_matches), "\n")
cat("Effect rows by model/status:\n")
print(records_labeled %>% dplyr::count(.data$model_method, .data$match_status), n = Inf)
cat("Resolved x limits:", paste(x_limits_resolved, collapse = ", "), "\n")
cat("Resolved y limits:", paste(y_limits_resolved, collapse = ", "), "\n")
```

# 4. 2D Scatter Plots

```{r scatter-plots, fig.width=10, fig.height=7}
p_scatter_all <- plot_scatter(
  records_labeled,
  title = "Effect diagnostic scatter: all refit effects",
  out_path = scatter_all_png
)
print(p_scatter_all)

p_scatter_status <- plot_scatter(
  records_labeled,
  title = "Effect diagnostic scatter split by accepted matching status",
  out_path = scatter_matched_unmatched_png,
  facet_var = "match_status"
)
print(p_scatter_status)

cat("Saved:", scatter_all_png, "\n")
cat("Saved:", scatter_matched_unmatched_png, "\n")
```

# 5. Differential 2D Density

```{r differential-density, fig.width=9, fig.height=6}
density_all_df <- kde2d_df(
  records_labeled,
  x_col = x_metric,
  y_col = y_metric,
  group_col = "model_method",
  xlim = x_limits_resolved,
  ylim = y_limits_resolved,
  n = density_grid_n,
  h = density_bandwidth
)

p_density_all <- plot_density_diff(
  density_all_df,
  title = "Differential joint density: all effects",
  subtitle = "Red: higher susine functional-mu density; blue: higher susie vanilla density",
  out_path = density_all_png
)
print(p_density_all)

for (status in c("matched", "unmatched")) {
  sub <- records_labeled %>% dplyr::filter(.data$match_status == status)
  if (dplyr::n_distinct(sub$model_method) < 2L || nrow(sub) < 6L) {
    cat("Skipping density for status", status, "- not enough rows.\n")
    next
  }

  dens <- kde2d_df(
    sub,
    x_col = x_metric,
    y_col = y_metric,
    group_col = "model_method",
    xlim = x_limits_resolved,
    ylim = y_limits_resolved,
    n = density_grid_n,
    h = density_bandwidth
  )

  p <- plot_density_diff(
    dens,
    title = paste("Differential joint density:", status, "effects"),
    subtitle = "Red: higher susine functional-mu density; blue: higher susie vanilla density",
    out_path = if (status == "matched") density_matched_png else density_unmatched_png
  )
  print(p)
}

cat("Saved:", density_all_png, "\n")
cat("Saved:", density_matched_png, "\n")
cat("Saved:", density_unmatched_png, "\n")
```

# 6. Matched Effect Flow Plot

```{r matched-flow, fig.width=10, fig.height=8}
matched_vanilla <- records_labeled %>%
  dplyr::filter(.data$model_method == "susie_vanilla", .data$match_status == "matched") %>%
  dplyr::transmute(
    match_id = .data$match_id,
    vanilla_dataset_bundle_id = .data$dataset_bundle_id,
    vanilla_effect = .data$effect,
    vanilla_x = .data[[x_metric]],
    vanilla_y = .data[[y_metric]]
  )

matched_mu <- records_labeled %>%
  dplyr::filter(.data$model_method == "susine_functional_mu", .data$match_status == "matched") %>%
  dplyr::transmute(
    match_id = .data$match_id,
    mu_dataset_bundle_id = .data$dataset_bundle_id,
    mu_effect = .data$effect,
    mu_x = .data[[x_metric]],
    mu_y = .data[[y_metric]]
  )

flow_df <- accepted_matches %>%
  dplyr::mutate(
    match_id = paste0(.data$vanilla_dataset_bundle_id, "_", .data$vanilla_effect, "_", .data$mu_effect)
  ) %>%
  dplyr::left_join(matched_vanilla, by = "match_id") %>%
  dplyr::left_join(matched_mu, by = "match_id") %>%
  dplyr::mutate(
    delta_x = .data$mu_x - .data$vanilla_x,
    delta_y = .data$mu_y - .data$vanilla_y,
    movement_distance = sqrt(.data$delta_x^2 + .data$delta_y^2),
    delta_direction = dplyr::case_when(
      .data$delta_y > 0 & .data$delta_x > 0 ~ "more accurate + more diffuse",
      .data$delta_y > 0 & .data$delta_x <= 0 ~ "more accurate + less diffuse",
      .data$delta_y <= 0 & .data$delta_x > 0 ~ "less accurate + more diffuse",
      TRUE ~ "less accurate + less diffuse"
    )
  ) %>%
  dplyr::arrange(dplyr::desc(.data$movement_distance))

if (is.finite(max_arrows_to_plot)) {
  flow_plot_df <- flow_df %>% dplyr::slice_head(n = as.integer(max_arrows_to_plot))
} else {
  flow_plot_df <- flow_df
}

readr::write_csv(flow_df, matched_delta_csv)

top_labels <- flow_df %>%
  dplyr::slice_head(n = label_top_discrepancies)

p_flow <- ggplot() +
  geom_point(
    data = flow_plot_df,
    aes(x = .data$vanilla_x, y = .data$vanilla_y),
    color = "black",
    alpha = 0.35,
    size = 1.0
  ) +
  geom_segment(
    data = flow_plot_df,
    aes(
      x = .data$vanilla_x,
      y = .data$vanilla_y,
      xend = .data$mu_x,
      yend = .data$mu_y,
      color = .data$delta_direction
    ),
    alpha = arrow_alpha,
    linewidth = arrow_linewidth,
    arrow = grid::arrow(length = grid::unit(arrow_head_length_inches, "inches"), type = "closed")
  ) +
  geom_point(
    data = flow_plot_df,
    aes(x = .data$mu_x, y = .data$mu_y, color = .data$delta_direction),
    alpha = 0.55,
    size = 1.1
  ) +
  geom_point(
    data = top_labels,
    aes(x = .data$mu_x, y = .data$mu_y),
    color = "red",
    size = 1.8,
    alpha = 0.75
  ) +
  keff_x_scale(xlim = x_limits_resolved) +
  coord_cartesian(xlim = x_limits_resolved, ylim = y_limits_resolved) +
  labs(
    x = x_label,
    y = y_label,
    color = "Movement",
    title = "Matched effect diagnostic movement",
    subtitle = "Arrows start at susie vanilla and end at susine functional-mu"
  ) +
  theme(legend.position = "bottom")

ggsave(flow_png, p_flow, width = 10, height = 8, dpi = 150)
print(p_flow)

cat("Saved:", flow_png, "\n")
cat("Saved matched deltas:", matched_delta_csv, "\n")
```

# 7. Matched Delta Density

```{r matched-delta-density, fig.width=10, fig.height=6}
if (nrow(flow_df)) {
  p_delta <- ggplot(flow_df, aes(x = .data$delta_x, y = .data$delta_y)) +
    geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
    geom_vline(xintercept = 0, color = "grey45", linetype = "dashed") +
    geom_density_2d_filled(
      alpha = 0.75,
      contour_var = "ndensity",
      breaks = delta_density_contour_breaks
    ) +
    geom_point(alpha = 0.25, size = 0.8, color = "black") +
    keff_ratio_x_scale() +
    labs(
      x = "Core-95 effective k ratio (mu / vanilla)",
      y = paste0("Delta ", y_label, " (mu - vanilla)"),
      fill = "Relative\ndensity",
      title = "Matched effect diagnostic displacement density"
    ) +
    theme(legend.position = "right")

  ggsave(delta_density_png, p_delta, width = 10, height = 6, dpi = 150)
  print(p_delta)

  delta_summary <- flow_df %>%
    dplyr::summarise(
      n = dplyr::n(),
      median_delta_x = stats::median(.data$delta_x, na.rm = TRUE),
      q05_delta_x = as.numeric(stats::quantile(.data$delta_x, 0.05, na.rm = TRUE)),
      q95_delta_x = as.numeric(stats::quantile(.data$delta_x, 0.95, na.rm = TRUE)),
      median_delta_y = stats::median(.data$delta_y, na.rm = TRUE),
      q05_delta_y = as.numeric(stats::quantile(.data$delta_y, 0.05, na.rm = TRUE)),
      q95_delta_y = as.numeric(stats::quantile(.data$delta_y, 0.95, na.rm = TRUE)),
      median_corr = stats::median(.data$corr, na.rm = TRUE),
      median_margin = stats::median(.data$margin, na.rm = TRUE),
      median_pair_var_pct = 100 * stats::median(.data$pair_var_explained_frac, na.rm = TRUE)
    )

  cat("Matched displacement summary:\n")
  print(delta_summary)
  cat("Saved:", delta_density_png, "\n")
}
```

# 8. Univariate Histograms

```{r univariate-histograms, fig.width=11, fig.height=5}
p_hist_diff <- plot_univariate_hist(
  records_labeled,
  metric = x_metric,
  xlab_text = x_label,
  out_path = hist_diffuseness_png,
  title = "Effect diffuseness: Core-95 effective k by match status",
  use_keff_x = TRUE,
  xlim = x_limits_resolved
)
print(p_hist_diff)

p_hist_acc <- plot_univariate_hist(
  records_labeled,
  metric = y_metric,
  xlab_text = y_label,
  out_path = hist_accuracy_png,
  title = "Effect accuracy ratio by match status",
  use_keff_x = FALSE,
  xlim = y_limits_resolved
)
print(p_hist_acc)

cat("Saved:", hist_diffuseness_png, "\n")
cat("Saved:", hist_accuracy_png, "\n")
```

# 9. AUPRC Attribution

Three per-dataset views that probe whether susine functional-mu is *refining*
existing model fits (high match counts, modest AUPRC gains, displacement
density tight around zero) or *unlocking* substantively different fits (low
match counts, large AUPRC swings, sparse / outlier displacements).

```{r auprc-attribution, fig.width=10, fig.height=6}
per_dataset_auprc <- records_labeled %>%
  dplyr::distinct(.data$dataset_bundle_id, .data$model_method, .data$model_AUPRC) %>%
  tidyr::pivot_wider(
    names_from = .data$model_method,
    values_from = .data$model_AUPRC,
    names_prefix = "auprc_"
  ) %>%
  dplyr::mutate(
    auprc_delta = .data$auprc_susine_functional_mu - .data$auprc_susie_vanilla
  )

matched_per_dataset <- accepted_matches %>%
  dplyr::count(.data$vanilla_dataset_bundle_id, name = "n_matched_pairs") %>%
  dplyr::rename(dataset_bundle_id = .data$vanilla_dataset_bundle_id)

# flow_df comes from joining accepted_matches with matched_vanilla / matched_mu,
# both of which redefine vanilla_dataset_bundle_id; dplyr suffixes the dupes
# to .x / .y, so the bare column name no longer exists. Pull movement_distance
# back to (clean) accepted_matches via match_id and aggregate from there.
movement_per_match <- flow_df %>%
  dplyr::select(.data$match_id, .data$movement_distance)

movement_per_dataset <- accepted_matches %>%
  dplyr::mutate(
    match_id = paste0(
      .data$vanilla_dataset_bundle_id, "_",
      .data$vanilla_effect, "_",
      .data$mu_effect
    )
  ) %>%
  dplyr::left_join(movement_per_match, by = "match_id") %>%
  dplyr::group_by(.data$vanilla_dataset_bundle_id) %>%
  dplyr::summarise(
    mean_movement_distance = mean(.data$movement_distance, na.rm = TRUE),
    median_movement_distance = stats::median(.data$movement_distance, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::rename(dataset_bundle_id = .data$vanilla_dataset_bundle_id)

per_dataset_summary <- per_dataset_auprc %>%
  dplyr::left_join(matched_per_dataset, by = "dataset_bundle_id") %>%
  dplyr::mutate(n_matched_pairs = tidyr::replace_na(.data$n_matched_pairs, 0L)) %>%
  dplyr::left_join(movement_per_dataset, by = "dataset_bundle_id")

readr::write_csv(per_dataset_summary, per_dataset_summary_csv)

n_total <- sum(is.finite(per_dataset_summary$auprc_delta))
n_pos <- sum(per_dataset_summary$auprc_delta > 0, na.rm = TRUE)
n_neg <- sum(per_dataset_summary$auprc_delta < 0, na.rm = TRUE)
n_zero <- sum(per_dataset_summary$auprc_delta == 0, na.rm = TRUE)
median_delta <- stats::median(per_dataset_summary$auprc_delta, na.rm = TRUE)

cat("Per-dataset AUPRC delta summary:\n")
cat(sprintf(
  "  total datasets: %d | gains: %d | losses: %d | ties: %d | median delta: %.4f\n",
  n_total, n_pos, n_neg, n_zero, median_delta
))

p_auprc_delta <- ggplot(per_dataset_summary, aes(x = .data$auprc_delta)) +
  geom_vline(xintercept = 0, color = "grey45", linetype = "dashed") +
  geom_histogram(bins = hist_bins, fill = "#2b8cbe", alpha = 0.7, color = NA) +
  labs(
    x = "AUPRC delta (mu - vanilla)",
    y = "Number of datasets",
    title = "Per-dataset AUPRC improvement from susine functional-mu",
    subtitle = sprintf(
      "n = %d datasets | %d gains, %d losses, %d ties | median delta = %.4f",
      n_total, n_pos, n_neg, n_zero, median_delta
    )
  )

ggsave(auprc_delta_hist_png, p_auprc_delta, width = 9, height = 5, dpi = 150)
print(p_auprc_delta)

p_match_count <- ggplot(per_dataset_summary, aes(x = .data$n_matched_pairs)) +
  geom_histogram(binwidth = 1, fill = "#2b8cbe", alpha = 0.7, color = NA) +
  scale_x_continuous(breaks = 0:10) +
  labs(
    x = "Number of accepted matched pairs per dataset (out of L = 10)",
    y = "Number of datasets",
    title = "Per-dataset matched-pair count distribution"
  )

ggsave(match_count_hist_png, p_match_count, width = 9, height = 5, dpi = 150)
print(p_match_count)

scatter_corr <- with(
  per_dataset_summary,
  suppressWarnings(stats::cor(
    n_matched_pairs,
    auprc_delta,
    use = "pairwise.complete.obs",
    method = "spearman"
  ))
)

p_auprc_match_scatter <- ggplot(
  per_dataset_summary,
  aes(
    x = .data$n_matched_pairs,
    y = .data$auprc_delta,
    color = .data$mean_movement_distance
  )
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_jitter(width = 0.25, height = 0, alpha = 0.6, size = 1.5) +
  geom_smooth(
    method = "loess",
    formula = y ~ x,
    se = FALSE,
    color = "black",
    linetype = "dashed",
    linewidth = 0.6
  ) +
  scale_x_continuous(breaks = 0:10) +
  scale_color_viridis_c(option = "C", na.value = "grey60") +
  labs(
    x = "Number of accepted matched pairs per dataset",
    y = "AUPRC delta (mu - vanilla)",
    color = "Mean matched\npair movement",
    title = "Per-dataset: matching density vs AUPRC improvement",
    subtitle = sprintf(
      "Spearman rho = %.3f. A predicts upper-right concentration; B predicts upper-left.",
      scatter_corr
    )
  )

ggsave(auprc_match_scatter_png, p_auprc_match_scatter, width = 10, height = 6, dpi = 150)
print(p_auprc_match_scatter)

cat("Saved:", auprc_delta_hist_png, "\n")
cat("Saved:", match_count_hist_png, "\n")
cat("Saved:", auprc_match_scatter_png, "\n")
cat("Saved:", per_dataset_summary_csv, "\n")
```

# 10. Confound Diagnostics

The negative match-count vs AUPRC-delta trend in section 9 could be driven by
dataset difficulty: hard datasets have low vanilla AUPRC (more headroom for
delta) and diffuse effects (fewer matched pairs), creating a spurious
negative correlation even if susine functional-mu were doing pure refinement
everywhere. The first plot below is the most direct view (per-dataset
baseline AUPRC vs susine AUPRC, with the y=x reference line); the next two
disentangle the match-count signal from the difficulty confound.

```{r baseline-vs-susine, fig.width=8, fig.height=8}
n_above_diag <- sum(per_dataset_summary$auprc_delta > 0, na.rm = TRUE)
n_below_diag <- sum(per_dataset_summary$auprc_delta < 0, na.rm = TRUE)
n_total_compared <- n_above_diag + n_below_diag

baseline_vs_susine_pearson <- with(
  per_dataset_summary,
  suppressWarnings(stats::cor(
    auprc_susie_vanilla,
    auprc_susine_functional_mu,
    method = "pearson",
    use = "pairwise.complete.obs"
  ))
)

p_baseline_vs_susine <- ggplot(
  per_dataset_summary,
  aes(
    x = .data$auprc_susie_vanilla,
    y = .data$auprc_susine_functional_mu,
    color = .data$n_matched_pairs
  )
) +
  geom_abline(slope = 1, intercept = 0, color = "grey45", linetype = "dashed") +
  geom_point(alpha = 0.65, size = 1.4) +
  geom_smooth(
    method = "lm", formula = y ~ x, se = FALSE,
    color = "black", linetype = "dashed", linewidth = 0.5
  ) +
  scale_color_viridis_c(option = "C") +
  coord_fixed() +
  labs(
    x = "Vanilla SuSiE AUPRC",
    y = "SuSiNE functional-mu AUPRC",
    color = "Matched pairs",
    title = "Per-dataset AUPRC: vanilla vs susine functional-mu",
    subtitle = sprintf(
      "Pearson r = %.3f | %d / %d datasets above the y=x line (susine wins)",
      baseline_vs_susine_pearson,
      n_above_diag,
      n_total_compared
    )
  )

ggsave(baseline_vs_susine_png, p_baseline_vs_susine, width = 8, height = 8, dpi = 150)
print(p_baseline_vs_susine)

cat("Saved:", baseline_vs_susine_png, "\n")
```

```{r confound-diagnostics, fig.width=11, fig.height=8}
quart_breaks <- stats::quantile(
  per_dataset_summary$auprc_susie_vanilla,
  probs = c(0, 0.25, 0.5, 0.75, 1.0),
  na.rm = TRUE,
  names = FALSE
)
quart_breaks <- unique(quart_breaks)

if (length(quart_breaks) >= 3L) {
  per_dataset_summary <- per_dataset_summary %>%
    dplyr::mutate(
      vanilla_auprc_quartile = cut(
        .data$auprc_susie_vanilla,
        breaks = quart_breaks,
        include.lowest = TRUE
      )
    )

  quartile_rho <- per_dataset_summary %>%
    dplyr::filter(!is.na(.data$vanilla_auprc_quartile)) %>%
    dplyr::group_by(.data$vanilla_auprc_quartile) %>%
    dplyr::summarise(
      n = dplyr::n(),
      rho = suppressWarnings(stats::cor(
        .data$n_matched_pairs,
        .data$auprc_delta,
        method = "spearman",
        use = "pairwise.complete.obs"
      )),
      label = sprintf("rho = %.3f (n=%d)", .data$rho, .data$n),
      .groups = "drop"
    )

  cat("Per-quartile match-count vs AUPRC-delta Spearman rho:\n")
  print(quartile_rho)

  p_strat <- ggplot(
    per_dataset_summary %>% dplyr::filter(!is.na(.data$vanilla_auprc_quartile)),
    aes(
      x = .data$n_matched_pairs,
      y = .data$auprc_delta,
      color = .data$mean_movement_distance
    )
  ) +
    geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
    geom_jitter(width = 0.25, height = 0, alpha = 0.6, size = 1.3) +
    geom_smooth(
      method = "loess", formula = y ~ x, se = FALSE,
      color = "black", linetype = "dashed", linewidth = 0.6
    ) +
    geom_text(
      data = quartile_rho,
      aes(x = -Inf, y = Inf, label = .data$label),
      hjust = -0.05, vjust = 1.4, size = 3.5,
      inherit.aes = FALSE
    ) +
    facet_wrap(~ vanilla_auprc_quartile, ncol = 2, scales = "free_y") +
    scale_x_continuous(breaks = 0:10) +
    scale_color_viridis_c(option = "C", na.value = "grey60") +
    labs(
      x = "Number of accepted matched pairs per dataset",
      y = "AUPRC delta (mu - vanilla)",
      color = "Mean matched\npair movement",
      title = "Match count vs AUPRC delta, stratified by vanilla AUPRC quartile",
      subtitle = "Lower quartiles = harder datasets; if rho stays negative within panels, the original trend is real"
    )

  ggsave(scatter_stratified_png, p_strat, width = 11, height = 8, dpi = 150)
  print(p_strat)
} else {
  cat("Not enough distinct vanilla AUPRC values to form quartiles; skipping.\n")
}

confound_rho <- with(
  per_dataset_summary,
  suppressWarnings(stats::cor(
    auprc_susie_vanilla,
    n_matched_pairs,
    method = "spearman",
    use = "pairwise.complete.obs"
  ))
)

p_confound <- ggplot(
  per_dataset_summary,
  aes(
    x = .data$auprc_susie_vanilla,
    y = .data$n_matched_pairs,
    color = .data$auprc_delta
  )
) +
  geom_jitter(width = 0, height = 0.25, alpha = 0.6, size = 1.3) +
  geom_smooth(
    method = "loess", formula = y ~ x, se = FALSE,
    color = "black", linetype = "dashed", linewidth = 0.6
  ) +
  scale_color_gradient2(
    low = "#2b6cb0", mid = "white", high = "#c53030", midpoint = 0
  ) +
  scale_y_continuous(breaks = 0:10) +
  labs(
    x = "Vanilla AUPRC (low = hard, high = easy)",
    y = "Number of accepted matched pairs",
    color = "AUPRC delta\n(mu - vanilla)",
    title = "Confound check: vanilla AUPRC vs match count",
    subtitle = sprintf(
      "Spearman rho = %.3f. Strong positive => confound chain is real.",
      confound_rho
    )
  )

ggsave(confound_check_png, p_confound, width = 10, height = 5, dpi = 150)
print(p_confound)

cat("Saved:", scatter_stratified_png, "\n")
cat("Saved:", confound_check_png, "\n")
```

# 11. Output Index

```{r output-index}
outputs <- tibble::tibble(
  artifact = c(
    "all effect scatter",
    "matched/unmatched scatter",
    "all effect differential density",
    "matched differential density",
    "unmatched differential density",
    "matched flow plot",
    "matched delta density",
    "univariate diffuseness histogram",
    "univariate accuracy histogram",
    "AUPRC delta histogram",
    "match count histogram",
    "AUPRC delta vs match count scatter",
    "AUPRC delta vs match count stratified by vanilla AUPRC",
    "vanilla AUPRC vs match count (confound check)",
    "baseline vs susine AUPRC scatter",
    "accepted matched pairs",
    "matched pair deltas",
    "per-dataset AUPRC and match summary"
  ),
  path = c(
    scatter_all_png,
    scatter_matched_unmatched_png,
    density_all_png,
    density_matched_png,
    density_unmatched_png,
    flow_png,
    delta_density_png,
    hist_diffuseness_png,
    hist_accuracy_png,
    auprc_delta_hist_png,
    match_count_hist_png,
    auprc_match_scatter_png,
    scatter_stratified_png,
    confound_check_png,
    baseline_vs_susine_png,
    matched_pairs_csv,
    matched_delta_csv,
    per_dataset_summary_csv
  )
)

kable(outputs)
```
