---
title: "Effect Drift Paper Figure (5-arm, cached plotting only)"
output:
  html_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: 72
---

This notebook is the lightweight plotting companion to
`effect_matching_drift_analysis_5arm.Rmd`. It assumes the heavy notebook has
already generated the cached CSV outputs in `study_dir`, then rebuilds only the
combined paper figure. Use this notebook for figure layout, titles, labels,
legends, and styling tweaks.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_drift_paper_figure_5arm.Rmd")

library(here)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(scales)
```

# Inputs

```{r paths}
output_root <- here("output")
study_name <- "baseline_sims_5arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
plot_dir <- file.path(study_dir, "drift_analysis_plots")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

drift_summary_csv <- file.path(study_dir, "drift_summary_5arm.csv")
auprc_per_dataset_csv <- file.path(study_dir, "auprc_per_dataset_5arm.csv")
pip_drift_summary_csv <- file.path(study_dir, "pip_drift_summary_5arm.csv")
pr_curves_pooled_csv <- file.path(study_dir, "pr_curves_pooled_top8_5arm.csv")
pr_auc_pooled_csv <- file.path(study_dir, "pr_auc_pooled_top8_5arm.csv")
effect_quality_pair_rows_csv <- file.path(study_dir, "effect_quality_pair_rows_5arm.csv")
effect_quality_density_shift_csv <- file.path(study_dir, "effect_quality_density_shift_5arm.csv")

paper_figure_png <- file.path(plot_dir, "paper_figure_effect_drift.png")

required_cache_files <- c(
  drift_summary_csv,
  auprc_per_dataset_csv,
  pip_drift_summary_csv,
  pr_curves_pooled_csv
)
missing_cache_files <- required_cache_files[!file.exists(required_cache_files)]
if (length(missing_cache_files)) {
  stop(
    "Missing required cached plotting inputs. Run ",
    "effect_matching_drift_analysis_5arm.Rmd once to generate them:\n",
    paste(missing_cache_files, collapse = "\n")
  )
}

cat("Study dir:", study_dir, "\n")
cat("Plot dir:", plot_dir, "\n")
```

# Plot Options

```{r options}
baseline_method <- "susie_vanilla"
mu_method <- "susine_functional_mu"
pi_method <- "susie_functional_pi"
warm_mu_method <- "susie_vanilla_warm_from_mu"
warm_pi_method <- "susie_vanilla_warm_from_pi"

method_labels <- c(
  susie_vanilla = "SuSiE-vanilla (cold)",
  susine_functional_mu = "SuSiNE-functional-mu (cold)",
  susie_vanilla_warm_from_mu = "SuSiE-vanilla (warm from mu)",
  susie_functional_pi = "SuSiE-functional-pi (cold)",
  susie_vanilla_warm_from_pi = "SuSiE-vanilla (warm from pi)"
)
method_legend_short <- c(
  susie_vanilla = "Vanilla cold",
  susine_functional_mu = "Functional-mu",
  susie_vanilla_warm_from_mu = "Warm vanilla (mu)",
  susie_functional_pi = "Functional-pi",
  susie_vanilla_warm_from_pi = "Warm vanilla (pi)"
)
method_color <- c(
  susie_vanilla = "black",
  susine_functional_mu = "#2b8cbe",
  susie_vanilla_warm_from_mu = "#9ecae1",
  susie_functional_pi = "#74c476",
  susie_vanilla_warm_from_pi = "#bae4b3"
)
methods_in_order <- c(
  baseline_method, mu_method, warm_mu_method, pi_method, warm_pi_method
)

pr_top_k <- 8L

x_keff_breaks <- c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000)
x_limits <- NULL
y_limits <- c(0, 1)
density_grid_n <- 140L
density_bandwidth <- c(0.45, 0.10)
density_trim_quantile <- c(0.005, 0.995)
density_fill_quantile <- 0.95
```

# Helpers

```{r helpers}
`%||%` <- function(x, y) if (is.null(x)) y else x

resolve_limits <- function(values, user_limits, trim = density_trim_quantile) {
  if (!is.null(user_limits)) {
    return(user_limits)
  }
  values <- values[is.finite(values)]
  if (!length(values)) {
    stop("Cannot infer limits from an empty/non-finite vector.")
  }
  stats::quantile(values, probs = trim, na.rm = TRUE, names = FALSE)
}

keff_x_scale <- function(breaks = x_keff_breaks, xlim = NULL) {
  finite_breaks <- breaks[breaks > 0]
  scale_x_continuous(
    breaks = log(finite_breaks),
    labels = formatC(finite_breaks, format = "fg", big.mark = ","),
    limits = xlim
  )
}

theme_effect_paper <- function(base_size = 8) {
  theme_bw(base_size = base_size) +
    theme(
      panel.grid.major = element_line(color = "grey91", linewidth = 0.22),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "grey35", fill = NA, linewidth = 0.35),
      axis.title = element_text(size = base_size),
      axis.text = element_text(size = base_size - 1, color = "grey25"),
      plot.title = element_text(face = "bold", size = base_size + 1.2,
                                margin = margin(b = 3)),
      strip.background = element_rect(fill = "grey88", color = "grey35",
                                      linewidth = 0.35),
      strip.text = element_text(size = base_size - 0.8, color = "grey20"),
      legend.title = element_text(size = base_size - 0.4),
      legend.text = element_text(size = base_size - 1.1),
      legend.key.height = grid::unit(3.2, "mm"),
      legend.key.width = grid::unit(6.0, "mm"),
      plot.margin = margin(2.5, 2.5, 2.5, 2.5)
    )
}

weighted_kde2d_df <- function(df, xlim, ylim, n, h) {
  group_values <- c("source", "destination")
  dens <- lapply(group_values, function(g) {
    sub <- df %>%
      dplyr::filter(.data$arm_role == g) %>%
      dplyr::filter(
        is.finite(.data$x_value),
        is.finite(.data$y_value),
        .data$x_value >= xlim[[1]],
        .data$x_value <= xlim[[2]],
        .data$y_value >= ylim[[1]],
        .data$y_value <= ylim[[2]]
      )
    if (nrow(sub) < 3L) return(NULL)
    w <- pmax(tidyr::replace_na(sub$pve_weight, 0), 0)
    if (sum(w) <= 0) {
      w <- rep(1 / nrow(sub), nrow(sub))
    } else {
      w <- w / sum(w)
    }
    x_grid <- seq(xlim[[1]], xlim[[2]], length.out = n)
    y_grid <- seq(ylim[[1]], ylim[[2]], length.out = n)
    kx <- outer(x_grid, sub$x_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[1]]) / h[[1]]
    })
    ky <- outer(y_grid, sub$y_value, function(grid, obs) {
      stats::dnorm((grid - obs) / h[[2]]) / h[[2]]
    })
    ky <- sweep(ky, 2L, w, `*`)
    z <- kx %*% t(ky)
    list(group = g, x = x_grid, y = y_grid, z = z,
         n_rows = nrow(sub), pve_mass = sum(sub$pve_weight, na.rm = TRUE))
  })
  names(dens) <- group_values
  if (is.null(dens$source) || is.null(dens$destination)) {
    return(tibble::tibble())
  }
  tidyr::expand_grid(x = dens$source$x, y = dens$source$y) %>%
    dplyr::mutate(
      density_source = as.vector(t(dens$source$z)),
      density_destination = as.vector(t(dens$destination$z)),
      density_diff = .data$density_destination - .data$density_source,
      n_source = dens$source$n_rows,
      n_destination = dens$destination$n_rows,
      pve_source = dens$source$pve_mass,
      pve_destination = dens$destination$pve_mass
    )
}

build_density_shift_from_pair_rows <- function(pair_rows) {
  x_limits_resolved <- resolve_limits(pair_rows$x_value, x_limits)
  y_limits_resolved <- resolve_limits(pair_rows$y_value, y_limits)
  density_shift_rows <- list()
  for (ch in levels(pair_rows$chain)) {
    for (cmp in levels(pair_rows$comparison_kind)) {
      sub <- pair_rows %>%
        dplyr::filter(.data$chain == ch, .data$comparison_kind == cmp)
      dens <- weighted_kde2d_df(
        sub,
        xlim = x_limits_resolved,
        ylim = y_limits_resolved,
        n = density_grid_n,
        h = density_bandwidth
      )
      if (!nrow(dens)) next
      density_shift_rows[[length(density_shift_rows) + 1L]] <- dens %>%
        dplyr::mutate(chain = ch, comparison_kind = cmp)
    }
  }
  dplyr::bind_rows(density_shift_rows)
}
```

# Read Cached Data

```{r read-cache}
drift_summary <- readr::read_csv(drift_summary_csv, show_col_types = FALSE)
auprc_per_dataset <- readr::read_csv(auprc_per_dataset_csv, show_col_types = FALSE) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))
pip_drift_summary <- readr::read_csv(pip_drift_summary_csv, show_col_types = FALSE) %>%
  dplyr::mutate(dataset_bundle_id = as.character(.data$dataset_bundle_id))

pr_curve_df <- readr::read_csv(pr_curves_pooled_csv, show_col_types = FALSE) %>%
  dplyr::mutate(
    model_method = factor(.data$model_method, levels = methods_in_order)
  )
if (file.exists(pr_auc_pooled_csv)) {
  pr_auc_df <- readr::read_csv(pr_auc_pooled_csv, show_col_types = FALSE) %>%
    dplyr::mutate(
      model_method = factor(.data$model_method, levels = methods_in_order)
    )
  prevalence <- unique(pr_auc_df$prevalence)
  if (!length(prevalence) || !is.finite(prevalence[[1]])) {
    prevalence <- 0.008
  } else {
    prevalence <- prevalence[[1]]
  }
} else {
  message("PR-AUC cache missing; estimating AUPRC labels from cached PR curves.")
  pr_auc_df <- pr_curve_df %>%
    dplyr::arrange(.data$model_method, .data$recall) %>%
    dplyr::group_by(.data$model_method) %>%
    dplyr::summarise(
      auprc_integral = sum(
        diff(.data$recall) *
          0.5 * (dplyr::lag(.data$precision)[-1] + .data$precision[-1]),
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      model_method = factor(.data$model_method, levels = methods_in_order)
    )
  prevalence <- 0.008
}

if (file.exists(effect_quality_density_shift_csv)) {
  density_shift_df <- readr::read_csv(
    effect_quality_density_shift_csv, show_col_types = FALSE
  )
} else {
  if (!file.exists(effect_quality_pair_rows_csv)) {
    stop(
      "Missing both density-shift cache and pair-row cache:\n",
      effect_quality_density_shift_csv, "\n", effect_quality_pair_rows_csv
    )
  }
  message("Density-shift cache missing; rebuilding once from effect-quality pair rows.")
  effect_quality_pair_rows <- readr::read_csv(
    effect_quality_pair_rows_csv, show_col_types = FALSE
  ) %>%
    dplyr::mutate(
      chain = factor(.data$chain, levels = c("mu", "pi")),
      arm_role = factor(.data$arm_role, levels = c("source", "destination")),
      comparison_kind = factor(
        .data$comparison_kind,
        levels = c("vanilla_to_treatment",
                   "treatment_to_warm",
                   "vanilla_to_warm")
      )
    )
  density_shift_df <- build_density_shift_from_pair_rows(effect_quality_pair_rows)
  readr::write_csv(density_shift_df, effect_quality_density_shift_csv)
}

density_shift_df <- density_shift_df %>%
  dplyr::mutate(
    chain = factor(.data$chain, levels = c("mu", "pi")),
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment",
                 "treatment_to_warm",
                 "vanilla_to_warm")
    )
  )

x_limits_resolved <- range(density_shift_df$x, na.rm = TRUE)
y_limits_resolved <- range(density_shift_df$y, na.rm = TRUE)
shared_abs_diff <- abs(density_shift_df$density_diff)
shared_abs_diff <- shared_abs_diff[is.finite(shared_abs_diff)]
shared_fill_lim <- stats::quantile(
  shared_abs_diff, probs = density_fill_quantile, na.rm = TRUE, names = FALSE
)
if (!is.finite(shared_fill_lim) || shared_fill_lim <= 0) {
  shared_fill_lim <- max(shared_abs_diff, na.rm = TRUE)
}
if (!is.finite(shared_fill_lim) || shared_fill_lim <= 0) {
  shared_fill_lim <- 1
}
```

# Combined Paper Figure

```{r paper-figure, fig.width=8.8, fig.height=9.9}
expr_basin_x <- expression(paste("Basin drift (", mu, " chain)"))
expr_basin_y <- expression(paste("Basin drift (", pi, " chain)"))
expr_jsd_x   <- expression(paste("PIP JSD (", mu, " chain)"))
expr_jsd_y   <- expression(paste("PIP JSD (", pi, " chain)"))
expr_dauprc  <- expression(paste(Delta, "AUPRC vs cold vanilla"))

chain_labeller <- c(mu = "mu chain", pi = "pi chain")
pair_kind_labeller <- c(
  "cold -> annotated" = "Cold -> annotated",
  "cold -> vanilla_warm" = "Cold -> warm vanilla"
)

method_legend_labels <- vapply(
  levels(pr_curve_df$model_method),
  function(m) {
    auprc_val <- pr_auc_df$auprc_integral[
      match(m, as.character(pr_auc_df$model_method))
    ]
    sprintf("%s (%.3f)",
            method_legend_short[[m]] %||% method_labels[[m]] %||% m,
            auprc_val)
  },
  character(1)
)

p_pr <- ggplot(pr_curve_df,
               aes(x = .data$recall, y = .data$precision,
                   color = .data$model_method)) +
  geom_hline(yintercept = prevalence, color = "grey60",
             linetype = "dashed", linewidth = 0.4) +
  geom_step(direction = "vh", linewidth = 0.7) +
  scale_color_manual(
    values = method_color[levels(pr_curve_df$model_method)],
    labels = method_legend_labels
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(x = "Recall", y = "Precision", color = "Method (pooled AUPRC)") +
  theme(legend.position = "right")

strip_chrome <- function(p) {
  p + labs(title = NULL, subtitle = NULL) +
    theme(plot.title = element_blank(), plot.subtitle = element_blank())
}

panel_A <- strip_chrome(p_pr) +
  labs(title = paste0(
    "A. Precision-recall on top-", pr_top_k,
    " causal SNPs (600 datasets)"
  )) +
  guides(color = guide_legend(
    nrow = 2, byrow = TRUE, title.position = "top",
    override.aes = list(linewidth = 0.7)
  )) +
  theme_effect_paper(base_size = 8) +
  theme(
    legend.position = "bottom",
    legend.justification = "left",
    legend.box.margin = margin(t = -4),
    legend.key.height = grid::unit(3.0, "mm"),
    legend.key.width = grid::unit(6, "mm"),
    legend.title = element_text(size = 6.8),
    legend.text = element_text(size = 6.2),
    panel.grid.minor = element_line(color = "grey95", linewidth = 0.18)
  )

basin_long_pf <- drift_summary %>%
  dplyr::filter(.data$pair_id %in% c(
    "vanilla__mu", "vanilla__pi",
    "vanilla__vanilla_warm_from_mu", "vanilla__vanilla_warm_from_pi"
  )) %>%
  dplyr::transmute(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    chain = dplyr::if_else(grepl("mu", .data$pair_id), "mu", "pi"),
    pair_kind = dplyr::if_else(
      .data$pair_id %in% c("vanilla__mu", "vanilla__pi"),
      "cold_ann", "cold_warm"
    ),
    basin_drift = .data$weighted_r2_drift
  )

jsd_long_pf <- pip_drift_summary %>%
  dplyr::filter(.data$pair_kind %in% c("vanilla_to_treatment",
                                       "vanilla_to_warm")) %>%
  dplyr::transmute(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    chain = .data$chain,
    pair_kind = dplyr::if_else(
      .data$pair_kind == "vanilla_to_treatment", "cold_ann", "cold_warm"
    ),
    pip_jsd = .data$pip_jsd
  )

basin_wide_pf <- basin_long_pf %>%
  tidyr::pivot_wider(names_from = chain, values_from = basin_drift,
                     names_prefix = "basin_")
jsd_wide_pf <- jsd_long_pf %>%
  tidyr::pivot_wider(names_from = chain, values_from = pip_jsd,
                     names_prefix = "jsd_")
joined_pf <- basin_wide_pf %>%
  dplyr::inner_join(jsd_wide_pf, by = c("dataset_bundle_id", "pair_kind"))

basin_max_pf <- max(0.05,
  max(joined_pf$basin_mu, joined_pf$basin_pi, na.rm = TRUE)) + 0.01
jsd_max_pf <- max(0.05,
  max(joined_pf$jsd_mu, joined_pf$jsd_pi, na.rm = TRUE)) + 0.02
basin_color_max <- max(joined_pf$basin_mu, na.rm = TRUE)
jsd_color_max <- max(joined_pf$jsd_mu, na.rm = TRUE)

make_b_panel <- function(df, x_col, y_col, color_col,
                         x_label, y_label, color_title, title_lab,
                         color_lim_max, max_val,
                         show_y = TRUE,
                         show_legend = TRUE) {
  ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]],
                 color = .data[[color_col]])) +
    geom_abline(slope = 1, intercept = 0,
                color = "grey50", linetype = "dashed", linewidth = 0.35) +
    geom_point(alpha = 0.62, size = 2.25, stroke = 0) +
    scale_color_viridis_c(
      option = "C",
      limits = c(0, color_lim_max),
      breaks = c(0, color_lim_max),
      labels = c("low", "high"),
      guide = if (show_legend) {
        guide_colorbar(
          title.position = "top",
          title.hjust = 0.5,
          label.position = "bottom",
          barwidth  = grid::unit(2.3, "cm"),
          barheight = grid::unit(0.24, "cm"),
          direction = "horizontal"
        )
      } else {
        "none"
      }
    ) +
    scale_x_continuous(expand = expansion(mult = c(0.015, 0.035))) +
    scale_y_continuous(expand = expansion(mult = c(0.015, 0.035))) +
    coord_cartesian(xlim = c(-0.02, max_val), ylim = c(-0.01, max_val),
                    expand = FALSE) +
    labs(x = x_label, y = y_label, color = color_title,
         title = title_lab) +
    theme_effect_paper(base_size = 7.2) +
    theme(
      plot.title = element_text(size = 7.6, face = "bold",
                                margin = margin(b = 0.5)),
      axis.title.x = element_text(size = 7.0, margin = margin(t = 3)),
      axis.title.y = element_text(size = 7.0, margin = margin(r = 3)),
      axis.text = element_text(size = 6.2, color = "grey25"),
      legend.margin = margin(t = -3, b = -4),
      legend.box.margin = margin(t = -4),
      plot.margin = margin(0.5, 2.5, 1.0, 2.5),
      aspect.ratio = 1
    ) +
    {
      if (show_y) {
        theme()
      } else {
        theme(axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank())
      }
    }
}

joined_pf_cold_ann  <- joined_pf %>% dplyr::filter(pair_kind == "cold_ann")
joined_pf_cold_warm <- joined_pf %>% dplyr::filter(pair_kind == "cold_warm")

basin_color_title <- expression(paste("PIP JSD (", mu, ")"))
jsd_color_title   <- expression(paste("Basin drift (", mu, ")"))

pB1 <- make_b_panel(
  joined_pf_cold_ann, "basin_mu", "basin_pi", "jsd_mu",
  expr_basin_x, expr_basin_y, basin_color_title,
  "Cold -> annotated",
  color_lim_max = jsd_color_max, max_val = basin_max_pf,
  show_y = TRUE,
  show_legend = TRUE
)
pB2 <- make_b_panel(
  joined_pf_cold_warm, "basin_mu", "basin_pi", "jsd_mu",
  expr_basin_x, expr_basin_y, basin_color_title,
  "Cold -> warm vanilla",
  color_lim_max = jsd_color_max, max_val = basin_max_pf,
  show_y = FALSE,
  show_legend = FALSE
)
pB3 <- make_b_panel(
  joined_pf_cold_ann, "jsd_mu", "jsd_pi", "basin_mu",
  expr_jsd_x, expr_jsd_y, jsd_color_title,
  "Cold -> annotated",
  color_lim_max = basin_color_max, max_val = jsd_max_pf,
  show_y = TRUE,
  show_legend = TRUE
)
pB4 <- make_b_panel(
  joined_pf_cold_warm, "jsd_mu", "jsd_pi", "basin_mu",
  expr_jsd_x, expr_jsd_y, jsd_color_title,
  "Cold -> warm vanilla",
  color_lim_max = basin_color_max, max_val = jsd_max_pf,
  show_y = FALSE,
  show_legend = FALSE
)

b_pair_header <- function(lbl) {
  ggplot() +
    annotate("text", x = 0.5, y = 0.38, label = lbl,
             size = 2.5, fontface = "bold") +
    xlim(0, 1) + ylim(0, 1) +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0))
}

b_panel_title <- function(lbl) {
  ggplot() +
    annotate("text", x = -0.055, y = 0.70, label = lbl,
             hjust = 0, size = 3.05, fontface = "bold") +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
    theme_void() +
    theme(plot.margin = margin(0, 0, 0, 0))
}

basin_pair_B <- (
  b_pair_header("Basin drift") /
    (pB1 | pB2) /
    guide_area()
) +
  plot_layout(
    heights = c(0.055, 1, 0.070),
    guides = "collect"
  ) &
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.title = element_text(size = 6.8),
    legend.text = element_text(size = 6.0),
    legend.margin = margin(t = -9, r = 0, b = -7, l = 0),
    legend.box.margin = margin(t = -12, r = 0, b = -6, l = 0)
  )

jsd_pair_B <- (
  b_pair_header("PIP JSD") /
    (pB3 | pB4) /
    guide_area()
) +
  plot_layout(
    heights = c(0.055, 1, 0.070),
    guides = "collect"
  ) &
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.title = element_text(size = 6.8),
    legend.text = element_text(size = 6.0),
    legend.margin = margin(t = -9, r = 0, b = -7, l = 0),
    legend.box.margin = margin(t = -12, r = 0, b = -6, l = 0)
  )

panel_C <- (
  b_panel_title("C. Effect-level drift diagnostics (mu vs pi)") /
    (basin_pair_B | jsd_pair_B)
) +
  plot_layout(heights = c(0.060, 1))

auprc_lookup_pf <- auprc_per_dataset %>%
  dplyr::transmute(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    delta_mu_treat = .data$auprc_delta_mu,
    delta_mu_warm  = .data$auprc_delta_warm_mu_vs_vanilla,
    delta_pi_treat = .data$auprc_delta_pi,
    delta_pi_warm  = .data$auprc_delta_warm_pi_vs_vanilla
  )

panel_B_long <- basin_long_pf %>%
  dplyr::left_join(auprc_lookup_pf, by = "dataset_bundle_id") %>%
  dplyr::mutate(
    auprc_delta = dplyr::case_when(
      chain == "mu" & pair_kind == "cold_ann"  ~ delta_mu_treat,
      chain == "mu" & pair_kind == "cold_warm" ~ delta_mu_warm,
      chain == "pi" & pair_kind == "cold_ann"  ~ delta_pi_treat,
      chain == "pi" & pair_kind == "cold_warm" ~ delta_pi_warm
    ),
    drift_bucket = cut(
      basin_drift,
      breaks = c(-Inf, 0.05, Inf),
      labels = c("<0.05", ">=0.05"),
      right = FALSE
    ),
    chain = factor(chain, levels = c("mu", "pi")),
    pair_kind = factor(
      pair_kind,
      levels = c("cold_ann", "cold_warm"),
      labels = c("cold -> annotated", "cold -> vanilla_warm")
    )
  ) %>%
  dplyr::filter(is.finite(.data$auprc_delta), !is.na(.data$drift_bucket))

bucket_counts <- panel_B_long %>%
  dplyr::count(.data$chain, .data$pair_kind, .data$drift_bucket)
y_top <- max(panel_B_long$auprc_delta, na.rm = TRUE) * 1.08

panel_B <- ggplot(
  panel_B_long,
  aes(x = .data$drift_bucket,
      y = .data$auprc_delta,
      fill = .data$drift_bucket)
) +
  geom_hline(yintercept = 0, color = "grey45", linetype = "dashed") +
  geom_boxplot(alpha = 0.65, outlier.size = 0.6, width = 0.65) +
  geom_text(
    data = bucket_counts,
    aes(x = .data$drift_bucket, y = y_top,
        label = sprintf("n=%d", .data$n)),
    inherit.aes = FALSE, size = 2.1, color = "grey25"
  ) +
  facet_grid(
    chain ~ pair_kind,
    labeller = labeller(chain = chain_labeller,
                        pair_kind = pair_kind_labeller)
  ) +
  scale_fill_manual(values = c("<0.05" = "#878787", ">=0.05" = "#c44e52")) +
  labs(
    x = "Basin drift bucket",
    y = expr_dauprc,
    fill = NULL,
    title = "B. AUPRC change by basin-drift bucket"
  ) +
  guides(fill = "none") +
  theme_effect_paper(base_size = 8) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 6.8),
    plot.title = element_text(face = "bold", size = 9.2,
                              margin = margin(b = 3))
  )

panel_D_df <- density_shift_df %>%
  dplyr::filter(.data$comparison_kind %in% c(
    "vanilla_to_treatment", "vanilla_to_warm"
  )) %>%
  dplyr::mutate(
    comparison_kind = factor(
      .data$comparison_kind,
      levels = c("vanilla_to_treatment", "vanilla_to_warm"),
      labels = c("cold -> annotated", "cold -> vanilla_warm")
    ),
    chain = factor(.data$chain, levels = c("mu", "pi"))
  )

x_label_paper <- expression(paste(
  "Effect diffuseness: effective top-95% support size  ", K[l]
))
y_label_paper <- expression(paste(
  "Effect accuracy: ", A[l], " = max causal PIP / max SNP PIP"
))

panel_D_core <- ggplot(panel_D_df,
                  aes(x = .data$x, y = .data$y,
                      fill = .data$density_diff)) +
  geom_raster(interpolate = TRUE) +
  geom_contour(aes(z = .data$density_diff),
               color = "grey20", alpha = 0.45, linewidth = 0.18) +
  facet_grid(
    chain ~ comparison_kind,
    labeller = labeller(chain = chain_labeller,
                        comparison_kind = pair_kind_labeller)
  ) +
  scale_fill_gradient2(
    low = "#2b6cb0", mid = "white", high = "#c53030",
    midpoint = 0, limits = c(-shared_fill_lim, shared_fill_lim),
    oob = scales::squish
  ) +
  keff_x_scale(xlim = x_limits_resolved) +
  coord_cartesian(xlim = x_limits_resolved,
                  ylim = y_limits_resolved,
                  expand = FALSE) +
  labs(
    x = x_label_paper, y = y_label_paper,
    fill = "destination - source\n(PVE-weighted density)",
    title = "D. Per-effect quality shifts"
  ) +
  guides(fill = guide_colorbar(
    title.position = "top",
    title.hjust = 0.5,
    barwidth = grid::unit(0.34, "cm"),
    barheight = grid::unit(2.4, "cm"),
    direction = "vertical"
  )) +
  theme_effect_paper(base_size = 8) +
  theme(
    legend.position = "right",
    legend.justification = "center",
    legend.margin = margin(0, 0, 0, 2),
    legend.box.margin = margin(0, 0, 0, 2),
    panel.spacing.y = grid::unit(1.25, "mm"),
    plot.title = element_text(face = "bold", size = 9.2,
                              margin = margin(b = 3))
  )

panel_D <- (panel_D_core + guide_area()) +
  plot_layout(widths = c(1, 0.13), guides = "collect") &
  theme(legend.position = "right")

top_summary_row <- (wrap_elements(full = panel_A) | wrap_elements(full = panel_B)) +
  plot_layout(widths = c(0.52, 0.48))

combined_fig <- (top_summary_row / panel_C / panel_D) +
  plot_layout(heights = c(1, 1, 1))

ggsave(
  paper_figure_png, combined_fig,
  width = 8.8, height = 9.9, dpi = 300, bg = "white",
  limitsize = FALSE
)
cat("Wrote combined paper figure to ", paper_figure_png, "\n", sep = "")
print(combined_fig)
```
