---
title: "Effect Matching Threshold Application (3-arm, same-X null, fixed threshold)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Applies the operating threshold (`corr_min = 0.99`, `margin_min = 0`,
`var_min = 0`) — pre-validated in
`effect_matching_threshold_calibration_same_X.Rmd` — to two pairwise matchings
in the 3-arm refit:

- `susie_vanilla` <-> `susine_functional_mu`
- `susie_vanilla` <-> `susie_functional_pi`

Mu<->pi matching is intentionally skipped for this iteration. The same-X
within-matrix derangement supplies the null. No grid sweep is performed; the
threshold is fixed. A real-vs-null correlation diagnostic per arm verifies
that the fixed threshold still cleanly separates the distributions.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/effect_matching_threshold_calibration_3arm.Rmd")
library(here)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

# 1. Inputs

```{r inputs}
output_root <- here("output")

study_name <- "baseline_sims_3arm_effect_matching"
study_dir <- file.path(output_root, "effect_matching_study", study_name)
records_rds <- file.path(study_dir, "effect_refit_records.rds")

baseline_method <- "susie_vanilla"
treatment_methods <- c("susine_functional_mu", "susie_functional_pi")
treatment_short <- c(
  susine_functional_mu = "mu",
  susie_functional_pi = "pi"
)

# Operating threshold (locked from the same-X 2-arm calibration).
fixed_corr_min <- 0.99
fixed_margin_min <- 0
fixed_var_min <- 0

near_zero_sd <- 1e-12

# Per-treatment output filenames are built inside the per-arm loop using
# treatment_short[[treatment]].

cat("Study dir:", study_dir, "\n")
cat("Records RDS:", records_rds, "\n")
cat("Baseline method:", baseline_method, "\n")
cat("Treatment methods:", paste(treatment_methods, collapse = ", "), "\n")
cat(sprintf(
  "Fixed threshold: corr_min = %.3f | margin_min = %.3f | var_min = %.4f\n",
  fixed_corr_min, fixed_margin_min, fixed_var_min
))
```

# 2. Helpers

```{r helpers}
format_duration <- function(seconds) {
  seconds <- as.numeric(seconds)
  if (!is.finite(seconds)) {
    return("unknown")
  }
  h <- floor(seconds / 3600)
  m <- floor((seconds %% 3600) / 60)
  s <- round(seconds %% 60)
  sprintf("%02d:%02d:%02d", h, m, s)
}

records_to_matrix <- function(rows) {
  rows <- rows %>% dplyr::arrange(.data$effect)
  mat <- do.call(cbind, rows$fitted_y_effect)
  storage.mode(mat) <- "double"
  colnames(mat) <- as.character(rows$effect)
  mat
}

safe_cor_matrix <- function(vanilla_mat, mu_mat, sd_eps = near_zero_sd) {
  if (nrow(vanilla_mat) != nrow(mu_mat)) {
    stop(
      "Mismatched n between baseline and treatment fitted-y matrices: ",
      nrow(vanilla_mat), " vs ", nrow(mu_mat),
      ". Same-X null assumes shared sample dimension."
    )
  }

  v_sd <- apply(vanilla_mat, 2L, stats::sd)
  m_sd <- apply(mu_mat, 2L, stats::sd)
  active_v <- is.finite(v_sd) & v_sd > sd_eps
  active_m <- is.finite(m_sd) & m_sd > sd_eps

  cor_mat <- matrix(
    0,
    nrow = ncol(vanilla_mat),
    ncol = ncol(mu_mat),
    dimnames = list(colnames(vanilla_mat), colnames(mu_mat))
  )

  if (any(active_v) && any(active_m)) {
    sub_cor <- suppressWarnings(stats::cor(
      vanilla_mat[, active_v, drop = FALSE],
      mu_mat[, active_m, drop = FALSE],
      use = "pairwise.complete.obs"
    ))
    sub_cor[!is.finite(sub_cor)] <- 0
    cor_mat[active_v, active_m] <- sub_cor
  }

  list(cor = cor_mat, active_vanilla = active_v, active_mu = active_m)
}

second_best_margin <- function(values, chosen_idx) {
  if (length(values) <= 1L) {
    return(Inf)
  }
  alternatives <- values[-chosen_idx]
  best_alt <- max(alternatives, na.rm = TRUE)
  values[[chosen_idx]] - best_alt
}

safe_min_numeric <- function(...) {
  x <- c(...)
  x <- x[is.finite(x)]
  if (!length(x)) {
    return(NA_real_)
  }
  min(x)
}

effect_metric_cols <- c(
  "size", "purity", "coverage",
  "effect_pip_entropy", "effect_pip_entropy_core95",
  "effect_k_eff_signal", "effect_k_eff_signal_core95",
  "tail_inflation_ratio", "tail_inflation_log",
  "accuracy_ratio", "var_y_explained_frac", "var_y_explained_pct"
)

add_effect_metric_columns <- function(out, vanilla_rows, mu_rows, vanilla_idx, mu_idx) {
  available <- intersect(effect_metric_cols, intersect(names(vanilla_rows), names(mu_rows)))
  for (nm in available) {
    out[[paste0(nm, "_vanilla")]] <- vanilla_rows[[nm]][vanilla_idx]
    out[[paste0(nm, "_mu")]] <- mu_rows[[nm]][mu_idx]
    out[[paste0(nm, "_delta_mu_minus_vanilla")]] <-
      mu_rows[[nm]][mu_idx] - vanilla_rows[[nm]][vanilla_idx]
  }
  out
}

compare_dataset_pair <- function(vanilla_rows,
                                 mu_rows,
                                 vanilla_dataset_id,
                                 mu_dataset_id,
                                 comparison_type,
                                 comparison_id,
                                 matrix_id = NA_character_,
                                 include_effect_metrics = FALSE) {
  vanilla_rows <- vanilla_rows %>% dplyr::arrange(.data$effect)
  mu_rows <- mu_rows %>% dplyr::arrange(.data$effect)

  vanilla_mat <- records_to_matrix(vanilla_rows)
  mu_mat <- records_to_matrix(mu_rows)
  cor_info <- safe_cor_matrix(vanilla_mat, mu_mat)
  cor_mat <- cor_info$cor

  row_best <- max.col(cor_mat, ties.method = "first")
  col_best <- max.col(t(cor_mat), ties.method = "first")

  candidate_rows <- vector("list", nrow(cor_mat))
  n_candidates <- 0L

  for (i in seq_len(nrow(cor_mat))) {
    j <- row_best[[i]]
    if (col_best[[j]] != i) {
      next
    }

    n_candidates <- n_candidates + 1L
    row_margin <- second_best_margin(cor_mat[i, ], j)
    col_margin <- second_best_margin(cor_mat[, j], i)

    out <- tibble::tibble(
      comparison_type = comparison_type,
      comparison_id = comparison_id,
      matrix_id = as.character(matrix_id),
      vanilla_dataset_bundle_id = vanilla_dataset_id,
      mu_dataset_bundle_id = mu_dataset_id,
      vanilla_effect = as.integer(rownames(cor_mat)[[i]]),
      mu_effect = as.integer(colnames(cor_mat)[[j]]),
      corr = as.numeric(cor_mat[i, j]),
      row_margin = as.numeric(row_margin),
      col_margin = as.numeric(col_margin),
      margin = min(row_margin, col_margin),
      var_y_explained_frac_vanilla = as.numeric(vanilla_rows$var_y_explained_frac[[i]]),
      var_y_explained_frac_mu = as.numeric(mu_rows$var_y_explained_frac[[j]]),
      pair_var_explained_frac = safe_min_numeric(
        as.numeric(vanilla_rows$var_y_explained_frac[[i]]),
        as.numeric(mu_rows$var_y_explained_frac[[j]])
      ),
      active_vanilla = as.logical(cor_info$active_vanilla[[i]]),
      active_mu = as.logical(cor_info$active_mu[[j]]),
      mutual_best = TRUE
    )

    if (include_effect_metrics) {
      out <- add_effect_metric_columns(out, vanilla_rows, mu_rows, i, j)
    }

    candidate_rows[[n_candidates]] <- out
  }

  if (!n_candidates) {
    return(tibble::tibble())
  }

  dplyr::bind_rows(candidate_rows[seq_len(n_candidates)])
}

run_pairwise_matching <- function(records,
                                  baseline_method,
                                  treatment_method,
                                  short_label) {
  needed_methods <- c(baseline_method, treatment_method)
  pair_records <- records %>%
    dplyr::filter(.data$model_method %in% needed_methods)

  method_counts <- pair_records %>%
    dplyr::count(.data$dataset_bundle_id, .data$model_method, name = "n_effects")

  complete_ids <- method_counts %>%
    dplyr::filter(.data$n_effects > 0L) %>%
    dplyr::count(.data$dataset_bundle_id, name = "n_methods") %>%
    dplyr::filter(.data$n_methods == length(needed_methods)) %>%
    dplyr::pull(.data$dataset_bundle_id) %>%
    sort()

  if (length(complete_ids) < 2L) {
    stop(
      "Need at least two complete datasets for ", treatment_method,
      "; found ", length(complete_ids), "."
    )
  }

  pair_records <- pair_records %>%
    dplyr::filter(.data$dataset_bundle_id %in% complete_ids)

  vanilla_by_dataset <- pair_records %>%
    dplyr::filter(.data$model_method == baseline_method) %>%
    split(.$dataset_bundle_id)

  treatment_by_dataset <- pair_records %>%
    dplyr::filter(.data$model_method == treatment_method) %>%
    split(.$dataset_bundle_id)

  dataset_to_matrix <- pair_records %>%
    dplyr::distinct(.data$dataset_bundle_id, .data$matrix_id) %>%
    dplyr::filter(.data$dataset_bundle_id %in% complete_ids)

  cat("\n--- Treatment arm:", treatment_method, "---\n")
  cat("Complete datasets:", length(complete_ids), "\n")
  cat("Distinct matrices:", dplyr::n_distinct(dataset_to_matrix$matrix_id), "\n")

  # Real candidates -------------------------------------------------------
  real_started <- proc.time()[["elapsed"]]
  real_candidates <- purrr::map_dfr(
    seq_along(complete_ids),
    function(i) {
      dataset_id <- complete_ids[[i]]
      matrix_id_i <- dataset_to_matrix$matrix_id[
        match(dataset_id, dataset_to_matrix$dataset_bundle_id)
      ]
      out <- compare_dataset_pair(
        vanilla_rows = vanilla_by_dataset[[dataset_id]],
        mu_rows = treatment_by_dataset[[dataset_id]],
        vanilla_dataset_id = dataset_id,
        mu_dataset_id = dataset_id,
        comparison_type = "real",
        comparison_id = paste0("real_", dataset_id),
        matrix_id = matrix_id_i,
        include_effect_metrics = TRUE
      )

      if (i %% 100L == 0L || i == length(complete_ids)) {
        elapsed <- proc.time()[["elapsed"]] - real_started
        eta <- elapsed / i * (length(complete_ids) - i)
        cat(sprintf(
          "[%s] Real candidates %d/%d | elapsed: %s | ETA: %s\n",
          short_label, i, length(complete_ids),
          format_duration(elapsed), format_duration(eta)
        ))
      }
      out
    }
  )

  # Null candidates: enumerated within-matrix mismatched pairs ------------
  null_pair_index <- dataset_to_matrix %>%
    dplyr::inner_join(
      dataset_to_matrix %>%
        dplyr::rename(mu_dataset_bundle_id = .data$dataset_bundle_id),
      by = "matrix_id",
      relationship = "many-to-many"
    ) %>%
    dplyr::filter(.data$dataset_bundle_id != .data$mu_dataset_bundle_id) %>%
    dplyr::transmute(
      matrix_id = .data$matrix_id,
      vanilla_dataset_bundle_id = as.character(.data$dataset_bundle_id),
      mu_dataset_bundle_id = as.character(.data$mu_dataset_bundle_id),
      pair_index = dplyr::row_number()
    )

  if (!nrow(null_pair_index)) {
    stop(
      "No within-matrix mismatched pairs available for ", treatment_method,
      "; either all matrices have one seed or matrix_id is missing."
    )
  }

  cat(sprintf("[%s] Within-matrix null pairs: %d\n",
              short_label, nrow(null_pair_index)))

  null_started <- proc.time()[["elapsed"]]
  null_candidate_chunks <- vector("list", nrow(null_pair_index))
  for (i in seq_len(nrow(null_pair_index))) {
    vanilla_id <- null_pair_index$vanilla_dataset_bundle_id[[i]]
    mu_id <- null_pair_index$mu_dataset_bundle_id[[i]]
    matrix_i <- null_pair_index$matrix_id[[i]]

    null_candidate_chunks[[i]] <- compare_dataset_pair(
      vanilla_rows = vanilla_by_dataset[[vanilla_id]],
      mu_rows = treatment_by_dataset[[mu_id]],
      vanilla_dataset_id = vanilla_id,
      mu_dataset_id = mu_id,
      comparison_type = "null",
      comparison_id = paste0("null_", matrix_i, "_", vanilla_id, "_", mu_id),
      matrix_id = matrix_i,
      include_effect_metrics = FALSE
    )

    if (i %% 200L == 0L || i == nrow(null_pair_index)) {
      elapsed <- proc.time()[["elapsed"]] - null_started
      eta <- elapsed / i * (nrow(null_pair_index) - i)
      cat(sprintf(
        "[%s] Null pairs %d/%d | elapsed: %s | ETA: %s\n",
        short_label, i, nrow(null_pair_index),
        format_duration(elapsed), format_duration(eta)
      ))
    }
  }

  null_candidates <- dplyr::bind_rows(null_candidate_chunks)
  n_fixed <- sum(
    null_candidates$vanilla_dataset_bundle_id == null_candidates$mu_dataset_bundle_id
  )
  if (n_fixed > 0L) {
    stop(n_fixed, " null candidate rows have vanilla_dataset == mu_dataset.")
  }

  list(
    real_candidates = real_candidates,
    null_candidates = null_candidates,
    null_pair_index = null_pair_index,
    n_complete = length(complete_ids)
  )
}

apply_fixed_threshold <- function(candidates, total_comparisons,
                                  corr_min, margin_min, var_min) {
  if (!nrow(candidates)) {
    return(list(
      n_matches = 0L,
      matches_per_pair = 0,
      any_match_rate = 0,
      filtered = candidates
    ))
  }
  keep <- candidates$corr >= corr_min &
    candidates$margin >= margin_min &
    candidates$pair_var_explained_frac >= var_min
  keep[is.na(keep)] <- FALSE
  filtered <- candidates[keep, , drop = FALSE]
  any_rate <- if (nrow(filtered) > 0L) {
    length(unique(filtered$comparison_id)) / total_comparisons
  } else {
    0
  }
  list(
    n_matches = nrow(filtered),
    matches_per_pair = nrow(filtered) / total_comparisons,
    any_match_rate = any_rate,
    filtered = filtered
  )
}
```

# 3. Load Records

```{r load-records}
stopifnot(file.exists(records_rds))

records <- readRDS(records_rds) %>%
  dplyr::mutate(
    dataset_bundle_id = as.character(.data$dataset_bundle_id),
    matrix_id = as.character(.data$matrix_id)
  )

methods_present <- sort(unique(records$model_method))
cat("Methods present in records:", paste(methods_present, collapse = ", "), "\n")

required_methods <- c(baseline_method, treatment_methods)
missing_methods <- setdiff(required_methods, methods_present)
if (length(missing_methods)) {
  stop(
    "Records are missing the following expected methods: ",
    paste(missing_methods, collapse = ", "),
    ". Re-run effect_matching_refit_export_3arm.Rmd first."
  )
}
```

# 4. Per-Treatment Matching, Threshold Application, And Diagnostics

```{r per-treatment-matching, fig.width=9, fig.height=4.5}
arm_summaries <- vector("list", length(treatment_methods))

for (k in seq_along(treatment_methods)) {
  treatment <- treatment_methods[[k]]
  short_label <- treatment_short[[treatment]]

  result <- run_pairwise_matching(
    records = records,
    baseline_method = baseline_method,
    treatment_method = treatment,
    short_label = short_label
  )

  real_candidates_csv <- file.path(
    study_dir,
    sprintf("real_match_candidates_vanilla_%s_same_X.csv", short_label)
  )
  null_candidates_csv <- file.path(
    study_dir,
    sprintf("null_match_candidates_vanilla_%s_same_X.csv", short_label)
  )
  null_pair_index_csv <- file.path(
    study_dir,
    sprintf("null_pair_index_vanilla_%s_same_X.csv", short_label)
  )
  accepted_real_csv <- file.path(
    study_dir,
    sprintf("accepted_real_matches_vanilla_%s_fixed_threshold.csv", short_label)
  )
  diagnostic_png <- file.path(
    study_dir,
    sprintf("frontier_real_vs_null_vanilla_%s_same_X_fixed.png", short_label)
  )

  readr::write_csv(result$real_candidates, real_candidates_csv)
  readr::write_csv(result$null_candidates, null_candidates_csv)
  readr::write_csv(result$null_pair_index, null_pair_index_csv)

  # Apply the fixed threshold to both real and null candidates so we can
  # report the operational match rates and (for null) the implied FDR.
  real_apply <- apply_fixed_threshold(
    result$real_candidates,
    total_comparisons = result$n_complete,
    corr_min = fixed_corr_min,
    margin_min = fixed_margin_min,
    var_min = fixed_var_min
  )
  null_apply <- apply_fixed_threshold(
    result$null_candidates,
    total_comparisons = nrow(result$null_pair_index),
    corr_min = fixed_corr_min,
    margin_min = fixed_margin_min,
    var_min = fixed_var_min
  )

  readr::write_csv(real_apply$filtered, accepted_real_csv)

  implied_fdr <- if (real_apply$n_matches > 0L) {
    null_apply$matches_per_pair * result$n_complete / real_apply$n_matches
  } else {
    NA_real_
  }

  arm_summaries[[k]] <- tibble::tibble(
    treatment = treatment,
    short_label = short_label,
    n_complete_datasets = result$n_complete,
    n_null_pairs = nrow(result$null_pair_index),
    n_real_candidates = nrow(result$real_candidates),
    n_null_candidates = nrow(result$null_candidates),
    n_real_accepted = real_apply$n_matches,
    real_matches_per_dataset = real_apply$matches_per_pair,
    null_matches_per_pair = null_apply$matches_per_pair,
    implied_fdr = implied_fdr
  )

  cat(sprintf(
    "[%s] Real candidates: %d | Null candidates: %d\n",
    short_label, nrow(result$real_candidates), nrow(result$null_candidates)
  ))
  cat(sprintf(
    "[%s] @ threshold (%.2f / %.2f / %.4f): real accepted = %d (%.3f per dataset) | null accepted/pair = %.4f | implied FDR = %.4f\n",
    short_label,
    fixed_corr_min, fixed_margin_min, fixed_var_min,
    real_apply$n_matches, real_apply$matches_per_pair,
    null_apply$matches_per_pair, implied_fdr
  ))

  # Diagnostic: real vs null correlation distribution (mutual-best candidates)
  corr_compare <- dplyr::bind_rows(
    result$real_candidates %>%
      dplyr::transmute(corr = .data$corr, comparison_type = "real"),
    result$null_candidates %>%
      dplyr::transmute(corr = .data$corr, comparison_type = "null")
  )

  p_corr <- ggplot(corr_compare, aes(x = .data$corr, fill = .data$comparison_type)) +
    geom_histogram(
      aes(y = after_stat(density)),
      position = "identity",
      bins = 60,
      alpha = 0.55
    ) +
    geom_vline(xintercept = fixed_corr_min, color = "black", linetype = "dashed") +
    scale_fill_manual(values = c(real = "#2b8cbe", null = "#c53030")) +
    labs(
      x = "Mutual-best fitted-y correlation",
      y = "Density",
      fill = "Comparison",
      title = sprintf(
        "vanilla vs %s: real vs same-X null correlations (mutual-best)",
        treatment
      ),
      subtitle = sprintf(
        "Fixed threshold corr_min = %.2f shown as dashed line",
        fixed_corr_min
      )
    ) +
    theme(legend.position = "bottom")

  ggsave(diagnostic_png, p_corr, width = 9, height = 4.5, dpi = 150)
  print(p_corr)
}

arm_summary <- dplyr::bind_rows(arm_summaries)
arm_summary_csv <- file.path(study_dir, "fixed_threshold_arm_summary.csv")
readr::write_csv(arm_summary, arm_summary_csv)
```

# 5. Cross-Arm Summary

```{r cross-arm-summary}
cat("Per-arm matching summary at fixed threshold:\n")
print(arm_summary)

kable(arm_summary, digits = 4)

cat("Saved per-arm CSVs and diagnostic plots to:", study_dir, "\n")
cat("Summary CSV:", arm_summary_csv, "\n")
```
