---
title: "JSD Distance Exploration — Categorical vs Bernoulli"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 6)
library(dplyr)
library(tidyr)
library(ggplot2)
devtools::load_all(here::here("..", "susine"))
devtools::load_all(here::here())
```

# Purpose

Build intuition for what categorical and Bernoulli JSD values mean for
pairs of PIP vectors. We construct synthetic PIP vectors with p=600 and
controlled differences, then compute both JSD types.

- **Categorical JSD** (`js_distance`): treats PIP vector as an unnormalized
  categorical distribution (adds epsilon, computes KL). Unbounded above.
- **Bernoulli JSD** (`js_distance_bernoulli`): treats each PIP as an
  independent Bernoulli parameter. Bounded in [0, ln(2)] ~ [0, 0.693].
  Averaged across all p variants.

```{r helpers}
p <- 600L

# Make a PIP vector: set specified indices to `value`, rest to 0
make_pips <- function(indices, value = 1.0, p = 600L) {
  v <- rep(0, p)
  v[indices] <- value
  v
}

# Compute both JSD types for a pair and return a 1-row tibble
compare_pair <- function(pips_a, pips_b, label = "") {
  tibble::tibble(
    label = label,
    n_diff = sum(pips_a != pips_b),
    cat_jsd = js_distance(pips_a, pips_b),
    bern_jsd = js_distance_bernoulli(pips_a, pips_b)
  )
}
```

# Group 1: Binary PIPs (0 or 1), 10 active variants

Each PIP vector has 10 variants at PIP=1, rest at 0 (sums to 10).
We swap 1, 2, 3, ..., 10 variant assignments between vectors.

```{r group1}
base_idx <- 1:10
results <- list()

for (n_swap in 0:10) {
  pips_a <- make_pips(base_idx, value = 1.0)
  # Swap: move n_swap indices from base_idx to new positions
  new_idx <- base_idx
  if (n_swap > 0) {
    new_idx[1:n_swap] <- (10 + 1):(10 + n_swap)
  }
  pips_b <- make_pips(new_idx, value = 1.0)
  results[[length(results) + 1]] <- compare_pair(
    pips_a, pips_b,
    label = sprintf("binary_10active_%dswap", n_swap)
  )
}

group1 <- bind_rows(results)
group1$group <- "Binary (PIP=1, 10 active)"
print(group1)
```

# Group 2: Half-weight PIPs (0 or 0.5), 20 active variants

Each vector has 20 variants at PIP=0.5 (sums to 10).
Swap 1, 2, 4, 6, 10, 20 variant assignments.

```{r group2}
base_idx <- 1:20
results <- list()

for (n_swap in c(0, 1, 2, 4, 6, 10, 20)) {
  pips_a <- make_pips(base_idx, value = 0.5)
  new_idx <- base_idx
  if (n_swap > 0) {
    new_idx[1:n_swap] <- (20 + 1):(20 + n_swap)
  }
  pips_b <- make_pips(new_idx, value = 0.5)
  results[[length(results) + 1]] <- compare_pair(
    pips_a, pips_b,
    label = sprintf("half_20active_%dswap", n_swap)
  )
}

group2 <- bind_rows(results)
group2$group <- "Half-weight (PIP=0.5, 20 active)"
print(group2)
```

# Group 3: Diffuse PIPs (0 or 0.1), 100 active variants

Each vector has 100 variants at PIP=0.1 (sums to 10).
Swap 2, 5, 10, 20, 50, 100 variant assignments.

```{r group3}
base_idx <- 1:100
results <- list()

for (n_swap in c(0, 2, 5, 10, 20, 50, 100)) {
  pips_a <- make_pips(base_idx, value = 0.1)
  new_idx <- base_idx
  if (n_swap > 0) {
    new_idx[1:n_swap] <- (100 + 1):(100 + n_swap)
  }
  pips_b <- make_pips(new_idx, value = 0.1)
  results[[length(results) + 1]] <- compare_pair(
    pips_a, pips_b,
    label = sprintf("diffuse_100active_%dswap", n_swap)
  )
}

group3 <- bind_rows(results)
group3$group <- "Diffuse (PIP=0.1, 100 active)"
print(group3)
```

# Group 4: Realistic SuSiE-like PIPs

Two CS-like structures: a few high-PIP variants with LD-spread tails.

```{r group4}
results <- list()

# Baseline: 3 "credible sets" each with 1 lead + 2 proxies
# CS1: variants 1 (PIP=0.8), 2-3 (PIP=0.05 each)
# CS2: variants 10 (PIP=0.7), 11-12 (PIP=0.10 each)
# CS3: variants 20 (PIP=0.9), 21 (PIP=0.05)
# Total PIP mass ~ 0.8+0.1+0.7+0.2+0.9+0.05 = 2.75
make_realistic_a <- function() {
  v <- rep(0, p)
  v[1] <- 0.8; v[2] <- 0.05; v[3] <- 0.05
  v[10] <- 0.7; v[11] <- 0.10; v[12] <- 0.10
  v[20] <- 0.9; v[21] <- 0.05
  # Add low background noise
  v[50:55] <- 0.02
  v
}

# Variant B1: CS1 lead swaps to a different variant (single CS swap)
make_realistic_b1 <- function() {
  v <- make_realistic_a()
  v[1] <- 0.05  # demote original lead
  v[4] <- 0.8   # new lead in same CS region
  v
}

# Variant B2: CS1 and CS2 both swap leads (two CS swaps)
make_realistic_b2 <- function() {
  v <- make_realistic_a()
  v[1] <- 0.05; v[4] <- 0.8    # CS1 swap
  v[10] <- 0.10; v[13] <- 0.7  # CS2 swap
  v
}

# Variant B3: completely different CS allocation (3 CS swaps)
make_realistic_b3 <- function() {
  v <- rep(0, p)
  v[4] <- 0.8; v[5] <- 0.05; v[6] <- 0.05
  v[13] <- 0.7; v[14] <- 0.10; v[15] <- 0.10
  v[25] <- 0.9; v[26] <- 0.05
  v[50:55] <- 0.02
  v
}

pips_a <- make_realistic_a()
results[[1]] <- compare_pair(pips_a, pips_a, "realistic_identical")
results[[2]] <- compare_pair(pips_a, make_realistic_b1(), "realistic_1cs_swap")
results[[3]] <- compare_pair(pips_a, make_realistic_b2(), "realistic_2cs_swap")
results[[4]] <- compare_pair(pips_a, make_realistic_b3(), "realistic_3cs_swap")

group4 <- bind_rows(results)
group4$group <- "Realistic (CS-like, mixed PIPs)"
print(group4)
```

# Summary table

```{r summary}
all_results <- bind_rows(group1, group2, group3, group4)
cat("All pairs:\n")
print(as.data.frame(all_results), digits = 6)
```

# Summary plot: Categorical vs Bernoulli JSD

```{r summary-plot, fig.width=10, fig.height=7}
p_summary <- ggplot(all_results, aes(x = bern_jsd, y = cat_jsd, color = group)) +
  geom_point(size = 3) +
  geom_text(aes(label = n_diff), hjust = -0.3, vjust = -0.3, size = 2.5) +
  scale_x_continuous("Bernoulli JSD (bounded [0, ln(2)])") +
  scale_y_continuous("Categorical JSD (unbounded)") +
  labs(
    title = "Categorical vs Bernoulli JSD across synthetic PIP pairs",
    subtitle = "Labels = number of differing variants. p = 600.",
    color = "Group"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
print(p_summary)
```

# Per-group scaling plots

```{r group-plots, fig.width=12, fig.height=5}
plot_data <- all_results %>%
  pivot_longer(cols = c(cat_jsd, bern_jsd),
               names_to = "jsd_type", values_to = "jsd_value") %>%
  mutate(jsd_type = ifelse(jsd_type == "cat_jsd", "Categorical", "Bernoulli"))

ggplot(plot_data, aes(x = n_diff, y = jsd_value, color = jsd_type)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~group, scales = "free") +
  labs(
    x = "Number of differing variants",
    y = "JSD value",
    color = "JSD type",
    title = "JSD scaling with number of PIP differences",
    subtitle = "Each group uses different PIP magnitudes and sparsity levels"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

# Clustering threshold implications

For each pair, show which Bernoulli JSD threshold bins would merge vs
separate them.

```{r threshold-table}
bjsd_cuts <- c(0.001, 0.002, 0.004, 0.006, 0.008, 0.01, 0.02, 0.05)
cat_cuts <- c(0.15, 0.50, 1.0, 2.0, 3.0, 5.0)

threshold_tbl <- all_results %>%
  select(label, group, n_diff, cat_jsd, bern_jsd)

for (th in bjsd_cuts) {
  col <- sprintf("bjsd_h%.3f", th)
  threshold_tbl[[col]] <- ifelse(all_results$bern_jsd <= th, "merged", "split")
}
for (th in cat_cuts) {
  col <- sprintf("cat_h%.2f", th)
  threshold_tbl[[col]] <- ifelse(all_results$cat_jsd <= th, "merged", "split")
}

cat("Threshold classification (merged = same cluster, split = different clusters):\n")
print(as.data.frame(threshold_tbl), right = FALSE)
```
