---
title: "Real-data LD-mismatch diligence: estimate_s_rss, kriging_rss, allele/strand audit"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Purpose

Reviewer-proofing the GTEx Lung real-data case study. The headline `YDJC`
finding is an `r = 0.98` variant pair that snaps to PIP near 1 with large
*opposite-sign* conditional effects. That is simultaneously the canonical
"signed annotation resolved a bidirectional cis-regulatory pair" success story
*and* the canonical reference-LD-mismatch / allele-flip artifact. The
warm-refit durability test varies the **prior** while holding `(z, R, n)`
fixed, so it cannot distinguish the two: both the annotated source fit and its
warm refit inhale the *same* reference `R`, and a shared `(z, R)`
misspecification is in the null space of any prior-only perturbation.

This workbook runs the three cheap, decisive `(z, R)`-consistency diagnostics
(no refitting, pure downstream functions of the synced bundle):

1. `susieR::estimate_s_rss` on all 20 loci -- a continuous, more sensitive
   version of the method-of-moments `sigma_hat^2` screen already in Methods.
2. `susieR::kriging_rss` on the responder loci -- per-variant expected `z`
   given LD neighbors. Its `logLR` is literally the log-LR for "this variant's
   `z` is sign-flipped relative to its LD" vs consistent, so it directly tests
   the YDJC opposite-sign story.
3. Allele orientation / strand-ambiguity audit. Upstream matching
   (`pipeline_ld.py`) keys on the exact `(chrom, pos, ref, alt)` tuple, which
   closes the orientation-swap flip pathway by construction; the only residual
   flip pathway is strand-ambiguous SNPs (A/T, C/G).

Run on HPC interactive, where the geometric_n20 bundles already exist.

```{r libs, include=FALSE}
devtools::load_all(here::here())
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(susieR)

manifest_path <- here::here(
  "data", "real_case_studies", "geometric_n20_loci", "locus_manifest.csv"
)
stopifnot(file.exists(manifest_path))

out_dir <- here::here(
  "output", "one_off_validations", "ld_mismatch_kriging_allele_audit"
)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
progress_log_path <- file.path(out_dir, "run_progress.log")
unlink(progress_log_path)

manifest <- readr::read_csv(manifest_path, show_col_types = FALSE)
all_loci <- manifest$locus_id

# Strongest + intermediate responders (off-zero weight) per the locus summary.
responder_loci <- all_loci[grepl(
  "^(ydjc|arsa|rrp7a|lgals9)_", all_loci, ignore.case = TRUE
)]

# The named YDJC headline pair, for a hard check.
ydjc_pair_ids <- c(
  "chr22_21715468_T_C_b38",
  "chr22_21715668_G_A_b38"
)

load_bundle <- function(locus_id) {
  test_susine:::load_real_data_locus_bundle(
    locus_id = locus_id,
    manifest_path = manifest_path,
    repo_root = here::here()
  )
}

safe_ld_z_diag <- function(bundle) {
  tryCatch(
    test_susine:::real_data_ld_z_consistency(
      R = bundle$R, z = bundle$z, n = bundle$n_sample
    ),
    error = function(e) list(
      mom_sigma2_hat = NA_real_,
      ld_chol_ok = NA,
      ld_z_inconsistency_flag = NA
    )
  )
}

is_strand_ambiguous <- function(ref, alt) {
  pair <- toupper(paste0(ref, alt))
  pair %in% c("AT", "TA", "CG", "GC")
}

safe_summary <- function(x, fn) {
  x <- x[is.finite(x)]
  if (!length(x)) return(NA_real_)
  fn(x)
}

fmt_duration <- function(seconds) {
  if (!is.finite(seconds)) return("NA")
  seconds <- max(0, as.numeric(seconds))
  if (seconds < 60) return(sprintf("%.1fs", seconds))
  if (seconds < 3600) {
    return(sprintf("%dm %.0fs", floor(seconds / 60), seconds %% 60))
  }
  sprintf("%dh %02dm", floor(seconds / 3600), floor((seconds %% 3600) / 60))
}

progress_line <- function(...) {
  line <- sprintf(...)
  cat(line, "\n", file = stderr())
  cat(line, "\n", file = progress_log_path, append = TRUE)
  flush.console()
}

progress_msg <- function(prefix, i, n, label, start_time, item_start = NULL) {
  elapsed <- as.numeric(difftime(Sys.time(), start_time, units = "secs"))
  mean_per_item <- elapsed / max(i, 1L)
  eta <- mean_per_item * max(n - i, 0L)
  item_txt <- ""
  if (!is.null(item_start)) {
    item_elapsed <- as.numeric(difftime(Sys.time(), item_start, units = "secs"))
    item_txt <- paste0(" | locus=", fmt_duration(item_elapsed))
  }
  progress_line(
    "%s [%d/%d] %s | elapsed=%s | ETA=%s%s",
    prefix, i, n, label, fmt_duration(elapsed), fmt_duration(eta), item_txt
  )
}

get_env_int <- function(name, default) {
  val <- suppressWarnings(as.integer(Sys.getenv(name, unset = NA_character_)))
  if (length(val) != 1L || is.na(val)) default else val
}

estimate_s_max_variants <- get_env_int("ESTIMATE_S_MAX_VARIANTS", 3000L)
ld_z_diag_max_variants <- get_env_int("LD_Z_DIAG_MAX_VARIANTS", estimate_s_max_variants)
estimate_s_subset_variants <- get_env_int("ESTIMATE_S_SUBSET_VARIANTS", 1500L)

subset_bundle_for_diagnostics <- function(bundle, locus_id, max_variants) {
  p <- length(bundle$z)
  required_ids <- character()
  if (grepl("^ydjc_", locus_id, ignore.case = TRUE)) {
    required_ids <- ydjc_pair_ids
  }
  required_idx <- which(bundle$variant_map$variant_id %in% required_ids)

  if (!is.finite(max_variants) || max_variants <= 0 || p <= max_variants) {
    keep <- seq_len(p)
    scope <- "full"
  } else {
    n_top <- max(0L, as.integer(max_variants) - length(required_idx))
    top_idx <- order(abs(bundle$z), decreasing = TRUE)[seq_len(min(n_top, p))]
    keep <- sort(unique(c(required_idx, top_idx)))
    scope <- sprintf("top_abs_z_subset_%d", length(keep))
  }

  list(
    z = bundle$z[keep],
    R = bundle$R[keep, keep, drop = FALSE],
    n_sample = bundle$n_sample,
    keep = keep,
    scope = scope,
    n_diagnostic_variants = length(keep)
  )
}
```

# 1. estimate_s_rss across all 20 loci

`s` in `[0, 1)` is the panel-LD inconsistency parameter: larger `s` => more
`z`-`R` disagreement. We report responders alongside non-responders so the
reader can see whether the loci carrying the annotation signal are *more*
inconsistent than the inert ones.

```{r estimate-s, include=FALSE}
estimate_s_start <- Sys.time()
n_loci <- length(all_loci)
progress_line(
  paste0(
    "estimate_s_rss: starting %d loci ",
    "(ESTIMATE_S_MAX_VARIANTS=%d, LD_Z_DIAG_MAX_VARIANTS=%d, ",
    "ESTIMATE_S_SUBSET_VARIANTS=%d)"
  ),
  n_loci, estimate_s_max_variants, ld_z_diag_max_variants,
  estimate_s_subset_variants
)

s_tbl <- map2_dfr(all_loci, seq_along(all_loci), function(lid, i) {
  locus_start <- Sys.time()
  progress_msg("estimate_s_rss start", i, n_loci, lid, estimate_s_start)

  progress_line("  [%d/%d] loading bundle: %s", i, n_loci, lid)
  bundle <- load_bundle(lid)
  p_locus <- length(bundle$z)
  progress_line("  [%d/%d] loaded bundle: %s p=%d", i, n_loci, lid, p_locus)

  diag_bundle <- subset_bundle_for_diagnostics(
    bundle = bundle,
    locus_id = lid,
    max_variants = if (p_locus <= estimate_s_max_variants) {
      p_locus
    } else {
      estimate_s_subset_variants
    }
  )
  progress_line(
    "  [%d/%d] diagnostic scope: %s n_diag=%d full_p=%d",
    i, n_loci, diag_bundle$scope, diag_bundle$n_diagnostic_variants, p_locus
  )

  if (diag_bundle$n_diagnostic_variants <= ld_z_diag_max_variants ||
      diag_bundle$scope != "full") {
    progress_line("  [%d/%d] computing MoM LD-z diagnostic: %s", i, n_loci, lid)
    ld_z_diag <- safe_ld_z_diag(diag_bundle)
  } else {
    progress_line(
      "  [%d/%d] skipping MoM LD-z diagnostic: %s p=%d > cap=%d",
      i, n_loci, lid, diag_bundle$n_diagnostic_variants, ld_z_diag_max_variants
    )
    ld_z_diag <- list(
      mom_sigma2_hat = NA_real_,
      ld_chol_ok = NA,
      ld_z_inconsistency_flag = NA
    )
  }

  if (diag_bundle$n_diagnostic_variants <= estimate_s_max_variants ||
      diag_bundle$scope != "full") {
    progress_line("  [%d/%d] computing estimate_s_rss: %s", i, n_loci, lid)
    s_val <- tryCatch(
      susieR::estimate_s_rss(
        z = diag_bundle$z, R = diag_bundle$R, n = diag_bundle$n_sample,
        method = "null-mle"
      ),
      error = function(e) {
        progress_line("  [%d/%d] estimate_s_rss ERROR: %s | %s", i, n_loci, lid,
                      conditionMessage(e))
        NA_real_
      }
    )
    estimate_s_status <- ifelse(is.na(s_val), "error_or_na", diag_bundle$scope)
  } else {
    progress_line(
      "  [%d/%d] skipping estimate_s_rss: %s p=%d > cap=%d",
      i, n_loci, lid, diag_bundle$n_diagnostic_variants, estimate_s_max_variants
    )
    s_val <- NA_real_
    estimate_s_status <- "skipped_p_gt_cap"
  }

  row <- tibble(
    locus_id = lid,
    n_variants = p_locus,
    n_diagnostic_variants = diag_bundle$n_diagnostic_variants,
    diagnostic_scope = diag_bundle$scope,
    n_sample = bundle$n_sample,
    n_sample_min = bundle$n_sample_min,
    n_sample_max = bundle$n_sample_max,
    n_sample_ratio = bundle$n_sample_max / bundle$n_sample_min,
    mom_sigma2_hat = ld_z_diag$mom_sigma2_hat,
    ld_chol_ok = ld_z_diag$ld_chol_ok,
    ld_z_inconsistency_flag = ld_z_diag$ld_z_inconsistency_flag,
    s = s_val,
    estimate_s_status = estimate_s_status,
    is_responder = lid %in% responder_loci
  )
  progress_msg(
    sprintf("estimate_s_rss done  p=%d s=%s status=%s", p_locus,
            ifelse(is.na(s_val), "NA", sprintf("%.4f", s_val)),
            estimate_s_status),
    i, n_loci, lid, estimate_s_start, locus_start
  )
  row
}) %>% arrange(desc(s))

progress_line(
  "estimate_s_rss: finished %d loci in %s",
  n_loci, fmt_duration(as.numeric(difftime(Sys.time(), estimate_s_start, units = "secs")))
)

readr::write_csv(s_tbl, file.path(out_dir, "estimate_s_rss_by_locus.csv"))

s_summary_tbl <- s_tbl %>%
  group_by(is_responder) %>%
  summarise(
    n = n(),
    median_s = safe_summary(s, median),
    max_s = safe_summary(s, max),
    median_n_ratio = safe_summary(n_sample_ratio, median),
    min_mom_sigma2_hat = safe_summary(mom_sigma2_hat, min),
    n_ld_z_flags = sum(ld_z_inconsistency_flag %in% TRUE, na.rm = TRUE),
    .groups = "drop"
  )
```

# 2. kriging_rss on the responder loci

For each responder we extract the per-variant conditional distribution. The
`logLR` column is the log-LR for an allele flip; `z_std_diff` is the
standardized observed-minus-expected `z`. The susieR convention flags a variant
when `logLR > 2 & abs(z) > 2` (it fits *better* sign-flipped and is a real
marginal signal) -- i.e. a candidate LD-inconsistency / flip.

```{r kriging, include=FALSE}
kriging_res <- map(responder_loci, function(lid) {
  bundle <- load_bundle(lid)
  kr <- susieR::kriging_rss(
    z = bundle$z, R = bundle$R, n = bundle$n_sample
  )
  cond <- kr$conditional_dist %>%
    mutate(
      locus_id = lid,
      ld_matrix_index = bundle$variant_map$ld_matrix_index,
      variant_id = bundle$variant_map$variant_id,
      chrom = bundle$variant_map$chrom,
      pos = bundle$variant_map$pos,
      ref = bundle$variant_map$ref,
      alt = bundle$variant_map$alt,
      annotation_a = bundle$a,
      strand_ambiguous = is_strand_ambiguous(ref, alt),
      flagged = logLR > 2 & abs(z) > 2
    )
  list(locus_id = lid, plot = kr$plot, cond = cond)
})
names(kriging_res) <- responder_loci

kriging_cond <- map_dfr(kriging_res, "cond")
readr::write_csv(
  kriging_cond, file.path(out_dir, "kriging_rss_conditional_by_variant.csv")
)

kriging_flagged_tbl <- kriging_cond %>%
  filter(flagged) %>%
  select(locus_id, variant_id, chrom, pos, ref, alt, z, condmean, z_std_diff,
         logLR, annotation_a, strand_ambiguous) %>%
  arrange(desc(logLR))

ydjc_lid <- responder_loci[grepl("^ydjc_", responder_loci, ignore.case = TRUE)]
if (length(ydjc_lid) == 1) {
  ydjc_bundle <- load_bundle(ydjc_lid)
  ydjc_pair_tbl <- kriging_res[[ydjc_lid]]$cond %>%
    filter(variant_id %in% ydjc_pair_ids) %>%
    arrange(match(variant_id, ydjc_pair_ids))
  if (nrow(ydjc_pair_tbl) != length(ydjc_pair_ids)) {
    stop(
      "Expected both YDJC headline variants in the loaded bundle; found ",
      nrow(ydjc_pair_tbl), "."
    )
  }
  ydjc_pair_r <- ydjc_bundle$R[
    ydjc_pair_tbl$ld_matrix_index[1],
    ydjc_pair_tbl$ld_matrix_index[2]
  ]
  ydjc_pair_tbl <- ydjc_pair_tbl %>%
    mutate(pairwise_r = ydjc_pair_r) %>%
    select(locus_id, variant_id, chrom, pos, ref, alt, pairwise_r, z,
           condmean, condvar, z_std_diff, logLR, annotation_a,
           strand_ambiguous, flagged)
} else {
  ydjc_pair_tbl <- tibble(
    locus_id = character(), variant_id = character(), chrom = character(),
    pos = integer(), ref = character(), alt = character(), pairwise_r = numeric(),
    z = numeric(), condmean = numeric(), condvar = numeric(),
    z_std_diff = numeric(), logLR = numeric(), annotation_a = numeric(),
    strand_ambiguous = logical(), flagged = logical()
  )
}

walk(responder_loci, function(lid) {
  ggsave(
    file.path(out_dir, paste0("kriging_obs_vs_exp_", lid, ".png")),
    kriging_res[[lid]]$plot, width = 6, height = 5, dpi = 200, bg = "white"
  )
})
```

# 3. Allele orientation / strand-ambiguity audit

Upstream `pipeline_ld.py` matches GTEx -> 1000G on the exact `(chrom, pos, ref,
alt)` tuple, so a ref/alt orientation swap cannot survive matching: orientation
is concordant by construction. The only residual silent-flip pathway is a
strand-ambiguous SNP (ref/alt in {A/T, C/G}), where the complementary-strand
allele set is identical and an opposite-strand record can match the tuple while
being the complement.

```{r strand-audit, include=FALSE}
strand_tbl <- map_dfr(responder_loci, function(lid) {
  bundle <- load_bundle(lid)
  bundle$variant_map %>%
    transmute(
      locus_id = lid, variant_id, ref, alt,
      strand_ambiguous = is_strand_ambiguous(ref, alt)
    )
})

readr::write_csv(
  strand_tbl, file.path(out_dir, "strand_ambiguity_responders.csv")
)

strand_summary_tbl <- strand_tbl %>%
  group_by(locus_id) %>%
  summarise(n_variants = n(),
            n_strand_ambiguous = sum(strand_ambiguous), .groups = "drop")

ydjc_strand_tbl <- strand_tbl %>%
  filter(variant_id %in% ydjc_pair_ids)

strand_flagged_tbl <- kriging_cond %>%
  filter(flagged) %>%
  select(locus_id, variant_id, ref, alt, logLR, strand_ambiguous) %>%
  arrange(desc(logLR))
```

# 4. Optional cross-check against high-PIP real-data result drivers

If the collected real-data ensemble outputs are present, this joins kriging
diagnostics to high-PIP variants. This is optional and downstream-only; if the
collection artifacts are not mounted, the table is empty and the core audit
above still runs.

```{r high-pip-crosscheck, include=FALSE}
candidate_aggregated_dirs <- file.path(
  here::here("output", "slurm_output", "real_data_ensemble_geometric_n20"),
  c("52906940", Sys.getenv("REAL_DATA_PARENT_JOB_ID", unset = NA_character_)),
  "aggregated"
)
candidate_aggregated_dirs <- unique(candidate_aggregated_dirs[!is.na(candidate_aggregated_dirs)])
aggregated_dir <- candidate_aggregated_dirs[
  file.exists(file.path(candidate_aggregated_dirs, "paper_real_data_ensemble_summary.csv"))
][1]

read_parquet_if_exists <- function(path) {
  if (length(path) != 1L || is.na(path) || !file.exists(path)) {
    return(tibble())
  }
  arrow::read_parquet(path) %>% tibble::as_tibble()
}

ensure_cols <- function(tbl, cols) {
  for (nm in setdiff(cols, names(tbl))) {
    tbl[[nm]] <- NA
  }
  tbl
}

if (!is.na(aggregated_dir)) {
  paper_summary <- readr::read_csv(
    file.path(aggregated_dir, "paper_real_data_ensemble_summary.csv"),
    show_col_types = FALSE
  )
  agg_pip_tbl <- read_parquet_if_exists(
    file.path(aggregated_dir, "aggregated_variant_pips_cluster_weight.parquet")
  ) %>%
    ensure_cols(c("locus_id", "variant_id", "aggregated_pip",
                  "aggregated_posterior_mean"))
  refit_pip_tbl <- read_parquet_if_exists(
    file.path(aggregated_dir, "highest_weight_refit_variant_posteriors.parquet")
  ) %>%
    ensure_cols(c("locus_id", "variant_id", "pip"))
  variant_posteriors <- tryCatch(
    arrow::open_dataset(file.path(aggregated_dir, "variant_posteriors_dataset")),
    error = function(e) NULL
  )

  anchor_tbl <- tibble()
  if (!is.null(variant_posteriors) &&
      all(c("locus_id", "run_id", "variant_id", "pip") %in%
          names(variant_posteriors)) &&
      "susie_anchor_run_id" %in% names(paper_summary)) {
    anchor_tbl <- variant_posteriors %>%
      filter(locus_id %in% responder_loci) %>%
      select(locus_id, run_id, variant_id, anchor_pip = pip) %>%
      collect() %>%
      inner_join(
        paper_summary %>% select(locus_id, susie_anchor_run_id),
        by = "locus_id"
      ) %>%
      filter(run_id == susie_anchor_run_id) %>%
      select(-run_id, -susie_anchor_run_id)
  }

  anchor_tbl <- ensure_cols(anchor_tbl, c("locus_id", "variant_id", "anchor_pip"))

  high_pip_driver_tbl <- kriging_cond %>%
    select(locus_id, variant_id, z, condmean, z_std_diff, logLR, flagged,
           strand_ambiguous, annotation_a) %>%
    left_join(
      agg_pip_tbl %>%
        filter(locus_id %in% responder_loci) %>%
        select(locus_id, variant_id, aggregated_pip,
               aggregated_posterior_mean),
      by = c("locus_id", "variant_id")
    ) %>%
    left_join(anchor_tbl, by = c("locus_id", "variant_id")) %>%
    left_join(
      refit_pip_tbl %>%
        filter(locus_id %in% responder_loci) %>%
        select(locus_id, variant_id, warm_refit_pip = pip),
      by = c("locus_id", "variant_id")
    ) %>%
    mutate(
      max_reported_pip = pmax(aggregated_pip, anchor_pip, warm_refit_pip,
                              na.rm = TRUE)
    ) %>%
    filter(is.finite(max_reported_pip), max_reported_pip >= 0.5) %>%
    arrange(desc(flagged), desc(max_reported_pip), desc(logLR))
} else {
  high_pip_driver_tbl <- tibble()
}

readr::write_csv(
  high_pip_driver_tbl,
  file.path(out_dir, "kriging_high_pip_result_drivers.csv")
)
```

# 5. Durable report exports

All final tables are written here before rendering. This makes the workbook
safe to run from a transient interactive session: the report artifacts persist
under `out_dir` even if the console disconnects.

```{r durable-report-exports, include=FALSE}
final_table_specs <- list(
  estimate_s_rss_by_locus = list(
    data = s_tbl,
    caption = paste(
      "All loci: estimate_s_rss, sample-size spread, and existing MoM",
      "LD-z consistency diagnostics"
    ),
    digits = 4
  ),
  estimate_s_rss_responder_summary = list(
    data = s_summary_tbl,
    caption = "Responder versus non-responder consistency summary",
    digits = 4
  ),
  kriging_flagged_variants = list(
    data = kriging_flagged_tbl,
    caption = "Responder variants flagged by kriging_rss (logLR > 2, |z| > 2)",
    digits = 3
  ),
  ydjc_headline_pair_kriging = list(
    data = ydjc_pair_tbl,
    caption = "YDJC headline pair: pairwise LD, kriging, annotation, and strand audit",
    digits = 3
  ),
  strand_ambiguity_summary = list(
    data = strand_summary_tbl,
    caption = "Strand-ambiguous SNP counts per responder locus",
    digits = NULL
  ),
  ydjc_headline_pair_strand = list(
    data = ydjc_strand_tbl,
    caption = "YDJC headline pair: strand-ambiguity check",
    digits = NULL
  ),
  kriging_flagged_x_strand = list(
    data = strand_flagged_tbl,
    caption = "kriging-flagged variants x strand ambiguity",
    digits = 3
  ),
  high_pip_result_drivers = list(
    data = high_pip_driver_tbl,
    caption = "High-PIP responder variants joined to kriging diagnostics",
    digits = 3
  )
)

write_report_csv <- function(name, spec) {
  readr::write_csv(spec$data, file.path(out_dir, paste0(name, ".csv")))
}

render_report_table <- function(spec) {
  tbl <- spec$data
  digits <- spec$digits
  table_txt <- if (is.null(digits)) {
    knitr::kable(tbl, format = "pipe")
  } else {
    knitr::kable(tbl, format = "pipe", digits = digits)
  }
  c("", paste0("## ", spec$caption), "", table_txt, "")
}

walk2(names(final_table_specs), final_table_specs, write_report_csv)

plot_files <- file.path(
  out_dir, paste0("kriging_obs_vs_exp_", responder_loci, ".png")
)

report_path <- file.path(out_dir, "ld_mismatch_copy_paste_report.md")
report_lines <- c(
  "# Real-data LD-mismatch diligence report",
  "",
  paste0("Generated: ", format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z")),
  paste0("Output directory: `", out_dir, "`"),
  paste0("Progress log: `", progress_log_path, "`"),
  "",
  "## Runtime settings",
  "",
  paste0("- `ESTIMATE_S_MAX_VARIANTS`: ", estimate_s_max_variants),
  paste0("- `LD_Z_DIAG_MAX_VARIANTS`: ", ld_z_diag_max_variants),
  paste0("- `ESTIMATE_S_SUBSET_VARIANTS`: ", estimate_s_subset_variants),
  "",
  "## Written files",
  "",
  paste0("- `", names(final_table_specs), ".csv`"),
  paste0("- `", basename(plot_files), "`"),
  "- `run_progress.log`",
  "- `ld_mismatch_copy_paste_report.md`",
  "",
  unlist(map(final_table_specs, render_report_table), use.names = FALSE),
  "",
  "## Decision tree",
  "",
  "- 1-3 pass: low/ordinary `s`, no kriging flag at the YDJC pair, and the YDJC pair is not strand-ambiguous. Keep YDJC as headline with a short supplement; caveat that in-sample LD remains the gold standard.",
  "- YDJC pair kriging-flagged: demote YDJC from the headline and lead with a responder that passes.",
  "- YDJC pair strand-ambiguous: fix harmonization and rerun before finalizing.",
  "- In all cases: warm-refit durability controls for prior dependence, not LD mismatch."
)

writeLines(report_lines, report_path)
progress_line("durable report written: %s", report_path)
```

# Decision tree (from ld_mismatch_diligence.md)

- **1-3 pass** (low `s`, no kriging flag at the YDJC pair, YDJC pair not
  strand-ambiguous): add a short supplementary subsection, keep YDJC as the
  headline, soften the Discussion caveat to "standard `z`-`R` consistency
  diagnostics flagged no problem at the responder loci; in-sample LD remains the
  gold standard and is the natural next step."
- **kriging flags the YDJC pair**: demote YDJC from the headline, lead with a
  responder that passes, state the mismatch honestly.
- **YDJC pair strand-ambiguous**: a real allele-harmonization bug -- fix and
  re-run upstream before finalizing anything (this may change the result).
- **In all cases**: state explicitly that the warm-refit durability test
  controls for prior dependence, not LD mismatch.

# Copy-paste report

All printed output is collected here.

```{r final-report-tables, echo=FALSE, results='asis', fig.width=6, fig.height=5}
emit_table <- function(x, ..., digits = NULL, caption = NULL) {
  cat("\n\n")
  if (is.null(digits)) {
    tbl <- knitr::kable(x, ..., caption = caption)
  } else {
    tbl <- knitr::kable(x, ..., digits = digits, caption = caption)
  }
  cat(tbl, sep = "\n")
  cat("\n\n")
}

cat("**Output directory:** `", out_dir, "`\n\n", sep = "")
cat("**Copy-paste Markdown report:** `", report_path, "`\n\n", sep = "")
cat("**Progress log:** `", progress_log_path, "`\n\n", sep = "")
cat("**Per-table CSV files:** `", out_dir, "/*.csv`\n\n", sep = "")

emit_table(
  s_tbl,
  digits = 4,
  caption = paste(
    "All loci: estimate_s_rss, sample-size spread, and existing MoM",
    "LD-z consistency diagnostics"
  )
)

emit_table(
  s_summary_tbl,
  digits = 4,
  caption = "Responder versus non-responder consistency summary"
)

emit_table(
  kriging_flagged_tbl,
  digits = 3,
  caption = "Responder variants flagged by kriging_rss (logLR > 2, |z| > 2)"
)

emit_table(
  ydjc_pair_tbl,
  digits = 3,
  caption = "YDJC headline pair: pairwise LD, kriging, annotation, and strand audit"
)

emit_table(
  strand_summary_tbl,
  caption = "Strand-ambiguous SNP counts per responder locus"
)

emit_table(
  ydjc_strand_tbl,
  caption = "YDJC headline pair: strand-ambiguity check"
)

emit_table(
  strand_flagged_tbl,
  digits = 3,
  caption = "kriging-flagged variants x strand ambiguity"
)

if (nrow(high_pip_driver_tbl)) {
  emit_table(
    high_pip_driver_tbl,
    digits = 3,
    caption = "High-PIP responder variants joined to kriging diagnostics"
  )
} else {
  cat(
    "\nHigh-PIP responder cross-check was empty or skipped because collected ",
    "real-data ensemble outputs were not found at parent job 52906940 or ",
    "REAL_DATA_PARENT_JOB_ID.\n\n",
    sep = ""
  )
}

walk(responder_loci, function(lid) {
  print(kriging_res[[lid]]$plot + ggtitle(lid))
})
```
