---
title: "mu_0 Scale Estimation: Experimental Validation"
date: "2026-03-05"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: show
editor_options:
  markdown:
    wrap: 72
---

# Overview

This workbook tests the theoretical predictions from
`refs/mu_0_estimation_theory.md` about why EB estimation works for
$\sigma_0^2$ but not for $c$ (the prior mean scale factor where
$\mu_{0j} = c \cdot a_j$).

**Key predictions to test:**

1.  `auto_scale_mu_0` works well at high annotation R^2, degrades at
    low R^2 and high L
2.  The $(c, \sigma_0^2)$ likelihood surface is unimodal in
    $\sigma_0^2$ but flat/multimodal in $c$
3.  Iterative "mean" mode is ineffective (wrong parameterization:
    uniform shift, not annotation-scaled)
4.  Iterative "both" mode collapses to
    $(\mu_0 = 0,\, \sigma_0^2 = \text{large})$
5.  Oracle $c$ gives best AUPRC; c-grid comes close

```{r setup, message=FALSE, warning=FALSE}
here::i_am(
  "vignettes/one_off_validations/mu_0_estimation_experiments.Rmd"
)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(tibble)
library(knitr)

devtools::load_all(here("../susine"), quiet = TRUE)
devtools::load_all(here("."), quiet = TRUE)

theme_set(theme_bw(base_size = 12))
set.seed(20260305)
```

# Synthetic Data Generation

Small-scale problems runnable locally. We generate multiple scenarios
varying annotation quality and signal strength.

```{r generate-data}
# --- Dimensions ---
n <- 400L
p <- 200L
L <- 5L

# --- Design matrix with AR1 LD structure ---
make_X <- function(n, p, rho = 0.5) {
  X <- matrix(rnorm(n * p), nrow = n)
  for (j in 2:p) {
    X[, j] <- rho * X[, j - 1] + sqrt(1 - rho^2) * X[, j]
  }
  # Standardize columns
  X <- scale(X, center = TRUE, scale = TRUE)
  X
}

# --- Generate one scenario ---
make_scenario <- function(n, p, p_star, c_true, annotation_r2,
                          noise_fraction = 0.8, rho = 0.5,
                          seed = NULL) {
  if (!is.null(seed)) set.seed(seed)

  X <- make_X(n, p, rho)
  causal_idx <- sort(sample.int(p, p_star))

  beta_true <- rep(0, p)
  beta_true[causal_idx] <- rnorm(p_star, sd = 1)

  # --- Annotation: a = beta / c_true + noise ---
  annotation <- rep(0, p)
  if (c_true != 0) {
    # Causal SNPs: annotation tracks true effect
    causal_signal <- beta_true[causal_idx] / c_true
    if (annotation_r2 > 0 && annotation_r2 < 1) {
      noise_sd <- sqrt(var(causal_signal) *
                          (1 - annotation_r2) / annotation_r2)
    } else if (annotation_r2 >= 1) {
      noise_sd <- 0
    } else {
      noise_sd <- sd(causal_signal) * 10
    }
    annotation[causal_idx] <- causal_signal + rnorm(p_star, sd = noise_sd)
  }
  # Noncausal: small background noise
  noncausal_idx <- setdiff(seq_len(p), causal_idx)
  annotation[noncausal_idx] <- rnorm(
    length(noncausal_idx),
    sd = sd(annotation[causal_idx]) * 0.3
  )

  # --- Phenotype ---
  signal <- drop(X %*% beta_true)
  sigma2 <- var(signal) * noise_fraction / (1 - noise_fraction)
  y <- signal + rnorm(n, sd = sqrt(sigma2))

  list(
    X = X, y = y, beta_true = beta_true,
    annotation = annotation,
    causal_idx = causal_idx,
    c_true = c_true,
    annotation_r2 = annotation_r2,
    noise_fraction = noise_fraction,
    sigma2 = sigma2,
    p_star = p_star
  )
}
```

```{r scenario-grid}
# Build a grid of scenarios
scenario_grid <- expand_grid(
  c_true = c(0.3, 0.7, 1.5),
  annotation_r2 = c(0.1, 0.4, 0.8),
  p_star = c(3, 5, 10),
  rep = 1:3
) %>%
  mutate(
    scenario_id = row_number(),
    seed = 20260305 + scenario_id
  )

cat("Total scenarios:", nrow(scenario_grid), "\n")
```

```{r generate-all}
# Generate all scenarios (this takes ~10-30 seconds)
scenarios <- scenario_grid %>%
  pmap(function(c_true, annotation_r2, p_star, rep, scenario_id, seed) {
    make_scenario(
      n = n, p = p, p_star = p_star,
      c_true = c_true, annotation_r2 = annotation_r2,
      noise_fraction = 0.8, seed = seed
    )
  })

cat("Generated", length(scenarios), "scenarios.\n")
```

# Experiment 1: auto_scale_mu_0 Validation

Test the one-shot `estimate_mu_0_scale_factor()` on known-$c$ data.

**Theory predicts:** works well when annotation R^2 is high; degrades
when annotation is noisy or L effects share LD.

```{r exp1-autoscale}
exp1_results <- scenario_grid %>%
  mutate(
    c_hat = map_dbl(scenarios, function(sc) {
      X_init <- susine:::initialize_X(sc$X)
      d <- attr(X_init, "d")
      xty <- susine:::compute_Xty(X_init, sc$y)
      beta_hat <- as.vector(xty / d)
      shat2 <- as.vector(sc$sigma2 / d)

      est <- susine::estimate_mu_0_scale_factor(
        beta_hat = beta_hat,
        shat2 = shat2,
        mu_0_unscaled = sc$annotation
      )
      est$c
    }),
    c_true = map_dbl(scenarios, ~.x$c_true)
  ) %>%
  mutate(
    c_error = c_hat - c_true,
    c_ratio = c_hat / c_true
  )
```

```{r exp1-scatter, fig.width=8, fig.height=6}
exp1_results %>%
  ggplot(aes(x = c_true, y = c_hat, color = factor(annotation_r2))) +
  geom_point(size = 2, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  facet_wrap(~p_star, labeller = label_both) +
  labs(
    title = "Experiment 1: auto_scale_mu_0 recovery of c",
    x = "True c", y = "Estimated c",
    color = "Annotation R²"
  )
```

```{r exp1-summary}
exp1_results %>%
  group_by(annotation_r2, p_star) %>%
  summarize(
    mean_bias = mean(c_error),
    rmse = sqrt(mean(c_error^2)),
    mean_ratio = mean(c_ratio),
    .groups = "drop"
  ) %>%
  kable(digits = 3,
        caption = "auto_scale_mu_0: bias and RMSE by annotation R² and p*")
```

# Experiment 2: Profile the $(c, \sigma_0^2)$ Likelihood Surface

Evaluate the SER negative log-likelihood on a fine 2D grid.

**Theory predicts:** unimodal in $\sigma_0^2$ at fixed $c$;
potentially flat or multimodal in $c$ at fixed $\sigma_0^2$.

```{r exp2-setup}
# Pick a representative scenario: medium annotation R2, moderate signal
exp2_idx <- which(
  scenario_grid$annotation_r2 == 0.4 &
    scenario_grid$c_true == 0.7 &
    scenario_grid$p_star == 5 &
    scenario_grid$rep == 1
)[1]

sc2 <- scenarios[[exp2_idx]]
cat("Experiment 2 scenario:", exp2_idx, "\n")
cat("  c_true =", sc2$c_true, ", annotation_r2 =", sc2$annotation_r2,
    ", p_star =", sc2$p_star, "\n")
```

```{r exp2-profile}
# Prepare X for susine internals
X2 <- susine:::initialize_X(sc2$X)
y2 <- sc2$y - mean(sc2$y)
sigma2_est <- sc2$sigma2

# Define the grid
c_grid <- seq(-1, 3, length.out = 50)
log_s0_grid <- seq(-4, 2, length.out = 50)

profile_2d <- expand_grid(
  c_val = c_grid,
  log_s0 = log_s0_grid
) %>%
  mutate(
    nll = pmap_dbl(list(c_val, log_s0), function(cv, ls) {
      mu_0_vec <- cv * sc2$annotation
      s0_val <- exp(ls)
      pi_vec <- rep(1 / p, p)

      priors <- data.frame(
        mu_0 = I(list(mu_0_vec)),
        sigma_0_2 = I(list(rep(s0_val, p))),
        prior_inclusion_weights = I(list(pi_vec))
      )

      tryCatch(
        susine:::neg_log_lik(priors, X2, y2, sigma2_est),
        error = function(e) NA_real_
      )
    })
  )
```

```{r exp2-contour, fig.width=9, fig.height=7}
profile_2d %>%
  filter(!is.na(nll)) %>%
  ggplot(aes(x = c_val, y = log_s0, z = nll, fill = nll)) +
  geom_tile() +
  geom_contour(color = "white", linewidth = 0.3, bins = 20) +
  geom_vline(xintercept = sc2$c_true, linetype = "dashed",
             color = "red", linewidth = 0.8) +
  scale_fill_viridis_c(option = "inferno", direction = -1,
                       name = "Neg log-lik") +
  labs(
    title = "SER Negative Log-Likelihood Surface",
    subtitle = sprintf("c_true = %.1f, ann_r2 = %.1f, p* = %d",
                       sc2$c_true, sc2$annotation_r2, sc2$p_star),
    x = "c (prior mean scale)", y = "log(sigma_0^2)"
  )
```

```{r exp2-slices, fig.width=10, fig.height=4}
# Slice at the optimal log_s0 to show c-direction profile
best_log_s0 <- profile_2d %>%
  filter(!is.na(nll)) %>%
  slice_min(nll, n = 1) %>%
  pull(log_s0)

p_cslice <- profile_2d %>%
  filter(abs(log_s0 - best_log_s0) < 0.1) %>%
  ggplot(aes(x = c_val, y = nll)) +
  geom_line() +
  geom_vline(xintercept = sc2$c_true, linetype = "dashed", color = "red") +
  labs(
    title = sprintf("NLL slice at log(s0^2) = %.1f", best_log_s0),
    x = "c", y = "Neg log-lik"
  )

# Slice at c_true to show sigma_0^2-direction profile
p_sslice <- profile_2d %>%
  filter(abs(c_val - sc2$c_true) < diff(c_grid)[1]) %>%
  ggplot(aes(x = log_s0, y = nll)) +
  geom_line() +
  labs(
    title = sprintf("NLL slice at c = %.1f (true)", sc2$c_true),
    x = "log(sigma_0^2)", y = "Neg log-lik"
  )

gridExtra::grid.arrange(p_cslice, p_sslice, ncol = 2)
```

```{r exp2-curvature}
# Compare curvature (second derivative) at the optimum
# in c-direction vs sigma_0_2-direction
opt_point <- profile_2d %>%
  filter(!is.na(nll)) %>%
  slice_min(nll, n = 1)

cat("Optimal point: c =", opt_point$c_val,
    ", log(s0^2) =", opt_point$log_s0,
    ", NLL =", opt_point$nll, "\n")
cat("True c =", sc2$c_true, "\n")
```

# Experiment 3: Iterative EB "mean" Mode vs. c-Grid

Test whether iterative EB with `prior_update_method = "mean"` converges
to a useful value.

**Theory predicts:** "mean" mode estimates a UNIFORM shift
(not annotation-scaled $c$), so it should converge near 0.

```{r exp3-setup}
# Pick a few representative scenarios
exp3_indices <- which(
  scenario_grid$annotation_r2 %in% c(0.4, 0.8) &
    scenario_grid$c_true == 0.7 &
    scenario_grid$p_star == 5 &
    scenario_grid$rep == 1
)
```

```{r exp3-iterative-eb}
exp3_results <- map_dfr(exp3_indices, function(idx) {
  sc <- scenarios[[idx]]
  mu_0_scaled <- sc$c_true * sc$annotation

  # Run susine with prior_update_method = "mean"
  fit_eb <- tryCatch(
    susine::susine(
      L = L, X = sc$X, y = sc$y,
      mu_0 = mu_0_scaled,
      sigma_0_2 = 0.2,
      prior_update_method = "mean",
      max_iter = 50L,
      tol = 1e-4
    ),
    error = function(e) {
      message("susine() failed for scenario ", idx, ": ", e$message)
      NULL
    }
  )

  # Run susine with fixed c (oracle)
  fit_oracle <- tryCatch(
    susine::susine(
      L = L, X = sc$X, y = sc$y,
      mu_0 = mu_0_scaled,
      sigma_0_2 = 0.2,
      prior_update_method = "none",
      max_iter = 50L,
      tol = 1e-4
    ),
    error = function(e) {
      message("susine() oracle failed for scenario ", idx, ": ", e$message)
      NULL
    }
  )

  # Run susine at c=0 (vanilla)
  fit_vanilla <- tryCatch(
    susine::susine(
      L = L, X = sc$X, y = sc$y,
      mu_0 = 0,
      sigma_0_2 = 0.2,
      prior_update_method = "none",
      max_iter = 50L,
      tol = 1e-4
    ),
    error = function(e) {
      message("susine() vanilla failed for scenario ", idx, ": ", e$message)
      NULL
    }
  )

  # Extract PIPs and compare
  extract_auprc <- function(fit, sc) {
    if (is.null(fit)) return(NA_real_)
    pips <- fit$model_fit$PIPs
    truth <- rep(0L, p)
    truth[sc$causal_idx] <- 1L
    # Simple AUPRC via trapezoidal rule
    ord <- order(pips, decreasing = TRUE)
    tp <- cumsum(truth[ord])
    fp <- cumsum(1L - truth[ord])
    precision <- tp / (tp + fp)
    recall <- tp / sum(truth)
    # Trapezoidal AUPRC
    sum(diff(recall) * (precision[-1] + precision[-length(precision)]) / 2)
  }

  tibble(
    scenario_id = idx,
    annotation_r2 = sc$annotation_r2,
    c_true = sc$c_true,
    auprc_eb_mean = extract_auprc(fit_eb, sc),
    auprc_oracle = extract_auprc(fit_oracle, sc),
    auprc_vanilla = extract_auprc(fit_vanilla, sc),
    # Check what mu_0 converged to under "mean" mode
    final_mu_0_eb = if (!is.null(fit_eb))
      mean(fit_eb$priors$mu_0[[1]]) else NA_real_,
    elbo_eb = if (!is.null(fit_eb))
      fit_eb$model_fit$elbo else NA_real_,
    elbo_oracle = if (!is.null(fit_oracle))
      fit_oracle$model_fit$elbo else NA_real_,
    elbo_vanilla = if (!is.null(fit_vanilla))
      fit_vanilla$model_fit$elbo else NA_real_
  )
})

kable(exp3_results, digits = 4,
      caption = "Experiment 3: Iterative EB 'mean' mode vs oracle vs vanilla")
```

# Experiment 4: Joint "both" Mode Pathology

Test `prior_update_method = "both"` from multiple initializations.

**Theory predicts:** convergence to $(mu_0 \approx 0, \sigma_0^2 = \text{large})$
or pathological values, because the 2D landscape has a flat c-direction.

```{r exp4-both-mode}
exp4_idx <- exp3_indices[1]  # Single scenario for detailed analysis
sc4 <- scenarios[[exp4_idx]]

# Multiple initializations via different mu_0 starting points
init_c_values <- c(0, 0.3, 0.7, 1.0, 1.5, 2.0)

exp4_results <- map_dfr(init_c_values, function(c_init) {
  mu_0_init <- c_init * sc4$annotation

  fit <- tryCatch(
    susine::susine(
      L = L, X = sc4$X, y = sc4$y,
      mu_0 = mu_0_init,
      sigma_0_2 = 0.2,
      prior_update_method = "both",
      max_iter = 50L,
      tol = 1e-4
    ),
    error = function(e) {
      message("both mode failed at c_init = ", c_init, ": ", e$message)
      NULL
    }
  )

  # Extract final priors
  if (!is.null(fit)) {
    # Average mu_0 and sigma_0_2 across effects to summarize
    final_mu0_vals <- map_dbl(seq_len(L), function(l) {
      mean(fit$priors$mu_0[[l]])
    })
    final_s02_vals <- map_dbl(seq_len(L), function(l) {
      mean(fit$priors$sigma_0_2[[l]])
    })

    tibble(
      c_init = c_init,
      final_mu0_mean = mean(final_mu0_vals),
      final_mu0_sd = sd(final_mu0_vals),
      final_s02_mean = mean(final_s02_vals),
      final_s02_sd = sd(final_s02_vals),
      elbo = fit$model_fit$elbo,
      converged = fit$model_fit$converged
    )
  } else {
    tibble(
      c_init = c_init,
      final_mu0_mean = NA_real_,
      final_mu0_sd = NA_real_,
      final_s02_mean = NA_real_,
      final_s02_sd = NA_real_,
      elbo = NA_real_,
      converged = NA
    )
  }
})

kable(exp4_results, digits = 4,
      caption = sprintf(
        "Experiment 4: 'both' mode from multiple inits (c_true = %.1f)",
        sc4$c_true
      ))
```

```{r exp4-analysis}
cat("True c =", sc4$c_true, "\n\n")
cat("Did 'both' mode find the right c from any initialization?\n")
cat("Did it collapse to mu_0 ~ 0?\n")
cat("Is ELBO sensitive to initialization?\n")
```

# Experiment 5: Oracle Comparison and c-Grid

Run a full c-grid at fine resolution and compare to oracle, auto_scale,
and vanilla.

```{r exp5-cgrid}
exp5_idx <- exp3_indices[1]
sc5 <- scenarios[[exp5_idx]]

c_grid_fine <- seq(0, 2, by = 0.1)

exp5_grid <- map_dfr(c_grid_fine, function(c_val) {
  mu_0_val <- c_val * sc5$annotation

  fit <- tryCatch(
    susine::susine(
      L = L, X = sc5$X, y = sc5$y,
      mu_0 = mu_0_val,
      sigma_0_2 = 0.2,
      prior_update_method = "none",
      max_iter = 50L,
      tol = 1e-4
    ),
    error = function(e) NULL
  )

  if (!is.null(fit)) {
    pips <- fit$model_fit$PIPs
    truth <- rep(0L, p)
    truth[sc5$causal_idx] <- 1L
    ord <- order(pips, decreasing = TRUE)
    tp <- cumsum(truth[ord])
    fp <- cumsum(1L - truth[ord])
    precision <- tp / (tp + fp)
    recall <- tp / sum(truth)
    auprc <- sum(diff(recall) *
                   (precision[-1] + precision[-length(precision)]) / 2)

    tibble(c_val = c_val, elbo = fit$model_fit$elbo, auprc = auprc)
  } else {
    tibble(c_val = c_val, elbo = NA_real_, auprc = NA_real_)
  }
})
```

```{r exp5-elbo-profile, fig.width=8, fig.height=4}
p5a <- exp5_grid %>%
  filter(!is.na(elbo)) %>%
  ggplot(aes(x = c_val, y = elbo)) +
  geom_line() +
  geom_point(size = 1.5) +
  geom_vline(xintercept = sc5$c_true, linetype = "dashed",
             color = "red") +
  annotate("text", x = sc5$c_true + 0.05, y = min(exp5_grid$elbo, na.rm = TRUE),
           label = "c_true", hjust = 0, color = "red") +
  labs(title = "ELBO across c-grid", x = "c", y = "ELBO")

p5b <- exp5_grid %>%
  filter(!is.na(auprc)) %>%
  ggplot(aes(x = c_val, y = auprc)) +
  geom_line() +
  geom_point(size = 1.5) +
  geom_vline(xintercept = sc5$c_true, linetype = "dashed",
             color = "red") +
  labs(title = "AUPRC across c-grid", x = "c", y = "AUPRC")

gridExtra::grid.arrange(p5a, p5b, ncol = 2)
```

```{r exp5-comparison}
# auto_scale estimate
X5 <- susine:::initialize_X(sc5$X)
d5 <- attr(X5, "d")
xty5 <- susine:::compute_Xty(X5, sc5$y)
beta_hat5 <- as.vector(xty5 / d5)
shat2_5 <- as.vector(sc5$sigma2 / d5)

auto_est <- susine::estimate_mu_0_scale_factor(
  beta_hat = beta_hat5,
  shat2 = shat2_5,
  mu_0_unscaled = sc5$annotation
)

# Best c from grid (max ELBO)
best_grid <- exp5_grid %>%
  filter(!is.na(elbo)) %>%
  slice_max(elbo, n = 1)

# Compile comparison
comparison <- tribble(
  ~method, ~c_used,
  "Oracle", sc5$c_true,
  "auto_scale", auto_est$c,
  "Best grid (max ELBO)", best_grid$c_val,
  "Vanilla (c=0)", 0
) %>%
  left_join(
    exp5_grid %>% rename(c_used = c_val),
    by = "c_used"
  )

kable(comparison, digits = 4,
      caption = "Experiment 5: Method comparison")
```

# Adversarial Synthesis

## Does the theory hold?

```{r synthesis}
cat("=== SYNTHESIS ===\n\n")

cat("Prediction 1: auto_scale works well at high ann R2.\n")
cat("  → Check exp1 summary table above. RMSE by annotation_r2.\n\n")

cat("Prediction 2: Likelihood surface unimodal in s0^2, flat in c.\n")
cat("  → Check exp2 contour and slice plots above.\n")
cat("  → Is the c-direction slice broader than the s0^2-direction?\n\n")

cat("Prediction 3: 'mean' mode converges to mu_0 ~ 0 (uniform shift).\n")
cat("  → Check exp3 final_mu_0_eb values.\n\n")

cat("Prediction 4: 'both' mode collapses or is initialization-dependent.\n")
cat("  → Check exp4 table for ELBO sensitivity to c_init.\n\n")

cat("Prediction 5: Oracle c gives best AUPRC; grid comes close.\n")
cat("  → Check exp5 comparison table.\n")
```

## Surprises and Open Questions

Fill in after running the workbook:

1.  Did `auto_scale_mu_0` over- or under-estimate c? By how much?
2.  Was the c-direction really flat, or did it have clear curvature?
3.  Did the "mean" mode do anything useful, or was it as predicted?
4.  Is the c-grid ensemble really close to oracle, or is there a big
    gap?
5.  **If auto_scale works well:** could we use it as a warm-start for
    the c-grid (narrowing the grid around c_hat)?
6.  **If auto_scale fails:** does it fail gracefully (c_hat near 0) or
    catastrophically (c_hat wildly wrong)?

## Implications for the Paper

-   If c-grid ELBO-softmax recovers near-oracle performance, the
    ensemble story is strong.
-   If `auto_scale_mu_0` is reliable, it could reduce the c-grid
    size needed (adaptive grid centered on c_hat).
-   If iterative EB for c is truly hopeless, the paper can cite this
    analysis as justification for the grid approach.

```{r session-info}
sessionInfo()
```
