---
title: "Posterior-Geometry Difficulty Metrics — Validation"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Validation of the posterior-geometry difficulty calculus
(`posterior-geometry.md`) against fine-mapping outcomes on the baseline
simulation study. The goal is to test whether the new metrics —
landscape ruggedness `M*`, top-1 dominance `D_1`, signal energy `S_T`,
1-swap decoy gap `G_top1`, and z-aware ruggedness `H` — predict AUPRC
and AUPRC-deltas (truth-warm vs cold, annotated vs cold) better than
the existing M1 + z-only family used by the stratified sampler.

Two stages:
- **Stage 1 (T-aware, simulation-only).** Use the true causal support
  T to compute `M*_T`, `D_1`, `S_T`, `G_top1`. Upper bound on what's
  recoverable.
- **Stage 2 (T-free, real-data analog).** Use only z and R to compute
  `M*_marg`, `D_1_free`, `S_max`, `G_pivot`, `H`. The R² gap between
  Stage 1 and Stage 2 is the cost of not knowing T.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/posterior_geometry_metrics.Rmd")
library(here)
library(devtools)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

theme_set(theme_bw(base_size = 12))
set.seed(1)
```

# 1. User Parameters

```{r params}
job_name <- "baseline_sims_screen"
parent_job_id <- "51509956"
output_root <- here("output")

# Knobs
mc_n_samples       <- 5000        # MC budget for M*, H, S_max
default_L_target   <- 5L          # used for T-free metrics where p_star unknown
ridge_lambda       <- 1e-4        # ridge for chol of R[γ, γ]

# Output
out_dir <- file.path(output_root, "one_off_validations", "posterior_geometry_metrics",
                     parent_job_id)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Locate Aggregated Output

```{r locate}
job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
aggregated_dir <- file.path(job_dir, "aggregated")
run_history_dir <- file.path(output_root, "run_history", job_name, parent_job_id)

stopifnot(dir.exists(aggregated_dir), dir.exists(run_history_dir))

run_metrics    <- readr::read_csv(file.path(aggregated_dir, "model_metrics.csv"),
                                  show_col_types = FALSE)
dataset_metric <- readr::read_csv(file.path(aggregated_dir, "dataset_metrics.csv"),
                                  show_col_types = FALSE)
dataset_bundles <- readr::read_csv(file.path(run_history_dir, "dataset_bundles.csv"),
                                   show_col_types = FALSE)

cat("run_metrics rows:    ", nrow(run_metrics),    "\n")
cat("dataset_metrics rows:", nrow(dataset_metric), "\n")
cat("dataset_bundles rows:", nrow(dataset_bundles), "\n")
```

# 3. Posterior-Geometry Metrics — Helpers

All metrics share the same computational primitive: `z_γ' (R_γ + λI)⁻¹ z_γ`
and `R_γ⁻¹` for small subsets γ ⊂ [p]. Cost per evaluation is O(L³).

```{r helpers}
# ----- core primitives -----------------------------------------------------

# Solve R[γ, γ] x = v (with ridge), return v' x.
pgm_qf <- function(R, idx, v, lambda = ridge_lambda) {
  L <- length(idx)
  Rgg <- R[idx, idx, drop = FALSE] + diag(lambda, L)
  ch <- tryCatch(suppressWarnings(chol(Rgg)), error = function(e) NULL)
  if (is.null(ch)) return(NA_real_)
  as.numeric(crossprod(backsolve(ch, v, transpose = TRUE)))
}

# tr(R[T, γ] R[γ, γ]⁻¹ R[γ, T]) = ||R[T, γ] R[γ, γ]^(-1/2)||²_F
pgm_subspace_energy <- function(R, T_idx, gamma_idx, lambda = ridge_lambda) {
  L <- length(gamma_idx)
  Rgg <- R[gamma_idx, gamma_idx, drop = FALSE] + diag(lambda, L)
  RTg <- R[T_idx, gamma_idx, drop = FALSE]
  M <- tryCatch(RTg %*% solve(Rgg, t(RTg)), error = function(e) NULL)
  if (is.null(M)) return(NA_real_)
  sum(diag(M))
}

# ----- Stage 1 (T-aware) ---------------------------------------------------

# M*_T: variance over random γ of subspace-energy with T fixed.
pgm_M_star_T <- function(R, T_idx, n_samples = mc_n_samples,
                         lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  L <- length(T_idx)
  vals <- vapply(seq_len(n_samples), function(i) {
    gamma <- sample.int(p, L)
    pgm_subspace_energy(R, T_idx, gamma, lambda)
  }, numeric(1))
  c(M_star_T = stats::var(vals, na.rm = TRUE),
    M_star_T_mean = mean(vals, na.rm = TRUE),
    M_star_T_n_finite = sum(is.finite(vals)))
}

# D_1: max_{j∈T} z_j² / max_j z_j².  Range (0, 1].  1 ⇔ marginally strongest is causal.
pgm_D1 <- function(z, T_idx) {
  z2 <- z^2
  max(z2[T_idx], na.rm = TRUE) / max(z2, na.rm = TRUE)
}

# S_T: log-posterior energy at the true support, in z'R⁻¹z scale.
pgm_S_T <- function(R, z, T_idx, lambda = ridge_lambda) {
  pgm_qf(R, T_idx, z[T_idx], lambda)
}

# G_top1: best 1-swap decoy gap.  max_{j∈T, k∉T} S(γ_jk) - S_T,
# where γ_jk = (T \ {j}) ∪ {k}.  Negative means T wins outright.
pgm_G_top1 <- function(R, z, T_idx, lambda = ridge_lambda) {
  S_T <- pgm_S_T(R, z, T_idx, lambda)
  if (!is.finite(S_T)) return(NA_real_)
  p <- nrow(R)
  others <- setdiff(seq_len(p), T_idx)
  best_decoy <- -Inf
  for (j in T_idx) {
    base <- setdiff(T_idx, j)
    for (k in others) {
      g <- c(base, k)
      v <- z[g]
      e <- pgm_qf(R, g, v, lambda)
      if (is.finite(e) && e > best_decoy) best_decoy <- e
    }
  }
  best_decoy - S_T
}

# ----- Stage 2 (T-free) ----------------------------------------------------

# M*_marg: marginalize over T uniformly.  Returns variance averaged over T.
pgm_M_star_marg <- function(R, L_target = default_L_target,
                            n_outer = 50L, n_inner = 200L,
                            lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  vals_var <- numeric(n_outer)
  for (t in seq_len(n_outer)) {
    T_t <- sample.int(p, L_target)
    inner <- vapply(seq_len(n_inner), function(i) {
      gamma <- sample.int(p, L_target)
      pgm_subspace_energy(R, T_t, gamma, lambda)
    }, numeric(1))
    vals_var[t] <- stats::var(inner, na.rm = TRUE)
  }
  c(M_star_marg = mean(vals_var, na.rm = TRUE))
}

# D_1_free: top-2 squared-z / top-1 squared-z.  Small ⇔ clear single winner.
pgm_D1_free <- function(z) {
  s <- sort(z^2, decreasing = TRUE)
  if (length(s) < 2) return(NA_real_)
  s[2] / s[1]
}

# H: variance of z_γ' R_γ⁻¹ z_γ over uniform random γ.
pgm_H <- function(R, z, L_target = default_L_target,
                  n_samples = mc_n_samples, lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  vals <- vapply(seq_len(n_samples), function(i) {
    gamma <- sample.int(p, L_target)
    pgm_qf(R, gamma, z[gamma], lambda)
  }, numeric(1))
  c(H = stats::var(vals, na.rm = TRUE),
    H_mean = mean(vals, na.rm = TRUE))
}

# S_max: best support-energy over MC-sampled γ's of size L_target.
pgm_S_max <- function(R, z, L_target = default_L_target,
                      n_samples = mc_n_samples, lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  vals <- vapply(seq_len(n_samples), function(i) {
    gamma <- sample.int(p, L_target)
    pgm_qf(R, gamma, z[gamma], lambda)
  }, numeric(1))
  c(S_max = max(vals, na.rm = TRUE))
}

# G_pivot: T-free analog of G_top1.  Pivot around the top-K marginal-z
# variants T_hat (call this our best guess at T).  Compute the 1-swap
# decoy gap relative to T_hat.  L = K.
pgm_G_pivot <- function(R, z, L_target = default_L_target,
                        lambda = ridge_lambda) {
  T_hat <- order(z^2, decreasing = TRUE)[seq_len(L_target)]
  pgm_G_top1(R, z, T_hat, lambda)
}

# ----- driver: compute every metric per dataset ---------------------------

pgm_compute_one <- function(R, z, T_idx, L_target,
                            mc_n_samples_local = mc_n_samples,
                            lambda = ridge_lambda, seed = 1L) {
  out <- list()
  # Stage 1
  out$S_T   <- pgm_S_T(R, z, T_idx, lambda)
  out$D_1   <- pgm_D1(z, T_idx)
  m_T       <- pgm_M_star_T(R, T_idx, mc_n_samples_local, lambda, seed = seed)
  out$M_star_T       <- unname(m_T["M_star_T"])
  out$M_star_T_mean  <- unname(m_T["M_star_T_mean"])
  out$G_top1 <- pgm_G_top1(R, z, T_idx, lambda)
  # Stage 2
  out$D_1_free <- pgm_D1_free(z)
  m_marg <- pgm_M_star_marg(R, L_target,
                            n_outer = max(50L, mc_n_samples_local %/% 100L),
                            n_inner = 100L, lambda = lambda, seed = seed + 1L)
  out$M_star_marg <- unname(m_marg["M_star_marg"])
  h <- pgm_H(R, z, L_target, mc_n_samples_local, lambda, seed = seed + 2L)
  out$H        <- unname(h["H"])
  out$H_mean   <- unname(h["H_mean"])
  s_max <- pgm_S_max(R, z, L_target, mc_n_samples_local, lambda, seed = seed + 3L)
  out$S_max    <- unname(s_max["S_max"])
  out$G_pivot  <- pgm_G_pivot(R, z, L_target, lambda)
  tibble::as_tibble(out)
}
```

# 4. Reconstruct R, z, T per Dataset

Re-generate each simulated dataset from its bundle row + the cached
genotype matrix. Cheaper than caching p×p matrices on disk.

```{r reconstruct}
# Identify the unique X matrices we need to load. Matrix paths live in the
# separate `data_matrices.csv` manifest, joined to bundles by matrix_id.
data_matrices <- readr::read_csv(file.path(run_history_dir, "data_matrices.csv"),
                                 show_col_types = FALSE)
needed_matrix_ids <- sort(unique(dataset_bundles$matrix_id))
matrix_paths <- data_matrices %>%
  dplyr::filter(.data$matrix_id %in% needed_matrix_ids) %>%
  dplyr::select(matrix_id, matrix_path) %>%
  dplyr::distinct()
stopifnot(all(needed_matrix_ids %in% matrix_paths$matrix_id))

# Load each X once.
load_X_cache <- function(matrix_paths) {
  out <- list()
  for (i in seq_len(nrow(matrix_paths))) {
    mid <- matrix_paths$matrix_id[i]
    path <- file.path(here(), matrix_paths$matrix_path[i])
    if (!file.exists(path)) {
      stop("Genotype matrix not found: ", path)
    }
    obj <- readRDS(path)
    out[[as.character(mid)]] <- as.matrix(obj)
  }
  out
}

X_cache <- load_X_cache(matrix_paths)

# Per-bundle: regenerate β, y, σ², compute z, T, R.
pgm_one_bundle <- function(bundle_row, X_cache) {
  X <- X_cache[[as.character(bundle_row$matrix_id)]]
  spec <- as.list(bundle_row)
  sim <- generate_simulation_data(spec, base_X = X)
  X_use <- sim$X
  y <- as.numeric(sim$y)
  beta <- as.numeric(sim$beta)
  sigma2 <- stats::var(sim$residuals %||% (y - X_use %*% beta))
  if (!is.finite(sigma2) || sigma2 <= 0) sigma2 <- 1
  T_idx <- which(beta != 0)
  L <- length(T_idx)
  # Standardize columns of X (test_susine convention) and y for z/R consistency
  Xc <- scale(X_use, center = TRUE, scale = TRUE)
  yc <- y - mean(y)
  n  <- nrow(Xc)
  R  <- crossprod(Xc) / n
  z  <- as.numeric(crossprod(Xc, yc) / sqrt(sigma2 * n))
  list(R = R, z = z, T_idx = T_idx, L_truth = L, sigma2 = sigma2)
}

# Per-bundle compute. Costs ~few seconds each (most is the regen + R).
compute_for_bundle <- function(bundle_row, X_cache) {
  data <- pgm_one_bundle(bundle_row, X_cache)
  L_target <- max(data$L_truth, 1L)
  m <- pgm_compute_one(data$R, data$z, data$T_idx, L_target,
                       mc_n_samples_local = mc_n_samples)
  m$L_truth <- data$L_truth
  m$sigma2  <- data$sigma2
  m
}

# Run across bundles. Use a small subset first to verify wiring before
# running the full set.
n_bundles_to_run <- nrow(dataset_bundles)
# n_bundles_to_run <- 20L   # uncomment to dry-run a subset
sub <- dataset_bundles %>% dplyr::slice_head(n = n_bundles_to_run)

cat("Computing posterior-geometry metrics for", nrow(sub), "bundles...\n")
t0 <- Sys.time()
metric_rows <- purrr::map_dfr(seq_len(nrow(sub)), function(i) {
  row <- sub[i, , drop = FALSE]
  m <- compute_for_bundle(row, X_cache)
  m$dataset_bundle_id <- row$dataset_bundle_id[[1]]
  m
})
cat("Done in", round(as.numeric(difftime(Sys.time(), t0, units = "secs")), 1), "sec.\n")

readr::write_csv(metric_rows,
                 file.path(out_dir, "posterior_geometry_metrics.csv"))
```

# 5. Build Outcome Table

For each dataset, the three AUPRC outcomes:
- `auprc_cold`: vanilla SuSiE cold IBSS
- `auprc_warm`: same model, truth-warm initialization
- `auprc_annotated`: best-AUPRC annotated arm (susine_functional_mu over
  any annotation_r2)

```{r outcomes}
# Inspect what's available before filtering.
cat("use_case_ids:\n"); print(sort(unique(run_metrics$use_case_id)))
if ("filtering" %in% names(run_metrics))
  cat("filtering values:", paste(sort(unique(run_metrics$filtering)), collapse = ", "), "\n")
if ("warm_method" %in% names(run_metrics))
  cat("warm_method values:", paste(sort(unique(stats::na.omit(run_metrics$warm_method))), collapse = ", "),
      " (NAs:", sum(is.na(run_metrics$warm_method)), ")\n")

# AUPRC is the same across `filtering = unfiltered/purity_filtered`
# (CS-filtering doesn't change classification metrics), so collapse to one row.
mm <- run_metrics
if ("filtering" %in% names(mm)) {
  mm <- mm %>% dplyr::filter(.data$filtering == "unfiltered" | is.na(.data$filtering))
}

# Cold baseline: vanilla SuSiE, no warm start.
cold_runs <- mm %>% dplyr::filter(.data$use_case_id == "susie_vanilla")
if ("warm_method" %in% names(cold_runs)) {
  cold_runs <- cold_runs %>% dplyr::filter(is.na(.data$warm_method))
}
cold_per_bundle <- cold_runs %>%
  dplyr::group_by(dataset_bundle_id) %>%
  dplyr::summarise(auprc_cold = mean(.data$AUPRC, na.rm = TRUE), .groups = "drop")

# Truth-warm: only present in jobs that include warm-start arms. Skip
# gracefully if absent (e.g. baseline_sims_screen does only cold fits).
has_warm <- "warm_method" %in% names(mm) &&
  any(!is.na(mm$warm_method))
warm_per_bundle <- if (has_warm) {
  mm %>%
    dplyr::filter(.data$use_case_id == "susie_vanilla", !is.na(.data$warm_method)) %>%
    dplyr::group_by(dataset_bundle_id) %>%
    dplyr::summarise(auprc_warm = mean(.data$AUPRC, na.rm = TRUE), .groups = "drop")
} else {
  cat("No truth-warm runs in this job; delta_warm regression will be skipped.\n")
  tibble::tibble(dataset_bundle_id = integer(), auprc_warm = numeric())
}

# Best annotated: max AUPRC across any susine_functional_mu fit.
annot_per_bundle <- mm %>%
  dplyr::filter(.data$use_case_id == "susine_functional_mu") %>%
  dplyr::group_by(dataset_bundle_id) %>%
  dplyr::summarise(auprc_annotated = max(.data$AUPRC, na.rm = TRUE), .groups = "drop")

outcomes <- cold_per_bundle %>%
  dplyr::left_join(warm_per_bundle,  by = "dataset_bundle_id") %>%
  dplyr::left_join(annot_per_bundle, by = "dataset_bundle_id") %>%
  dplyr::mutate(
    delta_warm      = if (has_warm) .data$auprc_warm - .data$auprc_cold else NA_real_,
    delta_annotated = .data$auprc_annotated - .data$auprc_cold
  )

cat("Outcome rows (datasets with cold AUPRC):", nrow(outcomes), "\n")
cat("  with non-NA warm:     ", sum(!is.na(outcomes$auprc_warm)), "\n")
cat("  with non-NA annotated:", sum(!is.na(outcomes$auprc_annotated)), "\n")
```

# 6. Existing Comparator Metric Set

The current real-data stratified sampler uses `M1, M2, M4,
z_count_abs_gt_3, z_eff_signals` (per `0_2_phase1_dataset_metrics_screening.ipynb`),
plus everything `compute_x_metrics_from_R()` already emits (M1, M2, M1_dist,
eigs summaries, off-diag means, row concentration, Toeplitz fit error).

```{r comparator-features}
existing_cols <- c(
  # primary sampler metrics
  "M1", "M2", "M4", "z_count_abs_gt_3", "z_eff_signals",
  # rest of the dataset-metric family present in dataset_metrics
  "M1_dist", "z_topk_ratio", "z_max_abs",
  "high_ld_count_095_per_snp", "high_ld_frac_095",
  "annotation_corr_beta_hat_std", "annotation_corr_z"
)
existing_cols <- intersect(existing_cols, names(dataset_metric))
cat("Existing comparator features:\n")
print(existing_cols)

existing_features <- dataset_metric %>%
  dplyr::select(dplyr::any_of(c("dataset_bundle_id", existing_cols)))
```

# 7. Three Regressions

For each outcome we fit three models:
- **Existing only**: comparator metrics from §6
- **Existing + Stage 2 (T-free)**: comparator + new T-free metrics
- **Existing + Stage 1 (T-aware)**: comparator + new T-aware metrics

```{r regressions}
master <- outcomes %>%
  dplyr::left_join(metric_rows,       by = "dataset_bundle_id") %>%
  dplyr::left_join(existing_features, by = "dataset_bundle_id")

stage1_cols <- c("S_T", "D_1", "M_star_T", "M_star_T_mean", "G_top1")
stage2_cols <- c("S_max", "D_1_free", "M_star_marg", "H", "H_mean", "G_pivot")
stage1_cols <- intersect(stage1_cols, names(master))
stage2_cols <- intersect(stage2_cols, names(master))

fit_one <- function(target, predictors, df) {
  df <- df %>% dplyr::select(dplyr::all_of(c(target, predictors))) %>% tidyr::drop_na()
  if (nrow(df) < length(predictors) + 5L) return(NULL)
  form <- stats::reformulate(predictors, response = target)
  stats::lm(form, data = df)
}

run_three <- function(target_col, label) {
  cat("\n===", label, "(target:", target_col, ") ===\n")
  m_existing <- fit_one(target_col, existing_cols, master)
  m_stage2   <- fit_one(target_col, c(existing_cols, stage2_cols), master)
  m_stage1   <- fit_one(target_col, c(existing_cols, stage1_cols, stage2_cols), master)

  r2 <- function(m) if (is.null(m)) NA_real_ else summary(m)$r.squared
  n_obs <- function(m) if (is.null(m)) 0L else stats::nobs(m)
  tibble::tibble(
    target = target_col,
    n_obs_existing = n_obs(m_existing),
    n_obs_stage2   = n_obs(m_stage2),
    n_obs_stage1_2 = n_obs(m_stage1),
    R2_existing    = r2(m_existing),
    R2_existing_stage2  = r2(m_stage2),
    R2_existing_full    = r2(m_stage1),
    delta_R2_stage2_real = r2(m_stage2) - r2(m_existing),
    delta_R2_stage1_upper = r2(m_stage1) - r2(m_existing)
  )
}

reg_targets <- list(
  list(col = "auprc_cold",      label = "AUPRC cold (baseline)"),
  list(col = "delta_annotated", label = "ΔAUPRC best-annotated vs cold")
)
if (any(!is.na(master$delta_warm))) {
  reg_targets <- append(reg_targets,
    list(list(col = "delta_warm", label = "ΔAUPRC truth-warm vs cold")), after = 1L)
}
reg_summary <- purrr::map_dfr(reg_targets,
                              function(t) run_three(t$col, t$label))

knitr::kable(reg_summary, digits = 3)
readr::write_csv(reg_summary, file.path(out_dir, "regression_summary.csv"))
```

# 8. Per-Predictor Coefficients (the full Stage-1 model)

```{r per-predictor}
report_coefs <- function(target_col) {
  m <- fit_one(target_col,
               c(existing_cols, stage1_cols, stage2_cols),
               master)
  if (is.null(m)) return(tibble::tibble())
  s <- summary(m)$coefficients %>% as.data.frame()
  s$predictor <- rownames(s); rownames(s) <- NULL
  s$target <- target_col
  s %>% dplyr::select(target, predictor, dplyr::everything())
}

coef_tbl <- purrr::map_dfr(reg_targets, function(t) report_coefs(t$col))
readr::write_csv(coef_tbl, file.path(out_dir, "regression_coefficients.csv"))
knitr::kable(head(coef_tbl, 60), digits = 3)
```

# 9. Diagnostic Scatterplots

```{r scatter, fig.width=10, fig.height=8}
make_scatter <- function(x_col, y_col, df, label) {
  ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]])) +
    geom_point(alpha = 0.5, size = 1.2) +
    geom_smooth(method = "loess", se = FALSE, formula = y ~ x) +
    labs(title = label, x = x_col, y = y_col)
}

scatter_targets <- vapply(reg_targets, `[[`, character(1), "col")
scatter_metrics <- intersect(c("S_T", "D_1", "M_star_T", "G_top1",
                               "S_max", "D_1_free", "M_star_marg", "H", "G_pivot",
                               "M1"),
                             names(master))

for (yc in scatter_targets) {
  for (xc in scatter_metrics) {
    p <- make_scatter(xc, yc, master, paste(yc, "vs", xc))
    ggsave(file.path(out_dir, sprintf("%s_vs_%s.png", yc, xc)),
           p, width = 6, height = 4, dpi = 130)
  }
}
cat("Saved", length(scatter_targets) * length(scatter_metrics), "scatter PNGs to", out_dir, "\n")
```

# 10. Take-Home

```{r take-home}
knitr::kable(reg_summary, digits = 3,
             caption = "R² for each target × predictor-set. delta_R2_stage2_real is the realistic upgrade for real-data stratification (T unknown). delta_R2_stage1_upper is the upper bound when T is known.")
```
