---
title: "Posterior-Geometry Difficulty Metrics — Validation"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Validation of the posterior-geometry difficulty calculus
(`posterior-geometry.md`) against fine-mapping outcomes on the baseline
simulation study. The goal is to test whether the new metrics —
landscape ruggedness `M*`, top-1 dominance `D_1`, signal energy `S_T`,
1-swap decoy gap `G_top1`, and z-aware ruggedness `H` — predict AUPRC
and AUPRC-deltas (truth-warm vs cold, annotated vs cold) better than
the existing M1 + z-only family used by the stratified sampler.

Two stages:
- **Stage 1 (T-aware, simulation-only).** Use the true causal support
  T to compute `M*_T`, `D_1`, `S_T`, `G_top1`. Upper bound on what's
  recoverable.
- **Stage 2 (T-free, real-data analog).** Use only z and R to compute
  `M*_marg`, `D_1_free`, `S_max`, `G_pivot`, `H`. The R² gap between
  Stage 1 and Stage 2 is the cost of not knowing T.

```{r setup, message=FALSE}
here::i_am("vignettes/one_off_validations/posterior_geometry_metrics.Rmd")
library(here)
library(devtools)
library(dplyr)
library(readr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

theme_set(theme_bw(base_size = 12))
set.seed(1)
```

# 1. User Parameters

```{r params}
# ---------------------------------------------------------------------------
# All user-tweakable knobs for this workbook live in this chunk.
# ---------------------------------------------------------------------------

# --- Data sources ----------------------------------------------------------

# Study to validate against. Source of AUPRC outcomes + true causal info.
# The 5-arm effect-matching study has both cold and warm-from-annotated fits
# broken out per dataset, which is what we need for the optimizer-barrier
# regression.
study_name <- "baseline_sims_5arm_effect_matching"
output_root <- here("output")
study_root <- file.path(output_root, "effect_matching_study", study_name)

# Source of the matrix catalog. We re-use the screening job's
# `data_matrices.csv` because the 5-arm study uses the same scenario_1
# matrices. Change this if you point the workbook at a study with a
# different matrix-id mapping.
matrices_run_history_dir <- file.path(output_root, "run_history",
                                      "baseline_sims_screen", "51509956")

# --- Compute knobs ---------------------------------------------------------

# Subset of bundles to run. `nrow(dataset_bundles)` for the full 600.
# Smaller value = quick wiring / sanity check.  Bump back up before drawing
# any conclusion: with too few bundles, the regressions become rank-deficient
# (predictors > observations).
n_bundles_to_run <- 600L

# MC budget per (M*, H, S_max) per dataset.  Cost scales linearly.
mc_n_samples <- 5000L

# L_target for the T-free Stage-2 metrics (M*_marg, H, S_max, G_pivot).  In
# real-data application this is the L you'd run susie with (harness default
# 10).  For sims with known p_star, 5 is a sensible default.
default_L_target <- 5L

# Ridge added to chol(R[γ, γ]) for numerical stability.  External LD
# matrices may need lambda ≥ 1e-3 if chol() complains; in-sample LD is
# fine at 1e-4.
ridge_lambda <- 1e-4

# --- Comparator metric set -------------------------------------------------

# Existing comparator metrics to regress against, mirroring the real-data
# stratified sampler in `0_2_phase1_dataset_metrics_screening.ipynb`.
# Names that aren't emitted by `compute_x_metrics_from_R()` /
# `z_score_metrics()` are silently dropped with a printout in §6.
existing_cols <- c(
  # primary sampler metrics
  "M1", "M2", "M4", "z_count_abs_gt_3", "z_eff_signals",
  # additional LD-only descriptors
  "M1_dist", "z_topk_ratio", "z_max_abs"
)

# --- Output ----------------------------------------------------------------

out_dir <- file.path(output_root, "one_off_validations",
                     "posterior_geometry_metrics", study_name)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Locate Aggregated Output

```{r locate}
stopifnot(dir.exists(study_root), dir.exists(matrices_run_history_dir))

# Per-dataset AUPRC outcomes (cold + 4 alternative arms + deltas)
auprc <- readr::read_csv(file.path(study_root, "auprc_per_dataset_5arm.csv"),
                         show_col_types = FALSE)

# True per-dataset causal info: list of length 600, names "1".."600";
# each entry has $causal_idx (integer T) and $beta_at_causal (numeric β at T).
causal_truth <- readRDS(file.path(study_root, "causal_truth_5arm.rds"))

# Bundle specs (matrix_id, phenotype_seed, architecture, ...)
dataset_bundles <- readr::read_csv(file.path(study_root, "dataset_bundles.csv"),
                                   show_col_types = FALSE)

# Matrix catalog: scenario_1 matrices, identical to those used by the
# screening study. We just borrow that file rather than re-building.
data_matrices <- readr::read_csv(file.path(matrices_run_history_dir, "data_matrices.csv"),
                                 show_col_types = FALSE)
job_config_json <- jsonlite::fromJSON(file.path(matrices_run_history_dir, "job_config.json"),
                                      simplifyVector = FALSE)
repo_root <- job_config_json$paths$repo_root %||% here()

cat("auprc rows:           ", nrow(auprc),           "\n")
cat("causal_truth length:  ", length(causal_truth),  "\n")
cat("dataset_bundles rows: ", nrow(dataset_bundles), "\n")
cat("data_matrices rows:   ", nrow(data_matrices),   "\n")
cat("repo_root:            ", repo_root,             "\n")
```

# 3. Posterior-Geometry Metrics — Helpers

All metrics share the same computational primitive: `z_γ' (R_γ + λI)⁻¹ z_γ`
and `R_γ⁻¹` for small subsets γ ⊂ [p]. Cost per evaluation is O(L³).

```{r helpers}
# ----- core primitives -----------------------------------------------------

# Solve R[γ, γ] x = v (with ridge), return v' x.
pgm_qf <- function(R, idx, v, lambda = ridge_lambda) {
  L <- length(idx)
  Rgg <- R[idx, idx, drop = FALSE] + diag(lambda, L)
  ch <- tryCatch(suppressWarnings(chol(Rgg)), error = function(e) NULL)
  if (is.null(ch)) return(NA_real_)
  as.numeric(crossprod(backsolve(ch, v, transpose = TRUE)))
}

# tr(R[T, γ] R[γ, γ]⁻¹ R[γ, T]) = ||R[T, γ] R[γ, γ]^(-1/2)||²_F
pgm_subspace_energy <- function(R, T_idx, gamma_idx, lambda = ridge_lambda) {
  L <- length(gamma_idx)
  Rgg <- R[gamma_idx, gamma_idx, drop = FALSE] + diag(lambda, L)
  RTg <- R[T_idx, gamma_idx, drop = FALSE]
  M <- tryCatch(RTg %*% solve(Rgg, t(RTg)), error = function(e) NULL)
  if (is.null(M)) return(NA_real_)
  sum(diag(M))
}

# ----- Stage 1 (T-aware) ---------------------------------------------------

# M*_T: variance over random γ of subspace-energy with T fixed.
pgm_M_star_T <- function(R, T_idx, n_samples = mc_n_samples,
                         lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  L <- length(T_idx)
  vals <- vapply(seq_len(n_samples), function(i) {
    gamma <- sample.int(p, L)
    pgm_subspace_energy(R, T_idx, gamma, lambda)
  }, numeric(1))
  c(M_star_T = stats::var(vals, na.rm = TRUE),
    M_star_T_mean = mean(vals, na.rm = TRUE),
    M_star_T_n_finite = sum(is.finite(vals)))
}

# D_1: max_{j∈T} z_j² / max_j z_j².  Range (0, 1].  1 ⇔ marginally strongest is causal.
pgm_D1 <- function(z, T_idx) {
  z2 <- z^2
  max(z2[T_idx], na.rm = TRUE) / max(z2, na.rm = TRUE)
}

# S_T: log-posterior energy at the true support, in z'R⁻¹z scale.
pgm_S_T <- function(R, z, T_idx, lambda = ridge_lambda) {
  pgm_qf(R, T_idx, z[T_idx], lambda)
}

# G_top1: best 1-swap decoy gap.  max_{j∈T, k∉T} S(γ_jk) - S_T,
# where γ_jk = (T \ {j}) ∪ {k}.  Negative means T wins outright.
pgm_G_top1 <- function(R, z, T_idx, lambda = ridge_lambda) {
  S_T <- pgm_S_T(R, z, T_idx, lambda)
  if (!is.finite(S_T)) return(NA_real_)
  p <- nrow(R)
  others <- setdiff(seq_len(p), T_idx)
  best_decoy <- -Inf
  for (j in T_idx) {
    base <- setdiff(T_idx, j)
    for (k in others) {
      g <- c(base, k)
      v <- z[g]
      e <- pgm_qf(R, g, v, lambda)
      if (is.finite(e) && e > best_decoy) best_decoy <- e
    }
  }
  best_decoy - S_T
}

# ----- Stage 2 (T-free) ----------------------------------------------------

# M*_marg: marginalize over T uniformly.  Returns variance averaged over T.
pgm_M_star_marg <- function(R, L_target = default_L_target,
                            n_outer = 50L, n_inner = 200L,
                            lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  vals_var <- numeric(n_outer)
  for (t in seq_len(n_outer)) {
    T_t <- sample.int(p, L_target)
    inner <- vapply(seq_len(n_inner), function(i) {
      gamma <- sample.int(p, L_target)
      pgm_subspace_energy(R, T_t, gamma, lambda)
    }, numeric(1))
    vals_var[t] <- stats::var(inner, na.rm = TRUE)
  }
  c(M_star_marg = mean(vals_var, na.rm = TRUE))
}

# D_1_free: top-2 squared-z / top-1 squared-z.  Small ⇔ clear single winner.
pgm_D1_free <- function(z) {
  s <- sort(z^2, decreasing = TRUE)
  if (length(s) < 2) return(NA_real_)
  s[2] / s[1]
}

# H: variance of z_γ' R_γ⁻¹ z_γ over uniform random γ.
pgm_H <- function(R, z, L_target = default_L_target,
                  n_samples = mc_n_samples, lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  vals <- vapply(seq_len(n_samples), function(i) {
    gamma <- sample.int(p, L_target)
    pgm_qf(R, gamma, z[gamma], lambda)
  }, numeric(1))
  c(H = stats::var(vals, na.rm = TRUE),
    H_mean = mean(vals, na.rm = TRUE))
}

# S_max: best support-energy over MC-sampled γ's of size L_target.
pgm_S_max <- function(R, z, L_target = default_L_target,
                      n_samples = mc_n_samples, lambda = ridge_lambda, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  p <- nrow(R)
  vals <- vapply(seq_len(n_samples), function(i) {
    gamma <- sample.int(p, L_target)
    pgm_qf(R, gamma, z[gamma], lambda)
  }, numeric(1))
  c(S_max = max(vals, na.rm = TRUE))
}

# G_pivot: T-free analog of G_top1.  Pivot around the top-K marginal-z
# variants T_hat (call this our best guess at T).  Compute the 1-swap
# decoy gap relative to T_hat.  L = K.
pgm_G_pivot <- function(R, z, L_target = default_L_target,
                        lambda = ridge_lambda) {
  T_hat <- order(z^2, decreasing = TRUE)[seq_len(L_target)]
  pgm_G_top1(R, z, T_hat, lambda)
}

# ----- driver: compute every metric per dataset ---------------------------

pgm_compute_one <- function(R, z, T_idx,
                            L_target_stage1, L_target_stage2,
                            mc_n_samples_local = mc_n_samples,
                            lambda = ridge_lambda, seed = 1L) {
  out <- list()
  # Stage 1: uses true T (length L_target_stage1).
  out$S_T   <- pgm_S_T(R, z, T_idx, lambda)
  out$D_1   <- pgm_D1(z, T_idx)
  m_T       <- pgm_M_star_T(R, T_idx, mc_n_samples_local, lambda, seed = seed)
  out$M_star_T       <- unname(m_T["M_star_T"])
  out$M_star_T_mean  <- unname(m_T["M_star_T_mean"])
  out$G_top1 <- pgm_G_top1(R, z, T_idx, lambda)
  # Stage 2: T-free, fixed L_target_stage2 (would be unknown in real data).
  out$D_1_free <- pgm_D1_free(z)
  m_marg <- pgm_M_star_marg(R, L_target_stage2,
                            n_outer = max(50L, mc_n_samples_local %/% 100L),
                            n_inner = 100L, lambda = lambda, seed = seed + 1L)
  out$M_star_marg <- unname(m_marg["M_star_marg"])
  h <- pgm_H(R, z, L_target_stage2, mc_n_samples_local, lambda, seed = seed + 2L)
  out$H        <- unname(h["H"])
  out$H_mean   <- unname(h["H_mean"])
  s_max <- pgm_S_max(R, z, L_target_stage2, mc_n_samples_local, lambda, seed = seed + 3L)
  out$S_max    <- unname(s_max["S_max"])
  out$G_pivot  <- pgm_G_pivot(R, z, L_target_stage2, lambda)
  tibble::as_tibble(out)
}
```

# 4. Reconstruct R, z, T per Dataset

Re-generate each simulated dataset from its bundle row + the cached
genotype matrix. Cheaper than caching p×p matrices on disk.

```{r reconstruct}
# 4a. Subset of matrices we actually need.
needed_matrix_ids <- sort(unique(dataset_bundles$matrix_id))
matrix_rows <- data_matrices %>%
  dplyr::filter(.data$matrix_id %in% needed_matrix_ids)
stopifnot(all(needed_matrix_ids %in% matrix_rows$matrix_id))

# 4b. Load each X once via the canonical pipeline loader.
cat("Loading", nrow(matrix_rows), "unique X matrices...\n")
X_cache <- list()
t_load <- Sys.time()
for (i in seq_len(nrow(matrix_rows))) {
  mid <- matrix_rows$matrix_id[i]
  X_cache[[as.character(mid)]] <- test_susine:::load_sampled_matrix(
    matrix_rows[i, , drop = FALSE], repo_root = repo_root)
  if (i %% 25 == 0 || i == nrow(matrix_rows))
    cat(sprintf("  %d / %d matrices (%.1f sec)\n",
                i, nrow(matrix_rows),
                as.numeric(difftime(Sys.time(), t_load, units = "secs"))))
}

# 4c. Per-bundle reconstruction.
#  T and β come from causal_truth_5arm.rds (no regen needed).
#  y comes from simulate_phenotype_h2(X, β, h2_total = 0.25, seed = phenotype_seed + 1L)
#  — same call generate_simulation_data() makes for susie2_oligogenic.
#  R = standardized X' X / n, z = X'y / sqrt(σ²·n).
pgm_one_bundle <- function(bundle_row, X_cache, causal_truth) {
  bid <- as.integer(bundle_row$dataset_bundle_id[[1]])
  truth <- causal_truth[[as.character(bid)]]
  if (is.null(truth)) stop("No causal_truth entry for dataset_bundle_id = ", bid)
  T_idx <- as.integer(truth$causal_idx)
  beta_at_T <- as.numeric(truth$beta_at_causal)

  X <- X_cache[[as.character(bundle_row$matrix_id)]]
  p <- ncol(X)
  beta <- numeric(p)
  beta[T_idx] <- beta_at_T

  architecture <- as.character(bundle_row$architecture[[1]])
  seed <- bundle_row$phenotype_seed[[1]]
  phenotype <- if (identical(architecture, "susie2_sparse")) {
    simulate_phenotype_h2(X = X, beta = beta,
                          h2_snp_per_causal = bundle_row$h2_snp_per_causal %||% 0.03,
                          h2_total = NA_real_, seed = seed + 1L)
  } else if (identical(architecture, "susie2_oligogenic")) {
    simulate_phenotype_h2(X = X, beta = beta,
                          h2_snp_per_causal = NA_real_,
                          h2_total = bundle_row$h2_total %||% 0.25, seed = seed + 1L)
  } else {
    simulate_phenotype(X = X, beta = beta,
                       noise_fraction = bundle_row$y_noise %||% 0.5,
                       seed = seed + 1L)
  }

  Xc <- standardize_x(X)
  yc <- phenotype$y - mean(phenotype$y)
  n  <- nrow(Xc)
  R  <- crossprod(Xc) / n
  sigma2 <- as.numeric(phenotype$sigma2)
  if (!is.finite(sigma2) || sigma2 <= 0) sigma2 <- stats::var(yc)
  z  <- as.numeric(crossprod(Xc, yc) / sqrt(sigma2 * n))
  list(R = R, z = z, T_idx = T_idx,
       L_truth = length(T_idx), sigma2 = sigma2)
}

# A small registry that stores existing comparator metrics alongside the new
# ones so we don't re-load X in §6. LD-only metrics (M1, M2, ...) are
# matrix-level so are cached by matrix_id; z-metrics are per-bundle.
x_metric_cache <- new.env(parent = emptyenv())
get_x_metrics_for_matrix <- function(R, mid) {
  key <- as.character(mid)
  if (exists(key, envir = x_metric_cache, inherits = FALSE))
    return(get(key, envir = x_metric_cache))
  m <- compute_x_metrics_from_R(R)
  assign(key, m, envir = x_metric_cache)
  m
}

compute_for_bundle <- function(bundle_row, X_cache, causal_truth,
                               L_target_stage2 = default_L_target) {
  data <- pgm_one_bundle(bundle_row, X_cache, causal_truth)
  if (length(data$T_idx) == 0L) {
    return(tibble::tibble(L_truth = 0L, sigma2 = data$sigma2))
  }
  pgm <- pgm_compute_one(
    R = data$R, z = data$z, T_idx = data$T_idx,
    L_target_stage1 = data$L_truth,
    L_target_stage2 = L_target_stage2,
    mc_n_samples_local = mc_n_samples,
    seed = bundle_row$dataset_bundle_id[[1]]
  )
  x_metrics <- get_x_metrics_for_matrix(data$R, bundle_row$matrix_id[[1]])
  z_metrics <- z_score_metrics(data$z, top_k = 10L) |> tibble::as_tibble()
  dplyr::bind_cols(
    pgm,
    tibble::tibble(L_truth = data$L_truth, sigma2 = data$sigma2),
    x_metrics,
    z_metrics
  )
}

# 4d. n_bundles_to_run is set in §1.
sub <- dataset_bundles %>% dplyr::slice_head(n = n_bundles_to_run)
cat("Computing metrics for", nrow(sub), "bundles...\n")
t0 <- Sys.time()
metric_rows <- purrr::map_dfr(seq_len(nrow(sub)), function(i) {
  row <- sub[i, , drop = FALSE]
  m <- compute_for_bundle(row, X_cache, causal_truth)
  m$dataset_bundle_id <- row$dataset_bundle_id[[1]]
  m$matrix_id <- row$matrix_id[[1]]
  if (i %% 5 == 0 || i == nrow(sub))
    cat(sprintf("  %d / %d bundles (%.1f sec)\n",
                i, nrow(sub),
                as.numeric(difftime(Sys.time(), t0, units = "secs"))))
  m
})
cat("Done.\n")

readr::write_csv(metric_rows,
                 file.path(out_dir, "posterior_geometry_metrics.csv"))
```

# 5. Build Outcome Table

For each dataset, the three AUPRC outcomes:
- `auprc_cold`: vanilla SuSiE cold IBSS
- `auprc_warm`: same model, truth-warm initialization
- `auprc_annotated`: best-AUPRC annotated arm (susine_functional_mu over
  any annotation_r2)

```{r outcomes}
# All three outcomes come straight off the per-dataset 5-arm AUPRC table.
# `auprc_delta_warm_mu_vs_vanilla` = warm-from-functional-μ minus cold; this
# is the optimizer-barrier signal (vanilla finds the basin only when
# initialized from the annotated-arm solution).
outcomes <- auprc %>%
  dplyr::transmute(
    dataset_bundle_id,
    auprc_cold      = .data$auprc_susie_vanilla,
    auprc_warm      = .data$auprc_susie_vanilla_warm_from_mu,
    auprc_annotated = .data$auprc_susine_functional_mu,
    delta_warm      = .data$auprc_delta_warm_mu_vs_vanilla,
    delta_annotated = .data$auprc_delta_mu
  )

cat("Outcome rows:", nrow(outcomes), "\n")
cat("  with non-NA warm:     ", sum(!is.na(outcomes$auprc_warm)), "\n")
cat("  with non-NA annotated:", sum(!is.na(outcomes$auprc_annotated)), "\n")
```

# 6. Existing Comparator Metric Set

The current real-data stratified sampler uses `M1, M2, M4,
z_count_abs_gt_3, z_eff_signals` (per `0_2_phase1_dataset_metrics_screening.ipynb`),
plus everything `compute_x_metrics_from_R()` already emits (M1, M2, M1_dist,
eigs summaries, off-diag means, row concentration, Toeplitz fit error).

```{r comparator-features}
# `existing_cols` is set in §1.  Comparator features were computed inline in
# §4 alongside the new metrics, so they're already columns of `metric_rows`.
# Filter to what's actually present and report what's missing.
requested <- existing_cols
existing_cols <- intersect(existing_cols, names(metric_rows))
missing_cols <- setdiff(requested, existing_cols)
cat("Existing comparator features present in metric_rows:\n")
print(existing_cols)
if (length(missing_cols)) {
  cat("Missing (skipped — not emitted by current helpers):\n")
  print(missing_cols)
}
```

# 7. Three Regressions

For each outcome we fit three models:
- **Existing only**: comparator metrics from §6
- **Existing + Stage 2 (T-free)**: comparator + new T-free metrics
- **Existing + Stage 1 (T-aware)**: comparator + new T-aware metrics

```{r regressions}
# Existing comparator features are already columns of metric_rows, so a
# single join brings everything in. existing_features (defined in §6) is
# kept around purely as a name list for the regressors.
master <- outcomes %>%
  dplyr::left_join(metric_rows, by = "dataset_bundle_id")

stage1_cols <- c("S_T", "D_1", "M_star_T", "M_star_T_mean", "G_top1")
stage2_cols <- c("S_max", "D_1_free", "M_star_marg", "H", "H_mean", "G_pivot")
stage1_cols <- intersect(stage1_cols, names(master))
stage2_cols <- intersect(stage2_cols, names(master))

fit_one <- function(target, predictors, df) {
  df <- df %>% dplyr::select(dplyr::all_of(c(target, predictors))) %>% tidyr::drop_na()
  if (nrow(df) < length(predictors) + 5L) return(NULL)
  form <- stats::reformulate(predictors, response = target)
  stats::lm(form, data = df)
}

run_three <- function(target_col, label) {
  cat("\n===", label, "(target:", target_col, ") ===\n")
  m_existing <- fit_one(target_col, existing_cols, master)
  m_stage2   <- fit_one(target_col, c(existing_cols, stage2_cols), master)
  m_stage1   <- fit_one(target_col, c(existing_cols, stage1_cols, stage2_cols), master)

  r2 <- function(m) if (is.null(m)) NA_real_ else summary(m)$r.squared
  n_obs <- function(m) if (is.null(m)) 0L else stats::nobs(m)
  tibble::tibble(
    target = target_col,
    n_obs_existing = n_obs(m_existing),
    n_obs_stage2   = n_obs(m_stage2),
    n_obs_stage1_2 = n_obs(m_stage1),
    R2_existing    = r2(m_existing),
    R2_existing_stage2  = r2(m_stage2),
    R2_existing_full    = r2(m_stage1),
    delta_R2_stage2_real = r2(m_stage2) - r2(m_existing),
    delta_R2_stage1_upper = r2(m_stage1) - r2(m_existing)
  )
}

reg_targets <- list(
  list(col = "auprc_cold",      label = "AUPRC cold (baseline)"),
  list(col = "delta_annotated", label = "ΔAUPRC best-annotated vs cold")
)
if (any(!is.na(master$delta_warm))) {
  reg_targets <- append(reg_targets,
    list(list(col = "delta_warm",
              label = "ΔAUPRC warm-from-μ vs cold (optimizer-barrier)")),
    after = 1L)
}
reg_summary <- purrr::map_dfr(reg_targets,
                              function(t) run_three(t$col, t$label))

knitr::kable(reg_summary, digits = 3)
readr::write_csv(reg_summary, file.path(out_dir, "regression_summary.csv"))
```

# 8. Per-Predictor Coefficients (the full Stage-1 model)

```{r per-predictor}
report_coefs <- function(target_col) {
  m <- fit_one(target_col,
               c(existing_cols, stage1_cols, stage2_cols),
               master)
  if (is.null(m)) return(tibble::tibble())
  s <- summary(m)$coefficients %>% as.data.frame()
  s$predictor <- rownames(s); rownames(s) <- NULL
  s$target <- target_col
  s %>% dplyr::select(target, predictor, dplyr::everything())
}

coef_tbl <- purrr::map_dfr(reg_targets, function(t) report_coefs(t$col))
readr::write_csv(coef_tbl, file.path(out_dir, "regression_coefficients.csv"))
knitr::kable(head(coef_tbl, 60), digits = 3)
```

# 9. Diagnostic Scatterplots

```{r scatter, fig.width=10, fig.height=8}
make_scatter <- function(x_col, y_col, df, label) {
  ggplot(df, aes(x = .data[[x_col]], y = .data[[y_col]])) +
    geom_point(alpha = 0.5, size = 1.2) +
    geom_smooth(method = "loess", se = FALSE, formula = y ~ x) +
    labs(title = label, x = x_col, y = y_col)
}

scatter_targets <- vapply(reg_targets, `[[`, character(1), "col")
scatter_metrics <- intersect(c("S_T", "D_1", "M_star_T", "G_top1",
                               "S_max", "D_1_free", "M_star_marg", "H", "G_pivot",
                               "M1"),
                             names(master))

for (yc in scatter_targets) {
  for (xc in scatter_metrics) {
    p <- make_scatter(xc, yc, master, paste(yc, "vs", xc))
    ggsave(file.path(out_dir, sprintf("%s_vs_%s.png", yc, xc)),
           p, width = 6, height = 4, dpi = 130)
  }
}
cat("Saved", length(scatter_targets) * length(scatter_metrics), "scatter PNGs to", out_dir, "\n")
```

# 10. Take-Home

```{r take-home}
knitr::kable(reg_summary, digits = 3,
             caption = "R² for each target × predictor-set. delta_R2_stage2_real is the realistic upgrade for real-data stratification (T unknown). delta_R2_stage1_upper is the upper bound when T is known.")
```
