---
title: "Verification: sigma_0_2 Bug Fix (susine vs susieR)"
date: "2026-03-04"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    code_folding: show
editor_options:
  markdown:
    wrap: 72
---

# Background

**Bug**: In `run_model.R`, the harness was passing
`sigma_0_2 = sigma_scalar * var(y)` to `susine::susine()`. But
`susine::initialize_priors()` (line 277 of `initialize.R`) internally
does `sigma_0_2 <- sigma_0_2 * var_y`. The result: susine received an
effective prior variance of $0.2 \times \text{var}(y)^2$ instead of
$0.2 \times \text{var}(y)$.

**Fix**: Changed `run_model.R` line 1138 to pass the raw scalar (e.g.,
0.2) without pre-multiplying by `var(y)`.
`susieR::susie()` uses `scaled_prior_variance` directly as the fraction
of `var(y)`, and `susine` does its own internal scaling — so both now
receive the same semantic value.

**This vignette** verifies the fix at three levels:

1.  **Direct engine calls** — match susine and susieR with identical
    arguments, no harness.
2.  **Through `run_use_case()`** — the actual harness function that
    dispatches to each backend.
3.  **Across multiple scenarios** — vary p\*, noise, sigma scalars, EB
    vs fixed, to build confidence the fix is robust.

```{r setup, message=FALSE, warning=FALSE}
here::i_am(
  "vignettes/one_off_validations/verify_sigma_0_2_bugfix.Rmd"
)
library(here)
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)

devtools::load_all(here("../susine"), quiet = TRUE)
devtools::load_all(here("../susieR"), quiet = TRUE)
devtools::load_all(here("."), quiet = TRUE)

set.seed(20260304)
```

# Level 1: Direct Engine Comparison

Same synthetic data, same hyperparameters, no harness. If this fails,
the engines themselves have diverged.

```{r direct-engine-data}
n <- 400L
p <- 200L
L <- 10L
scaled_prior_variance <- 0.2
tol <- 1e-5
max_iter <- 200L

# Correlated design (AR1-ish, rho = 0.6)
X <- matrix(rnorm(n * p), nrow = n)
for (j in 2:p) {
  X[, j] <- 0.6 * X[, j - 1] + sqrt(1 - 0.6^2) * X[, j]
}

causal_idx <- sort(sample(seq_len(p), 5))
beta_true <- rep(0, p)
beta_true[causal_idx] <- rnorm(length(causal_idx), sd = 0.7)

signal <- drop(X %*% beta_true)
noise_var <- var(signal) * (1 - 0.30) / 0.30
y <- signal + rnorm(n, sd = sqrt(noise_var))

cat(sprintf("n=%d, p=%d, L=%d, causal=%s\n", n, p, L,
            paste(causal_idx, collapse = ",")))
cat(sprintf("var(y)=%.4f, PVE≈0.30\n", var(y)))
```

## 1.1 Fixed Prior Variance

```{r direct-fixed}
# --- susieR ---
susie_formals <- names(formals(susieR::susie))
susie_args <- list(
  X = X, y = y, L = L,
  scaled_prior_variance = scaled_prior_variance,
  estimate_prior_variance = FALSE,
  estimate_residual_variance = TRUE,
  standardize = TRUE, intercept = TRUE,
  tol = tol, max_iter = max_iter
)
if ("convergence_method" %in% susie_formals)
  susie_args$convergence_method <- "elbo"
if ("refine" %in% susie_formals)
  susie_args$refine <- FALSE
susie_args <- susie_args[intersect(names(susie_args), susie_formals)]
fit_susieR <- do.call(susieR::susie, susie_args)

# --- susine ---
# Key: pass raw scalar (0.2), NOT 0.2 * var(y)
fit_susine <- susine::susine(
  L = L, X = X, y = y,
  mu_0 = 0,
  sigma_0_2 = scaled_prior_variance,
  prior_inclusion_weights = NULL,
  prior_update_method = "none",
  auto_scale_mu_0 = FALSE,
  auto_scale_sigma_0_2 = FALSE,
  tol = tol, max_iter = max_iter
)

pip_R <- fit_susieR$pip
pip_S <- fit_susine$model_fit$PIPs
```

```{r direct-fixed-checks}
check_agreement <- function(pip_a, pip_b, fit_a, fit_b,
                            label = "comparison",
                            max_pip_tol = 0.02,
                            mean_pip_tol = 0.005,
                            cor_tol = 0.995) {
  pip_diff <- abs(pip_a - pip_b)
  metrics <- tibble::tibble(
    metric = c("max_pip_diff", "mean_pip_diff", "pip_correlation",
               "top10_overlap", "sigma2_rel_diff", "elbo_rel_diff"),
    value = c(
      max(pip_diff),
      mean(pip_diff),
      cor(pip_a, pip_b),
      length(intersect(order(pip_a, decreasing = TRUE)[1:10],
                        order(pip_b, decreasing = TRUE)[1:10])) / 10,
      abs(tail(fit_b$model_fit$sigma_2, 1) - fit_a$sigma2) /
        fit_a$sigma2,
      abs(tail(fit_b$model_fit$elbo, 1) - tail(fit_a$elbo, 1)) /
        abs(tail(fit_a$elbo, 1))
    ),
    threshold = c(max_pip_tol, mean_pip_tol, NA, NA, 0.05, 0.01),
    pass = c(
      max(pip_diff) < max_pip_tol,
      mean(pip_diff) < mean_pip_tol,
      cor(pip_a, pip_b) > cor_tol,
      length(intersect(order(pip_a, decreasing = TRUE)[1:10],
                        order(pip_b, decreasing = TRUE)[1:10])) >= 7,
      abs(tail(fit_b$model_fit$sigma_2, 1) - fit_a$sigma2) /
        fit_a$sigma2 < 0.05,
      abs(tail(fit_b$model_fit$elbo, 1) - tail(fit_a$elbo, 1)) /
        abs(tail(fit_a$elbo, 1)) < 0.01
    )
  )
  cat(sprintf("\n=== %s ===\n", label))
  print(metrics, n = 10)
  cat(sprintf("%d / %d passed\n", sum(metrics$pass), nrow(metrics)))
  metrics
}

results_fixed <- check_agreement(
  pip_R, pip_S, fit_susieR, fit_susine,
  label = "Level 1: Fixed variance (direct calls)"
)
```

```{r direct-fixed-scatter, fig.width=6, fig.height=6}
plot_df <- tibble::tibble(susieR = pip_R, susine = pip_S)
ggplot(plot_df, aes(x = susieR, y = susine)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  geom_point(alpha = 0.5, size = 1.5) +
  labs(title = "Level 1: PIPs — susieR vs susine (fixed variance)",
       subtitle = sprintf("cor=%.6f, max|diff|=%.6f",
                          cor(pip_R, pip_S), max(abs(pip_R - pip_S))),
       x = "susieR PIP", y = "susine PIP") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_minimal(base_size = 13)
```

## 1.2 EB Prior Variance

```{r direct-eb}
susie_args_eb <- susie_args
susie_args_eb$estimate_prior_variance <- TRUE
susie_args_eb$scaled_prior_variance <- scaled_prior_variance
fit_susieR_eb <- do.call(susieR::susie, susie_args_eb)

fit_susine_eb <- susine::susine(
  L = L, X = X, y = y,
  mu_0 = 0,
  sigma_0_2 = scaled_prior_variance,
  prior_inclusion_weights = NULL,
  prior_update_method = "var",
  auto_scale_mu_0 = FALSE,
  auto_scale_sigma_0_2 = FALSE,
  tol = tol, max_iter = max_iter
)

results_eb <- check_agreement(
  fit_susieR_eb$pip, fit_susine_eb$model_fit$PIPs,
  fit_susieR_eb, fit_susine_eb,
  label = "Level 1: EB prior variance (direct calls)"
)
```

```{r direct-eb-scatter, fig.width=6, fig.height=6}
plot_df_eb <- tibble::tibble(
  susieR = fit_susieR_eb$pip,
  susine = fit_susine_eb$model_fit$PIPs
)
ggplot(plot_df_eb, aes(x = susieR, y = susine)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  geom_point(alpha = 0.5, size = 1.5) +
  labs(title = "Level 1: PIPs — susieR vs susine (EB)",
       subtitle = sprintf("cor=%.6f, max|diff|=%.6f",
                          cor(plot_df_eb$susieR, plot_df_eb$susine),
                          max(abs(plot_df_eb$susieR - plot_df_eb$susine))),
       x = "susieR PIP", y = "susine PIP") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_minimal(base_size = 13)
```

## 1.3 Demonstration: the OLD bug

Show what happens if you pass `sigma_0_2 = 0.2 * var(y)` (the
old buggy harness behavior). The PIPs should diverge badly.

```{r old-bug-demo}
fit_susine_buggy <- susine::susine(
  L = L, X = X, y = y,
  mu_0 = 0,
  sigma_0_2 = scaled_prior_variance * var(y),
  prior_inclusion_weights = NULL,
  prior_update_method = "none",
  auto_scale_mu_0 = FALSE,
  auto_scale_sigma_0_2 = FALSE,
  tol = tol, max_iter = max_iter
)

pip_buggy <- fit_susine_buggy$model_fit$PIPs
max_diff_buggy <- max(abs(pip_R - pip_buggy))
cor_buggy <- cor(pip_R, pip_buggy)

cat("=== Old bug: sigma_0_2 pre-multiplied by var(y) ===\n")
cat(sprintf("max |PIP diff| with susieR:  %.4f  (should be LARGE)\n", max_diff_buggy))
cat(sprintf("PIP correlation with susieR: %.6f  (should be LOW)\n", cor_buggy))
cat(sprintf("\nWith the FIX:  max diff = %.6f, cor = %.8f\n",
            max(abs(pip_R - pip_S)), cor(pip_R, pip_S)))
```

```{r old-bug-scatter, fig.width=10, fig.height=5}
plot_bug <- tibble::tibble(
  susieR = c(pip_R, pip_R),
  susine = c(pip_S, pip_buggy),
  version = rep(c("Fixed (correct)", "Buggy (old)"), each = p)
)

ggplot(plot_bug, aes(x = susieR, y = susine)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  geom_point(alpha = 0.4, size = 1.5) +
  facet_wrap(~version) +
  labs(title = "Bug demonstration: correct vs old (buggy) sigma_0_2 handling",
       x = "susieR PIP", y = "susine PIP") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_minimal(base_size = 13)
```

---

# Level 2: Through the Harness (`run_use_case`)

Call `run_use_case()` with mocked `use_case`, `run_row`, and
`job_config` objects for both "susine" and "susieR" backends.
This exercises the actual code path that the SLURM jobs use, including
`parse_sigma_0_2_scalar` and the arg list construction.

```{r harness-setup}
# Build a minimal data_bundle
data_bundle <- list(
  X = X,
  y = y,
  beta = beta_true,
  effect_sd = sd(beta_true[causal_idx]),
  priors_cache = new.env(parent = emptyenv())
)

# Build a run_row that corresponds to vanilla fixed-variance
run_row_fixed <- tibble::tibble(
  dataset_bundle_id = 1L,
  use_case_id = "susie_vanilla",
  matrix_id = "test_matrix_1",
  p_star = length(causal_idx),
  y_noise = round(1 - 0.30, 2),
  L = L,
  sigma_0_2_scalar = "0.2",
  annotation_r2 = NA_real_,
  inflate_match = NA_real_,
  c_value = NA_real_,
  tau_value = NA_real_,
  restart_id = NA_integer_,
  refine_step = NA_integer_,
  run_type = "cold",
  phenotype_seed = 42L,
  annotation_seed = NA_integer_
)

# Job config with explicit tol
job_config <- list(
  job = list(
    max_iter = max_iter,
    tol = tol
  )
)
```

## 2.1 susieR Backend via Harness

```{r harness-susieR}
uc_susieR <- tibble::tibble(
  use_case_id  = "susie_vanilla",
  backend      = "susieR",
  prior_mean_strategy     = "none",
  prior_variance_strategy = "fixed",
  inclusion_prior_strategy = "uniform",
  unmappable_effects = "none"
)

result_susieR <- test_susine:::run_use_case(
  use_case    = uc_susieR,
  run_row     = run_row_fixed,
  data_bundle = data_bundle,
  job_config  = job_config
)

pip_harness_R <- result_susieR$fits[[1]]$model_fit$PIPs
cat(sprintf("susieR via harness: %d PIPs, max=%.4f\n",
            length(pip_harness_R), max(pip_harness_R)))
```

## 2.2 susine Backend via Harness

```{r harness-susine}
uc_susine <- tibble::tibble(
  use_case_id  = "susine_vanilla",
  backend      = "susine",
  prior_mean_strategy     = "none",
  prior_variance_strategy = "fixed",
  inclusion_prior_strategy = "uniform",
  unmappable_effects = "none"
)

run_row_susine <- run_row_fixed %>%
  dplyr::mutate(use_case_id = "susine_vanilla")

result_susine <- test_susine:::run_use_case(
  use_case    = uc_susine,
  run_row     = run_row_susine,
  data_bundle = data_bundle,
  job_config  = job_config
)

pip_harness_S <- result_susine$fits[[1]]$model_fit$PIPs
cat(sprintf("susine via harness: %d PIPs, max=%.4f\n",
            length(pip_harness_S), max(pip_harness_S)))
```

## 2.3 Harness Comparison

```{r harness-compare}
results_harness <- check_agreement(
  pip_harness_R, pip_harness_S,
  result_susieR$fits[[1]]$raw, result_susine$fits[[1]],
  label = "Level 2: Through run_use_case() harness"
)
```

```{r harness-scatter, fig.width=6, fig.height=6}
plot_harness <- tibble::tibble(susieR = pip_harness_R, susine = pip_harness_S)
ggplot(plot_harness, aes(x = susieR, y = susine)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  geom_point(alpha = 0.5, size = 1.5) +
  labs(title = "Level 2: PIPs through the harness",
       subtitle = sprintf("cor=%.6f, max|diff|=%.6f",
                          cor(pip_harness_R, pip_harness_S),
                          max(abs(pip_harness_R - pip_harness_S))),
       x = "susieR PIP (via harness)", y = "susine PIP (via harness)") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_minimal(base_size = 13)
```

---

# Level 3: Multi-Scenario Sweep

Systematically vary signal density, noise level, and sigma scalar.
Each scenario fits both backends via `run_use_case()` and records
agreement metrics.

```{r sweep-scenarios}
scenarios <- tidyr::expand_grid(
  p_star = c(2, 5, 15),
  y_noise = c(0.5, 0.7, 0.9),
  sigma_0_2_scalar = c(0.1, 0.2, 0.5)
)

cat(sprintf("Running %d scenarios...\n", nrow(scenarios)))
```

```{r sweep-run, message=FALSE, warning=FALSE}
sweep_results <- purrr::pmap_dfr(scenarios, function(p_star, y_noise, sigma_0_2_scalar) {
  # --- Generate fresh data for this scenario ---
  loc_beta <- rep(0, p)
  loc_causal <- sort(sample(seq_len(p), p_star))
  loc_beta[loc_causal] <- rnorm(p_star, sd = 0.6)

  loc_signal <- drop(X %*% loc_beta)
  loc_pve <- 1 - y_noise
  if (var(loc_signal) < 1e-12) {
    loc_noise_var <- 1
  } else {
    loc_noise_var <- var(loc_signal) * y_noise / loc_pve
  }
  loc_y <- loc_signal + rnorm(n, sd = sqrt(loc_noise_var))

  loc_bundle <- list(
    X = X, y = loc_y,
    beta = loc_beta,
    effect_sd = if (p_star > 0) sd(loc_beta[loc_causal]) else 0.5,
    priors_cache = new.env(parent = emptyenv())
  )

  loc_row <- run_row_fixed %>%
    dplyr::mutate(
      p_star = !!p_star,
      y_noise = !!y_noise,
      sigma_0_2_scalar = as.character(!!sigma_0_2_scalar)
    )

  # susieR
  loc_row_R <- loc_row %>% dplyr::mutate(use_case_id = "susie_vanilla")
  res_R <- tryCatch(
    test_susine:::run_use_case(uc_susieR, loc_row_R, loc_bundle, job_config),
    error = function(e) NULL
  )

  # susine
  loc_row_S <- loc_row %>% dplyr::mutate(use_case_id = "susine_vanilla")
  res_S <- tryCatch(
    test_susine:::run_use_case(uc_susine, loc_row_S, loc_bundle, job_config),
    error = function(e) NULL
  )

  if (is.null(res_R) || is.null(res_S)) {
    return(tibble::tibble(
      p_star = p_star, y_noise = y_noise,
      sigma_0_2_scalar = sigma_0_2_scalar,
      max_pip_diff = NA_real_, mean_pip_diff = NA_real_,
      pip_cor = NA_real_, sigma2_rel_diff = NA_real_,
      status = "ERROR"
    ))
  }

  loc_pip_R <- res_R$fits[[1]]$model_fit$PIPs
  loc_pip_S <- res_S$fits[[1]]$model_fit$PIPs
  loc_diff <- abs(loc_pip_R - loc_pip_S)

  loc_sigma2_R <- res_R$fits[[1]]$raw$sigma2
  loc_sigma2_S <- tail(res_S$fits[[1]]$model_fit$sigma_2, 1)

  tibble::tibble(
    p_star = p_star,
    y_noise = y_noise,
    sigma_0_2_scalar = sigma_0_2_scalar,
    max_pip_diff = max(loc_diff),
    mean_pip_diff = mean(loc_diff),
    pip_cor = cor(loc_pip_R, loc_pip_S),
    sigma2_rel_diff = abs(loc_sigma2_S - loc_sigma2_R) / loc_sigma2_R,
    status = "OK"
  )
})

cat(sprintf("Completed %d scenarios.\n", nrow(sweep_results)))
```

## 3.1 Sweep Summary Table

```{r sweep-table}
sweep_results %>%
  dplyr::arrange(dplyr::desc(max_pip_diff)) %>%
  print(n = 40)

cat("\n=== Worst-case metrics across all scenarios ===\n")
cat(sprintf("max of max_pip_diff:  %.6f\n", max(sweep_results$max_pip_diff, na.rm = TRUE)))
cat(sprintf("max of mean_pip_diff: %.6f\n", max(sweep_results$mean_pip_diff, na.rm = TRUE)))
cat(sprintf("min pip_cor:          %.6f\n", min(sweep_results$pip_cor, na.rm = TRUE)))
cat(sprintf("max sigma2_rel_diff:  %.6f\n", max(sweep_results$sigma2_rel_diff, na.rm = TRUE)))
```

## 3.2 Agreement Heatmap

```{r sweep-heatmap, fig.width=10, fig.height=7}
ggplot(
  sweep_results %>% dplyr::filter(status == "OK"),
  aes(x = factor(sigma_0_2_scalar), y = factor(p_star), fill = max_pip_diff)
) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.4f", max_pip_diff)), size = 3.5) +
  facet_wrap(~paste("y_noise =", y_noise), ncol = 3) +
  scale_fill_gradient(low = "white", high = "firebrick",
                      limits = c(0, max(0.05, max(sweep_results$max_pip_diff, na.rm = TRUE)))) +
  labs(
    title = "Max |PIP diff| across susieR/susine (via harness)",
    subtitle = "All values should be < 0.02 — small numerical differences expected",
    x = expression(sigma[0]^2 ~ "scalar"), y = "p*", fill = "Max |PIP diff|"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "right")
```

## 3.3 Correlation Heatmap

```{r sweep-cor-heatmap, fig.width=10, fig.height=7}
ggplot(
  sweep_results %>% dplyr::filter(status == "OK"),
  aes(x = factor(sigma_0_2_scalar), y = factor(p_star), fill = pip_cor)
) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.5f", pip_cor)), size = 3.5) +
  facet_wrap(~paste("y_noise =", y_noise), ncol = 3) +
  scale_fill_gradient(low = "orange", high = "forestgreen",
                      limits = c(min(0.99, min(sweep_results$pip_cor, na.rm = TRUE)), 1)) +
  labs(
    title = "PIP Correlation across susieR/susine (via harness)",
    subtitle = "All values should be > 0.995",
    x = expression(sigma[0]^2 ~ "scalar"), y = "p*", fill = "PIP cor"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "right")
```

---

# Level 4: EB Variance via Harness

Repeat Level 2 with `prior_variance_strategy = "eb"` to confirm
the EB path also matches.

```{r harness-eb-setup}
uc_susieR_eb <- uc_susieR %>%
  dplyr::mutate(
    use_case_id = "susie_eb",
    prior_variance_strategy = "eb"
  )

uc_susine_eb <- uc_susine %>%
  dplyr::mutate(
    use_case_id = "susine_eb",
    prior_variance_strategy = "eb"
  )

run_row_eb <- run_row_fixed %>%
  dplyr::mutate(sigma_0_2_scalar = NA_character_)
```

```{r harness-eb-run}
result_R_eb <- test_susine:::run_use_case(
  uc_susieR_eb,
  run_row_eb %>% dplyr::mutate(use_case_id = "susie_eb"),
  data_bundle, job_config
)

result_S_eb <- test_susine:::run_use_case(
  uc_susine_eb,
  run_row_eb %>% dplyr::mutate(use_case_id = "susine_eb"),
  data_bundle, job_config
)

pip_h_R_eb <- result_R_eb$fits[[1]]$model_fit$PIPs
pip_h_S_eb <- result_S_eb$fits[[1]]$model_fit$PIPs

results_harness_eb <- check_agreement(
  pip_h_R_eb, pip_h_S_eb,
  result_R_eb$fits[[1]]$raw, result_S_eb$fits[[1]],
  label = "Level 4: EB variance through harness"
)
```

```{r harness-eb-scatter, fig.width=6, fig.height=6}
plot_eb <- tibble::tibble(susieR = pip_h_R_eb, susine = pip_h_S_eb)
ggplot(plot_eb, aes(x = susieR, y = susine)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  geom_point(alpha = 0.5, size = 1.5) +
  labs(title = "Level 4: PIPs — EB prior variance (via harness)",
       subtitle = sprintf("cor=%.6f, max|diff|=%.6f",
                          cor(pip_h_R_eb, pip_h_S_eb),
                          max(abs(pip_h_R_eb - pip_h_S_eb))),
       x = "susieR PIP", y = "susine PIP") +
  coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_minimal(base_size = 13)
```

---

# Final Verdict

```{r final-verdict}
all_results <- dplyr::bind_rows(
  results_fixed  %>% dplyr::mutate(level = "L1: Direct fixed"),
  results_eb     %>% dplyr::mutate(level = "L1: Direct EB"),
  results_harness %>% dplyr::mutate(level = "L2: Harness fixed"),
  results_harness_eb %>% dplyr::mutate(level = "L4: Harness EB")
)

cat("================================================================\n")
cat("  FINAL VERIFICATION SUMMARY\n")
cat("================================================================\n\n")

all_results %>%
  dplyr::select(level, metric, value, pass) %>%
  print(n = 30)

n_total <- nrow(all_results)
n_pass <- sum(all_results$pass)

cat(sprintf("\nDirect checks: %d / %d passed\n", n_pass, n_total))

# Sweep summary
sweep_ok <- sweep_results %>% dplyr::filter(status == "OK")
sweep_pass <- sum(sweep_ok$max_pip_diff < 0.02 & sweep_ok$pip_cor > 0.995)
cat(sprintf("Sweep scenarios: %d / %d within tolerance\n",
            sweep_pass, nrow(sweep_ok)))
cat(sprintf("Total: %d / %d all clear\n",
            n_pass + sweep_pass, n_total + nrow(sweep_ok)))

if (n_pass == n_total && sweep_pass == nrow(sweep_ok)) {
  cat("\n*** ALL CHECKS PASSED — sigma_0_2 bug is squashed. ***\n")
} else {
  cat("\n!!! SOME CHECKS FAILED — investigate above. !!!\n")
}
```
