---
title: "Annotation Contamination Sensitivity - Collect Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Post-hoc collection for the reviewer annotation-contamination sensitivity jobs.
This workbook aggregates the null-arm and causal-positive-control Slurm outputs,
validates the annotation construction gates, and writes combined tables for the
visualization workbook.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_annotation_contamination.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(knitr)
```

# Configuration

Set the parent Slurm array IDs returned by `sbatch` before rendering.

```{r user-config}
output_root <- here("output")

null_job_name <- "annotation_contamination_null"
null_parent_job_id <- "SET_NULL_PARENT_JOB_ID"

causal_job_name <- "annotation_contamination_causal_positive"
causal_parent_job_id <- "SET_CAUSAL_PARENT_JOB_ID"

force_reaggregate <- TRUE
write_snps_dataset <- FALSE
precision_threshold <- 0.75

combined_root <- file.path(output_root, "annotation_contamination_sensitivity")
combined_dir <- file.path(combined_root, "combined")
dir.create(combined_dir, recursive = TRUE, showWarnings = FALSE)

if (identical(null_parent_job_id, "SET_NULL_PARENT_JOB_ID") ||
    identical(causal_parent_job_id, "SET_CAUSAL_PARENT_JOB_ID")) {
  stop("Set null_parent_job_id and causal_parent_job_id before collecting.")
}
```

# Aggregate Arms

```{r aggregate-arms}
collect_arm <- function(job_name, parent_job_id, arm_label) {
  job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
  aggregated_dir <- file.path(job_dir, "aggregated")
  run_table_path <- file.path(output_root, "run_history", job_name, parent_job_id, "run_table.csv")

  if (force_reaggregate || !file.exists(file.path(aggregated_dir, "confusion_bins.csv"))) {
    test_susine::aggregate_staging_outputs(
      job_name = job_name,
      parent_job_id = parent_job_id,
      output_root = output_root,
      validate = TRUE,
      output_dir = aggregated_dir,
      write_snps_dataset = write_snps_dataset
    )
  }

  if (!file.exists(run_table_path)) {
    stop("Run table not found: ", run_table_path)
  }

  list(
    arm = arm_label,
    job_name = job_name,
    parent_job_id = parent_job_id,
    job_dir = job_dir,
    aggregated_dir = aggregated_dir,
    run_table = readr::read_csv(run_table_path, show_col_types = FALSE)
  )
}

arms <- list(
  null = collect_arm(null_job_name, null_parent_job_id, "null"),
  causal = collect_arm(causal_job_name, causal_parent_job_id, "causal")
)

arm_index <- tibble::tibble(
  arm = vapply(arms, `[[`, character(1), "arm"),
  job_name = vapply(arms, `[[`, character(1), "job_name"),
  parent_job_id = vapply(arms, `[[`, character(1), "parent_job_id"),
  aggregated_dir = vapply(arms, `[[`, character(1), "aggregated_dir")
)
readr::write_csv(arm_index, file.path(combined_dir, "arm_index.csv"))
kable(arm_index)
```

# Load Tables

```{r load-tables}
read_optional_csv <- function(dir, filename) {
  path <- file.path(dir, filename)
  if (!file.exists(path)) return(NULL)
  readr::read_csv(path, show_col_types = FALSE)
}

bind_arm_table <- function(filename, join_run_table = FALSE) {
  pieces <- lapply(arms, function(a) {
    tbl <- read_optional_csv(a$aggregated_dir, filename)
    if (is.null(tbl)) return(NULL)
    tbl <- dplyr::mutate(tbl, arm = a$arm, source_job_name = a$job_name, source_parent_job_id = a$parent_job_id)
    if (join_run_table && "run_id" %in% names(tbl)) {
      run_lookup <- a$run_table %>%
        dplyr::distinct(.data$run_id, .keep_all = TRUE) %>%
        dplyr::select(
          .data$run_id,
          dplyr::any_of(c(
            "spec_name", "use_case_id", "group_key", "annotation_contamination_arm",
            "annotation_contamination_lambda", "annotation_contamination_shuffle_z",
            "c_value", "sigma_0_2_scalar", "L", "dataset_bundle_id"
          ))
        )
      tbl <- tbl %>%
        dplyr::select(-dplyr::any_of(setdiff(names(run_lookup), "run_id"))) %>%
        dplyr::left_join(run_lookup, by = "run_id")
    }
    tbl
  })
  dplyr::bind_rows(pieces)
}

confusion_bins <- bind_arm_table("confusion_bins.csv")
model_metrics <- bind_arm_table("model_metrics.csv")
effect_metrics <- bind_arm_table("effect_metrics.csv")
effect_metrics_unfiltered <- bind_arm_table("effect_metrics_unfiltered.csv")
annotation_diagnostics <- bind_arm_table("annotation_diagnostics.csv")
prior_diagnostics <- bind_arm_table("prior_diagnostics.csv", join_run_table = TRUE)
validation <- bind_arm_table("validation.csv")

cat("Confusion-bin rows:", nrow(confusion_bins), "\n")
cat("Annotation diagnostic rows:", nrow(annotation_diagnostics), "\n")
cat("Prior diagnostic rows:", nrow(prior_diagnostics), "\n")
```

# Validation Gates

```{r validation-gates}
required_annotation_cols <- c(
  "annotation_contamination_arm", "annotation_contamination_lambda",
  "null_rms_rel_error", "causal_rms_rel_error", "causal_unchanged", "null_unchanged"
)
missing_annotation_cols <- setdiff(required_annotation_cols, names(annotation_diagnostics))
if (!nrow(annotation_diagnostics) || length(missing_annotation_cols)) {
  stop(
    "annotation_diagnostics.csv is missing or incomplete. Missing columns: ",
    paste(missing_annotation_cols, collapse = ", ")
  )
}

annotation_gate <- annotation_diagnostics %>%
  dplyr::mutate(
    active_rms_rel_error = dplyr::if_else(
      .data$annotation_contamination_arm == "null",
      .data$null_rms_rel_error,
      .data$causal_rms_rel_error
    ),
    untouched_block_ok = dplyr::if_else(
      .data$annotation_contamination_arm == "null",
      .data$causal_unchanged,
      .data$null_unchanged
    )
  ) %>%
  dplyr::group_by(.data$arm, .data$annotation_contamination_lambda) %>%
  dplyr::summarise(
    n_rows = dplyr::n(),
    max_active_rms_rel_error = max(.data$active_rms_rel_error, na.rm = TRUE),
    all_untouched_block_ok = all(.data$untouched_block_ok, na.rm = TRUE),
    max_abs_cor_anchor_low_pip = max(abs(.data$cor_a_z_anchor_low_pip), na.rm = TRUE),
    .groups = "drop"
  )

annotation_gate_overall <- annotation_gate %>%
  dplyr::summarise(
    n_rows = sum(.data$n_rows),
    max_active_rms_rel_error = max(.data$max_active_rms_rel_error, na.rm = TRUE),
    all_untouched_block_ok = all(.data$all_untouched_block_ok, na.rm = TRUE)
  )

required_prior_cols <- c("spec_name", "c_l")
missing_prior_cols <- setdiff(required_prior_cols, names(prior_diagnostics))
if (!nrow(prior_diagnostics) || length(missing_prior_cols)) {
  stop(
    "prior_diagnostics.csv is missing or incomplete. Set write_prior_diagnostics = TRUE before launch. Missing columns: ",
    paste(missing_prior_cols, collapse = ", ")
  )
}

prior_gate <- prior_diagnostics %>%
  dplyr::filter(.data$spec_name == "C-CS-annotation-contamination") %>%
  dplyr::summarise(
    n_rows = dplyr::n(),
    max_abs_c_l_minus_1 = max(abs(.data$c_l - 1), na.rm = TRUE),
    all_c_l_one = all(abs(.data$c_l - 1) < 1e-12, na.rm = TRUE)
  )

if (annotation_gate_overall$n_rows <= 0L ||
    annotation_gate_overall$max_active_rms_rel_error > 1e-8 ||
    !isTRUE(annotation_gate_overall$all_untouched_block_ok)) {
  stop("Annotation construction gate failed. Inspect annotation_gate.csv.")
}
if (prior_gate$n_rows <= 0L || !isTRUE(prior_gate$all_c_l_one)) {
  stop("Fixed-c prior gate failed: c_l differs from 1 or no C-CS prior rows were found.")
}

readr::write_csv(annotation_gate, file.path(combined_dir, "annotation_gate.csv"))
readr::write_csv(annotation_gate_overall, file.path(combined_dir, "annotation_gate_overall.csv"))
readr::write_csv(prior_gate, file.path(combined_dir, "prior_gate.csv"))

kable(annotation_gate_overall, digits = 4, caption = "Annotation construction gate")
kable(prior_gate, digits = 4, caption = "Fixed-c prior gate")
```

# Metrics

```{r metrics}
ccs_confusion <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "C-CS-annotation-contamination",
    .data$explore_method == "aggregation",
    .data$agg_method == "cluster_weight_credible"
  )

baseline_confusion <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "baseline-single",
    .data$explore_method == "base"
  )

if (!nrow(ccs_confusion)) {
  stop("No C-CS aggregate confusion-bin rows found.")
}
if (!nrow(baseline_confusion)) {
  stop("No baseline-single base confusion-bin rows found.")
}

metric_group_vars <- c("arm", "annotation_contamination_arm", "annotation_contamination_lambda")
baseline_group_vars <- "arm"

auprc_pooled <- test_susine::compute_auprc_from_confusion(
  ccs_confusion,
  group_vars = metric_group_vars
)

recall75_pooled <- test_susine::compute_recall_at_precision_from_confusion(
  ccs_confusion,
  group_vars = metric_group_vars,
  precision_threshold = precision_threshold
)

baseline_auprc <- test_susine::compute_auprc_from_confusion(
  baseline_confusion,
  group_vars = baseline_group_vars
)

baseline_recall75 <- test_susine::compute_recall_at_precision_from_confusion(
  baseline_confusion,
  group_vars = baseline_group_vars,
  precision_threshold = precision_threshold
)

metric_summary <- auprc_pooled %>%
  dplyr::left_join(recall75_pooled, by = metric_group_vars) %>%
  dplyr::mutate(series = "C-CS annotation contamination") %>%
  dplyr::arrange(.data$arm, .data$annotation_contamination_lambda)

baseline_summary <- baseline_auprc %>%
  dplyr::left_join(baseline_recall75, by = baseline_group_vars) %>%
  dplyr::mutate(
    series = "Annotation-free SuSiE-equivalent baseline",
    annotation_contamination_arm = NA_character_,
    annotation_contamination_lambda = NA_real_
  ) %>%
  dplyr::select(
    .data$arm,
    .data$annotation_contamination_arm,
    .data$annotation_contamination_lambda,
    dplyr::everything()
  )

lambda0_gate <- metric_summary %>%
  dplyr::filter(abs(.data$annotation_contamination_lambda) < .Machine$double.eps) %>%
  dplyr::summarise(
    n_lambda0_rows = dplyr::n(),
    auprc_range = max(.data$AUPRC, na.rm = TRUE) - min(.data$AUPRC, na.rm = TRUE),
    recall75_range = max(.data$recall_at_precision, na.rm = TRUE) - min(.data$recall_at_precision, na.rm = TRUE),
    gate = "cross-arm C-CS lambda-zero agreement"
  )

if (lambda0_gate$n_lambda0_rows < 2L ||
    lambda0_gate$auprc_range > 1e-12 || lambda0_gate$recall75_range > 1e-12) {
  stop("Lambda-zero cross-arm C-CS metric gate failed.")
}

readr::write_csv(metric_summary, file.path(combined_dir, "metric_summary.csv"))
readr::write_csv(baseline_summary, file.path(combined_dir, "baseline_summary.csv"))
readr::write_csv(lambda0_gate, file.path(combined_dir, "lambda0_gate.csv"))

kable(metric_summary, digits = 4, caption = "Pooled annotation sensitivity metrics")
kable(baseline_summary, digits = 4, caption = "Annotation-free baseline metrics")
kable(lambda0_gate, digits = 4, caption = "Lambda-zero cross-arm C-CS gate")
```

# Export Combined Raw Tables

```{r export}
write_if_present <- function(tbl, filename) {
  if (!is.null(tbl) && nrow(tbl)) {
    readr::write_csv(tbl, file.path(combined_dir, filename))
  }
}

write_if_present(confusion_bins, "confusion_bins_full.csv")
write_if_present(model_metrics, "model_metrics_full.csv")
write_if_present(effect_metrics, "effect_metrics_full.csv")
write_if_present(effect_metrics_unfiltered, "effect_metrics_unfiltered_full.csv")
write_if_present(annotation_diagnostics, "annotation_diagnostics_full.csv")
write_if_present(prior_diagnostics, "prior_diagnostics_full.csv")
write_if_present(validation, "validation_full.csv")

cat("Combined annotation sensitivity outputs written to:", combined_dir, "\n")
```