---
title: "Ensemble Scaling Experiment — Collect Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Post-hoc aggregation, subsampling analysis, oracle computation.
See `refs/ensemble_scaling_experiment_plan.md` for full spec.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_ensemble_scaling.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
}
library(dplyr)
library(tidyr)
library(readr)
library(arrow)
library(knitr)
```

# 1. Load Job Outputs

```{r load-config}
# TODO: update with actual job_id after SLURM run
job_name <- "ensemble_scaling_v1"
job_id <- "XXXXXX"

job_dir <- here("output", "run_history", job_name, job_id)
cfg <- jsonlite::fromJSON(file.path(job_dir, "job_config.json"),
                           simplifyVector = FALSE)
run_table <- readr::read_csv(file.path(job_dir, "run_table.csv"),
                              show_col_types = FALSE)

cat("Total runs:", nrow(run_table), "\n")
cat("Specs:", paste(unique(run_table$spec_name), collapse = ", "), "\n")
```

```{r load-staging}
# Aggregate staging outputs (flush files -> merged CSVs/parquets)
agg_dir <- file.path(job_dir, "aggregate")
if (!dir.exists(agg_dir)) {
  dir.create(agg_dir, recursive = TRUE)
}

# Run the standard staging aggregation if not already done
staging_dir <- file.path(job_dir, "staging")
if (dir.exists(staging_dir)) {
  aggregate_staging_outputs(
    staging_dir = staging_dir,
    output_dir = agg_dir,
    job_config_path = file.path(job_dir, "job_config.json")
  )
}

# Load merged outputs
model_metrics <- readr::read_csv(
  file.path(agg_dir, "model_metrics.csv"), show_col_types = FALSE
)
cat("Model metric rows:", nrow(model_metrics), "\n")
```

# 2. Load Per-Run PIP Vectors

```{r load-pips}
# Read per-run SNP-level PIPs from parquet
snps_dir <- file.path(agg_dir, "snps_dataset")
if (dir.exists(snps_dir)) {
  snps_files <- list.files(snps_dir, pattern = "\\.parquet$", full.names = TRUE)
  snps_all <- arrow::open_dataset(snps_dir) %>%
    dplyr::collect()
  cat("SNP-level rows:", nrow(snps_all), "\n")
} else {
  warning("snps_dataset directory not found. PIP-level analysis unavailable.")
  snps_all <- NULL
}
```

# 3. Grading: Top 8 as Causal

For susie2_oligogenic, grade using top 8 causal variants (3 sparse +
5 oligogenic effects) by absolute effect size. The 15 polygenic effects
are too weak to detect.

```{r grading}
grade_top_k <- 8L

# Build per-dataset grading index: top 8 causal by |beta|
grading_index <- model_metrics %>%
  dplyr::filter(!is.na(auprc)) %>%
  dplyr::distinct(dataset_bundle_id) %>%
  dplyr::mutate(top_k_causal = grade_top_k)

# The actual causal subsetting happens inside evaluation; the grading
# convention is enforced when computing AUPRC, power, FDR below.
```

# 4. Oracle Aggregation

Oracle = truth-aware ceiling. For each ensemble group (dataset x spec x
annotation_r2), select the single model with the best AUPRC.

```{r oracle}
# Oracle: best single model per group by AUPRC
oracle_rows <- model_metrics %>%
  dplyr::filter(!is.na(auprc), !isTRUE(is_agg)) %>%
  dplyr::group_by(dataset_bundle_id, spec_name, annotation_r2, group_key) %>%
  dplyr::slice_max(order_by = auprc, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(agg_method = "oracle")

cat("Oracle rows:", nrow(oracle_rows), "\n")
```

# 5. Subsampling Analysis

## 5a. Single-Lever Scaling (64x grids -> {4, 8, 16, 32, 64})

```{r subsample-64, eval=FALSE}
# For each pure 64-run spec, subsample ensembles at smaller sizes.
# This produces scaling curves without additional model fits.

pure_64_specs <- c("A-R", "A-F", "A-S", "B-R", "B-F", "C-C")
subsample_sizes <- c(4L, 8L, 16L, 32L, 64L)
n_draws <- 50L  # random subsets for restart-type specs

scaling_results <- list()

for (spec in pure_64_specs) {
  spec_runs <- model_metrics %>%
    dplyr::filter(spec_name == spec, !isTRUE(is_agg))

  spec_pips <- snps_all %>%
    dplyr::semi_join(spec_runs, by = "run_id")

  for (n_sub in subsample_sizes) {
    for (draw_i in seq_len(n_draws)) {
      # Subsample n_sub runs from this spec's full set per dataset
      # For grids (c, sigma_0_2): structured subsampling (every k-th)
      # For restarts: random subset
      # For refinements: prefix (first n_sub/2 BFS steps)

      # TODO: implement subsample logic per spec type
      # subsample_pips <- ...
      # agg_pips <- aggregate_use_case_pips(subsample_pips, ...)
      # auprc <- compute_auprc(agg_pips, truth)

      scaling_results[[length(scaling_results) + 1L]] <- tibble::tibble(
        spec_name = spec,
        subsample_size = n_sub,
        draw = draw_i,
        agg_method = "placeholder",
        auprc = NA_real_
      )
    }
  }
}

scaling_curves <- dplyr::bind_rows(scaling_results)
readr::write_csv(scaling_curves, file.path(agg_dir, "scaling_curves.csv"))
```

## 5b. Interaction Scaling (8x8 -> 4x4)

```{r subsample-8x8, eval=FALSE}
# For 8x8 interaction specs, subsample to 4x4 by taking every other
# point on each axis.

interaction_specs <- c("A-RF", "A-RS", "A-FS", "C-CS")

interaction_scaling <- list()
for (spec in interaction_specs) {
  # TODO: structured 4x4 subsampling of 8x8 grids
  # Take every other restart_id x every other refine_step (or sigma value)
}

interaction_scaling_df <- dplyr::bind_rows(interaction_scaling)
readr::write_csv(interaction_scaling_df,
                  file.path(agg_dir, "interaction_scaling.csv"))
```

# 6. Fixed-Budget Comparison

```{r budget-comparison, eval=FALSE}
# At budget B=64, compare performance across all specs.
# Each spec that uses exactly 64 runs is directly comparable.

budget_64_specs <- c("A-R", "A-S",
                     "A-RF", "A-RS", "A-FS", "A-RFS",
                     "B-R", "B-RF",
                     "C-C", "C-CS", "C-CFR")

budget_comparison <- model_metrics %>%
  dplyr::filter(spec_name %in% budget_64_specs, isTRUE(is_agg)) %>%
  dplyr::select(dataset_bundle_id, spec_name, annotation_r2,
                agg_method, auprc, power, mean_size, mean_purity)

readr::write_csv(budget_comparison,
                  file.path(agg_dir, "budget_comparison.csv"))
```

# 7. Aggregation Method Comparison

```{r aggregation-comparison}
# Compare all 5 aggregation methods across all specs and ensemble sizes.
agg_comparison <- model_metrics %>%
  dplyr::filter(isTRUE(is_agg) | agg_method == "oracle") %>%
  dplyr::bind_rows(oracle_rows) %>%
  dplyr::group_by(spec_name, agg_method, annotation_r2) %>%
  dplyr::summarise(
    mean_auprc = mean(auprc, na.rm = TRUE),
    sd_auprc = sd(auprc, na.rm = TRUE),
    n = dplyr::n(),
    .groups = "drop"
  )

readr::write_csv(agg_comparison,
                  file.path(agg_dir, "aggregation_comparison.csv"))
knitr::kable(agg_comparison, digits = 4,
             caption = "Aggregation method comparison")
```

# 8. Dataset Metrics

```{r dataset-metrics}
dataset_metrics_path <- file.path(agg_dir, "dataset_metrics.csv")
if (file.exists(dataset_metrics_path)) {
  dataset_metrics <- readr::read_csv(dataset_metrics_path,
                                      show_col_types = FALSE)
  cat("Dataset metric rows:", nrow(dataset_metrics), "\n")
} else {
  warning("dataset_metrics.csv not found.")
  dataset_metrics <- NULL
}
```

# 9. Export Consolidated Outputs

```{r export}
# Write oracle rows to consolidated model_metrics
model_metrics_with_oracle <- dplyr::bind_rows(model_metrics, oracle_rows)
readr::write_csv(model_metrics_with_oracle,
                  file.path(agg_dir, "model_metrics_with_oracle.csv"))
cat("Consolidated model_metrics written with",
    nrow(oracle_rows), "oracle rows.\n")
```

# 10. Runtime Summary

```{r runtime}
if ("wall_time_sec" %in% names(model_metrics)) {
  model_metrics %>%
    dplyr::filter(!is.na(wall_time_sec), !isTRUE(is_agg)) %>%
    dplyr::group_by(spec_name) %>%
    dplyr::summarise(
      mean_sec = mean(wall_time_sec, na.rm = TRUE),
      median_sec = median(wall_time_sec, na.rm = TRUE),
      total_hrs = sum(wall_time_sec, na.rm = TRUE) / 3600,
      n_runs = dplyr::n(),
      .groups = "drop"
    ) %>%
    knitr::kable(digits = 2, caption = "Runtime by spec")
}
```
