---
title: "Dirichlet Alpha Pilot — Collect & Visualize"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Collects staging outputs from the Dirichlet alpha pilot run,
then computes TPR @ FDR = 0.05 by alpha value across all datasets
and aggregation methods.

```{r setup, message=FALSE}
here::i_am(
  "vignettes/simulation pipeline/collect_results_workbook_dirichlet_pilot.Rmd"
)
library(here)
here()

library(devtools)
devtools::load_all(".")

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
theme_set(theme_bw(base_size = 13))
```

# 1. Job Settings

```{r settings}
job_name <- "dirichlet_alpha_pilot"
# Fill in your SLURM parent job ID after submission:
parent_job_id <- "50040324"
output_root <- here("output")
```

# 2. Aggregate Staging Outputs

```{r aggregate}
agg_dir <- file.path(
  output_root, "slurm_output", job_name, parent_job_id, "aggregated"
)

if (!dir.exists(agg_dir) || !file.exists(file.path(agg_dir, "confusion_bins.csv"))) {
  aggregate_staging_outputs(
    job_name = job_name,
    parent_job_id = parent_job_id,
    output_root = output_root,
    validate = TRUE
  )
}

stopifnot(file.exists(file.path(agg_dir, "confusion_bins.csv")))
cat("Aggregated outputs found at:", agg_dir, "\n")
```

# 3. Load Confusion Bins

```{r load-data}
conf <- readr::read_csv(
  file.path(agg_dir, "confusion_bins.csv"),
  show_col_types = FALSE
)

cat("Rows:", nrow(conf), "\n")
cat("Columns:", paste(names(conf), collapse = ", "), "\n")
cat("Alpha values:", sort(unique(conf$alpha_concentration)), "\n")

# Sanity check: alpha_concentration should be present from enrichment
stopifnot("alpha_concentration" %in% names(conf))
```

# 4. Compute TPR @ FDR = 0.05 by Alpha

Strategy: from confusion bins, reconstruct the cumulative TP/FP curve
by sweeping PIP thresholds from high to low. At each threshold, compute
FDR = FP / (TP + FP) and TPR = TP / total_causal. Find the threshold
where FDR <= 0.05 and report the corresponding TPR.

```{r tpr-at-fdr}
target_fdr <- 0.05

# Filter to purity-filtered, non-aggregated runs only
# (agg_method = NA means individual model fits)
conf_fits <- conf %>%
  filter(is.na(agg_method) | agg_method == "")

# Sum confusion bins across ALL runs per alpha value,
# collapsing across datasets, restarts, everything.
bins_by_alpha <- conf_fits %>%
  group_by(alpha_concentration, pip_threshold) %>%
  summarise(
    n_causal = sum(n_causal_at_bucket, na.rm = TRUE),
    n_noncausal = sum(n_noncausal_at_bucket, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(alpha_concentration, desc(pip_threshold))

# Cumulative TP/FP from high PIP threshold down
tpr_fdr_by_alpha <- bins_by_alpha %>%
  group_by(alpha_concentration) %>%
  mutate(
    cum_tp = cumsum(n_causal),
    cum_fp = cumsum(n_noncausal),
    total_causal = sum(n_causal),
    fdr = cum_fp / pmax(cum_tp + cum_fp, 1),
    tpr = cum_tp / pmax(total_causal, 1)
  ) %>%
  ungroup()

# Find TPR at the most aggressive threshold where FDR <= target
tpr_at_fdr <- tpr_fdr_by_alpha %>%
  filter(fdr <= target_fdr) %>%
  group_by(alpha_concentration) %>%
  slice_max(tpr, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(alpha_concentration, pip_threshold, tpr, fdr, cum_tp, cum_fp, total_causal)

cat("TPR @ FDR <=", target_fdr, "by alpha:\n")
print(as.data.frame(tpr_at_fdr))
```

# 5. Same Analysis on Aggregated PIPs

```{r tpr-at-fdr-aggregated}
# Now look at aggregated PIPs (where agg_method is set)
conf_agg <- conf %>%
  filter(!is.na(agg_method) & agg_method != "")

if (nrow(conf_agg) > 0) {
  bins_by_alpha_agg <- conf_agg %>%
    group_by(alpha_concentration, agg_method, pip_threshold) %>%
    summarise(
      n_causal = sum(n_causal_at_bucket, na.rm = TRUE),
      n_noncausal = sum(n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(alpha_concentration, agg_method, desc(pip_threshold))

  tpr_fdr_agg <- bins_by_alpha_agg %>%
    group_by(alpha_concentration, agg_method) %>%
    mutate(
      cum_tp = cumsum(n_causal),
      cum_fp = cumsum(n_noncausal),
      total_causal = sum(n_causal),
      fdr = cum_fp / pmax(cum_tp + cum_fp, 1),
      tpr = cum_tp / pmax(total_causal, 1)
    ) %>%
    ungroup()

  tpr_at_fdr_agg <- tpr_fdr_agg %>%
    filter(fdr <= target_fdr) %>%
    group_by(alpha_concentration, agg_method) %>%
    slice_max(tpr, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    select(alpha_concentration, agg_method, pip_threshold, tpr, fdr,
           cum_tp, cum_fp, total_causal)

  cat("TPR @ FDR <=", target_fdr, "by alpha x aggregation method:\n")
  print(as.data.frame(tpr_at_fdr_agg))
} else {
  cat("No aggregated runs found in confusion_bins.\n")
  tpr_at_fdr_agg <- tibble()
}
```

# 6. Plot

```{r plot-individual}
fig_dir <- file.path(output_root, "figures", job_name)
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

p1 <- ggplot(tpr_at_fdr, aes(x = factor(alpha_concentration), y = tpr)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", tpr)), vjust = -0.5, size = 3.5) +
  labs(
    x = "Dirichlet alpha",
    y = paste0("TPR @ FDR \u2264 ", target_fdr),
    title = "Individual Fits: TPR by Dirichlet Concentration",
    subtitle = "Pooled across all 450 datasets, all restarts"
  ) +
  ylim(0, NA) +
  theme(axis.text.x = element_text(size = 11))

ggsave(file.path(fig_dir, "tpr_by_alpha_individual.png"),
       p1, width = 7, height = 4, dpi = 150)
cat("Saved:", file.path(fig_dir, "tpr_by_alpha_individual.png"), "\n")
```

```{r plot-aggregated}
if (nrow(tpr_at_fdr_agg) > 0) {
  p2 <- ggplot(
    tpr_at_fdr_agg,
    aes(x = factor(alpha_concentration), y = tpr, fill = agg_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_text(
      aes(label = sprintf("%.3f", tpr)),
      position = position_dodge(width = 0.7),
      vjust = -0.5, size = 3
    ) +
    labs(
      x = "Dirichlet alpha",
      y = paste0("TPR @ FDR \u2264 ", target_fdr),
      fill = "Aggregation",
      title = "Aggregated PIPs: TPR by Alpha x Aggregation Method",
      subtitle = "Pooled across all 450 datasets"
    ) +
    ylim(0, NA) +
    theme(axis.text.x = element_text(size = 11))

  ggsave(file.path(fig_dir, "tpr_by_alpha_aggregated.png"),
         p2, width = 9, height = 5, dpi = 150)
  cat("Saved:", file.path(fig_dir, "tpr_by_alpha_aggregated.png"), "\n")
}
```
