---
title: "Ensemble Pilot (cluster-metric sweep) — Collect & AUPRC Sweep"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Post-hoc sweep: load the stored per-fit PIPs + per-fit `elbo_final`, then for
every (clustering metric x threshold) re-cluster the K=64 functional-grid fits,
aggregate them with **Method B** (frequency-free: cluster weight ~ exp(max-ELBO),
split within-cluster by ELBO-softmax), and compute the **overall AUPRC** of the
aggregated PIPs. No refitting — the SLURM job already did the only expensive
work. Re-run this as many times as you like with different grids.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_ensemble_pilot.Rmd")
library(here)
library(devtools)
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")
library(dplyr)
library(tidyr)
library(readr)
library(arrow)
library(ggplot2)
library(knitr)

# Method B everywhere in this sweep.
options(rd_preview.cluster_aggregation = "method_b")

# Headless PNG device for HPC (no X11/Cairo).
have_ragg <- requireNamespace("ragg", quietly = TRUE)
save_png <- function(filename, plot, width = 8, height = 5, dpi = 150) {
  dev <- if (have_ragg) ragg::agg_png else grDevices::png
  ggplot2::ggsave(filename, plot, width = width, height = height, dpi = dpi, device = dev)
}
```

# 1. Locate Output

```{r locate}
job_name      <- "ensemble_pilot_cluster_metric"
parent_job_id <- "REPLACE_WITH_SLURM_JOB_ID"
output_root   <- here("output")

stopifnot(parent_job_id != "REPLACE_WITH_SLURM_JOB_ID")
job_dir  <- file.path(output_root, "slurm_output", job_name, parent_job_id)
cfg_path <- file.path(output_root, "run_history", job_name, parent_job_id, "job_config.json")
cfg <- test_susine:::load_job_config(cfg_path)

out_dir <- file.path(job_dir, "consolidated")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
fig_dir <- file.path(job_dir, "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
cat("Job dir:", job_dir, "\n")
```

# 2. Sweep Grids (knobs)

```{r sweep-grids}
# First-pass WIDE log-spaced (exponential) grids. Ranges are set generously so
# each curve spans the full regime: low threshold -> ~64 singleton clusters
# (Method B ~ ELBO-softmax over all fits); high threshold -> 1 cluster (also
# ~ELBO-softmax). The interesting clustering behavior lives in between; refine
# with a denser grid there on the second pass. The whole sweep is post-hoc, so
# width is free.
exp_grid <- function(lo, hi, n = 24L) signif(exp(seq(log(lo), log(hi), length.out = n)), 3)
threshold_grids <- list(
  cosine        = exp_grid(1e-3, 1.05, 24),  # 1 - cos, distances in [0, 1]
  l2            = exp_grid(1e-2, 6.0,  24),  # raw Euclidean on PIP vectors
  bernoulli_jsd = exp_grid(1e-2, 15.0, 24)   # sum of per-variant binary JSDs
)
metrics <- names(threshold_grids)
cat("Metrics:", paste(metrics, collapse = ", "), "\n")
for (m in metrics) cat(sprintf("  %s: %d thresholds [%.3f, %.3f]\n",
                                m, length(threshold_grids[[m]]),
                                min(threshold_grids[[m]]), max(threshold_grids[[m]])))
```

# 3. Load Per-Fit PIPs, ELBOs, Causal Labels

```{r load}
staging_index <- test_susine::index_staging_outputs(job_name, parent_job_id, output_root = output_root)
snp_files <- staging_index %>% dplyr::filter(.data$type == "snps") %>% dplyr::pull(.data$path)
stopifnot(length(snp_files) > 0)

snps <- dplyr::bind_rows(lapply(snp_files, function(p)
  as.data.frame(arrow::read_parquet(p))))
# `causal_scored` = top-8 largest-effect causal mask (the paper's AUPRC
# positives). Older pilot runs predate it -> fall back to full causal.
if (!"causal_scored" %in% names(snps)) snps$causal_scored <- snps$causal
snps <- dplyr::select(snps, dplyr::any_of(
  c("run_id", "snp_index", "pip", "causal", "causal_scored")))
# The PIP-only lean parquet contains individual fits only (no aggregated rows).
cat("causal_scored present:", "causal_scored" %in% names(snps), "\n")

# Per-fit elbo_final from model_metrics (one row per run after de-dup on filtering).
task_dirs <- list.dirs(job_dir, recursive = FALSE, full.names = TRUE)
task_dirs <- task_dirs[grepl("task-", basename(task_dirs))]
mm_files <- sort(unlist(lapply(task_dirs, list.files, pattern = "model_metrics\\.csv$", full.names = TRUE)))
model_metrics <- dplyr::bind_rows(lapply(mm_files, readr::read_csv,
  show_col_types = FALSE, progress = FALSE))
elbo_by_run <- model_metrics %>%
  dplyr::distinct(.data$run_id, .keep_all = TRUE) %>%
  dplyr::select(.data$run_id, .data$elbo_final)

# Run metadata: locus (dataset_bundle_id) + grid coords.
run_info <- cfg$tables$runs %>%
  dplyr::select(.data$run_id, .data$dataset_bundle_id, .data$spec_name,
                .data$c_value, .data$sigma_0_2_scalar) %>%
  dplyr::filter(.data$spec_name == "C-CS")

snps <- snps %>%
  dplyr::inner_join(run_info, by = "run_id") %>%
  dplyr::inner_join(elbo_by_run, by = "run_id")

n_loci <- dplyr::n_distinct(snps$dataset_bundle_id)
cat("Loci:", n_loci, "\n")
cat("Fits per locus (median):",
    snps %>% dplyr::distinct(dataset_bundle_id, run_id) %>%
      dplyr::count(dataset_bundle_id) %>% dplyr::pull(n) %>% stats::median(), "\n")
```

```{r build-locus-structures}
# For each locus: pip_mat (K fits x p SNPs), elbo_vec (length K), causal (length p).
locus_ids <- sort(unique(snps$dataset_bundle_id))
pip_list <- list(); elbo_list <- list(); causal_list <- list()
for (lid in locus_ids) {
  s <- dplyr::filter(snps, .data$dataset_bundle_id == lid)
  runs_l <- sort(unique(s$run_id))
  p_order <- sort(unique(s$snp_index))
  pm <- matrix(NA_real_, nrow = length(runs_l), ncol = length(p_order),
               dimnames = list(as.character(runs_l), as.character(p_order)))
  for (ri in seq_along(runs_l)) {
    sr <- s[s$run_id == runs_l[[ri]], ]
    pm[ri, as.character(sr$snp_index)] <- sr$pip
  }
  ev <- elbo_by_run$elbo_final[match(runs_l, elbo_by_run$run_id)]
  # Label = top-8 scored causal mask (paper metric); identical across fits.
  cs <- s %>% dplyr::filter(.data$run_id == runs_l[[1]]) %>%
    dplyr::arrange(.data$snp_index) %>% dplyr::pull(.data$causal_scored)
  keep <- length(runs_l) >= 2 && all(is.finite(ev)) && length(cs) == ncol(pm)
  if (!keep) next
  pip_list[[as.character(lid)]]    <- pm
  elbo_list[[as.character(lid)]]   <- as.numeric(ev)
  causal_list[[as.character(lid)]] <- as.integer(cs)
}
cat("Usable loci:", length(pip_list), "\n")
cat("Pooled prevalence (scored positives / all SNPs):",
    round(mean(unlist(causal_list)), 4), "\n")
```

# 4. Reference Aggregations (metric/threshold-independent)

```{r reference}
auprc_pooled <- function(pip_by_locus) {
  pooled_pip <- unlist(pip_by_locus, use.names = FALSE)
  pooled_lab <- unlist(causal_list[names(pip_by_locus)], use.names = FALSE)
  test_susine:::auprc_average_precision(pooled_pip, pooled_lab)
}
auprc_mean_perlocus <- function(pip_by_locus) {
  mean(vapply(names(pip_by_locus), function(lid)
    test_susine:::auprc_average_precision(pip_by_locus[[lid]], causal_list[[lid]]),
    numeric(1)), na.rm = TRUE)
}

ref_methods <- list(
  single_best  = function(pm, ev) pm[which.max(ev), ],
  uniform      = function(pm, ev) colMeans(pm),
  elbo_softmax = function(pm, ev) as.numeric(crossprod(test_susine:::softmax_weights(ev), pm))
)
reference_tbl <- dplyr::bind_rows(lapply(names(ref_methods), function(nm) {
  pip_by_locus <- lapply(names(pip_list), function(lid) ref_methods[[nm]](pip_list[[lid]], elbo_list[[lid]]))
  names(pip_by_locus) <- names(pip_list)
  tibble::tibble(method = nm,
                 auprc_pooled = auprc_pooled(pip_by_locus),
                 auprc_mean_perlocus = auprc_mean_perlocus(pip_by_locus))
}))
cat("=== Reference (non-cluster) aggregations ===\n")
print(as.data.frame(reference_tbl), row.names = FALSE)
```

# 5. Metric x Threshold Sweep (Method B)

```{r sweep}
# Cache hclust per (metric, locus): distance depends on metric only, not threshold.
hc_cache <- list()
for (metric in metrics) {
  hc_cache[[metric]] <- lapply(names(pip_list), function(lid)
    test_susine:::prepare_pip_similarity_cache(pip_list[[lid]], metric = metric)$hc)
  names(hc_cache[[metric]]) <- names(pip_list)
}

sweep_rows <- list()
for (metric in metrics) {
  for (thr in threshold_grids[[metric]]) {
    agg_by_locus <- list(); n_clust <- numeric(0)
    for (lid in names(pip_list)) {
      hc <- hc_cache[[metric]][[lid]]
      agg_by_locus[[lid]] <- as.numeric(test_susine:::.cluster_weight_from_hc(
        hc, thr, elbo_list[[lid]], pip_list[[lid]]))
      n_clust <- c(n_clust, length(unique(stats::cutree(hc, h = thr))))
    }
    sweep_rows[[length(sweep_rows) + 1L]] <- tibble::tibble(
      metric = metric, threshold = thr,
      mean_n_clusters = mean(n_clust),
      auprc_pooled = auprc_pooled(agg_by_locus),
      auprc_mean_perlocus = auprc_mean_perlocus(agg_by_locus)
    )
  }
  cat("  swept", metric, "\n")
}
auprc_sweep <- dplyr::bind_rows(sweep_rows)
```

# 6. Results

```{r results}
best <- auprc_sweep %>%
  dplyr::group_by(.data$metric) %>%
  dplyr::slice_max(.data$auprc_pooled, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup()
cat("=== Best (metric, threshold) by pooled AUPRC ===\n")
print(as.data.frame(best), row.names = FALSE)

readr::write_csv(auprc_sweep, file.path(out_dir, "auprc_cluster_metric_sweep.csv"))
readr::write_csv(reference_tbl, file.path(out_dir, "auprc_reference_methods.csv"))

curve <- ggplot(auprc_sweep, aes(threshold, auprc_pooled, colour = metric)) +
  geom_line() + geom_point(size = 1) +
  geom_hline(data = reference_tbl, aes(yintercept = auprc_pooled, linetype = method),
             colour = "grey40") +
  facet_wrap(~ metric, scales = "free_x") +
  scale_x_log10() +
  labs(x = "clustering threshold (log scale)", y = "overall (pooled) AUPRC",
       title = "Method-B cluster-weight AUPRC vs threshold, by metric",
       subtitle = "grey lines: non-cluster reference aggregations") +
  theme_bw()
save_png(file.path(fig_dir, "auprc_vs_threshold_by_metric.png"), curve, width = 11, height = 4)

clust_curve <- ggplot(auprc_sweep, aes(mean_n_clusters, auprc_pooled, colour = metric)) +
  geom_path() + geom_point(size = 1) +
  labs(x = "mean # clusters per locus", y = "overall (pooled) AUPRC",
       title = "AUPRC vs ensemble fragmentation") + theme_bw()
save_png(file.path(fig_dir, "auprc_vs_nclusters.png"), clust_curve, width = 7, height = 5)

knitr::kable(auprc_sweep, digits = 4, caption = "Cluster-metric sweep (Method B)")
cat("\nWritten to:", out_dir, "and", fig_dir, "\n")
```
