---
title: "Ensemble Scaling Experiment — Collect Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Post-hoc collection: merge flush files, join metadata, compute AUPRC
from confusion bins, add oracle aggregation.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_ensemble_scaling.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
}
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 1. Locate Output

```{r locate}
job_name      <- "ensemble_scaling_full"
parent_job_id <- "50452623"
output_root   <- here("output")

# Job-level directory (contains task-001/, task-002/, etc.)
job_dir  <- file.path(output_root, "slurm_output", job_name, parent_job_id)

# Job config written by write_job_artifacts
cfg_path <- file.path(output_root, "run_history", job_name, parent_job_id, "job_config.json")
cfg <- test_susine:::load_job_config(cfg_path)

# Find all task directories
task_dirs <- list.dirs(job_dir, recursive = FALSE, full.names = TRUE)
task_dirs <- task_dirs[grepl("task-", basename(task_dirs))]
cat("Job dir:", job_dir, "\n")
cat("Task directories:", length(task_dirs), "\n")
cat(" ", paste(basename(task_dirs), collapse = ", "), "\n")
```

# 2. Merge Flush Files

```{r merge-flush}
log_stage <- function(msg) {
  cat(sprintf("[%s] %s\n", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), msg))
}

merge_flush_csv <- function(task_dirs,
                            pattern,
                            label,
                            col_select = NULL,
                            progress_every = 25L) {
  files <- sort(unlist(lapply(task_dirs, list.files, pattern = pattern,
                              full.names = TRUE)))
  if (!length(files)) return(tibble::tibble())

  log_stage(sprintf("%s: reading %d file(s)", label, length(files)))
  out <- vector("list", length(files))
  progress_every <- max(1L, as.integer(progress_every))

  for (i in seq_along(files)) {
    if (!is.null(col_select)) {
      out[[i]] <- readr::read_csv(
        files[[i]],
        show_col_types = FALSE,
        col_select = dplyr::any_of(col_select)
      )
    } else {
      out[[i]] <- readr::read_csv(
        files[[i]],
        show_col_types = FALSE
      )
    }
    if (i == 1L || i == length(files) || (i %% progress_every) == 0L) {
      cat(sprintf("  -> %s %d/%d\n", label, i, length(files)))
    }
  }

  dplyr::bind_rows(out)
}

confusion_flush_cols <- c(
  "run_id", "explore_method", "variant_id", "agg_method",
  "pip_threshold", "n_causal_at_bucket", "n_noncausal_at_bucket"
)

model_metrics <- merge_flush_csv(task_dirs, "model_metrics\\.csv$", "model_metrics")
confusion_bins <- merge_flush_csv(
  task_dirs,
  "confusion_bins\\.csv$",
  "confusion_bins",
  col_select = confusion_flush_cols
) %>%
  dplyr::select(dplyr::all_of(confusion_flush_cols))
dataset_metrics <- merge_flush_csv(task_dirs, "dataset_metrics\\.csv$", "dataset_metrics")
multimodal_metrics <- merge_flush_csv(task_dirs, "multimodal_metrics\\.csv$", "multimodal_metrics")
refine_depth <- merge_flush_csv(task_dirs, "refine_depth\\.csv$", "refine_depth")
effect_metrics <- merge_flush_csv(task_dirs, "effect_metrics\\.csv$", "effect_metrics")
tier_cs_metrics <- merge_flush_csv(task_dirs, "tier_cs_metrics\\.csv$", "tier_cs_metrics")
hg2_by_agg <- merge_flush_csv(task_dirs, "hg2_by_agg\\.csv$", "hg2_by_agg")
scaling_bins_raw <- merge_flush_csv(task_dirs, "scaling_bins\\.csv$", "scaling_bins")

cat("Model metrics:", nrow(model_metrics), "rows\n")
cat("Confusion bins:", nrow(confusion_bins), "rows\n")
cat("Dataset metrics:", nrow(dataset_metrics), "rows\n")
cat("Multimodal metrics:", nrow(multimodal_metrics), "rows\n")
cat("Refine depth:", nrow(refine_depth), "rows\n")
cat("Tier CS metrics:", nrow(tier_cs_metrics), "rows\n")
cat("hg2 by agg:", nrow(hg2_by_agg), "rows\n")
cat("Scaling bins:", nrow(scaling_bins_raw), "rows\n")
```

# 3. Join Run Table Metadata

The flush outputs only have `run_id`. Join with the run table for the
metadata the downstream summaries and plots actually use.

```{r join-metadata}
run_info <- cfg$tables$runs %>%
  dplyr::select(
    run_id, dataset_bundle_id, spec_name, use_case_id, prior_spec_id,
    exploration_methods, sigma_0_2_scalar, warm_method, c_value,
    refine_step, restart_id, annotation_r2, inflate_match, group_key
  )

confusion_run_info <- run_info %>%
  dplyr::select(
    run_id, dataset_bundle_id, spec_name, use_case_id,
    annotation_r2, group_key, c_value, sigma_0_2_scalar
  )

model_metrics <- model_metrics %>%
  dplyr::left_join(run_info, by = "run_id")

tier_cs_metrics <- tier_cs_metrics %>%
  dplyr::left_join(
    run_info %>% dplyr::select(run_id, dataset_bundle_id, spec_name,
                                use_case_id, group_key, annotation_r2),
    by = "run_id"
  )

cat("Model metrics with metadata:", nrow(model_metrics), "rows\n")
cat("Tier CS metrics with metadata:", nrow(tier_cs_metrics), "rows\n")
cat("Specs present:", paste(unique(na.omit(model_metrics$spec_name)), collapse = ", "), "\n")

# Enrich hg2_by_agg with spec_name via group_label -> group_key mapping
if (nrow(hg2_by_agg) > 0L) {
  if (!"spec_name" %in% names(hg2_by_agg) && "group_label" %in% names(hg2_by_agg)) {
    spec_map <- run_info %>%
      dplyr::distinct(.data$group_key, .data$spec_name)
    hg2_by_agg <- hg2_by_agg %>%
      dplyr::mutate(group_key = sub("^model_grid\\|", "", .data$group_label)) %>%
      dplyr::left_join(spec_map, by = "group_key") %>%
      dplyr::select(-group_key)
  }
  cat("hg2 by agg with metadata:", nrow(hg2_by_agg), "rows\n")
  cat("hg2 specs:", paste(unique(na.omit(hg2_by_agg$spec_name)), collapse = ", "), "\n")
  cat("hg2 agg methods:", paste(unique(na.omit(hg2_by_agg$agg_method)), collapse = ", "), "\n")
}
```

For confusion bins, aggregated rows share a `run_id` with the group's
representative run. Join only the fields the collect and visualize
workbooks use.

```{r join-confusion}
confusion_output_cols <- c(
  "run_id", "explore_method", "variant_id", "agg_method",
  "pip_threshold", "n_causal_at_bucket", "n_noncausal_at_bucket",
  "dataset_bundle_id", "spec_name", "use_case_id", "annotation_r2",
  "group_key", "c_value", "sigma_0_2_scalar"
)

# Individual fit rows (agg_method = NA) join directly
confusion_individual <- confusion_bins %>%
  dplyr::filter(is.na(agg_method)) %>%
  dplyr::left_join(confusion_run_info, by = "run_id") %>%
  dplyr::select(dplyr::all_of(confusion_output_cols))

# Aggregated rows: run_id is the group representative
confusion_agg <- confusion_bins %>%
  dplyr::filter(!is.na(agg_method)) %>%
  dplyr::left_join(confusion_run_info, by = "run_id") %>%
  dplyr::select(dplyr::all_of(confusion_output_cols))

cat("Confusion bins with metadata:", nrow(confusion_individual) + nrow(confusion_agg), "rows\n")
cat("  Individual:", nrow(confusion_individual), "\n")
cat("  Aggregated:", nrow(confusion_agg), "\n")
rm(confusion_bins)
invisible(gc())
```

# 4. Compute AUPRC from Confusion Bins

Uses pooled confusion-bin summaries plus the direct AP implementation in
`test_susine::compute_auprc_from_confusion()`.

```{r compute-auprc}
cat(
  "Confusion pooling backend:",
  if (requireNamespace("data.table", quietly = TRUE)) "data.table" else "dplyr fallback",
  "\n"
)

pool_confusion_summary <- function(bins,
                                   group_vars,
                                   label,
                                   chunk_var = NULL,
                                   chunk_size = 25L) {
  log_stage(sprintf("%s: pooling bins", label))

  if (is.null(chunk_var)) {
    preferred_chunk_vars <- c(
      "dataset_bundle_id", "spec_name", "use_case_id",
      "agg_method", "annotation_r2", "run_id"
    )
    available <- intersect(preferred_chunk_vars, intersect(group_vars, names(bins)))
    if (length(available)) {
      chunk_var <- available[[1]]
    } else if ("dataset_bundle_id" %in% names(bins)) {
      chunk_var <- "dataset_bundle_id"
    }
  }
  chunk_size <- max(1L, as.integer(chunk_size))

  if (is.null(chunk_var) || !chunk_var %in% names(bins) || !nrow(bins)) {
    pooled <- test_susine:::pool_confusion_bins(bins, group_vars)
    cat(label, "pooled rows:", nrow(pooled), "\n")
    return(pooled)
  }

  chunk_ids <- unique(bins[[chunk_var]])
  chunk_ids <- chunk_ids[order(is.na(chunk_ids), chunk_ids)]
  chunk_groups <- split(chunk_ids, ceiling(seq_along(chunk_ids) / chunk_size))
  partials <- vector("list", length(chunk_groups))

  for (i in seq_along(chunk_groups)) {
    ids <- chunk_groups[[i]]
    chunk_bins <- bins[bins[[chunk_var]] %in% ids, , drop = FALSE]
    partials[[i]] <- test_susine:::pool_confusion_bins(chunk_bins, group_vars)
    cat(sprintf("  -> %s chunk %d/%d (%d %s values)\n",
                label, i, length(chunk_groups), length(ids), chunk_var))
    rm(chunk_bins)
    invisible(gc())
  }

  combined_partials <- dplyr::bind_rows(partials)
  if (chunk_var %in% group_vars) {
    cat(sprintf("  -> %s chunk outputs are disjoint by %s; skipping final re-pool\n",
                label, chunk_var))
    pooled <- combined_partials
  } else {
    cat(sprintf("  -> %s binding partials complete; starting final re-pool\n", label))
    pooled <- test_susine:::pool_confusion_bins(
      combined_partials,
      group_vars
    )
    cat(sprintf("  -> %s final re-pool complete\n", label))
  }
  cat(label, "pooled rows:", nrow(pooled), "\n")
  pooled
}

compute_auprc_from_pooled <- function(pooled_bins,
                                      group_vars,
                                      label,
                                      progress = FALSE,
                                      progress_every = 1000L) {
  test_susine:::compute_auprc_from_pooled_confusion(
    pooled_bins,
    group_vars = group_vars,
    progress = progress,
    progress_every = progress_every,
    progress_label = label
  )
}

per_dataset_group_vars <- c(
  "dataset_bundle_id", "run_id", "spec_name", "use_case_id",
  "annotation_r2", "group_key", "explore_method", "agg_method", "variant_id"
)
pooled_agg_overall_vars <- c("spec_name", "use_case_id", "agg_method")
pooled_agg_by_r2_vars <- c("spec_name", "use_case_id", "agg_method", "annotation_r2")
pooled_individual_overall_vars <- c("spec_name", "use_case_id")
pooled_individual_by_r2_vars <- c("spec_name", "use_case_id", "annotation_r2")
pooled_curve_agg_vars <- c("spec_name", "agg_method")

# Per-dataset AUPRC (needed for oracle selection per dataset)
per_dataset_individual_bins <- pool_confusion_summary(
  confusion_individual,
  per_dataset_group_vars,
  "Per-dataset individual"
)
per_dataset_agg_bins <- pool_confusion_summary(
  confusion_agg,
  per_dataset_group_vars,
  "Per-dataset aggregated"
)
auprc_individual <- compute_auprc_from_pooled(
  per_dataset_individual_bins,
  group_vars = per_dataset_group_vars,
  label = "Per-dataset individual AUPRC",
  progress = TRUE,
  progress_every = 5000L
)
auprc_agg_per_ds <- compute_auprc_from_pooled(
  per_dataset_agg_bins,
  group_vars = per_dataset_group_vars,
  label = "Per-dataset aggregated AUPRC",
  progress = TRUE,
  progress_every = 5000L
)
auprc_per_dataset <- dplyr::bind_rows(auprc_individual, auprc_agg_per_ds)

cat("Per-dataset AUPRC:", nrow(auprc_per_dataset), "rows\n")
cat("  Individual fits:", nrow(auprc_individual), "\n")
cat("  Aggregated:", nrow(auprc_agg_per_ds), "\n")

# Pooled AUPRC: combine confusion bins across ALL datasets, then
# compute one AUPRC per (spec, agg_method). Used for overall barplot.
pooled_agg_bins_overall <- pool_confusion_summary(
  confusion_agg,
  pooled_agg_overall_vars,
  "Pooled aggregated overall"
)
auprc_pooled_agg <- compute_auprc_from_pooled(
  pooled_agg_bins_overall,
  group_vars = pooled_agg_overall_vars,
  label = "Pooled aggregated overall AUPRC"
)

# By-annotation-r2 pooled agg: one AUPRC per (spec, agg_method, annotation_r2).
# Used for per-r2 barplots; annotation-independent specs have annotation_r2=NA.
pooled_agg_bins_by_r2 <- pool_confusion_summary(
  confusion_agg,
  pooled_agg_by_r2_vars,
  "Pooled aggregated by r2"
)
auprc_pooled_agg_by_r2 <- compute_auprc_from_pooled(
  pooled_agg_bins_by_r2,
  group_vars = pooled_agg_by_r2_vars,
  label = "Pooled aggregated by-r2 AUPRC"
)

# Overall pooled individual (pooled across all r2 — for overall barplot truth-warm lines)
pooled_individual_bins_overall <- pool_confusion_summary(
  confusion_individual,
  pooled_individual_overall_vars,
  "Pooled individual overall"
)
auprc_pooled_individual_overall <- compute_auprc_from_pooled(
  pooled_individual_bins_overall,
  group_vars = pooled_individual_overall_vars,
  label = "Pooled individual overall AUPRC"
)

# By-r2 pooled individual (for per-r2 barplot truth-warm lines)
pooled_individual_bins_by_r2 <- pool_confusion_summary(
  confusion_individual,
  pooled_individual_by_r2_vars,
  "Pooled individual by r2"
)
auprc_pooled_individual <- compute_auprc_from_pooled(
  pooled_individual_bins_by_r2,
  group_vars = pooled_individual_by_r2_vars,
  label = "Pooled individual by-r2 AUPRC"
)

cat("Pooled aggregated (overall):", nrow(auprc_pooled_agg), "combos\n")
cat("Pooled aggregated (by r2):", nrow(auprc_pooled_agg_by_r2), "combos\n")
cat("Pooled individual (overall):", nrow(auprc_pooled_individual_overall), "combos\n")
cat("Pooled individual (by r2):", nrow(auprc_pooled_individual), "combos\n")

# Pooled TPR/FPR curves (for ROC-style plots)
pooled_curve_agg_bins <- pool_confusion_summary(
  confusion_agg,
  pooled_curve_agg_vars,
  "Pooled ROC aggregated"
)
curves_pooled_agg <- compute_power_fdr_curves(
  pooled_curve_agg_bins,
  group_vars = pooled_curve_agg_vars
)
invisible(gc())
```

# 4b. TPR @ FPR=0.05 from Confusion Bins

Uses the same pooling approach as AUPRC: bins combined across ALL datasets,
then scan thresholds to find TPR at the FPR = 0.05 crossing. No per-dataset
averaging.

```{r compute-tpr05}
compute_tpr05_from_pooled <- function(pooled_bins,
                                      group_vars,
                                      label,
                                      fpr_target = 0.05,
                                      progress = FALSE,
                                      progress_every = 1000L) {
  test_susine:::compute_tpr05_from_pooled_confusion(
    pooled_bins,
    group_vars = group_vars,
    fpr_target = fpr_target,
    progress = progress,
    progress_every = progress_every,
    progress_label = label
  )
}

tpr05_pooled_agg <- compute_tpr05_from_pooled(
  pooled_agg_bins_overall,
  group_vars = pooled_agg_overall_vars,
  label = "Pooled aggregated overall TPR@0.05"
)

tpr05_pooled_agg_by_r2 <- compute_tpr05_from_pooled(
  pooled_agg_bins_by_r2,
  group_vars = pooled_agg_by_r2_vars,
  label = "Pooled aggregated by-r2 TPR@0.05"
)

tpr05_pooled_individual_overall <- compute_tpr05_from_pooled(
  pooled_individual_bins_overall,
  group_vars = pooled_individual_overall_vars,
  label = "Pooled individual overall TPR@0.05"
)

tpr05_pooled_individual <- compute_tpr05_from_pooled(
  pooled_individual_bins_by_r2,
  group_vars = pooled_individual_by_r2_vars,
  label = "Pooled individual by-r2 TPR@0.05"
)

cat("TPR@0.05 pooled aggregated (overall):", nrow(tpr05_pooled_agg), "combos\n")
cat("TPR@0.05 pooled aggregated (by r2):", nrow(tpr05_pooled_agg_by_r2), "combos\n")
cat("TPR@0.05 pooled individual (overall):", nrow(tpr05_pooled_individual_overall), "combos\n")
cat("TPR@0.05 pooled individual (by r2):", nrow(tpr05_pooled_individual), "combos\n")
```

# 5. Oracle Aggregation

For each ensemble group, select the single best individual fit by AUPRC
(truth-aware upper bound on aggregation performance).

```{r oracle}
# Per-dataset oracle: best individual run per (dataset, spec, group_key)
oracle_run_ids <- auprc_individual %>%
  dplyr::filter(!is.na(spec_name), !spec_name %in% c("baseline-single", "truth-warm")) %>%
  dplyr::group_by(dataset_bundle_id, spec_name, use_case_id,
                   annotation_r2, group_key) %>%
  dplyr::slice_max(order_by = AUPRC, n = 1, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::select(run_id, dataset_bundle_id, spec_name, use_case_id,
                annotation_r2, group_key, AUPRC)

cat("Oracle rows (per-dataset):", nrow(oracle_run_ids), "\n")

# Per-dataset oracle AUPRC table (for auprc_combined)
oracle_per_ds <- oracle_run_ids %>%
  dplyr::mutate(agg_method = "oracle", explore_method = "aggregation")

# Combine all per-dataset AUPRC results
auprc_combined <- dplyr::bind_rows(auprc_agg_per_ds, oracle_per_ds)

# Pooled oracle AUPRC: pull confusion bins for the oracle-selected runs,
# pool across all datasets, then compute one AUPRC per spec.
oracle_confusion <- confusion_individual %>%
  dplyr::semi_join(oracle_run_ids, by = c("run_id", "spec_name", "use_case_id"))

pooled_oracle_bins_overall <- pool_confusion_summary(
  oracle_confusion,
  pooled_individual_overall_vars,
  "Pooled oracle overall"
)
pooled_oracle_bins_by_r2 <- pool_confusion_summary(
  oracle_confusion,
  pooled_individual_by_r2_vars,
  "Pooled oracle by r2"
)

auprc_pooled_oracle <- compute_auprc_from_pooled(
  pooled_oracle_bins_overall,
  group_vars = pooled_individual_overall_vars,
  label = "Pooled oracle overall AUPRC"
) %>%
  dplyr::mutate(agg_method = "oracle")

auprc_pooled_oracle_by_r2 <- compute_auprc_from_pooled(
  pooled_oracle_bins_by_r2,
  group_vars = pooled_individual_by_r2_vars,
  label = "Pooled oracle by-r2 AUPRC"
) %>%
  dplyr::mutate(agg_method = "oracle")

cat("Pooled oracle rows (overall):", nrow(auprc_pooled_oracle), "\n")
cat("Pooled oracle rows (by r2):", nrow(auprc_pooled_oracle_by_r2), "\n")

# Add oracle to both pooled aggregated tables
auprc_pooled_agg <- dplyr::bind_rows(auprc_pooled_agg, auprc_pooled_oracle)
auprc_pooled_agg_by_r2 <- dplyr::bind_rows(auprc_pooled_agg_by_r2, auprc_pooled_oracle_by_r2)

# Oracle TPR@0.05 (uses same oracle_confusion bins)
tpr05_pooled_oracle <- compute_tpr05_from_pooled(
  pooled_oracle_bins_overall,
  group_vars = pooled_individual_overall_vars,
  label = "Pooled oracle overall TPR@0.05"
) %>%
  dplyr::mutate(agg_method = "oracle")

tpr05_pooled_oracle_by_r2 <- compute_tpr05_from_pooled(
  pooled_oracle_bins_by_r2,
  group_vars = pooled_individual_by_r2_vars,
  label = "Pooled oracle by-r2 TPR@0.05"
) %>%
  dplyr::mutate(agg_method = "oracle")

tpr05_pooled_agg <- dplyr::bind_rows(tpr05_pooled_agg, tpr05_pooled_oracle)
tpr05_pooled_agg_by_r2 <- dplyr::bind_rows(tpr05_pooled_agg_by_r2, tpr05_pooled_oracle_by_r2)

cat("TPR@0.05 pooled oracle rows (overall):", nrow(tpr05_pooled_oracle), "\n")
cat("TPR@0.05 pooled oracle rows (by r2):", nrow(tpr05_pooled_oracle_by_r2), "\n")
```

# 6. Summary Tables

```{r summary-by-spec}
cat("=== AUPRC by spec x aggregation method ===\n")
auprc_combined %>%
  dplyr::group_by(spec_name, agg_method) %>%
  dplyr::summarise(
    mean_AUPRC = mean(AUPRC, na.rm = TRUE),
    sd_AUPRC = sd(AUPRC, na.rm = TRUE),
    n = dplyr::n(),
    .groups = "drop"
  ) %>%
  dplyr::arrange(spec_name, agg_method) %>%
  knitr::kable(digits = 4, caption = "Aggregated AUPRC by spec and method")
```

```{r summary-baselines}
cat("\n=== Baseline and truth-warm AUPRC ===\n")
auprc_pooled_individual %>%
  dplyr::filter(spec_name %in% c("baseline-single", "truth-warm")) %>%
  knitr::kable(digits = 4, caption = "Baseline and truth-warm pooled AUPRC")
```

# 6b. Subscaling Analysis

```{r load-scaling-bins}
# Scaling confusion bins were merged from flush files in Section 2.
# Report summary; set to NULL if empty.
if (nrow(scaling_bins_raw) > 0L) {
  cat("Scaling bins rows:", nrow(scaling_bins_raw), "\n")
  cat("Specs present:", paste(sort(unique(na.omit(scaling_bins_raw$spec_name))), collapse = ", "), "\n")
} else {
  scaling_bins_raw <- NULL
  message("No scaling_bins flush files found — subscaling analysis skipped.",
          "\nRe-run with write_scaling_confusion_bins = TRUE in the job config.")
}
```

```{r pool-scaling-bins}
# Pool confusion bins across datasets for each (spec_name, n_ensemble, agg_method, rep).
# Downstream AUPRC / TPR computation operates on these pooled bins.
if (!is.null(scaling_bins_raw) && nrow(scaling_bins_raw) > 0L) {
  scaling_bins_pooled <- scaling_bins_raw %>%
    dplyr::group_by(
      spec_name, annotation_r2, n_ensemble, resolution, agg_method, rep, pip_threshold
    ) %>%
    dplyr::summarise(
      n_causal_at_bucket    = sum(n_causal_at_bucket,    na.rm = TRUE),
      n_noncausal_at_bucket = sum(n_noncausal_at_bucket, na.rm = TRUE),
      n_datasets            = dplyr::n_distinct(dataset_bundle_id),
      .groups = "drop"
    )
  cat("Pooled scaling bins rows:", nrow(scaling_bins_pooled), "\n")
} else {
  scaling_bins_pooled <- NULL
}
```

# 7. Export Consolidated Outputs

```{r export-scaling}
# scaling_bins_pooled is written to consolidated in the export chunk below.
```

```{r export}
out_dir <- file.path(job_dir, "consolidated")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(auprc_combined, file.path(out_dir, "auprc_aggregated.csv"))
readr::write_csv(auprc_individual, file.path(out_dir, "auprc_individual.csv"))
readr::write_csv(auprc_pooled_agg, file.path(out_dir, "auprc_pooled_aggregated.csv"))
readr::write_csv(auprc_pooled_agg_by_r2, file.path(out_dir, "auprc_pooled_agg_by_r2.csv"))
readr::write_csv(auprc_pooled_individual_overall, file.path(out_dir, "auprc_pooled_individual_overall.csv"))
readr::write_csv(auprc_pooled_individual, file.path(out_dir, "auprc_pooled_individual.csv"))
readr::write_csv(model_metrics, file.path(out_dir, "model_metrics_full.csv"))
confusion_bins_full <- dplyr::bind_rows(confusion_individual, confusion_agg) %>%
  dplyr::select(dplyr::all_of(confusion_output_cols))
readr::write_csv(confusion_bins_full, file.path(out_dir, "confusion_bins_full.csv"))
if (nrow(dataset_metrics)) {
  readr::write_csv(dataset_metrics, file.path(out_dir, "dataset_metrics.csv"))
}
if (nrow(multimodal_metrics)) {
  readr::write_csv(multimodal_metrics, file.path(out_dir, "multimodal_metrics.csv"))
}
if (nrow(refine_depth)) {
  readr::write_csv(refine_depth, file.path(out_dir, "refine_depth.csv"))
}
if (nrow(tier_cs_metrics)) {
  readr::write_csv(tier_cs_metrics, file.path(out_dir, "tier_cs_metrics_full.csv"))
}
if (nrow(hg2_by_agg)) {
  readr::write_csv(hg2_by_agg, file.path(out_dir, "hg2_by_agg.csv"))
}
if (!is.null(scaling_bins_pooled) && nrow(scaling_bins_pooled) > 0L) {
  readr::write_csv(scaling_bins_pooled,
                   file.path(out_dir, "scaling_bins_pooled.csv"))
}

readr::write_csv(tpr05_pooled_agg, file.path(out_dir, "tpr05_pooled_aggregated.csv"))
readr::write_csv(tpr05_pooled_agg_by_r2, file.path(out_dir, "tpr05_pooled_agg_by_r2.csv"))
readr::write_csv(tpr05_pooled_individual_overall, file.path(out_dir, "tpr05_pooled_individual_overall.csv"))
readr::write_csv(tpr05_pooled_individual, file.path(out_dir, "tpr05_pooled_individual.csv"))
rm(confusion_bins_full)
invisible(gc())

cat("Consolidated outputs written to:", out_dir, "\n")
cat("Files:", paste(list.files(out_dir), collapse = ", "), "\n")
```

# 8. Runtime Summary

```{r runtime}
model_metrics %>%
  dplyr::filter(filtering == "purity_filtered", !is.na(spec_name)) %>%
  dplyr::group_by(spec_name) %>%
  dplyr::summarise(
    mean_sec = mean(wall_time_sec, na.rm = TRUE),
    median_sec = median(wall_time_sec, na.rm = TRUE),
    total_min = sum(wall_time_sec, na.rm = TRUE) / 60,
    n_runs = dplyr::n(),
    .groups = "drop"
  ) %>%
  knitr::kable(digits = 2, caption = "Runtime by spec")
```
