---
title: "Purity-Filter Minibatch - Collect Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
params:
  run_collection: false
editor_options:
  markdown:
    wrap: 72
---

Post-run collection for the purity-filter minibatch. Merges the staged
confusion-bin flush files (tagged by `filter_type`) and the per-effect diagnostic
records, then:

1. computes one pooled AUPRC per filter type (and per-dataset deltas vs the
   unfiltered PIP), the headline "AUPRC with vs without purity filtering" number;
2. builds the old-vs-new credible-set diagnostic figures that contrast the
   classical metrics (purity, size) with the effect-level metrics (core-95%
   $k_{\mathrm{eff}}$, accuracy ratio).

By default, this workbook is visualization-only and reads the consolidated CSVs
already written under the job output directory. Set `params$run_collection =
TRUE` only when the staging outputs need to be re-aggregated.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_purity_filter_minibatch.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(knitr)
```

# 1. Locate Output

```{r locate}
job_name <- "purity_filter_minibatch"
parent_job_id <- "53617896"
output_root <- here("output")
run_collection <- if (exists("params") && "run_collection" %in% names(params)) {
  isTRUE(params$run_collection)
} else {
  FALSE
}

job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
cfg_path <- file.path(output_root, "run_history", job_name, parent_job_id, "job_config.json")
cfg <- if (file.exists(cfg_path)) test_susine:::load_job_config(cfg_path) else NULL

aggregated_dir <- file.path(job_dir, "aggregated")
consolidated_dir <- file.path(job_dir, "consolidated")
plot_dir <- file.path(consolidated_dir, "plots")
dir.create(consolidated_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

task_dirs <- list.dirs(job_dir, recursive = FALSE, full.names = TRUE)
staging_flush_files <- list.files(
  task_dirs,
  pattern = "^flush-[0-9]+_",
  full.names = TRUE
)
summary_cache_files <- c(
  "auprc_by_filter_type.csv",
  "auprc_per_dataset_by_filter.csv",
  "auprc_per_dataset_delta_summary.csv",
  "effect_metrics_unfiltered_full.csv"
)
raw_cache_files <- c("confusion_bins.csv", "effect_metrics_unfiltered.csv")
consolidated_summary_cache_available <- all(file.exists(file.path(consolidated_dir, summary_cache_files)))
aggregated_summary_cache_available <- all(file.exists(file.path(aggregated_dir, summary_cache_files)))
consolidated_raw_cache_available <- all(file.exists(file.path(consolidated_dir, raw_cache_files)))
aggregated_raw_cache_available <- all(file.exists(file.path(aggregated_dir, raw_cache_files)))
summary_cache_available <- consolidated_summary_cache_available || aggregated_summary_cache_available
raw_cache_available <- consolidated_raw_cache_available || aggregated_raw_cache_available
cache_dir <- if (consolidated_summary_cache_available || consolidated_raw_cache_available) {
  consolidated_dir
} else if (aggregated_summary_cache_available || aggregated_raw_cache_available) {
  aggregated_dir
} else {
  consolidated_dir
}
use_consolidated_cache <- !run_collection

if (use_consolidated_cache && !summary_cache_available && !raw_cache_available) {
  stop(
    "Visualization-only mode requires either summary CSVs or raw aggregate CSVs in ",
    consolidated_dir,
    " or ",
    aggregated_dir,
    ". Re-knit with params$run_collection = TRUE to rebuild them from staging files."
  )
}

cat("Job dir:", job_dir, "\n")
cat("Run collection:", run_collection, "\n")
cat("Using consolidated cache only:", use_consolidated_cache, "\n")
cat("Reading cached tables from:", cache_dir, "\n")
```

# 2. Aggregate Staging Outputs

```{r aggregate}
if (!isTRUE(run_collection)) {
  message("Skipping staging aggregation; using cached recovery files.")
} else {
  if (!length(staging_flush_files)) {
    stop(
      "No staging flush files found and cached recovery files are incomplete: ",
      job_dir
    )
  }

  agg <- test_susine::aggregate_staging_outputs(
    job_name = job_name,
    parent_job_id = parent_job_id,
    output_root = output_root,
    validate = TRUE,
    output_dir = consolidated_dir,
    write_snps_dataset = FALSE
  )
}
```

# 3. Load Aggregated Tables

```{r load-aggregated}
read_optional_csv <- function(path) {
  if (!file.exists(path)) return(NULL)
  readr::read_csv(path, show_col_types = FALSE)
}

read_first_csv <- function(paths) {
  paths <- unique(paths)
  existing_path <- paths[file.exists(paths)][1]
  if (is.na(existing_path)) return(NULL)
  read_optional_csv(existing_path)
}

confusion_bins <- read_first_csv(file.path(
  c(consolidated_dir, aggregated_dir),
  "confusion_bins.csv"
))
effect_unfiltered <- read_first_csv(c(
  file.path(c(consolidated_dir, aggregated_dir), "effect_metrics_unfiltered.csv"),
  file.path(c(cache_dir, consolidated_dir, aggregated_dir), "effect_metrics_unfiltered_full.csv")
))

auprc_pooled_cached <- read_first_csv(file.path(
  c(cache_dir, consolidated_dir, aggregated_dir),
  "auprc_by_filter_type.csv"
))
auprc_per_dataset_cached <- read_first_csv(file.path(
  c(cache_dir, consolidated_dir, aggregated_dir),
  "auprc_per_dataset_by_filter.csv"
))
per_dataset_summary_cached <- read_first_csv(file.path(
  c(cache_dir, consolidated_dir, aggregated_dir),
  "auprc_per_dataset_delta_summary.csv"
))

filter_labels <- c(
  no_filter           = "No filter",
  purity_95           = "Purity ≥ 0.95",
  purity_90           = "Purity ≥ 0.90",
  purity_50           = "Purity ≥ 0.50",
  keff_core95_le_100  = "Core-95% effective support ≤ 100"
)
filter_level_template <- c("no_filter", "purity_95", "purity_90", "purity_50",
                           "keff_core95_le_100")
observed_filter_types <- unique(c(
  if (!is.null(confusion_bins) && "filter_type" %in% names(confusion_bins)) confusion_bins$filter_type else NULL,
  if (!is.null(auprc_pooled_cached) && "filter_type" %in% names(auprc_pooled_cached)) auprc_pooled_cached$filter_type else NULL,
  if (!is.null(auprc_per_dataset_cached) && "filter_type" %in% names(auprc_per_dataset_cached)) auprc_per_dataset_cached$filter_type else NULL
))
filter_levels <- intersect(filter_level_template, observed_filter_types)

if (!use_consolidated_cache || is.null(auprc_pooled_cached)) {
  stopifnot(!is.null(confusion_bins), "filter_type" %in% names(confusion_bins))
}

cat("Confusion-bin rows:", if (is.null(confusion_bins)) 0 else nrow(confusion_bins), "\n")
cat("Filter types:", paste(sort(filter_levels), collapse = ", "), "\n")
cat("Per-effect rows:", if (is.null(effect_unfiltered)) 0 else nrow(effect_unfiltered), "\n")
```

# 4. AUPRC by Filter Type (headline)

```{r auprc-by-filter}
# Pooled AUPRC across all 600 datasets, one value per filter type.
if (!is.null(auprc_pooled_cached)) {
  auprc_pooled <- auprc_pooled_cached
} else {
  auprc_pooled <- test_susine::compute_auprc_from_confusion(
    confusion_bins,
    group_vars = "filter_type"
  )
}

if (!is.null(auprc_per_dataset_cached)) {
  auprc_per_dataset <- auprc_per_dataset_cached
} else {
  auprc_per_dataset <- test_susine::compute_auprc_from_confusion(
    confusion_bins,
    group_vars = c("filter_type", "dataset_bundle_id")
  )
}

per_dataset_summary <- per_dataset_summary_cached

auprc_pooled <- auprc_pooled %>%
  dplyr::mutate(filter_type = factor(as.character(.data$filter_type), levels = filter_levels)) %>%
  dplyr::arrange(.data$filter_type)

auprc_no_filter <- auprc_pooled$AUPRC[as.character(auprc_pooled$filter_type) == "no_filter"]
if (!"delta_vs_unfiltered" %in% names(auprc_pooled)) {
  auprc_pooled <- auprc_pooled %>%
    dplyr::mutate(delta_vs_unfiltered = .data$AUPRC - auprc_no_filter)
}
auprc_pooled <- auprc_pooled %>%
  dplyr::mutate(
    filter_label = dplyr::coalesce(
      unname(filter_labels[as.character(.data$filter_type)]),
      as.character(.data$filter_type)
    ),
    filter_label_plot = gsub(
      " effective support ",
      " effective\nsupport ",
      .data$filter_label,
      fixed = TRUE
    )
  )


compute_pip_mass_ratio_from_bins <- function(confusion_df, group_vars = "filter_type") {
  required_cols <- c(group_vars, "pip_threshold", "n_causal_at_bucket", "n_noncausal_at_bucket")
  if (is.null(confusion_df) || !all(required_cols %in% names(confusion_df))) {
    return(tibble::tibble())
  }

  confusion_df %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::summarise(
      causal_pip_mass = sum(.data$pip_threshold * .data$n_causal_at_bucket, na.rm = TRUE),
      noncausal_pip_mass = sum(.data$pip_threshold * .data$n_noncausal_at_bucket, na.rm = TRUE),
      pip_mass_ratio = dplyr::if_else(
        .data$noncausal_pip_mass > 0,
        .data$causal_pip_mass / .data$noncausal_pip_mass,
        NA_real_
      ),
      .groups = "drop"
    )
}

pip_mass_summary <- compute_pip_mass_ratio_from_bins(confusion_bins, group_vars = "filter_type")
if (nrow(pip_mass_summary)) {
  auprc_pooled <- auprc_pooled %>%
    dplyr::select(-dplyr::any_of(c("causal_pip_mass", "noncausal_pip_mass", "pip_mass_ratio"))) %>%
    dplyr::left_join(pip_mass_summary, by = "filter_type")
} else {
  auprc_pooled <- auprc_pooled %>%
    dplyr::mutate(
      causal_pip_mass = NA_real_,
      noncausal_pip_mass = NA_real_,
      pip_mass_ratio = NA_real_
    )
}

kable(
  auprc_pooled %>%
    dplyr::select(.data$filter_label, .data$AUPRC, .data$delta_vs_unfiltered, .data$pip_mass_ratio),
  digits = 4,
  col.names = c("Filter", "Pooled AUPRC", "Δ vs unfiltered", "Causal/noncausal PIP mass"),
  caption = "Pooled causal-recovery AUPRC (top-8 positives) by credible-set filter, 600 datasets"
)

# Per-dataset AUPRC -> paired delta vs no_filter (sign test on the filtering cost).
if (is.null(per_dataset_summary)) {
  per_dataset_delta <- auprc_per_dataset %>%
    dplyr::select(.data$filter_type, .data$dataset_bundle_id, .data$AUPRC) %>%
    tidyr::pivot_wider(names_from = .data$filter_type, values_from = .data$AUPRC) %>%
    tidyr::pivot_longer(
      cols = dplyr::any_of(setdiff(filter_levels, "no_filter")),
      names_to = "filter_type", values_to = "AUPRC_filtered"
    ) %>%
    dplyr::mutate(delta = .data$AUPRC_filtered - .data$no_filter)

  per_dataset_summary <- per_dataset_delta %>%
    dplyr::filter(is.finite(.data$delta)) %>%
    dplyr::group_by(.data$filter_type) %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean_delta = mean(.data$delta),
      median_delta = stats::median(.data$delta),
      frac_hurt = mean(.data$delta < 0),
      frac_help = mean(.data$delta > 0),
      .groups = "drop"
    )
}

if (!is.null(per_dataset_summary)) {
  per_dataset_summary <- per_dataset_summary %>%
    dplyr::mutate(filter_type = factor(as.character(.data$filter_type), levels = filter_levels)) %>%
    dplyr::arrange(.data$filter_type)
}

kable(per_dataset_summary, digits = 4,
      caption = "Per-dataset paired AUPRC change vs unfiltered (negative = filter hurts)")
```

```{r auprc-plot}
p_auprc <- ggplot2::ggplot(
  auprc_pooled,
  ggplot2::aes(x = stats::reorder(.data$filter_label_plot, -.data$AUPRC), y = .data$AUPRC)
) +
  ggplot2::geom_col(fill = "#1F77B4", width = 0.7) +
  ggplot2::geom_hline(yintercept = auprc_no_filter, linetype = "dashed", color = "grey30") +
  ggplot2::geom_text(
    ggplot2::aes(label = sprintf("%.3f\nmass r=%.2f", .data$AUPRC, .data$pip_mass_ratio)),
    vjust = -0.25,
    size = 5.6,
    lineheight = 0.9
  ) +
  ggplot2::coord_cartesian(
    ylim = c(NA_real_, max(auprc_pooled$AUPRC, na.rm = TRUE) + 0.04)
  ) +
  ggplot2::labs(
    x = NULL, y = "Pooled AUPRC",
    title = "Credible-set filtering changes per-variant AUPRC",
    subtitle = "Dashed line: unfiltered PIP. Lower bars = the filter discarded useful signal."
  ) +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 20, hjust = 1))

ggplot2::ggsave(file.path(plot_dir, "auprc_by_filter_type.png"),
                p_auprc, width = 8, height = 5, dpi = 300)
p_auprc
```

```{r pr-curves-by-filter, eval = exists("confusion_bins") && !is.null(confusion_bins)}
pr_curves_by_filter <- confusion_bins %>%
  dplyr::group_by(.data$filter_type, .data$pip_threshold) %>%
  dplyr::summarise(
    n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
    n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(.data$filter_type) %>%
  dplyr::arrange(dplyr::desc(.data$pip_threshold), .by_group = TRUE) %>%
  dplyr::mutate(
    TP = cumsum(.data$n_causal_at_bucket),
    FP = cumsum(.data$n_noncausal_at_bucket),
    total_causal = sum(.data$n_causal_at_bucket, na.rm = TRUE),
    precision = dplyr::if_else(.data$TP + .data$FP > 0,
                               .data$TP / (.data$TP + .data$FP), NA_real_),
    recall = dplyr::if_else(.data$total_causal > 0,
                            .data$TP / .data$total_causal, NA_real_),
    filter_label = dplyr::coalesce(
      unname(filter_labels[as.character(.data$filter_type)]),
      as.character(.data$filter_type)
    )
  ) %>%
  dplyr::ungroup() %>%
  dplyr::filter(is.finite(.data$precision), is.finite(.data$recall)) %>%
  dplyr::mutate(
    filter_type = factor(as.character(.data$filter_type), levels = filter_levels),
    filter_label = factor(.data$filter_label, levels = unname(filter_labels[filter_levels]))
  )

pr_plot_filter_levels <- setdiff(filter_levels, "purity_90")
pr_curves_by_filter <- pr_curves_by_filter %>%
  dplyr::filter(
    as.character(.data$filter_type) %in% pr_plot_filter_levels,
    .data$pip_threshold > 0,
    .data$recall > 0
  ) %>%
  dplyr::mutate(
    filter_type = factor(as.character(.data$filter_type), levels = pr_plot_filter_levels),
    filter_label = factor(
      as.character(.data$filter_label),
      levels = unname(filter_labels[pr_plot_filter_levels])
    )
  )

p_pr_curves <- ggplot2::ggplot(
  pr_curves_by_filter,
  ggplot2::aes(x = .data$recall, y = .data$precision, color = .data$filter_label)
) +
  ggplot2::geom_line(linewidth = 1.1, alpha = 0.9) +
  ggplot2::geom_hline(yintercept = 0.5, linetype = "dotted", color = "grey45") +
  ggplot2::coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  ggplot2::labs(
    x = "Recall",
    y = "Precision",
    color = "Filter",
    title = "Pooled precision-recall curves by credible-set filter",
    subtitle = "Dotted line: 50% precision. Curves are pooled across all simulation datasets."
  ) +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::guides(color = ggplot2::guide_legend(nrow = 2, byrow = TRUE)) +
  ggplot2::theme(legend.position = "bottom")

ggplot2::ggsave(file.path(plot_dir, "pr_curves_by_filter_type.png"),
                p_pr_curves, width = 8, height = 5.5, dpi = 300)
p_pr_curves
```

# 5. Old vs New Credible-Set Diagnostics

The per-effect record contrasts the classical metrics (purity, size) with the
effect-level metrics introduced in the paper (core-95% $k_{\mathrm{eff}}$,
accuracy ratio). `coverage = 1` marks effects whose credible set actually
contains a true causal variant.

```{r effect-diagnostics, eval = exists("effect_unfiltered") && !is.null(effect_unfiltered)}
eff <- effect_unfiltered %>%
  dplyr::filter(.data$size > 0) %>%                 # active effects only
  dplyr::mutate(
    coverage_lab = factor(
      ifelse(.data$coverage > 0, "Contains causal variant", "No causal variant"),
      levels = c("Contains causal variant", "No causal variant")
    ),
    purity = as.numeric(.data$purity),
    accuracy_ratio = as.numeric(.data$accuracy_ratio),
    k_eff_core95 = as.numeric(.data$effect_k_eff_signal_core95),
    size = as.numeric(.data$size)
  )

cov_palette <- c("Contains causal variant" = "#0072B2", "No causal variant" = "#D55E00")
purity_cutoff <- 0.50
keff_filter_cutoff <- 100
scatter_point_size <- 3.2

# Classical CS diagnostics: purity, size, and binary 95% coverage.
p_old_cs <- ggplot2::ggplot(
  eff %>% dplyr::filter(is.finite(.data$purity), is.finite(.data$size), .data$size > 0),
  ggplot2::aes(x = .data$purity, y = .data$size, color = .data$coverage_lab)
) +
  ggplot2::geom_point(alpha = 0.45, size = scatter_point_size) +
  ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
  ggplot2::scale_y_log10() +
  ggplot2::scale_color_manual(values = cov_palette, name = "95% CS") +
  ggplot2::labs(
    x = "Credible-set purity (minimum pairwise |r|)",
    y = "95% credible-set size",
    title = "Classical credible-set summaries",
    subtitle = "Dashed line: 0.50 purity threshold."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")

# (A) Purity vs accuracy: are low-purity sets actually inaccurate?
p_purity_acc <- ggplot2::ggplot(
  eff %>% dplyr::filter(is.finite(.data$purity), is.finite(.data$accuracy_ratio)),
  ggplot2::aes(x = .data$purity, y = .data$accuracy_ratio, color = .data$coverage_lab)
) +
  ggplot2::geom_point(alpha = 0.45, size = scatter_point_size) +
  ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
  ggplot2::scale_color_manual(values = cov_palette, name = "Credible set") +
  ggplot2::labs(
    x = "Credible-set purity (minimum pairwise |r|)",
    y = "Accuracy ratio",
    title = "Purity vs accuracy",
    subtitle = "Dashed line: 0.50 purity threshold."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")

# (B) Core-95 k_eff vs accuracy: diffuse effects can still point to causals.
p_keff_acc <- ggplot2::ggplot(
  eff %>%
    dplyr::filter(is.finite(.data$k_eff_core95), is.finite(.data$accuracy_ratio)) %>%
    dplyr::arrange(.data$accuracy_ratio),
  ggplot2::aes(x = .data$k_eff_core95, y = .data$accuracy_ratio, color = .data$accuracy_ratio)
) +
  ggplot2::geom_point(alpha = 0.65, size = scatter_point_size) +
  ggplot2::geom_vline(xintercept = keff_filter_cutoff, linetype = "longdash", color = "grey35") +
  ggplot2::geom_hline(yintercept = 0.9, linetype = "dotted", color = "grey45") +
  ggplot2::scale_x_log10() +
  ggplot2::scale_color_gradientn(
    colors = c("#D55E00", "#F0E442", "#0072B2"),
    name = "Accuracy ratio",
    guide = ggplot2::guide_colorbar(
      title.position = "top",
      barwidth = grid::unit(4.0, "in"),
      barheight = grid::unit(0.18, "in")
    )
  ) +
  ggplot2::labs(
    x = expression("Core-95% effective support (" * k[eff] * ")"),
    y = "Accuracy ratio",
    title = "Effective support vs accuracy",
    subtitle = "Long dash: k_eff = 100 filter. Dotted line: A = 0.9."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")

purity_keff_plot_data <- eff %>%
  dplyr::filter(is.finite(.data$purity), is.finite(.data$k_eff_core95))

purity_side_summary <- purity_keff_plot_data %>%
  dplyr::mutate(
    purity_side = dplyr::if_else(.data$purity < purity_cutoff, "left", "right")
  ) %>%
  dplyr::group_by(.data$purity_side) %>%
  dplyr::summarise(
    n_effects = dplyr::n(),
    pct_no_causal = 100 * mean(.data$coverage <= 0),
    .groups = "drop"
  ) %>%
  dplyr::right_join(
    data.frame(purity_side = c("left", "right")),
    by = "purity_side"
  ) %>%
  dplyr::mutate(
    n_effects = dplyr::coalesce(.data$n_effects, 0L),
    pct_no_causal = dplyr::coalesce(.data$pct_no_causal, 0),
    x = dplyr::if_else(.data$purity_side == "left", purity_cutoff - 0.02, purity_cutoff + 0.02),
    hjust = dplyr::if_else(.data$purity_side == "left", 1, 0),
    label = sprintf(
      "%s\nn = %s\nNo causal: %.1f%%",
      dplyr::if_else(.data$purity_side == "left", "Purity < 0.50", "Purity >= 0.50"),
      format(.data$n_effects, big.mark = ",", scientific = FALSE),
      .data$pct_no_causal
    )
  )

# (C) Purity vs core-95 k_eff: purity is only weakly tied to diffuseness.
p_purity_keff <- ggplot2::ggplot(
  purity_keff_plot_data,
  ggplot2::aes(x = .data$purity, y = .data$k_eff_core95, color = .data$coverage_lab)
) +
  ggplot2::geom_point(alpha = 0.45, size = scatter_point_size) +
  ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
  ggplot2::geom_hline(yintercept = keff_filter_cutoff, linetype = "longdash", color = "grey35") +
  ggplot2::geom_label(
    data = purity_side_summary,
    ggplot2::aes(x = .data$x, y = Inf, label = .data$label, hjust = .data$hjust),
    inherit.aes = FALSE,
    vjust = 1.15,
    size = 5.4,
    label.size = 0,
    fill = "white",
    alpha = 0.85
  ) +
  ggplot2::scale_y_log10() +
  ggplot2::scale_color_manual(values = cov_palette, name = "Credible set") +
  ggplot2::labs(
    x = "Credible-set purity (minimum pairwise |r|)",
    y = expression("Core-95% effective support (" * k[eff] * ")"),
    title = "Purity vs diffuseness",
    subtitle = "Weak association: low-purity sets are not reliably more diffuse."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")

p_purity_keff_accuracy <- ggplot2::ggplot(
  purity_keff_plot_data %>%
    dplyr::filter(is.finite(.data$accuracy_ratio)) %>%
    dplyr::arrange(.data$accuracy_ratio),
  ggplot2::aes(x = .data$purity, y = .data$k_eff_core95, color = .data$accuracy_ratio)
) +
  ggplot2::geom_point(alpha = 0.65, size = scatter_point_size) +
  ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
  ggplot2::geom_hline(yintercept = keff_filter_cutoff, linetype = "longdash", color = "grey35") +
  ggplot2::scale_y_log10() +
  ggplot2::scale_color_gradientn(
    colors = c("#D55E00", "#F0E442", "#0072B2"),
    name = "Accuracy ratio",
    guide = ggplot2::guide_colorbar(
      title.position = "top",
      barwidth = grid::unit(4.0, "in"),
      barheight = grid::unit(0.18, "in")
    )
  ) +
  ggplot2::labs(
    x = "Credible-set purity (minimum pairwise |r|)",
    y = expression("Core-95% effective support (" * k[eff] * ")"),
    title = "Purity vs diffuseness, colored by accuracy",
    subtitle = "Dashed line: purity = 0.50. Long dash: k_eff = 100."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")
ggplot2::ggsave(file.path(plot_dir, "diag_old_cs_purity_size_coverage.png"),
                p_old_cs, width = 7, height = 5.5, dpi = 300)
ggplot2::ggsave(file.path(plot_dir, "diag_purity_vs_accuracy.png"),
                p_purity_acc, width = 7, height = 5.5, dpi = 300)
ggplot2::ggsave(file.path(plot_dir, "diag_keff_vs_accuracy.png"),
                p_keff_acc, width = 7, height = 5.5, dpi = 300)
ggplot2::ggsave(file.path(plot_dir, "diag_purity_vs_keff.png"),
                p_purity_keff, width = 7, height = 5.5, dpi = 300)
ggplot2::ggsave(file.path(plot_dir, "diag_purity_vs_keff_accuracy_ratio.png"),
                p_purity_keff_accuracy, width = 7, height = 5.5, dpi = 300)

p_old_cs
p_purity_acc
p_keff_acc
p_purity_keff
p_purity_keff_accuracy
```

```{r effect-diagnostics-summary, eval = exists("effect_unfiltered") && !is.null(effect_unfiltered)}
# Quantify the two purity failure modes and the direct k_eff filter target.
diag_summary <- eff %>%
  dplyr::filter(is.finite(.data$purity), is.finite(.data$accuracy_ratio)) %>%
  dplyr::summarise(
    n_effects = dplyr::n(),
    cor_purity_keff = suppressWarnings(stats::cor(.data$purity, .data$k_eff_core95,
                                                  use = "complete.obs", method = "spearman")),
    cor_purity_accuracy = suppressWarnings(stats::cor(.data$purity, .data$accuracy_ratio,
                                                      use = "complete.obs", method = "spearman")),
    frac_accurate_dropped_at_095 = mean(.data$purity < 0.95 & .data$accuracy_ratio >= 0.9 &
                                          .data$coverage > 0),
    frac_pure_but_inaccurate = mean(.data$purity >= 0.95 & .data$accuracy_ratio < 0.5),
    n_keff_dropped = sum(.data$k_eff_core95 > keff_filter_cutoff, na.rm = TRUE),
    frac_keff_dropped = mean(.data$k_eff_core95 > keff_filter_cutoff, na.rm = TRUE),
    frac_keff_dropped_covering_causal = mean(.data$coverage > 0 &
                                               .data$k_eff_core95 > keff_filter_cutoff,
                                             na.rm = TRUE),
    frac_keff_dropped_accurate = mean(.data$coverage > 0 & .data$accuracy_ratio >= 0.9 &
                                        .data$k_eff_core95 > keff_filter_cutoff,
                                      na.rm = TRUE),
    n_low_purity_concentrated_true = sum(.data$purity < purity_cutoff & .data$coverage > 0 &
                                           .data$accuracy_ratio >= 0.9 & .data$k_eff_core95 <= 5,
                                         na.rm = TRUE),
    n_zero_purity_concentrated_true = sum(.data$purity <= 1e-8 & .data$coverage > 0 &
                                            .data$accuracy_ratio >= 0.9 & .data$k_eff_core95 <= 5,
                                          na.rm = TRUE),
    n_high_purity_concentrated_noncausal = sum(.data$purity >= purity_cutoff & .data$coverage <= 0 &
                                                 .data$k_eff_core95 <= 5,
                                               na.rm = TRUE)
  )

keff_filter_effect_summary <- eff %>%
  dplyr::filter(is.finite(.data$k_eff_core95)) %>%
  dplyr::mutate(
    filter_group = dplyr::if_else(
      .data$k_eff_core95 <= keff_filter_cutoff,
      "kept_by_keff_filter",
      "dropped_by_keff_filter"
    )
  ) %>%
  dplyr::group_by(.data$filter_group) %>%
  dplyr::summarise(
    n_effects = dplyr::n(),
    frac_covering_causal = mean(.data$coverage > 0, na.rm = TRUE),
    frac_accuracy_ge_0_9 = mean(.data$accuracy_ratio >= 0.9, na.rm = TRUE),
    median_accuracy_ratio = stats::median(.data$accuracy_ratio, na.rm = TRUE),
    median_purity = stats::median(.data$purity, na.rm = TRUE),
    median_k_eff_core95 = stats::median(.data$k_eff_core95, na.rm = TRUE),
    .groups = "drop"
  )

purity_failure_mode_summary <- eff %>%
  dplyr::filter(is.finite(.data$purity), is.finite(.data$k_eff_core95)) %>%
  dplyr::mutate(
    failure_mode = dplyr::case_when(
      .data$purity <= 1e-8 & .data$coverage > 0 &
        .data$accuracy_ratio >= 0.9 & .data$k_eff_core95 <= 5 ~
        "zero_purity_concentrated_true",
      .data$purity < purity_cutoff & .data$coverage > 0 &
        .data$accuracy_ratio >= 0.9 & .data$k_eff_core95 <= 5 ~
        "low_purity_concentrated_true",
      .data$purity >= purity_cutoff & .data$coverage <= 0 & .data$k_eff_core95 <= 5 ~
        "high_purity_concentrated_noncausal",
      .data$k_eff_core95 > keff_filter_cutoff & .data$coverage > 0 ~
        "diffuse_causal_containing",
      .data$k_eff_core95 > keff_filter_cutoff & .data$coverage <= 0 ~
        "diffuse_noncausal",
      TRUE ~ "other"
    )
  ) %>%
  dplyr::count(.data$failure_mode, name = "n_effects") %>%
  dplyr::mutate(frac_effects = .data$n_effects / sum(.data$n_effects)) %>%
  dplyr::arrange(dplyr::desc(.data$n_effects))

kable(diag_summary, digits = 4,
      caption = "Decoupling of classical (purity) and effect-level (accuracy / k_eff) diagnostics")
kable(keff_filter_effect_summary, digits = 4,
      caption = "Effect-level composition of the direct k_eff filter")
kable(purity_failure_mode_summary, digits = 4,
      caption = "Effect counts for purity and diffuseness failure modes")
```

```{r filter-retention-summary, eval = exists("confusion_bins") && !is.null(confusion_bins) && all(c("filter_type", "n_effects_kept") %in% names(confusion_bins)) && any(c("dataset_bundle_id", "run_id") %in% names(confusion_bins))}
retention_unit_col <- if ("dataset_bundle_id" %in% names(confusion_bins)) {
  "dataset_bundle_id"
} else {
  "run_id"
}

filter_retention_summary <- confusion_bins %>%
  dplyr::distinct(
    .data$filter_type,
    retention_unit = .data[[retention_unit_col]],
    .data$n_effects_kept
  ) %>%
  dplyr::group_by(.data$filter_type) %>%
  dplyr::summarise(
    n_units = dplyr::n(),
    unit = retention_unit_col,
    n_all_effects_dropped = sum(.data$n_effects_kept == 0, na.rm = TRUE),
    frac_all_effects_dropped = mean(.data$n_effects_kept == 0, na.rm = TRUE),
    mean_effects_kept = mean(.data$n_effects_kept, na.rm = TRUE),
    median_effects_kept = stats::median(.data$n_effects_kept, na.rm = TRUE),
    min_effects_kept = min(.data$n_effects_kept, na.rm = TRUE),
    max_effects_kept = max(.data$n_effects_kept, na.rm = TRUE),
    .groups = "drop"
  )

kable(filter_retention_summary, digits = 4,
      caption = "Effect retention by filter; all-zero filtered fits have n_effects_kept = 0")
```

# 6. Paper Figure Composite

```{r paper-figure-composite}
if (exists("confusion_bins") && !is.null(confusion_bins)) {
  pr_required_cols <- c(
    "filter_type", "pip_threshold", "n_causal_at_bucket", "n_noncausal_at_bucket"
  )
  if (all(pr_required_cols %in% names(confusion_bins))) {
    pr_curves_by_filter <- confusion_bins %>%
      dplyr::group_by(.data$filter_type, .data$pip_threshold) %>%
      dplyr::summarise(
        n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
        n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(.data$filter_type) %>%
      dplyr::arrange(dplyr::desc(.data$pip_threshold), .by_group = TRUE) %>%
      dplyr::mutate(
        TP = cumsum(.data$n_causal_at_bucket),
        FP = cumsum(.data$n_noncausal_at_bucket),
        total_causal = sum(.data$n_causal_at_bucket, na.rm = TRUE),
        precision = dplyr::if_else(
          .data$TP + .data$FP > 0,
          .data$TP / (.data$TP + .data$FP),
          NA_real_
        ),
        recall = dplyr::if_else(
          .data$total_causal > 0,
          .data$TP / .data$total_causal,
          NA_real_
        ),
        filter_label = dplyr::coalesce(
          unname(filter_labels[as.character(.data$filter_type)]),
          as.character(.data$filter_type)
        )
      ) %>%
      dplyr::ungroup() %>%
      dplyr::filter(is.finite(.data$precision), is.finite(.data$recall)) %>%
      dplyr::mutate(
        filter_type = factor(as.character(.data$filter_type), levels = filter_levels),
        filter_label = factor(.data$filter_label, levels = unname(filter_labels[filter_levels]))
      )

    pr_plot_filter_levels <- setdiff(filter_levels, "purity_90")
    pr_curves_by_filter <- pr_curves_by_filter %>%
      dplyr::filter(
        as.character(.data$filter_type) %in% pr_plot_filter_levels,
        .data$pip_threshold > 0,
        .data$recall > 0
      ) %>%
      dplyr::mutate(
        filter_type = factor(as.character(.data$filter_type), levels = pr_plot_filter_levels),
        filter_label = factor(
          as.character(.data$filter_label),
          levels = unname(filter_labels[pr_plot_filter_levels])
        )
      )

    p_pr_curves <- ggplot2::ggplot(
      pr_curves_by_filter,
      ggplot2::aes(x = .data$recall, y = .data$precision, color = .data$filter_label)
    ) +
      ggplot2::geom_line(linewidth = 1.1, alpha = 0.9) +
      ggplot2::geom_hline(yintercept = 0.5, linetype = "dotted", color = "grey45") +
      ggplot2::coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      ggplot2::labs(
        x = "Recall",
        y = "Precision",
        color = "Filter",
        title = "Pooled precision-recall curves by credible-set filter",
        subtitle = "Dotted line: 50% precision. Curves are pooled across all simulation datasets."
      ) +
      ggplot2::theme_bw(base_size = 12) +
      ggplot2::guides(color = ggplot2::guide_legend(nrow = 2, byrow = TRUE)) +
      ggplot2::theme(legend.position = "bottom")

    pr_curve_path <- file.path(plot_dir, "pr_curves_by_filter_type.png")
    ggplot2::ggsave(pr_curve_path, p_pr_curves, width = 8, height = 5.5, dpi = 300)
    cat("PR curves written to:", pr_curve_path, "\n")
  } else {
    message(
      "Skipping PR curve plot; confusion_bins is missing required columns: ",
      paste(setdiff(pr_required_cols, names(confusion_bins)), collapse = ", ")
    )
  }
}
if ((!exists("p_purity_acc") || !exists("p_keff_acc") || !exists("p_purity_keff")) &&
    exists("effect_unfiltered") && !is.null(effect_unfiltered)) {
  eff <- effect_unfiltered %>%
    dplyr::filter(.data$size > 0) %>%
    dplyr::mutate(
      coverage_lab = factor(
        ifelse(.data$coverage > 0, "Contains causal variant", "No causal variant"),
        levels = c("Contains causal variant", "No causal variant")
      ),
      purity = as.numeric(.data$purity),
      accuracy_ratio = as.numeric(.data$accuracy_ratio),
      k_eff_core95 = as.numeric(.data$effect_k_eff_signal_core95),
      size = as.numeric(.data$size)
    )

  cov_palette <- c("Contains causal variant" = "#0072B2", "No causal variant" = "#D55E00")
  purity_cutoff <- 0.50
  keff_filter_cutoff <- 100
  scatter_point_size <- 3.2

  p_purity_acc <- ggplot2::ggplot(
    eff %>% dplyr::filter(is.finite(.data$purity), is.finite(.data$accuracy_ratio)),
    ggplot2::aes(x = .data$purity, y = .data$accuracy_ratio, color = .data$coverage_lab)
  ) +
    ggplot2::geom_point(alpha = 0.45, size = scatter_point_size) +
    ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
    ggplot2::scale_color_manual(values = cov_palette, name = "Credible set") +
    ggplot2::labs(
      x = "Credible-set purity (minimum pairwise |r|)",
      y = "Accuracy ratio",
      title = "Purity vs accuracy",
      subtitle = "Dashed line: 0.50 purity threshold."
    ) +
    ggplot2::theme_bw(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")

  p_keff_acc <- ggplot2::ggplot(
    eff %>%
      dplyr::filter(is.finite(.data$k_eff_core95), is.finite(.data$accuracy_ratio)) %>%
      dplyr::arrange(.data$accuracy_ratio),
    ggplot2::aes(x = .data$k_eff_core95, y = .data$accuracy_ratio, color = .data$accuracy_ratio)
  ) +
    ggplot2::geom_point(alpha = 0.65, size = scatter_point_size) +
    ggplot2::geom_vline(xintercept = keff_filter_cutoff, linetype = "longdash", color = "grey35") +
    ggplot2::geom_hline(yintercept = 0.9, linetype = "dotted", color = "grey45") +
    ggplot2::scale_x_log10() +
    ggplot2::scale_color_gradientn(
      colors = c("#D55E00", "#F0E442", "#0072B2"),
      name = "Accuracy ratio",
      guide = ggplot2::guide_colorbar(
        title.position = "top",
        barwidth = grid::unit(4.0, "in"),
        barheight = grid::unit(0.18, "in")
      )
    ) +
    ggplot2::labs(
      x = expression("Core-95% effective support (" * k[eff] * ")"),
      y = "Accuracy ratio",
      title = "Effective support vs accuracy",
      subtitle = "Long dash: k_eff = 100 filter. Dotted line: A = 0.9."
    ) +
    ggplot2::theme_bw(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")

  purity_keff_plot_data <- eff %>%
    dplyr::filter(is.finite(.data$purity), is.finite(.data$k_eff_core95))

  purity_side_summary <- purity_keff_plot_data %>%
    dplyr::mutate(
      purity_side = dplyr::if_else(.data$purity < purity_cutoff, "left", "right")
    ) %>%
    dplyr::group_by(.data$purity_side) %>%
    dplyr::summarise(
      n_effects = dplyr::n(),
      pct_no_causal = 100 * mean(.data$coverage <= 0),
      .groups = "drop"
    ) %>%
    dplyr::right_join(
      data.frame(purity_side = c("left", "right")),
      by = "purity_side"
    ) %>%
    dplyr::mutate(
      n_effects = dplyr::coalesce(.data$n_effects, 0L),
      pct_no_causal = dplyr::coalesce(.data$pct_no_causal, 0),
      x = dplyr::if_else(.data$purity_side == "left", purity_cutoff - 0.02, purity_cutoff + 0.02),
      hjust = dplyr::if_else(.data$purity_side == "left", 1, 0),
      label = sprintf(
        "%s\nn = %s\nNo causal: %.1f%%",
        dplyr::if_else(.data$purity_side == "left", "Purity < 0.50", "Purity >= 0.50"),
        format(.data$n_effects, big.mark = ",", scientific = FALSE),
        .data$pct_no_causal
      )
    )

  p_purity_keff <- ggplot2::ggplot(
    purity_keff_plot_data,
    ggplot2::aes(x = .data$purity, y = .data$k_eff_core95, color = .data$coverage_lab)
  ) +
    ggplot2::geom_point(alpha = 0.45, size = scatter_point_size) +
    ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
    ggplot2::geom_hline(yintercept = keff_filter_cutoff, linetype = "longdash", color = "grey35") +
    ggplot2::geom_label(
      data = purity_side_summary,
      ggplot2::aes(x = .data$x, y = Inf, label = .data$label, hjust = .data$hjust),
      inherit.aes = FALSE,
      vjust = 1.15,
      size = 5.4,
      label.size = 0,
      fill = "white",
      alpha = 0.85
    ) +
    ggplot2::scale_y_log10() +
    ggplot2::scale_color_manual(values = cov_palette, name = "Credible set") +
    ggplot2::labs(
      x = "Credible-set purity (minimum pairwise |r|)",
      y = expression("Core-95% effective support (" * k[eff] * ")"),
      title = "Purity vs diffuseness",
      subtitle = "Weak association: low-purity sets are not reliably more diffuse."
    ) +
    ggplot2::theme_bw(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")
}

if (!exists("p_old_cs") && exists("eff") && !is.null(eff)) {
  p_old_cs <- ggplot2::ggplot(
    eff %>% dplyr::filter(is.finite(.data$purity), is.finite(.data$size), .data$size > 0),
    ggplot2::aes(x = .data$purity, y = .data$size, color = .data$coverage_lab)
  ) +
    ggplot2::geom_point(alpha = 0.45, size = scatter_point_size) +
    ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
    ggplot2::scale_y_log10() +
    ggplot2::scale_color_manual(values = cov_palette, name = "95% CS") +
    ggplot2::labs(
      x = "Credible-set purity (minimum pairwise |r|)",
      y = "95% credible-set size",
      title = "Classical credible-set summaries",
      subtitle = "Dashed line: 0.50 purity threshold."
    ) +
    ggplot2::theme_bw(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")
}

if (exists("p_old_cs")) {
  old_cs_path <- file.path(plot_dir, "diag_old_cs_purity_size_coverage.png")
  ggplot2::ggsave(old_cs_path, p_old_cs, width = 7, height = 5.5, dpi = 300)
  cat("Old-CS diagnostic plot written to:", old_cs_path, "\n")
}

if (exists("purity_keff_plot_data") && !is.null(purity_keff_plot_data)) {
  p_purity_keff_accuracy <- ggplot2::ggplot(
    purity_keff_plot_data %>%
      dplyr::filter(is.finite(.data$accuracy_ratio)) %>%
      dplyr::arrange(.data$accuracy_ratio),
    ggplot2::aes(x = .data$purity, y = .data$k_eff_core95, color = .data$accuracy_ratio)
  ) +
    ggplot2::geom_point(alpha = 0.65, size = scatter_point_size) +
    ggplot2::geom_vline(xintercept = purity_cutoff, linetype = "dashed", color = "grey25") +
    ggplot2::geom_hline(yintercept = keff_filter_cutoff, linetype = "longdash", color = "grey35") +
    ggplot2::scale_y_log10() +
    ggplot2::scale_color_gradientn(
      colors = c("#D55E00", "#F0E442", "#0072B2"),
      name = "Accuracy ratio",
      guide = ggplot2::guide_colorbar(
        title.position = "top",
        barwidth = grid::unit(4.0, "in"),
        barheight = grid::unit(0.18, "in")
      )
    ) +
    ggplot2::labs(
      x = "Credible-set purity (minimum pairwise |r|)",
      y = expression("Core-95% effective support (" * k[eff] * ")"),
      title = "Purity vs diffuseness, colored by accuracy",
      subtitle = "Dashed line: purity = 0.50. Long dash: k_eff = 100."
    ) +
    ggplot2::theme_bw(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")

  purity_keff_accuracy_path <- file.path(plot_dir, "diag_purity_vs_keff_accuracy_ratio.png")
  ggplot2::ggsave(purity_keff_accuracy_path,
                  p_purity_keff_accuracy, width = 7, height = 5.5, dpi = 300)
  cat("Purity/effective-support accuracy plot written to:", purity_keff_accuracy_path, "\n")
}
required_paper_plot_objects <- c("p_pr_curves", "p_old_cs", "p_keff_acc", "p_purity_keff_accuracy")
missing_paper_plot_objects <- required_paper_plot_objects[
  !vapply(required_paper_plot_objects, exists, logical(1), inherits = TRUE)
]
if (length(missing_paper_plot_objects)) {
  stop(
    "Paper composite requires these plot objects; rerun the plot chunks first and ensure confusion_bins.csv plus effect_metrics_unfiltered_full.csv are available: ",
    paste(missing_paper_plot_objects, collapse = ", ")
  )
}

paper_panel_theme <- ggplot2::theme(
  plot.title = ggplot2::element_text(size = 24, face = "bold", hjust = 0.5),
  plot.subtitle = ggplot2::element_blank(),
  plot.caption = ggplot2::element_blank(),
  axis.title = ggplot2::element_text(size = 22),
  axis.text = ggplot2::element_text(size = 20),
  legend.title = ggplot2::element_text(size = 22),
  legend.text = ggplot2::element_text(size = 20),
  legend.position = "bottom",
  legend.direction = "horizontal",
  plot.margin = ggplot2::margin(10, 10, 10, 10)
)

extract_plot_legend <- function(plot) {
  plot_grob <- ggplot2::ggplotGrob(plot)
  guide_index <- which(vapply(plot_grob$grobs, function(grob) grob$name, character(1)) == "guide-box")
  if (!length(guide_index)) return(NULL)
  plot_grob$grobs[[guide_index[1]]]
}

open_png_device <- function(filename, width, height, dpi = 300) {
  if (requireNamespace("ragg", quietly = TRUE)) {
    ragg::agg_png(filename, width = width, height = height, units = "in",
                  res = dpi, background = "white")
  } else {
    grDevices::png(filename, width = width, height = height, units = "in",
                   res = dpi, bg = "white")
  }
}

save_paper_grid <- function(filename, panels, labels, legend_grob = NULL,
                            width = 19.8, height = 14.5, dpi = 300) {
  stopifnot(length(panels) == 4L, length(labels) == 4L)

  open_png_device(filename, width = width, height = height, dpi = dpi)
  layout_pushed <- FALSE
  on.exit({
    if (layout_pushed) grid::popViewport()
    grDevices::dev.off()
  }, add = TRUE)

  grid::grid.newpage()
  if (is.null(legend_grob)) {
    layout <- grid::grid.layout(
      nrow = 2, ncol = 2,
      widths = grid::unit(c(1, 1), "null"),
      heights = grid::unit(c(1, 1), "null")
    )
  } else {
    layout <- grid::grid.layout(
      nrow = 3, ncol = 2,
      widths = grid::unit(c(1, 1), "null"),
      heights = grid::unit.c(grid::unit(c(1, 1), "null"), grid::unit(0.85, "in"))
    )
  }
  grid::pushViewport(grid::viewport(layout = layout))
  layout_pushed <- TRUE

  positions <- data.frame(row = c(1, 1, 2, 2), col = c(1, 2, 1, 2))
  for (i in seq_along(panels)) {
    grid::pushViewport(grid::viewport(
      layout.pos.row = positions$row[i],
      layout.pos.col = positions$col[i]
    ))
    print(panels[[i]], newpage = FALSE)
    grid::grid.text(
      labels[[i]],
      x = grid::unit(0.01, "npc"),
      y = grid::unit(0.99, "npc"),
      just = c("left", "top"),
      gp = grid::gpar(fontsize = 30, fontface = "bold")
    )
    grid::popViewport()
  }

  if (!is.null(legend_grob)) {
    grid::pushViewport(grid::viewport(layout.pos.row = 3, layout.pos.col = 1:2))
    grid::grid.draw(legend_grob)
    grid::popViewport()
  }

  invisible(filename)
}

paper_shared_legend <- NULL

paper_panels <- list(
  p_pr_curves +
    ggplot2::labs(title = "Pooled precision-recall curves") +
    paper_panel_theme,
  p_old_cs +
    ggplot2::labs(title = "Classical CS summaries") +
    paper_panel_theme,
  p_keff_acc +
    ggplot2::labs(title = "Effective support versus accuracy", y = "Accuracy ratio") +
    paper_panel_theme,
  p_purity_keff_accuracy +
    ggplot2::labs(title = "Purity versus effective support") +
    paper_panel_theme
)

paper_purity_filter_path <- file.path(plot_dir, "paper_purity_filter_composite.png")
save_paper_grid(
  paper_purity_filter_path,
  panels = paper_panels,
  labels = LETTERS[1:4],
  legend_grob = paper_shared_legend
)

cat("Paper composite written to:", paper_purity_filter_path, "\n")
```

# 7. Export Consolidated Files

```{r export}
write_if_present <- function(df, filename) {
  if (!is.null(df) && nrow(df)) {
    readr::write_csv(df, file.path(consolidated_dir, filename))
  }
}

if (isTRUE(run_collection) || !consolidated_summary_cache_available) {
  write_if_present(auprc_pooled, "auprc_by_filter_type.csv")
  write_if_present(auprc_per_dataset, "auprc_per_dataset_by_filter.csv")
  write_if_present(per_dataset_summary, "auprc_per_dataset_delta_summary.csv")
  if (exists("effect_unfiltered")) write_if_present(effect_unfiltered, "effect_metrics_unfiltered_full.csv")

  cat("Consolidated files written to:", consolidated_dir, "\n")
} else {
  cat("Skipped consolidated data export; consolidated summary cache already exists.\n")
}
if (exists("diag_summary")) write_if_present(diag_summary, "effect_diagnostic_summary.csv")
if (exists("keff_filter_effect_summary")) write_if_present(keff_filter_effect_summary, "keff_filter_effect_summary.csv")
if (exists("purity_failure_mode_summary")) write_if_present(purity_failure_mode_summary, "purity_failure_mode_summary.csv")
if (exists("filter_retention_summary")) write_if_present(filter_retention_summary, "filter_retention_summary.csv")
if (exists("pip_mass_summary")) write_if_present(pip_mass_summary, "pip_mass_ratio_by_filter.csv")
if (exists("pr_curves_by_filter")) write_if_present(pr_curves_by_filter, "pr_curves_by_filter_type.csv")
if (exists("effect_unfiltered")) write_if_present(effect_unfiltered, "effect_metrics_unfiltered_full.csv")

cat("Plots written to:", plot_dir, "\n")
if (exists("paper_purity_filter_path")) {
  cat("Paper composite:", paper_purity_filter_path, "\n")
}

diagnostic_csvs <- c(
  "auprc_by_filter_type.csv",
  "auprc_per_dataset_by_filter.csv",
  "auprc_per_dataset_delta_summary.csv",
  "effect_diagnostic_summary.csv",
  "keff_filter_effect_summary.csv",
  "purity_failure_mode_summary.csv",
  "filter_retention_summary.csv",
  "pip_mass_ratio_by_filter.csv",
  "pr_curves_by_filter_type.csv",
  "effect_metrics_unfiltered_full.csv",
  "confusion_bins.csv"
)
diagnostic_paths <- tibble::tibble(
  diagnostic = diagnostic_csvs,
  path = file.path(consolidated_dir, diagnostic_csvs)
) %>%
  dplyr::mutate(exists = file.exists(.data$path))
cat("\nDiagnostic CSVs to download/send back if validating Figure 4:\n")
print(diagnostic_paths, n = Inf)
```
