---
title: "Architecture Sensitivity - Collect Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Post-hoc collection for the reviewer genetic-architecture sensitivity jobs. This
workbook aggregates the sparse and diffuse Slurm outputs, validates the fixed-c
prior gate, and writes combined tables for the visualization workbook.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_architecture_sensitivity.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(knitr)
```

# Configuration

Set the parent Slurm array IDs returned by `sbatch` before rendering.

```{r user-config}
output_root <- here("output")

sparse_job_name <- "architecture_sparse_sensitivity"
sparse_parent_job_id <- "SET_SPARSE_PARENT_JOB_ID"

diffuse_job_name <- "architecture_diffuse_sensitivity"
diffuse_parent_job_id <- "SET_DIFFUSE_PARENT_JOB_ID"

force_reaggregate <- TRUE
write_snps_dataset <- FALSE
precision_threshold <- 0.75

combined_root <- file.path(output_root, "architecture_sensitivity")
combined_dir <- file.path(combined_root, "combined")
dir.create(combined_dir, recursive = TRUE, showWarnings = FALSE)

if (identical(sparse_parent_job_id, "SET_SPARSE_PARENT_JOB_ID") ||
    identical(diffuse_parent_job_id, "SET_DIFFUSE_PARENT_JOB_ID")) {
  stop("Set sparse_parent_job_id and diffuse_parent_job_id before collecting.")
}
```

# Aggregate Jobs

```{r aggregate-jobs}
collect_job <- function(job_name, parent_job_id, analysis_label) {
  job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
  aggregated_dir <- file.path(job_dir, "aggregated")
  run_table_path <- file.path(output_root, "run_history", job_name, parent_job_id, "run_table.csv")

  if (force_reaggregate || !file.exists(file.path(aggregated_dir, "confusion_bins.csv"))) {
    test_susine::aggregate_staging_outputs(
      job_name = job_name,
      parent_job_id = parent_job_id,
      output_root = output_root,
      validate = TRUE,
      output_dir = aggregated_dir,
      write_snps_dataset = write_snps_dataset
    )
  }

  if (!file.exists(run_table_path)) {
    stop("Run table not found: ", run_table_path)
  }

  list(
    analysis = analysis_label,
    job_name = job_name,
    parent_job_id = parent_job_id,
    job_dir = job_dir,
    aggregated_dir = aggregated_dir,
    run_table = readr::read_csv(run_table_path, show_col_types = FALSE)
  )
}

jobs <- list(
  sparse = collect_job(sparse_job_name, sparse_parent_job_id, "sparse"),
  diffuse = collect_job(diffuse_job_name, diffuse_parent_job_id, "diffuse")
)

job_index <- tibble::tibble(
  analysis = vapply(jobs, `[[`, character(1), "analysis"),
  job_name = vapply(jobs, `[[`, character(1), "job_name"),
  parent_job_id = vapply(jobs, `[[`, character(1), "parent_job_id"),
  aggregated_dir = vapply(jobs, `[[`, character(1), "aggregated_dir")
)
readr::write_csv(job_index, file.path(combined_dir, "job_index.csv"))
kable(job_index)
```

# Load Tables

```{r load-tables}
read_optional_csv <- function(dir, filename) {
  path <- file.path(dir, filename)
  if (!file.exists(path)) return(NULL)
  readr::read_csv(path, show_col_types = FALSE)
}

bind_job_table <- function(filename) {
  pieces <- lapply(jobs, function(a) {
    tbl <- read_optional_csv(a$aggregated_dir, filename)
    if (is.null(tbl)) return(NULL)
    dplyr::mutate(
      tbl,
      analysis = a$analysis,
      source_job_name = a$job_name,
      source_parent_job_id = a$parent_job_id
    )
  })
  dplyr::bind_rows(pieces)
}

confusion_bins <- bind_job_table("confusion_bins.csv")
model_metrics <- bind_job_table("model_metrics.csv")
effect_metrics <- bind_job_table("effect_metrics.csv")
effect_metrics_unfiltered <- bind_job_table("effect_metrics_unfiltered.csv")
prior_diagnostics <- bind_job_table("prior_diagnostics.csv")
validation <- bind_job_table("validation.csv")

cat("Confusion-bin rows:", nrow(confusion_bins), "\n")
cat("Prior diagnostic rows:", nrow(prior_diagnostics), "\n")
```

# Validation Gates

```{r validation-gates}
required_confusion_cols <- c(
  "analysis", "spec_name", "explore_method", "architecture", "p_star",
  "h2_snp_per_causal", "h2_total", "diffuse_k"
)
missing_confusion_cols <- setdiff(required_confusion_cols, names(confusion_bins))
if (!nrow(confusion_bins) || length(missing_confusion_cols)) {
  stop(
    "confusion_bins.csv is missing or incomplete. Missing columns: ",
    paste(missing_confusion_cols, collapse = ", ")
  )
}

required_prior_cols <- c("spec_name", "c_l")
missing_prior_cols <- setdiff(required_prior_cols, names(prior_diagnostics))
if (!nrow(prior_diagnostics) || length(missing_prior_cols)) {
  stop(
    "prior_diagnostics.csv is missing or incomplete. Set write_prior_diagnostics = TRUE before launch. Missing columns: ",
    paste(missing_prior_cols, collapse = ", ")
  )
}

ccs_specs <- c("C-CS-architecture-sparse", "C-CS-architecture-diffuse")
prior_gate <- prior_diagnostics %>%
  dplyr::filter(.data$spec_name %in% ccs_specs) %>%
  dplyr::group_by(.data$analysis, .data$spec_name) %>%
  dplyr::summarise(
    n_rows = dplyr::n(),
    max_abs_c_l_minus_1 = max(abs(.data$c_l - 1), na.rm = TRUE),
    all_c_l_one = all(abs(.data$c_l - 1) < 1e-12, na.rm = TRUE),
    .groups = "drop"
  )

if (!nrow(prior_gate) || any(prior_gate$n_rows <= 0L) || !all(prior_gate$all_c_l_one)) {
  stop("Fixed-c prior gate failed: c_l differs from 1 or no C-CS prior rows were found.")
}

confusion_presence <- confusion_bins %>%
  dplyr::filter(.data$spec_name %in% c(ccs_specs, "baseline-single")) %>%
  dplyr::group_by(
    .data$analysis, .data$spec_name, .data$architecture,
    .data$p_star, .data$h2_snp_per_causal, .data$h2_total, .data$diffuse_k
  ) %>%
  dplyr::summarise(n_rows = dplyr::n(), .groups = "drop")

if (!nrow(confusion_presence)) {
  stop("No architecture-sensitivity confusion bins were found.")
}

readr::write_csv(prior_gate, file.path(combined_dir, "prior_gate.csv"))
readr::write_csv(confusion_presence, file.path(combined_dir, "confusion_presence.csv"))

kable(prior_gate, digits = 4, caption = "Fixed-c prior gate")
kable(confusion_presence, caption = "Confusion-bin presence by architecture setting")
```

# Metrics

```{r metrics}
ccs_confusion <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name %in% ccs_specs,
    .data$explore_method == "aggregation",
    .data$agg_method == "cluster_weight_credible"
  )

baseline_confusion <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "baseline-single",
    .data$explore_method == "base"
  )

if (!nrow(ccs_confusion)) {
  stop("No C-CS aggregate confusion-bin rows found.")
}
if (!nrow(baseline_confusion)) {
  stop("No baseline-single base confusion-bin rows found.")
}

metric_group_vars <- c(
  "analysis", "architecture", "p_star", "h2_snp_per_causal", "h2_total", "diffuse_k"
)

auprc_pooled <- test_susine::compute_auprc_from_confusion(
  ccs_confusion,
  group_vars = metric_group_vars
)

recall75_pooled <- test_susine::compute_recall_at_precision_from_confusion(
  ccs_confusion,
  group_vars = metric_group_vars,
  precision_threshold = precision_threshold
)

baseline_auprc <- test_susine::compute_auprc_from_confusion(
  baseline_confusion,
  group_vars = metric_group_vars
)

baseline_recall75 <- test_susine::compute_recall_at_precision_from_confusion(
  baseline_confusion,
  group_vars = metric_group_vars,
  precision_threshold = precision_threshold
)

metric_summary <- auprc_pooled %>%
  dplyr::left_join(recall75_pooled, by = metric_group_vars) %>%
  dplyr::mutate(
    series = "C-CS",
    total_h2 = dplyr::if_else(
      .data$analysis == "sparse",
      as.numeric(.data$p_star) * as.numeric(.data$h2_snp_per_causal),
      as.numeric(.data$h2_total)
    )
  ) %>%
  dplyr::arrange(.data$analysis, .data$h2_snp_per_causal, .data$p_star, .data$h2_total)

baseline_summary <- baseline_auprc %>%
  dplyr::left_join(baseline_recall75, by = metric_group_vars) %>%
  dplyr::mutate(
    series = "Annotation-free baseline",
    total_h2 = dplyr::if_else(
      .data$analysis == "sparse",
      as.numeric(.data$p_star) * as.numeric(.data$h2_snp_per_causal),
      as.numeric(.data$h2_total)
    )
  ) %>%
  dplyr::arrange(.data$analysis, .data$h2_snp_per_causal, .data$p_star, .data$h2_total)

readr::write_csv(metric_summary, file.path(combined_dir, "metric_summary.csv"))
readr::write_csv(baseline_summary, file.path(combined_dir, "baseline_summary.csv"))

kable(metric_summary, digits = 4, caption = "Pooled architecture sensitivity C-CS metrics")
kable(baseline_summary, digits = 4, caption = "Annotation-free baseline metrics")
```

# Export Combined Raw Tables

```{r export}
write_if_present <- function(tbl, filename) {
  if (!is.null(tbl) && nrow(tbl)) {
    readr::write_csv(tbl, file.path(combined_dir, filename))
  }
}

write_if_present(confusion_bins, "confusion_bins_full.csv")
write_if_present(model_metrics, "model_metrics_full.csv")
write_if_present(effect_metrics, "effect_metrics_full.csv")
write_if_present(effect_metrics_unfiltered, "effect_metrics_unfiltered_full.csv")
write_if_present(prior_diagnostics, "prior_diagnostics_full.csv")
write_if_present(validation, "validation_full.csv")

cat("Combined architecture sensitivity outputs written to:", combined_dir, "\n")
```