---
title: "Warm Start Pilot v3 — Collect & Visualize"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Collects staging outputs from the warm start pilot v3
(init_alpha + prior_refit), then computes TPR @ FPR and multimodal
metrics split by warm_method and alpha_concentration.

```{r setup, message=FALSE}
here::i_am(
  "vignettes/simulation pipeline/collect_results_workbook_dirichlet_pilot_v3.Rmd"
)
library(here)
here()

library(devtools)
devtools::load_all(".")

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(stringr)
theme_set(theme_bw(base_size = 13))
```

# 1. Job Settings

```{r settings}
job_name <- "warm_start_pilot_dual"
parent_job_id <- "FILL_IN"
output_root <- here("output")

agg_dir <- file.path(
  output_root, "slurm_output", job_name, parent_job_id, "aggregated"
)
fig_dir <- file.path(agg_dir, "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Aggregate Staging Outputs

```{r aggregate}
if (!dir.exists(agg_dir) || !file.exists(file.path(agg_dir, "confusion_bins.csv"))) {
  aggregate_staging_outputs(
    job_name = job_name,
    parent_job_id = parent_job_id,
    output_root = output_root,
    validate = TRUE
  )
}

stopifnot(file.exists(file.path(agg_dir, "confusion_bins.csv")))
cat("Aggregated outputs found at:", agg_dir, "\n")
```

# 3. Load Data

```{r load-data}
conf <- readr::read_csv(
  file.path(agg_dir, "confusion_bins.csv"),
  show_col_types = FALSE
)

# Parse warm_method from group_key if not already a column
if (!"warm_method" %in% names(conf) && "group_key" %in% names(conf)) {
  conf <- conf %>%
    mutate(warm_method = stringr::str_extract(group_key, "(?<=wm=)[a-z_]+"))
}

stopifnot("alpha_concentration" %in% names(conf))
cat("Confusion bins rows:", nrow(conf), "\n")
cat("Warm methods:", sort(unique(conf$warm_method)), "\n")
cat("Alpha values:", sort(unique(conf$alpha_concentration)), "\n")
```

# 4. Baseline Reference

```{r baseline-reference}
# Default runs are identical across all (warm_method, alpha) groups — take any one
baseline_alpha <- sort(unique(conf$alpha_concentration))[1L]
conf_baseline <- conf %>%
  filter(
    (is.na(agg_method) | agg_method == ""),
    run_type == "default"
  ) %>%
  # Deduplicate: baseline is shared, take first alpha/method combo
  filter(alpha_concentration == baseline_alpha) %>%
  filter(warm_method == first(warm_method) | is.na(warm_method))

curves_baseline <- compute_power_fdr_curves(conf_baseline)

baseline_tpr_fpr005 <- {
  df <- arrange(distinct(curves_baseline, fpr, power), fpr)
  val <- stats::approx(df$fpr, df$power, xout = 0.05, ties = "ordered", rule = 2)$y
  if (is.na(val)) 0 else val
}
baseline_tpr_fpr010 <- {
  df <- arrange(distinct(curves_baseline, fpr, power), fpr)
  val <- stats::approx(df$fpr, df$power, xout = 0.10, ties = "ordered", rule = 2)$y
  if (is.na(val)) 0 else val
}

cat(sprintf("Baseline (uniform prior, no warm start) reference:\n"))
cat(sprintf("  TPR @ FPR=0.05: %.4f\n", baseline_tpr_fpr005))
cat(sprintf("  TPR @ FPR=0.10: %.4f\n", baseline_tpr_fpr010))
```

# 5. TPR @ FPR — Individual Fits by Method

```{r tpr-at-fpr-individual}
conf_fits <- conf %>%
  filter(is.na(agg_method) | agg_method == "")

curves_individual <- compute_power_fdr_curves(
  conf_fits,
  group_vars = c("warm_method", "alpha_concentration")
)

tpr_at_fpr_individual <- curves_individual %>%
  group_by(warm_method, alpha_concentration) %>%
  summarise(
    tpr_fpr005 = {
      df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
      val <- stats::approx(df$fpr, df$power, xout = 0.05, ties = "ordered", rule = 2)$y
      if (is.na(val)) 0 else val
    },
    tpr_fpr010 = {
      df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
      val <- stats::approx(df$fpr, df$power, xout = 0.10, ties = "ordered", rule = 2)$y
      if (is.na(val)) 0 else val
    },
    .groups = "drop"
  )

cat("Individual fits — TPR @ FPR thresholds:\n")
print(as.data.frame(tpr_at_fpr_individual))
```

```{r plot-fpr-individual}
p1 <- ggplot(tpr_at_fpr_individual,
             aes(x = factor(alpha_concentration), y = tpr_fpr005,
                 fill = warm_method)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
             color = "red", linewidth = 0.7) +
  annotate("text", x = Inf, y = baseline_tpr_fpr005,
           label = sprintf("baseline = %.3f", baseline_tpr_fpr005),
           hjust = 1.1, vjust = -0.5, color = "red", size = 3) +
  labs(
    x = "Dirichlet alpha",
    y = "TPR @ FPR = 0.05",
    fill = "Warm method",
    title = "Individual Fits: TPR @ FPR = 0.05 by Method and Alpha",
    subtitle = "Pooled across all datasets and restarts"
  ) +
  ylim(0, NA)

ggsave(file.path(fig_dir, "tpr_fpr005_individual_by_method.png"),
       p1, width = 9, height = 5, dpi = 150)
cat("Saved:", file.path(fig_dir, "tpr_fpr005_individual_by_method.png"), "\n")
```

# 6. TPR @ FPR — Aggregated PIPs by Method

```{r tpr-at-fpr-aggregated}
conf_agg <- conf %>%
  filter(!is.na(agg_method) & agg_method != "")

if (nrow(conf_agg) > 0) {
  curves_agg <- compute_power_fdr_curves(
    conf_agg,
    group_vars = c("warm_method", "alpha_concentration", "agg_method")
  )

  tpr_at_fpr_agg <- curves_agg %>%
    group_by(warm_method, alpha_concentration, agg_method) %>%
    summarise(
      tpr_fpr005 = {
        df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
        val <- stats::approx(df$fpr, df$power, xout = 0.05, ties = "ordered", rule = 2)$y
        if (is.na(val)) 0 else val
      },
      tpr_fpr010 = {
        df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
        val <- stats::approx(df$fpr, df$power, xout = 0.10, ties = "ordered", rule = 2)$y
        if (is.na(val)) 0 else val
      },
      .groups = "drop"
    )

  cat("Aggregated — TPR @ FPR thresholds:\n")
  print(as.data.frame(tpr_at_fpr_agg))
} else {
  tpr_at_fpr_agg <- tibble()
}
```

```{r plot-fpr-aggregated}
if (nrow(tpr_at_fpr_agg) > 0) {
  p2 <- ggplot(
    tpr_at_fpr_agg,
    aes(x = factor(alpha_concentration), y = tpr_fpr005, fill = agg_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    facet_wrap(~warm_method, ncol = 1,
               labeller = labeller(warm_method = c(
                 init_alpha = "init_alpha (trajectory perturbation)",
                 prior_refit = "prior_refit (explore + refit)"
               ))) +
    labs(
      x = "Dirichlet alpha",
      y = "TPR @ FPR = 0.05",
      fill = "Aggregation",
      title = "Aggregated PIPs: TPR @ FPR = 0.05 by Method and Alpha",
      subtitle = "Pooled across all datasets"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "tpr_fpr005_aggregated_by_method.png"),
         p2, width = 10, height = 8, dpi = 150)
  cat("Saved:", file.path(fig_dir, "tpr_fpr005_aggregated_by_method.png"), "\n")
}
```

```{r plot-fpr010-aggregated}
if (nrow(tpr_at_fpr_agg) > 0) {
  p2b <- ggplot(
    tpr_at_fpr_agg,
    aes(x = factor(alpha_concentration), y = tpr_fpr010, fill = agg_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fpr010, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    facet_wrap(~warm_method, ncol = 1,
               labeller = labeller(warm_method = c(
                 init_alpha = "init_alpha (trajectory perturbation)",
                 prior_refit = "prior_refit (explore + refit)"
               ))) +
    labs(
      x = "Dirichlet alpha",
      y = "TPR @ FPR = 0.10",
      fill = "Aggregation",
      title = "Aggregated PIPs: TPR @ FPR = 0.10 by Method and Alpha",
      subtitle = "Pooled across all datasets"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "tpr_fpr010_aggregated_by_method.png"),
         p2b, width = 10, height = 8, dpi = 150)
  cat("Saved:", file.path(fig_dir, "tpr_fpr010_aggregated_by_method.png"), "\n")
}
```

# 7. Head-to-Head: Best Alpha per Method

```{r best-alpha-comparison}
if (nrow(tpr_at_fpr_agg) > 0) {
  best_per_method <- tpr_at_fpr_agg %>%
    group_by(warm_method, agg_method) %>%
    slice_max(tpr_fpr005, n = 1, with_ties = FALSE) %>%
    ungroup()

  cat("Best alpha per (warm_method, agg_method):\n")
  print(as.data.frame(best_per_method))

  p_h2h <- ggplot(
    best_per_method,
    aes(x = agg_method, y = tpr_fpr005, fill = warm_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    geom_text(aes(label = sprintf("a=%.0e", alpha_concentration)),
              position = position_dodge(width = 0.7), vjust = -0.3, size = 2.5) +
    labs(
      x = "Aggregation Method",
      y = "TPR @ FPR = 0.05",
      fill = "Warm method",
      title = "Best Alpha per Method: TPR @ FPR = 0.05",
      subtitle = "Labels show the best alpha_concentration for each bar"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "head_to_head_best_alpha.png"),
         p_h2h, width = 9, height = 5, dpi = 150)
  cat("Saved:", file.path(fig_dir, "head_to_head_best_alpha.png"), "\n")
}
```

# 8. Multimodal Metrics by Method

```{r load-multimodal}
mm_path <- file.path(agg_dir, "multimodal_metrics.csv")
if (file.exists(mm_path)) {
  mm <- readr::read_csv(mm_path, show_col_types = FALSE)

  # Parse warm_method and alpha from group_label
  mm <- mm %>%
    mutate(
      warm_method = stringr::str_extract(group_label, "(?<=wm=)[a-z_]+"),
      alpha_concentration = as.numeric(
        stringr::str_extract(group_label, "(?<=alpha=)[0-9.e-]+")
      )
    )

  cat("Multimodal metrics rows:", nrow(mm), "\n")
  cat("Warm methods:", sort(unique(mm$warm_method)), "\n")
  cat("Alpha values:", sort(unique(mm$alpha_concentration)), "\n")
} else {
  mm <- tibble()
  cat("No multimodal_metrics.csv found.\n")
}
```

```{r multimodal-summary}
if (nrow(mm) > 0) {
  cat("=== Mean JSD by warm_method and alpha ===\n")
  mm %>%
    filter(!is.na(alpha_concentration)) %>%
    group_by(warm_method, alpha_concentration) %>%
    summarise(
      n = n(),
      mean_jsd = mean(mean_jsd, na.rm = TRUE),
      mean_max_jsd = mean(max_jsd, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(warm_method, alpha_concentration) %>%
    print(n = Inf)
}
```

```{r plot-multimodal-jsd-by-method}
if (nrow(mm) > 0 && any(!is.na(mm$warm_method))) {
  mm_jsd <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, mean_jsd, max_jsd) %>%
    tidyr::pivot_longer(
      cols = c(mean_jsd, max_jsd),
      names_to = "metric",
      values_to = "value"
    ) %>%
    mutate(metric = dplyr::recode(metric,
      mean_jsd = "Mean pairwise JSD",
      max_jsd = "Max pairwise JSD"
    ))

  p_jsd <- ggplot(mm_jsd, aes(x = factor(alpha_concentration), y = value,
                               fill = warm_method)) +
    geom_boxplot(outlier.size = 0.5) +
    facet_wrap(~metric, scales = "free_y", ncol = 1) +
    labs(x = "Dirichlet alpha", fill = "Warm method",
         title = "PIP Diversity (JSD) by Method and Alpha") +
    theme(axis.text.x = element_text(size = 9))

  ggsave(file.path(fig_dir, "multimodal_jsd_by_method.png"),
         p_jsd, width = 10, height = 7, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_jsd_by_method.png"), "\n")
}
```

```{r plot-multimodal-agnostic-by-method}
if (nrow(mm) > 0 && any(!is.na(mm$warm_method))) {
  mm_agnostic <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, jaccard_top10, mean_pip_var) %>%
    tidyr::pivot_longer(
      cols = c(jaccard_top10, mean_pip_var),
      names_to = "metric",
      values_to = "value"
    )

  p_agnostic <- ggplot(mm_agnostic,
                        aes(x = factor(alpha_concentration), y = value,
                            fill = warm_method)) +
    geom_boxplot(outlier.size = 0.5) +
    facet_wrap(~metric, scales = "free_y", ncol = 2) +
    labs(x = "Dirichlet alpha", fill = "Warm method",
         title = "JSD-Agnostic Multimodal Metrics by Method and Alpha") +
    theme(axis.text.x = element_text(size = 9))

  ggsave(file.path(fig_dir, "multimodal_agnostic_by_method.png"),
         p_agnostic, width = 10, height = 5, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_agnostic_by_method.png"), "\n")
}
```

```{r plot-multimodal-n-clusters}
if (nrow(mm) > 0) {
  jsd_clust_cols <- grep("^n_clusters_jsd_", names(mm), value = TRUE)
  bjsd_clust_cols <- grep("^n_clusters_bjsd_", names(mm), value = TRUE)

  mm_clust_cat <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, all_of(jsd_clust_cols)) %>%
    tidyr::pivot_longer(
      cols = -c(warm_method, alpha_concentration),
      names_to = "col_name",
      values_to = "n_clusters"
    ) %>%
    mutate(
      threshold = dplyr::case_match(col_name,
        "n_clusters_jsd_015" ~ 0.15,
        "n_clusters_jsd_050" ~ 0.50,
        "n_clusters_jsd_100" ~ 1.00,
        "n_clusters_jsd_200" ~ 2.00,
        "n_clusters_jsd_300" ~ 3.00,
        "n_clusters_jsd_500" ~ 5.00
      ),
      jsd_type = "Categorical"
    )

  mm_clust_bern <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, all_of(bjsd_clust_cols)) %>%
    tidyr::pivot_longer(
      cols = -c(warm_method, alpha_concentration),
      names_to = "col_name",
      values_to = "n_clusters"
    ) %>%
    mutate(
      threshold = dplyr::case_match(col_name,
        "n_clusters_bjsd_0001" ~ 0.001,
        "n_clusters_bjsd_0002" ~ 0.002,
        "n_clusters_bjsd_0004" ~ 0.004,
        "n_clusters_bjsd_0006" ~ 0.006,
        "n_clusters_bjsd_0008" ~ 0.008,
        "n_clusters_bjsd_000" ~ 0.001,
        "n_clusters_bjsd_001" ~ 0.01,
        "n_clusters_bjsd_005" ~ 0.05,
        "n_clusters_bjsd_010" ~ 0.10,
        "n_clusters_bjsd_020" ~ 0.20,
        "n_clusters_bjsd_040" ~ 0.40
      ),
      jsd_type = "Bernoulli"
    )

  mm_clust <- bind_rows(mm_clust_cat, mm_clust_bern) %>%
    filter(!is.na(threshold)) %>%
    mutate(threshold_label = sprintf("h = %s", threshold))

  p_clust <- ggplot(mm_clust,
                     aes(x = factor(alpha_concentration), y = n_clusters,
                         fill = warm_method)) +
    geom_boxplot(outlier.size = 0.5) +
    facet_grid(jsd_type ~ threshold_label, scales = "free") +
    labs(x = "Dirichlet alpha", y = "Number of clusters",
         fill = "Warm method",
         title = "Cluster Count by Method, JSD Type, and Threshold") +
    theme(axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
          strip.text = element_text(size = 8))

  ggsave(file.path(fig_dir, "multimodal_n_clusters_by_method.png"),
         p_clust, width = 16, height = 7, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_n_clusters_by_method.png"), "\n")
}
```
