---
title: "Dirichlet Alpha Pilot v2 — Collect & Visualize"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Collects staging outputs from the Dirichlet alpha pilot v2,
then computes TPR @ FPR and TPR @ FDR by alpha value across all
datasets and aggregation methods. Includes both categorical and
Bernoulli JSD multimodal metrics.

```{r setup, message=FALSE}
here::i_am(
  "vignettes/simulation pipeline/collect_results_workbook_dirichlet_pilot.Rmd"
)
library(here)
here()

library(devtools)
devtools::load_all(".")

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(stringr)
theme_set(theme_bw(base_size = 13))

```

# 1. Job Settings

```{r settings}
job_name <- "dirichlet_alpha_pilot_v2"
parent_job_id <- "FILL_IN"
output_root <- here("output")

agg_dir <- file.path(
  output_root, "slurm_output", job_name, parent_job_id, "aggregated"
)
fig_dir <- file.path(agg_dir, "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Aggregate Staging Outputs

```{r aggregate}

if (!dir.exists(agg_dir) || !file.exists(file.path(agg_dir, "confusion_bins.csv"))) {
  aggregate_staging_outputs(
    job_name = job_name,
    parent_job_id = parent_job_id,
    output_root = output_root,
    validate = TRUE
  )
}

stopifnot(file.exists(file.path(agg_dir, "confusion_bins.csv")))
cat("Aggregated outputs found at:", agg_dir, "\n")
```

# 3. Load Data

```{r load-data}
conf <- readr::read_csv(
  file.path(agg_dir, "confusion_bins.csv"),
  show_col_types = FALSE
)

stopifnot("alpha_concentration" %in% names(conf))
cat("Confusion bins rows:", nrow(conf), "\n")
cat("Alpha values:", sort(unique(conf$alpha_concentration)), "\n")
```

# 4. TPR @ FPR (matching EB ablation convention)

Uses `compute_power_fdr_curves()` which produces both `fpr` and `fdr`
columns. FPR = FP / total_noncausal (lenient). FDR = FP / (TP + FP)
(strict).

```{r baseline-reference}
# Compute baseline (default uniform prior) TPR @ FPR and FDR
# Default runs are identical across alpha values — take any one
baseline_alpha <- sort(unique(conf$alpha_concentration))[1L]
conf_baseline <- conf %>%
  filter(
    (is.na(agg_method) | agg_method == ""),
    run_type == "default",
    alpha_concentration == baseline_alpha
  )

curves_baseline <- compute_power_fdr_curves(conf_baseline)

# TPR @ FPR thresholds
baseline_tpr_fpr005 <- {
  idx <- which(curves_baseline$fpr >= 0.05)[1L]
  if (is.na(idx)) 0 else curves_baseline$power[idx]
}
baseline_tpr_fpr010 <- {
  idx <- which(curves_baseline$fpr >= 0.10)[1L]
  if (is.na(idx)) 0 else curves_baseline$power[idx]
}

# TPR @ FDR thresholds (interpolated)
baseline_tpr_fdr005 <- {
  df <- arrange(distinct(curves_baseline, fdr, power), fdr)
  val <- stats::approx(df$fdr, df$power, xout = 0.05,
                       ties = "ordered", rule = 2)$y
  if (is.na(val)) 0 else val
}
baseline_tpr_fdr010 <- {
  df <- arrange(distinct(curves_baseline, fdr, power), fdr)
  val <- stats::approx(df$fdr, df$power, xout = 0.10,
                       ties = "ordered", rule = 2)$y
  if (is.na(val)) 0 else val
}

cat(sprintf("Baseline (uniform prior) reference:\n"))
cat(sprintf("  TPR @ FPR=0.05: %.4f\n", baseline_tpr_fpr005))
cat(sprintf("  TPR @ FPR=0.10: %.4f\n", baseline_tpr_fpr010))
cat(sprintf("  TPR @ FDR=0.05: %.4f\n", baseline_tpr_fdr005))
cat(sprintf("  TPR @ FDR=0.10: %.4f\n", baseline_tpr_fdr010))
```

```{r tpr-at-fpr-individual}
# Individual fits only (no aggregation)
conf_fits <- conf %>%
  filter(is.na(agg_method) | agg_method == "")

curves_individual <- compute_power_fdr_curves(
  conf_fits,
  group_vars = "alpha_concentration"
)

# TPR @ FPR = 0.05 (EB ablation convention)
tpr_at_fpr_individual <- curves_individual %>%
  group_by(alpha_concentration) %>%
  summarise(
    tpr_fpr005 = {
      idx <- which(fpr >= 0.05)[1L]
      if (is.na(idx)) NA_real_ else power[idx]
    },
    tpr_fpr010 = {
      idx <- which(fpr >= 0.10)[1L]
      if (is.na(idx)) NA_real_ else power[idx]
    },
    .groups = "drop"
  )

cat("Individual fits — TPR @ FPR thresholds:\n")
print(as.data.frame(tpr_at_fpr_individual))
```

```{r tpr-at-fpr-aggregated}
# Aggregated PIPs
conf_agg <- conf %>%
  filter(!is.na(agg_method) & agg_method != "")

if (nrow(conf_agg) > 0) {
  curves_agg <- compute_power_fdr_curves(
    conf_agg,
    group_vars = c("alpha_concentration", "agg_method")
  )

  tpr_at_fpr_agg <- curves_agg %>%
    group_by(alpha_concentration, agg_method) %>%
    summarise(
      tpr_fpr005 = {
        idx <- which(fpr >= 0.05)[1L]
        if (is.na(idx)) NA_real_ else power[idx]
      },
      tpr_fpr010 = {
        idx <- which(fpr >= 0.10)[1L]
        if (is.na(idx)) NA_real_ else power[idx]
      },
      .groups = "drop"
    )

  # Fill missing combos with NA
  all_combos <- tidyr::expand_grid(
    alpha_concentration = unique(curves_agg$alpha_concentration),
    agg_method = unique(curves_agg$agg_method)
  )
  tpr_at_fpr_agg <- all_combos %>%
    left_join(tpr_at_fpr_agg, by = c("alpha_concentration", "agg_method")) %>%
    mutate(across(starts_with("tpr_"), ~coalesce(., 0)))

  cat("Aggregated — TPR @ FPR thresholds:\n")
  print(as.data.frame(tpr_at_fpr_agg))
} else {
  tpr_at_fpr_agg <- tibble()
}
```

# 5. TPR @ FDR (stricter criterion)

```{r tpr-at-fdr-aggregated}
if (nrow(conf_agg) > 0) {
  # Use interpolation to handle bin-resolution issues
  tpr_at_fdr_agg <- curves_agg %>%
    group_by(alpha_concentration, agg_method) %>%
    summarise(
      tpr_fdr005 = {
        df <- arrange(distinct(cur_data_all(), fdr, power), fdr)
        val <- stats::approx(df$fdr, df$power, xout = 0.05,
                             ties = "ordered", rule = 2)$y
        if (is.na(val)) 0 else val
      },
      tpr_fdr010 = {
        df <- arrange(distinct(cur_data_all(), fdr, power), fdr)
        val <- stats::approx(df$fdr, df$power, xout = 0.10,
                             ties = "ordered", rule = 2)$y
        if (is.na(val)) 0 else val
      },
      .groups = "drop"
    )

  all_combos <- tidyr::expand_grid(
    alpha_concentration = unique(curves_agg$alpha_concentration),
    agg_method = unique(curves_agg$agg_method)
  )
  tpr_at_fdr_agg <- all_combos %>%
    left_join(tpr_at_fdr_agg, by = c("alpha_concentration", "agg_method")) %>%
    mutate(across(starts_with("tpr_"), ~coalesce(., 0)))

  cat("Aggregated — TPR @ FDR thresholds (interpolated):\n")
  print(as.data.frame(tpr_at_fdr_agg))
}
```

# 6. Plots — TPR @ FPR

```{r plot-fpr-individual}
p1 <- ggplot(tpr_at_fpr_individual,
             aes(x = factor(alpha_concentration), y = tpr_fpr005)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
             color = "red", linewidth = 0.7) +
  annotate("text", x = Inf, y = baseline_tpr_fpr005,
           label = sprintf("baseline = %.3f", baseline_tpr_fpr005),
           hjust = 1.1, vjust = -0.5, color = "red", size = 3) +
  geom_text(aes(label = sprintf("%.3f", tpr_fpr005)), vjust = -0.5, size = 3.5) +
  labs(
    x = "Dirichlet alpha",
    y = "TPR @ FPR = 0.05",
    title = "Individual Fits: TPR @ FPR = 0.05 by Alpha",
    subtitle = "Pooled across all datasets and restarts"
  ) +
  ylim(0, NA)

ggsave(file.path(fig_dir, "tpr_fpr005_individual.png"),
       p1, width = 7, height = 4, dpi = 150)
cat("Saved:", file.path(fig_dir, "tpr_fpr005_individual.png"), "\n")
```

```{r plot-fpr-aggregated}
if (nrow(tpr_at_fpr_agg) > 0) {
  p2 <- ggplot(
    tpr_at_fpr_agg,
    aes(x = factor(alpha_concentration), y = tpr_fpr005, fill = agg_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    annotate("text", x = Inf, y = baseline_tpr_fpr005,
             label = sprintf("baseline = %.3f", baseline_tpr_fpr005),
             hjust = 1.1, vjust = -0.5, color = "red", size = 3) +
    geom_text(
      aes(label = sprintf("%.3f", tpr_fpr005)),
      position = position_dodge(width = 0.7),
      vjust = -0.5, size = 2.5
    ) +
    labs(
      x = "Dirichlet alpha",
      y = "TPR @ FPR = 0.05",
      fill = "Aggregation",
      title = "Aggregated PIPs: TPR @ FPR = 0.05 by Alpha",
      subtitle = "Pooled across all datasets"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "tpr_fpr005_aggregated.png"),
         p2, width = 9, height = 5, dpi = 150)
  cat("Saved:", file.path(fig_dir, "tpr_fpr005_aggregated.png"), "\n")
}
```

```{r plot-fdr-aggregated}
if (nrow(tpr_at_fdr_agg) > 0) {
  p2b <- ggplot(
    tpr_at_fdr_agg,
    aes(x = factor(alpha_concentration), y = tpr_fdr010, fill = agg_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fdr010, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    annotate("text", x = Inf, y = baseline_tpr_fdr010,
             label = sprintf("baseline = %.3f", baseline_tpr_fdr010),
             hjust = 1.1, vjust = -0.5, color = "red", size = 3) +
    geom_text(
      aes(label = sprintf("%.3f", tpr_fdr010)),
      position = position_dodge(width = 0.7),
      vjust = -0.5, size = 2.5
    ) +
    labs(
      x = "Dirichlet alpha",
      y = "TPR @ FDR = 0.10",
      fill = "Aggregation",
      title = "Aggregated PIPs: TPR @ FDR = 0.10 by Alpha (interpolated)",
      subtitle = "Pooled across all datasets"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "tpr_fdr010_aggregated.png"),
         p2b, width = 9, height = 5, dpi = 150)
  cat("Saved:", file.path(fig_dir, "tpr_fdr010_aggregated.png"), "\n")
}
```

# 7. Multimodal Metrics by Alpha

```{r load-multimodal}
mm_path <- file.path(agg_dir, "multimodal_metrics.csv")
if (file.exists(mm_path)) {
  mm <- readr::read_csv(mm_path, show_col_types = FALSE)

  # Parse alpha from group_label (e.g., "model_grid|...|alpha=0.01")
  mm <- mm %>%
    mutate(
      alpha_concentration = as.numeric(
        stringr::str_extract(group_label, "(?<=alpha=)[0-9.]+")
      )
    )

  cat("Multimodal metrics rows:", nrow(mm), "\n")
  cat("Alpha values:", sort(unique(mm$alpha_concentration)), "\n")
} else {
  mm <- tibble()
  cat("No multimodal_metrics.csv found.\n")
}
```

```{r multimodal-summary}
if (nrow(mm) > 0) {
  # Categorical JSD summary
  cat("=== Categorical JSD (unnormalized PIP vector) ===\n")
  mm %>%
    filter(!is.na(alpha_concentration)) %>%
    group_by(alpha_concentration) %>%
    summarise(
      n = n(),
      mean_jsd = mean(mean_jsd, na.rm = TRUE),
      mean_max_jsd = mean(max_jsd, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    print()

  # Bernoulli JSD summary (if present)
  if ("mean_bjsd" %in% names(mm)) {
    cat("\n=== Bernoulli JSD (per-variant, bounded [0, 0.693]) ===\n")
    mm %>%
      filter(!is.na(alpha_concentration)) %>%
      group_by(alpha_concentration) %>%
      summarise(
        n = n(),
        mean_bjsd = mean(mean_bjsd, na.rm = TRUE),
        mean_max_bjsd = mean(max_bjsd, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      print()
  }
}
```

```{r plot-multimodal-agnostic}
# Plot 1: JSD-agnostic metrics (jaccard, pip_var)
if (nrow(mm) > 0 && any(!is.na(mm$alpha_concentration))) {
  mm_agnostic <- mm %>%
    filter(!is.na(alpha_concentration)) %>%
    select(alpha_concentration, jaccard_top10, mean_pip_var) %>%
    tidyr::pivot_longer(
      cols = -alpha_concentration,
      names_to = "metric",
      values_to = "value"
    )

  p_agnostic <- ggplot(mm_agnostic, aes(x = factor(alpha_concentration), y = value)) +
    geom_boxplot(fill = "lightblue", outlier.size = 0.5) +
    facet_wrap(~metric, scales = "free_y", ncol = 2) +
    labs(x = "Dirichlet alpha",
         title = "JSD-Agnostic Multimodal Metrics by Alpha") +
    theme(axis.text.x = element_text(size = 9))

  ggsave(file.path(fig_dir, "multimodal_agnostic.png"),
         p_agnostic, width = 8, height = 4, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_agnostic.png"), "\n")
}
```

```{r plot-multimodal-mean-jsd}
# Plot 2: Mean JSD — categorical vs Bernoulli (2 rows)
if (nrow(mm) > 0 && "mean_bjsd" %in% names(mm)) {
  mm_mean <- mm %>%
    filter(!is.na(alpha_concentration)) %>%
    select(alpha_concentration, mean_jsd, mean_bjsd) %>%
    tidyr::pivot_longer(
      cols = c(mean_jsd, mean_bjsd),
      names_to = "jsd_type",
      values_to = "value"
    ) %>%
    mutate(jsd_type = dplyr::recode(jsd_type,
      mean_jsd = "Categorical (unnormalized)",
      mean_bjsd = "Bernoulli (bounded [0, 0.693])"
    ))

  p_mean <- ggplot(mm_mean, aes(x = factor(alpha_concentration), y = value)) +
    geom_boxplot(fill = "lightblue", outlier.size = 0.5) +
    facet_wrap(~jsd_type, scales = "free_y", ncol = 1) +
    labs(x = "Dirichlet alpha", y = "Mean pairwise JSD",
         title = "Mean JSD by Alpha") +
    theme(axis.text.x = element_text(size = 9))

  ggsave(file.path(fig_dir, "multimodal_mean_jsd.png"),
         p_mean, width = 7, height = 6, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_mean_jsd.png"), "\n")
}
```

```{r plot-multimodal-max-jsd}
# Plot 3: Max JSD — categorical vs Bernoulli (2 rows)
if (nrow(mm) > 0 && "max_bjsd" %in% names(mm)) {
  mm_max <- mm %>%
    filter(!is.na(alpha_concentration)) %>%
    select(alpha_concentration, max_jsd, max_bjsd) %>%
    tidyr::pivot_longer(
      cols = c(max_jsd, max_bjsd),
      names_to = "jsd_type",
      values_to = "value"
    ) %>%
    mutate(jsd_type = dplyr::recode(jsd_type,
      max_jsd = "Categorical (unnormalized)",
      max_bjsd = "Bernoulli (bounded [0, 0.693])"
    ))

  p_max <- ggplot(mm_max, aes(x = factor(alpha_concentration), y = value)) +
    geom_boxplot(fill = "lightyellow", outlier.size = 0.5) +
    facet_wrap(~jsd_type, scales = "free_y", ncol = 1) +
    labs(x = "Dirichlet alpha", y = "Max pairwise JSD",
         title = "Max JSD by Alpha") +
    theme(axis.text.x = element_text(size = 9))

  ggsave(file.path(fig_dir, "multimodal_max_jsd.png"),
         p_max, width = 7, height = 6, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_max_jsd.png"), "\n")
}
```

```{r plot-multimodal-n-clusters}
# Plot 4: n_clusters — rows = JSD type, columns = threshold
if (nrow(mm) > 0) {
  # Gather categorical JSD cluster columns
  jsd_clust_cols <- grep("^n_clusters_jsd_", names(mm), value = TRUE)
  bjsd_clust_cols <- grep("^n_clusters_bjsd_", names(mm), value = TRUE)

  mm_clust_cat <- mm %>%
    filter(!is.na(alpha_concentration)) %>%
    select(alpha_concentration, all_of(jsd_clust_cols)) %>%
    tidyr::pivot_longer(
      cols = -alpha_concentration,
      names_to = "col_name",
      values_to = "n_clusters"
    ) %>%
    mutate(
      threshold = as.numeric(sub("n_clusters_jsd_", "", col_name)) / 100,
      # Fix: 015 -> 0.15, 050 -> 0.50, 100 -> 1.00, etc.
      threshold = dplyr::case_match(col_name,
        "n_clusters_jsd_015" ~ 0.15,
        "n_clusters_jsd_050" ~ 0.50,
        "n_clusters_jsd_100" ~ 1.00,
        "n_clusters_jsd_200" ~ 2.00,
        "n_clusters_jsd_300" ~ 3.00,
        "n_clusters_jsd_500" ~ 5.00
      ),
      jsd_type = "Categorical"
    )

  mm_clust_bern <- mm %>%
    filter(!is.na(alpha_concentration)) %>%
    select(alpha_concentration, all_of(bjsd_clust_cols)) %>%
    tidyr::pivot_longer(
      cols = -alpha_concentration,
      names_to = "col_name",
      values_to = "n_clusters"
    ) %>%
    mutate(
      threshold = dplyr::case_match(col_name,
        "n_clusters_bjsd_001" ~ 0.01,
        "n_clusters_bjsd_005" ~ 0.05,
        "n_clusters_bjsd_010" ~ 0.10,
        "n_clusters_bjsd_020" ~ 0.20,
        "n_clusters_bjsd_040" ~ 0.40
      ),
      jsd_type = "Bernoulli"
    )

  mm_clust <- bind_rows(mm_clust_cat, mm_clust_bern) %>%
    mutate(threshold_label = sprintf("h = %s", threshold))

  p_clust <- ggplot(mm_clust, aes(x = factor(alpha_concentration), y = n_clusters)) +
    geom_boxplot(fill = "lavender", outlier.size = 0.5) +
    facet_grid(jsd_type ~ threshold_label, scales = "free") +
    labs(x = "Dirichlet alpha", y = "Number of clusters",
         title = "Cluster Count by JSD Type and Threshold") +
    theme(axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
          strip.text = element_text(size = 8))

  ggsave(file.path(fig_dir, "multimodal_n_clusters.png"),
         p_clust, width = 14, height = 6, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_n_clusters.png"), "\n")
}
```
