---
title: "SuSiNE Pilot Results Collection"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Aggregates pilot outputs for the current refactored harness
workflow. Uses staged task flush files, writes job-level aggregated outputs,
builds confusion matrices, and runs metrics/seed quality checks.

# Section 1: Setup (Always Run First)

Load libraries, select the pilot job, and resolve paths.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_pilot.Rmd")
library(here)
here()

library(devtools)
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
library(purrr)
library(fs)
library(arrow)
```

# 1.1 Pilot Job Inputs

```{r job-inputs}
# Prefilled pilot defaults: all-use-case, all-exploration pilot at K=12/L=10.
# Set parent_job_id to the SLURM parent array id for your run.
job_name <- "pilot_all_methods_k12_l10"
parent_job_id <- "REPLACE_WITH_PARENT_JOB_ID"
output_root <- here("output")

skip_snps_dataset_aggregation <- FALSE
require_multimodal_metrics <- TRUE
run_quality_checks <- TRUE
```

# 1.2 Resolve Paths and Load Tables

```{r load-paths-and-tables}
paths <- test_susine::resolve_job_paths(
  job_name = job_name,
  parent_job_id = parent_job_id,
  output_root = output_root
)

if (!dir.exists(paths$run_history_dir)) {
  stop("Run history directory not found: ", paths$run_history_dir)
}

run_table <- readr::read_csv(
  file.path(paths$run_history_dir, "run_table.csv"),
  show_col_types = FALSE
)

dataset_bundles_path <- file.path(paths$run_history_dir, "dataset_bundles.csv")
if (!file.exists(dataset_bundles_path)) {
  dataset_bundles_path <- file.path(output_root, "temp", job_name, "dataset_bundles.csv")
}
dataset_bundles <- readr::read_csv(dataset_bundles_path, show_col_types = FALSE)
use_cases_path <- file.path(paths$run_history_dir, "use_cases.csv")

cat("Job:", job_name, "/", parent_job_id, "\n")
cat("Run history dir:", paths$run_history_dir, "\n")
cat("Staging dir:", paths$staging_dir, "\n")
cat("Aggregated dir:", paths$aggregated_dir, "\n")
cat("Loaded run_table with", nrow(run_table), "rows\n")
cat("Loaded dataset_bundles with", nrow(dataset_bundles), "rows\n")
```

---

# Section 2: Staging to Aggregated

Auto-skips aggregation if outputs already exist.

```{r check-aggregation-status}
required_files <- c(
  "model_metrics.csv",
  "effect_metrics.csv",
  "validation.csv",
  "confusion_bins.csv",
  "dataset_metrics.csv"
)
if (isTRUE(require_multimodal_metrics)) {
  required_files <- c(required_files, "multimodal_metrics.csv")
}
required_exist <- all(file.exists(file.path(paths$aggregated_dir, required_files)))

snps_dataset_exists <- dir.exists(paths$snps_dataset_dir) &&
  length(list.files(paths$snps_dataset_dir, recursive = TRUE)) > 0

staging_exists <- dir.exists(paths$staging_dir)
run_aggregation <- staging_exists && (!required_exist || (!skip_snps_dataset_aggregation && !snps_dataset_exists))

cat("Task output directory exists:", staging_exists, "\n")
cat("Required aggregated files exist:", required_exist, "\n")
cat("Aggregated SNPs dataset exists:", snps_dataset_exists, "\n")
cat("Will run aggregation:", run_aggregation, "\n")
```

## 2.1 Index and Validate Staging Files

```{r index-and-validate-staging}
if (run_aggregation) {
  staging_index <- test_susine::index_staging_outputs(
    job_name = job_name,
    parent_job_id = parent_job_id,
    output_root = output_root
  )

  cat("Indexed", nrow(staging_index), "staging files across",
      dplyr::n_distinct(staging_index$task_id), "tasks\n")

  validation <- test_susine::validate_staging_outputs(staging_index)
  bad <- dplyr::filter(validation, !ok)
  cat("Unreadable staging files:", nrow(bad), "\n")

  if (nrow(bad) > 0) {
    print(bad)
    stop("Fix unreadable staging files before aggregation.")
  }
} else {
  message("Skipping staging validation because aggregation is not needed.")
}
```

## 2.2 Aggregate Job Outputs

```{r aggregate-staging}
if (run_aggregation) {
  agg <- test_susine::aggregate_staging_outputs(
    job_name = job_name,
    parent_job_id = parent_job_id,
    output_root = output_root,
    validate = FALSE,
    output_dir = paths$aggregated_dir,
    write_snps_dataset = !skip_snps_dataset_aggregation
  )

  cat("Aggregated outputs written to:", agg$output_dir, "\n")
  print(sort(list.files(agg$output_dir, recursive = TRUE)))
} else {
  message("Using existing aggregated outputs.")
}
```

---

# Section 3: Build Confusion Matrices

```{r confusion-config}
pip_bucket_width <- 0.01
detailed_grouping_vars <- c(
  "matrix_id", "use_case_id", "variant_type", "agg_method",
  "exploration_mode", "exploration_methods", "exploration_group",
  "p_star", "L", "y_noise", "architecture",
  "annotation_r2", "inflate_match", "sigma_0_2_scalar",
  "c_value", "tau_value", "init_type", "refine_step"
)

dataset_grouping_vars <- c(
  "matrix_id", "use_case_id", "variant_type", "agg_method"
)
```

```{r confusion-input-discovery}
confusion_bins_path <- file.path(paths$aggregated_dir, "confusion_bins.csv")
confusion_bins_exists <- file.exists(confusion_bins_path)
staging_confusion_files <- list.files(
  paths$staging_dir,
  pattern = "_confusion_bins\\.csv$",
  recursive = TRUE,
  full.names = TRUE
)

snps_dataset_exists <- dir.exists(paths$snps_dataset_dir) &&
  length(list.files(paths$snps_dataset_dir, recursive = TRUE)) > 0
staging_snp_files <- list.files(
  paths$staging_dir,
  pattern = "_snps\\.parquet$",
  recursive = TRUE,
  full.names = TRUE
)

if (!confusion_bins_exists && !length(staging_confusion_files) &&
    !snps_dataset_exists && !length(staging_snp_files)) {
  stop("No confusion bins or SNP sources found. Run Section 2 first.")
}

cat("confusion_bins.csv exists:", confusion_bins_exists, "\n")
cat("staging confusion files:", length(staging_confusion_files), "\n")
cat("snps_dataset exists:", snps_dataset_exists, "\n")
cat("staging snp fragments:", length(staging_snp_files), "\n")
```

```{r build-confusion-detailed, message=FALSE}
detailed_output_path <- file.path(paths$aggregated_dir, "confusion_matrix_detailed.csv")
use_confusion_bins <- confusion_bins_exists || length(staging_confusion_files) > 0
stream_confusion <- !use_confusion_bins && !snps_dataset_exists

confusion_detailed <- test_susine::build_confusion_matrix(
  confusion_bins_path = if (confusion_bins_exists) confusion_bins_path else NULL,
  confusion_bins_files = if (!confusion_bins_exists && length(staging_confusion_files)) staging_confusion_files else NULL,
  snps_dataset_path = if (!use_confusion_bins && snps_dataset_exists) paths$snps_dataset_dir else NULL,
  snp_files = if (!use_confusion_bins && !snps_dataset_exists) staging_snp_files else NULL,
  run_table = run_table,
  grouping_vars = detailed_grouping_vars,
  pip_bucket_width = pip_bucket_width,
  use_cases_path = use_cases_path,
  output_path = detailed_output_path,
  stream = stream_confusion
)

cat("Detailed confusion rows:", nrow(confusion_detailed), "\n")
```

```{r build-confusion-dataset, message=FALSE}
dataset_output_path <- file.path(paths$aggregated_dir, "confusion_matrix_by_dataset.csv")

confusion_by_dataset <- test_susine::build_confusion_matrix(
  confusion_bins_path = if (confusion_bins_exists) confusion_bins_path else NULL,
  confusion_bins_files = if (!confusion_bins_exists && length(staging_confusion_files)) staging_confusion_files else NULL,
  snps_dataset_path = if (!use_confusion_bins && snps_dataset_exists) paths$snps_dataset_dir else NULL,
  snp_files = if (!use_confusion_bins && !snps_dataset_exists) staging_snp_files else NULL,
  run_table = run_table,
  grouping_vars = dataset_grouping_vars,
  pip_bucket_width = pip_bucket_width,
  use_cases_path = NULL,
  output_path = dataset_output_path,
  stream = stream_confusion
)

cat("Dataset confusion rows:", nrow(confusion_by_dataset), "\n")
cat("Unique matrices:", dplyr::n_distinct(confusion_by_dataset$matrix_id), "\n")
```

---

# Section 4: Quality Checks

```{r quality-checks}
if (isTRUE(run_quality_checks)) {
  metric_check <- test_susine::validate_metrics_coverage(
    aggregated_dir = paths$aggregated_dir,
    require_multimodal = require_multimodal_metrics
  )
  print(metric_check$checks)
  if (!metric_check$ok) {
    stop("Metrics coverage check failed.")
  } else {
    cat("Metrics coverage check passed.\n")
  }

  seed_check <- test_susine::seed_management_report(
    job_config = list(tables = list(
      runs = run_table,
      dataset_bundles = dataset_bundles
    ))
  )
  print(seed_check)
  if (!all(seed_check$pass)) {
    warning("Seed management checks reported failures.")
  } else {
    cat("Seed management checks passed.\n")
  }
} else {
  message("Skipping quality checks.")
}
```

# Section 5: Output Summary

```{r summary}
cat("Aggregated output directory:", paths$aggregated_dir, "\n")
print(sort(list.files(paths$aggregated_dir, recursive = TRUE)))
if (dir.exists(paths$snps_dataset_dir)) {
  n_parts <- length(list.files(paths$snps_dataset_dir, pattern = "\\.parquet$", recursive = TRUE))
  cat("snps_dataset parquet fragments:", n_parts, "\n")
}
```

# Section 6: Optional Cleanup (Manual)

```{r cleanup-staging, eval=FALSE}
# DANGER: enable manually after confirming aggregation output quality.
if (dir.exists(paths$staging_dir)) {
  task_dirs <- list.dirs(paths$staging_dir, recursive = FALSE, full.names = TRUE)
  task_dirs <- task_dirs[grepl("task-", basename(task_dirs))]
  purrr::walk(task_dirs, fs::dir_delete)
  cat("Deleted", length(task_dirs), "task directories from staging.\n")
}
```
