---
title: "Warm Start Pilot v3 — Collect & Visualize"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Collects staging outputs from the warm start pilot v3
(init_alpha + prior_refit), then computes TPR @ FPR and multimodal
metrics split by warm_method and alpha_concentration.

```{r setup, message=FALSE}
here::i_am(
  "vignettes/simulation pipeline/collect_results_workbook_dirichlet_pilot_v3.Rmd"
)
library(here)
here()

library(devtools)
devtools::load_all(".")

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(stringr)
theme_set(theme_bw(base_size = 13))
```

# 1. Job Settings

```{r settings}
job_name <- "warm_start_pilot_dual"
parent_job_id <- "50110509"
output_root <- here("output")

agg_dir <- file.path(
  output_root, "slurm_output", job_name, parent_job_id, "aggregated"
)
fig_dir <- file.path(agg_dir, "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
```

# 2. Aggregate Staging Outputs

```{r aggregate}
if (!dir.exists(agg_dir) || !file.exists(file.path(agg_dir, "confusion_bins.csv"))) {
  aggregate_staging_outputs(
    job_name = job_name,
    parent_job_id = parent_job_id,
    output_root = output_root,
    validate = TRUE
  )
}

stopifnot(file.exists(file.path(agg_dir, "confusion_bins.csv")))
cat("Aggregated outputs found at:", agg_dir, "\n")
```

# 3. Load Data

```{r load-data}
conf <- readr::read_csv(
  file.path(agg_dir, "confusion_bins.csv"),
  show_col_types = FALSE
)

# Parse warm_method from group_key if not already a column
if (!"warm_method" %in% names(conf) && "group_key" %in% names(conf)) {
  conf <- conf %>%
    mutate(warm_method = stringr::str_extract(group_key, "(?<=wm=)[a-z_]+"))
}

stopifnot("alpha_concentration" %in% names(conf))
cat("Confusion bins rows:", nrow(conf), "\n")
cat("Warm methods:", sort(unique(conf$warm_method)), "\n")
cat("Alpha values:", sort(unique(conf$alpha_concentration)), "\n")
```

# 4. Baseline Reference

```{r baseline-reference}
# Default runs are identical across all (warm_method, alpha) groups — take any one
baseline_alpha <- sort(unique(conf$alpha_concentration))[1L]
conf_baseline <- conf %>%
  filter(
    (is.na(agg_method) | agg_method == ""),
    run_type == "default"
  ) %>%
  # Deduplicate: baseline is shared, take first alpha/method combo
  filter(alpha_concentration == baseline_alpha) %>%
  filter(warm_method == first(warm_method) | is.na(warm_method))

curves_baseline <- compute_power_fdr_curves(conf_baseline)

baseline_tpr_fpr005 <- {
  df <- arrange(distinct(curves_baseline, fpr, power), fpr)
  val <- stats::approx(df$fpr, df$power, xout = 0.05, ties = "ordered", rule = 2)$y
  if (is.na(val)) 0 else val
}
baseline_tpr_fpr010 <- {
  df <- arrange(distinct(curves_baseline, fpr, power), fpr)
  val <- stats::approx(df$fpr, df$power, xout = 0.10, ties = "ordered", rule = 2)$y
  if (is.na(val)) 0 else val
}

cat(sprintf("Baseline (uniform prior, no warm start) reference:\n"))
cat(sprintf("  TPR @ FPR=0.05: %.4f\n", baseline_tpr_fpr005))
cat(sprintf("  TPR @ FPR=0.10: %.4f\n", baseline_tpr_fpr010))
```

# 5. TPR @ FPR — Individual Fits by Method

```{r tpr-at-fpr-individual}
conf_fits <- conf %>%
  filter(is.na(agg_method) | agg_method == "")

curves_individual <- compute_power_fdr_curves(
  conf_fits,
  group_vars = c("warm_method", "alpha_concentration")
)

tpr_at_fpr_individual <- curves_individual %>%
  group_by(warm_method, alpha_concentration) %>%
  summarise(
    tpr_fpr005 = {
      df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
      val <- stats::approx(df$fpr, df$power, xout = 0.05, ties = "ordered", rule = 2)$y
      if (is.na(val)) 0 else val
    },
    tpr_fpr010 = {
      df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
      val <- stats::approx(df$fpr, df$power, xout = 0.10, ties = "ordered", rule = 2)$y
      if (is.na(val)) 0 else val
    },
    .groups = "drop"
  )

cat("Individual fits — TPR @ FPR thresholds:\n")
print(as.data.frame(tpr_at_fpr_individual))
```

```{r plot-fpr-individual}
p1 <- ggplot(tpr_at_fpr_individual,
             aes(x = factor(alpha_concentration), y = tpr_fpr005,
                 fill = warm_method)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
             color = "red", linewidth = 0.7) +
  annotate("text", x = Inf, y = baseline_tpr_fpr005,
           label = sprintf("baseline = %.3f", baseline_tpr_fpr005),
           hjust = 1.1, vjust = -0.5, color = "red", size = 3) +
  labs(
    x = "Dirichlet alpha",
    y = "TPR @ FPR = 0.05",
    fill = "Warm method",
    title = "Individual Fits: TPR @ FPR = 0.05 by Method and Alpha",
    subtitle = "Pooled across all datasets and restarts"
  ) +
  ylim(0, NA)

ggsave(file.path(fig_dir, "tpr_fpr005_individual_by_method.png"),
       p1, width = 9, height = 5, dpi = 150)
cat("Saved:", file.path(fig_dir, "tpr_fpr005_individual_by_method.png"), "\n")
```

# 6. TPR @ FPR — Aggregated PIPs by Method

```{r tpr-at-fpr-aggregated}
conf_agg <- conf %>%
  filter(!is.na(agg_method) & agg_method != "")

if (nrow(conf_agg) > 0) {
  curves_agg <- compute_power_fdr_curves(
    conf_agg,
    group_vars = c("warm_method", "alpha_concentration", "agg_method")
  )

  tpr_at_fpr_agg <- curves_agg %>%
    group_by(warm_method, alpha_concentration, agg_method) %>%
    summarise(
      tpr_fpr005 = {
        df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
        val <- stats::approx(df$fpr, df$power, xout = 0.05, ties = "ordered", rule = 2)$y
        if (is.na(val)) 0 else val
      },
      tpr_fpr010 = {
        df <- arrange(distinct(cur_data_all(), fpr, power), fpr)
        val <- stats::approx(df$fpr, df$power, xout = 0.10, ties = "ordered", rule = 2)$y
        if (is.na(val)) 0 else val
      },
      .groups = "drop"
    )

  cat("Aggregated — TPR @ FPR thresholds:\n")
  print(as.data.frame(tpr_at_fpr_agg))
} else {
  tpr_at_fpr_agg <- tibble()
}
```

```{r plot-fpr-aggregated}
if (nrow(tpr_at_fpr_agg) > 0) {
  p2 <- ggplot(
    tpr_at_fpr_agg,
    aes(x = factor(alpha_concentration), y = tpr_fpr005, fill = agg_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    facet_wrap(~warm_method, ncol = 1,
               labeller = labeller(warm_method = c(
                 init_alpha = "init_alpha (trajectory perturbation)",
                 prior_refit = "prior_refit (explore + refit)"
               ))) +
    labs(
      x = "Dirichlet alpha",
      y = "TPR @ FPR = 0.05",
      fill = "Aggregation",
      title = "Aggregated PIPs: TPR @ FPR = 0.05 by Method and Alpha",
      subtitle = "Pooled across all datasets"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "tpr_fpr005_aggregated_by_method.png"),
         p2, width = 10, height = 8, dpi = 150)
  cat("Saved:", file.path(fig_dir, "tpr_fpr005_aggregated_by_method.png"), "\n")
}
```

```{r plot-fpr010-aggregated}
if (nrow(tpr_at_fpr_agg) > 0) {
  p2b <- ggplot(
    tpr_at_fpr_agg,
    aes(x = factor(alpha_concentration), y = tpr_fpr010, fill = agg_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fpr010, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    facet_wrap(~warm_method, ncol = 1,
               labeller = labeller(warm_method = c(
                 init_alpha = "init_alpha (trajectory perturbation)",
                 prior_refit = "prior_refit (explore + refit)"
               ))) +
    labs(
      x = "Dirichlet alpha",
      y = "TPR @ FPR = 0.10",
      fill = "Aggregation",
      title = "Aggregated PIPs: TPR @ FPR = 0.10 by Method and Alpha",
      subtitle = "Pooled across all datasets"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "tpr_fpr010_aggregated_by_method.png"),
         p2b, width = 10, height = 8, dpi = 150)
  cat("Saved:", file.path(fig_dir, "tpr_fpr010_aggregated_by_method.png"), "\n")
}
```

# 7. Precision-Recall Curves

Full precision-recall curves from confusion bins, with baseline overlay.
Shows where on the curve each method differs — far more informative than
a single AUPRC number.

```{r pr-helper}
# Build PR curve from confusion bins, returning one row per threshold
build_pr_curve <- function(confusion_df, label) {
  confusion_df %>%
    arrange(desc(pip_threshold)) %>%
    mutate(
      TP = cumsum(n_causal_at_bucket),
      FP = cumsum(n_noncausal_at_bucket),
      precision = TP / (TP + FP),
      recall = TP / sum(n_causal_at_bucket)
    ) %>%
    filter(!is.na(precision), !is.na(recall)) %>%
    select(pip_threshold, precision, recall) %>%
    mutate(source = label)
}

keep_agg_methods <- c("max_elbo", "elbo_softmax", "uniform", "cluster_weight")
```

```{r pr-baseline}
pr_baseline <- build_pr_curve(conf_baseline, "Baseline (default)")
baseline_auprc <- compute_auprc_single(pr_baseline$precision, pr_baseline$recall)
cat(sprintf("Baseline AUPRC: %.4f\n", baseline_auprc))
```

```{r pr-individual}
# Individual fits: one PR curve per (warm_method, alpha), baseline overlaid
for (wm in sort(unique(conf_fits$warm_method))) {
  conf_wm <- conf_fits %>% filter(warm_method == wm)
  alphas <- sort(unique(conf_wm$alpha_concentration))

  pr_curves <- purrr::map_dfr(alphas, function(a) {
    conf_wm %>%
      filter(alpha_concentration == a) %>%
      build_pr_curve(sprintf("a = %s", a))
  })

  p <- ggplot(pr_curves, aes(x = recall, y = precision, color = source)) +
    geom_line(linewidth = 0.7, alpha = 0.8) +
    geom_line(data = pr_baseline, aes(x = recall, y = precision),
              color = "red", linewidth = 1, linetype = "dashed",
              inherit.aes = FALSE) +
    geom_hline(yintercept = 0.50, linetype = "dotted", color = "grey50",
               linewidth = 0.4) +
    labs(
      x = "Recall (TPR)",
      y = "Precision",
      color = "Alpha",
      title = sprintf("Precision-Recall: Individual Fits (%s)", wm),
      subtitle = "Red dashed = baseline; dotted = 50%% precision"
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
    theme(legend.position = "right")

  fname <- sprintf("pr_curve_individual_%s.png", wm)
  ggsave(file.path(fig_dir, fname), p, width = 9, height = 6, dpi = 150)
  cat("Saved:", file.path(fig_dir, fname), "\n")
}
```

```{r pr-aggregated}
# Aggregated: one plot per (warm_method, alpha), lines colored by agg_method
conf_agg_core <- conf_agg %>% filter(agg_method %in% keep_agg_methods)

if (nrow(conf_agg_core) > 0) {
  for (wm in sort(unique(conf_agg_core$warm_method))) {
    conf_wm <- conf_agg_core %>% filter(warm_method == wm)
    alphas <- sort(unique(conf_wm$alpha_concentration))

    pr_curves <- purrr::map_dfr(alphas, function(a) {
      purrr::map_dfr(keep_agg_methods, function(m) {
        sub <- conf_wm %>% filter(alpha_concentration == a, agg_method == m)
        if (nrow(sub) == 0) return(tibble())
        build_pr_curve(sub, m) %>% mutate(alpha_concentration = a)
      })
    })

    if (nrow(pr_curves) == 0) next

    # Add baseline to each alpha facet
    pr_base_facet <- purrr::map_dfr(alphas, function(a) {
      pr_baseline %>% mutate(alpha_concentration = a)
    })

    p <- ggplot(pr_curves, aes(x = recall, y = precision, color = source)) +
      geom_line(linewidth = 0.7, alpha = 0.8) +
      geom_line(data = pr_base_facet, aes(x = recall, y = precision),
                color = "red", linewidth = 0.8, linetype = "dashed",
                inherit.aes = FALSE) +
      geom_hline(yintercept = 0.50, linetype = "dotted", color = "grey50",
                 linewidth = 0.4) +
      facet_wrap(~alpha_concentration, labeller = label_both) +
      labs(
        x = "Recall (TPR)",
        y = "Precision",
        color = "Aggregation",
        title = sprintf("Precision-Recall: Aggregated PIPs (%s)", wm),
        subtitle = "Red dashed = baseline; dotted = 50%% precision"
      ) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      theme(legend.position = "bottom")

    fname <- sprintf("pr_curve_aggregated_%s.png", wm)
    ggsave(file.path(fig_dir, fname), p, width = 12, height = 10, dpi = 150)
    cat("Saved:", file.path(fig_dir, fname), "\n")
  }
}
```

# 8. AUPRC Summary Bar Plots

```{r auprc-barplots}
# Compute AUPRC per (warm_method, alpha, agg_method) from confusion bins
auprc_agg <- compute_auprc_from_confusion(
  conf_agg %>% filter(agg_method %in% keep_agg_methods),
  group_vars = c("warm_method", "alpha_concentration", "agg_method")
)

p_auprc_bar <- ggplot(
  auprc_agg,
  aes(x = factor(alpha_concentration), y = AUPRC, fill = agg_method)
) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(yintercept = baseline_auprc, linetype = "dashed",
             color = "red", linewidth = 0.7) +
  annotate("text", x = Inf, y = baseline_auprc,
           label = sprintf("baseline = %.3f", baseline_auprc),
           hjust = 1.1, vjust = -0.5, color = "red", size = 3) +
  facet_wrap(~warm_method, ncol = 1, scales = "free_x",
             labeller = labeller(warm_method = c(
               init_alpha = "init_alpha (trajectory perturbation)",
               prior_refit = "prior_refit (explore + refit)"
             ))) +
  labs(
    x = "Dirichlet alpha",
    y = "AUPRC",
    fill = "Aggregation",
    title = "Aggregated PIPs: AUPRC by Warm Method, Alpha, and Aggregation",
    subtitle = "Pooled across all datasets"
  ) +
  ylim(0, NA)

ggsave(file.path(fig_dir, "auprc_barplot_by_method.png"),
       p_auprc_bar, width = 10, height = 8, dpi = 150)
cat("Saved:", file.path(fig_dir, "auprc_barplot_by_method.png"), "\n")
```

# 8b. Bucket-Aware PR / AUPRC Diagnostics

These diagnostics keep the existing figures unchanged and add a second,
more bucket-aware view of PR performance:

- exact PR step curves from the stored confusion bins
- exact step-area AUPRC from the stored confusion bins
- a separate `0.01` re-binned sensitivity check, using a uniform-within-bin
  approximation only for diagnosis

```{r bucket-aware-pr-helpers}
build_pr_curve_step_single <- function(confusion_df, label) {
  curve_df <- confusion_df %>%
    dplyr::group_by(.data$pip_threshold) %>%
    dplyr::summarise(
      n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
      n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::arrange(dplyr::desc(.data$pip_threshold)) %>%
    dplyr::mutate(
      TP = cumsum(.data$n_causal_at_bucket),
      FP = cumsum(.data$n_noncausal_at_bucket),
      total_causal = sum(.data$n_causal_at_bucket),
      precision = dplyr::if_else(.data$TP + .data$FP > 0,
                                 .data$TP / (.data$TP + .data$FP),
                                 NA_real_),
      recall = dplyr::if_else(.data$total_causal > 0,
                              .data$TP / .data$total_causal,
                              NA_real_)
    ) %>%
    dplyr::filter(!is.na(.data$precision), !is.na(.data$recall)) %>%
    dplyr::select(.data$pip_threshold, .data$precision, .data$recall) %>%
    dplyr::arrange(.data$recall, dplyr::desc(.data$pip_threshold))

  if (nrow(curve_df) == 0) {
    return(tibble())
  }

  if (min(curve_df$recall) > 0) {
    curve_df <- dplyr::bind_rows(
      tibble::tibble(pip_threshold = Inf, precision = 1, recall = 0),
      curve_df
    )
  }

  curve_df %>%
    dplyr::mutate(source = label)
}

compute_auprc_step_single <- function(precision, recall) {
  df <- tibble::tibble(precision = precision, recall = recall) %>%
    dplyr::filter(!is.na(.data$precision), !is.na(.data$recall)) %>%
    dplyr::arrange(.data$recall)

  if (nrow(df) == 0) return(NA_real_)

  if (min(df$recall) > 0) {
    df <- dplyr::bind_rows(
      tibble::tibble(recall = 0, precision = 1),
      df
    )
  }

  # Average-precision style step area: precision is held constant between
  # recall jumps at each stored threshold.
  sum(diff(df$recall) * utils::tail(df$precision, -1))
}

compute_auprc_step_from_confusion <- function(confusion_df, group_vars) {
  curves <- compute_power_fdr_curves(confusion_df, group_vars = group_vars)

  curves %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::summarise(
      AUPRC_step = compute_auprc_step_single(.data$precision, .data$recall),
      .groups = "drop"
    )
}

rebin_confusion_uniform <- function(confusion_df, group_vars,
                                    target_width = 0.01) {
  stopifnot(target_width > 0, target_width <= 1)

  target_breaks <- seq(0, 1, by = target_width)
  target_starts <- round(target_breaks[target_breaks < 1], 10)

  aggregated <- confusion_df %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(c(group_vars, "pip_threshold")))) %>%
    dplyr::summarise(
      n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
      n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    )

  if (nrow(aggregated) == 0) {
    return(aggregated)
  }

  if (length(group_vars) > 0) {
    grouped <- aggregated %>%
      dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
      tidyr::nest()
  } else {
    grouped <- tibble::tibble(data = list(aggregated))
  }

  rebinned <- lapply(seq_len(nrow(grouped)), function(i) {
    source_df <- grouped$data[[i]] %>%
      dplyr::arrange(.data$pip_threshold)
    upper_edge <- dplyr::lead(source_df$pip_threshold, default = 1)

    pieces <- lapply(seq_len(nrow(source_df)), function(j) {
      lower <- source_df$pip_threshold[j]
      upper <- upper_edge[j]
      n_causal <- source_df$n_causal_at_bucket[j]
      n_noncausal <- source_df$n_noncausal_at_bucket[j]

      if (!is.finite(lower) || !is.finite(upper)) {
        return(NULL)
      }

      if (upper > lower) {
        overlap <- pmax(
          0,
          pmin(target_starts + target_width, upper) -
            pmax(target_starts, lower)
        )
        keep <- overlap > 0
        if (!any(keep)) {
          return(NULL)
        }
        weights <- overlap[keep] / sum(overlap[keep])
        tibble::tibble(
          pip_threshold = target_starts[keep],
          n_causal_at_bucket = n_causal * weights,
          n_noncausal_at_bucket = n_noncausal * weights
        )
      } else {
        target_bin <- if (lower >= 1) {
          1 - target_width
        } else {
          floor(lower / target_width) * target_width
        }
        tibble::tibble(
          pip_threshold = round(target_bin, 10),
          n_causal_at_bucket = n_causal,
          n_noncausal_at_bucket = n_noncausal
        )
      }
    })

    piece_df <- dplyr::bind_rows(pieces) %>%
      dplyr::group_by(.data$pip_threshold) %>%
      dplyr::summarise(
        n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
        n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
        .groups = "drop"
      )

    if (length(group_vars) > 0) {
      dplyr::bind_cols(grouped[i, group_vars, drop = FALSE], piece_df)
    } else {
      piece_df
    }
  })

  dplyr::bind_rows(rebinned)
}
```

```{r pr-aggregated-step}
pr_baseline_step <- build_pr_curve_step_single(conf_baseline, "Baseline (default)")
baseline_auprc_step <- compute_auprc_step_single(
  pr_baseline_step$precision,
  pr_baseline_step$recall
)
cat(sprintf("Baseline AUPRC (stored bins, step): %.4f\n", baseline_auprc_step))

if (nrow(conf_agg_core) > 0) {
  for (wm in sort(unique(conf_agg_core$warm_method))) {
    conf_wm <- conf_agg_core %>% filter(warm_method == wm)
    alphas <- sort(unique(conf_wm$alpha_concentration))

    pr_curves_step <- purrr::map_dfr(alphas, function(a) {
      purrr::map_dfr(keep_agg_methods, function(m) {
        sub <- conf_wm %>% filter(alpha_concentration == a, agg_method == m)
        if (nrow(sub) == 0) return(tibble())
        build_pr_curve_step_single(sub, m) %>%
          mutate(alpha_concentration = a)
      })
    })

    if (nrow(pr_curves_step) == 0) next

    pr_base_facet_step <- purrr::map_dfr(alphas, function(a) {
      pr_baseline_step %>% mutate(alpha_concentration = a)
    })

    p_step <- ggplot(pr_curves_step, aes(x = recall, y = precision, color = source)) +
      geom_step(direction = "vh", linewidth = 0.7, alpha = 0.85) +
      geom_step(
        data = pr_base_facet_step,
        aes(x = recall, y = precision),
        color = "red",
        linewidth = 0.8,
        linetype = "dashed",
        direction = "vh",
        inherit.aes = FALSE
      ) +
      geom_hline(yintercept = 0.50, linetype = "dotted", color = "grey50",
                 linewidth = 0.4) +
      facet_wrap(~alpha_concentration, labeller = label_both) +
      labs(
        x = "Recall (TPR)",
        y = "Precision",
        color = "Aggregation",
        title = sprintf("Precision-Recall Step Curves: Aggregated PIPs (%s)", wm),
        subtitle = "Stored confusion bins only; red dashed = baseline"
      ) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      theme(legend.position = "bottom")

    fname <- sprintf("pr_curve_step_aggregated_%s.png", wm)
    ggsave(file.path(fig_dir, fname), p_step, width = 12, height = 10, dpi = 150)
    cat("Saved:", file.path(fig_dir, fname), "\n")
  }
}
```

```{r auprc-bucket-aware-diagnostics}
group_vars_agg <- c("warm_method", "alpha_concentration", "agg_method")

auprc_agg_step <- compute_auprc_step_from_confusion(
  conf_agg_core,
  group_vars = group_vars_agg
)

conf_baseline_rebin001 <- rebin_confusion_uniform(
  conf_baseline,
  group_vars = character(),
  target_width = 0.01
)
pr_baseline_rebin001 <- build_pr_curve_step_single(
  conf_baseline_rebin001,
  "Baseline (default)"
)
baseline_auprc_step_rebin001 <- compute_auprc_step_single(
  pr_baseline_rebin001$precision,
  pr_baseline_rebin001$recall
)

conf_agg_rebin001 <- rebin_confusion_uniform(
  conf_agg_core,
  group_vars = group_vars_agg,
  target_width = 0.01
)
auprc_agg_step_rebin001 <- compute_auprc_step_from_confusion(
  conf_agg_rebin001,
  group_vars = group_vars_agg
) %>%
  dplyr::rename(AUPRC_step_rebin001 = .data$AUPRC_step)

auprc_diagnostics <- auprc_agg %>%
  dplyr::rename(AUPRC_trapezoid = .data$AUPRC) %>%
  dplyr::left_join(auprc_agg_step, by = group_vars_agg) %>%
  dplyr::left_join(auprc_agg_step_rebin001, by = group_vars_agg) %>%
  dplyr::mutate(
    delta_step_minus_trapezoid = .data$AUPRC_step - .data$AUPRC_trapezoid,
    delta_rebin001_minus_step = .data$AUPRC_step_rebin001 - .data$AUPRC_step
  ) %>%
  dplyr::arrange(.data$warm_method, .data$alpha_concentration, .data$agg_method)

diag_path <- file.path(fig_dir, "auprc_metric_diagnostics.csv")
readr::write_csv(auprc_diagnostics, diag_path)
cat("Saved:", diag_path, "\n")

baseline_metric_df <- tidyr::crossing(
  warm_method = sort(unique(conf_agg_core$warm_method)),
  metric = c(
    "Stored bins: trapezoid (current)",
    "Stored bins: step (recommended)",
    "0.01 re-bin: step (sensitivity)"
  )
) %>%
  dplyr::mutate(
    baseline_auprc = dplyr::case_when(
      .data$metric == "Stored bins: trapezoid (current)" ~ baseline_auprc,
      .data$metric == "Stored bins: step (recommended)" ~ baseline_auprc_step,
      TRUE ~ baseline_auprc_step_rebin001
    )
  )

auprc_diag_plot_df <- auprc_diagnostics %>%
  dplyr::select(
    .data$warm_method,
    .data$alpha_concentration,
    .data$agg_method,
    .data$AUPRC_trapezoid,
    .data$AUPRC_step,
    .data$AUPRC_step_rebin001
  ) %>%
  tidyr::pivot_longer(
    cols = c(.data$AUPRC_trapezoid, .data$AUPRC_step, .data$AUPRC_step_rebin001),
    names_to = "metric",
    values_to = "AUPRC"
  ) %>%
  dplyr::mutate(
    metric = dplyr::recode(
      .data$metric,
      AUPRC_trapezoid = "Stored bins: trapezoid (current)",
      AUPRC_step = "Stored bins: step (recommended)",
      AUPRC_step_rebin001 = "0.01 re-bin: step (sensitivity)"
    )
  )

p_auprc_diag <- ggplot(
  auprc_diag_plot_df,
  aes(x = factor(alpha_concentration), y = AUPRC, fill = agg_method)
) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_hline(
    data = baseline_metric_df,
    aes(yintercept = baseline_auprc),
    linetype = "dashed",
    color = "red",
    linewidth = 0.7,
    inherit.aes = FALSE
  ) +
  facet_grid(metric ~ warm_method, scales = "free_x") +
  labs(
    x = "Dirichlet alpha",
    y = "AUPRC",
    fill = "Aggregation",
    title = "AUPRC Sensitivity to PR Integration and Re-binning",
    subtitle = "Current trapezoid vs exact step area; 0.01 re-bin shown only as a sensitivity check"
  ) +
  ylim(0, NA) +
  theme(legend.position = "bottom")

ggsave(file.path(fig_dir, "auprc_metric_sensitivity_by_method.png"),
       p_auprc_diag, width = 12, height = 10, dpi = 150)
cat("Saved:", file.path(fig_dir, "auprc_metric_sensitivity_by_method.png"), "\n")
```

# 9. Calibration Curves


PIP calibration = observed causal fraction per PIP bin vs. predicted PIP.
We re-bin the variable-width confusion bins into uniform 0.05 increments
and pool counts across all datasets. The baseline (default uniform prior)
calibration is overlaid on every panel as a red line for reference.

```{r calibration-helper}
# Re-bin confusion bins into 0.05 PIP increments, pooling across datasets.
# Returns one row per (group_vars..., pip_bin).
aggregate_calibration <- function(confusion_df, group_vars, bin_width = 0.05) {
  pip_breaks <- seq(0, 1, by = bin_width)
  confusion_df %>%
    dplyr::mutate(
      pip_bin = pip_breaks[findInterval(pip_threshold, pip_breaks,
                                        rightmost.closed = TRUE,
                                        left.open = FALSE)]
    ) %>%
    dplyr::group_by(
      dplyr::across(dplyr::all_of(group_vars)),
      pip_bin
    ) %>%
    dplyr::summarise(
      n_causal = sum(n_causal_at_bucket, na.rm = TRUE),
      n_noncausal = sum(n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      n_total = n_causal + n_noncausal,
      observed_causal_frac = dplyr::if_else(
        n_total > 0, n_causal / n_total, NA_real_
      ),
      pip_bin_mid = pip_bin + bin_width / 2
    ) %>%
    dplyr::filter(n_total > 0)
}
```

```{r calibration-data}
# Baseline calibration (pooled across datasets, for overlay)
cal_baseline <- aggregate_calibration(conf_baseline, group_vars = character(0))

# Individual fits: pooled calibration per (warm_method, alpha)
cal_individual <- aggregate_calibration(
  conf_fits,
  group_vars = c("warm_method", "alpha_concentration")
)

# Aggregated: filter to core agg methods, pooled calibration
conf_agg_core <- conf_agg %>%
  filter(agg_method %in% keep_agg_methods)

if (nrow(conf_agg_core) > 0) {
  cal_agg <- aggregate_calibration(
    conf_agg_core,
    group_vars = c("warm_method", "alpha_concentration", "agg_method")
  )
} else {
  cal_agg <- tibble()
}
```

```{r plot-calibration-individual}
# One page per warm_method, since each has its own alpha grid
for (wm in sort(unique(cal_individual$warm_method))) {
  cal_wm <- cal_individual %>% filter(warm_method == wm)

  p <- ggplot(cal_wm,
              aes(x = pip_bin_mid, y = observed_causal_frac)) +
    geom_line(linewidth = 0.7, color = "steelblue") +
    geom_point(size = 1.5, color = "steelblue") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed",
                color = "grey40") +
    geom_line(data = cal_baseline,
              aes(x = pip_bin_mid, y = observed_causal_frac),
              color = "red", linewidth = 0.8, linetype = "dashed",
              inherit.aes = FALSE) +
    geom_point(data = cal_baseline,
               aes(x = pip_bin_mid, y = observed_causal_frac),
               color = "red", size = 1.5, inherit.aes = FALSE) +
    facet_wrap(~alpha_concentration, labeller = label_both) +
    labs(
      x = "Predicted PIP (bin midpoint, width = 0.05)",
      y = "Observed causal fraction",
      title = sprintf("PIP Calibration: Individual Fits (%s)", wm),
      subtitle = "Red dashed = baseline; grey dashed = perfect calibration"
    ) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))

  fname <- sprintf("calibration_individual_%s.png", wm)
  ggsave(file.path(fig_dir, fname), p, width = 12, height = 10, dpi = 150)
  cat("Saved:", file.path(fig_dir, fname), "\n")
}
```

```{r plot-calibration-aggregated}
if (nrow(cal_agg) > 0) {
  for (wm in sort(unique(cal_agg$warm_method))) {
    cal_wm <- cal_agg %>% filter(warm_method == wm)

    p <- ggplot(cal_wm,
                aes(x = pip_bin_mid, y = observed_causal_frac,
                    color = agg_method)) +
      geom_line(linewidth = 0.7) +
      geom_point(size = 1.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed",
                  color = "grey40") +
      geom_line(data = cal_baseline,
                aes(x = pip_bin_mid, y = observed_causal_frac),
                color = "red", linewidth = 0.8, linetype = "dashed",
                inherit.aes = FALSE) +
      geom_point(data = cal_baseline,
                 aes(x = pip_bin_mid, y = observed_causal_frac),
                 color = "red", size = 1.5, inherit.aes = FALSE) +
      facet_wrap(~alpha_concentration, labeller = label_both) +
      labs(
        x = "Predicted PIP (bin midpoint, width = 0.05)",
        y = "Observed causal fraction",
        color = "Aggregation",
        title = sprintf("PIP Calibration: Aggregated PIPs (%s)", wm),
        subtitle = "Red dashed = baseline; grey dashed = perfect calibration"
      ) +
      coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
      theme(legend.position = "bottom")

    fname <- sprintf("calibration_aggregated_%s.png", wm)
    ggsave(file.path(fig_dir, fname), p, width = 12, height = 10, dpi = 150)
    cat("Saved:", file.path(fig_dir, fname), "\n")
  }
}
```

# 10. Precision-Recall: Baseline vs Best Ensemble

```{r precision-recall-comparison}
# Helper: build PR curve from confusion bins
build_pr_curve <- function(confusion_df, label) {
  confusion_df %>%
    arrange(desc(pip_threshold)) %>%
    mutate(
      TP = cumsum(n_causal_at_bucket),
      FP = cumsum(n_noncausal_at_bucket),
      precision = TP / (TP + FP),
      recall = TP / sum(n_causal_at_bucket)
    ) %>%
    filter(!is.na(precision), !is.na(recall)) %>%
    select(pip_threshold, precision, recall) %>%
    mutate(source = label)
}

pr_baseline <- build_pr_curve(conf_baseline, "Baseline (default)")

pr_ensemble <- conf %>%
  filter(
    warm_method == "init_alpha",
    alpha_concentration == 0,
    agg_method == "cluster_weight"
  ) %>%
  build_pr_curve("init_alpha, a=0, cluster_weight")

pr_combined <- bind_rows(pr_baseline, pr_ensemble)

p_pr <- ggplot(pr_combined, aes(x = recall, y = precision, color = source)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 0.5, alpha = 0.5) +
  geom_hline(yintercept = 0.50, linetype = "dotted", color = "grey50") +
  annotate("text", x = 0.95, y = 0.52, label = "50% precision",
           color = "grey50", size = 3, hjust = 1) +
  labs(
    x = "Recall (TPR)",
    y = "Precision",
    color = NULL,
    title = "Precision-Recall: Baseline vs Ensemble (init_alpha, a=0, cluster_weight)",
    subtitle = "Pooled across all datasets"
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme(legend.position = "bottom")

ggsave(file.path(fig_dir, "precision_recall_baseline_vs_ensemble.png"),
       p_pr, width = 8, height = 6, dpi = 150)
cat("Saved:", file.path(fig_dir, "precision_recall_baseline_vs_ensemble.png"), "\n")
```

# 11. Head-to-Head: Best Alpha per Method

```{r best-alpha-comparison}
if (nrow(tpr_at_fpr_agg) > 0) {
  best_per_method <- tpr_at_fpr_agg %>%
    group_by(warm_method, agg_method) %>%
    slice_max(tpr_fpr005, n = 1, with_ties = FALSE) %>%
    ungroup()

  cat("Best alpha per (warm_method, agg_method):\n")
  print(as.data.frame(best_per_method))

  p_h2h <- ggplot(
    best_per_method,
    aes(x = agg_method, y = tpr_fpr005, fill = warm_method)
  ) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_hline(yintercept = baseline_tpr_fpr005, linetype = "dashed",
               color = "red", linewidth = 0.7) +
    geom_text(aes(label = sprintf("a=%.0e", alpha_concentration)),
              position = position_dodge(width = 0.7), vjust = -0.3, size = 2.5) +
    labs(
      x = "Aggregation Method",
      y = "TPR @ FPR = 0.05",
      fill = "Warm method",
      title = "Best Alpha per Method: TPR @ FPR = 0.05",
      subtitle = "Labels show the best alpha_concentration for each bar"
    ) +
    ylim(0, NA)

  ggsave(file.path(fig_dir, "head_to_head_best_alpha.png"),
         p_h2h, width = 9, height = 5, dpi = 150)
  cat("Saved:", file.path(fig_dir, "head_to_head_best_alpha.png"), "\n")
}
```

# 12. Multimodal Metrics by Method

```{r load-multimodal}
mm_path <- file.path(agg_dir, "multimodal_metrics.csv")
if (file.exists(mm_path)) {
  mm <- readr::read_csv(mm_path, show_col_types = FALSE)

  # Parse warm_method and alpha from group_label
  mm <- mm %>%
    mutate(
      warm_method = stringr::str_extract(group_label, "(?<=wm=)[a-z_]+"),
      alpha_concentration = as.numeric(
        stringr::str_extract(group_label, "(?<=alpha=)[0-9.e-]+")
      )
    )

  cat("Multimodal metrics rows:", nrow(mm), "\n")
  cat("Warm methods:", sort(unique(mm$warm_method)), "\n")
  cat("Alpha values:", sort(unique(mm$alpha_concentration)), "\n")
} else {
  mm <- tibble()
  cat("No multimodal_metrics.csv found.\n")
}
```

```{r multimodal-summary}
if (nrow(mm) > 0) {
  cat("=== Mean JSD by warm_method and alpha ===\n")
  mm %>%
    filter(!is.na(alpha_concentration)) %>%
    group_by(warm_method, alpha_concentration) %>%
    summarise(
      n = n(),
      mean_jsd = mean(mean_jsd, na.rm = TRUE),
      mean_max_jsd = mean(max_jsd, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    arrange(warm_method, alpha_concentration) %>%
    print(n = Inf)
}
```

```{r plot-multimodal-jsd-by-method}
if (nrow(mm) > 0 && any(!is.na(mm$warm_method))) {
  mm_jsd <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, mean_jsd, max_jsd) %>%
    tidyr::pivot_longer(
      cols = c(mean_jsd, max_jsd),
      names_to = "metric",
      values_to = "value"
    ) %>%
    mutate(metric = dplyr::recode(metric,
      mean_jsd = "Mean pairwise JSD",
      max_jsd = "Max pairwise JSD"
    ))

  p_jsd <- ggplot(mm_jsd, aes(x = factor(alpha_concentration), y = value,
                               fill = warm_method)) +
    geom_boxplot(outlier.size = 0.5) +
    facet_wrap(~metric, scales = "free_y", ncol = 1) +
    labs(x = "Dirichlet alpha", fill = "Warm method",
         title = "PIP Diversity (JSD) by Method and Alpha") +
    theme(axis.text.x = element_text(size = 9))

  ggsave(file.path(fig_dir, "multimodal_jsd_by_method.png"),
         p_jsd, width = 10, height = 7, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_jsd_by_method.png"), "\n")
}
```

```{r plot-multimodal-agnostic-by-method}
if (nrow(mm) > 0 && any(!is.na(mm$warm_method))) {
  mm_agnostic <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, jaccard_top10, mean_pip_var) %>%
    tidyr::pivot_longer(
      cols = c(jaccard_top10, mean_pip_var),
      names_to = "metric",
      values_to = "value"
    )

  p_agnostic <- ggplot(mm_agnostic,
                        aes(x = factor(alpha_concentration), y = value,
                            fill = warm_method)) +
    geom_boxplot(outlier.size = 0.5) +
    facet_wrap(~metric, scales = "free_y", ncol = 2) +
    labs(x = "Dirichlet alpha", fill = "Warm method",
         title = "JSD-Agnostic Multimodal Metrics by Method and Alpha") +
    theme(axis.text.x = element_text(size = 9))

  ggsave(file.path(fig_dir, "multimodal_agnostic_by_method.png"),
         p_agnostic, width = 10, height = 5, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_agnostic_by_method.png"), "\n")
}
```

```{r plot-multimodal-n-clusters}
if (nrow(mm) > 0) {
  jsd_clust_cols <- grep("^n_clusters_jsd_", names(mm), value = TRUE)
  bjsd_clust_cols <- grep("^n_clusters_bjsd_", names(mm), value = TRUE)

  mm_clust_cat <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, all_of(jsd_clust_cols)) %>%
    tidyr::pivot_longer(
      cols = -c(warm_method, alpha_concentration),
      names_to = "col_name",
      values_to = "n_clusters"
    ) %>%
    mutate(
      threshold = dplyr::case_match(col_name,
        "n_clusters_jsd_015" ~ 0.15,
        "n_clusters_jsd_050" ~ 0.50,
        "n_clusters_jsd_100" ~ 1.00,
        "n_clusters_jsd_200" ~ 2.00,
        "n_clusters_jsd_300" ~ 3.00,
        "n_clusters_jsd_500" ~ 5.00
      ),
      jsd_type = "Categorical"
    )

  mm_clust_bern <- mm %>%
    filter(!is.na(alpha_concentration), !is.na(warm_method)) %>%
    select(warm_method, alpha_concentration, all_of(bjsd_clust_cols)) %>%
    tidyr::pivot_longer(
      cols = -c(warm_method, alpha_concentration),
      names_to = "col_name",
      values_to = "n_clusters"
    ) %>%
    mutate(
      threshold = dplyr::case_match(col_name,
        "n_clusters_bjsd_0001" ~ 0.001,
        "n_clusters_bjsd_0002" ~ 0.002,
        "n_clusters_bjsd_0004" ~ 0.004,
        "n_clusters_bjsd_0006" ~ 0.006,
        "n_clusters_bjsd_0008" ~ 0.008,
        "n_clusters_bjsd_000" ~ 0.001,
        "n_clusters_bjsd_001" ~ 0.01,
        "n_clusters_bjsd_005" ~ 0.05,
        "n_clusters_bjsd_010" ~ 0.10,
        "n_clusters_bjsd_020" ~ 0.20,
        "n_clusters_bjsd_040" ~ 0.40
      ),
      jsd_type = "Bernoulli"
    )

  mm_clust <- bind_rows(mm_clust_cat, mm_clust_bern) %>%
    filter(!is.na(threshold)) %>%
    mutate(threshold_label = sprintf("h = %s", threshold))

  p_clust <- ggplot(mm_clust,
                     aes(x = factor(alpha_concentration), y = n_clusters,
                         fill = warm_method)) +
    geom_boxplot(outlier.size = 0.5) +
    facet_grid(jsd_type ~ threshold_label, scales = "free") +
    labs(x = "Dirichlet alpha", y = "Number of clusters",
         fill = "Warm method",
         title = "Cluster Count by Method, JSD Type, and Threshold") +
    theme(axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
          strip.text = element_text(size = 8))

  ggsave(file.path(fig_dir, "multimodal_n_clusters_by_method.png"),
         p_clust, width = 16, height = 7, dpi = 150)
  cat("Saved:", file.path(fig_dir, "multimodal_n_clusters_by_method.png"), "\n")
}
```
