---
title: "Baseline Sims Screening - Collect Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Post-hoc collection for the baseline screening job. This merges staged
flush files, derives baseline-specific identifiers, and writes a
consolidated directory for the baseline visualization workbook.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_baseline_sims.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(knitr)
```

# 1. Locate Output

```{r locate}
job_name <- "baseline_sims_screen"
parent_job_id <- "51509956"
output_root <- here("output")

job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
cfg_path <- file.path(output_root, "run_history", job_name, parent_job_id, "job_config.json")
cfg <- test_susine:::load_job_config(cfg_path)

aggregated_dir <- file.path(job_dir, "aggregated")
consolidated_dir <- file.path(job_dir, "consolidated")
dir.create(consolidated_dir, recursive = TRUE, showWarnings = FALSE)

cat("Job dir:", job_dir, "\n")
cat("Aggregated dir:", aggregated_dir, "\n")
cat("Consolidated dir:", consolidated_dir, "\n")
```

# 2. Aggregate Staging Outputs

```{r aggregate}
agg <- test_susine::aggregate_staging_outputs(
  job_name = job_name,
  parent_job_id = parent_job_id,
  output_root = output_root,
  validate = TRUE,
  output_dir = aggregated_dir,
  write_snps_dataset = FALSE
)

if (!is.null(agg$validation) && nrow(agg$validation)) {
  cat("Validated staged files:", nrow(agg$validation), "\n")
}
```

# 3. Load Aggregated Tables

```{r load-aggregated}
read_optional_csv <- function(path) {
  if (!file.exists(path)) {
    return(NULL)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

model_metrics <- read_optional_csv(file.path(aggregated_dir, "model_metrics.csv"))
effect_metrics <- read_optional_csv(file.path(aggregated_dir, "effect_metrics.csv"))
effect_metrics_unfiltered <- read_optional_csv(file.path(aggregated_dir, "effect_metrics_unfiltered.csv"))
confusion_bins <- read_optional_csv(file.path(aggregated_dir, "confusion_bins.csv"))
dataset_metrics <- read_optional_csv(file.path(aggregated_dir, "dataset_metrics.csv"))
tier_cs_metrics <- read_optional_csv(file.path(aggregated_dir, "tier_cs_metrics.csv"))
validation <- read_optional_csv(file.path(aggregated_dir, "validation.csv"))

cat("Model metrics rows:", if (is.null(model_metrics)) 0 else nrow(model_metrics), "\n")
cat("Filtered effect rows:", if (is.null(effect_metrics)) 0 else nrow(effect_metrics), "\n")
cat("Unfiltered effect rows:", if (is.null(effect_metrics_unfiltered)) 0 else nrow(effect_metrics_unfiltered), "\n")
cat("Confusion-bin rows:", if (is.null(confusion_bins)) 0 else nrow(confusion_bins), "\n")
cat("Dataset-metric rows:", if (is.null(dataset_metrics)) 0 else nrow(dataset_metrics), "\n")
cat("Tier-CS rows:", if (is.null(tier_cs_metrics)) 0 else nrow(tier_cs_metrics), "\n")
```

# 4. Derive Baseline Identifiers

```{r identifiers}
fmt_num <- function(x, digits = 3) {
  out <- format(signif(as.numeric(x), digits), trim = TRUE, scientific = FALSE)
  out
}

annot_label <- function(annotation_r2, inflate_match) {
  dplyr::if_else(
    is.na(annotation_r2) | is.na(inflate_match),
    "no_annotation",
    paste0("r2=", fmt_num(annotation_r2, 2), " | inflate=", fmt_num(inflate_match, 2))
  )
}

derive_baseline_identifiers <- function(df) {
  if (is.null(df) || !nrow(df)) {
    return(df)
  }
  if (!"use_case_id" %in% names(df)) {
    return(df)
  }

  df %>%
    dplyr::mutate(
      method_family = dplyr::case_when(
        .data$use_case_id == "susie_vanilla" ~ "susie_vanilla",
        .data$use_case_id == "susine_vanilla" ~ "susine_vanilla",
        .data$use_case_id == "susie_eb" ~ "susie_eb",
        .data$use_case_id == "susie_inf" ~ "susie_inf",
        .data$use_case_id == "susie_ash_fixed" ~ "susie_ash_fixed",
        .data$use_case_id == "susine_functional_mu" ~ "susine_functional_mu",
        .data$use_case_id == "susine_eb_clamped_scale_var_nonneg" ~ "susine_eb_clamped_scale_var_nonneg",
        .data$use_case_id == "susie_functional_pi" ~ "susie_functional_pi",
        TRUE ~ .data$use_case_id
      ),
      annotation_label = annot_label(.data$annotation_r2, .data$inflate_match),
      setting_label = dplyr::case_when(
        .data$method_family %in% c("susie_vanilla", "susine_vanilla", "susie_inf", "susie_ash_fixed") ~
          paste0(.data$method_family, " | sigma=", .data$sigma_0_2_scalar),
        .data$method_family == "susie_functional_pi" ~
          paste0(.data$method_family, " | tau=", fmt_num(.data$tau_value, 3)),
        .data$method_family == "susine_functional_mu" ~
          paste0(
            .data$method_family,
            " | c=", fmt_num(.data$c_value, 3),
            " | sigma=", .data$sigma_0_2_scalar
          ),
        TRUE ~ .data$method_family
      )
    )
}

model_metrics <- derive_baseline_identifiers(model_metrics)
effect_metrics <- derive_baseline_identifiers(effect_metrics)
effect_metrics_unfiltered <- derive_baseline_identifiers(effect_metrics_unfiltered)
confusion_bins <- derive_baseline_identifiers(confusion_bins)
tier_cs_metrics <- derive_baseline_identifiers(tier_cs_metrics)
```

# 5. Compute Baseline Screening Metrics

```{r screening-metrics}
per_dataset_group_vars <- c(
  "dataset_bundle_id", "run_id", "spec_name", "use_case_id",
  "method_family", "setting_label", "annotation_label",
  "annotation_r2", "inflate_match",
  "sigma_0_2_scalar", "c_value", "tau_value"
)
pooled_group_vars <- c(
  "spec_name", "use_case_id", "method_family", "setting_label",
  "annotation_label", "annotation_r2", "inflate_match",
  "sigma_0_2_scalar", "c_value", "tau_value"
)
annotation_group_vars <- c(
  "spec_name", "use_case_id", "method_family", "setting_label",
  "annotation_label", "sigma_0_2_scalar", "c_value", "tau_value"
)

auprc_per_dataset <- if (!is.null(confusion_bins) && nrow(confusion_bins)) {
  test_susine::compute_auprc_from_confusion(
    confusion_bins,
    group_vars = per_dataset_group_vars
  )
} else {
  tibble::tibble()
}

auprc_pooled_overall <- if (!is.null(confusion_bins) && nrow(confusion_bins)) {
  test_susine::compute_auprc_from_confusion(
    confusion_bins,
    group_vars = pooled_group_vars
  )
} else {
  tibble::tibble()
}

auprc_pooled_by_annotation <- if (!is.null(confusion_bins) && nrow(confusion_bins)) {
  test_susine::compute_auprc_from_confusion(
    confusion_bins,
    group_vars = annotation_group_vars
  )
} else {
  tibble::tibble()
}
tpr05_per_dataset <- if (!is.null(confusion_bins) && nrow(confusion_bins)) {
  pooled <- test_susine:::pool_confusion_bins(confusion_bins, per_dataset_group_vars)
  test_susine:::compute_tpr05_from_pooled_confusion(
    pooled,
    group_vars = per_dataset_group_vars
  )
} else {
  tibble::tibble()
}

tpr05_pooled_overall <- if (!is.null(confusion_bins) && nrow(confusion_bins)) {
  pooled <- test_susine:::pool_confusion_bins(confusion_bins, pooled_group_vars)
  test_susine:::compute_tpr05_from_pooled_confusion(
    pooled,
    group_vars = pooled_group_vars
  )
} else {
  tibble::tibble()
}

tpr05_pooled_by_annotation <- if (!is.null(confusion_bins) && nrow(confusion_bins)) {
  pooled <- test_susine:::pool_confusion_bins(confusion_bins, annotation_group_vars)
  test_susine:::compute_tpr05_from_pooled_confusion(
    pooled,
    group_vars = annotation_group_vars
  )
} else {
  tibble::tibble()
}

screening_summary <- auprc_pooled_overall %>%
  dplyr::left_join(
    tpr05_pooled_overall,
    by = pooled_group_vars
  ) %>%
  dplyr::arrange(dplyr::desc(.data$AUPRC), dplyr::desc(.data$tpr_fpr05))

kable(head(screening_summary, 20), digits = 4, caption = "Top pooled baseline settings")
```

# 6. Export Consolidated Files

```{r export}
write_if_present <- function(df, filename) {
  if (!is.null(df) && nrow(df)) {
    readr::write_csv(df, file.path(consolidated_dir, filename))
  }
}

write_if_present(model_metrics, "model_metrics_full.csv")
write_if_present(effect_metrics, "effect_metrics_full.csv")
write_if_present(effect_metrics_unfiltered, "effect_metrics_unfiltered_full.csv")
write_if_present(confusion_bins, "confusion_bins_full.csv")
write_if_present(dataset_metrics, "dataset_metrics_full.csv")
write_if_present(tier_cs_metrics, "tier_cs_metrics_full.csv")
write_if_present(validation, "validation_full.csv")

write_if_present(auprc_per_dataset, "auprc_per_dataset.csv")
write_if_present(auprc_pooled_overall, "auprc_pooled_overall.csv")
write_if_present(auprc_pooled_by_annotation, "auprc_pooled_by_annotation.csv")
write_if_present(tpr05_per_dataset, "tpr05_per_dataset.csv")
write_if_present(tpr05_pooled_overall, "tpr05_pooled_overall.csv")
write_if_present(tpr05_pooled_by_annotation, "tpr05_pooled_by_annotation.csv")
write_if_present(screening_summary, "screening_summary.csv")

cat("Consolidated files written to:", consolidated_dir, "\n")
```
