---
title: "Purity-Filter Minibatch - Collect Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Post-run collection for the purity-filter minibatch. Merges the staged
confusion-bin flush files (tagged by `filter_type`) and the per-effect diagnostic
records, then:

1. computes one pooled AUPRC per filter type (and per-dataset deltas vs the
   unfiltered PIP), the headline "AUPRC with vs without purity filtering" number;
2. builds the old-vs-new credible-set diagnostic figures that contrast the
   classical metrics (purity, size) with the effect-level metrics (core-95%
   $k_{\mathrm{eff}}$, accuracy ratio).

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/collect_results_workbook_purity_filter_minibatch.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(knitr)
```

# 1. Locate Output

```{r locate}
job_name <- "purity_filter_minibatch"
parent_job_id <- Sys.getenv("PURITY_FILTER_PARENT_JOB_ID", unset = "53615664")
output_root <- here("output")

job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
cfg_path <- file.path(output_root, "run_history", job_name, parent_job_id, "job_config.json")
cfg <- if (file.exists(cfg_path)) test_susine:::load_job_config(cfg_path) else NULL

aggregated_dir <- file.path(job_dir, "aggregated")
consolidated_dir <- file.path(job_dir, "consolidated")
plot_dir <- file.path(consolidated_dir, "plots")
dir.create(consolidated_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

task_dirs <- list.dirs(job_dir, recursive = FALSE, full.names = TRUE)
staging_flush_files <- list.files(
  task_dirs,
  pattern = "^flush-[0-9]+_",
  full.names = TRUE
)
consolidated_cache_available <- all(file.exists(file.path(
  consolidated_dir,
  c("auprc_by_filter_type.csv", "auprc_per_dataset_by_filter.csv",
    "auprc_per_dataset_delta_summary.csv", "effect_metrics_unfiltered_full.csv")
)))
use_consolidated_cache <- !length(staging_flush_files) && consolidated_cache_available

cat("Job dir:", job_dir, "\n")
cat("Using consolidated cache only:", use_consolidated_cache, "\n")
```

# 2. Aggregate Staging Outputs

```{r aggregate, eval = !use_consolidated_cache}
if (!length(staging_flush_files)) {
  stop(
    "No staging flush files found and consolidated recovery files are incomplete: ",
    job_dir
  )
}

agg <- test_susine::aggregate_staging_outputs(
  job_name = job_name,
  parent_job_id = parent_job_id,
  output_root = output_root,
  validate = TRUE,
  output_dir = aggregated_dir,
  write_snps_dataset = FALSE
)
```

# 3. Load Aggregated Tables

```{r load-aggregated}
read_optional_csv <- function(path) {
  if (!file.exists(path)) return(NULL)
  readr::read_csv(path, show_col_types = FALSE)
}

confusion_bins <- read_optional_csv(file.path(aggregated_dir, "confusion_bins.csv"))
effect_unfiltered <- read_optional_csv(file.path(aggregated_dir, "effect_metrics_unfiltered.csv"))
if (is.null(effect_unfiltered)) {
  effect_unfiltered <- read_optional_csv(file.path(consolidated_dir, "effect_metrics_unfiltered_full.csv"))
}

auprc_pooled_cached <- read_optional_csv(file.path(consolidated_dir, "auprc_by_filter_type.csv"))
auprc_per_dataset_cached <- read_optional_csv(file.path(consolidated_dir, "auprc_per_dataset_by_filter.csv"))
per_dataset_summary_cached <- read_optional_csv(file.path(consolidated_dir, "auprc_per_dataset_delta_summary.csv"))

filter_labels <- c(
  no_filter           = "No filter",
  purity_95           = "Purity ≥ 0.95",
  purity_90           = "Purity ≥ 0.90",
  purity_50           = "Purity ≥ 0.50",
  keff_core95_le_100  = "k_eff(core95) ≤ 100"
)
filter_level_template <- c("no_filter", "purity_95", "purity_90", "purity_50",
                           "keff_core95_le_100")
observed_filter_types <- unique(c(
  if (!is.null(confusion_bins) && "filter_type" %in% names(confusion_bins)) confusion_bins$filter_type else NULL,
  if (!is.null(auprc_pooled_cached) && "filter_type" %in% names(auprc_pooled_cached)) auprc_pooled_cached$filter_type else NULL,
  if (!is.null(auprc_per_dataset_cached) && "filter_type" %in% names(auprc_per_dataset_cached)) auprc_per_dataset_cached$filter_type else NULL
))
filter_levels <- intersect(filter_level_template, observed_filter_types)

if (!use_consolidated_cache) {
  stopifnot(!is.null(confusion_bins), "filter_type" %in% names(confusion_bins))
}
if (use_consolidated_cache && is.null(auprc_pooled_cached)) {
  stop("Consolidated recovery mode requires auprc_by_filter_type.csv")
}

cat("Confusion-bin rows:", if (is.null(confusion_bins)) 0 else nrow(confusion_bins), "\n")
cat("Filter types:", paste(sort(filter_levels), collapse = ", "), "\n")
cat("Per-effect rows:", if (is.null(effect_unfiltered)) 0 else nrow(effect_unfiltered), "\n")
```

# 4. AUPRC by Filter Type (headline)

```{r auprc-by-filter}
# Pooled AUPRC across all 600 datasets, one value per filter type.
if (use_consolidated_cache) {
  auprc_pooled <- auprc_pooled_cached
  auprc_per_dataset <- auprc_per_dataset_cached
  per_dataset_summary <- per_dataset_summary_cached
} else {
  auprc_pooled <- test_susine::compute_auprc_from_confusion(
    confusion_bins,
    group_vars = "filter_type"
  )
}

auprc_pooled <- auprc_pooled %>%
  dplyr::mutate(filter_type = factor(as.character(.data$filter_type), levels = filter_levels)) %>%
  dplyr::arrange(.data$filter_type)

auprc_no_filter <- auprc_pooled$AUPRC[as.character(auprc_pooled$filter_type) == "no_filter"]
if (!"delta_vs_unfiltered" %in% names(auprc_pooled)) {
  auprc_pooled <- auprc_pooled %>%
    dplyr::mutate(delta_vs_unfiltered = .data$AUPRC - auprc_no_filter)
}
if (!"filter_label" %in% names(auprc_pooled)) {
  auprc_pooled <- auprc_pooled %>%
    dplyr::mutate(filter_label = filter_labels[as.character(.data$filter_type)])
}

kable(
  auprc_pooled %>%
    dplyr::select(.data$filter_label, .data$AUPRC, .data$delta_vs_unfiltered),
  digits = 4,
  col.names = c("Filter", "Pooled AUPRC", "Δ vs unfiltered"),
  caption = "Pooled causal-recovery AUPRC (top-8 positives) by credible-set filter, 600 datasets"
)

# Per-dataset AUPRC -> paired delta vs no_filter (sign test on the filtering cost).
if (!use_consolidated_cache) {
  auprc_per_dataset <- test_susine::compute_auprc_from_confusion(
    confusion_bins,
    group_vars = c("filter_type", "dataset_bundle_id")
  )

  per_dataset_delta <- auprc_per_dataset %>%
    dplyr::select(.data$filter_type, .data$dataset_bundle_id, .data$AUPRC) %>%
    tidyr::pivot_wider(names_from = .data$filter_type, values_from = .data$AUPRC) %>%
    tidyr::pivot_longer(
      cols = dplyr::any_of(setdiff(filter_levels, "no_filter")),
      names_to = "filter_type", values_to = "AUPRC_filtered"
    ) %>%
    dplyr::mutate(delta = .data$AUPRC_filtered - .data$no_filter)

  per_dataset_summary <- per_dataset_delta %>%
    dplyr::filter(is.finite(.data$delta)) %>%
    dplyr::group_by(.data$filter_type) %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean_delta = mean(.data$delta),
      median_delta = stats::median(.data$delta),
      frac_hurt = mean(.data$delta < 0),
      frac_help = mean(.data$delta > 0),
      .groups = "drop"
    )
}

if (!is.null(per_dataset_summary)) {
  per_dataset_summary <- per_dataset_summary %>%
    dplyr::mutate(filter_type = factor(as.character(.data$filter_type), levels = filter_levels)) %>%
    dplyr::arrange(.data$filter_type)
}

kable(per_dataset_summary, digits = 4,
      caption = "Per-dataset paired AUPRC change vs unfiltered (negative = filter hurts)")
```

```{r auprc-plot}
p_auprc <- ggplot2::ggplot(
  auprc_pooled,
  ggplot2::aes(x = stats::reorder(.data$filter_label, -.data$AUPRC), y = .data$AUPRC)
) +
  ggplot2::geom_col(fill = "#1F77B4", width = 0.7) +
  ggplot2::geom_hline(yintercept = auprc_no_filter, linetype = "dashed", color = "grey30") +
  ggplot2::geom_text(ggplot2::aes(label = sprintf("%.3f", .data$AUPRC)),
                     vjust = -0.4, size = 3.5) +
  ggplot2::labs(
    x = NULL, y = "Pooled AUPRC",
    title = "Credible-set filtering changes per-variant AUPRC",
    subtitle = "Dashed line: unfiltered PIP. Lower bars = the filter discarded useful signal."
  ) +
  ggplot2::theme_bw(base_size = 12) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 20, hjust = 1))

ggplot2::ggsave(file.path(plot_dir, "auprc_by_filter_type.png"),
                p_auprc, width = 8, height = 5, dpi = 300)
p_auprc
```

# 5. Old vs New Credible-Set Diagnostics

The per-effect record contrasts the classical metrics (purity, size) with the
effect-level metrics introduced in the paper (core-95% $k_{\mathrm{eff}}$,
accuracy ratio). `coverage = 1` marks effects whose credible set actually
contains a true causal variant.

```{r effect-diagnostics, eval = !is.null(effect_unfiltered)}
eff <- effect_unfiltered %>%
  dplyr::filter(.data$size > 0) %>%                 # active effects only
  dplyr::mutate(
    coverage_lab = factor(ifelse(.data$coverage > 0, "CS contains causal", "CS has no causal"),
                          levels = c("CS contains causal", "CS has no causal")),
    purity = as.numeric(.data$purity),
    accuracy_ratio = as.numeric(.data$accuracy_ratio),
    k_eff_core95 = as.numeric(.data$effect_k_eff_signal_core95),
    size = as.numeric(.data$size)
  )

cov_palette <- c("CS contains causal" = "#2CA02C", "CS has no causal" = "#D62728")

# (A) Purity vs accuracy: are low-purity sets actually inaccurate?
p_purity_acc <- ggplot2::ggplot(
  eff %>% dplyr::filter(is.finite(.data$purity), is.finite(.data$accuracy_ratio)),
  ggplot2::aes(x = .data$purity, y = .data$accuracy_ratio, color = .data$coverage_lab)
) +
  ggplot2::geom_point(alpha = 0.25, size = 0.8) +
  ggplot2::geom_vline(xintercept = 0.95, linetype = "dashed", color = "grey25") +
  ggplot2::scale_color_manual(values = cov_palette, name = NULL) +
  ggplot2::labs(
    x = "Credible-set purity (min pairwise |r|)",
    y = "Accuracy ratio A (top causal PIP / top PIP)",
    title = "Purity vs accuracy",
    subtitle = "Dashed line: 0.95 purity filter. Green points left of it are accurate sets the filter would discard."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")

# (B) Size vs core-95 k_eff: size overstates diffuseness.
p_size_keff <- ggplot2::ggplot(
  eff %>% dplyr::filter(is.finite(.data$size), is.finite(.data$k_eff_core95)),
  ggplot2::aes(x = .data$size, y = .data$k_eff_core95, color = .data$coverage_lab)
) +
  ggplot2::geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "grey50") +
  ggplot2::geom_point(alpha = 0.25, size = 0.8) +
  ggplot2::scale_x_log10() +
  ggplot2::scale_y_log10() +
  ggplot2::scale_color_manual(values = cov_palette, name = NULL) +
  ggplot2::labs(
    x = "Credible-set size (count)",
    y = "Core-95% effective support k_eff",
    title = "Size vs core-95% diffuseness",
    subtitle = "Dotted line: y = x. Points below it have a peaked core despite a large set."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")

# (C) Purity vs core-95 k_eff: purity is only weakly tied to diffuseness.
p_purity_keff <- ggplot2::ggplot(
  eff %>% dplyr::filter(is.finite(.data$purity), is.finite(.data$k_eff_core95)),
  ggplot2::aes(x = .data$purity, y = .data$k_eff_core95, color = .data$coverage_lab)
) +
  ggplot2::geom_point(alpha = 0.25, size = 0.8) +
  ggplot2::geom_vline(xintercept = 0.95, linetype = "dashed", color = "grey25") +
  ggplot2::scale_y_log10() +
  ggplot2::scale_color_manual(values = cov_palette, name = NULL) +
  ggplot2::labs(
    x = "Credible-set purity (min pairwise |r|)",
    y = "Core-95% effective support k_eff",
    title = "Purity vs diffuseness",
    subtitle = "Weak association: low-purity sets are not reliably more diffuse."
  ) +
  ggplot2::theme_bw(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")

ggplot2::ggsave(file.path(plot_dir, "diag_purity_vs_accuracy.png"),
                p_purity_acc, width = 7, height = 5.5, dpi = 300)
ggplot2::ggsave(file.path(plot_dir, "diag_size_vs_keff.png"),
                p_size_keff, width = 7, height = 5.5, dpi = 300)
ggplot2::ggsave(file.path(plot_dir, "diag_purity_vs_keff.png"),
                p_purity_keff, width = 7, height = 5.5, dpi = 300)

p_purity_acc
p_size_keff
p_purity_keff
```

```{r effect-diagnostics-summary, eval = !is.null(effect_unfiltered)}
# Quantify "high-quality but low-purity" and "high-purity but inaccurate" effects.
diag_summary <- eff %>%
  dplyr::filter(is.finite(.data$purity), is.finite(.data$accuracy_ratio)) %>%
  dplyr::summarise(
    n_effects = dplyr::n(),
    cor_purity_keff = suppressWarnings(stats::cor(.data$purity, .data$k_eff_core95,
                                                  use = "complete.obs", method = "spearman")),
    cor_purity_accuracy = suppressWarnings(stats::cor(.data$purity, .data$accuracy_ratio,
                                                      use = "complete.obs", method = "spearman")),
    # Accurate (A >= 0.9) AND covers a causal, yet would be dropped at purity 0.95:
    frac_accurate_dropped_at_095 = mean(.data$purity < 0.95 & .data$accuracy_ratio >= 0.9 &
                                          .data$coverage > 0),
    # High purity (>= 0.95) but inaccurate (A < 0.5):
    frac_pure_but_inaccurate = mean(.data$purity >= 0.95 & .data$accuracy_ratio < 0.5)
  )

kable(diag_summary, digits = 4,
      caption = "Decoupling of classical (purity) and effect-level (accuracy / k_eff) diagnostics")
```

# 6. Paper Figure Composite

```{r paper-figure-composite, eval = exists("p_auprc") && exists("p_purity_acc") && exists("p_size_keff") && exists("p_purity_keff")}
paper_panel_theme <- ggplot2::theme(
  plot.title = ggplot2::element_text(size = 11, face = "bold", hjust = 0.5),
  plot.subtitle = ggplot2::element_blank(),
  plot.caption = ggplot2::element_blank(),
  axis.title = ggplot2::element_text(size = 10),
  axis.text = ggplot2::element_text(size = 9),
  legend.title = ggplot2::element_text(size = 9),
  legend.text = ggplot2::element_text(size = 8),
  plot.margin = ggplot2::margin(7, 7, 7, 7)
)

open_png_device <- function(filename, width, height, dpi = 300) {
  if (requireNamespace("ragg", quietly = TRUE)) {
    ragg::agg_png(filename, width = width, height = height, units = "in",
                  res = dpi, background = "white")
  } else {
    grDevices::png(filename, width = width, height = height, units = "in",
                   res = dpi, bg = "white")
  }
}

save_paper_grid <- function(filename, panels, labels, width = 13.5,
                            height = 9.0, dpi = 300) {
  stopifnot(length(panels) == 4L, length(labels) == 4L)

  open_png_device(filename, width = width, height = height, dpi = dpi)
  on.exit(grDevices::dev.off(), add = TRUE)

  grid::grid.newpage()
  layout <- grid::grid.layout(
    nrow = 2, ncol = 2,
    widths = grid::unit(c(1, 1), "null"),
    heights = grid::unit(c(1, 1), "null")
  )
  grid::pushViewport(grid::viewport(layout = layout))
  on.exit(grid::popViewport(), add = TRUE)

  positions <- data.frame(row = c(1, 1, 2, 2), col = c(1, 2, 1, 2))
  for (i in seq_along(panels)) {
    grid::pushViewport(grid::viewport(
      layout.pos.row = positions$row[i],
      layout.pos.col = positions$col[i]
    ))
    print(panels[[i]], newpage = FALSE)
    grid::grid.text(
      labels[[i]],
      x = grid::unit(0.01, "npc"),
      y = grid::unit(0.99, "npc"),
      just = c("left", "top"),
      gp = grid::gpar(fontsize = 17, fontface = "bold")
    )
    grid::popViewport()
  }

  invisible(filename)
}

paper_panels <- list(
  p_auprc +
    ggplot2::labs(title = "AUPRC by credible-set filter") +
    paper_panel_theme +
    ggplot2::theme(legend.position = "bottom"),
  p_purity_acc +
    ggplot2::labs(title = "Purity versus causal coverage") +
    paper_panel_theme,
  p_size_keff +
    ggplot2::labs(title = "Credible-set size versus k_eff") +
    paper_panel_theme +
    ggplot2::theme(legend.position = "none"),
  p_purity_keff +
    ggplot2::labs(title = "Purity versus k_eff") +
    paper_panel_theme +
    ggplot2::theme(legend.position = "none")
)

paper_purity_filter_path <- file.path(plot_dir, "paper_purity_filter_composite.png")
save_paper_grid(
  paper_purity_filter_path,
  panels = paper_panels,
  labels = LETTERS[1:4]
)

cat("Paper composite written to:", paper_purity_filter_path, "\n")
```

# 7. Export Consolidated Files

```{r export}
write_if_present <- function(df, filename) {
  if (!is.null(df) && nrow(df)) {
    readr::write_csv(df, file.path(consolidated_dir, filename))
  }
}

write_if_present(auprc_pooled, "auprc_by_filter_type.csv")
write_if_present(auprc_per_dataset, "auprc_per_dataset_by_filter.csv")
write_if_present(per_dataset_summary, "auprc_per_dataset_delta_summary.csv")
if (exists("diag_summary")) write_if_present(diag_summary, "effect_diagnostic_summary.csv")
write_if_present(effect_unfiltered, "effect_metrics_unfiltered_full.csv")

cat("Consolidated files written to:", consolidated_dir, "\n")
cat("Plots written to:", plot_dir, "\n")
```
