---
title: "Export Matrix-Level AUPRC Difficulty Data"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Build a matrix-level modeling table for difficulty analysis.

This workbook:

1. Reads the consolidated ensemble-scaling outputs.
2. Re-attaches `matrix_id` from the job config if needed.
3. Pools confusion bins across all `dataset_bundle_id` values sharing the
   same `matrix_id`.
4. Computes matrix-level pooled AUPRC for four methods:
   - oracle warm-start vanilla (`truth-warm` + `susine_vanilla`)
   - baseline vanilla (`baseline-single` + `susine_vanilla`)
   - `C-CS` cluster-weight aggregation
   - `C-CS` oracle aggregation
5. Aggregates the non-LD dataset metrics to the matrix level.
6. Writes a single CSV for downstream modeling.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/export_matrix_level_auprc_difficulty_data.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

## 1. Job Inputs

```{r inputs}
job_name <- "ensemble_scaling_full"
parent_job_id <- "50339226"

# For C-CS summaries, pick the annotation R2 slice to export.
target_annotation_r2 <- 0.5

output_root <- here("output")
job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
consolidated_dir <- file.path(job_dir, "consolidated")
cfg_path <- file.path(output_root, "run_history", job_name, parent_job_id, "job_config.json")

out_csv <- file.path(
  consolidated_dir,
  sprintf("matrix_level_auprc_difficulty_data_r2_%s.csv",
          gsub("\\.", "p", as.character(target_annotation_r2)))
)

cat("Job dir:", job_dir, "\n")
cat("Consolidated dir:", consolidated_dir, "\n")
cat("Config path:", cfg_path, "\n")
cat("Output CSV:", out_csv, "\n")
```

## 2. Load Data

```{r load}
stopifnot(dir.exists(consolidated_dir))
stopifnot(file.exists(cfg_path))

cfg <- test_susine:::load_job_config(cfg_path)

bundle_lookup <- tibble::as_tibble(cfg$tables$dataset_bundles) %>%
  dplyr::mutate(dataset_bundle_id = as.integer(.data$dataset_bundle_id)) %>%
  dplyr::distinct(.data$dataset_bundle_id, .keep_all = TRUE)

confusion_bins <- readr::read_csv(
  file.path(consolidated_dir, "confusion_bins_full.csv"),
  show_col_types = FALSE
) %>%
  dplyr::mutate(dataset_bundle_id = as.integer(.data$dataset_bundle_id))

dataset_metrics <- readr::read_csv(
  file.path(consolidated_dir, "dataset_metrics.csv"),
  show_col_types = FALSE
) %>%
  dplyr::mutate(dataset_bundle_id = as.integer(.data$dataset_bundle_id))

cat("Bundle lookup rows:", nrow(bundle_lookup), "\n")
cat("Confusion rows:", nrow(confusion_bins), "\n")
cat("Dataset metric rows:", nrow(dataset_metrics), "\n")
```

## 3. Helpers

```{r helpers}
add_bundle_cols <- function(df, cols) {
  if (!"dataset_bundle_id" %in% names(df)) {
    stop("Expected dataset_bundle_id in input data.")
  }
  cols <- intersect(cols, names(bundle_lookup))
  if (!length(cols)) {
    return(df)
  }
  available <- setdiff(cols, names(df))
  if (!length(available)) {
    return(df)
  }
  df %>%
    dplyr::left_join(
      bundle_lookup %>% dplyr::select(.data$dataset_bundle_id, dplyr::all_of(available)),
      by = "dataset_bundle_id"
    )
}

mean_or_na <- function(x) {
  if (all(is.na(x))) {
    return(NA_real_)
  }
  mean(x, na.rm = TRUE)
}

extract_method_bins <- function(data,
                                method_label,
                                spec_name,
                                use_case_id,
                                agg_method = NULL,
                                annotation_r2 = NULL) {
  out <- data %>%
    dplyr::filter(
      .data$spec_name == !!spec_name,
      .data$use_case_id == !!use_case_id
    )

  if (is.null(agg_method)) {
    if ("agg_method" %in% names(out)) {
      out <- out %>% dplyr::filter(is.na(.data$agg_method))
    }
  } else {
    out <- out %>% dplyr::filter(.data$agg_method == !!agg_method)
  }

  if (!is.null(annotation_r2)) {
    out <- out %>% dplyr::filter(.data$annotation_r2 == !!annotation_r2)
  }

  out %>%
    dplyr::mutate(method_label = method_label)
}
```

## 4. Join Matrix Metadata

```{r join-metadata}
bundle_cols <- c("matrix_id", "data_scenario", "y_noise", "p_star", "phenotype_seed")

confusion_bins <- confusion_bins %>%
  add_bundle_cols(bundle_cols)

dataset_metrics <- dataset_metrics %>%
  add_bundle_cols(bundle_cols) %>%
  dplyr::distinct(.data$dataset_bundle_id, .keep_all = TRUE)

if (!"matrix_id" %in% names(confusion_bins) || all(is.na(confusion_bins$matrix_id))) {
  stop("matrix_id could not be attached to confusion_bins. Check dataset_bundles in the job config.")
}

if (!"matrix_id" %in% names(dataset_metrics) || all(is.na(dataset_metrics$matrix_id))) {
  stop("matrix_id could not be attached to dataset_metrics. Check dataset_bundles in the job config.")
}

cat("Unique matrix IDs in confusion bins:", dplyr::n_distinct(confusion_bins$matrix_id), "\n")
cat("Unique matrix IDs in dataset metrics:", dplyr::n_distinct(dataset_metrics$matrix_id), "\n")
```

## 5. Compute Matrix-Level AUPRC

```{r matrix-auprc}
method_bins <- dplyr::bind_rows(
  extract_method_bins(
    confusion_bins,
    method_label = "auprc_oracle_warm_start_baseline",
    spec_name = "truth-warm",
    use_case_id = "susine_vanilla"
  ),
  extract_method_bins(
    confusion_bins,
    method_label = "auprc_baseline",
    spec_name = "baseline-single",
    use_case_id = "susine_vanilla"
  ),
  extract_method_bins(
    confusion_bins,
    method_label = "auprc_ccs_cluster_weight",
    spec_name = "C-CS",
    use_case_id = "susine_functional_mu",
    agg_method = "cluster_weight",
    annotation_r2 = target_annotation_r2
  ),
  extract_method_bins(
    confusion_bins,
    method_label = "auprc_ccs_oracle",
    spec_name = "C-CS",
    use_case_id = "susine_functional_mu",
    agg_method = "oracle",
    annotation_r2 = target_annotation_r2
  )
)

expected_methods <- c(
  "auprc_oracle_warm_start_baseline",
  "auprc_baseline",
  "auprc_ccs_cluster_weight",
  "auprc_ccs_oracle"
)

missing_methods <- setdiff(expected_methods, unique(method_bins$method_label))
if (length(missing_methods)) {
  stop(
    "Missing confusion-bin rows for method(s): ",
    paste(missing_methods, collapse = ", "),
    ". Check spec filters and target_annotation_r2."
  )
}

cat("Rows retained for target methods:", nrow(method_bins), "\n")
cat("Methods present:", paste(sort(unique(method_bins$method_label)), collapse = ", "), "\n")

matrix_auprc_long <- test_susine::compute_auprc_from_confusion(
  method_bins,
  group_vars = c("matrix_id", "method_label")
)

matrix_auprc_wide <- matrix_auprc_long %>%
  tidyr::pivot_wider(
    id_cols = "matrix_id",
    names_from = "method_label",
    values_from = "AUPRC"
  )

kable(matrix_auprc_wide %>% dplyr::arrange(.data$matrix_id))
```

## 6. Aggregate Dataset Metrics to Matrix Level

```{r matrix-metrics}
metric_cols <- c(
  "M1",
  "z_topk_ratio",
  "z_max_abs",
  "z_count_abs_gt_3",
  "z_eff_signals"
)

missing_metric_cols <- setdiff(metric_cols, names(dataset_metrics))
if (length(missing_metric_cols)) {
  stop(
    "dataset_metrics is missing required columns: ",
    paste(missing_metric_cols, collapse = ", ")
  )
}

matrix_metrics <- dataset_metrics %>%
  dplyr::group_by(.data$matrix_id) %>%
  dplyr::summarise(
    data_scenario = dplyr::first(.data$data_scenario),
    n_dataset_bundles = dplyr::n_distinct(.data$dataset_bundle_id),
    dplyr::across(dplyr::all_of(metric_cols), mean_or_na),
    .groups = "drop"
  )

kable(matrix_metrics %>% dplyr::arrange(.data$matrix_id))
```

## 7. Build Final Export Table

```{r export}
matrix_level_df <- matrix_metrics %>%
  dplyr::left_join(matrix_auprc_wide, by = "matrix_id") %>%
  dplyr::mutate(
    warm_start_gain_vs_baseline =
      .data$auprc_oracle_warm_start_baseline - .data$auprc_baseline,
    ccs_oracle_gain_vs_baseline =
      .data$auprc_ccs_oracle - .data$auprc_baseline,
    ccs_cluster_weight_gain_vs_baseline =
      .data$auprc_ccs_cluster_weight - .data$auprc_baseline
  ) %>%
  dplyr::arrange(.data$matrix_id)

readr::write_csv(matrix_level_df, out_csv)

cat("Wrote:", out_csv, "\n")
cat("Rows:", nrow(matrix_level_df), "\n")
cat("Columns:", ncol(matrix_level_df), "\n")

kable(matrix_level_df)
```
