---
title: "Ensemble Pilot (cluster-metric sweep) — Run Control"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

A minimal pilot whose only job is to **fit and store per-fit PIPs** for a
single functional-grid ensemble (C-CS: c-grid x sigma-grid, K = 64) on 50 loci,
one prior-quality setting (annotation r2 = 0.3, noncausal inflation = 0.95),
one seed each. The expensive step (the 50 x 64 = 3200 SuSiNE fits) runs once on
SLURM. Every clustering-metric x threshold x Method-B aggregation choice is then
a cheap post-hoc sweep in `collect_results_workbook_ensemble_pilot.Rmd` — no
refitting. `write_snps_parquet = TRUE` is what makes that possible (per-fit PIPs
+ causal labels are stored; per-fit `elbo_final` comes from `model_metrics`).

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_ensemble_pilot.Rmd")
library(here)
library(devtools)
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 0. User Configuration

```{r user-config}
job_name <- "ensemble_pilot_cluster_metric"
email    <- "mgc5166@psu.edu"

# HPC resources per SLURM array task
slurm_time <- "03:59:59"
slurm_mem  <- "8G"

# Pilot dataset grid
n_matrices_target <- 50L        # 50 loci
seeds             <- 1L         # one phenotype seed per locus

# Single prior-quality cell
annotation_r2_level        <- 0.3
noncausal_inflation_factor <- 0.95

# Simulation design
L_grid            <- 10L
architecture_grid <- "susie2_oligogenic"
sigma_0_2_default <- 0.2
max_iter          <- 100L

# Task packing
task_unit            <- "dataset"
bundles_per_task     <- 4L
task_assignment_seed <- 42L

# Output: per-fit PIPs are the whole point. Confusion/scaling bins are NOT
# needed — all aggregation/AUPRC happens post-hoc in the collect workbook.
verbose_file_output          <- FALSE
write_snps_parquet           <- TRUE
write_snps_pip_only          <- TRUE   # lean: only run_id, snp_index, pip, causal
write_confusion_bins         <- FALSE
write_scaling_confusion_bins <- FALSE
write_tier_cs_metrics        <- FALSE

credible_set_rho <- 0.95
purity_threshold <- 0.50
softmax_temperature <- 1
jsd_threshold <- 0.15  # run-time default; irrelevant, sweep happens in collect

pip_breaks <- sort(unique(c(
  seq(0.00, 0.05, by = 0.005),
  seq(0.05, 0.10, by = 0.01),
  seq(0.10, 1.00, by = 0.02)
)))
z_top_k <- 10
```

# 1. Ensemble Spec (C-CS: c-grid x sigma-grid, K = 64)

```{r spec}
c_grid_8     <- seq(0, 1.5, length.out = 8)
sigma_grid_8 <- c(0.01, 0.03, 0.07, 0.1, 0.2, 0.4, 0.7, 1.0)
stopifnot(0.2 %in% sigma_grid_8)

spec_C_CS <- list(
  name = "C-CS",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 64L,
  c_grid_values = c_grid_8,
  sigma_0_2_grid_values = sigma_grid_8
)
all_specs <- list(spec_C_CS)
cat("Ensemble spec:", spec_C_CS$name, "K =", spec_C_CS$K, "\n")
```

# 2. Matrix Selection (M1-stratified, 50 loci from scenario_1)

```{r matrix-selection}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here("data", "sampled_simulated_genotypes",
                      "scenario_sampling_summary.csv")
)
cat("Total available genotype matrices:", nrow(data_matrix_catalog_full), "\n")

m1_summary_path <- here("data", "sampled_simulated_genotypes",
                        "scenario_sampling_summary_with_m1.csv")
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>%
      dplyr::select(matrix_id, matrix_path),
    base_dir = here()
  ) %>% dplyr::select(matrix_id, M1)
  catalog_with_m1 <- data_matrix_catalog_full %>%
    dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}
if (!"matrix_id" %in% names(catalog_with_m1)) {
  join_col <- intersect(c("dataset_label", "matrix_path"), names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(data_matrix_catalog_full %>%
                       dplyr::select(matrix_id, dplyr::all_of(join_col)), by = join_col)
}
catalog_with_m1 <- catalog_with_m1 %>%
  dplyr::filter(!is.na(M1)) %>% dplyr::arrange(M1)

n_target <- min(n_matrices_target, nrow(catalog_with_m1))
idx      <- unique(round(seq(1, nrow(catalog_with_m1), length.out = n_target)))
selected <- catalog_with_m1[idx, , drop = FALSE]
data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(dplyr::select(selected, matrix_id), by = "matrix_id")

cat("Matrices selected (M1-stratified):", nrow(data_matrix_catalog), "\n")
cat("Seeds:", paste(seeds, collapse = ", "), "\n")
```

# 3. Build Job Config

```{r build-config}
cfg <- make_job_config(
  job_name = job_name,
  HPC = TRUE,
  time = slurm_time,
  mem = slurm_mem,
  email = email,
  output_root = here("output"),
  grid_mode = "full",

  use_case_ids = "susine_functional_mu",
  exploration_specs = all_specs,

  L_grid = L_grid,
  y_noise_grid = NA_real_,
  p_star_grid = NA_integer_,
  architecture_grid = architecture_grid,
  seeds = seeds,

  prior_quality = tibble::tibble(
    annotation_r2 = annotation_r2_level,
    inflate_match = noncausal_inflation_factor
  ),

  sigma_0_2_scalars = sigma_0_2_default,

  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here("data", "sampled_simulated_genotypes",
                                  "scenario_sampling_summary.csv"),

  task_unit = task_unit,
  bundles_per_task = bundles_per_task,
  task_assignment_seed = task_assignment_seed,

  max_iter = max_iter,

  aggregation_methods = "elbo_softmax",   # placeholder; real sweep is post-hoc
  include_overall_pool = FALSE,
  softmax_temperature = softmax_temperature,

  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = verbose_file_output,
  write_snps_parquet = write_snps_parquet,
  write_snps_pip_only = write_snps_pip_only,
  write_confusion_bins = write_confusion_bins,
  write_tier_cs_metrics = write_tier_cs_metrics,
  write_scaling_confusion_bins = write_scaling_confusion_bins,

  metrics_settings = list(
    pip_breaks = pip_breaks,
    z_top_k = z_top_k,
    jsd_threshold = jsd_threshold
  )
)

runs <- cfg$tables$runs
cat("Total runs:", nrow(runs), "\n")
cat("Datasets (loci x seed):", length(unique(runs$dataset_bundle_id)), "\n")
cat("Fits per dataset (expect ~64):",
    round(nrow(runs) / length(unique(runs$dataset_bundle_id)), 1), "\n")
```

# 4. Write Artifacts

```{r write-artifacts}
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)
cat("Job config written to:", arts$job_config, "\n")
cat("SLURM script:", arts$slurm_script, "\n")
```

# 5. Submit

```{r submit, eval=FALSE}
# sbatch <slurm_script_path>
# Then collect + sweep with collect_results_workbook_ensemble_pilot.Rmd
```
