---
title: "Annotation Contamination Sensitivity -- Causal-block positive control"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

Builds the reviewer-requested annotation-contamination sensitivity job for the
Causal-block positive control arm. The job uses the standard oligogenic simulation substrate,
central annotation setting (`annotation_r2 = 0.3`, `inflate_match = 0.95`), a
4 x 4 C-CS grid, and a single annotation-free baseline.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_annotation_contamination_causal.Rmd")
library(here)
library(devtools)
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# Configuration

```{r user-config}
job_name <- "annotation_contamination_causal_positive"
email    <- "mgc5166@psu.edu"

slurm_time <- "11:59:59"
slurm_mem  <- "8G"

n_matrices_target <- 150L
seeds             <- 1:4

L_grid            <- 10L
y_noise_grid      <- NA_real_
p_star_grid       <- NA_integer_
architecture_grid <- "susie2_oligogenic"

annotation_r2 <- 0.3
inflate_match <- 0.95
lambda_grid <- c(0, 0.1, 0.2, 0.3, 0.5, 0.7, 1.0)
annotation_arm <- "causal"

c_grid_4 <- c(0.00, 0.43, 1.07, 1.50)
sigma_grid_4 <- c(0.01, 0.10, 0.40, 1.00)

max_iter <- 100L
bundles_per_task <- 4L
task_assignment_seed <- 42L
aggregation_methods <- "cluster_weight_credible"
include_overall_pool <- FALSE
softmax_temperature <- 1

credible_set_rho <- 0.95
purity_threshold <- 0.50
verbose_file_output <- FALSE
write_snps_parquet <- FALSE
write_confusion_bins <- TRUE
write_tier_cs_metrics <- FALSE
write_scaling_confusion_bins <- FALSE

pip_breaks <- sort(unique(c(
  seq(0.00, 0.05, by = 0.005),
  seq(0.05, 0.10, by = 0.01),
  seq(0.10, 1.00, by = 0.02)
)))
```

# Specs

```{r specs}
spec_C_CS_4x4 <- list(
  name = "C-CS-annotation-contamination",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 16L,
  c_grid_values = c_grid_4,
  sigma_0_2_grid_values = sigma_grid_4
)

spec_baseline_single <- list(
  name = "baseline-single",
  use_case_ids = "susine_vanilla",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

all_specs <- list(spec_C_CS_4x4, spec_baseline_single)
```

# Matrix Selection

```{r matrix-selection}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here("data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv")
)

m1_summary_path <- here("data", "sampled_simulated_genotypes", "scenario_sampling_summary_with_m1.csv")
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>% dplyr::select(matrix_id, matrix_path),
    base_dir = here()
  ) %>% dplyr::select(matrix_id, M1)
  catalog_with_m1 <- data_matrix_catalog_full %>% dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}
if (!"matrix_id" %in% names(catalog_with_m1)) {
  if ("gene" %in% names(catalog_with_m1)) {
    catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = gene)
  }
  join_col <- intersect(c("dataset_label", "matrix_path"), names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(
      data_matrix_catalog_full %>% dplyr::select(matrix_id, dplyr::all_of(join_col)),
      by = join_col
    )
}
catalog_with_m1 <- catalog_with_m1 %>% dplyr::filter(!is.na(M1)) %>% dplyr::arrange(M1)
idx <- unique(round(seq(1, nrow(catalog_with_m1), length.out = min(n_matrices_target, nrow(catalog_with_m1)))))
selected <- catalog_with_m1[idx, , drop = FALSE]
data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(dplyr::select(selected, matrix_id), by = "matrix_id")
```

# Build Job

```{r build-config}
prior_quality <- tidyr::expand_grid(
  annotation_r2 = annotation_r2,
  inflate_match = inflate_match,
  annotation_contamination_arm = annotation_arm,
  annotation_contamination_lambda = lambda_grid,
  annotation_contamination_shuffle_z = FALSE
)

cfg <- make_job_config(
  job_name = job_name,
  HPC = TRUE,
  time = slurm_time,
  mem = slurm_mem,
  email = email,
  output_root = here("output"),
  grid_mode = "full",
  use_case_ids = c("susine_vanilla", "susine_functional_mu"),
  exploration_specs = all_specs,
  L_grid = L_grid,
  y_noise_grid = y_noise_grid,
  p_star_grid = p_star_grid,
  architecture_grid = architecture_grid,
  seeds = seeds,
  prior_quality = prior_quality,
  sigma_0_2_scalars = 0.2,
  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here("data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"),
  task_unit = "dataset",
  bundles_per_task = bundles_per_task,
  task_assignment_seed = task_assignment_seed,
  max_iter = max_iter,
  aggregation_methods = aggregation_methods,
  include_overall_pool = include_overall_pool,
  softmax_temperature = softmax_temperature,
  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = verbose_file_output,
  write_snps_parquet = write_snps_parquet,
  write_confusion_bins = write_confusion_bins,
  write_prior_diagnostics = TRUE,
  write_tier_cs_metrics = write_tier_cs_metrics,
  write_scaling_confusion_bins = write_scaling_confusion_bins,
  metrics_settings = list(
    pip_breaks = pip_breaks,
    z_top_k = 10,
    jsd_threshold = 0.15
  )
)
```

# Budget

```{r budget}
runs <- cfg$tables$runs
cat("Total run rows:", nrow(runs), "\n")
cat("Datasets:", dplyr::n_distinct(runs$dataset_bundle_id), "\n")
runs %>%
  dplyr::group_by(spec_name, annotation_contamination_arm, annotation_contamination_lambda) %>%
  dplyr::summarise(n_runs = dplyr::n(), n_datasets = dplyr::n_distinct(dataset_bundle_id), .groups = "drop") %>%
  knitr::kable()
```

# Write Artifacts

```{r write-artifacts}
artifacts <- write_job_artifacts(
  cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)
artifacts
cat("Submit with:\n")
cat("sbatch", normalizePath(artifacts$slurm_script, winslash = "/", mustWork = FALSE), "\n")
```
# Reporting Gates

Before reporting this run, validate the collected outputs:

- `annotation_diagnostics.csv`: the contaminated block RMS relative error should be below `1e-8`, and the untouched block flag should be `TRUE` for every row.
- Lambda zero: compare the `annotation_contamination_lambda == 0` C-CS aggregate against the clean C-CS run with the same dataset bundles; pooled AUPRC and recall at precision 0.75 should match to numerical tolerance.
- `prior_diagnostics.csv`: fixed-c C-CS contaminated fits should have `c_l == 1`, confirming that `estimate_mu_0_scale_factor` did not run.
- Anchor diagnostic: `anchor_sigma_0_2_scalar` records which `c = 0` fit defines the low-PIP anchor set; the pipeline chooses the smallest `sigma_0_2_scalar` when multiple anchors are available.
