---
title: "Ensemble Scaling — Heritability Estimand Hotfix (Panel E)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Minimal re-run to regenerate **only** what Panel E (estimated heritability) of
the ensemble-scaling composite needs, with the corrected local-genetic-variance
estimand (`hg2_expected_pve`) and the per-dataset realized truth
(`true_hg2_realized`). See
`refs/decisions/heritability_estimand_decision_2026-06-09.md` (sec 5.4).

Scope vs the full ensemble job:

- Specs: **C-CS** (the SuSiNE ensemble, `c_grid_8 x sigma_grid_8`) and
  **baseline-single / `susine_vanilla`** (the SuSiE box). Nothing else.
- Annotation: **`annotation_r2 = 0.3`** only (the paper focus level).
- Same M1-stratified matrices + seeds as the full job, so the dataset population
  (and therefore the boxplot distribution) matches exactly.
- Scaling-confusion analysis OFF; aggregation trimmed to the two methods Panel E
  consumes (`max_elbo`, `cluster_weight_credible`).

Downstream: collect with `collect_results_workbook_ensemble_scaling.Rmd` (point
`parent_job_id` at this job), then regenerate Panel E via the prepare/visualize
paper workbooks pointed at this job and splice Panel E into the full composite
(other panels are not produced here).

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_ensemble_scaling_hg2_hotfix.Rmd")
library(here)
library(devtools)
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 0. User Configuration

```{r user-config}
# Fresh job_name so outputs do not collide with the full ensemble job.
job_name <- "ensemble_scaling_hg2_hotfix"
email    <- "mgc5166@psu.edu"

slurm_time <- "11:59:59"
slurm_mem  <- "8G"

# MUST match the full job for apples-to-apples datasets.
n_matrices_target <- 150L
seeds             <- 1:4

L_grid            <- 10L
y_noise_grid      <- NA_real_
p_star_grid       <- NA_integer_
architecture_grid <- "susie2_oligogenic"

# Focus annotation level only.
annotation_r2_levels <- c(0.3)
noncausal_inflation_factor <- 0.95

sigma_0_2_default <- 0.2
max_iter <- 100L

task_unit            <- "dataset"
bundles_per_task     <- 4L
task_assignment_seed <- 42L

# Only the two aggregations Panel E uses.
aggregation_methods  <- c("max_elbo", "cluster_weight_credible")
include_overall_pool <- FALSE
softmax_temperature  <- 1

credible_set_rho     <- 0.95
purity_threshold     <- 0.50
verbose_file_output  <- FALSE
write_snps_parquet   <- FALSE
write_confusion_bins <- TRUE
write_tier_cs_metrics <- FALSE

# Scaling analysis OFF (not needed for Panel E).
write_scaling_confusion_bins <- FALSE

pip_breaks <- sort(unique(c(
  seq(0.00, 0.05, by = 0.005),
  seq(0.05, 0.10, by = 0.01),
  seq(0.10, 1.00, by = 0.02)
)))
z_top_k       <- 10
jsd_threshold  <- 0.15
```

# 1. Grid + Specs (C-CS + vanilla baseline only)

```{r specs}
c_grid_8  <- seq(0, 1.5, length.out = 8)
sigma_grid_8 <- c(0.01, 0.03, 0.07, 0.1, 0.2, 0.4, 0.7, 1.0)
stopifnot(0.2 %in% sigma_grid_8)

# C-CS: SuSiNE ensemble (c x sigma grid), the panel-E ensemble + max-ELBO boxes.
spec_C_CS <- list(
  name = "C-CS",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 64L,
  c_grid_values = c_grid_8,
  sigma_0_2_grid_values = sigma_grid_8
)

# Baseline single SuSiE fit (the panel-E "SuSiE" box). vanilla only.
spec_baseline_single <- list(
  name = "baseline-single",
  use_case_ids = "susine_vanilla",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

all_specs <- list(spec_C_CS, spec_baseline_single)
cat("Specs:", paste(vapply(all_specs, `[[`, character(1), "name"), collapse = ", "), "\n")
```

# 2. Matrix Selection (identical to the full job)

```{r matrix-selection}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)

m1_summary_path <- here(
  "data", "sampled_simulated_genotypes", "scenario_sampling_summary_with_m1.csv"
)
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>%
      dplyr::select(matrix_id, matrix_path),
    base_dir = here()
  ) %>%
    dplyr::select(matrix_id, M1)
  catalog_with_m1 <- data_matrix_catalog_full %>%
    dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}
if (!"matrix_id" %in% names(catalog_with_m1)) {
  if ("gene" %in% names(catalog_with_m1)) {
    catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = gene)
  }
  join_col <- intersect(c("dataset_label", "matrix_path"), names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(
      data_matrix_catalog_full %>% dplyr::select(matrix_id, dplyr::all_of(join_col)),
      by = join_col
    )
}
catalog_with_m1 <- catalog_with_m1 %>% dplyr::filter(!is.na(M1)) %>% dplyr::arrange(M1)

n_target <- min(n_matrices_target, nrow(catalog_with_m1))
idx      <- unique(round(seq(1, nrow(catalog_with_m1), length.out = n_target)))
selected <- catalog_with_m1[idx, , drop = FALSE]
data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(dplyr::select(selected, matrix_id), by = "matrix_id")
cat("Matrices selected:", nrow(data_matrix_catalog), " Seeds:", paste(seeds, collapse=","), "\n")
```

# 3. Build Job Config

```{r build-config}
cfg <- make_job_config(
  job_name = job_name,
  HPC = TRUE,
  time = slurm_time,
  mem = slurm_mem,
  email = email,
  output_root = here("output"),
  grid_mode = "full",

  use_case_ids = c("susine_vanilla", "susine_functional_mu"),
  exploration_specs = all_specs,

  L_grid = L_grid,
  y_noise_grid = y_noise_grid,
  p_star_grid = p_star_grid,
  architecture_grid = architecture_grid,
  seeds = seeds,

  prior_quality = tibble::tibble(
    annotation_r2 = annotation_r2_levels,
    inflate_match = rep(noncausal_inflation_factor, length(annotation_r2_levels))
  ),
  sigma_0_2_scalars = sigma_0_2_default,

  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),

  task_unit = task_unit,
  bundles_per_task = bundles_per_task,
  task_assignment_seed = task_assignment_seed,
  max_iter = max_iter,

  aggregation_methods = aggregation_methods,
  include_overall_pool = include_overall_pool,
  softmax_temperature = softmax_temperature,

  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = verbose_file_output,
  write_snps_parquet = write_snps_parquet,
  write_confusion_bins = write_confusion_bins,
  write_tier_cs_metrics = write_tier_cs_metrics,
  write_scaling_confusion_bins = write_scaling_confusion_bins,

  metrics_settings = list(
    pip_breaks = pip_breaks,
    z_top_k = z_top_k,
    jsd_threshold = jsd_threshold
  )
)
```

# 4. Budget Summary

```{r budget}
runs <- cfg$tables$runs
cat("Total runs (gross):", nrow(runs), "\n")
cat("Datasets:", length(unique(runs$dataset_bundle_id)), "\n")
if ("spec_name" %in% names(runs)) {
  runs %>%
    dplyr::group_by(spec_name) %>%
    dplyr::summarise(
      n_runs = dplyr::n(),
      n_use_cases = dplyr::n_distinct(use_case_id),
      n_datasets = dplyr::n_distinct(dataset_bundle_id),
      .groups = "drop"
    ) %>%
    knitr::kable(caption = "Runs per spec (hotfix)")
}
```

# 5. Write Artifacts

```{r write-artifacts, eval=FALSE}
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)
cat("Job config written to:", arts$job_config, "\n")
cat("SLURM script:", arts$slurm_script, "\n")
```

# 6. Submit

```{r submit, eval=FALSE}
# sbatch <slurm_script_path>
```
