---
title: "Ensemble Scaling Experiment — Run Control"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Experiment plan: `refs/ensemble_scaling_experiment_plan.md`

Builds a single heterogeneous job config with all specs (A-R through
C-CSFR, plus C-CSR) plus standalone truth-warm-start baselines.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_ensemble_scaling.Rmd")
library(here)
here()

library(devtools)
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 0. User Configuration

All run-level knobs in one place. Edit these before knitting.

```{r user-config}
# ── Job identity ──
job_name <- "ensemble_scaling_full"
email    <- "mgc5166@psu.edu"

# ── HPC resources (per SLURM array task) ──
slurm_time <- "05:59:59"
slurm_mem  <- "8G"

# ── Dataset grid ──
n_matrices_target <- 150L  # M1-stratified genotype matrices
seeds             <- 1:4   # phenotype seeds per matrix

# ── Simulation design ──
L_grid            <- 10L
y_noise_grid      <- NA_real_
p_star_grid       <- NA_integer_
architecture_grid <- "susie2_oligogenic"

# ── Annotation quality levels ──
annotation_r2_levels <- c(0, 0.2, 0.5)

# ── Prior variance default (for non-gridded runs) ──
sigma_0_2_default <- 0.2

# ── Model fitting ──
max_iter <- 100L

# ── Task packing ──
task_unit            <- "dataset"
bundles_per_task     <- 4L
task_assignment_seed <- 42L

# ── Aggregation ──
aggregation_methods  <- c("max_elbo", "uniform", "elbo_softmax",
                          "cluster_weight", "cluster_weight_jsd_050")
include_overall_pool <- FALSE
softmax_temperature  <- 1

# ── Output / metrics ──
credible_set_rho     <- 0.95
purity_threshold     <- 0.50
verbose_file_output  <- FALSE
write_snps_parquet   <- FALSE
write_confusion_bins <- TRUE
write_tier_cs_metrics <- TRUE

# ── Scaling analysis ──
write_scaling_confusion_bins <- TRUE
scaling_n_ens_sizes          <- c(4L, 8L, 16L, 32L, 64L)
scaling_restart_reps         <- 50L

# Variable-resolution PIP breaks:
#   0.02 steps: 1.00 to 0.10 (46 breaks)
#   0.01 steps: 0.10 to 0.05 (5 breaks)
#   0.005 steps: 0.05 to 0.00 (10 breaks)
pip_breaks <- sort(unique(c(
  seq(0.00, 0.05, by = 0.005),
  seq(0.05, 0.10, by = 0.01),
  seq(0.10, 1.00, by = 0.02)
)))
z_top_k       <- 10
jsd_threshold  <- 0.15
```

# 1. Grid Value Definitions

```{r grid-values}
# c_grid values: seq(0, 1.5, length.out = N)
c_grid_64 <- seq(0, 1.5, length.out = 64)
c_grid_8  <- seq(0, 1.5, length.out = 8)
c_grid_4  <- seq(0, 1.5, length.out = 4)
c_grid_3  <- seq(0, 1.5, length.out = 3)

# sigma_0_2 grid values: hand-picked with 0.2 ensured
sigma_grid_64 <- {
  v <- 10^seq(-2, 0, length.out = 64)
  v[which.min(abs(v - 0.2))] <- 0.2
  sort(unique(v))
}
sigma_grid_8 <- c(0.01, 0.03, 0.07, 0.1, 0.2, 0.4, 0.7, 1.0)
sigma_grid_4 <- c(0.02, 0.1, 0.2, 0.5)
sigma_grid_3 <- c(0.05, 0.2, 1.0)

# Verify 0.2 is included in all grids
stopifnot(0.2 %in% sigma_grid_64)
stopifnot(0.2 %in% sigma_grid_8)
stopifnot(0.2 %in% sigma_grid_4)
stopifnot(0.2 %in% sigma_grid_3)
```

# 2. Exploration Specs

Each spec defines one row in the method catalog from the experiment plan.

```{r exploration-specs}
# Shared restart / refine settings
restart_64 <- list(n_inits = 64L, alpha_concentration = 0.001)
restart_8  <- list(n_inits = 8L,  alpha_concentration = 0.001)
restart_4  <- list(n_inits = 4L,  alpha_concentration = 0.001)
restart_3  <- list(n_inits = 3L,  alpha_concentration = 0.001)

refine_32 <- list(n_steps = 32L, cs_source = "filtered", purity_threshold = 0.95)
refine_8  <- list(n_steps = 8L,  cs_source = "filtered", purity_threshold = 0.95)
refine_4  <- list(n_steps = 4L,  cs_source = "filtered", purity_threshold = 0.95)
refine_3  <- list(n_steps = 3L,  cs_source = "filtered", purity_threshold = 0.95)
refine_2  <- list(n_steps = 2L,  cs_source = "filtered", purity_threshold = 0.95)

# ── Group A: Baseline SuSiE (no annotations) ──
spec_A_R <- list(
  name = "A-R",
  use_case_ids = "susine_vanilla",
  exploration_methods = "restart",
  exploration_mode = "separate",
  K = 64L,
  restart_settings = restart_64
)

spec_A_F <- list(
  name = "A-F",
  use_case_ids = "susine_vanilla",
  exploration_methods = "refine",
  exploration_mode = "separate",
  K = 32L,
  refine_settings = refine_32
)

spec_A_S <- list(
  name = "A-S",
  use_case_ids = "susine_vanilla",
  exploration_methods = "sigma_0_2_grid",
  exploration_mode = "separate",
  K = 64L,
  sigma_0_2_grid_values = sigma_grid_64
)

spec_A_RF <- list(
  name = "A-RF",
  use_case_ids = "susine_vanilla",
  exploration_methods = c("restart", "refine"),
  exploration_mode = "intersect",
  K = 64L,
  restart_settings = restart_8,
  refine_settings = refine_4   # 8 restarts × 4 refine steps × 2 fits/step ≈ 64 fits
)

spec_A_RS <- list(
  name = "A-RS",
  use_case_ids = "susine_vanilla",
  exploration_methods = c("restart", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 64L,
  restart_settings = restart_8,
  sigma_0_2_grid_values = sigma_grid_8
)

spec_A_FS <- list(
  name = "A-FS",
  use_case_ids = "susine_vanilla",
  exploration_methods = c("refine", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 64L,
  refine_settings = refine_4,   # 8 sigma × 4 refine steps × 2 fits/step ≈ 64 fits
  sigma_0_2_grid_values = sigma_grid_8
)

spec_A_RFS <- list(
  name = "A-RFS",
  use_case_ids = "susine_vanilla",
  exploration_methods = c("restart", "refine", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 64L,
  restart_settings = restart_4,
  refine_settings = refine_2,   # 4 restarts × 4 sigma × 2 refine steps × 2 fits/step = 64 fits
  sigma_0_2_grid_values = sigma_grid_4
)

# ── Group B: EB SuSiNE (annotations via EB) ──
spec_B_R <- list(
  name = "B-R",
  use_case_ids = "susine_eb_clamped_scale_var_nonneg",
  exploration_methods = "restart",
  exploration_mode = "separate",
  K = 64L,
  restart_settings = restart_64
)

spec_B_F <- list(
  name = "B-F",
  use_case_ids = "susine_eb_clamped_scale_var_nonneg",
  exploration_methods = "refine",
  exploration_mode = "separate",
  K = 32L,
  refine_settings = refine_32
)

spec_B_RF <- list(
  name = "B-RF",
  use_case_ids = "susine_eb_clamped_scale_var_nonneg",
  exploration_methods = c("restart", "refine"),
  exploration_mode = "intersect",
  K = 64L,
  restart_settings = restart_8,
  refine_settings = refine_4   # 8 restarts × 4 refine steps × 2 fits/step ≈ 64 fits
)

# ── Group C: Grid SuSiNE (annotations, no EB) ──
spec_C_C <- list(
  name = "C-C",
  use_case_ids = "susine_functional_mu",
  exploration_methods = "c_grid",
  exploration_mode = "separate",
  K = 64L,
  c_grid_values = c_grid_64
)

spec_C_CS <- list(
  name = "C-CS",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 64L,
  c_grid_values = c_grid_8,
  sigma_0_2_grid_values = sigma_grid_8
)

spec_C_CSR <- list(
  name = "C-CSR",
  use_case_ids = "susine_functional_mu",
  exploration_methods = "cs_grid_refit",
  exploration_mode = "separate",
  K = 64L,
  c_grid_values = c_grid_8,
  sigma_0_2_grid_values = sigma_grid_8
)

spec_C_CSFR <- list(
  name = "C-CSFR",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid", "refine", "restart"),
  exploration_mode = "intersect",
  K = 81L,
  c_grid_values = c_grid_3,
  sigma_0_2_grid_values = sigma_grid_3,
  refine_settings = refine_3,
  restart_settings = restart_3
)

spec_C_CFR <- list(
  name = "C-CFR",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "refine", "restart"),
  exploration_mode = "intersect",
  K = 64L,
  c_grid_values = c_grid_4,
  refine_settings = refine_2,   # 4 c × 4 restart × 2 refine steps × 2 fits/step = 64 fits
  restart_settings = restart_4
)

spec_C_CF <- list(
  name = "C-CF",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "refine"),
  exploration_mode = "intersect",
  K = 64L,
  c_grid_values = c_grid_8,
  refine_settings = refine_4    # 8 c × 4 refine steps × 2 fits/step = 64 fits
  # sigma fixed to 0.2 via global sigma_0_2_scalars; no restart axis
)

# ── Standalone single-fit baselines ──
spec_baseline_single <- list(
  name = "baseline-single",
  use_case_ids = c("susine_vanilla", "susine_eb_clamped_scale_var_nonneg",
                   "susine_functional_mu"),
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

all_specs <- list(
  spec_A_R, spec_A_F, spec_A_S,
  spec_A_RF, spec_A_RS, spec_A_FS, spec_A_RFS,
  spec_B_R, spec_B_F, spec_B_RF,
  spec_C_C, spec_C_CS, spec_C_CSR, spec_C_CSFR, spec_C_CFR, spec_C_CF,
  spec_baseline_single
)

expected_spec_names <- c(
  "A-R", "A-F", "A-S", "A-RF", "A-RS", "A-FS", "A-RFS",
  "B-R", "B-F", "B-RF",
  "C-C", "C-CS", "C-CSR", "C-CSFR", "C-CFR", "C-CF",
  "baseline-single"
)
actual_spec_names <- vapply(all_specs, function(s) s$name, character(1))
stopifnot(setequal(actual_spec_names, expected_spec_names))

cat("Total specs:", length(all_specs), "\n")
for (s in all_specs) {
  cat(sprintf("  %-20s K=%-4d methods=%s\n",
              s$name, s$K, paste(s$exploration_methods, collapse="x")))
}
```

# 3. Build Job Config

```{r matrix-selection}
# n_matrices_target and seeds are set in Section 0 (User Configuration).

data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)
cat("Total available genotype matrices:", nrow(data_matrix_catalog_full), "\n")

# M1-stratified selection: sort by M1, pick evenly-spaced matrices
m1_summary_path <- here(
  "data", "sampled_simulated_genotypes", "scenario_sampling_summary_with_m1.csv"
)
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>%
      dplyr::select(matrix_id, matrix_path),
    base_dir = here()
  ) %>%
    dplyr::select(matrix_id, M1)
  catalog_with_m1 <- data_matrix_catalog_full %>%
    dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}

# Recover matrix_id if the cached CSV omitted it (handles gene / dataset_label variants)
if (!"matrix_id" %in% names(catalog_with_m1)) {
  if ("gene" %in% names(catalog_with_m1)) {
    catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = gene)
  }
  join_col <- intersect(c("dataset_label", "matrix_path"),
                        names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(
      data_matrix_catalog_full %>%
        dplyr::select(matrix_id, dplyr::all_of(join_col)),
      by = join_col
    )
}

catalog_with_m1 <- catalog_with_m1 %>%
  dplyr::filter(!is.na(M1)) %>%
  dplyr::arrange(M1)

n_target <- min(n_matrices_target, nrow(catalog_with_m1))
idx      <- unique(round(seq(1, nrow(catalog_with_m1), length.out = n_target)))
selected <- catalog_with_m1[idx, , drop = FALSE]

data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(dplyr::select(selected, matrix_id), by = "matrix_id")

cat("Matrices selected (M1-stratified):", nrow(data_matrix_catalog), "\n")
cat("Seeds:", paste(seeds, collapse = ", "), "\n")
```

```{r build-config}

cfg <- make_job_config(
  job_name = job_name,
  HPC = TRUE,
  time = slurm_time,
  mem = slurm_mem,
  email = email,
  output_root = here("output"),
  grid_mode = "full",

  # Use case IDs (union across all specs — overridden per spec)
  use_case_ids = c("susine_vanilla",
                   "susine_eb_clamped_scale_var_nonneg",
                   "susine_functional_mu"),

  # Heterogeneous exploration specs
  exploration_specs = all_specs,

  # Dataset grid
  L_grid = L_grid,
  y_noise_grid = y_noise_grid,
  p_star_grid = p_star_grid,
  architecture_grid = architecture_grid,
  seeds = seeds,

  # Annotation quality
  prior_quality = tibble::tibble(
    annotation_r2 = annotation_r2_levels,
    inflate_match = rep(1, length(annotation_r2_levels))
  ),

  # Prior variance (default for non-gridded runs)
  sigma_0_2_scalars = sigma_0_2_default,

  # Data source
  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),

  # Task packing
  task_unit = task_unit,
  bundles_per_task = bundles_per_task,
  task_assignment_seed = task_assignment_seed,

  # Model settings
  max_iter = max_iter,

  # Aggregation
  aggregation_methods = aggregation_methods,
  include_overall_pool = include_overall_pool,
  softmax_temperature = softmax_temperature,

  # Output flags
  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = verbose_file_output,
  write_snps_parquet = write_snps_parquet,
  write_confusion_bins = write_confusion_bins,
  write_tier_cs_metrics = write_tier_cs_metrics,

  # Scaling analysis
  write_scaling_confusion_bins = write_scaling_confusion_bins,
  scaling_n_ens_sizes = scaling_n_ens_sizes,
  scaling_restart_reps = scaling_restart_reps,

  # Metrics
  metrics_settings = list(
    pip_breaks = pip_breaks,
    z_top_k = z_top_k,
    jsd_threshold = jsd_threshold
  )
)
```

# 3b. Append Truth-Warm-Start Standalone Baselines

These are single-run benchmarks warm-started on the true top-10 causal
variants (one variant per effect). Not part of the 4-lever scaling grids.
Added post-hoc to the run table.

```{r truth-warm-baselines}
# Build truth-warm rows for each dataset bundle x use case combo.
# Three baselines:
# 1. susine_vanilla + truth warm start
# 2. susine_eb_clamped_scale_var_nonneg + truth warm start
# 3. susine_functional_mu (c=0.5, sigma_0_2=0.2) + truth warm start

runs <- cfg$tables$runs
bundles <- cfg$tables$dataset_bundles

truth_warm_use_cases <- tibble::tribble(
  ~use_case_id,                          ~c_value, ~sigma_0_2_scalar,
  "susine_vanilla",                      NA_real_, "0.2",
  "susine_eb_clamped_scale_var_nonneg",  NA_real_, NA_character_,
  "susine_functional_mu",               0.5,       "0.2"
)

# Template from existing runs for column structure
template <- runs[1, , drop = FALSE]

truth_warm_rows <- list()
for (bi in seq_len(nrow(bundles))) {
  bundle <- bundles[bi, , drop = FALSE]
  for (ti in seq_len(nrow(truth_warm_use_cases))) {
    tw <- truth_warm_use_cases[ti, , drop = FALSE]
    # For annotated use cases, expand across annotation R-squared levels
    uc_spec <- dplyr::filter(cfg$tables$use_cases,
                              prior_spec_id == tw$use_case_id)
    needs_ann <- isTRUE(uc_spec$supports_annotation[1])

    if (needs_ann) {
      ann_levels <- c(0, 0.2, 0.5)
    } else {
      ann_levels <- NA_real_
    }
    for (ann_r2 in ann_levels) {
      row <- template
      row$dataset_bundle_id <- bundle$dataset_bundle_id
      row$use_case_id <- tw$use_case_id
      row$prior_spec_id <- tw$use_case_id
      row$L <- 10L
      row$phenotype_seed <- bundle$phenotype_seed
      row$exploration_methods <- "single"
      row$exploration_mode <- "separate"
      row$run_type <- "warm"
      row$warm_method <- "truth_warm"
      row$restart_id <- NA_integer_
      row$refine_step <- NA_integer_
      row$c_value <- tw$c_value
      row$sigma_0_2_scalar <- tw$sigma_0_2_scalar
      row$annotation_r2 <- ann_r2
      row$inflate_match <- if (is.na(ann_r2)) NA_real_ else 1
      row$spec_name <- "truth-warm"
      row$group_key <- paste0("L=10|r2=", ann_r2, "|inflate=",
                              if (is.na(ann_r2)) "NA" else "1")
      truth_warm_rows[[length(truth_warm_rows) + 1L]] <- row
    }
  }
}

truth_warm_tbl <- dplyr::bind_rows(truth_warm_rows)
# Assign run_ids continuing from existing max
max_run_id <- max(runs$run_id, na.rm = TRUE)
truth_warm_tbl$run_id <- seq(max_run_id + 1L, length.out = nrow(truth_warm_tbl))

runs <- dplyr::bind_rows(runs, truth_warm_tbl)
cfg$tables$runs <- runs
cat("Truth-warm baseline rows added:", nrow(truth_warm_tbl), "\n")
cat("Total runs after truth-warm:", nrow(runs), "\n")
```

# 4. Budget Summary

```{r budget-summary}
runs <- cfg$tables$runs

cat("Total runs (gross):", nrow(runs), "\n")
cat("Datasets:", length(unique(runs$dataset_bundle_id)), "\n")

if ("spec_name" %in% names(runs)) {
  runs %>%
    dplyr::group_by(spec_name) %>%
    dplyr::summarise(
      n_runs = dplyr::n(),
      n_use_cases = dplyr::n_distinct(use_case_id),
      n_datasets = dplyr::n_distinct(dataset_bundle_id),
      .groups = "drop"
    ) %>%
    knitr::kable(caption = "Runs per spec")
}
```

# 5. Write Artifacts

```{r write-artifacts, eval=FALSE}
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)
cat("Job config written to:", arts$job_config, "\n")
cat("SLURM script:", arts$slurm_script, "\n")
```

# 6. Submit

```{r submit, eval=FALSE}
# sbatch <slurm_script_path>
```
