---
title: "Ensemble Scaling — Smoke Test"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Minimal smoke test: 2 matrices, 1 seed, 1 R², small K values. Tests
the full pipeline end-to-end: heterogeneous specs, two-step refinement,
sigma_0_2_grid, truth-warm baselines, all aggregation methods.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_ensemble_scaling_smoke.Rmd")
library(here)
here()

library(devtools)
devtools::load_all(".")

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
} else {
  stop("susine package not found at: ", susine_path)
}

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 1. Smoke Test Grid Values

Small grids (K=4) to keep runtime under a few minutes.

```{r grid-values}
c_grid_4 <- seq(0, 1.5, length.out = 4)    # 0, 0.5, 1.0, 1.5
c_grid_2 <- seq(0, 1.5, length.out = 2)    # 0, 1.5

sigma_grid_4 <- c(0.02, 0.1, 0.2, 0.5)
sigma_grid_2 <- c(0.1, 0.2)

stopifnot(0.2 %in% sigma_grid_4)
stopifnot(0.2 %in% sigma_grid_2)

cat("c_grid_4:", c_grid_4, "\n")
cat("sigma_grid_4:", sigma_grid_4, "\n")
```

# 2. Smoke Test Specs

Subset of the full experiment: 1 pure spec per group + 1 interaction +
baseline + truth-warm. Tests every code path.

```{r specs}
restart_4 <- list(n_inits = 4L, alpha_concentration = 0.001)
restart_2 <- list(n_inits = 2L, alpha_concentration = 0.001)

refine_2 <- list(n_steps = 2L, cs_source = "filtered", purity_threshold = 0.95)

# ── Group A: one pure lever + one interaction ──
spec_A_R <- list(
  name = "A-R",
  use_case_ids = "susine_vanilla",
  exploration_methods = "restart",
  exploration_mode = "separate",
  K = 4L,
  restart_settings = restart_4
)

spec_A_F <- list(
  name = "A-F",
  use_case_ids = "susine_vanilla",
  exploration_methods = "refine",
  exploration_mode = "separate",
  K = 2L,
  refine_settings = refine_2
)

spec_A_S <- list(
  name = "A-S",
  use_case_ids = "susine_vanilla",
  exploration_methods = "sigma_0_2_grid",
  exploration_mode = "separate",
  K = 4L,
  sigma_0_2_grid_values = sigma_grid_4
)

spec_A_RF <- list(
  name = "A-RF",
  use_case_ids = "susine_vanilla",
  exploration_methods = c("restart", "refine"),
  exploration_mode = "intersect",
  K = 4L,
  restart_settings = restart_2,
  refine_settings = refine_2
)

# ── Group B: one pure lever ──
spec_B_R <- list(
  name = "B-R",
  use_case_ids = "susine_eb_clamped_scale_var_nonneg",
  exploration_methods = "restart",
  exploration_mode = "separate",
  K = 4L,
  restart_settings = restart_4
)

# ── Group C: one pure lever + one interaction ──
spec_C_C <- list(
  name = "C-C",
  use_case_ids = "susine_functional_mu",
  exploration_methods = "c_grid",
  exploration_mode = "separate",
  K = 4L,
  c_grid_values = c_grid_4
)

spec_C_CS <- list(
  name = "C-CS",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 4L,
  c_grid_values = c_grid_2,
  sigma_0_2_grid_values = sigma_grid_2
)

# ── Single-fit baselines ──
spec_baseline <- list(
  name = "baseline-single",
  use_case_ids = c("susine_vanilla", "susine_eb_clamped_scale_var_nonneg",
                   "susine_functional_mu"),
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

smoke_specs <- list(
  spec_A_R, spec_A_F, spec_A_S, spec_A_RF,
  spec_B_R,
  spec_C_C, spec_C_CS,
  spec_baseline
)

cat("Smoke test specs:\n")
for (s in smoke_specs) {
  cat(sprintf("  %-20s K=%-4d methods=%s\n",
              s$name, s$K, paste(s$exploration_methods, collapse="x")))
}
```

# 3. Build Smoke Test Config

```{r build-config}
# Use first 2 matrices only
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)
cat("Full catalog:", nrow(data_matrix_catalog_full), "matrices\n")

# Take first 2
data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::slice_head(n = 2)
cat("Smoke test matrices:", nrow(data_matrix_catalog), "\n")

cfg <- make_job_config(
  job_name = "ensemble_scaling_smoke",
  HPC = FALSE,
  time = "00:30:00",
  mem = "4G",
  email = "mgc5166@psu.edu",
  output_root = here("output"),
  grid_mode = "full",

  use_case_ids = c("susine_vanilla",
                   "susine_eb_clamped_scale_var_nonneg",
                   "susine_functional_mu"),

  exploration_specs = smoke_specs,

  # Minimal dataset grid: 2 matrices, 1 seed, 1 architecture
  L_grid = 10L,
  y_noise_grid = NA_real_,
  p_star_grid = NA_integer_,
  architecture_grid = "susie2_oligogenic",
  seeds = 1L,

  # Single annotation R-squared level for smoke
  prior_quality = tibble::tibble(
    annotation_r2 = 0.2,
    inflate_match = 1
  ),

  sigma_0_2_scalars = 0.2,

  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),

  task_unit = "dataset",
  bundles_per_task = 2L,
  task_assignment_seed = 1L,

  # Fast convergence for smoke test
  max_iter = 50L,

  aggregation_methods = c("max_elbo", "uniform", "elbo_softmax", "cluster_weight", "cluster_weight_jsd_050"),
  include_overall_pool = FALSE,
  softmax_temperature = 1,

  credible_set_rho = 0.95,
  purity_threshold = 0.50,
  verbose_file_output = FALSE,
  write_snps_parquet = TRUE,
  write_confusion_bins = TRUE,
  write_tier_cs_metrics = TRUE,

  metrics_settings = list(
    pip_breaks = sort(unique(c(
      seq(0.00, 0.05, by = 0.005),
      seq(0.05, 0.10, by = 0.01),
      seq(0.10, 1.00, by = 0.02)
    ))),
    z_top_k = 10,
    jsd_threshold = 0.15
  )
)
```

# 3b. Truth-Warm Baselines

```{r truth-warm}
runs <- cfg$tables$runs
bundles <- cfg$tables$dataset_bundles

truth_warm_use_cases <- tibble::tribble(
  ~use_case_id,                          ~c_value, ~sigma_0_2_scalar,
  "susine_vanilla",                      NA_real_, "0.2",
  "susine_eb_clamped_scale_var_nonneg",  NA_real_, NA_character_,
  "susine_functional_mu",               0.5,       "0.2"
)

template <- runs[1, , drop = FALSE]

truth_warm_rows <- list()
for (bi in seq_len(nrow(bundles))) {
  bundle <- bundles[bi, , drop = FALSE]
  for (ti in seq_len(nrow(truth_warm_use_cases))) {
    tw <- truth_warm_use_cases[ti, , drop = FALSE]
    uc_spec <- dplyr::filter(cfg$tables$use_cases,
                              prior_spec_id == tw$use_case_id)
    needs_ann <- isTRUE(uc_spec$supports_annotation[1])
    ann_levels <- if (needs_ann) 0.2 else NA_real_

    for (ann_r2 in ann_levels) {
      row <- template
      row$dataset_bundle_id <- bundle$dataset_bundle_id
      row$use_case_id <- tw$use_case_id
      row$prior_spec_id <- tw$use_case_id
      row$L <- 10L
      row$phenotype_seed <- bundle$phenotype_seed
      row$exploration_methods <- "single"
      row$exploration_mode <- "separate"
      row$run_type <- "warm"
      row$warm_method <- "truth_warm"
      row$restart_id <- NA_integer_
      row$refine_step <- NA_integer_
      row$c_value <- tw$c_value
      row$sigma_0_2_scalar <- tw$sigma_0_2_scalar
      row$annotation_r2 <- ann_r2
      row$inflate_match <- if (is.na(ann_r2)) NA_real_ else 1
      row$spec_name <- "truth-warm"
      row$group_key <- paste0("L=10|r2=", ann_r2, "|inflate=",
                              if (is.na(ann_r2)) "NA" else "1")
      truth_warm_rows[[length(truth_warm_rows) + 1L]] <- row
    }
  }
}

truth_warm_tbl <- dplyr::bind_rows(truth_warm_rows)
max_run_id <- max(runs$run_id, na.rm = TRUE)
truth_warm_tbl$run_id <- seq(max_run_id + 1L, length.out = nrow(truth_warm_tbl))

runs <- dplyr::bind_rows(runs, truth_warm_tbl)
cfg$tables$runs <- runs
cat("Truth-warm rows added:", nrow(truth_warm_tbl), "\n")
cat("Total runs:", nrow(runs), "\n")
```

# 4. Budget Summary

```{r budget}
runs <- cfg$tables$runs

cat("\n=== SMOKE TEST BUDGET ===\n")
cat("Total runs:", nrow(runs), "\n")
cat("Datasets:", length(unique(runs$dataset_bundle_id)), "\n")

if ("spec_name" %in% names(runs)) {
  runs %>%
    dplyr::group_by(spec_name) %>%
    dplyr::summarise(
      n_runs = dplyr::n(),
      n_use_cases = dplyr::n_distinct(use_case_id),
      n_datasets = dplyr::n_distinct(dataset_bundle_id),
      .groups = "drop"
    ) %>%
    knitr::kable(caption = "Smoke test runs per spec")
}
```

# 5. Run Locally

Execute the smoke test directly (no SLURM). Runs all tasks sequentially.

```{r run-smoke, eval=FALSE}
# Write artifacts first
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)

# Run all tasks locally
n_tasks <- cfg$job$n_tasks
cat("Running", n_tasks, "tasks locally...\n")

for (tid in seq_len(n_tasks)) {
  cat("\n--- Task", tid, "of", n_tasks, "---\n")
  run_task(
    job_name = cfg$job$name,
    task_id = tid,
    job_root = here("output"),
    config_path = arts$job_config,
    quiet = FALSE
  )
}
cat("\nSmoke test complete.\n")
```

# 6. Quick Validation

```{r validate, eval=FALSE}
# Check that outputs were produced
staging_dir <- file.path(cfg$paths$temp_dir, "staging")
if (dir.exists(staging_dir)) {
  staging_files <- list.files(staging_dir, recursive = TRUE)
  cat("Staging files:", length(staging_files), "\n")
  cat("File types:\n")
  print(table(tools::file_ext(staging_files)))
} else {
  cat("No staging directory found — check for errors above.\n")
}

# Verify key outputs exist
model_metrics_files <- list.files(staging_dir, pattern = "model_metrics",
                                   recursive = TRUE)
cat("\nModel metrics files:", length(model_metrics_files), "\n")

confusion_files <- list.files(staging_dir, pattern = "confusion",
                               recursive = TRUE)
cat("Confusion bin files:", length(confusion_files), "\n")

snp_files <- list.files(staging_dir, pattern = "\\.parquet$",
                         recursive = TRUE)
cat("SNP parquet files:", length(snp_files), "\n")

# Check for per-spec coverage
if (length(model_metrics_files)) {
  mm_sample <- readr::read_csv(
    file.path(staging_dir, model_metrics_files[1]),
    show_col_types = FALSE
  )
  cat("\nSpecs represented in first flush file:\n")
  if ("spec_name" %in% names(mm_sample)) {
    print(table(mm_sample$spec_name))
  }
  cat("\nExplore methods:\n")
  if ("explore_method" %in% names(mm_sample)) {
    print(table(mm_sample$explore_method))
  }
}
```

# 7. Smoke Test Checks

```{r checks, eval=FALSE}
# After running, verify:
# 1. Every spec_name in the run table produced at least one output row
# 2. Refinement specs (A-F, A-RF) produced both perturb (not stored) and
#    refit (stored) — check that run count matches expected
# 3. sigma_0_2_grid runs (A-S, C-CS) have distinct sigma values
# 4. Truth-warm runs converged and produced reasonable PIPs
# 5. Aggregation methods all ran without error

cat("=== Smoke test validation checklist ===\n")
cat("[ ] All specs produced output rows\n")
cat("[ ] Refinement two-step (perturb+refit) worked\n")
cat("[ ] sigma_0_2_grid has distinct sigma values per run\n")
cat("[ ] Truth-warm baselines converged\n")
cat("[ ] All 4 aggregation methods produced results\n")
cat("[ ] PIP breaks match variable-resolution config\n")
```
