---
title: "Dirichlet Alpha Calibration Pilot"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Calibrate the Dirichlet concentration parameter (alpha) for
random restart pi sampling. Tests alpha = {0.001, 0.01, 0.1, 0.5, 1, 2, 5, 10}
with K=11 fits (1 baseline + 10 warm starts) per alpha, using
susine_vanilla with fixed sigma_0_2 = 0.2 on all 150 scenario_1
matrices x 3 seeds = 450 datasets. Architecture: SuSiE 2 paper
Scenario 1 (3 sparse / 5 oligogenic / 15 polygenic, h2_total = 0.25).

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_dirichlet_pilot.Rmd")
library(here)
here()

library(devtools)
devtools::load_all(".")

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 1. Pilot Inputs

```{r job-inputs}
job_inputs <- list(
  job_name = "dirichlet_alpha_pilot_v2",
  HPC = TRUE,
  mem = "6G",
  time = "02:59:00",
  email = "mgc5166@psu.edu",
  output_root = here("output"),

  grid_mode = "full",

  # Single use case: susine vanilla (no EB, no annotations)
  use_case_ids = "susine_vanilla",

  # Restart exploration: 11 fits = 1 baseline + 10 warm starts
  exploration_methods = "restart",
  exploration_mode = "separate",
  K = 11L,

  # Fixed simulation settings
  L_grid = 10L,
  architecture_grid = "susie2_oligogenic",  # SuSiE 2 paper Scenario 1: 3/5/15, h2=0.25
  pair_L_p_star = FALSE,
  # y_noise and p_star are unused — h2_total=0.25 and tier counts hardcoded

  # No annotations for this study
  annotation_r2 = 0.0,
  inflate_match = 1.0,

  # Restart settings: n_inits = 11 gives 1 baseline + 10 warm starts.
  # alpha_concentration here is IGNORED when alpha_concentration_grid is

  # set -- each run gets its per-run alpha from the grid instead.
  restart_settings = list(
    n_inits = 11L,
    alpha_concentration = 1  # fallback only; overridden by grid below
  ),

  # Alpha grid: each restart group is duplicated per alpha value,
  # so total fits = 450 datasets x 8 alphas x 11 = 39,600
  alpha_concentration_grid = c(0.001, 0.01, 0.1, 0.5, 1.0, 2.0, 5.0, 10.0),

  # Fixed prior variance
  sigma_0_2_scalars = 0.2,

  # Seeds: 3 per matrix

  seeds = 1:3,
  task_unit = "dataset",
  bundles_per_task = 5L,
  task_assignment_seed = 42L,

  # Data source: all 150 scenario_1 matrices
  data_scenarios = "scenario_1",
  repo_root = here(),
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),

  # Output settings
  credible_set_rho = 0.95,
  purity_threshold = 0.50,
  verbose_file_output = FALSE,
  write_legacy_snp_csv = FALSE,
  write_snps_parquet = FALSE,
  write_confusion_bins = TRUE,

  # All aggregation methods
  aggregation_methods = c("max_elbo", "uniform", "elbo_softmax", "cluster_weight"),
  include_overall_pool = FALSE,
  softmax_temperature = 1,

  metrics_settings = list(
    pip_breaks = c(
      seq(0.000, 0.048, by = 0.002),   # 25 bins: [0.00, 0.05) — fine for FPR resolution
      seq(0.05,  0.19,  by = 0.01),    # 15 bins: [0.05, 0.20)
      seq(0.20,  1.00,  by = 0.05)     # 17 bins: [0.20, 1.00]
    ),
    z_top_k = 10,
    jsd_threshold = 0.15
  )
)

stopifnot(job_inputs$restart_settings$n_inits == job_inputs$K)
```

# 2. Build Matrix Catalog

```{r matrix-catalog}
data_matrix_catalog <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = unique(job_inputs$data_scenarios),
  repo_root = job_inputs$repo_root,
  summary_path = job_inputs$sampled_scenario_summary
)

cat("Matrices available:", nrow(data_matrix_catalog), "\n")
```

# 3. Build Job Config

```{r build-config}
prior_quality <- prior_quality_grid(
  annotation_r2_levels = job_inputs$annotation_r2,
  inflate_match_levels = job_inputs$inflate_match
)

job_config <- make_job_config(
  job_name = job_inputs$job_name,
  HPC = job_inputs$HPC,
  time = job_inputs$time,
  mem = job_inputs$mem,
  use_case_ids = job_inputs$use_case_ids,
  exploration_methods = job_inputs$exploration_methods,
  exploration_mode = job_inputs$exploration_mode,
  K = job_inputs$K,
  L_grid = job_inputs$L_grid,
  y_noise_grid = job_inputs$y_noise_grid,
  prior_quality = prior_quality,
  p_star_grid = job_inputs$p_star_grid,
  seeds = job_inputs$seeds,
  architecture_grid = job_inputs$architecture_grid,
  data_scenarios = job_inputs$data_scenarios,
  repo_root = job_inputs$repo_root,
  sampled_scenario_summary = job_inputs$sampled_scenario_summary,
  data_matrix_catalog = data_matrix_catalog,
  task_unit = job_inputs$task_unit,
  bundles_per_task = job_inputs$bundles_per_task,
  email = job_inputs$email,
  output_root = job_inputs$output_root,
  credible_set_rho = job_inputs$credible_set_rho,
  purity_threshold = job_inputs$purity_threshold,
  verbose_file_output = job_inputs$verbose_file_output,
  write_legacy_snp_csv = job_inputs$write_legacy_snp_csv,
  write_snps_parquet = job_inputs$write_snps_parquet,
  write_confusion_bins = job_inputs$write_confusion_bins,
  task_assignment_seed = job_inputs$task_assignment_seed,
  grid_mode = job_inputs$grid_mode,
  pair_L_p_star = job_inputs$pair_L_p_star,
  sigma_0_2_scalars = job_inputs$sigma_0_2_scalars,
  restart_settings = job_inputs$restart_settings,
  alpha_concentration_grid = job_inputs$alpha_concentration_grid,
  aggregation_methods = job_inputs$aggregation_methods,
  include_overall_pool = job_inputs$include_overall_pool,
  softmax_temperature = job_inputs$softmax_temperature,
  metrics_settings = job_inputs$metrics_settings
)
```

# 4. Inspect Run Design

```{r run-accounting}
runs <- job_config$tables$runs

cat("Total runs:", nrow(runs), "\n")
cat("Unique datasets:", n_distinct(runs$dataset_bundle_id), "\n")
cat("Alpha values in run table:",
    sort(unique(runs$alpha_concentration[!is.na(runs$alpha_concentration)])), "\n")

# Runs per alpha x dataset
runs %>%
  filter(!is.na(alpha_concentration)) %>%
  count(alpha_concentration, name = "runs_per_alpha") %>%
  mutate(datasets = runs_per_alpha / job_inputs$K) %>%
  knitr::kable(caption = "Runs per alpha value")
```

```{r preview-tables, eval=FALSE}
View(as.data.frame(job_config$tables$dataset_bundles))
View(as.data.frame(job_config$tables$runs))
```

# 5. Write Job Artifacts

```{r write-artifacts}
commit_artifacts <- TRUE
if (commit_artifacts) {
  artifacts <- write_job_artifacts(
    job_config,
    run_task_script = here("inst", "scripts", "run_task.R")
  )
  str(artifacts)
}
```

# 6. Submit

```{r submit-hpc, eval=FALSE}
# On cluster:
# sbatch output/slurm_scripts/dirichlet_alpha_pilot.slurm
```

```{r local-smoke-single-task, eval=FALSE}
# Optional local smoke test (1 dataset bundle)
run_task(
  job_name = job_inputs$job_name,
  task_id = 1L,
  job_root = here("output"),
  config_path = here("output", "temp", job_inputs$job_name, "job_config.json"),
  quiet = TRUE
)
```
