---
title: "Baseline Sims Screening - Smoke Test"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Minimal local smoke test for the baseline-screening pipeline. This keeps
the same baseline model families as the full baseline workbook, but
shrinks to 1 dataset, 1 phenotype seed, 1 annotation setting, and small
tuning grids so the whole job fits in a single local task.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_baseline_sims_smoke.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(knitr)
library(tidyr)
```

# 1. Smoke Grid Values

```{r grid-values}
job_name <- "baseline_sims_screen_smoke"

sigma_0_2_default <- 0.2
sigma_grid_vanilla <- c(0.1, 0.2)
sigma_grid_functional_mu <- c(0.1, 0.2)
c_grid_functional_mu <- c(0, 1.5)
tau_grid_functional_pi <- exp(seq(log(0.01), log(10), length.out = 4))

annotation_r2_levels <- 0.3
inflate_match_levels <- 1

L_grid <- 10L
seeds <- 1L
n_matrices_target <- 1L

max_iter <- 50L
task_unit <- "dataset"
bundles_per_task <- 1L
task_assignment_seed <- 1L

cat("sigma_grid_vanilla:", sigma_grid_vanilla, "\n")
cat("sigma_grid_functional_mu:", sigma_grid_functional_mu, "\n")
cat("c_grid_functional_mu:", c_grid_functional_mu, "\n")
cat("tau_grid_functional_pi:", signif(tau_grid_functional_pi, 3), "\n")
```

# 2. Smoke Spec Catalog

```{r smoke-specs}
spec_susie_vanilla_sigma <- list(
  name = "susie_vanilla_sigma_smoke",
  use_case_ids = "susie_vanilla",
  exploration_methods = "sigma_0_2_grid",
  exploration_mode = "separate",
  K = length(sigma_grid_vanilla),
  sigma_0_2_grid_values = sigma_grid_vanilla
)

spec_susine_vanilla_sigma <- list(
  name = "susine_vanilla_sigma_smoke",
  use_case_ids = "susine_vanilla",
  exploration_methods = "sigma_0_2_grid",
  exploration_mode = "separate",
  K = length(sigma_grid_vanilla),
  sigma_0_2_grid_values = sigma_grid_vanilla
)

spec_susie_eb_single <- list(
  name = "susie_eb_single_smoke",
  use_case_ids = "susie_eb",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

spec_susie_inf_single <- list(
  name = "susie_inf_single_smoke",
  use_case_ids = "susie_inf",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

spec_susie_ash_fixed_single <- list(
  name = "susie_ash_fixed_single_smoke",
  use_case_ids = "susie_ash_fixed",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

spec_susine_functional_mu_grid <- list(
  name = "susine_functional_mu_grid_smoke",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = length(c_grid_functional_mu) * length(sigma_grid_functional_mu),
  c_grid_values = c_grid_functional_mu,
  sigma_0_2_grid_values = sigma_grid_functional_mu
)

spec_susine_eb_clamped_single <- list(
  name = "susine_eb_clamped_scale_var_nonneg_single_smoke",
  use_case_ids = "susine_eb_clamped_scale_var_nonneg",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

spec_susie_functional_pi_tau <- list(
  name = "susie_functional_pi_tau_smoke",
  use_case_ids = "susie_functional_pi",
  exploration_methods = "tau_grid",
  exploration_mode = "separate",
  K = length(tau_grid_functional_pi),
  tau_grid_values = tau_grid_functional_pi
)

all_specs <- list(
  spec_susie_vanilla_sigma,
  spec_susine_vanilla_sigma,
  spec_susie_eb_single,
  spec_susie_inf_single,
  spec_susie_ash_fixed_single,
  spec_susine_functional_mu_grid,
  spec_susine_eb_clamped_single,
  spec_susie_functional_pi_tau
)

cat("Smoke specs:", length(all_specs), "\n")
for (s in all_specs) {
  cat(sprintf("  %-52s K=%-2d methods=%s\n",
              s$name, s$K, paste(s$exploration_methods, collapse = "x")))
}
```

# 3. Matrix Selection

```{r matrix-selection}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)
cat("Total available genotype matrices:", nrow(data_matrix_catalog_full), "\n")

data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::slice_head(n = n_matrices_target)

cat("Smoke-test matrices selected:", nrow(data_matrix_catalog), "\n")
cat("Phenotype seeds:", paste(seeds, collapse = ", "), "\n")
```

# 4. Build Smoke Config

```{r build-config}
cfg <- make_job_config(
  job_name = job_name,
  HPC = FALSE,
  time = "00:30:00",
  mem = "4G",
  email = "mgc5166@psu.edu",
  output_root = here("output"),
  grid_mode = "full",
  use_case_ids = c(
    "susie_vanilla",
    "susine_vanilla",
    "susie_eb",
    "susie_inf",
    "susie_ash_fixed",
    "susine_functional_mu",
    "susine_eb_clamped_scale_var_nonneg",
    "susie_functional_pi"
  ),
  exploration_specs = all_specs,
  L_grid = L_grid,
  y_noise_grid = NA_real_,
  p_star_grid = NA_integer_,
  architecture_grid = "susie2_oligogenic",
  seeds = seeds,
  prior_quality = tidyr::expand_grid(
    annotation_r2 = annotation_r2_levels,
    inflate_match = inflate_match_levels
  ),
  sigma_0_2_scalars = sigma_0_2_default,
  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),
  task_unit = task_unit,
  bundles_per_task = bundles_per_task,
  task_assignment_seed = task_assignment_seed,
  max_iter = max_iter,
  aggregation_methods = character(0),
  include_overall_pool = FALSE,
  credible_set_rho = 0.95,
  purity_threshold = 0.50,
  verbose_file_output = FALSE,
  write_snps_parquet = FALSE,
  write_confusion_bins = TRUE,
  write_tier_cs_metrics = TRUE,
  write_scaling_confusion_bins = FALSE,
  metrics_settings = list(
    pip_breaks = sort(unique(c(
      seq(0.00, 0.05, by = 0.005),
      seq(0.05, 0.10, by = 0.01),
      seq(0.10, 1.00, by = 0.02)
    ))),
    z_top_k = 10L,
    jsd_threshold = 0.15
  )
)
```

# 5. Budget Summary

```{r budget-summary}
runs <- cfg$tables$runs

cat("Total runs:", nrow(runs), "\n")
cat("Tasks:", cfg$job$n_tasks, "\n")
cat("Dataset bundles:", dplyr::n_distinct(runs$dataset_bundle_id), "\n")

runs %>%
  dplyr::group_by(.data$spec_name, .data$use_case_id) %>%
  dplyr::summarise(
    n_runs = dplyr::n(),
    n_datasets = dplyr::n_distinct(.data$dataset_bundle_id),
    .groups = "drop"
  ) %>%
  knitr::kable(caption = "Runs per smoke-test baseline spec")
```

# 6. Smoke Checks

```{r smoke-checks}
stopifnot(cfg$job$n_tasks == 1L)
stopifnot(nrow(cfg$tables$dataset_bundles) == 1L)
stopifnot(length(unique(runs$phenotype_seed)) == 1L)
stopifnot(length(unique(runs$dataset_bundle_id)) == 1L)

expected_specs <- c(
  "susie_vanilla_sigma_smoke",
  "susine_vanilla_sigma_smoke",
  "susie_eb_single_smoke",
  "susie_inf_single_smoke",
  "susie_ash_fixed_single_smoke",
  "susine_functional_mu_grid_smoke",
  "susine_eb_clamped_scale_var_nonneg_single_smoke",
  "susie_functional_pi_tau_smoke"
)
stopifnot(all(expected_specs %in% unique(runs$spec_name)))

annotated_specs <- c(
  "susine_functional_mu_grid_smoke",
  "susine_eb_clamped_scale_var_nonneg_single_smoke",
  "susie_functional_pi_tau_smoke"
)
annotated_runs <- runs %>%
  dplyr::filter(.data$spec_name %in% annotated_specs)

stopifnot(all(unique(annotated_runs$annotation_r2) == 0.3))
stopifnot(all(unique(annotated_runs$inflate_match) == 1))

cat("Smoke checks passed.\n")
```

# 7. Write Artifacts

```{r write-artifacts, eval=FALSE}
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)
cat("Job config written to:", arts$job_config, "\n")
cat("SLURM script written to:", arts$slurm_script, "\n")
```

# 8. Run Locally

```{r run-locally, eval=FALSE}
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)

run_task(
  job_name = cfg$job$name,
  task_id = 1L,
  job_root = here("output"),
  config_path = arts$job_config,
  quiet = FALSE
)
```

# 9. Quick Validation

```{r validate, eval=FALSE}
staging_dir <- file.path(cfg$paths$temp_dir, "staging")
if (dir.exists(staging_dir)) {
  staging_files <- list.files(staging_dir, recursive = TRUE)
  cat("Staging files:", length(staging_files), "\n")
  print(table(tools::file_ext(staging_files)))
} else {
  cat("No staging directory found -- check the local run above.\n")
}

model_metrics_files <- list.files(
  staging_dir,
  pattern = "model_metrics",
  recursive = TRUE
)
cat("Model metrics files:", length(model_metrics_files), "\n")

effect_metrics_files <- list.files(
  staging_dir,
  pattern = "effect_metrics",
  recursive = TRUE
)
cat("Effect metrics files:", length(effect_metrics_files), "\n")

confusion_files <- list.files(
  staging_dir,
  pattern = "confusion",
  recursive = TRUE
)
cat("Confusion bin files:", length(confusion_files), "\n")
```
