---
title: "SuSiNE EB Method Ablation — Comprehensive Study"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Comprehensive ablation study to determine the best EB
prior-update strategy for (1) prior variance only, (2) prior mean
(scalar or annotation-scaled c) only, and (3) joint mean + variance.
Compares 23 use cases across 3 genetic architectures (sparse,
susie2_sparse, susie2_oligogenic), multiple annotation qualities,
and sigma_0_2 grids for scale-only EB methods.

Key question: if you had to pick one use case for each of the three
EB slots (variance, scale, both), which performs best?

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_eb_ablation.Rmd")
library(here)
here()

library(devtools)
devtools::load_all(here("..", "susine"))
devtools::load_all(here("."))

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# 1. Ablation Inputs

```{r job-inputs}
job_inputs <- list(
  job_name = "eb_ablation_v2",
  HPC = TRUE,
  mem = "6G",
  time = "3:59:59",
  email = "mgc5166@psu.edu",
  output_root = here("output"),

  # Full factorial grid.
  grid_mode = "full",

  # --- 13 use cases (nonneg-only for annotation EB methods) ---
  use_case_ids = c(
    # Baseline, no annotation (2)
    "susie_vanilla",
    "susine_vanilla",

    # EB on mean and/or variance, no annotations (4)
    "susie_eb",                # susieR EB variance
    "susine_eb",               # susine EB variance
    "susine_eb_mean",          # susine EB scalar mean only (sigma fixed)
    "susine_eb_mu_var",        # susine EB scalar mean + variance jointly

    # Both EB (c + sigma), with annotations — nonneg only (5)
    "susine_eb_scale_var_nonneg",
    "susine_eb_penalized_scale_var_nonneg",
    "susine_eb_scale_var_outer_nonneg",
    "susine_eb_clamped_scale_var_nonneg",
    "susine_eb_alternating_scale_var_nonneg",

    # Scale only EB — nonneg only (1)
    "susine_eb_scale_nonneg",

    # Annotations, no EB — fixed c + fixed sigma (1)
    # Uses c_grid x sigma_0_2_scalars for exploration
    "susine_functional_mu"
  ),

  # Mixed exploration: "single" applies to all use cases;
  # "c_grid" only applies to susine_functional_mu (functional_mu strategy).
  # exploration_mode = "separate" generates separate groups for each method.
  exploration_methods = c("single", "c_grid"),
  exploration_mode = "separate",
  K = 5L,

  # Core simulation settings.
  L_grid = 10L,

  # y_noise only used by "sparse" architecture; susie2_* use h2-calibrated noise.
  y_noise_grid = c(0.5, 0.8),

  # p_star: used by sparse + susie2_sparse (not susie2_oligogenic, which has
  # fixed tier counts of 3/5/15 = 23 total causal).
  p_star_grid = c(1L, 3L, 5L, 10L),

  # Two SuSiE 2.0 architectures only (legacy sparse dropped).
  architecture_grid = c("susie2_sparse", "susie2_oligogenic"),
  pair_L_p_star = FALSE,

  # Annotation quality: wide range to stress-test EB methods.
  annotation_r2 = c(0.1, 0.3, 0.5),
  inflate_match = c(0.1, 0.5, 1.0, 1.5),

  # Prior variance scalars: used by fixed-sigma methods AND "scale" EB
  # (which fixes sigma and optimizes c only). Three values.
  sigma_0_2_scalars = c(0.1, 0.2, 0.4),

  # c_grid for susine_functional_mu (K = 5 values).
  c_grid_values = c(0, 0.33, 0.67, 1.0, 1.33),

  # Seeds and task packing.
  seeds = 1, #1:3
  task_unit = "dataset",
  bundles_per_task = 1L,
  task_assignment_seed = 1L,

  # --- Matrix selection ---
  # n3 finemapping matrix (built-in) + M1-stratified simulated genotypes.
  data_scenarios = c("simulation_n3", "scenario_1"),
  repo_root = here(),
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),
  pilot_selection_mode = "M1_stratified",
  pilot_n_matrices = 1L,#10L
  pilot_matrix_ids_path = here("data", "pilot_matrix_ids.csv"),
  pilot_m1_cache_path = here("output", "pilot_m1_metrics.csv"),
  pilot_catalog_with_m1_path = here(
    "data", "sampled_simulated_genotypes",
    "scenario_sampling_summary_with_m1.csv"
  ),

  # --- Output flags ---
  credible_set_rho = 0.95,
  purity_threshold = 0.50,
  verbose_file_output = FALSE,
  write_legacy_snp_csv = FALSE,
  write_snps_parquet = FALSE,
  write_confusion_bins = TRUE,
  write_tier_cs_metrics = TRUE,
  write_prior_diagnostics = TRUE,

  # Aggregation: for single exploration, aggregation is trivial (1 fit).
  # For c_grid exploration (susine_functional_mu), use ELBO-softmax BMA.
  aggregation_methods = c("max_elbo", "elbo_softmax", "uniform", "cluster_weight"),
  include_overall_pool = FALSE,
  overall_aggregation_methods = c("max_elbo"),
  softmax_temperature = 1,

  metrics_settings = list(
    pip_bucket_width = 0.01,
    pip_breaks = pip_breaks_variable(),
    z_top_k = 10,
    jsd_threshold = 0.15
  ),

  # tau_grid not used (no functional_pi use cases).
  tau_grid_values = 1,
  restart_settings = list(
    n_inits = 1L,
    alpha_concentration = 0.001,
    debug_alpha = FALSE
  ),
  refine_settings = list(
    n_steps = 1L,
    cs_source = "filtered"
  )
)

# Validation: c_grid size must match K.
stopifnot(length(job_inputs$c_grid_values) == job_inputs$K)
```

# 2. (Optional) Restrict to Pilot Matrices

```{r pilot-matrix-selection}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = unique(job_inputs$data_scenarios),
  repo_root = job_inputs$repo_root,
  summary_path = job_inputs$sampled_scenario_summary
)

pilot_mode <- job_inputs$pilot_selection_mode

if (pilot_mode == "csv") {
  pilot_matrix_ids <- readr::read_csv(
    job_inputs$pilot_matrix_ids_path, show_col_types = FALSE
  )
  data_matrix_catalog <- data_matrix_catalog_full %>%
    dplyr::semi_join(pilot_matrix_ids, by = "matrix_id")
} else if (pilot_mode == "M1_stratified") {
  if (file.exists(job_inputs$pilot_catalog_with_m1_path)) {
    catalog_with_m1 <- readr::read_csv(
      job_inputs$pilot_catalog_with_m1_path, show_col_types = FALSE
    )
    if (!"matrix_id" %in% names(catalog_with_m1)) {
      if ("gene" %in% names(catalog_with_m1) &&
          !"dataset_label" %in% names(catalog_with_m1)) {
        catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = gene)
      }
      if ("dataset_label" %in% names(catalog_with_m1)) {
        catalog_with_m1 <- catalog_with_m1 %>%
          dplyr::left_join(
            data_matrix_catalog_full %>%
              dplyr::select(matrix_id, dataset_label),
            by = "dataset_label"
          )
      }
    }
  } else {
    m1_tbl <- test_susine::build_x_metrics_table(
      matrix_paths_df = data_matrix_catalog_full %>%
        dplyr::select(matrix_id, matrix_path),
      base_dir = job_inputs$repo_root
    ) %>%
      dplyr::select(matrix_id, M1)

    catalog_with_m1 <- data_matrix_catalog_full %>%
      dplyr::left_join(m1_tbl, by = "matrix_id")

    dir.create(
      dirname(job_inputs$pilot_catalog_with_m1_path),
      recursive = TRUE, showWarnings = FALSE
    )
    readr::write_csv(catalog_with_m1, job_inputs$pilot_catalog_with_m1_path)

    if (!is.null(job_inputs$pilot_m1_cache_path)) {
      readr::write_csv(m1_tbl, job_inputs$pilot_m1_cache_path)
    }
  }

  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::filter(!is.na(M1)) %>%
    dplyr::arrange(M1)

  n_target <- min(job_inputs$pilot_n_matrices, nrow(catalog_with_m1))
  if (n_target < 1) stop("No matrices available after M1 filtering.")

  idx <- unique(round(seq(1, nrow(catalog_with_m1), length.out = n_target)))
  selected <- catalog_with_m1[idx, , drop = FALSE]

  data_matrix_catalog <- data_matrix_catalog_full %>%
    dplyr::semi_join(selected %>% dplyr::select(matrix_id), by = "matrix_id")
} else {
  data_matrix_catalog <- data_matrix_catalog_full
}

cat("Pilot matrix selection mode:", pilot_mode, "\n")
cat("Matrices selected:", nrow(data_matrix_catalog), "\n")
```

# 3. Build Job Config

```{r build-config}
prior_quality <- prior_quality_grid(
  annotation_r2_levels = job_inputs$annotation_r2,
  inflate_match_levels = job_inputs$inflate_match
)

job_config <- make_job_config(
  job_name = job_inputs$job_name,
  HPC = job_inputs$HPC,
  time = job_inputs$time,
  mem = job_inputs$mem,
  use_case_ids = job_inputs$use_case_ids,
  exploration_methods = job_inputs$exploration_methods,
  exploration_mode = job_inputs$exploration_mode,
  K = job_inputs$K,
  L_grid = job_inputs$L_grid,
  y_noise_grid = job_inputs$y_noise_grid,
  prior_quality = prior_quality,
  p_star_grid = job_inputs$p_star_grid,
  seeds = job_inputs$seeds,
  architecture_grid = job_inputs$architecture_grid,
  data_scenarios = job_inputs$data_scenarios,
  repo_root = job_inputs$repo_root,
  sampled_scenario_summary = job_inputs$sampled_scenario_summary,
  data_matrix_catalog = data_matrix_catalog,
  task_unit = job_inputs$task_unit,
  bundles_per_task = job_inputs$bundles_per_task,
  email = job_inputs$email,
  output_root = job_inputs$output_root,
  credible_set_rho = job_inputs$credible_set_rho,
  purity_threshold = job_inputs$purity_threshold,
  verbose_file_output = job_inputs$verbose_file_output,
  write_legacy_snp_csv = job_inputs$write_legacy_snp_csv,
  write_snps_parquet = job_inputs$write_snps_parquet,
  write_confusion_bins = job_inputs$write_confusion_bins,
  write_tier_cs_metrics = job_inputs$write_tier_cs_metrics,
  write_prior_diagnostics = job_inputs$write_prior_diagnostics,
  task_assignment_seed = job_inputs$task_assignment_seed,
  grid_mode = job_inputs$grid_mode,
  pair_L_p_star = job_inputs$pair_L_p_star,
  sigma_0_2_scalars = job_inputs$sigma_0_2_scalars,
  c_grid_values = job_inputs$c_grid_values,
  tau_grid_values = job_inputs$tau_grid_values,
  restart_settings = job_inputs$restart_settings,
  refine_settings = job_inputs$refine_settings,
  aggregation_methods = job_inputs$aggregation_methods,
  overall_aggregation_methods = job_inputs$overall_aggregation_methods,
  include_overall_pool = job_inputs$include_overall_pool,
  softmax_temperature = job_inputs$softmax_temperature,
  metrics_settings = job_inputs$metrics_settings
)
```

# 4. Inspect Run Design

```{r run-accounting}
lines <- test_susine::run_accounting_lines(
  job_config,
  seeds_per_matrix = length(unique(job_inputs$seeds)),
  use_case_ids = job_inputs$use_case_ids
)
cat(paste(lines, collapse = "\n"), "\n")
```

```{r preview-bundles}
# Verify architecture-specific grid collapsing:
# - susie2_* architectures should have y_noise = NA
# - susie2_oligogenic should have p_star = NA
job_config$tables$dataset_bundles %>%
  dplyr::count(architecture, y_noise, p_star) %>%
  knitr::kable(caption = "Dataset bundle grid (architecture x y_noise x p_star)")
```

```{r preview-tables}
cat("Dataset bundles:", nrow(job_config$tables$dataset_bundles), "\n")
cat("Total runs:", nrow(job_config$tables$runs), "\n")
cat("SLURM tasks:", nrow(job_config$tables$tasks), "\n")
```

```{r use-case-summary}
# Verify which use cases get sigma_0_2 expansion and c_grid exploration.
job_config$tables$runs %>%
  dplyr::group_by(use_case_id, exploration_methods) %>%
  dplyr::summarise(
    n_runs = dplyr::n(),
    sigma_values = paste(sort(unique(sigma_0_2_scalar)), collapse = ", "),
    c_values = paste(sort(unique(c_value)), collapse = ", "),
    .groups = "drop"
  ) %>%
  knitr::kable(caption = "Runs per use case x exploration method")
```

# 5. Write Job Artifacts

```{r write-artifacts}
commit_artifacts <- TRUE
if (commit_artifacts) {
  artifacts <- write_job_artifacts(
    job_config,
    run_task_script = here("inst", "scripts", "run_task.R")
  )
  str(artifacts)
}
```

# 6. Submit / Local Smoke

```{r submit-hpc, eval=FALSE}
# On cluster:
# sbatch output/slurm_scripts/eb_ablation_v2.slurm
```

```{r local-smoke-single-task, eval=FALSE}
# Optional local smoke run for task 1.
run_task(
  job_name = job_inputs$job_name,
  task_id = 1L,
  job_root = here("output"),
  config_path = here("output", "temp", job_inputs$job_name, "job_config.json"),
  quiet = FALSE
)
```
