---
title: "Architecture Sensitivity -- Diffuse Causal Architecture"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

Builds the reviewer-requested diffuse genetic-architecture sensitivity job. The
job uses the standard genotype substrate, central annotation setting
(`annotation_r2 = 0.3`, `inflate_match = 0.95`), `k = 100` causal variants,
total locus h2 `{0.05, 0.15, 0.25}`, a 4 x 4 C-CS grid, and a single
annotation-free baseline.

The existing output path marks the top eight causal variants by absolute effect
as positives when more than eight causal variants are present, so this arm uses
the requested top-8 diffuse target without adding a separate truth definition.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_architecture_diffuse.Rmd")
library(here)
library(devtools)
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
} else {
  warning("susine package not found at: ", susine_path)
}
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```

# Configuration

```{r user-config}
job_name <- "architecture_diffuse_sensitivity"
email    <- "mgc5166@psu.edu"

slurm_time <- "11:59:59"
slurm_mem  <- "8G"

n_matrices_target <- 150L
seeds             <- 1:4

L_grid            <- 10L
y_noise_grid      <- NA_real_
p_star_grid       <- NA_integer_
h2_snp_per_causal_grid <- NULL
h2_total_grid     <- c(0.05, 0.15, 0.25)
diffuse_k_grid    <- 100L
architecture_grid <- "susie2_diffuse"

annotation_r2 <- 0.3
inflate_match <- 0.95

c_grid_4 <- c(0.00, 0.43, 1.07, 1.50)
sigma_grid_4 <- c(0.01, 0.10, 0.40, 1.00)

max_iter <- 100L
bundles_per_task <- 12L
task_assignment_seed <- 42L
buffer_flush_interval <- 4L
aggregation_methods <- "cluster_weight_credible"
include_overall_pool <- FALSE
softmax_temperature <- 1

credible_set_rho <- 0.95
purity_threshold <- 0.50
verbose_file_output <- FALSE
write_snps_parquet <- FALSE
write_confusion_bins <- TRUE
write_tier_cs_metrics <- FALSE
write_scaling_confusion_bins <- FALSE

pip_breaks <- sort(unique(c(
  seq(0.00, 0.05, by = 0.005),
  seq(0.05, 0.10, by = 0.01),
  seq(0.10, 1.00, by = 0.02)
)))
```

# Specs

```{r specs}
spec_C_CS_4x4 <- list(
  name = "C-CS-architecture-diffuse",
  use_case_ids = "susine_functional_mu",
  exploration_methods = c("c_grid", "sigma_0_2_grid"),
  exploration_mode = "intersect",
  K = 16L,
  c_grid_values = c_grid_4,
  sigma_0_2_grid_values = sigma_grid_4
)

spec_baseline_single <- list(
  name = "baseline-single",
  use_case_ids = "susine_vanilla",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

all_specs <- list(spec_C_CS_4x4, spec_baseline_single)
```

# Matrix Selection

```{r matrix-selection}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here("data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv")
)

m1_summary_path <- here("data", "sampled_simulated_genotypes", "scenario_sampling_summary_with_m1.csv")
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>% dplyr::select(matrix_id, matrix_path),
    base_dir = here()
  ) %>% dplyr::select(matrix_id, M1)
  catalog_with_m1 <- data_matrix_catalog_full %>% dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}
if (!"matrix_id" %in% names(catalog_with_m1)) {
  if ("gene" %in% names(catalog_with_m1)) {
    catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = gene)
  }
  join_col <- intersect(c("dataset_label", "matrix_path"), names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(
      data_matrix_catalog_full %>% dplyr::select(matrix_id, dplyr::all_of(join_col)),
      by = join_col
    )
}
catalog_with_m1 <- catalog_with_m1 %>% dplyr::filter(!is.na(M1)) %>% dplyr::arrange(M1)
idx <- unique(round(seq(1, nrow(catalog_with_m1), length.out = min(n_matrices_target, nrow(catalog_with_m1)))))
selected <- catalog_with_m1[idx, , drop = FALSE]
data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(dplyr::select(selected, matrix_id), by = "matrix_id")
```

# Build Job

```{r build-config}
prior_quality <- tidyr::expand_grid(
  annotation_r2 = annotation_r2,
  inflate_match = inflate_match
)

cfg <- make_job_config(
  job_name = job_name,
  HPC = TRUE,
  time = slurm_time,
  mem = slurm_mem,
  email = email,
  output_root = here("output"),
  grid_mode = "full",
  use_case_ids = c("susine_vanilla", "susine_functional_mu"),
  exploration_specs = all_specs,
  L_grid = L_grid,
  y_noise_grid = y_noise_grid,
  p_star_grid = p_star_grid,
  h2_snp_per_causal_grid = h2_snp_per_causal_grid,
  h2_total_grid = h2_total_grid,
  diffuse_k_grid = diffuse_k_grid,
  architecture_grid = architecture_grid,
  seeds = seeds,
  prior_quality = prior_quality,
  sigma_0_2_scalars = 0.2,
  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here("data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"),
  task_unit = "dataset",
  bundles_per_task = bundles_per_task,
  task_assignment_seed = task_assignment_seed,
  buffer_flush_interval = buffer_flush_interval,
  max_iter = max_iter,
  aggregation_methods = aggregation_methods,
  include_overall_pool = include_overall_pool,
  softmax_temperature = softmax_temperature,
  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = verbose_file_output,
  write_snps_parquet = write_snps_parquet,
  write_confusion_bins = write_confusion_bins,
  write_prior_diagnostics = TRUE,
  write_tier_cs_metrics = write_tier_cs_metrics,
  write_scaling_confusion_bins = write_scaling_confusion_bins,
  metrics_settings = list(
    pip_breaks = pip_breaks,
    z_top_k = 10,
    jsd_threshold = 0.15
  )
)
```

# Budget

```{r budget}
runs <- cfg$tables$runs
cat("Total run rows:", nrow(runs), "\n")
cat("Datasets:", dplyr::n_distinct(runs$dataset_bundle_id), "\n")
runs %>%
  dplyr::group_by(spec_name, h2_total, diffuse_k) %>%
  dplyr::summarise(n_runs = dplyr::n(), n_datasets = dplyr::n_distinct(dataset_bundle_id), .groups = "drop") %>%
  knitr::kable()
```

# Write Artifacts

```{r write-artifacts}
artifacts <- write_job_artifacts(
  cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)
artifacts
cat("Submit with:\n")
cat("sbatch", normalizePath(artifacts$slurm_script, winslash = "/", mustWork = FALSE), "\n")
```

# Reporting Gates

Before reporting this run, validate the collected outputs:

- `prior_diagnostics.csv`: fixed-c C-CS fits should have `c_l == 1`, confirming that `estimate_mu_0_scale_factor` did not run.
- `confusion_bins.csv`: C-CS aggregate rows and baseline rows should both be present for every total-h2 setting.
- Positive labels are the top eight causal variants by absolute effect for this arm.