---
title: "Purity-Filter Minibatch - Run Control"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Slim minibatch that quantifies how credible-set filtering changes per-variant
AUPRC on the **same 600 oligogenic datasets** as the baseline / ensemble studies.
A single annotation-agnostic SuSiE-vanilla fit ($c = 0$, $\sigma_0^2 = 0.2$) is run
per dataset; the run-time evaluator then recomputes the combined PIP after
dropping credible sets under several filters and writes confusion bins tagged by
`filter_type`:

- `no_filter` (the standard unfiltered PIP),
- `purity_95`, `purity_90`, `purity_50` (drop CSs with min-pairwise-$|r|$ below
  the threshold),
- `keff_core95_le_100` (drop diffuse effects whose core-95% effective support
  size exceeds 100).

Outputs are deliberately slim: per-filter confusion bins (for pooled AUPRC) plus
the per-effect diagnostic record (purity, coverage, size, core-95% $k_{\mathrm{eff}}$,
accuracy ratio) that the standard pipeline already emits. The post-run collect
workbook turns the confusion bins into one AUPRC per filter type.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/run_control_workbook_purity_filter_minibatch.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(knitr)
library(tidyr)
```

# 0. User Configuration

```{r user-config}
job_name <- "purity_filter_minibatch"
email <- "mgc5166@psu.edu"

slurm_time <- "02:59:59"
slurm_mem <- "8G"

# Match the baseline / ensemble substrate exactly so the 600 datasets are
# identical (same matrices, seeds, architecture, task assignment).
n_matrices_target <- 150L
seeds <- 1:4

L_grid <- 10L
y_noise_grid <- NA_real_
p_star_grid <- NA_integer_
architecture_grid <- "susie2_oligogenic"

# SuSiE-vanilla single fit, c = 0, sigma_0_2 = 0.2 (baseline-SuSiE equivalent).
sigma_0_2_default <- 0.2

max_iter <- 100L

task_unit <- "dataset"
bundles_per_task <- 4L
task_assignment_seed <- 42L

credible_set_rho <- 0.95
# This per-run purity_threshold only gates the *filtered* CS view that feeds
# power/size/coverage; the AUPRC sweep below uses its own thresholds.
purity_threshold <- 0.50
verbose_file_output <- FALSE
write_snps_parquet <- FALSE
write_confusion_bins <- FALSE        # the sweep replaces the single set
write_tier_cs_metrics <- FALSE

# Purity-filter sweep specification (drives confusion bins by filter_type).
purity_filter_sweep <- TRUE
purity_filter_specs <- list(
  purity_thresholds = c(0.95, 0.90, 0.50),
  keff_core95_max = 100
)

pip_breaks <- sort(unique(c(
  seq(0.00, 0.05, by = 0.005),
  seq(0.05, 0.10, by = 0.01),
  seq(0.10, 1.00, by = 0.02)
)))
z_top_k <- 10L
jsd_threshold <- 0.15
```

# 1. Spec Catalog (single SuSiE-vanilla fit)

```{r specs}
spec_susie_vanilla_single <- list(
  name = "susie_vanilla_single",
  use_case_ids = "susie_vanilla",
  exploration_methods = "single",
  exploration_mode = "separate",
  K = 1L
)

all_specs <- list(spec_susie_vanilla_single)

cat("Total specs:", length(all_specs), "\n")
for (s in all_specs) {
  cat(sprintf("  %-30s methods=%s\n", s$name,
              paste(s$exploration_methods, collapse = "x")))
}
```

# 2. Matrix Selection (identical to baseline / ensemble)

```{r matrix-selection}
data_matrix_catalog_full <- test_susine:::build_data_matrix_catalog(
  requested_scenarios = "scenario_1",
  repo_root = here(),
  summary_path = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  )
)
cat("Total available genotype matrices:", nrow(data_matrix_catalog_full), "\n")

m1_summary_path <- here(
  "data", "sampled_simulated_genotypes", "scenario_sampling_summary_with_m1.csv"
)
if (file.exists(m1_summary_path)) {
  catalog_with_m1 <- readr::read_csv(m1_summary_path, show_col_types = FALSE)
} else {
  m1_tbl <- test_susine::build_x_metrics_table(
    matrix_paths_df = data_matrix_catalog_full %>%
      dplyr::select(matrix_id, matrix_path),
    base_dir = here()
  ) %>%
    dplyr::select(matrix_id, M1)
  catalog_with_m1 <- data_matrix_catalog_full %>%
    dplyr::left_join(m1_tbl, by = "matrix_id")
  readr::write_csv(catalog_with_m1, m1_summary_path)
}

if (!"matrix_id" %in% names(catalog_with_m1)) {
  if ("gene" %in% names(catalog_with_m1)) {
    catalog_with_m1 <- dplyr::rename(catalog_with_m1, dataset_label = gene)
  }
  join_col <- intersect(c("dataset_label", "matrix_path"), names(catalog_with_m1))[[1L]]
  catalog_with_m1 <- catalog_with_m1 %>%
    dplyr::left_join(
      data_matrix_catalog_full %>%
        dplyr::select(matrix_id, dplyr::all_of(join_col)),
      by = join_col
    )
}

catalog_with_m1 <- catalog_with_m1 %>%
  dplyr::filter(!is.na(.data$M1)) %>%
  dplyr::arrange(.data$M1)

n_target <- min(n_matrices_target, nrow(catalog_with_m1))
idx <- unique(round(seq(1, nrow(catalog_with_m1), length.out = n_target)))
selected <- catalog_with_m1[idx, , drop = FALSE]

data_matrix_catalog <- data_matrix_catalog_full %>%
  dplyr::semi_join(dplyr::select(selected, .data$matrix_id), by = "matrix_id")

cat("Matrices selected (M1-stratified):", nrow(data_matrix_catalog), "\n")
cat("Phenotype seeds:", paste(seeds, collapse = ", "), "\n")
```

# 3. Build Job Config

```{r build-config}
cfg <- make_job_config(
  job_name = job_name,
  HPC = TRUE,
  time = slurm_time,
  mem = slurm_mem,
  email = email,
  output_root = here("output"),
  grid_mode = "full",
  use_case_ids = "susie_vanilla",
  exploration_specs = all_specs,
  L_grid = L_grid,
  y_noise_grid = y_noise_grid,
  p_star_grid = p_star_grid,
  architecture_grid = architecture_grid,
  seeds = seeds,
  # Annotation-agnostic: a single (annotation_r2, inflate_match) row yields one
  # fit per dataset = 600 runs total.
  prior_quality = tibble::tibble(annotation_r2 = 0, inflate_match = 1),
  sigma_0_2_scalars = sigma_0_2_default,
  data_scenarios = "scenario_1",
  repo_root = here(),
  data_matrix_catalog = data_matrix_catalog,
  sampled_scenario_summary = here(
    "data", "sampled_simulated_genotypes", "scenario_sampling_summary.csv"
  ),
  task_unit = task_unit,
  bundles_per_task = bundles_per_task,
  task_assignment_seed = task_assignment_seed,
  max_iter = max_iter,
  aggregation_methods = character(0),
  include_overall_pool = FALSE,
  credible_set_rho = credible_set_rho,
  purity_threshold = purity_threshold,
  verbose_file_output = verbose_file_output,
  write_snps_parquet = write_snps_parquet,
  write_confusion_bins = write_confusion_bins,
  write_tier_cs_metrics = write_tier_cs_metrics,
  write_scaling_confusion_bins = FALSE,
  purity_filter_sweep = purity_filter_sweep,
  purity_filter_specs = purity_filter_specs,
  metrics_settings = list(
    pip_breaks = pip_breaks,
    z_top_k = z_top_k,
    jsd_threshold = jsd_threshold
  )
)
```

# 4. Budget Summary

```{r budget-summary}
runs <- cfg$tables$runs

cat("Total runs:", nrow(runs), "\n")
cat("Datasets:", dplyr::n_distinct(runs$dataset_bundle_id), "\n")
cat("Use cases:", paste(sort(unique(runs$use_case_id)), collapse = ", "), "\n")
cat("Purity-filter sweep enabled:",
    isTRUE(cfg$job$compute$purity_filter_sweep), "\n")
cat("Filter specs:\n")
str(cfg$job$compute$purity_filter_specs)
```

# 5. Smoke Checks

```{r smoke-checks}
stopifnot(nrow(runs) == 600L)
stopifnot(identical(sort(unique(runs$use_case_id)), "susie_vanilla"))
stopifnot(isTRUE(cfg$job$compute$purity_filter_sweep))
stopifnot(dplyr::n_distinct(runs$dataset_bundle_id) == 600L)
cat("Smoke checks passed.\n")
```

# 6. Write Artifacts

```{r write-artifacts, eval=FALSE}
arts <- write_job_artifacts(
  job_config = cfg,
  run_task_script = here("inst", "scripts", "run_task.R")
)
cat("Job config written to:", arts$job_config, "\n")
cat("SLURM script:", arts$slurm_script, "\n")
```

# 7. Submit

```{r submit, eval=FALSE}
# sbatch <slurm_script_path>
```
