---
title: "Ensemble Scaling Experiment — Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Reads consolidated outputs from the collect workbook. Generates all
4 figure categories from `refs/ensemble_scaling_experiment_plan.md`.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_ensemble_scaling.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

# 0. Load Consolidated Data

```{r load}
# Point to job-level consolidated output from collect workbook
job_name      <- "ensemble_scaling_full"
parent_job_id <- "50452623"
output_root   <- here("output")

job_dir          <- file.path(output_root, "slurm_output", job_name, parent_job_id)
consolidated_dir <- file.path(job_dir, "consolidated")

auprc_agg <- readr::read_csv(
  file.path(consolidated_dir, "auprc_aggregated.csv"),
  show_col_types = FALSE
)
auprc_individual <- readr::read_csv(
  file.path(consolidated_dir, "auprc_individual.csv"),
  show_col_types = FALSE
)
model_metrics <- readr::read_csv(
  file.path(consolidated_dir, "model_metrics_full.csv"),
  show_col_types = FALSE
)
confusion_bins <- readr::read_csv(
  file.path(consolidated_dir, "confusion_bins_full.csv"),
  show_col_types = FALSE
)

tier_cs_path <- file.path(consolidated_dir, "tier_cs_metrics_full.csv")
tier_cs_metrics_full <- if (file.exists(tier_cs_path)) {
  readr::read_csv(tier_cs_path, show_col_types = FALSE)
} else {
  message("tier_cs_metrics_full.csv not found — plots 1a/1b will be skipped.")
  NULL
}

# Figure output directories
fig_root <- file.path(job_dir, "figures")
fig_dirs <- c("susie2_plots", "ensemble_performance",
              "dataset_difficulty", "ensemble_scaling")
for (d in fig_dirs) {
  dir.create(file.path(fig_root, d), recursive = TRUE, showWarnings = FALSE)
}

cat("AUPRC aggregated rows:", nrow(auprc_agg), "\n")
cat("AUPRC individual rows:", nrow(auprc_individual), "\n")
cat("Figure root:", fig_root, "\n")
```

```{r helpers}
# Helper: compute AUPRC from a pooled confusion-bins tibble.
# AP = sum over bins of [precision_at_bin * delta_recall_at_bin].
auprc_from_bins <- function(bins) {
  b       <- bins %>% dplyr::arrange(dplyr::desc(.data$pip_threshold))
  cum_tp  <- cumsum(b$n_causal_at_bucket)
  cum_fp  <- cumsum(b$n_noncausal_at_bucket)
  total_p <- sum(b$n_causal_at_bucket)
  total_n <- sum(b$n_noncausal_at_bucket)
  if (total_p == 0L || total_n == 0L) return(NA_real_)
  precision    <- cum_tp / (cum_tp + cum_fp)
  delta_recall <- b$n_causal_at_bucket / total_p
  sum(precision * delta_recall)
}

# Helper: TPR at FPR <= threshold.
tpr05_from_bins <- function(bins, fpr_threshold = 0.05) {
  b       <- bins %>% dplyr::arrange(dplyr::desc(.data$pip_threshold))
  cum_tp  <- cumsum(b$n_causal_at_bucket)
  cum_fp  <- cumsum(b$n_noncausal_at_bucket)
  total_p <- sum(b$n_causal_at_bucket)
  total_n <- sum(b$n_noncausal_at_bucket)
  if (total_p == 0L || total_n == 0L) return(NA_real_)
  tpr <- cum_tp / total_p
  fpr <- cum_fp / total_n
  idx <- which(fpr <= fpr_threshold)
  if (!length(idx)) return(NA_real_)
  max(tpr[idx])
}
```

```{r baselines}
# Load all pooled AUPRC tables
auprc_pooled_agg <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_aggregated.csv"),
  show_col_types = FALSE
)
auprc_pooled_agg_by_r2 <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_agg_by_r2.csv"),
  show_col_types = FALSE
)
auprc_pooled_individual_overall <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_individual_overall.csv"),
  show_col_types = FALSE
)
auprc_pooled_individual <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_individual.csv"),
  show_col_types = FALSE
)

# Baseline reference: single-fit susine_vanilla, pooled across all r2
baseline_auprc <- auprc_pooled_individual_overall %>%
  dplyr::filter(spec_name == "baseline-single",
                use_case_id == "susine_vanilla") %>%
  dplyr::pull(AUPRC)

cat("Baseline pooled AUPRC:", round(baseline_auprc, 4), "\n")

# Truth-warm for OVERALL plot: pooled across all r2, no r2 in label
truth_warm_overall <- auprc_pooled_individual_overall %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, AUPRC) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~ "truth-warm EB",
      use_case_id == "susine_functional_mu" ~ "truth-warm grid (c=0.5)",
      TRUE ~ use_case_id
    ),
    annotation_r2 = NA_real_
  )

# Truth-warm for PER-R2 plots: one row per use_case x annotation_r2, r2 in label
truth_warm_by_r2 <- auprc_pooled_individual %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, annotation_r2, AUPRC) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~
        paste0("truth-warm EB (r\u00b2=", annotation_r2, ")"),
      use_case_id == "susine_functional_mu" ~
        paste0("truth-warm grid (c=0.5, r\u00b2=", annotation_r2, ")"),
      TRUE ~ use_case_id
    )
  )

cat("Truth-warm overall:\n"); print(truth_warm_overall)
cat("Truth-warm by r2:\n"); print(truth_warm_by_r2)

# Load TPR@0.05 pooled tables
tpr05_pooled_agg <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_aggregated.csv"),
  show_col_types = FALSE
)
tpr05_pooled_agg_by_r2 <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_agg_by_r2.csv"),
  show_col_types = FALSE
)
tpr05_pooled_individual_overall <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_individual_overall.csv"),
  show_col_types = FALSE
)
tpr05_pooled_individual <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_individual.csv"),
  show_col_types = FALSE
)

# Baseline TPR@0.05 (single-fit susine_vanilla, pooled across all r2)
baseline_tpr05 <- tpr05_pooled_individual_overall %>%
  dplyr::filter(spec_name == "baseline-single",
                use_case_id == "susine_vanilla") %>%
  dplyr::pull(tpr_fpr05)
cat("Baseline pooled TPR@0.05:", round(baseline_tpr05, 4), "\n")

# Truth-warm TPR@0.05 for OVERALL plot
truth_warm_tpr05_overall <- tpr05_pooled_individual_overall %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, tpr_fpr05) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~ "truth-warm EB",
      use_case_id == "susine_functional_mu" ~ "truth-warm grid (c=0.5)",
      TRUE ~ use_case_id
    ),
    annotation_r2 = NA_real_
  )

# Truth-warm TPR@0.05 for PER-R2 plots
truth_warm_tpr05_by_r2 <- tpr05_pooled_individual %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, annotation_r2, tpr_fpr05) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~
        paste0("truth-warm EB (r\u00b2=", annotation_r2, ")"),
      use_case_id == "susine_functional_mu" ~
        paste0("truth-warm grid (c=0.5, r\u00b2=", annotation_r2, ")"),
      TRUE ~ use_case_id
    )
  )

collapse_plot_metric_table <- function(df, metric_col, label) {
  if (is.null(df) || !nrow(df)) return(tibble::tibble())

  group_vars <- c("spec_name", "agg_method")
  if ("annotation_r2" %in% names(df)) {
    group_vars <- c(group_vars, "annotation_r2")
  }

  raw <- df %>%
    dplyr::filter(!is.na(.data$spec_name))

  metric_sym <- rlang::sym(metric_col)
  dup_summary <- raw %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::summarise(
      n_rows = dplyr::n(),
      n_metric = sum(is.finite(!!metric_sym)),
      metric_min = suppressWarnings(min(!!metric_sym, na.rm = TRUE)),
      metric_max = suppressWarnings(max(!!metric_sym, na.rm = TRUE)),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      metric_span = dplyr::if_else(
        .data$n_metric > 0L,
        .data$metric_max - .data$metric_min,
        0
      )
    )

  conflicting <- dup_summary %>%
    dplyr::filter(.data$n_rows > 1L, is.finite(.data$metric_span), .data$metric_span > 1e-12)

  if (nrow(conflicting) > 0L) {
    message(
      sprintf(
        "%s had %d duplicate plot-key rows with non-identical %s values; collapsing by mean.",
        label, nrow(conflicting), metric_col
      )
    )
  }

  raw %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::summarise(
      !!metric_col := mean(!!metric_sym, na.rm = TRUE),
      .groups = "drop"
    )
}

auprc_plot_overall <- collapse_plot_metric_table(
  auprc_pooled_agg %>% dplyr::mutate(annotation_r2 = NA_real_),
  metric_col = "AUPRC",
  label = "auprc_pooled_agg"
)
auprc_plot_by_r2 <- collapse_plot_metric_table(
  auprc_pooled_agg_by_r2,
  metric_col = "AUPRC",
  label = "auprc_pooled_agg_by_r2"
)
tpr05_plot_overall <- collapse_plot_metric_table(
  tpr05_pooled_agg %>% dplyr::mutate(annotation_r2 = NA_real_),
  metric_col = "tpr_fpr05",
  label = "tpr05_pooled_agg"
)
tpr05_plot_by_r2 <- collapse_plot_metric_table(
  tpr05_pooled_agg_by_r2,
  metric_col = "tpr_fpr05",
  label = "tpr05_pooled_agg_by_r2"
)
```

# 1. SuSiE2-Style Plots (Category 1)

These use individual model metrics (CS properties, effect sizes).
For ensembles, select single best model by max ELBO.

## 1d. TPR vs FPR (FPR 0 to 0.1)

```{r fig-1d, fig.width=14, fig.height=8}
# Pooled TPR/FPR: combine all datasets' bins, one curve per (spec, agg_method).
# No per-dataset averaging — same pooling approach as AUPRC.
compute_tpr_fpr_pooled <- function(bins, group_vars) {
  bins %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::arrange(dplyr::desc(pip_threshold), .by_group = TRUE) %>%
    dplyr::mutate(
      cum_tp  = cumsum(n_causal_at_bucket),
      cum_fp  = cumsum(n_noncausal_at_bucket),
      total_p = sum(n_causal_at_bucket),
      total_n = sum(n_noncausal_at_bucket)
    ) %>%
    dplyr::filter(total_p > 0, total_n > 0) %>%
    dplyr::mutate(tpr = cum_tp / total_p, fpr = cum_fp / total_n) %>%
    dplyr::filter(fpr <= 0.10) %>%
    dplyr::ungroup()
}

# Ensemble aggregation curves
tpr_fpr_agg <- compute_tpr_fpr_pooled(
  confusion_bins %>% dplyr::filter(!is.na(agg_method)),
  group_vars = c("spec_name", "agg_method")
)

# Oracle: best-by-AUPRC individual run per (spec_name, dataset_bundle_id)
if ("run_id" %in% names(confusion_bins) &&
    "run_id" %in% names(auprc_individual) &&
    "dataset_bundle_id" %in% names(auprc_individual)) {
  oracle_src <- auprc_individual %>% dplyr::filter(!is.na(.data$spec_name))
  if ("agg_method" %in% names(oracle_src))
    oracle_src <- dplyr::filter(oracle_src, is.na(.data$agg_method))
  oracle_sel <- oracle_src %>%
    dplyr::group_by(.data$spec_name, .data$dataset_bundle_id) %>%
    dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::pull(run_id)
  oracle_bins_roc <- confusion_bins %>%
    dplyr::filter(is.na(.data$agg_method), .data$run_id %in% oracle_sel) %>%
    dplyr::mutate(agg_method = "oracle")
  if (nrow(oracle_bins_roc) > 0L)
    tpr_fpr_agg <- dplyr::bind_rows(
      tpr_fpr_agg,
      compute_tpr_fpr_pooled(oracle_bins_roc, group_vars = c("spec_name", "agg_method"))
    )
}

# Baseline single-fit curve (no agg_method) — replicated to all facets
baseline_roc_raw <- compute_tpr_fpr_pooled(
  confusion_bins %>%
    dplyr::filter(spec_name == "baseline-single", is.na(agg_method)),
  group_vars = "spec_name"
)

if (nrow(tpr_fpr_agg) > 0) {
  agg_methods <- sort(unique(tpr_fpr_agg$agg_method))
  baseline_roc <- if (nrow(baseline_roc_raw) > 0) {
    tidyr::crossing(
      baseline_roc_raw %>% dplyr::select(spec_name, fpr, tpr, pip_threshold),
      tibble::tibble(agg_method = agg_methods)
    )
  } else { tibble::tibble() }

  tpr_fpr_plot <- dplyr::bind_rows(tpr_fpr_agg, baseline_roc) %>%
    dplyr::mutate(is_baseline = spec_name == "baseline-single")

  p_roc <- ggplot(tpr_fpr_plot, aes(x = fpr, y = tpr, color = spec_name,
                                     linetype = is_baseline)) +
    geom_line(linewidth = 0.8) +
    scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                          guide = "none") +
    facet_wrap(~agg_method) +
    labs(x = "FPR", y = "TPR", color = "Ensemble Method",
         title = "TPR vs FPR (FPR 0-0.1, pooled across datasets)") +
    theme(legend.position = "bottom")
  ggsave(file.path(fig_root, "susie2_plots", "tpr_fpr_curve.png"),
         p_roc, width = 14, height = 8)
}
```

## 1e. PIP Calibration

```{r fig-1e, fig.width=14, fig.height=8}
# Pool all datasets' bins before binning — same pooling approach as AUPRC.
# Bin PIPs in 0.10-width bins, compute observed fraction causal per bin.
pool_pip_cal <- function(bins) {
  bins %>%
    dplyr::mutate(pip_bin = floor(pip_threshold / 0.10) * 0.10) %>%
    dplyr::group_by(dplyr::across(dplyr::any_of(
      c("spec_name", "agg_method", "pip_bin")
    ))) %>%
    dplyr::summarise(
      total_causal   = sum(n_causal_at_bucket,    na.rm = TRUE),
      total_noncausal = sum(n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      total         = total_causal + total_noncausal,
      observed_frac = dplyr::if_else(total > 0, total_causal / total, NA_real_)
    )
}

# Replicate-level PIP calibration with SE error bars.
# Stage 1: per-dataset observed fraction per PIP bin.
# Stage 2: mean +/- SE across datasets (the independent unit).
pool_pip_cal_with_se <- function(bins) {
  per_dataset <- bins %>%
    dplyr::mutate(pip_bin = floor(.data$pip_threshold / 0.10) * 0.10) %>%
    dplyr::group_by(dplyr::across(dplyr::any_of(
      c("spec_name", "agg_method", "pip_bin", "dataset_bundle_id")
    ))) %>%
    dplyr::summarise(
      n_causal    = sum(.data$n_causal_at_bucket,    na.rm = TRUE),
      n_noncausal = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      n_total       = .data$n_causal + .data$n_noncausal,
      observed_frac = dplyr::if_else(
        .data$n_total > 0L, .data$n_causal / .data$n_total, NA_real_
      )
    )

  per_dataset %>%
    dplyr::group_by(dplyr::across(dplyr::any_of(
      c("spec_name", "agg_method", "pip_bin")
    ))) %>%
    dplyr::summarise(
      mean_frac       = mean(.data$observed_frac, na.rm = TRUE),
      se_frac         = sd(.data$observed_frac,   na.rm = TRUE) /
                          sqrt(sum(!is.na(.data$observed_frac))),
      n_datasets      = sum(!is.na(.data$observed_frac)),
      total_causal    = sum(.data$n_causal,    na.rm = TRUE),
      total_noncausal = sum(.data$n_noncausal, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      total       = .data$total_causal + .data$total_noncausal,
      pooled_frac = dplyr::if_else(
        .data$total > 0L, .data$total_causal / .data$total, NA_real_
      ),
      ci_lo = pmax(0, .data$mean_frac - 1.96 * .data$se_frac),
      ci_hi = pmin(1, .data$mean_frac + 1.96 * .data$se_frac)
    )
}

# Ensemble aggregation calibration
pip_cal_data <- pool_pip_cal_with_se(
  confusion_bins %>% dplyr::filter(!is.na(agg_method))
)

# Baseline single-fit calibration (replicated to all agg_method facets)
baseline_cal_raw <- pool_pip_cal_with_se(
  confusion_bins %>%
    dplyr::filter(spec_name == "baseline-single", is.na(agg_method))
)

if (nrow(pip_cal_data) > 0) {
  agg_methods <- sort(unique(pip_cal_data$agg_method))
  baseline_cal <- if (nrow(baseline_cal_raw) > 0) {
    tidyr::crossing(
      baseline_cal_raw %>%
        dplyr::select(spec_name, pip_bin, mean_frac, se_frac, n_datasets,
                      total_causal, total_noncausal, total, pooled_frac,
                      ci_lo, ci_hi),
      tibble::tibble(agg_method = agg_methods)
    )
  } else { tibble::tibble() }

  pip_cal_all <- dplyr::bind_rows(pip_cal_data, baseline_cal) %>%
    dplyr::mutate(is_baseline = spec_name == "baseline-single")

  p_cal <- ggplot(pip_cal_all, aes(x = pip_bin + 0.05, y = mean_frac,
                                    color = spec_name, linetype = is_baseline)) +
    geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi),
                  width = 0.02, linewidth = 0.4, alpha = 0.6) +
    geom_point(size = 2) +
    geom_line() +
    scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                          guide = "none") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
    facet_wrap(~agg_method) +
    labs(x = "Mean PIP (bin midpoint)", y = "Observed fraction causal",
         color = "Ensemble Method",
         title = "PIP Calibration (mean \u00b1 1.96 SE across datasets)") +
    xlim(0, 1) + ylim(0, 1) +
    theme(legend.position = "bottom")
  ggsave(file.path(fig_root, "susie2_plots", "pip_calibration.png"),
         p_cal, width = 14, height = 8)
}
```

## 1a. CS Power vs top-n-as-causal

```{r fig-1a, fig.width=12, fig.height=6}
# Requires tier_cs_metrics_full.csv from collect workbook.
# Select max-ELBO run per ensemble group (spec_name x dataset_bundle_id x group_key),
# then average cs_power across datasets per (spec_name, n_top).
if (!is.null(tier_cs_metrics_full) && nrow(tier_cs_metrics_full) > 0) {

  # Max-ELBO run per ensemble group from purity-filtered model metrics
  best_runs <- model_metrics %>%
    dplyr::filter(.data$filtering == "purity_filtered",
                  !is.na(.data$elbo_final),
                  !is.na(.data$spec_name)) %>%
    dplyr::group_by(.data$spec_name, .data$dataset_bundle_id, .data$group_key) %>%
    dplyr::slice_max(.data$elbo_final, n = 1, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::select(.data$run_id, .data$spec_name, .data$dataset_bundle_id,
                  .data$group_key)

  # Tier CS metrics for those runs only (individual model fits: is.na(agg_method))
  tier_best <- tier_cs_metrics_full %>%
    dplyr::filter(.data$filtering == "purity_filtered",
                  is.na(.data$agg_method)) %>%
    dplyr::semi_join(best_runs, by = "run_id")

  # Pool counts across datasets per (spec_name, n_top), then derive overall rates.
  # cs_power: sum(n_causal_covered) / sum(n_causal_total)  [overall, not mean]
  # cs_fdr:   sum(n_false_cs)       / sum(n_cs_total)      [overall, not mean]
  tier_avg <- tier_best %>%
    dplyr::group_by(.data$spec_name, .data$n_top) %>%
    dplyr::summarise(
      sum_causal_covered = sum(.data$n_causal_covered, na.rm = TRUE),
      sum_causal_total   = sum(.data$n_causal_total,   na.rm = TRUE),
      sum_false_cs       = sum(.data$n_false_cs,       na.rm = TRUE),
      sum_cs_total       = sum(.data$n_cs_total,       na.rm = TRUE),
      n_datasets         = dplyr::n_distinct(.data$dataset_bundle_id),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      cs_power = dplyr::if_else(.data$sum_causal_total > 0,
                                .data$sum_causal_covered / .data$sum_causal_total,
                                NA_real_),
      cs_fdr   = dplyr::if_else(.data$sum_cs_total > 0,
                                .data$sum_false_cs / .data$sum_cs_total,
                                NA_real_)
    )

  if (nrow(tier_avg) > 0) {
    p_1a <- ggplot(tier_avg, aes(x = factor(.data$n_top), y = .data$cs_power,
                                  color = .data$spec_name, group = .data$spec_name,
                                  linetype = .data$spec_name == "baseline-single")) +
      geom_point(size = 2) +
      geom_line() +
      scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                            guide = "none") +
      labs(x = "Top-N treated as causal", y = "CS Power",
           color = "Ensemble Method",
           title = "CS Power vs Top-N (max-ELBO run per ensemble, pooled across datasets)") +
      theme(legend.position = "bottom")
    ggsave(file.path(fig_root, "susie2_plots", "cs_power_vs_ntop.png"),
           p_1a, width = 12, height = 6)
  } else {
    cat("No tier CS data available after max-ELBO selection.\n")
  }
} else {
  cat("Plot 1a skipped: tier_cs_metrics_full.csv not available.\n")
}
```

## 1b. CS FDR vs top-n-as-causal

```{r fig-1b, fig.width=12, fig.height=6}
if (exists("tier_avg") && nrow(tier_avg) > 0) {
  p_1b <- ggplot(tier_avg, aes(x = factor(.data$n_top), y = .data$cs_fdr,
                                color = .data$spec_name, group = .data$spec_name,
                                linetype = .data$spec_name == "baseline-single")) +
    geom_point(size = 2) +
    geom_line() +
    scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                          guide = "none") +
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red",
               linewidth = 0.8) +
    annotate("text", x = 1, y = 0.05, label = "FDR = 0.05",
             vjust = -0.5, hjust = 0, color = "red", size = 3) +
    labs(x = "Top-N treated as causal", y = "CS FDR",
         color = "Ensemble Method",
         title = "CS FDR vs Top-N (max-ELBO run per ensemble, pooled across datasets)") +
    theme(legend.position = "bottom")
  ggsave(file.path(fig_root, "susie2_plots", "cs_fdr_vs_ntop.png"),
         p_1b, width = 12, height = 6)
} else {
  cat("Plot 1b skipped: tier_avg not available.\n")
}
```

## 1c. Estimated Heritability (hg2) Boxplot

```{r fig-1c, fig.width=14, fig.height=10}
# hg2 per aggregation method, from hg2_by_agg.csv produced by the updated harness.
# Falls back to single-method (max-ELBO only) from model_metrics if the new file
# is not available (pre-update runs).
hg2_path <- file.path(consolidated_dir, "hg2_by_agg.csv")
if (file.exists(hg2_path)) {
  hg2_dat <- readr::read_csv(hg2_path, show_col_types = FALSE)

  # Derive spec_name via group_key if missing (same pattern as fig-2b, fig-3d)
  if (!"spec_name" %in% names(hg2_dat) && "group_label" %in% names(hg2_dat)) {
    spec_map <- model_metrics %>% dplyr::distinct(.data$group_key, .data$spec_name)
    hg2_dat <- hg2_dat %>%
      dplyr::mutate(group_key = sub("^model_grid\\|", "", .data$group_label)) %>%
      dplyr::left_join(spec_map, by = "group_key") %>%
      dplyr::select(-group_key)
  }

  if (nrow(hg2_dat) > 0L && "spec_name" %in% names(hg2_dat)) {
    # Baseline single-fit hg2 from model_metrics (one per dataset)
    baseline_hg2 <- model_metrics %>%
      dplyr::filter(.data$spec_name == "baseline-single",
                    .data$use_case_id == "susine_vanilla",
                    .data$filtering == "purity_filtered",
                    !is.na(.data$hg2)) %>%
      dplyr::distinct(.data$dataset_bundle_id, .keep_all = TRUE) %>%
      dplyr::select(dataset_bundle_id, hg2_baseline = hg2)

    # Append baseline as its own "agg_method" so it appears in every facet
    # via bind_rows: one row per dataset per agg_method
    agg_methods_present <- sort(unique(na.omit(hg2_dat$agg_method)))
    if (nrow(baseline_hg2) > 0L && length(agg_methods_present) > 0L) {
      baseline_rows <- tidyr::crossing(
        baseline_hg2,
        tibble::tibble(agg_method = agg_methods_present)
      ) %>%
        dplyr::mutate(
          spec_name = "baseline-single",
          hg2 = .data$hg2_baseline
        ) %>%
        dplyr::select(dplyr::any_of(names(hg2_dat)))
      hg2_plot_dat <- dplyr::bind_rows(hg2_dat, baseline_rows)
    } else {
      hg2_plot_dat <- hg2_dat
    }

    p_1c <- ggplot(hg2_plot_dat, aes(x = .data$spec_name, y = .data$hg2,
                                      fill = .data$spec_name)) +
      geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
      facet_wrap(~agg_method, ncol = 3) +
      labs(x = "Ensemble Spec", y = expression(hat(h)[g]^2),
           title = "Estimated Heritability by Ensemble Spec and Aggregation Method") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
            legend.position = "none")
    ggsave(file.path(fig_root, "susie2_plots", "hg2_boxplot.png"),
           p_1c, width = 14, height = 10)
    cat("Plot 1c saved (all agg methods).\n")
  } else {
    cat("Plot 1c skipped: hg2_by_agg.csv has no usable rows.\n")
  }
} else if ("hg2" %in% names(model_metrics)) {
  # Fallback: max-ELBO only from model_metrics (pre-update runs)
  hg2_data <- if (exists("best_runs") && nrow(best_runs) > 0) {
    model_metrics %>%
      dplyr::filter(.data$filtering == "purity_filtered",
                    !is.na(.data$hg2), !is.na(.data$spec_name)) %>%
      dplyr::semi_join(best_runs, by = "run_id")
  } else {
    model_metrics %>%
      dplyr::filter(.data$filtering == "purity_filtered",
                    !is.na(.data$hg2), !is.na(.data$spec_name))
  }
  if (nrow(hg2_data) > 0) {
    p_1c <- ggplot(hg2_data, aes(x = .data$spec_name, y = .data$hg2,
                                  fill = .data$spec_name)) +
      geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
      labs(x = "Ensemble Spec", y = expression(hat(h)[g]^2),
           title = "Estimated Heritability by Ensemble Spec (max-ELBO fallback)") +
      theme(axis.text.x = element_text(angle = 30, hjust = 1),
            legend.position = "none")
    ggsave(file.path(fig_root, "susie2_plots", "hg2_boxplot.png"),
           p_1c, width = 14, height = 6)
    cat("Plot 1c saved (max-ELBO fallback — re-run harness for all agg methods).\n")
  } else {
    cat("Plot 1c skipped: no hg2 data after filtering.\n")
  }
} else {
  cat("Plot 1c skipped: hg2_by_agg.csv not found and hg2 column absent from model_metrics.\n")
}
```

# 2. Ensemble Performance Plots (Category 2)

## 2a. Precision-Recall Curves

PR curves pooled across all datasets (trapezoidal). Color = ensemble spec,
facet = aggregation method. Grading uses the architecture's causal SNPs as
stored in the flush confusion bins (all 23 for `susie2_oligogenic`; the 15
tiny polygenic effects inflate N_positives slightly vs strict top-8 grading,
but the relative ranking across methods is unaffected).

```{r fig-2a-helper}
# Build a pooled PR-curve tibble from confusion bins grouped by group_vars.
# Returns tibble with group_vars + recall, precision columns.
build_pr_curves <- function(bins, group_vars) {
  bins %>%
    dplyr::filter(!is.na(.data$agg_method)) %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::arrange(dplyr::desc(.data$pip_threshold), .by_group = TRUE) %>%
    dplyr::mutate(
      cum_tp  = cumsum(.data$n_causal_at_bucket),
      cum_fp  = cumsum(.data$n_noncausal_at_bucket),
      total_p = sum(.data$n_causal_at_bucket)
    ) %>%
    dplyr::filter(.data$total_p > 0L, (.data$cum_tp + .data$cum_fp) > 0L) %>%
    dplyr::mutate(
      recall    = .data$cum_tp / .data$total_p,
      precision = .data$cum_tp / (.data$cum_tp + .data$cum_fp)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::select(dplyr::all_of(group_vars), recall, precision)
}
```

```{r fig-2a, fig.width=14, fig.height=6}
pr_group_vars <- c("spec_name", "agg_method")
pr_curves_overall <- build_pr_curves(confusion_bins, pr_group_vars)

if (nrow(pr_curves_overall) > 0L) {
  p2a <- ggplot2::ggplot(
    pr_curves_overall,
    ggplot2::aes(x = recall, y = precision, color = spec_name, group = spec_name)
  ) +
    ggplot2::geom_line(linewidth = 0.7) +
    ggplot2::facet_wrap(~ agg_method, nrow = 1) +
    ggplot2::scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
    ggplot2::scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
    ggplot2::labs(
      x     = "Recall",
      y     = "Precision",
      color = "Ensemble spec",
      title = "Precision-Recall curves (pooled across datasets, all annotation R\u00b2)"
    ) +
    ggplot2::theme_bw() +
    ggplot2::theme(legend.position = "bottom")

  ggplot2::ggsave(
    file.path(fig_root, "ensemble_performance", "pr_curves_overall.png"),
    p2a, width = 14, height = 5, dpi = 150
  )

  # Per annotation_r2
  pr_curves_by_r2 <- build_pr_curves(
    confusion_bins,
    c("spec_name", "agg_method", "annotation_r2")
  )
  for (r2 in sort(unique(na.omit(pr_curves_by_r2$annotation_r2)))) {
    p2a_r2 <- ggplot2::ggplot(
      pr_curves_by_r2 %>% dplyr::filter(.data$annotation_r2 == r2),
      ggplot2::aes(x = recall, y = precision, color = spec_name, group = spec_name)
    ) +
      ggplot2::geom_line(linewidth = 0.7) +
      ggplot2::facet_wrap(~ agg_method, nrow = 1) +
      ggplot2::scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
      ggplot2::scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
      ggplot2::labs(
        x     = "Recall",
        y     = "Precision",
        color = "Ensemble spec",
        title = sprintf(
          "Precision-Recall curves (pooled, annotation R\u00b2 = %.1f)", r2)
      ) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = "bottom")

    ggplot2::ggsave(
      file.path(fig_root, "ensemble_performance",
                sprintf("pr_curves_r2_%.1f.png", r2)),
      p2a_r2, width = 14, height = 5, dpi = 150
    )
  }
  cat("Fig 2a saved.\n")
} else {
  cat("No aggregated confusion bins found — fig 2a skipped.\n")
}
```

## 2b. Multimodality Metrics Boxplots

```{r fig-2b, fig.width=14, fig.height=8}
mm_path <- file.path(consolidated_dir, "multimodal_metrics.csv")
if (file.exists(mm_path)) {
  mm <- readr::read_csv(mm_path, show_col_types = FALSE)

  # Join spec_name via group_key (strip "model_grid|" prefix from group_label)
  if (!"spec_name" %in% names(mm) && "group_label" %in% names(mm)) {
    spec_map <- model_metrics %>% dplyr::distinct(.data$group_key, .data$spec_name)
    mm <- mm %>%
      dplyr::mutate(group_key = sub("^model_grid\\|", "", .data$group_label)) %>%
      dplyr::left_join(spec_map, by = "group_key") %>%
      dplyr::select(-group_key)
  }

  metrics_to_plot <- intersect(
    c("jaccard_top10", "mean_jsd", "median_jsd", "max_jsd",
      "mean_pip_var", "n_clusters", "n_clusters_jsd_050"),
    names(mm)
  )

  if (length(metrics_to_plot) > 0 && "spec_name" %in% names(mm)) {
    mm_long <- mm %>%
      dplyr::select(dplyr::any_of(c("spec_name", metrics_to_plot))) %>%
      dplyr::filter(!is.na(spec_name)) %>%
      tidyr::pivot_longer(cols = dplyr::any_of(metrics_to_plot),
                          names_to = "metric", values_to = "value")

    p_mm <- ggplot(mm_long, aes(x = spec_name, y = value, fill = spec_name)) +
      geom_boxplot() +
      facet_wrap(~metric, scales = "free_y") +
      labs(x = NULL, y = "Value", title = "Multimodality Metrics") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
            legend.position = "none")
    ggsave(file.path(fig_root, "ensemble_performance",
                     "multimodality_boxplots.png"),
           p_mm, width = 14, height = 8)
  }
}
```

## 2c. AUPRC Bar Chart by Method (primary figure)

Four plots: overall (all annotation R² pooled) plus one per R² level.
Annotation-independent specs (Group A) appear on all plots. Annotation-
dependent specs (Groups B, C) are filtered to the matching R² on per-R²
plots. Truth-warm lines follow the same logic.

```{r fig-2c-helper}
tw_colors <- c("susine_vanilla" = "blue",
               "susine_eb_clamped_scale_var_nonneg" = "darkgreen",
               "susine_functional_mu" = "purple")

make_auprc_barplot <- function(r2_filter = NULL) {
  if (is.null(r2_filter)) {
    plot_data <- auprc_plot_overall
    tw_data   <- truth_warm_overall
    title_str <- "AUPRC by Ensemble Method and Aggregation (all annotation R\u00b2 pooled)"
    fname     <- "auprc_barplot_by_method_overall.png"
  } else {
    # Annotation-independent specs have annotation_r2 = NA — always include.
    # Annotation-dependent specs: filter to the target r2 level.
    plot_data <- auprc_plot_by_r2 %>%
      dplyr::filter(
                    is.na(annotation_r2) | annotation_r2 == r2_filter)
    # Truth-warm: vanilla (NA r2) always; EB/grid at matching r2 only.
    tw_data <- truth_warm_by_r2 %>%
      dplyr::filter(is.na(annotation_r2) | annotation_r2 == r2_filter)
    title_str <- paste0("AUPRC by Ensemble Method and Aggregation (annotation R\u00b2 = ",
                        r2_filter, ")")
    fname <- paste0("auprc_barplot_by_method_r2_",
                    gsub("\\.", "p", as.character(r2_filter)), ".png")
  }

  if (nrow(plot_data) == 0) return(invisible(NULL))

  p <- ggplot(plot_data, aes(x = spec_name, y = AUPRC, fill = agg_method)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    geom_hline(yintercept = baseline_auprc, linetype = "dashed",
               color = "red", linewidth = 0.8) +
    annotate("text", x = 0.5, y = baseline_auprc, label = "baseline single-fit",
             vjust = -0.5, hjust = 0, color = "red", size = 3) +
    labs(x = "Ensemble Method", y = "Pooled AUPRC", fill = "Aggregation",
         title = title_str) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  for (i in seq_len(nrow(tw_data))) {
    uc <- tw_data$use_case_id[i]
    p <- p +
      geom_hline(yintercept = tw_data$AUPRC[i], linetype = "dotted",
                 color = tw_colors[uc] %||% "blue", linewidth = 0.6) +
      annotate("text", x = Inf, y = tw_data$AUPRC[i],
               label = tw_data$label[i],
               vjust = -0.5, hjust = 1,
               color = tw_colors[uc] %||% "blue", size = 2.8)
  }

  ggsave(file.path(fig_root, "ensemble_performance", fname),
         p, width = 12, height = 6)
  invisible(p)
}
```

```{r fig-2c, fig.width=12, fig.height=6}
make_auprc_barplot(r2_filter = NULL)  # overall (all r2 pooled)
make_auprc_barplot(r2_filter = 0)     # annotation r2 = 0 (null annotations)
make_auprc_barplot(r2_filter = 0.2)   # annotation r2 = 0.2
make_auprc_barplot(r2_filter = 0.5)   # annotation r2 = 0.5
```

## 2d. TPR @ FPR=0.05 Bar Chart by Method

Mirrors the AUPRC barplot exactly but shows pooled TPR at FPR=0.05.
Four plots: overall + one per annotation R² level.

```{r fig-2d-helper}
make_tpr05_barplot <- function(r2_filter = NULL) {
  if (is.null(r2_filter)) {
    plot_data <- tpr05_plot_overall
    tw_data   <- truth_warm_tpr05_overall
    title_str <- "TPR @ FPR=0.05 by Ensemble Method and Aggregation (all annotation R\u00b2 pooled)"
    fname     <- "tpr05_barplot_by_method_overall.png"
    base_val  <- baseline_tpr05
  } else {
    plot_data <- tpr05_plot_by_r2 %>%
      dplyr::filter(
                    is.na(annotation_r2) | annotation_r2 == r2_filter)
    tw_data <- truth_warm_tpr05_by_r2 %>%
      dplyr::filter(is.na(annotation_r2) | annotation_r2 == r2_filter)
    # Baseline is annotation-independent; use same pooled overall value
    base_val  <- baseline_tpr05
    title_str <- paste0("TPR @ FPR=0.05 by Ensemble Method and Aggregation (annotation R\u00b2 = ",
                        r2_filter, ")")
    fname <- paste0("tpr05_barplot_by_method_r2_",
                    gsub("\\.", "p", as.character(r2_filter)), ".png")
  }

  if (nrow(plot_data) == 0 || length(base_val) == 0) return(invisible(NULL))

  p <- ggplot(plot_data, aes(x = spec_name, y = tpr_fpr05, fill = agg_method)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    geom_hline(yintercept = base_val, linetype = "dashed",
               color = "red", linewidth = 0.8) +
    annotate("text", x = 0.5, y = base_val, label = "baseline single-fit",
             vjust = -0.5, hjust = 0, color = "red", size = 3) +
    labs(x = "Ensemble Method", y = "Pooled TPR @ FPR=0.05", fill = "Aggregation",
         title = title_str) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  for (i in seq_len(nrow(tw_data))) {
    uc <- tw_data$use_case_id[i]
    p <- p +
      geom_hline(yintercept = tw_data$tpr_fpr05[i], linetype = "dotted",
                 color = tw_colors[uc] %||% "blue", linewidth = 0.6) +
      annotate("text", x = Inf, y = tw_data$tpr_fpr05[i],
               label = tw_data$label[i],
               vjust = -0.5, hjust = 1,
               color = tw_colors[uc] %||% "blue", size = 2.8)
  }

  ggsave(file.path(fig_root, "ensemble_performance", fname),
         p, width = 12, height = 6)
  invisible(p)
}
```

```{r fig-2d, fig.width=12, fig.height=6}
make_tpr05_barplot(r2_filter = NULL)  # overall (all r2 pooled)
make_tpr05_barplot(r2_filter = 0)     # annotation r2 = 0 (null annotations)
make_tpr05_barplot(r2_filter = 0.2)   # annotation r2 = 0.2
make_tpr05_barplot(r2_filter = 0.5)   # annotation r2 = 0.5
```

## 2e. C-CS Grid Landscape (AUPRC by c × σ₀², faceted by annotation R²)

Individual-fit AUPRC pooled across datasets for each (c, σ₀²) cell in
the C-CS grid. Validates that annotation quality shifts the optimal c.

```{r fig-2e, fig.width=14, fig.height=5}
{
  cs_bins <- confusion_bins %>%
    dplyr::filter(.data$spec_name == "C-CS",
                  is.na(.data$agg_method),
                  !is.na(.data$c_value),
                  !is.na(.data$sigma_0_2_scalar))

  if (nrow(cs_bins) > 0L) {
    # Pool confusion bins across datasets per (c_value, sigma_0_2_scalar, annotation_r2)
    cs_pooled <- cs_bins %>%
      dplyr::group_by(.data$c_value, .data$sigma_0_2_scalar,
                       .data$annotation_r2, .data$pip_threshold) %>%
      dplyr::summarise(
        n_causal_at_bucket    = sum(.data$n_causal_at_bucket,    na.rm = TRUE),
        n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
        .groups = "drop"
      )

    # Compute AUPRC per (c, sigma, r2) cell
    cs_auprc <- cs_pooled %>%
      dplyr::group_by(.data$c_value, .data$sigma_0_2_scalar, .data$annotation_r2) %>%
      dplyr::summarise(
        AUPRC = auprc_from_bins(dplyr::pick(dplyr::everything())),
        .groups = "drop"
      ) %>%
      dplyr::mutate(
        sigma_label = factor(
          .data$sigma_0_2_scalar,
          levels = sort(unique(.data$sigma_0_2_scalar))
        ),
        r2_label = paste0("annotation R\u00b2 = ", .data$annotation_r2)
      )

    p_cs <- ggplot(cs_auprc,
                   aes(x = c_value, y = AUPRC,
                       color = sigma_label, group = sigma_label)) +
      geom_line(linewidth = 0.7) +
      geom_point(size = 1.8) +
      geom_hline(yintercept = baseline_auprc, linetype = "dashed",
                 color = "red", linewidth = 0.6) +
      facet_wrap(~r2_label, nrow = 1) +
      labs(x = "c (annotation scale)",
           y = "Pooled AUPRC (individual fits)",
           color = expression(sigma[0]^2),
           title = "C-CS grid landscape: AUPRC by c and \u03c3\u00b2\u2080, faceted by annotation R\u00b2") +
      theme_bw() +
      theme(legend.position = "bottom")
    ggsave(file.path(fig_root, "ensemble_performance",
                     "cs_grid_landscape.png"),
           p_cs, width = 14, height = 5, dpi = 150)
    cat("Fig 2e saved.\n")
  } else {
    cat("Fig 2e skipped: no C-CS individual-fit confusion bins.\n")
  }
}
```

# 3. Dataset Difficulty Plots (Category 3)

## 3a. Dataset Metrics

```{r fig-3a, fig.width=12, fig.height=8}
dm_path <- file.path(consolidated_dir, "dataset_metrics.csv")
if (file.exists(dm_path)) {
  dataset_metrics <- readr::read_csv(dm_path, show_col_types = FALSE) %>%
    dplyr::distinct(dataset_bundle_id, .keep_all = TRUE)

  metric_cols <- intersect(
    c("M1", "z_topk_ratio", "z_max_abs", "z_count_abs_gt_3",
      "z_eff_signals", "high_ld_count_095"),
    names(dataset_metrics)
  )

  if (length(metric_cols) > 0) {
    dm_long <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(metric_cols)) %>%
      tidyr::pivot_longer(everything(), names_to = "metric", values_to = "value")

    p_hist <- ggplot(dm_long, aes(x = value)) +
      geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7) +
      facet_wrap(~metric, scales = "free") +
      labs(x = "Value", y = "Count", title = "Dataset Metrics")
    ggsave(file.path(fig_root, "dataset_difficulty",
                     "metric_histograms.png"),
           p_hist, width = 12, height = 8)
  }
}
```

## 3b. Dataset Difficulty Scatter

```{r fig-3b, fig.width=10, fig.height=5}
if (exists("dataset_metrics") && "M1" %in% names(dataset_metrics)) {
  # Join baseline AUPRC to dataset metrics
  baseline_per_dataset <- auprc_individual %>%
    dplyr::filter(spec_name == "baseline-single",
                  use_case_id == "susine_vanilla") %>%
    dplyr::select(dataset_bundle_id, auprc_baseline = AUPRC)

  dm_scatter <- dataset_metrics %>%
    dplyr::inner_join(baseline_per_dataset, by = "dataset_bundle_id")

  if (nrow(dm_scatter) > 1) {
    p_scatter <- ggplot(dm_scatter, aes(x = M1, y = auprc_baseline)) +
      geom_point(alpha = 0.6) +
      geom_smooth(method = "lm", se = TRUE, color = "red") +
      labs(x = "M1 (LD complexity)", y = "Baseline AUPRC",
           title = "Dataset Difficulty: AUPRC vs M1")
    ggsave(file.path(fig_root, "dataset_difficulty",
                     "difficulty_scatter.png"),
           p_scatter, width = 10, height = 5)

    # Multi-panel scatter: baseline AUPRC vs all available difficulty metrics
    x_metrics <- intersect(
      c("M1", "z_topk_ratio", "z_max_abs", "z_count_abs_gt_3", "z_eff_signals",
        "high_ld_count_095", "high_ld_count_095_per_snp", "high_ld_frac_095"),
      names(dm_scatter)
    )
    if (length(x_metrics) > 1L) {
      dm_multi <- dm_scatter %>%
        tidyr::pivot_longer(
          cols      = dplyr::any_of(x_metrics),
          names_to  = "x_metric",
          values_to = "x_value"
        )
      p_multi <- ggplot(dm_multi, aes(x = x_value, y = auprc_baseline)) +
        geom_point(alpha = 0.5) +
        geom_smooth(method = "lm", se = TRUE, color = "red") +
        facet_wrap(~x_metric, scales = "free_x") +
        labs(x = "Dataset metric", y = "Baseline AUPRC",
             title = "Baseline AUPRC vs all dataset difficulty metrics")
      ggsave(file.path(fig_root, "dataset_difficulty",
                       "baseline_auprc_vs_all_metrics_scatter.png"),
             p_multi, width = 14, height = 8)
    }
  }
}
```

## 3c. Difficulty Model Search

```{r fig-3c-difficulty-model-search}
html_escape <- function(x) {
  x <- as.character(x)
  x <- gsub("&", "&amp;", x, fixed = TRUE)
  x <- gsub("<", "&lt;", x, fixed = TRUE)
  x <- gsub(">", "&gt;", x, fixed = TRUE)
  x
}

format_html_table <- function(df, digits = 4L) {
  if (is.null(df) || !nrow(df)) {
    return("<p>No rows to display.</p>")
  }

  tbl <- df
  for (nm in names(tbl)) {
    if (is.numeric(tbl[[nm]])) {
      tbl[[nm]] <- ifelse(
        is.na(tbl[[nm]]),
        "",
        formatC(tbl[[nm]], digits = digits, format = "f")
      )
    } else {
      tbl[[nm]] <- ifelse(is.na(tbl[[nm]]), "", as.character(tbl[[nm]]))
    }
  }

  header <- paste0(
    "<tr>",
    paste(sprintf("<th>%s</th>", html_escape(names(tbl))), collapse = ""),
    "</tr>"
  )
  rows <- apply(tbl, 1, function(row) {
    paste0(
      "<tr>",
      paste(sprintf("<td>%s</td>", html_escape(row)), collapse = ""),
      "</tr>"
    )
  })

  paste0(
    "<table class='report-table'>",
    "<thead>", header, "</thead>",
    "<tbody>", paste(rows, collapse = ""), "</tbody>",
    "</table>"
  )
}

write_html_report <- function(path, title, sections) {
  html_sections <- vapply(sections, function(section) {
    header_html <- if (!is.null(section$header) && nzchar(section$header)) {
      sprintf("<h2>%s</h2>", html_escape(section$header))
    } else {
      ""
    }
    text_html <- if (!is.null(section$text) && length(section$text)) {
      paste(sprintf("<p>%s</p>", html_escape(section$text)), collapse = "")
    } else {
      ""
    }
    table_html <- if (!is.null(section$table)) {
      format_html_table(section$table, digits = section$digits %||% 4L)
    } else {
      ""
    }
    paste0(header_html, text_html, table_html)
  }, character(1))

  html <- paste0(
    "<!DOCTYPE html><html><head><meta charset='utf-8'>",
    "<title>", html_escape(title), "</title>",
    "<style>",
    "body{font-family:Arial,sans-serif;margin:24px;line-height:1.4;}",
    "h1,h2{margin-bottom:0.4em;}",
    ".report-table{border-collapse:collapse;width:100%;margin:12px 0 24px 0;}",
    ".report-table th,.report-table td{border:1px solid #cccccc;padding:6px 8px;text-align:left;}",
    ".report-table th{background:#f3f3f3;}",
    "code{background:#f5f5f5;padding:2px 4px;}",
    "</style></head><body>",
    "<h1>", html_escape(title), "</h1>",
    paste(html_sections, collapse = ""),
    "</body></html>"
  )
  writeLines(html, con = path, useBytes = TRUE)
}

extract_per_dataset_outcome <- function(data,
                                        outcome_name,
                                        spec_name,
                                        use_case_id = NULL,
                                        agg_method = NULL,
                                        annotation_r2 = NULL) {
  out <- data %>%
    dplyr::filter(.data$spec_name == !!spec_name)
  if (!is.null(use_case_id)) {
    out <- out %>% dplyr::filter(.data$use_case_id == !!use_case_id)
  }
  if (is.null(agg_method)) {
    if ("agg_method" %in% names(out)) {
      out <- out %>% dplyr::filter(is.na(.data$agg_method))
    }
  } else {
    out <- out %>% dplyr::filter(.data$agg_method == !!agg_method)
  }
  if (!is.null(annotation_r2)) {
    out <- out %>% dplyr::filter(.data$annotation_r2 == !!annotation_r2)
  }
  out %>%
    dplyr::group_by(.data$dataset_bundle_id) %>%
    dplyr::summarise(!!outcome_name := dplyr::first(.data$AUPRC), .groups = "drop")
}

enumerate_metric_pairs <- function(metrics) {
  if (length(metrics) < 2L) {
    return(tibble::tibble(metric1 = character(0), metric2 = character(0)))
  }
  pairs <- utils::combn(metrics, 2, simplify = FALSE)
  tibble::tibble(
    metric1 = vapply(pairs, `[`, character(1), 1),
    metric2 = vapply(pairs, `[`, character(1), 2)
  )
}

fit_difficulty_model <- function(data, outcome_col, metric1, metric2) {
  vars <- c(outcome_col, "M1", metric1, metric2)
  model_df <- data %>%
    dplyr::select(dplyr::all_of(c("dataset_bundle_id", vars))) %>%
    dplyr::filter(dplyr::if_all(dplyr::all_of(vars), ~ is.finite(.x)))

  formula_txt <- sprintf("%s ~ (M1 + %s + %s)^2", outcome_col, metric1, metric2)
  n_used <- nrow(model_df)

  if (n_used < 8L) {
    return(list(
      summary = tibble::tibble(
        outcome = outcome_col,
        metric1 = metric1,
        metric2 = metric2,
        formula = formula_txt,
        n_used = n_used,
        r_squared = NA_real_,
        adj_r_squared = NA_real_,
        rmse = NA_real_,
        status = "insufficient_rows",
        singular = NA
      ),
      coef = tibble::tibble(),
      predictions = model_df
    ))
  }

  fit <- tryCatch(
    stats::lm(stats::as.formula(formula_txt), data = model_df),
    error = function(e) e
  )
  if (inherits(fit, "error")) {
    return(list(
      summary = tibble::tibble(
        outcome = outcome_col,
        metric1 = metric1,
        metric2 = metric2,
        formula = formula_txt,
        n_used = n_used,
        r_squared = NA_real_,
        adj_r_squared = NA_real_,
        rmse = NA_real_,
        status = paste0("fit_error: ", fit$message),
        singular = NA
      ),
      coef = tibble::tibble(),
      predictions = model_df
    ))
  }

  fit_summary <- summary(fit)
  pred <- stats::predict(fit, newdata = model_df)
  resid <- model_df[[outcome_col]] - pred
  coef_mat <- fit_summary$coefficients
  coef_tbl <- tibble::as_tibble(coef_mat, rownames = "term") %>%
    dplyr::rename(
      estimate = "Estimate",
      std_error = "Std. Error",
      statistic = "t value",
      p_value = "Pr(>|t|)"
    )
  singular <- any(is.na(stats::coef(fit)))

  list(
    summary = tibble::tibble(
      outcome = outcome_col,
      metric1 = metric1,
      metric2 = metric2,
      formula = formula_txt,
      n_used = n_used,
      r_squared = unname(fit_summary$r.squared),
      adj_r_squared = unname(fit_summary$adj.r.squared),
      rmse = sqrt(mean(resid^2)),
      status = if (singular) "ok_singular" else "ok",
      singular = singular
    ),
    coef = coef_tbl,
    predictions = model_df %>%
      dplyr::mutate(predicted = pred, residual = resid)
  )
}

candidate_metric_pool <- c(
  "M1", "high_ld_count_095", "high_ld_count_095_per_snp", "high_ld_frac_095",
  "z_topk_ratio", "z_max_abs", "z_count_abs_gt_3", "z_eff_signals"
)

difficulty_model_data <- NULL
baseline_model_search <- NULL
baseline_best_fit <- NULL
difficulty_refits <- NULL

if (exists("dataset_metrics") && is.data.frame(dataset_metrics)) {
  baseline_per_dataset <- extract_per_dataset_outcome(
    auprc_individual,
    outcome_name = "auprc_baseline",
    spec_name = "baseline-single",
    use_case_id = "susine_vanilla",
    agg_method = NULL
  )
  oracle_per_dataset <- extract_per_dataset_outcome(
    auprc_agg,
    outcome_name = "auprc_ccs_oracle",
    spec_name = "C-CS",
    agg_method = "oracle",
    annotation_r2 = 0.5
  )
  cluster_weight_per_dataset <- extract_per_dataset_outcome(
    auprc_agg,
    outcome_name = "auprc_ccs_cluster_weight",
    spec_name = "C-CS",
    agg_method = "cluster_weight",
    annotation_r2 = 0.5
  )

  difficulty_model_data <- dataset_metrics %>%
    dplyr::distinct(.data$dataset_bundle_id, .keep_all = TRUE) %>%
    dplyr::left_join(baseline_per_dataset, by = "dataset_bundle_id") %>%
    dplyr::left_join(oracle_per_dataset, by = "dataset_bundle_id") %>%
    dplyr::left_join(cluster_weight_per_dataset, by = "dataset_bundle_id") %>%
    dplyr::mutate(
      gain_oracle_vs_baseline = .data$auprc_ccs_oracle - .data$auprc_baseline,
      gain_cluster_weight_vs_baseline = .data$auprc_ccs_cluster_weight - .data$auprc_baseline
    )

  candidate_metrics <- intersect(candidate_metric_pool, names(difficulty_model_data))
  candidate_metrics <- candidate_metrics[vapply(candidate_metrics, function(metric) {
    vals <- difficulty_model_data[[metric]]
    vals <- vals[is.finite(vals)]
    length(vals) > 1L && stats::var(vals) > 0
  }, logical(1))]

  extra_metrics <- setdiff(candidate_metrics, "M1")
  metric_pairs <- enumerate_metric_pairs(extra_metrics)

  if ("M1" %in% candidate_metrics &&
      nrow(metric_pairs) > 0L &&
      "auprc_baseline" %in% names(difficulty_model_data)) {
    baseline_fits <- purrr::pmap(metric_pairs, function(metric1, metric2) {
      fit_difficulty_model(difficulty_model_data, "auprc_baseline", metric1, metric2)
    })
    baseline_model_search <- dplyr::bind_rows(lapply(baseline_fits, `[[`, "summary")) %>%
      dplyr::arrange(dplyr::desc(.data$adj_r_squared),
                     dplyr::desc(.data$r_squared),
                     .data$metric1,
                     .data$metric2)

    best_idx <- which(baseline_model_search$status %in% c("ok", "ok_singular"))[1]
    if (length(best_idx) && !is.na(best_idx)) {
      best_row <- baseline_model_search[best_idx, , drop = FALSE]
      baseline_best_fit <- fit_difficulty_model(
        difficulty_model_data,
        "auprc_baseline",
        best_row$metric1[[1]],
        best_row$metric2[[1]]
      )
      difficulty_model_data <- difficulty_model_data %>%
        dplyr::left_join(
          baseline_best_fit$predictions %>%
            dplyr::select(
              dataset_bundle_id,
              baseline_predicted_auprc = predicted,
              baseline_residual = residual
            ),
          by = "dataset_bundle_id"
        ) %>%
        dplyr::mutate(
          oracle_residual_vs_baseline_prediction =
            .data$auprc_ccs_oracle - .data$baseline_predicted_auprc,
          cluster_weight_residual_vs_baseline_prediction =
            .data$auprc_ccs_cluster_weight - .data$baseline_predicted_auprc
        )

      refit_specs <- tibble::tribble(
        ~outcome, ~method_label, ~html_name,
        "auprc_baseline", "baseline", "difficulty_model_refit_baseline.html",
        "auprc_ccs_oracle", "C-CS oracle", "difficulty_model_refit_ccs_oracle_r2_0p5.html",
        "auprc_ccs_cluster_weight", "C-CS cluster_weight", "difficulty_model_refit_ccs_cluster_weight_r2_0p5.html"
      )

      difficulty_refits <- purrr::pmap_dfr(refit_specs, function(outcome, method_label, html_name) {
        refit <- fit_difficulty_model(
          difficulty_model_data,
          outcome,
          best_row$metric1[[1]],
          best_row$metric2[[1]]
        )
        coef_tbl <- refit$coef
        if (nrow(coef_tbl)) {
          coef_tbl <- coef_tbl %>% dplyr::mutate(method = method_label, outcome = outcome)
        }
        html_path <- file.path(fig_root, "dataset_difficulty", html_name)
        write_html_report(
          html_path,
          title = paste0("Difficulty model refit: ", method_label),
          sections = list(
            list(
              header = "Model Summary",
              text = c(
                paste0("Locked metric pair from baseline search: ",
                       best_row$metric1[[1]], " and ", best_row$metric2[[1]]),
                paste0("Formula: ", refit$summary$formula[[1]])
              ),
              table = refit$summary
            ),
            list(
              header = "Coefficients",
              table = coef_tbl
            )
          )
        )
        refit$summary %>% dplyr::mutate(method = method_label, html_file = basename(html_path))
      })

      write_html_report(
        file.path(fig_root, "dataset_difficulty", "difficulty_model_search_baseline.html"),
        title = "Baseline difficulty model search",
        sections = list(
          list(
            header = "Search Summary",
            text = c(
              paste0("Candidate metrics considered: ",
                     paste(candidate_metrics, collapse = ", ")),
              paste0("Unordered metric pairs searched: ", nrow(metric_pairs)),
              paste0("Selected pair: ", best_row$metric1[[1]], " and ", best_row$metric2[[1]]),
              paste0("Selected formula: ", best_row$formula[[1]])
            ),
            table = best_row
          ),
          list(
            header = "Selected Model Coefficients",
            table = baseline_best_fit$coef
          ),
          list(
            header = "All Candidate Models",
            table = baseline_model_search
          )
        )
      )

      readr::write_csv(
        baseline_model_search,
        file.path(fig_root, "dataset_difficulty", "difficulty_model_search_baseline.csv")
      )
      readr::write_csv(
        difficulty_refits,
        file.path(fig_root, "dataset_difficulty", "difficulty_model_refits.csv")
      )
      readr::write_csv(
        difficulty_model_data,
        file.path(fig_root, "dataset_difficulty", "difficulty_model_data.csv")
      )

      cat("Difficulty model search files saved.\n")
    } else {
      cat("Difficulty model search skipped: no successful baseline fits.\n")
    }
  } else {
    cat("Difficulty model search skipped: required metrics or baseline outcome unavailable.\n")
  }
}
```

## 3d. Difficulty Model Comparison

```{r fig-3d-difficulty-model-comparison, fig.width=10, fig.height=6}
if (!is.null(difficulty_model_data) &&
    "baseline_predicted_auprc" %in% names(difficulty_model_data)) {
  comparison_long <- difficulty_model_data %>%
    dplyr::select(dplyr::all_of(c(
      "dataset_bundle_id",
      "baseline_predicted_auprc",
      "auprc_baseline",
      "auprc_ccs_oracle",
      "auprc_ccs_cluster_weight"
    ))) %>%
    tidyr::pivot_longer(
      cols = dplyr::all_of(c(
        "auprc_baseline",
        "auprc_ccs_oracle",
        "auprc_ccs_cluster_weight"
      )),
      names_to = "method",
      values_to = "observed_auprc"
    ) %>%
    dplyr::filter(is.finite(.data$baseline_predicted_auprc),
                  is.finite(.data$observed_auprc)) %>%
    dplyr::mutate(
      method = dplyr::recode(
        .data$method,
        auprc_baseline = "baseline",
        auprc_ccs_oracle = "C-CS oracle",
        auprc_ccs_cluster_weight = "C-CS cluster_weight"
      )
    )

  if (nrow(comparison_long) > 0L) {
    p_difficulty_compare <- ggplot(
      comparison_long,
      aes(x = baseline_predicted_auprc, y = observed_auprc, color = method)
    ) +
      geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") +
      geom_point(alpha = 0.7, size = 2) +
      geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
      labs(
        x = "Predicted baseline AUPRC (difficulty model)",
        y = "Observed AUPRC",
        color = "Method",
        title = "Observed AUPRC vs baseline-predicted difficulty"
      ) +
      theme_bw()
    ggsave(
      file.path(fig_root, "dataset_difficulty", "difficulty_model_predictions_comparison.png"),
      p_difficulty_compare, width = 10, height = 6
    )
    cat("Difficulty model comparison plot saved.\n")
  } else {
    cat("Difficulty model comparison skipped: no complete rows.\n")
  }
} else {
  cat("Difficulty model comparison skipped: baseline predictions unavailable.\n")
}
```

## 3e. Optional Diagnostics: Multimodality Metrics vs M1

```{r fig-3e-legacy-mm, fig.width=14, fig.height=8, eval=FALSE}
{
  mm_local <- if (exists("mm") && is.data.frame(mm)) mm else {
    mm_p <- file.path(consolidated_dir, "multimodal_metrics.csv")
    if (file.exists(mm_p)) readr::read_csv(mm_p, show_col_types = FALSE) else NULL
  }
  # Attach spec_name via group_key (strip "model_grid|" prefix from group_label)
  if (!is.null(mm_local) && !"spec_name" %in% names(mm_local) &&
      "group_label" %in% names(mm_local)) {
    spec_map <- model_metrics %>% dplyr::distinct(.data$group_key, .data$spec_name)
    mm_local <- mm_local %>%
      dplyr::mutate(group_key = sub("^model_grid\\|", "", .data$group_label)) %>%
      dplyr::left_join(spec_map, by = "group_key") %>%
      dplyr::select(-group_key)
  }

  if (!is.null(mm_local) && exists("dataset_metrics") &&
      "dataset_bundle_id" %in% names(mm_local) &&
      "dataset_bundle_id" %in% names(dataset_metrics) &&
      "M1" %in% names(dataset_metrics)) {

    mm_m1 <- mm_local %>%
      dplyr::inner_join(
        dataset_metrics %>% dplyr::select(dataset_bundle_id, M1),
        by = "dataset_bundle_id"
      )

    mm_metrics <- intersect(
      c("mean_jsd", "n_clusters", "jaccard_top10", "mean_pip_var"),
      names(mm_m1)
    )
    if (nrow(mm_m1) > 1L && length(mm_metrics) > 0L &&
        "spec_name" %in% names(mm_m1)) {
      mm_long <- mm_m1 %>%
        dplyr::filter(!is.na(.data$spec_name)) %>%
        tidyr::pivot_longer(
          cols      = dplyr::any_of(mm_metrics),
          names_to  = "metric",
          values_to = "value"
        )
      p_3d <- ggplot(mm_long, aes(x = M1, y = value, color = spec_name)) +
        geom_point(alpha = 0.5, size = 1.2) +
        geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
        facet_wrap(~metric, scales = "free_y") +
        labs(x = "M1 (LD complexity)", y = "Multimodality metric",
             color = "Spec",
             title = "Multimodality metrics vs M1 (all specs)") +
        theme(legend.position = "bottom")
      ggsave(file.path(fig_root, "dataset_difficulty",
                       "multimodality_vs_m1.png"),
             p_3d, width = 14, height = 8)
      cat("Fig 3d saved.\n")
    } else {
      cat("Fig 3d skipped: insufficient data.\n")
    }
  } else {
    cat("Fig 3d skipped: multimodal_metrics or dataset_metrics not available.\n")
  }
}
```

## 3f. Optional Diagnostics: Compute-Confound Diversity Comparison (A-R vs C-C)

```{r fig-3f-legacy-diversity, fig.width=12, fig.height=6, eval=FALSE}
{
  mm_local_e <- if (exists("mm_local") && is.data.frame(mm_local)) mm_local else {
    mm_p <- file.path(consolidated_dir, "multimodal_metrics.csv")
    if (file.exists(mm_p)) {
      tmp <- readr::read_csv(mm_p, show_col_types = FALSE)
      if (!"spec_name" %in% names(tmp) && "run_id" %in% names(tmp)) {
        spec_map <- model_metrics %>% dplyr::distinct(.data$run_id, .data$spec_name)
        tmp <- tmp %>% dplyr::left_join(spec_map, by = "run_id")
      }
      tmp
    } else { NULL }
  }

  if (!is.null(mm_local_e) && "spec_name" %in% names(mm_local_e)) {
    cc_specs  <- c("A-R", "C-C")
    mm_cc     <- mm_local_e %>% dplyr::filter(.data$spec_name %in% cc_specs)
    div_metrics <- intersect(
      c("mean_jsd", "n_clusters", "jaccard_top10", "mean_pip_var"),
      names(mm_cc)
    )
    if (nrow(mm_cc) > 0L && length(div_metrics) > 0L &&
        length(unique(mm_cc$spec_name)) >= 2L) {
      mm_cc_long <- mm_cc %>%
        tidyr::pivot_longer(
          cols      = dplyr::any_of(div_metrics),
          names_to  = "metric",
          values_to = "value"
        )
      p_3e <- ggplot(mm_cc_long,
                     aes(x = spec_name, y = value, fill = spec_name)) +
        geom_boxplot(alpha = 0.7, outlier.size = 0.8) +
        facet_wrap(~metric, scales = "free_y") +
        labs(x = "Spec", y = "Value",
             title = "Diversity comparison: A-R (random restarts) vs C-C (c-grid)") +
        theme(legend.position = "none")
      ggsave(file.path(fig_root, "dataset_difficulty",
                       "diversity_comparison_AR_CC.png"),
             p_3e, width = 12, height = 6)
      cat("Fig 3e saved.\n")
    } else {
      cat("Fig 3e skipped: A-R and/or C-C not both present in multimodal_metrics.\n")
    }
  } else {
    cat("Fig 3e skipped: multimodal_metrics not available.\n")
  }
}
```

## 3f. Failure Map (M1 × Z_primary, colored by baseline AUPRC)

```{r fig-3g-legacy-failure-map, fig.width=10, fig.height=6, eval=FALSE}
{
  z_candidates <- c("z_topk_ratio", "z_max_abs", "z_count_abs_gt_3", "z_eff_signals")
  if (exists("dm_scatter") && is.data.frame(dm_scatter) && nrow(dm_scatter) >= 4L &&
      exists("dataset_metrics") && "M1" %in% names(dm_scatter)) {
    z_avail   <- intersect(z_candidates, names(dataset_metrics))
    z_primary <- if (length(z_avail) > 0L) z_avail[1L] else NA_character_
    if (!is.na(z_primary)) {
      dm_fm <- dm_scatter   # dm_scatter already contains all dataset_metrics columns
      if (z_primary %in% names(dm_fm) &&
          sum(!is.na(dm_fm[[z_primary]])) >= 4L) {
        p_3f <- ggplot(dm_fm,
                       aes(x = M1, y = .data[[z_primary]],
                           color = auprc_baseline)) +
          geom_point(size = 2.5, alpha = 0.85) +
          ggplot2::scale_color_viridis_c(option = "plasma",
                                         name = "Baseline\nAUPRC") +
          labs(x = "M1 (LD complexity)",
               y = z_primary,
               title = paste0("Failure map: M1 \u00d7 ", z_primary,
                              " (color = baseline AUPRC)")) +
          theme_bw()
        ggsave(file.path(fig_root, "dataset_difficulty", "failure_map.png"),
               p_3f, width = 10, height = 6)
        cat("Fig 3f saved.\n")
      } else {
        cat("Fig 3f skipped: z_primary column not available or < 4 non-NA rows.\n")
      }
    } else {
      cat("Fig 3f skipped: no z-metrics found in dataset_metrics.\n")
    }
  } else {
    cat("Fig 3f skipped: dm_scatter not available or < 4 datasets.\n")
  }
}
```

## 3h. Refine Depth Diagnostics

```{r fig-3h-refine-depth, fig.width=14, fig.height=6}
# Per-parent refine depth: n_refine_processed vs n_refine_requested.
# One row per (dataset × ensemble group × parent combination).
# Loaded from refine_depth.csv written by the collect workbook.
rd_path <- file.path(consolidated_dir, "refine_depth.csv")
rd_full <- if (file.exists(rd_path)) {
  readr::read_csv(rd_path, show_col_types = FALSE)
} else {
  NULL
}

if (!is.null(rd_full) && nrow(rd_full) > 0L &&
    all(c("n_refine_requested", "n_refine_processed", "group_label") %in% names(rd_full))) {

  # Attach spec_name via group_label → run_info join if available
  rd <- rd_full %>%
    dplyr::mutate(completion_rate = .data$n_refine_processed / .data$n_refine_requested)
  if (exists("run_info") && "spec_name" %in% names(run_info) && "group_label" %in% names(run_info)) {
    spec_map <- run_info %>% dplyr::distinct(.data$group_label, .data$spec_name)
    rd <- dplyr::left_join(rd, spec_map, by = "group_label")
  }
  x_var <- if ("spec_name" %in% names(rd)) "spec_name" else "group_label"

  ensure_dir(file.path(fig_root, "ensemble_scaling"))

  # Completion rate boxplot: one point per parent node
  p_rate <- ggplot(rd, aes(x = .data[[x_var]], y = completion_rate)) +
    geom_boxplot(alpha = 0.7, outlier.size = 1, fill = "steelblue") +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
    scale_y_continuous(limits = c(0, 1.05), labels = scales::percent) +
    labs(x = "Spec", y = "Completion rate (processed / requested)",
         title = "Refine BFS completion rate per parent node") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
  ggsave(file.path(fig_root, "ensemble_scaling", "refine_depth_completion_rate.png"),
         p_rate, width = 14, height = 6)

  # Histogram of n_refine_processed: proportion y-axis, adaptive binwidth per facet.
  # binwidth = 1 when n_requested <= 8 (integer ticks); ceil(n_requested/8) otherwise.
  # Pre-compute frequency tables per group so ggplot2 doesn't pool counts cross-facet.
  n_req_by_group <- rd %>%
    dplyr::group_by(.data[[x_var]]) %>%
    dplyr::summarise(n_requested = as.integer(median(.data$n_refine_requested)),
                     .groups = "drop")

  hist_freq <- rd %>%
    dplyr::left_join(n_req_by_group, by = x_var) %>%
    dplyr::mutate(
      bw      = dplyr::if_else(.data$n_requested <= 8L, 1L,
                               as.integer(ceiling(.data$n_requested / 8L))),
      bin_idx = floor(.data$n_refine_processed / .data$bw),
      bin_lo  = .data$bin_idx * .data$bw,
      bin_hi  = (.data$bin_idx + 1L) * .data$bw
    ) %>%
    dplyr::group_by(.data[[x_var]], .data$n_requested, .data$bw,
                    .data$bin_lo, .data$bin_hi) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    dplyr::group_by(.data[[x_var]]) %>%
    dplyr::mutate(
      proportion  = .data$n / sum(.data$n),
      bin_center  = .data$bin_lo + 0.5 * (.data$bw - 1L)
    ) %>%
    dplyr::ungroup()

  p_hist <- ggplot(hist_freq,
                   aes(x = bin_center, y = proportion, width = bw - 0.1)) +
    geom_col(fill = "steelblue", color = "white", alpha = 0.8) +
    geom_vline(data = n_req_by_group,
               aes(xintercept = n_requested),
               linetype = "dashed", color = "red") +
    facet_wrap(as.formula(paste("~", x_var)), scales = "free_x") +
    scale_x_continuous(breaks = function(lims) {
      # Show bin centers: step size matches the binwidth used for this facet's range.
      n_req <- max(lims)
      bw_val <- if (n_req <= 8) 1L else as.integer(ceiling(n_req / 8L))
      seq(0L, ceiling(lims[2]), by = bw_val)
    }) +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
    labs(x = "Refine steps processed (per parent node)", y = "Proportion",
         title = "Refine steps processed per parent (red dashed = n_requested)") +
    theme_bw()
  ggsave(file.path(fig_root, "ensemble_scaling", "refine_depth_histogram.png"),
         p_hist, width = 14, height = 6)

} else {
  message("refine_depth.csv not found or empty; skipping fig-3c.")
}
```

# 4. Ensemble Scaling Plots (Category 4)

Figures 4a–4d show AUPRC and TPR@0.05 as a function of ensemble size.
Scaling confusion bins are computed during the SLURM run and pooled in the
collect workbook; loaded here from `scaling_bins_pooled.csv`.

```{r load-scaling}
scaling_bins_path <- file.path(consolidated_dir, "scaling_bins_pooled.csv")

if (file.exists(scaling_bins_path)) {
  scaling_bins_pooled <- readr::read_csv(scaling_bins_path, show_col_types = FALSE)
  cat("Scaling bins pooled rows:", nrow(scaling_bins_pooled), "\n")

  # For specs with non-NA annotation_r2, create overall rows by pooling bins
  # across annotation_r2 levels (so the "overall" plot includes all specs).
  has_r2 <- scaling_bins_pooled %>% dplyr::filter(!is.na(.data$annotation_r2))
  if (nrow(has_r2) > 0L) {
    overall_pooled <- has_r2 %>%
      dplyr::group_by(spec_name, n_ensemble, resolution, agg_method, rep,
                       pip_threshold) %>%
      dplyr::summarise(
        n_causal_at_bucket    = sum(.data$n_causal_at_bucket,    na.rm = TRUE),
        n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::mutate(annotation_r2 = NA_real_)
    scaling_bins_pooled <- dplyr::bind_rows(scaling_bins_pooled, overall_pooled)
  }

  # Compute per-rep AUPRC and TPR@0.05 from bins pooled across datasets.
  # Pure-lever specs: group by n_ensemble (the subsample size).
  # Interaction specs: group by resolution (the grid label); n_ensemble may
  #   vary across datasets due to BFS early stopping, so we pool bins at the
  #   resolution level and let the pooled counts weight datasets naturally.

  scaling_pure <- scaling_bins_pooled %>% dplyr::filter(is.na(.data$resolution))
  scaling_inter <- scaling_bins_pooled %>% dplyr::filter(!is.na(.data$resolution))

  metrics_pure <- if (nrow(scaling_pure) > 0L) {
    scaling_pure %>%
      dplyr::group_by(spec_name, annotation_r2, n_ensemble, agg_method, rep) %>%
      dplyr::summarise(
        AUPRC = auprc_from_bins(dplyr::pick(dplyr::everything())),
        tpr05 = tpr05_from_bins(dplyr::pick(dplyr::everything())),
        .groups = "drop"
      ) %>%
      dplyr::mutate(resolution = NA_character_)
  } else { tibble::tibble() }

  metrics_inter <- if (nrow(scaling_inter) > 0L) {
    scaling_inter %>%
      dplyr::group_by(spec_name, annotation_r2, resolution, agg_method, rep,
                       pip_threshold) %>%
      dplyr::summarise(
        n_causal_at_bucket    = sum(.data$n_causal_at_bucket,    na.rm = TRUE),
        n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(spec_name, annotation_r2, resolution, agg_method, rep) %>%
      dplyr::summarise(
        AUPRC = auprc_from_bins(dplyr::pick(dplyr::everything())),
        tpr05 = tpr05_from_bins(dplyr::pick(dplyr::everything())),
        .groups = "drop"
      ) %>%
      dplyr::mutate(n_ensemble = NA_integer_)
  } else { tibble::tibble() }

  scaling_metrics_by_rep <- dplyr::bind_rows(metrics_pure, metrics_inter)

  # Summarise over reps: mean +/- SE.  SE is NA when only 1 rep (deterministic specs).
  scaling_summary <- scaling_metrics_by_rep %>%
    dplyr::group_by(spec_name, annotation_r2, n_ensemble, resolution, agg_method) %>%
    dplyr::summarise(
      AUPRC_mean = mean(AUPRC, na.rm = TRUE),
      AUPRC_se   = ifelse(dplyr::n() > 1L,
                          sd(AUPRC, na.rm = TRUE) / sqrt(sum(!is.na(AUPRC))),
                          NA_real_),
      tpr05_mean = mean(tpr05, na.rm = TRUE),
      tpr05_se   = ifelse(dplyr::n() > 1L,
                          sd(tpr05, na.rm = TRUE) / sqrt(sum(!is.na(tpr05))),
                          NA_real_),
      n_reps     = dplyr::n(),
      .groups = "drop"
    )

  # Anchor full-size scaling endpoints to the exact pooled full-ensemble
  # summaries used in the main barplots. This guarantees that the maximal
  # ensemble-size / full-resolution endpoint reconciles exactly.
  parse_resolution_area <- function(x) {
    parts <- strsplit(as.character(x), "x", fixed = TRUE)
    vapply(parts, function(p) {
      nums <- suppressWarnings(as.numeric(p))
      if (length(nums) < 1L || any(!is.finite(nums))) return(NA_real_)
      prod(nums)
    }, numeric(1L))
  }

  anchor_scaling_full_endpoints <- function(summary_df, full_auprc_df, full_tpr_df) {
    if (!nrow(summary_df)) return(summary_df)

    full_targets <- full_auprc_df %>%
      dplyr::rename(AUPRC_full = AUPRC) %>%
      dplyr::left_join(
        full_tpr_df %>% dplyr::rename(tpr05_full = tpr_fpr05),
        by = c("spec_name", "agg_method", "annotation_r2")
      )

    anchored <- summary_df %>%
      dplyr::left_join(full_targets,
                       by = c("spec_name", "agg_method", "annotation_r2")) %>%
      dplyr::group_by(.data$spec_name, .data$agg_method, .data$annotation_r2) %>%
      dplyr::mutate(
        .full_idx = dplyr::case_when(
          all(is.na(.data$resolution)) ~ .data$n_ensemble == max(.data$n_ensemble, na.rm = TRUE),
          all(!is.na(.data$resolution)) ~ parse_resolution_area(.data$resolution) ==
            max(parse_resolution_area(.data$resolution), na.rm = TRUE),
          TRUE ~ FALSE
        ),
        AUPRC_mean = dplyr::if_else(.data$.full_idx & is.finite(.data$AUPRC_full),
                                    .data$AUPRC_full, .data$AUPRC_mean),
        tpr05_mean = dplyr::if_else(.data$.full_idx & is.finite(.data$tpr05_full),
                                    .data$tpr05_full, .data$tpr05_mean),
        AUPRC_se = dplyr::if_else(.data$.full_idx & is.finite(.data$AUPRC_full),
                                  NA_real_, .data$AUPRC_se),
        tpr05_se = dplyr::if_else(.data$.full_idx & is.finite(.data$tpr05_full),
                                  NA_real_, .data$tpr05_se),
        n_reps = dplyr::if_else(.data$.full_idx & is.finite(.data$AUPRC_full),
                                1L, .data$n_reps)
      ) %>%
      dplyr::ungroup() %>%
      dplyr::select(-.data$AUPRC_full, -.data$tpr05_full, -.data$.full_idx)

    anchored
  }

  scaling_summary <- anchor_scaling_full_endpoints(
    scaling_summary,
    full_auprc_df = dplyr::bind_rows(auprc_plot_overall, auprc_plot_by_r2),
    full_tpr_df = dplyr::bind_rows(tpr05_plot_overall, tpr05_plot_by_r2)
  )

  # Separate pure-lever and interaction subsets.
  scaling_curves      <- scaling_summary %>% dplyr::filter(is.na(.data$resolution))
  interaction_scaling <- scaling_summary %>% dplyr::filter(!is.na(.data$resolution))

  cat("Scaling curves rows:", nrow(scaling_curves), "\n")
  cat("Interaction scaling rows:", nrow(interaction_scaling), "\n")
} else {
  scaling_bins_pooled <- NULL
  scaling_curves      <- NULL
  interaction_scaling <- NULL
  message("scaling_bins_pooled.csv not found — Category 4 plots will be skipped.")
}
```

```{r fig-4a}
# AUPRC scaling curves: ensemble size vs pooled AUPRC, faceted by spec.
# Ribbon shows ±1 SE for restart specs (50 reps); no ribbon for deterministic specs.
if (!is.null(scaling_curves) && nrow(scaling_curves) > 0L) {
  make_scaling_plot <- function(data, y_col, se_col, y_label, title,
                                baseline_val = NULL) {
    p <- ggplot2::ggplot(
      data,
      ggplot2::aes(
        x     = n_ensemble,
        y     = .data[[y_col]],
        color = agg_method,
        group = agg_method
      )
    ) +
      ggplot2::geom_line() +
      ggplot2::geom_point(size = 1.8)

    # Add SE ribbon only where se is non-NA
    ribbon_data <- data %>% dplyr::filter(!is.na(.data[[se_col]]))
    if (nrow(ribbon_data) > 0L) {
      p <- p + ggplot2::geom_ribbon(
        data = ribbon_data,
        ggplot2::aes(
          ymin = .data[[y_col]] - .data[[se_col]],
          ymax = .data[[y_col]] + .data[[se_col]],
          fill = agg_method
        ),
        alpha = 0.15, color = NA
      )
    }

    # Baseline single-fit reference line
    if (!is.null(baseline_val) && is.finite(baseline_val)) {
      p <- p +
        ggplot2::geom_hline(yintercept = baseline_val, linetype = "dashed",
                            color = "red", linewidth = 0.6) +
        ggplot2::annotate("text", x = min(data$n_ensemble, na.rm = TRUE),
                          y = baseline_val, label = "baseline single-fit",
                          vjust = -0.5, hjust = 0, color = "red", size = 2.8)
    }

    p +
      ggplot2::scale_x_continuous(breaks = c(4, 8, 16, 32, 64)) +
      ggplot2::facet_wrap(~ spec_name, nrow = 2) +
      ggplot2::labs(
        x     = "Ensemble size (# fits)",
        y     = y_label,
        color = "Aggregation",
        fill  = "Aggregation",
        title = title
      ) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = "bottom")
  }

  # Overall (pooled across all annotation_r2)
  p4a_overall <- make_scaling_plot(
    scaling_curves %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col    = "AUPRC_mean",
    se_col   = "AUPRC_se",
    y_label  = "AUPRC (pooled across datasets)",
    title    = "AUPRC scaling \u2014 all annotation R\u00b2",
    baseline_val = baseline_auprc
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "auprc_scaling_curves_overall.png"),
    p4a_overall, width = 12, height = 6, dpi = 150
  )


  # Per annotation_r2
  for (r2 in sort(unique(na.omit(scaling_curves$annotation_r2)))) {
    p4a_r2 <- make_scaling_plot(
      scaling_curves %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col    = "AUPRC_mean",
      se_col   = "AUPRC_se",
      y_label  = "AUPRC (pooled across datasets)",
      title    = sprintf("AUPRC scaling \u2014 annotation R\u00b2 = %.1f", r2),
      baseline_val = baseline_auprc
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("auprc_scaling_curves_r2_%.1f.png", r2)),
      p4a_r2, width = 12, height = 6, dpi = 150
    )
  }
  cat("Fig 4a saved.\n")
} else {
  cat("scaling_bins_pooled.csv not found — fig 4a skipped.\n")
}
```

```{r fig-4b}
# TPR@FPR=0.05 scaling curves: same structure as 4a.
if (!is.null(scaling_curves) && nrow(scaling_curves) > 0L) {
  p4b_overall <- make_scaling_plot(
    scaling_curves %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col    = "tpr05_mean",
    se_col   = "tpr05_se",
    y_label  = "TPR at FPR \u2264 0.05 (pooled)",
    title    = "TPR@FPR=0.05 scaling \u2014 all annotation R\u00b2",
    baseline_val = baseline_tpr05
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "tpr05_scaling_curves_overall.png"),
    p4b_overall, width = 12, height = 6, dpi = 150
  )


  for (r2 in sort(unique(na.omit(scaling_curves$annotation_r2)))) {
    p4b_r2 <- make_scaling_plot(
      scaling_curves %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col    = "tpr05_mean",
      se_col   = "tpr05_se",
      y_label  = "TPR at FPR \u2264 0.05 (pooled)",
      title    = sprintf("TPR@FPR=0.05 scaling \u2014 annotation R\u00b2 = %.1f", r2),
      baseline_val = baseline_tpr05
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("tpr05_scaling_curves_r2_%.1f.png", r2)),
      p4b_r2, width = 12, height = 6, dpi = 150
    )
  }
  cat("Fig 4b saved.\n")
} else {
  cat("scaling_curves.csv not found — fig 4b skipped.\n")
}
```

```{r fig-4c}
# 2-lever interaction AUPRC: compare 4x4 vs 8x8 grid resolutions.
if (!is.null(interaction_scaling) && nrow(interaction_scaling) > 0L) {
  make_interaction_plot <- function(data, y_col, y_label, title,
                                    baseline_val = NULL) {
    p <- ggplot2::ggplot(
      data,
      ggplot2::aes(
        x     = resolution,
        y     = .data[[y_col]],
        color = agg_method,
        group = agg_method
      )
    ) +
      ggplot2::geom_line() +
      ggplot2::geom_point(size = 2.5)

    if (!is.null(baseline_val) && is.finite(baseline_val)) {
      p <- p +
        ggplot2::geom_hline(yintercept = baseline_val, linetype = "dashed",
                            color = "red", linewidth = 0.6) +
        ggplot2::annotate("text", x = 1, y = baseline_val,
                          label = "baseline single-fit",
                          vjust = -0.5, hjust = 0, color = "red", size = 2.8)
    }

    p +
      ggplot2::facet_wrap(~ spec_name, nrow = 2) +
      ggplot2::labs(
        x     = "Grid resolution",
        y     = y_label,
        color = "Aggregation",
        title = title
      ) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = "bottom",
                     axis.text.x = element_text(angle = 45, hjust = 1))
  }

  p4c_overall <- make_interaction_plot(
    interaction_scaling %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col   = "AUPRC_mean",
    y_label = "AUPRC (pooled across datasets)",
    title   = "Interaction grid scaling \u2014 AUPRC",
    baseline_val = baseline_auprc
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "auprc_interaction_scaling_overall.png"),
    p4c_overall, width = 10, height = 4, dpi = 150
  )


  for (r2 in sort(unique(na.omit(interaction_scaling$annotation_r2)))) {
    p4c_r2 <- make_interaction_plot(
      interaction_scaling %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col   = "AUPRC_mean",
      y_label = "AUPRC (pooled across datasets)",
      title   = sprintf("Interaction grid scaling \u2014 AUPRC (R\u00b2=%.1f)", r2),
      baseline_val = baseline_auprc
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("auprc_interaction_scaling_r2_%.1f.png", r2)),
      p4c_r2, width = 10, height = 4, dpi = 150
    )
  }
  cat("Fig 4c saved.\n")
} else {
  cat("scaling_bins_pooled.csv not found — fig 4c skipped.\n")
}
```

```{r fig-4d}
# 2-lever interaction TPR@0.05: same structure as 4c.
if (!is.null(interaction_scaling) && nrow(interaction_scaling) > 0L) {
  p4d_overall <- make_interaction_plot(
    interaction_scaling %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col   = "tpr05_mean",
    y_label = "TPR at FPR \u2264 0.05 (pooled)",
    title   = "Interaction grid scaling \u2014 TPR@FPR=0.05",
    baseline_val = baseline_tpr05
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "tpr05_interaction_scaling_overall.png"),
    p4d_overall, width = 10, height = 4, dpi = 150
  )


  for (r2 in sort(unique(na.omit(interaction_scaling$annotation_r2)))) {
    p4d_r2 <- make_interaction_plot(
      interaction_scaling %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col   = "tpr05_mean",
      y_label = "TPR at FPR \u2264 0.05 (pooled)",
      title   = sprintf(
        "Interaction grid scaling \u2014 TPR@FPR=0.05 (R\u00b2=%.1f)", r2),
      baseline_val = baseline_tpr05
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("tpr05_interaction_scaling_r2_%.1f.png", r2)),
      p4d_r2, width = 10, height = 4, dpi = 150
    )
  }
  cat("Fig 4d saved.\n")
} else {
  cat("scaling_bins_pooled.csv not found — fig 4d skipped.\n")
}
```

# 5. Compute vs AUPRC

Total ensemble compute time (sum of wall_time_sec across all runs in
the ensemble, averaged over datasets) on the x-axis, pooled AUPRC on
the y-axis. Color = ensemble method (spec_name), shape = aggregation
method. Baseline single-fit shown as a black star.

```{r compute-vs-auprc, fig.width=12, fig.height=7}
{
  # Build per-ensemble total compute time: sum wall_time across all

  # individual (non-agg) runs within each (spec_name, dataset_bundle_id),
  # then average over datasets.
  wt_available <- "wall_time_sec" %in% names(model_metrics)

  if (wt_available) {
    # Individual runs only (exclude aggregation pseudo-rows)
    wt_per_ensemble <- model_metrics %>%
      dplyr::filter(!is.na(.data$spec_name),
                    .data$filtering == "purity_filtered",
                    is.na(.data$agg_method) |
                      .data$explore_method != "aggregation") %>%
      dplyr::group_by(.data$spec_name, .data$dataset_bundle_id) %>%
      dplyr::summarise(
        total_wall_sec = sum(.data$wall_time_sec, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::group_by(.data$spec_name) %>%
      dplyr::summarise(
        mean_total_wall_sec = mean(.data$total_wall_sec, na.rm = TRUE),
        .groups = "drop"
      )

    # Join with pooled AUPRC (one row per spec x agg_method)
    compute_auprc <- auprc_plot_overall %>%
      dplyr::filter(.data$spec_name != "baseline-single") %>%
      dplyr::inner_join(wt_per_ensemble, by = "spec_name") %>%
      dplyr::mutate(
        compute_min = .data$mean_total_wall_sec / 60
      )

    # Baseline single-fit: average wall_time across datasets (1 run each)
    baseline_wt <- model_metrics %>%
      dplyr::filter(.data$spec_name == "baseline-single",
                    .data$filtering == "purity_filtered",
                    is.na(.data$agg_method) |
                      .data$explore_method != "aggregation",
                    .data$use_case_id == "susine_vanilla") %>%
      dplyr::group_by(.data$dataset_bundle_id) %>%
      dplyr::summarise(
        total_wall_sec = sum(.data$wall_time_sec, na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::summarise(
        mean_total_wall_sec = mean(.data$total_wall_sec, na.rm = TRUE)
      ) %>%
      dplyr::pull(mean_total_wall_sec)
    baseline_compute_min <- baseline_wt / 60

    if (nrow(compute_auprc) > 0L) {
      # Aggregation method shapes
      agg_shapes <- c(
        "max_elbo"       = 15,  # filled square
        "uniform"        = 16,  # filled circle
        "elbo_softmax"   = 17,  # filled triangle
        "cluster_weight" = 18   # filled diamond
      )

      p_compute <- ggplot(
        compute_auprc,
        aes(x = compute_min, y = AUPRC,
            color = spec_name, shape = agg_method)
      ) +
        geom_point(size = 3, alpha = 0.85) +
        scale_shape_manual(
          values = agg_shapes,
          name = "Aggregation"
        ) +
        # Baseline star
        annotate(
          "point",
          x = baseline_compute_min,
          y = baseline_auprc,
          shape = 8, size = 4, color = "black", stroke = 1.2
        ) +
        annotate(
          "text",
          x = baseline_compute_min,
          y = baseline_auprc,
          label = "baseline single-fit",
          hjust = -0.15, vjust = -0.8, size = 3, color = "black"
        ) +
        labs(
          x = "Total compute time (minutes, mean over datasets)",
          y = "Pooled AUPRC",
          color = "Ensemble Method",
          title = "Compute vs. AUPRC (all annotation R\u00b2 pooled)"
        ) +
        theme_bw() +
        theme(legend.position = "bottom",
              legend.box = "vertical")
      ggsave(file.path(fig_root, "ensemble_scaling",
                       "compute_vs_auprc.png"),
             p_compute, width = 12, height = 7, dpi = 150)
      cat("Fig compute_vs_auprc saved.\n")

      # Per annotation_r2 variant
      for (r2 in sort(unique(na.omit(
        auprc_plot_by_r2$annotation_r2)))) {
        compute_auprc_r2 <- auprc_plot_by_r2 %>%
          dplyr::filter(.data$spec_name != "baseline-single",
                        is.na(.data$annotation_r2) |
                          .data$annotation_r2 == r2) %>%
          dplyr::inner_join(wt_per_ensemble, by = "spec_name") %>%
          dplyr::mutate(compute_min = .data$mean_total_wall_sec / 60)

        if (nrow(compute_auprc_r2) > 0L) {
          # Baseline AUPRC at this r2 level
          bl_auprc_r2 <- auprc_plot_by_r2 %>%
            dplyr::filter(.data$spec_name == "baseline-single",
                          is.na(.data$annotation_r2) |
                            .data$annotation_r2 == r2) %>%
            dplyr::pull(AUPRC)
          bl_auprc_r2 <- if (length(bl_auprc_r2)) bl_auprc_r2[1] else baseline_auprc

          p_r2 <- ggplot(
            compute_auprc_r2,
            aes(x = compute_min, y = AUPRC,
                color = spec_name, shape = agg_method)
          ) +
            geom_point(size = 3, alpha = 0.85) +
            scale_shape_manual(values = agg_shapes, name = "Aggregation") +
            annotate("point", x = baseline_compute_min, y = bl_auprc_r2,
                     shape = 8, size = 4, color = "black", stroke = 1.2) +
            annotate("text", x = baseline_compute_min, y = bl_auprc_r2,
                     label = "baseline single-fit",
                     hjust = -0.15, vjust = -0.8, size = 3, color = "black") +
            labs(
              x = "Total compute time (minutes, mean over datasets)",
              y = "Pooled AUPRC",
              color = "Ensemble Method",
              title = sprintf(
                "Compute vs. AUPRC (annotation R\u00b2 = %.1f)", r2)
            ) +
            theme_bw() +
            theme(legend.position = "bottom", legend.box = "vertical")
          ggsave(file.path(fig_root, "ensemble_scaling",
                           sprintf("compute_vs_auprc_r2_%.1f.png", r2)),
                 p_r2, width = 12, height = 7, dpi = 150)
        }
      }
    } else {
      cat("compute_vs_auprc skipped: no data after joining wall_time.\n")
    }
  } else {
    cat("compute_vs_auprc skipped: wall_time_sec not in model_metrics.\n")
  }
}
```

# 6. Bootstrap AUPRC Stability

Bootstrap analysis to demonstrate that pooled AUPRC estimates are
stable at the available sample size. For each subsample size $n$, we
draw $B$ bootstrap resamples (with replacement) from the full set of
datasets, pool confusion bins within each resample, compute AUPRC,
and report the bootstrap mean $\pm$ 1.96 SE.

```{r bootstrap-auprc-stability, fig.width=12, fig.height=10}
{
  # ── Helper: pooled AUPRC from confusion bins ──
  # Computes average precision directly from binned counts (descending

  # pip_threshold order). AP = (1/P) * sum over positive-containing bins
  # of [ cum_tp_at_bin * n_causal_in_bin / (cum_tp + cum_fp)_at_bin ].
  # This is the bin-level analogue of the per-SNP AP formula.
  pooled_auprc_counts <- function(causal_counts, noncausal_counts) {
    cum_tp  <- cumsum(causal_counts)
    cum_fp  <- cumsum(noncausal_counts)
    total_p <- sum(causal_counts)
    total_n <- sum(noncausal_counts)
    if (total_p == 0L || total_n == 0L) return(NA_real_)
    # For each bin with causal SNPs, the average precision contribution is:
    # precision_at_bin * (n_causal_in_bin / total_positives)
    # This correctly weights by the number of positives discovered at each
    # threshold, matching the per-SNP AP formula.
    precision <- cum_tp / (cum_tp + cum_fp)
    delta_recall <- causal_counts / total_p
    sum(precision * delta_recall)
  }

  # ── Bootstrap stability function ──
  # bins:        confusion_bins tibble (must have dataset_bundle_id,

  #              pip_threshold, n_causal_at_bucket, n_noncausal_at_bucket)
  # dataset_ids: character vector of unique dataset IDs to resample from
  # sizes:       integer vector of subsample sizes
  # B:           number of bootstrap replicates per size
  bootstrap_auprc_stability <- function(bins, dataset_ids, sizes, B = 300L,
                                        progress_label = "bootstrap") {
    cat(sprintf("[%s] preparing dataset x threshold matrices\n", progress_label))
    bins_compact <- bins %>%
      dplyr::select(
        .data$dataset_bundle_id,
        .data$pip_threshold,
        .data$n_causal_at_bucket,
        .data$n_noncausal_at_bucket
      ) %>%
      dplyr::group_by(.data$dataset_bundle_id, .data$pip_threshold) %>%
      dplyr::summarise(
        n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
        n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
        .groups = "drop"
      )

    pip_levels <- sort(unique(bins_compact$pip_threshold), decreasing = TRUE)
    bins_compact <- bins_compact %>%
      dplyr::mutate(
        dataset_bundle_id = factor(.data$dataset_bundle_id, levels = dataset_ids),
        pip_threshold = factor(.data$pip_threshold, levels = pip_levels)
      )

    causal_mat <- stats::xtabs(
      n_causal_at_bucket ~ dataset_bundle_id + pip_threshold,
      data = bins_compact
    )
    noncausal_mat <- stats::xtabs(
      n_noncausal_at_bucket ~ dataset_bundle_id + pip_threshold,
      data = bins_compact
    )

    n_dataset_ids <- length(dataset_ids)
    cat(sprintf("[%s] matrix prep complete: %d datasets x %d thresholds\n",
                progress_label, nrow(causal_mat), ncol(causal_mat)))

    results <- vector("list", length(sizes))
    for (si in seq_along(sizes)) {
      n <- sizes[si]
      cat(sprintf("[%s] size %d/%d (n=%d, B=%d)\n",
                  progress_label, si, length(sizes), n, B))

      weight_mat <- replicate(
        B,
        tabulate(sample.int(n_dataset_ids, size = n, replace = TRUE),
                 nbins = n_dataset_ids)
      )
      pooled_causal <- crossprod(causal_mat, weight_mat)
      pooled_noncausal <- crossprod(noncausal_mat, weight_mat)
      auprcs <- vapply(
        seq_len(B),
        function(b) {
          pooled_auprc_counts(
            causal_counts = pooled_causal[, b],
            noncausal_counts = pooled_noncausal[, b]
          )
        },
        numeric(1)
      )

      results[[si]] <- tibble::tibble(
        n_datasets = n,
        auprc_mean = mean(auprcs, na.rm = TRUE),
        auprc_se   = sd(auprcs, na.rm = TRUE)
      )
    }
    cat(sprintf("[%s] complete\n", progress_label))
    dplyr::bind_rows(results)
  }

  # ── Run bootstrap for baseline single-fit (susine_vanilla only) ──
  baseline_bins <- confusion_bins %>%
    dplyr::filter(.data$spec_name == "baseline-single",
                  .data$use_case_id == "susine_vanilla",
                  is.na(.data$agg_method))
  baseline_datasets <- sort(unique(baseline_bins$dataset_bundle_id))
  n_total <- length(baseline_datasets)
  cat("Bootstrap stability: total datasets =", n_total, "\n")

  if (n_total >= 20L) {
    # Subsample sizes: steps of ~10% of total, from ~10% up to 100%
    step <- max(10L, as.integer(round(n_total / 10)))
    sizes <- sort(unique(c(
      seq(step, n_total, by = step),
      n_total
    )))
    # Ensure we don't go below a reasonable minimum
    sizes <- sizes[sizes >= 10L]

    B <- 300L
    set.seed(42)
    cat("Running bootstrap: sizes =", paste(sizes, collapse = ", "),
        ", B =", B, "\n")
    cat("[baseline bootstrap] starting\n")

    boot_baseline <- bootstrap_auprc_stability(
      bins        = baseline_bins,
      dataset_ids = baseline_datasets,
      sizes       = sizes,
      B           = B,
      progress_label = "baseline bootstrap"
    ) %>%
      dplyr::mutate(
        ci_lo = .data$auprc_mean - 1.96 * .data$auprc_se,
        ci_hi = .data$auprc_mean + 1.96 * .data$auprc_se,
        spec_label = "baseline single-fit (vanilla)"
      )

    # ── Combined baseline figure (faceted: AUPRC + SE panels) ──
    boot_long <- boot_baseline %>%
      tidyr::pivot_longer(
        cols = c(auprc_mean, auprc_se),
        names_to = "metric",
        values_to = "value"
      ) %>%
      dplyr::mutate(
        ci_lo = dplyr::if_else(.data$metric == "auprc_mean", .data$ci_lo, NA_real_),
        ci_hi = dplyr::if_else(.data$metric == "auprc_mean", .data$ci_hi, NA_real_),
        metric_label = dplyr::if_else(
          .data$metric == "auprc_mean",
          "Pooled AUPRC (mean \u00b1 1.96 SE)",
          "Bootstrap SE"
        )
      )

    p_boot <- ggplot(boot_long, aes(x = n_datasets, y = value)) +
      geom_ribbon(data = dplyr::filter(boot_long, .data$metric == "auprc_mean"),
                  aes(ymin = ci_lo, ymax = ci_hi),
                  fill = "steelblue", alpha = 0.25) +
      geom_line(color = "steelblue", linewidth = 0.8) +
      geom_point(color = "steelblue", size = 2) +
      facet_wrap(~metric_label, scales = "free_y", ncol = 1) +
      labs(x = "Number of datasets in bootstrap resample", y = NULL,
           title = "AUPRC estimation stability (baseline single-fit)") +
      theme_bw()

    ggsave(file.path(fig_root, "ensemble_scaling",
                     "bootstrap_auprc_stability.png"),
           p_boot, width = 10, height = 8, dpi = 150)

    # ── Optional: run for a few ensemble specs too ──
    # Pick the top-performing spec per agg_method to show stability
    # is not unique to the baseline.
    agg_bins <- confusion_bins %>%
      dplyr::filter(!is.na(.data$agg_method),
                    !is.na(.data$spec_name),
                    .data$spec_name != "baseline-single")
    agg_combos <- agg_bins %>%
      dplyr::distinct(.data$spec_name, .data$agg_method)

    # Select up to 3 representative (spec, agg) combos via highest AUPRC
    if (nrow(agg_combos) > 0L && exists("auprc_plot_overall")) {
      top_combos <- auprc_plot_overall %>%
        dplyr::filter(.data$spec_name != "baseline-single") %>%
        dplyr::arrange(dplyr::desc(.data$AUPRC)) %>%
        dplyr::slice_head(n = 3L)

      boot_ensemble_list <- list()
      for (ri in seq_len(nrow(top_combos))) {
        sn <- top_combos$spec_name[ri]
        am <- top_combos$agg_method[ri]
        combo_label <- sprintf("ensemble bootstrap %d/%d: %s (%s)",
                               ri, nrow(top_combos), sn, am)
        combo_bins <- confusion_bins %>%
          dplyr::filter(.data$spec_name == sn,
                        .data$agg_method == am)
        combo_datasets <- sort(unique(combo_bins$dataset_bundle_id))
        if (length(combo_datasets) < 20L) {
          cat(sprintf("[%s] skipped: only %d datasets available\n",
                      combo_label, length(combo_datasets)))
          next
        }

        cat(sprintf("[%s] starting\n", combo_label))

        boot_ensemble_list[[ri]] <- bootstrap_auprc_stability(
          bins        = combo_bins,
          dataset_ids = combo_datasets,
          sizes       = sizes[sizes <= length(combo_datasets)],
          B           = B,
          progress_label = combo_label
        ) %>%
          dplyr::mutate(
            ci_lo = .data$auprc_mean - 1.96 * .data$auprc_se,
            ci_hi = .data$auprc_mean + 1.96 * .data$auprc_se,
            spec_label = paste0(sn, " (", am, ")")
          )
      }

      if (length(boot_ensemble_list) > 0L) {
        boot_all <- dplyr::bind_rows(
          boot_baseline,
          dplyr::bind_rows(boot_ensemble_list)
        )

        boot_all_long <- boot_all %>%
          tidyr::pivot_longer(
            cols = c(auprc_mean, auprc_se),
            names_to = "metric",
            values_to = "value"
          ) %>%
          dplyr::mutate(
            ci_lo = dplyr::if_else(.data$metric == "auprc_mean", .data$ci_lo, NA_real_),
            ci_hi = dplyr::if_else(.data$metric == "auprc_mean", .data$ci_hi, NA_real_),
            metric_label = dplyr::if_else(
              .data$metric == "auprc_mean",
              "Pooled AUPRC (mean \u00b1 1.96 SE)",
              "Bootstrap SE"
            )
          )

        p_boot_multi <- ggplot(boot_all_long,
                               aes(x = n_datasets, y = value,
                                   color = spec_label, fill = spec_label)) +
          geom_ribbon(
            data = dplyr::filter(boot_all_long, .data$metric == "auprc_mean"),
            aes(ymin = ci_lo, ymax = ci_hi),
            alpha = 0.15, color = NA) +
          geom_line(linewidth = 0.7) +
          geom_point(size = 1.8) +
          facet_wrap(~metric_label, scales = "free_y", ncol = 1) +
          labs(x = "Number of datasets in bootstrap resample", y = NULL,
               color = "Method", fill = "Method",
               title = "AUPRC estimation stability (selected methods)") +
          theme_bw() +
          theme(legend.position = "bottom")

        ggsave(file.path(fig_root, "ensemble_scaling",
                         "bootstrap_auprc_stability_multi.png"),
               p_boot_multi, width = 12, height = 10, dpi = 150)
      }
    }

    # Print summary
    cat("\nBootstrap stability summary (baseline):\n")
    print(knitr::kable(
      boot_baseline %>%
        dplyr::mutate(
          auprc_mean = round(auprc_mean, 4),
          auprc_se   = round(auprc_se, 5),
          ci_lo      = round(ci_lo, 4),
          ci_hi      = round(ci_hi, 4)
        ) %>%
        dplyr::select(n_datasets, auprc_mean, auprc_se, ci_lo, ci_hi),
      col.names = c("N datasets", "Mean AUPRC", "SE", "95% CI lo", "95% CI hi")
    ))
    cat("Fig bootstrap_auprc_stability saved.\n")
  } else {
    cat("Bootstrap stability skipped: fewer than 20 datasets.\n")
  }
}
```

# Summary Table

```{r summary-table}
if (exists("auprc_plot_overall") && nrow(auprc_plot_overall) > 0L) {
  summary_tbl <- auprc_plot_overall %>%
    dplyr::select(spec_name, agg_method, AUPRC) %>%
    dplyr::left_join(
      tpr05_plot_overall %>%
        dplyr::select(spec_name, agg_method, tpr_fpr05),
      by = c("spec_name", "agg_method")
    ) %>%
    dplyr::arrange(.data$spec_name, .data$agg_method) %>%
    dplyr::mutate(
      AUPRC     = round(.data$AUPRC,     4L),
      tpr_fpr05 = round(.data$tpr_fpr05, 4L)
    )

  print(knitr::kable(
    summary_tbl,
    col.names = c("Spec", "Aggregation", "Pooled AUPRC", "TPR @ FPR=0.05"),
    caption   = "Pooled AUPRC and TPR@FPR=0.05 by ensemble spec and aggregation method"
  ))
  readr::write_csv(summary_tbl,
                   file.path(consolidated_dir, "summary_table.csv"))
  cat("Summary table written.\n")
} else {
  cat("Summary table skipped: auprc_plot_overall not available.\n")
}
```

# Session Info

```{r session}
cat("\nFigures written to:", fig_root, "\n")
cat("Files:\n")
for (d in fig_dirs) {
  figs <- list.files(file.path(fig_root, d))
  if (length(figs)) cat(sprintf("  %s/: %s\n", d, paste(figs, collapse = ", ")))
}
sessionInfo()
```
