---
title: "Baseline Sims Screening - Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Visualization workbook for the baseline screening job. This keeps the
`susie2_plots` and `dataset_difficulty` sections, replaces ensemble
performance with baseline screening views, removes scaling analysis, and
adds `effect_diagnostics`.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_baseline_sims.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(knitr)

paper_theme <- theme_bw(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey92"),
    strip.background = element_rect(fill = "grey95", color = NA),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.key = element_rect(fill = NA, color = NA),
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 9),
    plot.title = element_text(size = 11, face = "bold")
  )
theme_set(paper_theme)

method_pretty <- c(
  susie_vanilla                       = "SuSiE",
  susine_vanilla                      = "SuSiNE (mu=0)",
  susie_inf                           = "SuSiE-inf",
  susie_ash_fixed                     = "SuSiE-ash",
  susie_eb                            = "SuSiE (EB sigma^2)",
  susie_functional_pi                 = "SuSiE (functional pi)",
  susine_functional_pi                = "SuSiNE (functional pi)",
  susine_functional_mu                = "SuSiNE (functional mu)",
  susine_eb_clamped_scale_var_nonneg  = "SuSiNE-EB"
)
```

# 1. Load Consolidated Data

```{r load}
job_name <- "baseline_sims_screen"
parent_job_id <- "51509956"
output_root <- here("output")

job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
consolidated_dir <- file.path(job_dir, "consolidated")
fig_root <- file.path(job_dir, "figures")

fig_dirs <- c(
  "susie2_plots",
  "baseline_performance",
  file.path("baseline_performance", "confirmation"),
  "dataset_difficulty",
  "effect_diagnostics"
)
for (d in fig_dirs) {
  dir.create(file.path(fig_root, d), recursive = TRUE, showWarnings = FALSE)
}

read_req <- function(filename) {
  readr::read_csv(file.path(consolidated_dir, filename), show_col_types = FALSE)
}

read_opt <- function(filename) {
  path <- file.path(consolidated_dir, filename)
  if (!file.exists(path)) {
    return(NULL)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

model_metrics <- read_req("model_metrics_full.csv")
effect_metrics <- read_opt("effect_metrics_full.csv")
effect_metrics_unfiltered <- read_opt("effect_metrics_unfiltered_full.csv")
confusion_bins <- read_req("confusion_bins_full.csv")
dataset_metrics <- read_opt("dataset_metrics_full.csv")
tier_cs_metrics <- read_opt("tier_cs_metrics_full.csv")
auprc_per_dataset <- read_req("auprc_per_dataset.csv")
auprc_pooled_overall <- read_req("auprc_pooled_overall.csv")
auprc_pooled_by_annotation <- read_req("auprc_pooled_by_annotation.csv")
tpr05_per_dataset <- read_req("tpr05_per_dataset.csv")
tpr05_pooled_overall <- read_req("tpr05_pooled_overall.csv")
tpr05_pooled_by_annotation <- read_req("tpr05_pooled_by_annotation.csv")
screening_summary <- read_req("screening_summary.csv")

cat("Figure root:", fig_root, "\n")
cat("Pooled AUPRC settings:", nrow(auprc_pooled_overall), "\n")
cat("Unfiltered effect rows:", if (is.null(effect_metrics_unfiltered)) 0 else nrow(effect_metrics_unfiltered), "\n")
```

```{r helpers}
fmt_num <- function(x, digits = 3) {
  format(signif(as.numeric(x), digits), trim = TRUE, scientific = FALSE)
}

top_settings_by_family <- screening_summary %>%
  dplyr::group_by(.data$method_family) %>%
  dplyr::slice_max(.data$AUPRC, n = 3L, with_ties = FALSE) %>%
  dplyr::ungroup()

best_settings_by_family <- screening_summary %>%
  dplyr::group_by(.data$method_family) %>%
  dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
  dplyr::ungroup()
```

# 2. Susie2 Plots

## 2.1 TPR vs FPR

```{r tpr-fpr, fig.width=12, fig.height=8}
plot_settings <- top_settings_by_family %>%
  dplyr::select(.data$setting_label, .data$annotation_label)

roc_df <- confusion_bins %>%
  dplyr::semi_join(plot_settings, by = c("setting_label", "annotation_label")) %>%
  dplyr::group_by(
    .data$method_family,
    .data$setting_label,
    .data$annotation_label,
    .data$pip_threshold
  ) %>%
  dplyr::summarise(
    n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
    n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::group_by(.data$method_family, .data$setting_label, .data$annotation_label) %>%
  dplyr::arrange(dplyr::desc(.data$pip_threshold), .by_group = TRUE) %>%
  dplyr::mutate(
    cum_tp = cumsum(.data$n_causal_at_bucket),
    cum_fp = cumsum(.data$n_noncausal_at_bucket),
    total_p = sum(.data$n_causal_at_bucket),
    total_n = sum(.data$n_noncausal_at_bucket),
    tpr = dplyr::if_else(.data$total_p > 0, .data$cum_tp / .data$total_p, NA_real_),
    fpr = dplyr::if_else(.data$total_n > 0, .data$cum_fp / .data$total_n, NA_real_)
  ) %>%
  dplyr::ungroup()

p_roc <- ggplot(roc_df, aes(x = .data$fpr, y = .data$tpr, color = .data$setting_label)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~method_family, scales = "free_y") +
  labs(
    x = "FPR",
    y = "TPR",
    color = "Setting",
    title = "TPR vs FPR for top baseline settings by method family"
  ) +
  theme(legend.position = "bottom")

ggsave(
  file.path(fig_root, "susie2_plots", "tpr_fpr_top_settings.png"),
  p_roc,
  width = 12,
  height = 8,
  dpi = 150
)
```

## 2.2 PIP Calibration

```{r pip-calibration, fig.width=11, fig.height=7}
calib_screening <- screening_summary %>%
  dplyr::mutate(sigma_value = as.numeric(.data$sigma_0_2_scalar))

calib_annotation_settings <- calib_screening %>%
  dplyr::filter(.data$method_family %in% c(
    "susine_functional_mu",
    "susie_functional_pi",
    "susine_functional_pi",
    "susine_eb_clamped_scale_var_nonneg"
  )) %>%
  dplyr::filter(!is.na(.data$annotation_r2), !is.na(.data$inflate_match)) %>%
  dplyr::filter(abs(.data$annotation_r2 - 0.3) < 1e-6,
                abs(.data$inflate_match - 0.9) < 1e-6) %>%
  dplyr::group_by(.data$method_family) %>%
  dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
  dplyr::ungroup()

calib_nonann_settings <- calib_screening %>%
  dplyr::filter(.data$method_family %in% c("susie_inf", "susie_ash_fixed", "susie_eb")) %>%
  dplyr::group_by(.data$method_family) %>%
  dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
  dplyr::ungroup()

calib_vanilla_setting <- calib_screening %>%
  dplyr::filter(.data$method_family == "susie_vanilla",
                !is.na(.data$sigma_value),
                abs(.data$sigma_value - 0.2) < 1e-6) %>%
  dplyr::slice(1L)

calib_settings <- dplyr::bind_rows(
  calib_annotation_settings,
  calib_nonann_settings,
  calib_vanilla_setting
) %>%
  dplyr::select(.data$method_family, .data$setting_label, .data$annotation_label)

calib_join_keys <- intersect(
  c("method_family", "setting_label", "annotation_label"),
  names(confusion_bins)
)

cal_df <- confusion_bins %>%
  dplyr::semi_join(calib_settings, by = calib_join_keys) %>%
  dplyr::mutate(pip_bin = floor(.data$pip_threshold / 0.10) * 0.10) %>%
  dplyr::group_by(
    .data$method_family,
    .data$setting_label,
    .data$annotation_label,
    .data$dataset_bundle_id,
    .data$pip_bin
  ) %>%
  dplyr::summarise(
    n_causal = sum(.data$n_causal_at_bucket, na.rm = TRUE),
    n_noncausal = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    n_total = .data$n_causal + .data$n_noncausal,
    observed_frac = dplyr::if_else(.data$n_total > 0, .data$n_causal / .data$n_total, NA_real_)
  ) %>%
  dplyr::group_by(.data$method_family, .data$setting_label, .data$annotation_label, .data$pip_bin) %>%
  dplyr::summarise(
    mean_frac = mean(.data$observed_frac, na.rm = TRUE),
    se_frac = sd(.data$observed_frac, na.rm = TRUE) / sqrt(sum(!is.na(.data$observed_frac))),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    ci_lo = pmax(0, .data$mean_frac - 1.96 * .data$se_frac),
    ci_hi = pmin(1, .data$mean_frac + 1.96 * .data$se_frac),
    method_family = factor(.data$method_family,
      levels = c("susie_vanilla", "susie_eb", "susie_inf", "susie_ash_fixed",
                 "susie_functional_pi", "susine_functional_pi",
                 "susine_eb_clamped_scale_var_nonneg",
                 "susine_functional_mu"))
  )

p_cal <- ggplot(cal_df, aes(x = .data$pip_bin + 0.05, y = .data$mean_frac)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_errorbar(aes(ymin = .data$ci_lo, ymax = .data$ci_hi),
                width = 0.025, color = "grey35", alpha = 0.75) +
  geom_line(color = "#08519c", linewidth = 0.7) +
  geom_point(color = "#08519c", size = 1.7) +
  facet_wrap(~method_family, ncol = 4,
             labeller = labeller(method_family = method_pretty)) +
  scale_x_continuous(breaks = seq(0, 1, 0.25)) +
  scale_y_continuous(breaks = seq(0, 1, 0.25)) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    x = "Mean PIP bin midpoint",
    y = "Observed fraction causal",
    title = NULL,
    caption = "Annotation-consuming methods at R^2(a, beta) = 0.3, sign match = 0.9 (best AUPRC). SuSiE at sigma^2 = 0.2."
  ) +
  theme(legend.position = "none",
        plot.caption = element_text(size = 8, color = "grey30", hjust = 0))

ggsave(
  file.path(fig_root, "susie2_plots", "pip_calibration_top_settings.png"),
  p_cal,
  width = 11,
  height = 6.5,
  dpi = 200
)
```

## 2.3 CS Power and FDR

```{r cs-power-fdr, fig.width=12, fig.height=10}
if (!is.null(tier_cs_metrics) && nrow(tier_cs_metrics)) {
  tier_summary <- tier_cs_metrics %>%
    dplyr::semi_join(plot_settings, by = c("setting_label", "annotation_label")) %>%
    dplyr::group_by(.data$method_family, .data$setting_label, .data$n_top) %>%
    dplyr::summarise(
      sum_causal_covered = sum(.data$n_causal_covered, na.rm = TRUE),
      sum_causal_total = sum(.data$n_causal_total, na.rm = TRUE),
      sum_false_cs = sum(.data$n_false_cs, na.rm = TRUE),
      sum_cs_total = sum(.data$n_cs_total, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      cs_power = dplyr::if_else(.data$sum_causal_total > 0, .data$sum_causal_covered / .data$sum_causal_total, NA_real_),
      cs_fdr = dplyr::if_else(.data$sum_cs_total > 0, .data$sum_false_cs / .data$sum_cs_total, NA_real_)
    )

  p_power <- ggplot(tier_summary, aes(x = factor(.data$n_top), y = .data$cs_power, color = .data$setting_label, group = .data$setting_label)) +
    geom_point() +
    geom_line() +
    facet_wrap(~method_family) +
    labs(x = "Top-N treated as causal", y = "CS power", color = "Setting",
         title = "CS power for top baseline settings") +
    theme(legend.position = "bottom")

  p_fdr <- ggplot(tier_summary, aes(x = factor(.data$n_top), y = .data$cs_fdr, color = .data$setting_label, group = .data$setting_label)) +
    geom_point() +
    geom_line() +
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
    facet_wrap(~method_family) +
    labs(x = "Top-N treated as causal", y = "CS FDR", color = "Setting",
         title = "CS FDR for top baseline settings") +
    theme(legend.position = "bottom")

  ggsave(file.path(fig_root, "susie2_plots", "cs_power_top_settings.png"), p_power, width = 12, height = 5, dpi = 150)
  ggsave(file.path(fig_root, "susie2_plots", "cs_fdr_top_settings.png"), p_fdr, width = 12, height = 5, dpi = 150)
}
```

# 3. Baseline Performance

## 3.1 Method groups

```{r overall-screening, fig.width=14, fig.height=8}
obsolete_baseline_plots <- c(
  "auprc_overall_all_settings.png",
  "tpr05_overall_all_settings.png",
  "sigma_sweeps.png",
  "susine_functional_mu_heatmap.png",
  "tpr05_by_annotation.png"
)
unlink(file.path(fig_root, "baseline_performance", obsolete_baseline_plots))

functional_pi_methods <- intersect(
  c("susie_functional_pi", "susine_functional_pi"),
  unique(screening_summary$method_family)
)
annotation_methods <- c("susine_functional_mu", functional_pi_methods)
confirmation_methods <- c("susie_vanilla", "susine_vanilla")
non_annotation_methods <- intersect(
  c("susie_vanilla", "susie_inf", "susie_ash_fixed", "susie_eb"),
  unique(screening_summary$method_family)
)

metric_param_label <- function(df) {
  df %>%
    dplyr::mutate(
      sigma_value = as.numeric(.data$sigma_0_2_scalar),
      param_label = dplyr::case_when(
        !is.na(.data$sigma_value) ~ paste0("sigma=", fmt_num(.data$sigma_value, 3)),
        !is.na(.data$c_value) ~ paste0("c=", fmt_num(.data$c_value, 3)),
        !is.na(.data$tau_value) ~ paste0("tau=", fmt_num(.data$tau_value, 3)),
        TRUE ~ "default"
      ),
      annotation_r2_f = ifelse(is.na(.data$annotation_r2), "none", fmt_num(.data$annotation_r2, 3)),
      inflate_match_f = ifelse(is.na(.data$inflate_match), "none", fmt_num(.data$inflate_match, 3))
    )
}

screening_plot_df <- metric_param_label(screening_summary)
```

## 3.2 Vanilla confirmation

```{r vanilla-confirmation, fig.width=10, fig.height=5}
vanilla_confirmation <- screening_plot_df %>%
  dplyr::filter(.data$method_family %in% confirmation_methods)

p_confirm_auprc <- ggplot(
  vanilla_confirmation,
  aes(x = .data$sigma_value, y = .data$AUPRC, color = .data$method_family, group = .data$method_family)
) +
  geom_point(size = 2) +
  geom_line(linewidth = 0.7) +
  labs(
    x = expression(sigma[0]^2 / var(y)),
    y = "Pooled AUPRC",
    color = "Method family",
    title = "Vanilla implementation confirmation"
  )

p_confirm_tpr <- ggplot(
  vanilla_confirmation,
  aes(x = .data$sigma_value, y = .data$tpr_fpr05, color = .data$method_family, group = .data$method_family)
) +
  geom_point(size = 2) +
  geom_line(linewidth = 0.7) +
  labs(
    x = expression(sigma[0]^2 / var(y)),
    y = "TPR@FPR=0.05",
    color = "Method family",
    title = "Vanilla implementation confirmation"
  )

ggsave(file.path(fig_root, "baseline_performance", "confirmation", "susie_vs_susine_vanilla_auprc.png"), p_confirm_auprc, width = 10, height = 5, dpi = 150)
ggsave(file.path(fig_root, "baseline_performance", "confirmation", "susie_vs_susine_vanilla_tpr05.png"), p_confirm_tpr, width = 10, height = 5, dpi = 150)
```

## 3.3 Non-annotation baselines

```{r non-annotation-baselines, fig.width=12, fig.height=6}
non_annotation_df <- screening_plot_df %>%
  dplyr::filter(.data$method_family %in% non_annotation_methods)

non_annotation_plot_df <- non_annotation_df %>%
  dplyr::mutate(
    method_label = factor(
      unname(method_pretty[.data$method_family]),
      levels = unname(method_pretty[intersect(names(method_pretty), non_annotation_methods)])
    )
  )

non_ann_palette <- grDevices::colorRampPalette(
  c("#08306b", "#2171b5", "#6baed6", "#bdd7e7")
)(length(unique(non_annotation_plot_df$param_label)))

p_non_ann_auprc <- ggplot(
  non_annotation_plot_df,
  aes(x = .data$method_label, y = .data$AUPRC, fill = .data$param_label)
) +
  geom_col(position = position_dodge2(width = 0.8, preserve = "single"),
           color = "grey20", linewidth = 0.2) +
  scale_fill_manual(values = non_ann_palette) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Pooled AUPRC",
    fill = "Parameter",
    title = NULL
  ) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

p_non_ann_tpr <- ggplot(
  non_annotation_plot_df,
  aes(x = .data$method_label, y = .data$tpr_fpr05, fill = .data$param_label)
) +
  geom_col(position = position_dodge2(width = 0.8, preserve = "single"),
           color = "grey20", linewidth = 0.2) +
  scale_fill_manual(values = non_ann_palette) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = expression("TPR @ FPR = 0.05"),
    fill = "Parameter",
    title = NULL
  ) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

ggsave(file.path(fig_root, "baseline_performance", "auprc_non_annotation_methods.png"), p_non_ann_auprc, width = 7, height = 5, dpi = 200)
ggsave(file.path(fig_root, "baseline_performance", "tpr05_non_annotation_methods.png"), p_non_ann_tpr, width = 7, height = 5, dpi = 200)
```

## 3.4 Annotation response

```{r annotation-response, fig.width=14, fig.height=9}
annotation_df <- screening_plot_df %>%
  dplyr::filter(.data$method_family %in% annotation_methods) %>%
  dplyr::filter(!is.na(.data$annotation_r2), !is.na(.data$inflate_match))

non_annotation_reference <- non_annotation_df %>%
  dplyr::group_by(.data$method_family) %>%
  dplyr::summarise(
    AUPRC = max(.data$AUPRC, na.rm = TRUE),
    tpr_fpr05 = max(.data$tpr_fpr05, na.rm = TRUE),
    .groups = "drop"
  )

make_annotation_response_plot <- function(metric_col, y_label, title = NULL) {
  mu_df <- annotation_df %>%
    dplyr::filter(.data$method_family == "susine_functional_mu") %>%
    dplyr::mutate(
      c_value = as.numeric(.data$c_value),
      plot_value = .data[[metric_col]]
    )

  sigma_levels <- sort(unique(mu_df$sigma_value))
  sigma_label_fn <- function(s) {
    paste0("sigma^2 = ", formatC(s, format = "fg", digits = 2))
  }
  mu_df <- mu_df %>%
    dplyr::mutate(
      sigma_label = factor(sigma_label_fn(.data$sigma_value),
                           levels = sigma_label_fn(sigma_levels))
    )
  mu_palette <- grDevices::colorRampPalette(
    c("#08306b", "#2171b5", "#6baed6", "#74c476", "#238b45")
  )(length(sigma_levels))
  names(mu_palette) <- sigma_label_fn(sigma_levels)

  pi_best <- annotation_df %>%
    dplyr::filter(.data$method_family %in% functional_pi_methods) %>%
    dplyr::filter(is.na(.data$tau_value) | .data$tau_value >= 0.01) %>%
    dplyr::mutate(plot_value = .data[[metric_col]]) %>%
    dplyr::group_by(.data$annotation_r2_f, .data$inflate_match_f) %>%
    dplyr::slice_max(.data$plot_value, n = 1L, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::transmute(.data$annotation_r2_f, .data$inflate_match_f,
                     plot_value = .data$plot_value,
                     ref_label = "SuSiE functional pi (best tau)")

  susine_eb_df <- screening_plot_df %>%
    dplyr::filter(.data$method_family == "susine_eb_clamped_scale_var_nonneg") %>%
    dplyr::filter(!is.na(.data$annotation_r2), !is.na(.data$inflate_match)) %>%
    dplyr::mutate(plot_value = .data[[metric_col]]) %>%
    dplyr::group_by(.data$annotation_r2_f, .data$inflate_match_f) %>%
    dplyr::summarise(plot_value = max(.data$plot_value, na.rm = TRUE),
                     .groups = "drop") %>%
    dplyr::mutate(ref_label = "SuSiNE-EB (mu, sigma^2 estimated)")

  vanilla_value <- non_annotation_df %>%
    dplyr::filter(.data$method_family == "susie_vanilla") %>%
    dplyr::summarise(v = max(.data[[metric_col]], na.rm = TRUE)) %>%
    dplyr::pull(v)

  facet_refs <- dplyr::bind_rows(pi_best, susine_eb_df)
  vanilla_df <- tibble::tibble(plot_value = vanilla_value,
                               ref_label = "SuSiE vanilla (best sigma^2)")

  ref_levels <- c(
    "SuSiE vanilla (best sigma^2)",
    "SuSiE functional pi (best tau)",
    "SuSiNE-EB (mu, sigma^2 estimated)"
  )
  ref_palette <- stats::setNames(
    c("grey25", "#d94801", "#54278f"),
    ref_levels
  )
  facet_refs$ref_label <- factor(facet_refs$ref_label, levels = ref_levels)
  vanilla_df$ref_label <- factor(vanilla_df$ref_label, levels = ref_levels)

  ggplot() +
    geom_hline(
      data = vanilla_df,
      aes(yintercept = .data$plot_value, linetype = .data$ref_label),
      color = ref_palette[["SuSiE vanilla (best sigma^2)"]],
      linewidth = 0.55
    ) +
    geom_hline(
      data = dplyr::filter(facet_refs, .data$ref_label == "SuSiE functional pi (best tau)"),
      aes(yintercept = .data$plot_value, linetype = .data$ref_label),
      color = ref_palette[["SuSiE functional pi (best tau)"]],
      linewidth = 0.7
    ) +
    geom_hline(
      data = dplyr::filter(facet_refs, .data$ref_label == "SuSiNE-EB (mu, sigma^2 estimated)"),
      aes(yintercept = .data$plot_value, linetype = .data$ref_label),
      color = ref_palette[["SuSiNE-EB (mu, sigma^2 estimated)"]],
      linewidth = 0.7
    ) +
    scale_linetype_manual(
      name = "Reference",
      values = c(
        "SuSiE vanilla (best sigma^2)" = "solid",
        "SuSiE functional pi (best tau)" = "dashed",
        "SuSiNE-EB (mu, sigma^2 estimated)" = "dotted"
      ),
      breaks = ref_levels,
      guide = guide_legend(
        order = 2, nrow = 1,
        override.aes = list(color = unname(ref_palette[ref_levels]))
      )
    ) +
    geom_line(
      data = mu_df,
      aes(x = .data$c_value, y = .data$plot_value,
          color = .data$sigma_label,
          group = interaction(.data$sigma_label, .data$annotation_r2_f, .data$inflate_match_f)),
      linewidth = 0.8
    ) +
    geom_point(
      data = mu_df,
      aes(x = .data$c_value, y = .data$plot_value, color = .data$sigma_label),
      size = 1.5
    ) +
    scale_color_manual(
      values = mu_palette,
      name = "SuSiNE functional mu prior variance",
      guide = guide_legend(order = 1, nrow = 1)
    ) +
    facet_grid(
      rows = vars(.data$inflate_match_f),
      cols = vars(.data$annotation_r2_f),
      labeller = labeller(
        annotation_r2_f = function(x) paste0("R^2(a, beta) = ", x),
        inflate_match_f = function(x) paste0("Sign match = ", x)
      )
    ) +
    labs(
      x = expression("Annotation scale " * c),
      y = y_label,
      title = title
    ) +
    theme(legend.justification = "center")
}

p_ann_auprc <- make_annotation_response_plot(
  metric_col = "AUPRC",
  y_label = "Pooled AUPRC"
)

p_ann_tpr <- make_annotation_response_plot(
  metric_col = "tpr_fpr05",
  y_label = "TPR @ FPR = 0.05"
)

ggsave(file.path(fig_root, "baseline_performance", "auprc_by_annotation.png"), p_ann_auprc, width = 12, height = 9, dpi = 200)
ggsave(file.path(fig_root, "baseline_performance", "tpr05_annotation_consuming.png"), p_ann_tpr, width = 12, height = 9, dpi = 200)
```

## 3.5 Tau sweep

```{r tau-sweep, fig.width=10, fig.height=5}
tau_sweeps <- screening_plot_df %>%
  dplyr::filter(.data$method_family %in% functional_pi_methods) %>%
  dplyr::filter(!is.na(.data$inflate_match)) %>%
  dplyr::filter(is.na(.data$tau_value) | .data$tau_value >= 0.01) %>%
  dplyr::filter(!is.na(.data$annotation_r2))

mu_best_per_facet <- screening_plot_df %>%
  dplyr::filter(.data$method_family == "susine_functional_mu") %>%
  dplyr::filter(!is.na(.data$annotation_r2), !is.na(.data$inflate_match)) %>%
  dplyr::group_by(.data$annotation_r2_f, .data$inflate_match_f) %>%
  dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
  dplyr::ungroup()

vanilla_auprc <- screening_plot_df %>%
  dplyr::filter(.data$method_family == "susie_vanilla") %>%
  dplyr::summarise(AUPRC = max(.data$AUPRC, na.rm = TRUE)) %>%
  dplyr::pull(.data$AUPRC)

r2_levels <- sort(unique(tau_sweeps$annotation_r2_f))
r2_palette <- stats::setNames(
  grDevices::colorRampPalette(c("#d94801", "#54278f"))(length(r2_levels)),
  r2_levels
)

p_tau <- ggplot() +
  geom_hline(
    aes(yintercept = vanilla_auprc, linetype = "SuSiE vanilla (best sigma^2)"),
    color = "grey25", linewidth = 0.55
  ) +
  geom_hline(
    data = mu_best_per_facet,
    aes(yintercept = .data$AUPRC, color = .data$annotation_r2_f,
        linetype = "SuSiNE functional mu (best c, sigma^2)"),
    linewidth = 0.6
  ) +
  geom_line(
    data = tau_sweeps,
    aes(x = .data$tau_value, y = .data$AUPRC,
        color = .data$annotation_r2_f, group = .data$annotation_r2_f),
    linewidth = 0.7
  ) +
  geom_point(
    data = tau_sweeps,
    aes(x = .data$tau_value, y = .data$AUPRC, color = .data$annotation_r2_f),
    size = 1.8
  ) +
  scale_x_log10() +
  scale_color_manual(values = r2_palette, name = expression(R^2 * "(a, " * beta * ")")) +
  scale_linetype_manual(
    name = "Reference",
    values = c(
      "SuSiE vanilla (best sigma^2)" = "solid",
      "SuSiNE functional mu (best c, sigma^2)" = "dashed"
    )
  ) +
  facet_wrap(
    ~inflate_match_f,
    labeller = labeller(inflate_match_f = function(x) paste0("Sign match = ", x))
  ) +
  guides(
    color = guide_legend(order = 1),
    linetype = guide_legend(order = 2,
                            override.aes = list(color = c("grey25", "grey25")))
  ) +
  labs(
    x = expression(tau),
    y = "Pooled AUPRC",
    title = NULL
  )

ggsave(file.path(fig_root, "baseline_performance", "tau_sweep.png"),
       p_tau, width = 10, height = 5, dpi = 200)
```

## 3.6 Top settings table

```{r top-settings-table}
top_table <- screening_summary %>%
  dplyr::group_by(.data$method_family) %>%
  dplyr::slice_max(.data$AUPRC, n = 5L, with_ties = FALSE) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    .data$method_family,
    .data$setting_label,
    .data$annotation_label,
    .data$AUPRC,
    .data$tpr_fpr05
  )

readr::write_csv(top_table, file.path(fig_root, "baseline_performance", "top_settings_by_family.csv"))
kable(top_table, digits = 4, caption = "Top pooled settings by method family")
```

# 4. Dataset Difficulty

```{r dataset-difficulty, fig.width=12, fig.height=8}
if (!is.null(dataset_metrics) && nrow(dataset_metrics)) {
  best_run_ids <- auprc_per_dataset %>%
    dplyr::group_by(.data$method_family, .data$dataset_bundle_id) %>%
    dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::select(.data$dataset_bundle_id, .data$method_family, .data$AUPRC, .data$annotation_label)

  difficulty_df <- dataset_metrics %>%
    dplyr::distinct(.data$dataset_bundle_id, .keep_all = TRUE) %>%
    dplyr::inner_join(best_run_ids, by = "dataset_bundle_id")

  p_m1 <- ggplot(difficulty_df, aes(x = .data$M1, y = .data$AUPRC, color = .data$method_family)) +
    geom_point(alpha = 0.7) +
    geom_smooth(se = FALSE) +
    facet_wrap(~annotation_label) +
    labs(x = "M1", y = "Per-dataset AUPRC", color = "Method family",
         title = "Dataset difficulty vs AUPRC for best per-family settings")

  z_metric <- if ("z_eff_signals" %in% names(difficulty_df)) "z_eff_signals" else "z_topk_ratio"
  p_z <- ggplot(difficulty_df, aes(x = .data[[z_metric]], y = .data$AUPRC, color = .data$method_family)) +
    geom_point(alpha = 0.7) +
    geom_smooth(se = FALSE) +
    facet_wrap(~annotation_label) +
    labs(x = z_metric, y = "Per-dataset AUPRC", color = "Method family",
         title = "Z-score difficulty metric vs AUPRC")

  ggsave(file.path(fig_root, "dataset_difficulty", "difficulty_vs_M1.png"), p_m1, width = 12, height = 8, dpi = 150)
  ggsave(file.path(fig_root, "dataset_difficulty", "difficulty_vs_z_metric.png"), p_z, width = 12, height = 8, dpi = 150)
}
```

# 5. Effect Diagnostics

```{r effect-diagnostics, fig.width=12, fig.height=8}
if (!is.null(effect_metrics_unfiltered) && nrow(effect_metrics_unfiltered)) {
  diag_methods <- c("susie_vanilla", "susine_functional_mu", "susie_functional_pi")
  effect_diag <- effect_metrics_unfiltered %>%
    dplyr::filter(.data$method_family %in% diag_methods)

  p_main <- ggplot(
    effect_diag,
    aes(x = .data$effect_k_eff_signal_core95, y = .data$accuracy_ratio, color = .data$method_family)
  ) +
    geom_point(alpha = 0.35, size = 1.2) +
    facet_wrap(~annotation_label) +
    labs(
      x = "effect_k_eff_signal_core95",
      y = "accuracy_ratio",
      color = "Method family",
      title = "Effect diffuseness vs causal-centering accuracy"
    )

  p_tail <- ggplot(
    effect_diag,
    aes(x = .data$tail_inflation_ratio, y = .data$accuracy_ratio, color = .data$method_family)
  ) +
    geom_point(alpha = 0.35, size = 1.2) +
    facet_wrap(~annotation_label) +
    labs(
      x = "tail_inflation_ratio",
      y = "accuracy_ratio",
      color = "Method family",
      title = "Tail inflation ratio vs accuracy ratio"
    )

  p_tail_log <- ggplot(
    effect_diag,
    aes(x = .data$tail_inflation_log, y = .data$accuracy_ratio, color = .data$method_family)
  ) +
    geom_point(alpha = 0.35, size = 1.2) +
    facet_wrap(~annotation_label) +
    labs(
      x = "tail_inflation_log",
      y = "accuracy_ratio",
      color = "Method family",
      title = "Tail inflation log-ratio vs accuracy ratio"
    )

  p_entropy <- ggplot(
    effect_diag,
    aes(x = .data$effect_pip_entropy_core95, y = .data$accuracy_ratio, color = .data$method_family)
  ) +
    geom_point(alpha = 0.35, size = 1.2) +
    facet_wrap(~annotation_label) +
    labs(
      x = "effect_pip_entropy_core95",
      y = "accuracy_ratio",
      color = "Method family",
      title = "Core-95 entropy vs accuracy ratio"
    )

  p_bin2d <- ggplot(
    effect_diag,
    aes(x = .data$effect_k_eff_signal_core95, y = .data$accuracy_ratio)
  ) +
    geom_bin2d() +
    facet_grid(rows = vars(method_family), cols = vars(annotation_label)) +
    scale_fill_viridis_c() +
    labs(
      x = "effect_k_eff_signal_core95",
      y = "accuracy_ratio",
      fill = "Count",
      title = "Effect diffuseness vs accuracy ratio density"
    )

  annotation_grid <- screening_plot_df %>%
    dplyr::filter(.data$method_family %in% annotation_methods) %>%
    dplyr::filter(!is.na(.data$annotation_r2), !is.na(.data$inflate_match)) %>%
    dplyr::distinct(
      .data$annotation_r2,
      .data$inflate_match,
      .data$annotation_r2_f,
      .data$inflate_match_f
    )

  best_annotation_settings <- screening_plot_df %>%
    dplyr::filter(.data$method_family %in% annotation_methods) %>%
    dplyr::filter(!is.na(.data$annotation_r2), !is.na(.data$inflate_match)) %>%
    dplyr::filter(
      !(.data$method_family %in% functional_pi_methods) |
        is.na(.data$tau_value) |
        .data$tau_value >= 0.01
    ) %>%
    dplyr::group_by(.data$method_family, .data$annotation_r2, .data$inflate_match) %>%
    dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::select(
      .data$method_family,
      .data$setting_label,
      .data$annotation_label,
      .data$annotation_r2,
      .data$inflate_match
    )

  best_vanilla_setting <- screening_plot_df %>%
    dplyr::filter(.data$method_family == "susie_vanilla") %>%
    dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
    dplyr::select(.data$method_family, .data$setting_label)

  annotation_effect_density <- effect_metrics_unfiltered %>%
    dplyr::semi_join(
      best_annotation_settings,
      by = c("method_family", "setting_label", "annotation_label", "annotation_r2", "inflate_match")
    ) %>%
    dplyr::mutate(
      annotation_r2_f = ifelse(is.na(.data$annotation_r2), "none", fmt_num(.data$annotation_r2, 3)),
      inflate_match_f = ifelse(is.na(.data$inflate_match), "none", fmt_num(.data$inflate_match, 3)),
      density_label = .data$method_family
    )

  vanilla_effect_density <- effect_metrics_unfiltered %>%
    dplyr::semi_join(best_vanilla_setting, by = c("method_family", "setting_label")) %>%
    dplyr::select(
      -dplyr::any_of(c("annotation_r2", "inflate_match", "annotation_label"))
    ) %>%
    dplyr::mutate(density_label = "susie_vanilla") %>%
    { tidyr::expand_grid(annotation_grid, .) }

  effect_density_df <- dplyr::bind_rows(
    vanilla_effect_density,
    annotation_effect_density
  ) %>%
    dplyr::mutate(
      density_label = factor(
        .data$density_label,
        levels = c("susie_vanilla", "susine_functional_mu", functional_pi_methods)
      )
    )

  density_colors <- c(
    susie_vanilla = "black",
    susine_functional_mu = "#2c7fb8",
    stats::setNames(rep("#d73027", length(functional_pi_methods)), functional_pi_methods)
  )

  density_selection_counts <- effect_density_df %>%
    dplyr::count(.data$annotation_r2_f, .data$inflate_match_f, .data$density_label, name = "n_effect_rows")

  readr::write_csv(
    density_selection_counts,
    file.path(fig_root, "effect_diagnostics", "effect_density_selection_counts.csv")
  )
  kable(density_selection_counts, caption = "Effect-density rows selected by facet and model")

  p_accuracy_density <- ggplot(
    effect_density_df,
    aes(x = .data$accuracy_ratio, color = .data$density_label, fill = .data$density_label)
  ) +
    geom_density(alpha = 0.12, linewidth = 0.8, na.rm = TRUE) +
    facet_grid(rows = vars(inflate_match_f), cols = vars(annotation_r2_f)) +
    scale_color_manual(values = density_colors, drop = FALSE) +
    scale_fill_manual(values = density_colors, drop = FALSE) +
    coord_cartesian(xlim = c(0, 1)) +
    labs(
      x = "accuracy_ratio",
      y = "Density",
      color = "Model",
      fill = "Model",
      title = "Effect accuracy density for best annotation settings"
    ) +
    theme(legend.position = "bottom")

  p_entropy_density <- ggplot(
    effect_density_df,
    aes(x = .data$effect_pip_entropy_core95, color = .data$density_label, fill = .data$density_label)
  ) +
    geom_density(alpha = 0.12, linewidth = 0.8, na.rm = TRUE) +
    facet_grid(rows = vars(inflate_match_f), cols = vars(annotation_r2_f)) +
    scale_color_manual(values = density_colors, drop = FALSE) +
    scale_fill_manual(values = density_colors, drop = FALSE) +
    labs(
      x = "effect_pip_entropy_core95",
      y = "Density",
      color = "Model",
      fill = "Model",
      title = "Effect core-95 entropy density for best annotation settings"
    ) +
    theme(legend.position = "bottom")

  make_kde_grid <- function(df, x_col, y_col, x_range, y_range, n = 90L) {
    out_grid <- tibble::as_tibble(
      base::expand.grid(
        x = seq(x_range[[1]], x_range[[2]], length.out = n),
        y = seq(y_range[[1]], y_range[[2]], length.out = n)
      )
    )

    df <- df %>%
      dplyr::filter(is.finite(.data[[x_col]]), is.finite(.data[[y_col]]))
    if (nrow(df) < 5L ||
        dplyr::n_distinct(df[[x_col]]) < 2L ||
        dplyr::n_distinct(df[[y_col]]) < 2L) {
      return(dplyr::mutate(out_grid, density = 0))
    }

    kde <- MASS::kde2d(
      x = df[[x_col]],
      y = df[[y_col]],
      n = n,
      lims = c(x_range, y_range)
    )
    dplyr::mutate(out_grid, density = as.vector(kde$z))
  }

  diff_x_range <- range(effect_density_df$effect_pip_entropy_core95, finite = TRUE, na.rm = TRUE)
  if (!all(is.finite(diff_x_range)) || diff_x_range[[1]] == diff_x_range[[2]]) {
    diff_x_range <- c(0, 1)
  }
  diff_y_range <- c(0, 1)

  annotation_density_methods <- effect_density_df %>%
    dplyr::filter(.data$density_label != "susie_vanilla") %>%
    dplyr::filter(!is.na(.data$density_label)) %>%
    dplyr::distinct(.data$density_label) %>%
    dplyr::pull(.data$density_label) %>%
    as.character()

  differential_density <- dplyr::bind_rows(lapply(annotation_density_methods, function(method_id) {
    dplyr::bind_rows(lapply(seq_len(nrow(annotation_grid)), function(i) {
      facet_row <- annotation_grid[i, , drop = FALSE]
      baseline_rows <- effect_density_df %>%
        dplyr::filter(
          .data$density_label == "susie_vanilla",
          .data$annotation_r2_f == facet_row$annotation_r2_f[[1]],
          .data$inflate_match_f == facet_row$inflate_match_f[[1]]
        )
      annotation_rows <- effect_density_df %>%
        dplyr::filter(
          as.character(.data$density_label) == method_id,
          .data$annotation_r2_f == facet_row$annotation_r2_f[[1]],
          .data$inflate_match_f == facet_row$inflate_match_f[[1]]
        )

      baseline_kde <- make_kde_grid(
        baseline_rows,
        x_col = "effect_pip_entropy_core95",
        y_col = "accuracy_ratio",
        x_range = diff_x_range,
        y_range = diff_y_range
      )
      annotation_kde <- make_kde_grid(
        annotation_rows,
        x_col = "effect_pip_entropy_core95",
        y_col = "accuracy_ratio",
        x_range = diff_x_range,
        y_range = diff_y_range
      )

      tibble::tibble(
        density_label = method_id,
        annotation_r2_f = facet_row$annotation_r2_f[[1]],
        inflate_match_f = facet_row$inflate_match_f[[1]],
        x = annotation_kde$x,
        y = annotation_kde$y,
        density_diff = annotation_kde$density - baseline_kde$density
      )
    }))
  }))

  if (nrow(differential_density)) {
    readr::write_csv(
      differential_density,
      file.path(fig_root, "effect_diagnostics", "effect_entropy_accuracy_differential_density.csv")
    )

    diff_limit <- max(abs(differential_density$density_diff), na.rm = TRUE)
    if (!is.finite(diff_limit) || diff_limit <= 0) {
      diff_limit <- 1
    }

    for (method_id in annotation_density_methods) {
      method_diff <- differential_density %>%
        dplyr::filter(.data$density_label == method_id)
      p_diff_density <- ggplot(
        method_diff,
        aes(x = .data$x, y = .data$y, fill = .data$density_diff)
      ) +
        geom_raster(interpolate = TRUE) +
        geom_contour(aes(z = .data$density_diff), breaks = 0, color = "black", linewidth = 0.25, alpha = 0.6) +
        facet_grid(rows = vars(inflate_match_f), cols = vars(annotation_r2_f)) +
        scale_fill_gradient2(
          low = "#2c7bb6",
          mid = "white",
          high = "#d7191c",
          midpoint = 0,
          limits = c(-diff_limit, diff_limit),
          name = "Density difference"
        ) +
        coord_cartesian(xlim = diff_x_range, ylim = diff_y_range, expand = FALSE) +
        labs(
          x = "effect_pip_entropy_core95",
          y = "accuracy_ratio",
          title = paste0(method_id, " differential effect density vs susie_vanilla")
        )

      safe_method_id <- gsub("[^A-Za-z0-9]+", "_", method_id)
      ggsave(
        file.path(fig_root, "effect_diagnostics", paste0("effect_entropy_accuracy_differential_density_", safe_method_id, ".png")),
        p_diff_density,
        width = 14,
        height = 9,
        dpi = 150
      )
    }
  }

  ggsave(file.path(fig_root, "effect_diagnostics", "effect_keff_core95_vs_accuracy_ratio.png"), p_main, width = 12, height = 8, dpi = 150)
  ggsave(file.path(fig_root, "effect_diagnostics", "tail_inflation_ratio_vs_accuracy_ratio.png"), p_tail, width = 12, height = 8, dpi = 150)
  ggsave(file.path(fig_root, "effect_diagnostics", "tail_inflation_log_vs_accuracy_ratio.png"), p_tail_log, width = 12, height = 8, dpi = 150)
  ggsave(file.path(fig_root, "effect_diagnostics", "effect_entropy_core95_vs_accuracy_ratio.png"), p_entropy, width = 12, height = 8, dpi = 150)
  ggsave(file.path(fig_root, "effect_diagnostics", "effect_keff_core95_vs_accuracy_ratio_bin2d.png"), p_bin2d, width = 14, height = 10, dpi = 150)
  ggsave(file.path(fig_root, "effect_diagnostics", "accuracy_ratio_density_best_annotation_settings.png"), p_accuracy_density, width = 14, height = 9, dpi = 150)
  ggsave(file.path(fig_root, "effect_diagnostics", "effect_entropy_core95_density_best_annotation_settings.png"), p_entropy_density, width = 14, height = 9, dpi = 150)
}
```
