---
title: "Ensemble Scaling Experiment - Paper Figures"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Paper-focused figure renderer for the ensemble scaling simulation. Run
`prepare_results_workbook_ensemble_scaling_paper.Rmd` first on the HPC
consolidated outputs; this workbook then reads the compact plot-data RDS and
writes PNG figures only.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_ensemble_scaling_paper.Rmd")
library(here)
library(dplyr)
library(ggplot2)
library(readr)
library(knitr)

theme_set(theme_bw(base_size = 11))
```

# 0. User Configuration

```{r config}
job_name <- "ensemble_scaling_full"
parent_job_id <- "51250228"
output_root <- here("output")

job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
fig_dir <- file.path(job_dir, "figures", "paper_ensemble_scaling")
plot_data_dir <- file.path(fig_dir, "plot_data")
plot_data_rds <- file.path(plot_data_dir, "paper_ensemble_plot_data.rds")

if (!file.exists(plot_data_rds)) {
  stop(
    "Prepared paper plot-data RDS not found: ", plot_data_rds, "\n",
    "Run prepare_results_workbook_ensemble_scaling_paper.Rmd first.",
    call. = FALSE
  )
}
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

cat("Plot-data RDS:", plot_data_rds, "\n")
cat("Figure dir:", fig_dir, "\n")
```

# 1. Load Plot Data

```{r load-plot-data}
paper_plot_data <- readRDS(plot_data_rds)

annotation_r2_focus <- paper_plot_data$config$annotation_r2_focus
drop_aggs <- paper_plot_data$config$drop_aggs
bootstrap_B <- paper_plot_data$config$bootstrap_B
bootstrap_seed <- paper_plot_data$config$bootstrap_seed
true_h2 <- paper_plot_data$config$true_h2

list2env(paper_plot_data$labels, envir = environment())
list2env(paper_plot_data$refs, envir = environment())
list2env(paper_plot_data$data, envir = environment())
list2env(paper_plot_data$diagnostics, envir = environment())

cat("Prepared for annotation_r2 focus:", annotation_r2_focus, "\n")
cat("Baseline SuSiE AUPRC:", round(baseline_auprc, 4), "\n")
cat("Truth-warm vanilla AUPRC:", round(truth_warm_vanilla, 4), "\n")
cat("Baseline model-fit wall time:", round(baseline_compute_sec, 3), "sec\n")
```

```{r helpers}
format_r2_tag <- function(x) {
  gsub("\\.", "p", format(x, trim = TRUE, scientific = FALSE))
}

open_png_device <- function(path, width, height, res = 300, bg = "white") {
  if (requireNamespace("ragg", quietly = TRUE)) {
    return(ragg::agg_png(
      filename = path,
      width = width,
      height = height,
      units = "in",
      res = res,
      background = bg
    ))
  }

  if (isTRUE(grDevices::capabilities("cairo"))) {
    return(grDevices::png(
      filename = path,
      width = width,
      height = height,
      units = "in",
      res = res,
      bg = bg,
      type = "cairo"
    ))
  }

  stop(
    "No headless PNG device is available. Install the R package 'ragg' ",
    "or use an R build with Cairo support.",
    call. = FALSE
  )
}

save_paper_png <- function(plot, filename, width, height) {
  path <- file.path(fig_dir, filename)
  open_png_device(path, width = width, height = height, res = 300, bg = "white")
  on.exit(grDevices::dev.off(), add = TRUE)
  print(plot)
  invisible(path)
}

paper_theme <- function(base_size = 11) {
  ggplot2::theme_bw(base_size = base_size) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      plot.title.position = "plot",
      legend.position = "bottom",
      legend.box = "vertical"
    )
}

gain_fill_scale <- function(values, name = "Delta AUPRC") {
  if (length(values) > 0L && min(values, na.rm = TRUE) >= 0) {
    return(ggplot2::scale_fill_gradientn(
      colors = c("#F7FCF0", "#D9F0A3", "#ADDD8E", "#78C679",
                 "#41AB5D", "#238443", "#005A32"),
      name = name
    ))
  }

  ggplot2::scale_fill_gradient2(
    low = "#2166AC",
    mid = "white",
    high = "#B2182B",
    midpoint = 0,
    name = name
  )
}

heatmap_text_color <- function(values) {
  if (!length(values) || all(!is.finite(values))) {
    return(rep("black", length(values)))
  }
  rng <- range(values, na.rm = TRUE)
  if (!is.finite(diff(rng)) || diff(rng) <= 0) {
    return(rep("black", length(values)))
  }
  ifelse(values >= rng[[1]] + 0.75 * diff(rng), "white", "black")
}

reference_label_layer <- function(x, y, label, hjust = 1.02, vjust = -0.35,
                                  size = 3) {
  if (!is.finite(y)) return(NULL)
  ggplot2::annotate(
    "label",
    x = x,
    y = y,
    label = label,
    hjust = hjust,
    vjust = vjust,
    size = size,
    label.size = 0,
    label.padding = grid::unit(0.08, "lines"),
    fill = "white",
    alpha = 0.85,
    color = "grey20"
  )
}

shape_values_for <- function(labels) {
  shape_pool <- c(16, 17, 15, 18, 3, 4, 8, 7, 0, 1, 2, 5, 6,
                  9, 10, 11, 12, 13, 14, 19, 20, 21, 22, 23, 24, 25)
  stats::setNames(rep_len(shape_pool, length(labels)), labels)
}

sort_subscale_labels <- function(labels) {
  labels <- unique(as.character(labels))
  first_dim <- suppressWarnings(as.integer(sub("x.*$", "", labels)))
  second_dim <- suppressWarnings(as.integer(sub("^.*x", "", labels)))
  labels[order(second_dim, first_dim, labels, na.last = TRUE)]
}

tag_panel <- function(plot, tag) {
  plot +
    ggplot2::labs(tag = tag) +
    ggplot2::theme(
      plot.tag = ggplot2::element_text(face = "bold", size = 18),
      plot.tag.position = c(0.01, 0.99),
      plot.margin = grid::unit(c(5.5, 5.5, 5.5, 5.5), "pt")
    )
}

save_composite_png <- function(filename, panels, layout_matrix,
                               widths = NULL, heights = NULL,
                               width = 15, height = 20) {
  stopifnot(is.matrix(layout_matrix))
  if (is.null(widths)) widths <- rep(1, ncol(layout_matrix))
  if (is.null(heights)) heights <- rep(1, nrow(layout_matrix))

  path <- file.path(fig_dir, filename)
  open_png_device(path, width = width, height = height, res = 300, bg = "white")
  on.exit(grDevices::dev.off(), add = TRUE)

  grid::grid.newpage()
  grid::pushViewport(grid::viewport(
    layout = grid::grid.layout(
      nrow(layout_matrix),
      ncol(layout_matrix),
      widths = grid::unit(widths, "null"),
      heights = grid::unit(heights, "null")
    )
  ))

  panel_ids <- unique(as.vector(layout_matrix))
  panel_ids <- panel_ids[!is.na(panel_ids)]
  for (id in panel_ids) {
    pos <- which(layout_matrix == id, arr.ind = TRUE)
    print(
      panels[[id]],
      vp = grid::viewport(
        layout.pos.row = range(pos[, "row"]),
        layout.pos.col = range(pos[, "col"])
      ),
      newpage = FALSE
    )
  }
  invisible(path)
}
```

# 2. Manuscript Numbers

```{r manuscript-numbers}
knitr::kable(spec_key, caption = "Ensemble abbreviation key")

knitr::kable(
  ccs_compute_coordinates,
  digits = 4,
  caption = "C-CS compute plot coordinates for 4x4 and 8x8"
)

knitr::kable(
  pr_precision50_points,
  digits = 4,
  caption = "PR operating points: maximum recall with precision >= 50%"
)

knitr::kable(
  pr_same_recall_points,
  digits = 4,
  caption = "PR operating points: non-baseline rows at baseline 50%-precision recall"
)
```

# 3. Overall Ensemble Performance Heatmap

```{r fig-delta-heatmap, fig.width=11, fig.height=8}
if (nrow(heatmap_data) > 0L && is.finite(baseline_auprc)) {
  heatmap_plot_data <- heatmap_data %>%
    dplyr::mutate(
      tile_text_color = heatmap_text_color(.data$delta_auprc),
      spec_name = factor(.data$spec_name, levels = rev(spec_order)),
      agg_method = factor(.data$agg_method, levels = agg_order)
    )

  y_levels <- levels(heatmap_plot_data$spec_name)
  y_prefix <- sub("-.*$", "", y_levels)
  y_separators <- if (length(y_prefix) > 1L) {
    which(y_prefix[-length(y_prefix)] != y_prefix[-1L]) + 0.5
  } else {
    numeric(0L)
  }
  oracle_x <- match("oracle", agg_order) - 0.5

  p_heatmap <- ggplot2::ggplot(
    heatmap_plot_data,
    ggplot2::aes(x = .data$agg_method, y = .data$spec_name, fill = .data$delta_auprc)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.4) +
    ggplot2::geom_text(
      ggplot2::aes(label = .data$tile_label, color = .data$tile_text_color),
      size = 3.6,
      lineheight = 0.9
    ) +
    ggplot2::geom_hline(
      yintercept = y_separators,
      linewidth = 0.6,
      color = "grey20"
    ) +
    ggplot2::geom_vline(
      xintercept = oracle_x,
      linewidth = 0.6,
      color = "grey20"
    ) +
    ggplot2::scale_x_discrete(labels = agg_labels, drop = FALSE) +
    ggplot2::scale_color_identity() +
    gain_fill_scale(heatmap_plot_data$delta_auprc) +
    ggplot2::labs(
      x = "Aggregation method",
      y = "Ensemble method",
      title = "AUPRC gain over baseline SuSiE",
      subtitle = paste0(
        "annotation_r2 = ", annotation_r2_focus,
        "; labels show delta AUPRC and percent gain"
      )
    ) +
    paper_theme(base_size = 15) +
    ggplot2::theme(
      axis.text.x = ggplot2::element_text(size = 13),
      axis.text.y = ggplot2::element_text(size = 13),
      legend.text = ggplot2::element_text(size = 12),
      legend.title = ggplot2::element_text(size = 13)
    )

  save_paper_png(
    p_heatmap,
    paste0("paper_delta_auprc_heatmap_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 11,
    height = 8
  )
  invisible(NULL)
} else {
  message("Skipping heatmap: no rows after filtering.")
}
```

# 4. Best Methods vs Annotation Quality

```{r fig-ccs-by-r2, fig.width=7, fig.height=4.8}
ccs_by_r2_plot_data <- ccs_by_r2 %>%
  dplyr::mutate(agg_method = factor(.data$agg_method, levels = agg_order))

if (nrow(ccs_by_r2_plot_data) > 0L) {
  ccs_r2_label_x <- max(ccs_by_r2_plot_data$annotation_r2, na.rm = TRUE)

  p_ccs_r2 <- ggplot2::ggplot(
    ccs_by_r2_plot_data,
    ggplot2::aes(
      x = .data$annotation_r2,
      y = .data$AUPRC,
      color = .data$agg_method,
      group = .data$agg_method
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.55
    ) +
    ggplot2::geom_hline(
      yintercept = truth_warm_vanilla,
      linetype = "dotted",
      color = "grey25",
      linewidth = 0.55
    ) +
    ggplot2::geom_line(linewidth = 0.75) +
    ggplot2::geom_point(size = 2) +
    reference_label_layer(ccs_r2_label_x, baseline_auprc, "SuSiE baseline") +
    reference_label_layer(
      ccs_r2_label_x,
      truth_warm_vanilla,
      "truth-warm vanilla",
      vjust = 1.2
    ) +
    ggplot2::coord_cartesian(ylim = c(0.2, 0.4)) +
    ggplot2::scale_x_continuous(breaks = sort(unique(ccs_by_r2_plot_data$annotation_r2))) +
    ggplot2::scale_color_discrete(labels = agg_labels[levels(ccs_by_r2_plot_data$agg_method)]) +
    ggplot2::labs(
      x = "Annotation R2",
      y = "Pooled AUPRC",
      color = "Aggregation",
      title = "C-CS aggregation performance across annotation quality"
    ) +
    paper_theme()

  save_paper_png(
    p_ccs_r2,
    "paper_ccs_aggregation_by_annotation_r2.png",
    width = 7,
    height = 4.8
  )
  invisible(NULL)
} else {
  message("Skipping C-CS by annotation quality plot: no rows.")
}
```

# 5. Bootstrap AUPRC CI

```{r fig-bootstrap, fig.width=7, fig.height=4.8}
if (nrow(boot_plot_data) > 0L) {
  p_boot <- ggplot2::ggplot(
    boot_plot_data,
    ggplot2::aes(x = .data$delta_AUPRC)
  ) +
    ggplot2::geom_histogram(
      ggplot2::aes(y = ggplot2::after_stat(density)),
      bins = 45,
      fill = "#4C78A8",
      color = "white",
      linewidth = 0.2
    ) +
    ggplot2::geom_vline(
      xintercept = 0,
      linetype = "dotted",
      color = "grey35",
      linewidth = 0.55
    ) +
    ggplot2::geom_vline(
      xintercept = delta_ci,
      linetype = "dashed",
      color = "grey20",
      linewidth = 0.55
    ) +
    ggplot2::geom_vline(
      xintercept = observed_delta,
      color = "#1B4F72",
      linewidth = 0.7
    ) +
    ggplot2::labs(
      x = "Paired bootstrap delta AUPRC (C-CS cluster weight - SuSiE baseline)",
      y = "Density",
      title = "Paired dataset bootstrap gain",
      subtitle = sprintf(
        "annotation_r2 = %.1f; n = %d datasets; observed delta = %.3f; 95%% CI: [%.3f, %.3f]",
        annotation_r2_focus,
        shared_boot_n,
        observed_delta,
        delta_ci[[1]],
        delta_ci[[2]]
      )
    ) +
    paper_theme()

  save_paper_png(
    p_boot,
    paste0("paper_bootstrap_auprc_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 7,
    height = 4.8
  )
  invisible(NULL)
} else {
  message("Skipping bootstrap plot: insufficient shared dataset IDs.")
}
```

# 6. PR Curves

```{r fig-pr-curves, fig.width=8, fig.height=5}
pr_plot_data <- pr_curve_counts %>%
  dplyr::select(method, annotation_r2, recall, precision) %>%
  dplyr::mutate(
    method = factor(.data$method, levels = c("SuSiE baseline", "C-CS cluster weight"))
  )

if (nrow(pr_plot_data) > 0L) {
  p_pr <- ggplot2::ggplot(
    pr_plot_data,
    ggplot2::aes(x = .data$recall, y = .data$precision, color = .data$method)
  ) +
    ggplot2::geom_path(linewidth = 0.75) +
    ggplot2::facet_wrap(~annotation_r2, labeller = ggplot2::label_both) +
    ggplot2::scale_color_manual(
      values = c("SuSiE baseline" = "grey35", "C-CS cluster weight" = "#4C78A8")
    ) +
    ggplot2::coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
    ggplot2::labs(
      x = "Recall",
      y = "Precision",
      color = NULL,
      title = "Precision-recall curves for C-CS cluster weight"
    ) +
    paper_theme()

  save_paper_png(
    p_pr,
    "paper_pr_curves_ccs_cluster_weight_by_r2.png",
    width = 8,
    height = 5
  )
  invisible(NULL)
} else {
  message("Skipping PR curves: no rows.")
}
```

# 7. Estimated Heritability

```{r fig-hg2, fig.width=8.5, fig.height=5}
if (nrow(hg2_plot_data) > 0L) {
  hg2_plot_data <- hg2_plot_data %>%
    dplyr::mutate(
      spec_name = factor(.data$spec_name, levels = c("baseline", spec_order)),
      series = factor(.data$series, levels = c("baseline", "ensemble", "C-CS"))
    )

  p_hg2 <- ggplot2::ggplot(
    hg2_plot_data,
    ggplot2::aes(x = .data$spec_name, y = .data$hg2, fill = .data$series)
  ) +
    ggplot2::geom_boxplot(outlier.size = 0.4, linewidth = 0.35) +
    ggplot2::geom_hline(
      yintercept = true_h2,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.55
    ) +
    reference_label_layer(
      tail(levels(hg2_plot_data$spec_name), 1L),
      true_h2,
      sprintf("true h2 = %.2f", true_h2)
    ) +
    ggplot2::scale_fill_manual(
      values = c(baseline = "grey70", ensemble = "#9BBAD7", `C-CS` = "#D95F02"),
      guide = "none"
    ) +
    ggplot2::labs(
      x = "Method",
      y = "Estimated h_g^2",
      title = "Estimated heritability under cluster-weight aggregation",
      subtitle = paste0("annotation_r2 = ", annotation_r2_focus)
    ) +
    paper_theme() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))

  save_paper_png(
    p_hg2,
    paste0("paper_hg2_cluster_weight_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 8.5,
    height = 5
  )
  invisible(NULL)
} else {
  message("Skipping hg2 plot: no rows after filtering.")
}
```

# 8. PIP Calibration

```{r fig-pip-calibration, fig.width=6.8, fig.height=5}
if (nrow(pip_cal_data) > 0L) {
  p_cal <- ggplot2::ggplot(
    pip_cal_data,
    ggplot2::aes(x = .data$pip_mid, y = .data$mean_frac, color = .data$method)
  ) +
    ggplot2::geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey35") +
    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = .data$lo, ymax = .data$hi),
      width = 0.015,
      linewidth = 0.35
    ) +
    ggplot2::geom_point(size = 2) +
    ggplot2::geom_line(linewidth = 0.65) +
    ggplot2::coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
    ggplot2::scale_color_manual(
      values = c("SuSiE baseline" = "grey35", "C-CS cluster weight" = "#4C78A8")
    ) +
    ggplot2::labs(
      x = "Mean PIP bin",
      y = "Observed causal fraction",
      color = NULL,
      title = "PIP calibration",
      subtitle = paste0("annotation_r2 = ", annotation_r2_focus)
    ) +
    paper_theme()

  save_paper_png(
    p_cal,
    paste0(
      "paper_pip_calibration_ccs_cluster_weight_r2_",
      format_r2_tag(annotation_r2_focus),
      ".png"
    ),
    width = 6.8,
    height = 5
  )
  invisible(NULL)
} else {
  message("Skipping PIP calibration plot: no rows.")
}
```

# 9. C-CS 4x4 vs 8x8 Scaling

```{r fig-ccs-resolution, fig.width=6.5, fig.height=4.7}
if (nrow(ccs_resolution_plot_data) > 0L) {
  ccs_resolution_plot_data <- ccs_resolution_plot_data %>%
    dplyr::mutate(
      resolution = factor(.data$resolution, levels = c("4x4", "8x8")),
      method = factor(.data$method, levels = c("cluster weight", "oracle"))
    )

  ccs_resolution_label_x <- tail(levels(ccs_resolution_plot_data$resolution), 1L)

  p_res <- ggplot2::ggplot(
    ccs_resolution_plot_data,
    ggplot2::aes(
      x = .data$resolution,
      y = .data$AUPRC,
      color = .data$method,
      group = .data$method
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.55
    ) +
    ggplot2::geom_line(linewidth = 0.75) +
    ggplot2::geom_point(size = 2.5) +
    reference_label_layer(
      ccs_resolution_label_x,
      baseline_auprc,
      "SuSiE baseline"
    ) +
    ggplot2::labs(
      x = "C-CS grid resolution",
      y = "Pooled AUPRC",
      color = NULL,
      title = "C-CS grid resolution scaling",
      subtitle = paste0("annotation_r2 = ", annotation_r2_focus)
    ) +
    paper_theme()

  save_paper_png(
    p_res,
    paste0("paper_ccs_4x4_8x8_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 6.5,
    height = 4.7
  )
  invisible(NULL)
} else {
  message("Skipping C-CS 4x4/8x8 plot: no rows.")
}
```

# 10. Compute vs AUPRC

```{r fig-compute-vs-auprc, fig.width=8.8, fig.height=5}
if (nrow(compute_plot_data) > 0L) {
  subscale_levels <- sort_subscale_labels(compute_plot_data$subscale_label)
  compute_plot_data <- compute_plot_data %>%
    dplyr::mutate(
      subscale_label = factor(.data$subscale_label, levels = subscale_levels),
      spec_name = factor(.data$spec_name, levels = spec_order)
    )
  compute_label_x <- max(compute_plot_data$compute_min, na.rm = TRUE)

  p_compute <- ggplot2::ggplot(
    compute_plot_data,
    ggplot2::aes(
      x = .data$compute_min,
      y = .data$AUPRC_mean,
      color = .data$spec_name,
      shape = .data$subscale_label
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.55
    ) +
    ggplot2::geom_point(size = 2.4, alpha = 0.9, stroke = 0.8) +
    ggplot2::annotate(
      "point",
      x = baseline_compute_min,
      y = baseline_auprc,
      shape = 8,
      size = 3.2,
      color = "black",
      stroke = 1
    ) +
    reference_label_layer(compute_label_x, baseline_auprc, "SuSiE baseline") +
    (if (is.finite(baseline_compute_min)) {
      reference_label_layer(
        baseline_compute_min,
        baseline_auprc,
        sprintf("%.1f sec", baseline_compute_sec),
        hjust = -0.1,
        vjust = -0.45,
        size = 2.8
      )
    } else {
      NULL
    }) +
    ggplot2::labs(
      x = "Mean model-fit wall time per dataset (minutes)",
      y = "Pooled AUPRC",
      color = "Ensemble",
      shape = "Subscale shape",
      title = "Model-fit time versus AUPRC for cluster-weight aggregation",
      subtitle = paste0(
        "A specs are annotation-independent; B/C specs use annotation_r2 = ",
        annotation_r2_focus
      ),
      caption = if (is.finite(baseline_compute_min)) {
        sprintf(
          paste0(
            "Baseline marker uses mean wall_time_sec = %.1f sec (%.4f min); ",
            "wall_time_sec times model fitting only, excluding data generation, ",
            "evaluation, aggregation, I/O, and scheduler overhead."
          ),
          baseline_compute_sec,
          baseline_compute_min
        )
      } else {
        NULL
      }
    ) +
    ggplot2::scale_shape_manual(values = shape_values_for(subscale_levels)) +
    ggplot2::guides(
      color = ggplot2::guide_legend(
        nrow = 2,
        byrow = TRUE,
        order = 1,
        override.aes = list(shape = 16, size = 3)
      ),
      shape = ggplot2::guide_legend(
        nrow = 1,
        byrow = TRUE,
        order = 2,
        override.aes = list(color = "grey20", size = 3)
      )
    ) +
    paper_theme() +
    ggplot2::theme(
      legend.spacing.y = grid::unit(0.03, "cm"),
      legend.box.spacing = grid::unit(0.08, "cm")
    )

  save_paper_png(
    p_compute,
    paste0(
      "paper_compute_vs_auprc_cluster_weight_r2_",
      format_r2_tag(annotation_r2_focus),
      ".png"
    ),
    width = 8.8,
    height = 5
  )
  invisible(NULL)
} else {
  message("Skipping compute plot: no rows after joining compute times.")
}
```

# 11. Composite Main Figure

The bootstrap panel is intentionally left out of the main composite. It is
better summarized in text using the paired delta CI from the prep workbook,
while the figure space is used for performance patterns, operating behavior,
calibration, heritability, and model-fit time.

```{r fig-composite-main, fig.width=15, fig.height=20}
composite_required <- c(
  "p_heatmap", "p_ccs_r2", "p_res", "p_pr",
  "p_cal", "p_hg2", "p_compute"
)
missing_composite <- composite_required[!vapply(
  composite_required,
  exists,
  logical(1L),
  envir = knitr::knit_global()
)]

if (length(missing_composite) == 0L) {
  composite_panels <- list(
    A = tag_panel(p_heatmap, "A"),
    B = tag_panel(p_ccs_r2, "B"),
    C = tag_panel(p_res, "C"),
    D = tag_panel(p_pr, "D"),
    E = tag_panel(p_cal, "E"),
    F = tag_panel(p_hg2, "F"),
    G = tag_panel(p_compute, "G")
  )

  composite_layout <- rbind(
    c("A", "A"),
    c("B", "C"),
    c("D", "E"),
    c("F", "F"),
    c("G", "G")
  )

  composite_path <- save_composite_png(
    filename = "paper_ensemble_scaling_composite_main.png",
    panels = composite_panels,
    layout_matrix = composite_layout,
    widths = c(1, 1),
    heights = c(2.35, 1.25, 1.45, 1.55, 1.75),
    width = 15,
    height = 20
  )

  cat("Saved composite main figure:", composite_path, "\n")
  invisible(NULL)
} else {
  message(
    "Skipping composite figure; missing component plots: ",
    paste(missing_composite, collapse = ", ")
  )
}
```

# 12. Output Check

```{r output-check}
expected_figures <- c(
  paste0("paper_delta_auprc_heatmap_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  "paper_ccs_aggregation_by_annotation_r2.png",
  paste0("paper_bootstrap_auprc_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  "paper_pr_curves_ccs_cluster_weight_by_r2.png",
  paste0("paper_hg2_cluster_weight_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  paste0(
    "paper_pip_calibration_ccs_cluster_weight_r2_",
    format_r2_tag(annotation_r2_focus),
    ".png"
  ),
  paste0("paper_ccs_4x4_8x8_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  paste0(
    "paper_compute_vs_auprc_cluster_weight_r2_",
    format_r2_tag(annotation_r2_focus),
    ".png"
  ),
  "paper_ensemble_scaling_composite_main.png"
)

figure_paths <- file.path(fig_dir, expected_figures)
figure_check <- tibble::tibble(
  figure = expected_figures,
  exists = file.exists(figure_paths),
  bytes = ifelse(file.exists(figure_paths), file.info(figure_paths)$size, NA_real_)
)

knitr::kable(figure_check, caption = "Expected paper figure outputs")
```
