---
title: "SuSiNE EB Method Ablation — Visualization (SuSiE2-Style)"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Visualizes EB ablation study results using SuSiE 2.0 paper-style plots
(CS power, CS FDR, hg2, TPR/FPR curves). Structured around 5 method-comparison groups:

  1. Baseline — no annotation, no EB (susine_vanilla)
  2. Annotations, no EB (susine_functional_mu, c-grid)
  3. EB, no annotations (susie_eb / susine_eb / susine_eb_mean / susine_eb_mu_var)
  4. EB with annotations (all annotation-based EB methods)
  5. Master summary table
  6. Prior diagnostics

Architectures: susie2_sparse + susie2_oligogenic only.
Only real model fits (variant_type == "base") are used — no ensemble aggregations.
Each comparison section includes both sparse (B-E) and oligogenic (F-I) panels
modeled after SuSiE 2.0 paper Figure 1.

Run collect_results_workbook.Rmd first to produce aggregated CSVs.

# Section 1: Setup

```{r setup, message=FALSE, warning=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_eb_ablation.Rmd")
library(here)
library(devtools)
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(stringr)
library(purrr)
library(scales)
library(knitr)
```

## 1.1 Job Configuration

```{r job-config}
# UPDATE these two values after each run.
job_name      <- "eb_ablation_v2"
parent_job_id <- "49603067"
output_root   <- here("output")

paths <- test_susine::resolve_job_paths(
  job_name      = job_name,
  parent_job_id = parent_job_id,
  output_root   = output_root
)
```

## 1.2 Plot Directories

```{r plot-dirs}
save_plots_to_file    <- TRUE
show_in_rstudio_plots <- TRUE

if (isTRUE(getOption("knitr.in.progress"))) {
  knitr::opts_chunk$set(fig.ext = "png")
} else if (show_in_rstudio_plots && interactive()) {
  knitr::opts_chunk$set(fig.show = "hide", dev = "RStudioGD", fig.ext = "png")
} else {
  knitr::opts_chunk$set(fig.ext = "png")
}

plot_base_dir <- file.path(paths$aggregated_dir, "plots_eb_ablation")
plot_dirs <- list(
  baseline    = file.path(plot_base_dir, "s4_baseline"),
  annot_noeb  = file.path(plot_base_dir, "s5_annot_noeb"),
  eb_noannot  = file.path(plot_base_dir, "s6_eb_noannot"),
  eb_annot    = file.path(plot_base_dir, "s7_eb_annot"),
  master      = file.path(plot_base_dir, "s8_master_table"),
  diagnostics = file.path(plot_base_dir, "s9_prior_diagnostics")
)
for (d in plot_dirs) dir.create(d, showWarnings = FALSE, recursive = TRUE)
cat("Plot directories created under:", plot_base_dir, "\n")

save_or_show <- function(p, filename, width = 12, height = 7,
                          dpi = 300, subdir = "baseline") {
  if (show_in_rstudio_plots && interactive()) {
    try(print(p), silent = TRUE)
  }
  if (save_plots_to_file) {
    fp <- file.path(plot_dirs[[subdir]], filename)
    tryCatch(
      { ggsave(fp, p, width = width, height = height, dpi = dpi, bg = "white")
        cat("Saved:", filename, "\n") },
      error = function(e) cat("Error saving:", filename, "-", conditionMessage(e), "\n")
    )
  } else {
    tryCatch(print(p), error = function(e) message("Cannot display plot."))
  }
}
```

## 1.3 Use Case Metadata

```{r uc-metadata}
# EB strategy category
uc_category <- c(
  susie_vanilla                          = "baseline",
  susine_vanilla                         = "baseline",
  susie_eb                               = "eb_variance",
  susine_eb                              = "eb_variance",
  susine_eb_mean                         = "eb_mean",
  susine_eb_mu_var                       = "eb_joint_no_ann",
  susine_eb_scale                        = "scale_only",
  susine_eb_scale_nonneg                 = "scale_only",
  susine_eb_scale_var                    = "ann_eb_joint",
  susine_eb_scale_var_nonneg             = "ann_eb_joint",
  susine_eb_penalized_scale_var          = "ann_eb_joint",
  susine_eb_penalized_scale_var_nonneg   = "ann_eb_joint",
  susine_eb_scale_var_outer              = "ann_eb_joint",
  susine_eb_scale_var_outer_nonneg       = "ann_eb_joint",
  susine_eb_clamped_scale_var            = "ann_eb_joint",
  susine_eb_clamped_scale_var_nonneg     = "ann_eb_joint",
  susine_eb_alternating_scale_var        = "ann_eb_joint",
  susine_eb_alternating_scale_var_nonneg = "ann_eb_joint",
  susine_functional_mu                   = "fixed_c_grid"
)

# Whether annotation input is used
uc_needs_annotation <- c(
  susie_vanilla                          = FALSE,
  susine_vanilla                         = FALSE,
  susie_eb                               = FALSE,
  susine_eb                              = FALSE,
  susine_eb_mean                         = FALSE,
  susine_eb_mu_var                       = FALSE,
  susine_eb_scale                        = TRUE,
  susine_eb_scale_nonneg                 = TRUE,
  susine_eb_scale_var                    = TRUE,
  susine_eb_scale_var_nonneg             = TRUE,
  susine_eb_penalized_scale_var          = TRUE,
  susine_eb_penalized_scale_var_nonneg   = TRUE,
  susine_eb_scale_var_outer              = TRUE,
  susine_eb_scale_var_outer_nonneg       = TRUE,
  susine_eb_clamped_scale_var            = TRUE,
  susine_eb_clamped_scale_var_nonneg     = TRUE,
  susine_eb_alternating_scale_var        = TRUE,
  susine_eb_alternating_scale_var_nonneg = TRUE,
  susine_functional_mu                   = TRUE
)

# Whether EB is used at all
uc_has_eb <- c(
  susie_vanilla                          = FALSE,
  susine_vanilla                         = FALSE,
  susie_eb                               = TRUE,
  susine_eb                              = TRUE,
  susine_eb_mean                         = TRUE,
  susine_eb_mu_var                       = TRUE,
  susine_eb_scale                        = TRUE,
  susine_eb_scale_nonneg                 = TRUE,
  susine_eb_scale_var                    = TRUE,
  susine_eb_scale_var_nonneg             = TRUE,
  susine_eb_penalized_scale_var          = TRUE,
  susine_eb_penalized_scale_var_nonneg   = TRUE,
  susine_eb_scale_var_outer              = TRUE,
  susine_eb_scale_var_outer_nonneg       = TRUE,
  susine_eb_clamped_scale_var            = TRUE,
  susine_eb_clamped_scale_var_nonneg     = TRUE,
  susine_eb_alternating_scale_var        = TRUE,
  susine_eb_alternating_scale_var_nonneg = TRUE,
  susine_functional_mu                   = FALSE
)

# What EB optimizes: none / var / mu / both
uc_eb_target <- c(
  susie_vanilla                          = "none",
  susine_vanilla                         = "none",
  susie_eb                               = "var",
  susine_eb                              = "var",
  susine_eb_mean                         = "mu",
  susine_eb_mu_var                       = "both",
  susine_eb_scale                        = "mu",
  susine_eb_scale_nonneg                 = "mu",
  susine_eb_scale_var                    = "both",
  susine_eb_scale_var_nonneg             = "both",
  susine_eb_penalized_scale_var          = "both",
  susine_eb_penalized_scale_var_nonneg   = "both",
  susine_eb_scale_var_outer              = "both",
  susine_eb_scale_var_outer_nonneg       = "both",
  susine_eb_clamped_scale_var            = "both",
  susine_eb_clamped_scale_var_nonneg     = "both",
  susine_eb_alternating_scale_var        = "both",
  susine_eb_alternating_scale_var_nonneg = "both",
  susine_functional_mu                   = "none"
)

# Short display labels
uc_short_label <- c(
  susie_vanilla                          = "SuSiE",
  susine_vanilla                         = "SuSiNE",
  susie_eb                               = "SuSiE+EB_var",
  susine_eb                              = "SuSiNE+EB_var",
  susine_eb_mean                         = "EB_mean",
  susine_eb_mu_var                       = "EB_mu+var",
  susine_eb_scale                        = "EB_scale",
  susine_eb_scale_nonneg                 = "EB_scale+",
  susine_eb_scale_var                    = "scale_var",
  susine_eb_scale_var_nonneg             = "scale_var+",
  susine_eb_penalized_scale_var          = "pen_sv",
  susine_eb_penalized_scale_var_nonneg   = "pen_sv+",
  susine_eb_scale_var_outer              = "outer_sv",
  susine_eb_scale_var_outer_nonneg       = "outer_sv+",
  susine_eb_clamped_scale_var            = "clamp_sv",
  susine_eb_clamped_scale_var_nonneg     = "clamp_sv+",
  susine_eb_alternating_scale_var        = "alt_sv",
  susine_eb_alternating_scale_var_nonneg = "alt_sv+",
  susine_functional_mu                   = "fixed_c_grid"
)

# Category-level palette
category_palette <- c(
  baseline        = "#636363",
  eb_variance     = "#2171b5",
  eb_mean         = "#74c476",
  eb_joint_no_ann = "#238b45",
  scale_only      = "#fd8d3c",
  ann_eb_joint    = "#cb181d",
  fixed_c_grid    = "#984ea3"
)

arch_labels <- c(
  susie2_sparse     = "SuSiE2-Sparse",
  susie2_oligogenic = "SuSiE2-Oligogenic"
)
```

## 1.4 Plot Helpers

```{r plot-helpers}
# Shared ggplot theme for SuSiE2-style panels
theme_susie2 <- function() {
  theme_bw() +
    theme(
      legend.position = "bottom",
      strip.background = element_rect(fill = "gray92"),
      strip.text = element_text(size = 9),
      panel.grid.minor = element_blank()
    )
}

# Extract TPR at specified FPR thresholds from a compute_power_fdr_curves() output.
# Uses TRUE false positive rate (FP / total_noncausal), NOT false discovery rate.
# Returns one row per group_key with tpr_fpr05 and tpr_fpr10 columns.
tpr_from_pfd <- function(pfd_tbl) {
  group_cols <- setdiff(names(pfd_tbl),
    c("pip_threshold", "TP", "FP", "power", "fdr", "precision",
      "n_causal_at_bucket", "n_noncausal_at_bucket",
      "total_causal", "total_noncausal", "FN", "TN", "recall", "fpr"))
  pfd_tbl %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_cols))) %>%
    dplyr::arrange(.data$fpr, .by_group = TRUE) %>%
    dplyr::summarise(
      tpr_fpr05 = {
        idx <- which(.data$fpr >= 0.05)[1L]
        if (is.na(idx)) NA_real_ else .data$power[idx]
      },
      tpr_fpr10 = {
        idx <- which(.data$fpr >= 0.10)[1L]
        if (is.na(idx)) NA_real_ else .data$power[idx]
      },
      .groups = "drop"
    )
}

# Add use-case metadata columns
add_uc_cols <- function(df) {
  df %>%
    dplyr::mutate(
      eb_category      = dplyr::recode(.data$use_case_id, !!!uc_category,    .default = "other"),
      short_label      = dplyr::recode(.data$use_case_id, !!!uc_short_label, .default = .data$use_case_id),
      needs_annotation = uc_needs_annotation[.data$use_case_id],
      has_eb           = uc_has_eb[.data$use_case_id],
      nonneg           = stringr::str_ends(.data$use_case_id, "_nonneg"),
      method_base      = stringr::str_remove(.data$use_case_id, "_nonneg$")
    )
}

# Produce three SuSiE2-style CS panels (power, FDR, hg2) for a given subset.
# x_var: column for x-axis (e.g. "p_star_lbl" or "short_label")
# color_var: column for color (e.g. "sigma_lbl" or "short_label")
# title_prefix: string prepended to each panel title
plot_cs_panels <- function(cs_sub, x_var, color_var, facet_var = "architecture",
                            title_prefix = "", is_discrete_x = TRUE) {
  if (is.null(cs_sub) || nrow(cs_sub) == 0) {
    message("plot_cs_panels: empty data — skipping")
    return(invisible(NULL))
  }
  mk <- function(y_var, y_label, add_hline = FALSE, hline_y = NULL) {
    p <- ggplot(cs_sub, aes(x = .data[[x_var]], y = .data[[y_var]],
                             color = .data[[color_var]],
                             group = .data[[color_var]])) +
      geom_point(size = 2) + geom_line() +
      facet_wrap(as.formula(paste("~", facet_var)),
                 labeller = labeller(.cols = arch_labels)) +
      labs(title = paste0(title_prefix, y_label),
           x = x_var, y = y_label, color = color_var) +
      theme_susie2()
    if (is_discrete_x) {
      p <- p + scale_x_discrete()
    }
    if (add_hline && !is.null(hline_y)) {
      p <- p + geom_hline(yintercept = hline_y, linetype = "dashed", color = "red", linewidth = 0.7)
    }
    p
  }
  list(
    cs_power = mk("mean_cs_power", "95% CS Power"),
    cs_fdr   = mk("mean_cs_fdr",   "95% CS FDR",  add_hline = TRUE, hline_y = 0.05),
    hg2      = mk("mean_hg2",      "hg2 (estimated variance explained)")
  )
}

# TPR vs FPR curve for one or two xlim ranges.
# Uses TRUE false positive rate (FP / total_noncausal) on x-axis.
plot_tpr_fpr <- function(pfd_sub, color_var, title = "Variant-level TPR vs FPR",
                          facet_var = "architecture") {
  if (is.null(pfd_sub) || nrow(pfd_sub) == 0) {
    message("plot_tpr_fpr: empty data — skipping")
    return(invisible(NULL))
  }
  base_p <- ggplot(pfd_sub, aes(x = fpr, y = power, color = .data[[color_var]])) +
    geom_line() +
    geom_vline(xintercept = 0.05, linetype = "dashed", color = "red", linewidth = 0.5) +
    facet_wrap(as.formula(paste("~", facet_var)),
               labeller = labeller(.cols = arch_labels)) +
    scale_x_continuous(expand = expansion(c(0, 0.02))) +
    labs(x = "Variant-level FPR", y = "TPR (Power)", color = color_var, title = title) +
    theme_susie2()
  p_narrow <- base_p + coord_cartesian(xlim = c(0, 0.1))
  p_wide   <- base_p + coord_cartesian(xlim = c(0, 0.5))
  list(narrow = p_narrow, wide = p_wide)
}

# Plot tier CS panels for oligogenic (CS power/FDR vs n_top)
plot_tier_cs <- function(tier_sub, color_var, title_prefix = "") {
  if (is.null(tier_sub) || nrow(tier_sub) == 0) {
    message("plot_tier_cs: empty data — skipping")
    return(invisible(NULL))
  }
  p_power <- ggplot(tier_sub, aes(x = factor(.data$n_top), y = mean_cs_power,
                                   color = .data[[color_var]],
                                   group = .data[[color_var]])) +
    geom_point(size = 2.5) + geom_line() +
    labs(title = paste0(title_prefix, "Oligogenic CS Power vs top-N causal"),
         x = "Top-N causal variants", y = "95% CS Power", color = color_var) +
    theme_susie2()

  p_fdr <- ggplot(tier_sub, aes(x = factor(.data$n_top), y = mean_cs_fdr,
                                 color = .data[[color_var]],
                                 group = .data[[color_var]])) +
    geom_point(size = 2.5) + geom_line() +
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red") +
    labs(title = paste0(title_prefix, "Oligogenic CS FDR vs top-N causal"),
         x = "Top-N causal variants", y = "95% CS FDR", color = color_var) +
    theme_susie2()

  list(cs_power = p_power, cs_fdr = p_fdr)
}
```

---

# Section 2: Load Data

```{r load-data, message=FALSE}
agg <- paths$aggregated_dir

conf_path        <- file.path(agg, "confusion_matrix_detailed.csv")
model_path       <- file.path(agg, "model_metrics.csv")
prior_diag_path  <- file.path(agg, "prior_diagnostics.csv")
tier_cs_path     <- file.path(agg, "tier_cs_metrics.csv")

if (!file.exists(conf_path))  stop("Missing confusion_matrix_detailed.csv")
if (!file.exists(model_path)) stop("Missing model_metrics.csv")

conf_raw      <- readr::read_csv(conf_path,  show_col_types = FALSE)
model_metrics <- readr::read_csv(model_path, show_col_types = FALSE)

prior_diagnostics <- if (file.exists(prior_diag_path)) {
  readr::read_csv(prior_diag_path, show_col_types = FALSE)
} else {
  message("prior_diagnostics.csv not found — Section 9 will be skipped.")
  NULL
}

tier_cs_metrics <- if (file.exists(tier_cs_path)) {
  readr::read_csv(tier_cs_path, show_col_types = FALSE) %>% add_uc_cols()
} else {
  message("tier_cs_metrics.csv not found — oligogenic top-N panels will be skipped.")
  NULL
}

cat("Confusion rows:      ", nrow(conf_raw), "\n")
cat("Model metric rows:   ", nrow(model_metrics), "\n")
```

```{r peek-columns}
cat("--- confusion_matrix_detailed columns ---\n")
print(names(conf_raw))
if ("variant_type" %in% names(conf_raw)) {
  cat("\nvariant_type counts:\n")
  print(table(conf_raw$variant_type, useNA = "ifany"))
}
cat("\n--- model_metrics columns ---\n")
print(names(model_metrics))
```

---

# Section 3: Data Preparation

## 3.1 Filter to susie2_* architectures + real fits only

```{r filter-base}
TARGET_ARCHS <- c("susie2_sparse", "susie2_oligogenic")

conf_base <- conf_raw %>%
  dplyr::filter(
    .data$variant_type == "base",
    .data$architecture %in% TARGET_ARCHS
  ) %>%
  add_uc_cols()

cat("Base fit confusion rows (susie2_* only):", nrow(conf_base), "\n")
cat("use_case_ids present:\n")
print(sort(unique(conf_base$use_case_id)))
cat("architectures:\n")
print(table(conf_base$architecture))
```

## 3.2 Best-c for `susine_functional_mu`

```{r best-c-lookup}
best_c_keys <- intersect(
  c("matrix_id", "architecture", "annotation_r2", "inflate_match",
    "sigma_0_2_scalar", "y_noise", "p_star", "L"),
  names(model_metrics)
)

if ("c_value" %in% names(model_metrics) && "elbo_final" %in% names(model_metrics)) {
  best_c_lookup <- model_metrics %>%
    dplyr::filter(
      .data$use_case_id == "susine_functional_mu",
      .data$filtering   == "purity_filtered",
      .data$architecture %in% TARGET_ARCHS
    ) %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(c(best_c_keys, "c_value")))) %>%
    dplyr::summarise(mean_elbo = mean(.data$elbo_final, na.rm = TRUE), .groups = "drop") %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(best_c_keys))) %>%
    dplyr::slice_max(.data$mean_elbo, n = 1, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::select(-"mean_elbo")
  cat("best_c_lookup rows:", nrow(best_c_lookup), "\n")
  if (nrow(best_c_lookup) > 0) {
    cat("Best c_value distribution:\n")
    print(table(best_c_lookup$c_value, useNA = "ifany"))
  }
} else {
  message("c_value or elbo_final not in model_metrics — susine_functional_mu uses all c values.")
  best_c_lookup <- NULL
}
```

## 3.3 Build `conf_fair`

```{r conf-fair}
conf_functional_best <- conf_base %>%
  dplyr::filter(.data$use_case_id == "susine_functional_mu")

if (!is.null(best_c_lookup) && "c_value" %in% names(conf_base)) {
  conf_join_keys <- intersect(names(best_c_lookup), names(conf_base))
  conf_join_keys <- conf_join_keys[conf_join_keys != "c_value"]
  all_join_keys  <- c(conf_join_keys, "c_value")
  conf_functional_best <- conf_functional_best %>%
    dplyr::semi_join(best_c_lookup, by = all_join_keys)
}

conf_fair <- conf_base %>%
  dplyr::filter(.data$use_case_id != "susine_functional_mu") %>%
  dplyr::bind_rows(conf_functional_best)

cat("conf_fair rows:", nrow(conf_fair), "\n")
```

## 3.4 Compute pooled power-FDR curves

```{r pfd-curves}
# One curve per (use_case_id, architecture): pools all datasets + annotation conditions
overall_pfd <- test_susine::compute_power_fdr_curves(
  conf_fair,
  group_vars = c("use_case_id", "architecture")
) %>% add_uc_cols()

# Also stratified by annotation_r2 + inflate_match (for annotation sections)
if (all(c("annotation_r2", "inflate_match") %in% names(conf_fair))) {
  annot_pfd <- test_susine::compute_power_fdr_curves(
    conf_fair %>% dplyr::filter(!is.na(.data$annotation_r2)),
    group_vars = c("use_case_id", "architecture", "annotation_r2", "inflate_match")
  ) %>% add_uc_cols()
} else {
  annot_pfd <- NULL
}

# Oligogenic top-8 curve (uses top_n_causal == 8 rows if available)
if ("top_n_causal" %in% names(conf_fair)) {
  oligo_top8_pfd <- test_susine::compute_power_fdr_curves(
    conf_fair %>% dplyr::filter(
      .data$architecture == "susie2_oligogenic",
      !is.na(.data$top_n_causal) & .data$top_n_causal == 8L
    ),
    group_vars = c("use_case_id", "architecture")
  ) %>% add_uc_cols()
} else {
  oligo_top8_pfd <- NULL
}

cat("overall_pfd rows:", nrow(overall_pfd), "\n")
cat("fpr column present:", "fpr" %in% names(overall_pfd), "\n")
```

## 3.5 Extract TPR @ FPR thresholds

```{r tpr-overall}
tpr_overall <- tpr_from_pfd(overall_pfd)
cat("tpr_overall rows:", nrow(tpr_overall), "\n")
cat("tpr_fpr10 range:", range(tpr_overall$tpr_fpr10, na.rm = TRUE), "\n")
```

## 3.6 CS metrics from model_metrics

```{r cs-metrics}
mm_base <- model_metrics %>%
  dplyr::filter(
    .data$filtering   == "purity_filtered",
    .data$architecture %in% TARGET_ARCHS
  )

# Best-c filter for susine_functional_mu in model_metrics
mm_functional_best <- mm_base %>%
  dplyr::filter(.data$use_case_id == "susine_functional_mu")

if (!is.null(best_c_lookup) && "c_value" %in% names(mm_base)) {
  mm_join_keys <- intersect(names(best_c_lookup), names(mm_base))
  mm_all_join  <- c(mm_join_keys[mm_join_keys != "c_value"], "c_value")
  mm_functional_best <- mm_functional_best %>%
    dplyr::semi_join(best_c_lookup, by = intersect(mm_all_join, names(.)))
}

mm_fair <- mm_base %>%
  dplyr::filter(.data$use_case_id != "susine_functional_mu") %>%
  dplyr::bind_rows(mm_functional_best) %>%
  add_uc_cols() %>%
  dplyr::mutate(cs_fdr = 1 - .data$mean_coverage)

# Per-(use_case, architecture, p_star) CS summary
# p_star varies for sparse; is NA for oligogenic (use NA as a dummy level)
cs_metrics <- mm_fair %>%
  dplyr::mutate(
    p_star_lbl = dplyr::if_else(!is.na(.data$p_star), as.character(.data$p_star), "fixed"),
    sigma_lbl  = dplyr::if_else(!is.na(.data$sigma_0_2_scalar),
                                as.character(.data$sigma_0_2_scalar), "EB")
  ) %>%
  dplyr::group_by(.data$use_case_id, .data$short_label, .data$eb_category,
                  .data$has_eb, .data$needs_annotation,
                  .data$architecture, .data$p_star_lbl, .data$sigma_lbl) %>%
  dplyr::summarise(
    mean_cs_power = mean(.data$power,   na.rm = TRUE),
    mean_cs_fdr   = mean(.data$cs_fdr,  na.rm = TRUE),
    mean_hg2      = mean(.data$hg2,     na.rm = TRUE),
    n_fits        = dplyr::n(),
    .groups = "drop"
  )

cat("cs_metrics rows:", nrow(cs_metrics), "\n")
```

## 3.7 Per-run ranks (for master table)

```{r ranks}
rank_id_col <- if ("dataset_bundle_id" %in% names(mm_fair)) "dataset_bundle_id" else "matrix_id"

ranks_tbl <- mm_fair %>%
  dplyr::group_by(.data[[rank_id_col]], .data$architecture) %>%
  dplyr::mutate(
    auprc_rank  = rank(-.data$AUPRC,        ties.method = "min"),
    tpr10_rank  = if ("tpr_at_fpr_10" %in% names(.))
                    rank(-.data$tpr_at_fpr_10, ties.method = "min")
                  else NA_integer_,
    n_competing = dplyr::n()
  ) %>%
  dplyr::ungroup()

cat("Ranked rows:", nrow(ranks_tbl), "\n")
```

## 3.8 Tier CS summary (computed once, filtered per section)

```{r tier-summary}
tier_summary <- if (!is.null(tier_cs_metrics) && nrow(tier_cs_metrics) > 0) {
  tier_cs_metrics %>%
    dplyr::filter(.data$filtering == "purity_filtered") %>%
    dplyr::group_by(.data$use_case_id, .data$short_label, .data$eb_category,
                    .data$architecture, .data$n_top) %>%
    dplyr::summarise(
      mean_cs_power = mean(.data$cs_power, na.rm = TRUE),
      mean_cs_fdr   = mean(.data$cs_fdr,   na.rm = TRUE),
      .groups = "drop"
    )
} else {
  NULL
}

if (!is.null(tier_summary)) {
  cat("tier_summary rows:", nrow(tier_summary), "\n")
  cat("n_top values:", sort(unique(tier_summary$n_top)), "\n")
} else {
  message("No tier_cs_metrics — oligogenic top-N panels will be skipped in all sections.")
}
```

---

# Section 4: Baseline — No Annotations, No EB

Uses `susine_vanilla` only (susie and susine replicate; color by sigma_0_2_scalar).

## 4.1 Sparse high-level panels (SuSiE2 Fig 1B-E)

```{r s4-prep}
s4_cs <- cs_metrics %>%
  dplyr::filter(.data$use_case_id == "susine_vanilla")

s4_pfd <- overall_pfd %>%
  dplyr::filter(.data$use_case_id == "susine_vanilla")

cat("S4 cs rows:", nrow(s4_cs), "\n")
```

```{r s4-cs-panels, message=FALSE, fig.width=12, fig.height=5}
panels_s4 <- plot_cs_panels(
  cs_sub       = s4_cs,
  x_var        = "p_star_lbl",
  color_var    = "sigma_lbl",
  facet_var    = "architecture",
  title_prefix = "Baseline (susine_vanilla): "
)
for (nm in names(panels_s4)) {
  if (!is.null(panels_s4[[nm]])) {
    save_or_show(panels_s4[[nm]], sprintf("s4_%s.png", nm),
                 width = 12, height = 5, subdir = "baseline")
  }
}
```

```{r s4-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
curves_s4 <- plot_tpr_fpr(s4_pfd, color_var = "sigma_lbl",
                           title = "Baseline (susine_vanilla): Variant-level TPR vs FPR")
if (!is.null(curves_s4$narrow)) {
  save_or_show(curves_s4$narrow, "s4_tpr_fpr_narrow.png", width = 10, height = 5, subdir = "baseline")
  save_or_show(curves_s4$wide,   "s4_tpr_fpr_wide.png",   width = 10, height = 5, subdir = "baseline")
}
```

## 4.2 Oligogenic high-level panels (SuSiE2 Fig 1F-I)

```{r s4-oligo-tier, message=FALSE, fig.width=10, fig.height=6}
if (!is.null(tier_summary)) {
  s4_tier <- tier_summary %>%
    dplyr::filter(.data$use_case_id == "susine_vanilla",
                  .data$architecture == "susie2_oligogenic")

  tier_panels_s4 <- plot_tier_cs(s4_tier, color_var = "short_label",
                                  title_prefix = "Baseline: ")
  if (!is.null(tier_panels_s4)) {
    for (nm in names(tier_panels_s4)) {
      save_or_show(tier_panels_s4[[nm]], sprintf("s4_oligo_%s.png", nm),
                   width = 10, height = 6, subdir = "baseline")
    }
  }
} else {
  message("Tier CS data not available — oligogenic CS panels skipped for baseline.")
}
```

```{r s4-oligo-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
if (!is.null(oligo_top8_pfd)) {
  s4_top8 <- oligo_top8_pfd %>%
    dplyr::filter(.data$use_case_id == "susine_vanilla")

  if (nrow(s4_top8) > 0) {
    curves_s4_top8 <- plot_tpr_fpr(
      s4_top8, color_var = "short_label",
      title = "Baseline: Oligogenic TPR vs FPR (top-8 causal)"
    )
    if (!is.null(curves_s4_top8$narrow)) {
      save_or_show(curves_s4_top8$narrow, "s4_oligo_top8_tpr_fpr_narrow.png",
                   width = 10, height = 5, subdir = "baseline")
      save_or_show(curves_s4_top8$wide,   "s4_oligo_top8_tpr_fpr_wide.png",
                   width = 10, height = 5, subdir = "baseline")
    }
  }
} else {
  message("Top-8 confusion bins not available — oligogenic TPR/FPR skipped for baseline.")
}
```

---

# Section 5: Annotations, No EB (`susine_functional_mu`)

Uses all c_values (not best-c restricted) to show how c affects performance.

## 5.1 Sparse high-level panels (best-c, lumped over annotation settings)

```{r s5-prep}
s5_cs <- cs_metrics %>%
  dplyr::filter(.data$use_case_id == "susine_functional_mu")

s5_pfd_fair <- overall_pfd %>%
  dplyr::filter(.data$use_case_id == "susine_functional_mu")
```

```{r s5-cs-panels, message=FALSE, fig.width=12, fig.height=5}
panels_s5 <- plot_cs_panels(
  cs_sub       = s5_cs,
  x_var        = "p_star_lbl",
  color_var    = "sigma_lbl",
  facet_var    = "architecture",
  title_prefix = "Annot, no EB (susine_functional_mu): "
)
for (nm in names(panels_s5)) {
  if (!is.null(panels_s5[[nm]])) {
    save_or_show(panels_s5[[nm]], sprintf("s5_%s.png", nm),
                 width = 12, height = 5, subdir = "annot_noeb")
  }
}
```

```{r s5-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
curves_s5 <- plot_tpr_fpr(s5_pfd_fair, color_var = "short_label",
                           title = "Annot no EB (best-c): Variant-level TPR vs FPR")
if (!is.null(curves_s5$narrow)) {
  save_or_show(curves_s5$narrow, "s5_tpr_fpr_narrow.png", width = 10, height = 5, subdir = "annot_noeb")
  save_or_show(curves_s5$wide,   "s5_tpr_fpr_wide.png",   width = 10, height = 5, subdir = "annot_noeb")
}
```

## 5.2 Oligogenic high-level panels

```{r s5-oligo-tier, message=FALSE, fig.width=10, fig.height=6}
if (!is.null(tier_summary)) {
  s5_tier <- tier_summary %>%
    dplyr::filter(.data$use_case_id == "susine_functional_mu",
                  .data$architecture == "susie2_oligogenic")

  tier_panels_s5 <- plot_tier_cs(s5_tier, color_var = "short_label",
                                  title_prefix = "Annot no EB: ")
  if (!is.null(tier_panels_s5)) {
    for (nm in names(tier_panels_s5)) {
      save_or_show(tier_panels_s5[[nm]], sprintf("s5_oligo_%s.png", nm),
                   width = 10, height = 6, subdir = "annot_noeb")
    }
  }
}
```

```{r s5-oligo-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
if (!is.null(oligo_top8_pfd)) {
  s5_top8 <- oligo_top8_pfd %>%
    dplyr::filter(.data$use_case_id == "susine_functional_mu")

  if (nrow(s5_top8) > 0) {
    curves_s5_top8 <- plot_tpr_fpr(
      s5_top8, color_var = "short_label",
      title = "Annot no EB: Oligogenic TPR vs FPR (top-8 causal)"
    )
    if (!is.null(curves_s5_top8$narrow)) {
      save_or_show(curves_s5_top8$narrow, "s5_oligo_top8_tpr_fpr_narrow.png",
                   width = 10, height = 5, subdir = "annot_noeb")
      save_or_show(curves_s5_top8$wide,   "s5_oligo_top8_tpr_fpr_wide.png",
                   width = 10, height = 5, subdir = "annot_noeb")
    }
  }
}
```

## 5.3 Overall TPR @ FPR=0.05 vs c_value (faceted by annotation_r2 x inflate_match)

```{r s5-tpr-vs-c, message=FALSE, fig.width=14, fig.height=8}
# Use conf_base (all c values) for this section, not conf_fair
s5_pfd_all_c <- if ("c_value" %in% names(conf_base)) {
  test_susine::compute_power_fdr_curves(
    conf_base %>% dplyr::filter(.data$use_case_id == "susine_functional_mu"),
    group_vars = intersect(
      c("use_case_id", "architecture", "annotation_r2", "inflate_match", "c_value"),
      names(conf_base)
    )
  ) %>% add_uc_cols()
} else {
  test_susine::compute_power_fdr_curves(
    conf_base %>% dplyr::filter(.data$use_case_id == "susine_functional_mu"),
    group_vars = intersect(c("use_case_id", "architecture"), names(conf_base))
  ) %>% add_uc_cols()
}

if (!is.null(s5_pfd_all_c) && "c_value" %in% names(s5_pfd_all_c)) {
  tpr_s5 <- tpr_from_pfd(s5_pfd_all_c)

  if (all(c("annotation_r2", "inflate_match") %in% names(tpr_s5))) {
    for (arch in unique(tpr_s5$architecture)) {
      p_s5 <- tpr_s5 %>%
        dplyr::filter(.data$architecture == arch) %>%
        ggplot(aes(x = .data$c_value, y = tpr_fpr05, color = use_case_id)) +
        geom_point(size = 2) + geom_line() +
        facet_grid(annotation_r2 ~ inflate_match, labeller = label_both) +
        labs(
          title = sprintf("Annot, no EB: Overall TPR @ FPR=0.05 vs c_value [%s]", arch),
          x = "c_value (fixed annotation scale)", y = "TPR @ FPR=0.05", color = "Use case"
        ) +
        theme_susie2()
      save_or_show(p_s5, sprintf("s5_tpr_fpr05_vs_c_%s.png", arch),
                   width = 14, height = 8, subdir = "annot_noeb")
    }
  }
}
```

---

# Section 6: EB, No Annotations

Comparison: `susie_vanilla`, `susine_vanilla`, `susie_eb`, `susine_eb`,
`susine_eb_mean`, `susine_eb_mu_var`.

## 6.1 Sparse high-level panels

```{r s6-prep}
S6_USE_CASES <- c("susie_vanilla", "susine_vanilla", "susie_eb",
                  "susine_eb", "susine_eb_mean", "susine_eb_mu_var")

s6_cs <- cs_metrics %>%
  dplyr::filter(.data$use_case_id %in% S6_USE_CASES)

s6_pfd <- overall_pfd %>%
  dplyr::filter(.data$use_case_id %in% S6_USE_CASES)

cat("S6 use cases present:", paste(sort(unique(s6_cs$use_case_id)), collapse = ", "), "\n")
```

```{r s6-cs-panels, message=FALSE, fig.width=12, fig.height=5}
panels_s6 <- plot_cs_panels(
  cs_sub       = s6_cs,
  x_var        = "p_star_lbl",
  color_var    = "short_label",
  facet_var    = "architecture",
  title_prefix = "EB no annot: "
)
for (nm in names(panels_s6)) {
  if (!is.null(panels_s6[[nm]])) {
    save_or_show(panels_s6[[nm]], sprintf("s6_%s.png", nm),
                 width = 12, height = 5, subdir = "eb_noannot")
  }
}
```

```{r s6-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
curves_s6 <- plot_tpr_fpr(s6_pfd, color_var = "short_label",
                           title = "EB no annot: Variant-level TPR vs FPR")
if (!is.null(curves_s6$narrow)) {
  save_or_show(curves_s6$narrow, "s6_tpr_fpr_narrow.png", width = 10, height = 5, subdir = "eb_noannot")
  save_or_show(curves_s6$wide,   "s6_tpr_fpr_wide.png",   width = 10, height = 5, subdir = "eb_noannot")
}
```

## 6.2 Oligogenic high-level panels

```{r s6-oligo-tier, message=FALSE, fig.width=10, fig.height=6}
if (!is.null(tier_summary)) {
  s6_tier <- tier_summary %>%
    dplyr::filter(.data$use_case_id %in% S6_USE_CASES,
                  .data$architecture == "susie2_oligogenic")

  tier_panels_s6 <- plot_tier_cs(s6_tier, color_var = "short_label",
                                  title_prefix = "EB no annot: ")
  if (!is.null(tier_panels_s6)) {
    for (nm in names(tier_panels_s6)) {
      save_or_show(tier_panels_s6[[nm]], sprintf("s6_oligo_%s.png", nm),
                   width = 10, height = 6, subdir = "eb_noannot")
    }
  }
}
```

```{r s6-oligo-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
if (!is.null(oligo_top8_pfd)) {
  s6_top8 <- oligo_top8_pfd %>%
    dplyr::filter(.data$use_case_id %in% S6_USE_CASES)

  if (nrow(s6_top8) > 0) {
    curves_s6_top8 <- plot_tpr_fpr(
      s6_top8, color_var = "short_label",
      title = "EB no annot: Oligogenic TPR vs FPR (top-8 causal)"
    )
    if (!is.null(curves_s6_top8$narrow)) {
      save_or_show(curves_s6_top8$narrow, "s6_oligo_top8_tpr_fpr_narrow.png",
                   width = 10, height = 5, subdir = "eb_noannot")
      save_or_show(curves_s6_top8$wide,   "s6_oligo_top8_tpr_fpr_wide.png",
                   width = 10, height = 5, subdir = "eb_noannot")
    }
  }
}
```

---

# Section 7: EB with Annotations

All annotation-based EB methods plus `susine_functional_mu` (best-c) as baseline.

## 7.1 Sparse high-level panels (lumped over annotation conditions)

```{r s7-prep}
S7_USE_CASES <- c(
  "susine_functional_mu",
  "susine_eb_scale", "susine_eb_scale_nonneg",
  "susine_eb_scale_var", "susine_eb_scale_var_nonneg",
  "susine_eb_penalized_scale_var", "susine_eb_penalized_scale_var_nonneg",
  "susine_eb_scale_var_outer", "susine_eb_scale_var_outer_nonneg",
  "susine_eb_clamped_scale_var", "susine_eb_clamped_scale_var_nonneg",
  "susine_eb_alternating_scale_var", "susine_eb_alternating_scale_var_nonneg"
)

s7_cs <- cs_metrics %>%
  dplyr::filter(.data$use_case_id %in% S7_USE_CASES)

s7_pfd <- overall_pfd %>%
  dplyr::filter(.data$use_case_id %in% S7_USE_CASES)

s7_annot_pfd <- if (!is.null(annot_pfd)) {
  annot_pfd %>% dplyr::filter(.data$use_case_id %in% S7_USE_CASES)
} else {
  NULL
}

cat("S7 use cases present:", paste(sort(unique(s7_cs$use_case_id)), collapse = ", "), "\n")
```

```{r s7-cs-panels, message=FALSE, fig.width=14, fig.height=5}
panels_s7 <- plot_cs_panels(
  cs_sub       = s7_cs,
  x_var        = "p_star_lbl",
  color_var    = "short_label",
  facet_var    = "architecture",
  title_prefix = "EB with annot: "
)
for (nm in names(panels_s7)) {
  if (!is.null(panels_s7[[nm]])) {
    save_or_show(panels_s7[[nm]], sprintf("s7_%s.png", nm),
                 width = 14, height = 5, subdir = "eb_annot")
  }
}
```

```{r s7-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
curves_s7 <- plot_tpr_fpr(s7_pfd, color_var = "short_label",
                           title = "EB with annot: Variant-level TPR vs FPR")
if (!is.null(curves_s7$narrow)) {
  save_or_show(curves_s7$narrow, "s7_tpr_fpr_narrow.png", width = 10, height = 5, subdir = "eb_annot")
  save_or_show(curves_s7$wide,   "s7_tpr_fpr_wide.png",   width = 10, height = 5, subdir = "eb_annot")
}
```

## 7.2 Oligogenic high-level panels

```{r s7-oligo-tier, message=FALSE, fig.width=12, fig.height=6}
if (!is.null(tier_summary)) {
  s7_tier <- tier_summary %>%
    dplyr::filter(.data$use_case_id %in% S7_USE_CASES,
                  .data$architecture == "susie2_oligogenic")

  tier_panels_s7 <- plot_tier_cs(s7_tier, color_var = "short_label",
                                  title_prefix = "EB with annot: ")
  if (!is.null(tier_panels_s7)) {
    for (nm in names(tier_panels_s7)) {
      save_or_show(tier_panels_s7[[nm]], sprintf("s7_oligo_%s.png", nm),
                   width = 12, height = 6, subdir = "eb_annot")
    }
  }
}
```

```{r s7-oligo-tpr-fpr, message=FALSE, fig.width=10, fig.height=5}
if (!is.null(oligo_top8_pfd)) {
  s7_top8 <- oligo_top8_pfd %>%
    dplyr::filter(.data$use_case_id %in% S7_USE_CASES)

  if (nrow(s7_top8) > 0) {
    curves_s7_top8 <- plot_tpr_fpr(
      s7_top8, color_var = "short_label",
      title = "EB with annot: Oligogenic TPR vs FPR (top-8 causal)"
    )
    if (!is.null(curves_s7_top8$narrow)) {
      save_or_show(curves_s7_top8$narrow, "s7_oligo_top8_tpr_fpr_narrow.png",
                   width = 10, height = 5, subdir = "eb_annot")
      save_or_show(curves_s7_top8$wide,   "s7_oligo_top8_tpr_fpr_wide.png",
                   width = 10, height = 5, subdir = "eb_annot")
    }
  }
}
```

## 7.3 Overall TPR @ FPR=0.05 vs annotation_r2 (faceted by architecture x inflate_match)

```{r s7-tpr-vs-annot, message=FALSE, fig.width=14, fig.height=8}
if (!is.null(s7_annot_pfd) && nrow(s7_annot_pfd) > 0) {
  tpr_s7 <- tpr_from_pfd(s7_annot_pfd)

  p_s7_tpr <- ggplot(tpr_s7, aes(x = annotation_r2, y = tpr_fpr05, color = short_label)) +
    geom_point(size = 2) + geom_line() +
    facet_grid(architecture ~ inflate_match,
               labeller = labeller(architecture = arch_labels, inflate_match = label_both)) +
    scale_x_continuous(breaks = sort(unique(tpr_s7$annotation_r2))) +
    labs(
      title = "EB with annot: Overall TPR @ FPR=0.05 vs annotation_r2",
      x = "Annotation R\u00b2", y = "TPR @ FPR=0.05", color = "Method"
    ) +
    theme_susie2()
  save_or_show(p_s7_tpr, "s7_tpr_fpr05_vs_annot_r2.png",
               width = 14, height = 8, subdir = "eb_annot")
}
```

## 7.4 Nonneg constraint comparison

```{r s7-nonneg, message=FALSE, fig.width=10, fig.height=6}
tpr_nonneg <- tpr_overall %>%
  dplyr::filter(.data$use_case_id %in% S7_USE_CASES,
                .data$needs_annotation) %>%
  dplyr::mutate(
    nonneg      = stringr::str_ends(.data$use_case_id, "_nonneg"),
    method_base = stringr::str_remove(.data$use_case_id, "_nonneg$")
  )

if (nrow(tpr_nonneg) > 0 && any(tpr_nonneg$nonneg)) {
  tpr_nonneg_wide <- tpr_nonneg %>%
    dplyr::select("method_base", "architecture", "nonneg", "tpr_fpr10") %>%
    tidyr::pivot_wider(names_from = "nonneg", values_from = "tpr_fpr10",
                       names_prefix = "nonneg_") %>%
    dplyr::rename(tpr_unconstrained = nonneg_FALSE, tpr_nonneg = nonneg_TRUE)

  if ("tpr_unconstrained" %in% names(tpr_nonneg_wide) &&
      "tpr_nonneg" %in% names(tpr_nonneg_wide)) {
    p_nonneg <- ggplot(tpr_nonneg_wide,
                       aes(x = tpr_unconstrained, y = tpr_nonneg, color = method_base)) +
      geom_point(size = 2.5) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
      facet_wrap(~ architecture, labeller = labeller(.cols = arch_labels)) +
      labs(
        title = "Nonneg constraint effect: TPR@FPR=0.1",
        x = "Unconstrained (TPR@FPR=0.1)", y = "Nonneg constrained (TPR@FPR=0.1)",
        color = "Method base"
      ) +
      theme_susie2()
    save_or_show(p_nonneg, "s7_nonneg_scatter.png", width = 10, height = 6, subdir = "eb_annot")
  }
}
```

---

# Section 8: Master Summary Table

One row per (use_case_id, architecture). Primary sort: overall TPR@FPR=0.1 descending.

```{r master-table}
# Overall AUPRC pooled across all datasets
overall_auprc <- test_susine::compute_auprc_from_confusion(
  conf_fair, group_vars = c("use_case_id", "architecture")
) %>% add_uc_cols() %>%
  dplyr::rename(overall_AUPRC = "AUPRC")

# Rank stats
rank_stats <- ranks_tbl %>%
  dplyr::group_by(.data$use_case_id, .data$architecture) %>%
  dplyr::summarise(
    pct_rank1_tpr10 = if (all(is.na(.data$tpr10_rank))) NA_real_
                      else mean(.data$tpr10_rank == 1L, na.rm = TRUE),
    pct_top3_tpr10  = if (all(is.na(.data$tpr10_rank))) NA_real_
                      else mean(.data$tpr10_rank <= 3L, na.rm = TRUE),
    n_fits = dplyr::n(),
    .groups = "drop"
  )

# Sigma / c fixed values (first non-NA per use_case, if applicable)
fixed_params <- mm_fair %>%
  dplyr::group_by(.data$use_case_id, .data$architecture) %>%
  dplyr::summarise(
    sigma_fixed = {
      vals <- unique(na.omit(.data$sigma_0_2_scalar))
      if (length(vals) == 1L) vals else NA_character_
    },
    c_fixed = if ("c_value" %in% names(.)) {
      vals <- unique(na.omit(.data$c_value))
      if (length(vals) == 1L) as.character(vals) else NA_character_
    } else NA_character_,
    .groups = "drop"
  )

master_table <- overall_auprc %>%
  dplyr::left_join(tpr_overall,   by = c("use_case_id", "architecture")) %>%
  dplyr::left_join(rank_stats,    by = c("use_case_id", "architecture")) %>%
  dplyr::left_join(fixed_params,  by = c("use_case_id", "architecture")) %>%
  dplyr::mutate(
    has_annotation = uc_needs_annotation[.data$use_case_id],
    has_eb         = uc_has_eb[.data$use_case_id],
    eb_type        = uc_eb_target[.data$use_case_id]
  ) %>%
  dplyr::select(
    "use_case_id", "short_label", "architecture",
    "has_annotation", "has_eb", "eb_type",
    "c_fixed", "sigma_fixed",
    "overall_AUPRC", "tpr_fpr05", "tpr_fpr10",
    "pct_rank1_tpr10", "pct_top3_tpr10",
    "n_fits"
  ) %>%
  dplyr::arrange(dplyr::desc(.data$tpr_fpr10), dplyr::desc(.data$overall_AUPRC))

readr::write_csv(master_table, file.path(plot_dirs$master, "master_summary.csv"))
cat("Wrote master_summary.csv\n")

master_table %>%
  dplyr::mutate(dplyr::across(dplyr::where(is.numeric), ~ round(.x, 3))) %>%
  knitr::kable(caption = "Master summary (ranked by overall TPR@FPR=0.1, then AUPRC)")
```

---

# Section 9: Prior Diagnostics

What do EB methods actually learn? Distributions of c_l and sigma_0_2_l per effect.

```{r prior-diag-check}
if (!is.null(prior_diagnostics) && nrow(prior_diagnostics) > 0) {
  cat("prior_diagnostics rows:", nrow(prior_diagnostics), "\n")
  cat("Columns:", paste(names(prior_diagnostics), collapse = ", "), "\n")
  cat("use_case_id counts:\n")
  print(table(prior_diagnostics$use_case_id, useNA = "ifany"))
} else {
  message("No prior_diagnostics available — skipping Section 9.")
  knitr::knit_exit()
}
```

```{r prior-diag-c, message=FALSE}
c_col   <- intersect(c("c_l", "c_value", "c"), names(prior_diagnostics))[1]
sig_col <- intersect(c("sigma_0_2_l", "sigma_0_2", "sigma_l"), names(prior_diagnostics))[1]

if (!is.na(c_col)) {
  pd <- prior_diagnostics %>%
    add_uc_cols() %>%
    dplyr::filter(!is.na(.data[[c_col]]),
                  .data$eb_category != "baseline",
                  .data$eb_category != "fixed_c_grid")

  if (nrow(pd) > 0) {
    p_c_dist <- ggplot(pd, aes(x = .data[[c_col]], fill = eb_category)) +
      geom_histogram(bins = 80, alpha = 0.8) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "gray30") +
      scale_x_continuous(
        trans = scales::pseudo_log_trans(sigma = 1, base = 10),
        breaks = c(-100, -10, -1, -0.1, 0, 0.1, 1, 10, 100),
        labels = scales::label_number(drop0trailing = TRUE)
      ) +
      scale_fill_manual(values = category_palette) +
      facet_wrap(~ short_label, scales = "free_y") +
      labs(
        title = sprintf("Learned %s distribution by EB method (pseudo-log x-axis)", c_col),
        x = sprintf("Learned c per effect [%s, pseudo-log scale, sigma=1]", c_col),
        y = "Count", fill = "Category"
      ) +
      theme_minimal() +
      theme(legend.position = "bottom", strip.text = element_text(size = 8),
            axis.text.x = element_text(angle = 30, hjust = 1))

    save_or_show(p_c_dist, "c_l_distribution_by_method.png",
                 width = 16, height = 12, subdir = "diagnostics")
  }
}
```

```{r prior-diag-sigma, message=FALSE}
if (!is.na(sig_col)) {
  pd_sig <- prior_diagnostics %>%
    add_uc_cols() %>%
    dplyr::filter(!is.na(.data[[sig_col]]),
                  .data$eb_category != "baseline") %>%
    dplyr::mutate(
      log_sigma = pmax(-10, pmin(5, log(.data[[sig_col]] + 1e-12)))
    )

  if (nrow(pd_sig) > 0) {
    p_sig_dist <- ggplot(pd_sig, aes(x = log_sigma, fill = short_label)) +
      geom_histogram(bins = 50, alpha = 0.8, show.legend = FALSE) +
      facet_wrap(~ short_label, scales = "free_y") +
      labs(
        title = sprintf("Learned %s distribution by EB method (log scale)", sig_col),
        subtitle = "log(sigma_0_2_l), floored at -10, ceilinged at 5",
        x = "log(sigma_0_2_l)  [floor = -10, ceil = 5]",
        y = "Count"
      ) +
      theme_minimal() +
      theme(strip.text = element_text(size = 8))

    save_or_show(p_sig_dist, "sigma_0_2_l_distribution_by_method.png",
                 width = 16, height = 12, subdir = "diagnostics")
  }
}
```

---

# Section 10: Output Summary

```{r output-summary}
cat("All plots saved under:", plot_base_dir, "\n\n")
purrr::walk(names(plot_dirs), function(subdir) {
  files <- list.files(plot_dirs[[subdir]], full.names = FALSE)
  if (length(files) > 0) {
    cat(sprintf("[%s] %d files\n", subdir, length(files)))
    purrr::walk(files, function(f) cat("  ", f, "\n"))
  }
})
```
