---
title: "Ensemble Scaling Experiment — Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Reads consolidated outputs from the collect workbook. Generates all
4 figure categories from `refs/ensemble_scaling_experiment_plan.md`.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_ensemble_scaling.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

# 0. Load Consolidated Data

```{r load}
# Point to job-level consolidated output from collect workbook
job_dir <- here("output", "slurm_output", "local_test")
consolidated_dir <- file.path(job_dir, "consolidated")

auprc_agg <- readr::read_csv(
  file.path(consolidated_dir, "auprc_aggregated.csv"),
  show_col_types = FALSE
)
auprc_individual <- readr::read_csv(
  file.path(consolidated_dir, "auprc_individual.csv"),
  show_col_types = FALSE
)
model_metrics <- readr::read_csv(
  file.path(consolidated_dir, "model_metrics_full.csv"),
  show_col_types = FALSE
)
confusion_bins <- readr::read_csv(
  file.path(consolidated_dir, "confusion_bins_full.csv"),
  show_col_types = FALSE
)

tier_cs_path <- file.path(consolidated_dir, "tier_cs_metrics_full.csv")
tier_cs_metrics_full <- if (file.exists(tier_cs_path)) {
  readr::read_csv(tier_cs_path, show_col_types = FALSE)
} else {
  message("tier_cs_metrics_full.csv not found — plots 1a/1b will be skipped.")
  NULL
}

# Figure output directories
fig_root <- file.path(job_dir, "figures")
fig_dirs <- c("susie2_plots", "ensemble_performance",
              "dataset_difficulty", "ensemble_scaling")
for (d in fig_dirs) {
  dir.create(file.path(fig_root, d), recursive = TRUE, showWarnings = FALSE)
}

cat("AUPRC aggregated rows:", nrow(auprc_agg), "\n")
cat("AUPRC individual rows:", nrow(auprc_individual), "\n")
cat("Figure root:", fig_root, "\n")
```

```{r baselines}
# Load all pooled AUPRC tables
auprc_pooled_agg <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_aggregated.csv"),
  show_col_types = FALSE
)
auprc_pooled_agg_by_r2 <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_agg_by_r2.csv"),
  show_col_types = FALSE
)
auprc_pooled_individual_overall <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_individual_overall.csv"),
  show_col_types = FALSE
)
auprc_pooled_individual <- readr::read_csv(
  file.path(consolidated_dir, "auprc_pooled_individual.csv"),
  show_col_types = FALSE
)

# Baseline reference: single-fit susine_vanilla, pooled across all r2
baseline_auprc <- auprc_pooled_individual_overall %>%
  dplyr::filter(spec_name == "baseline-single",
                use_case_id == "susine_vanilla") %>%
  dplyr::pull(AUPRC)

cat("Baseline pooled AUPRC:", round(baseline_auprc, 4), "\n")

# Truth-warm for OVERALL plot: pooled across all r2, no r2 in label
truth_warm_overall <- auprc_pooled_individual_overall %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, AUPRC) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~ "truth-warm EB",
      use_case_id == "susine_functional_mu" ~ "truth-warm grid (c=0.5)",
      TRUE ~ use_case_id
    ),
    annotation_r2 = NA_real_
  )

# Truth-warm for PER-R2 plots: one row per use_case x annotation_r2, r2 in label
truth_warm_by_r2 <- auprc_pooled_individual %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, annotation_r2, AUPRC) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~
        paste0("truth-warm EB (r\u00b2=", annotation_r2, ")"),
      use_case_id == "susine_functional_mu" ~
        paste0("truth-warm grid (c=0.5, r\u00b2=", annotation_r2, ")"),
      TRUE ~ use_case_id
    )
  )

cat("Truth-warm overall:\n"); print(truth_warm_overall)
cat("Truth-warm by r2:\n"); print(truth_warm_by_r2)

# Load TPR@0.05 pooled tables
tpr05_pooled_agg <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_aggregated.csv"),
  show_col_types = FALSE
)
tpr05_pooled_agg_by_r2 <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_agg_by_r2.csv"),
  show_col_types = FALSE
)
tpr05_pooled_individual_overall <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_individual_overall.csv"),
  show_col_types = FALSE
)
tpr05_pooled_individual <- readr::read_csv(
  file.path(consolidated_dir, "tpr05_pooled_individual.csv"),
  show_col_types = FALSE
)

# Baseline TPR@0.05 (single-fit susine_vanilla, pooled across all r2)
baseline_tpr05 <- tpr05_pooled_individual_overall %>%
  dplyr::filter(spec_name == "baseline-single",
                use_case_id == "susine_vanilla") %>%
  dplyr::pull(tpr_fpr05)
cat("Baseline pooled TPR@0.05:", round(baseline_tpr05, 4), "\n")

# Truth-warm TPR@0.05 for OVERALL plot
truth_warm_tpr05_overall <- tpr05_pooled_individual_overall %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, tpr_fpr05) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~ "truth-warm EB",
      use_case_id == "susine_functional_mu" ~ "truth-warm grid (c=0.5)",
      TRUE ~ use_case_id
    ),
    annotation_r2 = NA_real_
  )

# Truth-warm TPR@0.05 for PER-R2 plots
truth_warm_tpr05_by_r2 <- tpr05_pooled_individual %>%
  dplyr::filter(spec_name == "truth-warm") %>%
  dplyr::select(use_case_id, annotation_r2, tpr_fpr05) %>%
  dplyr::mutate(
    label = dplyr::case_when(
      use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
      use_case_id == "susine_eb_clamped_scale_var_nonneg" ~
        paste0("truth-warm EB (r\u00b2=", annotation_r2, ")"),
      use_case_id == "susine_functional_mu" ~
        paste0("truth-warm grid (c=0.5, r\u00b2=", annotation_r2, ")"),
      TRUE ~ use_case_id
    )
  )
```

# 1. SuSiE2-Style Plots (Category 1)

These use individual model metrics (CS properties, effect sizes).
For ensembles, select single best model by max ELBO.

## 1d. TPR vs FPR (FPR 0 to 0.1)

```{r fig-1d, fig.width=14, fig.height=8}
# Pooled TPR/FPR: combine all datasets' bins, one curve per (spec, agg_method).
# No per-dataset averaging — same pooling approach as AUPRC.
compute_tpr_fpr_pooled <- function(bins, group_vars) {
  bins %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::arrange(dplyr::desc(pip_threshold), .by_group = TRUE) %>%
    dplyr::mutate(
      cum_tp  = cumsum(n_causal_at_bucket),
      cum_fp  = cumsum(n_noncausal_at_bucket),
      total_p = sum(n_causal_at_bucket),
      total_n = sum(n_noncausal_at_bucket)
    ) %>%
    dplyr::filter(total_p > 0, total_n > 0) %>%
    dplyr::mutate(tpr = cum_tp / total_p, fpr = cum_fp / total_n) %>%
    dplyr::filter(fpr <= 0.10) %>%
    dplyr::ungroup()
}

# Ensemble aggregation curves
tpr_fpr_agg <- compute_tpr_fpr_pooled(
  confusion_bins %>% dplyr::filter(!is.na(agg_method)),
  group_vars = c("spec_name", "agg_method")
)

# Baseline single-fit curve (no agg_method) — replicated to all facets
baseline_roc_raw <- compute_tpr_fpr_pooled(
  confusion_bins %>%
    dplyr::filter(spec_name == "baseline-single", is.na(agg_method)),
  group_vars = "spec_name"
)

if (nrow(tpr_fpr_agg) > 0) {
  agg_methods <- sort(unique(tpr_fpr_agg$agg_method))
  baseline_roc <- if (nrow(baseline_roc_raw) > 0) {
    tidyr::crossing(
      baseline_roc_raw %>% dplyr::select(spec_name, fpr, tpr, pip_threshold),
      tibble::tibble(agg_method = agg_methods)
    )
  } else { tibble::tibble() }

  tpr_fpr_plot <- dplyr::bind_rows(tpr_fpr_agg, baseline_roc) %>%
    dplyr::mutate(is_baseline = spec_name == "baseline-single")

  p_roc <- ggplot(tpr_fpr_plot, aes(x = fpr, y = tpr, color = spec_name,
                                     linetype = is_baseline)) +
    geom_line(linewidth = 0.8) +
    scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                          guide = "none") +
    facet_wrap(~agg_method) +
    labs(x = "FPR", y = "TPR", color = "Ensemble Method",
         title = "TPR vs FPR (FPR 0-0.1, pooled across datasets)") +
    theme(legend.position = "bottom")
  ggsave(file.path(fig_root, "susie2_plots", "tpr_fpr_curve.png"),
         p_roc, width = 14, height = 8)
}
```

## 1e. PIP Calibration

```{r fig-1e, fig.width=14, fig.height=8}
# Pool all datasets' bins before binning — same pooling approach as AUPRC.
# Bin PIPs in 0.10-width bins, compute observed fraction causal per bin.
pool_pip_cal <- function(bins) {
  bins %>%
    dplyr::mutate(pip_bin = floor(pip_threshold / 0.10) * 0.10) %>%
    dplyr::group_by(dplyr::across(dplyr::any_of(
      c("spec_name", "agg_method", "pip_bin")
    ))) %>%
    dplyr::summarise(
      total_causal   = sum(n_causal_at_bucket,    na.rm = TRUE),
      total_noncausal = sum(n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      total         = total_causal + total_noncausal,
      observed_frac = dplyr::if_else(total > 0, total_causal / total, NA_real_)
    )
}

# Ensemble aggregation calibration
pip_cal_data <- pool_pip_cal(
  confusion_bins %>% dplyr::filter(!is.na(agg_method))
)

# Baseline single-fit calibration (replicated to all agg_method facets)
baseline_cal_raw <- pool_pip_cal(
  confusion_bins %>%
    dplyr::filter(spec_name == "baseline-single", is.na(agg_method))
)

if (nrow(pip_cal_data) > 0) {
  agg_methods <- sort(unique(pip_cal_data$agg_method))
  baseline_cal <- if (nrow(baseline_cal_raw) > 0) {
    tidyr::crossing(
      baseline_cal_raw %>%
        dplyr::select(spec_name, pip_bin, total_causal, total_noncausal,
                      total, observed_frac),
      tibble::tibble(agg_method = agg_methods)
    )
  } else { tibble::tibble() }

  pip_cal_all <- dplyr::bind_rows(pip_cal_data, baseline_cal) %>%
    dplyr::mutate(is_baseline = spec_name == "baseline-single")

  p_cal <- ggplot(pip_cal_all, aes(x = pip_bin + 0.05, y = observed_frac,
                                    color = spec_name, linetype = is_baseline)) +
    geom_point(size = 2) +
    geom_line() +
    scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                          guide = "none") +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
    facet_wrap(~agg_method) +
    labs(x = "Mean PIP (bin midpoint)", y = "Observed fraction causal",
         color = "Ensemble Method",
         title = "PIP Calibration (pooled across datasets)") +
    xlim(0, 1) + ylim(0, 1) +
    theme(legend.position = "bottom")
  ggsave(file.path(fig_root, "susie2_plots", "pip_calibration.png"),
         p_cal, width = 14, height = 8)
}
```

## 1a. CS Power vs top-n-as-causal

```{r fig-1a, fig.width=12, fig.height=6}
# Requires tier_cs_metrics_full.csv from collect workbook.
# Select max-ELBO run per ensemble group (spec_name x dataset_bundle_id x group_key),
# then average cs_power across datasets per (spec_name, n_top).
if (!is.null(tier_cs_metrics_full) && nrow(tier_cs_metrics_full) > 0) {

  # Max-ELBO run per ensemble group from purity-filtered model metrics
  best_runs <- model_metrics %>%
    dplyr::filter(.data$filtering == "purity_filtered",
                  !is.na(.data$elbo_final),
                  !is.na(.data$spec_name)) %>%
    dplyr::group_by(.data$spec_name, .data$dataset_bundle_id, .data$group_key) %>%
    dplyr::slice_max(.data$elbo_final, n = 1, with_ties = FALSE) %>%
    dplyr::ungroup() %>%
    dplyr::select(.data$run_id, .data$spec_name, .data$dataset_bundle_id,
                  .data$group_key)

  # Tier CS metrics for those runs only (individual model fits: is.na(agg_method))
  tier_best <- tier_cs_metrics_full %>%
    dplyr::filter(.data$filtering == "purity_filtered",
                  is.na(.data$agg_method)) %>%
    dplyr::semi_join(best_runs, by = "run_id")

  # Pool counts across datasets per (spec_name, n_top), then derive overall rates.
  # cs_power: sum(n_causal_covered) / sum(n_causal_total)  [overall, not mean]
  # cs_fdr:   sum(n_false_cs)       / sum(n_cs_total)      [overall, not mean]
  tier_avg <- tier_best %>%
    dplyr::group_by(.data$spec_name, .data$n_top) %>%
    dplyr::summarise(
      sum_causal_covered = sum(.data$n_causal_covered, na.rm = TRUE),
      sum_causal_total   = sum(.data$n_causal_total,   na.rm = TRUE),
      sum_false_cs       = sum(.data$n_false_cs,       na.rm = TRUE),
      sum_cs_total       = sum(.data$n_cs_total,       na.rm = TRUE),
      n_datasets         = dplyr::n_distinct(.data$dataset_bundle_id),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      cs_power = dplyr::if_else(.data$sum_causal_total > 0,
                                .data$sum_causal_covered / .data$sum_causal_total,
                                NA_real_),
      cs_fdr   = dplyr::if_else(.data$sum_cs_total > 0,
                                .data$sum_false_cs / .data$sum_cs_total,
                                NA_real_)
    )

  if (nrow(tier_avg) > 0) {
    p_1a <- ggplot(tier_avg, aes(x = factor(.data$n_top), y = .data$cs_power,
                                  color = .data$spec_name, group = .data$spec_name,
                                  linetype = .data$spec_name == "baseline-single")) +
      geom_point(size = 2) +
      geom_line() +
      scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                            guide = "none") +
      labs(x = "Top-N treated as causal", y = "CS Power",
           color = "Ensemble Method",
           title = "CS Power vs Top-N (max-ELBO run per ensemble, pooled across datasets)") +
      theme(legend.position = "bottom")
    ggsave(file.path(fig_root, "susie2_plots", "cs_power_vs_ntop.png"),
           p_1a, width = 12, height = 6)
  } else {
    cat("No tier CS data available after max-ELBO selection.\n")
  }
} else {
  cat("Plot 1a skipped: tier_cs_metrics_full.csv not available.\n")
}
```

## 1b. CS FDR vs top-n-as-causal

```{r fig-1b, fig.width=12, fig.height=6}
if (exists("tier_avg") && nrow(tier_avg) > 0) {
  p_1b <- ggplot(tier_avg, aes(x = factor(.data$n_top), y = .data$cs_fdr,
                                color = .data$spec_name, group = .data$spec_name,
                                linetype = .data$spec_name == "baseline-single")) +
    geom_point(size = 2) +
    geom_line() +
    scale_linetype_manual(values = c("FALSE" = "solid", "TRUE" = "dashed"),
                          guide = "none") +
    geom_hline(yintercept = 0.05, linetype = "dashed", color = "red",
               linewidth = 0.8) +
    annotate("text", x = 1, y = 0.05, label = "FDR = 0.05",
             vjust = -0.5, hjust = 0, color = "red", size = 3) +
    labs(x = "Top-N treated as causal", y = "CS FDR",
         color = "Ensemble Method",
         title = "CS FDR vs Top-N (max-ELBO run per ensemble, pooled across datasets)") +
    theme(legend.position = "bottom")
  ggsave(file.path(fig_root, "susie2_plots", "cs_fdr_vs_ntop.png"),
         p_1b, width = 12, height = 6)
} else {
  cat("Plot 1b skipped: tier_avg not available.\n")
}
```

## 1c. Estimated Heritability (hg2) Boxplot

```{r fig-1c, fig.width=14, fig.height=6}
# hg2 from model_metrics, max-ELBO run per ensemble group.
# truth-warm is split by use_case_id (vanilla / EB / grid) since each reflects
# a distinct prior specification, not a single model.
if ("hg2" %in% names(model_metrics)) {
  hg2_data <- if (exists("best_runs") && nrow(best_runs) > 0) {
    model_metrics %>%
      dplyr::filter(.data$filtering == "purity_filtered",
                    !is.na(.data$hg2), !is.na(.data$spec_name)) %>%
      dplyr::semi_join(best_runs, by = "run_id")
  } else {
    # Fallback: use all purity-filtered runs
    model_metrics %>%
      dplyr::filter(.data$filtering == "purity_filtered",
                    !is.na(.data$hg2), !is.na(.data$spec_name))
  }

  if (nrow(hg2_data) > 0) {
    # Build x-axis label: truth-warm -> use_case_id; all others -> spec_name
    hg2_data <- hg2_data %>%
      dplyr::mutate(
        hg2_label = dplyr::if_else(
          .data$spec_name == "truth-warm" & "use_case_id" %in% names(.),
          paste0("truth-warm\n(", .data$use_case_id, ")"),
          .data$spec_name
        )
      )

    p_1c <- ggplot(hg2_data, aes(x = .data$hg2_label, y = .data$hg2,
                                  fill = .data$hg2_label)) +
      geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
      labs(x = "Ensemble Method", y = "hg2 (estimated heritability)",
           title = "Estimated Heritability by Ensemble Method (max-ELBO run per group)") +
      theme(axis.text.x = element_text(angle = 30, hjust = 1),
            legend.position = "none")
    ggsave(file.path(fig_root, "susie2_plots", "hg2_boxplot.png"),
           p_1c, width = 14, height = 6)
  } else {
    cat("Plot 1c skipped: no hg2 data after max-ELBO selection.\n")
  }
} else {
  cat("Plot 1c skipped: hg2 column not found in model_metrics.\n")
}
```

# 2. Ensemble Performance Plots (Category 2)

## 2a. Precision-Recall Curves

PR curves pooled across all datasets (trapezoidal). Color = ensemble spec,
facet = aggregation method. Grading uses the architecture's causal SNPs as
stored in the flush confusion bins (all 23 for `susie2_oligogenic`; the 15
tiny polygenic effects inflate N_positives slightly vs strict top-8 grading,
but the relative ranking across methods is unaffected).

```{r fig-2a-helper}
# Build a pooled PR-curve tibble from confusion bins grouped by group_vars.
# Returns tibble with group_vars + recall, precision columns.
build_pr_curves <- function(bins, group_vars) {
  bins %>%
    dplyr::filter(!is.na(.data$agg_method)) %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::arrange(dplyr::desc(.data$pip_threshold), .by_group = TRUE) %>%
    dplyr::mutate(
      cum_tp  = cumsum(.data$n_causal_at_bucket),
      cum_fp  = cumsum(.data$n_noncausal_at_bucket),
      total_p = sum(.data$n_causal_at_bucket)
    ) %>%
    dplyr::filter(.data$total_p > 0L, (.data$cum_tp + .data$cum_fp) > 0L) %>%
    dplyr::mutate(
      recall    = .data$cum_tp / .data$total_p,
      precision = .data$cum_tp / (.data$cum_tp + .data$cum_fp)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::select(dplyr::all_of(group_vars), recall, precision)
}
```

```{r fig-2a, fig.width=14, fig.height=6}
pr_group_vars <- c("spec_name", "agg_method")
pr_curves_overall <- build_pr_curves(confusion_bins, pr_group_vars)

if (nrow(pr_curves_overall) > 0L) {
  p2a <- ggplot2::ggplot(
    pr_curves_overall,
    ggplot2::aes(x = recall, y = precision, color = spec_name, group = spec_name)
  ) +
    ggplot2::geom_line(linewidth = 0.7) +
    ggplot2::facet_wrap(~ agg_method, nrow = 1) +
    ggplot2::scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
    ggplot2::scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
    ggplot2::labs(
      x     = "Recall",
      y     = "Precision",
      color = "Ensemble spec",
      title = "Precision-Recall curves (pooled across datasets, all annotation R\u00b2)"
    ) +
    ggplot2::theme_bw() +
    ggplot2::theme(legend.position = "bottom")

  ggplot2::ggsave(
    file.path(fig_root, "ensemble_performance", "pr_curves_overall.png"),
    p2a, width = 14, height = 5, dpi = 150
  )
  print(p2a)

  # Per annotation_r2
  pr_curves_by_r2 <- build_pr_curves(
    confusion_bins,
    c("spec_name", "agg_method", "annotation_r2")
  )
  for (r2 in sort(unique(na.omit(pr_curves_by_r2$annotation_r2)))) {
    p2a_r2 <- ggplot2::ggplot(
      pr_curves_by_r2 %>% dplyr::filter(.data$annotation_r2 == r2),
      ggplot2::aes(x = recall, y = precision, color = spec_name, group = spec_name)
    ) +
      ggplot2::geom_line(linewidth = 0.7) +
      ggplot2::facet_wrap(~ agg_method, nrow = 1) +
      ggplot2::scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
      ggplot2::scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
      ggplot2::labs(
        x     = "Recall",
        y     = "Precision",
        color = "Ensemble spec",
        title = sprintf(
          "Precision-Recall curves (pooled, annotation R\u00b2 = %.1f)", r2)
      ) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = "bottom")

    ggplot2::ggsave(
      file.path(fig_root, "ensemble_performance",
                sprintf("pr_curves_r2_%.1f.png", r2)),
      p2a_r2, width = 14, height = 5, dpi = 150
    )
  }
  cat("Fig 2a saved.\n")
} else {
  cat("No aggregated confusion bins found — fig 2a skipped.\n")
}
```

## 2b. Multimodality Metrics Boxplots

```{r fig-2b, fig.width=14, fig.height=8}
mm_path <- file.path(consolidated_dir, "multimodal_metrics.csv")
if (file.exists(mm_path)) {
  mm <- readr::read_csv(mm_path, show_col_types = FALSE)

  # Join spec_name from run_info if needed
  if (!"spec_name" %in% names(mm) && "run_id" %in% names(mm)) {
    run_info <- readr::read_csv(
      file.path(consolidated_dir, "model_metrics_full.csv"),
      show_col_types = FALSE
    ) %>%
      dplyr::distinct(run_id, spec_name)
    mm <- mm %>% dplyr::left_join(run_info, by = "run_id")
  }

  metrics_to_plot <- intersect(
    c("jaccard_top10", "mean_jsd", "median_jsd", "max_jsd",
      "mean_pip_var", "n_clusters"),
    names(mm)
  )

  if (length(metrics_to_plot) > 0 && "spec_name" %in% names(mm)) {
    mm_long <- mm %>%
      dplyr::select(dplyr::any_of(c("spec_name", metrics_to_plot))) %>%
      dplyr::filter(!is.na(spec_name)) %>%
      tidyr::pivot_longer(cols = dplyr::any_of(metrics_to_plot),
                          names_to = "metric", values_to = "value")

    p_mm <- ggplot(mm_long, aes(x = spec_name, y = value, fill = spec_name)) +
      geom_boxplot() +
      facet_wrap(~metric, scales = "free_y") +
      labs(x = NULL, y = "Value", title = "Multimodality Metrics") +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
            legend.position = "none")
    ggsave(file.path(fig_root, "ensemble_performance",
                     "multimodality_boxplots.png"),
           p_mm, width = 14, height = 8)
  }
}
```

## 2c. AUPRC Bar Chart by Method (primary figure)

Four plots: overall (all annotation R² pooled) plus one per R² level.
Annotation-independent specs (Group A) appear on all plots. Annotation-
dependent specs (Groups B, C) are filtered to the matching R² on per-R²
plots. Truth-warm lines follow the same logic.

```{r fig-2c-helper}
tw_colors <- c("susine_vanilla" = "blue",
               "susine_eb_clamped_scale_var_nonneg" = "darkgreen",
               "susine_functional_mu" = "purple")

make_auprc_barplot <- function(r2_filter = NULL) {
  if (is.null(r2_filter)) {
    plot_data <- auprc_pooled_agg %>% dplyr::filter(!is.na(spec_name))
    tw_data   <- truth_warm_overall
    title_str <- "AUPRC by Ensemble Method and Aggregation (all annotation R\u00b2 pooled)"
    fname     <- "auprc_barplot_by_method_overall.png"
  } else {
    # Annotation-independent specs have annotation_r2 = NA — always include.
    # Annotation-dependent specs: filter to the target r2 level.
    plot_data <- auprc_pooled_agg_by_r2 %>%
      dplyr::filter(!is.na(spec_name),
                    is.na(annotation_r2) | annotation_r2 == r2_filter)
    # Truth-warm: vanilla (NA r2) always; EB/grid at matching r2 only.
    tw_data <- truth_warm_by_r2 %>%
      dplyr::filter(is.na(annotation_r2) | annotation_r2 == r2_filter)
    title_str <- paste0("AUPRC by Ensemble Method and Aggregation (annotation R\u00b2 = ",
                        r2_filter, ")")
    fname <- paste0("auprc_barplot_by_method_r2_",
                    gsub("\\.", "p", as.character(r2_filter)), ".png")
  }

  if (nrow(plot_data) == 0) return(invisible(NULL))

  p <- ggplot(plot_data, aes(x = spec_name, y = AUPRC, fill = agg_method)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    geom_hline(yintercept = baseline_auprc, linetype = "dashed",
               color = "red", linewidth = 0.8) +
    annotate("text", x = 0.5, y = baseline_auprc, label = "baseline single-fit",
             vjust = -0.5, hjust = 0, color = "red", size = 3) +
    labs(x = "Ensemble Method", y = "Pooled AUPRC", fill = "Aggregation",
         title = title_str) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  for (i in seq_len(nrow(tw_data))) {
    uc <- tw_data$use_case_id[i]
    p <- p +
      geom_hline(yintercept = tw_data$AUPRC[i], linetype = "dotted",
                 color = tw_colors[uc] %||% "blue", linewidth = 0.6) +
      annotate("text", x = Inf, y = tw_data$AUPRC[i],
               label = tw_data$label[i],
               vjust = -0.5, hjust = 1,
               color = tw_colors[uc] %||% "blue", size = 2.8)
  }

  ggsave(file.path(fig_root, "ensemble_performance", fname),
         p, width = 12, height = 6)
  invisible(p)
}
```

```{r fig-2c, fig.width=12, fig.height=6}
make_auprc_barplot(r2_filter = NULL)  # overall (all r2 pooled)
make_auprc_barplot(r2_filter = 0)     # annotation r2 = 0 (null annotations)
make_auprc_barplot(r2_filter = 0.2)   # annotation r2 = 0.2
make_auprc_barplot(r2_filter = 0.5)   # annotation r2 = 0.5
```

## 2d. TPR @ FPR=0.05 Bar Chart by Method

Mirrors the AUPRC barplot exactly but shows pooled TPR at FPR=0.05.
Four plots: overall + one per annotation R² level.

```{r fig-2d-helper}
make_tpr05_barplot <- function(r2_filter = NULL) {
  if (is.null(r2_filter)) {
    plot_data <- tpr05_pooled_agg %>% dplyr::filter(!is.na(spec_name))
    tw_data   <- truth_warm_tpr05_overall
    title_str <- "TPR @ FPR=0.05 by Ensemble Method and Aggregation (all annotation R\u00b2 pooled)"
    fname     <- "tpr05_barplot_by_method_overall.png"
    base_val  <- baseline_tpr05
  } else {
    plot_data <- tpr05_pooled_agg_by_r2 %>%
      dplyr::filter(!is.na(spec_name),
                    is.na(annotation_r2) | annotation_r2 == r2_filter)
    tw_data <- truth_warm_tpr05_by_r2 %>%
      dplyr::filter(is.na(annotation_r2) | annotation_r2 == r2_filter)
    # Baseline is annotation-independent; use same pooled overall value
    base_val  <- baseline_tpr05
    title_str <- paste0("TPR @ FPR=0.05 by Ensemble Method and Aggregation (annotation R\u00b2 = ",
                        r2_filter, ")")
    fname <- paste0("tpr05_barplot_by_method_r2_",
                    gsub("\\.", "p", as.character(r2_filter)), ".png")
  }

  if (nrow(plot_data) == 0 || length(base_val) == 0) return(invisible(NULL))

  p <- ggplot(plot_data, aes(x = spec_name, y = tpr_fpr05, fill = agg_method)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    geom_hline(yintercept = base_val, linetype = "dashed",
               color = "red", linewidth = 0.8) +
    annotate("text", x = 0.5, y = base_val, label = "baseline single-fit",
             vjust = -0.5, hjust = 0, color = "red", size = 3) +
    labs(x = "Ensemble Method", y = "Pooled TPR @ FPR=0.05", fill = "Aggregation",
         title = title_str) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  for (i in seq_len(nrow(tw_data))) {
    uc <- tw_data$use_case_id[i]
    p <- p +
      geom_hline(yintercept = tw_data$tpr_fpr05[i], linetype = "dotted",
                 color = tw_colors[uc] %||% "blue", linewidth = 0.6) +
      annotate("text", x = Inf, y = tw_data$tpr_fpr05[i],
               label = tw_data$label[i],
               vjust = -0.5, hjust = 1,
               color = tw_colors[uc] %||% "blue", size = 2.8)
  }

  ggsave(file.path(fig_root, "ensemble_performance", fname),
         p, width = 12, height = 6)
  invisible(p)
}
```

```{r fig-2d, fig.width=12, fig.height=6}
make_tpr05_barplot(r2_filter = NULL)  # overall (all r2 pooled)
make_tpr05_barplot(r2_filter = 0)     # annotation r2 = 0 (null annotations)
make_tpr05_barplot(r2_filter = 0.2)   # annotation r2 = 0.2
make_tpr05_barplot(r2_filter = 0.5)   # annotation r2 = 0.5
```

# 3. Dataset Difficulty Plots (Category 3)

## 3a. Dataset Metrics

```{r fig-3a, fig.width=12, fig.height=8}
dm_path <- file.path(consolidated_dir, "dataset_metrics.csv")
if (file.exists(dm_path)) {
  dataset_metrics <- readr::read_csv(dm_path, show_col_types = FALSE)

  metric_cols <- intersect(
    c("M1", "z_topk_ratio", "z_max_abs", "z_count_abs_gt_3",
      "z_eff_signals", "high_ld_count_095"),
    names(dataset_metrics)
  )

  if (length(metric_cols) > 0) {
    dm_long <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(metric_cols)) %>%
      tidyr::pivot_longer(everything(), names_to = "metric", values_to = "value")

    p_hist <- ggplot(dm_long, aes(x = value)) +
      geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7) +
      facet_wrap(~metric, scales = "free") +
      labs(x = "Value", y = "Count", title = "Dataset Metrics")
    ggsave(file.path(fig_root, "dataset_difficulty",
                     "metric_histograms.png"),
           p_hist, width = 12, height = 8)
  }
}
```

## 3b. Dataset Difficulty Scatter

```{r fig-3b, fig.width=10, fig.height=5}
if (exists("dataset_metrics") && "M1" %in% names(dataset_metrics)) {
  # Join baseline AUPRC to dataset metrics
  baseline_per_dataset <- auprc_individual %>%
    dplyr::filter(spec_name == "baseline-single",
                  use_case_id == "susine_vanilla") %>%
    dplyr::select(dataset_bundle_id, auprc_baseline = AUPRC)

  dm_scatter <- dataset_metrics %>%
    dplyr::inner_join(baseline_per_dataset, by = "dataset_bundle_id")

  if (nrow(dm_scatter) > 1) {
    p_scatter <- ggplot(dm_scatter, aes(x = M1, y = auprc_baseline)) +
      geom_point(alpha = 0.6) +
      geom_smooth(method = "lm", se = TRUE, color = "red") +
      labs(x = "M1 (LD complexity)", y = "Baseline AUPRC",
           title = "Dataset Difficulty: AUPRC vs M1")
    ggsave(file.path(fig_root, "dataset_difficulty",
                     "difficulty_scatter.png"),
           p_scatter, width = 10, height = 5)
  }
}
```

## 3c. Refine Depth Diagnostics

```{r fig-3c-refine-depth, fig.width=14, fig.height=6}
# Per-parent refine depth: n_refine_processed vs n_refine_requested.
# One row per (dataset × ensemble group × parent combination).
# Loaded from refine_depth.csv written by the collect workbook.
rd_path <- file.path(consolidated_dir, "refine_depth.csv")
rd_full <- if (file.exists(rd_path)) {
  readr::read_csv(rd_path, show_col_types = FALSE)
} else {
  NULL
}

if (!is.null(rd_full) && nrow(rd_full) > 0L &&
    all(c("n_refine_requested", "n_refine_processed", "group_label") %in% names(rd_full))) {

  # Attach spec_name via group_label → run_info join if available
  rd <- rd_full %>%
    dplyr::mutate(completion_rate = .data$n_refine_processed / .data$n_refine_requested)
  if (exists("run_info") && "spec_name" %in% names(run_info) && "group_label" %in% names(run_info)) {
    spec_map <- run_info %>% dplyr::distinct(.data$group_label, .data$spec_name)
    rd <- dplyr::left_join(rd, spec_map, by = "group_label")
  }
  x_var <- if ("spec_name" %in% names(rd)) "spec_name" else "group_label"

  ensure_dir(file.path(fig_root, "ensemble_scaling"))

  # Completion rate boxplot: one point per parent node
  p_rate <- ggplot(rd, aes(x = .data[[x_var]], y = completion_rate)) +
    geom_boxplot(alpha = 0.7, outlier.size = 1, fill = "steelblue") +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
    scale_y_continuous(limits = c(0, 1.05), labels = scales::percent) +
    labs(x = "Spec", y = "Completion rate (processed / requested)",
         title = "Refine BFS completion rate per parent node") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
  print(p_rate)
  ggsave(file.path(fig_root, "ensemble_scaling", "refine_depth_completion_rate.png"),
         p_rate, width = 14, height = 6)

  # Histogram of n_refine_processed, faceted by spec; red dashed = median n_requested
  n_req_vals <- rd %>%
    dplyr::group_by(.data[[x_var]]) %>%
    dplyr::summarise(n_refine_requested = median(.data$n_refine_requested), .groups = "drop")

  p_hist <- ggplot(rd, aes(x = n_refine_processed)) +
    geom_histogram(binwidth = 1, fill = "steelblue", color = "white", alpha = 0.8) +
    geom_vline(data = n_req_vals,
               aes(xintercept = n_refine_requested),
               linetype = "dashed", color = "red") +
    facet_wrap(as.formula(paste("~", x_var)), scales = "free_x") +
    labs(x = "n_refine_processed", y = "Count",
         title = "Refine depth per parent (red dashed = n_requested)") +
    theme_bw()
  print(p_hist)
  ggsave(file.path(fig_root, "ensemble_scaling", "refine_depth_histogram.png"),
         p_hist, width = 14, height = 6)

} else {
  message("refine_depth.csv not found or empty; skipping fig-3c.")
}
```

# 4. Ensemble Scaling Plots (Category 4)

Figures 4a–4d show AUPRC and TPR@0.05 as a function of ensemble size.
Computed by `compute_scaling_curves()` and `compute_interaction_scaling()`
in the collect workbook; loaded from `scaling_curves.csv` and
`interaction_scaling.csv` in the consolidated directory.

```{r load-scaling}
scaling_path     <- file.path(consolidated_dir, "scaling_curves.csv")
interaction_path <- file.path(consolidated_dir, "interaction_scaling.csv")

scaling_curves <- if (file.exists(scaling_path)) {
  readr::read_csv(scaling_path, show_col_types = FALSE)
} else {
  message("scaling_curves.csv not found — Category 4 plots will be skipped.")
  NULL
}

interaction_scaling <- if (file.exists(interaction_path)) {
  readr::read_csv(interaction_path, show_col_types = FALSE)
} else {
  message("interaction_scaling.csv not found — Category 4c/4d plots will be skipped.")
  NULL
}
```

```{r fig-4a}
# AUPRC scaling curves: ensemble size vs pooled AUPRC, faceted by spec.
# Ribbon shows ±1 SE for restart specs (50 reps); no ribbon for deterministic specs.
if (!is.null(scaling_curves) && nrow(scaling_curves) > 0L) {
  make_scaling_plot <- function(data, y_col, se_col, y_label, title) {
    p <- ggplot2::ggplot(
      data,
      ggplot2::aes(
        x     = n_ensemble,
        y     = .data[[y_col]],
        color = agg_method,
        group = agg_method
      )
    ) +
      ggplot2::geom_line() +
      ggplot2::geom_point(size = 1.8)

    # Add SE ribbon only where se is non-NA
    ribbon_data <- data %>% dplyr::filter(!is.na(.data[[se_col]]))
    if (nrow(ribbon_data) > 0L) {
      p <- p + ggplot2::geom_ribbon(
        data = ribbon_data,
        ggplot2::aes(
          ymin = .data[[y_col]] - .data[[se_col]],
          ymax = .data[[y_col]] + .data[[se_col]],
          fill = agg_method
        ),
        alpha = 0.15, color = NA
      )
    }

    p +
      ggplot2::scale_x_continuous(breaks = c(4, 8, 16, 32, 64)) +
      ggplot2::facet_wrap(~ spec_name, nrow = 2) +
      ggplot2::labs(
        x     = "Ensemble size (# fits)",
        y     = y_label,
        color = "Aggregation",
        fill  = "Aggregation",
        title = title
      ) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = "bottom")
  }

  # Overall (pooled across all annotation_r2)
  p4a_overall <- make_scaling_plot(
    scaling_curves %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col    = "AUPRC_mean",
    se_col   = "AUPRC_se",
    y_label  = "AUPRC (pooled across datasets)",
    title    = "AUPRC scaling — all annotation R\u00b2"
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "auprc_scaling_curves_overall.png"),
    p4a_overall, width = 12, height = 6, dpi = 150
  )
  print(p4a_overall)

  # Per annotation_r2
  for (r2 in sort(unique(na.omit(scaling_curves$annotation_r2)))) {
    p4a_r2 <- make_scaling_plot(
      scaling_curves %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col    = "AUPRC_mean",
      se_col   = "AUPRC_se",
      y_label  = "AUPRC (pooled across datasets)",
      title    = sprintf("AUPRC scaling \u2014 annotation R\u00b2 = %.1f", r2)
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("auprc_scaling_curves_r2_%.1f.png", r2)),
      p4a_r2, width = 12, height = 6, dpi = 150
    )
  }
  cat("Fig 4a saved.\n")
} else {
  cat("scaling_curves.csv not found — fig 4a skipped.\n")
}
```

```{r fig-4b}
# TPR@FPR=0.05 scaling curves: same structure as 4a.
if (!is.null(scaling_curves) && nrow(scaling_curves) > 0L) {
  p4b_overall <- make_scaling_plot(
    scaling_curves %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col    = "tpr05_mean",
    se_col   = "tpr05_se",
    y_label  = "TPR at FPR \u2264 0.05 (pooled)",
    title    = "TPR@FPR=0.05 scaling \u2014 all annotation R\u00b2"
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "tpr05_scaling_curves_overall.png"),
    p4b_overall, width = 12, height = 6, dpi = 150
  )
  print(p4b_overall)

  for (r2 in sort(unique(na.omit(scaling_curves$annotation_r2)))) {
    p4b_r2 <- make_scaling_plot(
      scaling_curves %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col    = "tpr05_mean",
      se_col   = "tpr05_se",
      y_label  = "TPR at FPR \u2264 0.05 (pooled)",
      title    = sprintf("TPR@FPR=0.05 scaling \u2014 annotation R\u00b2 = %.1f", r2)
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("tpr05_scaling_curves_r2_%.1f.png", r2)),
      p4b_r2, width = 12, height = 6, dpi = 150
    )
  }
  cat("Fig 4b saved.\n")
} else {
  cat("scaling_curves.csv not found — fig 4b skipped.\n")
}
```

```{r fig-4c}
# 2-lever interaction AUPRC: compare 4x4 vs 8x8 grid resolutions.
if (!is.null(interaction_scaling) && nrow(interaction_scaling) > 0L) {
  make_interaction_plot <- function(data, y_col, y_label, title) {
    ggplot2::ggplot(
      data,
      ggplot2::aes(
        x     = resolution,
        y     = .data[[y_col]],
        color = agg_method,
        group = agg_method
      )
    ) +
      ggplot2::geom_line() +
      ggplot2::geom_point(size = 2.5) +
      ggplot2::facet_wrap(~ spec_name, nrow = 1) +
      ggplot2::labs(
        x     = "Grid resolution",
        y     = y_label,
        color = "Aggregation",
        title = title
      ) +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = "bottom")
  }

  p4c_overall <- make_interaction_plot(
    interaction_scaling %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col   = "AUPRC",
    y_label = "AUPRC (pooled across datasets)",
    title   = "Interaction grid scaling \u2014 AUPRC at 4\u00d74 vs 8\u00d78"
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "auprc_interaction_scaling_overall.png"),
    p4c_overall, width = 10, height = 4, dpi = 150
  )
  print(p4c_overall)

  for (r2 in sort(unique(na.omit(interaction_scaling$annotation_r2)))) {
    p4c_r2 <- make_interaction_plot(
      interaction_scaling %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col   = "AUPRC",
      y_label = "AUPRC (pooled across datasets)",
      title   = sprintf("Interaction grid scaling \u2014 AUPRC (R\u00b2=%.1f)", r2)
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("auprc_interaction_scaling_r2_%.1f.png", r2)),
      p4c_r2, width = 10, height = 4, dpi = 150
    )
  }
  cat("Fig 4c saved.\n")
} else {
  cat("interaction_scaling.csv not found — fig 4c skipped.\n")
}
```

```{r fig-4d}
# 2-lever interaction TPR@0.05: same structure as 4c.
if (!is.null(interaction_scaling) && nrow(interaction_scaling) > 0L) {
  p4d_overall <- make_interaction_plot(
    interaction_scaling %>% dplyr::filter(is.na(.data$annotation_r2)),
    y_col   = "tpr05",
    y_label = "TPR at FPR \u2264 0.05 (pooled)",
    title   = "Interaction grid scaling \u2014 TPR@FPR=0.05 at 4\u00d74 vs 8\u00d78"
  )
  ggplot2::ggsave(
    file.path(fig_root, "ensemble_scaling", "tpr05_interaction_scaling_overall.png"),
    p4d_overall, width = 10, height = 4, dpi = 150
  )
  print(p4d_overall)

  for (r2 in sort(unique(na.omit(interaction_scaling$annotation_r2)))) {
    p4d_r2 <- make_interaction_plot(
      interaction_scaling %>% dplyr::filter(.data$annotation_r2 == r2),
      y_col   = "tpr05",
      y_label = "TPR at FPR \u2264 0.05 (pooled)",
      title   = sprintf(
        "Interaction grid scaling \u2014 TPR@FPR=0.05 (R\u00b2=%.1f)", r2)
    )
    ggplot2::ggsave(
      file.path(fig_root, "ensemble_scaling",
                sprintf("tpr05_interaction_scaling_r2_%.1f.png", r2)),
      p4d_r2, width = 10, height = 4, dpi = 150
    )
  }
  cat("Fig 4d saved.\n")
} else {
  cat("interaction_scaling.csv not found — fig 4d skipped.\n")
}
```

# Session Info

```{r session}
cat("\nFigures written to:", fig_root, "\n")
cat("Files:\n")
for (d in fig_dirs) {
  figs <- list.files(file.path(fig_root, d))
  if (length(figs)) cat(sprintf("  %s/: %s\n", d, paste(figs, collapse = ", ")))
}
sessionInfo()
```
