---
title: "Ensemble Scaling Experiment - Paper Figures"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Paper-focused figures for the ensemble scaling simulation. This workbook reads
the consolidated CSV files produced by
`collect_results_workbook_ensemble_scaling.Rmd` and writes PNG figures only.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_ensemble_scaling_paper.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(
    susine_path, winslash = "/", mustWork = TRUE
  ))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 11))

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0L) y else x
}
```

# 0. User Configuration

```{r config}
job_name <- "ensemble_scaling_full"
parent_job_id <- "51250228"
output_root <- here("output")

job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
consolidated_dir <- file.path(job_dir, "consolidated")

annotation_r2_focus <- 0.2
drop_aggs <- "cluster_weight_jsd_050"
bootstrap_B <- 2000L
bootstrap_seed <- 20260505L
true_h2 <- 0.25

fig_dir <- file.path(job_dir, "figures", "paper_ensemble_scaling")

if (!dir.exists(consolidated_dir)) {
  stop(
    "Consolidated output directory not found: ", consolidated_dir, "\n",
    "Run collect_results_workbook_ensemble_scaling.Rmd on the HPC output first, ",
    "or update job_name/parent_job_id/consolidated_dir at the top of this workbook.",
    call. = FALSE
  )
}
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

cat("Consolidated dir:", consolidated_dir, "\n")
cat("Figure dir:", fig_dir, "\n")
```

```{r load-data}
read_required_csv <- function(filename) {
  path <- file.path(consolidated_dir, filename)
  if (!file.exists(path)) {
    stop("Required consolidated CSV missing: ", path, call. = FALSE)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

read_optional_csv <- function(filename) {
  path <- file.path(consolidated_dir, filename)
  if (!file.exists(path)) {
    message("Optional consolidated CSV missing: ", filename)
    return(NULL)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

auprc_agg <- read_required_csv("auprc_aggregated.csv") %>%
  test_susine:::normalize_agg_method_df()
auprc_individual <- read_required_csv("auprc_individual.csv")
auprc_pooled_agg <- read_required_csv("auprc_pooled_aggregated.csv") %>%
  test_susine:::normalize_agg_method_df()
auprc_pooled_agg_by_r2 <- read_required_csv("auprc_pooled_agg_by_r2.csv") %>%
  test_susine:::normalize_agg_method_df()
auprc_pooled_individual_overall <- read_required_csv(
  "auprc_pooled_individual_overall.csv"
)
auprc_pooled_individual <- read_required_csv("auprc_pooled_individual.csv")

confusion_bins <- read_required_csv("confusion_bins_full.csv") %>%
  test_susine:::normalize_agg_method_df()
model_metrics <- read_required_csv("model_metrics_full.csv") %>%
  test_susine:::normalize_agg_method_df()

hg2_by_agg <- read_optional_csv("hg2_by_agg.csv")
if (!is.null(hg2_by_agg)) {
  hg2_by_agg <- test_susine:::normalize_agg_method_df(hg2_by_agg)
}

scaling_bins_pooled <- read_optional_csv("scaling_bins_pooled.csv")
if (!is.null(scaling_bins_pooled)) {
  scaling_bins_pooled <- scaling_bins_pooled %>%
    test_susine:::normalize_agg_method_df()
}

cat("AUPRC aggregated rows:", nrow(auprc_agg), "\n")
cat("AUPRC individual rows:", nrow(auprc_individual), "\n")
cat("Confusion-bin rows:", nrow(confusion_bins), "\n")
cat("Model-metric rows:", nrow(model_metrics), "\n")
```

# 1. Helpers

```{r helpers}
approx_equal <- function(x, y, tol = 1e-8) {
  is.finite(x) & is.finite(y) & abs(x - y) <= tol
}

near_focus <- function(x, focus = annotation_r2_focus) {
  approx_equal(as.numeric(x), focus)
}

is_group_a <- function(spec_name) {
  startsWith(as.character(spec_name), "A-")
}

is_focus_annotation_row <- function(spec_name, annotation_r2,
                                    focus = annotation_r2_focus) {
  is_group_a(spec_name) & is.na(annotation_r2) |
    !is_group_a(spec_name) & near_focus(annotation_r2, focus)
}

auprc_from_bins <- function(bins) {
  test_susine:::auprc_from_pooled_bins(bins)
}

save_paper_png <- function(plot, filename, width, height) {
  ggplot2::ggsave(
    filename = file.path(fig_dir, filename),
    plot = plot,
    width = width,
    height = height,
    dpi = 300,
    bg = "white"
  )
}

paper_theme <- function(base_size = 11) {
  ggplot2::theme_bw(base_size = base_size) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      plot.title.position = "plot",
      legend.position = "bottom",
      legend.box = "vertical"
    )
}

spec_order <- c(
  "A-R", "A-F", "A-S", "A-RF", "A-RS", "A-FS", "A-RFS",
  "B-R", "B-F", "B-RF",
  "C-C", "C-CS", "C-CSR", "C-CSFR", "C-CFR", "C-CF"
)

spec_key <- tibble::tribble(
  ~spec_name, ~family, ~description,
  "A-R", "A: vanilla SuSiE", "Restarts",
  "A-F", "A: vanilla SuSiE", "Refinement",
  "A-S", "A: vanilla SuSiE", "Prior-variance grid",
  "A-RF", "A: vanilla SuSiE", "Restarts x refinement",
  "A-RS", "A: vanilla SuSiE", "Restarts x prior-variance grid",
  "A-FS", "A: vanilla SuSiE", "Refinement x prior-variance grid",
  "A-RFS", "A: vanilla SuSiE", "Restarts x refinement x prior-variance grid",
  "B-R", "B: EB SuSiNE", "Restarts",
  "B-F", "B: EB SuSiNE", "Refinement",
  "B-RF", "B: EB SuSiNE", "Restarts x refinement",
  "C-C", "C: grid SuSiNE", "Annotation-scale grid",
  "C-CS", "C: grid SuSiNE", "Annotation-scale x prior-variance grid",
  "C-CSR", "C: grid SuSiNE", "C-CS search, then default-prior refit",
  "C-CSFR", "C: grid SuSiNE", "C x sigma x refinement x restarts",
  "C-CFR", "C: grid SuSiNE", "C x refinement x restarts",
  "C-CF", "C: grid SuSiNE", "C x refinement"
)

agg_order <- c("max_elbo", "uniform", "elbo_softmax", "cluster_weight", "oracle")
agg_labels <- c(
  max_elbo = "max ELBO",
  uniform = "uniform",
  elbo_softmax = "ELBO softmax",
  cluster_weight = "cluster weight",
  oracle = "oracle"
)

format_r2_tag <- function(x) {
  gsub("\\.", "p", format(x, trim = TRUE, scientific = FALSE))
}

extract_baseline_metric <- function(plot_table, metric_col) {
  plot_table %>%
    dplyr::filter(
      .data$series_type == "baseline_single",
      .data$spec_name == "baseline-single",
      .data$use_case_id == "susine_vanilla",
      is.na(.data$annotation_r2)
    ) %>%
    dplyr::pull(!!rlang::sym(metric_col)) %>%
    dplyr::first()
}

extract_truth_warm_refs <- function(plot_table, metric_col) {
  plot_table %>%
    dplyr::filter(.data$series_type == "truth_warm") %>%
    dplyr::select(use_case_id, annotation_r2, dplyr::all_of(metric_col)) %>%
    dplyr::mutate(
      label = dplyr::case_when(
        .data$use_case_id == "susine_vanilla" ~ "truth-warm vanilla",
        .data$use_case_id == "susine_eb_clamped_scale_var_nonneg" &
          is.na(.data$annotation_r2) ~ "truth-warm EB",
        .data$use_case_id == "susine_functional_mu" &
          is.na(.data$annotation_r2) ~ "truth-warm grid (c=0.5)",
        .data$use_case_id == "susine_eb_clamped_scale_var_nonneg" ~
          paste0("truth-warm EB (R2=", .data$annotation_r2, ")"),
        .data$use_case_id == "susine_functional_mu" ~
          paste0("truth-warm grid (c=0.5, R2=", .data$annotation_r2, ")"),
        TRUE ~ .data$use_case_id
      )
    )
}

build_pr_curves <- function(bins, group_vars) {
  bins %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::arrange(dplyr::desc(.data$pip_threshold), .by_group = TRUE) %>%
    dplyr::mutate(
      cum_tp = cumsum(.data$n_causal_at_bucket),
      cum_fp = cumsum(.data$n_noncausal_at_bucket),
      total_p = sum(.data$n_causal_at_bucket)
    ) %>%
    dplyr::filter(.data$total_p > 0L, (.data$cum_tp + .data$cum_fp) > 0L) %>%
    dplyr::mutate(
      recall = .data$cum_tp / .data$total_p,
      precision = .data$cum_tp / (.data$cum_tp + .data$cum_fp)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::select(dplyr::all_of(group_vars), recall, precision)
}

pool_bins <- function(bins, group_vars) {
  bins %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(c(group_vars, "pip_threshold")))) %>%
    dplyr::summarise(
      n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
      n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    )
}

paper_select_axis_subset <- function(vals, n_keep, axis_col) {
  vals <- sort(unique(vals))
  vals <- vals[!is.na(vals)]
  if (!length(vals)) return(vals)
  n_keep <- min(as.integer(n_keep), length(vals))
  if (n_keep >= length(vals)) return(vals)
  if (axis_col %in% c("restart_id", "refine_step")) {
    return(vals[seq_len(n_keep)])
  }
  vals[unique(round(seq(1, length(vals), length.out = n_keep)))]
}

paper_select_subsample_run_ids <- function(run_meta, n_ens, lever_type) {
  n_sel <- min(as.integer(n_ens), nrow(run_meta))
  switch(lever_type,
    restart = {
      run_meta %>%
        dplyr::arrange(.data$restart_id, .data$run_id) %>%
        dplyr::slice_head(n = n_sel) %>%
        dplyr::pull("run_id")
    },
    refine = {
      run_meta %>%
        dplyr::arrange(.data$refine_step, .data$run_id) %>%
        dplyr::slice_head(n = n_sel) %>%
        dplyr::pull("run_id")
    },
    sigma_0_2 = {
      vals <- sort(unique(run_meta$sigma_0_2_scalar_num))
      keep <- vals[unique(round(seq(1, length(vals), length.out = n_sel)))]
      run_meta %>%
        dplyr::filter(.data$sigma_0_2_scalar_num %in% keep) %>%
        dplyr::pull("run_id")
    },
    c_grid = {
      vals <- sort(unique(run_meta$c_value))
      keep <- vals[unique(round(seq(1, length(vals), length.out = n_sel)))]
      run_meta %>%
        dplyr::filter(.data$c_value %in% keep) %>%
        dplyr::pull("run_id")
    },
    stop("Unknown lever_type: ", lever_type)
  )
}

prepare_individual_run_rows <- function(model_metrics) {
  model_metrics %>%
    dplyr::filter(
      .data$filtering == "purity_filtered",
      !is.na(.data$spec_name),
      !.data$spec_name %in% c("baseline-single", "truth-warm"),
      is.na(.data$agg_method) | .data$explore_method != "aggregation"
    ) %>%
    dplyr::mutate(
      sigma_0_2_scalar_num = suppressWarnings(as.numeric(.data$sigma_0_2_scalar)),
      c_value = suppressWarnings(as.numeric(.data$c_value)),
      restart_id = suppressWarnings(as.integer(.data$restart_id)),
      refine_step = suppressWarnings(as.integer(.data$refine_step)),
      wall_time_sec = as.numeric(.data$wall_time_sec)
    ) %>%
    dplyr::distinct(run_id, .keep_all = TRUE)
}

select_subscale_run_rows <- function(run_rows, n_ensemble, resolution) {
  run_rows <- run_rows %>%
    dplyr::arrange(.data$run_id)
  if (!nrow(run_rows)) return(run_rows)

  method_ids <- unique(trimws(strsplit(
    as.character(run_rows$exploration_methods[[1]] %||% ""),
    "x",
    fixed = TRUE
  )[[1]]))
  is_cs_grid_refit <- "cs_grid_refit" %in% method_ids

  if (!is.na(resolution) && nzchar(as.character(resolution))) {
    axis_info <- list(
      if ("restart" %in% method_ids) list(col = "restart_id") else NULL,
      if ("refine" %in% method_ids) list(col = "refine_step") else NULL,
      if ("c_grid" %in% method_ids || is_cs_grid_refit) list(col = "c_value") else NULL,
      if ("sigma_0_2_grid" %in% method_ids || is_cs_grid_refit) {
        list(col = "sigma_0_2_scalar_num")
      } else {
        NULL
      }
    )
    axis_info <- Filter(Negate(is.null), axis_info)
    if (!length(axis_info)) return(run_rows[0, , drop = FALSE])

    axis_sizes <- vapply(axis_info, function(ax) {
      length(unique(stats::na.omit(run_rows[[ax$col]])))
    }, integer(1L))
    sort_order <- order(axis_sizes, decreasing = TRUE)
    axis_info <- axis_info[sort_order]

    res_sizes <- suppressWarnings(as.integer(strsplit(
      as.character(resolution), "x", fixed = TRUE
    )[[1]]))
    if (length(res_sizes) != length(axis_info) || any(is.na(res_sizes))) {
      return(run_rows[0, , drop = FALSE])
    }

    selected <- run_rows
    for (i in seq_along(axis_info)) {
      col <- axis_info[[i]]$col
      keep_vals <- paper_select_axis_subset(selected[[col]], res_sizes[[i]], col)
      selected <- selected %>% dplyr::filter(.data[[col]] %in% keep_vals)
    }
    return(selected)
  }

  lever_type <- if ("restart" %in% method_ids) {
    "restart"
  } else if ("refine" %in% method_ids) {
    "refine"
  } else if ("sigma_0_2_grid" %in% method_ids) {
    "sigma_0_2"
  } else if ("c_grid" %in% method_ids) {
    "c_grid"
  } else {
    NA_character_
  }
  if (is.na(lever_type)) return(run_rows[0, , drop = FALSE])

  ids <- paper_select_subsample_run_ids(run_rows, n_ensemble, lever_type)
  run_rows %>% dplyr::filter(.data$run_id %in% ids)
}

compute_scaling_metrics <- function(scaling_bins) {
  if (is.null(scaling_bins) || !nrow(scaling_bins)) {
    return(tibble::tibble())
  }

  scaling_bins %>%
    dplyr::filter(!.data$agg_method %in% drop_aggs) %>%
    dplyr::group_by(
      .data$spec_name, .data$annotation_r2, .data$n_ensemble,
      .data$resolution, .data$agg_method, .data$rep
    ) %>%
    dplyr::summarise(
      AUPRC = auprc_from_bins(dplyr::pick(dplyr::everything())),
      .groups = "drop"
    ) %>%
    dplyr::group_by(
      .data$spec_name, .data$annotation_r2, .data$n_ensemble,
      .data$resolution, .data$agg_method
    ) %>%
    dplyr::summarise(
      AUPRC_mean = mean(.data$AUPRC, na.rm = TRUE),
      AUPRC_se = ifelse(
        dplyr::n() > 1L,
        stats::sd(.data$AUPRC, na.rm = TRUE) / sqrt(sum(!is.na(.data$AUPRC))),
        NA_real_
      ),
      n_reps = dplyr::n(),
      .groups = "drop"
    )
}

estimate_subscale_compute <- function(points, model_metrics) {
  if (!nrow(points) || !"wall_time_sec" %in% names(model_metrics)) {
    return(tibble::tibble())
  }

  base_runs <- prepare_individual_run_rows(model_metrics)
  out <- lapply(seq_len(nrow(points)), function(i) {
    point <- points[i, , drop = FALSE]
    run_rows <- base_runs %>%
      dplyr::filter(.data$spec_name == point$spec_name[[1]])

    if (is_group_a(point$spec_name[[1]])) {
      run_rows <- run_rows %>% dplyr::filter(is.na(.data$annotation_r2))
    } else {
      run_rows <- run_rows %>%
        dplyr::filter(near_focus(.data$annotation_r2, annotation_r2_focus))
    }

    if (!nrow(run_rows)) {
      return(tibble::tibble(
        spec_name = point$spec_name,
        annotation_r2 = point$annotation_r2,
        n_ensemble = point$n_ensemble,
        resolution = point$resolution,
        mean_total_wall_sec = NA_real_,
        n_datasets_compute = 0L
      ))
    }

    dataset_groups <- split(run_rows, run_rows$dataset_bundle_id)
    wall_times <- vapply(dataset_groups, function(g) {
      selected <- select_subscale_run_rows(g, point$n_ensemble[[1]], point$resolution[[1]])
      sum(selected$wall_time_sec, na.rm = TRUE)
    }, numeric(1L))

    tibble::tibble(
      spec_name = point$spec_name,
      annotation_r2 = point$annotation_r2,
      n_ensemble = point$n_ensemble,
      resolution = point$resolution,
      mean_total_wall_sec = mean(wall_times, na.rm = TRUE),
      n_datasets_compute = length(wall_times)
    )
  })

  dplyr::bind_rows(out)
}

baseline_compute_minutes <- function(model_metrics) {
  if (!"wall_time_sec" %in% names(model_metrics)) return(NA_real_)
  model_metrics %>%
    dplyr::filter(
      .data$spec_name == "baseline-single",
      .data$use_case_id == "susine_vanilla",
      .data$filtering == "purity_filtered"
    ) %>%
    dplyr::distinct(dataset_bundle_id, run_id, .keep_all = TRUE) %>%
    dplyr::summarise(mean_wall_sec = mean(.data$wall_time_sec, na.rm = TRUE)) %>%
    dplyr::pull("mean_wall_sec") / 60
}
```

```{r shared-tables}
auprc_plot_table <- test_susine:::build_shared_plot_metric_table(
  auprc_pooled_agg,
  auprc_pooled_agg_by_r2,
  auprc_pooled_individual_overall,
  auprc_pooled_individual,
  metric_col = "AUPRC",
  label = "Paper AUPRC plot table"
) %>%
  dplyr::filter(is.na(.data$agg_method) | !.data$agg_method %in% drop_aggs)

baseline_auprc <- extract_baseline_metric(auprc_plot_table, "AUPRC")
truth_warm_refs <- extract_truth_warm_refs(auprc_plot_table, "AUPRC")
truth_warm_vanilla_vec <- truth_warm_refs %>%
  dplyr::filter(.data$use_case_id == "susine_vanilla", is.na(.data$annotation_r2)) %>%
  dplyr::pull("AUPRC")
truth_warm_vanilla <- if (length(truth_warm_vanilla_vec)) {
  truth_warm_vanilla_vec[[1]]
} else {
  NA_real_
}

cat("Baseline SuSiE AUPRC:", round(baseline_auprc, 4), "\n")
cat("Truth-warm vanilla AUPRC:", round(truth_warm_vanilla, 4), "\n")

knitr::kable(spec_key, caption = "Ensemble abbreviation key")
```

# 2. Overall Ensemble Performance Heatmap

```{r fig-delta-heatmap, fig.width=10, fig.height=7}
heatmap_data <- test_susine:::filter_shared_plot_metric_rows(
  auprc_plot_table,
  r2_filter = annotation_r2_focus
) %>%
  dplyr::filter(
    !.data$agg_method %in% drop_aggs,
    .data$agg_method %in% agg_order,
    .data$spec_name %in% spec_order
  ) %>%
  dplyr::mutate(
    delta_auprc = .data$AUPRC - baseline_auprc,
    spec_name = factor(.data$spec_name, levels = rev(spec_order)),
    agg_method = factor(.data$agg_method, levels = agg_order)
  )

if (nrow(heatmap_data) > 0L && is.finite(baseline_auprc)) {
  heatmap_check <- heatmap_data %>%
    dplyr::mutate(check_delta = .data$AUPRC - baseline_auprc - .data$delta_auprc)
  stopifnot(max(abs(heatmap_check$check_delta), na.rm = TRUE) < 1e-12)

  y_levels <- levels(heatmap_data$spec_name)
  y_prefix <- sub("-.*$", "", y_levels)
  y_separators <- if (length(y_prefix) > 1L) {
    which(y_prefix[-length(y_prefix)] != y_prefix[-1L]) + 0.5
  } else {
    numeric(0L)
  }
  oracle_x <- match("oracle", agg_order) - 0.5

  p_heatmap <- ggplot2::ggplot(
    heatmap_data,
    ggplot2::aes(x = .data$agg_method, y = .data$spec_name, fill = .data$delta_auprc)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.4) +
    ggplot2::geom_text(
      ggplot2::aes(label = sprintf("%+.3f", .data$delta_auprc)),
      size = 2.7
    ) +
    ggplot2::geom_hline(
      yintercept = y_separators,
      linewidth = 0.6,
      color = "grey20"
    ) +
    ggplot2::geom_vline(
      xintercept = oracle_x,
      linewidth = 0.6,
      color = "grey20"
    ) +
    ggplot2::scale_x_discrete(labels = agg_labels, drop = FALSE) +
    ggplot2::scale_fill_gradient2(
      low = "#2166AC",
      mid = "white",
      high = "#B2182B",
      midpoint = 0,
      name = "Delta AUPRC"
    ) +
    ggplot2::labs(
      x = "Aggregation method",
      y = "Ensemble method",
      title = "AUPRC gain over baseline SuSiE",
      subtitle = paste0("annotation_r2 = ", annotation_r2_focus)
    ) +
    paper_theme()

  save_paper_png(
    p_heatmap,
    paste0("paper_delta_auprc_heatmap_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 10,
    height = 7
  )
  p_heatmap
} else {
  message("Skipping heatmap: no rows after filtering.")
}
```

# 3. Best Methods vs Annotation Quality

```{r fig-ccs-by-r2, fig.width=7, fig.height=4.8}
ccs_by_r2 <- auprc_plot_table %>%
  dplyr::filter(
    .data$series_type == "aggregated",
    .data$spec_name == "C-CS",
    !is.na(.data$annotation_r2),
    !.data$agg_method %in% drop_aggs,
    .data$agg_method %in% agg_order
  ) %>%
  dplyr::mutate(
    agg_method = factor(.data$agg_method, levels = agg_order)
  )

if (nrow(ccs_by_r2) > 0L) {
  p_ccs_r2 <- ggplot2::ggplot(
    ccs_by_r2,
    ggplot2::aes(
      x = .data$annotation_r2,
      y = .data$AUPRC,
      color = .data$agg_method,
      group = .data$agg_method
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.55
    ) +
    ggplot2::geom_hline(
      yintercept = truth_warm_vanilla,
      linetype = "dotted",
      color = "grey25",
      linewidth = 0.55
    ) +
    ggplot2::geom_line(linewidth = 0.75) +
    ggplot2::geom_point(size = 2) +
    ggplot2::coord_cartesian(ylim = c(0.2, 0.4)) +
    ggplot2::scale_x_continuous(breaks = sort(unique(ccs_by_r2$annotation_r2))) +
    ggplot2::scale_color_discrete(labels = agg_labels[levels(ccs_by_r2$agg_method)]) +
    ggplot2::labs(
      x = "Annotation R2",
      y = "Pooled AUPRC",
      color = "Aggregation",
      title = "C-CS aggregation performance across annotation quality"
    ) +
    paper_theme()

  save_paper_png(
    p_ccs_r2,
    "paper_ccs_aggregation_by_annotation_r2.png",
    width = 7,
    height = 4.8
  )
  p_ccs_r2
} else {
  message("Skipping C-CS by annotation quality plot: no rows.")
}
```

# 4. Bootstrap AUPRC CI

```{r fig-bootstrap, fig.width=7, fig.height=4.8}
bootstrap_auprc_from_bins <- function(bins, sample_ids) {
  weights <- tibble::tibble(dataset_bundle_id = sample_ids) %>%
    dplyr::count(.data$dataset_bundle_id, name = "boot_weight")

  bins %>%
    dplyr::inner_join(weights, by = "dataset_bundle_id") %>%
    dplyr::group_by(.data$pip_threshold) %>%
    dplyr::summarise(
      n_causal_at_bucket = sum(.data$n_causal_at_bucket * .data$boot_weight, na.rm = TRUE),
      n_noncausal_at_bucket = sum(
        .data$n_noncausal_at_bucket * .data$boot_weight,
        na.rm = TRUE
      ),
      .groups = "drop"
    ) %>%
    auprc_from_bins()
}

baseline_boot_bins <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "baseline-single",
    .data$use_case_id == "susine_vanilla",
    is.na(.data$agg_method)
  )

ccs_boot_bins <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "C-CS",
    near_focus(.data$annotation_r2, annotation_r2_focus),
    .data$agg_method == "cluster_weight"
  )

shared_boot_ids <- intersect(
  unique(baseline_boot_bins$dataset_bundle_id),
  unique(ccs_boot_bins$dataset_bundle_id)
)

if (length(shared_boot_ids) > 1L) {
  set.seed(bootstrap_seed)
  boot_samples <- replicate(
    bootstrap_B,
    sample(shared_boot_ids, size = length(shared_boot_ids), replace = TRUE),
    simplify = FALSE
  )

  baseline_boot <- vapply(
    boot_samples,
    function(ids) bootstrap_auprc_from_bins(baseline_boot_bins, ids),
    numeric(1L)
  )
  ccs_boot <- vapply(
    boot_samples,
    function(ids) bootstrap_auprc_from_bins(ccs_boot_bins, ids),
    numeric(1L)
  )
  delta_boot <- ccs_boot - baseline_boot

  method_ci <- dplyr::bind_rows(
    tibble::tibble(method = "SuSiE baseline", AUPRC = baseline_boot),
    tibble::tibble(method = "C-CS cluster weight", AUPRC = ccs_boot)
  ) %>%
    dplyr::group_by(.data$method) %>%
    dplyr::summarise(
      ci_lo = stats::quantile(.data$AUPRC, 0.025, na.rm = TRUE),
      ci_hi = stats::quantile(.data$AUPRC, 0.975, na.rm = TRUE),
      .groups = "drop"
    )
  delta_ci <- stats::quantile(delta_boot, c(0.025, 0.975), na.rm = TRUE)

  boot_plot_data <- dplyr::bind_rows(
    tibble::tibble(method = "SuSiE baseline", AUPRC = baseline_boot),
    tibble::tibble(method = "C-CS cluster weight", AUPRC = ccs_boot)
  )

  p_boot <- ggplot2::ggplot(
    boot_plot_data,
    ggplot2::aes(x = .data$AUPRC, fill = .data$method, color = .data$method)
  ) +
    ggplot2::geom_histogram(
      ggplot2::aes(y = ggplot2::after_stat(density)),
      bins = 45,
      alpha = 0.35,
      position = "identity"
    ) +
    ggplot2::geom_vline(
      data = method_ci,
      ggplot2::aes(xintercept = .data$ci_lo, color = .data$method),
      linetype = "dashed",
      linewidth = 0.5
    ) +
    ggplot2::geom_vline(
      data = method_ci,
      ggplot2::aes(xintercept = .data$ci_hi, color = .data$method),
      linetype = "dashed",
      linewidth = 0.5
    ) +
    ggplot2::labs(
      x = "Bootstrap pooled AUPRC",
      y = "Density",
      fill = NULL,
      color = NULL,
      title = "Dataset bootstrap AUPRC",
      subtitle = sprintf(
        "annotation_r2 = %.1f; n = %d datasets; paired delta 95%% CI: [%.3f, %.3f]",
        annotation_r2_focus,
        length(shared_boot_ids),
        delta_ci[[1]],
        delta_ci[[2]]
      )
    ) +
    paper_theme()

  save_paper_png(
    p_boot,
    paste0("paper_bootstrap_auprc_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 7,
    height = 4.8
  )
  p_boot
} else {
  message("Skipping bootstrap plot: insufficient shared dataset IDs.")
}
```

# 5. PR Curves

```{r fig-pr-curves, fig.width=9, fig.height=4.8}
ccs_pr_bins <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "C-CS",
    .data$agg_method == "cluster_weight",
    !is.na(.data$annotation_r2)
  ) %>%
  dplyr::mutate(method = "C-CS cluster weight")

pr_r2_values <- sort(unique(ccs_pr_bins$annotation_r2))

baseline_pr_bins_raw <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "baseline-single",
    .data$use_case_id == "susine_vanilla",
    is.na(.data$agg_method)
  )

baseline_pr_bins <- dplyr::bind_rows(lapply(pr_r2_values, function(r2) {
  baseline_pr_bins_raw %>%
    dplyr::mutate(annotation_r2 = r2, method = "SuSiE baseline")
}))

pr_curve_data <- dplyr::bind_rows(baseline_pr_bins, ccs_pr_bins) %>%
  build_pr_curves(c("method", "annotation_r2"))

if (nrow(pr_curve_data) > 0L) {
  p_pr <- ggplot2::ggplot(
    pr_curve_data,
    ggplot2::aes(
      x = .data$recall,
      y = .data$precision,
      color = .data$method,
      group = .data$method
    )
  ) +
    ggplot2::geom_line(linewidth = 0.75) +
    ggplot2::facet_wrap(~ annotation_r2, nrow = 1, labeller = ggplot2::label_both) +
    ggplot2::scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
    ggplot2::scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
    ggplot2::labs(
      x = "Recall",
      y = "Precision",
      color = NULL,
      title = "Precision-recall curves for baseline SuSiE and C-CS cluster weight"
    ) +
    paper_theme()

  save_paper_png(
    p_pr,
    "paper_pr_curves_ccs_cluster_weight_by_r2.png",
    width = 9,
    height = 4.8
  )
  p_pr
} else {
  message("Skipping PR curves: no rows.")
}
```

# 6. Estimated Heritability

```{r fig-hg2, fig.width=8.5, fig.height=5}
if (!is.null(hg2_by_agg) && nrow(hg2_by_agg) > 0L) {
  if (!"spec_name" %in% names(hg2_by_agg) && "group_label" %in% names(hg2_by_agg)) {
    spec_map <- model_metrics %>%
      dplyr::distinct(group_key, spec_name)
    hg2_by_agg <- hg2_by_agg %>%
      dplyr::mutate(group_key = sub("^model_grid\\|", "", .data$group_label)) %>%
      dplyr::left_join(spec_map, by = "group_key") %>%
      dplyr::select(-dplyr::any_of("group_key"))
  }

  hg2_focus <- hg2_by_agg %>%
    dplyr::filter(
      .data$agg_method == "cluster_weight",
      .data$spec_name %in% spec_order,
      is_focus_annotation_row(.data$spec_name, .data$annotation_r2)
    ) %>%
    dplyr::transmute(
      spec_name = as.character(.data$spec_name),
      dataset_bundle_id = .data$dataset_bundle_id,
      hg2 = .data$hg2,
      series = "ensemble"
    )

  baseline_hg2 <- model_metrics %>%
    dplyr::filter(
      .data$spec_name == "baseline-single",
      .data$use_case_id == "susine_vanilla",
      .data$filtering == "purity_filtered",
      !is.na(.data$hg2)
    ) %>%
    dplyr::distinct(dataset_bundle_id, run_id, .keep_all = TRUE) %>%
    dplyr::transmute(
      spec_name = "baseline",
      dataset_bundle_id = .data$dataset_bundle_id,
      hg2 = .data$hg2,
      series = "baseline"
    )

  hg2_plot_data <- dplyr::bind_rows(hg2_focus, baseline_hg2) %>%
    dplyr::mutate(
      spec_name = factor(.data$spec_name, levels = c("baseline", spec_order)),
      series = factor(.data$series, levels = c("baseline", "ensemble"))
    )

  if (nrow(hg2_plot_data) > 0L) {
    p_hg2 <- ggplot2::ggplot(
      hg2_plot_data,
      ggplot2::aes(x = .data$spec_name, y = .data$hg2, fill = .data$series)
    ) +
      ggplot2::geom_boxplot(outlier.size = 0.4, linewidth = 0.35) +
      ggplot2::geom_hline(
        yintercept = true_h2,
        linetype = "dashed",
        color = "grey25",
        linewidth = 0.55
      ) +
      ggplot2::scale_fill_manual(
        values = c(baseline = "grey70", ensemble = "#4C78A8"),
        guide = "none"
      ) +
      ggplot2::labs(
        x = "Method",
        y = "Estimated h_g^2",
        title = "Estimated heritability under cluster-weight aggregation",
        subtitle = paste0("annotation_r2 = ", annotation_r2_focus, "; dashed line = true h2")
      ) +
      paper_theme() +
      ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))

    save_paper_png(
      p_hg2,
      paste0("paper_hg2_cluster_weight_r2_", format_r2_tag(annotation_r2_focus), ".png"),
      width = 8.5,
      height = 5
    )
    p_hg2
  } else {
    message("Skipping hg2 plot: no rows after filtering.")
  }
} else {
  message("Skipping hg2 plot: hg2_by_agg.csv is unavailable.")
}
```

# 7. PIP Calibration

```{r fig-pip-calibration, fig.width=6.8, fig.height=5}
pip_cal_with_se <- function(bins) {
  bins %>%
    dplyr::mutate(
      pip_bin = pmin(0.9, floor(.data$pip_threshold / 0.10) * 0.10)
    ) %>%
    dplyr::group_by(.data$method, .data$dataset_bundle_id, .data$pip_bin) %>%
    dplyr::summarise(
      n_causal = sum(.data$n_causal_at_bucket, na.rm = TRUE),
      n_noncausal = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      n_total = .data$n_causal + .data$n_noncausal,
      observed_frac = dplyr::if_else(
        .data$n_total > 0L,
        .data$n_causal / .data$n_total,
        NA_real_
      )
    ) %>%
    dplyr::group_by(.data$method, .data$pip_bin) %>%
    dplyr::summarise(
      mean_frac = mean(.data$observed_frac, na.rm = TRUE),
      se_frac = stats::sd(.data$observed_frac, na.rm = TRUE) /
        sqrt(sum(!is.na(.data$observed_frac))),
      n_datasets = sum(!is.na(.data$observed_frac)),
      total_causal = sum(.data$n_causal, na.rm = TRUE),
      total_noncausal = sum(.data$n_noncausal, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(
      total = .data$total_causal + .data$total_noncausal,
      ci_lo = pmax(0, .data$mean_frac - 1.96 * .data$se_frac),
      ci_hi = pmin(1, .data$mean_frac + 1.96 * .data$se_frac),
      pip_mid = .data$pip_bin + 0.05
    )
}

baseline_cal_bins <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "baseline-single",
    .data$use_case_id == "susine_vanilla",
    is.na(.data$agg_method)
  ) %>%
  dplyr::mutate(method = "SuSiE baseline")

ccs_cal_bins <- confusion_bins %>%
  dplyr::filter(
    .data$spec_name == "C-CS",
    near_focus(.data$annotation_r2, annotation_r2_focus),
    .data$agg_method == "cluster_weight"
  ) %>%
  dplyr::mutate(method = "C-CS cluster weight")

pip_cal_data <- dplyr::bind_rows(baseline_cal_bins, ccs_cal_bins) %>%
  pip_cal_with_se()

if (nrow(pip_cal_data) > 0L) {
  p_cal <- ggplot2::ggplot(
    pip_cal_data,
    ggplot2::aes(x = .data$pip_mid, y = .data$mean_frac, color = .data$method)
  ) +
    ggplot2::geom_abline(
      slope = 1,
      intercept = 0,
      linetype = "dashed",
      color = "grey45"
    ) +
    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = .data$ci_lo, ymax = .data$ci_hi),
      width = 0.018,
      linewidth = 0.35,
      alpha = 0.75
    ) +
    ggplot2::geom_line(linewidth = 0.7) +
    ggplot2::geom_point(size = 1.8) +
    ggplot2::coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
    ggplot2::labs(
      x = "PIP bin midpoint",
      y = "Observed causal fraction",
      color = NULL,
      title = "PIP calibration",
      subtitle = paste0("annotation_r2 = ", annotation_r2_focus)
    ) +
    paper_theme()

  save_paper_png(
    p_cal,
    paste0(
      "paper_pip_calibration_ccs_cluster_weight_r2_",
      format_r2_tag(annotation_r2_focus),
      ".png"
    ),
    width = 6.8,
    height = 5
  )
  p_cal
} else {
  message("Skipping PIP calibration: no rows.")
}
```

# 8. C-CS 4x4 vs 8x8 Scaling

```{r fig-ccs-resolution-scaling, fig.width=6.5, fig.height=4.7}
compute_ccs_resolution_oracle <- function(confusion_bins, focus = annotation_r2_focus) {
  ccs_individual <- confusion_bins %>%
    dplyr::filter(
      .data$spec_name == "C-CS",
      near_focus(.data$annotation_r2, focus),
      is.na(.data$agg_method),
      !is.na(.data$c_value),
      !is.na(.data$sigma_0_2_scalar)
    ) %>%
    dplyr::mutate(
      c_value = suppressWarnings(as.numeric(.data$c_value)),
      sigma_0_2_scalar_num = suppressWarnings(as.numeric(.data$sigma_0_2_scalar))
    )

  if (!nrow(ccs_individual)) return(tibble::tibble())

  out <- lapply(split(ccs_individual, ccs_individual$dataset_bundle_id), function(ds_bins) {
    c_vals <- sort(unique(stats::na.omit(ds_bins$c_value)))
    s_vals <- sort(unique(stats::na.omit(ds_bins$sigma_0_2_scalar_num)))
    if (!length(c_vals) || !length(s_vals)) return(tibble::tibble())

    resolution_rows <- lapply(c("4x4", "8x8"), function(resolution) {
      res_sizes <- as.integer(strsplit(resolution, "x", fixed = TRUE)[[1]])
      keep_c <- paper_select_axis_subset(c_vals, res_sizes[[1]], "c_value")
      keep_s <- paper_select_axis_subset(s_vals, res_sizes[[2]], "sigma_0_2_scalar_num")

      selected <- ds_bins %>%
        dplyr::filter(
          .data$c_value %in% keep_c,
          .data$sigma_0_2_scalar_num %in% keep_s
        )
      if (!nrow(selected)) return(tibble::tibble())

      per_run <- selected %>%
        dplyr::group_by(.data$dataset_bundle_id, .data$run_id) %>%
        dplyr::summarise(
          AUPRC = auprc_from_bins(dplyr::pick(dplyr::everything())),
          .groups = "drop"
        ) %>%
        dplyr::slice_max(.data$AUPRC, n = 1L, with_ties = FALSE) %>%
        dplyr::mutate(resolution = resolution)

      selected %>%
        dplyr::inner_join(
          per_run %>% dplyr::select(dataset_bundle_id, run_id, resolution),
          by = c("dataset_bundle_id", "run_id")
        )
    })
    dplyr::bind_rows(resolution_rows)
  })

  oracle_bins <- dplyr::bind_rows(out)
  if (!nrow(oracle_bins)) return(tibble::tibble())

  oracle_bins %>%
    dplyr::group_by(.data$resolution, .data$pip_threshold) %>%
    dplyr::summarise(
      n_causal_at_bucket = sum(.data$n_causal_at_bucket, na.rm = TRUE),
      n_noncausal_at_bucket = sum(.data$n_noncausal_at_bucket, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::group_by(.data$resolution) %>%
    dplyr::summarise(
      AUPRC = auprc_from_bins(dplyr::pick(dplyr::everything())),
      .groups = "drop"
    ) %>%
    dplyr::mutate(method = "oracle")
}

scaling_metrics <- compute_scaling_metrics(scaling_bins_pooled)

ccs_resolution_cluster <- scaling_metrics %>%
  dplyr::filter(
    .data$spec_name == "C-CS",
    near_focus(.data$annotation_r2, annotation_r2_focus),
    .data$agg_method == "cluster_weight",
    .data$resolution %in% c("4x4", "8x8")
  ) %>%
  dplyr::transmute(
    resolution = .data$resolution,
    method = "cluster weight",
    AUPRC = .data$AUPRC_mean
  )

ccs_resolution_oracle <- compute_ccs_resolution_oracle(confusion_bins)
ccs_resolution_plot_data <- dplyr::bind_rows(
  ccs_resolution_cluster,
  ccs_resolution_oracle
) %>%
  dplyr::mutate(
    resolution = factor(.data$resolution, levels = c("4x4", "8x8")),
    method = factor(.data$method, levels = c("cluster weight", "oracle"))
  )

if (nrow(ccs_resolution_plot_data) > 0L) {
  p_res <- ggplot2::ggplot(
    ccs_resolution_plot_data,
    ggplot2::aes(
      x = .data$resolution,
      y = .data$AUPRC,
      color = .data$method,
      group = .data$method
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.55
    ) +
    ggplot2::geom_line(linewidth = 0.75) +
    ggplot2::geom_point(size = 2.5) +
    ggplot2::labs(
      x = "C-CS grid resolution",
      y = "Pooled AUPRC",
      color = NULL,
      title = "C-CS grid resolution scaling",
      subtitle = paste0("annotation_r2 = ", annotation_r2_focus)
    ) +
    paper_theme()

  save_paper_png(
    p_res,
    paste0("paper_ccs_4x4_8x8_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 6.5,
    height = 4.7
  )
  p_res
} else {
  message("Skipping C-CS 4x4/8x8 plot: no rows.")
}
```

# 9. Compute vs AUPRC

```{r fig-compute-vs-auprc, fig.width=7.5, fig.height=5.2}
if (nrow(scaling_metrics) > 0L && "wall_time_sec" %in% names(model_metrics)) {
  compute_scaling_points <- scaling_metrics %>%
    dplyr::filter(
      .data$agg_method == "cluster_weight",
      .data$spec_name %in% spec_order,
      is_focus_annotation_row(.data$spec_name, .data$annotation_r2)
    )

  compute_times <- estimate_subscale_compute(compute_scaling_points, model_metrics)

  compute_plot_data <- compute_scaling_points %>%
    dplyr::left_join(
      compute_times,
      by = c("spec_name", "annotation_r2", "n_ensemble", "resolution")
    ) %>%
    dplyr::filter(is.finite(.data$mean_total_wall_sec)) %>%
    dplyr::mutate(
      compute_min = .data$mean_total_wall_sec / 60,
      scaling_point = dplyr::if_else(
        is.na(.data$resolution),
        paste0("n=", .data$n_ensemble),
        .data$resolution
      ),
      spec_name = factor(.data$spec_name, levels = spec_order)
    )

  baseline_compute_min <- baseline_compute_minutes(model_metrics)

  if (nrow(compute_plot_data) > 0L) {
    p_compute <- ggplot2::ggplot(
      compute_plot_data,
      ggplot2::aes(
        x = .data$compute_min,
        y = .data$AUPRC_mean,
        color = .data$spec_name
      )
    ) +
      ggplot2::geom_hline(
        yintercept = baseline_auprc,
        linetype = "dashed",
        color = "grey25",
        linewidth = 0.55
      ) +
      ggplot2::geom_point(size = 2, alpha = 0.85) +
      ggplot2::geom_point(
        data = dplyr::filter(compute_plot_data, !is.na(.data$resolution)),
        ggplot2::aes(shape = .data$scaling_point),
        size = 2.6,
        stroke = 0.8
      ) +
      ggplot2::annotate(
        "point",
        x = baseline_compute_min,
        y = baseline_auprc,
        shape = 8,
        size = 3.2,
        color = "black",
        stroke = 1
      ) +
      ggplot2::labs(
        x = "Mean compute time per dataset (minutes)",
        y = "Pooled AUPRC",
        color = "Ensemble",
        shape = "Grid resolution",
        title = "Compute versus AUPRC for cluster-weight aggregation",
        subtitle = paste0("A specs are annotation-independent; B/C specs use annotation_r2 = ",
                          annotation_r2_focus)
      ) +
      paper_theme()

    save_paper_png(
      p_compute,
      paste0(
        "paper_compute_vs_auprc_cluster_weight_r2_",
        format_r2_tag(annotation_r2_focus),
        ".png"
      ),
      width = 7.5,
      height = 5.2
    )
    p_compute
  } else {
    message("Skipping compute plot: no rows after joining compute times.")
  }
} else {
  message("Skipping compute plot: scaling bins or wall_time_sec unavailable.")
}
```

# 10. Output Check

```{r output-check}
expected_figures <- c(
  paste0("paper_delta_auprc_heatmap_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  "paper_ccs_aggregation_by_annotation_r2.png",
  paste0("paper_bootstrap_auprc_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  "paper_pr_curves_ccs_cluster_weight_by_r2.png",
  paste0("paper_hg2_cluster_weight_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  paste0(
    "paper_pip_calibration_ccs_cluster_weight_r2_",
    format_r2_tag(annotation_r2_focus),
    ".png"
  ),
  paste0("paper_ccs_4x4_8x8_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  paste0(
    "paper_compute_vs_auprc_cluster_weight_r2_",
    format_r2_tag(annotation_r2_focus),
    ".png"
  )
)

figure_paths <- file.path(fig_dir, expected_figures)
figure_exists <- file.exists(figure_paths)
figure_bytes <- rep(NA_real_, length(figure_paths))
figure_bytes[figure_exists] <- file.info(figure_paths[figure_exists])$size

figure_check <- tibble::tibble(
  figure = expected_figures,
  path = figure_paths,
  exists = figure_exists,
  bytes = figure_bytes
)

knitr::kable(figure_check, caption = "Expected paper figure outputs")

if (any(!figure_check$exists | is.na(figure_check$bytes) | figure_check$bytes <= 0)) {
  warning("Some expected figures were not written or are empty. See table above.")
}
```
