---
title: "Ensemble Scaling Experiment - Paper Figures"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    dev: pdf
editor_options:
  markdown:
    wrap: 72
---

Paper-focused figure renderer for the ensemble scaling simulation. Run
`prepare_results_workbook_ensemble_scaling_paper.Rmd` first on the HPC
consolidated outputs; this workbook then reads the compact plot-data RDS and
writes PNG figures only.

```{r setup, message=FALSE, fig.keep='none', dev='pdf'}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_ensemble_scaling_paper.Rmd")
library(here)
library(dplyr)
library(ggplot2)
library(readr)
library(knitr)

knitr::opts_chunk$set(dev = "pdf")
theme_set(theme_bw(base_size = 14))
```

# 0. User Configuration

```{r config}
job_name <- "ensemble_scaling_full"
# Must match the parent_job_id used in the prepare workbook.
parent_job_id <- "53520405"
output_root <- here("output")
write_component_pngs <- FALSE

job_dir <- file.path(output_root, "slurm_output", job_name, parent_job_id)
fig_dir <- file.path(job_dir, "figures", "paper_ensemble_scaling")
plot_data_dir <- file.path(fig_dir, "plot_data")
plot_data_rds <- file.path(plot_data_dir, "paper_ensemble_plot_data.rds")

if (!file.exists(plot_data_rds)) {
  stop(
    "Prepared paper plot-data RDS not found: ", plot_data_rds, "\n",
    "Run prepare_results_workbook_ensemble_scaling_paper.Rmd first.",
    call. = FALSE
  )
}
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

cat("Plot-data RDS:", plot_data_rds, "\n")
cat("Figure dir:", fig_dir, "\n")
```

# 1. Load Plot Data

```{r load-plot-data}
paper_plot_data <- readRDS(plot_data_rds)

annotation_r2_focus <- paper_plot_data$config$annotation_r2_focus
drop_aggs <- paper_plot_data$config$drop_aggs
bootstrap_B <- paper_plot_data$config$bootstrap_B
bootstrap_seed <- paper_plot_data$config$bootstrap_seed
true_h2 <- paper_plot_data$config$true_h2

list2env(paper_plot_data$labels, envir = environment())
list2env(paper_plot_data$refs, envir = environment())
list2env(paper_plot_data$data, envir = environment())
list2env(paper_plot_data$diagnostics, envir = environment())

cat("Prepared for causal annotation signal focus:", annotation_r2_focus, "\n")
cat("Baseline SuSiE AUPRC:", round(baseline_auprc, 4), "\n")
cat("Truth-warm vanilla AUPRC:", round(truth_warm_vanilla, 4), "\n")
cat("Baseline model-fit wall time:", round(baseline_compute_sec, 3), "sec\n")
```

```{r helpers}
format_r2_tag <- function(x) {
  gsub("\\.", "p", format(x, trim = TRUE, scientific = FALSE))
}

open_png_device <- function(path, width, height, res = 300, bg = "white") {
  if (requireNamespace("ragg", quietly = TRUE)) {
    return(ragg::agg_png(
      filename = path,
      width = width,
      height = height,
      units = "in",
      res = res,
      background = bg
    ))
  }

  if (isTRUE(grDevices::capabilities("cairo"))) {
    return(grDevices::png(
      filename = path,
      width = width,
      height = height,
      units = "in",
      res = res,
      bg = bg,
      type = "cairo"
    ))
  }

  stop(
    "No headless PNG device is available. Install the R package 'ragg' ",
    "or use an R build with Cairo support.",
    call. = FALSE
  )
}

save_paper_png <- function(plot, filename, width, height) {
  path <- file.path(fig_dir, filename)
  if (!isTRUE(write_component_pngs)) {
    return(invisible(path))
  }
  open_png_device(path, width = width, height = height, res = 300, bg = "white")
  on.exit(grDevices::dev.off(), add = TRUE)
  print(plot)
  invisible(path)
}

font_sizes <- list(
  base = 14,
  title = 25,
  subtitle = 13,
  axis_title = 23,
  axis_text = 22,
  strip = 11,
  legend_title = 22,
  legend_text = 20,
  caption = 10,
  tag = 30
)

paper_theme <- function(base_size = font_sizes$base) {
  ggplot2::theme_bw(base_size = base_size) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      plot.title.position = "plot",
      plot.title = ggplot2::element_text(
        size = font_sizes$title,
        hjust = 0.5
      ),
      plot.subtitle = ggplot2::element_text(size = font_sizes$subtitle),
      axis.title = ggplot2::element_text(size = font_sizes$axis_title),
      axis.text = ggplot2::element_text(size = font_sizes$axis_text),
      axis.title.x = ggplot2::element_text(
        size = font_sizes$axis_title,
        margin = ggplot2::margin(t = 10)
      ),
      axis.title.y = ggplot2::element_text(
        size = font_sizes$axis_title,
        margin = ggplot2::margin(r = 10)
      ),
      strip.text = ggplot2::element_text(size = font_sizes$strip),
      legend.title = ggplot2::element_text(size = font_sizes$legend_title),
      legend.text = ggplot2::element_text(size = font_sizes$legend_text),
      plot.caption = ggplot2::element_text(size = font_sizes$caption),
      legend.position = "bottom",
      legend.box = "vertical"
    )
}

compact_bottom_legend_theme <- function() {
  ggplot2::theme(
    legend.margin = ggplot2::margin(t = 0.5, r = 0, b = 1, l = 0),
    legend.box.margin = ggplot2::margin(t = -2, r = 0, b = -2, l = 0),
    legend.spacing.y = grid::unit(0.01, "cm"),
    legend.key.height = grid::unit(0.32, "cm")
  )
}

paper_orange <- "#D95F02"
paper_blue <- "#1F77B4"
paper_green <- "#2CA02C"
paper_red <- "#D62728"

gain_fill_scale <- function(values, name = "Delta AUPRC") {
  if (length(values) > 0L && min(values, na.rm = TRUE) >= 0) {
    return(ggplot2::scale_fill_gradientn(
      colors = c("#F7FBFF", "#DEEBF7", "#C6DBEF", "#9ECAE1",
                 "#6BAED6", "#3182BD", "#08519C"),
      name = name
    ))
  }

  ggplot2::scale_fill_gradient2(
    low = paper_orange,
    mid = "white",
    high = paper_blue,
    midpoint = 0,
    name = name
  )
}

heatmap_text_color <- function(values) {
  if (!length(values) || all(!is.finite(values))) {
    return(rep("black", length(values)))
  }
  rng <- range(values, na.rm = TRUE)
  if (!is.finite(diff(rng)) || diff(rng) <= 0) {
    return(rep("black", length(values)))
  }
  ifelse(values >= rng[[1]] + 0.75 * diff(rng), "white", "black")
}

reference_label_layer <- function(x, y, label, hjust = 1.02, vjust = -0.35,
                                  size = 3.2) {
  if (!is.finite(y)) return(NULL)
  ggplot2::annotate(
    "label",
    x = x,
    y = y,
    label = label,
    hjust = hjust,
    vjust = vjust,
    size = size,
    label.size = 0,
    label.padding = grid::unit(0.08, "lines"),
    fill = "white",
    alpha = 0.85,
    color = "grey20"
  )
}

shape_values_for <- function(labels) {
  shape_pool <- c(16, 17, 15, 18, 3, 4, 8, 7, 0, 1, 2, 5, 6,
                  9, 10, 11, 12, 13, 14, 19, 20, 21, 22, 23, 24, 25)
  stats::setNames(rep_len(shape_pool, length(labels)), labels)
}

sort_subscale_labels <- function(labels) {
  labels <- unique(as.character(labels))
  first_dim <- suppressWarnings(as.integer(sub("x.*$", "", labels)))
  second_dim <- suppressWarnings(as.integer(sub("^.*x", "", labels)))
  labels[order(second_dim, first_dim, labels, na.last = TRUE)]
}

tag_panel <- function(plot, tag) {
  plot +
    ggplot2::theme(
      plot.margin = grid::unit(c(2, 2, 2, 2), "pt")
    )
}

save_composite_png <- function(filename, panels, layout_matrix,
                               widths = NULL, heights = NULL,
                               width = 15, height = 20,
                               bottom_note = NULL,
                               bottom_note_size = font_sizes$caption,
                               bottom_note_height = NULL) {
  stopifnot(is.matrix(layout_matrix))
  if (is.null(widths)) widths <- rep(1, ncol(layout_matrix))
  if (is.null(heights)) heights <- rep(1, nrow(layout_matrix))

  path <- file.path(fig_dir, filename)
  open_png_device(path, width = width, height = height, res = 300, bg = "white")
  on.exit(grDevices::dev.off(), add = TRUE)

  grid::grid.newpage()
  note_height <- if (is.null(bottom_note)) {
    0
  } else if (is.null(bottom_note_height)) {
    0.08
  } else {
    bottom_note_height
  }
  grid::pushViewport(grid::viewport(
    y = grid::unit(1, "npc"),
    height = grid::unit(1 - note_height, "npc"),
    just = "top",
    layout = grid::grid.layout(
      nrow(layout_matrix),
      ncol(layout_matrix),
      widths = grid::unit(widths, "null"),
      heights = grid::unit(heights, "null")
    )
  ))

  panel_ids <- unique(as.vector(layout_matrix))
  panel_ids <- panel_ids[!is.na(panel_ids)]
  for (id in panel_ids) {
    pos <- which(layout_matrix == id, arr.ind = TRUE)
    grid::pushViewport(grid::viewport(
      layout.pos.row = range(pos[, "row"]),
      layout.pos.col = range(pos[, "col"])
    ))
    print(
      panels[[id]],
      vp = grid::viewport(
        x = grid::unit(0.5, "npc"),
        y = grid::unit(0.5, "npc"),
        width = grid::unit(0.95, "npc"),
        height = grid::unit(0.95, "npc")
      ),
      newpage = FALSE
    )
    grid::grid.text(
      id,
      x = grid::unit(0.02, "npc"),
      y = grid::unit(0.995, "npc"),
      just = c("left", "top"),
      gp = grid::gpar(fontface = "bold", fontsize = font_sizes$tag)
    )
    grid::popViewport()
  }
  grid::popViewport()
  if (!is.null(bottom_note)) {
    grid::grid.text(
      bottom_note,
      x = grid::unit(0.02, "npc"),
      y = grid::unit(0.012, "npc"),
      just = c("left", "bottom"),
      gp = grid::gpar(fontsize = bottom_note_size, lineheight = 1.05)
    )
  }
  invisible(path)
}
```

# 2. Manuscript Numbers

```{r manuscript-numbers}
knitr::kable(spec_key, caption = "Ensemble abbreviation key")

knitr::kable(
  ccs_compute_coordinates,
  digits = 4,
  caption = "C-CS compute plot coordinates for 4x4 and 8x8"
)

knitr::kable(
  pr_precision50_points,
  digits = 4,
  caption = "PR operating points: maximum recall with precision >= 50%"
)

knitr::kable(
  pr_same_recall_points,
  digits = 4,
  caption = "PR operating points: non-baseline rows at baseline 50%-precision recall"
)
```

# 3. Overall Ensemble Performance Heatmap

```{r fig-delta-heatmap, fig.width=11, fig.height=8}
if (nrow(heatmap_data) > 0L && is.finite(baseline_auprc)) {
  heatmap_plot_data <- heatmap_data %>%
    dplyr::mutate(
      tile_label = dplyr::if_else(
        is.finite(.data$delta_pct),
        sprintf("%+.3f (%+.0f%%)", .data$delta_auprc, .data$delta_pct),
        sprintf("%+.3f", .data$delta_auprc)
      ),
      tile_text_color = heatmap_text_color(.data$delta_auprc),
      spec_name = factor(.data$spec_name, levels = rev(spec_order)),
      agg_method = factor(.data$agg_method, levels = agg_order)
    )
  heatmap_highlight_data <- heatmap_plot_data %>%
    dplyr::filter(
      as.character(.data$spec_name) == "C-CS",
      as.character(.data$agg_method) %in% c("cluster_weight_credible", "cluster_weight_max")
    )

  y_levels <- levels(heatmap_plot_data$spec_name)
  y_prefix <- sub("-.*$", "", y_levels)
  y_separators <- if (length(y_prefix) > 1L) {
    which(y_prefix[-length(y_prefix)] != y_prefix[-1L]) + 0.5
  } else {
    numeric(0L)
  }
  oracle_x <- match("oracle", agg_order) - 0.5

  p_heatmap <- ggplot2::ggplot(
    heatmap_plot_data,
    ggplot2::aes(x = .data$agg_method, y = .data$spec_name, fill = .data$delta_auprc)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.65) +
    ggplot2::geom_tile(
      data = heatmap_highlight_data,
      ggplot2::aes(x = .data$agg_method, y = .data$spec_name),
      fill = NA,
      color = paper_red,
      linewidth = 2.2,
      inherit.aes = FALSE
    ) +
    ggplot2::geom_text(
      ggplot2::aes(label = .data$tile_label, color = .data$tile_text_color),
      size = 5
    ) +
    ggplot2::geom_hline(
      yintercept = y_separators,
      linewidth = 0.95,
      color = "grey20"
    ) +
    ggplot2::geom_vline(
      xintercept = oracle_x,
      linewidth = 0.95,
      color = "grey20"
    ) +
    ggplot2::scale_x_discrete(labels = agg_labels, drop = FALSE) +
    ggplot2::scale_color_identity() +
    gain_fill_scale(heatmap_plot_data$delta_auprc) +
    ggplot2::guides(
      fill = ggplot2::guide_colorbar(
        direction = "vertical",
        barheight = grid::unit(3.8, "cm"),
        barwidth = grid::unit(0.45, "cm"),
        title.position = "top"
      )
    ) +
    ggplot2::labs(
      x = "Aggregation method",
      y = "Ensemble method",
      title = "AUPRC gain over baseline SuSiE"
    ) +
    paper_theme() +
    ggplot2::theme(
      legend.position = "right",
      legend.direction = "vertical",
      legend.box = "vertical",
      legend.justification = "center"
    )

  save_paper_png(
    p_heatmap,
    paste0("paper_delta_auprc_heatmap_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 12,
    height = 8
  )
  invisible(NULL)
} else {
  message("Skipping heatmap: no rows after filtering.")
}
```

# 4. Best Methods vs Annotation Accuracy

```{r fig-ccs-by-r2, fig.width=7, fig.height=4.8}
# credible is rendered LAST so the leftover `p_ccs_r2` (reused as panel B of the
# composite) is the credible version, matching the credible-based panels C-F.
for (cm_suffix in c("max", "credible")) {
  cm <- paste0("cluster_weight_", cm_suffix)
  ccs_by_r2_keep <- c(cm, "oracle")
  ccs_by_r2_plot_data <- ccs_by_r2 %>%
    dplyr::filter(.data$agg_method %in% ccs_by_r2_keep) %>%
    dplyr::mutate(agg_method = factor(.data$agg_method, levels = ccs_by_r2_keep))

  if (!nrow(ccs_by_r2_plot_data)) {
    message("Skipping C-CS by annotation accuracy plot (", cm, "): no rows.")
    next
  }
  ccs_r2_label_x <- max(ccs_by_r2_plot_data$annotation_r2, na.rm = TRUE)

  p_ccs_r2 <- ggplot2::ggplot(
    ccs_by_r2_plot_data,
    ggplot2::aes(
      x = .data$annotation_r2,
      y = .data$AUPRC,
      color = .data$agg_method,
      group = .data$agg_method
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "solid",
      color = paper_orange,
      linewidth = 0.9
    ) +
    ggplot2::geom_hline(
      yintercept = truth_warm_vanilla,
      linetype = "dashed",
      color = paper_orange,
      linewidth = 0.9
    ) +
    ggplot2::geom_line(ggplot2::aes(linetype = .data$agg_method), linewidth = 1.2) +
    ggplot2::geom_point(size = 3.2, show.legend = FALSE) +
    reference_label_layer(
      ccs_r2_label_x,
      baseline_auprc,
      "SuSiE baseline",
      size = 6.4
    ) +
    reference_label_layer(
      ccs_r2_label_x,
      truth_warm_vanilla,
      "truth-warm SuSiE",
      # vjust = 0,
      size = 6.4
    ) +
    ggplot2::coord_cartesian(ylim = c(0.2, 0.4)) +
    ggplot2::scale_x_continuous(breaks = sort(unique(ccs_by_r2_plot_data$annotation_r2))) +
    ggplot2::scale_color_manual(
      values = stats::setNames(c(paper_blue, paper_blue), c(cm, "oracle")),
      labels = agg_labels[ccs_by_r2_keep],
      drop = FALSE
    ) +
    ggplot2::scale_linetype_manual(
      values = stats::setNames(c("solid", "dashed"), c(cm, "oracle")),
      labels = agg_labels[ccs_by_r2_keep],
      drop = FALSE
    ) +
    ggplot2::labs(
      x = expression("Causal annotation accuracy " * phi[a]),
      y = "AUPRC",
      color = "Aggregation",
      linetype = "Aggregation",
      title = "AUPRC by Annotation Accuracy"
    ) +
    paper_theme() +
    compact_bottom_legend_theme() +
    ggplot2::guides(
      color = "none",
      linetype = ggplot2::guide_legend(
        override.aes = list(color = paper_blue, linewidth = 1.8)
      )
    ) +
    ggplot2::theme(
      legend.key.width = grid::unit(2.6, "cm")
    )

  save_paper_png(
    p_ccs_r2,
    paste0("paper_ccs_aggregation_by_annotation_r2_", cm_suffix, ".png"),
    width = 7,
    height = 4.8
  )
}
```

# 5. Bootstrap AUPRC CI

```{r fig-bootstrap, fig.width=7, fig.height=4.8}
if (isTRUE(write_component_pngs) && nrow(boot_plot_data) > 0L) {
  p_boot <- ggplot2::ggplot(
    boot_plot_data,
    ggplot2::aes(x = .data$delta_AUPRC)
  ) +
    ggplot2::geom_histogram(
      ggplot2::aes(y = ggplot2::after_stat(density)),
      bins = 45,
      fill = paper_blue,
      color = "white",
      linewidth = 0.35
    ) +
    ggplot2::geom_vline(
      xintercept = 0,
      linetype = "dotted",
      color = "grey35",
      linewidth = 0.9
    ) +
    ggplot2::geom_vline(
      xintercept = delta_ci,
      linetype = "dashed",
      color = "grey20",
      linewidth = 0.9
    ) +
    ggplot2::geom_vline(
      xintercept = observed_delta,
      color = paper_blue,
      linewidth = 1.1
    ) +
    ggplot2::labs(
      x = "Paired bootstrap delta AUPRC (SuSiNE (C-CS) - SuSiE baseline)",
      y = "Density",
      title = "Paired dataset bootstrap gain",
      subtitle = sprintf(
        "n = %d datasets; observed delta = %.3f; 95%% CI: [%.3f, %.3f]",
        shared_boot_n,
        observed_delta,
        delta_ci[[1]],
        delta_ci[[2]]
      )
    ) +
    paper_theme()

  save_paper_png(
    p_boot,
    paste0("paper_bootstrap_auprc_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 7,
    height = 4.8
  )
  invisible(NULL)
} else {
  message("Skipping bootstrap plot: component PNG output is disabled or data are unavailable.")
}
```

# 6. PR Curves

```{r fig-pr-curves, fig.width=8, fig.height=5}
pr_plot_data <- pr_curve_counts %>%
  dplyr::select(method, annotation_r2, recall, precision) %>%
  dplyr::bind_rows(
    pr_curve_counts %>%
      dplyr::distinct(.data$method, .data$annotation_r2) %>%
      dplyr::mutate(recall = 0, precision = 1)
  ) %>%
  dplyr::mutate(
    method = dplyr::recode(
      .data$method,
      "C-CS cluster weight" = "SuSiNE (C-CS)"
    ),
    method = factor(.data$method, levels = c("SuSiE baseline", "SuSiNE (C-CS)")),
    annotation_label = sprintf("phi[a] == %.2f", as.numeric(.data$annotation_r2))
  ) %>%
  dplyr::arrange(.data$method, .data$annotation_r2, .data$recall, dplyr::desc(.data$precision))

if (nrow(pr_plot_data) > 0L) {
  p_pr <- ggplot2::ggplot(
    pr_plot_data,
    ggplot2::aes(x = .data$recall, y = .data$precision, color = .data$method)
  ) +
    ggplot2::geom_path(linewidth = 1.2) +
    ggplot2::facet_wrap(
      ~annotation_label,
      labeller = ggplot2::label_parsed
    ) +
    ggplot2::scale_color_manual(
      values = c("SuSiE baseline" = paper_orange, "SuSiNE (C-CS)" = paper_blue)
    ) +
    ggplot2::coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
    ggplot2::scale_x_continuous(
      breaks = c(0, 0.25, 0.5, 0.75, 1),
      labels = function(x) ifelse(x %in% c(0, 1), sprintf("%.1f", x), sprintf("%.2f", x))
    ) +
    ggplot2::labs(
      x = "Recall",
      y = "Precision",
      color = NULL,
      title = expression("Precision-recall curves by Annotation Accuracy " * phi[a])
    ) +
    paper_theme() +
    compact_bottom_legend_theme() +
    ggplot2::theme(
    strip.text = ggplot2::element_text(size = 22)
    )
  
  save_paper_png(
    p_pr,
    "paper_pr_curves_ccs_cluster_weight_by_r2.png",
    width = 8,
    height = 5
  )
  invisible(NULL)
} else {
  message("Skipping PR curves: no rows.")
}
```

# 7. Estimated Heritability

```{r fig-hg2, fig.width=4.8, fig.height=4.8}
if (nrow(hg2_plot_data) > 0L) {
  series_levels <- c("susie", "susine_perfit", "susine_ensemble")
  series_labels <- c(susie = "SuSiE",
                     susine_perfit = "SuSiNE\n(max-ELBO fit)",
                     susine_ensemble = "SuSiNE\n(ensemble)")
  hg2_plot_data <- hg2_plot_data %>%
    dplyr::filter(.data$series %in% series_levels) %>%
    dplyr::mutate(
      method_label = factor(
        unname(series_labels[as.character(.data$series)]),
        levels = unname(series_labels[series_levels])
      ),
      series = factor(.data$series, levels = series_levels)
    )

  p_hg2 <- ggplot2::ggplot(
    hg2_plot_data,
    ggplot2::aes(x = .data$method_label, y = .data$hg2, fill = .data$series)
  ) +
    ggplot2::geom_boxplot(outlier.size = 1.3, linewidth = 0.75) +
    ggplot2::geom_hline(
      yintercept = true_h2,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.9
    ) +
    reference_label_layer(
      tail(levels(hg2_plot_data$method_label), 1L),
      true_h2,
      sprintf("true h2 = %.2f", true_h2),
      size = 6.4,
      hjust = 1.2
    ) +
    ggplot2::scale_fill_manual(
      values = c(susie = paper_orange, susine_perfit = "#9BBAD7",
                 susine_ensemble = paper_blue),
      guide = "none"
    ) +
    ggplot2::labs(
      x = "Method",
      y = expression("Estimated " * h[g]^2),
      title = "Estimated heritability"
    ) +
    paper_theme() +
    ggplot2::theme(
      axis.text.x = ggplot2::element_text(
        angle = 0,
        hjust = 0.5,
        size = font_sizes$axis_text
      )
    )

  save_paper_png(
    p_hg2,
    paste0("paper_hg2_cluster_weight_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 4.8,
    height = 4.8
  )
  invisible(NULL)
} else {
  message("Skipping hg2 plot: no rows after filtering.")
}
```

# 8. PIP Calibration

```{r fig-pip-calibration, fig.width=6.8, fig.height=5}
if (nrow(pip_cal_data) > 0L) {
  pip_cal_plot_data <- pip_cal_data %>%
    dplyr::mutate(
      method = dplyr::recode(
        .data$method,
        "C-CS cluster weight" = "SuSiNE (C-CS)"
      ),
      method = factor(.data$method, levels = c("SuSiE baseline", "SuSiNE (C-CS)"))
    )

  p_cal <- ggplot2::ggplot(
    pip_cal_plot_data,
    ggplot2::aes(x = .data$pip_mid, y = .data$mean_frac, color = .data$method)
  ) +
    ggplot2::geom_abline(
      slope = 1,
      intercept = 0,
      linetype = "dashed",
      color = "grey35",
      linewidth = 0.9
    ) +
    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = .data$lo, ymax = .data$hi),
      width = 0.015,
      linewidth = 0.6
    ) +
    ggplot2::geom_point(size = 3.2) +
    ggplot2::geom_line(linewidth = 1.05) +
    ggplot2::coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
    ggplot2::scale_color_manual(
      values = c("SuSiE baseline" = paper_orange, "SuSiNE (C-CS)" = paper_blue)
    ) +
    ggplot2::labs(
      x = "Mean PIP bin",
      y = "Observed causal",
      color = NULL,
      title = "PIP calibration"
    ) +
    paper_theme() +
    compact_bottom_legend_theme()

  save_paper_png(
    p_cal,
    paste0(
      "paper_pip_calibration_ccs_cluster_weight_r2_",
      format_r2_tag(annotation_r2_focus),
      ".png"
    ),
    width = 6.8,
    height = 5
  )
  invisible(NULL)
} else {
  message("Skipping PIP calibration plot: no rows.")
}
```

# 9. C-CS 4x4 vs 8x8 Scaling

```{r fig-ccs-resolution, fig.width=6.5, fig.height=4.7}
if (isTRUE(write_component_pngs) && nrow(ccs_resolution_plot_data) > 0L) {
  ccs_resolution_plot_data <- ccs_resolution_plot_data %>%
    dplyr::mutate(
      resolution = factor(.data$resolution, levels = c("4x4", "8x8")),
      method = factor(.data$method, levels = c("cluster weight", "oracle"))
    )

  ccs_resolution_label_x <- tail(levels(ccs_resolution_plot_data$resolution), 1L)

  p_res <- ggplot2::ggplot(
    ccs_resolution_plot_data,
    ggplot2::aes(
      x = .data$resolution,
      y = .data$AUPRC,
      color = .data$method,
      group = .data$method
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.9
    ) +
    ggplot2::geom_line(linewidth = 1.2) +
    ggplot2::geom_point(size = 4) +
    reference_label_layer(
      ccs_resolution_label_x,
      baseline_auprc,
      "SuSiE baseline",
      size = 6.4
    ) +
    ggplot2::labs(
      x = "C-CS grid resolution",
      y = "AUPRC",
      color = NULL,
      title = "AUPRC by Ensemble Size"
    ) +
    paper_theme() +
    compact_bottom_legend_theme()

  save_paper_png(
    p_res,
    paste0("paper_ccs_4x4_8x8_r2_", format_r2_tag(annotation_r2_focus), ".png"),
    width = 6.5,
    height = 4.7
  )
  invisible(NULL)
} else {
  message("Skipping C-CS 4x4/8x8 plot: component PNG output is disabled or data are unavailable.")
}
```

# 10. Compute vs AUPRC

```{r fig-compute-vs-auprc, fig.width=8.8, fig.height=5}
if (nrow(compute_plot_data) > 0L) {
  subscale_fraction_map <- c(
    "4x1" = "1/16",
    "8x1" = "1/8",
    "16x1" = "1/4",
    "32x1" = "1/2",
    "64x1" = "1",
    "4x2" = "1/4",
    "4x4" = "1/4",
    "8x4" = "1",
    "8x8" = "1"
  )
  subscale_fraction_order <- c("1/16", "1/8", "1/4", "1/2", "1")
  compute_second_breaks <- c(0, 1, 5, 10, 30, 60, 120, 240)
  compute_plot_data <- compute_plot_data %>%
    dplyr::mutate(
      compute_sec = 60 * .data$compute_min,
      subscale_label = as.character(.data$subscale_label),
      subscale_fraction = dplyr::recode(
        .data$subscale_label,
        !!!subscale_fraction_map,
        .default = .data$subscale_label
      ),
      spec_name = factor(.data$spec_name, levels = spec_order)
    )
  subscale_fraction_levels <- c(
    subscale_fraction_order,
    sort(setdiff(unique(compute_plot_data$subscale_fraction), subscale_fraction_order))
  )
  fraction_shape_values <- c(
    "1/16" = 16,
    "1/8" = 17,
    "1/4" = 15,
    "1/2" = 18,
    "1" = 0
  )
  if (length(setdiff(subscale_fraction_levels, names(fraction_shape_values))) > 0L) {
    extra_fraction_levels <- setdiff(subscale_fraction_levels, names(fraction_shape_values))
    fraction_shape_values <- c(
      fraction_shape_values,
      shape_values_for(extra_fraction_levels)
    )
  }
  fraction_shape_values <- fraction_shape_values[subscale_fraction_levels]
  compute_ccs_label_data <- compute_plot_data %>%
    dplyr::filter(as.character(.data$spec_name) == "C-CS") %>%
    dplyr::mutate(
      dot_label = paste0("C-CS, ", as.character(.data$subscale_fraction))
    )
  compute_label_x <- max(compute_plot_data$compute_sec, na.rm = TRUE)

  p_compute <- ggplot2::ggplot(
    compute_plot_data,
    ggplot2::aes(
      x = .data$compute_sec,
      y = .data$AUPRC_mean,
      color = .data$spec_name,
      shape = .data$subscale_fraction
    )
  ) +
    ggplot2::geom_hline(
      yintercept = baseline_auprc,
      linetype = "dashed",
      color = "grey25",
      linewidth = 0.9
    ) +
    ggplot2::geom_point(size = 5, alpha = 0.9, stroke = 1.5) +
    ggplot2::annotate(
      "point",
      x = baseline_compute_sec,
      y = baseline_auprc,
      shape = 8,
      size = 5,
      color = "black",
      stroke = 1.4
    ) +
    ggplot2::geom_text(
      data = compute_ccs_label_data,
      ggplot2::aes(
        x = .data$compute_sec,
        y = .data$AUPRC_mean,
        label = .data$dot_label
      ),
      color = "grey20",
      hjust = -0.08,
      vjust = 1.7,
      size = 7,
      show.legend = FALSE,
      check_overlap = TRUE,
      inherit.aes = FALSE
    ) +
    reference_label_layer(
      compute_label_x,
      baseline_auprc,
      "SuSiE baseline",
      size = 7
    ) +
    ggplot2::labs(
      x = "Mean wall time per dataset (seconds)",
      y = "AUPRC",
      color = "Ensemble",
      shape = "Fraction of full ensemble",
      title = "AUPRC by Compute Cost"
    ) +
    ggplot2::scale_x_continuous(
      trans = scales::pseudo_log_trans(base = 10, sigma = 1),
      breaks = compute_second_breaks,
      labels = as.character(compute_second_breaks)
    ) +
    ggplot2::scale_shape_manual(
      values = fraction_shape_values,
      breaks = subscale_fraction_levels
    ) +
    ggplot2::guides(
      color = ggplot2::guide_legend(
        ncol = 2,
        byrow = TRUE,
        order = 1,
        override.aes = list(shape = 16, size = 5)
      ),
      shape = ggplot2::guide_legend(
        ncol = 2,
        byrow = TRUE,
        order = 2,
        override.aes = list(color = "grey20", size = 5)
      )
    ) +
    paper_theme() +
    ggplot2::theme(
      legend.position = "right",
      legend.box = "vertical",
      legend.direction = "vertical",
      legend.title = ggplot2::element_text(size = font_sizes$legend_title),
      legend.text = ggplot2::element_text(size = font_sizes$legend_text),
      legend.key.size = grid::unit(0.75, "cm"),
      legend.spacing.y = grid::unit(0.03, "cm"),
      legend.box.spacing = grid::unit(0.08, "cm"),
      plot.title = ggplot2::element_text(
        size = font_sizes$title,
        hjust = 0.35
      )
    )

  save_paper_png(
    p_compute,
    paste0(
      "paper_compute_vs_auprc_cluster_weight_r2_",
      format_r2_tag(annotation_r2_focus),
      ".png"
    ),
    width = 9.8,
    height = 5.2
  )
  invisible(NULL)
} else {
  message("Skipping compute plot: no rows after joining compute times.")
}
```

# 11. Composite Main Figure

The bootstrap panel is intentionally left out of the main composite. It is
better summarized in text using the paired delta CI from the prep workbook,
while the figure space is used for performance patterns, operating behavior,
calibration, heritability, and model-fit time.

```{r fig-composite-main, fig.width=17.55, fig.height=25.48}
composite_required <- c(
  "p_heatmap", "p_ccs_r2", "p_pr",
  "p_cal", "p_hg2", "p_compute"
)
missing_composite <- composite_required[!vapply(
  composite_required,
  exists,
  logical(1L),
  envir = knitr::knit_global()
)]

if (length(missing_composite) == 0L) {
  composite_note <- paste0(
    "Method key: \n",
    "Prefixes: A = annotation-agnostic SuSiE; ",
    "B = SuSiNE-EB; C = SuSiNE fixed-priors.\n",
    "Suffixes: R = restarts; F = refinement; S = prior-variance grid; ",
    "C = annotation-scale grid. \n",
    "*Exception: C-CSR, see caption"
  )

  heatmap_panels <- list(
    A = tag_panel(p_heatmap, "A")
  )

  heatmap_layout <- rbind(rep("A", 1L))

  heatmap_path <- save_composite_png(
    filename = "paper_ensemble_scaling_composite_heatmap.png",
    panels = heatmap_panels,
    layout_matrix = heatmap_layout,
    widths = 1,
    heights = 1,
    width = 18,
    height = 10.2,
    bottom_note = composite_note,
    bottom_note_size = 24,
    bottom_note_height = 0.15
  )

  performance_panels <- list(
    B = tag_panel(p_ccs_r2, "B"),
    C = tag_panel(p_cal, "C"),
    D = tag_panel(p_pr, "D"),
    E = tag_panel(p_hg2, "E"),
    F = tag_panel(p_compute, "F")
  )

  composite_cols <- 25L
  performance_layout <- rbind(
    c(rep("B", 12L), rep("C", 13L)),
    rep(NA_character_, composite_cols),
    rep("D", composite_cols),
    rep(NA_character_, composite_cols),
    c(rep("E", 10L), rep("F", 15L))
  )

  performance_path <- save_composite_png(
    filename = "paper_ensemble_scaling_composite_performance.png",
    panels = performance_panels,
    layout_matrix = performance_layout,
    widths = rep(1, composite_cols),
    heights = c(1.75, 0.12, 2.25, 0.12, 2.22),
    width = 19.8,
    height = 20.7,
    bottom_note = composite_note,
    bottom_note_size = 24,
    bottom_note_height = 0.10
  )

  cat("Saved heatmap composite figure:", heatmap_path, "\n")
  cat("Saved performance composite figure:", performance_path, "\n")
  invisible(NULL)
} else {
  message(
    "Skipping composite figure; missing component plots: ",
    paste(missing_composite, collapse = ", ")
  )
}
```

# 11b. Ensemble Multimodality

Per-ensemble (per-locus) 5-number summaries of how multimodal each ensemble is:
cluster counts under the two locked-in cluster-weight metrics, top-10 Jaccard
overlap, and the largest pairwise distance between any two fits. Read directly
from `consolidated/multimodal_metrics.csv`; falls back to whatever columns the
run produced (credible/max columns require a run with the extended
`compute_multimodal_metrics`).

```{r fig-multimodality, fig.width=13, fig.height=4.5}
consolidated_dir <- file.path(job_dir, "consolidated")
mm_path <- file.path(consolidated_dir, "multimodal_metrics.csv")
mm_spec_path <- file.path(consolidated_dir, "model_metrics_full.csv")

if (file.exists(mm_path)) {
  mm <- readr::read_csv(mm_path, show_col_types = FALSE)

  # Attach spec_name via group_key (strip "model_grid|" from group_label).
  if (!"spec_name" %in% names(mm) && "group_label" %in% names(mm) &&
      file.exists(mm_spec_path)) {
    spec_map <- readr::read_csv(mm_spec_path, show_col_types = FALSE) %>%
      dplyr::distinct(.data$group_key, .data$spec_name)
    mm <- mm %>%
      dplyr::mutate(group_key = sub("^model_grid\\|", "", .data$group_label)) %>%
      dplyr::left_join(spec_map, by = "group_key")
  }

  # Use the credible/max metric set when the run produced it; otherwise fall
  # back to the JSD-based columns. Either way it's a tight 5-facet panel.
  have_new <- all(c("n_clusters_credible", "n_clusters_max") %in% names(mm))
  metric_labels <- if (have_new) c(
    n_clusters_credible = "# clusters (credible 0.05)",
    n_clusters_max      = "# clusters (max-shift 0.10)",
    jaccard_top10       = "Jaccard top-10",
    max_credible_dist   = "max credible dist.",
    max_maxshift_dist   = "max PIP L-inf dist."
  ) else c(
    n_clusters          = "# clusters (JSD 0.15)",
    n_clusters_jsd_050  = "# clusters (JSD 0.50)",
    jaccard_top10       = "Jaccard top-10",
    max_jsd             = "max JSD dist.",
    mean_pip_var        = "mean PIP var."
  )
  metrics_present <- intersect(names(metric_labels), names(mm))

  if (length(metrics_present) && "spec_name" %in% names(mm)) {
    mm_long <- mm %>%
      dplyr::filter(!is.na(.data$spec_name)) %>%
      dplyr::select(dplyr::any_of(c("spec_name", metrics_present))) %>%
      tidyr::pivot_longer(dplyr::any_of(metrics_present),
                          names_to = "metric", values_to = "value") %>%
      dplyr::filter(!is.na(.data$value)) %>%
      dplyr::mutate(
        spec_name = factor(.data$spec_name, levels = spec_order),
        metric = factor(unname(metric_labels[.data$metric]),
                        levels = unname(metric_labels[metrics_present])),
        group = sub("-.*$", "", as.character(.data$spec_name))
      ) %>%
      dplyr::filter(!is.na(.data$spec_name))

    # Stacked wide panels (one metric per row), specs as horizontal abbreviations
    # along the bottom, boxes tinted by ensemble family (A/B/C).
    p_mm <- ggplot2::ggplot(
        mm_long, ggplot2::aes(x = .data$spec_name, y = .data$value)) +
      ggplot2::geom_boxplot(
        ggplot2::aes(fill = .data$group),
        alpha = 0.6, width = 0.66, outlier.size = 0.45, linewidth = 0.45) +
      ggplot2::facet_wrap(~ metric, ncol = 1, scales = "free_y") +
      ggplot2::scale_fill_manual(
        values = c(A = paper_orange, B = "#9BBAD7", C = paper_blue),
        labels = c(A = "A: SuSiE", B = "B: EB SuSiNE", C = "C: grid SuSiNE"),
        name = NULL, drop = TRUE,
        guide = ggplot2::guide_legend(nrow = 1)) +
      ggplot2::labs(x = NULL, y = NULL,
                    title = "Ensemble multimodality across loci") +
      paper_theme() +
      ggplot2::theme(
        legend.position = "top",
        axis.text.x = ggplot2::element_text(angle = 0, hjust = 0.5, size = 9),
        panel.spacing.y = grid::unit(0.6, "lines"),
        panel.grid.minor = ggplot2::element_blank(),
        strip.text = ggplot2::element_text(face = "bold")
      )

    ggplot2::ggsave(
      file.path(fig_dir, "paper_ensemble_multimodality.png"),
      p_mm, width = 12, height = 11.5, dpi = 300,
      device = if (requireNamespace("ragg", quietly = TRUE)) ragg::agg_png else NULL
    )
    cat("Wrote paper_ensemble_multimodality.png; metrics:",
        paste(metrics_present, collapse = ", "), "\n")
  } else {
    message("Multimodality figure skipped: no usable metrics / spec_name.")
  }
} else {
  message("Multimodality figure skipped: ", mm_path, " not found.")
}
```

# 12. Final Printed Summaries

```{r final-summaries}
bootstrap_summary <- tibble::tibble()
if (nrow(boot_plot_data) > 0L && is.finite(observed_delta)) {
  delta_boot <- boot_plot_data$delta_AUPRC
  p_two_sided <- 2 * min(
    mean(delta_boot <= 0, na.rm = TRUE),
    mean(delta_boot >= 0, na.rm = TRUE)
  )
  bootstrap_summary <- tibble::tibble(
    comparison = "SuSiNE C-CS cluster weight - SuSiE baseline",
    annotation_r2 = annotation_r2_focus,
    shared_datasets = shared_boot_n,
    bootstrap_reps = length(delta_boot),
    observed_delta_AUPRC = observed_delta,
    ci_low = unname(delta_ci[[1]]),
    ci_high = unname(delta_ci[[2]]),
    p_two_sided = min(1, p_two_sided),
    significant_95pct = is.finite(delta_ci[[1]]) &&
      is.finite(delta_ci[[2]]) &&
      (delta_ci[[1]] > 0 || delta_ci[[2]] < 0)
  )
}

if (nrow(bootstrap_summary) > 0L) {
  knitr::kable(
    bootstrap_summary,
    digits = 4,
    caption = "Paired bootstrap AUPRC delta test"
  )
} else {
  cat("Paired bootstrap AUPRC delta test unavailable: no bootstrap data.\n")
}

hg2_comparison <- tibble::tibble()
hg2_summary <- tibble::tibble()
if (exists("hg2_plot_data") && nrow(hg2_plot_data) > 0L) {
  hg2_wide <- hg2_plot_data %>%
    dplyr::mutate(
      method = dplyr::case_when(
        as.character(.data$series) == "susie" ~ "baseline_susie",
        as.character(.data$series) == "susine_ensemble" ~ "susine_ccs",
        TRUE ~ NA_character_
      )
    ) %>%
    dplyr::filter(!is.na(.data$method), is.finite(.data$hg2)) %>%
    dplyr::group_by(.data$dataset_bundle_id, .data$method) %>%
    dplyr::summarise(mean_hg2 = mean(.data$hg2, na.rm = TRUE), .groups = "drop") %>%
    tidyr::pivot_wider(names_from = method, values_from = mean_hg2)

  for (nm in c("baseline_susie", "susine_ccs")) {
    if (!nm %in% names(hg2_wide)) hg2_wide[[nm]] <- NA_real_
  }

  hg2_comparison <- hg2_wide %>%
    dplyr::filter(is.finite(.data$baseline_susie), is.finite(.data$susine_ccs)) %>%
    dplyr::mutate(
      delta_hg2 = .data$susine_ccs - .data$baseline_susie,
      category = dplyr::case_when(
        .data$delta_hg2 > 0.01 ~ "higher than baseline SuSiE",
        abs(.data$delta_hg2) <= 0.01 ~ "near baseline SuSiE (+/- 0.01)",
        .data$delta_hg2 < -0.01 ~ "less than baseline SuSiE",
        TRUE ~ "unclassified"
      )
    )

  hg2_summary <- hg2_comparison %>%
    dplyr::count(.data$category, name = "n_datasets") %>%
    dplyr::mutate(
      pct_datasets = 100 * .data$n_datasets / sum(.data$n_datasets),
      total_datasets = sum(.data$n_datasets)
    ) %>%
    dplyr::arrange(factor(
      .data$category,
      levels = c(
        "higher than baseline SuSiE",
        "near baseline SuSiE (+/- 0.01)",
        "less than baseline SuSiE",
        "unclassified"
      )
    ))
}

if (nrow(hg2_summary) > 0L) {
  knitr::kable(
    hg2_summary,
    digits = 2,
    caption = paste0(
      "Dataset-level mean ensemble heritability versus baseline SuSiE",
      " (C-CS; mean over available ensemble/annotation rows)"
    )
  )
} else {
  cat("Heritability comparison unavailable: no matched baseline/C-CS hg2 rows.\n")
}
```

# 13. Output Check

```{r output-check}
component_figures <- c(
  paste0("paper_delta_auprc_heatmap_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  "paper_ccs_aggregation_by_annotation_r2.png",
  paste0("paper_bootstrap_auprc_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  "paper_pr_curves_ccs_cluster_weight_by_r2.png",
  paste0("paper_hg2_cluster_weight_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  paste0(
    "paper_pip_calibration_ccs_cluster_weight_r2_",
    format_r2_tag(annotation_r2_focus),
    ".png"
  ),
  paste0("paper_ccs_4x4_8x8_r2_", format_r2_tag(annotation_r2_focus), ".png"),
  paste0(
    "paper_compute_vs_auprc_cluster_weight_r2_",
    format_r2_tag(annotation_r2_focus),
    ".png"
  )
)
composite_figures <- c(
  "paper_ensemble_scaling_composite_heatmap.png",
  "paper_ensemble_scaling_composite_performance.png"
)
expected_figures <- c(
  if (isTRUE(write_component_pngs)) component_figures else character(0L),
  composite_figures
)

figure_paths <- file.path(fig_dir, expected_figures)
figure_check <- tibble::tibble(
  figure = expected_figures,
  exists = file.exists(figure_paths),
  bytes = ifelse(file.exists(figure_paths), file.info(figure_paths)$size, NA_real_)
)

knitr::kable(figure_check, caption = "Expected paper figure outputs")
```
