---
title: "Annotation Contamination Sensitivity - Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Visualization workbook for the reviewer annotation-contamination sensitivity
analysis. It reads the combined tables written by
`collect_results_workbook_annotation_contamination.Rmd`, generates the primary
metric curves, and saves paper-facing figures and CSVs.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_annotation_contamination.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(knitr)

theme_set(
  theme_bw(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey92"),
      strip.background = element_rect(fill = "grey95", color = NA),
      strip.text = element_text(face = "bold"),
      legend.position = "bottom",
      plot.title = element_text(face = "bold")
    )
)
```

# Load Data

```{r load}
output_root <- here("output")
combined_dir <- file.path(output_root, "annotation_contamination_sensitivity", "combined")
fig_dir <- file.path(output_root, "annotation_contamination_sensitivity", "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

read_req <- function(filename) {
  path <- file.path(combined_dir, filename)
  if (!file.exists(path)) stop("Missing required file: ", path)
  readr::read_csv(path, show_col_types = FALSE)
}

metric_summary <- read_req("metric_summary.csv")
baseline_summary <- read_req("baseline_summary.csv")
annotation_diagnostics <- read_req("annotation_diagnostics_full.csv")
annotation_gate <- read_req("annotation_gate.csv")
lambda0_gate <- read_req("lambda0_gate.csv")
prior_gate <- read_req("prior_gate.csv")

arm_labels <- c(
  null = "Null-block contamination",
  causal = "Causal-block positive control"
)

metric_summary <- metric_summary %>%
  dplyr::mutate(
    arm_label = dplyr::recode(.data$arm, !!!arm_labels),
    lambda = as.numeric(.data$annotation_contamination_lambda)
  )

baseline_summary <- baseline_summary %>%
  dplyr::mutate(
    arm_label = dplyr::recode(.data$arm, !!!arm_labels)
  )

annotation_diagnostics <- annotation_diagnostics %>%
  dplyr::mutate(
    arm_label = dplyr::recode(.data$arm, !!!arm_labels),
    lambda = as.numeric(.data$annotation_contamination_lambda)
  )

cat("Combined dir:", combined_dir, "\n")
cat("Figure dir:", fig_dir, "\n")
```

# Validation Summary

```{r validation-summary}
kable(annotation_gate, digits = 4, caption = "Annotation construction gate by arm and lambda")
kable(lambda0_gate, digits = 4, caption = "Lambda-zero cross-arm metric gate")
kable(prior_gate, digits = 4, caption = "Fixed-c prior gate")
```

# Metric Curves

```{r metric-curves, fig.width=8, fig.height=5}
p_auprc <- ggplot(
  metric_summary,
  aes(x = .data$lambda, y = .data$AUPRC, color = .data$arm_label, group = .data$arm_label)
) +
  geom_hline(
    data = baseline_summary,
    aes(yintercept = .data$AUPRC, color = .data$arm_label),
    linetype = "dashed",
    linewidth = 0.65,
    show.legend = FALSE
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = expression(lambda),
    y = "Pooled AUPRC",
    color = "Arm",
    title = "Annotation contamination sensitivity: pooled AUPRC",
    caption = "Dashed horizontal lines show the annotation-free SuSiE-equivalent baseline."
  )

ggsave(file.path(fig_dir, "auprc_by_lambda.png"), p_auprc, width = 8, height = 5, dpi = 220)
p_auprc
```

```{r recall75-curve, fig.width=8, fig.height=5}
p_recall <- ggplot(
  metric_summary,
  aes(x = .data$lambda, y = .data$recall_at_precision, color = .data$arm_label, group = .data$arm_label)
) +
  geom_hline(
    data = baseline_summary,
    aes(yintercept = .data$recall_at_precision, color = .data$arm_label),
    linetype = "dashed",
    linewidth = 0.65,
    show.legend = FALSE
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  labs(
    x = expression(lambda),
    y = "Recall at precision >= 0.75",
    color = "Arm",
    title = "Annotation contamination sensitivity: recall at precision 0.75",
    caption = "Dashed horizontal lines show the annotation-free SuSiE-equivalent baseline."
  )

ggsave(file.path(fig_dir, "recall75_by_lambda.png"), p_recall, width = 8, height = 5, dpi = 220)
p_recall
```

# Combined Metric Figure

```{r combined-metric-figure, fig.width=9, fig.height=5}
metric_long <- metric_summary %>%
  dplyr::select(.data$arm_label, .data$lambda, .data$AUPRC, .data$recall_at_precision) %>%
  tidyr::pivot_longer(
    cols = c("AUPRC", "recall_at_precision"),
    names_to = "metric",
    values_to = "value"
  ) %>%
  dplyr::mutate(
    metric = dplyr::recode(
      .data$metric,
      AUPRC = "Pooled AUPRC",
      recall_at_precision = "Recall at precision >= 0.75"
    )
  )

baseline_long <- baseline_summary %>%
  dplyr::select(.data$arm_label, .data$AUPRC, .data$recall_at_precision) %>%
  tidyr::pivot_longer(
    cols = c("AUPRC", "recall_at_precision"),
    names_to = "metric",
    values_to = "value"
  ) %>%
  dplyr::mutate(
    metric = dplyr::recode(
      .data$metric,
      AUPRC = "Pooled AUPRC",
      recall_at_precision = "Recall at precision >= 0.75"
    )
  )

p_metrics <- ggplot(
  metric_long,
  aes(x = .data$lambda, y = .data$value, color = .data$arm_label, group = .data$arm_label)
) +
  geom_hline(
    data = baseline_long,
    aes(yintercept = .data$value, color = .data$arm_label),
    linetype = "dashed",
    linewidth = 0.65,
    show.legend = FALSE
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~metric, scales = "free_y") +
  labs(
    x = expression(lambda),
    y = NULL,
    color = "Arm",
    title = "Annotation sensitivity metrics",
    caption = "Dashed horizontal lines show the annotation-free SuSiE-equivalent baseline."
  )

ggsave(file.path(fig_dir, "annotation_sensitivity_metrics.png"), p_metrics, width = 9, height = 5, dpi = 240)
p_metrics
```

# Annotation-Z Diagnostics

```{r annotation-z-diagnostics, fig.width=9, fig.height=5}
diag_summary <- annotation_diagnostics %>%
  dplyr::group_by(.data$arm_label, .data$lambda) %>%
  dplyr::summarise(
    cor_a_z_null = mean(.data$cor_a_z_null, na.rm = TRUE),
    cor_a_z_causal = mean(.data$cor_a_z_causal, na.rm = TRUE),
    cor_a_z_anchor_low_pip = mean(.data$cor_a_z_anchor_low_pip, na.rm = TRUE),
    .groups = "drop"
  )

diag_long <- diag_summary %>%
  tidyr::pivot_longer(
    cols = c("cor_a_z_null", "cor_a_z_causal", "cor_a_z_anchor_low_pip"),
    names_to = "diagnostic",
    values_to = "correlation"
  ) %>%
  dplyr::mutate(
    diagnostic = dplyr::recode(
      .data$diagnostic,
      cor_a_z_null = "Null variants",
      cor_a_z_causal = "Causal variants",
      cor_a_z_anchor_low_pip = "Anchor low-PIP variants"
    )
  )

p_diag <- ggplot(
  diag_long,
  aes(x = .data$lambda, y = .data$correlation, color = .data$diagnostic, group = .data$diagnostic)
) +
  geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8) +
  facet_wrap(~arm_label) +
  labs(
    x = expression(lambda),
    y = "Mean cor(annotation, z)",
    color = "Block",
    title = "Annotation/z alignment diagnostics"
  )

ggsave(file.path(fig_dir, "annotation_z_correlations.png"), p_diag, width = 9, height = 5, dpi = 220)
p_diag
```

# RMS And Untouched-Block Diagnostics

```{r construction-diagnostics, fig.width=8, fig.height=5}
construction_long <- annotation_gate %>%
  dplyr::mutate(
    arm_label = dplyr::recode(.data$arm, !!!arm_labels),
    lambda = as.numeric(.data$annotation_contamination_lambda),
    max_active_rms_rel_error_plot = pmax(.data$max_active_rms_rel_error, 1e-16)
  )

p_rms <- ggplot(
  construction_long,
  aes(x = .data$lambda, y = .data$max_active_rms_rel_error_plot, color = .data$arm_label, group = .data$arm_label)
) +
  geom_hline(yintercept = 1e-8, linetype = "dashed", color = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_y_log10() +
  labs(
    x = expression(lambda),
    y = "Max active-block RMS relative error",
    color = "Arm",
    title = "RMS construction invariant"
  )

ggsave(file.path(fig_dir, "rms_invariant_by_lambda.png"), p_rms, width = 8, height = 5, dpi = 220)
p_rms
```

# Export Plot Data

```{r export-plot-data}
readr::write_csv(metric_summary, file.path(fig_dir, "metric_summary_plot_data.csv"))
readr::write_csv(baseline_summary, file.path(fig_dir, "baseline_summary_plot_data.csv"))
readr::write_csv(diag_summary, file.path(fig_dir, "annotation_z_diagnostic_plot_data.csv"))
readr::write_csv(construction_long, file.path(fig_dir, "construction_gate_plot_data.csv"))

cat("Saved figures and plot data to:", fig_dir, "\n")
```