---
title: "SuSiNE Pilot Phase A — Results Visualization"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
editor_options:
  markdown:
    wrap: 72
---

Summary: Phase A screening results. Evaluates 9 use cases (single fits
+ small c/tau grids for functional arms) across 10 M1-stratified
matrices. Uses confusion bins and dataset metrics from the completed
SLURM run to answer key questions driving Phase B design.

Primary questions:

1. Which non-functional use cases perform best?
2. How does EB compare with fixed-variance counterparts?
3. Which z-score metric best complements M1 for predicting difficulty?
4. Do use case rankings change across dataset-difficulty strata?
5. Do use case rankings change across simulation settings?
6. How do functional priors (mu vs pi) compare to no-prior baselines?
7. Does max-ELBO aggregation help functional priors? Where are gains largest?

# Section 1: Setup

```{r setup, message=FALSE, warning=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_phase_a.Rmd")
library(here)
library(devtools)
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(purrr)
library(broom)
```

## 1.1 User Options

```{r user-options}
save_plots_to_file <- TRUE
show_in_rstudio_plots <- TRUE
fdr_threshold <- 0.5  # FDR level for Power@FDR summaries
# If TRUE, error when functional_mu rows with c_value = NA are present.
strict_fmu_c_na <- FALSE

if (isTRUE(getOption("knitr.in.progress"))) {
  knitr::opts_chunk$set(fig.ext = "png")
} else if (show_in_rstudio_plots && interactive()) {
  knitr::opts_chunk$set(fig.show = "hide", dev = "RStudioGD", fig.ext = "png")
} else {
  knitr::opts_chunk$set(fig.ext = "png")
}

maybe_show_plot <- function(p) {
  if (show_in_rstudio_plots && interactive()) {
    try(print(p), silent = TRUE)
  }
}

save_or_show <- function(p, filename, width = 10, height = 7, dpi = 300,
                         subdir = "phase_a") {
  maybe_show_plot(p)
  if (save_plots_to_file) {
    filepath <- file.path(plot_dirs[[subdir]], filename)
    tryCatch(
      {
        ggsave(filepath, p, width = width, height = height, dpi = dpi, bg = "white")
        cat("Saved:", filename, "\n")
      },
      error = function(e) cat("Error saving:", filename, "-", conditionMessage(e), "\n")
    )
  } else {
    tryCatch(print(p), error = function(e) {
      cat("Cannot display plot. Set save_plots_to_file = TRUE.\n")
    })
  }
}

#' Extract Power at a given FDR threshold from power-FDR curves.
#' For each group, finds the highest power where FDR <= fdr_threshold.
power_at_fdr <- function(curves_df, fdr_thresh = fdr_threshold,
                         group_cols = "use_case_id") {
  curves_df %>%
    dplyr::filter(fdr <= fdr_thresh) %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_cols))) %>%
    dplyr::summarise(
      power_at_fdr = max(power, na.rm = TRUE),
      pip_at_fdr = min(pip_threshold[power == max(power)], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(fdr_threshold = fdr_thresh)
}
```

## 1.2 Job Configuration

```{r job-config}
# UPDATE THESE for your completed Phase A run
job_name <- Sys.getenv("TS_JOB_NAME", unset = "pilot_phase_a")
parent_job_id <- Sys.getenv("TS_PARENT_JOB_ID", unset = "48997267")
output_root <- here("output")

paths <- test_susine::resolve_job_paths(
  job_name = job_name,
  parent_job_id = parent_job_id,
  output_root = output_root
)
```

## 1.3 Plot Directories

```{r create-plot-dirs}
plot_base_dir <- file.path(paths$aggregated_dir, "plots_phase_a")
plot_dirs <- list(
  phase_a = file.path(plot_base_dir, "main"),
  difficulty = file.path(plot_base_dir, "difficulty"),
  functional = file.path(plot_base_dir, "functional"),
  stratified = file.path(plot_base_dir, "stratified")
)

for (d in plot_dirs) {
  dir.create(d, showWarnings = FALSE, recursive = TRUE)
}
cat("Plot directories created under:", plot_base_dir, "\n")
```

## 1.4 Load Data

```{r load-data, message=FALSE}
# --- Confusion bins (primary data source) ---
confusion_bins_path <- file.path(paths$aggregated_dir, "confusion_bins.csv")
confusion_detailed_path <- file.path(paths$aggregated_dir, "confusion_matrix_detailed.csv")

# Prefer the detailed confusion matrix if already built; else use raw bins
if (file.exists(confusion_detailed_path)) {
  confusion <- readr::read_csv(confusion_detailed_path, show_col_types = FALSE)
  cat("Loaded pre-built confusion_matrix_detailed:", nrow(confusion), "rows\n")
} else if (file.exists(confusion_bins_path)) {
  confusion <- readr::read_csv(confusion_bins_path, show_col_types = FALSE)
  cat("Loaded raw confusion_bins:", nrow(confusion), "rows\n")
} else {
  stop("No confusion data found. Run collect_results_workbook.Rmd first.")
}

# --- Dataset metrics ---
dataset_metrics_path <- file.path(paths$aggregated_dir, "dataset_metrics.csv")
if (file.exists(dataset_metrics_path)) {
  dataset_metrics <- readr::read_csv(dataset_metrics_path, show_col_types = FALSE)
  cat("Loaded dataset_metrics:", nrow(dataset_metrics), "rows\n")
} else {
  warning("dataset_metrics.csv not found — difficulty analyses will be skipped.")
  dataset_metrics <- NULL
}

# --- Model metrics (for ELBO-based analysis) ---
model_metrics_path <- file.path(paths$aggregated_dir, "model_metrics.csv")
if (file.exists(model_metrics_path)) {
  model_metrics <- readr::read_csv(model_metrics_path, show_col_types = FALSE)
  cat("Loaded model_metrics:", nrow(model_metrics), "rows\n")
} else {
  warning("model_metrics.csv not found — ELBO-based analyses will use confusion bins only.")
  model_metrics <- NULL
}

# --- Run table (for joining extra columns when needed) ---
run_table_path <- file.path(paths$run_history_dir, "run_table.csv")
if (file.exists(run_table_path)) {
  run_table <- readr::read_csv(run_table_path, show_col_types = FALSE)
  cat("Loaded run_table:", nrow(run_table), "rows\n")
} else {
  run_table <- NULL
}
```

## 1.5 Define Use Case Groups and Palettes

```{r use-case-definitions}
# --- Use case groupings ---
# Non-functional (no annotation dependency)
nonfunctional_ids <- c(
  "susie_vanilla", "susie_ash_fixed", "susie_ash_eb",
  "susie_inf", "susine_vanilla", "susie_eb", "susine_eb"
)

# Fixed-variance non-functional (exclude EB use cases)
fixed_nonfunctional_ids <- c(
  "susie_vanilla", "susie_ash_fixed", "susie_inf", "susine_vanilla"
)

# EB use cases and their fixed counterparts
eb_pairs <- tibble::tribble(
  ~eb_id,        ~fixed_id,         ~pair_label,
  "susie_eb",    "susie_vanilla",   "SuSiE vanilla",
  "susine_eb",   "susine_vanilla",  "SuSiNE vanilla",
  "susie_ash_eb","susie_ash_fixed", "SuSiE-ash"
)

# Functional use cases
functional_ids <- c("susine_functional_pi", "susine_functional_mu")

# All use case IDs
all_ids <- c(nonfunctional_ids, functional_ids)

# --- Labels from catalog ---
all_labels <- test_susine::get_use_case_labels(all_ids)

# --- Color palette (9 use cases) ---
# Blues: vanilla SuSiE family
# Greens: ash family
# Purples: inf
# Oranges: functional
use_case_palette <- c(
  "susie_vanilla"       = "#2171b5",
  "susine_vanilla"      = "#6baed6",
  "susie_eb"            = "#08306b",
  "susine_eb"           = "#4292c6",
  "susie_ash_fixed"     = "#238b45",
  "susie_ash_eb"        = "#74c476",
  "susie_inf"           = "#6a3d9a",
  "susine_functional_pi"= "#e6550d",
  "susine_functional_mu"= "#d62728"
)

# Short labels for compact plots
short_labels <- c(
  "susie_vanilla"       = "Vanilla",
  "susine_vanilla"      = "Vanilla (susine)",
  "susie_eb"            = "EB",
  "susine_eb"           = "EB (susine)",
  "susie_ash_fixed"     = "Ash-fixed",
  "susie_ash_eb"        = "Ash-EB",
  "susie_inf"           = "Inf",
  "susine_functional_pi"= "Func-pi",
  "susine_functional_mu"= "Func-mu"
)

# --- Shared ggplot theme ---
theme_phase_a <- function(base_size = 13) {
  ggplot2::theme_minimal(base_size = base_size) +
    ggplot2::theme(
      legend.position = "bottom",
      legend.title = ggplot2::element_text(size = 12, face = "bold"),
      legend.text = ggplot2::element_text(size = 11),
      strip.text = ggplot2::element_text(size = 12, face = "bold"),
      axis.title = ggplot2::element_text(size = 13),
      plot.title = ggplot2::element_text(size = 15, face = "bold", hjust = 0.5),
      plot.subtitle = ggplot2::element_text(size = 11, hjust = 0.5, color = "gray40")
    )
}
```

## 1.6 Data Prep — Filter to Base Fits

Phase A has `single`, `c_grid`, and `tau_grid` exploration methods. For
most analyses (Q1–Q5), we focus on **base fits** (`agg_method` is NA or
`explore_method == "base"`). The functional prior analyses (Q6–Q7) use
both base fits and max-ELBO aggregated fits.

```{r data-prep}
# Base individual fits only (no aggregated rows)
confusion_base <- confusion %>%
  dplyr::filter(is.na(agg_method) | agg_method == "none")

# For non-functional use cases: single exploration only (no c/tau grid)
confusion_nonfunc <- confusion_base %>%
  dplyr::filter(use_case_id %in% nonfunctional_ids)

# All fits including aggregated (max_elbo rows) for functional analysis
confusion_with_agg <- confusion %>%
  dplyr::filter(
    is.na(agg_method) | agg_method == "none" | agg_method == "max_elbo"
  )

cat("Base fits:", nrow(confusion_base), "rows\n")
cat("Non-functional base fits:", nrow(confusion_nonfunc), "rows\n")
cat("With aggregation rows:", nrow(confusion_with_agg), "rows\n")

# Quick sanity: use case counts
confusion_base %>%
  dplyr::count(use_case_id) %>%
  dplyr::arrange(dplyr::desc(n)) %>%
  print()
```

---

# Section 2: Overview Heatmaps

Top-level view of all 9 use cases across p\* and y_noise. Shows both
AUPRC and Power @ FDR to orient the reader before detailed analyses.

## 2.1 AUPRC Heatmap: Use Case × p_star × y_noise

```{r heatmap-auprc, fig.width=16, fig.height=10}
auprc_heat_data <- test_susine::compute_auprc_from_confusion(
  confusion_base %>% dplyr::filter(use_case_id %in% all_ids),
  group_vars = c("use_case_id", "p_star", "y_noise")
)

if (nrow(auprc_heat_data) > 0) {
  p_heat_auprc <- ggplot2::ggplot(
    auprc_heat_data,
    ggplot2::aes(x = factor(p_star), y = use_case_id, fill = AUPRC)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.5) +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.3f", AUPRC)),
                       size = 3, color = "white") +
    ggplot2::facet_wrap(~paste0("y_noise = ", y_noise), ncol = 3) +
    ggplot2::scale_fill_viridis_c(option = "magma", direction = -1) +
    ggplot2::scale_y_discrete(labels = short_labels) +
    ggplot2::labs(
      title = "AUPRC: All Use Cases \u00d7 p* \u00d7 y_noise",
      x = "p*", y = NULL, fill = "AUPRC"
    ) +
    theme_phase_a() +
    ggplot2::theme(legend.position = "right",
                   axis.text.y = ggplot2::element_text(size = 10))

  save_or_show(p_heat_auprc, "heatmap_auprc_pstar_ynoise.png",
               width = 16, height = 10, subdir = "phase_a")
}
```

## 2.2 Power @ FDR Heatmap: Use Case × p_star × y_noise

```{r heatmap-power-fdr, fig.width=16, fig.height=10}
curves_all <- test_susine::compute_power_fdr_curves(
  confusion_base %>% dplyr::filter(use_case_id %in% all_ids),
  group_vars = c("use_case_id", "p_star", "y_noise")
)

power_fdr_heat <- power_at_fdr(curves_all, fdr_thresh = fdr_threshold,
                                group_cols = c("use_case_id", "p_star", "y_noise"))

if (nrow(power_fdr_heat) > 0) {
  p_heat_power <- ggplot2::ggplot(
    power_fdr_heat,
    ggplot2::aes(x = factor(p_star), y = use_case_id, fill = power_at_fdr)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.5) +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.3f", power_at_fdr)),
                       size = 3, color = "white") +
    ggplot2::facet_wrap(~paste0("y_noise = ", y_noise), ncol = 3) +
    ggplot2::scale_fill_viridis_c(option = "viridis", direction = -1) +
    ggplot2::scale_y_discrete(labels = short_labels) +
    ggplot2::labs(
      title = paste0("Power @ FDR \u2264 ", fdr_threshold,
                     ": All Use Cases \u00d7 p* \u00d7 y_noise"),
      x = "p*", y = NULL, fill = "Power"
    ) +
    theme_phase_a() +
    ggplot2::theme(legend.position = "right",
                   axis.text.y = ggplot2::element_text(size = 10))

  save_or_show(p_heat_power, "heatmap_power_at_fdr_pstar_ynoise.png",
               width = 16, height = 10, subdir = "phase_a")
}
```

---

# Section 3: Q1 — Non-Functional Use Case Ranking

Which non-functional use cases perform best, split by simulation
settings?

## 3.1 AUPRC by p_star

```{r q1-auprc-pstar, fig.width=12, fig.height=6}
auprc_by_pstar <- test_susine::compute_auprc_from_confusion(
  confusion_nonfunc,
  group_vars = c("use_case_id", "p_star")
)

p_q1a <- ggplot2::ggplot(
  auprc_by_pstar,
  ggplot2::aes(x = factor(p_star), y = AUPRC, color = use_case_id,
               group = use_case_id)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_manual(
    values = use_case_palette[nonfunctional_ids],
    labels = short_labels[nonfunctional_ids],
    name = "Use Case"
  ) +
  ggplot2::labs(
    title = "AUPRC by p* — Non-Functional Use Cases",
    x = "p* (number of causal effects)", y = "AUPRC"
  ) +
  theme_phase_a() +
  ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

save_or_show(p_q1a, "q1_auprc_by_pstar.png", width = 12, height = 6,
             subdir = "phase_a")
```

## 3.2 AUPRC by y_noise

```{r q1-auprc-ynoise, fig.width=12, fig.height=6}
auprc_by_ynoise <- test_susine::compute_auprc_from_confusion(
  confusion_nonfunc,
  group_vars = c("use_case_id", "y_noise")
)

p_q1b <- ggplot2::ggplot(
  auprc_by_ynoise,
  ggplot2::aes(x = y_noise, y = AUPRC, color = use_case_id,
               group = use_case_id)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_manual(
    values = use_case_palette[nonfunctional_ids],
    labels = short_labels[nonfunctional_ids],
    name = "Use Case"
  ) +
  ggplot2::labs(
    title = "AUPRC by Noise Level — Non-Functional Use Cases",
    x = "y_noise (residual variance fraction)", y = "AUPRC"
  ) +
  theme_phase_a() +
  ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

save_or_show(p_q1b, "q1_auprc_by_ynoise.png", width = 12, height = 6,
             subdir = "phase_a")
```

## 3.3 Power @ FDR by p_star

```{r q1-power-pstar, fig.width=12, fig.height=6}
curves_nonfunc_settings <- test_susine::compute_power_fdr_curves(
  confusion_nonfunc,
  group_vars = c("use_case_id", "p_star")
)

power_q1 <- power_at_fdr(curves_nonfunc_settings, fdr_thresh = fdr_threshold,
                          group_cols = c("use_case_id", "p_star"))

p_q1c <- ggplot2::ggplot(
  power_q1,
  ggplot2::aes(x = factor(p_star), y = power_at_fdr, color = use_case_id,
               group = use_case_id)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_manual(
    values = use_case_palette[nonfunctional_ids],
    labels = short_labels[nonfunctional_ids],
    name = "Use Case"
  ) +
  ggplot2::labs(
    title = paste0("Power @ FDR \u2264 ", fdr_threshold, " by p*"),
    x = "p*", y = paste0("Power (FDR \u2264 ", fdr_threshold, ")")
  ) +
  theme_phase_a() +
  ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

save_or_show(p_q1c, "q1_power_at_fdr_by_pstar.png", width = 12, height = 6,
             subdir = "phase_a")
```

## 3.4 Power vs FDR Curves Faceted by p_star × y_noise

```{r q1-power-fdr-faceted, fig.width=14, fig.height=10}
curves_by_settings <- test_susine::compute_power_fdr_curves(
  confusion_nonfunc,
  group_vars = c("use_case_id", "p_star", "y_noise")
) %>%
  dplyr::filter(pip_threshold >= 0.05, pip_threshold <= 0.95)

pstar_show <- c(1, 5, 20)
curves_facet <- curves_by_settings %>%
  dplyr::filter(p_star %in% pstar_show) %>%
  dplyr::mutate(
    p_star_label = paste0("p* = ", p_star),
    y_noise_label = paste0("y_noise = ", y_noise)
  )

if (nrow(curves_facet) > 0) {
  p_q1d <- ggplot2::ggplot(
    curves_facet,
    ggplot2::aes(x = fdr, y = power, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.8) +
    ggplot2::facet_grid(p_star_label ~ y_noise_label, scales = "free_y") +
    ggplot2::scale_color_manual(
      values = use_case_palette[nonfunctional_ids],
      labels = short_labels[nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "Power vs FDR — Non-Functional, by p* and y_noise",
      x = "FDR", y = "Power"
    ) +
    ggplot2::coord_cartesian(xlim = c(0, 1)) +
    theme_phase_a() +
    ggplot2::guides(color = ggplot2::guide_legend(ncol = 3))

  save_or_show(p_q1d, "q1_power_fdr_faceted_pstar_ynoise.png",
               width = 14, height = 10, subdir = "stratified")
}
```

---

# Section 4: Q2 — EB vs Fixed-Variance Comparison

Don't anchor y-axis at 0 — the range from min to max AUPRC matters more
than the distance from 0. Also show Power @ FDR.

```{r q2-eb-data}
# For each EB pair: compute AUPRC and Power@FDR at each sigma_0_2_scalar
eb_comparison_data <- list()

for (i in seq_len(nrow(eb_pairs))) {
  pair <- eb_pairs[i, ]

  # Fixed-variance: group by sigma_0_2_scalar
  fixed_conf <- confusion_nonfunc %>%
    dplyr::filter(use_case_id == pair$fixed_id)

  fixed_auprc <- test_susine::compute_auprc_from_confusion(
    fixed_conf, group_vars = c("use_case_id", "sigma_0_2_scalar")
  )

  fixed_curves <- test_susine::compute_power_fdr_curves(
    fixed_conf, group_vars = c("use_case_id", "sigma_0_2_scalar")
  )
  fixed_power <- power_at_fdr(fixed_curves, fdr_thresh = fdr_threshold,
                               group_cols = c("use_case_id", "sigma_0_2_scalar"))

  fixed_combined <- dplyr::left_join(fixed_auprc, fixed_power,
                                      by = c("use_case_id", "sigma_0_2_scalar")) %>%
    dplyr::mutate(
      prior_var_type = paste0("fixed (", sigma_0_2_scalar, ")"),
      pair_label = pair$pair_label
    )

  # EB: single value
  eb_conf <- confusion_nonfunc %>%
    dplyr::filter(use_case_id == pair$eb_id)

  eb_auprc <- test_susine::compute_auprc_from_confusion(
    eb_conf, group_vars = "use_case_id"
  )
  eb_curves <- test_susine::compute_power_fdr_curves(
    eb_conf, group_vars = "use_case_id"
  )
  eb_power <- power_at_fdr(eb_curves, fdr_thresh = fdr_threshold,
                            group_cols = "use_case_id")

  eb_combined <- dplyr::left_join(eb_auprc, eb_power, by = "use_case_id") %>%
    dplyr::mutate(
      prior_var_type = "EB",
      pair_label = pair$pair_label,
      sigma_0_2_scalar = NA_real_
    )

  eb_comparison_data[[i]] <- dplyr::bind_rows(fixed_combined, eb_combined)
}

eb_comparison <- dplyr::bind_rows(eb_comparison_data)

cat("=== EB vs Fixed Variance Comparison ===\n")
eb_comparison %>%
  dplyr::select(pair_label, prior_var_type, AUPRC, power_at_fdr) %>%
  dplyr::arrange(pair_label, prior_var_type) %>%
  print(n = 30)
```

```{r q2-eb-plot-auprc, fig.width=14, fig.height=5}
# AUPRC comparison — dot plot with range-based y-axis (no zero anchor)
p_q2a <- ggplot2::ggplot(
  eb_comparison,
  ggplot2::aes(x = prior_var_type, y = AUPRC, color = pair_label)
) +
  ggplot2::geom_point(size = 4) +
  ggplot2::geom_line(ggplot2::aes(group = pair_label), linewidth = 0.5,
                     linetype = "dotted") +
  ggplot2::facet_wrap(~pair_label, scales = "free_x") +
  ggplot2::labs(
    title = "AUPRC: EB vs Fixed Prior Variance",
    x = "Prior Variance Setting", y = "AUPRC"
  ) +
  theme_phase_a() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, hjust = 1),
                 legend.position = "none")

save_or_show(p_q2a, "q2_eb_vs_fixed_auprc.png", width = 14, height = 5,
             subdir = "phase_a")
```

```{r q2-eb-plot-power, fig.width=14, fig.height=5}
p_q2b <- ggplot2::ggplot(
  eb_comparison,
  ggplot2::aes(x = prior_var_type, y = power_at_fdr, color = pair_label)
) +
  ggplot2::geom_point(size = 4) +
  ggplot2::geom_line(ggplot2::aes(group = pair_label), linewidth = 0.5,
                     linetype = "dotted") +
  ggplot2::facet_wrap(~pair_label, scales = "free_x") +
  ggplot2::labs(
    title = paste0("Power @ FDR \u2264 ", fdr_threshold,
                   ": EB vs Fixed Prior Variance"),
    x = "Prior Variance Setting",
    y = paste0("Power (FDR \u2264 ", fdr_threshold, ")")
  ) +
  theme_phase_a() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, hjust = 1),
                 legend.position = "none")

save_or_show(p_q2b, "q2_eb_vs_fixed_power.png", width = 14, height = 5,
             subdir = "phase_a")
```

---

# Section 5: Q3 — Dataset Difficulty Metrics

Aggregate per dataset_bundle (not per matrix) to get ~450 observation
points (10 matrices x 5 p_star x 3 y_noise x 3 seeds). Use:

1. Overall per-bundle AUPRC (single curve from combined confusion bins
   across all base fits)
2. Mean per-bundle AUPRC across use cases (for reference)

Then test `M1 + high-LD + z-score` combinations and visualize each
predictor marginally (after accounting for the other two).

```{r q3-difficulty-setup}
# Per-bundle AUPRC:
# 1) overall_AUPRC from combined confusion bins across all base fits.
# 2) mean_AUPRC across use-case-specific AUPRC values.
if ("dataset_bundle_id" %in% names(confusion_base)) {
  auprc_per_bundle_usecase <- test_susine::compute_auprc_from_confusion(
    confusion_base,
    group_vars = c("dataset_bundle_id", "use_case_id")
  )
  auprc_bundle_mean <- auprc_per_bundle_usecase %>%
    dplyr::group_by(dataset_bundle_id) %>%
    dplyr::summarise(mean_AUPRC = mean(AUPRC, na.rm = TRUE), .groups = "drop")

  auprc_bundle_overall <- test_susine::compute_auprc_from_confusion(
    confusion_base,
    group_vars = c("dataset_bundle_id")
  ) %>%
    dplyr::rename(overall_AUPRC = AUPRC)

  cat("Per-bundle use-case AUPRC rows:", nrow(auprc_per_bundle_usecase), "\n")
  cat("Unique bundles (overall AUPRC):", nrow(auprc_bundle_overall), "\n")
} else {
  warning("dataset_bundle_id not in confusion data; falling back to matrix_id aggregation.")
  auprc_per_bundle_usecase <- test_susine::aggregate_auprc_by_dataset_usecase(confusion_base)
  auprc_bundle_mean <- auprc_per_bundle_usecase %>%
    dplyr::group_by(matrix_id) %>%
    dplyr::summarise(mean_AUPRC = mean(AUPRC, na.rm = TRUE), .groups = "drop")
  auprc_bundle_overall <- test_susine::compute_auprc_from_confusion(
    confusion_base,
    group_vars = c("matrix_id")
  ) %>%
    dplyr::rename(overall_AUPRC = AUPRC)
}
```

```{r q3-join-metrics}
if (!is.null(dataset_metrics)) {
  dm_cols <- c(
    "M1",
    "high_ld_count_095",
    "high_ld_count_095_per_snp",
    "high_ld_frac_095",
    "z_topk_ratio",
    "z_max_abs",
    "z_count_abs_gt_3",
    "z_eff_signals"
  )

  if ("dataset_bundle_id" %in% names(dataset_metrics) &&
      "dataset_bundle_id" %in% names(auprc_bundle_overall)) {
    ds_diff_bundle <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(c("dataset_bundle_id", dm_cols))) %>%
      dplyr::inner_join(auprc_bundle_overall, by = "dataset_bundle_id") %>%
      dplyr::left_join(auprc_bundle_mean, by = "dataset_bundle_id")
  } else if ("matrix_id" %in% names(dataset_metrics)) {
    ds_diff_bundle <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(c("matrix_id", dm_cols))) %>%
      dplyr::group_by(matrix_id) %>%
      dplyr::summarise(dplyr::across(dplyr::any_of(dm_cols), ~ mean(.x, na.rm = TRUE)),
                       .groups = "drop") %>%
      dplyr::inner_join(auprc_bundle_overall, by = "matrix_id") %>%
      dplyr::left_join(auprc_bundle_mean, by = "matrix_id")
  } else {
    ds_diff_bundle <- NULL
    warning("Cannot join dataset_metrics.")
  }

  if (!is.null(ds_diff_bundle)) {
    cat("Difficulty table (per-bundle):", nrow(ds_diff_bundle), "rows\n")
  }
} else {
  ds_diff_bundle <- NULL
  message("No dataset_metrics — skipping difficulty scatter.")
}
```

## 5.1 Trio R2 Screen

```{r q3-r2-screen}
if (!is.null(ds_diff_bundle) && nrow(ds_diff_bundle) >= 10) {
  z_metrics <- c("z_topk_ratio", "z_max_abs", "z_count_abs_gt_3", "z_eff_signals")
  high_ld_metrics <- c("high_ld_count_095", "high_ld_count_095_per_snp", "high_ld_frac_095")
  z_metrics <- intersect(z_metrics, names(ds_diff_bundle))
  high_ld_metrics <- intersect(high_ld_metrics, names(ds_diff_bundle))

  r2_m1 <- tryCatch(
    summary(stats::lm(overall_AUPRC ~ M1, data = ds_diff_bundle))$r.squared,
    error = function(e) NA_real_
  )

  if (length(z_metrics) > 0 && length(high_ld_metrics) > 0) {
    trio_results <- tidyr::expand_grid(
      high_ld_metric = high_ld_metrics,
      z_metric = z_metrics
    ) %>%
      dplyr::rowwise() %>%
      dplyr::mutate(
        R2_M1_only = r2_m1,
        R2_trio = tryCatch(
          summary(stats::lm(
            stats::as.formula(
              paste("overall_AUPRC ~ M1 +", high_ld_metric, "+", z_metric)
            ),
            data = ds_diff_bundle
          ))$r.squared,
          error = function(e) NA_real_
        ),
        R2_gain = R2_trio - R2_M1_only
      ) %>%
      dplyr::ungroup() %>%
      dplyr::arrange(dplyr::desc(R2_trio))

    cat("=== Trio screen: M1 + high-LD + z-metric -> overall AUPRC ===\n")
    print(trio_results)

    trio_table_path <- file.path(paths$aggregated_dir, "q3_trio_r2_table.csv")
    readr::write_csv(trio_results, trio_table_path)
    cat("Saved trio R2 table:", trio_table_path, "\n")

    best_row <- trio_results %>% dplyr::slice_max(R2_trio, n = 1, with_ties = FALSE)
    best_high_ld <- best_row$high_ld_metric[1]
    best_z <- best_row$z_metric[1]
    best_r2 <- best_row$R2_trio[1]
    cat("\nBest high-LD metric:", best_high_ld, "\n")
    cat("Best z-metric:", best_z, "\n")
    cat("Best trio R2:", sprintf("%.4f", best_r2), "\n")
  } else {
    best_high_ld <- "high_ld_count_095_per_snp"
    best_z <- "z_max_abs"
    best_r2 <- NA_real_
    message("Missing high-LD or z metrics in dataset_metrics; defaulting metric choices.")
  }
} else {
  best_high_ld <- "high_ld_count_095_per_snp"
  best_z <- "z_max_abs"
  best_r2 <- NA_real_
  message("Insufficient data for R2 screen; defaulting to high_ld_count_095_per_snp + z_max_abs.")
}
```

## 5.2 Marginal Predictor Effects: M1, High-LD, and Best z-Metric

```{r q3-scatter, fig.width=15, fig.height=5}
if (!is.null(ds_diff_bundle) && nrow(ds_diff_bundle) >= 3) {
  make_partial_df <- function(df, y_var, x_var, controls = character()) {
    keep_cols <- unique(c(y_var, x_var, controls))
    dfx <- df %>%
      dplyr::select(dplyr::any_of(keep_cols)) %>%
      tidyr::drop_na()
    if (!nrow(dfx)) {
      return(NULL)
    }
    y_resid <- if (length(controls)) {
      stats::residuals(stats::lm(stats::as.formula(
        paste(y_var, "~", paste(controls, collapse = " + "))
      ), data = dfx))
    } else {
      dfx[[y_var]]
    }
    x_resid <- if (length(controls)) {
      stats::residuals(stats::lm(stats::as.formula(
        paste(x_var, "~", paste(controls, collapse = " + "))
      ), data = dfx))
    } else {
      dfx[[x_var]]
    }
    tibble::tibble(x = x_resid, y = y_resid)
  }

  partial_m1 <- make_partial_df(
    ds_diff_bundle, "overall_AUPRC", "M1", controls = c(best_high_ld, best_z)
  )
  partial_high_ld <- make_partial_df(
    ds_diff_bundle, "overall_AUPRC", best_high_ld, controls = c("M1", best_z)
  )
  partial_z <- make_partial_df(
    ds_diff_bundle, "overall_AUPRC", best_z, controls = c("M1", best_high_ld)
  )

  plot_partial <- function(df, title, x_label, color) {
    ggplot2::ggplot(df, ggplot2::aes(x = x, y = y)) +
      ggplot2::geom_point(size = 1.8, alpha = 0.55, color = color) +
      ggplot2::geom_smooth(method = "lm", se = TRUE, color = "black", linewidth = 0.7) +
      ggplot2::labs(title = title, x = x_label, y = "Overall AUPRC (residualized)") +
      theme_phase_a()
  }

  p_m1 <- plot_partial(partial_m1, "Marginal Effect: M1", "M1 residual", "#1f78b4")
  p_high_ld <- plot_partial(
    partial_high_ld,
    paste("Marginal Effect:", best_high_ld),
    paste(best_high_ld, "residual"),
    "#33a02c"
  )
  p_z <- plot_partial(
    partial_z,
    paste("Marginal Effect:", best_z),
    paste(best_z, "residual"),
    "#e31a1c"
  )

  p_q3 <- cowplot::plot_grid(p_m1, p_high_ld, p_z, ncol = 3)
  save_or_show(p_q3, "q3_difficulty_partial_effects.png",
               width = 15, height = 5, subdir = "difficulty")

  if (!is.na(best_r2)) {
    cat("Best trio model R2:", sprintf("%.4f", best_r2), "\n")
  }
}
```

---

# Section 6: Q4 — Use Case Performance vs M1 (Line Graph)

Colored line graph: x = M1, y = metric, color = use case.
Per-matrix AUPRC and Power @ FDR for each non-functional use case.

```{r q4-line-setup}
# Per-matrix AUPRC by use case
auprc_per_matrix_uc <- test_susine::aggregate_auprc_by_dataset_usecase(
  confusion_nonfunc
) %>%
  dplyr::filter(use_case_id %in% nonfunctional_ids)

# Add M1 to each matrix
if (!is.null(dataset_metrics)) {
  m1_by_matrix <- dataset_metrics %>%
    dplyr::select(dplyr::any_of(c("matrix_id", "M1"))) %>%
    dplyr::distinct() %>%
    dplyr::group_by(matrix_id) %>%
    dplyr::summarise(M1 = mean(M1, na.rm = TRUE), .groups = "drop")

  if (!"matrix_id" %in% names(m1_by_matrix) && !is.null(run_table)) {
    m1_by_matrix <- dataset_metrics %>%
      dplyr::left_join(
        run_table %>% dplyr::distinct(dataset_bundle_id, matrix_id),
        by = "dataset_bundle_id"
      ) %>%
      dplyr::group_by(matrix_id) %>%
      dplyr::summarise(M1 = mean(M1, na.rm = TRUE), .groups = "drop")
  }

  auprc_m1 <- auprc_per_matrix_uc %>%
    dplyr::inner_join(m1_by_matrix, by = "matrix_id")
} else {
  auprc_m1 <- NULL
}
```

## 6.1 AUPRC vs M1 by Use Case

```{r q4-auprc-m1, fig.width=12, fig.height=7}
if (!is.null(auprc_m1) && nrow(auprc_m1) > 0) {
  p_q4a <- ggplot2::ggplot(
    auprc_m1,
    ggplot2::aes(x = M1, y = AUPRC, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 2.5) +
    ggplot2::scale_color_manual(
      values = use_case_palette[nonfunctional_ids],
      labels = short_labels[nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "AUPRC vs LD Complexity (M1) — Non-Functional",
      x = "M1 (LD complexity)", y = "AUPRC"
    ) +
    theme_phase_a() +
    ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

  save_or_show(p_q4a, "q4_auprc_vs_m1_lines.png", width = 12, height = 7,
               subdir = "difficulty")
}
```

## 6.2 Power @ FDR vs M1 by Use Case

```{r q4-power-m1, fig.width=12, fig.height=7}
if (!is.null(auprc_m1) && nrow(auprc_m1) > 0) {
  # Compute power@FDR per matrix × use_case
  curves_per_matrix <- test_susine::compute_power_fdr_curves(
    confusion_nonfunc,
    group_vars = c("use_case_id", "matrix_id")
  )
  power_per_matrix <- power_at_fdr(curves_per_matrix, fdr_thresh = fdr_threshold,
                                    group_cols = c("use_case_id", "matrix_id"))

  power_m1 <- power_per_matrix %>%
    dplyr::inner_join(m1_by_matrix, by = "matrix_id") %>%
    dplyr::filter(use_case_id %in% nonfunctional_ids)

  p_q4b <- ggplot2::ggplot(
    power_m1,
    ggplot2::aes(x = M1, y = power_at_fdr, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 2.5) +
    ggplot2::scale_color_manual(
      values = use_case_palette[nonfunctional_ids],
      labels = short_labels[nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = paste0("Power @ FDR \u2264 ", fdr_threshold,
                     " vs M1 — Non-Functional"),
      x = "M1 (LD complexity)",
      y = paste0("Power (FDR \u2264 ", fdr_threshold, ")")
    ) +
    theme_phase_a() +
    ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

  save_or_show(p_q4b, "q4_power_at_fdr_vs_m1_lines.png", width = 12, height = 7,
               subdir = "difficulty")
}
```

---

# Section 7: Q5 — Functional Prior Comparison

Compare functional priors (mu_0 = c\*a, pi = softmax(|a|/tau)) to the
vanilla (non-functional) baseline across c, tau, annotation_r2,
inflate_match.

## 7.1 AUPRC by c_value (Functional mu)

```{r q6-c-value, fig.width=10, fig.height=6}
confusion_mu <- confusion_base %>%
  dplyr::filter(use_case_id == "susine_functional_mu")

format_param_val <- function(x) {
  trimws(format(x, digits = 3, scientific = FALSE))
}

# Diagnostics: c-value counts before filtering
cat("=== functional_mu row counts by c_value (raw) ===\n")
print(confusion_mu %>% dplyr::count(c_value, sort = TRUE))

if (any(is.na(confusion_mu$c_value))) {
  msg <- "functional_mu has rows with c_value = NA. Excluding these rows from Q6 plot."
  if (isTRUE(strict_fmu_c_na)) {
    stop(msg)
  } else {
    warning(msg)
  }
}

confusion_mu_plot <- confusion_mu %>%
  dplyr::filter(!is.na(c_value), !is.na(annotation_r2), !is.na(inflate_match))

auprc_baseline <- test_susine::compute_auprc_from_confusion(
  confusion_base %>% dplyr::filter(use_case_id == "susine_vanilla"),
  group_vars = "use_case_id"
)
baseline_val <- auprc_baseline$AUPRC[1]

auprc_by_c <- test_susine::compute_auprc_from_confusion(
  confusion_mu_plot,
  group_vars = c("c_value")
)

auprc_by_c_detail <- test_susine::compute_auprc_from_confusion(
  confusion_mu_plot,
  group_vars = c("c_value", "annotation_r2", "inflate_match")
)

q6_numeric <- auprc_by_c %>%
  dplyr::mutate(
    AUPRC_susine_baseline = baseline_val,
    diff_vs_susine_baseline = AUPRC - AUPRC_susine_baseline
  ) %>%
  dplyr::arrange(c_value)

cat("=== AUPRC by c_value (functional mu) ===\n")
print(q6_numeric)

q6_table_path <- file.path(paths$aggregated_dir, "q6_auprc_by_c_table.csv")
readr::write_csv(q6_numeric, q6_table_path)
cat("Saved Q6 numeric table:", q6_table_path, "\n")

q6_detail_path <- file.path(paths$aggregated_dir, "q6_auprc_by_c_table_stratified.csv")
readr::write_csv(auprc_by_c_detail, q6_detail_path)
cat("Saved Q6 stratified c table:", q6_detail_path, "\n")

if (nrow(auprc_by_c_detail) > 0) {
  p_q6c <- ggplot2::ggplot(
    auprc_by_c_detail,
    ggplot2::aes(
      x = c_value,
      y = AUPRC,
      color = factor(annotation_r2),
      group = factor(annotation_r2)
    )
  ) +
    ggplot2::geom_line(linewidth = 1.0) +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_hline(
      yintercept = baseline_val,
      linetype = "dashed",
      color = use_case_palette["susine_vanilla"],
      linewidth = 0.8
    ) +
    ggplot2::facet_wrap(~paste0("inflate_match = ", inflate_match), ncol = 3) +
    ggplot2::scale_color_brewer(palette = "Dark2", name = "annotation_r2") +
    ggplot2::labs(
      title = "AUPRC by c (Functional mu): color = annotation_r2, facet = inflate_match",
      subtitle = "mu_0 = c * a; dashed = vanilla SuSiNE baseline",
      x = "c (annotation scale)", y = "AUPRC"
    ) + theme_phase_a()

  save_or_show(p_q6c, "q6_auprc_by_c_value.png", width = 10, height = 6,
               subdir = "functional")

  p_q6c_alt <- ggplot2::ggplot(
    auprc_by_c_detail,
    ggplot2::aes(
      x = c_value,
      y = AUPRC,
      color = factor(inflate_match),
      group = factor(inflate_match)
    )
  ) +
    ggplot2::geom_line(linewidth = 1.0) +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_hline(
      yintercept = baseline_val,
      linetype = "dashed",
      color = use_case_palette["susine_vanilla"],
      linewidth = 0.8
    ) +
    ggplot2::facet_wrap(~paste0("annotation_r2 = ", annotation_r2), ncol = 3) +
    ggplot2::scale_color_brewer(palette = "Set1", name = "inflate_match") +
    ggplot2::labs(
      title = "AUPRC by c (Functional mu): color = inflate_match, facet = annotation_r2",
      subtitle = "mu_0 = c * a; dashed = vanilla SuSiNE baseline",
      x = "c (annotation scale)", y = "AUPRC"
    ) + theme_phase_a()

  save_or_show(p_q6c_alt, "q6_auprc_by_c_value_by_inflate_color.png",
               width = 10, height = 6, subdir = "functional")
}
```

## 7.2 AUPRC by tau_value (Functional pi)

```{r q6-tau-value, fig.width=10, fig.height=6}
confusion_pi <- confusion_base %>%
  dplyr::filter(use_case_id == "susine_functional_pi") %>%
  dplyr::filter(!is.na(tau_value), !is.na(annotation_r2), !is.na(inflate_match))

auprc_by_tau <- test_susine::compute_auprc_from_confusion(
  confusion_pi,
  group_vars = c("tau_value")
)

auprc_by_tau_detail <- test_susine::compute_auprc_from_confusion(
  confusion_pi,
  group_vars = c("tau_value", "annotation_r2", "inflate_match")
)

cat("=== AUPRC by tau_value (functional pi) ===\n")
print(auprc_by_tau)

q6_tau_table_path <- file.path(paths$aggregated_dir, "q6_auprc_by_tau_table.csv")
readr::write_csv(auprc_by_tau, q6_tau_table_path)
q6_tau_detail_path <- file.path(paths$aggregated_dir, "q6_auprc_by_tau_table_stratified.csv")
readr::write_csv(auprc_by_tau_detail, q6_tau_detail_path)

if (nrow(auprc_by_tau_detail) > 0) {
  p_q6t <- ggplot2::ggplot(
    auprc_by_tau_detail,
    ggplot2::aes(
      x = tau_value,
      y = AUPRC,
      color = factor(annotation_r2),
      group = factor(annotation_r2)
    )
  ) +
    ggplot2::geom_line(linewidth = 1.0) +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_hline(
      yintercept = baseline_val,
      linetype = "dashed",
      color = use_case_palette["susine_vanilla"],
      linewidth = 0.8
    ) +
    ggplot2::facet_wrap(~paste0("inflate_match = ", inflate_match), ncol = 3) +
    ggplot2::scale_color_brewer(palette = "Dark2", name = "annotation_r2") +
    ggplot2::labs(
      title = "AUPRC by tau (Functional pi): color = annotation_r2, facet = inflate_match",
      subtitle = "pi = softmax(|a|/tau); dashed = vanilla SuSiNE baseline",
      x = "tau (softmax temperature)", y = "AUPRC"
    ) + theme_phase_a()

  save_or_show(p_q6t, "q6_auprc_by_tau_value.png", width = 10, height = 6,
               subdir = "functional")

  p_q6t_alt <- ggplot2::ggplot(
    auprc_by_tau_detail,
    ggplot2::aes(
      x = tau_value,
      y = AUPRC,
      color = factor(inflate_match),
      group = factor(inflate_match)
    )
  ) +
    ggplot2::geom_line(linewidth = 1.0) +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_hline(
      yintercept = baseline_val,
      linetype = "dashed",
      color = use_case_palette["susine_vanilla"],
      linewidth = 0.8
    ) +
    ggplot2::facet_wrap(~paste0("annotation_r2 = ", annotation_r2), ncol = 3) +
    ggplot2::scale_color_brewer(palette = "Set1", name = "inflate_match") +
    ggplot2::labs(
      title = "AUPRC by tau (Functional pi): color = inflate_match, facet = annotation_r2",
      subtitle = "pi = softmax(|a|/tau); dashed = vanilla SuSiNE baseline",
      x = "tau (softmax temperature)", y = "AUPRC"
    ) + theme_phase_a()

  save_or_show(p_q6t_alt, "q6_auprc_by_tau_value_by_inflate_color.png",
               width = 10, height = 6, subdir = "functional")
}
```

## 7.3 Functional AUPRC by annotation_r2, Faceted by inflate_match

Baseline (susine_vanilla) has inflate_match = NA so it appears as a
horizontal reference line within each facet, not a separate panel.

```{r q6-annotation-quality, fig.width=14, fig.height=6}
func_conf <- confusion_base %>%
  dplyr::filter(use_case_id %in% functional_ids) %>%
  dplyr::filter(!is.na(annotation_r2), !is.na(inflate_match)) %>%
  dplyr::mutate(
    param_group = dplyr::case_when(
      use_case_id == "susine_functional_mu" & !is.na(c_value) ~
        paste0("func-mu c=", format_param_val(c_value)),
      use_case_id == "susine_functional_pi" & !is.na(tau_value) ~
        paste0("func-pi tau=", format_param_val(tau_value)),
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::filter(!is.na(param_group))

auprc_func_annot <- test_susine::compute_auprc_from_confusion(
  func_conf,
  group_vars = c("param_group", "annotation_r2", "inflate_match")
)

if (nrow(auprc_func_annot) > 0) {
  mu_vals <- sort(unique(confusion_mu_plot$c_value))
  pi_vals <- sort(unique(confusion_pi$tau_value))
  mu_groups <- paste0("func-mu c=", format_param_val(mu_vals))
  pi_groups <- paste0("func-pi tau=", format_param_val(pi_vals))
  param_levels <- c(mu_groups, pi_groups)
  param_levels <- param_levels[param_levels %in% unique(auprc_func_annot$param_group)]

  base_cols <- c("#1b9e77", "#1f78b4", "#6a3d9a", "#e31a1c", "#ff7f00", "#b15928")
  if (length(param_levels) > length(base_cols)) {
    param_cols <- scales::hue_pal()(length(param_levels))
  } else {
    param_cols <- base_cols[seq_along(param_levels)]
  }
  param_palette <- stats::setNames(param_cols, param_levels)

  p_q6ann <- ggplot2::ggplot(
    dplyr::mutate(auprc_func_annot, param_group = factor(param_group, levels = param_levels)),
    ggplot2::aes(x = annotation_r2, y = AUPRC, color = param_group, group = param_group)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_hline(
      yintercept = baseline_val,
      linetype = "dashed",
      color = use_case_palette["susine_vanilla"],
      linewidth = 0.7
    ) +
    ggplot2::facet_wrap(~paste("inflate_match =", inflate_match), ncol = 3) +
    ggplot2::scale_color_manual(
      values = param_palette,
      name = "Functional setting"
    ) +
    ggplot2::labs(
      title = "Functional AUPRC by Annotation Quality",
      subtitle = "Colors split by func-mu c and func-pi tau; dashed line = vanilla baseline",
      x = "annotation_r2", y = "AUPRC"
    ) +
    theme_phase_a()

  save_or_show(p_q6ann, "q6_functional_by_annotation_quality.png",
               width = 14, height = 6, subdir = "functional")
}
```

## 7.4 Functional Gain Breakdown by annotation_r2 and c/tau

```{r q6-breakdown-gain, fig.width=14, fig.height=8}
auprc_func_param <- test_susine::compute_auprc_from_confusion(
  func_conf %>%
    dplyr::mutate(
      param_axis = dplyr::if_else(
        use_case_id == "susine_functional_mu",
        "c",
        "tau"
      ),
      param_value = dplyr::if_else(
        use_case_id == "susine_functional_mu",
        c_value,
        tau_value
      )
    ),
  group_vars = c("use_case_id", "param_axis", "param_value", "annotation_r2", "inflate_match")
) %>%
  dplyr::mutate(
    AUPRC_gain_vs_vanilla = AUPRC - baseline_val,
    use_case_label = dplyr::coalesce(short_labels[use_case_id], use_case_id),
    param_value_label = format_param_val(param_value)
  )

if (nrow(auprc_func_param) > 0) {
  gain_table_path <- file.path(paths$aggregated_dir, "q6_functional_gain_breakdown_table.csv")
  readr::write_csv(auprc_func_param, gain_table_path)
  cat("Saved functional gain breakdown table:", gain_table_path, "\n")

  param_levels <- auprc_func_param %>%
    dplyr::arrange(param_value) %>%
    dplyr::pull(param_value_label) %>%
    unique()

  p_q6gain <- ggplot2::ggplot(
    auprc_func_param %>%
      dplyr::mutate(param_value_label = factor(param_value_label, levels = param_levels)),
    ggplot2::aes(
      x = factor(annotation_r2),
      y = param_value_label,
      fill = AUPRC_gain_vs_vanilla
    )
  ) +
    ggplot2::geom_tile(color = "white") +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.3f", AUPRC_gain_vs_vanilla)), size = 3) +
    ggplot2::facet_grid(use_case_label ~ paste0("inflate_match = ", inflate_match)) +
    ggplot2::scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac", midpoint = 0) +
    ggplot2::labs(
      title = "Functional Gain over Vanilla by Annotation Quality and c/tau",
      subtitle = "Negative values identify underperformance regimes",
      x = "annotation_r2",
      y = "c or tau",
      fill = "AUPRC gain"
    ) +
    theme_phase_a()

  save_or_show(p_q6gain, "q6_functional_gain_by_annotation_param.png",
               width = 14, height = 8, subdir = "functional")
}
```

## 7.5 Functional-mu p*=1 Diagnostic Slice

```{r q6-functional-mu-pstar1, fig.width=14, fig.height=6}
mu_by_setting <- test_susine::compute_auprc_from_confusion(
  confusion_mu_plot,
  group_vars = c("p_star", "y_noise", "c_value", "annotation_r2", "inflate_match")
) %>%
  dplyr::filter(p_star == 1)

baseline_by_setting <- test_susine::compute_auprc_from_confusion(
  confusion_base %>% dplyr::filter(use_case_id == "susine_vanilla"),
  group_vars = c("p_star", "y_noise")
) %>%
  dplyr::rename(AUPRC_baseline = AUPRC) %>%
  dplyr::filter(p_star == 1) %>%
  dplyr::select(-p_star)

mu_pstar1_diag <- mu_by_setting %>%
  dplyr::left_join(baseline_by_setting, by = "y_noise") %>%
  dplyr::mutate(
    AUPRC_gain_vs_baseline = AUPRC - AUPRC_baseline,
    c_label = format_param_val(c_value)
  )

if (nrow(mu_pstar1_diag) > 0) {
  pstar1_table_path <- file.path(paths$aggregated_dir, "q6_functional_mu_pstar1_diagnostic.csv")
  readr::write_csv(mu_pstar1_diag, pstar1_table_path)
  cat("Saved p*=1 diagnostic table:", pstar1_table_path, "\n")

  c_levels <- mu_pstar1_diag %>%
    dplyr::arrange(c_value) %>%
    dplyr::pull(c_label) %>%
    unique()

  p_mu_pstar1 <- ggplot2::ggplot(
    mu_pstar1_diag %>%
      dplyr::mutate(c_label = factor(c_label, levels = c_levels)),
    ggplot2::aes(
      x = factor(annotation_r2),
      y = c_label,
      fill = AUPRC_gain_vs_baseline
    )
  ) +
    ggplot2::geom_tile(color = "white") +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.3f", AUPRC_gain_vs_baseline)), size = 3) +
    ggplot2::facet_grid(y_noise ~ paste0("inflate_match = ", inflate_match)) +
    ggplot2::scale_fill_gradient2(low = "#b2182b", mid = "white", high = "#2166ac", midpoint = 0) +
    ggplot2::labs(
      title = "Functional-mu Diagnostic at p* = 1",
      subtitle = "Cell value = AUPRC(functional-mu) - AUPRC(susine_vanilla)",
      x = "annotation_r2",
      y = "c value",
      fill = "AUPRC gain"
    ) +
    theme_phase_a()

  save_or_show(p_mu_pstar1, "q6_functional_mu_pstar1_diagnostic.png",
               width = 14, height = 6, subdir = "functional")
}
```

---

# Section 8: Q7 — Max-ELBO Functional vs Non-Functional Baseline

The question: compared to a baseline model with NO functional
information (susie_vanilla), how does max-ELBO aggregated performance
of func-mu and func-pi scale with annotation_r2?

## 8.1 Max-ELBO vs Baseline: Overall

```{r q7-setup}
# Max-ELBO aggregated rows for functional use cases
confusion_agg_func <- confusion_with_agg %>%
  dplyr::filter(use_case_id %in% functional_ids, agg_method == "max_elbo")

# Non-functional baseline: susie_vanilla base fits
confusion_baseline <- confusion_base %>%
  dplyr::filter(use_case_id == "susie_vanilla")

cat("Max-ELBO functional rows:", nrow(confusion_agg_func), "\n")
cat("Baseline rows:", nrow(confusion_baseline), "\n")
```

```{r q7-overall-gain}
auprc_maxelbo <- test_susine::compute_auprc_from_confusion(
  confusion_agg_func, group_vars = "use_case_id"
) %>%
  dplyr::rename(AUPRC_max_elbo = AUPRC)

auprc_baseline_val <- test_susine::compute_auprc_from_confusion(
  confusion_baseline, group_vars = "use_case_id"
)$AUPRC[1]

gain_overall <- auprc_maxelbo %>%
  dplyr::mutate(
    AUPRC_baseline = auprc_baseline_val,
    AUPRC_gain = AUPRC_max_elbo - AUPRC_baseline,
    pct_gain = 100 * AUPRC_gain / AUPRC_baseline,
    label = short_labels[use_case_id]
  )

cat("=== Max-ELBO Functional vs Vanilla Baseline ===\n")
gain_overall %>%
  dplyr::select(label, AUPRC_baseline, AUPRC_max_elbo, AUPRC_gain, pct_gain) %>%
  print()
```

## 8.2 Gain vs annotation_r2

```{r q7-gain-by-annot, fig.width=12, fig.height=6}
auprc_maxelbo_annot <- test_susine::compute_auprc_from_confusion(
  confusion_agg_func,
  group_vars = c("use_case_id", "annotation_r2")
) %>%
  dplyr::rename(AUPRC_max_elbo = AUPRC) %>%
  dplyr::filter(!is.na(annotation_r2))

# Baseline is constant across annotation_r2
gain_annot <- auprc_maxelbo_annot %>%
  dplyr::mutate(
    AUPRC_baseline = auprc_baseline_val,
    AUPRC_gain = AUPRC_max_elbo - AUPRC_baseline,
    label = short_labels[use_case_id]
  )

if (nrow(gain_annot) > 0) {
  p_q7a <- ggplot2::ggplot(
    gain_annot,
    ggplot2::aes(x = annotation_r2, y = AUPRC_gain, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 1.0) +
    ggplot2::geom_point(size = 3.5) +
    ggplot2::geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
    ggplot2::scale_color_manual(
      values = use_case_palette[functional_ids],
      labels = short_labels[functional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "Max-ELBO AUPRC Gain over Vanilla Baseline",
      subtitle = "Gain = AUPRC(max-ELBO functional) - AUPRC(vanilla SuSiE)",
      x = "annotation_r2", y = "AUPRC Gain"
    ) +
    theme_phase_a()

  save_or_show(p_q7a, "q7_maxelbo_vs_baseline_by_annotation.png",
               width = 12, height = 6, subdir = "functional")
}
```

## 8.3 Gain by p_star and y_noise

```{r q7-gain-by-settings, fig.width=14, fig.height=8}
auprc_maxelbo_settings <- test_susine::compute_auprc_from_confusion(
  confusion_agg_func,
  group_vars = c("use_case_id", "p_star", "y_noise")
) %>%
  dplyr::rename(AUPRC_max_elbo = AUPRC)

# Baseline by the same settings
auprc_baseline_settings <- test_susine::compute_auprc_from_confusion(
  confusion_baseline,
  group_vars = c("use_case_id", "p_star", "y_noise")
) %>%
  dplyr::rename(AUPRC_baseline = AUPRC) %>%
  dplyr::select(-use_case_id)

gain_settings <- auprc_maxelbo_settings %>%
  dplyr::left_join(auprc_baseline_settings, by = c("p_star", "y_noise")) %>%
  dplyr::mutate(
    AUPRC_gain = AUPRC_max_elbo - AUPRC_baseline,
    label = short_labels[use_case_id]
  )

if (nrow(gain_settings) > 0) {
  p_q7b <- ggplot2::ggplot(
    gain_settings,
    ggplot2::aes(x = factor(p_star), y = AUPRC_gain, fill = use_case_id)
  ) +
    ggplot2::geom_col(position = ggplot2::position_dodge(width = 0.8),
                      width = 0.7, alpha = 0.85) +
    ggplot2::facet_wrap(~paste("y_noise =", y_noise), ncol = 3) +
    ggplot2::geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
    ggplot2::scale_fill_manual(
      values = use_case_palette[functional_ids],
      labels = short_labels[functional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "Max-ELBO Functional Gain over Vanilla by p* and y_noise",
      subtitle = "Positive = functional max-ELBO beats vanilla baseline",
      x = "p*", y = "AUPRC Gain"
    ) +
    theme_phase_a()

  save_or_show(p_q7b, "q7_maxelbo_vs_baseline_by_pstar_ynoise.png",
               width = 14, height = 8, subdir = "functional")
}
```

---

# Section 9: Additional Analyses

## 9.1 Verification: susine vs susieR Baselines

```{r q_extra_verification}
verification_pairs <- tibble::tribble(
  ~susie_id,        ~susine_id,       ~pair_label,
  "susie_vanilla",  "susine_vanilla", "Vanilla",
  "susie_eb",       "susine_eb",      "EB"
)

for (i in seq_len(nrow(verification_pairs))) {
  pair <- verification_pairs[i, ]
  auprc_pair <- test_susine::aggregate_auprc_by_dataset_usecase(
    confusion_base %>%
      dplyr::filter(use_case_id %in% c(pair$susie_id, pair$susine_id))
  )

  if (nrow(auprc_pair) > 0) {
    wide <- auprc_pair %>%
      tidyr::pivot_wider(id_cols = matrix_id, names_from = use_case_id,
                         values_from = AUPRC)
    if (ncol(wide) >= 3) {
      vals <- wide[[pair$susie_id]]
      ref <- wide[[pair$susine_id]]
      corr <- cor(vals, ref, use = "complete.obs")
      max_diff <- max(abs(vals - ref), na.rm = TRUE)
      cat(sprintf("Verification (%s): corr=%.5f, max|diff|=%.6f\n",
                  pair$pair_label, corr, max_diff))
    }
  }
}
```

## 9.2 Sigma Sensitivity (Fixed-Variance Use Cases)

```{r q_extra_sigma, fig.width=12, fig.height=7}
fixed_nonfunctional_ids <- c("susie_vanilla", "susie_ash_fixed",
                              "susie_inf", "susine_vanilla")

auprc_by_sigma <- test_susine::compute_auprc_from_confusion(
  confusion_nonfunc %>% dplyr::filter(use_case_id %in% fixed_nonfunctional_ids),
  group_vars = c("use_case_id", "sigma_0_2_scalar")
) %>%
  dplyr::filter(!is.na(sigma_0_2_scalar))

if (nrow(auprc_by_sigma) > 0) {
  p_sigma <- ggplot2::ggplot(
    auprc_by_sigma,
    ggplot2::aes(x = sigma_0_2_scalar, y = AUPRC, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 3) +
    ggplot2::scale_color_manual(
      values = use_case_palette[fixed_nonfunctional_ids],
      labels = short_labels[fixed_nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "AUPRC Sensitivity to Prior Variance",
      x = expression(sigma[0]^2 ~ "(fraction of var(y))"), y = "AUPRC"
    ) +
    theme_phase_a()

  save_or_show(p_sigma, "extra_sigma_sensitivity.png", width = 12, height = 7,
               subdir = "phase_a")
}
```

---

# Section 10: Summary and Phase B Recommendations

```{r summary-recommendations}
cat("==============================================================\n")
cat("  Phase A Results Summary\n")
cat("==============================================================\n\n")

cat(sprintf("FDR threshold used for Power@FDR: %.2f\n\n", fdr_threshold))

# Heatmap summaries
cat("--- Per-setting heatmaps in plots_phase_a/main/ ---\n")
cat("  heatmap_auprc_pstar_ynoise.png\n")
cat("  heatmap_power_at_fdr_pstar_ynoise.png\n\n")

# EB comparison
cat("--- EB vs Fixed comparison ---\n")
if (exists("eb_comparison")) {
  eb_comparison %>%
    dplyr::select(pair_label, prior_var_type, AUPRC, power_at_fdr) %>%
    dplyr::arrange(pair_label, prior_var_type) %>%
    print(n = 30)
}

# Difficulty
if (exists("best_z") && exists("best_high_ld")) {
  cat(
    "\n--- Best trio for difficulty: M1 +",
    best_high_ld,
    "+",
    best_z,
    "---\n"
  )
  if (exists("trio_results")) print(trio_results)
}

# Max-ELBO functional gains
cat("\n--- Max-ELBO Functional vs Baseline ---\n")
if (exists("gain_overall")) {
  gain_overall %>%
    dplyr::select(label, AUPRC_baseline, AUPRC_max_elbo, AUPRC_gain, pct_gain) %>%
    print()
}

cat("\n==============================================================\n")
cat("  Phase B Design Guidance\n")
cat("==============================================================\n")
cat("
Based on Phase A screening:
1. Verify susine/susieR consistency after sigma_0_2 bug fix.
2. Confirm sigma_0^2 sensitivity (pick 1-2 values or keep 3).
3. Set c-grid and tau-grid ranges based on Q6 results.
4. Use best high-LD + z-metric pair for difficulty stratification.
5. Decide which use cases to keep/drop for exploration budget.
")
```
