---
title: "Ensemble Scaling Experiment — Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

All figures from `refs/ensemble_scaling_experiment_plan.md`.
Organized into 4 categories with subfolder output.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_ensemble_scaling.Rmd")
library(here)
library(devtools)
devtools::load_all(".")
susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  devtools::load_all(susine_path)
}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(knitr)

theme_set(theme_bw(base_size = 12))
```

# 0. Load Data

```{r load-data}
job_name <- "ensemble_scaling_v1"
job_id <- "XXXXXX"  # TODO: update after SLURM run

agg_dir <- here("output", "run_history", job_name, job_id, "aggregate")
fig_root <- file.path(agg_dir, "figures")

# Create figure subdirectories
fig_dirs <- c("susie2_plots", "ensemble_performance",
              "dataset_difficulty", "ensemble_scaling")
for (d in fig_dirs) {
  dir.create(file.path(fig_root, d), recursive = TRUE, showWarnings = FALSE)
}

# Load consolidated outputs
model_metrics <- readr::read_csv(
  file.path(agg_dir, "model_metrics_with_oracle.csv"),
  show_col_types = FALSE
)

# Baseline reference line: single-fit susine at sigma_0_2 = 0.2
baseline_auprc <- model_metrics %>%
  dplyr::filter(
    spec_name == "baseline-single",
    use_case_id == "susine_vanilla",
    !isTRUE(is_agg)
  ) %>%
  dplyr::summarise(median_auprc = median(auprc, na.rm = TRUE)) %>%
  dplyr::pull(median_auprc)

# Truth-warm baselines
truth_warm <- model_metrics %>%
  dplyr::filter(spec_name == "truth-warm")

# Scaling curves (from collect workbook)
scaling_path <- file.path(agg_dir, "scaling_curves.csv")
if (file.exists(scaling_path)) {
  scaling_curves <- readr::read_csv(scaling_path, show_col_types = FALSE)
} else {
  scaling_curves <- NULL
}

# Dataset metrics
dm_path <- file.path(agg_dir, "dataset_metrics.csv")
if (file.exists(dm_path)) {
  dataset_metrics <- readr::read_csv(dm_path, show_col_types = FALSE)
} else {
  dataset_metrics <- NULL
}
```

# 1. SuSiE2-Style Plots (Category 1)

Individual model metrics using max-ELBO model selection per ensemble.

```{r cat1-setup}
# For Category 1 plots: select single best model per ensemble by max ELBO
max_elbo_models <- model_metrics %>%
  dplyr::filter(agg_method == "max_elbo" | is.na(agg_method), !isTRUE(is_agg))
```

## 1a. CS Power vs top-n-as-causal

```{r fig-1a, fig.width=8, fig.height=5}
# TODO: implement sweep over top_n grading thresholds.
# For each ensemble, select max-ELBO model, compute CS power at each n.
# Color by ensemble method (spec_name), full ensembles only.

# Placeholder structure:
# ggplot(cs_power_data, aes(x = top_n, y = cs_power, color = spec_name)) +
#   geom_line() +
#   labs(x = "Top n variants treated as causal", y = "CS Power") +
#   theme(legend.position = "bottom")
# ggsave(file.path(fig_root, "susie2_plots", "cs_power_vs_topn.png"),
#        width = 8, height = 5)
```

## 1b. CS FDR vs top-n-as-causal

```{r fig-1b, fig.width=8, fig.height=5}
# Same structure as 1a but Y = CS FDR
# ggsave(file.path(fig_root, "susie2_plots", "cs_fdr_vs_topn.png"), ...)
```

## 1c. Pearson R-squared (effect size prediction)

```{r fig-1c, fig.width=8, fig.height=5}
# Y: R-squared between estimated and true effect sizes
# Color by ensemble method, full ensembles only
# ggsave(file.path(fig_root, "susie2_plots", "effect_r2.png"), ...)
```

## 1d. TPR vs FPR (FPR 0 to 0.1)

```{r fig-1d, fig.width=12, fig.height=8}
# Uses PIPs -> all aggregation methods (including oracle)
# Facet by aggregation method, color by ensemble method
# Grade using top 8 as causal
# ggsave(file.path(fig_root, "susie2_plots", "tpr_fpr_curve.png"), ...)
```

## 1e. PIP Calibration

```{r fig-1e, fig.width=12, fig.height=8}
# Bin PIPs in 0.10-width bins (10 bins)
# Plot observed fraction causal vs mean PIP per bin
# Facet by aggregation method, color by ensemble method
# ggsave(file.path(fig_root, "susie2_plots", "pip_calibration.png"), ...)
```

# 2. Ensemble Performance Plots (Category 2)

PIP-based metrics, all aggregation methods.

## 2a. Precision-Recall Curves

```{r fig-2a, fig.width=12, fig.height=8}
# Grade using top 8 as causal
# Facet by aggregation method (including oracle)
# Color by ensemble method (spec_name)
# ggsave(file.path(fig_root, "ensemble_performance", "pr_curves.png"), ...)
```

## 2b. Multimodality Metrics Boxplots

```{r fig-2b, fig.width=14, fig.height=10}
# Load multimodal metrics if available
mm_path <- file.path(agg_dir, "multimodal_metrics.csv")
if (file.exists(mm_path)) {
  mm <- readr::read_csv(mm_path, show_col_types = FALSE)

  metrics_to_plot <- c("jaccard_top10", "mean_jsd", "median_jsd",
                       "max_jsd", "mean_pip_var", "n_clusters")

  mm_long <- mm %>%
    dplyr::select(dplyr::any_of(c("spec_name", metrics_to_plot))) %>%
    tidyr::pivot_longer(cols = dplyr::any_of(metrics_to_plot),
                        names_to = "metric", values_to = "value")

  p_mm <- ggplot(mm_long, aes(x = spec_name, y = value, fill = spec_name)) +
    geom_boxplot() +
    facet_wrap(~metric, scales = "free_y") +
    labs(x = NULL, y = "Value") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none")
  print(p_mm)
  ggsave(file.path(fig_root, "ensemble_performance",
                   "multimodality_boxplots.png"),
         p_mm, width = 14, height = 10)
}
```

## 2c. AUPRC Bar Chart by Method (primary figure)

```{r fig-2c, fig.width=12, fig.height=6}
# X = ensemble method, color = aggregation method (including oracle)
# Red dashed line = baseline single-fit susine sigma_0_2 = 0.2
# Include truth-warm baselines

agg_summary <- model_metrics %>%
  dplyr::filter(isTRUE(is_agg) | agg_method == "oracle") %>%
  dplyr::group_by(spec_name, agg_method) %>%
  dplyr::summarise(
    mean_auprc = mean(auprc, na.rm = TRUE),
    se_auprc = sd(auprc, na.rm = TRUE) / sqrt(dplyr::n()),
    .groups = "drop"
  )

if (nrow(agg_summary) > 0) {
  p_bar <- ggplot(agg_summary, aes(x = spec_name, y = mean_auprc,
                                    fill = agg_method)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    geom_errorbar(aes(ymin = mean_auprc - se_auprc,
                       ymax = mean_auprc + se_auprc),
                  position = position_dodge(width = 0.8), width = 0.2) +
    geom_hline(yintercept = baseline_auprc, linetype = "dashed",
               color = "red", linewidth = 0.8) +
    labs(x = "Ensemble Method", y = "Mean AUPRC", fill = "Aggregation") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(p_bar)
  ggsave(file.path(fig_root, "ensemble_performance",
                   "auprc_barplot_by_method.png"),
         p_bar, width = 12, height = 6)
}
```

# 3. Dataset Difficulty Plots (Category 3)

## 3a. Dataset Metrics Histograms + Correlation

```{r fig-3a, fig.width=12, fig.height=10}
if (!is.null(dataset_metrics)) {
  metric_cols <- c("M1", "z_topk_ratio", "z_max_abs",
                   "z_count_abs_gt_3", "z_eff_signals", "high_ld_count_095")
  available <- intersect(metric_cols, names(dataset_metrics))

  if (length(available) > 0) {
    dm_long <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(available)) %>%
      tidyr::pivot_longer(everything(), names_to = "metric", values_to = "value")

    p_hist <- ggplot(dm_long, aes(x = value)) +
      geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
      facet_wrap(~metric, scales = "free") +
      labs(x = "Value", y = "Count")
    print(p_hist)
    ggsave(file.path(fig_root, "dataset_difficulty",
                     "metric_histograms.png"),
           p_hist, width = 12, height = 10)

    # Correlation matrix
    dm_mat <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(available)) %>%
      stats::cor(use = "pairwise.complete.obs")
    png(file.path(fig_root, "dataset_difficulty", "metric_correlations.png"),
        width = 800, height = 800)
    corrplot::corrplot(dm_mat, method = "color", type = "upper",
                       addCoef.col = "black", number.cex = 0.8,
                       tl.cex = 0.9)
    dev.off()
  }
}
```

## 3b. Dataset Difficulty Scatter (M1 as primary predictor)

```{r fig-3b, fig.width=14, fig.height=5}
if (!is.null(dataset_metrics)) {
  # Baseline performance: single-fit susine at sigma_0_2=0.2
  baseline_perf <- model_metrics %>%
    dplyr::filter(
      spec_name == "baseline-single",
      use_case_id == "susine_vanilla",
      !isTRUE(is_agg)
    ) %>%
    dplyr::select(dataset_bundle_id, auprc_baseline = auprc)

  dm_with_perf <- dataset_metrics %>%
    dplyr::inner_join(baseline_perf, by = "dataset_bundle_id")

  if ("M1" %in% names(dm_with_perf) && nrow(dm_with_perf) > 0) {
    # Find best supplementary metrics via multiple regression
    z_candidates <- intersect(
      c("z_topk_ratio", "z_max_abs", "z_count_abs_gt_3", "z_eff_signals"),
      names(dm_with_perf)
    )
    ld_candidates <- intersect(
      c("high_ld_count_095"),
      names(dm_with_perf)
    )

    best_r2 <- 0
    best_z <- z_candidates[1]
    best_ld <- ld_candidates[1]

    for (zc in z_candidates) {
      for (lc in ld_candidates) {
        fml <- as.formula(paste("auprc_baseline ~ M1 +", zc, "+", lc))
        fit <- lm(fml, data = dm_with_perf)
        r2 <- summary(fit)$r.squared
        if (r2 > best_r2) {
          best_r2 <- r2
          best_z <- zc
          best_ld <- lc
        }
      }
    }

    cat("Best 3-predictor model: M1 +", best_z, "+", best_ld,
        "-> R2 =", round(best_r2, 4), "\n")

    # Fit combined model for regression lines
    fml_best <- as.formula(
      paste("auprc_baseline ~ M1 +", best_z, "+", best_ld)
    )
    lm_best <- lm(fml_best, data = dm_with_perf)
    dm_with_perf$predicted <- predict(lm_best, dm_with_perf)

    # Find best ensemble methods (no-annotation and with-annotation)
    ens_perf <- model_metrics %>%
      dplyr::filter(isTRUE(is_agg)) %>%
      dplyr::group_by(spec_name, agg_method) %>%
      dplyr::summarise(mean_auprc = mean(auprc, na.rm = TRUE),
                       .groups = "drop")

    # Group A = no annotations, Group B/C = with annotations
    best_no_ann <- ens_perf %>%
      dplyr::filter(grepl("^A-", spec_name)) %>%
      dplyr::slice_max(mean_auprc, n = 1)
    best_ann <- ens_perf %>%
      dplyr::filter(grepl("^[BC]-", spec_name)) %>%
      dplyr::slice_max(mean_auprc, n = 1)

    # Get per-dataset AUPRC for best methods
    best_no_ann_perf <- model_metrics %>%
      dplyr::filter(spec_name == best_no_ann$spec_name[1],
                    agg_method == best_no_ann$agg_method[1]) %>%
      dplyr::select(dataset_bundle_id, auprc_no_ann = auprc)

    best_ann_perf <- model_metrics %>%
      dplyr::filter(spec_name == best_ann$spec_name[1],
                    agg_method == best_ann$agg_method[1]) %>%
      dplyr::select(dataset_bundle_id, auprc_ann = auprc)

    dm_scatter <- dm_with_perf %>%
      dplyr::left_join(best_no_ann_perf, by = "dataset_bundle_id") %>%
      dplyr::left_join(best_ann_perf, by = "dataset_bundle_id")

    # 3 scatter panels: M1, best_z, best_ld
    scatter_vars <- c("M1", best_z, best_ld)
    scatter_plots <- list()
    for (sv in scatter_vars) {
      p <- ggplot(dm_scatter, aes(x = .data[[sv]])) +
        geom_point(aes(y = auprc_baseline), color = "grey50", alpha = 0.4) +
        geom_point(aes(y = auprc_no_ann), color = "dodgerblue", alpha = 0.4) +
        geom_point(aes(y = auprc_ann), color = "darkgreen", alpha = 0.4) +
        geom_line(aes(y = predicted), color = "red", linewidth = 0.8) +
        labs(x = sv, y = "AUPRC") +
        ylim(0, 1)
      scatter_plots[[sv]] <- p
    }
    p_combined <- patchwork::wrap_plots(scatter_plots, nrow = 1)
    print(p_combined)
    ggsave(file.path(fig_root, "dataset_difficulty",
                     "difficulty_scatter.png"),
           p_combined, width = 14, height = 5)
  }
}
```

# 4. Ensemble Scaling Plots (Category 4)

## 4a. 64-Grid Scaling: AUPRC

```{r fig-4a, fig.width=14, fig.height=8}
if (!is.null(scaling_curves)) {
  pure_specs <- c("A-R", "A-F", "A-S", "B-R", "B-F", "C-C")
  sc_pure <- scaling_curves %>%
    dplyr::filter(spec_name %in% pure_specs)

  if (nrow(sc_pure) > 0) {
    sc_summary <- sc_pure %>%
      dplyr::group_by(spec_name, subsample_size, agg_method) %>%
      dplyr::summarise(
        mean_auprc = mean(auprc, na.rm = TRUE),
        se_auprc = sd(auprc, na.rm = TRUE) / sqrt(dplyr::n()),
        .groups = "drop"
      )

    p_scale <- ggplot(sc_summary,
                      aes(x = subsample_size, y = mean_auprc,
                          color = agg_method)) +
      geom_line() +
      geom_point() +
      geom_errorbar(aes(ymin = mean_auprc - se_auprc,
                         ymax = mean_auprc + se_auprc), width = 0.05) +
      facet_wrap(~spec_name, scales = "free_y") +
      scale_x_log10() +
      labs(x = "Ensemble Size", y = "AUPRC", color = "Aggregation") +
      theme(legend.position = "bottom")
    print(p_scale)
    ggsave(file.path(fig_root, "ensemble_scaling",
                     "scaling_64_auprc.png"),
           p_scale, width = 14, height = 8)
  }
}
```

## 4b. 64-Grid Scaling: TPR at FPR=0.05

```{r fig-4b, fig.width=14, fig.height=8}
# Same as 4a but Y = TPR at fixed FPR=0.05
# Requires per-subsample TPR computation from confusion bins
# ggsave(file.path(fig_root, "ensemble_scaling",
#                  "scaling_64_tpr_fpr05.png"), ...)
```

## 4c. 8x8 Interaction Scaling: AUPRC

```{r fig-4c, fig.width=12, fig.height=6}
# Y: AUPRC, X: grid resolution {4x4, 8x8}
# Color by aggregation method
# Facet by interaction spec (A-RF, A-RS, A-FS, C-CS)

interaction_path <- file.path(agg_dir, "interaction_scaling.csv")
if (file.exists(interaction_path)) {
  int_scaling <- readr::read_csv(interaction_path, show_col_types = FALSE)

  if (nrow(int_scaling) > 0) {
    p_int <- ggplot(int_scaling,
                    aes(x = factor(grid_resolution), y = mean_auprc,
                        fill = agg_method)) +
      geom_col(position = position_dodge(width = 0.8), width = 0.7) +
      facet_wrap(~spec_name) +
      labs(x = "Grid Resolution", y = "AUPRC", fill = "Aggregation")
    print(p_int)
    ggsave(file.path(fig_root, "ensemble_scaling",
                     "scaling_8x8_auprc.png"),
           p_int, width = 12, height = 6)
  }
}
```

## 4d. 8x8 Interaction Scaling: TPR at FPR=0.05

```{r fig-4d, fig.width=12, fig.height=6}
# Same as 4c but Y = TPR at fixed FPR=0.05
# ggsave(file.path(fig_root, "ensemble_scaling",
#                  "scaling_8x8_tpr_fpr05.png"), ...)
```

# Session Info

```{r session}
sessionInfo()
```
