---
title: "Architecture Sensitivity - Visualize Results"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Visualization workbook for the reviewer genetic-architecture sensitivity
analysis. It reads the combined tables written by
`collect_results_workbook_architecture_sensitivity.Rmd`, generates the sparse and
diffuse metric figures, and saves paper-facing figures and CSVs.

```{r setup, message=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_architecture_sensitivity.Rmd")
library(here)
library(devtools)

susine_path <- here("..", "susine")
if (file.exists(file.path(susine_path, "DESCRIPTION"))) {
  options(test_susine.local_susine_path = normalizePath(susine_path, winslash = "/", mustWork = TRUE))
  devtools::load_all(susine_path)
}
devtools::load_all(".")

library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(knitr)

readr::local_edition(2)

theme_set(
  theme_bw(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "grey92"),
      strip.background = element_rect(fill = "grey95", color = NA),
      strip.text = element_text(face = "bold"),
      legend.position = "bottom",
      plot.title = element_text(face = "bold")
    )
)
```

# Load Data

```{r load}
output_root <- here("output")
combined_dir <- file.path(output_root, "architecture_sensitivity", "combined")
fig_dir <- file.path(output_root, "architecture_sensitivity", "figures")
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

read_req <- function(filename) {
  path <- file.path(combined_dir, filename)
  if (!file.exists(path)) stop("Missing required file: ", path)
  readr::read_csv(path, show_col_types = FALSE)
}

metric_summary <- read_req("metric_summary.csv")
baseline_summary <- read_req("baseline_summary.csv")
prior_gate <- read_req("prior_gate.csv")
confusion_presence <- read_req("confusion_presence.csv")

plot_data <- dplyr::bind_rows(metric_summary, baseline_summary) %>%
  dplyr::mutate(
    series = factor(.data$series, levels = c("C-CS", "Annotation-free baseline")),
    p_star = as.integer(.data$p_star),
    h2_snp_per_causal = as.numeric(.data$h2_snp_per_causal),
    h2_total = as.numeric(.data$h2_total),
    diffuse_k = as.integer(.data$diffuse_k),
    total_h2 = as.numeric(.data$total_h2),
    pve_label = paste0("Per-causal PVE = ", formatC(.data$h2_snp_per_causal, format = "f", digits = 2))
  )

cat("Combined dir:", combined_dir, "\n")
cat("Figure dir:", fig_dir, "\n")
```

# Validation Summary

```{r validation-summary}
kable(prior_gate, digits = 4, caption = "Fixed-c prior gate")
kable(confusion_presence, caption = "Confusion-bin presence by architecture setting")
```

# Sparse Architecture Metrics

```{r sparse-metrics, fig.width=10, fig.height=6}
sparse_long <- plot_data %>%
  dplyr::filter(.data$analysis == "sparse") %>%
  dplyr::select(
    .data$series, .data$p_star, .data$h2_snp_per_causal, .data$pve_label,
    .data$total_h2, .data$AUPRC, .data$recall_at_precision
  ) %>%
  tidyr::pivot_longer(
    cols = c("AUPRC", "recall_at_precision"),
    names_to = "metric",
    values_to = "value"
  ) %>%
  dplyr::mutate(
    metric = dplyr::recode(
      .data$metric,
      AUPRC = "Pooled AUPRC",
      recall_at_precision = "Recall at precision >= 0.75"
    ),
    p_star_label = factor(.data$p_star, levels = sort(unique(.data$p_star)))
  )

if (!nrow(sparse_long)) {
  stop("No sparse architecture metrics found.")
}

p_sparse <- ggplot(
  sparse_long,
  aes(
    x = .data$p_star_label,
    y = .data$value,
    color = .data$series,
    linetype = .data$series,
    group = .data$series
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_grid(metric ~ pve_label, scales = "free_y") +
  labs(
    x = "Number of causal variants (k)",
    y = NULL,
    color = "Series",
    linetype = "Series",
    title = "Sparse architecture sensitivity",
    caption = "Positive set is all causal variants for k <= 5."
  )

ggsave(file.path(fig_dir, "sparse_architecture_metrics.png"), p_sparse, width = 10, height = 6, dpi = 240)
p_sparse
```

```{r sparse-total-h2, fig.width=8, fig.height=5}
p_sparse_h2 <- ggplot(
  sparse_long,
  aes(
    x = .data$total_h2,
    y = .data$value,
    color = .data$series,
    linetype = .data$series,
    group = interaction(.data$series, .data$h2_snp_per_causal)
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~metric, scales = "free_y") +
  labs(
    x = "Approximate total locus h2 (k x per-causal PVE)",
    y = NULL,
    color = "Series",
    linetype = "Series",
    title = "Sparse architecture sensitivity by total signal"
  )

ggsave(file.path(fig_dir, "sparse_architecture_by_total_h2.png"), p_sparse_h2, width = 8, height = 5, dpi = 240)
p_sparse_h2
```

# Diffuse Architecture Metrics

```{r diffuse-metrics, fig.width=8, fig.height=5}
diffuse_long <- plot_data %>%
  dplyr::filter(.data$analysis == "diffuse") %>%
  dplyr::select(
    .data$series, .data$h2_total, .data$diffuse_k,
    .data$AUPRC, .data$recall_at_precision
  ) %>%
  tidyr::pivot_longer(
    cols = c("AUPRC", "recall_at_precision"),
    names_to = "metric",
    values_to = "value"
  ) %>%
  dplyr::mutate(
    metric = dplyr::recode(
      .data$metric,
      AUPRC = "Pooled AUPRC",
      recall_at_precision = "Recall at precision >= 0.75"
    )
  )

if (!nrow(diffuse_long)) {
  stop("No diffuse architecture metrics found.")
}

p_diffuse <- ggplot(
  diffuse_long,
  aes(
    x = .data$h2_total,
    y = .data$value,
    color = .data$series,
    linetype = .data$series,
    group = .data$series
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_wrap(~metric, scales = "free_y") +
  labs(
    x = "Total locus h2",
    y = NULL,
    color = "Series",
    linetype = "Series",
    title = "Diffuse architecture sensitivity",
    caption = "Positive set is the top eight causal variants by absolute effect."
  )

ggsave(file.path(fig_dir, "diffuse_architecture_metrics.png"), p_diffuse, width = 8, height = 5, dpi = 240)
p_diffuse
```

# Export Plot Data

```{r export-plot-data}
readr::write_csv(plot_data, file.path(fig_dir, "architecture_metric_plot_data.csv"))
readr::write_csv(sparse_long, file.path(fig_dir, "sparse_architecture_plot_data.csv"))
readr::write_csv(diffuse_long, file.path(fig_dir, "diffuse_architecture_plot_data.csv"))

cat("Saved figures and plot data to:", fig_dir, "\n")
```