---
title: "SuSiNE Pilot Phase A — Results Visualization"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
editor_options:
  markdown:
    wrap: 72
---

Summary: Phase A screening results. Evaluates 9 use cases (single fits
+ small c/tau grids for functional arms) across 10 M1-stratified
matrices. Uses confusion bins and dataset metrics from the completed
SLURM run to answer key questions driving Phase B design.

Primary questions:

1. Which non-functional use cases perform best?
2. How does EB compare with fixed-variance counterparts?
3. Which z-score metric best complements M1 for predicting difficulty?
4. Do use case rankings change across dataset-difficulty strata?
5. Do use case rankings change across simulation settings?
6. How do functional priors (mu vs pi) compare to no-prior baselines?
7. Does max-ELBO aggregation help functional priors? Where are gains largest?

# Section 1: Setup

```{r setup, message=FALSE, warning=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_phase_a.Rmd")
library(here)
library(devtools)
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(purrr)
library(broom)
```

## 1.1 User Options

```{r user-options}
save_plots_to_file <- TRUE
show_in_rstudio_plots <- TRUE
fdr_threshold <- 0.5  # FDR level for Power@FDR summaries

if (isTRUE(getOption("knitr.in.progress"))) {
  knitr::opts_chunk$set(fig.ext = "png")
} else if (show_in_rstudio_plots && interactive()) {
  knitr::opts_chunk$set(fig.show = "hide", dev = "RStudioGD", fig.ext = "png")
} else {
  knitr::opts_chunk$set(fig.ext = "png")
}

maybe_show_plot <- function(p) {
  if (show_in_rstudio_plots && interactive()) {
    try(print(p), silent = TRUE)
  }
}

save_or_show <- function(p, filename, width = 10, height = 7, dpi = 300,
                         subdir = "phase_a") {
  maybe_show_plot(p)
  if (save_plots_to_file) {
    filepath <- file.path(plot_dirs[[subdir]], filename)
    tryCatch(
      {
        ggsave(filepath, p, width = width, height = height, dpi = dpi, bg = "white")
        cat("Saved:", filename, "\n")
      },
      error = function(e) cat("Error saving:", filename, "-", conditionMessage(e), "\n")
    )
  } else {
    tryCatch(print(p), error = function(e) {
      cat("Cannot display plot. Set save_plots_to_file = TRUE.\n")
    })
  }
}

#' Extract Power at a given FDR threshold from power-FDR curves.
#' For each group, finds the highest power where FDR <= fdr_threshold.
power_at_fdr <- function(curves_df, fdr_thresh = fdr_threshold,
                         group_cols = "use_case_id") {
  curves_df %>%
    dplyr::filter(fdr <= fdr_thresh) %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_cols))) %>%
    dplyr::summarise(
      power_at_fdr = max(power, na.rm = TRUE),
      pip_at_fdr = min(pip_threshold[power == max(power)], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(fdr_threshold = fdr_thresh)
}
```

## 1.2 Job Configuration

```{r job-config}
# UPDATE THESE for your completed Phase A run
job_name <- "pilot_phase_a"
parent_job_id <- "48988192"
output_root <- here("output")

paths <- test_susine::resolve_job_paths(
  job_name = job_name,
  parent_job_id = parent_job_id,
  output_root = output_root
)
```

## 1.3 Plot Directories

```{r create-plot-dirs}
plot_base_dir <- file.path(paths$aggregated_dir, "plots_phase_a")
plot_dirs <- list(
  phase_a = file.path(plot_base_dir, "main"),
  difficulty = file.path(plot_base_dir, "difficulty"),
  functional = file.path(plot_base_dir, "functional"),
  stratified = file.path(plot_base_dir, "stratified")
)

for (d in plot_dirs) {
  dir.create(d, showWarnings = FALSE, recursive = TRUE)
}
cat("Plot directories created under:", plot_base_dir, "\n")
```

## 1.4 Load Data

```{r load-data, message=FALSE}
# --- Confusion bins (primary data source) ---
confusion_bins_path <- file.path(paths$aggregated_dir, "confusion_bins.csv")
confusion_detailed_path <- file.path(paths$aggregated_dir, "confusion_matrix_detailed.csv")

# Prefer the detailed confusion matrix if already built; else use raw bins
if (file.exists(confusion_detailed_path)) {
  confusion <- readr::read_csv(confusion_detailed_path, show_col_types = FALSE)
  cat("Loaded pre-built confusion_matrix_detailed:", nrow(confusion), "rows\n")
} else if (file.exists(confusion_bins_path)) {
  confusion <- readr::read_csv(confusion_bins_path, show_col_types = FALSE)
  cat("Loaded raw confusion_bins:", nrow(confusion), "rows\n")
} else {
  stop("No confusion data found. Run collect_results_workbook.Rmd first.")
}

# --- Dataset metrics ---
dataset_metrics_path <- file.path(paths$aggregated_dir, "dataset_metrics.csv")
if (file.exists(dataset_metrics_path)) {
  dataset_metrics <- readr::read_csv(dataset_metrics_path, show_col_types = FALSE)
  cat("Loaded dataset_metrics:", nrow(dataset_metrics), "rows\n")
} else {
  warning("dataset_metrics.csv not found — difficulty analyses will be skipped.")
  dataset_metrics <- NULL
}

# --- Model metrics (for ELBO-based analysis) ---
model_metrics_path <- file.path(paths$aggregated_dir, "model_metrics.csv")
if (file.exists(model_metrics_path)) {
  model_metrics <- readr::read_csv(model_metrics_path, show_col_types = FALSE)
  cat("Loaded model_metrics:", nrow(model_metrics), "rows\n")
} else {
  warning("model_metrics.csv not found — ELBO-based analyses will use confusion bins only.")
  model_metrics <- NULL
}

# --- Run table (for joining extra columns when needed) ---
run_table_path <- file.path(paths$run_history_dir, "run_table.csv")
if (file.exists(run_table_path)) {
  run_table <- readr::read_csv(run_table_path, show_col_types = FALSE)
  cat("Loaded run_table:", nrow(run_table), "rows\n")
} else {
  run_table <- NULL
}
```

## 1.5 Define Use Case Groups and Palettes

```{r use-case-definitions}
# --- Use case groupings ---
# Non-functional (no annotation dependency)
nonfunctional_ids <- c(
  "susie_vanilla", "susie_ash_fixed", "susie_ash_eb",
  "susie_inf", "susine_vanilla", "susie_eb", "susine_eb"
)

# Fixed-variance non-functional (exclude EB use cases)
fixed_nonfunctional_ids <- c(
  "susie_vanilla", "susie_ash_fixed", "susie_inf", "susine_vanilla"
)

# EB use cases and their fixed counterparts
eb_pairs <- tibble::tribble(
  ~eb_id,        ~fixed_id,         ~pair_label,
  "susie_eb",    "susie_vanilla",   "SuSiE vanilla",
  "susine_eb",   "susine_vanilla",  "SuSiNE vanilla",
  "susie_ash_eb","susie_ash_fixed", "SuSiE-ash"
)

# Functional use cases
functional_ids <- c("susine_functional_pi", "susine_functional_mu")

# All use case IDs
all_ids <- c(nonfunctional_ids, functional_ids)

# --- Labels from catalog ---
all_labels <- test_susine::get_use_case_labels(all_ids)

# --- Color palette (9 use cases) ---
# Blues: vanilla SuSiE family
# Greens: ash family
# Purples: inf
# Oranges: functional
use_case_palette <- c(
  "susie_vanilla"       = "#2171b5",
  "susine_vanilla"      = "#6baed6",
  "susie_eb"            = "#08306b",
  "susine_eb"           = "#4292c6",
  "susie_ash_fixed"     = "#238b45",
  "susie_ash_eb"        = "#74c476",
  "susie_inf"           = "#6a3d9a",
  "susine_functional_pi"= "#e6550d",
  "susine_functional_mu"= "#d62728"
)

# Short labels for compact plots
short_labels <- c(
  "susie_vanilla"       = "Vanilla",
  "susine_vanilla"      = "Vanilla (susine)",
  "susie_eb"            = "EB",
  "susine_eb"           = "EB (susine)",
  "susie_ash_fixed"     = "Ash-fixed",
  "susie_ash_eb"        = "Ash-EB",
  "susie_inf"           = "Inf",
  "susine_functional_pi"= "Func-pi",
  "susine_functional_mu"= "Func-mu"
)

# --- Shared ggplot theme ---
theme_phase_a <- function(base_size = 13) {
  ggplot2::theme_minimal(base_size = base_size) +
    ggplot2::theme(
      legend.position = "bottom",
      legend.title = ggplot2::element_text(size = 12, face = "bold"),
      legend.text = ggplot2::element_text(size = 11),
      strip.text = ggplot2::element_text(size = 12, face = "bold"),
      axis.title = ggplot2::element_text(size = 13),
      plot.title = ggplot2::element_text(size = 15, face = "bold", hjust = 0.5),
      plot.subtitle = ggplot2::element_text(size = 11, hjust = 0.5, color = "gray40")
    )
}
```

## 1.6 Data Prep — Filter to Base Fits

Phase A has `single`, `c_grid`, and `tau_grid` exploration methods. For
most analyses (Q1–Q5), we focus on **base fits** (`agg_method` is NA or
`explore_method == "base"`). The functional prior analyses (Q6–Q7) use
both base fits and max-ELBO aggregated fits.

```{r data-prep}
# Base individual fits only (no aggregated rows)
confusion_base <- confusion %>%
  dplyr::filter(is.na(agg_method) | agg_method == "none")

# For non-functional use cases: single exploration only (no c/tau grid)
confusion_nonfunc <- confusion_base %>%
  dplyr::filter(use_case_id %in% nonfunctional_ids)

# All fits including aggregated (max_elbo rows) for functional analysis
confusion_with_agg <- confusion %>%
  dplyr::filter(
    is.na(agg_method) | agg_method == "none" | agg_method == "max_elbo"
  )

cat("Base fits:", nrow(confusion_base), "rows\n")
cat("Non-functional base fits:", nrow(confusion_nonfunc), "rows\n")
cat("With aggregation rows:", nrow(confusion_with_agg), "rows\n")

# Quick sanity: use case counts
confusion_base %>%
  dplyr::count(use_case_id) %>%
  dplyr::arrange(dplyr::desc(n)) %>%
  print()
```

---

# Section 2: Overview Heatmaps

Top-level view of all 9 use cases across p\* and y_noise. Shows both
AUPRC and Power @ FDR to orient the reader before detailed analyses.

## 2.1 AUPRC Heatmap: Use Case × p_star × y_noise

```{r heatmap-auprc, fig.width=16, fig.height=10}
auprc_heat_data <- test_susine::compute_auprc_from_confusion(
  confusion_base %>% dplyr::filter(use_case_id %in% all_ids),
  group_vars = c("use_case_id", "p_star", "y_noise")
)

if (nrow(auprc_heat_data) > 0) {
  p_heat_auprc <- ggplot2::ggplot(
    auprc_heat_data,
    ggplot2::aes(x = factor(p_star), y = use_case_id, fill = AUPRC)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.5) +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.3f", AUPRC)),
                       size = 3, color = "white") +
    ggplot2::facet_wrap(~paste0("y_noise = ", y_noise), ncol = 3) +
    ggplot2::scale_fill_viridis_c(option = "magma", direction = -1) +
    ggplot2::scale_y_discrete(labels = short_labels) +
    ggplot2::labs(
      title = "AUPRC: All Use Cases \u00d7 p* \u00d7 y_noise",
      x = "p*", y = NULL, fill = "AUPRC"
    ) +
    theme_phase_a() +
    ggplot2::theme(legend.position = "right",
                   axis.text.y = ggplot2::element_text(size = 10))

  save_or_show(p_heat_auprc, "heatmap_auprc_pstar_ynoise.png",
               width = 16, height = 10, subdir = "phase_a")
}
```

## 2.2 Power @ FDR Heatmap: Use Case × p_star × y_noise

```{r heatmap-power-fdr, fig.width=16, fig.height=10}
curves_all <- test_susine::compute_power_fdr_curves(
  confusion_base %>% dplyr::filter(use_case_id %in% all_ids),
  group_vars = c("use_case_id", "p_star", "y_noise")
)

power_fdr_heat <- power_at_fdr(curves_all, fdr_thresh = fdr_threshold,
                                group_cols = c("use_case_id", "p_star", "y_noise"))

if (nrow(power_fdr_heat) > 0) {
  p_heat_power <- ggplot2::ggplot(
    power_fdr_heat,
    ggplot2::aes(x = factor(p_star), y = use_case_id, fill = power_at_fdr)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.5) +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.3f", power_at_fdr)),
                       size = 3, color = "white") +
    ggplot2::facet_wrap(~paste0("y_noise = ", y_noise), ncol = 3) +
    ggplot2::scale_fill_viridis_c(option = "viridis", direction = -1) +
    ggplot2::scale_y_discrete(labels = short_labels) +
    ggplot2::labs(
      title = paste0("Power @ FDR \u2264 ", fdr_threshold,
                     ": All Use Cases \u00d7 p* \u00d7 y_noise"),
      x = "p*", y = NULL, fill = "Power"
    ) +
    theme_phase_a() +
    ggplot2::theme(legend.position = "right",
                   axis.text.y = ggplot2::element_text(size = 10))

  save_or_show(p_heat_power, "heatmap_power_at_fdr_pstar_ynoise.png",
               width = 16, height = 10, subdir = "phase_a")
}
```

---

# Section 3: Q1 — Non-Functional Use Case Ranking

Which non-functional use cases perform best, split by simulation
settings?

## 3.1 AUPRC by p_star

```{r q1-auprc-pstar, fig.width=12, fig.height=6}
auprc_by_pstar <- test_susine::compute_auprc_from_confusion(
  confusion_nonfunc,
  group_vars = c("use_case_id", "p_star")
)

p_q1a <- ggplot2::ggplot(
  auprc_by_pstar,
  ggplot2::aes(x = factor(p_star), y = AUPRC, color = use_case_id,
               group = use_case_id)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_manual(
    values = use_case_palette[nonfunctional_ids],
    labels = short_labels[nonfunctional_ids],
    name = "Use Case"
  ) +
  ggplot2::labs(
    title = "AUPRC by p* — Non-Functional Use Cases",
    x = "p* (number of causal effects)", y = "AUPRC"
  ) +
  theme_phase_a() +
  ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

save_or_show(p_q1a, "q1_auprc_by_pstar.png", width = 12, height = 6,
             subdir = "phase_a")
```

## 3.2 AUPRC by y_noise

```{r q1-auprc-ynoise, fig.width=12, fig.height=6}
auprc_by_ynoise <- test_susine::compute_auprc_from_confusion(
  confusion_nonfunc,
  group_vars = c("use_case_id", "y_noise")
)

p_q1b <- ggplot2::ggplot(
  auprc_by_ynoise,
  ggplot2::aes(x = y_noise, y = AUPRC, color = use_case_id,
               group = use_case_id)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_manual(
    values = use_case_palette[nonfunctional_ids],
    labels = short_labels[nonfunctional_ids],
    name = "Use Case"
  ) +
  ggplot2::labs(
    title = "AUPRC by Noise Level — Non-Functional Use Cases",
    x = "y_noise (residual variance fraction)", y = "AUPRC"
  ) +
  theme_phase_a() +
  ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

save_or_show(p_q1b, "q1_auprc_by_ynoise.png", width = 12, height = 6,
             subdir = "phase_a")
```

## 3.3 Power @ FDR by p_star

```{r q1-power-pstar, fig.width=12, fig.height=6}
curves_nonfunc_settings <- test_susine::compute_power_fdr_curves(
  confusion_nonfunc,
  group_vars = c("use_case_id", "p_star")
)

power_q1 <- power_at_fdr(curves_nonfunc_settings, fdr_thresh = fdr_threshold,
                          group_cols = c("use_case_id", "p_star"))

p_q1c <- ggplot2::ggplot(
  power_q1,
  ggplot2::aes(x = factor(p_star), y = power_at_fdr, color = use_case_id,
               group = use_case_id)
) +
  ggplot2::geom_line(linewidth = 0.9) +
  ggplot2::geom_point(size = 3) +
  ggplot2::scale_color_manual(
    values = use_case_palette[nonfunctional_ids],
    labels = short_labels[nonfunctional_ids],
    name = "Use Case"
  ) +
  ggplot2::labs(
    title = paste0("Power @ FDR \u2264 ", fdr_threshold, " by p*"),
    x = "p*", y = paste0("Power (FDR \u2264 ", fdr_threshold, ")")
  ) +
  theme_phase_a() +
  ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

save_or_show(p_q1c, "q1_power_at_fdr_by_pstar.png", width = 12, height = 6,
             subdir = "phase_a")
```

## 3.4 Power vs FDR Curves Faceted by p_star × y_noise

```{r q1-power-fdr-faceted, fig.width=14, fig.height=10}
curves_by_settings <- test_susine::compute_power_fdr_curves(
  confusion_nonfunc,
  group_vars = c("use_case_id", "p_star", "y_noise")
) %>%
  dplyr::filter(pip_threshold >= 0.05, pip_threshold <= 0.95)

pstar_show <- c(1, 5, 20)
curves_facet <- curves_by_settings %>%
  dplyr::filter(p_star %in% pstar_show) %>%
  dplyr::mutate(
    p_star_label = paste0("p* = ", p_star),
    y_noise_label = paste0("y_noise = ", y_noise)
  )

if (nrow(curves_facet) > 0) {
  p_q1d <- ggplot2::ggplot(
    curves_facet,
    ggplot2::aes(x = fdr, y = power, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.8) +
    ggplot2::facet_grid(p_star_label ~ y_noise_label) +
    ggplot2::scale_color_manual(
      values = use_case_palette[nonfunctional_ids],
      labels = short_labels[nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "Power vs FDR — Non-Functional, by p* and y_noise",
      x = "FDR", y = "Power"
    ) +
    ggplot2::coord_cartesian(xlim = c(0, 1)) +
    theme_phase_a() +
    ggplot2::guides(color = ggplot2::guide_legend(ncol = 3))

  save_or_show(p_q1d, "q1_power_fdr_faceted_pstar_ynoise.png",
               width = 14, height = 10, subdir = "stratified")
}
```

---

# Section 4: Q2 — EB vs Fixed-Variance Comparison

Don't anchor y-axis at 0 — the range from min to max AUPRC matters more
than the distance from 0. Also show Power @ FDR.

```{r q2-eb-data}
# For each EB pair: compute AUPRC and Power@FDR at each sigma_0_2_scalar
eb_comparison_data <- list()

for (i in seq_len(nrow(eb_pairs))) {
  pair <- eb_pairs[i, ]

  # Fixed-variance: group by sigma_0_2_scalar
  fixed_conf <- confusion_nonfunc %>%
    dplyr::filter(use_case_id == pair$fixed_id)

  fixed_auprc <- test_susine::compute_auprc_from_confusion(
    fixed_conf, group_vars = c("use_case_id", "sigma_0_2_scalar")
  )

  fixed_curves <- test_susine::compute_power_fdr_curves(
    fixed_conf, group_vars = c("use_case_id", "sigma_0_2_scalar")
  )
  fixed_power <- power_at_fdr(fixed_curves, fdr_thresh = fdr_threshold,
                               group_cols = c("use_case_id", "sigma_0_2_scalar"))

  fixed_combined <- dplyr::left_join(fixed_auprc, fixed_power,
                                      by = c("use_case_id", "sigma_0_2_scalar")) %>%
    dplyr::mutate(
      prior_var_type = paste0("fixed (", sigma_0_2_scalar, ")"),
      pair_label = pair$pair_label
    )

  # EB: single value
  eb_conf <- confusion_nonfunc %>%
    dplyr::filter(use_case_id == pair$eb_id)

  eb_auprc <- test_susine::compute_auprc_from_confusion(
    eb_conf, group_vars = "use_case_id"
  )
  eb_curves <- test_susine::compute_power_fdr_curves(
    eb_conf, group_vars = "use_case_id"
  )
  eb_power <- power_at_fdr(eb_curves, fdr_thresh = fdr_threshold,
                            group_cols = "use_case_id")

  eb_combined <- dplyr::left_join(eb_auprc, eb_power, by = "use_case_id") %>%
    dplyr::mutate(
      prior_var_type = "EB",
      pair_label = pair$pair_label,
      sigma_0_2_scalar = NA_real_
    )

  eb_comparison_data[[i]] <- dplyr::bind_rows(fixed_combined, eb_combined)
}

eb_comparison <- dplyr::bind_rows(eb_comparison_data)

cat("=== EB vs Fixed Variance Comparison ===\n")
eb_comparison %>%
  dplyr::select(pair_label, prior_var_type, AUPRC, power_at_fdr) %>%
  dplyr::arrange(pair_label, prior_var_type) %>%
  print(n = 30)
```

```{r q2-eb-plot-auprc, fig.width=14, fig.height=5}
# AUPRC comparison — dot plot with range-based y-axis (no zero anchor)
p_q2a <- ggplot2::ggplot(
  eb_comparison,
  ggplot2::aes(x = prior_var_type, y = AUPRC, color = pair_label)
) +
  ggplot2::geom_point(size = 4) +
  ggplot2::geom_line(ggplot2::aes(group = pair_label), linewidth = 0.5,
                     linetype = "dotted") +
  ggplot2::facet_wrap(~pair_label, scales = "free_x") +
  ggplot2::labs(
    title = "AUPRC: EB vs Fixed Prior Variance",
    x = "Prior Variance Setting", y = "AUPRC"
  ) +
  theme_phase_a() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, hjust = 1),
                 legend.position = "none")

save_or_show(p_q2a, "q2_eb_vs_fixed_auprc.png", width = 14, height = 5,
             subdir = "phase_a")
```

```{r q2-eb-plot-power, fig.width=14, fig.height=5}
p_q2b <- ggplot2::ggplot(
  eb_comparison,
  ggplot2::aes(x = prior_var_type, y = power_at_fdr, color = pair_label)
) +
  ggplot2::geom_point(size = 4) +
  ggplot2::geom_line(ggplot2::aes(group = pair_label), linewidth = 0.5,
                     linetype = "dotted") +
  ggplot2::facet_wrap(~pair_label, scales = "free_x") +
  ggplot2::labs(
    title = paste0("Power @ FDR \u2264 ", fdr_threshold,
                   ": EB vs Fixed Prior Variance"),
    x = "Prior Variance Setting",
    y = paste0("Power (FDR \u2264 ", fdr_threshold, ")")
  ) +
  theme_phase_a() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, hjust = 1),
                 legend.position = "none")

save_or_show(p_q2b, "q2_eb_vs_fixed_power.png", width = 14, height = 5,
             subdir = "phase_a")
```

---

# Section 5: Q3 — Dataset Difficulty Metrics

Aggregate per dataset_bundle (not per matrix) to get ~450 observation
points (10 matrices × 5 p_star × 3 y_noise × 3 seeds). Show both AUPRC
and Power @ FDR.

```{r q3-difficulty-setup}
# Per-bundle AUPRC and power: aggregate confusion bins by everything
# EXCEPT dataset_bundle. This gives one AUPRC per (bundle × use_case).
# We then average across use cases for the difficulty scatter.

# dataset_bundle_id should be in confusion_bins from run_enrich_full
if ("dataset_bundle_id" %in% names(confusion_nonfunc)) {
  auprc_per_bundle <- test_susine::compute_auprc_from_confusion(
    confusion_nonfunc,
    group_vars = c("dataset_bundle_id", "use_case_id")
  )

  # Average across use cases per bundle
  auprc_bundle_mean <- auprc_per_bundle %>%
    dplyr::group_by(dataset_bundle_id) %>%
    dplyr::summarise(mean_AUPRC = mean(AUPRC, na.rm = TRUE), .groups = "drop")

  cat("Per-bundle AUPRCs computed:", nrow(auprc_per_bundle), "rows\n")
  cat("Unique bundles:", nrow(auprc_bundle_mean), "\n")
} else {
  warning("dataset_bundle_id not in confusion data; falling back to matrix_id aggregation.")
  auprc_per_bundle <- test_susine::aggregate_auprc_by_dataset_usecase(confusion_nonfunc)
  auprc_bundle_mean <- auprc_per_bundle %>%
    dplyr::group_by(matrix_id) %>%
    dplyr::summarise(mean_AUPRC = mean(AUPRC, na.rm = TRUE), .groups = "drop")
}
```

```{r q3-join-metrics}
if (!is.null(dataset_metrics)) {
  # dataset_metrics has one row per (dataset_bundle_id × phenotype_seed) or per bundle
  dm_cols <- c("M1", "z_topk_ratio", "z_max_abs", "z_count_abs_gt_3", "z_eff_signals")

  if ("dataset_bundle_id" %in% names(dataset_metrics) &&
      "dataset_bundle_id" %in% names(auprc_bundle_mean)) {
    ds_diff_bundle <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(c("dataset_bundle_id", dm_cols))) %>%
      dplyr::inner_join(auprc_bundle_mean, by = "dataset_bundle_id")
  } else if ("matrix_id" %in% names(dataset_metrics)) {
    ds_diff_bundle <- dataset_metrics %>%
      dplyr::select(dplyr::any_of(c("matrix_id", dm_cols))) %>%
      dplyr::group_by(matrix_id) %>%
      dplyr::summarise(dplyr::across(dplyr::everything(), ~ mean(.x, na.rm = TRUE)),
                       .groups = "drop") %>%
      dplyr::inner_join(auprc_bundle_mean, by = "matrix_id")
  } else {
    ds_diff_bundle <- NULL
    warning("Cannot join dataset_metrics.")
  }

  if (!is.null(ds_diff_bundle)) {
    cat("Difficulty table (per-bundle):", nrow(ds_diff_bundle), "rows\n")
  }
} else {
  ds_diff_bundle <- NULL
  message("No dataset_metrics — skipping difficulty scatter.")
}
```

## 5.1 Bivariate R² Screen

```{r q3-r2-screen}
if (!is.null(ds_diff_bundle) && nrow(ds_diff_bundle) >= 10) {
  z_metrics <- c("z_topk_ratio", "z_max_abs", "z_count_abs_gt_3", "z_eff_signals")

  r2_results <- purrr::map_dfr(z_metrics, function(z_var) {
    f_z <- stats::as.formula(paste("mean_AUPRC ~", z_var))
    r2_z <- tryCatch(summary(stats::lm(f_z, data = ds_diff_bundle))$r.squared,
                     error = function(e) NA_real_)
    r2_m1 <- tryCatch(summary(stats::lm(mean_AUPRC ~ M1, data = ds_diff_bundle))$r.squared,
                      error = function(e) NA_real_)
    f_both <- stats::as.formula(paste("mean_AUPRC ~ M1 +", z_var))
    r2_both <- tryCatch(summary(stats::lm(f_both, data = ds_diff_bundle))$r.squared,
                        error = function(e) NA_real_)

    tibble::tibble(z_metric = z_var, R2_z_only = r2_z, R2_M1_only = r2_m1,
                   R2_M1_plus_z = r2_both, R2_gain = r2_both - r2_m1)
  })

  cat("=== Bivariate R² Screen: M1 + z-metric → mean AUPRC ===\n")
  r2_results %>% dplyr::arrange(dplyr::desc(R2_M1_plus_z)) %>% print()

  best_z <- r2_results %>%
    dplyr::slice_max(R2_M1_plus_z, n = 1) %>%
    dplyr::pull(z_metric)
  cat("\nBest complementary z-metric:", best_z, "\n")
} else {
  best_z <- "z_max_abs"
  message("Insufficient data for R² screen; defaulting to z_max_abs.")
}
```

## 5.2 Per-Bundle Scatter: M1 and Best z-Metric vs AUPRC

```{r q3-scatter, fig.width=12, fig.height=5}
if (!is.null(ds_diff_bundle) && nrow(ds_diff_bundle) >= 3) {
  p_m1 <- ggplot2::ggplot(ds_diff_bundle,
                           ggplot2::aes(x = M1, y = mean_AUPRC)) +
    ggplot2::geom_point(size = 1.8, alpha = 0.5) +
    ggplot2::geom_smooth(method = "lm", se = TRUE, color = "steelblue") +
    ggplot2::labs(title = "M1 vs Mean AUPRC (per bundle)",
                  x = "M1", y = "Mean AUPRC") +
    theme_phase_a()

  p_z <- ggplot2::ggplot(ds_diff_bundle,
                          ggplot2::aes(x = .data[[best_z]], y = mean_AUPRC)) +
    ggplot2::geom_point(size = 1.8, alpha = 0.5) +
    ggplot2::geom_smooth(method = "lm", se = TRUE, color = "darkred") +
    ggplot2::labs(title = paste(best_z, "vs Mean AUPRC (per bundle)"),
                  x = best_z, y = "Mean AUPRC") +
    theme_phase_a()

  p_q3 <- cowplot::plot_grid(p_m1, p_z, ncol = 2)
  save_or_show(p_q3, "q3_difficulty_scatter_per_bundle.png",
               width = 12, height = 5, subdir = "difficulty")
}
```

---

# Section 6: Q4 — Use Case Performance vs M1 (Line Graph)

Colored line graph: x = M1, y = metric, color = use case.
Per-matrix AUPRC and Power @ FDR for each non-functional use case.

```{r q4-line-setup}
# Per-matrix AUPRC by use case
auprc_per_matrix_uc <- test_susine::aggregate_auprc_by_dataset_usecase(
  confusion_nonfunc
) %>%
  dplyr::filter(use_case_id %in% nonfunctional_ids)

# Add M1 to each matrix
if (!is.null(dataset_metrics)) {
  m1_by_matrix <- dataset_metrics %>%
    dplyr::select(dplyr::any_of(c("matrix_id", "M1"))) %>%
    dplyr::distinct() %>%
    dplyr::group_by(matrix_id) %>%
    dplyr::summarise(M1 = mean(M1, na.rm = TRUE), .groups = "drop")

  if (!"matrix_id" %in% names(m1_by_matrix) && !is.null(run_table)) {
    m1_by_matrix <- dataset_metrics %>%
      dplyr::left_join(
        run_table %>% dplyr::distinct(dataset_bundle_id, matrix_id),
        by = "dataset_bundle_id"
      ) %>%
      dplyr::group_by(matrix_id) %>%
      dplyr::summarise(M1 = mean(M1, na.rm = TRUE), .groups = "drop")
  }

  auprc_m1 <- auprc_per_matrix_uc %>%
    dplyr::inner_join(m1_by_matrix, by = "matrix_id")
} else {
  auprc_m1 <- NULL
}
```

## 6.1 AUPRC vs M1 by Use Case

```{r q4-auprc-m1, fig.width=12, fig.height=7}
if (!is.null(auprc_m1) && nrow(auprc_m1) > 0) {
  p_q4a <- ggplot2::ggplot(
    auprc_m1,
    ggplot2::aes(x = M1, y = AUPRC, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 2.5) +
    ggplot2::scale_color_manual(
      values = use_case_palette[nonfunctional_ids],
      labels = short_labels[nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "AUPRC vs LD Complexity (M1) — Non-Functional",
      x = "M1 (LD complexity)", y = "AUPRC"
    ) +
    theme_phase_a() +
    ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

  save_or_show(p_q4a, "q4_auprc_vs_m1_lines.png", width = 12, height = 7,
               subdir = "difficulty")
}
```

## 6.2 Power @ FDR vs M1 by Use Case

```{r q4-power-m1, fig.width=12, fig.height=7}
if (!is.null(auprc_m1) && nrow(auprc_m1) > 0) {
  # Compute power@FDR per matrix × use_case
  curves_per_matrix <- test_susine::compute_power_fdr_curves(
    confusion_nonfunc,
    group_vars = c("use_case_id", "matrix_id")
  )
  power_per_matrix <- power_at_fdr(curves_per_matrix, fdr_thresh = fdr_threshold,
                                    group_cols = c("use_case_id", "matrix_id"))

  power_m1 <- power_per_matrix %>%
    dplyr::inner_join(m1_by_matrix, by = "matrix_id") %>%
    dplyr::filter(use_case_id %in% nonfunctional_ids)

  p_q4b <- ggplot2::ggplot(
    power_m1,
    ggplot2::aes(x = M1, y = power_at_fdr, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 2.5) +
    ggplot2::scale_color_manual(
      values = use_case_palette[nonfunctional_ids],
      labels = short_labels[nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = paste0("Power @ FDR \u2264 ", fdr_threshold,
                     " vs M1 — Non-Functional"),
      x = "M1 (LD complexity)",
      y = paste0("Power (FDR \u2264 ", fdr_threshold, ")")
    ) +
    theme_phase_a() +
    ggplot2::guides(color = ggplot2::guide_legend(ncol = 2))

  save_or_show(p_q4b, "q4_power_at_fdr_vs_m1_lines.png", width = 12, height = 7,
               subdir = "difficulty")
}
```

---

# Section 7: Q5 — Functional Prior Comparison

Compare functional priors (mu_0 = c\*a, pi = softmax(|a|/tau)) to the
vanilla (non-functional) baseline across c, tau, annotation_r2,
inflate_match.

## 7.1 AUPRC by c_value (Functional mu)

```{r q6-c-value, fig.width=10, fig.height=6}
confusion_mu <- confusion_base %>%
  dplyr::filter(use_case_id == "susine_functional_mu")

auprc_by_c <- test_susine::compute_auprc_from_confusion(
  confusion_mu, group_vars = c("use_case_id", "c_value")
)

auprc_baseline <- test_susine::compute_auprc_from_confusion(
  confusion_base %>% dplyr::filter(use_case_id == "susie_vanilla"),
  group_vars = "use_case_id"
)

cat("=== AUPRC by c_value (functional mu) ===\n")
print(auprc_by_c)

if (nrow(auprc_by_c) > 0) {
  p_q6c <- ggplot2::ggplot(auprc_by_c, ggplot2::aes(x = c_value, y = AUPRC)) +
    ggplot2::geom_line(linewidth = 1.0,
                       color = use_case_palette["susine_functional_mu"]) +
    ggplot2::geom_point(size = 3.5,
                        color = use_case_palette["susine_functional_mu"]) +
    ggplot2::geom_hline(yintercept = auprc_baseline$AUPRC[1], linetype = "dashed",
                        color = use_case_palette["susie_vanilla"], linewidth = 0.8) +
    ggplot2::annotate("text", x = max(auprc_by_c$c_value, na.rm = TRUE),
                      y = auprc_baseline$AUPRC[1],
                      label = "Vanilla baseline", hjust = 1, vjust = -0.5,
                      color = use_case_palette["susie_vanilla"], size = 3.5) +
    ggplot2::labs(
      title = "AUPRC by Prior Mean Scale (c) — Functional mu",
      subtitle = "mu_0 = c * a; dashed = vanilla SuSiE baseline",
      x = "c (annotation scale)", y = "AUPRC"
    ) + theme_phase_a()

  save_or_show(p_q6c, "q6_auprc_by_c_value.png", width = 10, height = 6,
               subdir = "functional")
}
```

## 7.2 AUPRC by tau_value (Functional pi)

```{r q6-tau-value, fig.width=10, fig.height=6}
confusion_pi <- confusion_base %>%
  dplyr::filter(use_case_id == "susine_functional_pi")

auprc_by_tau <- test_susine::compute_auprc_from_confusion(
  confusion_pi, group_vars = c("use_case_id", "tau_value")
)

cat("=== AUPRC by tau_value (functional pi) ===\n")
print(auprc_by_tau)

if (nrow(auprc_by_tau) > 0) {
  p_q6t <- ggplot2::ggplot(auprc_by_tau, ggplot2::aes(x = tau_value, y = AUPRC)) +
    ggplot2::geom_line(linewidth = 1.0,
                       color = use_case_palette["susine_functional_pi"]) +
    ggplot2::geom_point(size = 3.5,
                        color = use_case_palette["susine_functional_pi"]) +
    ggplot2::geom_hline(yintercept = auprc_baseline$AUPRC[1], linetype = "dashed",
                        color = use_case_palette["susie_vanilla"], linewidth = 0.8) +
    ggplot2::annotate("text", x = max(auprc_by_tau$tau_value, na.rm = TRUE),
                      y = auprc_baseline$AUPRC[1],
                      label = "Vanilla baseline", hjust = 1, vjust = -0.5,
                      color = use_case_palette["susie_vanilla"], size = 3.5) +
    ggplot2::labs(
      title = "AUPRC by Softmax Temperature (tau) — Functional pi",
      subtitle = "pi = softmax(|a|/tau); dashed = vanilla SuSiE baseline",
      x = "tau (softmax temperature)", y = "AUPRC"
    ) + theme_phase_a()

  save_or_show(p_q6t, "q6_auprc_by_tau_value.png", width = 10, height = 6,
               subdir = "functional")
}
```

## 7.3 Functional AUPRC by annotation_r2, Faceted by inflate_match

Baseline (susie_vanilla) has inflate_match = NA so it appears as a
horizontal reference line within each facet, not a separate panel.

```{r q6-annotation-quality, fig.width=14, fig.height=6}
func_conf <- confusion_base %>%
  dplyr::filter(use_case_id %in% functional_ids)

auprc_func_annot <- test_susine::compute_auprc_from_confusion(
  func_conf,
  group_vars = c("use_case_id", "annotation_r2", "inflate_match")
) %>%
  dplyr::filter(!is.na(annotation_r2), !is.na(inflate_match))

if (nrow(auprc_func_annot) > 0) {
  # Baseline AUPRC (inflate_match agnostic)
  baseline_val <- auprc_baseline$AUPRC[1]

  p_q6ann <- ggplot2::ggplot(
    auprc_func_annot,
    ggplot2::aes(x = annotation_r2, y = AUPRC, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_hline(yintercept = baseline_val, linetype = "dashed",
                        color = use_case_palette["susie_vanilla"],
                        linewidth = 0.7) +
    ggplot2::annotate("text", x = 0, y = baseline_val,
                      label = "Vanilla (no annotation)", hjust = 0, vjust = -0.5,
                      color = use_case_palette["susie_vanilla"], size = 3) +
    ggplot2::facet_wrap(~paste("inflate_match =", inflate_match), ncol = 3) +
    ggplot2::scale_color_manual(
      values = use_case_palette[functional_ids],
      labels = short_labels[functional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "Functional AUPRC by Annotation Quality",
      subtitle = "Dashed line = vanilla baseline (no functional info)",
      x = "annotation_r2", y = "AUPRC"
    ) +
    theme_phase_a()

  save_or_show(p_q6ann, "q6_functional_by_annotation_quality.png",
               width = 14, height = 6, subdir = "functional")
}
```

---

# Section 8: Q7 — Max-ELBO Functional vs Non-Functional Baseline

The question: compared to a baseline model with NO functional
information (susie_vanilla), how does max-ELBO aggregated performance
of func-mu and func-pi scale with annotation_r2?

## 8.1 Max-ELBO vs Baseline: Overall

```{r q7-setup}
# Max-ELBO aggregated rows for functional use cases
confusion_agg_func <- confusion_with_agg %>%
  dplyr::filter(use_case_id %in% functional_ids, agg_method == "max_elbo")

# Non-functional baseline: susie_vanilla base fits
confusion_baseline <- confusion_base %>%
  dplyr::filter(use_case_id == "susie_vanilla")

cat("Max-ELBO functional rows:", nrow(confusion_agg_func), "\n")
cat("Baseline rows:", nrow(confusion_baseline), "\n")
```

```{r q7-overall-gain}
auprc_maxelbo <- test_susine::compute_auprc_from_confusion(
  confusion_agg_func, group_vars = "use_case_id"
) %>%
  dplyr::rename(AUPRC_max_elbo = AUPRC)

auprc_baseline_val <- test_susine::compute_auprc_from_confusion(
  confusion_baseline, group_vars = "use_case_id"
)$AUPRC[1]

gain_overall <- auprc_maxelbo %>%
  dplyr::mutate(
    AUPRC_baseline = auprc_baseline_val,
    AUPRC_gain = AUPRC_max_elbo - AUPRC_baseline,
    pct_gain = 100 * AUPRC_gain / AUPRC_baseline,
    label = short_labels[use_case_id]
  )

cat("=== Max-ELBO Functional vs Vanilla Baseline ===\n")
gain_overall %>%
  dplyr::select(label, AUPRC_baseline, AUPRC_max_elbo, AUPRC_gain, pct_gain) %>%
  print()
```

## 8.2 Gain vs annotation_r2

```{r q7-gain-by-annot, fig.width=12, fig.height=6}
auprc_maxelbo_annot <- test_susine::compute_auprc_from_confusion(
  confusion_agg_func,
  group_vars = c("use_case_id", "annotation_r2")
) %>%
  dplyr::rename(AUPRC_max_elbo = AUPRC) %>%
  dplyr::filter(!is.na(annotation_r2))

# Baseline is constant across annotation_r2
gain_annot <- auprc_maxelbo_annot %>%
  dplyr::mutate(
    AUPRC_baseline = auprc_baseline_val,
    AUPRC_gain = AUPRC_max_elbo - AUPRC_baseline,
    label = short_labels[use_case_id]
  )

if (nrow(gain_annot) > 0) {
  p_q7a <- ggplot2::ggplot(
    gain_annot,
    ggplot2::aes(x = annotation_r2, y = AUPRC_gain, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 1.0) +
    ggplot2::geom_point(size = 3.5) +
    ggplot2::geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
    ggplot2::scale_color_manual(
      values = use_case_palette[functional_ids],
      labels = short_labels[functional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "Max-ELBO AUPRC Gain over Vanilla Baseline",
      subtitle = "Gain = AUPRC(max-ELBO functional) - AUPRC(vanilla SuSiE)",
      x = "annotation_r2", y = "AUPRC Gain"
    ) +
    theme_phase_a()

  save_or_show(p_q7a, "q7_maxelbo_vs_baseline_by_annotation.png",
               width = 12, height = 6, subdir = "functional")
}
```

## 8.3 Gain by p_star and y_noise

```{r q7-gain-by-settings, fig.width=14, fig.height=8}
auprc_maxelbo_settings <- test_susine::compute_auprc_from_confusion(
  confusion_agg_func,
  group_vars = c("use_case_id", "p_star", "y_noise")
) %>%
  dplyr::rename(AUPRC_max_elbo = AUPRC)

# Baseline by the same settings
auprc_baseline_settings <- test_susine::compute_auprc_from_confusion(
  confusion_baseline,
  group_vars = c("use_case_id", "p_star", "y_noise")
) %>%
  dplyr::rename(AUPRC_baseline = AUPRC) %>%
  dplyr::select(-use_case_id)

gain_settings <- auprc_maxelbo_settings %>%
  dplyr::left_join(auprc_baseline_settings, by = c("p_star", "y_noise")) %>%
  dplyr::mutate(
    AUPRC_gain = AUPRC_max_elbo - AUPRC_baseline,
    label = short_labels[use_case_id]
  )

if (nrow(gain_settings) > 0) {
  p_q7b <- ggplot2::ggplot(
    gain_settings,
    ggplot2::aes(x = factor(p_star), y = AUPRC_gain, fill = use_case_id)
  ) +
    ggplot2::geom_col(position = ggplot2::position_dodge(width = 0.8),
                      width = 0.7, alpha = 0.85) +
    ggplot2::facet_wrap(~paste("y_noise =", y_noise), ncol = 3) +
    ggplot2::geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
    ggplot2::scale_fill_manual(
      values = use_case_palette[functional_ids],
      labels = short_labels[functional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "Max-ELBO Functional Gain over Vanilla by p* and y_noise",
      subtitle = "Positive = functional max-ELBO beats vanilla baseline",
      x = "p*", y = "AUPRC Gain"
    ) +
    theme_phase_a()

  save_or_show(p_q7b, "q7_maxelbo_vs_baseline_by_pstar_ynoise.png",
               width = 14, height = 8, subdir = "functional")
}
```

---

# Section 9: Additional Analyses

## 9.1 Verification: susine vs susieR Baselines

```{r q_extra_verification}
verification_pairs <- tibble::tribble(
  ~susie_id,        ~susine_id,       ~pair_label,
  "susie_vanilla",  "susine_vanilla", "Vanilla",
  "susie_eb",       "susine_eb",      "EB"
)

for (i in seq_len(nrow(verification_pairs))) {
  pair <- verification_pairs[i, ]
  auprc_pair <- test_susine::aggregate_auprc_by_dataset_usecase(
    confusion_base %>%
      dplyr::filter(use_case_id %in% c(pair$susie_id, pair$susine_id))
  )

  if (nrow(auprc_pair) > 0) {
    wide <- auprc_pair %>%
      tidyr::pivot_wider(id_cols = matrix_id, names_from = use_case_id,
                         values_from = AUPRC)
    if (ncol(wide) >= 3) {
      vals <- wide[[pair$susie_id]]
      ref <- wide[[pair$susine_id]]
      corr <- cor(vals, ref, use = "complete.obs")
      max_diff <- max(abs(vals - ref), na.rm = TRUE)
      cat(sprintf("Verification (%s): corr=%.5f, max|diff|=%.6f\n",
                  pair$pair_label, corr, max_diff))
    }
  }
}
```

## 9.2 Sigma Sensitivity (Fixed-Variance Use Cases)

```{r q_extra_sigma, fig.width=12, fig.height=7}
fixed_nonfunctional_ids <- c("susie_vanilla", "susie_ash_fixed",
                              "susie_inf", "susine_vanilla")

auprc_by_sigma <- test_susine::compute_auprc_from_confusion(
  confusion_nonfunc %>% dplyr::filter(use_case_id %in% fixed_nonfunctional_ids),
  group_vars = c("use_case_id", "sigma_0_2_scalar")
) %>%
  dplyr::filter(!is.na(sigma_0_2_scalar))

if (nrow(auprc_by_sigma) > 0) {
  p_sigma <- ggplot2::ggplot(
    auprc_by_sigma,
    ggplot2::aes(x = sigma_0_2_scalar, y = AUPRC, color = use_case_id)
  ) +
    ggplot2::geom_line(linewidth = 0.9) +
    ggplot2::geom_point(size = 3) +
    ggplot2::scale_color_manual(
      values = use_case_palette[fixed_nonfunctional_ids],
      labels = short_labels[fixed_nonfunctional_ids],
      name = "Use Case"
    ) +
    ggplot2::labs(
      title = "AUPRC Sensitivity to Prior Variance",
      x = expression(sigma[0]^2 ~ "(fraction of var(y))"), y = "AUPRC"
    ) +
    theme_phase_a()

  save_or_show(p_sigma, "extra_sigma_sensitivity.png", width = 12, height = 7,
               subdir = "phase_a")
}
```

---

# Section 10: Summary and Phase B Recommendations

```{r summary-recommendations}
cat("==============================================================\n")
cat("  Phase A Results Summary\n")
cat("==============================================================\n\n")

cat(sprintf("FDR threshold used for Power@FDR: %.2f\n\n", fdr_threshold))

# Heatmap summaries
cat("--- Per-setting heatmaps in plots_phase_a/main/ ---\n")
cat("  heatmap_auprc_pstar_ynoise.png\n")
cat("  heatmap_power_at_fdr_pstar_ynoise.png\n\n")

# EB comparison
cat("--- EB vs Fixed comparison ---\n")
if (exists("eb_comparison")) {
  eb_comparison %>%
    dplyr::select(pair_label, prior_var_type, AUPRC, power_at_fdr) %>%
    dplyr::arrange(pair_label, prior_var_type) %>%
    print(n = 30)
}

# Difficulty
if (exists("best_z")) {
  cat("\n--- Best z-metric for difficulty:", best_z, "---\n")
  if (exists("r2_results")) print(r2_results)
}

# Max-ELBO functional gains
cat("\n--- Max-ELBO Functional vs Baseline ---\n")
if (exists("gain_overall")) {
  gain_overall %>%
    dplyr::select(label, AUPRC_baseline, AUPRC_max_elbo, AUPRC_gain, pct_gain) %>%
    print()
}

cat("\n==============================================================\n")
cat("  Phase B Design Guidance\n")
cat("==============================================================\n")
cat("
Based on Phase A screening:
1. Verify susine/susieR consistency after sigma_0_2 bug fix.
2. Confirm sigma_0^2 sensitivity (pick 1-2 values or keep 3).
3. Set c-grid and tau-grid ranges based on Q6 results.
4. Use best z-metric for difficulty stratification.
5. Decide which use cases to keep/drop for exploration budget.
")
```
