---
title: "SuSiNE EB Method Ablation — Visualization"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
editor_options:
  markdown:
    wrap: 72
---

Summary: Visualizes aggregated outputs from the EB ablation study.
Key question: which EB strategy is best for (1) prior variance only,
(2) annotation scale (c) only, and (3) joint mean + variance?
Moderators: annotation quality (annotation_r2, inflate_match),
architecture (sparse / susie2_sparse / susie2_oligogenic), and
the nonneg constraint on c.

Run collect_results_workbook.Rmd first to produce aggregated CSVs.

# Section 1: Setup

```{r setup, message=FALSE, warning=FALSE}
here::i_am("vignettes/simulation pipeline/visualize_results_workbook_eb_ablation.Rmd")
library(here)
library(devtools)
devtools::load_all(".")

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(stringr)
library(purrr)
library(scales)
```

## 1.1 Job Configuration

```{r job-config}
# UPDATE these two values after each run.
job_name      <- "eb_ablation_v2"
parent_job_id <- "REPLACE_WITH_PARENT_JOB_ID"
output_root   <- here("output")

paths <- test_susine::resolve_job_paths(
  job_name      = job_name,
  parent_job_id = parent_job_id,
  output_root   = output_root
)
```

## 1.2 Plot Directories

```{r plot-dirs}
save_plots_to_file    <- TRUE
show_in_rstudio_plots <- TRUE

if (isTRUE(getOption("knitr.in.progress"))) {
  knitr::opts_chunk$set(fig.ext = "png")
} else if (show_in_rstudio_plots && interactive()) {
  knitr::opts_chunk$set(fig.show = "hide", dev = "RStudioGD", fig.ext = "png")
} else {
  knitr::opts_chunk$set(fig.ext = "png")
}

plot_base_dir <- file.path(paths$aggregated_dir, "plots_eb_ablation")
plot_dirs <- list(
  overview    = file.path(plot_base_dir, "overview"),
  no_ann      = file.path(plot_base_dir, "no_annotation"),
  ann_eb      = file.path(plot_base_dir, "annotation_eb"),
  nonneg      = file.path(plot_base_dir, "nonneg"),
  diagnostics = file.path(plot_base_dir, "prior_diagnostics"),
  tables      = file.path(plot_base_dir, "tables")
)
for (d in plot_dirs) dir.create(d, showWarnings = FALSE, recursive = TRUE)
cat("Plot directories created under:", plot_base_dir, "\n")

maybe_show_plot <- function(p) {
  if (show_in_rstudio_plots && interactive()) try(print(p), silent = TRUE)
}

save_or_show <- function(p, filename, width = 12, height = 7,
                          dpi = 300, subdir = "overview") {
  maybe_show_plot(p)
  if (save_plots_to_file) {
    fp <- file.path(plot_dirs[[subdir]], filename)
    tryCatch(
      { ggsave(fp, p, width = width, height = height, dpi = dpi, bg = "white")
        cat("Saved:", filename, "\n") },
      error = function(e) cat("Error saving:", filename, "-", conditionMessage(e), "\n")
    )
  } else {
    tryCatch(print(p), error = function(e) message("Cannot display plot."))
  }
}
```

## 1.3 Use Case Categories and Palettes

```{r uc-categories}
# 23 use cases grouped by EB strategy role.
uc_category <- c(
  susie_vanilla                         = "baseline",
  susine_vanilla                        = "baseline",
  susie_eb                              = "eb_variance",
  susine_eb                             = "eb_variance",
  susine_eb_mean                        = "eb_mean",
  susine_eb_mu_var                      = "eb_joint_no_ann",
  susine_eb_scale                       = "scale_only",
  susine_eb_scale_nonneg                = "scale_only",
  susine_eb_scale_var                   = "ann_eb_joint",
  susine_eb_scale_var_nonneg            = "ann_eb_joint",
  susine_eb_penalized_scale_var         = "ann_eb_joint",
  susine_eb_penalized_scale_var_nonneg  = "ann_eb_joint",
  susine_eb_scale_var_outer             = "ann_eb_joint",
  susine_eb_scale_var_outer_nonneg      = "ann_eb_joint",
  susine_eb_clamped_scale_var           = "ann_eb_joint",
  susine_eb_clamped_scale_var_nonneg    = "ann_eb_joint",
  susine_eb_alternating_scale_var       = "ann_eb_joint",
  susine_eb_alternating_scale_var_nonneg = "ann_eb_joint",
  susine_functional_mu                  = "fixed_c_grid"
)

# Whether the use case requires annotation
uc_needs_annotation <- c(
  susie_vanilla                         = FALSE,
  susine_vanilla                        = FALSE,
  susie_eb                              = FALSE,
  susine_eb                             = FALSE,
  susine_eb_mean                        = FALSE,
  susine_eb_mu_var                      = FALSE,
  susine_eb_scale                       = TRUE,
  susine_eb_scale_nonneg                = TRUE,
  susine_eb_scale_var                   = TRUE,
  susine_eb_scale_var_nonneg            = TRUE,
  susine_eb_penalized_scale_var         = TRUE,
  susine_eb_penalized_scale_var_nonneg  = TRUE,
  susine_eb_scale_var_outer             = TRUE,
  susine_eb_scale_var_outer_nonneg      = TRUE,
  susine_eb_clamped_scale_var           = TRUE,
  susine_eb_clamped_scale_var_nonneg    = TRUE,
  susine_eb_alternating_scale_var       = TRUE,
  susine_eb_alternating_scale_var_nonneg = TRUE,
  susine_functional_mu                  = TRUE
)

# Category-level palette
category_palette <- c(
  baseline        = "#636363",
  eb_variance     = "#2171b5",
  eb_mean         = "#74c476",
  eb_joint_no_ann = "#238b45",
  scale_only      = "#fd8d3c",
  ann_eb_joint    = "#cb181d",
  fixed_c_grid    = "#984ea3"
)

# Per-use-case short labels for axis display
uc_short_label <- c(
  susie_vanilla                          = "SuSiE",
  susine_vanilla                         = "SuSiNE",
  susie_eb                               = "SuSiE+EB_var",
  susine_eb                              = "SuSiNE+EB_var",
  susine_eb_mean                         = "EB_mean",
  susine_eb_mu_var                       = "EB_mu+var",
  susine_eb_scale                        = "EB_scale",
  susine_eb_scale_nonneg                 = "EB_scale+",
  susine_eb_scale_var                    = "scale_var",
  susine_eb_scale_var_nonneg             = "scale_var+",
  susine_eb_penalized_scale_var          = "pen_sv",
  susine_eb_penalized_scale_var_nonneg   = "pen_sv+",
  susine_eb_scale_var_outer              = "outer_sv",
  susine_eb_scale_var_outer_nonneg       = "outer_sv+",
  susine_eb_clamped_scale_var            = "clamp_sv",
  susine_eb_clamped_scale_var_nonneg     = "clamp_sv+",
  susine_eb_alternating_scale_var        = "alt_sv",
  susine_eb_alternating_scale_var_nonneg = "alt_sv+",
  susine_functional_mu                   = "fixed_c_grid"
)
```

---

# Section 2: Load Data

```{r load-data, message=FALSE}
agg <- paths$aggregated_dir

conf_path        <- file.path(agg, "confusion_matrix_detailed.csv")
dataset_path     <- file.path(agg, "dataset_metrics.csv")
multimodal_path  <- file.path(agg, "multimodal_metrics.csv")
model_path       <- file.path(agg, "model_metrics.csv")
prior_diag_path  <- file.path(agg, "prior_diagnostics.csv")

if (!file.exists(conf_path))    stop("Missing confusion_matrix_detailed.csv — run collect workbook first.")
if (!file.exists(dataset_path)) stop("Missing dataset_metrics.csv.")

conf_raw        <- readr::read_csv(conf_path,       show_col_types = FALSE)
dataset_metrics <- readr::read_csv(dataset_path,    show_col_types = FALSE)

multimodal_metrics <- if (file.exists(multimodal_path)) {
  readr::read_csv(multimodal_path, show_col_types = FALSE)
} else { message("multimodal_metrics.csv not found."); NULL }

model_metrics <- if (file.exists(model_path)) {
  readr::read_csv(model_path, show_col_types = FALSE)
} else { message("model_metrics.csv not found."); NULL }

prior_diagnostics <- if (file.exists(prior_diag_path)) {
  readr::read_csv(prior_diag_path, show_col_types = FALSE)
} else { message("prior_diagnostics.csv not found."); NULL }

cat("Confusion rows:", nrow(conf_raw), "\n")
cat("Dataset rows:  ", nrow(dataset_metrics), "\n")
```

## 2.1 Peek at Column Structure

```{r peek-columns}
# Inspect key columns before filtering — adjust filters below if values differ.
cat("--- confusion_matrix_detailed columns ---\n")
print(names(conf_raw))

if (!is.null(conf_raw) && "variant_type" %in% names(conf_raw)) {
  cat("\nvariant_type values:\n")
  print(table(conf_raw$variant_type, useNA = "ifany"))
}
if (!is.null(conf_raw) && "agg_method" %in% names(conf_raw)) {
  cat("\nagg_method values:\n")
  print(table(conf_raw$agg_method, useNA = "ifany"))
}
if (!is.null(prior_diagnostics)) {
  cat("\n--- prior_diagnostics columns ---\n")
  print(names(prior_diagnostics))
}
```

---

# Section 3: Data Preparation

```{r data-prep}
# Primary confusion filter:
#  - variant_type: keep causal-SNP rows (adjust if column has different name/value)
#  - agg_method: for single-fit use cases pick "single" or "max_elbo";
#                for c_grid (susine_functional_mu) pick "elbo_softmax"
#  Adjust these strings to match what you see in 2.1 above.
primary_variant_type <- "causal"   # or "all" / "purity_filtered" — check peek
primary_agg_single   <- "single"   # single-run methods
primary_agg_cgrid    <- "elbo_softmax"

conf <- conf_raw %>%
  dplyr::filter(
    .data$variant_type == primary_variant_type | !("variant_type" %in% names(.))
  ) %>%
  dplyr::filter(
    (.data$use_case_id == "susine_functional_mu" & .data$agg_method == primary_agg_cgrid) |
    (.data$use_case_id != "susine_functional_mu" & .data$agg_method == primary_agg_single) |
    !("agg_method" %in% names(.))
  ) %>%
  dplyr::mutate(
    eb_category = dplyr::recode(.data$use_case_id, !!!uc_category, .default = "other"),
    needs_annotation = uc_needs_annotation[.data$use_case_id],
    short_label = dplyr::recode(.data$use_case_id, !!!uc_short_label, .default = .data$use_case_id),
    nonneg = stringr::str_ends(.data$use_case_id, "_nonneg"),
    # Paired method name for nonneg comparisons
    method_base = stringr::str_remove(.data$use_case_id, "_nonneg$")
  )

cat("Rows after primary filter:", nrow(conf), "\n")
cat("use_case_id values present:\n")
print(sort(unique(conf$use_case_id)))
```

```{r compute-auprc}
# AUPRC per (use_case_id, architecture, annotation_r2, inflate_match, sigma_0_2_scalar)
# — this is the core per-dataset-scenario summary metric.
auprc_group_vars <- intersect(
  c("use_case_id", "architecture", "annotation_r2", "inflate_match",
    "sigma_0_2_scalar", "matrix_id", "y_noise", "p_star", "L"),
  names(conf)
)

auprc_tbl <- test_susine::compute_auprc_from_confusion(conf, group_vars = auprc_group_vars) %>%
  dplyr::mutate(
    eb_category  = dplyr::recode(.data$use_case_id, !!!uc_category,     .default = "other"),
    short_label  = dplyr::recode(.data$use_case_id, !!!uc_short_label,  .default = .data$use_case_id),
    nonneg       = stringr::str_ends(.data$use_case_id, "_nonneg"),
    method_base  = stringr::str_remove(.data$use_case_id, "_nonneg$"),
    needs_annotation = uc_needs_annotation[.data$use_case_id]
  )

cat("AUPRC rows:", nrow(auprc_tbl), "\n")
```

---

# Section 4: Overview — AUPRC by EB Category

```{r auprc-overview, message=FALSE}
# Collapse over annotation_r2/inflate_match for overview (annotation-free methods
# don't have these; annotation-based methods are averaged over quality levels here).
p_overview <- ggplot(
  auprc_tbl,
  aes(x = reorder(short_label, AUPRC, FUN = median, na.rm = TRUE),
      y = AUPRC,
      fill = eb_category)
) +
  geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
  scale_fill_manual(values = category_palette) +
  facet_wrap(~ architecture, ncol = 1) +
  coord_flip() +
  labs(
    title = "AUPRC by use case (primary agg method)",
    x = NULL, y = "AUPRC", fill = "EB category"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", strip.text = element_text(face = "bold"))

save_or_show(p_overview, "auprc_overview_by_architecture.png",
             width = 10, height = 14, subdir = "overview")
```

---

# Section 5: No-Annotation EB Methods

Compares: vanilla → EB variance → EB mean → EB joint (no annotations).

```{r no-ann-subset}
no_ann_ids <- c("susie_vanilla", "susine_vanilla", "susie_eb",
                "susine_eb", "susine_eb_mean", "susine_eb_mu_var")

auprc_no_ann <- auprc_tbl %>%
  dplyr::filter(.data$use_case_id %in% no_ann_ids)

no_ann_order <- c("SuSiE", "SuSiNE", "SuSiE+EB_var", "SuSiNE+EB_var",
                  "EB_mean", "EB_mu+var")
auprc_no_ann <- auprc_no_ann %>%
  dplyr::mutate(
    short_label = factor(.data$short_label, levels = no_ann_order)
  )
```

```{r no-ann-plot, message=FALSE}
p_no_ann <- ggplot(
  auprc_no_ann,
  aes(x = short_label, y = AUPRC, fill = eb_category)
) +
  geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
  scale_fill_manual(values = category_palette) +
  facet_wrap(~ architecture) +
  labs(
    title = "AUPRC: no-annotation EB methods",
    subtitle = "Does EB variance/mean/joint estimation improve over vanilla?",
    x = NULL, y = "AUPRC", fill = "EB category"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "bottom")

save_or_show(p_no_ann, "auprc_no_annotation_eb.png", subdir = "no_ann")
```

```{r no-ann-by-pstar, message=FALSE}
# Facet by p_star to see if EB methods help more in oligogenic settings.
if ("p_star" %in% names(auprc_no_ann) && any(!is.na(auprc_no_ann$p_star))) {
  p_no_ann_pstar <- ggplot(
    auprc_no_ann %>% dplyr::filter(!is.na(.data$p_star)),
    aes(x = short_label, y = AUPRC, fill = eb_category)
  ) +
    geom_boxplot(outlier.size = 0.5, alpha = 0.8) +
    scale_fill_manual(values = category_palette) +
    facet_grid(architecture ~ p_star, labeller = label_both) +
    labs(
      title = "AUPRC vs p* (no-annotation EB methods)",
      x = NULL, y = "AUPRC", fill = "EB category"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 35, hjust = 1),
          legend.position = "bottom")

  save_or_show(p_no_ann_pstar, "auprc_no_annotation_by_pstar.png",
               width = 16, height = 10, subdir = "no_ann")
}
```

---

# Section 6: Annotation-Based EB Methods (scale_var variants)

Compares all annotation-based joint EB methods against susine_vanilla and
susine_functional_mu (fixed c-grid) baselines.

```{r ann-eb-subset}
ann_eb_ids <- c(
  "susine_vanilla",
  "susine_functional_mu",
  "susine_eb_scale_var",          "susine_eb_scale_var_nonneg",
  "susine_eb_penalized_scale_var", "susine_eb_penalized_scale_var_nonneg",
  "susine_eb_scale_var_outer",    "susine_eb_scale_var_outer_nonneg",
  "susine_eb_clamped_scale_var",  "susine_eb_clamped_scale_var_nonneg",
  "susine_eb_alternating_scale_var", "susine_eb_alternating_scale_var_nonneg"
)

auprc_ann_eb <- auprc_tbl %>%
  dplyr::filter(
    .data$use_case_id %in% ann_eb_ids,
    !is.na(.data$annotation_r2) | .data$use_case_id == "susine_vanilla"
  )
```

```{r ann-eb-by-r2, message=FALSE}
# AUPRC by method × annotation_r2, faceted by inflate_match.
p_ann_r2 <- ggplot(
  auprc_ann_eb %>% dplyr::filter(!is.na(.data$annotation_r2)),
  aes(x = factor(annotation_r2), y = AUPRC,
      fill = short_label, color = short_label)
) +
  geom_boxplot(alpha = 0.5, outlier.size = 0.4, position = position_dodge(0.8)) +
  facet_grid(architecture ~ inflate_match, labeller = label_both) +
  labs(
    title = "Annotation-based EB methods: AUPRC vs annotation quality",
    subtitle = "Faceted by inflate_match (annotation inflation factor)",
    x = "annotation_r2", y = "AUPRC",
    fill = "Method", color = "Method"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        strip.text = element_text(face = "bold"))

save_or_show(p_ann_r2, "auprc_ann_eb_by_r2_inflate.png",
             width = 16, height = 12, subdir = "ann_eb")
```

```{r ann-eb-winner, message=FALSE}
# Median AUPRC per method × annotation_r2 — which method wins most often?
winner_tbl <- auprc_ann_eb %>%
  dplyr::filter(!is.na(.data$annotation_r2)) %>%
  dplyr::group_by(.data$use_case_id, .data$short_label,
                  .data$architecture, .data$annotation_r2, .data$inflate_match) %>%
  dplyr::summarise(
    median_AUPRC = median(.data$AUPRC, na.rm = TRUE),
    n            = dplyr::n(),
    .groups = "drop"
  )

readr::write_csv(winner_tbl,
  file.path(plot_dirs$tables, "ann_eb_winner_table.csv"))
cat("Wrote ann_eb_winner_table.csv\n")

# Heatmap: method × annotation_r2 median AUPRC (best architecture only)
best_arch <- winner_tbl %>%
  dplyr::group_by(.data$architecture) %>%
  dplyr::summarise(mean_auprc = mean(.data$median_AUPRC, na.rm = TRUE)) %>%
  dplyr::slice_max(.data$mean_auprc, n = 1) %>%
  dplyr::pull(.data$architecture)

p_heatmap <- winner_tbl %>%
  dplyr::filter(.data$architecture == best_arch[1]) %>%
  ggplot(aes(x = factor(annotation_r2), y = short_label, fill = median_AUPRC)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "plasma", name = "Median\nAUPRC") +
  facet_wrap(~ inflate_match, labeller = label_both) +
  labs(
    title = sprintf("Median AUPRC heatmap — annotation EB methods (%s)", best_arch[1]),
    x = "annotation_r2", y = NULL
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))

save_or_show(p_heatmap, "auprc_heatmap_ann_eb.png",
             width = 14, height = 8, subdir = "ann_eb")
```

---

# Section 7: Scale-Only EB (sigma grid)

`susine_eb_scale` fixes mu_0 and optimizes only c; run across sigma_0_2_scalar grid.

```{r scale-only-plot, message=FALSE}
auprc_scale <- auprc_tbl %>%
  dplyr::filter(.data$use_case_id %in% c("susine_eb_scale", "susine_eb_scale_nonneg",
                                          "susine_vanilla"))

if (nrow(auprc_scale) > 0 && "sigma_0_2_scalar" %in% names(auprc_scale)) {
  p_scale <- ggplot(
    auprc_scale %>% dplyr::filter(!is.na(.data$annotation_r2) |
                                    .data$use_case_id == "susine_vanilla"),
    aes(x = factor(sigma_0_2_scalar), y = AUPRC, fill = short_label)
  ) +
    geom_boxplot(alpha = 0.7, outlier.size = 0.4) +
    facet_grid(architecture ~ annotation_r2, labeller = label_both) +
    labs(
      title = "Scale-only EB (c optimized, sigma fixed): AUPRC vs sigma_0_2",
      x = "sigma_0_2_scalar", y = "AUPRC", fill = "Method"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")

  save_or_show(p_scale, "auprc_scale_only_eb.png",
               width = 14, height = 10, subdir = "ann_eb")
}
```

---

# Section 8: nonneg Constraint Effect

Paired comparison: for each annotation-based method, does the nonneg
constraint on c help or hurt?

```{r nonneg-comparison, message=FALSE}
# Pivot wide: AUPRC for unconstrained vs nonneg, matched by dataset scenario.
paired_vars <- intersect(
  c("architecture", "annotation_r2", "inflate_match",
    "sigma_0_2_scalar", "matrix_id", "y_noise", "p_star", "L", "method_base"),
  names(auprc_tbl)
)

auprc_nonneg <- auprc_tbl %>%
  dplyr::filter(.data$eb_category == "ann_eb_joint") %>%
  dplyr::select(dplyr::all_of(c(paired_vars, "nonneg", "AUPRC"))) %>%
  tidyr::pivot_wider(
    names_from  = "nonneg",
    values_from = "AUPRC",
    names_prefix = "nonneg_"
  ) %>%
  dplyr::rename(AUPRC_unconstrained = nonneg_FALSE,
                AUPRC_nonneg        = nonneg_TRUE) %>%
  dplyr::mutate(delta_nonneg = .data$AUPRC_nonneg - .data$AUPRC_unconstrained)

if (nrow(auprc_nonneg) > 0 && "AUPRC_unconstrained" %in% names(auprc_nonneg)) {
  p_nonneg_scatter <- ggplot(
    auprc_nonneg,
    aes(x = AUPRC_unconstrained, y = AUPRC_nonneg,
        color = method_base)
  ) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
    geom_point(alpha = 0.5, size = 0.8) +
    facet_wrap(~ architecture) +
    labs(
      title = "nonneg constraint: AUPRC head-to-head per method",
      subtitle = "Above diagonal = nonneg helps",
      x = "AUPRC (unconstrained)", y = "AUPRC (nonneg)",
      color = "Base method"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 8))

  save_or_show(p_nonneg_scatter, "nonneg_head_to_head.png",
               width = 12, height = 8, subdir = "nonneg")

  p_nonneg_delta <- ggplot(
    auprc_nonneg,
    aes(x = delta_nonneg, fill = method_base)
  ) +
    geom_histogram(bins = 30, alpha = 0.7) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    facet_grid(method_base ~ architecture, scales = "free_y") +
    labs(
      title = "AUPRC delta (nonneg - unconstrained) by method",
      x = "Delta AUPRC", y = "Count", fill = "Base method"
    ) +
    theme_minimal() +
    theme(legend.position = "none",
          strip.text.y = element_text(size = 7))

  save_or_show(p_nonneg_delta, "nonneg_delta_histogram.png",
               width = 14, height = 10, subdir = "nonneg")
}
```

---

# Section 9: Power-FDR Curves

```{r pwr-fdr-no-ann, message=FALSE}
# No-annotation methods overall.
conf_no_ann <- conf %>%
  dplyr::filter(.data$use_case_id %in% no_ann_ids)

if (nrow(conf_no_ann) > 0) {
  curves_no_ann <- test_susine::compute_power_fdr_curves(
    conf_no_ann,
    group_vars = c("use_case_id", "architecture")
  )

  no_ann_pal <- setNames(
    scales::hue_pal()(length(no_ann_ids)),
    no_ann_ids
  )

  p_curves_no_ann <- ggplot(
    curves_no_ann,
    aes(x = fdr, y = power, color = use_case_id)
  ) +
    geom_line(linewidth = 0.8) +
    facet_wrap(~ architecture) +
    labs(
      title = "Power vs FDR — no-annotation EB methods",
      x = "FDR (1 - Precision)", y = "Power (Recall)",
      color = "Use case"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")

  save_or_show(p_curves_no_ann, "power_fdr_no_annotation.png",
               subdir = "overview")
}
```

```{r pwr-fdr-ann-eb, message=FALSE}
# Annotation EB methods at best annotation quality.
best_r2 <- if ("annotation_r2" %in% names(conf))
  max(conf$annotation_r2, na.rm = TRUE) else NA

conf_ann_best <- conf %>%
  dplyr::filter(
    .data$use_case_id %in% ann_eb_ids,
    is.na(.data$annotation_r2) | .data$annotation_r2 == best_r2
  )

if (nrow(conf_ann_best) > 0) {
  curves_ann <- test_susine::compute_power_fdr_curves(
    conf_ann_best,
    group_vars = c("use_case_id", "architecture")
  )

  p_curves_ann <- ggplot(curves_ann, aes(x = fdr, y = power, color = use_case_id)) +
    geom_line(linewidth = 0.7) +
    facet_wrap(~ architecture) +
    labs(
      title = sprintf("Power vs FDR — annotation EB methods (r2 = %.1f)", best_r2),
      x = "FDR", y = "Power", color = "Use case"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 7))

  save_or_show(p_curves_ann, "power_fdr_annotation_eb_best_r2.png",
               subdir = "ann_eb")
}
```

---

# Section 10: Prior Diagnostics

Visualizes what EB methods actually learned: distribution of c_l and
sigma_0_2_l per effect across datasets.

```{r prior-diag-check}
if (!is.null(prior_diagnostics) && nrow(prior_diagnostics) > 0) {
  cat("prior_diagnostics columns:\n")
  print(names(prior_diagnostics))
  cat("\nRows:", nrow(prior_diagnostics), "\n")
  cat("use_case_id values:\n")
  print(table(prior_diagnostics$use_case_id, useNA = "ifany"))
}
```

```{r prior-diag-c, message=FALSE}
if (!is.null(prior_diagnostics)) {
  c_col <- intersect(c("c_l", "c_value", "c"), names(prior_diagnostics))[1]
  sig_col <- intersect(c("sigma_0_2_l", "sigma_0_2", "sigma_l"), names(prior_diagnostics))[1]

  if (!is.na(c_col)) {
    pd <- prior_diagnostics %>%
      dplyr::mutate(
        short_label = dplyr::recode(.data$use_case_id, !!!uc_short_label,
                                    .default = .data$use_case_id),
        eb_category = dplyr::recode(.data$use_case_id, !!!uc_category, .default = "other")
      ) %>%
      dplyr::filter(!is.na(.data[[c_col]]))

    p_c_dist <- ggplot(pd, aes(x = .data[[c_col]], fill = eb_category)) +
      geom_histogram(bins = 40, alpha = 0.8) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "gray30") +
      scale_fill_manual(values = category_palette) +
      facet_wrap(~ short_label, scales = "free_y") +
      labs(
        title = sprintf("Learned c_l distribution ('%s') by EB method", c_col),
        x = "Learned c per effect", y = "Count",
        fill = "Category"
      ) +
      theme_minimal() +
      theme(legend.position = "bottom",
            strip.text = element_text(size = 8))

    save_or_show(p_c_dist, "c_l_distribution_by_method.png",
                 width = 16, height = 12, subdir = "diagnostics")
  } else {
    message("No c_l-like column found in prior_diagnostics. Columns: ",
            paste(names(prior_diagnostics), collapse = ", "))
  }

  if (!is.na(sig_col)) {
    pd_sig <- prior_diagnostics %>%
      dplyr::filter(!is.na(.data[[sig_col]])) %>%
      dplyr::mutate(
        short_label = dplyr::recode(.data$use_case_id, !!!uc_short_label,
                                    .default = .data$use_case_id)
      )

    p_sig_dist <- ggplot(pd_sig, aes(x = .data[[sig_col]], fill = short_label)) +
      geom_histogram(bins = 40, alpha = 0.8, show.legend = FALSE) +
      facet_wrap(~ short_label, scales = "free") +
      labs(
        title = sprintf("Learned sigma_0_2 distribution ('%s') by EB method", sig_col),
        x = "Learned sigma_0_2 per effect", y = "Count"
      ) +
      theme_minimal() +
      theme(strip.text = element_text(size = 8))

    save_or_show(p_sig_dist, "sigma_0_2_l_distribution_by_method.png",
                 width = 16, height = 12, subdir = "diagnostics")
  }
}
```

---

# Section 11: Summary Tables

```{r summary-table}
auprc_summary <- auprc_tbl %>%
  dplyr::group_by(.data$use_case_id, .data$short_label, .data$eb_category,
                  .data$architecture) %>%
  dplyr::summarise(
    n             = dplyr::n(),
    median_AUPRC  = median(.data$AUPRC, na.rm = TRUE),
    mean_AUPRC    = mean(.data$AUPRC,   na.rm = TRUE),
    sd_AUPRC      = sd(.data$AUPRC,     na.rm = TRUE),
    q25_AUPRC     = quantile(.data$AUPRC, 0.25, na.rm = TRUE),
    q75_AUPRC     = quantile(.data$AUPRC, 0.75, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(.data$architecture, dplyr::desc(.data$median_AUPRC))

readr::write_csv(auprc_summary, file.path(plot_dirs$tables, "auprc_summary.csv"))
cat("Wrote auprc_summary.csv\n")

# Print top methods per architecture
auprc_summary %>%
  dplyr::group_by(.data$architecture) %>%
  dplyr::slice_max(.data$median_AUPRC, n = 5) %>%
  dplyr::select(.data$architecture, .data$short_label, .data$eb_category,
                .data$median_AUPRC, .data$n) %>%
  knitr::kable(digits = 4, caption = "Top 5 methods by median AUPRC per architecture")
```

```{r annotation-eb-summary}
# Best annotation-based EB method per (annotation_r2, inflate_match, architecture).
if ("annotation_r2" %in% names(auprc_tbl)) {
  best_ann_eb <- auprc_tbl %>%
    dplyr::filter(.data$eb_category == "ann_eb_joint",
                  !is.na(.data$annotation_r2)) %>%
    dplyr::group_by(.data$use_case_id, .data$short_label, .data$architecture,
                    .data$annotation_r2, .data$inflate_match) %>%
    dplyr::summarise(
      median_AUPRC = median(.data$AUPRC, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::group_by(.data$architecture, .data$annotation_r2, .data$inflate_match) %>%
    dplyr::slice_max(.data$median_AUPRC, n = 1) %>%
    dplyr::ungroup()

  readr::write_csv(best_ann_eb, file.path(plot_dirs$tables, "best_ann_eb_by_quality.csv"))
  cat("Wrote best_ann_eb_by_quality.csv\n")
  print(best_ann_eb)
}
```
